### 2019-07-22 main.Rmd
updating section on linear granularity

---
title: Exploring probability distributions for bivariate temporal granularities
authors:
- name: Sayani Gupta
  affiliation: Department of Econometrics and Business Statistics, Monash University
  email: earo.wang@monash.edu

- name: Rob J Hyndman
  affiliation: Department of Econometrics and Business Statistics, Monash University

- name: Dianne Cook
  affiliation: Department of Econometrics and Business Statistics, Monash University

keywords:
- data visualization
- statistical distributions
- time granularities
- periodicties
- grammar of graphics
- R

abstract: |
  Smart meters measure energy usage at fine temporal scales, and are now installed in many households around the world. We propose some new tools to explore this type of data, which deconstruct time in many different ways. There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week. Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.

  The hierarchical structure of many granularities creates a natural nested ordering. For example, hours are nested within days, days within weeks, weeks within months, and so on. We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities. For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.

   Visualizing data across various granularities helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated within the tidy workflow, so that probability distributions can be examined using the range of graphics available in the [ggplot2](https://cran.r-project.org/package=ggplot2) package. 
bibliography: bibliography.bib
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
  cache = FALSE, cache.path = 'cache/', 
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

```{r external, include = FALSE}
#read_chunk('scripts/main.R')
```

```{r load}
```

# Introduction

<!--temporal granularities and why should we care --> 
Temporal data can be available at various resolution depending on the context. Social and economic data are often collected and reported at coarses temporal scales like monthly, quarterly or annually. But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible. For example, it might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimisation every second.  As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases. For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters. This approach requires deconstructing time in various possible ways.

A temporal granularity which results from such a deconstruction may be intuitively described as a sequence of time granules, each one consisting of a set of time instants. There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week. Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.

It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the observed data. This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.

<!--motivation -->

The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world. Smart meters currently provide half-hourly use in kwh for each household, from the time that they were installed, some as early as 2012. Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.

It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate. But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.

We propose that the analysis of probability distributions of smart meter data at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA). EDA calls for utilizing visualization and transformation to explore data systematically. It is a process of generating hypothesis, testing them and consequently refining them through investigations.

<!--details abt time gran --> 
The hierarchical structure of many granularities creates a natural nested ordering. For example, hours are nested within days, days within weeks, weeks within months, and so on. We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities. For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.

<!--existing closely related work and how ours relate --> 
[**lubridate**]([https://cran.r-project.org/package=lubridate)#acad paper in R allows creation of granularities that are mostly single-order-up like hour of the day, second of the minute. This paper utilises the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.

<!--actions on multiple order up - usage --> 
Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. However, this approach can lead to a myriad of choices all of which are not useful. Analysts are expected to iteratively visualise these choices for exploring possible patterns in the data. But too many choices might leave him bewildered.

<!-- challenges and how you are dealing wih it --> 
This work provides tools for systematically exploring bivariate granularities within the tidy workflow through proper study of what can be considered a prospective graphic for exploration. Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis. Probability distributions can be examined using the range of graphics available in the [ggplot2](https://cran.r-project.org/package=ggplot2) package.

In particular, this work provides the following tools.

  * Functions to create multiple-order-up time granularities. This is an extension to the [lubridate]([https://cran.r-project.org/package=lubridate) package, which allows for the creation of some calendar categorizations, usually single-order-up.

  * Checks on the feasibility of creating plots or drawing inferences from two granularities together. Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
```

(ref:ped-map-cap) Google map of the Melbourne city area, dots indicate all the sensor locations. The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.

```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot)"}
```

(ref:time-series-plot) Time series plots showing 2016 pedestrian counts, measured by three different sensors in the city of Melbourne. Small multiples of lines show that the foot traffic varies at one location to another. The spike in counts at the State Library corresponds to the timing of the event "White Night", where there were many people taking part in activities in the city throughout the night. A relatively persistent pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather noisy and spiky, with several runs of missing records.

```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
```

(ref:facet-time-cap) Hourly pedestrian counts for 2016, faceted by sensors, and days of the week. The focus is on time of day and day of week across the sensors. Daily commuter patterns at Flagstaff Station, the variability of the foot traffic at Birrarung Marr, and the consistent pedestrian behavior at the State Library, can be seen.

The work is inspired by @Wickham2012glyph, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps. It is also related to recent work by @R-geofacet which provides methods in the **geofacet** R package to arrange data plots into a grid, while preserving the geographical position. Both of these show data in a spatial context.

In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events. There are some existing works in this area. For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days. It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office. @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively. However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day. Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation. Color-encoding is also low on the hierarchy of optimal variable mapping [@cleveland1984graphical; @lam2007overview].

The proposed algorithm goes beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers. The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)). Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes. An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

# Computing multiple order-up granularities recursively from single order-up granularities {#sec:algorithm}

## Time granularities {#sec:timegran}

Time can be represented at varied levels of abstraction depending on the accuracy required for the context. Time granularity can be defined as the resolution power of the temporal qualification of a statement. Providing a formalism with the concept
of time granularity makes it possible to model time information with respect to differently
grained temporal domains.
### Linear time granularities{#linear-gran-def}

Linear granularities are defined in @Bettini1998-ed, as follows.

**Definition:** A **time domain** is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.

A time domain can be **discrete** (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be **dense** (if it is an infinite set). A time domain is assumed to be discrete for the purpose of our discussion.

**Definition:** A linear **granularity** is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:

  (C1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
than all elements of $G(j)$, and  
  (C2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.

**Definition:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.

The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The notation for creating these cells is as follows:

* $k = 1, \dots , 7$ is the day of the week, that is the first day of the month.
* $d = 28, 29, 30$ or $31$ representing the number of days in any month.
* $(i, j)$ is the grid position where $1 \le i \le 5$, is week within the month, $1 \le j \le 7$, is day of the week.
* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.

The grid position for any day in the month is given by
\begin{equation}
  \begin{aligned}
  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
  j &= g \text{ mod } 7.
  \end{aligned}
  (\#eq:grid)
\end{equation}

\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.

```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
include_graphics("img/month.png")
```

### Formal conceptualisation of circular time granularities{#cirgran-def}

Suppose we have a tsibble [@Wang2019-aw] with a time index in one column and keys and variables in other columns.  A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index.  A linear granularity is a mapping of row numbers to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years.

For circular granularities, we need a symbolic representation of time to represent periodicity, which we call 'calendar categorization'. Anything that maps a time index to a categorical variable can be considered a calendar categorization. The number of categories is essentially the periodicity of a circular time granularity.

We want to use modular arithmetic on the domain of the circular granularity to define the calendar categorization. Hence, we start with the definition of equivalence classes and then move on to define a circular granularity.

**Definition: Equivalence class** Let $m \in N \backslash{0}$. For any $a \in Z$ (set of integers), $[a]$ = {$b \in Z | a \equiv (b \mod m)$} where $[a]$ is defined as the equivalence class to which a belongs.

The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$. Thus $Z_m = \{[0], [1], ..., [m-1]\}$. However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.

<!-- A **calendar domain** is a triplet $(T; \le; M, \le)$ where $T$ is a non-empty set of time instants, m is a modular period, M is the ring of integers modulo m denoted by the set ${1,2,\dot,(m-1)}$ $\le$ is a total order on the set $M$ and   -->

<!-- (Notes: a totally ordered set requires that every element in the set is comparable) -->

<!-- A **calendar domain** is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a **total** order on $T$. -->

**Definition:** A **circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.

For example, suppose $C$ is a circular granularity denoting Hour-of-Day and we have hourly data for 100 hours. The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.

**Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once.

**Definition:** A **circular granule** represents an equivalence class inside each cycle.

A few categorizations are listed below using modular arithmetic with appropriate period to illustrate further. We assume that the time index is in hours, and $n_i$ is the number of categories created by $C_i$. Then the following categorizations can be computed.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lr@{~}lr@{~}r}
\toprule
HOD: & $C_1(s)$ & $= s \mod 24$ & $n_1 =$&$24$ \\
HOW: & $C_2(s)$ & $= s \mod `r 24*7`$ & $n_2=$&$`r 24*7`$\\
HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\
HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\
DOW: & $C_5(s)$ & $= \lfloor s/24\rfloor \mod 7$ & $n_5=$&$7$\\
DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\
DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\
WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\
WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\
MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\
\bottomrule
\end{tabular}
\end{center}
\caption{Illustrative calendar categorizations with hourly time index}
\label{definitions}
\end{table}

Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year.

## Reasons to use calendar-based graphics

The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization. It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live. Comparing the conventional displays (Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time)) with the new display (Figure \@ref(fig:facet)), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events. For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.

\newpage

# Case study {#sec:case}

```{r load-elec, cache.extra = tools::md5sum("data/elec.rds")}
```

The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia. Individuals can download their own data from the energy supplier, and the data used in this section is sourced from four colleagues of the authors. The calendar display is useful to help people understand their energy use. The data contains half-hourly electricity consumption in the first half of 2018. The analysis begins by looking at the distribution over days of week, then time of day split by work days and non-work days, followed by the calendar display to inspect the daily schedules.

Figure \@ref(fig:dow) shows the energy use across days of week using boxplots. Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the week days than other households. The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends. The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns.

```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households. Suggested by the medians, household 3 uses more energy than the others on the week days. By contrast, household 2 sees considerably larger variability."}
```

Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend. Household 1 is an early riser, starting their day before 6am and going back home around 6pm on week days. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between week days and weekends. As for household 4, their home routine starts after 6pm on week days. Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

```{r hod, fig.height = 6, fig.cap = "(ref:hod-cap)"}
```

(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. All households have different week days versus weekends daily routines. On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.

```{r calendar-elec, fig.height = 10.5, fig.cap = "(ref:calendar-elec-cap)"}
```

(ref:calendar-elec-cap) Calendar displays faceted by each household using global scales. Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.

Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage. Glancing over household 1, their overall energy use is relatively low. Their week day energy use is distinguishable from their weekends, indicating a working household. The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends. In contrast, household 2 keeps a cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in mid-January. These observations help to explain the stripes and clusters of household 2 in Figure \@ref(fig:hod). It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months. Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture. However, the calendar plots speak the stories that are untold by previous plots, for example, their vacation time. Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June. Further, household 3 takes one short trip in January and the another one starting in the fourth week of June. Household 4 is away over two or three weeks in early April and late June. They all tend to take breaks during June probably due to the fact that the University winter break starts in June.

# Harmony and clashes based on conditioning - issue of data structure

# Visualisation

# Case study: Analysis on cricket {#cricket}

# Case study: Analysis on smart meter data {#smartmeter}

# Discussion {#sec:discussion}

# Acknowledgements {-}

We would like to thank Stuart Lee and Heike Hofmann for their feedback on earlier versions of this work. We thank Thomas Lin Pedersen for pointing out some critical **ggplot2** internals, which makes the `facet_calendar()` implementation possible. We are very grateful to anonymous reviewers for helpful comments that have led to many improvements of the paper. The **sugrrants** R package is available from CRAN <https://CRAN.R-project.org/package=sugrrants> and the development version is available on Github <https://github.com/earowang/sugrrants>. All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.

# Bibliography

<!-- ```{r write-bib} -->
<!-- write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
<!-- ``` -->

### 2019-07-28 main.Rmd
adding to the section how to obtain multiple order up

@@
 
-# Computing multiple order-up granularities recursively from single order-up granularities {#sec:algorithm}
-
-## Time granularities {#sec:timegran}
+# Time granularities{#sec:algorithm}
 
@@
 Time granularity can be defined as the resolution power of the temporal qualification of a statement.
-Providing a formalism with the concept
-of time granularity makes it possible to model time information with respect to differently
-grained temporal domains.
-### Linear time granularities{#linear-gran-def}
+Providing a formalism with the concept of time granularity is important to model time information across differently grained temporal domains{#Euzenat2005-de}.
+
+## Linear time granularities{#linear-gran-def}
 
@@
 
-The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph].
-To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year.
-Each month conforms to a layout of 5 rows and 7 columns, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default.
-These cells provide a micro canvas on which to plot the data.
-The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar.
-Months are of different lengths, ranging from 28 to 31 days.
-Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention.
-The notation for creating these cells is as follows:
+<!--need to include this \autoref{fig:linear-time} shows the implication of this condition.
+If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor s/(24*7)\rfloor$ weeks and so on.
+-->
 
-* $k = 1, \dots , 7$ is the day of the week, that is the first day of the month.
-* $d = 28, 29, 30$ or $31$ representing the number of days in any month.
-* $(i, j)$ is the grid position where $1 \le i \le 5$, is week within the month, $1 \le j \le 7$, is day of the week.
-* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.
+<!-- ```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="\\label{fig:linear-time} The time domain distributed as linear granularities"} -->
+<!-- knitr::include_graphics("Figs/linear-time.png") -->
+<!-- ``` -->
 
-The grid position for any day in the month is given by
-\begin{equation}
-  \begin{aligned}
-  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
-  j &= g \text{ mod } 7.
-  \end{aligned}
-  (\#eq:grid)
-\end{equation}
-
-\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k=5$.
-
-```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
-include_graphics("img/month.png")
-```
-
-### Formal conceptualisation of circular time granularities{#cirgran-def}
+## Formal conceptualisation of circular time granularities{#cirgran-def}
 
@@
 
-## Reasons to use calendar-based graphics
+## Computing multiple order-up granularities from single order-up granularities recursively
 
-The purpose of the calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is consistent with why analysts should and do use data visualization.
-It complements the traditional graphical toolbox used to understand general trends, and better profiles vivid and detailed data stories about the way we live.
-Comparing the conventional displays (Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time)) with the new display (Figure \@ref(fig:facet)), it can be seen that the calendar display is more informatively compelling: when special events happened, and on what day of the week, and whether they were day or night events.
-For example, Figure \@ref(fig:facet) informs the reader that many events were held in Birrarung Marr on weekend days, while September's events took place on Friday evenings, which is difficult to discern from the conventional displays.
+<!-- Often sets come equipped with a natural hierarchical structure.
+For example, the set of natural numbers N is equipped with a natural pre-order structure, where $n\leq n'$ whenever we can find some other number m so that $n+m=n'$.
+Similarly, -->
 
-\newpage
+The hierarchical structure of time creates a natural nested ordering, where hours are nested within days, days within weeks, weeks within months, and so on.
+We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
+For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
 
-# Case study {#sec:case}
+Consider a hierarchy table laying out the structure of all temporal granularities we are interested to look at. The hierarchy table consists of three columns: 
+ - The first column represents the temporal granularities in ascending order. 
+ - The second column represents the constant which relates subsequent granularities.
+ - The third column provides the functional form for single-order up-granularities
 
-```{r load-elec, cache.extra = tools::md5sum("data/elec.rds")}
-```
+Order: Order is defined as the position of the granularity in the hierarchical table.
+If the two granularties have three intermediate granularities in the hierarchical table, then their order difference is 3.
 
-The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia.
-Individuals can download their own data from the energy supplier, and the data used in this section is sourced from four colleagues of the authors.
-The calendar display is useful to help people understand their energy use.
-The data contains half-hourly electricity consumption in the first half of 2018.
-The analysis begins by looking at the distribution over days of week, then time of day split by work days and non-work days, followed by the calendar display to inspect the daily schedules.
+Then it is possible to obtain all multiple-order-up granularities from this hierarchy table using a recursive relation.
 
-Figure \@ref(fig:dow) shows the energy use across days of week using boxplots.
-Inspecting the medians across households tells us that household 3, a family size of one couple and two children, uses more energy over the week days than other households.
-The relatively larger boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends.
-The other two households (1 and 4) tend to consume more energy with more variation on the weekends relative to the week days, reflecting work and leisure patterns.
+Suppose, for $order(x) <  order(y)$ \\
+$f(x, y)$ represents the functional form for computing granularity x_y
+$c(x, y)$ is a constant which relates x and y.
 
-```{r dow, fig.height = 7, fig.cap = "Boxplots of daily energy usage against day of week for four households.
-Suggested by the medians, household 3 uses more energy than the others on the week days.
-By contrast, household 2 sees considerably larger variability."}
-```
+Then, 
+$f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] $, for for $order(x+1) < order(y)$ can be used recursively to obtain any multiple order-up granularities.
+For $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
 
-Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend.
-Household 1 is an early riser, starting their day before 6am and going back home around 6pm on week days.
-They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh.
-On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1.
-A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school.
-They also have a consistent energy pattern between week days and weekends.
-As for household 4, their home routine starts after 6pm on week days.
-Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.
+\begin{equation} \label{eq1}
+\begin{split}
+f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] \\
+  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)[f(x+2, y) - 1] - 1]\\
+  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) - 1] + c(x, x+1)c(x+1, x+2)[f(x+2, y) - 1]
+\end{split}
+\end{equation}
 
-```{r hod, fig.height = 6, fig.cap = "(ref:hod-cap)"}
-```
+# Obtaining multiple-order-up granularities from just granularity which relates the lowest temporal unit to the highest one in the hierarchical table.
 
-(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day.
-All households have different week days versus weekends daily routines.
-On week days, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4.
-The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.
+  In the above equations, it is shown how to obtain f(x,z) given f(x,y) and the hierarchy table,  provided $order(z) > order(y) \neq order(x)$. In general,
 
-```{r calendar-elec, fig.height = 10.5, fig.cap = "(ref:calendar-elec-cap)"}
-```
+  $f(x,z) = \sum_{i=0}^{order(z) - order(y) + 1}c(x, x+i)f(x+1, x+i)$
 
-(ref:calendar-elec-cap) Calendar displays faceted by each household using global scales.
-Long flat low energy usage indicates vacation time, and high energy consumption by household 2 is visible in January and February.
+    We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x)$.
 
-Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage.
-Glancing over household 1, their overall energy use is relatively low.
-Their week day energy use is distinguishable from their weekends, indicating a working household.
-The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends.
-In contrast, household 2 keeps a cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in mid-January.
-These observations help to explain the stripes and clusters of household 2 in Figure \@ref(fig:hod).
-It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months.
-Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture.
-However, the calendar plots speak the stories that are untold by previous plots, for example, their vacation time.
-Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June.
-Further, household 3 takes one short trip in January and the another one starting in the fourth week of June.
-Household 4 is away over two or three weeks in early April and late June.
-They all tend to take breaks during June probably due to the fact that the University winter break starts in June.
+  $f(w,y) = f(x,y) \mod c(x,w)$, for $f(x,y) \mod c(x,w)!=0$ \\
+  $= c(x,w))$, otherwise.
 
-# Harmony and clashes based on conditioning - issue of data structure
+Let us investigate if that is true:
+
+  f(ball, over) = 1
+f(over, quarter) = 4
+f(quarter, semester) = 2
+f(semester, match) = `1
+
+then, f(ball_quarter) = 1 + 6*4 = 25, 
+f(ball, semester) = 1 + 6*4 + 5*6*1 = 55
+f(ball, semester) = 1 + 6*4 + 5*6*1 + 5*6*2*1 = 115
+
+# Harmony and Clashes
 
@@
 
+# Case study: Analysis on smart meter data {#smartmeter}
+
 # Case study: Analysis on cricket {#cricket}
-
-# Case study: Analysis on smart meter data {#smartmeter}
 

### 2019-07-28 main.Rmd
adding brief on visualisation and harmony/clash

@@
 
-Consider a hierarchy table laying out the structure of all temporal granularities we are interested to look at. The hierarchy table consists of three columns: 
+Consider a hierarchy table laying out the structure of all temporal granularities we are interested to look at, consisting of the following  three columns: 
  - The first column represents the temporal granularities in ascending order.   
  - The second column represents the constant which relates subsequent granularities.  
- - The third column provides the functional form for single-order up-granularities
+ - The third column provides the functional form for single-order up-granularities.
 
@@
 f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] \\  
-  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)[f(x+2, y) - 1] - 1]\\
+  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)[f(x+2, y) - 1] - 1]\\ (expanding f(x+1, y))  
   & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) - 1] + c(x, x+1)c(x+1, x+2)[f(x+2, y) - 1]
@@
 
+Suppose, we are not given the intermediate single-order-up granularities but still want all possible multiple order-up granularties.
+In that case, it is necessary to atleast have the have the functional form that relates the smallest temporal unit to the highest temporal unit.
+In that case, we are able to form single-order-up from this functional relationship and still use the recursive relation to get the multiple order-up granularities.
+
   In the above equations, it is shown how to obtain f(x,z) given f(x,y) and the hierarchy table,  provided $order(z) > order(y) \neq order(x)$. In general,
@@
 
-    We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x)$.
+  We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x).
 
@@
 
+We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.
+The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
+
+Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+These are structurally empty sets in that it is impossible for them to have any observations.
+
+Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
+
 # Visualisation
+
+With huge amount of data currently available, it is important to visualize the distribution of the time series variable rather than any particular statistic.
+But which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
+
+Offcourse, the time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
+That neccessiates plotting temporal granularity along the x-axis and the other one across facets.
+
+Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration.
+The harmonies when plotted together can help exploration, however, the clashes were plotted against each other is not appropriate for exploration.
+
+We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.
+Since we are restricting ourselves to 2D space, we assume that the circular granularities are plotted across the aesthetics and facets.
+The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
 

### 2019-07-29 main.Rmd
adding section on visualisation

@@
 All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
+So graphics involving $C_1$ and $C_2$ aare potentially useful.
 
@@
 
-With huge amount of data currently available, it is important to visualize the distribution of the time series variable rather than any particular statistic.
-But which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
+<!-- Why visualise -->
+Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
+However, improper choice of models can lead to wrong predictions.
+One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
+
+<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
+
+<!-- Why distribution plot -->
+<!-- Huge data difficult to summarise using one statistic-->
+But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a dataset.
+Soon enough following questions like below become more interesting:
+
+ - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?
+
+ - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
+More generally, how the variation is the dataset changes across different percentiles/deciles?
+
+  - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median?
+
+This is when displaying a probability distribution becomes a potentially useful approach.
+
+<!-- What all can you see from a distribution plot-->
+
+The entire distribution can be visualized or some contextual summary statistics can be visualised to emphasize certain properties of the distribution.
+These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution.
+They can also be useful in detecting extreme behavior or anolmalies in the dataset.
+
+hich ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
 

### 2019-07-29 main.Rmd
adding bibliography

@@
 
-```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
-```
-
-(ref:ped-map-cap) Google map of the Melbourne city area, dots indicate all the sensor locations.
-The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.
-
-```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot)"}
-```
-
-(ref:time-series-plot) Time series plots showing 2016 pedestrian counts, measured by three different sensors in the city of Melbourne.
-Small multiples of lines show that the foot traffic varies at one location to another.
-The spike in counts at the State Library corresponds to the timing of the event "White Night", where there were many people taking part in activities in the city throughout the night.
-A relatively persistent pattern repeats from one week to another at Flagstaff Station.
-Birrarung Marr looks rather noisy and spiky, with several runs of missing records.
-
-```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
-```
-
-(ref:facet-time-cap) Hourly pedestrian counts for 2016, faceted by sensors, and days of the week.
-The focus is on time of day and day of week across the sensors.
-Daily commuter patterns at Flagstaff Station, the variability of the foot traffic at Birrarung Marr, and the consistent pedestrian behavior at the State Library, can be seen.
-
-The work is inspired by @Wickham2012glyph, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps.
-It is also related to recent work by @R-geofacet which provides methods in the **geofacet** R package to arrange data plots into a grid, while preserving the geographical position.
-Both of these show data in a spatial context.
-
-In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events.
-There are some existing works in this area.
-For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days.
-It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office.
-@Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively.
-However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day.
-Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation.
-Color-encoding is also low on the hierarchy of optimal variable mapping [@cleveland1984graphical; @lam2007overview].
-
-The proposed algorithm goes beyond the calendar-based heatmap.
-The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers.
-The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.
-
-The remainder of the paper is organized as follows.
+<!-- The remainder of the paper is organized as follows.
 Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
@@
 Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.
+-->
 
@@
 The entire distribution can be visualized or some contextual summary statistics can be visualised to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution.
-They can also be useful in detecting extreme behavior or anolmalies in the dataset.
+These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anolmalies in the dataset.
 
-hich ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
+Which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
 

### 2019-07-29 main.Rmd
adding visualisation lit

@@
 
+<!-- Lit review on statistical distribution plot-->
+Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+The box plot is a compact distributional summary, displaying less detail than a histogram.
+Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
+Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
+
+The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.
+Violin plots [@Hintze1998-zi, 1998] display the density for all data points and not only the box.
+The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
+The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
+The probabilities for which the summarization is required can be chosen based on the requirement.
+These regions do not need to be contiguous and help identify multi-modality.
+The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+Because this display just adds extra letter values, it suffers from the same problems as the original box plot, and multimodality is almost impossible to spot[@Wickham_undated-vr].
+
+Finally, a density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution.
+Also, a Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
+Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
+
 Which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.

### 2019-07-30 main.Rmd
minor changes

@@
 
-# Obtaining multiple-order-up granularities from just granularity which relates the lowest temporal unit to the highest one in the hierarchical table.
+Obtaining multiple-order-up granularities from just granularity which relates the lowest temporal unit to the highest one in the hierarchical table.
 
@@
 <!-- Huge data difficult to summarise using one statistic-->
-But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a dataset.
-Soon enough following questions like below become more interesting:
+But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a dataset. Soon enough following questions become more interesting:
 
@@
 <!-- Lit review on statistical distribution plot-->
+## Statistical distribution plots
+
 Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
@@
 
+Moreover, much like the quartiles divide the dataset equally into four equal parts, extensions might include dividing the dataset even further.
+The deciles plots consist of 9 values that split the dataset into ten parts and the percentile plot consists of 99 values that split the dataset into hundred parts.
+A large dataset is required before the extreme percentiles can be estimated with any accuracy.
+
 Finally, a density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution.
@@
 
-Which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
+## Distributions conditional on bivariate granularities
 
-Offcourse, the time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
-That neccessiates plotting temporal granularity along the x-axis and the other one across facets.
+The time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
+That neccessiates plotting one temporal granularity along the x-axis and the other one across facets.
 
-Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration.
-The harmonies when plotted together can help exploration, however, the clashes were plotted against each other is not appropriate for exploration.
+Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
+The harmonies when plotted together can help exploration.
+But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
+This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
 
-We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.
+## Advice algorithm for exploring conditional probability distributions
+
+Recommendations for distribution plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
+They will vary depending on which granularity is placed on the x-axis and which one across facets.
+Assumptions are made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.
+Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot.
+
+<!-- We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.
 Since we are restricting ourselves to 2D space, we assume that the circular granularities are plotted across the aesthetics and facets.
 The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
+-->
 

### 2019-08-05 main.Rmd
minor chnages

@@
   affiliation: Department of Econometrics and Business Statistics, Monash University
-  email: earo.wang@monash.edu
+  email: Sayani.Gupta@monash.edu
 
@@
 Time granularity can be defined as the resolution power of the temporal qualification of a statement.
-Providing a formalism with the concept of time granularity is important to model time information across differently grained temporal domains{#Euzenat2005-de}.
+Providing a formalism with the concept of time granularity is important to model time information across differently grained temporal domains [@Euzenat2005-de].
 
@@
 
-Linear granularities are defined in @Bettini1998-ed, as follows.
+There has been several attempts to provide the framework for formally characterizing time-granularities and identifying their structural properties, relationships and symbolic representations.
+One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
 
@@
 The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+\autoref{fig:linear-time} shows the implication of this condition..
+If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor s/(24*7)\rfloor$ weeks and so on.
 
-<!--need to include this \autoref{fig:linear-time} shows the implication of this condition.
-If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor s/(24*7)\rfloor$ weeks and so on.
--->
+*These definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.*
 
-<!-- ```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="\\label{fig:linear-time} The time domain distributed as linear granularities"} -->
-<!-- knitr::include_graphics("Figs/linear-time.png") -->
-<!-- ``` -->
+```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="\\label{fig:linear-time} The time domain distributed as linear granularities"}
+knitr::include_graphics("Figs/linear-time.png")
+```
 
@@
 Suppose we have a tsibble [@Wang2019-aw] with a time index in one column and keys and variables in other columns.
-A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index.
+A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
 A linear granularity is a mapping of row numbers to subsets of the time domain.
@@
 **Definition: Equivalence class** Let $m \in N \backslash{0}$.
-For any $a \in Z$ (set of integers), $[a]$ = {$b \in Z | a \equiv (b \mod m)$} where $[a]$ is defined as the equivalence class to which a belongs.
+For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
 
@@
 
-## Computing multiple order-up granularities from single order-up granularities recursively
+## Computation of mnultiple order-up granularities from single order-up granularities
 
@@
 
+<!-- What is multiple and single order granularities -->
 The hierarchical structure of time creates a natural nested ordering, where hours are nested within days, days within weeks, weeks within months, and so on.
@@
 
-Consider a hierarchy table laying out the structure of all temporal granularities we are interested to look at, consisting of the following  three columns: 
- - The first column represents the temporal granularities in ascending order.   
- - The second column represents the constant which relates subsequent granularities.  
+In section \@ref(cirgran-def), the definition of calendar categorisation considers index set as set of integers to be able to perform the modular operations.
+However, the index of a tsibble can be any date-time object which comes in different formats.
+[@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data.
+It not only helps to identify and parse different date-time data, it also helps to extract each date-time element with its own accessor function.
+However, table 5 shows that it mostly allows for the creation of some calendar categorizations, usually single-order-up.
+
+##Reproduce Table 5 from the paper
+
+Hence, this section will talk about creating multiple-order up granularities from these single-order-up granularities.
+The hierarchical structure of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.
+
+Consider a hierarchy table laying out the structure of all temporal units we are interested to look at, consisting of the following  three columns: 
+ - The first column represents the temporal units in ascending order of hierarchy.   
+ - The second column represents the constant which relates subsequent temporal units.  
  - The third column provides the functional form for single-order up-granularities.
 
-Order: Order is defined as the position of the granularity in the hierarchical table.
+Order: Order is defined as the position of the temporal units in the hierarchical table.
 If the two granularties have three intermediate granularities in the hierarchical table, then their order difference is 3.
 
-Then it is possible to obtain all multiple-order-up granularities from this hierarchy table using a recursive relation.
-
-Suppose, for $order(x) <  order(y)$ \\
+Suppose, $order(x) <  order(y)$  
 $f(x, y)$ represents the functional form for computing granularity x_y

### 2019-08-05 main.Rmd
multiple from single

@@
 The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
-\autoref{fig:linear-time} shows the implication of this condition..
+\autoref{fig:linear-time} shows the implication of this condition.
 If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor s/(24*7)\rfloor$ weeks and so on.
-
-*These definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.*
 
@@
 
+*These definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.*
+
 ## Formal conceptualisation of circular time granularities{#cirgran-def}
 
-Suppose we have a tsibble [@Wang2019-aw] with a time index in one column and keys and variables in other columns.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
+Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
+A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index.
 A linear granularity is a mapping of row numbers to subsets of the time domain.
@@
 
-For circular granularities, we need a symbolic representation of time to represent periodicity, which we call 'calendar categorization'.
+What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
+Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc.
+Many of these are circular, such as day-of-week, time-of-day.
+Some are nearly circular such as day-of-month.
+Some are irregular such as workingday/non-working day.
+Let's call all of these "calendar categorizations".
 Anything that maps a time index to a categorical variable can be considered a calendar categorization.
+We can specify the categorizations corresponding to circular granularities using modular arithmetic.
 The number of categories is essentially the periodicity of a circular time granularity.
 
-We want to use modular arithmetic on the domain of the circular granularity to define the calendar categorization.
-Hence, we start with the definition of equivalence classes and then move on to define a circular granularity.
+For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by $C_i$. Then the following categorizations can be computed.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lr@{~}lr@{~}r}
+\toprule
+MOH: & $C_1(s)$ & $= s \mod 60$ & $n_1 =$&$60$ \\
+MOD: & $C_2(s)$ & $= s \mod `r 60*24`$ & $n_2=$&$`r 60*24`$\\
+MOW: & $C_3(s)$ & $= s \mod `r 60*24*7`$ & $n_2=$&$`r 60*24*7`$\\
+HOD: & $C_4(s)$ & $= \lfloor s/60\rfloor\mod 24$  & $n_4 =$&$24$ \\
+HOW: & $C_6(s)$ & $= \lfloor s/60\rfloor\mod 24*7$  & $n_5=$&$`r 24*7`$\\
+DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_5=$&$7$\\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Illustrative calendar categorizations with hourly time index}
+\label{definitions}
+\end{table}
+
+We start with the definition of an equivalence class and then move on to define a circular granularity.
 
@@
 
-**Definition:** A **circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.
+**Definition:** A **periodic circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.
 
@@
 
-A few categorizations are listed below using modular arithmetic with appropriate period to illustrate further.
-We assume that the time index is in hours, and $n_i$ is the number of categories created by $C_i$.
-Then the following categorizations can be computed.
+**Definition:** A **aperiodic circular granularity** can not be defined in terms of modular arithmetic since each coarser unit is not always made up of equal number of finer units.
 
@@
 
-## Computation of mnultiple order-up granularities from single order-up granularities
+## Computation of multiple order-up granularities from single order-up granularities
 
@@
 Then, 
-$f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] $, for for $order(x+1) < order(y)$ can be used recursively to obtain any multiple order-up granularities.
+$f(x, y) = f(x, x+1) + c(x, x+1)[f(x+1, y)-1]$, for $order(x+1) < order(y)$ can be used recursively to obtain any multiple order-up granularities.
 For $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
@@
 \begin{split}
-f(x, y) & = f(x, x+1) + c(x, x+1)*[f(x+1, y)-1] \\  
-  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)[f(x+2, y) - 1] - 1]\\ (expanding f(x+1, y))  
-  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) - 1] + c(x, x+1)c(x+1, x+2)[f(x+2, y) - 1]
+f(x, y) & = f(x, x+1) + c(x, x+1)(f(x+1, y)-1) \\  
+  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
+  &\vdots\\
+  & = \sum_{i=0}^{order(y) - order(x) + 1} c(x, x+i)(f(x
+  +i, x+i+1) - 1)\\
+  & = \sum_{i=0}^{order(y) - order(x) + 1} c(x, x+i)(f_1(x+i) - 1)
 \end{split}
 \end{equation}
+
+This is valid in a scenario when the relationship between x and (x+1) is periodic (Define periodic in Definition section).
+However, when x and x+1 is not periodic, then some modification is required in Equation(1).
+
+Consider, a hierarchy table of the form below:
+
+```{r, hierarchy_tbl}
+hierarchy_model <- tibble::tibble(units = c("minute", "hour", "day", "month", "quarter", "semester", "year"), conversion_fac  = c(60, 24, NA, 3, 2, 2, 1))
+knitr::kable(hierarchy_model)
+```
+
+The table suggests that one hour is composed of 60 minutes, 1 day is composed of 24 hours, 1 quarter is composed of 3 months, 1 semester is consists of 2 quarters and 1 year is made up of 2 semesters.
+These partitions are periodic, since the number of finer units within each of these coarses units are same always.
 
@@
 
-  $f(w,y) = f(x,y) \mod c(x,w)$, for $f(x,y) \mod c(x,w)!=0$ \\
+  $f(w,y) = f(x,y) \mod c(x,w)$, for $f(x,y) \mod c(x,w)!=0$
   $= c(x,w))$, otherwise.

### 2019-08-05 main.Rmd
adding section on computation

@@
   echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
-  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
-  cache = FALSE, cache.path = 'cache/', 
+  fig.path = "figure/", fig.align = "center", fig.show = "hold",
+  cache = FALSE, cache.path = "cache/",
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 ```{r external, include = FALSE}
-#read_chunk('scripts/main.R')
+# read_chunk('scripts/main.R')
 ```
@@
 ```{r load}
+
 ```
@@
 
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by $C_i$. Then the following categorizations can be computed.
+For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$. Then the following categorizations can be computed.
 
@@
 MOD: & $C_2(s)$ & $= s \mod `r 60*24`$ & $n_2=$&$`r 60*24`$\\
-MOW: & $C_3(s)$ & $= s \mod `r 60*24*7`$ & $n_2=$&$`r 60*24*7`$\\
+
 HOD: & $C_4(s)$ & $= \lfloor s/60\rfloor\mod 24$  & $n_4 =$&$24$ \\
 HOW: & $C_6(s)$ & $= \lfloor s/60\rfloor\mod 24*7$  & $n_5=$&$`r 24*7`$\\
-DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_5=$&$7$\\
+
+DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_6=$&$7$\\
 \bottomrule
@@
 \end{center}
-\caption{Illustrative calendar categorizations with hourly time index}
+\caption{Illustrative circular granularities with time index in minutes}
 \label{definitions}
@@
 
-We start with the definition of an equivalence class and then move on to define a circular granularity.
+It is easy to note here that due to unequal length of months or years, these formulae can not be used for computing week-of-month or day-of-quarter.
+Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
 
@@
 **Definition:** A **circular granule** represents an equivalence class inside each cycle.
-
-**Definition:** A **aperiodic circular granularity** can not be defined in terms of modular arithmetic since each coarser unit is not always made up of equal number of finer units.
 
@@
 
-<!-- Often sets come equipped with a natural hierarchical structure.
-For example, the set of natural numbers N is equipped with a natural pre-order structure, where $n\leq n'$ whenever we can find some other number m so that $n+m=n'$.
-Similarly, -->
-
-<!-- What is multiple and single order granularities -->
+<!-- What is multiple and single order up granularities -->
 The hierarchical structure of time creates a natural nested ordering, where hours are nested within days, days within weeks, weeks within months, and so on.
@@
 
-In section \@ref(cirgran-def), the definition of calendar categorisation considers index set as set of integers to be able to perform the modular operations.
-However, the index of a tsibble can be any date-time object which comes in different formats.
-[@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data.
-It not only helps to identify and parse different date-time data, it also helps to extract each date-time element with its own accessor function.
-However, table 5 shows that it mostly allows for the creation of some calendar categorizations, usually single-order-up.
+The index of a tsibble can be any date-time object which comes in different formats.
+[@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data and extracting each date-time element with its own accessor function.
+However, table shows that it mostly allows for the creation of single-order-up granularities.
 
-##Reproduce Table 5 from the paper
+```{r table_accessor}
+kable(tibble::tibble(
+  `Categorisation` = c("Second-of-minute", "Minute-of-hour", "Hour-of-day", "Day-of-week", "Day-of-month", "Week-of-year", "Day-of-year","Month-of-year"),
+ Type = c("Periodic", "Periodic", "Periodic", "Periodic", "Aperiodic", "Aperiodic", "Aperiodic", "Periodic"),
+Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "latex", booktabs = TRUE)
+```
+
+Hence, the idea is to create multiple-order-granularities from single-order-up in order to incorporate and be able to access different calendar categorisation.
+The computation of multiple-order-up granularities from single-order-up granularities can differ if single-order-up granularities are circular or aperiodic.
+
+### All single-order-up granularities are periodic:
+
+<!-- In section \@ref(cirgran-def), the definition of calendar categorisation considers index set as set of integers to be able to perform the modular operations. However, -->
 
@@
 \end{equation}
+
+### Single-order-up granularities are mixed - that is circular or periodic:
 

### 2019-08-05 main.Rmd
differentiating circular and aperiodic granularities

@@
 
-## Formal conceptualisation of circular time granularities{#cirgran-def}
+## Formal conceptualisation of calendar categorisations{#cirgran-def}
 
+<!-- Representation with tsibble -->
 Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
@@
 
+<!-- What is calendar categorisation -->
 What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
@@
 Let's call all of these "calendar categorizations".
-Anything that maps a time index to a categorical variable can be considered a calendar categorization.
-We can specify the categorizations corresponding to circular granularities using modular arithmetic.
+Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
+
+<!-- What is circular granularity and some examples -->
+We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
 The number of categories is essentially the periodicity of a circular time granularity.
-
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$. Then the following categorizations can be computed.
+For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
+Then the following categorizations can be computed.
 
@@
 
-It is easy to note here that due to unequal length of months or years, these formulae can not be used for computing week-of-month or day-of-quarter.
+It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
 Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
 
+<!-- Formal definition of circular granularity -->
 **Definition: Equivalence class** Let $m \in N \backslash{0}$.
@@
 
-<!-- A **calendar domain** is a triplet $(T; \le; M, \le)$ where $T$ is a non-empty set of time instants, m is a modular period, M is the ring of integers modulo m denoted by the set ${1,2,\dot,(m-1)}$ $\le$ is a total order on the set $M$ and   -->
-
-<!-- (Notes: a totally ordered set requires that every element in the set is comparable) -->
-
-<!-- A **calendar domain** is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a **total** order on $T$. -->
-
-**Definition:** A **periodic circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.
+**Definition:** A **circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.
 
@@
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-HOD: & $C_1(s)$ & $= s \mod 24$ & $n_1 =$&$24$ \\
-HOW: & $C_2(s)$ & $= s \mod `r 24*7`$ & $n_2=$&$`r 24*7`$\\
-HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\
-HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\
-DOW: & $C_5(s)$ & $= \lfloor s/24\rfloor \mod 7$ & $n_5=$&$7$\\
-DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\
-DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\
-WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\
-WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\
-MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Illustrative calendar categorizations with hourly time index}
-\label{definitions}
-\end{table}
+<!-- Formal definition of aperiodic circular granularity -->
 
-Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year.
+**Defintion:** An **Aperiodic circular granularity** can not be defined using modular arithmetic.
+The modulus for these type of calendar categorisations are not constant due to unequal length of some linear granularities.
+
+<!-- \begin{table}[ht] -->
+<!-- \begin{center} -->
+<!-- \begin{tabular}{lr@{~}lr@{~}r} -->
+<!-- \toprule -->
+<!-- HOD: & $C_1(s)$ & $= s \mod 24$ & $n_1 =$&$24$ \\ -->
+<!-- HOW: & $C_2(s)$ & $= s \mod `r 24*7`$ & $n_2=$&$`r 24*7`$\\ -->
+<!-- HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\ -->
+<!-- HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\ -->
+<!-- DOW: & $C_5(s)$ & $= \lfloor s/24\rfloor \mod 7$ & $n_5=$&$7$\\ -->
+<!-- DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\ -->
+<!-- DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\ -->
+<!-- WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\ -->
+<!-- WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\ -->
+<!-- MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\ -->
+<!-- \bottomrule -->
+<!-- \end{tabular} -->
+<!-- \end{center} -->
+<!-- \caption{Illustrative calendar categorizations with hourly time index} -->
+<!-- \label{definitions} -->
+<!-- \end{table} -->
+
+<!-- Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year. -->
 
@@
 The hierarchical structure of time creates a natural nested ordering, where hours are nested within days, days within weeks, weeks within months, and so on.
-We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
+We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities.
+For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are **single-order-up**.
 
@@
 [@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data and extracting each date-time element with its own accessor function.
-However, table shows that it mostly allows for the creation of single-order-up granularities.
+However, table shows the accessor functions mostly allows for the creation of single-order-up granularities.
 
@@
  Type = c("Periodic", "Periodic", "Periodic", "Periodic", "Aperiodic", "Aperiodic", "Aperiodic", "Periodic"),
-Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "latex", booktabs = TRUE)
+Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "markdown")
 ```
 
-Hence, the idea is to create multiple-order-granularities from single-order-up in order to incorporate and be able to access different calendar categorisation.
-The computation of multiple-order-up granularities from single-order-up granularities can differ if single-order-up granularities are circular or aperiodic.
+The hierarchical nature of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.The idea is to incorporate and access different calendar categorisations to account for varied levels of periodicities in the data.
+The computation of multiple-order-up granularities from single-order-up granularities can differ basis the single-order-up granularities are circular or aperiodic.
 
-### All single-order-up granularities are periodic:
+### All single-order-up granularities are periodic
 
-<!-- In section \@ref(cirgran-def), the definition of calendar categorisation considers index set as set of integers to be able to perform the modular operations. However, -->
+**Definition**: A **hierarchy table** has three columns:  
+ - The first column represents the linear granularities in ascending order of temporal hierarchy.   
+ - The second column represents the constant which relates subsequent linear granularities.  
+ - The third column provides the accessor function for single-order up circular granularities.
 
-Hence, this section will talk about creating multiple-order up granularities from these single-order-up granularities.
-The hierarchical structure of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.
+**Definition**: **Order** is defined as the position of the linear granularities in the hierarchical table. 
+<!-- For example, if two linear granularties have three intermediate granularities in the hierarchical table, then their order difference is 3.  -->
 
-Consider a hierarchy table laying out the structure of all temporal units we are interested to look at, consisting of the following  three columns: 
- - The first column represents the temporal units in ascending order of hierarchy.   
- - The second column represents the constant which relates subsequent temporal units.  
- - The third column provides the functional form for single-order up-granularities.
+<!-- , "month", "quarter","semester", "year"), -->
 
-Order: Order is defined as the position of the temporal units in the hierarchical table.
-If the two granularties have three intermediate granularities in the hierarchical table, then their order difference is 3.
+Suppose, z is the index set of the tsibble, x,y are two linear granularities with  $order(x) <  order(y)$.
+Also, $f_{(x, y)}$ denotes the accessor function for computing circular granularity "x_y" and $c(x, y)$ is a constant which relates x and y.
+It is easy to see that for $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
 
-Suppose, $order(x) <  order(y)$  
-$f(x, y)$ represents the functional form for computing granularity x_y
-$c(x, y)$ is a constant which relates x and y.
-
-Then, 
-$f(x, y) = f(x, x+1) + c(x, x+1)[f(x+1, y)-1]$, for $order(x+1) < order(y)$ can be used recursively to obtain any multiple order-up granularities.
-For $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
+Then, the accessor function f can be used recursively to obtain any multiple-order-up granularities as follows:
 
@@
 \begin{split}
-f(x, y) & = f(x, x+1) + c(x, x+1)(f(x+1, y)-1) \\  
+f_{(x,y)}(z) & = f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, y)}(z)-1) \\  
   & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\
@@
   &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) + 1} c(x, x+i)(f(x
+  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
   +i, x+i+1) - 1)\\
-  & = \sum_{i=0}^{order(y) - order(x) + 1} c(x, x+i)(f_1(x+i) - 1)
+\end{split}
+\end{equation}
+
+To elaborate more, let us consider the following hierarchy table.
+
+```{r table_accessor1}
+kable(tibble::tibble(
+  `linear granularities`  = c("second", "minute", "hhour", "hour", "day", "week"),
+  `Conversion factor` = c(60, 30, 2, 24, 7, 1), 
+  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", 1)), format = "markdown")
+```
+
+From Equation(1), we have
+\begin{equation} \label{eq1}
+\begin{split}
+f_(hhour, week)(z) = f_(hhour, hour)(z) + c(hhour, hour)f_(hour,day) + c(hhour, day)f(day, week)\\
+               = hhour_hour + 2hour_day + 2*24day_week
 \end{split}
@@
 
-We would like to thank Stuart Lee and Heike Hofmann for their feedback on earlier versions of this work.
-We thank Thomas Lin Pedersen for pointing out some critical **ggplot2** internals, which makes the `facet_calendar()` implementation possible.
-We are very grateful to anonymous reviewers for helpful comments that have led to many improvements of the paper.
-The **sugrrants** R package is available from CRAN <https://CRAN.R-project.org/package=sugrrants> and the development version is available on Github <https://github.com/earowang/sugrrants>.
-All materials required to reproduce this article and a history of the changes can be found at the project's Github repository <https://github.com/earowang/paper-calendar-vis>.
-
 # Bibliography

### 2019-08-05 main.Rmd
spell check and mixed single gran intro

@@
 f_{(x,y)}(z) & = f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, y)}(z)-1) \\  
-  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
+  & =  f_{(x, x+1)}(z) + c(x, x+1)[f_{(x+1, x+2)}(z) + c(x+1, x+2)(f_{(x+2, y)}(z) - 1) - 1]\\
+  & =  f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, x+2)}(z) - 1) + c(x, x+1)c(x+1, x+2)(f_{(x+2, y)}(z) - 1)\\
+  & =  f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, x+2)}(z) - 1) + c(x, x+2)(f_{(x+2, y)}(z) - 1)\\
   &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
-  +i, x+i+1) - 1)\\
+  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f_{(x
+  +i, x+i+1)}(z) - 1)\\
 \end{split}
@@
 
-To elaborate more, let us consider the following hierarchy table.
+For illustration, let us consider the following hierarchy table.
 
@@
 \begin{split}
-f_(hhour, week)(z) = f_(hhour, hour)(z) + c(hhour, hour)f_(hour,day) + c(hhour, day)f(day, week)\\
-               = hhour_hour + 2hour_day + 2*24day_week
+f_{(hhour, week)}(z) & = f_{(hhour, hour)}(z) + c(hhour, hour)f_{(hour,day)}(z) + c(hhour, day)f_{(day, week)}(z) \\
+              & = {hhour\_hour}(z) + 2{hour\_day}(z) + 2*24{day\_week}(z) \\
 \end{split}
@@
 
-This is valid in a scenario when the relationship between x and (x+1) is periodic (Define periodic in Definition section).
+<!-- This is valid in a scenario when the relationship between x and (x+1) is periodic (Define periodic in Definition section).
 However, when x and x+1 is not periodic, then some modification is required in Equation(1).
+-->
 
-Consider, a hierarchy table of the form below:
+Suppose we extend the hierarchy table to include the linear granularities month and quarter.
+Since months consists of unequal number of days, any linear granularity which is higher in order than months will also have unequal number of days.
+Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
+However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
 
-```{r, hierarchy_tbl}
-hierarchy_model <- tibble::tibble(units = c("minute", "hour", "day", "month", "quarter", "semester", "year"), conversion_fac = c(60, 24, NA, 3, 2, 2, 1))
-knitr::kable(hierarchy_model)
+```{r table_accessor2}
+kable(tibble::tibble(
+  `linear granularities`  = c("second", "minute", "hhour", "hour", "day", "week", "month", "quarter"),
+  `Conversion factor` = c(60, 30, 2, 24, 7, "aperiodic", 3, 1), 
+  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", "week_month", "month_quarter", 1)), format = "markdown")
 ```
@@
 The table suggests that one hour is composed of 60 minutes, 1 day is composed of 24 hours, 1 quarter is composed of 3 months, 1 semester is consists of 2 quarters and 1 year is made up of 2 semesters.
-These partitions are periodic, since the number of finer units within each of these coarses units are same always.
+These partitions are periodic, since the number of finer units within each of these coarser units are same always.
 
@@
 Suppose, we are not given the intermediate single-order-up granularities but still want all possible multiple order-up granularties.
-In that case, it is necessary to atleast have the have the functional form that relates the smallest temporal unit to the highest temporal unit.
+In that case, it is necessary to at least have the have the functional form that relates the smallest temporal unit to the highest temporal unit.
 In that case, we are able to form single-order-up from this functional relationship and still use the recursive relation to get the multiple order-up granularities.
@@
 All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ aare potentially useful.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
 
-# Visualisation
+# Visualization
 
-<!-- Why visualise -->
+<!-- Why visualize -->
 Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
@@
 <!-- Why distribution plot -->
-<!-- Huge data difficult to summarise using one statistic-->
-But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a dataset. Soon enough following questions become more interesting:
+<!-- Huge data difficult to summaries using one statistic-->
+But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a data set. Soon enough following questions become more interesting:
 
@@
  - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
-More generally, how the variation is the dataset changes across different percentiles/deciles?
+More generally, how the variation is the data set changes across different percentiles/deciles?
 
@@
 
-The entire distribution can be visualized or some contextual summary statistics can be visualised to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anolmalies in the dataset.
+The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
+These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
 
@@
 
-Moreover, much like the quartiles divide the dataset equally into four equal parts, extensions might include dividing the dataset even further.
-The deciles plots consist of 9 values that split the dataset into ten parts and the percentile plot consists of 99 values that split the dataset into hundred parts.
-A large dataset is required before the extreme percentiles can be estimated with any accuracy.
+Moreover, much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
+The deciles plots consist of 9 values that split the data set into ten parts and the percentile plot consists of 99 values that split the data set into hundred parts.
+A large data set is required before the extreme percentiles can be estimated with any accuracy.
 

### 2019-08-05 main.Rmd
Restructing and getting sections sorted to flow the story better

@@
 library(knitr)
+library(tidyverse)
+library(lubridate)
+library(lvplot)
+library(ggridges)
 opts_chunk$set(
@@
 <!--existing closely related work and how ours relate --> 
-[**lubridate**]([https://cran.r-project.org/package=lubridate)#acad paper in R allows creation of granularities that are mostly single-order-up like hour of the day, second of the minute.
+[@Grolemund2011-qf] allows creation of granularities that are mostly single-order-up like hour of the day, second of the minute.
 This paper utilises the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.
@@
 
-## Computation of multiple order-up granularities from single order-up granularities
+## Single-order-up and multiple-order-up granularities
 
@@
 For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are **single-order-up**.
+
+# Computation of multiple order-up granularities from single order-up granularities
 
@@
 
-### All single-order-up granularities are periodic
+## Periodic single-order-up granularities
 
@@
 
-### Single-order-up granularities are mixed - that is circular or periodic:
+## Mixed single-order-up granularities - circular or periodic
 
@@
 kable(tibble::tibble(
-  `linear granularities`  = c("second", "minute", "hhour", "hour", "day", "week", "month", "quarter"),
-  `Conversion factor` = c(60, 30, 2, 24, 7, "aperiodic", 3, 1), 
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", "week_month", "month_quarter", 1)), format = "markdown")
+  `linear granularities`  = c("second", "minute", "hhour", "hour", "day", "month", "quarter"),
+  `Conversion factor` = c(60, 30, 2, 24, "aperiodic", 3, 1), 
+  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_month", "month_quarter", 1)), format = "markdown")
 ```
 
-The table suggests that one hour is composed of 60 minutes, 1 day is composed of 24 hours, 1 quarter is composed of 3 months, 1 semester is consists of 2 quarters and 1 year is made up of 2 semesters.
-These partitions are periodic, since the number of finer units within each of these coarser units are same always.
+There can be three scenarios for obtaining calendar categorisation here:
+- calendar categorisation consisting of two linear granularities whose orders are less than day  
+- calendar categorisation consisting of two linear granularities whose orders are more than month  
+- calendar categorisation consisting of one linear granularity with order at most day and another with order at least month
 
-Obtaining multiple-order-up granularities from just granularity which relates the lowest temporal unit to the highest one in the hierarchical table.
+The calendar categorisation resulting from the first two cases are periodic and has been handled in the earlier section.
 
-Suppose, we are not given the intermediate single-order-up granularities but still want all possible multiple order-up granularties.
+The calendar categorisation resulting from the last case are aperiodic.
+Examples might include day of the quarter or hour of the month.
+In this section, we will see how to obtain aperiodic circular granularities of these types.
+
+\begin{equation} \label{eq1}
+\begin{split}
+f_{(hour, month)}(z) = f_{(hour, day)}(z) + c(hour, day)f_{(day, month)}(z)\\
+\end{split}
+\end{equation}
+
+<!-- Suppose, we are not given the intermediate single-order-up granularities but still want all possible multiple order-up granularties.
 In that case, it is necessary to at least have the have the functional form that relates the smallest temporal unit to the highest temporal unit.
 In that case, we are able to form single-order-up from this functional relationship and still use the recursive relation to get the multiple order-up granularities.
+-->
 
-  In the above equations, it is shown how to obtain f(x,z) given f(x,y) and the hierarchy table,  provided $order(z) > order(y) \neq order(x)$. In general,
+  <!-- In the above equations, it is shown how to obtain f(x,z) given f(x,y) and the hierarchy table,  provided $order(z) > order(y) \neq order(x)$. In general, -->
 
-  $f(x,z) = \sum_{i=0}^{order(z) - order(y) + 1}c(x, x+i)f(x+1, x+i)$
+  <!-- $f(x,z) = \sum_{i=0}^{order(z) - order(y) + 1}c(x, x+i)f(x+1, x+i)$ -->
 
-  We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x).
+  <!-- We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x). -->
 
-  $f(w,y) = f(x,y) \mod c(x,w)$, for $f(x,y) \mod c(x,w)!=0$
-  $= c(x,w))$, otherwise.
+  <!-- $f(w,y) = f(x,y) \mod c(x,w)$, for $f(x,y) \mod c(x,w)!=0$ -->
+  <!-- $= c(x,w))$, otherwise. -->
 
-Let us investigate if that is true:
+### Computing single-order-up granularities:
 
-f(ball, over) = 1
-f(over, quarter) = 4
-f(quarter, semester) = 2
-f(semester, match) = `1
+<!-- f(ball, over) = 1 -->
+<!-- f(over, quarter) = 4 -->
+<!-- f(quarter, semester) = 2 -->
+<!-- f(semester, match) = `1 -->
 
-then, f(ball_quarter) = 1 + 6*4 = 25, 
-f(ball, semester) = 1 + 6*4 + 5*6*1 = 55
-f(ball, semester) = 1 + 6*4 + 5*6*1 + 5*6*2*1 = 115
-
-# Harmony and Clashes
-
-We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.
-The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
-
-Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
-
-Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
+<!-- then, f(ball_quarter) = 1 + 6*4 = 25,  -->
+<!-- f(ball, semester) = 1 + 6*4 + 5*6*1 = 55 -->
+<!-- f(ball, semester) = 1 + 6*4 + 5*6*1 + 5*6*2*1 = 115 -->
 
@@
 
-## Distributions conditional on bivariate granularities
+## Harmony and Clashes
+
+We investigate some combinations of circular time granularities (periodic/aperiodic) which facilitate or hinder exploratory analysis.
+The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
+
+<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+These are structurally empty sets in that it is impossible for them to have any observations.
+-->
+
+<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
+-->
+
+\autoref{fig:allFig} (a) shows the letter value plot of electricity consumption of Victoria across days of the month for few months like January, February, April and December.
+Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
+Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
+b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
+c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
+We can conclude that Month-of-Year and Day-of-Month are harmonies.
+
+\autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
+The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
+All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
+Hence, Day-of-Month and Day-of-Year are clashes.
+
+Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+Violin plots are a combination of box plot and density plot.
+Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
+These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
+
+In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
+The y-axis represents week of the year and the x-axis represents electricity consumption.
+Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
+They are designed to bring out changes in distributions over time.
+The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
+So, Week-of-Month and Week-of-Year are clashes.
+
+\autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
+Decile plots are useful in displaying distribution through deciles without much clutter.
+This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
+So Month-of-Year and Day-of-Year are clashes as well.
+
+From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
+Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
+
+```{r moy-dom,echo=FALSE, out.width = "70%", fig.align = 'center',fig.cap ="Letter-value plot of electricity consumption in Victoria across days of the month\n faceted by months January, February, April and December.
+"}
+
+# change all of these to gravitas function
+VIC <- read.csv("VIC2015/Data/demand_VIC.csv")
+VIC$Date <- as.Date(VIC$Date,origin = "1899-12-30")
+
+first_day_of_month_wday <- function(dx) {
+  day(dx) <- 1
+  wday(dx)
+}
+
+VIC <- VIC %>%mutate(Indx_Year = year(Date),
+                     Indx_Month = month(Date, label = FALSE, abbr = TRUE),
+                     Indx_Wk_Yr = week(Date),
+                     Indx_Wk_Month = ceiling((day(Date) + first_day_of_month_wday(Date) - 1) / 7),
+                     Indx_Day_Week = wday(Date, label=FALSE, abbr=TRUE,
+                                          week_start=1),
+
+                     Indx_Day_Month = day(Date),
+
+                     Indx_Day_Year =  yday(Date),
+
+                     Indx_Weekend=if_else(Indx_Day_Week %in% c(6,7),1,0),
+
+                     Indx_HlHr_day = Period,
+                     month = month(Date, label = FALSE, abbr = TRUE),
+         year =  year(Date),
+         yday =yday(Date),
+         wday = wday(Date, label=FALSE, abbr=TRUE,
+                     week_start=1),
+         bow = (wday - 1) * 48 + Period,
+         dom = day(Date),
+         bom = (dom - 1) * 48 + Period,
+         Weekend=if_else(wday %in% c(6,7),1,0),
+         Indx_hour = ceiling(Period/2),
+         Indx_Hour_Yr = Indx_hour + 24*(yday-1),
+         Indx_Hour_Month = Indx_hour + 24*(Indx_Day_Month-1),
+         Indx_Hour_Wk = Indx_hour + 24*(wday-1))
+
+VIC <- as_tibble(VIC)
+
+```
+
+```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
+(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
+Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
+
+par(mfrow = c(3, 2))
+
+VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Month %in% c(1,2,4,12))%>% ggplot(aes(Indx_Day_Month,OperationalLessIndustrial,group = Indx_Day_Month)) + geom_lv(aes(fill=..LV..),outlier.colour = "red",outlier.shape = 1) + scale_fill_brewer()+ facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks=seq(0,31,5)) + ggtitle("(a) Letter value plot")
+
+VIC_hod_dow <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
+  group_by(wday,Indx_hour) %>%
+  do({x <- .$OperationalLessIndustrial
+  map_dfr(.x = seq(0.1,0.9,0.1),
+          .f = ~ tibble(Quantile = .x,
+                            Value = quantile(x, probs = .x,na.rm=TRUE)))
+  }) %>%  filter(wday %in% c(1,2,6,7))
+
+VIC_hod_dow %>%ggplot(aes(x=Indx_hour,y=Value,col=as.factor(Quantile))) + geom_line() + facet_wrap(~wday)+ scale_x_continuous(breaks=seq(1, 24,5)) + ylab("")  + xlab("Day of the Week") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot")
+
+VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Day_Month %in% c(1,15,29,31))%>% ggplot(aes(as.factor(yday),OperationalLessIndustrial,group = yday)) + geom_boxplot()+ facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
+  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Boxplot")
+
+VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,4))%>% ggplot(aes(as.factor(Indx_Day_Month),OperationalLessIndustrial)) + geom_violin(alpha = 0.03)+ facet_wrap(~Indx_Wk_Month,nrow=3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks=seq(0,31,5))+ scale_y_continuous(breaks = seq(2000,9000,2000))+ ggtitle("(d) Violin plot")
+
+VIC%>% dplyr:::filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,5),Indx_Wk_Yr <20)%>% ggplot(aes(x=OperationalLessIndustrial,y=as.factor(Indx_Wk_Yr),group=Indx_Wk_Yr)) + geom_density_ridges2() +facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000,10000,3000)) + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot")
+
+VIC_moy_doy <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
+  group_by(Indx_Month,yday) %>%
+  do({x <- .$OperationalLessIndustrial
+  map_dfr(.x = seq(0.1,0.9,0.1),
+          .f = ~ tibble(Quantile = .x,
+                            Value = quantile(x, probs = .x,na.rm=TRUE)))
+  }) %>%  filter(Indx_Month %in% c(1,7,11))
+
+VIC_moy_doy %>%ggplot(aes(x=yday,y=Value,col=as.factor(Quantile),group=yday)) + geom_line() + facet_wrap(~Indx_Month)+ scale_x_continuous(breaks=seq(1, 336, 60)) + ylab("")  + xlab("Day of the Year") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
+
+# grid.arrange(p1,p2,p3,p4,nrow = 2)
+
+# 
+# x <- gridExtra::arrangeGrob (p5, p6, p1, p2, p3, p4,newpage = TRUE)
+# grid::grid.draw(x)
+
+# 
+# gridExtra::grid.arrange(
+#   p1,
+#   p2,
+#   p3,
+#   p4,
+#   p5,
+#   p6,
+#   nrow = 3,
+#   top = "Title of the page",
+#   bottom = grid::textGrob(
+#     "this footnote is right-justified",
+#     gp = grid::gpar(fontface = 3, fontsize = 9),
+#     hjust = 1,
+#     x = 1
+#   )
+# )
+# plot.new()
+# mtext('Various probability distribution plot of electricity consumption data of Victoria from 2012 to 2014 (top left) Letter Value plot by DoM and MoY (top right) Decile Plot by HoD and DoW (middle left) Box plot by DoY and DoM (middle right) Violin plot of DoM and WoM (bottom left) Ridge plot by WoM and WoY (bottom right) Decile plot by DoY and MoY', side = 3, line = - 10, outer = TRUE)
+```
+
+In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
+Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
+Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
+
+In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
+Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
+There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
+
+More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
+When C1 and C2 are used as aesthetics or facets, problems arise when we encounter empty combinations.
+Let $S_{i,j}$ be the set of combination of the levels of C1 and C2.
+In this example, we have 12 such sets $S_{i,j}$ because i can take 4 values and j can take 3 values.
+The graphs that don't work are those where many of these 12 sets are empty.
+In other words, if there are levels of the x-axis which are not spanned by levels of the faceted variable or vice versa we will have structurally empty sets leading to potential ineffective graphs.
+
+Thus, the common link that differentiates harmonies from clashes are:
+
+  a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis.
+
+  b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets.
+<!-- ## Distributions conditional on bivariate granularities -->
 

### 2019-08-12 main.Rmd
minor changes
why do you define single and multiple order up

@@
 
-**Defintion:** An **Aperiodic circular granularity** can not be defined using modular arithmetic.
+**Defintion:** An **Aperiodic circular granularity** can not be defined using modular arithmetic in a similar fashion.
 The modulus for these type of calendar categorisations are not constant due to unequal length of some linear granularities.
+For example, please refer to the table below:
 
-<!-- \begin{table}[ht] -->
-<!-- \begin{center} -->
-<!-- \begin{tabular}{lr@{~}lr@{~}r} -->
-<!-- \toprule -->
-<!-- HOD: & $C_1(s)$ & $= s \mod 24$ & $n_1 =$&$24$ \\ -->
-<!-- HOW: & $C_2(s)$ & $= s \mod `r 24*7`$ & $n_2=$&$`r 24*7`$\\ -->
-<!-- HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\ -->
-<!-- HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\ -->
-<!-- DOW: & $C_5(s)$ & $= \lfloor s/24\rfloor \mod 7$ & $n_5=$&$7$\\ -->
-<!-- DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\ -->
-<!-- DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\ -->
-<!-- WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\ -->
-<!-- WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\ -->
-<!-- MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\ -->
-<!-- \bottomrule -->
-<!-- \end{tabular} -->
-<!-- \end{center} -->
-<!-- \caption{Illustrative calendar categorizations with hourly time index} -->
-<!-- \label{definitions} -->
-<!-- \end{table} -->
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lr@{~}lr@{~}r}
+\toprule
 
-<!-- Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year. -->
+HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\
+HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\
+DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\
+DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\
+WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\
+WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\
+MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Illustrative aperiodic circular granularities with time index in hours}
+\label{definitions}
+\end{table}
+
+Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year.
+
+ <!-- Why we need single order up and multiple order up granularities -->
+<!-- A time series which is characterised by being composed of repeating cycles. -->
+Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
+Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
+To consider the exhaustive set of temporal regularities that might exist in the data, we can categorise the set of temporal granularities to single or multiple order up granularities.
 
@@
 
-## Mixed single-order-up granularities - circular or periodic
+## Mixed single-order-up granularities - circular or aperiodic
 

### 2019-08-12 main.Rmd
Removed smart meters from abstract to make it more general

@@
 abstract: |
-  Smart meters measure energy usage at fine temporal scales, and are now installed in many households around the world.
+ Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
+As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases.
 We propose some new tools to explore this type of data, which deconstruct time in many different ways.
@@
 
-  The hierarchical structure of many granularities creates a natural nested ordering.
+  The hierarchical structure of linear granularities creates a natural nested ordering.
 For example, hours are nested within days, days within weeks, weeks within months, and so on.
@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated within the tidy workflow, so that probability distributions can be examined using the range of graphics available in the [ggplot2](https://cran.r-project.org/package=ggplot2) package.
+This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in [@Wickham2009-pk].
 bibliography: bibliography.bib
@@
 Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
-To consider the exhaustive set of temporal regularities that might exist in the data, we can categorise the set of temporal granularities to single or multiple order up granularities.
+To consider the exhaustive set of temporal regularities that might exist ein the data, we can categorise the set of temporal granularities to single or multiple order up granularities.
 

### 2019-08-12 main.Rmd
Introduction of linear time granularities

@@
 
-  The hierarchical structure of linear granularities creates a natural nested ordering.
-For example, hours are nested within days, days within weeks, weeks within months, and so on.
-We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
+  The hierarchical structure of linear granularities creates a natural nested ordering resulting in single-order-up and multiple-order-up granularities.
 For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
 
-   Visualizing data across various granularities helps us to understand periodicities, pattern and anomalies in the data.
+   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data.
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
@@
 But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible.
-For example, it might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimisation every second.
+It might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimisation every second.
 As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases.
@@
 <!--existing closely related work and how ours relate --> 
-[@Grolemund2011-qf] allows creation of granularities that are mostly single-order-up like hour of the day, second of the minute.
  This paper utilises the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.
@@
 
-Time can be represented at varied levels of abstraction depending on the accuracy required for the context.
-Time granularity can be defined as the resolution power of the temporal qualification of a statement.
-Providing a formalism with the concept of time granularity is important to model time information across differently grained temporal domains [@Euzenat2005-de].
+Often we partition time into years, months, weeks, days and so on in a hierarchical manner to relate it to data.
+Such discrete abstractions of time can be thought of as linear time granularities [@aigner2011visualization].
+Providing a formalism to this abstraction is important to model a time series across differently grained temporal domains.
 

### 2019-08-12 main.Rmd
defining calendar categorisation first and then move on to linear or circular

@@
 
-# Time granularities{#sec:algorithm}
+# Formal conceptualisation of calendar categorisations{#cirgran-def}
 
-Often we partition time into years, months, weeks, days and so on in a hierarchical manner to relate it to data.
-Such discrete abstractions of time can be thought of as linear time granularities [@aigner2011visualization].
+<!-- general calendar categorisation -->
+Often we partition time into months, weeks, days and so on in a hierarchical manner to relate it to data.
+Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
 Providing a formalism to this abstraction is important to model a time series across differently grained temporal domains.
+Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc which are useful to represent periodicity in the data.
+Many of these are circular, such as day-of-week, time-of-day.
+Some are nearly circular such as day-of-month.
+Some are irregular such as workingday/non-working day.
+Let us call all of these different abstractions of time as "calendar categorizations".
+
+These calendar categorisations can be linear, circular or aperiodic.
+We call the calendar categorisations as "linear" if they respect the linear progression of time.
+We call these _*linear time granularities*_.
+Examples include hours, days, weeks and months.
+_*Circular time granularities*_ accommodate periodicities in time such as hour of the day, and day of the week.
+_*Aperiodic circular granularities*_ are neither linear nor circular, such as day of the month or public holidays.
+
+<!-- <!-- Representation with tsibble -->
+<!-- Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
+A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
+A linear granularity is a mapping of row numbers to subsets of the time domain.
+For example, if the time index is days, then a linear granularity might be weeks, months or years.
+-->
+
+<!-- <!-- What is calendar categorisation --> 
+<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
+Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc.
+Many of these are circular, such as day-of-week, time-of-day.
+Some are nearly circular such as day-of-month.
+Some are irregular such as workingday/non-working day.
+Let's call all of these "calendar categorizations".
+Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
+-->
+
+<!-- What is circular granularity and some examples -->
+<!-- We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
+The number of categories is essentially the periodicity of a circular time granularity.
+For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
+Then the following categorizations can be computed.
+-->
 
@@
 ## Formal conceptualisation of calendar categorisations{#cirgran-def}
-
-<!-- Representation with tsibble -->
-Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-
-<!-- What is calendar categorisation -->
-What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc.
-Many of these are circular, such as day-of-week, time-of-day.
-Some are nearly circular such as day-of-month.
-Some are irregular such as workingday/non-working day.
-Let's call all of these "calendar categorizations".
-Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
-
-<!-- What is circular granularity and some examples -->
-We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
-The number of categories is essentially the periodicity of a circular time granularity.
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
-Then the following categorizations can be computed.
 

### 2019-08-12 main.Rmd
changed intro of calendar categorisation

@@
 Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Providing a formalism to this abstraction is important to model a time series across differently grained temporal domains.
-Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc which are useful to represent periodicity in the data.
-Many of these are circular, such as day-of-week, time-of-day.
-Some are nearly circular such as day-of-month.
-Some are irregular such as workingday/non-working day.
+Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc which are useful to represent different periodicities in the data.
 Let us call all of these different abstractions of time as "calendar categorizations".
@@
 These calendar categorisations can be linear, circular or aperiodic.
-We call the calendar categorisations as "linear" if they respect the linear progression of time.
-We call these _*linear time granularities*_.
+The calendar categorisations as "linear" if they respect the linear progression of time.
+We call these **linear time granularities**.
 Examples include hours, days, weeks and months.
-_*Circular time granularities*_ accommodate periodicities in time such as hour of the day, and day of the week.
-_*Aperiodic circular granularities*_ are neither linear nor circular, such as day of the month or public holidays.
+**Circular time granularities** accommodate periodicities in time such as hour of the day, and day of the week.
+**Aperiodic circular granularities** are neither linear nor circular, such as day of the month or public holidays.
 
@@
 -->
+
+Providing a formalism to these abstractions is important to model a time series across differently grained temporal domains.
 

### 2019-08-12 main.Rmd
changing structure in Section 2 to make it flow better

@@
 **Circular time granularities** accommodate periodicities in time such as hour of the day, and day of the week.
-**Aperiodic circular granularities** are neither linear nor circular, such as day of the month or public holidays.
+**Aperiodic time granularities** are neither linear nor circular, such as day of the month or public holidays.
 
@@
 
-*These definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.*
+The definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.
+Hence, there is a need to define circular time granularities in a different approach.
 
-## Formal conceptualisation of calendar categorisations{#cirgran-def}
+## Circular time granularities{#circular-gran-def}
+
+Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
+A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
+A linear granularity is a mapping of row numbers to subsets of the time domain.
+For example, if the time index is days, then a linear granularity might be weeks, months or years.
+What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
 
@@
 
-It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
+<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
 Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
+-->
+
+We want to use modular arithmetic to define circular granularity. Hence, we start with the definition of equivalence classes and then move on to define a circular granularity.
 
@@
 
+## Aperiodic time granularities{#aperiodic-gran-def}
 <!-- Formal definition of aperiodic circular granularity -->
@@
 
-Note that most of the formulas are approximations only due to unequal month and year lengths, and due to the fact that there are not an integer number of weeks per month or weeks per year.
-
  <!-- Why we need single order up and multiple order up granularities -->
 <!-- A time series which is characterised by being composed of repeating cycles. -->
+
 Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
@@
 
-## Single-order-up and multiple-order-up granularities
+# Single-order-up and multiple-order-up granularities
 
@@
 
-# Computation of multiple order-up granularities from single order-up granularities
+## Computation of single-order-up granularities
+
+## Computation of multiple order-up granularities
 
@@
 
-## Periodic single-order-up granularities
+### Periodic single-order-up granularities
 
@@
 
-## Mixed single-order-up granularities - circular or aperiodic
+### Mixed single-order-up granularities - circular or aperiodic
 
@@
   <!-- $= c(x,w))$, otherwise. -->
-
-### Computing single-order-up granularities:
 

### 2019-08-12 main.Rmd
Changing advice algorithm

@@
 
+Levels are categorised as very high/high/medium/low each for the facet variable and the x-axis variable.
+Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
+For example, any levels above 31 is considered as very high, any levels between 14 to 31 aere taken as high and that between 7 to 14 is taken as medium and below 7 is low.
+31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
+
 <!-- We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.

### 2019-08-12 main.Rmd
spell check and making implememtnation and concept different for multiple gran construction

--- 2019-08-12 main.Rmd
@@
 Temporal data can be available at various resolution depending on the context.
-Social and economic data are often collected and reported at coarses temporal scales like monthly, quarterly or annually.
+Social and economic data are often collected and reported at coarser temporal scales like monthly, quarterly or annually.
 But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible.
-It might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimisation every second.
+It might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimization every second.
 As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases.
@@
 The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world.
-Smart meters currently provide half-hourly use in kwh for each household, from the time that they were installed, some as early as 2012.
+Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
 Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
@@
 
-It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
@@
 <!--existing closely related work and how ours relate --> 
- This paper utilises the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.
+ This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.
 
@@
 However, this approach can lead to a myriad of choices all of which are not useful.
-Analysts are expected to iteratively visualise these choices for exploring possible patterns in the data.
+Analysts are expected to iteratively visualize these choices for exploring possible patterns in the data.
 But too many choices might leave him bewildered.
@@
 
-# Formal conceptualisation of calendar categorisations{#cirgran-def}
+# Formal conceptualization of calendar categorizations{#cirgran-def}
 
-<!-- general calendar categorisation -->
+<!-- general calendar categorization -->
 Often we partition time into months, weeks, days and so on in a hierarchical manner to relate it to data.
 Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc which are useful to represent different periodicities in the data.
+Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
 Let us call all of these different abstractions of time as "calendar categorizations".
 
-These calendar categorisations can be linear, circular or aperiodic.
-The calendar categorisations as "linear" if they respect the linear progression of time.
+These calendar categorizations can be linear, circular or aperiodic.
+The calendar categorizations as "linear" if they respect the linear progression of time.
 We call these **linear time granularities**.
@@
 
-<!-- <!-- What is calendar categorisation --> 
+<!-- <!-- What is calendar categorization --> 
 <!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc.
+Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc.
 Many of these are circular, such as day-of-week, time-of-day.
 Some are nearly circular such as day-of-month.
-Some are irregular such as workingday/non-working day.
+Some are irregular such as working day/non-working day.
 Let's call all of these "calendar categorizations".
@@
 
-**Defintion:** An **Aperiodic circular granularity** can not be defined using modular arithmetic in a similar fashion.
-The modulus for these type of calendar categorisations are not constant due to unequal length of some linear granularities.
+**Definition:** An **Aperiodic circular granularity** can not be defined using modular arithmetic in a similar fashion.
+The modulus for these type of calendar categorizations are not constant due to unequal length of some linear granularities.
 For example, please refer to the table below:
@@
  <!-- Why we need single order up and multiple order up granularities -->
-<!-- A time series which is characterised by being composed of repeating cycles. -->
+<!-- A time series which is characterized by being composed of repeating cycles. -->
 
@@
 Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
-To consider the exhaustive set of temporal regularities that might exist ein the data, we can categorise the set of temporal granularities to single or multiple order up granularities.
+To consider the exhaustive set of temporal regularities that might exist in the data, we can categorize the set of temporal granularities to single or multiple order up granularities.
 
@@
 
-The index of a tsibble can be any date-time object which comes in different formats.
+<!-- The index of a tsibble can be any date-time object which comes in different formats.
 [@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data and extracting each date-time element with its own accessor function.
 However, table shows the accessor functions mostly allows for the creation of single-order-up granularities.
+-->
 
-```{r table_accessor}
-kable(tibble::tibble(
-  `Categorisation` = c("Second-of-minute", "Minute-of-hour", "Hour-of-day", "Day-of-week", "Day-of-month", "Week-of-year", "Day-of-year","Month-of-year"),
- Type = c("Periodic", "Periodic", "Periodic", "Periodic", "Aperiodic", "Aperiodic", "Aperiodic", "Periodic"),
-Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "markdown")
-```
+<!-- ```{r table_accessor} -->
+<!-- kable(tibble::tibble( -->
+<!--   `Categorization` = c("Second-of-minute", "Minute-of-hour", "Hour-of-day", "Day-of-week", "Day-of-month", "Week-of-year", "Day-of-year","Month-of-year"), -->
+<!--  Type = c("Periodic", "Periodic", "Periodic", "Periodic", "Aperiodic", "Aperiodic", "Aperiodic", "Periodic"), -->
+<!-- Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "markdown") -->
+<!-- ``` -->
 
-The hierarchical nature of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.The idea is to incorporate and access different calendar categorisations to account for varied levels of periodicities in the data.
+The hierarchical nature of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.The idea is to incorporate and access all possible calendar categorizations to account for different periodicities in the data.
 The computation of multiple-order-up granularities from single-order-up granularities can differ basis the single-order-up granularities are circular or aperiodic.
@@
 
-There can be three scenarios for obtaining calendar categorisation here:
-- calendar categorisation consisting of two linear granularities whose orders are less than day  
-- calendar categorisation consisting of two linear granularities whose orders are more than month  
-- calendar categorisation consisting of one linear granularity with order at most day and another with order at least month
+There can be three scenarios for obtaining calendar categorization here:
+- calendar categorization consisting of two linear granularities whose orders are less than day  
+- calendar categorization consisting of two linear granularities whose orders are more than month  
+- calendar categorization consisting of one linear granularity with order at most day and another with order at least month
 
-The calendar categorisation resulting from the first two cases are periodic and has been handled in the earlier section.
+The calendar categorization resulting from the first two cases are periodic and has been handled in the earlier section.
 
-The calendar categorisation resulting from the last case are aperiodic.
+The calendar categorization resulting from the last case are aperiodic.
 Examples might include day of the quarter or hour of the month.
@@
 VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Day_Month %in% c(1,15,29,31))%>% ggplot(aes(as.factor(yday),OperationalLessIndustrial,group = yday)) + geom_boxplot()+ facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
-  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Boxplot")
+  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot")
 
@@
 
-<!-- We investigate some combinations of circular time granularities which facilitate or hinder exploratory analysis.
-Since we are restricting ourselves to 2D space, we assume that the circular granularities are plotted across the aesthetics and facets.
-The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
--->
+Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
+Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
+For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
+31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
+These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
+Let us consider case by case and see which plots are better suitable in which scenarios.
+
+ -  very high facet and x-axis levels
+
+ X-axis variable is treated as a categorical variable and hence any plots which
 

### 2019-08-21 main.Rmd
including description of short calendars

@@
 
-Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
+A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index in a tsibble[@Wang2019-aw] for a given key.
 A linear granularity is a mapping of row numbers to subsets of the time domain.
@@
 
-## Computation of single-order-up granularities
+## Single-order-up granularities
+
+<!-- some discussions on primary units of calendar -->
+The day is the basic unit of time underlying all calendars.
+The most common  way of defining days are from sunrise to sunrise (Hindu calendar, for example) or from sunset to sunset (Islamic and Hebrew calendars).
+Further, various calendars use different conventions to structure days into larger units: weeks, months, years and cycle of years.
+Any civil day is divided into 24 hours and each hour into 60 minutes and each minute into 60 seconds.
+There are exceptions where Greeks and Romans separate the "day" and "night" into 12 hours each.
+In London, for example, length of each "hour" varies from about 39 minutes in December to 83 minutes in June.
+The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
 
@@
 
-<!-- Suppose, we are not given the intermediate single-order-up granularities but still want all possible multiple order-up granularties.
-In that case, it is necessary to at least have the have the functional form that relates the smallest temporal unit to the highest temporal unit.
-In that case, we are able to form single-order-up from this functional relationship and still use the recursive relation to get the multiple order-up granularities.
--->
-
-  <!-- In the above equations, it is shown how to obtain f(x,z) given f(x,y) and the hierarchy table,  provided $order(z) > order(y) \neq order(x)$. In general, -->
-
-  <!-- $f(x,z) = \sum_{i=0}^{order(z) - order(y) + 1}c(x, x+i)f(x+1, x+i)$ -->
-
-  <!-- We could also derive f(w, y) given f(x, y) and the hierarchy table, for $order(y)>=order(w) > order(x). -->
-
-  <!-- $f(w,y) = f(x,y) \mod c(x,w)$, for $f(x,y) \mod c(x,w)!=0$ -->
-  <!-- $= c(x,w))$, otherwise. -->
-
-<!-- f(ball, over) = 1 -->
-<!-- f(over, quarter) = 4 -->
-<!-- f(quarter, semester) = 2 -->
-<!-- f(semester, match) = `1 -->
-
-<!-- then, f(ball_quarter) = 1 + 6*4 = 25,  -->
-<!-- f(ball, semester) = 1 + 6*4 + 5*6*1 = 55 -->
-<!-- f(ball, semester) = 1 + 6*4 + 5*6*1 + 5*6*2*1 = 115 -->
+Here, the first part of the equation is a single-order-up granularity which can be obtained using last section.
+The second part is not single-order-up and can not be broken down further since each month consists of different number of days.
+In this case, it is important for us to know which month of the yearnand if the year is a leap year to obtain day of the month.
 

### 2019-08-23 main.Rmd
case study for smart metter graph 1 added

@@
 
+Smart meters provide large quantities of measurements on energy usage for households across Australia, and indeed many parts of the world.
+Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+
+<!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
+It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
+
+One of the customer trial [smart-meter]z conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in NewCastle, New South Wales and some parts of Sydney provides customer wise data on half hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
+It would be interesting to explore the energy consumption distribution for these customers across different temporal granularities.
+
+```{r search_gran}
+
+library(gravitas)
+library(tsibble)
+load("data/smart_meter_13_cust1.rda")
+load("data/smart_meter_13_cust2.rda")
+load("data/smart_meter_13_cust3.rda")
+
+data_sm1 <- smart_meter_13_cust1 %>% as_tsibble()
+data_sm2 <- smart_meter_13_cust2 %>% as_tsibble()
+data_sm3 <- smart_meter_13_cust3 %>% as_tsibble()
+```
+
+<!-- Refine your search of possible temporal granularities by altering arguments in search_gran() -->
+
+<!-- The default for search gran in this case, provides temporal granularities ranging from half-hour to year.
+If these options are considered too many, the default options can be modified to limit the possibilities.
+For example, the most coarce temporal unit can be set to be “month”.
+-->
+
+<!-- # ```{r} -->
+<!-- # library(gravitas) -->
+<!-- # library(dplyr) -->
+<!-- # data_sm %>% search_gran(ugran = "year", filter_in = "event_key", filter_out = c("quarter", "semester")) -->
+<!-- # ``` -->
+
+<!-- # ```{r} -->
+<!-- # data_sm %>% harmony(ugran = "month", filter_in = "event_key", filter_out = c("quarter", "semester"), facet_h = 31) -->
+<!-- # ``` -->
+
+```{r}
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+
+data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+
+```
+
 # Case study: Analysis on cricket {#cricket}

### 2019-08-24 main.Rmd
loading data  for two households

@@
 
+Let us compare the energy consumption for two households from the customer trials.
+
 ```{r search_gran}
@@
 
+For the purpose of this analysis.
+We only take the data for the year 2013.
+Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
 <!-- Refine your search of possible temporal granularities by altering arguments in search_gran() -->
@@
 
-```{r}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+```{r cust1_hour-day_day-week}
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") +  ggtitle("Boxplot of energy consumption for Customer 1")
+```
+<!-- # ``````{r cust2_hour-day_day-week} -->
+<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
+<!-- # ``` -->
 
-data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+``````{r cust3_hour-day_day-week}
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 3")
+```
 
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+```{r cust1_hhour-day_day-week}
+data_sm1 %>% granplot("day_week", "week_month", "general_supply_kwh", plot_type = "decile") +  ggtitle("Boxplot of energy consumption for Customer 1")
+```
+
+<!-- # ```{r cust2_hhour-day_day-week} -->
+<!-- # data_sm2 %>% granplot("day_week", "week_month", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
+<!-- # ``` -->
+
+```{r cust3_hhour-day_day-week}
+data_sm3 %>% granplot("day_week", "week_month", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 3")
+```
 

### 2019-08-25 main.Rmd
case study updating

--- 2019-08-23 main.Rmd
@@
     fig_width: 8
+    fig_caption: yes
     dev: "pdf"
@@
   fig.path = "figure/", fig.align = "center", fig.show = "hold",
-  cache = FALSE, cache.path = "cache/",
+  cache = TRUE, cache.path = "cache/",
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 
-One of the customer trial [smart-meter]z conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in NewCastle, New South Wales and some parts of Sydney provides customer wise data on half hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-It would be interesting to explore the energy consumption distribution for these customers across different temporal granularities.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in NewCastle, New South Wales and some parts of Sydney provides customer wise data on half hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
+It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggreagation or looking only at coarser temporal units.
+
+Let us compare the energy consumption distribution for two households from the customer trials.
 
@@
 ```
+
+For the purpose of this analysis.
+We only take the data for the year 2013.
+Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
 
@@
 
-```{r}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+ -  Hourly usage across months of the year
 
-data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+(in Figure \@ref(fig:cust1_hd_my)) shows the distribution of hourly comsumption of the households across different months of the year.
+The two households are similar when it comes to your usage in the morning hours.
+wThe second household has an evening peak across all months in contrast to the first household who has a steep evening peak only in the month of May and June.
+This suggests an use of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months.
+Let us dive deeper and see the hourly usage across different days of the week for these two households.
 
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile")
+```{r cust1_hd_my, fig.cap="(ref:cust1_hd_my)"}
+
+library(ggplot2)
+data_sm1 %>% granplot("month_year", "hour_day", response = "general_supply_kwh", plot_type = "quantile") +  ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+(ref:cust1_hd_my-cap) Something that I am yet to write
+
+```{r cust3_hour_day-month_year}
+data_sm3 %>% granplot("month_year", "hour_day", "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+-  Hourly usage across days of the week
+
+Hourly usage for two households across different days of the week are shown through decile plots.
+It is clear that energy conusmption only goes up for few hours in the evening.
+It goes much higher in the weekends implying that there are hardly any activities that take place during weekdays.
+Whereas, for the second household, there are two peaks - once during the day and one in the evening.
+Also, the difference between the energy consumption is not very different between weekdays and weekends.
+This suggests that the household might uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
+
+```{r cust1_hour-day_day-week}
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("Boxplot of energy consumption for Customer 1")
+```
+<!-- # ``````{r cust2_hour-day_day-week} -->
+<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
+<!-- # ``` -->
+
+``````{r cust3_hour-day_day-week}
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 3")
+```
+
+-  Half-Hourly usage across days of the week
+
+Now, let us check which half hours are the most active for Household 1.
+For most days, half hours between 34th and 38th  half hour of the day are higher than other days.
+The energy consumption in these hours on Fridays match with that os Saturday and Sunday.
+
+<!-- ```{r cust1_hhour-day_day-week} -->
+<!-- data_sm1 %>% -->
+<!--    filter(lubridate::hour(reading_datetime)>=23) %>% -->
+<!--   granplot("hhour_day","day_week", response = "general_sxupply_kwh", plot_type = "lv") + -->
+<!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->
+<!-- ``` -->
+
+<!-- <!-- # ```{r cust2_hhour-day_day-week} --> -->
+<!-- <!-- # data_sm2 %>% granplot("day_week", "week_month", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 2") --> -->
+<!-- <!-- # ``` --> -->
+
+<!-- ```{r cust3_hhour-day_day-week} -->
+<!-- data_sm3 %>% create_gran("hhour_day") %>% -->
+<!--    filter(hhour_day>=23) %>% -->
+<!--   granplot("hhour_day","day_week",response = "general_supply_kwh", plot_type = "lv") + -->
+<!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->
+<!-- ``` -->
 

### 2019-08-25 main.Rmd
adding on the first analysis of case study

@@
 
-(in Figure \@ref(fig:cust1_hd_my)) shows the distribution of hourly comsumption of the households across different months of the year.
-The two households are similar when it comes to your usage in the morning hours.
-wThe second household has an evening peak across all months in contrast to the first household who has a steep evening peak only in the month of May and June.
-This suggests an use of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months.
+\autoref{fig:cust1_hd_my} shows the distribution of hourly comsumption of the households across different months of the year.
+The two households are similar when it comes to their usage in the morning hours.
+The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage).
+However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar.
+This implies that consuming more energy in the evening is more typical for household 2.
+For the evenings of OCtober, November and December, it is clear that household 1 typically consumes much less energy than household 2.
+Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH.
+This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely.
+All of these can be explained with more demographics of the househols and weather conditions in that area.
+
 Let us dive deeper and see the hourly usage across different days of the week for these two households.
 
-```{r cust1_hd_my, fig.cap="(ref:cust1_hd_my)"}
+<!-- ```{r cust1_hd_my} -->
+
+<!-- library(ggplot2) -->
+<!-- data_sm1 %>%  -->
+<!--   group_by(month_year = lubridate::month(reading_datetime), -->
+<!--            hour_day = lubridate::hour(reading_datetime)) %>%  -->
+<!--   summarise(hourly_tot = mean(general_supply_kwh)) %>%  -->
+<!--   ggplot(aes(x = hour_day, y = month_year)) + facet_wrap(~ month_year)+ -->
+<!--   geom_point() +  -->
+<!--   ggtitle("Bar plot of energy consumption across hours of day by month of years for Household 1") -->
+<!-- ``` -->
+
+```{r cust1_hd_my}
 
 library(ggplot2)
-data_sm1 %>% granplot("month_year", "hour_day", response = "general_supply_kwh", plot_type = "quantile") +  ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm1 %>% granplot("month_year", "hour_day", response = "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
 
-(ref:cust1_hd_my-cap) Something that I am yet to write
-
-```{r cust3_hour_day-month_year}
+```{r cust3_hd_my}
 data_sm3 %>% granplot("month_year", "hour_day", "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))

### 2019-08-25 main.Rmd
adding on the second  analysis of case study

@@
 ```{r cust1_hd_my}
-
 library(ggplot2)
@@
 
-Hourly usage for two households across different days of the week are shown through decile plots.
-It is clear that energy conusmption only goes up for few hours in the evening.
-It goes much higher in the weekends implying that there are hardly any activities that take place during weekdays.
-Whereas, for the second household, there are two peaks - once during the day and one in the evening.
-Also, the difference between the energy consumption is not very different between weekdays and weekends.
-This suggests that the household might uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
+\autoref{fig:cust1_hd_my} shows distribution of hourly eenrgy usage for two households across different days of the week through decile plots.
+It is clear that household 1 deviates from their typical behavior only on weekends.
+A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
+Was it only for the months May and June, that this household consumed more energy than they usually do?
+\autoref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
+In fact, months like November nad December saw higher ebergy usage than typical even on other weekdays.
+Still for this household, it can safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
+The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.
+
+/autoref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
+The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December.
+Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.
+
+```{r cust3_dw_my}
+library(ggplot2)
+
+data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+```{r cust3_dw_my}
+library(ggplot2)
+
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across hours of day by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
 
 ```{r cust1_hour-day_day-week}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("Boxplot of energy consumption for Customer 1")
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile") +  ggtitle("Boxplot of energy consumption for Customer 1")
 ```

### 2019-08-25 main.Rmd
updating analysis 2 of case study

@@
   fig.path = "figure/", fig.align = "center", fig.show = "hold",
-  cache = TRUE, cache.path = "cache/",
+  cache = FALSE, cache.path = "cache/",
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggreagation or looking only at coarser temporal units.
+The idea here is to show how looking at time across different granularities together can help identify different behavioral patterns.
 
 Let us compare the energy consumption distribution for two households from the customer trials.
+For the purpose of this analysis, we only take the data for the year 2013.
+Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
 
@@
 ```
-
-For the purpose of this analysis.
-We only take the data for the year 2013.
-Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
 
@@
 
+```{r cust1_hour-day_day-week}
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Boxplot of energy consumption for Customer 1")
+```
+<!-- # ``````{r cust2_hour-day_day-week} -->
+<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
+<!-- # ``` -->
+
+``````{r cust3_hour-day_day-week}
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Boxplot of energy consumption for Customer 3")
+```
+
 \autoref{fig:cust1_hd_my} shows distribution of hourly eenrgy usage for two households across different days of the week through decile plots.
@@
 
-```{r cust3_dw_my}
+-  Across days of the week by months
+
+```{r cust1_dw_my}
 library(ggplot2)
 
-data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
@@
 
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across hours of day by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
-```{r cust1_hour-day_day-week}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile") +  ggtitle("Boxplot of energy consumption for Customer 1")
-```
-<!-- # ``````{r cust2_hour-day_day-week} -->
-<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
-<!-- # ``` -->
-
-``````{r cust3_hour-day_day-week}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 3")
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```

### 2019-08-25 main.Rmd
Spell check

@@
 The second part is not single-order-up and can not be broken down further since each month consists of different number of days.
-In this case, it is important for us to know which month of the yearnand if the year is a leap year to obtain day of the month.
+In this case, it is important for us to know which month of the year and if the year is a leap year to obtain day of the month.
 
@@
 One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in NewCastle, New South Wales and some parts of Sydney provides customer wise data on half hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggreagation or looking only at coarser temporal units.
+It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
 The idea here is to show how looking at time across different granularities together can help identify different behavioral patterns.
@@
 If these options are considered too many, the default options can be modified to limit the possibilities.
-For example, the most coarce temporal unit can be set to be “month”.
+For example, the most coarse temporal unit can be set to be “month”.
 -->
@@
 
-\autoref{fig:cust1_hd_my} shows the distribution of hourly comsumption of the households across different months of the year.
+\autoref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year.
 The two households are similar when it comes to their usage in the morning hours.
@@
 This implies that consuming more energy in the evening is more typical for household 2.
-For the evenings of OCtober, November and December, it is clear that household 1 typically consumes much less energy than household 2.
+For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2.
 Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH.
 This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely.
-All of these can be explained with more demographics of the househols and weather conditions in that area.
+All of these can be explained with more demographics of the household and weather conditions in that area.
 
@@
 
-\autoref{fig:cust1_hd_my} shows distribution of hourly eenrgy usage for two households across different days of the week through decile plots.
+\autoref{fig:cust1_hd_my} shows distribution of hourly energy usage for two households across different days of the week through decile plots.
 It is clear that household 1 deviates from their typical behavior only on weekends.
@@
 \autoref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
-In fact, months like November nad December saw higher ebergy usage than typical even on other weekdays.
+In fact, months like November and December saw higher energy usage than typical even on other weekdays.
 Still for this household, it can safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
@@
 <!--    filter(lubridate::hour(reading_datetime)>=23) %>% -->
-<!--   granplot("hhour_day","day_week", response = "general_sxupply_kwh", plot_type = "lv") + -->
+<!--   granplot("hhour_day","day_week", response = "general_supply_kwh", plot_type = "lv") + -->
 <!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->

### 2019-08-25 main.Rmd
added text for the case study

@@
 library(ggplot2)
-data_sm1 %>% granplot("month_year", "hour_day", response = "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm1 %>% 
+  granplot("month_year", "hour_day", 
+           response = "general_supply_kwh", 
+           plot_type = "quantile", 
+           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
+  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
+  scale_x_discrete(breaks = seq(0,23,2))
 ```
@@
 ```{r cust3_hd_my}
-data_sm3 %>% granplot("month_year", "hour_day", "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm3 %>% 
+  create_gran("month_year") %>% 
+  filter(month_year != "Mar") %>% 
+    granplot("month_year", "hour_day", "general_supply_kwh",
+             plot_type = "quantile", 
+             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + 
+  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
+  scale_x_discrete(breaks = seq(0,23,2))
 ```
@@
 
-```{r cust1_hour-day_day-week}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Boxplot of energy consumption for Customer 1")
+```{r cust1_hd_dw}
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
 ```
@@
 
-``````{r cust3_hour-day_day-week}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Boxplot of energy consumption for Customer 3")
+``````{r cust3_hd_dw}
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
 ```
 
-\autoref{fig:cust1_hd_my} shows distribution of hourly energy usage for two households across different days of the week through decile plots.
-It is clear that household 1 deviates from their typical behavior only on weekends.
+\autoref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1.
+Energy consumption typically goes up between 16 to 18 hours.
+On weekends, the energy consumption for these hours are even higher.
+The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour.
+This suggests that these hours are extremely volatile based on weather conditions/months.
+Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range.
+This is not true for the 17th or 18th hour of the day.
+The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home professionals.
+
+\autoref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2.
+Contrary to Household 1, Household 2's energy consumption increases through noon through 7pm.
+We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption from the households.
+This indicates that this household might have kids or someone working from home.
+
 A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
@@
 In fact, months like November and December saw higher energy usage than typical even on other weekdays.
-Still for this household, it can safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
+Still for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
 The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.
@@
 
--  Half-Hourly usage across days of the week
+To summarise, we can posit the following for the two households:
 
-Now, let us check which half hours are the most active for Household 1.
-For most days, half hours between 34th and 38th  half hour of the day are higher than other days.
-The energy consumption in these hours on Fridays match with that os Saturday and Sunday.
+ - First household tpically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June).
+They are strictly 9 to 5 professsionals, hence their energy consumption during day time is pretty consistent and only peaks up during few hours in the evening.
+They do all energy intensive chores (e.g.
+laundry, vacuum cleaning) mostly in the evenings of weekends.
 
-<!-- ```{r cust1_hhour-day_day-week} -->
-<!-- data_sm1 %>% -->
-<!--    filter(lubridate::hour(reading_datetime)>=23) %>% -->
-<!--   granplot("hhour_day","day_week", response = "general_supply_kwh", plot_type = "lv") + -->
-<!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->
-<!-- ``` -->
+ - Second household tpically use air conditioners in summer months and heaters in winter months.
+This might be a family with children with stay-at-home parents/ work-from-home professionals with extremely varying energy consumption peaking up from noon.
+Their weekday behavior is not distinctly different from weekends.
+Also, they are more likely to distribute their energy intensive chores through out the week and not leave all them for the weekends.
 
-<!-- <!-- # ```{r cust2_hhour-day_day-week} --> -->
-<!-- <!-- # data_sm2 %>% granplot("day_week", "week_month", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 2") --> -->
-<!-- <!-- # ``` --> -->
-
-<!-- ```{r cust3_hhour-day_day-week} -->
-<!-- data_sm3 %>% create_gran("hhour_day") %>% -->
-<!--    filter(hhour_day>=23) %>% -->
-<!--   granplot("hhour_day","day_week",response = "general_supply_kwh", plot_type = "lv") + -->
-<!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->
-<!-- ``` -->
-
-```
+We can keep on slicing and dicing the data in a similar fashion and derive more insights about the behavior of these households.
 

### 2019-08-25 smart_meter.Rmd
adding fig captions in case study

--- 2019-08-25 main.Rmd
@@
 ---
-title: Exploring probability distributions for bivariate temporal granularities
-authors:
-- name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-  email: Sayani.Gupta@monash.edu
-
-- name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-keywords:
-- data visualization
-- statistical distributions
-- time granularities
-- periodicties
-- grammar of graphics
-- R
-
-abstract: |
- Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
-As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases.
-We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week.
-Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.
-
-  The hierarchical structure of linear granularities creates a natural nested ordering resulting in single-order-up and multiple-order-up granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
-
-   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in [@Wickham2009-pk].
-bibliography: bibliography.bib
+title: "Case study: Analysis on smart meter data"
+author: "Sayani Gupta"
+date: "25/08/2019"
 output: 
-  bookdown::pdf_book:
-    #base_format: rticles::asa_article
-    fig_height: 5
-    fig_width: 8
+  pdf_document: 
     fig_caption: yes
-    dev: "pdf"
 ---
@@
 ```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
-options("knitr.graphics.auto_pdf" = TRUE)
 library(knitr)
@@
 library(ggridges)
+library(gravitas)
+options("knitr.graphics.auto_pdf" = TRUE)
+library(knitr)
 opts_chunk$set(
   echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
-  fig.path = "figure/", fig.align = "center", fig.show = "hold",
-  cache = FALSE, cache.path = "cache/",
+  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
+  cache = TRUE, cache.path = 'cache/', 
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
@@
 ```
-
-```{r external, include = FALSE}
-# read_chunk('scripts/main.R')
-```
-
-```{r load}
-
-```
-
-# Introduction
-
-<!--temporal granularities and why should we care --> 
-Temporal data can be available at various resolution depending on the context.
-Social and economic data are often collected and reported at coarser temporal scales like monthly, quarterly or annually.
-But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible.
-It might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimization every second.
-As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases.
-For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
-This approach requires deconstructing time in various possible ways.
-
-A temporal granularity which results from such a deconstruction may be intuitively described as a sequence of time granules, each one consisting of a set of time instants.
-There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week.
-Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.
-
-It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the observed data.
-This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-
-<!--motivation -->
-
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world.
-Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-
-It is common to see aggregates of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-
-We propose that the analysis of probability distributions of smart meter data at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA).
-EDA calls for utilizing visualization and transformation to explore data systematically.
-It is a process of generating hypothesis, testing them and consequently refining them through investigations.
-
-<!--details abt time gran --> 
-The hierarchical structure of many granularities creates a natural nested ordering.
-For example, hours are nested within days, days within weeks, weeks within months, and so on.
-We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
-
-<!--existing closely related work and how ours relate --> 
- This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.
-
-<!--actions on multiple order up - usage --> 
-Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-However, this approach can lead to a myriad of choices all of which are not useful.
-Analysts are expected to iteratively visualize these choices for exploring possible patterns in the data.
-But too many choices might leave him bewildered.
-
-<!-- challenges and how you are dealing wih it --> 
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow through proper study of what can be considered a prospective graphic for exploration.
-Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-Probability distributions can be examined using the range of graphics available in the [ggplot2](https://cran.r-project.org/package=ggplot2) package.
-
-In particular, this work provides the following tools.
-
-  * Functions to create multiple-order-up time granularities.
-This is an extension to the [lubridate]([https://cran.r-project.org/package=lubridate) package, which allows for the creation of some calendar categorizations, usually single-order-up.
-
-  * Checks on the feasibility of creating plots or drawing inferences from two granularities together.
-Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-
-<!-- The remainder of the paper is organized as follows.
-Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)).
-Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes.
-An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case).
-Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.
--->
-
-# Formal conceptualization of calendar categorizations{#cirgran-def}
-
-<!-- general calendar categorization -->
-Often we partition time into months, weeks, days and so on in a hierarchical manner to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
-Let us call all of these different abstractions of time as "calendar categorizations".
-
-These calendar categorizations can be linear, circular or aperiodic.
-The calendar categorizations as "linear" if they respect the linear progression of time.
-We call these **linear time granularities**.
-Examples include hours, days, weeks and months.
-**Circular time granularities** accommodate periodicities in time such as hour of the day, and day of the week.
-**Aperiodic time granularities** are neither linear nor circular, such as day of the month or public holidays.
-
-<!-- <!-- Representation with tsibble -->
-<!-- Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
--->
-
-<!-- <!-- What is calendar categorization --> 
-<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc.
-Many of these are circular, such as day-of-week, time-of-day.
-Some are nearly circular such as day-of-month.
-Some are irregular such as working day/non-working day.
-Let's call all of these "calendar categorizations".
-Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
--->
-
-<!-- What is circular granularity and some examples -->
-<!-- We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
-The number of categories is essentially the periodicity of a circular time granularity.
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
-Then the following categorizations can be computed.
--->
-
-Providing a formalism to these abstractions is important to model a time series across differently grained temporal domains.
-
-## Linear time granularities{#linear-gran-def}
-
-There has been several attempts to provide the framework for formally characterizing time-granularities and identifying their structural properties, relationships and symbolic representations.
-One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
-
-**Definition:** A **time domain** is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
-
-A time domain can be **discrete** (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be **dense** (if it is an infinite set).
-A time domain is assumed to be discrete for the purpose of our discussion.
-
-**Definition:** A linear **granularity** is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
-
-  (C1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
-than all elements of $G(j)$, and  
-  (C2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
-
-**Definition:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
-
-The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
-\autoref{fig:linear-time} shows the implication of this condition.
-If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor s/(24*7)\rfloor$ weeks and so on.
-
-```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="\\label{fig:linear-time} The time domain distributed as linear granularities"}
-knitr::include_graphics("Figs/linear-time.png")
-```
-
-The definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.
-Hence, there is a need to define circular time granularities in a different approach.
-
-## Circular time granularities{#circular-gran-def}
-
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index in a tsibble[@Wang2019-aw] for a given key.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-MOH: & $C_1(s)$ & $= s \mod 60$ & $n_1 =$&$60$ \\
-MOD: & $C_2(s)$ & $= s \mod `r 60*24`$ & $n_2=$&$`r 60*24`$\\
-
-HOD: & $C_4(s)$ & $= \lfloor s/60\rfloor\mod 24$  & $n_4 =$&$24$ \\
-HOW: & $C_6(s)$ & $= \lfloor s/60\rfloor\mod 24*7$  & $n_5=$&$`r 24*7`$\\
-
-DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_6=$&$7$\\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Illustrative circular granularities with time index in minutes}
-\label{definitions}
-\end{table}
-
-<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
-Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
--->
-
-We want to use modular arithmetic to define circular granularity. Hence, we start with the definition of equivalence classes and then move on to define a circular granularity.
-
-<!-- Formal definition of circular granularity -->
-**Definition: Equivalence class** Let $m \in N \backslash{0}$.
-For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
-
-The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$.
-Thus $Z_m = \{[0], [1], ..., [m-1]\}$.
-However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.
-
-**Definition:** A **circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.
-
-For example, suppose $C$ is a circular granularity denoting Hour-of-Day and we have hourly data for 100 hours.
-The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
-
-**Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once.
-
-**Definition:** A **circular granule** represents an equivalence class inside each cycle.
-
-## Aperiodic time granularities{#aperiodic-gran-def}
-<!-- Formal definition of aperiodic circular granularity -->
-
-**Definition:** An **Aperiodic circular granularity** can not be defined using modular arithmetic in a similar fashion.
-The modulus for these type of calendar categorizations are not constant due to unequal length of some linear granularities.
-For example, please refer to the table below:
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-
-HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\
-HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\
-DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\
-DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\
-WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\
-WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\
-MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Illustrative aperiodic circular granularities with time index in hours}
-\label{definitions}
-\end{table}
-
- <!-- Why we need single order up and multiple order up granularities -->
-<!-- A time series which is characterized by being composed of repeating cycles. -->
-
-Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
-Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
-To consider the exhaustive set of temporal regularities that might exist in the data, we can categorize the set of temporal granularities to single or multiple order up granularities.
-
-# Single-order-up and multiple-order-up granularities
-
-<!-- What is multiple and single order up granularities -->
-The hierarchical structure of time creates a natural nested ordering, where hours are nested within days, days within weeks, weeks within months, and so on.
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are **single-order-up**.
-
-## Single-order-up granularities
-
-<!-- some discussions on primary units of calendar -->
-The day is the basic unit of time underlying all calendars.
-The most common  way of defining days are from sunrise to sunrise (Hindu calendar, for example) or from sunset to sunset (Islamic and Hebrew calendars).
-Further, various calendars use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-Any civil day is divided into 24 hours and each hour into 60 minutes and each minute into 60 seconds.
-There are exceptions where Greeks and Romans separate the "day" and "night" into 12 hours each.
-In London, for example, length of each "hour" varies from about 39 minutes in December to 83 minutes in June.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
-
-## Computation of multiple order-up granularities
-
-<!-- The index of a tsibble can be any date-time object which comes in different formats.
-[@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data and extracting each date-time element with its own accessor function.
-However, table shows the accessor functions mostly allows for the creation of single-order-up granularities.
--->
-
-<!-- ```{r table_accessor} -->
-<!-- kable(tibble::tibble( -->
-<!--   `Categorization` = c("Second-of-minute", "Minute-of-hour", "Hour-of-day", "Day-of-week", "Day-of-month", "Week-of-year", "Day-of-year","Month-of-year"), -->
-<!--  Type = c("Periodic", "Periodic", "Periodic", "Periodic", "Aperiodic", "Aperiodic", "Aperiodic", "Periodic"), -->
-<!-- Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "markdown") -->
-<!-- ``` -->
-
-The hierarchical nature of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.The idea is to incorporate and access all possible calendar categorizations to account for different periodicities in the data.
-The computation of multiple-order-up granularities from single-order-up granularities can differ basis the single-order-up granularities are circular or aperiodic.
-
-### Periodic single-order-up granularities
-
-**Definition**: A **hierarchy table** has three columns:  
- - The first column represents the linear granularities in ascending order of temporal hierarchy.  
- - The second column represents the constant which relates subsequent linear granularities.  
- - The third column provides the accessor function for single-order up circular granularities.
-
-**Definition**: **Order** is defined as the position of the linear granularities in the hierarchical table. 
-<!-- For example, if two linear granularties have three intermediate granularities in the hierarchical table, then their order difference is 3.  -->
-
-<!-- , "month", "quarter","semester", "year"), -->
-
-Suppose, z is the index set of the tsibble, x,y are two linear granularities with  $order(x) <  order(y)$.
-Also, $f_{(x, y)}$ denotes the accessor function for computing circular granularity "x_y" and $c(x, y)$ is a constant which relates x and y.
-It is easy to see that for $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
-
-Then, the accessor function f can be used recursively to obtain any multiple-order-up granularities as follows:
-
-\begin{equation} \label{eq1}
-\begin{split}
-f_{(x,y)}(z) & = f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, y)}(z)-1) \\  
-  & =  f_{(x, x+1)}(z) + c(x, x+1)[f_{(x+1, x+2)}(z) + c(x+1, x+2)(f_{(x+2, y)}(z) - 1) - 1]\\
-  & =  f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, x+2)}(z) - 1) + c(x, x+1)c(x+1, x+2)(f_{(x+2, y)}(z) - 1)\\
-  & =  f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, x+2)}(z) - 1) + c(x, x+2)(f_{(x+2, y)}(z) - 1)\\
-  &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f_{(x
-  +i, x+i+1)}(z) - 1)\\
-\end{split}
-\end{equation}
-
-For illustration, let us consider the following hierarchy table.
-
-```{r table_accessor1}
-kable(tibble::tibble(
-  `linear granularities`  = c("second", "minute", "hhour", "hour", "day", "week"),
-  `Conversion factor` = c(60, 30, 2, 24, 7, 1), 
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", 1)), format = "markdown")
-```
-
-From Equation(1), we have
-\begin{equation} \label{eq1}
-\begin{split}
-f_{(hhour, week)}(z) & = f_{(hhour, hour)}(z) + c(hhour, hour)f_{(hour,day)}(z) + c(hhour, day)f_{(day, week)}(z) \\
-              & = {hhour\_hour}(z) + 2{hour\_day}(z) + 2*24{day\_week}(z) \\
-\end{split}
-\end{equation}
-
-### Mixed single-order-up granularities - circular or aperiodic
-
-<!-- This is valid in a scenario when the relationship between x and (x+1) is periodic (Define periodic in Definition section).
-However, when x and x+1 is not periodic, then some modification is required in Equation(1).
--->
-
-Suppose we extend the hierarchy table to include the linear granularities month and quarter.
-Since months consists of unequal number of days, any linear granularity which is higher in order than months will also have unequal number of days.
-Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
-However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
-
-```{r table_accessor2}
-kable(tibble::tibble(
-  `linear granularities`  = c("second", "minute", "hhour", "hour", "day", "month", "quarter"),
-  `Conversion factor` = c(60, 30, 2, 24, "aperiodic", 3, 1), 
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_month", "month_quarter", 1)), format = "markdown")
-```
-
-There can be three scenarios for obtaining calendar categorization here:
-- calendar categorization consisting of two linear granularities whose orders are less than day  
-- calendar categorization consisting of two linear granularities whose orders are more than month  
-- calendar categorization consisting of one linear granularity with order at most day and another with order at least month
-
-The calendar categorization resulting from the first two cases are periodic and has been handled in the earlier section.
-
-The calendar categorization resulting from the last case are aperiodic.
-Examples might include day of the quarter or hour of the month.
-In this section, we will see how to obtain aperiodic circular granularities of these types.
-
-\begin{equation} \label{eq1}
-\begin{split}
-f_{(hour, month)}(z) = f_{(hour, day)}(z) + c(hour, day)f_{(day, month)}(z)\\
-\end{split}
-\end{equation}
-
-Here, the first part of the equation is a single-order-up granularity which can be obtained using last section.
-The second part is not single-order-up and can not be broken down further since each month consists of different number of days.
-In this case, it is important for us to know which month of the year and if the year is a leap year to obtain day of the month.
-
-# Visualization
-
-<!-- Why visualize -->
-Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
-However, improper choice of models can lead to wrong predictions.
-One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
-
-<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
-
-<!-- Why distribution plot -->
-<!-- Huge data difficult to summaries using one statistic-->
-But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a data set. Soon enough following questions become more interesting:
-
- - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?
-
- - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
-More generally, how the variation is the data set changes across different percentiles/deciles?
-
-  - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median?
-
-This is when displaying a probability distribution becomes a potentially useful approach.
-
-<!-- What all can you see from a distribution plot-->
-
-The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
-
-<!-- Lit review on statistical distribution plot-->
-## Statistical distribution plots
-
-Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-The box plot is a compact distributional summary, displaying less detail than a histogram.
-Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
-
-The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.
-Violin plots [@Hintze1998-zi, 1998] display the density for all data points and not only the box.
-The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
-The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
-The probabilities for which the summarization is required can be chosen based on the requirement.
-These regions do not need to be contiguous and help identify multi-modality.
-The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
-Because this display just adds extra letter values, it suffers from the same problems as the original box plot, and multimodality is almost impossible to spot[@Wickham_undated-vr].
-
-Moreover, much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
-The deciles plots consist of 9 values that split the data set into ten parts and the percentile plot consists of 99 values that split the data set into hundred parts.
-A large data set is required before the extreme percentiles can be estimated with any accuracy.
-
-Finally, a density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution.
-Also, a Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
-Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
-
-## Harmony and Clashes
-
-We investigate some combinations of circular time granularities (periodic/aperiodic) which facilitate or hinder exploratory analysis.
-The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
-
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-\autoref{fig:allFig} (a) shows the letter value plot of electricity consumption of Victoria across days of the month for few months like January, February, April and December.
-Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
-
-\autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
-
-Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
-
-In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
-
-\autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
-
-From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
-
-```{r moy-dom,echo=FALSE, out.width = "70%", fig.align = 'center',fig.cap ="Letter-value plot of electricity consumption in Victoria across days of the month\n faceted by months January, February, April and December.
-"}
-
-# change all of these to gravitas function
-VIC <- read.csv("VIC2015/Data/demand_VIC.csv")
-VIC$Date <- as.Date(VIC$Date,origin = "1899-12-30")
-
-first_day_of_month_wday <- function(dx) {
-  day(dx) <- 1
-  wday(dx)
-}
-
-VIC <- VIC %>%mutate(Indx_Year = year(Date),
-                     Indx_Month = month(Date, label = FALSE, abbr = TRUE),
-                     Indx_Wk_Yr = week(Date),
-                     Indx_Wk_Month = ceiling((day(Date) + first_day_of_month_wday(Date) - 1) / 7),
-                     Indx_Day_Week = wday(Date, label=FALSE, abbr=TRUE,
-                                          week_start=1),
-
-                     Indx_Day_Month = day(Date),
-
-                     Indx_Day_Year =  yday(Date),
-
-                     Indx_Weekend=if_else(Indx_Day_Week %in% c(6,7),1,0),
-
-                     Indx_HlHr_day = Period,
-                     month = month(Date, label = FALSE, abbr = TRUE),
-         year =  year(Date),
-         yday =yday(Date),
-         wday = wday(Date, label=FALSE, abbr=TRUE,
-                     week_start=1),
-         bow = (wday - 1) * 48 + Period,
-         dom = day(Date),
-         bom = (dom - 1) * 48 + Period,
-         Weekend=if_else(wday %in% c(6,7),1,0),
-         Indx_hour = ceiling(Period/2),
-         Indx_Hour_Yr = Indx_hour + 24*(yday-1),
-         Indx_Hour_Month = Indx_hour + 24*(Indx_Day_Month-1),
-         Indx_Hour_Wk = Indx_hour + 24*(wday-1))
-
-VIC <- as_tibble(VIC)
-
-```
-
-```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
-Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
-
-par(mfrow = c(3, 2))
-
-VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Month %in% c(1,2,4,12))%>% ggplot(aes(Indx_Day_Month,OperationalLessIndustrial,group = Indx_Day_Month)) + geom_lv(aes(fill=..LV..),outlier.colour = "red",outlier.shape = 1) + scale_fill_brewer()+ facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks=seq(0,31,5)) + ggtitle("(a) Letter value plot")
-
-VIC_hod_dow <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
-  group_by(wday,Indx_hour) %>%
-  do({x <- .$OperationalLessIndustrial
-  map_dfr(.x = seq(0.1,0.9,0.1),
-          .f = ~ tibble(Quantile = .x,
-                            Value = quantile(x, probs = .x,na.rm=TRUE)))
-  }) %>%  filter(wday %in% c(1,2,6,7))
-
-VIC_hod_dow %>%ggplot(aes(x=Indx_hour,y=Value,col=as.factor(Quantile))) + geom_line() + facet_wrap(~wday)+ scale_x_continuous(breaks=seq(1, 24,5)) + ylab("")  + xlab("Day of the Week") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot")
-
-VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Day_Month %in% c(1,15,29,31))%>% ggplot(aes(as.factor(yday),OperationalLessIndustrial,group = yday)) + geom_boxplot()+ facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
-  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot")
-
-VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,4))%>% ggplot(aes(as.factor(Indx_Day_Month),OperationalLessIndustrial)) + geom_violin(alpha = 0.03)+ facet_wrap(~Indx_Wk_Month,nrow=3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks=seq(0,31,5))+ scale_y_continuous(breaks = seq(2000,9000,2000))+ ggtitle("(d) Violin plot")
-
-VIC%>% dplyr:::filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,5),Indx_Wk_Yr <20)%>% ggplot(aes(x=OperationalLessIndustrial,y=as.factor(Indx_Wk_Yr),group=Indx_Wk_Yr)) + geom_density_ridges2() +facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000,10000,3000)) + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot")
-
-VIC_moy_doy <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
-  group_by(Indx_Month,yday) %>%
-  do({x <- .$OperationalLessIndustrial
-  map_dfr(.x = seq(0.1,0.9,0.1),
-          .f = ~ tibble(Quantile = .x,
-                            Value = quantile(x, probs = .x,na.rm=TRUE)))
-  }) %>%  filter(Indx_Month %in% c(1,7,11))
-
-VIC_moy_doy %>%ggplot(aes(x=yday,y=Value,col=as.factor(Quantile),group=yday)) + geom_line() + facet_wrap(~Indx_Month)+ scale_x_continuous(breaks=seq(1, 336, 60)) + ylab("")  + xlab("Day of the Year") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
-
-# grid.arrange(p1,p2,p3,p4,nrow = 2)
-
-# 
-# x <- gridExtra::arrangeGrob (p5, p6, p1, p2, p3, p4,newpage = TRUE)
-# grid::grid.draw(x)
-
-# 
-# gridExtra::grid.arrange(
-#   p1,
-#   p2,
-#   p3,
-#   p4,
-#   p5,
-#   p6,
-#   nrow = 3,
-#   top = "Title of the page",
-#   bottom = grid::textGrob(
-#     "this footnote is right-justified",
-#     gp = grid::gpar(fontface = 3, fontsize = 9),
-#     hjust = 1,
-#     x = 1
-#   )
-# )
-# plot.new()
-# mtext('Various probability distribution plot of electricity consumption data of Victoria from 2012 to 2014 (top left) Letter Value plot by DoM and MoY (top right) Decile Plot by HoD and DoW (middle left) Box plot by DoY and DoM (middle right) Violin plot of DoM and WoM (bottom left) Ridge plot by WoM and WoY (bottom right) Decile plot by DoY and MoY', side = 3, line = - 10, outer = TRUE)
-```
-
-In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
-
-In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
-
-More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
-When C1 and C2 are used as aesthetics or facets, problems arise when we encounter empty combinations.
-Let $S_{i,j}$ be the set of combination of the levels of C1 and C2.
-In this example, we have 12 such sets $S_{i,j}$ because i can take 4 values and j can take 3 values.
-The graphs that don't work are those where many of these 12 sets are empty.
-In other words, if there are levels of the x-axis which are not spanned by levels of the faceted variable or vice versa we will have structurally empty sets leading to potential ineffective graphs.
-
-Thus, the common link that differentiates harmonies from clashes are:
-
-  a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis.
-
-  b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets.
-<!-- ## Distributions conditional on bivariate granularities -->
-
-The time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
-That neccessiates plotting one temporal granularity along the x-axis and the other one across facets.
-
-Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
-The harmonies when plotted together can help exploration.
-But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
-This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
-
-## Advice algorithm for exploring conditional probability distributions
-
-Recommendations for distribution plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
-They will vary depending on which granularity is placed on the x-axis and which one across facets.
-Assumptions are made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.
-Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot.
-
-Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
-These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
-Let us consider case by case and see which plots are better suitable in which scenarios.
-
- -  very high facet and x-axis levels
-
- X-axis variable is treated as a categorical variable and hence any plots which
 
@@
 
-\autoref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year.
+Figure \ref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year.
 The two households are similar when it comes to their usage in the morning hours.
@@
 
-```{r cust1_hd_my}
+```{r cust1_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the first household.
+10th, 25th, 50th, 75th and 90th percentile are shown.
+For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior.
+Evening hours generally exhibits a peak.
+Some extreme behavior is exhibited in the evening hours of May and June, however, overall energy comsumption does not seem to vary much across months.
+Median and 75th percentile is less than 0.25 KwH across most months."}
 library(ggplot2)
@@
 
-```{r cust3_hd_my}
+(ref:cust1_hd_my-cap) Something that I want to describe.
+
+```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household.
+10th, 25th, 50th, 75th and 90th percentile are shown.
+Distinct evening peaks for few hours across all months.
+The median comsumption also varies across months.
+This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January.
+This is evident from the 75th and 90th percentile of consumption."}
 data_sm3 %>% 
@@
 
-```{r cust1_hd_dw}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
-```
-<!-- # ``````{r cust2_hour-day_day-week} -->
-<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
-<!-- # ``` -->
-
-``````{r cust3_hd_dw}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
-```
-
-\autoref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1.
+Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1.
 Energy consumption typically goes up between 16 to 18 hours.
@@
 
-\autoref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2.
+Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2.
 Contrary to Household 1, Household 2's energy consumption increases through noon through 7pm.
@@
 
+```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the first household.
+A long tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears.
+If an hour has a distribution that seems like a straight line, they are extremely volatile.
+On weekdays, hours between 17 to 19 hours seems to be extremely volatile comapred to other hours.
+On weekends, energy consumption is volatile even in morning hours.
+"}
+
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
+```
+<!-- # ``````{r cust2_hour-day_day-week} -->
+<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
+<!-- # ``` -->
+
+```{r cust3_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the second household."}
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
+```
+
 A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
 Was it only for the months May and June, that this household consumed more energy than they usually do?
-\autoref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
+Figure \ref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
 In fact, months like November and December saw higher energy usage than typical even on other weekdays.
@@
 
-/autoref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
+Figure \ref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
 The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December.
@@
 
-```{r cust1_dw_my}
+```{r cust1_dw_my,fig.cap="this is a caption"}
 library(ggplot2)
@@
 
-```{r cust3_dw_my}
+```{r cust3_dw_my, fig.cap="this is a caption"}
 library(ggplot2)
@@
 We can keep on slicing and dicing the data in a similar fashion and derive more insights about the behavior of these households.
-
-# Case study: Analysis on cricket {#cricket}
-
-# Discussion {#sec:discussion}
-
-# Acknowledgements {-}
-
-# Bibliography
-
-<!-- ```{r write-bib} -->
-<!-- write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
-<!-- ``` -->

### 2019-08-25 smart_meter.Rmd
smart meter analysis added

@@
 
-```{r cust3_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the second household."}
+```{r cust3_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the second household.
+Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent.
+Hours in between these have a long right tail which are even higher for weekends."}
 data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
@@
 
-```{r cust1_dw_my,fig.cap="this is a caption"}
+```{r cust1_my_dw, fig.cap="Letter value plots of energy consumption across days of the week and months for the first household.
+Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.Outliers are colored in red.
+It can be noted that they do not typically consume more than 0.25 KwH in the weekdays.
+Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}
+
 library(ggplot2)
 
-data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days-week by month-year for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+
 ```
 
-```{r cust3_dw_my, fig.cap="this is a caption"}
+```{r cust3_my_dw, fig.cap="Letter value plots of energy consumption across days of the week and months for the first household.
+Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH.
+The extreme values in the tail of the distribution are represented by different letter values.
+In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays (represented by letter value D), which is far more higher than subsequent months and comparable to summer month December.
+For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}
+
 library(ggplot2)
 
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```

### 2019-08-25 smart_meter.Rmd
fixing captioning error

@@
 
+-  Across days of the week by months
+
 A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
 Was it only for the months May and June, that this household consumed more energy than they usually do?
-Figure \ref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
+Figure \ref{fig:cust1_my_dw} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
 In fact, months like November and December saw higher energy usage than typical even on other weekdays.
@@
 
-Figure \ref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
+Figure \ref{fig:cust3_my_dw} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
 The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December.
 Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.
-
--  Across days of the week by months
-
-```{r cust1_my_dw, fig.cap="Letter value plots of energy consumption across days of the week and months for the first household.
-Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.Outliers are colored in red.
-It can be noted that they do not typically consume more than 0.25 KwH in the weekdays.
-Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}
-
-library(ggplot2)
-
-data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days-week by month-year for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-
-```
-
-```{r cust3_my_dw, fig.cap="Letter value plots of energy consumption across days of the week and months for the first household.
-Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH.
-The extreme values in the tail of the distribution are represented by different letter values.
-In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays (represented by letter value D), which is far more higher than subsequent months and comparable to summer month December.
-For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}
-
-library(ggplot2)
-
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
 
@@
 We can keep on slicing and dicing the data in a similar fashion and derive more insights about the behavior of these households.
+
+```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
+Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.Outliers are colored in red.
+It can be noted that they do not typically consume more than 0.25 KwH in the weekdays.
+Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}
+
+data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days-week by month-year for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
+Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH.
+The extreme values in the tail of the distribution are represented by different letter values.
+In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays (represented by letter value D), which is far more higher than subsequent months and comparable to summer month December.
+For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}
+
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```

### 2019-08-25 main.Rmd
figure referencinf correctedremoving special characters

@@
 
-\autoref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year.
+Figure \ref{fig:cust1_hd_my} and \ref{fig:cust3_hd_my} shows the distribution of hourly consumption of household 1 and 2 across different months of the year.
 The two households are similar when it comes to their usage in the morning hours.
@@
 
-```{r cust3_hd_my}
+```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household.
+10th, 25th, 50th, 75th and 90th percentile are shown.
+Distinct evening peaks for few hours across all months.
+The median comsumption also varies across months.
+This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January.
+This is evident from the 75th and 90th percentile of consumption."}
 data_sm3 %>% 
@@
 
-```{r cust1_hd_dw}
+Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for household 1.
+Energy consumption typically goes up between 16 to 18 hours.
+On weekends, the energy consumption for these hours are even higher.
+The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour.
+This suggests that these hours are extremely volatile based on weather conditions/months.
+Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range.
+This is not true for the 17th or 18th hour of the day.
+The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home parents/ working from home professionals.
+
+Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for Household 2.
+Contrary to household 1, household 2's energy consumption increases through noon to 7pm.
+We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption for this household.
+This indicates that this household might have kids or someone working from home.
+
+```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the first household.
+A long tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears.
+If an hour has a distribution that seems like a straight line, they are extremely volatile.
+On weekdays, hours between 17 to 19 hours seems to be extremely volatile comapred to other hours.
+On weekends, energy consumption is volatile even in morning hours.
+"}
+
 data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
@@
 
-``````{r cust3_hd_dw}
+```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household.
+Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent.
+Hours in between these have a long right tail which are even higher for weekends."}
 data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
@@
 
-\autoref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1.
-Energy consumption typically goes up between 16 to 18 hours.
-On weekends, the energy consumption for these hours are even higher.
-The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour.
-This suggests that these hours are extremely volatile based on weather conditions/months.
-Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range.
-This is not true for the 17th or 18th hour of the day.
-The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home professionals.
-
-\autoref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2.
-Contrary to Household 1, Household 2's energy consumption increases through noon through 7pm.
-We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption from the households.
-This indicates that this household might have kids or someone working from home.
+-  Across days of the week by months
 
 A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
-Was it only for the months May and June, that this household consumed more energy than they usually do?
-\autoref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
+Was it only for the months May and June, that this household consumed more energy than they usually do \?
+Figure \ref{fig:cust1_my_dw} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
 In fact, months like November and December saw higher energy usage than typical even on other weekdays.
@@
 
-/autoref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
+Figure \ref{fig:cust3_my_dw} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
 The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December.
 Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.
-
--  Across days of the week by months
-
-```{r cust1_dw_my}
-library(ggplot2)
-
-data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
-```{r cust3_dw_my}
-library(ggplot2)
-
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
 
@@
 They do all energy intensive chores (e.g.
-laundry, vacuum cleaning) mostly in the evenings of weekends.
+laundry, vacuum cleaning) mostly in the weekends.
 
@@
 We can keep on slicing and dicing the data in a similar fashion and derive more insights about the behavior of these households.
+
+```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
+Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.Outliers are colored in red.
+It can be noted that they do not typically consume more than 0.25 KwH in the weekdays.
+Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}
+
+data_sm1 %>% granplot("month_year", 
+                      "day_week", 
+                      response = "general_supply_kwh", plot_type = "lv") +    ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
+Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH.
+The extreme values in the tail of the distribution are represented by different letter values.
+In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December.
+For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}
+
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
 

### 2019-08-25 smart_meter.Rmd
figure referencinf correctedremoving special characters

@@
 
-Figure \ref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year.
+Figure \ref{fig:cust1_hd_my} and \ref{fig:cust3_hd_my} shows the distribution of hourly consumption of household 1 and 2 across different months of the year.
 The two households are similar when it comes to their usage in the morning hours.
@@
 
-(ref:cust1_hd_my-cap) Something that I want to describe.
-
 ```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household.
@@
 
-Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1.
+Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for household 1.
 Energy consumption typically goes up between 16 to 18 hours.
@@
 This is not true for the 17th or 18th hour of the day.
-The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home professionals.
+The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home parents/ working from home professionals.
 
-Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2.
-Contrary to Household 1, Household 2's energy consumption increases through noon through 7pm.
-We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption from the households.
+Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for Household 2.
+Contrary to household 1, household 2's energy consumption increases through noon to 7pm.
+We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption for this household.
 This indicates that this household might have kids or someone working from home.
@@
 
-```{r cust3_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the second household.
+```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household.
 Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent.
@@
 A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
-Was it only for the months May and June, that this household consumed more energy than they usually do?
+Was it only for the months May and June, that this household consumed more energy than they usually do \?
 Figure \ref{fig:cust1_my_dw} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
@@
 Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.Outliers are colored in red.
+M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.Outliers are colored in red.
 It can be noted that they do not typically consume more than 0.25 KwH in the weekdays.
@@
 
-data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days-week by month-year for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm1 %>% granplot("month_year", 
+                      "day_week", 
+                      response = "general_supply_kwh", plot_type = "lv") +    ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
@@
 The extreme values in the tail of the distribution are represented by different letter values.
-In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays (represented by letter value D), which is far more higher than subsequent months and comparable to summer month December.
+In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December.
 For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}

### 2019-08-25 smart_meter.Rmd
spell grammar check

@@
 ---
-title: "Case study: Analysis on smart meter data"
+title: "Case study: Analysis of smart meter data"
 author: "Sayani Gupta"
@@
 Smart meters provide large quantities of measurements on energy usage for households across Australia, and indeed many parts of the world.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
@@
 <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
-It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
 
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in NewCastle, New South Wales and some parts of Sydney provides customer wise data on half hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
 It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at time across different granularities together can help identify different behavioral patterns.
+The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns.
 
 Let us compare the energy consumption distribution for two households from the customer trials.
-For the purpose of this analysis, we only take the data for the year 2013.
+For this analysis, we only take the data for the year 2013.
 Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
@@
 The two households are similar when it comes to their usage in the morning hours.
-The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage).
+The first household's extreme behavior is different from that of the second household in the evening hours of May and June (represented by higher 90th percentile energy usage).
 However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar.
 This implies that consuming more energy in the evening is more typical for household 2.
-For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2.
+For the evenings of October, November, and December, it is clear that household 1 typically consumes much less energy than household 2.
 Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH.
-This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely.
+This might be an indication that of usage of a heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely.
 All of these can be explained with more demographics of the household and weather conditions in that area.
@@
 
-Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for household 1.
+Figure \ref{fig:cust1_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for household 1.
 Energy consumption typically goes up between 16 to 18 hours.
-On weekends, the energy consumption for these hours are even higher.
+On weekends, the energy consumption for these hours is even higher.
 The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour.
 This suggests that these hours are extremely volatile based on weather conditions/months.
-Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range.
+Energy consumption distribution in the morning hours of weekdays has a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical weekday, we can say with reasonable certainty that the energy consumption would fall in that range.
 This is not true for the 17th or 18th hour of the day.
@@
 
-Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for Household 2.
-Contrary to household 1, household 2's energy consumption increases through noon to 7pm.
-We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption for this household.
+Figure \ref{fig:cust3_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for Household 2.
+Contrary to household 1, household 2's energy consumption increases through noon to 7 pm.
+We can't say with certainty if the morning hours from 10 am would manifest a higher or lower level of energy consumption for this household.
 This indicates that this household might have kids or someone working from home.
@@
 ```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the first household.
-A long tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears.
+A long-tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears.
 If an hour has a distribution that seems like a straight line, they are extremely volatile.
-On weekdays, hours between 17 to 19 hours seems to be extremely volatile comapred to other hours.
+On weekdays, hours between 17 to 19 hours seems to be extremely volatile compared to other hours.
 On weekends, energy consumption is volatile even in morning hours.
@@
 
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3)) +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
@@
 Hours in between these have a long right tail which are even higher for weekends."}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3)) +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
@@
 In fact, months like November and December saw higher energy usage than typical even on other weekdays.
-Still for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
+Still, for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
 The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.
 
-Figure \ref{fig:cust3_my_dw} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
-The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December.
+Figure \ref{fig:cust3_my_dw} shows that the second household uniformly distributes their regular activity like using dishwasher or heater/air conditioner throughout the week.
+The distribution of energy consumption on Monday, Friday and Sunday look quite similar for December.
 Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.
@@
 
- - First household tpically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June).
-They are strictly 9 to 5 professsionals, hence their energy consumption during day time is pretty consistent and only peaks up during few hours in the evening.
-They do all energy intensive chores (e.g.
-laundry, vacuum cleaning) mostly in the weekends.
+ - First household typically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June).
+They are strictly 9 to 5 professionals, hence their energy consumption during day time is pretty consistent and only peaks up during a few hours in the evening.
+They do all energy-intensive chores (e.g.
+laundry, vacuum cleaning) mostly on the weekends.
 
- - Second household tpically use air conditioners in summer months and heaters in winter months.
+ - Second household typically use air conditioners in summer months and heaters in winter months.
 This might be a family with children with stay-at-home parents/ work-from-home professionals with extremely varying energy consumption peaking up from noon.
 Their weekday behavior is not distinctly different from weekends.
-Also, they are more likely to distribute their energy intensive chores through out the week and not leave all them for the weekends.
+Also, they are more likely to distribute their energy-intensive chores throughout the week and not leave all them for the weekends.
 
-We can keep on slicing and dicing the data in a similar fashion and derive more insights about the behavior of these households.
+We can keep on slicing and dicing the data similarly and derive more insights about the behavior of these households.
 
@@
 
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```

### 2019-08-26 smart_meter.Rmd
only 9 months analysis included
adding analysis of weekends across weeks of the months

@@
 
-# Case study: Analysis on smart meter data {#smartmeter}
-
 Smart meters provide large quantities of measurements on energy usage for households across Australia, and indeed many parts of the world.
@@
 For this analysis, we only take the data for the year 2013.
+The first household has data from April and the second household has data from March.
+For the purpose of comparison, data for both households have been observed from April.
 Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
@@
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
-  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
-  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
+  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +ggtitle("") +
   scale_x_discrete(breaks = seq(0,23,2))
@@
              plot_type = "quantile", 
-             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + 
-  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
+             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") +
+  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
   scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
@@
 
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3)) +
+data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("") +
+  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") 
+  scale_x_discrete(breaks = seq(0,23,3)) +
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
@@
 Hours in between these have a long right tail which are even higher for weekends."}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3)) +
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +
+  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") +
+  scale_x_discrete(breaks = seq(0,23,3)) +
+  ggtitle("") +
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
@@
 
-To summarise, we can posit the following for the two households:
-
- - First household typically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June).
-They are strictly 9 to 5 professionals, hence their energy consumption during day time is pretty consistent and only peaks up during a few hours in the evening.
-They do all energy-intensive chores (e.g.
-laundry, vacuum cleaning) mostly on the weekends.
-
- - Second household typically use air conditioners in summer months and heaters in winter months.
-This might be a family with children with stay-at-home parents/ work-from-home professionals with extremely varying energy consumption peaking up from noon.
-Their weekday behavior is not distinctly different from weekends.
-Also, they are more likely to distribute their energy-intensive chores throughout the week and not leave all them for the weekends.
-
-We can keep on slicing and dicing the data similarly and derive more insights about the behavior of these households.
-
 ```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
@@
                       "day_week", 
-                      response = "general_supply_kwh", plot_type = "lv") +    ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
+                      response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +  #ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
@@
 
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
+  #ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + 
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
+
+-  Across week of the month and weekends
+
+Another aspect which might of interest is how these households distribute their monthly chores between Saturday and Sunday across weeks of the month.
+"Monthly" chores might include washing mattress, comforters, bed and pillow covers, vacuum cloth furniture or cleaning dishwasher or oven.
+Also some households might visit friends and family or have them come over on weekends at least once a month.
+Figures \ref{fig:cust1_wm_wndwd} and \ref{fig:cust3_wm_wndwd} shows that the letter value plot of energy consumption for Saturday and Sunday across weeks of the months.
+
+Figure \ref{fig:cust1_wm_wndwd} shows that median consumption of energy is higher on Sundays.
+Also Sundays have longer right-tails than Saturdays across all weeks of the month implying that Sunday is a little more active day for the household.
+And the forth weekend of the month, either they take it easy on themselves or stay out of their house for trips/visit friends.
+
+Figure \ref{fig:cust3_wm_wndwd} shows that household 2 has all-absorbing Sundays.
+They are either busy with weekly chores or have the entire family stay at home every Sunday.
+The first Saturday of the month looks pretty consistent compared to any other weeks of the month, implying either they are out for work/trips/visiting family.
+Also, 4th Sunday of the month has a long right tail implying they choose the last week of the month for doing their "monthly" chores.
+
+```{r cust1_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 1.
+The 4th week of the month sees low energy consumption for both Saturdays and Sundays.Typically Sundays are heavier on energy than Saturdays."}
+
+data_sm1 %>% 
+   create_gran("day_week") %>% 
+  filter(day_week %in%c("Sat", "Sun")) %>% 
+  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
+  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+```{r cust3_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 2.
+The forth  Sunday has high energy consumption and long right tails implying they choose the last Sunday of the month either to do monthly chores or have friends come over.
+The first Saturday of the week seems quite and consistent for this household."}
+
+data_sm3 %>%
+  create_gran("day_week") %>% 
+  filter(day_week %in%c("Sat", "Sun")) %>% 
+  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") + ggtitle("") +
+  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") + 
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
+To summarise, we can posit the following for the two households:
+
+ - First household typically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June).
+They are strictly 9 to 5 professionals, hence their energy consumption during day time is pretty consistent and only peaks up during a few hours in the evening.
+They do all energy-intensive chores (e.g.
+laundry, vacuum cleaning) mostly on the weekends.
+Also, part of the forth weekend of the month is spent out of the house visiting friends/family or going out for leisure trips.
+
+ - Second household typically use air conditioners in summer months and heaters in winter months.
+This might be a family with children with stay-at-home parents/ work-from-home professionals with extremely varying energy consumption peaking up from noon.
+Their weekday behavior is not distinctly different from weekends.
+Also, they are more likely to distribute their energy-intensive chores throughout the week and not leave all them for the weekends.
+Sundays are still pretty heavy on energy for this household when all members stay at home.
+Also, they sometimes spend the last Sunday of the month doing "monthly" chores or have people come over.
+
+We can keep on slicing and dicing the data similarly and derive more insights about the behavior of these households.

### 2019-08-26 smart_meter.Rmd
smart meter analysis uploaded

@@
 
-Figure \ref{fig:cust1_hd_my} and \ref{fig:cust3_hd_my} shows the distribution of hourly consumption of household 1 and 2 across different months of the year.
+Figure \ref{fig:cust1_hd_my} shows the distribution of hourly consumption of both households across different months of the year.
 The two households are similar when it comes to their usage in the morning hours.
@@
 For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior.
-Evening hours generally exhibits a peak.
-Some extreme behavior is exhibited in the evening hours of May and June, however, overall energy comsumption does not seem to vary much across months.
-Median and 75th percentile is less than 0.25 KwH across most months."}
+Some extreme behavior is exhibited in the evening hours of May and June for household 1, however, overall energy comsumption does not seem to vary much across months.
+Median and 75th percentile is less than 0.25 KwH across most months.
+For household 2, distinct evening peaks for observed for few evening hours across all months.
+The median comsumption also varies across months.
+They are likely to be using heaters in colder months and air conditioners in hot months like January."}
 library(ggplot2)
@@
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
-  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
+  ggtitle("Quantile plot across hour-of-day by month-ofyear for Household 1") +
   scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +ggtitle("") +
   scale_x_discrete(breaks = seq(0,23,2))
-```
 
-```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household.
-10th, 25th, 50th, 75th and 90th percentile are shown.
-Distinct evening peaks for few hours across all months.
-The median comsumption also varies across months.
-This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January.
-This is evident from the 75th and 90th percentile of consumption."}
 data_sm3 %>% 
@@
              plot_type = "quantile", 
-             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") +
-  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
+             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  +
+  ggtitle("Quantile plot across hour-of-day by month-ofyear for Household 1") + 
   scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
@@
 
+<!-- # ```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household.
+10th, 25th, 50th, 75th and 90th percentile are shown.
+Distinct evening peaks for few hours across all months.
+The median comsumption also varies across months.
+This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January.
+This is evident from the 75th and 90th percentile of consumption."} -->
+<!-- #  -->
+<!-- # ```  -->
+
 -  Hourly usage across days of the week
 
-Figure \ref{fig:cust1_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for household 1.
+Figure \ref{fig:cust1_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for both households.
 Energy consumption typically goes up between 16 to 18 hours.
@@
 
-Figure \ref{fig:cust3_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for Household 2.
+For the second household, the distribution of hourly energy usage across different days of the week through violin plots for Household 2.
 Contrary to household 1, household 2's energy consumption increases through noon to 7 pm.
@@
 
-```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the first household.
-A long-tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears.
+```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the two households  A long-tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears.
 If an hour has a distribution that seems like a straight line, they are extremely volatile.
-On weekdays, hours between 17 to 19 hours seems to be extremely volatile compared to other hours.
+For household 1, hours between 17 to 19 hours (weekdays) seem to be volatile compared to other hours.
 On weekends, energy consumption is volatile even in morning hours.
-"}
+For the second one, hours before 10am and beyond 10pm have a bulge implying those hours are pretty consistent.
+Hours in between these are variable."}
 
 data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("") +
-  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") 
+  ggtitle("Violin plot across hour-of-day and day-of-week for Household 1") +
   scale_x_discrete(breaks = seq(0,23,3)) +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+
+data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +
+  ggtitle("Violin plot across hour-of-day and day-of-week for Household 2") +
+  scale_x_discrete(breaks = seq(0,23,3)) +
+  ggtitle("") +
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
@@
 
-```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household.
+<!-- # ```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household.
 Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent.
-Hours in between these have a long right tail which are even higher for weekends."}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +
-  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") +
-  scale_x_discrete(breaks = seq(0,23,3)) +
-  ggtitle("") +
-  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
+Hours in between these have a long right tail which are even higher for weekends."} -->
+<!-- #  -->
+<!-- # ``` -->
 
@@
 
-```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
-Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.Outliers are colored in red.
-It can be noted that they do not typically consume more than 0.25 KwH in the weekdays.
-Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}
+```{r cust1_my_dw, fig.cap = "LV plots across days of the week and months.
+They convey detailed information about tails where outliers (colored in red) are unexpected observations rather than extreme ones.
+M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.
+Household 1's tail behaviors are not distinctly visible because of the consistency in behavior on weekdays.
+For the second household, the F letter value seem to range between 0.75 to 1.25 KwH on weekends and 0.25 to 0.75 KwH on weekdays in Mar, which is very different from Dec.
+This reinforces the fact that their behavior changes across months but not so much between days-of-week."}
 
@@
                       "day_week", 
-                      response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +  #ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
+                      response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +  
+  ggtitle("LV plot across day-of-week by month-of-year for Household 1") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+
+data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
+  ggtitle("LV plot across day-of-week by month-of-year for Household 2")  + 
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
@@
 
-```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
+<!-- # ```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household.
 Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH.
@@
 In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December.
-For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}
-
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
-  #ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + 
-  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
+For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."} -->
+<!-- #  -->
+<!-- #  -->
+<!-- # ``` -->
 
@@
 
-Figure \ref{fig:cust1_wm_wndwd} shows that median consumption of energy is higher on Sundays.
+Figure \ref{fig:cust1_wm_wndwd} shows that median consumption of energy for both households.
+For household 1,  it is higher on Sundays.
 Also Sundays have longer right-tails than Saturdays across all weeks of the month implying that Sunday is a little more active day for the household.
 And the forth weekend of the month, either they take it easy on themselves or stay out of their house for trips/visit friends.
-
-Figure \ref{fig:cust3_wm_wndwd} shows that household 2 has all-absorbing Sundays.
+Household 2 has all-absorbing Sundays.
 They are either busy with weekly chores or have the entire family stay at home every Sunday.
@@
 
-```{r cust1_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 1.
-The 4th week of the month sees low energy consumption for both Saturdays and Sundays.Typically Sundays are heavier on energy than Saturdays."}
+```{r cust1_wm_wndwd, fig.cap = "LV plot for Saturday and Sunday across weeks-of-month.
+For household 1, the 4th week of the month sees low energy consumption for both days.
+Typically Sundays are heavier on energy than Saturdays.
+For household 2, the forth  Sunday has high energy consumption and long right tails implying more variable behavior.
+The first Saturday of the week seems quiet and consistent for this household."}
 
@@
   granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
-  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") +
+  ggtitle("Letter value plot across day-of-week by week-of-month for Household 1") +
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
-```{r cust3_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 2.
-The forth  Sunday has high energy consumption and long right tails implying they choose the last Sunday of the month either to do monthly chores or have friends come over.
-The first Saturday of the week seems quite and consistent for this household."}
 
@@
   granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") + ggtitle("") +
-  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") + 
+    ggtitle("Letter value plot across day-of-week by week-of-month for Household 2") + 
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
 ```
+
+<!-- # ```{r cust3_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 2.
+The forth  Sunday has high energy consumption and long right tails implying they choose the last Sunday of the month either to do monthly chores or have friends come over.
+The first Saturday of the week seems quite and consistent for this household."} -->
+<!-- # -->
+<!-- # -->
+<!-- # ``` -->
+<!-- # -->
 
@@
 Also, they sometimes spend the last Sunday of the month doing "monthly" chores or have people come over.
-
-We can keep on slicing and dicing the data similarly and derive more insights about the behavior of these households.

### 2019-08-27 main.Rmd
single order up granularity updated

--- 2019-08-25 main.Rmd
@@
 The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
+
+Let us define the arrangement of these coarser temporal units from finer temporal units in an hierarchy table.
+For example, in "Mayan" calendar, one day is referred to as 1 kin and they believed that the universe is destroyed at the start of every cycle (2,880,000 days).
+The Mayan calendar was structured as follows:
+
+ - 1 kin = 1 day
+ - 1 uinal = 20 kin
+ - 1 tun  = 18 uinal
+ - 1 katun = 20 tun
+ - 1 baktun = 20 katun
+
+Thus, the hierarchy table for the "Mayan" calendar would look like the following:
+
+| Temporal units	| Conversion factor 	| conversion in days 	| single-order-up accessors 	|
+|------------	|-------------------	|--------------------	|---------------------------	|
+| 1 kin      	| 1 day             	|                    	| kin_uinal                 	|
+| 1 uinal    	| 20 kin            	| 20 days            	| uinal_tun                 	|
+| 1 tun      	| 18 uinal          	| 360 days           	| tun_katun                 	|
+| 1 katun    	| 20 tun            	| 7200 days          	| katun_baktun              	|
+| 1 baktun   	| 20 katun          	| 144,000 days       	| 1                         	|
+
+Single-order-up granularities like kin_uinal(kin of the uinal), uinal_tun(uinal of the tun) can be computed using modular arithmatic in this periodic set up.
 
@@
 
-Smart meters provide large quantities of measurements on energy usage for households across Australia, and indeed many parts of the world.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-
-<!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
-It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in NewCastle, New South Wales and some parts of Sydney provides customer wise data on half hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at time across different granularities together can help identify different behavioral patterns.
-
-Let us compare the energy consumption distribution for two households from the customer trials.
-For the purpose of this analysis, we only take the data for the year 2013.
-Let us look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
-
-```{r search_gran}
-
-library(gravitas)
-library(tsibble)
-load("data/smart_meter_13_cust1.rda")
-load("data/smart_meter_13_cust2.rda")
-load("data/smart_meter_13_cust3.rda")
-
-data_sm1 <- smart_meter_13_cust1 %>% as_tsibble()
-data_sm2 <- smart_meter_13_cust2 %>% as_tsibble()
-data_sm3 <- smart_meter_13_cust3 %>% as_tsibble()
-```
-
-<!-- Refine your search of possible temporal granularities by altering arguments in search_gran() -->
-
-<!-- The default for search gran in this case, provides temporal granularities ranging from half-hour to year.
-If these options are considered too many, the default options can be modified to limit the possibilities.
-For example, the most coarse temporal unit can be set to be “month”.
--->
-
-<!-- # ```{r} -->
-<!-- # library(gravitas) -->
-<!-- # library(dplyr) -->
-<!-- # data_sm %>% search_gran(ugran = "year", filter_in = "event_key", filter_out = c("quarter", "semester")) -->
-<!-- # ``` -->
-
-<!-- # ```{r} -->
-<!-- # data_sm %>% harmony(ugran = "month", filter_in = "event_key", filter_out = c("quarter", "semester"), facet_h = 31) -->
-<!-- # ``` -->
-
- -  Hourly usage across months of the year
-
-\autoref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year.
-The two households are similar when it comes to their usage in the morning hours.
-The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage).
-However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar.
-This implies that consuming more energy in the evening is more typical for household 2.
-For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2.
-Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH.
-This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely.
-All of these can be explained with more demographics of the household and weather conditions in that area.
-
-Let us dive deeper and see the hourly usage across different days of the week for these two households.
-
-<!-- ```{r cust1_hd_my} -->
-
-<!-- library(ggplot2) -->
-<!-- data_sm1 %>%  -->
-<!--   group_by(month_year = lubridate::month(reading_datetime), -->
-<!--            hour_day = lubridate::hour(reading_datetime)) %>%  -->
-<!--   summarise(hourly_tot = mean(general_supply_kwh)) %>%  -->
-<!--   ggplot(aes(x = hour_day, y = month_year)) + facet_wrap(~ month_year)+ -->
-<!--   geom_point() +  -->
-<!--   ggtitle("Bar plot of energy consumption across hours of day by month of years for Household 1") -->
-<!-- ``` -->
-
-```{r cust1_hd_my}
-library(ggplot2)
-data_sm1 %>% granplot("month_year", "hour_day", response = "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
-```{r cust3_hd_my}
-data_sm3 %>% granplot("month_year", "hour_day", "general_supply_kwh", plot_type = "quantile", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("Letter value plot of energy consumption across hours of day by month of years for Household 2") + scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
--  Hourly usage across days of the week
-
-```{r cust1_hour-day_day-week}
-data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Boxplot of energy consumption for Customer 1")
-```
-<!-- # ``````{r cust2_hour-day_day-week} -->
-<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
-<!-- # ``` -->
-
-``````{r cust3_hour-day_day-week}
-data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Boxplot of energy consumption for Customer 3")
-```
-
-\autoref{fig:cust1_hd_my} shows distribution of hourly energy usage for two households across different days of the week through decile plots.
-It is clear that household 1 deviates from their typical behavior only on weekends.
-A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June.
-Was it only for the months May and June, that this household consumed more energy than they usually do?
-\autoref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays.
-In fact, months like November and December saw higher energy usage than typical even on other weekdays.
-Still for this household, it can safely said that most of their activities which involve higher energy consumption are concentrated on weekends.
-The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.
-
-/autoref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week.
-The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December.
-Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.
-
--  Across days of the week by months
-
-```{r cust1_dw_my}
-library(ggplot2)
-
-data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
-```{r cust3_dw_my}
-library(ggplot2)
-
-data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-```
-
--  Half-Hourly usage across days of the week
-
-Now, let us check which half hours are the most active for Household 1.
-For most days, half hours between 34th and 38th  half hour of the day are higher than other days.
-The energy consumption in these hours on Fridays match with that os Saturday and Sunday.
-
-<!-- ```{r cust1_hhour-day_day-week} -->
-<!-- data_sm1 %>% -->
-<!--    filter(lubridate::hour(reading_datetime)>=23) %>% -->
-<!--   granplot("hhour_day","day_week", response = "general_supply_kwh", plot_type = "lv") + -->
-<!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->
-<!-- ``` -->
-
-<!-- <!-- # ```{r cust2_hhour-day_day-week} --> -->
-<!-- <!-- # data_sm2 %>% granplot("day_week", "week_month", "general_supply_kwh", plot_type = "decile") + ggtitle("Boxplot of energy consumption for Customer 2") --> -->
-<!-- <!-- # ``` --> -->
-
-<!-- ```{r cust3_hhour-day_day-week} -->
-<!-- data_sm3 %>% create_gran("hhour_day") %>% -->
-<!--    filter(hhour_day>=23) %>% -->
-<!--   granplot("hhour_day","day_week",response = "general_supply_kwh", plot_type = "lv") + -->
-<!--   ggtitle("Letter value plot of energy consumption across half-hour of the day across days of the week for Household 1") -->
-<!-- ``` -->
-
-```
-
 # Case study: Analysis on cricket {#cricket}

### 2019-08-27 smart_meter.Rmd
single order up granularity updated

@@
 ---
-title: "Case study: Analysis of smart meter data"
+title: 'Case study: Analysis of smart meter data'
 author: "Sayani Gupta"
@@
 output:
+  html_document:
+    df_print: paged
   pdf_document:

### 2019-09-09 main.Rmd
adding cricket example

@@
 library(ggridges)
+library(viridis)
+library(tsibble)
+library(gravitas)
 opts_chunk$set(
@@
 
+Th application is not only restricted to temporal data.
+We provide an example of cricket to illustrate how this can be generalised in other applications.
+The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
+With eight teams, each team plays each other twice in a home-and-away round-robin format in the league phase.
+In a Twenty20 game the two teams have a single innings each, which is restricted to a maximum of 20 overs.
+Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls.
+We have sourced the the ball by ball data for IPL 2008 from https://www.kaggle.com/littleraj30/indian-premier-league-2019-ball-by-ball.
+There are many interesting questions that can possibly be answered with such a dataset, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
+
+The dataset contains the information on match_id, inning, batting team, bowling team, over of the innings, balls of the over, total runs and many more.
+A hierarchy table like [] can be construed for this game format:
+
+```{r hierarchy, echo=TRUE}
+hierarchy_model <- tibble::tibble(units = c("index", "ball", "over", "inning", "match"), convert_fct  = c(1, 6, 20, 2, 1))
+hierarchy_model
+```
+
+Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
+
+Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that.
+
+a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions?
+
+b) How the scores vary per each over each periodic granularities like ball of the over, ball of the innings, over of the inning, over of the match, inning of the match and others.
+
+We will look at the ball by ball data for all batting teams.Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
+
+```{r read_data, echo=TRUE}
+# reading ball by ball data
+cricket <-  read_csv("deliveries.csv")
+
+# removing cases where length of overs were more or less than 6 
+# extra run is wide_runs + noball_runs + bye_runs +  legbye_runs + penalty_runs
+
+# cricket_filter <- data %>% 
+#   filter(extra_runs == 0)%>%
+#   group_by(batting_team, match_id, inning, over) %>% 
+#   summarise(n = length(ball)) %>% 
+#   filter(n==6) %>% 
+#   mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
+# 
+# 
+# cricket_periodic <- cricket %>% 
+#   mutate(cricket_key = paste(batting_team, match_id, inning, over, sep = "_")) %>% 
+#   filter(cricket_key %in% cricket_filter$key) %>% 
+#   mutate(balls = purrr::rep_along(match_id, 1:6))
+
+# 
+# match_select <- cricket %>% 
+#    filter(ball<=6) %>%
+#   group_by(batting_team, match_id, inning)%>%
+#   summarise(n = length(over)) %>% filter(n==120) %>% 
+#   mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
+# 
+# cricket %>% 
+#   group_by(match_id, inning, batting_team) %>% 
+#   summarise(n_over = length(over))
+
+# each over needs to have 6 balls
+
+cricket_fltr_ball <- cricket %>% 
+    filter(wide_runs + noball_runs== 0) %>%    # filtering out scores that are gained by no balls or wide
+  # still have to filter out overs for which number of balls per over less than 6
+    group_by(batting_team, match_id, inning, over) %>% 
+    summarise(n_over = length(ball)) %>% 
+    filter(n_over==6) %>% 
+    mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
+
+cricket_over_crctd <- cricket %>% 
+   mutate(cricket_key = paste(batting_team, match_id, inning, over, sep = "_")) %>% 
+   filter(cricket_key %in% cricket_fltr_ball$key) %>% 
+    filter(wide_runs + noball_runs== 0) %>% 
+   mutate(ball_per_over = purrr::rep_along(match_id, 1:6))
+
+# each innings need to have 20 overs
+
+cricket_fltr_over <- cricket_over_crctd %>%
+  filter(inning %in% c(1,2)) %>% 
+  group_by(inning, match_id) %>%
+           summarise(n = length(over)) %>% filter(n==120) %>% 
+           mutate(key = paste(match_id, inning, sep = "_"))
+
+cricket_over_inning_crctd <- cricket_over_crctd %>% 
+   mutate(over_key = paste(match_id, inning, sep = "_")) %>% 
+   filter(over_key %in% cricket_fltr_over$key) %>% 
+    filter(wide_runs + noball_runs== 0) %>% 
+   mutate(ball_per_over = purrr::rep_along(match_id, 1:6)) %>% 
+   filter(inning %in% c(1,2))
+
+# each match needs to have 2 innings
+
+cricket_filtr_inning <- cricketdata %>% group_by(match_id) %>% summarise(length(inning)) %>% filter(`length(inning)`==240)
+
+cricketdata <- cricket_over_inning_crctd %>% 
+         filter(match_id %in% cricket_filtr_inning$match_id) %>% 
+    filter(wide_runs + noball_runs== 0) %>% 
+   mutate(ball_per_over = purrr::rep_along(match_id, 1:6)) %>% 
+   filter(inning %in% c(1,2))
+
+save(cricketdata, file = "data/cricketdata.rda")
+
+cricket_tsibble <- cricketdata %>% 
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+
+cricketdata %>% mutate(data_index = row_number()) %>% as_tsibble(index = data_index) %>% dynamic_create_gran(hierarchy_model, "ball_over")
+
+```
+
+data_MI_dec <- data_MI %>%
+  group_by(balls_game) %>%
+  do({x <- .$total_runs
+  map_dfr(.x = c(0.25,0.5, 0.75, 0.9),
+          .f = ~ tibble(Quantile = .x,
+                            Value = quantile(x, probs = .x,na.rm=TRUE)))
+  })
+
+data_MI_dec %>%
+  ggplot(aes(x = balls_game,y = Value,col = as.factor(Quantile))) +
+  geom_step() +
+  xlab("Balls per match") + scale_x_continuous(breaks = seq(1, 120, 12)) +
+  ggtitle("Quartile plot of total runs per ball across balls per match")
+
+```
+
+The scores/ball increases from 1st to 120th ball in the 90th percentile.
+In the last 2 overs, players are more vulnerable to get more scores, which is evident from the fact that in 90 percent of cases, their scores vary between 4 and 6.
+Till around 50th ball, players play safe and may get no runs per ball in 25% of the times, after which they more likely to get one score per ball.
+
+```{r over, echo=TRUE}
+data_MI %>% ggplot(aes(x = as.factor(over), y= total_runs)) + ggtitle("Letter value plot of runs per ball across each over of a match") + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1, k = 5) + xlab("overs per match")
+
+# data_MI %>% ggplot(aes(y =as.factor(over), x= total_runs)) + geom_density_ridges2() + ylab("Balls per  over")
+```
+
+In the first two overs, players play really safe and make no runs 50% of the times.
+The median is constant at 1 for all overs after the 2nd over.
+It is interesting to see the distribution of tails in this plot.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
+Till the first four overs, scoring a 4 or a 6 is extreme.
+The region which covers the tail 3.13% falls between 4 and 6 till up to 13th over.
+It is interesting to see how the letter values of scores (D, E and F) moves up, as we move towards the end of the game.
+
+```{r quarter, echo=TRUE}
+data_MI %>% ggplot(aes(x = as.factor(quarter_match), y= total_runs)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per ball across each quarter of a match")
+
+# data_MI %>% ggplot(aes(y =as.factor(quarter), x= total_runs)) + geom_density_ridges2() + ylab("quarters per match") + xlab(" Total runs per ball")
+```
+
+4th quarter of the match witnesses low density(mass) at scores 0, 1 and more at 4/6. The density at score 6 increases with the quarters of the match.
+
+```{r semester, echo=TRUE}
+# data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match")
+
+data_MI %>% ggplot(aes(x = total_runs, y= as.factor(semester_match))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per ball across each semesters of a match")
+
+# data_MI %>% ggplot(aes(y =as.factor(semester), x= total_runs, fill=factor(..quantile..))) +   stat_density_ridges(
+#     geom = "density_ridges_gradient", calc_ecdf = TRUE,
+#     quantiles = 4, quantile_lines = TRUE
+#   )  + ylab("semester") + xlab(" Total runs per ball") +
+#   scale_fill_viridis(discrete = TRUE, name = "Quartiles")
+```
+
+In the second semester of the game, there is more probability to score more than zero as can be seen from the drop in peaks of the distribution at 0 and increased peak at 1, 2, 4 and 6.
+
+Now, it will be interesting to see the distribution of runs per ball across two categorisation together:
+
+```{r bivariate1, echo=TRUE}
+# data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match")
+
+data_MI %>% ggplot(aes(x = as.factor(over_quarter), y = total_runs))+  geom_violin()+
+                     facet_wrap(~quarter_semester) + xlab("overs per quarter")+ ylab(" Total runs per ball") + ggtitle("Boxplot of runs per ball across each over_quarter and quarter_semester") 
+```
+
+# runs per over - as variation would be more
+
+```{r checking}
+data_MI_over <- data_MI %>% group_by(match_id, over) %>% summarise(runs_over = sum(total_runs)) %>% 
+    mutate(quarter_match = case_when(
+
+    between(over,1,5) ~ 1,
+    between(over,6,10) ~ 2,
+    between(over,11,15) ~ 3,
+    between(over,16,20) ~ 4
+  )) %>% 
+   mutate(semester_match = case_when(
+
+    between(quarter_match,1,2) ~ 1,
+    between(quarter_match,3,4) ~ 2
+  ))  %>% 
+   mutate(over_quarter = 
+  if_else(over%%5==0,5, over%%5)
+   ) %>%  mutate(quarter_semester = 
+  if_else(quarter_match%%2==0,2, quarter_match%%2)
+   ) %>% 
+  rename(over_match = over)
+
+```
+
+```{r over_match, echo=TRUE}
+data_MI_over %>% ggplot(aes(x = as.factor(over_match), y= runs_over)) + ggtitle("Letter value plot of runs per over across each over of a match") + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1, k = 5) + xlab("overs per match")
+
+# data_MI %>% ggplot(aes(y =as.factor(over), x= total_runs)) + geom_density_ridges2() + ylab("Balls per  over")
+```
+
+```{r over_quarter, echo=TRUE}
+data_MI_over %>% ggplot(aes(x = as.factor(over_quarter), y= runs_over)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per over across each quarter of a match")
+
+data_MI_over %>% ggplot(aes(x = as.factor(over_quarter), y= runs_over)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per over across each quarter of a match") + facet_grid(~quarter_match)
+
+# data_MI %>% ggplot(aes(y =as.factor(quarter), x= total_runs)) + geom_density_ridges2() + ylab("quarters per match") + xlab(" Total runs per ball")
+```
+
+```{r quarter_match, echo=TRUE}
+# data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match")
+
+data_MI_over %>% ggplot(aes(x = runs_over, y= as.factor(quarter_match))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per overs across each semesters of a match")
+
+data_MI_over %>% ggplot(aes(x = runs_over, y= as.factor(over_quarter))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per overs across each semesters of a match") + facet_wrap(~quarter_match)
+
+# data_MI %>% ggplot(aes(y =as.factor(semester), x= total_runs, fill=factor(..quantile..))) +   stat_density_ridges(
+#     geom = "density_ridges_gradient", calc_ecdf = TRUE,
+#     quantiles = 4, quantile_lines = TRUE
+#   )  + ylab("semester") + xlab(" Total runs per ball") +
+#   scale_fill_viridis(discrete = TRUE, name = "Quartiles")
+```
+
 # Discussion {#sec:discussion}

### 2019-09-13 main.Rmd
cleaning up folder and main file

--- 2019-08-27 main.Rmd
@@
   `Conversion factor` = c(60, 30, 2, 24, 7, 1),
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", 1)), format = "markdown")
+  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", 1)
+), format = "markdown")
 ```
@@
   `Conversion factor` = c(60, 30, 2, 24, "aperiodic", 3, 1),
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_month", "month_quarter", 1)), format = "markdown")
+  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_month", "month_quarter", 1)
+), format = "markdown")
 ```
@@
 VIC <- read.csv("VIC2015/Data/demand_VIC.csv")
-VIC$Date <- as.Date(VIC$Date,origin = "1899-12-30")
+VIC$Date <- as.Date(VIC$Date, origin = "1899-12-30")
 
@@
 
-VIC <- VIC %>%mutate(Indx_Year = year(Date),
+VIC <- VIC %>% mutate(
+  Indx_Year = year(Date),
   Indx_Month = month(Date, label = FALSE, abbr = TRUE),
@@
   Indx_Wk_Month = ceiling((day(Date) + first_day_of_month_wday(Date) - 1) / 7),
-                     Indx_Day_Week = wday(Date, label=FALSE, abbr=TRUE,
-                                          week_start=1),
+  Indx_Day_Week = wday(Date,
+    label = FALSE, abbr = TRUE,
+    week_start = 1
+  ),
 
@@
 
-                     Indx_Weekend=if_else(Indx_Day_Week %in% c(6,7),1,0),
+  Indx_Weekend = if_else(Indx_Day_Week %in% c(6, 7), 1, 0),
 
@@
   year = year(Date),
-         yday =yday(Date),
-         wday = wday(Date, label=FALSE, abbr=TRUE,
-                     week_start=1),
+  yday = yday(Date),
+  wday = wday(Date,
+    label = FALSE, abbr = TRUE,
+    week_start = 1
+  ),
   bow = (wday - 1) * 48 + Period,
@@
   bom = (dom - 1) * 48 + Period,
-         Weekend=if_else(wday %in% c(6,7),1,0),
-         Indx_hour = ceiling(Period/2),
-         Indx_Hour_Yr = Indx_hour + 24*(yday-1),
-         Indx_Hour_Month = Indx_hour + 24*(Indx_Day_Month-1),
-         Indx_Hour_Wk = Indx_hour + 24*(wday-1))
+  Weekend = if_else(wday %in% c(6, 7), 1, 0),
+  Indx_hour = ceiling(Period / 2),
+  Indx_Hour_Yr = Indx_hour + 24 * (yday - 1),
+  Indx_Hour_Month = Indx_hour + 24 * (Indx_Day_Month - 1),
+  Indx_Hour_Wk = Indx_hour + 24 * (wday - 1)
+)
 
 VIC <- as_tibble(VIC)
-
 ```
@@
 
-VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Month %in% c(1,2,4,12))%>% ggplot(aes(Indx_Day_Month,OperationalLessIndustrial,group = Indx_Day_Month)) + geom_lv(aes(fill=..LV..),outlier.colour = "red",outlier.shape = 1) + scale_fill_brewer()+ facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks=seq(0,31,5)) + ggtitle("(a) Letter value plot")
+VIC %>% filter(year %in% c(2012, 2013, 2014), Indx_Month %in% c(1, 2, 4, 12)) %>% ggplot(aes(Indx_Day_Month, OperationalLessIndustrial, group = Indx_Day_Month)) + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1) + scale_fill_brewer() + facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks = seq(0, 31, 5)) + ggtitle("(a) Letter value plot")
 
-VIC_hod_dow <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
-  group_by(wday,Indx_hour) %>%
-  do({x <- .$OperationalLessIndustrial
-  map_dfr(.x = seq(0.1,0.9,0.1),
-          .f = ~ tibble(Quantile = .x,
-                            Value = quantile(x, probs = .x,na.rm=TRUE)))
-  }) %>%  filter(wday %in% c(1,2,6,7))
+VIC_hod_dow <- VIC %>%
+  filter(year %in% c(2012, 2013, 2014)) %>%
+  group_by(wday, Indx_hour) %>%
+  do({
+    x <- .$OperationalLessIndustrial
+    map_dfr(
+      .x = seq(0.1, 0.9, 0.1),
+      .f = ~ tibble(
+        Quantile = .x,
+        Value = quantile(x, probs = .x, na.rm = TRUE)
+      )
+    )
+  }) %>%
+  filter(wday %in% c(1, 2, 6, 7))
 
-VIC_hod_dow %>%ggplot(aes(x=Indx_hour,y=Value,col=as.factor(Quantile))) + geom_line() + facet_wrap(~wday)+ scale_x_continuous(breaks=seq(1, 24,5)) + ylab("")  + xlab("Day of the Week") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot")
+VIC_hod_dow %>% ggplot(aes(x = Indx_hour, y = Value, col = as.factor(Quantile))) + geom_line() + facet_wrap(~wday) + scale_x_continuous(breaks = seq(1, 24, 5)) + ylab("") + xlab("Day of the Week") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot")
 
-VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Day_Month %in% c(1,15,29,31))%>% ggplot(aes(as.factor(yday),OperationalLessIndustrial,group = yday)) + geom_boxplot()+ facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
-  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot")
+VIC %>%
+  filter(year %in% c(2012, 2013, 2014), Indx_Day_Month %in% c(1, 15, 29, 31)) %>%
+  ggplot(aes(as.factor(yday), OperationalLessIndustrial, group = yday)) + geom_boxplot() + facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
+  xlab("Days of the Year") + scale_x_discrete(breaks = seq(0, 366, 60)) + theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot")
 
-VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,4))%>% ggplot(aes(as.factor(Indx_Day_Month),OperationalLessIndustrial)) + geom_violin(alpha = 0.03)+ facet_wrap(~Indx_Wk_Month,nrow=3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks=seq(0,31,5))+ scale_y_continuous(breaks = seq(2000,9000,2000))+ ggtitle("(d) Violin plot")
+VIC %>% filter(year %in% c(2012, 2013, 2014), Indx_Wk_Month %in% c(1, 2, 4)) %>% ggplot(aes(as.factor(Indx_Day_Month), OperationalLessIndustrial)) + geom_violin(alpha = 0.03) + facet_wrap(~Indx_Wk_Month, nrow = 3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + scale_y_continuous(breaks = seq(2000, 9000, 2000)) + ggtitle("(d) Violin plot")
 
-VIC%>% dplyr:::filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,5),Indx_Wk_Yr <20)%>% ggplot(aes(x=OperationalLessIndustrial,y=as.factor(Indx_Wk_Yr),group=Indx_Wk_Yr)) + geom_density_ridges2() +facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000,10000,3000)) + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot")
+VIC %>% dplyr:::filter(year %in% c(2012, 2013, 2014), Indx_Wk_Month %in% c(1, 2, 5), Indx_Wk_Yr < 20) %>% ggplot(aes(x = OperationalLessIndustrial, y = as.factor(Indx_Wk_Yr), group = Indx_Wk_Yr)) + geom_density_ridges2() + facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000, 10000, 3000)) + theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot")
 
-VIC_moy_doy <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
-  group_by(Indx_Month,yday) %>%
-  do({x <- .$OperationalLessIndustrial
-  map_dfr(.x = seq(0.1,0.9,0.1),
-          .f = ~ tibble(Quantile = .x,
-                            Value = quantile(x, probs = .x,na.rm=TRUE)))
-  }) %>%  filter(Indx_Month %in% c(1,7,11))
+VIC_moy_doy <- VIC %>%
+  filter(year %in% c(2012, 2013, 2014)) %>%
+  group_by(Indx_Month, yday) %>%
+  do({
+    x <- .$OperationalLessIndustrial
+    map_dfr(
+      .x = seq(0.1, 0.9, 0.1),
+      .f = ~ tibble(
+        Quantile = .x,
+        Value = quantile(x, probs = .x, na.rm = TRUE)
+      )
+    )
+  }) %>%
+  filter(Indx_Month %in% c(1, 7, 11))
 
-VIC_moy_doy %>%ggplot(aes(x=yday,y=Value,col=as.factor(Quantile),group=yday)) + geom_line() + facet_wrap(~Indx_Month)+ scale_x_continuous(breaks=seq(1, 336, 60)) + ylab("")  + xlab("Day of the Year") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
+VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
 
@@
 
+Smart meters provide large quantities of measurements on energy usage for households across Australia, and indeed many parts of the world.
+Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+
+<!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
+It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
+
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
+It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
+The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns.
+
+Let us see the behavior of typical and extreme behaviors of 50 households of this trial.
+We want to  look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
+
+```{r search_gran}
+
+library(gravitas)
+library(tsibble)
+load("data/sm_cust50.RData")
+
+library(ggplot2)
+sm_cust50 %>%
+  granplot("month_year", "hour_day",
+    response = "general_supply_kwh",
+    plot_type = "quantile",
+    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  ggtitle("Quantile plot across hour-of-day by month-ofyear") + ggtitle("") +
+  scale_x_discrete(breaks = seq(0, 23, 2)) + scale_color_brewer(type = "div",
+                                                                palette = "Dark2", direction = -1)
+
+#observation: the two peaks of the day are more ditinct in winter months compared to summaer months, where the energy profile is quite flat.
+
+# Hourly usage across days of the week
+sm_cust50 %>%
+  granplot("day_week", 
+           "hour_day",
+           "general_supply_kwh", 
+           plot_type = "quantile", 
+          quantile_prob = c(0.25, 0.5, 0.75),
+           response = "general_supply_kwh") +
+  ggtitle("") +
+  ggtitle("Quantile plot across hour-of-day and day-of-week") +
+  scale_x_discrete(breaks = seq(0, 23, 3)) +
+  scale_color_brewer(type = "div",
+                     palette = "Dark2", direction = -1)
+#observation: on saturday and sunday morning peaks are higher and little late (in absolte time) than other weekdays
+
+# Across days of the week by months
+sm_cust50 %>%
+  granplot("month_year",
+    "day_week",
+    response = "general_supply_kwh", plot_type = "quantile", 
+          quantile_prob = c(0.25, 0.5, 0.75)) + ggtitle("") +
+  ggtitle("Quantile plot across day-of-week by month-of-year")
+
+# Across week of the month and weekends
+
+sm_cust50 %>%
+  create_gran("day_week") %>%
+  filter(day_week %in% c("Sat", "Sun")) %>%
+  granplot("day_week", "week_month",
+           response = "general_supply_kwh",
+           plot_type = "lv") + 
+  ggtitle("") +
+  ggtitle("Letter value plot across day-of-week by week-of-month") +
+  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+```
+
 # Case study: Analysis on cricket {#cricket}
+
+Th application is not only restricted to temporal data.
+We provide an example of cricket to illustrate how this can be generalised in other applications.
+The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
+With eight teams, each team plays each other twice in a home-and-away round-robin format in the league phase.
+In a Twenty20 game the two teams have a single innings each, which is restricted to a maximum of 20 overs.
+Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls.
+We have sourced the the ball by ball data for IPL 2008 from https://www.kaggle.com/littleraj30/indian-premier-league-2019-ball-by-ball.
+There are many interesting questions that can possibly be answered with such a dataset, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
+
+The dataset contains the information on match_id, inning, batting team, bowling team, over of the innings, balls of the over, total runs and many more.
+A hierarchy table like [] can be construed for this game format:
+
+```{r hierarchy, echo=TRUE}
+hierarchy_model <- tibble::tibble(units = c("index", "ball", "over", "inning", "match"), convert_fct = c(1, 6, 20, 2, 1))
+hierarchy_model
+```
+
+Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
+
+Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that.
+
+a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions?
+
+b) How the scores vary per each over each periodic granularities like ball of the over, ball of the innings, over of the inning, over of the match, inning of the match and others.
+
+We will look at the ball by ball data for all batting teams.Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
+
+```{r read_data, echo=TRUE}
+
+cricket_tsibble <- cricketdata %>%
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+
+cricket_tsibble %>% dynamic_create_gran(
+  "ball_over",
+  hierarchy_model
+)
+
+cricket_tsibble %>% granplot("ball_over", "inning_match",
+  hierarchy_model,
+  response = "total_runs",
+  plot_type = "boxplot"
+)
+
+cricket_tsibble %>% granplot("over_inning",
+  "inning_match",
+  hierarchy_model,
+  response = "total_runs",
+  plot_type = "quantile"
+)
+
+cricket_tsibble %>%
+  dynamic_create_gran(
+    "over_inning",
+    hierarchy_model
+  ) %>%
+  dynamic_create_gran(
+    "inning_match",
+    hierarchy_model
+  ) %>%
+  granplot(
+    gran1 = "inning_match",
+    gran2 = "over_inning",
+    hierarchy_model,
+    response = "total_runs",
+    plot_type = "quantile"
+  )
+
+cricket_tsibble %>%
+  dynamic_create_gran(
+    "ball_over",
+    hierarchy_model
+  ) %>%
+  dynamic_create_gran(
+    "over_match",
+    hierarchy_model
+  ) %>%
+  # as_tibble() %>%
+  # group_by(match_id, inning, batting_team, over_inning,inning_match) %>%
+  # summarise(total_over = sum(total_runs)) %>%
+  granplot(
+    gran1 = "ball_over",
+    gran2 = "over_match",
+    hierarchy_model,
+    response = "total_runs",
+    plot_type = "quantile"
+  )
+```
+
+The scores/ball increases from 1st to 120th ball in the 90th percentile.
+In the last 2 overs, players are more vulnerable to get more scores, which is evident from the fact that in 90 percent of cases, their scores vary between 4 and 6.
+Till around 50th ball, players play safe and may get no runs per ball in 25% of the times, after which they more likely to get one score per ball.
+
+<!-- # ```{r over, echo=TRUE} -->
+<!-- # data_MI %>% ggplot(aes(x = as.factor(over), y= total_runs)) + ggtitle("Letter value plot of runs per ball across each over of a match") + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1, k = 5) + xlab("overs per match") -->
+
+<!-- # data_MI %>% ggplot(aes(y =as.factor(over), x= total_runs)) + geom_density_ridges2() + ylab("Balls per  over") -->
+<!-- ``` -->
+
+<!-- In the first two overs, players play really safe and make no runs 50% of the times.
+The median is constant at 1 for all overs after the 2nd over.
+It is interesting to see the distribution of tails in this plot.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
+Till the first four overs, scoring a 4 or a 6 is extreme.
+The region which covers the tail 3.13% falls between 4 and 6 till up to 13th over.
+It is interesting to see how the letter values of scores (D, E and F) moves up, as we move towards the end of the game.
+-->
+
+<!-- ```{r quarter, echo=TRUE} -->
+<!-- data_MI %>% ggplot(aes(x = as.factor(quarter_match), y= total_runs)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per ball across each quarter of a match") -->
+
+<!-- # data_MI %>% ggplot(aes(y =as.factor(quarter), x= total_runs)) + geom_density_ridges2() + ylab("quarters per match") + xlab(" Total runs per ball") -->
+<!-- ``` -->
+
+<!-- 4th quarter of the match witnesses low density(mass) at scores 0, 1 and more at 4/6. The density at score 6 increases with the quarters of the match. -->
+
+<!-- ```{r semester, echo=TRUE} -->
+<!-- # data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match") -->
+
+<!-- data_MI %>% ggplot(aes(x = total_runs, y= as.factor(semester_match))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per ball across each semesters of a match") -->
+
+<!-- # data_MI %>% ggplot(aes(y =as.factor(semester), x= total_runs, fill=factor(..quantile..))) +   stat_density_ridges( -->
+<!-- #     geom = "density_ridges_gradient", calc_ecdf = TRUE, -->
+<!-- #     quantiles = 4, quantile_lines = TRUE -->
+<!-- #   )  + ylab("semester") + xlab(" Total runs per ball") + -->
+<!-- #   scale_fill_viridis(discrete = TRUE, name = "Quartiles") -->
+<!-- ``` -->
+
+<!-- In the second semester of the game, there is more probability to score more than zero as can be seen from the drop in peaks of the distribution at 0 and increased peak at 1, 2, 4 and 6. -->
+
+<!-- Now, it will be interesting to see the distribution of runs per ball across two categorisation together: -->
+
+<!-- ```{r bivariate1, echo=TRUE} -->
+<!-- # data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match") -->
+
+<!-- data_MI %>% ggplot(aes(x = as.factor(over_quarter), y = total_runs))+  geom_violin()+ -->
+<!--                      facet_wrap(~quarter_semester) + xlab("overs per quarter")+ ylab(" Total runs per ball") + ggtitle("Boxplot of runs per ball across each over_quarter and quarter_semester")  -->
+<!-- ``` -->
+
+<!-- # runs per over - as variation would be more -->
+
+<!-- ```{r checking} -->
+<!-- data_MI_over <- data_MI %>% group_by(match_id, over) %>% summarise(runs_over = sum(total_runs)) %>%  -->
+<!--     mutate(quarter_match = case_when( -->
+
+<!--     between(over,1,5) ~ 1, -->
+<!--     between(over,6,10) ~ 2, -->
+<!--     between(over,11,15) ~ 3, -->
+<!--     between(over,16,20) ~ 4 -->
+<!--   )) %>%  -->
+<!--    mutate(semester_match = case_when( -->
+
+<!--     between(quarter_match,1,2) ~ 1, -->
+<!--     between(quarter_match,3,4) ~ 2 -->
+<!--   ))  %>%  -->
+<!--    mutate(over_quarter =  -->
+<!--   if_else(over%%5==0,5, over%%5) -->
+<!--    ) %>%  mutate(quarter_semester =  -->
+<!--   if_else(quarter_match%%2==0,2, quarter_match%%2) -->
+<!--    ) %>%  -->
+<!--   rename(over_match = over) -->
+
+<!-- ``` -->
+
+<!-- ```{r over_match, echo=TRUE} -->
+<!-- data_MI_over %>% ggplot(aes(x = as.factor(over_match), y= runs_over)) + ggtitle("Letter value plot of runs per over across each over of a match") + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1, k = 5) + xlab("overs per match") -->
+
+<!-- # data_MI %>% ggplot(aes(y =as.factor(over), x= total_runs)) + geom_density_ridges2() + ylab("Balls per  over") -->
+<!-- ``` -->
+
+<!-- ```{r over_quarter, echo=TRUE} -->
+<!-- data_MI_over %>% ggplot(aes(x = as.factor(over_quarter), y= runs_over)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per over across each quarter of a match") -->
+
+<!-- data_MI_over %>% ggplot(aes(x = as.factor(over_quarter), y= runs_over)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per over across each quarter of a match") + facet_grid(~quarter_match) -->
+
+<!-- # data_MI %>% ggplot(aes(y =as.factor(quarter), x= total_runs)) + geom_density_ridges2() + ylab("quarters per match") + xlab(" Total runs per ball") -->
+<!-- ``` -->
+
+<!-- ```{r quarter_match, echo=TRUE} -->
+<!-- # data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match") -->
+
+<!-- data_MI_over %>% ggplot(aes(x = runs_over, y= as.factor(quarter_match))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per overs across each semesters of a match") -->
+
+<!-- data_MI_over %>% ggplot(aes(x = runs_over, y= as.factor(over_quarter))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per overs across each semesters of a match") + facet_wrap(~quarter_match) -->
+
+<!-- # data_MI %>% ggplot(aes(y =as.factor(semester), x= total_runs, fill=factor(..quantile..))) +   stat_density_ridges( -->
+<!-- #     geom = "density_ridges_gradient", calc_ecdf = TRUE, -->
+<!-- #     quantiles = 4, quantile_lines = TRUE -->
+<!-- #   )  + ylab("semester") + xlab(" Total runs per ball") + -->
+<!-- #   scale_fill_viridis(discrete = TRUE, name = "Quartiles") -->
+<!-- ``` -->
 

### 2019-09-13 main.Rmd
adding all harmonies in cricket example

@@
 # change all of these to gravitas function
-VIC <- read.csv("VIC2015/Data/demand_VIC.csv")
+VIC <- read.csv("data/VIC2015/Data/demand_VIC.csv")
 VIC$Date <- as.Date(VIC$Date, origin = "1899-12-30")
@@
 
-```{r hierarchy, echo=TRUE}
-hierarchy_model <- tibble::tibble(units = c("index", "ball", "over", "inning", "match"), convert_fct = c(1, 6, 20, 2, 1))
-hierarchy_model
+```{r hierarchy, echo=FALSE}
+hierarchy_model <- tibble::tibble(units = c("index", "ball", "over", "inning", "match"), 
+                                  convert_fct = c(1, 6, 20, 2, 1))
+knitr::kable(hierarchy_model, caption = "T20 hierarchy table")
 ```
@@
 
-```{r read_data, echo=TRUE}
+```{r read_data, echo=FALSE}
 
@@
 
-cricket_tsibble %>% dynamic_create_gran(
-  "ball_over",
-  hierarchy_model
-)
-
-cricket_tsibble %>% granplot("ball_over", "inning_match",
+# cricket_tsibble %>% dynamic_create_gran(
+#   "ball_over",
+#   hierarchy_model
+# )
+
+#over_inning    ball_over 
+cricket_tsibble %>% granplot("over_inning", "ball_over",
   hierarchy_model,
   response = "total_runs",
-  plot_type = "boxplot"
-)
-
-cricket_tsibble %>% granplot("over_inning",
-  "inning_match",
+  plot_type = "violin"
+
+)
+
+#inning_match   ball_over
+
+cricket_tsibble %>% granplot("inning_match", "ball_over",
   hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile"
+  response = "total_runs"
+)
+
+# inning_match   ball_inning
+
+cricket_tsibble %>% granplot("inning_match", "ball_inning",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      over_inning
+
+cricket_tsibble %>% granplot("ball_over", "over_inning",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# inning_match   over_inning
+
+cricket_tsibble %>% granplot("inning_match", "over_inning",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      over_match
+
+cricket_tsibble %>% granplot("ball_over", "over_match",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      inning_match
+
+cricket_tsibble %>% granplot("ball_over", "over_match",
+  hierarchy_model,
+  response = "total_runs"
 )
@@
 cricket_tsibble %>%
-  dynamic_create_gran(
-    "over_inning",
-    hierarchy_model
-  ) %>%
-  dynamic_create_gran(
-    "inning_match",
-    hierarchy_model
-  ) %>%
   granplot(
@@
 
-cricket_tsibble %>%
-  dynamic_create_gran(
-    "ball_over",
-    hierarchy_model
-  ) %>%
-  dynamic_create_gran(
-    "over_match",
-    hierarchy_model
-  ) %>%
-  # as_tibble() %>%
-  # group_by(match_id, inning, batting_team, over_inning,inning_match) %>%
-  # summarise(total_over = sum(total_runs)) %>%
-  granplot(
-    gran1 = "ball_over",
-    gran2 = "over_match",
-    hierarchy_model,
-    response = "total_runs",
-    plot_type = "quantile"
-  )
 ```

### 2019-09-13 main.Rmd
trying out graphics relevant for analysis

@@
   response = "total_runs",
-  plot_type = "violin"
+  plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
 
@@
   hierarchy_model,
-  response = "total_runs"
+  response = "total_runs",
+   plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
 )
@@
 # inning_match   ball_inning 
-
+#interesting plot
 cricket_tsibble %>% granplot("inning_match", "ball_inning",
   hierarchy_model,
-  response = "total_runs"
-)
+  response = "total_runs",
+   plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
+  scale_x_discrete(breaks = seq(1,120,5))
 
@@
   hierarchy_model,
-  response = "total_runs"
+  response = "total_runs",
+  plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
 )
@@
   response = "total_runs"
-)
+) + scale_color_brewer(type = "div", palette = "Dark2")
 

### 2019-09-15 main.Rmd
section on single and multiple order granularity  updated example of  mayan included

@@
 <!-- What is multiple and single order up granularities -->
-The hierarchical structure of time creates a natural nested ordering, where hours are nested within days, days within weeks, weeks within months, and so on.
+The hierarchical structure of time creates a natural nested ordering.
+Hence, given a hierarchy we can define **single-order-up** and **multiple-order-up** granularities.
 We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are **single-order-up**.
+We take the help of Mayan calendar to illustrate this.
+In "Mayan" calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+
+ - 1 kin = 1 day
+ - 1 uinal = 20 kin
+ - 1 tun  = 18 uinal
+ - 1 katun = 20 tun
+ - 1 baktun = 20 katun
+
+Thus, the hierarchy table for the "Mayan" calendar would look like the following:
+
+| Units	      | Conversion factor
+|------------	|-------------------	|
+|  kin      	| 20             	    |
+|  uinal    	| 18            	    |
+|  tun      	| 20           	      |
+|  katun    	| 20            	    |
+|  baktun   	| 1          	        |
+
+Examples of **multiple-order-up**  granularities can be kin of the tun or kin of the baktun whereas examples of single-order-up granularities, which are nested within single levels may include kin of the uinal, uinal of the tun etc.
+
+In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up.
 
@@
 
-| Temporal units	| Conversion factor 	| conversion in days 	| single-order-up accessors 	|
-|------------	|-------------------	|--------------------	|---------------------------	|
-| 1 kin      	| 1 day             	|                    	| kin_uinal                 	|
-| 1 uinal    	| 20 kin            	| 20 days            	| uinal_tun                 	|
-| 1 tun      	| 18 uinal          	| 360 days           	| tun_katun                 	|
-| 1 katun    	| 20 tun            	| 7200 days          	| katun_baktun              	|
-| 1 baktun   	| 20 katun          	| 144,000 days       	| 1                         	|
+| Units	      | Conversion factor
+|------------	|-------------------	|
+|  index      | 1                   |
+|  kin      	| 20             	    |
+|  uinal    	| 18            	    |
+|  tun      	| 20           	      |
+|  katun    	| 20            	    |
+|  baktun   	| 1          	        |
 
 Single-order-up granularities like kin_uinal(kin of the uinal), uinal_tun(uinal of the tun) can be computed using modular arithmatic in this periodic set up.
+
+Suppose, the index of the tsibble is
 
@@
  - The second column represents the constant which relates subsequent linear granularities.
- - The third column provides the accessor function for single-order up circular granularities.
 
@@
 sm_cust50 %>%
-  create_gran("day_week") %>%
+  dynamic_create_gran("day_week") %>%
   filter(day_week %in% c("Sat", "Sun")) %>%
@@
   scale_y_continuous(breaks = seq(0, 1.5, 0.25))
+
+sm_cust50 %>%
+  granplot("fortnight_month", "hour_day",
+           response = "general_supply_kwh",
+           plot_type = "lv") + 
+  ggtitle("")
 ```
@@
   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-)
 
@@
 
-cricket_tsibble %>% granplot("ball_over", "over_match",
+cricket_tsibble %>% granplot("ball_over", "inning_match",
   hierarchy_model,
@@
 
-The scores/ball increases from 1st to 120th ball in the 90th percentile.
-In the last 2 overs, players are more vulnerable to get more scores, which is evident from the fact that in 90 percent of cases, their scores vary between 4 and 6.
-Till around 50th ball, players play safe and may get no runs per ball in 25% of the times, after which they more likely to get one score per ball.
+To see more variation in the response variable, we can also see the distribution of runs per over.
 
-<!-- # ```{r over, echo=TRUE} -->
-<!-- # data_MI %>% ggplot(aes(x = as.factor(over), y= total_runs)) + ggtitle("Letter value plot of runs per ball across each over of a match") + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1, k = 5) + xlab("overs per match") -->
+```{r runs_per-over}
+cricket_per_over <- cricketdata %>% 
+  group_by(match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(runs_per_over = sum(total_runs))
 
-<!-- # data_MI %>% ggplot(aes(y =as.factor(over), x= total_runs)) + geom_density_ridges2() + ylab("Balls per  over") -->
-<!-- ``` -->
+cricket_tsibble <- cricket_per_over %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
 
-<!-- In the first two overs, players play really safe and make no runs 50% of the times.
-The median is constant at 1 for all overs after the 2nd over.
-It is interesting to see the distribution of tails in this plot.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Till the first four overs, scoring a 4 or a 6 is extreme.
-The region which covers the tail 3.13% falls between 4 and 6 till up to 13th over.
-It is interesting to see how the letter values of scores (D, E and F) moves up, as we move towards the end of the game.
--->
+  cricket_tsibble %>% granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
 
-<!-- ```{r quarter, echo=TRUE} -->
-<!-- data_MI %>% ggplot(aes(x = as.factor(quarter_match), y= total_runs)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per ball across each quarter of a match") -->
+  )
 
-<!-- # data_MI %>% ggplot(aes(y =as.factor(quarter), x= total_runs)) + geom_density_ridges2() + ylab("quarters per match") + xlab(" Total runs per ball") -->
-<!-- ``` -->
+  cricket_tsibble %>%
+    filter(batting_team == "Chennai Super Kings") %>% 
+    granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+  )
 
-<!-- 4th quarter of the match witnesses low density(mass) at scores 0, 1 and more at 4/6. The density at score 6 increases with the quarters of the match. -->
+    cricket_tsibble %>%
+    filter(batting_team == "Kolkata Knight Riders") %>% 
+    granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+  )
 
-<!-- ```{r semester, echo=TRUE} -->
-<!-- # data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match") -->
-
-<!-- data_MI %>% ggplot(aes(x = total_runs, y= as.factor(semester_match))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per ball across each semesters of a match") -->
-
-<!-- # data_MI %>% ggplot(aes(y =as.factor(semester), x= total_runs, fill=factor(..quantile..))) +   stat_density_ridges( -->
-<!-- #     geom = "density_ridges_gradient", calc_ecdf = TRUE, -->
-<!-- #     quantiles = 4, quantile_lines = TRUE -->
-<!-- #   )  + ylab("semester") + xlab(" Total runs per ball") + -->
-<!-- #   scale_fill_viridis(discrete = TRUE, name = "Quartiles") -->
-<!-- ``` -->
-
-<!-- In the second semester of the game, there is more probability to score more than zero as can be seen from the drop in peaks of the distribution at 0 and increased peak at 1, 2, 4 and 6. -->
-
-<!-- Now, it will be interesting to see the distribution of runs per ball across two categorisation together: -->
-
-<!-- ```{r bivariate1, echo=TRUE} -->
-<!-- # data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match") -->
-
-<!-- data_MI %>% ggplot(aes(x = as.factor(over_quarter), y = total_runs))+  geom_violin()+ -->
-<!--                      facet_wrap(~quarter_semester) + xlab("overs per quarter")+ ylab(" Total runs per ball") + ggtitle("Boxplot of runs per ball across each over_quarter and quarter_semester")  -->
-<!-- ``` -->
-
-<!-- # runs per over - as variation would be more -->
-
-<!-- ```{r checking} -->
-<!-- data_MI_over <- data_MI %>% group_by(match_id, over) %>% summarise(runs_over = sum(total_runs)) %>%  -->
-<!--     mutate(quarter_match = case_when( -->
-
-<!--     between(over,1,5) ~ 1, -->
-<!--     between(over,6,10) ~ 2, -->
-<!--     between(over,11,15) ~ 3, -->
-<!--     between(over,16,20) ~ 4 -->
-<!--   )) %>%  -->
-<!--    mutate(semester_match = case_when( -->
-
-<!--     between(quarter_match,1,2) ~ 1, -->
-<!--     between(quarter_match,3,4) ~ 2 -->
-<!--   ))  %>%  -->
-<!--    mutate(over_quarter =  -->
-<!--   if_else(over%%5==0,5, over%%5) -->
-<!--    ) %>%  mutate(quarter_semester =  -->
-<!--   if_else(quarter_match%%2==0,2, quarter_match%%2) -->
-<!--    ) %>%  -->
-<!--   rename(over_match = over) -->
-
-<!-- ``` -->
-
-<!-- ```{r over_match, echo=TRUE} -->
-<!-- data_MI_over %>% ggplot(aes(x = as.factor(over_match), y= runs_over)) + ggtitle("Letter value plot of runs per over across each over of a match") + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1, k = 5) + xlab("overs per match") -->
-
-<!-- # data_MI %>% ggplot(aes(y =as.factor(over), x= total_runs)) + geom_density_ridges2() + ylab("Balls per  over") -->
-<!-- ``` -->
-
-<!-- ```{r over_quarter, echo=TRUE} -->
-<!-- data_MI_over %>% ggplot(aes(x = as.factor(over_quarter), y= runs_over)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per over across each quarter of a match") -->
-
-<!-- data_MI_over %>% ggplot(aes(x = as.factor(over_quarter), y= runs_over)) +  geom_violin()+ xlab("quarters per match")+ ylab(" Total runs per ball") + ggtitle("Violin plot of runs per over across each quarter of a match") + facet_grid(~quarter_match) -->
-
-<!-- # data_MI %>% ggplot(aes(y =as.factor(quarter), x= total_runs)) + geom_density_ridges2() + ylab("quarters per match") + xlab(" Total runs per ball") -->
-<!-- ``` -->
-
-<!-- ```{r quarter_match, echo=TRUE} -->
-<!-- # data_MI %>% ggplot(aes(x = as.factor(semester_match), y= total_runs)) +  geom_violin()+ xlab("semester per match")+ ylab(" Total runs per ball") +ggtitle("Violin plot of runs per ball across each semesters of a match") -->
-
-<!-- data_MI_over %>% ggplot(aes(x = runs_over, y= as.factor(quarter_match))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per overs across each semesters of a match") -->
-
-<!-- data_MI_over %>% ggplot(aes(x = runs_over, y= as.factor(over_quarter))) +  geom_density_ridges2()+ ylab("semesters per match")+ xlab(" Total runs per ball")+ ggtitle("Ridge plot of runs per overs across each semesters of a match") + facet_wrap(~quarter_match) -->
-
-<!-- # data_MI %>% ggplot(aes(y =as.factor(semester), x= total_runs, fill=factor(..quantile..))) +   stat_density_ridges( -->
-<!-- #     geom = "density_ridges_gradient", calc_ecdf = TRUE, -->
-<!-- #     quantiles = 4, quantile_lines = TRUE -->
-<!-- #   )  + ylab("semester") + xlab(" Total runs per ball") + -->
-<!-- #   scale_fill_viridis(discrete = TRUE, name = "Quartiles") -->
-<!-- ``` -->
+```
 

### 2019-09-15 main.Rmd
intro for single and multiple order up

@@
 
+<!-- some discussions on primary units of calendar -->
+So far we have used the Gregorian calendar as it is the most widely used calendar.
+But it is far from being the only one.
+All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars.
+Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
+Any civil day is divided into 24 hours and each hour into 60 minutes and each minute into 60 seconds.
+There are exceptions where in London, for example, length of each "hour" varies from about 39 minutes in December to 83 minutes in June.
+The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
+Nevertheless, for any calendar, the hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
+We shall use the notion of a hierarchy table to define single-order-up and multiple-order-up granularities.
+
+Consider a **hierarchy table** to be consisting of two columns:  
+ - The first column represents the (temporal) units in ascending order of hierarchy.  
+ - The second column represents the constant which relates subsequent temporal(units).
+
+and, **Order**, is defined as the position of the linear granularities in the hierarchical table.
+
+We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concering a single level as **single-order-up** granularities.
+
 <!-- What is multiple and single order up granularities -->
-The hierarchical structure of time creates a natural nested ordering.
-Hence, given a hierarchy we can define **single-order-up** and **multiple-order-up** granularities.
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities.
-We take the help of Mayan calendar to illustrate this.
-In "Mayan" calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+ We take the help of Mayan calendar to illustrate this. In "Mayan" calendar, one day was referred to as 1 kin and the calendar was structured as follows:
 
@@
 
-## Single-order-up granularities
+## Computation of single-order-up granularities
 
-<!-- some discussions on primary units of calendar -->
-The day is the basic unit of time underlying all calendars.
-The most common  way of defining days are from sunrise to sunrise (Hindu calendar, for example) or from sunset to sunset (Islamic and Hebrew calendars).
-Further, various calendars use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-Any civil day is divided into 24 hours and each hour into 60 minutes and each minute into 60 seconds.
-There are exceptions where Greeks and Romans separate the "day" and "night" into 12 hours each.
-In London, for example, length of each "hour" varies from about 39 minutes in December to 83 minutes in June.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
+Suppose, z is the index set of the tsibble, x,y are two units in the hierarchy table  with  $order(x) <  order(y)$.
+Let $f_{(x, y)}$ denotes the accessor function for computing the single-order-up granularity x-of-y and $c(x, y)$ is a constant which relates x and y.
 
-Let us define the arrangement of these coarser temporal units from finer temporal units in an hierarchy table.
-For example, in "Mayan" calendar, one day is referred to as 1 kin and they believed that the universe is destroyed at the start of every cycle (2,880,000 days).
-The Mayan calendar was structured as follows:
+<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
 
- - 1 kin = 1 day
- - 1 uinal = 20 kin
- - 1 tun  = 18 uinal
- - 1 katun = 20 tun
- - 1 baktun = 20 katun
+Then $f_{(x, y)}$ can be  computed using modular arithmetic as follows:
 
-Thus, the hierarchy table for the "Mayan" calendar would look like the following:
+$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$ where $y = x+1$
 
-| Units	      | Conversion factor
-|------------	|-------------------	|
-|  index      | 1                   |
-|  kin      	| 20             	    |
-|  uinal    	| 18            	    |
-|  tun      	| 20           	      |
-|  katun    	| 20            	    |
-|  baktun   	| 1          	        |
+See table[] for an illustration of computing single order up granularities for Mayan calendar.
 
-Single-order-up granularities like kin_uinal(kin of the uinal), uinal_tun(uinal of the tun) can be computed using modular arithmatic in this periodic set up.
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lr@{~}lr@{~}r}
+\toprule
 
-Suppose, the index of the tsibble is
+$kin-uinal:   = z \mod 20$  \\
+$uinal-tun:   = \lfloor z/20\rfloor \mod 18 $\\
+$tun-katun:    = \lfloor z/20*18\rfloor \mod 20 $\\
+$katun-baktun:   = \lfloor z/20*18*20\rfloor \mod 20 $\\
+
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Illustrative aperiodic circular granularities with time index in hours}
+\label{definitions}
+\end{table}
 
@@
 
-<!-- The index of a tsibble can be any date-time object which comes in different formats.
-[@Grolemund2011-qf] goes a long way in facilitating analysis of date-time data and extracting each date-time element with its own accessor function.
-However, table shows the accessor functions mostly allows for the creation of single-order-up granularities.
--->
-
-<!-- ```{r table_accessor} -->
-<!-- kable(tibble::tibble( -->
-<!--   `Categorization` = c("Second-of-minute", "Minute-of-hour", "Hour-of-day", "Day-of-week", "Day-of-month", "Week-of-year", "Day-of-year","Month-of-year"), -->
-<!--  Type = c("Periodic", "Periodic", "Periodic", "Periodic", "Aperiodic", "Aperiodic", "Aperiodic", "Periodic"), -->
-<!-- Accessor = c("second()", "minute()", "hour()","wday()", "mday()","week()", "yday()", "month()")), format = "markdown") -->
-<!-- ``` -->
-
-The hierarchical nature of time helps to create a framework where the single-order-up granularities can be used recursively to create multiple-order-up granularities.The idea is to incorporate and access all possible calendar categorizations to account for different periodicities in the data.
-The computation of multiple-order-up granularities from single-order-up granularities can differ basis the single-order-up granularities are circular or aperiodic.
-
 ### Periodic single-order-up granularities
-
-**Definition**: A **hierarchy table** has three columns:  
- - The first column represents the linear granularities in ascending order of temporal hierarchy.  
- - The second column represents the constant which relates subsequent linear granularities.
-
-**Definition**: **Order** is defined as the position of the linear granularities in the hierarchical table. 
-<!-- For example, if two linear granularties have three intermediate granularities in the hierarchical table, then their order difference is 3.  -->
-
-<!-- , "month", "quarter","semester", "year"), -->
 
@@
 \begin{split}
-f_{(x,y)}(z) & = f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, y)}(z)-1) \\  
-  & =  f_{(x, x+1)}(z) + c(x, x+1)[f_{(x+1, x+2)}(z) + c(x+1, x+2)(f_{(x+2, y)}(z) - 1) - 1]\\
-  & =  f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, x+2)}(z) - 1) + c(x, x+1)c(x+1, x+2)(f_{(x+2, y)}(z) - 1)\\
-  & =  f_{(x, x+1)}(z) + c(x, x+1)(f_{(x+1, x+2)}(z) - 1) + c(x, x+2)(f_{(x+2, y)}(z) - 1)\\
+f(x,y) & = f(x, x+1) + c(x, x+1)(f(x+1, y)-1) \\  
+  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
   &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f_{(x
-  +i, x+i+1)}(z) - 1)\\
+  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
+  +i, x+i+1) - 1)\\
 \end{split}

### 2019-09-15 main.Rmd
minor changes

@@
 Nevertheless, for any calendar, the hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
-We shall use the notion of a hierarchy table to define single-order-up and multiple-order-up granularities.
+We shall use the notion of a hierarchy table to define them.
 
@@
  - The first column represents the (temporal) units in ascending order of hierarchy.  
- - The second column represents the constant which relates subsequent temporal(units).
+ - The second column represents the constant which relates subsequent (temporal) units.
 
-and, **Order**, is defined as the position of the linear granularities in the hierarchical table.
+and, **order**, is defined as the position of the linear granularities in the hierarchical table.
 
 We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concering a single level as **single-order-up** granularities.
-
-<!-- What is multiple and single order up granularities -->
- We take the help of Mayan calendar to illustrate this. In "Mayan" calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+We take the help of Mayan calendar to illustrate this.
+In "Mayan" calendar, one day was referred to as 1 kin and the calendar was structured as follows:
 
@@
 
-Thus, the hierarchy table for the "Mayan" calendar would look like the following:
+Thus, the hierarchy table for the Mayan calendar would look like the following:
 
@@
 
-Examples of **multiple-order-up**  granularities can be kin of the tun or kin of the baktun whereas examples of single-order-up granularities, which are nested within single levels may include kin of the uinal, uinal of the tun etc.
+Examples of multiple-order-up  granularities can be kin of the tun or kin of the baktun whereas examples of single-order-up granularities may include kin of the uinal, uinal of the tun etc.
 

### 2019-09-15 main.Rmd
multiple order up example changed

@@
 
-and, **order**, is defined as the position of the linear granularities in the hierarchical table.
+and, **order**, is defined as the position of the units in the hierarchical table.
 
 We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concering a single level as **single-order-up** granularities.
-We take the help of Mayan calendar to illustrate this.
-In "Mayan" calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+"Mayan" calendar is used to illustrate this.
+In Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
 
@@
 
-Suppose, z is the index set of the tsibble, x,y are two units in the hierarchy table  with  $order(x) <  order(y)$.
-Let $f_{(x, y)}$ denotes the accessor function for computing the single-order-up granularity x-of-y and $c(x, y)$ is a constant which relates x and y.
+Suppose, z denotes the index of the tsibble, x,y are two units in the hierarchy table  with  $order(x) <  order(y)$.
+Let $f(x, y)$ denotes the accessor function for computing the single-order-up granularity x-of-y and $c(x, y)$ is a constant which relates x and y.
 
@@
 
-Then $f_{(x, y)}$ can be  computed using modular arithmetic as follows:
+Then $f(x, y)$ can be  computed using modular arithmetic as follows:
 
@@
 \toprule
+\textbf{Single-order-up granularities}
 
-$kin-uinal:   = z \mod 20$  \\
-$uinal-tun:   = \lfloor z/20\rfloor \mod 18 $\\
-$tun-katun:    = \lfloor z/20*18\rfloor \mod 20 $\\
-$katun-baktun:   = \lfloor z/20*18*20\rfloor \mod 20 $\\
+$kin\_uinal:   = z \mod 20$  \\
+$uinal\_tun:   = \lfloor z/20\rfloor \mod 18 $\\
+$tun\_katun:    = \lfloor z/20*18\rfloor \mod 20 $\\
+$katun\_baktun:   = \lfloor z/20*18*20\rfloor \mod 20 $\\
 
@@
 \end{center}
-\caption{Illustrative aperiodic circular granularities with time index in hours}
-\label{definitions}
+\caption{Illustrative single-order-up granularities for Mayan calendar with kin as the index}
+\label{single_order}
 \end{table}
+
+### Aperiodic single-order-up granularities
+
+Aperiodic single-order-up granularities refers to the granularities which are formed with units that does not repeat its values in regular intervals or periods.
+In Gregorian calendar, examples may include days of the month where each month may consists of 28, 29, 30 or 31 days.
+So there is no single number that can be used for converting days to months.
+We cannot compute the aperiodic single-order-up granularities using the index of the tsibble and modular arithmetic like in the periodic case.
 
@@
 
-### Periodic single-order-up granularities
+Computation of multiple-order-up granularities will differ if the units in the hierarchy table are periodic or aperiodic.
+We split the method of computation into two cases - one concerining all periodic single-order-up granularities and the second with mixed single-order-up granularities.
 
-Suppose, z is the index set of the tsibble, x,y are two linear granularities with  $order(x) <  order(y)$.
-Also, $f_{(x, y)}$ denotes the accessor function for computing circular granularity "x_y" and $c(x, y)$ is a constant which relates x and y.
+### All single-order-up granularities are periodic
+
+z is the index set of the tsibble, x,y are two linear granularities with  $order(x) <  order(y)$.
+Also, $f(x, y)$ denotes the accessor function for computing circular granularity x_y and $c(x, y)$ is a constant which relates x and y.
 It is easy to see that for $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
@@
 
-For illustration, let us consider the following hierarchy table.
-
-```{r table_accessor1}
-kable(tibble::tibble(
-  `linear granularities` = c("second", "minute", "hhour", "hour", "day", "week"),
-  `Conversion factor` = c(60, 30, 2, 24, 7, 1),
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_week", 1)
-), format = "markdown")
-```
+Let us use the equation to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.
 
@@
 \begin{split}
-f_{(hhour, week)}(z) & = f_{(hhour, hour)}(z) + c(hhour, hour)f_{(hour,day)}(z) + c(hhour, day)f_{(day, week)}(z) \\
-              & = {hhour\_hour}(z) + 2{hour\_day}(z) + 2*24{day\_week}(z) \\
+f(uinal, baktun) & = f(uinal, tun) + c(uinal, tun)f(tun,katun) + c(uinal, katun)f(katun, baktun) \\
+              & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\
 \end{split}
 \end{equation}
+             (using equations \ref{single_order})
 
@@
 
-<!-- This is valid in a scenario when the relationship between x and (x+1) is periodic (Define periodic in Definition section).
-However, when x and x+1 is not periodic, then some modification is required in Equation(1).
+Let us turn to Gregorian calendar for addressing this case.
+
+Suppose we have a hierarchy table for  Gregorian calendar as \ref{table_accessor2}.
+Since months consists of unequal number of days, any temporal unit which is higher in order than months will also have unequal number of days.
+This is an example of a hierarchy table which has both periodic and aperiodic single-order-up granularities.
+The single-order-up granularity day_month is aperiodic.
+Any single-order-up granularities which are formed by units below days are periodic.
+Similarly, all single-order-up granularities which are formed using units whose orders are higher than months are also periodic.
+
+<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
+However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
 -->
-
-Suppose we extend the hierarchy table to include the linear granularities month and quarter.
-Since months consists of unequal number of days, any linear granularity which is higher in order than months will also have unequal number of days.
-Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
-However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
 
@@
 kable(tibble::tibble(
-  `linear granularities` = c("second", "minute", "hhour", "hour", "day", "month", "quarter"),
-  `Conversion factor` = c(60, 30, 2, 24, "aperiodic", 3, 1),
-  `Single-order-up accessors` = c("second_minute", "minute_hhour", "hhour_hour", "hour_day", "day_month", "month_quarter", 1)
-), format = "markdown")
+  `linear granularities` = c("minute", "hour", "day", "month", "quarter", "year"),
+  `Conversion factor` = c(60,24, "aperiodic", 3, 4, 1), format = "markdown")
 ```
@@
 There can be three scenarios for obtaining calendar categorization here:
-- calendar categorization consisting of two linear granularities whose orders are less than day  
-- calendar categorization consisting of two linear granularities whose orders are more than month  
-- calendar categorization consisting of one linear granularity with order at most day and another with order at least month
+- granularities consisting of two units whose orders are less than day  
+- granularities consisting of two units whose orders are more than month  
+- granularities consisting of one unit with order at most day and another with order at least month
 

### 2019-09-16 main.Rmd
playing with cricekt example

@@
 
-z is the index set of the tsibble, x,y are two linear granularities with  $order(x) <  order(y)$.
+z is the index set of the tsibble and x, y are two linear granularities with  $order(x) <  order(y)$.
 Also, $f(x, y)$ denotes the accessor function for computing circular granularity x_y and $c(x, y)$ is a constant which relates x and y.
@@
   `linear granularities` = c("minute", "hour", "day", "month", "quarter", "year"),
-  `Conversion factor` = c(60,24, "aperiodic", 3, 4, 1), format = "markdown")
+  `Conversion factor` = c(60,24, "aperiodic", 3, 4, 1), format = "markdown"))
 ```
@@
 
+Only for 2008
 ```{r read_data, echo=FALSE}
@@
 
+```{r read_data_all_seas, echo=FALSE}
+
+load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda")
+
+cricket_tsibble_all <- cricket_all_seas %>%
+  filter(season %in% c(2012:2015)) %>% 
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+
+# cricket_tsibble %>% dynamic_create_gran(
+#   "ball_over",
+#   hierarchy_model
+# )
+
+#over_inning    ball_over 
+cricket_tsibble_all %>% granplot("over_inning", "ball_over",
+  hierarchy_model,
+  response = "total_runs",
+  plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+)
+
+#inning_match   ball_over
+
+cricket_tsibble_all %>% granplot("inning_match", "ball_over",
+  hierarchy_model,
+  response = "total_runs",
+   plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+)
+
+# inning_match   ball_inning 
+#interesting plot
+cricket_tsibble_all %>% granplot("inning_match", "ball_inning",
+  hierarchy_model,
+  response = "total_runs",
+   plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
+  scale_x_discrete(breaks = seq(1,120,5))
+
+# ball_over      over_inning
+
+cricket_tsibble_all %>% granplot("ball_over", "over_inning",
+  hierarchy_model,
+  response = "total_runs",
+  plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
+
+# inning_match   over_inning
+
+cricket_tsibble_all %>% granplot("inning_match", "over_inning",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      over_match
+
+cricket_tsibble_all %>% granplot("ball_over", "over_match",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      inning_match
+
+cricket_tsibble_all %>% granplot("ball_over", "inning_match",
+  hierarchy_model,
+  response = "total_runs"
+) + scale_color_brewer(type = "div", palette = "Dark2")
+
+cricket_tsibble_all %>%
+  granplot(
+    gran1 = "inning_match",
+    gran2 = "over_inning",
+    hierarchy_model,
+    response = "total_runs",
+    plot_type = "quantile"
+  )
+```
+
 To see more variation in the response variable, we can also see the distribution of runs per over.
@@
 
+```{r runs_per-over}
+
+# all season per over runs
+cricket_per_over_all <- cricket_all_seas %>% 
+  group_by(season, match_id, batting_team,inning, over) %>%
+  summarise(runs_per_over = sum(total_runs))
+
+# all season per over runs tsibble
+cricket_tsibble_all <- cricket_per_over_all %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
+
+  cricket_tsibble_all %>% granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+  )
+
+  cricket_tsibble_all %>% granplot("inning_match", "season",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+  )
+
+  cricket_tsibble_all %>% granplot("season", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+  )
+
+  cricket_tsibble %>%
+    filter(batting_team == "Chennai Super Kings") %>% 
+    granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+  )
+
+    cricket_tsibble %>%
+    filter(batting_team == "Kolkata Knight Riders") %>% 
+    granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+  )
+
+```
+
 # Discussion {#sec:discussion}

### 2019-09-16 main.Rmd
just keeping relevant information on cricket runs per over

@@
 Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls.
-We have sourced the the ball by ball data for IPL 2008 from https://www.kaggle.com/littleraj30/indian-premier-league-2019-ball-by-ball.
+We have sourced the the ball by ball data for IPL season 2008 to 2016 from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
+
 There are many interesting questions that can possibly be answered with such a dataset, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
@@
 
-a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions?
-
-b) How the scores vary per each over each periodic granularities like ball of the over, ball of the innings, over of the inning, over of the match, inning of the match and others.
-
-We will look at the ball by ball data for all batting teams.Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
-
-Only for 2008
+<!-- a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions? -->
+
+How the scores per over vary across granularities/categorisations like innings of a match or matches of a season?
+Are these different for the winning teams in that season and the teams that couldn;t qualify for the playoffs?
+
+We will look at the ball by ball data for all batting teams.
+Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
+
 ```{r read_data, echo=FALSE}
@@
 
-  )
+  )  +
 

### 2019-09-16 main.Rmd
just keeping relevant information on cricket runs per over

@@
 Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
-
-```{r read_data, echo=FALSE}
-
-cricket_tsibble <- cricketdata %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-# cricket_tsibble %>% dynamic_create_gran(
-#   "ball_over",
-#   hierarchy_model
-# )
-
-#over_inning    ball_over 
-cricket_tsibble %>% granplot("over_inning", "ball_over",
-  hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-)
-
-#inning_match   ball_over
-
-cricket_tsibble %>% granplot("inning_match", "ball_over",
-  hierarchy_model,
-  response = "total_runs",
-   plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-)
-
-# inning_match   ball_inning 
-#interesting plot
-cricket_tsibble %>% granplot("inning_match", "ball_inning",
-  hierarchy_model,
-  response = "total_runs",
-   plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
-  scale_x_discrete(breaks = seq(1,120,5))
-
-# ball_over      over_inning
-
-cricket_tsibble %>% granplot("ball_over", "over_inning",
-  hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-
-# inning_match   over_inning
-
-cricket_tsibble %>% granplot("inning_match", "over_inning",
-  hierarchy_model,
-  response = "total_runs"
-)
-
-# ball_over      over_match
-
-cricket_tsibble %>% granplot("ball_over", "over_match",
-  hierarchy_model,
-  response = "total_runs"
-)
-
-# ball_over      inning_match
-
-cricket_tsibble %>% granplot("ball_over", "inning_match",
-  hierarchy_model,
-  response = "total_runs"
-) + scale_color_brewer(type = "div", palette = "Dark2")
-
-cricket_tsibble %>%
-  granplot(
-    gran1 = "inning_match",
-    gran2 = "over_inning",
-    hierarchy_model,
-    response = "total_runs",
-    plot_type = "quantile"
-  )
-
-```
-
-```{r read_data_all_seas, echo=FALSE}
-
-load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda")
-
-cricket_tsibble_all <- cricket_all_seas %>%
-  filter(season %in% c(2012:2015)) %>% 
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-# cricket_tsibble %>% dynamic_create_gran(
-#   "ball_over",
-#   hierarchy_model
-# )
-
-#over_inning    ball_over 
-cricket_tsibble_all %>% granplot("over_inning", "ball_over",
-  hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-)
-
-#inning_match   ball_over
-
-cricket_tsibble_all %>% granplot("inning_match", "ball_over",
-  hierarchy_model,
-  response = "total_runs",
-   plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-)
-
-# inning_match   ball_inning 
-#interesting plot
-cricket_tsibble_all %>% granplot("inning_match", "ball_inning",
-  hierarchy_model,
-  response = "total_runs",
-   plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
-  scale_x_discrete(breaks = seq(1,120,5))
-
-# ball_over      over_inning
-
-cricket_tsibble_all %>% granplot("ball_over", "over_inning",
-  hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-
-# inning_match   over_inning
-
-cricket_tsibble_all %>% granplot("inning_match", "over_inning",
-  hierarchy_model,
-  response = "total_runs"
-)
-
-# ball_over      over_match
-
-cricket_tsibble_all %>% granplot("ball_over", "over_match",
-  hierarchy_model,
-  response = "total_runs"
-)
-
-# ball_over      inning_match
-
-cricket_tsibble_all %>% granplot("ball_over", "inning_match",
-  hierarchy_model,
-  response = "total_runs"
-) + scale_color_brewer(type = "div", palette = "Dark2")
-
-cricket_tsibble_all %>%
-  granplot(
-    gran1 = "inning_match",
-    gran2 = "over_inning",
-    hierarchy_model,
-    response = "total_runs",
-    plot_type = "quantile"
-  )
-```
-
-To see more variation in the response variable, we can also see the distribution of runs per over.
+Also, we look at runs per over as that would have more variability compared to runs per ball and it be easier to observe the strategies of the winning team through that.
 
@@
   as_tsibble(index = data_index)
+```
+
+First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas_over_inning}.
+The distribution of runs per over has not significantly changed from 2008 to 2016.
+There is no clear trend that we can observe for any over of the innings for any seasons.
+So we can work with any season and
+
+```{r seas_over_inning}
+ cricket_tsibble_all %>% granplot("over_inning", "season",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+  )  + 
+    theme( # remove the vertical grid lines
+           panel.grid.major.x = element_blank() ,
+           # explicitly set the horizontal lines (or they will disappear too)
+           panel.grid.major.y = element_line( size=.1, color="black" ) 
+    )
+
+```
 
@@
                                plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
+    theme( # remove the vertical grid lines
+           panel.grid.major.x = element_blank() ,
+           # explicitly set the horizontal lines (or they will disappear too)
+           panel.grid.major.y = element_line( size=.1, color="black" ) 
     )
@@
 
+<!-- ```{r read_data, echo=FALSE} -->
+
+<!-- cricket_tsibble <- cricketdata %>% -->
+<!--   mutate(data_index = row_number()) %>% -->
+<!--   as_tsibble(index = data_index) -->
+
+<!-- # cricket_tsibble %>% dynamic_create_gran( -->
+<!-- #   "ball_over", -->
+<!-- #   hierarchy_model -->
+<!-- # ) -->
+
+<!-- #over_inning    ball_over -->
+<!-- cricket_tsibble %>% granplot("over_inning", "ball_over", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--   plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!-- ) -->
+
+<!-- #inning_match   ball_over -->
+
+<!-- cricket_tsibble %>% granplot("inning_match", "ball_over", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--    plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+<!-- ) -->
+
+<!-- # inning_match   ball_inning -->
+<!-- #interesting plot -->
+<!-- cricket_tsibble %>% granplot("inning_match", "ball_inning", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--    plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
+<!--   scale_x_discrete(breaks = seq(1,120,5)) -->
+
+<!-- # ball_over      over_inning -->
+
+<!-- cricket_tsibble %>% granplot("ball_over", "over_inning", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--   plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
+
+<!-- # inning_match   over_inning -->
+
+<!-- cricket_tsibble %>% granplot("inning_match", "over_inning", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs" -->
+<!-- ) -->
+
+<!-- # ball_over      over_match -->
+
+<!-- cricket_tsibble %>% granplot("ball_over", "over_match", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs" -->
+<!-- ) -->
+
+<!-- # ball_over      inning_match -->
+
+<!-- cricket_tsibble %>% granplot("ball_over", "inning_match", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs" -->
+<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->
+
+<!-- cricket_tsibble %>% -->
+<!--   granplot( -->
+<!--     gran1 = "inning_match", -->
+<!--     gran2 = "over_inning", -->
+<!--     hierarchy_model, -->
+<!--     response = "total_runs", -->
+<!--     plot_type = "quantile" -->
+<!--   ) -->
+
+<!-- ``` -->
+
+```{r read_data_all_seas, echo=FALSE}
+
+load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda")
+
+cricket_tsibble_all <- cricket_all_seas %>%
+  filter(season %in% c(2012:2015)) %>% 
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+
+# cricket_tsibble %>% dynamic_create_gran(
+#   "ball_over",
+#   hierarchy_model
+# )
+
+#over_inning    ball_over 
+cricket_tsibble_all %>% granplot("over_inning", "ball_over",
+  hierarchy_model,
+  response = "total_runs",
+  plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+)
+
+#inning_match   ball_over
+
+cricket_tsibble_all %>% granplot("inning_match", "ball_over",
+  hierarchy_model,
+  response = "total_runs",
+   plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+)
+
+# inning_match   ball_inning 
+#interesting plot
+cricket_tsibble_all %>% granplot("inning_match", "ball_inning",
+  hierarchy_model,
+  response = "total_runs",
+   plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
+  scale_x_discrete(breaks = seq(1,120,5))
+
+# ball_over      over_inning
+
+cricket_tsibble_all %>% granplot("ball_over", "over_inning",
+  hierarchy_model,
+  response = "total_runs",
+  plot_type = "quantile",
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
+
+# inning_match   over_inning
+
+cricket_tsibble_all %>% granplot("inning_match", "over_inning",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      over_match
+
+cricket_tsibble_all %>% granplot("ball_over", "over_match",
+  hierarchy_model,
+  response = "total_runs"
+)
+
+# ball_over      inning_match
+
+cricket_tsibble_all %>% granplot("ball_over", "inning_match",
+  hierarchy_model,
+  response = "total_runs"
+) + scale_color_brewer(type = "div", palette = "Dark2")
+
+cricket_tsibble_all %>%
+  granplot(
+    gran1 = "inning_match",
+    gran2 = "over_inning",
+    hierarchy_model,
+    response = "total_runs",
+    plot_type = "quantile"
+  )
+```
+
 # Discussion {#sec:discussion}

### 2019-09-23 main.Rmd
cricket examples updated

@@
           quantile_prob = c(0.25, 0.5, 0.75),
-           response = "general_supply_kwh") +
+           response = "general_supply_kwh", overlay = FALSE) +
   ggtitle("") +
@@
     response = "general_supply_kwh", plot_type = "quantile", 
-          quantile_prob = c(0.25, 0.5, 0.75)) + ggtitle("") +
+          quantile_prob = c(0.25, 0.5, 0.75), overlay = FALSE) + ggtitle("") +
   ggtitle("Quantile plot across day-of-week by month-of-year")
@@
 
+Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings?
+
 We will look at the ball by ball data for all batting teams.
@@
 ```{r runs_per-over}
-cricket_per_over <- cricketdata %>% 
-  group_by(match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs))
-
-cricket_tsibble <- cricket_per_over %>% 
+
+cricket <-  read_csv("data-raw/deliveries_all.csv")
+matches <- read_csv("data-raw/matches_all.csv")
+
+cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id"))
+
+cricket_per_over <- cricket_season %>% 
+  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))
+
+cricket_tsibble_all <- cricket_per_over %>% 
   ungroup() %>% 
@@
 There is no clear trend that we can observe for any over of the innings for any seasons.
-So we can work with any season and
+So we can work with any subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.
 
 ```{r seas_over_inning}
- cricket_tsibble_all %>% granplot("over_inning", "season",
+ cricket_tsibble_all %>% granplot("over", "season",
                                hierarchy_model,
@@
            # explicitly set the horizontal lines (or they will disappear too)
-           panel.grid.major.y = element_line( size=.1, color="black" ) 
-    )
-
-```
-
-  cricket_tsibble %>% granplot("inning_match", "over_inning",
+           panel.grid.major.y = element_blank() 
+    ) + scale_x_discrete(breaks = seq(2008, 2016, 3))
+
+```
+
+Mumbai Indians(MI) won against Chennai Super kings(CSK) in 2013 and 2015.
+So it would be interesting to see the difference in the strategies between these two teams tgroughout all matches in the two seasons.
+In \autoref{fig:2013_15_MI}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
+It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
+CSK has been more uniform in their approach maintaining consistent scores per over, whereas MI's scores increase as they proceed with the game nothwithstanding if they bat first or second.
+CSK's run rate increase as it reached the last over when it bats in the 2nd innings, but doesn't increase much when it bats first.
+
+<!-- ```{r 2013_15_MI} -->
+<!--  cricket_tsibble_all %>%  -->
+<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
+<!--            granplot("inning", "over", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "run_rate", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
+<!--     theme( # remove the vertical grid lines -->
+<!--            panel.grid.major.x = element_blank() , -->
+<!--            # explicitly set the horizontal lines (or they will disappear too) -->
+<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
+<!--     ) -->
+<!--  cricket_tsibble_all %>%  -->
+<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
+<!--            granplot("inning", "over", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "run_rate", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
+<!--     theme( # remove the vertical grid lines -->
+<!--            panel.grid.major.x = element_blank() , -->
+<!--            # explicitly set the horizontal lines (or they will disappear too) -->
+<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
+<!--     ) -->
+
+<!-- ``` -->
+
+```{r lv_MI_CSK}
+# fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (a) Mumbai Indians and (b) Chennai Super Kings.", fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Letter value plot of run rate across overs of an inning and innings of the match for (a) Mumbai Indians and (b) Chennai Super Kings.", fig.show = 'hold'
+
+ cricket_tsibble_all %>% 
+  filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>% 
+           granplot("inning", "over",
                                hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
-    theme( # remove the vertical grid lines
-           panel.grid.major.x = element_blank() ,
-           # explicitly set the horizontal lines (or they will disappear too)
-           panel.grid.major.y = element_line( size=.1, color="black" ) 
-    )
-
-  cricket_tsibble %>%
-    filter(batting_team == "Chennai Super Kings") %>% 
-    granplot("inning_match", "over_inning",
+                               response = "run_rate",
+                               plot_type = "lv",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")
+
+ cricket_tsibble_all %>% 
+  filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>% 
+           granplot("inning", "over",
                                hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-  )
-
-    cricket_tsibble %>%
-    filter(batting_team == "Kolkata Knight Riders") %>% 
-    granplot("inning_match", "over_inning",
+                               response = "run_rate",
+                               plot_type = "lv",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
+
+```
+
+```{r MI_CSK_violin}
+
+matches_select <- matches %>% select(id, toss_winner, winner)
+
+cricket_winner <- cricket_tsibble_all %>% 
+  left_join(matches_select, by = c("match_id" = "id"))
+
+MI_winner <- cricket_winner %>% 
+             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians") %>% 
+             mutate(win = if_else(winner == "Mumbai Indians", 1, 0))
+
+CSK_winner <- cricket_winner %>% 
+             filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings") %>% 
+             mutate(win = if_else(winner == "Chennai Super Kings", 1, 0))
+
+ MI_winner %>% 
+            granplot( "win","inning",
                                 hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-  )
-
-```
-
-```{r runs_per-over}
-
-# all season per over runs
-cricket_per_over_all <- cricket_all_seas %>% 
-  group_by(season, match_id, batting_team,inning, over) %>%
-  summarise(runs_per_over = sum(total_runs))
-
-# all season per over runs tsibble
-cricket_tsibble_all <- cricket_per_over_all %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-  cricket_tsibble_all %>% granplot("inning_match", "over_inning",
+                                response = "run_rate",
+                                plot_type = "violin",
+                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
+                     overlay = FALSE) + ggtitle("") + geom_boxplot(width = 0.1)
+
+ CSK_winner %>% filter(inning %in% c(1,2)) %>% 
+           granplot( "win","inning",
                                hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-  )
-
-  cricket_tsibble_all %>% granplot("inning_match", "season",
-                               hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-  )
-
-  cricket_tsibble_all %>% granplot("season", "over_inning",
-                               hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-  )  +
-
-  cricket_tsibble %>%
-    filter(batting_team == "Chennai Super Kings") %>% 
-    granplot("inning_match", "over_inning",
-                               hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-  )
-
-    cricket_tsibble %>%
-    filter(batting_team == "Kolkata Knight Riders") %>% 
-    granplot("inning_match", "over_inning",
-                               hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-  )
+                               response = "run_rate",
+                               plot_type = "violin",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
+                    overlay = FALSE) + ggtitle("") + geom_boxplot(width = 0.1)
+
+```
+
+<!-- For MI, cases when they lost, distribution of run rates not very different depending on if they bat first or second.
+However, cases when they won, distribution of run rate is very skewed when they bat in the 2nd innings implying winning by short margins.
+-->
+
+For CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
+Run rate can go up to 6 runs per over depending on the situation.
+For MI, this is not true.
+When they won, their run rates in the 2nd innings went up to 6 and led to their win, but stayed within 4 whenever they lost.
+This implies CSK is more of a fighter than MI when it comes to defending in the 2nd innings.
+
+<!--   cricket_tsibble %>% granplot("inning_match", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + -->
+<!--     theme( # remove the vertical grid lines -->
+<!--            panel.grid.major.x = element_blank() , -->
+<!--            # explicitly set the horizontal lines (or they will disappear too) -->
+<!--            panel.grid.major.y = element_line( size=.1, color="black" )  -->
+<!--     ) -->
+
+<!--   cricket_tsibble %>% -->
+<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
+<!--     granplot("inning_match", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+<!--   ) -->
+
+<!--     cricket_tsibble %>% -->
+<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
+<!--     granplot("inning_match", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+<!--   ) -->
+
+<!-- ``` -->
+
+<!-- ```{r runs_per-over} -->
+
+<!-- # all season per over runs -->
+<!-- cricket_per_over_all <- cricket_all_seas %>%  -->
+<!--   group_by(season, match_id, batting_team,inning, over) %>% -->
+<!--   summarise(runs_per_over = sum(total_runs))  -->
+
+<!-- # all season per over runs tsibble -->
+<!-- cricket_tsibble_all <- cricket_per_over_all %>%  -->
+<!--   ungroup() %>%  -->
+<!--   mutate(data_index = row_number()) %>%  -->
+<!--   as_tsibble(index = data_index) -->
+
+<!--   cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   ) -->
+
+<!--   cricket_tsibble_all %>% granplot("inning_match", "season", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   )   -->
+
+<!--   cricket_tsibble_all %>% granplot("season", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   )  +  -->
+
+<!--   cricket_tsibble %>% -->
+<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
+<!--     granplot("inning_match", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+<!--   ) -->
+
+<!--     cricket_tsibble %>% -->
+<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
+<!--     granplot("inning_match", "over_inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "runs_per_over", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+<!--   ) -->
 
@@
 
-```{r read_data_all_seas, echo=FALSE}
-
-load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda")
-
-cricket_tsibble_all <- cricket_all_seas %>%
-  filter(season %in% c(2012:2015)) %>% 
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-# cricket_tsibble %>% dynamic_create_gran(
-#   "ball_over",
-#   hierarchy_model
-# )
-
-#over_inning    ball_over 
-cricket_tsibble_all %>% granplot("over_inning", "ball_over",
-  hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-
-)
-
-#inning_match   ball_over
-
-cricket_tsibble_all %>% granplot("inning_match", "ball_over",
-  hierarchy_model,
-  response = "total_runs",
-   plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-)
-
-# inning_match   ball_inning 
-#interesting plot
-cricket_tsibble_all %>% granplot("inning_match", "ball_inning",
-  hierarchy_model,
-  response = "total_runs",
-   plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
-  scale_x_discrete(breaks = seq(1,120,5))
-
-# ball_over      over_inning
-
-cricket_tsibble_all %>% granplot("ball_over", "over_inning",
-  hierarchy_model,
-  response = "total_runs",
-  plot_type = "quantile",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-
-# inning_match   over_inning
-
-cricket_tsibble_all %>% granplot("inning_match", "over_inning",
-  hierarchy_model,
-  response = "total_runs"
-)
-
-# ball_over      over_match
-
-cricket_tsibble_all %>% granplot("ball_over", "over_match",
-  hierarchy_model,
-  response = "total_runs"
-)
-
-# ball_over      inning_match
-
-cricket_tsibble_all %>% granplot("ball_over", "inning_match",
-  hierarchy_model,
-  response = "total_runs"
-) + scale_color_brewer(type = "div", palette = "Dark2")
-
-cricket_tsibble_all %>%
-  granplot(
-    gran1 = "inning_match",
-    gran2 = "over_inning",
-    hierarchy_model,
-    response = "total_runs",
-    plot_type = "quantile"
-  )
-```
+<!-- ```{r read_data_all_seas, echo=FALSE} -->
+
+<!-- load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda") -->
+
+<!-- cricket_tsibble_all <- cricket_all_seas %>% -->
+<!--   filter(season %in% c(2012:2015)) %>%  -->
+<!--   mutate(data_index = row_number()) %>% -->
+<!--   as_tsibble(index = data_index) -->
+
+<!-- # cricket_tsibble %>% dynamic_create_gran( -->
+<!-- #   "ball_over", -->
+<!-- #   hierarchy_model -->
+<!-- # ) -->
+
+<!-- #over_inning    ball_over  -->
+<!-- cricket_tsibble_all %>% granplot("over_inning", "ball_over", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--   plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!-- ) -->
+
+<!-- #inning_match   ball_over  -->
+
+<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_over", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--    plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+<!-- ) -->
+
+<!-- # inning_match   ball_inning  -->
+<!-- #interesting plot -->
+<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_inning", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--    plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
+<!--   scale_x_discrete(breaks = seq(1,120,5)) -->
+
+<!-- # ball_over      over_inning   -->
+
+<!-- cricket_tsibble_all %>% granplot("ball_over", "over_inning", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs", -->
+<!--   plot_type = "quantile", -->
+<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
+
+<!-- # inning_match   over_inning   -->
+
+<!-- cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs" -->
+<!-- ) -->
+
+<!-- # ball_over      over_match   -->
+
+<!-- cricket_tsibble_all %>% granplot("ball_over", "over_match", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs" -->
+<!-- ) -->
+
+<!-- # ball_over      inning_match   -->
+
+<!-- cricket_tsibble_all %>% granplot("ball_over", "inning_match", -->
+<!--   hierarchy_model, -->
+<!--   response = "total_runs" -->
+<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->
+
+<!-- cricket_tsibble_all %>% -->
+<!--   granplot( -->
+<!--     gran1 = "inning_match", -->
+<!--     gran2 = "over_inning", -->
+<!--     hierarchy_model, -->
+<!--     response = "total_runs", -->
+<!--     plot_type = "quantile" -->
+<!--   ) -->
+<!-- ``` -->
 

### 2019-09-23 cricket_per_over.R

@@ -1,0 +1,22 @@
+cricket_per_over <- cricketdata %>% 
+  group_by(match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(runs_per_over = sum(total_runs)) 
+
+cricket_tsibble <- cricket_per_over %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
+
+
+harmony(cricket_tsibble, lowest_unit = "over", highest_unit = "match", hierarchy_tbl =  hierarchy_model)
+cricket_per_over %>% 
+  
+  
+  cricket_tsibble %>% granplot("inning_match", "over_inning",
+                               hierarchy_model,
+                               response = "runs_per_over",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+                               
+                               
+  )

### 2019-09-23 main.Rmd
changing season to make analysis more interesting

@@
 
-# Case study: Analysis on smart meter data {#smartmeter}
+# Case studies
+
+## Smart meter data of Australia {#smartmeter}
 
@@
 
-# Case study: Analysis on cricket {#cricket}
+## T20 cricket data of Indian Premiere League {#cricket}
 
@@
 We have sourced the the ball by ball data for IPL season 2008 to 2016 from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-
-There are many interesting questions that can possibly be answered with such a dataset, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
-
-The dataset contains the information on match_id, inning, batting team, bowling team, over of the innings, balls of the over, total runs and many more.
-A hierarchy table like [] can be construed for this game format:
+The dataset contains the information on batting team, bowling team,  balls of the over,  over of the innings, innings of the match and total runs per ball for 577 matches spanning over 9 seasons (2008 to 2016).
+It also has information on which team won, winning margin and several other useful information.
+
+A hierarchy like table \ref{tab:hierarchy} can be construed for this game format, where index here would refer to ball as the data set contains ball-by-ball data.
 
@@
                                   convert_fct = c(1, 6, 20, 2, 1))
-knitr::kable(hierarchy_model, caption = "T20 hierarchy table")
-```
-
-Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
+knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket")
+```
+
+There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
+
+<!-- Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
+-->
 
@@
 How the scores per over vary across granularities/categorisations like innings of a match or matches of a season?
-Are these different for the winning teams in that season and the teams that couldn;t qualify for the playoffs?
+Are these different for the winning teams in that season and the teams that couldn't qualify for the playoffs?
 
@@
 
-First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas_over_inning}.
+First, we look at the distribution of runs per over across over of the innings and seasons in \ref{fig:seas-over-inning}.
 The distribution of runs per over has not significantly changed from 2008 to 2016.
-There is no clear trend that we can observe for any over of the innings for any seasons.
-So we can work with any subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.
-
-```{r seas_over_inning}
- cricket_tsibble_all %>% granplot("over", "season",
+There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
+Hence, we work with subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.
+
+```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons."}
+  cricket_tsibble_all %>% 
+  granplot("over", "season",
                                 hierarchy_model,
@@
                                 plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+                                quantile_prob = c(0.25, 0.5, 0.75), overlay = FALSE
 
@@
      ) + scale_x_discrete(breaks = seq(2008, 2016, 3))
-
-```
-
-Mumbai Indians(MI) won against Chennai Super kings(CSK) in 2013 and 2015.
-So it would be interesting to see the difference in the strategies between these two teams tgroughout all matches in the two seasons.
-In \autoref{fig:2013_15_MI}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
-It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
-CSK has been more uniform in their approach maintaining consistent scores per over, whereas MI's scores increase as they proceed with the game nothwithstanding if they bat first or second.
-CSK's run rate increase as it reached the last over when it bats in the 2nd innings, but doesn't increase much when it bats first.
-
-<!-- ```{r 2013_15_MI} -->
-<!--  cricket_tsibble_all %>%  -->
-<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
-<!--            granplot("inning", "over", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
-<!--     ) -->
-<!--  cricket_tsibble_all %>%  -->
-<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
-<!--            granplot("inning", "over", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
-<!--     ) -->
-
-<!-- ``` -->
-
-```{r lv_MI_CSK}
-# fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (a) Mumbai Indians and (b) Chennai Super Kings.", fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Letter value plot of run rate across overs of an inning and innings of the match for (a) Mumbai Indians and (b) Chennai Super Kings.", fig.show = 'hold'
-
- cricket_tsibble_all %>% 
-  filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>% 
-           granplot("inning", "over",
-                               hierarchy_model,
-                               response = "run_rate",
-                               plot_type = "lv",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")
-
- cricket_tsibble_all %>% 
-  filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>% 
-           granplot("inning", "over",
-                               hierarchy_model,
-                               response = "run_rate",
-                               plot_type = "lv",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
-
-```
-
-```{r MI_CSK_violin}
-
-matches_select <- matches %>% select(id, toss_winner, winner)
+```
+
+Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles.
+MI won against CSK both in 2013 and 2015 finals.
+It would be interesting to see the difference in their strategies throughout all matches in the two seasons.
+We can understand difference in strategy by answering these two questions.
+
+- Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first?
+
+- Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings?
+
+```{r violin, fig.cap= "Violin and box plots of run rate across innings of the match and matches won or lost for Mumbai Indians(top) and Chennai Super Kings(bottom).The violins for CSK has long right tails in 2nd innings for cases when they lost or won.
+MI has long right tail only when they won the match."}
+
+matches_select <- matches %>% select(id, toss_winner, winner, win_by_runs)
 
@@
 MI_winner <- cricket_winner %>% 
-             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians") %>% 
+             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians",
+                    season %in% c(2010:2015)) %>% 
              mutate(win = if_else(winner == "Mumbai Indians", 1, 0))
@@
 CSK_winner <- cricket_winner %>% 
-             filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings") %>% 
+             filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings",
+                    season %in% c(2010:2015)) %>% 
              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0))
@@
                                 quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-                     overlay = FALSE) + ggtitle("") + geom_boxplot(width = 0.1)
+                     overlay = FALSE) + ggtitle("") + 
+    geom_boxplot(width = 0.1, color = "blue")
 
@@
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-                    overlay = FALSE) + ggtitle("") + geom_boxplot(width = 0.1)
-
+                    overlay = FALSE) + ggtitle("") + 
+   geom_boxplot(width = 0.1, color = "blue")
+```
+
+```{r}
+cricket_winner %>% as_tibble() %>% 
+             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians",
+                    winner!= "Mumbai Indians",
+                    win_by_runs!=0) %>% select(win_by_runs) %>% summary()
+
+cricket_winner %>% as_tibble() %>% 
+             filter(batting_team ==  "Chennai Super Kings",
+                    winner!= "Chennai Super Kings",
+                    win_by_runs!=0,
+                    inning ==2) %>% select(win_by_runs) %>% summary()
 ```
@@
 
-For CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
+Figure \ref{fig:violin} shows that for CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
 Run rate can go up to 6 runs per over depending on the situation.
@@
 This implies CSK is more of a fighter than MI when it comes to defending in the 2nd innings.
+However, we can delve deeper and try to see which overs are the most volatile for CSK and if it is different to that of MI?
+
+In Figure \ref{fig:lv}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
+It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
+There is an upward shift of the distribution in the second part of the first innings (after 10th over) and more variability (longer and distinct letter values) in the second part of 2nd innings.
+CSK has been more uniform in their approach maintaining consistent scores per over, whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second.
+CSK's run rate distibution is extremely right tailed in the 2nd innings, but not in the first.
+Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}.
+
+<!-- ```{r 2013_15_MI} -->
+<!--  cricket_tsibble_all %>%  -->
+<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
+<!--            granplot("inning", "over", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "run_rate", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
+<!--     theme( # remove the vertical grid lines -->
+<!--            panel.grid.major.x = element_blank() , -->
+<!--            # explicitly set the horizontal lines (or they will disappear too) -->
+<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
+<!--     ) -->
+<!--  cricket_tsibble_all %>%  -->
+<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
+<!--            granplot("inning", "over", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "run_rate", -->
+<!--                                plot_type = "quantile", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
+
+<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
+<!--     theme( # remove the vertical grid lines -->
+<!--            panel.grid.major.x = element_blank() , -->
+<!--            # explicitly set the horizontal lines (or they will disappear too) -->
+<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
+<!--     ) -->
+
+<!-- ``` -->
+
+```{r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."}
+
+ cricket_tsibble_all %>% 
+  filter(season %in% c(2010:2015), batting_team=="Mumbai Indians") %>% 
+           granplot("inning", "over",
+                               hierarchy_model,
+                               response = "run_rate",
+                               plot_type = "lv",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")
+
+ cricket_tsibble_all %>% 
+  filter(season %in% c(2010:2015), batting_team=="Chennai Super Kings",
+         inning %in% c(1,2)) %>% 
+           granplot("inning", "over",
+                               hierarchy_model,
+                               response = "run_rate",
+                               plot_type = "lv",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
+
+```
 

### 2019-09-23 main.Rmd
changing season to make analysis more interesting

@@
              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians",
-                    season %in% c(2010:2015)) %>% 
+                    season %in% c(2013, 2015)) %>% 
              mutate(win = if_else(winner == "Mumbai Indians", 1, 0))
@@
              filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings",
-                    season %in% c(2010:2015)) %>% 
+                    season %in% c(2013, 2015)) %>% 
              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0))
@@
                     winner!= "Mumbai Indians",
-                    win_by_runs!=0) %>% select(win_by_runs) %>% summary()
+                    win_by_runs!=0,
+                    season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary()
 
@@
                     win_by_runs!=0,
-                    inning ==2) %>% select(win_by_runs) %>% summary()
+                    inning ==2, 
+                    season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary()
 ```
@@
  cricket_tsibble_all %>% 
-  filter(season %in% c(2010:2015), batting_team=="Mumbai Indians") %>% 
+  filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>% 
            granplot("inning", "over",
@@
  cricket_tsibble_all %>% 
-  filter(season %in% c(2010:2015), batting_team=="Chennai Super Kings",
-         inning %in% c(1,2)) %>% 
+  filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>% 
            granplot("inning", "over",

### 2019-09-23 main.Rmd
idea 1: with good fielding( caught and caught and bowled, run rate decreases

@@
 
+MI_winner %>% 
+  gran_obs("win", "inning", hierarchy_model)
+
 CSK_winner <- cricket_winner %>% 
@@
              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0))
+
+CSK_winner %>% 
+  gran_obs("win", "inning", hierarchy_model)
 
@@
 
- cricket_tsibble_all %>% 
-  filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>% 
+MI <- cricket_tsibble_all %>% 
+  filter(season %in% c(2010:2015), batting_team=="Mumbai Indians")
+
+CSK <- cricket_tsibble_all %>% 
+  filter(season %in% c(2010:2015),
+         batting_team=="Chennai Super Kings",
+         inning %in% c(1,2))
+
+ MI %>% 
+   gran_obs("inning", "over", hierarchy_model) %>%
+   knitr::kable(format = "markdown")
+
+ CSK %>% 
+   gran_obs("inning", "over", hierarchy_model) %>%
+   knitr::kable(format = "markdown")
+
+MI %>% 
   granplot("inning", "over",
@@
 
- cricket_tsibble_all %>% 
-  filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>% 
+CSK %>% 
   granplot("inning", "over",
@@
 
+### Overs in which fielding is good, run rate in that over is set ti reduce
+
+```{r}
+cricket_fielding <- cricket_season %>% 
+  mutate(fielding_proxy = if_else(dismissal_kind %in% c("caught", "caught and bowled"),1,0)) %>% 
+  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs),
+            fielding_wckts = sum(fielding_proxy))
+
+cricket_tsibble_fld <- cricket_fielding %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
+
+cricket_tsibble_fld %>% gran_obs("over", "fielding_wckts")
+
+cricket_tsibble_fld %>% 
+  filter(fielding_wckts %in% c(0,1,2)) %>% 
+  granplot( "over","fielding_wckts",
+                               hierarchy_model,
+                               response = "run_rate",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
+           overlay = FALSE)  + ggtitle("")
+
+```
+
 <!--   cricket_tsibble %>% granplot("inning_match", "over_inning", -->

### 2019-09-23 main.Rmd
cricket added 4 analysis

@@
 
-Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that.
+<!-- Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that. -->
 
@@
 
-How the scores per over vary across granularities/categorisations like innings of a match or matches of a season?
+<!-- How the scores per over vary across granularities/categorisations like innings of a match or matches of a season?
 Are these different for the winning teams in that season and the teams that couldn't qualify for the playoffs?
-
-Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings?
-
-We will look at the ball by ball data for all batting teams.
+-->
+
+<!-- Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings? -->
+
+<!-- We will look at the ball by ball data for all batting teams.
 Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
 Also, we look at runs per over as that would have more variability compared to runs per ball and it be easier to observe the strategies of the winning team through that.
+-->
+
+First, we look at the distribution of runs per over across over of the innings and seasons in \ref{fig:seas-over-inning}.
+The distribution of runs per over has not significantly changed from 2008 to 2016.
+There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
+Hence, we work with subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.
+
+Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final 4 from 2010 to 2015.
+It would be interesting to see the difference in their strategies throughout all matches in the two seasons.
+The following two questions might help us partially understand their strategies.
+
+- Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first?
+
+- Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings?
 
 ```{r runs_per-over}
-
 cricket <-  read_csv("data-raw/deliveries_all.csv")
@@
 ```
-
-First, we look at the distribution of runs per over across over of the innings and seasons in \ref{fig:seas-over-inning}.
-The distribution of runs per over has not significantly changed from 2008 to 2016.
-There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
-Hence, we work with subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.
 
@@
 ```
-
-Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles.
-MI won against CSK both in 2013 and 2015 finals.
-It would be interesting to see the difference in their strategies throughout all matches in the two seasons.
-We can understand difference in strategy by answering these two questions.
-
-- Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first?
-
-- Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings?
 
@@
 It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
-There is an upward shift of the distribution in the second part of the first innings (after 10th over) and more variability (longer and distinct letter values) in the second part of 2nd innings.
-CSK has been more uniform in their approach maintaining consistent scores per over, whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second.
+There is an upward shift of the distribution in the second part of the innings (after 10th over) and more variability (longer and distinct letter values) as well.
+CSK has been more uniform in their approach maintaining consistent scores per over.
+Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}.
+
+<!-- whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second.
 CSK's run rate distibution is extremely right tailed in the 2nd innings, but not in the first.
-Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}.
+.
+--> -->
 
@@
    gran_obs("inning", "over", hierarchy_model) %>%
-   knitr::kable(format = "markdown")
+   knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
 
@@
 
-### Overs in which fielding is good, run rate in that over is set ti reduce
-
-```{r}
+Q3: Is run rate set to reduce in overs where fielding is good?
+
+For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
+If a batsman is bowled out, it does not necessarily signify good fielding.
+So we only include catches and run out as a measure of fielding.
+Difference in run rates should be negative if fielding is good.
+Let us see if this fact is true.Thus, Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over.
+
+```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and dismissals by catch and catch and bowled."}
 cricket_fielding <- cricket_season %>% 
@@
   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs),
-            fielding_wckts = sum(fielding_proxy))
+  summarise(runs_per_over = sum(total_runs),
+            run_rate = sum(total_runs)/length(total_runs),
+            fielding_wckts = sum(fielding_proxy)) %>%
+            mutate(lagged_run_rate = c(0,diff(run_rate)))
 
@@
   filter(fielding_wckts %in% c(0,1,2)) %>% 
-  granplot( "over","fielding_wckts",
+  granplot( "fielding_wckts","over",
+                               hierarchy_model,
+                               response = "lagged_run_rate",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
+           overlay = TRUE)  + ggtitle("")
+
+```
+
+Q4: Is run rate set to reduce in overs where number of dot balls are more/number of wickets are more?
+
+A dot ball is a delivery bowled without any runs scored off it.
+The number of dot balls is reflective of the quality of bowling in the game.
+Run rate of an over should ideally decrease if the number of dot balls increase or the batsman can likely to go for big shots and get dismissed.
+Figure \ref{fig:dot} shows that for any over, increase in dot balls lead to decrease in run rate.
+
+```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."}
+cricket_dot <- cricket_season %>% 
+  mutate(wicket = if_else(is.na(dismissal_kind),0,1),
+         dot_balls = if_else(total_runs==0,1, 0)
+         ) %>% 
+  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs),
+            wicket = sum(wicket),
+            dot_balls = sum(dot_balls))
+
+cricket_tsibble_dot <- cricket_dot %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
+
+#cricket_tsibble_dot %>% gran_obs("over", "fielding_wckts")
+
+cricket_tsibble_dot  %>% filter(dot_balls %in% c(1,2,3,4)) %>% 
+  granplot("over","dot_balls",
                                hierarchy_model,
@@
            overlay = FALSE)  + ggtitle("")
-
 ```

### 2019-09-24 main.Rmd
energy analysis added

@@
 The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns.
-
-Let us see the behavior of typical and extreme behaviors of 50 households of this trial.
-We want to  look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities.
-
-```{r search_gran}
+We look at the behavior of typical and extreme behaviors of 50 households from that trial.
+
+<!-- We want to  look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities. -->
+
+Figure \ref{fig:my-hd} shows the overlay quantile plot of energy consumption of 50 households.
+The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+Each facet represents a month and energy consumption across each hours of the day is shown inside each facet.
+It can be observed that the median is very close (and hence not distinctly visible) to the lower boundaries of all the other bands implying energy consumption for these households are extremely left skewed.75% of the households has distinct higher evening peaks compared to morning peaks, especially in the winter months.
+The top 5% households behave very differently than even the top 10% households in terms of total energy consumption across hours of the day.
+
+```{r my-hd, echo=TRUE}
 
@@
     plot_type = "quantile",
-    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
   ) +
-  ggtitle("Quantile plot across hour-of-day by month-ofyear") + ggtitle("") +
-  scale_x_discrete(breaks = seq(0, 23, 2)) + scale_color_brewer(type = "div",
-                                                                palette = "Dark2", direction = -1)
-
-#observation: the two peaks of the day are more ditinct in winter months compared to summaer months, where the energy profile is quite flat.
+  ggtitle("") 
+```
+
+Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
+Thsi is true for both top 50% and 75% of the customers.
+Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
+All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
+
+```{r dw-hd}
 
@@
            response = "general_supply_kwh", overlay = FALSE) +
-  ggtitle("") +
-  ggtitle("Quantile plot across hour-of-day and day-of-week") +
-  scale_x_discrete(breaks = seq(0, 23, 3)) +
-  scale_color_brewer(type = "div",
-                     palette = "Dark2", direction = -1)
-#observation: on saturday and sunday morning peaks are higher and little late (in absolte time) than other weekdays
-
+  ggtitle("")
+
+```
+
+Figure \ref{fig:hhd-dm} shows distribution of energy consumption across hour hour of the day across different days of the month.
+We can't see much variation across days of week implying that the energy behavior doesn't change depending on which day of the month it is.
+
+```{r hhd-dm}
 # Across days of the week by months
+# sm_cust50 %>%
+#   granplot("month_year",
+#     "day_week",
+#     response = "general_supply_kwh", plot_type = "quantile", 
+#           quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)) + ggtitle("") +
+#   ggtitle("Quantile plot across day-of-week by month-of-year")
+
+# Across week of the month and weekends
+
+# sm_cust50 %>%
+#   dynamic_create_gran("day_week") %>%
+#   filter(day_week %in% c("Sat", "Sun")) %>%
+#   granplot("day_week", "week_month",
+#            response = "general_supply_kwh",
+#            plot_type = "lv") + 
+#   ggtitle("")
+
+# 
+# sm_cust50 %>%
+#   granplot( "hhour_hour","hour_day",
+#            response = "general_supply_kwh",
+#            plot_type = "quantile", overlay = TRUE) + 
+#   ggtitle("")
+
 sm_cust50 %>%
-  granplot("month_year",
-    "day_week",
-    response = "general_supply_kwh", plot_type = "quantile", 
-          quantile_prob = c(0.25, 0.5, 0.75), overlay = FALSE) + ggtitle("") +
-  ggtitle("Quantile plot across day-of-week by month-of-year")
-
-# Across week of the month and weekends
-
-sm_cust50 %>%
-  dynamic_create_gran("day_week") %>%
-  filter(day_week %in% c("Sat", "Sun")) %>%
-  granplot("day_week", "week_month",
+  granplot( "hhour_day","day_month",
            response = "general_supply_kwh",
-           plot_type = "lv") + 
-  ggtitle("") +
-  ggtitle("Letter value plot across day-of-week by week-of-month") +
-  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
-
-sm_cust50 %>%
-  granplot("fortnight_month", "hour_day",
-           response = "general_supply_kwh",
-           plot_type = "lv") + 
+           plot_type = "quantile", overlay = TRUE) + 
   ggtitle("")

### 2019-10-28 main.Rmd
rmd update with prob_plot

@@
 Social and economic data are often collected and reported at coarser temporal scales like monthly, quarterly or annually.
-But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible.
-It might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimization every second.
+But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales.
+Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be collected at every second.
 As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases.
@@
 
-# Formal conceptualization of calendar categorizations{#cirgran-def}
+# Definitions of time granularities{#cirgran-def}
 
@@
 
-## Linear time granularities{#linear-gran-def}
+## Linear {#linear-gran-def}
 
@@
 \autoref{fig:linear-time} shows the implication of this condition.
-If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor s/(24*7)\rfloor$ weeks and so on.
+If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
 
@@
 
-## Circular time granularities{#circular-gran-def}
-
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index in a tsibble[@Wang2019-aw] for a given key.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-MOH: & $C_1(s)$ & $= s \mod 60$ & $n_1 =$&$60$ \\
-MOD: & $C_2(s)$ & $= s \mod `r 60*24`$ & $n_2=$&$`r 60*24`$\\
-
-HOD: & $C_4(s)$ & $= \lfloor s/60\rfloor\mod 24$  & $n_4 =$&$24$ \\
-HOW: & $C_6(s)$ & $= \lfloor s/60\rfloor\mod 24*7$  & $n_5=$&$`r 24*7`$\\
-
-DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_6=$&$7$\\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Illustrative circular granularities with time index in minutes}
-\label{definitions}
-\end{table}
+## Circular {#circular-gran-def}
+
+Periodicity is very common in all kinds of data and circular granularities can be construed to reflect such periodicities.
+Circular granularities can be constructed by putting to use two linear granularities.
+However, mappings between linear granularities can be regular or irregular.
+For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
+On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
+Hence, circular granularities can also be "periodic" or "aperiodic" depending on
+
+A linear granularity, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to subsets of the bottom granularity.
+For example, if the bottom granularity is days, then a linear granularity might be weeks, months or years.
+Here, the domain is the index set and range is subsets of bottom granularity.
+Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of bottom granularity.
+They also accomodate periodicity and hence the smaller subsets of bottom granularity are nested within the larger subsets of bottom granularity.
+We use modular arithmetic to define circular granularity and use the following definitions:
+
+<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity. -->
+
+<!-- \begin{table}[ht] -->
+<!-- \begin{center} -->
+<!-- \begin{tabular}{lr@{~}lr@{~}r} -->
+<!-- \toprule -->
+<!-- MOH: & $C_1(s)$ & $= s \mod 60$ & $n_1 =$&$60$ \\ -->
+<!-- MOD: & $C_2(s)$ & $= s \mod `r 60*24`$ & $n_2=$&$`r 60*24`$\\ -->
+
+<!-- HOD: & $C_4(s)$ & $= \lfloor s/60\rfloor\mod 24$  & $n_4 =$&$24$ \\ -->
+<!-- HOW: & $C_6(s)$ & $= \lfloor s/60\rfloor\mod 24*7$  & $n_5=$&$`r 24*7`$\\ -->
+
+<!-- DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_6=$&$7$\\ -->
+<!-- \bottomrule -->
+<!-- \end{tabular} -->
+<!-- \end{center} -->
+<!-- \caption{Illustrative circular granularities with time index in minutes} -->
+<!-- \label{definitions} -->
+<!-- \end{table} -->
 
@@
 
-We want to use modular arithmetic to define circular granularity. Hence, we start with the definition of equivalence classes and then move on to define a circular granularity.
-
 <!-- Formal definition of circular granularity -->
+
 **Definition: Equivalence class** Let $m \in N \backslash{0}$.
@@
 
-For example, suppose $C$ is a circular granularity denoting Hour-of-Day and we have hourly data for 100 hours.
+For example, suppose $C$ is a circular granularity denoting hour-of-day and we have hourly data for 100 hours.
 The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
@@
 
-## Aperiodic time granularities{#aperiodic-gran-def}
+## Aperiodic {#aperiodic-gran-def}
 <!-- Formal definition of aperiodic circular granularity -->
@@
 
-# Single-order-up and multiple-order-up granularities
+# Computation of time granularities
+
+## Definitions of time granularities based on order
+
+The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
+We shall use the notion of a hierarchy table to define them.
+
+Consider a **hierarchy table** to be consisting of two columns:
+
+ -  The first column represents the (temporal) units in ascending order of hierarchy. Any two units can join together to form a granularity.
+
+ - The second column represents the relationship between subsequent (temporal) units.
+For periodic granularities, this could be represented by a constant, whereas, for a aperiodic one, there is no single constant which can be used to represent their relationship.
+
+and, **order**, is defined as the position of the units in the hierarchical table.
+
+We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concering a single level as **single-order-up** granularities.
+Let us look at few calendars to see examples of single and multiple order-up granularities.
 
@@
 The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
-Nevertheless, for any calendar, the hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
-We shall use the notion of a hierarchy table to define them.
-
-Consider a **hierarchy table** to be consisting of two columns:  
- - The first column represents the (temporal) units in ascending order of hierarchy.  
- - The second column represents the constant which relates subsequent (temporal) units.
-
-and, **order**, is defined as the position of the units in the hierarchical table.
-
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concering a single level as **single-order-up** granularities.
-"Mayan" calendar is used to illustrate this.
-In Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+Nevertheless, for any calendar a hierarchy can be defined.
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
 
@@
 
-## Computation of single-order-up granularities
+### Single-order-up granularities
 
 Suppose, z denotes the index of the tsibble, x,y are two units in the hierarchy table  with  $order(x) <  order(y)$.
-Let $f(x, y)$ denotes the accessor function for computing the single-order-up granularity x-of-y and $c(x, y)$ is a constant which relates x and y.
+Let $f(x, y)$ denotes the accessor function for computing the single-order-up granularity relating x and y and $c(x, y)$ is a constant which relates x and y.
 
@@
 
-### Aperiodic single-order-up granularities
+#### Aperiodic single-order-up granularities
 
@@
 So there is no single number that can be used for converting days to months.
-We cannot compute the aperiodic single-order-up granularities using the index of the tsibble and modular arithmetic like in the periodic case.
-
-## Computation of multiple order-up granularities
+We cannot compute the aperiodic single-order-up granularities using the index of the tsibble and modular arithmetic like in the periodic case using just the hierarchy table and relationships of units.
+
+### Computation of multiple order-up granularities
 
@@
 
-### All single-order-up granularities are periodic
+#### All single-order-up granularities are periodic
 
 z is the index set of the tsibble and x, y are two linear granularities with  $order(x) <  order(y)$.
-Also, $f(x, y)$ denotes the accessor function for computing circular granularity x_y and $c(x, y)$ is a constant which relates x and y.
+Also, $f(x, y)$ denotes the accessor function for computing circular granularity which relates x and y and $c(x, y)$ is a constant which relates x and y.
 It is easy to see that for $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
@@
 
-### Mixed single-order-up granularities - circular or aperiodic
+#### Mixed single-order-up granularities - circular or aperiodic
 
@@
 
-# Visualization
-
-<!-- Why visualize -->
-Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
-However, improper choice of models can lead to wrong predictions.
-One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
-
-<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
-
-<!-- Why distribution plot -->
-<!-- Huge data difficult to summaries using one statistic-->
-But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a data set. Soon enough following questions become more interesting:
-
- - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?
-
- - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
-More generally, how the variation is the data set changes across different percentiles/deciles?
-
-  - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median?
-
-This is when displaying a probability distribution becomes a potentially useful approach.
-
-<!-- What all can you see from a distribution plot-->
-
-The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
-
-<!-- Lit review on statistical distribution plot-->
-## Statistical distribution plots
-
-Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-The box plot is a compact distributional summary, displaying less detail than a histogram.
-Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
-
-The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.
-Violin plots [@Hintze1998-zi, 1998] display the density for all data points and not only the box.
-The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
-The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
-The probabilities for which the summarization is required can be chosen based on the requirement.
-These regions do not need to be contiguous and help identify multi-modality.
-The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
-Because this display just adds extra letter values, it suffers from the same problems as the original box plot, and multimodality is almost impossible to spot[@Wickham_undated-vr].
-
-Moreover, much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
-The deciles plots consist of 9 values that split the data set into ten parts and the percentile plot consists of 99 values that split the data set into hundred parts.
-A large data set is required before the extreme percentiles can be estimated with any accuracy.
-
-Finally, a density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution.
-Also, a Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
-Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
+# Interaction
 
@@
   b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets.
+
+# Visualization
+
+<!-- Why visualize -->
+Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
+However, improper choice of models can lead to wrong predictions.
+One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
+
+<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
+
+<!-- Why distribution plot -->
+<!-- Huge data difficult to summaries using one statistic-->
+But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a data set. Soon enough following questions become more interesting:
+
+ - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?
+
+ - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
+More generally, how the variation is the data set changes across different percentiles/deciles?
+
+  - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median?
+
+This is when displaying a probability distribution becomes a potentially useful approach.
+
+<!-- What all can you see from a distribution plot-->
+
+The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
+These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
+
+<!-- Lit review on statistical distribution plot-->
+## Statistical distribution plots
+
+Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+The box plot is a compact distributional summary, displaying less detail than a histogram.
+Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
+Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
+
+The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.
+Violin plots [@Hintze1998-zi, 1998] display the density for all data points and not only the box.
+The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
+The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
+The probabilities for which the summarization is required can be chosen based on the requirement.
+These regions do not need to be contiguous and help identify multi-modality.
+The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+Because this display just adds extra letter values, it suffers from the same problems as the original box plot, and multimodality is almost impossible to spot[@Wickham_undated-vr].
+
+Moreover, much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
+The deciles plots consist of 9 values that split the data set into ten parts and the percentile plot consists of 99 values that split the data set into hundred parts.
+A large data set is required before the extreme percentiles can be estimated with any accuracy.
+
+Finally, a density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution.
+Also, a Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
+Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
+
 <!-- ## Distributions conditional on bivariate granularities -->
@@
 sm_cust50 %>%
-  granplot("month_year", "hour_day",
+  prob_plot("month_year", "hour_day",
     response = "general_supply_kwh",
@@
 sm_cust50 %>%
-  granplot("day_week", 
+  prob_plot("day_week", 
            "hour_day",
@@
 # sm_cust50 %>%
-#   granplot("month_year",
+#   prob_plot("month_year",
 #     "day_week",
@@
 #   filter(day_week %in% c("Sat", "Sun")) %>%
-#   granplot("day_week", "week_month",
+#   prob_plot("day_week", "week_month",
 #            response = "general_supply_kwh",
@@
 # sm_cust50 %>%
-#   granplot( "hhour_hour","hour_day",
+#   prob_plot( "hhour_hour","hour_day",
 #            response = "general_supply_kwh",
@@
 sm_cust50 %>%
-  granplot( "hhour_day","day_month",
+  prob_plot( "hhour_day","day_month",
            response = "general_supply_kwh",
@@
   cricket_tsibble_all %>% 
-  granplot("over", "season",
+  prob_plot("over", "season",
                                 hierarchy_model,
@@
  MI_winner %>% 
-            granplot( "win","inning",
+            prob_plot( "win","inning",
                                 hierarchy_model,
@@
  CSK_winner %>% filter(inning %in% c(1,2)) %>% 
-           granplot( "win","inning",
+           prob_plot( "win","inning",
                                hierarchy_model,
@@
 <!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
-<!--            granplot("inning", "over", -->
+<!--            prob_plot("inning", "over", -->
 <!--                                hierarchy_model, -->
@@
 <!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
-<!--            granplot("inning", "over", -->
+<!--            prob_plot("inning", "over", -->
 <!--                                hierarchy_model, -->
@@
   filter(season %in% c(2010:2015),
-         batting_team=="Chennai Super Kings",
+         batting_team=="Kolkata Knight Riders",
          inning %in% c(1,2))
@@
 MI %>% 
-  granplot("inning", "over",
+  prob_plot("inning", "over",
                                hierarchy_model,
@@
 CSK %>% 
-  granplot("inning", "over",
+  prob_plot("inning", "over",
                                hierarchy_model,
@@
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
+
+cricket_tsibble_all %>% distinct(batting_team)
 
@@
   filter(fielding_wckts %in% c(0,1,2)) %>% 
-  granplot( "fielding_wckts","over",
+  prob_plot( "fielding_wckts","over",
                                hierarchy_model,
@@
 ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."}
+
 cricket_dot <- cricket_season %>% 
@@
 cricket_tsibble_dot  %>% filter(dot_balls %in% c(1,2,3,4)) %>% 
-  granplot("over","dot_balls",
+  prob_plot("over","dot_balls",
                                hierarchy_model,
@@
 
-<!--   cricket_tsibble %>% granplot("inning_match", "over_inning", -->
+<!--   cricket_tsibble %>% prob_plot("inning_match", "over_inning", -->
 <!--                                hierarchy_model, -->
@@
 <!--     filter(batting_team == "Chennai Super Kings") %>%  -->
-<!--     granplot("inning_match", "over_inning", -->
+<!--     prob_plot("inning_match", "over_inning", -->
 <!--                                hierarchy_model, -->
@@
 <!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
-<!--     granplot("inning_match", "over_inning", -->
+<!--     prob_plot("inning_match", "over_inning", -->
 <!--                                hierarchy_model, -->
@@
 
-<!--   cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
+<!--   cricket_tsibble_all %>% prob_plot("inning_match", "over_inning", -->
 <!--                                hierarchy_model, -->
@@
 
-<!--   cricket_tsibble_all %>% granplot("inning_match", "season", -->
+<!--   cricket_tsibble_all %>% prob_plot("inning_match", "season", -->
 <!--                                hierarchy_model, -->
@@
 
-<!--   cricket_tsibble_all %>% granplot("season", "over_inning", -->
+<!--   cricket_tsibble_all %>% prob_plot("season", "over_inning", -->
 <!--                                hierarchy_model, -->

### 2019-10-29 main.Rmd
relating linear granularity with circular one

@@
 A linear granularity, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to subsets of the bottom granularity.
-For example, if the bottom granularity is days, then a linear granularity might be weeks, months or years.
 Here, the domain is the index set and range is subsets of bottom granularity.
 Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of bottom granularity.
-They also accomodate periodicity and hence the smaller subsets of bottom granularity are nested within the larger subsets of bottom granularity.
+They also accomodate periodicity and hence the smaller subsets of bottom granularity are repeating themselves within the larger subsets of bottom granularity.
 We use modular arithmetic to define circular granularity and use the following definitions:

### 2019-10-30 main.Rmd
making changes to circular granularity definition

@@
 
-A linear granularity, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to subsets of the bottom granularity.
-Here, the domain is the index set and range is subsets of bottom granularity.
-Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of bottom granularity.
-They also accomodate periodicity and hence the smaller subsets of bottom granularity are repeating themselves within the larger subsets of bottom granularity.
+Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
+
+A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index.
+
+A linear granularity is a mapping of row numbers to subsets of the time domain.
+For example, if the time index is days, then a linear granularity might be weeks, months or years.
+Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain.
 We use modular arithmetic to define circular granularity and use the following definitions:
+
+<!-- A linear granularity, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to subsets of linear bottom granularity.
+Here, the domain is the index set and range contains subsets of linear bottom granularity.
+Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of linear bottom granularity.
+We use modular arithmetic to define circular granularity and use the following definitions: -->
+
+<!-- They also accomodate periodicity and hence the smaller subsets of bottom granularity are repeating themselves within the larger subsets of bottom granularity.  -->
 
@@
 
-**Definition:** A **circular granularity** $C$ with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that $C(s) = (s\mod m)$ for $s \in Z$.
+**Definition:** A **circular granularity** $C$, joining two linear granularity with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that
+
+$C(s) = (s\mod m)$ for $s \in Z$.
 

### 2019-10-30 main.Rmd
harmony and clash section update

@@
 
-## Harmony and Clashes
+When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
+It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
+
+For example, suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A,B,C,D$\}, and $C_2$ maps row numbers to a set $\{X,Y,Z\}$.
+That is, let $S_{ij}$ be the set of row numbers such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
+In this example, we have 12 such sets $S_{ij}$ because $i$ can take 4 values and $j$ can take 3 values.
+The graphs that don't work are those where many of these 12 sets are empty.
+Now, many situations can lead to any of these sets being empty.
+Let us discuss the following cases, where one or more of these sets can be empty.
+
+Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
+These are called structurally empty combinations.
+Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+These are structurally empty sets in that it is impossible for them to have any observations.
+
+Secondly, empty combinations can turn up due to difference in occurance of event location or duration in a calendar.
+These are called event-driven empty combinations.
+Again, let us consider a specific example to illustrate this.
+Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
+Here $C_1$ can take 7 values while $C_2$ can take 2 values.
+So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
+While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
+For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
+
+<!-- However, this is different from the first example because these are not structurally empty.
+Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
+-->
+
+Thirdly, empty combinations can be a result of how granularities are constructed.
+Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
+Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
+this is different from the first case because these are not structurally empty combinations.
+The structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
+Hence, they are referred to as "build-based" empty combinations.
+
+Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
 
@@
 The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
+
+## Harmony and Clashes
 

### 2019-10-30 main.Rmd
interaction section update

@@
 
-When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
+When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems might arise when we have empty combinations.
 It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
@@
 Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called structurally empty combinations.
+These are called "structurally" empty combinations.
 Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
@@
 Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
+For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
 In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
@@
 
-Secondly, empty combinations can turn up due to difference in occurance of event location or duration in a calendar.
-These are called event-driven empty combinations.
+Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
+These are called "event-driven" empty combinations.
 Again, let us consider a specific example to illustrate this.
@@
 Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
-this is different from the first case because these are not structurally empty combinations.
-The structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
+This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
 Hence, they are referred to as "build-based" empty combinations.
 
-Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
 Here $C_1$ can take 7 values while $C_2$ can take 12 values.
@@
 
-We investigate some combinations of circular time granularities (periodic/aperiodic) which facilitate or hinder exploratory analysis.
-The combinations of circular granularities which promote the exploratory analysis through visualization are referred to as **harmonies** and the ones which impede the analysis are referred to as **clashes**.
+The combinations of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
+And the ones that ones that do not lead to any missing combinations are called **harmonies** as they promote the exploratory analysis through visualization.
+
+Even with harmonies, problems might occur due to rarely occuring categories.
+suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
+Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
+There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(n_1n_2)$.
+If we require at least $k$ observations to create a meaningful panel, then provided $T \ge n_1n_2k$, the visualization will be acceptable.
+The value of $k$ will depend on what type of visualization we are producing.
+For scatterplots, even $k=3$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
+
+The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations.
+
+Problems will also occur when we have rarely occurring categories such as the 366th day of the year, or the 31st day of the month.
 

### 2019-11-03 main.Rmd
cleaning old ggplot2 codes to prob_plot

@@
 <!--temporal granularities and why should we care -->
-Temporal data can be available at various resolution depending on the context.
-Social and economic data are often collected and reported at coarser temporal scales like monthly, quarterly or annually.
-But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales.
+
+<!-- problem of analyzing data with such finer scales and classical ways to deal with it -->
+Temporal data are available at various resolution depending on the context.
+Social and economic data like GDP are often collected and reported at coarser temporal scales like monthly, quarterly or annually.
+With recent advancement in technology, more and more data are recorded and stored at much finer temporal scales.
 Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be collected at every second.
-As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases.
+As the frequency of data increases, the number of questions about periodicity of the observed variable also increases.
 For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
 This approach requires deconstructing time in various possible ways.
-
-A temporal granularity which results from such a deconstruction may be intuitively described as a sequence of time granules, each one consisting of a set of time instants.
-There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
+Calendar-based graphics[] can unpack the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
+They mainly pick out patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
+Any sub-daily resolution temporal data can also be displayed using this type of faceting [reference trellis plot] with days of the week, month of the year and another sub-daily deconstruction of time.
+
+<!-- # challenges of classical ways of displaying -->
+But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
+A temporal granularity which results from such a deconstruction may be intuitively described as a sequence of time granules, each one consisting of a set of time instants[].
+There can be several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
@@
 Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.
-
+Also, <!--details abt time gran -->
+the hierarchical structure of time creates a natural nested ordering.
+For example, hours are nested within days, days within weeks, weeks within months, and so on.
+Hence, we can construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
+
+<!-- We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
+For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and are single-order-up.
+-->
+
+<!-- TODO -->
 It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the observed data.
@@
 It is a process of generating hypothesis, testing them and consequently refining them through investigations.
-
-<!--details abt time gran --> 
-The hierarchical structure of many granularities creates a natural nested ordering.
-For example, hours are nested within days, days within weeks, weeks within months, and so on.
-We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
 
@@
 
+<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+These are structurally empty sets in that it is impossible for them to have any observations.
+-->
+
+<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
+-->
+
+\autoref{fig:allFig} (a) shows the letter value plot of electricity consumption of Victoria across days of the month for few months like January, February, April and December.
+Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
+Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
+b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
+c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
+We can conclude that Month-of-Year and Day-of-Month are harmonies.
+
+\autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
+The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
+All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
+Hence, Day-of-Month and Day-of-Year are clashes.
+
+Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+Violin plots are a combination of box plot and density plot.
+Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
+These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
+
+In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
+The y-axis represents week of the year and the x-axis represents electricity consumption.
+Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
+They are designed to bring out changes in distributions over time.
+The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
+So, Week-of-Month and Week-of-Year are clashes.
+
+\autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
+Decile plots are useful in displaying distribution through deciles without much clutter.
+This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
+So Month-of-Year and Day-of-Year are clashes as well.
+
+From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
+Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
+
+```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
+(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
+Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
+
+# change all of these to gravitas function
+library(gravitas)
+library(tsibbledata)
+library(dplyr)
+library(ggplot2)
+
+par(mfrow = c(3, 2))
+
+VIC <- vic_elec
+
+VIC %>% 
+  create_gran("month_year") %>% 
+  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>% 
+  prob_plot("month_year","day_month",
+            response = "Demand",
+            plot_type = "lv") + 
+  ylab("Electricity Demand [KWh]") +
+  ggtitle("(a) Letter value plot")
+
+VIC %>% 
+  create_gran("day_week") %>% 
+  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>% 
+  prob_plot("day_week","hour_day",
+            response = "Demand",
+            plot_type = "quantile",
+            symmetric = FALSE) + 
+  ylab("Electricity Demand [KWh]") +
+  ggtitle("(b) Decile plot")
+
+VIC %>%
+  create_gran("day_month") %>% 
+  filter(day_month %in% c(1, 15, 29, 31)) %>%
+  prob_plot("day_month",
+                "day_year",
+                response = "Demand",
+                plot_type = "boxplot")+
+  ylab("Electricity Demand [KWh]") +
+  scale_x_discrete(breaks = seq(0, 366, 60)) + 
+  theme(legend.position = "bottom", 
+        strip.text = element_text(size = 7, margin = margin())) +
+  ggtitle("(c) Box plot")
+
+VIC %>%
+    create_gran("week_month") %>% 
+  filter(week_month %in% c(1, 2, 4)) %>%
+   prob_plot("week_month",
+                "day_month",
+                response = "Demand",
+                plot_type = "violin", alpha = 0.3) + ylab("") + 
+  xlab("Days of the Month") + 
+  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + 
+  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
+  ggtitle("(d) Violin plot")
+
+VIC %>%  create_gran("week_month") %>%
+  create_gran("week_year") %>%
+  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>% 
+  prob_plot("week_year", 
+            "week_month",
+            response = "Demand",
+            plot_type = "ridge")  + 
+  xlab("Electricity Demand [KWh]") +
+  ylab("Weeks of the Year") +
+  scale_x_continuous(breaks = seq(2000, 10000, 3000)) + 
+  theme(legend.position = "bottom", 
+        strip.text = element_text(size = 7, margin = margin())) +
+  ggtitle("(e) Ridge plot")
+
+VIC %>%
+  create_gran("month_year") %>% 
+  filter(month_year %in% c("Jan", "Jul", "Nov")) %>% 
+   prob_plot("month_year", 
+            "day_year",
+            response = "Demand",
+            plot_type = "quantile",
+            quantile_prob = seq(0.1, 0.9, 0.1),
+            symmetric = FALSE) +
+   #scale_x_discrete(breaks = seq(1, 336, 60)) + 
+   ylab("") + 
+   xlab("Day of the Year") + theme(legend.position = "none", 
+                                   strip.text = element_text(size = 7, 
+                                                             margin = margin())) +
+  ggtitle("(f) Decile plot")
+
+# VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
+```
+
 When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems might arise when we have empty combinations.
@@
 Even with harmonies, problems might occur due to rarely occuring categories.
+
 suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
@@
 
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-\autoref{fig:allFig} (a) shows the letter value plot of electricity consumption of Victoria across days of the month for few months like January, February, April and December.
-Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
-
-\autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
-
-Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
-
-In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
-
-\autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
-
-From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
-
-```{r moy-dom,echo=FALSE, out.width = "70%", fig.align = 'center',fig.cap ="Letter-value plot of electricity consumption in Victoria across days of the month\n faceted by months January, February, April and December.
-"}
-
-# change all of these to gravitas function
-VIC <- read.csv("data/VIC2015/Data/demand_VIC.csv")
-VIC$Date <- as.Date(VIC$Date, origin = "1899-12-30")
-
-first_day_of_month_wday <- function(dx) {
-  day(dx) <- 1
-  wday(dx)
-}
-
-VIC <- VIC %>% mutate(
-  Indx_Year = year(Date),
-  Indx_Month = month(Date, label = FALSE, abbr = TRUE),
-  Indx_Wk_Yr = week(Date),
-  Indx_Wk_Month = ceiling((day(Date) + first_day_of_month_wday(Date) - 1) / 7),
-  Indx_Day_Week = wday(Date,
-    label = FALSE, abbr = TRUE,
-    week_start = 1
-  ),
-
-  Indx_Day_Month = day(Date),
-
-  Indx_Day_Year = yday(Date),
-
-  Indx_Weekend = if_else(Indx_Day_Week %in% c(6, 7), 1, 0),
-
-  Indx_HlHr_day = Period,
-  month = month(Date, label = FALSE, abbr = TRUE),
-  year = year(Date),
-  yday = yday(Date),
-  wday = wday(Date,
-    label = FALSE, abbr = TRUE,
-    week_start = 1
-  ),
-  bow = (wday - 1) * 48 + Period,
-  dom = day(Date),
-  bom = (dom - 1) * 48 + Period,
-  Weekend = if_else(wday %in% c(6, 7), 1, 0),
-  Indx_hour = ceiling(Period / 2),
-  Indx_Hour_Yr = Indx_hour + 24 * (yday - 1),
-  Indx_Hour_Month = Indx_hour + 24 * (Indx_Day_Month - 1),
-  Indx_Hour_Wk = Indx_hour + 24 * (wday - 1)
-)
-
-VIC <- as_tibble(VIC)
-```
-
-```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
-Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
-
-par(mfrow = c(3, 2))
-
-VIC %>% filter(year %in% c(2012, 2013, 2014), Indx_Month %in% c(1, 2, 4, 12)) %>% ggplot(aes(Indx_Day_Month, OperationalLessIndustrial, group = Indx_Day_Month)) + geom_lv(aes(fill = ..LV..), outlier.colour = "red", outlier.shape = 1) + scale_fill_brewer() + facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks = seq(0, 31, 5)) + ggtitle("(a) Letter value plot")
-
-VIC_hod_dow <- VIC %>%
-  filter(year %in% c(2012, 2013, 2014)) %>%
-  group_by(wday, Indx_hour) %>%
-  do({
-    x <- .$OperationalLessIndustrial
-    map_dfr(
-      .x = seq(0.1, 0.9, 0.1),
-      .f = ~ tibble(
-        Quantile = .x,
-        Value = quantile(x, probs = .x, na.rm = TRUE)
-      )
-    )
-  }) %>%
-  filter(wday %in% c(1, 2, 6, 7))
-
-VIC_hod_dow %>% ggplot(aes(x = Indx_hour, y = Value, col = as.factor(Quantile))) + geom_line() + facet_wrap(~wday) + scale_x_continuous(breaks = seq(1, 24, 5)) + ylab("") + xlab("Day of the Week") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot")
-
-VIC %>%
-  filter(year %in% c(2012, 2013, 2014), Indx_Day_Month %in% c(1, 15, 29, 31)) %>%
-  ggplot(aes(as.factor(yday), OperationalLessIndustrial, group = yday)) + geom_boxplot() + facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
-  xlab("Days of the Year") + scale_x_discrete(breaks = seq(0, 366, 60)) + theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot")
-
-VIC %>% filter(year %in% c(2012, 2013, 2014), Indx_Wk_Month %in% c(1, 2, 4)) %>% ggplot(aes(as.factor(Indx_Day_Month), OperationalLessIndustrial)) + geom_violin(alpha = 0.03) + facet_wrap(~Indx_Wk_Month, nrow = 3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + scale_y_continuous(breaks = seq(2000, 9000, 2000)) + ggtitle("(d) Violin plot")
-
-VIC %>% dplyr:::filter(year %in% c(2012, 2013, 2014), Indx_Wk_Month %in% c(1, 2, 5), Indx_Wk_Yr < 20) %>% ggplot(aes(x = OperationalLessIndustrial, y = as.factor(Indx_Wk_Yr), group = Indx_Wk_Yr)) + geom_density_ridges2() + facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000, 10000, 3000)) + theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot")
-
-VIC_moy_doy <- VIC %>%
-  filter(year %in% c(2012, 2013, 2014)) %>%
-  group_by(Indx_Month, yday) %>%
-  do({
-    x <- .$OperationalLessIndustrial
-    map_dfr(
-      .x = seq(0.1, 0.9, 0.1),
-      .f = ~ tibble(
-        Quantile = .x,
-        Value = quantile(x, probs = .x, na.rm = TRUE)
-      )
-    )
-  }) %>%
-  filter(Indx_Month %in% c(1, 7, 11))
-
-VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
-
+```{r}
 # grid.arrange(p1,p2,p3,p4,nrow = 2)

### 2019-11-03 main.Rmd
revising section headers

@@
 <!-- general calendar categorization -->
-Often we partition time into months, weeks, days and so on in a hierarchical manner to relate it to data.
+Often we partition time into months, weeks or days to relate it to data.
 Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
 Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
-Let us call all of these different abstractions of time as "calendar categorizations".
-
-These calendar categorizations can be linear, circular or aperiodic.
-The calendar categorizations as "linear" if they respect the linear progression of time.
-We call these **linear time granularities**.
+
+## Arrangement: Linear vs. Circular vs. aperiodic
+
+The arrangement of the time domain can result in deconstructing time in linear, circular or nearly circular ways.
+Time granularities are **linear** if they respect the linear progression of time.
 Examples include hours, days, weeks and months.
-**Circular time granularities** accommodate periodicities in time such as hour of the day, and day of the week.
-**Aperiodic time granularities** are neither linear nor circular, such as day of the month or public holidays.
+**Circular** time granularities accommodate periodicities in time such as hour of the day, and day of the week.
+**Aperiodic** time granularities are neither linear nor circular, such as day of the month or public holidays.
 
@@
 
-## Linear {#linear-gran-def}
+### Linear {#linear-gran-def}
 
@@
 
-## Circular {#circular-gran-def}
-
-Periodicity is very common in all kinds of data and circular granularities can be construed to reflect such periodicities.
-Circular granularities can be constructed by putting to use two linear granularities.
+### Circular {#circular-gran-def}
+
+Periodicity is very common in all kinds of data.
+Circular granularities, which are constructed by using two linear granularities, can be utilized to explain such periodicities.
 However, mappings between linear granularities can be regular or irregular.
@@
 On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Hence, circular granularities can also be "periodic" or "aperiodic" depending on
+Hence, circular granularities can also be "periodic" or "aperiodic" depending on if the mapping is regular or irregular.
 
@@
 
-## Aperiodic {#aperiodic-gran-def}
+### Aperiodic {#aperiodic-gran-def}
 <!-- Formal definition of aperiodic circular granularity -->
@@
 
-Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
+<!-- Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
 Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
-To consider the exhaustive set of temporal regularities that might exist in the data, we can categorize the set of temporal granularities to single or multiple order up granularities.
-
-# Computation of time granularities
-
-## Definitions of time granularities based on order
+-->
+
+To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar.
+
+## Order: Single vs. Multiple
 
@@
 
- -  The first column represents the (temporal) units in ascending order of hierarchy. Any two units can join together to form a granularity.
+ -  The first column represents the (temporal) units in ascending order of hierarchy.
+Any two units, which are essentially the linear granularities, can join together to form a circular/aperiodic granularity.
 
@@
 
-### Single-order-up granularities
+# Computation of time granularities
+
+Computation of time granularities would differ in a periodic and non-periodic set up.
+
+## Single-order-up granularities
 
@@
 
-### Computation of multiple order-up granularities
+## Multiple order-up granularities
 
@@
 
-#### All single-order-up granularities are periodic
+### All single-order-up granularities are periodic
 
@@
 
-#### Mixed single-order-up granularities - circular or aperiodic
+### Mixed single-order-up granularities - circular or aperiodic
 
@@
 
-# Interaction
+# Interaction of time granularities
 
@@
 
-\autoref{fig:allFig} (a) shows the letter value plot of electricity consumption of Victoria across days of the month for few months like January, February, April and December.
-Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+<!-- What  is interaction and some examples why it is important to study interaction -->
+
+Before exploring the behavior of a "dependent variable" across time granularities,
+it is important to know how these granularities interact with each other.
+If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable, For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
+\autoref{fig:allFig}.
+\autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
+\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
+\autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig} (f) shows decile plots across  day of the year and month of the year.
+
+Clearly, in \autoref{fig:allFig}, we observe that some choices of time granularities work and others do not.
+In \autoref{fig:allFig} (c), there will be no observations for some combinations "Day-of-Month" and "Day-of-Year".
+In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
+On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every "Day-of-Week" can occur in any "Month-of-Year".
+Thus the graphs that don't work are those where many of the combination sets are empty.
+In other words, if there are levels of time granularities plotted across x-axis which are not spanned by levels of the time granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
+We hypothesize that the interaction of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
+
+<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
+
+<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
 M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
@@
 We can conclude that Month-of-Year and Day-of-Month are harmonies.
-
-\autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
+-->
+
+<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
 The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
@@
 Hence, Day-of-Month and Day-of-Year are clashes.
-
-Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+-->
+
+<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
 Violin plots are a combination of box plot and density plot.
@@
 These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
-
-In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
+-->
+
+<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
 The y-axis represents week of the year and the x-axis represents electricity consumption.
@@
 So, Week-of-Month and Week-of-Year are clashes.
-
-\autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
+-->
+
+<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
 Decile plots are useful in displaying distribution through deciles without much clutter.
@@
 So Month-of-Year and Day-of-Year are clashes as well.
-
-From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
+-->
+
+<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
+Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
+Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
+-->
+
+<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
+Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
+There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
+-->
+
+<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
 Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
+-->
 
@@
   filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>% 
-  prob_plot("month_year","day_month",
+  prob_plot("month_year","day_week",
             response = "Demand",
@@
 
-When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems might arise when we have empty combinations.
+To define a general framework, we define harmony and clashes.
+Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in a non-compatible ways and hinder the process of exploration of the dependent variable if plotted together.
+
+## Harmony and Clashes
+
+<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
 It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
-
-For example, suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A,B,C,D$\}, and $C_2$ maps row numbers to a set $\{X,Y,Z\}$.
+-->
+
+Suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps row numbers to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
 That is, let $S_{ij}$ be the set of row numbers such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-In this example, we have 12 such sets $S_{ij}$ because $i$ can take 4 values and $j$ can take 3 values.
+Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
 The graphs that don't work are those where many of these 12 sets are empty.
@@
 
-The combinations of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that ones that do not lead to any missing combinations are called **harmonies** as they promote the exploratory analysis through visualization.
-
-Even with harmonies, problems might occur due to rarely occuring categories.
-
-suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
-Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
-There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(n_1n_2)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge n_1n_2k$, the visualization will be acceptable.
-The value of $k$ will depend on what type of visualization we are producing.
-For scatterplots, even $k=3$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
-
-The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations.
-
-Problems will also occur when we have rarely occurring categories such as the 366th day of the year, or the 31st day of the month.
-
-## Harmony and Clashes
+Therefore, combinations of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
+And the ones that do not lead to any missing combinations are called **harmonies** as they promote the exploratory analysis through visualization.
 
@@
 
-In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
-
-In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
-
-More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
+<!-- More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
 When C1 and C2 are used as aesthetics or facets, problems arise when we encounter empty combinations.
@@
 In this example, we have 12 such sets $S_{i,j}$ because i can take 4 values and j can take 3 values.
-The graphs that don't work are those where many of these 12 sets are empty.
-In other words, if there are levels of the x-axis which are not spanned by levels of the faceted variable or vice versa we will have structurally empty sets leading to potential ineffective graphs.
-
-Thus, the common link that differentiates harmonies from clashes are:
-
-  a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis.
-
-  b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets.
+-->
+
+<!-- Thus, the common link that differentiates harmonies from clashes are: -->
+
+<!--   a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis. -->
+
+<!--   b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets. -->
 
@@
 
-## Advice algorithm for exploring conditional probability distributions
+## Plotting harmonies and clashes
+
+## Homogeneity across and within facets
 
@@
  X-axis variable is treated as a categorical variable and hence any plots which
+
+Even with harmonies, problems might occur due to rarely occuring categories.
+
+suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
+Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
+There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(n_1n_2)$.
+If we require at least $k$ observations to create a meaningful panel, then provided $T \ge n_1n_2k$, the visualization will be acceptable.
+The value of $k$ will depend on what type of visualization we are producing.
+For scatterplots, even $k=3$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
+
+The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations.
+
+Problems will also occur when we have rarely occurring categories such as the 366th day of the year, or the 31st day of the month.
+
+## Number of levels of the facet variable
 
@@
 
+```{r}
+
 <!--   cricket_tsibble %>% prob_plot("inning_match", "over_inning", -->
@@
 
-```
-
-<!-- ```{r read_data, echo=FALSE} -->
+{r read_data, echo=FALSE}
 
@@
 <!--   ) -->
-<!-- ``` -->
+<!-- ``` --
+```
 

### 2019-11-03 main.Rmd
revising subsections for visualization section

@@
 
+<!-- ## Distributions conditional on bivariate granularities -->
+
+The time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
+That neccessiates plotting one temporal granularity along the x-axis and the other one across facets.
+
+Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
+The harmonies when plotted together can help exploration.
+But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
+This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
+
+## Choice of Plots
+
+Recommendations for distribution plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
+They will vary depending on which granularity is placed on the x-axis and which one across facets.
+Assumptions are made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.
+Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot.
+
+Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
+Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
+For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
+31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
+These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
+Let us consider case by case and see which plots are better suitable in which scenarios.
+
+<!-- Graphical output devices are flat.
+If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
+As soon as the data move to three variables, we must be content in attempting to infer multi-dimensional structure by a two-dimensional medium.
+-->
+
+Facetting creates small multiples of data each representing a different subset of the data.
+Small multiples are powerful for exploratory data analysis where we can rapidly compare patterns in different parts of the data and see whether they are the same or different.
+But space and resolution quickly become a problem if the number of levels of facet variables is too high.
+Also, Cleveland suggested that a graphic should display as much information as it can, with the lowest possible cognitive strain to the viewer.
+
+## Effect of interaction
+
+We should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
+
+## Effect of number of observation
+
+### Rarely occuring events
+
+Even with harmonies, problems might occur due to rarely occuring categories.
+
+Suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
+Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
+There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(n_1n_2)$.
+If we require at least $k$ observations to create a meaningful panel, then provided $T \ge n_1n_2k$, the visualization will be acceptable.
+The value of $k$ will depend on what type of visualization we are producing.
+For scatterplots, even $k=3$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
+
+The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations.
+
+Problems will also occur when we have rarely occurring categories such as the 366th day of the year, or the 31st day of the month.
+
+### Homogeneity within and across facets
+
+We should be cautious to plot two granularities which vary largely across or within facets.
+Even if for each combinations, the number is good enough for plotting distribution, large differences between number of observations between and across facets, might have an impact on the emperical dfistribution.
+We use Gini's measure to compute if differences in number of observations across or within facets are significantly different.
+
 <!-- Lit review on statistical distribution plot-->
-## Statistical distribution plots
+ Statistical distribution plots
 
@@
 
-<!-- ## Distributions conditional on bivariate granularities -->
-
-The time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
-That neccessiates plotting one temporal granularity along the x-axis and the other one across facets.
-
-Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
-The harmonies when plotted together can help exploration.
-But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
-This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
-
-## Plotting harmonies and clashes
-
-## Homogeneity across and within facets
-
-Recommendations for distribution plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
-They will vary depending on which granularity is placed on the x-axis and which one across facets.
-Assumptions are made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.
-Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot.
-
-Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
-These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
-Let us consider case by case and see which plots are better suitable in which scenarios.
-
- -  very high facet and x-axis levels
-
- X-axis variable is treated as a categorical variable and hence any plots which
-
-Even with harmonies, problems might occur due to rarely occuring categories.
-
-suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
-Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
-There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(n_1n_2)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge n_1n_2k$, the visualization will be acceptable.
-The value of $k$ will depend on what type of visualization we are producing.
-For scatterplots, even $k=3$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
-
-The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations.
-
-Problems will also occur when we have rarely occurring categories such as the 366th day of the year, or the 31st day of the month.
-
-## Number of levels of the facet variable
+Recommendation system that aid exploration
 
@@
 
-```{r}
-
 <!--   cricket_tsibble %>% prob_plot("inning_match", "over_inning", -->

### 2019-11-05 main.Rmd
visualization section rough draft update

@@
 
-Before exploring the behavior of a "dependent variable" across time granularities,
-it is important to know how these granularities interact with each other.
-If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable, For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
-\autoref{fig:allFig}.
-\autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
-\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
-\autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig} (f) shows decile plots across  day of the year and month of the year.
-
-Clearly, in \autoref{fig:allFig}, we observe that some choices of time granularities work and others do not.
-In \autoref{fig:allFig} (c), there will be no observations for some combinations "Day-of-Month" and "Day-of-Year".
-In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every "Day-of-Week" can occur in any "Month-of-Year".
-Thus the graphs that don't work are those where many of the combination sets are empty.
-In other words, if there are levels of time granularities plotted across x-axis which are not spanned by levels of the time granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
-We hypothesize that the interaction of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
-
 <!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
@@
 
-```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
-Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
-
-# change all of these to gravitas function
-library(gravitas)
-library(tsibbledata)
-library(dplyr)
-library(ggplot2)
-
-par(mfrow = c(3, 2))
-
-VIC <- vic_elec
-
-VIC %>% 
-  create_gran("month_year") %>% 
-  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>% 
-  prob_plot("month_year","day_week",
-            response = "Demand",
-            plot_type = "lv") + 
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(a) Letter value plot")
-
-VIC %>% 
-  create_gran("day_week") %>% 
-  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>% 
-  prob_plot("day_week","hour_day",
-            response = "Demand",
-            plot_type = "quantile",
-            symmetric = FALSE) + 
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(b) Decile plot")
-
-VIC %>%
-  create_gran("day_month") %>% 
-  filter(day_month %in% c(1, 15, 29, 31)) %>%
-  prob_plot("day_month",
-                "day_year",
-                response = "Demand",
-                plot_type = "boxplot")+
-  ylab("Electricity Demand [KWh]") +
-  scale_x_discrete(breaks = seq(0, 366, 60)) + 
-  theme(legend.position = "bottom", 
-        strip.text = element_text(size = 7, margin = margin())) +
-  ggtitle("(c) Box plot")
-
-VIC %>%
-    create_gran("week_month") %>% 
-  filter(week_month %in% c(1, 2, 4)) %>%
-   prob_plot("week_month",
-                "day_month",
-                response = "Demand",
-                plot_type = "violin", alpha = 0.3) + ylab("") + 
-  xlab("Days of the Month") + 
-  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + 
-  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
-  ggtitle("(d) Violin plot")
-
-VIC %>%  create_gran("week_month") %>%
-  create_gran("week_year") %>%
-  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>% 
-  prob_plot("week_year", 
-            "week_month",
-            response = "Demand",
-            plot_type = "ridge")  + 
-  xlab("Electricity Demand [KWh]") +
-  ylab("Weeks of the Year") +
-  scale_x_continuous(breaks = seq(2000, 10000, 3000)) + 
-  theme(legend.position = "bottom", 
-        strip.text = element_text(size = 7, margin = margin())) +
-  ggtitle("(e) Ridge plot")
-
-VIC %>%
-  create_gran("month_year") %>% 
-  filter(month_year %in% c("Jan", "Jul", "Nov")) %>% 
-   prob_plot("month_year", 
-            "day_year",
-            response = "Demand",
-            plot_type = "quantile",
-            quantile_prob = seq(0.1, 0.9, 0.1),
-            symmetric = FALSE) +
-   #scale_x_discrete(breaks = seq(1, 336, 60)) + 
-   ylab("") + 
-   xlab("Day of the Year") + theme(legend.position = "none", 
-                                   strip.text = element_text(size = 7, 
-                                                             margin = margin())) +
-  ggtitle("(f) Decile plot")
-
-# VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
-```
-
+Before exploring the behavior of a "dependent variable" across time granularities,
+it is important to know how these granularities interact with each other.
+If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
 To define a general framework, we define harmony and clashes.
@@
 <!-- Why visualize -->
-Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
+<!-- Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
 However, improper choice of models can lead to wrong predictions.
 One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
+-->
 
@@
 <!-- Huge data difficult to summaries using one statistic-->
-But with huge amount of data being available, sometimes mean, median or any one summary statistic is not enough to understand a data set. Soon enough following questions become more interesting:
+With huge amount of data being available, mean, median or any one summary statistic might not enough to understand a data set. Soon enough following questions become more interesting:
 
@@
 This is when displaying a probability distribution becomes a potentially useful approach.
-
-<!-- What all can you see from a distribution plot-->
-
 The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
@@
 
-The time series variable can be plotted against many time granularities to get more understanding of the underlying periodicty, however, we will restrict ourselves to see the distribution of the time series across bivariate temporal granularities.
-That neccessiates plotting one temporal granularity along the x-axis and the other one across facets.
-
-Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
+<!-- Why 2D -->
+Graphical output devices are flat.
+If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
+As soon as the data move to three variables, we need to infer multi-dimensional structure by a two-dimensional medium[Cleveland, elements of graphing data].
+Any attempt to encode all these temporal granularities together to develop insights on periodicity might fail or become clumsy.
+Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
+
+<!-- Why facet-ing -->
+`ggplot2` (Wickham, 2016) is an excellent tool that facilitates the process of mapping different variables to a 2D frame through grammar of graphics.
+One aspect of ggplot2 is faceting, inspired by Trellis plots (Cleveland, 1993).
+These are based on conditioning the values taken on by one or more of the variables in a data set.
+This means creating same plots using data subsets corresponding to each level of the conditioned variable.
+This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
+
+<!-- What are we visualizing -->
+We will see the distribution of the time series across bivariate temporal granularities through faceting approach with one temporal granularity plotted along the x-axis, the other one across facets and the dependent time series variable on the y-axis.
+
+Now, different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
+We will look at each of these aspects separately and analyze the effect of each on visualization.
+
+<!-- Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
 The harmonies when plotted together can help exploration.
@@
 This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
+-->
 
@@
 
-Recommendations for distribution plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
-They will vary depending on which granularity is placed on the x-axis and which one across facets.
-Assumptions are made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.
-Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot.
+<!-- Lit review on statistical distribution plot-->
+
+<!-- Description and lit review of boxplot -->
+There are several ways to plot statistical distributions.
+The displayed probabilities in these plots are either computed using kernel density estimation methods or emperical statistical methods.
+We will discuss the benefits and challenges of few conventional and recent ways to plot distributions using both of these methods.
+
+### Emperical methods
+
+Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+The box plot is a compact distributional summary, displaying less detail than a histogram.
+Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
+Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
+However, we do not have an idea if the distribution is multimodal.
+Moreover, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
+Also, the number of outliers is large for larger data set since number of outliers is proportional to the number of observations.
+
+<!-- Description and lit review of letter-value plots -->
+The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.It shows only actual data values and no smoothed summaries.
+However, the letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles.
+These might lead to a lot of letter values being shown and leading to overload of information in one plot.
+Also, multimodality is difficult to spot.
+
+<!-- Description and lit review of quantile plots -->
+
+Quantile plots visually portray the quantiles of the distribution of sample data.
+Much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
+These plots are referred to as "quantile" plots with, where number of quantiles 9 for deciles and 99 for percentiles.
+No distributional assumptions are made about the data since empirical deciles are plotted.
+It avoids much clutter and just enable us to focus on specific probabilities.
+These plots do not allow us to see the shape, skewness or nature of the tail of the distribution with so much clarity as violin plots, ridge plots or letter value plots.
+Also, a large data set is required for the extreme percentiles to be estimated with any accuracy.
+
+<!-- Description and lit review of density plots -->
+
+<!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  -->
+
+### Kernel density estimation methods
+
+A density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution, unlike discretizing the distribution and showing only parts of it.
+
+Violin plots [@Hintze1998-zi, 1998] adds the information available from local density estimates to summary statistics provided by box plots.
+The shape of the violin represents the density estimate of the variable.
+The more data points in a specific range, the larger the violin is for that range.
+Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories.
+Howevr, the density in violin are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
+As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
+
+The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
+This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
+
+<!-- Description and lit review of density plots -->
+ A Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
+Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
+A clear advantage over boxplots is that these plots allow us to see multimodality in the distribution.
+However, these plots can be obscuring when there is overlap of distribution for two or more categories of the y-axis.
+Also, if there are lot of categories, it is difficult to compare the height of the densities across categories.
+
+<!-- Description and lit review of HDR plots -->
+The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
+The probabilities for which the summarization is required can be chosen based on the requirement.
+These regions do not need to be contiguous and help identify multi-modality.
+
+As a general rule, the quantile plots are really useful for picking patterns, whereas, other more involved methods of plotting are useful for studying anomalies or outlier behavior.
+It is good to be aware of the benefits and challenges while choosing any distribution plots, depending on the context.
+
+### Levels of temporal granularities
+
+Choice of plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
+The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.Space and resolution become a problem if the number of levels of facet variables is too high.
+Hence, plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.
 
 Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
+Default values for these levels are chosen based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
+Any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.
+31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
+
+Space and cognitive power both are constrained if the number of levels of facet variables is very high.
+So as a general rule, we should avoid plotting a temporal granularity with very high levels across facets.
+
+ - If levels of both the granularities are low, then any distribution plots might be chosen depending on which feature of the distribution who is interested to look at.
+
+ - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories.
+
+ - If level of the granularity plotted across x-axis is low, then any distribution plots might be chosen depending on which feature of the distribution who is interested to look at, irrespective of the number of the levels of the facet variable.
+
+ - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.
+
+<!-- Let us consider case by case and see which plots are better suitable in which scenarios. -->
+
+<!--  Assumptions should be made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.  -->
+<!-- Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot. -->
+
+<!-- Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
 Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
@@
 Let us consider case by case and see which plots are better suitable in which scenarios.
-
-<!-- Graphical output devices are flat.
-If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
-As soon as the data move to three variables, we must be content in attempting to infer multi-dimensional structure by a two-dimensional medium.
--->
-
-Facetting creates small multiples of data each representing a different subset of the data.
+-->
+
+<!-- Facetting creates small multiples of data each representing a different subset of the data.
 Small multiples are powerful for exploratory data analysis where we can rapidly compare patterns in different parts of the data and see whether they are the same or different.
@@
 Also, Cleveland suggested that a graphic should display as much information as it can, with the lowest possible cognitive strain to the viewer.
+-->
 
@@
 
+For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
+\autoref{fig:allFig}.
+\autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
+\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
+\autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig} (f) shows decile plots across  day of the year and month of the year.
+
+Clearly, in \autoref{fig:allFig}, we observe that some choices of time granularities work and others do not.
+In \autoref{fig:allFig} (c), there will be no observations for some combinations "Day-of-Month" and "Day-of-Year".
+In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
+On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every "Day-of-Week" can occur in any "Month-of-Year".
+Thus the graphs that don't work are those where many of the combination sets are empty.
+In other words, if there are levels of time granularities plotted across x-axis which are not spanned by levels of the time granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
+We hypothesize that the interaction of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
+
 We should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
 
+```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "h", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
+(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
+Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
+
+# change all of these to gravitas function
+library(gravitas)
+library(tsibbledata)
+library(dplyr)
+library(ggplot2)
+
+par(mfrow = c(3, 2))
+
+VIC <- vic_elec
+
+VIC %>% 
+  create_gran("month_year") %>% 
+  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>% 
+  prob_plot("month_year","day_week",
+            response = "Demand",
+            plot_type = "lv") + 
+  ylab("Electricity Demand [KWh]") +
+  ggtitle("(a) Letter value plot")
+
+VIC %>% 
+  create_gran("day_week") %>% 
+  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>% 
+  prob_plot("day_week","hour_day",
+            response = "Demand",
+            plot_type = "quantile",
+            symmetric = FALSE) + 
+  ylab("Electricity Demand [KWh]") +
+  ggtitle("(b) Decile plot")
+
+VIC %>%
+  create_gran("day_month") %>% 
+  filter(day_month %in% c(1, 15, 29, 31)) %>%
+  prob_plot("day_month",
+                "day_year",
+                response = "Demand",
+                plot_type = "boxplot")+
+  ylab("Electricity Demand [KWh]") +
+  scale_x_discrete(breaks = seq(0, 366, 60)) + 
+  theme(legend.position = "bottom", 
+        strip.text = element_text(size = 7, margin = margin())) +
+  ggtitle("(c) Box plot")
+
+VIC %>%
+    create_gran("week_month") %>% 
+  filter(week_month %in% c(1, 2, 4)) %>%
+   prob_plot("week_month",
+                "day_month",
+                response = "Demand",
+                plot_type = "violin", alpha = 0.3) + ylab("") + 
+  xlab("Days of the Month") + 
+  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + 
+  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
+  ggtitle("(d) Violin plot")
+
+VIC %>%  create_gran("week_month") %>%
+  create_gran("week_year") %>%
+  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>% 
+  prob_plot("week_year", 
+            "week_month",
+            response = "Demand",
+            plot_type = "ridge")  + 
+  xlab("Electricity Demand [KWh]") +
+  ylab("Weeks of the Year") +
+  scale_x_continuous(breaks = seq(2000, 10000, 3000)) + 
+  theme(legend.position = "bottom", 
+        strip.text = element_text(size = 7, margin = margin())) +
+  ggtitle("(e) Ridge plot")
+
+VIC %>%
+  create_gran("month_year") %>% 
+  filter(month_year %in% c("Jan", "Jul", "Nov")) %>% 
+   prob_plot("month_year", 
+            "day_year",
+            response = "Demand",
+            plot_type = "quantile",
+            quantile_prob = seq(0.1, 0.9, 0.1),
+            symmetric = FALSE) +
+   #scale_x_discrete(breaks = seq(1, 336, 60)) + 
+   ylab("") + 
+   xlab("Day of the Year") + theme(legend.position = "none", 
+                                   strip.text = element_text(size = 7, 
+                                                             margin = margin())) +
+  ggtitle("(f) Decile plot")
+
+# VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
+```
+
 ## Effect of number of observation
@@
 
-Suppose we have $T$ hourly observations, and $C_1$ is HOD and $C_2$ is DOW.
+ Suppose we have $T$ hourly observations, and $C_1$ with $n_1$ categories and $C_2$ with $n_1$ categories.
 Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
@@
 The value of $k$ will depend on what type of visualization we are producing.
-For scatterplots, even $k=3$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
-
-The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations.
-
-Problems will also occur when we have rarely occurring categories such as the 366th day of the year, or the 31st day of the month.
-
-### Homogeneity within and across facets
-
-We should be cautious to plot two granularities which vary largely across or within facets.
-Even if for each combinations, the number is good enough for plotting distribution, large differences between number of observations between and across facets, might have an impact on the emperical dfistribution.
+For emperical estimates, even $k=10$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
+
+<!-- The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations. -->
+
+Let us take a specific example to illustrate this. Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
+
+### Unevenly distributed events
+
+Even when there are no rarely occuring events, when number of observations are not evenly distributed between different categories of the time granularities.Large differences between number of observations between and across facets, might have an impact on the estimation of dfistribution.
 We use Gini's measure to compute if differences in number of observations across or within facets are significantly different.
-
-<!-- Lit review on statistical distribution plot-->
- Statistical distribution plots
-
-Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-The box plot is a compact distributional summary, displaying less detail than a histogram.
-Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
-
-The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.
-Violin plots [@Hintze1998-zi, 1998] display the density for all data points and not only the box.
-The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
-The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
-The probabilities for which the summarization is required can be chosen based on the requirement.
-These regions do not need to be contiguous and help identify multi-modality.
-The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
-Because this display just adds extra letter values, it suffers from the same problems as the original box plot, and multimodality is almost impossible to spot[@Wickham_undated-vr].
-
-Moreover, much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
-The deciles plots consist of 9 values that split the data set into ten parts and the percentile plot consists of 99 values that split the data set into hundred parts.
-A large data set is required before the extreme percentiles can be estimated with any accuracy.
-
-Finally, a density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution.
-Also, a Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
-Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
-
-Recommendation system that aid exploration
 

### 2019-11-07 main.Rmd
smart meter study update

@@
 
-Smart meters provide large quantities of measurements on energy usage for households across Australia, and indeed many parts of the world.
+Smart meters provide large quantities of measurements on energy usage for households across Australia.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
+It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
+The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns and identify the extreme and regular households amongst these 50 households.
+
+<!-- typical and extreme behaviors of 50 households from that trial.  -->
+
+<!-- , and indeed many parts of the world.
 Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-
-<!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
-It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+-->
+
+<!-- <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions --> -->
+<!-- It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns.
-We look at the behavior of typical and extreme behaviors of 50 households from that trial.
+-->
 
@@
 
-Figure \ref{fig:my-hd} shows the overlay quantile plot of energy consumption of 50 households.
+In Figure \ref{fig:day_fortnight} and \ref{fig:day_fortnight_agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy beahvior of these two customers.
+In Figure \ref{fig:day_fortnight_agg} it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
+However, Figure \ref{fig:day_fortnight} adds to the beahvior by looking at the distribution.
+If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
+However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
+This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
+
+```{r day_fortnight_agg}
+library(lvplot)
+library(ggplot2)
+library(dplyr)
+library(tibble)
+library(gravitas)
+library(tidyr)
+# smart_meter10 %>% 
+#   filter(customer_id %in% c(10006704, 10017936)) %>% 
+#   create_gran("day_fortnight") %>%
+#   as_tibble() %>% 
+#   group_by(customer_id, day_fortnight) %>% 
+#   summarise(median= log(median(general_supply_kwh)),
+#             total = log(sum(general_supply_kwh))) %>% 
+#   gather("daily_consumption", "value", - c(customer_id, day_fortnight)) %>% 
+#   ggplot2::ggplot(aes(
+#     x = as.factor(day_fortnight),
+#     y = value, 
+#     color = daily_consumption,
+#     group = daily_consumption)) +
+#   geom_line() +
+#   xlab("day_fortnight") +
+#   ylab("general_supply_kwh") + 
+#   facet_wrap(~customer_id) + 
+#   theme_bw()
+
+smart_meter10 %>% 
+  filter(customer_id %in% c(10006704, 10017936)) %>% 
+  create_gran("day_fortnight") %>%
+  as_tibble() %>% 
+  group_by(customer_id, day_fortnight) %>% 
+  summarise(median_consumption= median(general_supply_kwh)) %>%
+  ggplot2::ggplot(aes(
+    x = as.factor(day_fortnight),
+    y = median_consumption,
+    color =  customer_id,
+    group = customer_id)) +
+  geom_line() +
+  xlab("day_fortnight") +
+  ylab("general_supply_kwh") + 
+  theme_bw() 
+```
+
+```{r day_fortnight}
+library(lvplot)
+library(ggplot2)
+library(dplyr)
+library(tibble)
+library(gravitas)
+smart_meter10 %>%
+  filter(customer_id %in% c(10006704, 10017936)) %>% 
+  create_gran("day_fortnight") %>%
+  ggplot2::ggplot(aes(
+    x = as.factor(day_fortnight),
+    y = general_supply_kwh)) +
+  xlab("day_fortnight") +
+  geom_lv(
+    outlier.colour = "red",
+    aes(fill = ..LV..),
+    k = 5) +
+  facet_wrap(~customer_id) + 
+  scale_fill_lv() +
+  theme_bw() +
+  scale_y_sqrt() 
+```
+
+<!-- Figure \autoref{fig:my-hd} shows the overlay quantile plot of energy consumption of 50 households.
 The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
@@
 The top 5% households behave very differently than even the top 10% households in terms of total energy consumption across hours of the day.
+-->
+
+While trying to explore the energy behavior of these customers systematically, the first thing we should have at our disposal for examining periodicities of energy behavior across time granularities is to know the number of time granularities we can look at exhaustively.
+If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
+
+```{r}
+library(tsibble)
+interval(smart_meter10)
+smart_meter10 %>% search_gran()
+```
+
+The interval of this tsibble is 30 minutes, and hence the default for search_gran in this case, provides temporal granularities ranging from half-hour to year.
+If these options are considered too many, the default options can be modified to limit the possibilities.
+For example, the most coarse temporal unit can be set to be a “month”.
+
+```{r}
+smart_meter10 %>% 
+  search_gran(highest_unit = "month")
+```
+
+This looks better. However, some intermediate temporal units might not be pertinent to the analysis and we might want to remove them from the list of granularities.
+
+```{r search_gran_limit2}
+
+smart_meter10 %>% search_gran(highest_unit = "month",
+                              filter_out = c("hhour", "fortnight")
+)
+```
+
+Now that we have the list of granularities to look at, we can visualize the distribution.
+From the search list, we found that we can look at six granularities, that amounts to analyzing six graphics.
+However, what happens if we want to see the distribution of energy across two granularities at a time?
+This is equivalent to looking at the distribution of energy consumption across one granularity conditional on another one.
+One way can be to plot one of the granularities on the x-axis and another on the facet.
+Different perspectives of the data can be derived depending on where the granularities are placed.
+
+So, what is the number of pairs of granularities we can look at?
+It is equivalent to taking 2 granularities from 6, which essentially means we need to examine 30 plots.
+The good news is, not all time granularities can be plotted together and we do not have to analyze so many plots!
+
+Harmony/clash can be identified to considerably reduce the number of visualizations that can aid exploratory analysis.
+
+```{r is_harmony, echo= TRUE}
+smart_meter10 %>% 
+  is_harmony(gran1 = "hour_day", 
+             gran2 = "day_week")
+
+smart_meter10 %>%
+  is_harmony(gran1 = "hour_day", 
+             gran2 = "day_week", 
+             facet_h = 14)
+
+smart_meter10 %>% 
+  is_harmony(gran1 = "day_month",
+             gran2 = "week_month")
+```
+
+Let us now look at all the harmonies that we can examine. Fortunately, we are left with only 13 out of 30 visualizations.
+
+```{r harmony, echo=TRUE}
+smart_meter10 %>% harmony(
+  ugran = "month",
+  filter_out = c("hhour", "fortnight")
+)
+```
+
+In \ref{fig:bothcust} We visualize the harmony pair (wknd_wday, hour_day) through a box plot.
+Boxplot of energy consumption is shown across wknd_wday (facet) and hour-day (x-axis) for the same two households.
+For the second household, outliers are less prominent implying their regular behavior is more stable.
+For the first household, energy behavior is not significantly different between weekdays and weekends.
+For the second household, median energy consumption for the early morning hours is extremely high for weekends compared to weekdays.
+
+ For the second household, outliers are less prominent implying their regular behavior is more stable.
+For the first household, energy behavior is not significantly different between weekdays and weekends.
+For the second household, median energy consumption for the early morning hours is extremely high for weekends compared to weekdays.
+
+```{r}
+cust1 <- smart_meter10 %>% 
+  filter(customer_id %in% c(10006704)) %>% 
+  prob_plot("wknd_wday",
+                            "hour_day",
+                            response = "general_supply_kwh",
+                            plot_type = "boxplot") +
+  scale_y_sqrt() +
+  ggtitle("Energy consumption distribution for customer id: 10006704")
+
+cust2 <- smart_meter10 %>% 
+  filter(customer_id %in% c(10017936))%>% 
+  prob_plot("wknd_wday",
+                            "hour_day",
+                            response = "general_supply_kwh",
+                            plot_type = "boxplot") +
+  scale_y_sqrt() + 
+  ggtitle("Energy consumption distribution for customer id: 10017936")
+```
+
+```{r bothcust}
+cust1
+cust2
+```
+
+Now, we would like to see how these customers' behavior relate to the rest of the 
+50 households across these two measures -
+
+- if the regular behavior is regular and extreme behavior extreme or vice versa
+- if the weekend and weekday behavior are different for most of these households.
+
+Figure \ref{fig:wknd_wday50} shows the overlay quantile plot of energy consumption of 10 households.
+The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+It can be observed that the median is very closer to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
+Weekend and weekday behavior is not very different for these ten customers, implying the first customer is more typical than the second customer when compared to these ten customers.
+
+```{r wknd_wday50}
+smart_meter10 %>% 
+  prob_plot("wknd_wday",
+                            "hour_day",
+                            response = "general_supply_kwh",
+                            plot_type = "quantile",
+            symmetric = TRUE, 
+            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
+  scale_y_sqrt()
+```
 

### 2019-11-11 main.Rmd
changing time granulariites start

@@
 Examples include hours, days, weeks and months.
-**Circular** time granularities accommodate periodicities in time such as hour of the day, and day of the week.
-**Aperiodic** time granularities are neither linear nor circular, such as day of the month or public holidays.
+However, periodicity is very common in all kinds of data.
+Time granularities, which are constructed by using two linear granularities, can be utilized to explain such periodicities.
+The mappings between linear granularities can be regular or irregular.
+For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
+On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
+Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
+Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.
 
@@
                             plot_type = "quantile",
+            symmetric = FALSE, 
+            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
+  scale_y_sqrt()
+```
+
+```{r wknd_wday50_df}
+smart_meter10 %>% 
+  prob_plot("month_year",
+                            "day_fortnight",
+                            response = "general_supply_kwh",
+                            plot_type = "quantile",
             symmetric = TRUE, 
             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
-  scale_y_sqrt()
+  scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3))
 ```
@@
 library(tsibble)
-load("data/sm_cust50.RData")
 
@@
     plot_type = "quantile",
-    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
-  ) +
+    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99),
+    symmetric = TRUE) +
   ggtitle("") 
@@
 
-Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
-Thsi is true for both top 50% and 75% of the customers.
+<!-- Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
+This is true for both top 50% and 75% of the customers.
 Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
 All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
-
-```{r dw-hd}
-
-# Hourly usage across days of the week
-sm_cust50 %>%
-  prob_plot("day_week", 
-           "hour_day",
-           "general_supply_kwh", 
-           plot_type = "quantile", 
-          quantile_prob = c(0.25, 0.5, 0.75),
-           response = "general_supply_kwh", overlay = FALSE) +
-  ggtitle("")
-
-```
-
-Figure \ref{fig:hhd-dm} shows distribution of energy consumption across hour hour of the day across different days of the month.
-We can't see much variation across days of week implying that the energy behavior doesn't change depending on which day of the month it is.
-
-```{r hhd-dm}
-# Across days of the week by months
-# sm_cust50 %>%
-#   prob_plot("month_year",
-#     "day_week",
-#     response = "general_supply_kwh", plot_type = "quantile", 
-#           quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)) + ggtitle("") +
-#   ggtitle("Quantile plot across day-of-week by month-of-year")
-
-# Across week of the month and weekends
-
-# sm_cust50 %>%
-#   dynamic_create_gran("day_week") %>%
-#   filter(day_week %in% c("Sat", "Sun")) %>%
-#   prob_plot("day_week", "week_month",
-#            response = "general_supply_kwh",
-#            plot_type = "lv") + 
-#   ggtitle("")
-
-# 
-# sm_cust50 %>%
-#   prob_plot( "hhour_hour","hour_day",
-#            response = "general_supply_kwh",
-#            plot_type = "quantile", overlay = TRUE) + 
-#   ggtitle("")
-
-sm_cust50 %>%
-  prob_plot( "hhour_day","day_month",
-           response = "general_supply_kwh",
-           plot_type = "quantile", overlay = TRUE) + 
-  ggtitle("")
-```
+-->
+
+<!-- ```{r dw-hd} -->
+
+<!-- # Hourly usage across days of the week -->
+<!-- sm_cust50 %>% -->
+<!--   prob_plot("day_week",  -->
+<!--            "hour_day", -->
+<!--            "general_supply_kwh",  -->
+<!--            plot_type = "quantile",  -->
+<!--           quantile_prob = c(0.25, 0.5, 0.75), -->
+<!--            response = "general_supply_kwh", overlay = FALSE) + -->
+<!--   ggtitle("") -->
+
+<!-- ``` -->
 

### 2019-11-11 main.Rmd
inputenc code error sorted and literature review of periodic granularity added

@@
 
-Providing a formalism to these abstractions is important to model a time series across differently grained temporal domains.
+Providing a formalism to some of these abstractions is important to model a time series across differently grained temporal domains.
 
@@
 
-Periodicity is very common in all kinds of data.
-Circular granularities, which are constructed by using two linear granularities, can be utilized to explain such periodicities.
-However, mappings between linear granularities can be regular or irregular.
-For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
-On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Hence, circular granularities can also be "periodic" or "aperiodic" depending on if the mapping is regular or irregular.
-
-Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-
-A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index.
-
-A linear granularity is a mapping of row numbers to subsets of the time domain.
+Relationship between two linear granularities can be expressed in numerous ways. 
+**Definition: Groups Into** : A linear granularity G groups into a linear granularity H, denoted
+$$G \trianglelefteq H,$$
+if for each index j there exists a (possibly infinite) subset S of the integers such
+that
+\begin{equation}
+H(j) = \bigcup_{i \in S}G(i)
+\end{equation}
+
+According to this definition, $$G \trianglelefteq H,$$ if each granule H (i) is the union of some
+granules of G.
+For example, $$Days \trianglelefteq Months,$$ since a week is composed of 7 days.
+This relationship, however, is not sufficient to fully describe a granularity in terms of another one.
+For example, it is clear that $$Days \trianglelefteq Months,$$, since each month is a grouping of a number of days, however it is also a periodical grouping.
+If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
+Considering leap years and all their exceptions, the period becomes 400 years, but the grouping is always periodic.
+In order to define a granularity in terms of another granularity including the notion of periodicity, a particular case of the groups into relationship is considered.
+
+**Definition: Periodical** A granularity H is periodical with respect to a granularity G if
+(1) $$G \trianglelefteq H,$$, and
+(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
+that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
+$H (i + R) = \bigcup_{j \in S} G(j + P)$.
+
+A granularity H which is periodical with respect to G is specified by: (i) the
+R sets of indexes of G S0,...,SR-1 describing the granules of H within one period; (ii) the value of P; (iii) the indexes of first and last granules in H, if their value is not infinite.
+Then, if S0,...,SR-1 are the sets of indexes of G describing
+H (0), . . . , H (R - 1), respectively, then the description of an arbitrary granule H (j ) is given by: $\bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$.
+
+This formula can be explained observing that H (j ) is either one of H (0), .
+.
+.
+, H (R - 1) or, for the periodicity of H, is one of these granules “shifted” ahead or behind on the time line of a finite number of periods.
+If $H (j ) \epsilon {H (0), .
+.
+.
+, H (R - 1)}$, then the formula ‘j mod R’ defines the index (among those in {0,...,R - 1}) of the granule that must be shifted to obtain H (j).The number of periods each granule of G composing H (j mod R) should be shifted is given by $\lfloor j/R \rfloor$.
+
+Many common granularities are in this kind of relationship, for example, Years is
+periodical with respect to both Days and Months.
+
+As mentioned in the introduction, we are also interested in granularities which
+are periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way.
+
+**Definition: Quasi Periodical** A granularity H is quasi-periodical with respect to a granularity G if
+(1) $$G \trianglelefteq H,$$, and
+(2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
+integers R,P, where R is less than the minimum of the number of granules of H
+between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
+
+Intuitively, the definition requires that all granules of H within the span of time
+between two exceptions have the same periodical behavior, characterized by R and P.
+
+Most practical problems seem to require only a granularity system containing a
+set of time granularities which are all periodical (or quasi-periodical) with respect to
+a basic granularity.The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
+
+The formalism of circular granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
+
+A linear granularity is a mapping of the index set to subsets of the time domain.
 For example, if the time index is days, then a linear granularity might be weeks, months or years.
@@
 
-**Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once.
-
-**Definition:** A **circular granule** represents an equivalence class inside each cycle.
+<!-- **Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once. -->
+
+<!-- **Definition:** A **circular granule** represents an equivalence class inside each cycle. -->
 
@@
 
-Computation of time granularities would differ in a periodic and non-periodic set up.
+An algebraic representation for time granularities, which
+we call the calendar algebra. It is assumed that there exists a “bottom”
+granularity known to the system. Calendar algebra operations are designed to generate
+new granularities from the bottom one or recursively, from those already generated.
+Thus, the relationship between the operand(s) and the resulting granularities are encoded
+in the operations.
+
+The calendar algebra consists of two kinds of operations: grouping-oriented and
+granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity, while the
+granule-oriented operations do not change the granules of a granularity, but rather make
+choices of which granules should remain in the new granularity.
+
+- **grouping-oriented** : Let G be a full-integer labeled granularity, and m a positive integer. The grouping
+operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G
+into m-granule groups and making each group a granule of the resulting granularity.
+More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
+each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$. Granularity week is defined by $week = Group_7(day)$. Note that, we assume that the day labeled 1 starts a
+week.
+
+-**granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
+The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
+Since G2 partitions G1, each granule of G1 consists of some contiguous granules of G2.
+The granules of G1 can be partitioned into m-granule groups such that G1(1) to G1(m) are in one group, G1(m + 1) to G1(2m) are in the following group, and so on.
+The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
+For example, if G1 represents 30-day groups (i.e., G1 = Group_30(day)) and we want to add a day (i.e., k = 1) to the 5th one in each group of 12 months (i.e., l = 5 and m = 12), we may have $Alter^{12}_{5,1}(day, G1)$.
+Intuitively, this will make May have 31 days.
+
+<!-- Computation of time granularities would differ in a periodic and non-periodic set up. -->
+
+Now, let us dicuss the computation of circular granularities.
 
@@
 library(ggplot2)
-sm_cust50 %>%
+smart_meter10 %>%
   prob_plot("month_year", "hour_day",

### 2019-11-14 main.Rmd
changing to mapping of graphical element

@@
 library(gravitas)
+library(tsibble)
 opts_chunk$set(
@@
 
-### Circular {#circular-gran-def}
-
-Relationship between two linear granularities can be expressed in numerous ways. 
+#### Relationships between two linear granularities and association to Periodicity
+
+[@Bettini1998-ed] talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
+
 **Definition: Groups Into** : A linear granularity G groups into a linear granularity H, denoted
@@
 granules of G.
-For example, $$Days \trianglelefteq Months,$$ since a week is composed of 7 days.
+For example, $$Days \trianglelefteq Months,$$.
 This relationship, however, is not sufficient to fully describe a granularity in terms of another one.
@@
 
-**Definition: Periodical** A granularity H is periodical with respect to a granularity G if
+**Definition: Groups periodically into** A granularity H is periodical with respect to a granularity G if
 (1) $$G \trianglelefteq H,$$, and
@@
 
-A granularity H which is periodical with respect to G is specified by: (i) the
-R sets of indexes of G S0,...,SR-1 describing the granules of H within one period; (ii) the value of P; (iii) the indexes of first and last granules in H, if their value is not infinite.
-Then, if S0,...,SR-1 are the sets of indexes of G describing
-H (0), . . . , H (R - 1), respectively, then the description of an arbitrary granule H (j ) is given by: $\bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$.
+A granularity H which is periodical with respect to G is specified by:
+(i) the R sets of indexes of G, ${S_0,...,S_(R-1)}$ describing the granules of H within one period;
+(ii) the value of P; 
+(iii) the indexes of first and last granules in H, if their value is not infinite. 
+Then, if ${S_0,...,S_(R-1)}$ are the sets of indexes of G describing
+$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$.
 
@@
 
-Many common granularities are in this kind of relationship, for example, Years is
-periodical with respect to both Days and Months.
-
-As mentioned in the introduction, we are also interested in granularities which
-are periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way.
+Granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
 
@@
 set of time granularities which are all periodical (or quasi-periodical) with respect to
-a basic granularity.The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
-
-The formalism of circular granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+a basic granularity. The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
+
+### Circular {#circular-gran-def}
+
+For mapping it in graphical elements, we need to consider all the periodic granules as similar.
+The formalism  is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
 Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
+
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the time index is days, then a linear granularity might be weeks, months or years.
+Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain.
+We use modular arithmetic for manipulating the index of the periodic granularities so that they are treated as similar across the coarser granularity.
+
+Since the assumption in [@Bettini2000-vy] is that all granularities can be generated from the bottom granularity by calendar algebraic operations - we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time.
+
+Suppose,
+z denotes the index of the tsibble,
+x,y are two units in the hierarchy table with  $order(x) <  order(y)$. 
+Let $f(x, y)$ denotes the accessor function for representing the finer unit x in terms of a coarser unit y and $c(x, y)$ is a constant which relates x and y.
+
+<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
+
+Then $f(x, y)$ can be  computed using modular arithmetic as follows:
+
+$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$
+
+<!-- The formalism of circular granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
+-->
 
@@
 
--**granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
+- **granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
 The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
@@
   +i, x+i+1) - 1)\\
-\end{split}
+\end{split}x
 \end{equation}
@@
 
-# Interaction of time granularities
+# Synergy of time granularities
 
@@
 
-## Effect of interaction
+## Effect of synergy
 
@@
 In other words, if there are levels of time granularities plotted across x-axis which are not spanned by levels of the time granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
-We hypothesize that the interaction of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
+We hypothesize that the synergy of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
 
@@
 
-Th application is not only restricted to temporal data.
+The application is not only restricted to temporal data.
 We provide an example of cricket to illustrate how this can be generalised in other applications.
@@
 Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls.
-We have sourced the the ball by ball data for IPL season 2008 to 2016 from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The dataset contains the information on batting team, bowling team,  balls of the over,  over of the innings, innings of the match and total runs per ball for 577 matches spanning over 9 seasons (2008 to 2016).
-It also has information on which team won, winning margin and several other useful information.
-
-A hierarchy like table \ref{tab:hierarchy} can be construed for this game format, where index here would refer to ball as the data set contains ball-by-ball data.
+A hierarchy like table \ref{tab:hierarchy} can be construed for this game format.
+
+The ball by ball data for IPL season 2008 to 2016 from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
+The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data cross overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016).
 
 ```{r hierarchy, echo=FALSE}
-hierarchy_model <- tibble::tibble(units = c("index", "ball", "over", "inning", "match"), 
-                                  convert_fct = c(1, 6, 20, 2, 1))
+hierarchy_model <- tibble::tibble(units = c("ball", "over", "inning", "match"), 
+                                  convert_fct = c(6, 20, 2, 1))
 knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket")
+```
+
+```{r readdata}
+library(gravitas)
+library(tibble)
+glimpse(cricket)
+```
+
+Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, where each over, which represents an ordering from past to future, can form the index of the tsibble.
+The hierarchy table would look like the following:
+
+```{r hierarchy2}
+
+hierarchy_model <- tibble::tibble(
+  units = c("index", "over", "inning", "match"),
+  convert_fct = c(1, 20, 2, 1))
+
+knitr::kable(hierarchy_model, format = "markdown")
+
 ```
@@
 Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final 4 from 2010 to 2015.
-It would be interesting to see the difference in their strategies throughout all matches in the two seasons.
+It would be interesting to observe their strategies throughout all matches in the two seasons.
 The following two questions might help us partially understand their strategies.
 
-- Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first?
-
-- Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings?
+<!-- - Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first? -->
+
+<!-- - Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings? -->
+
+Q1: How run rates vary depending on if a team bats first or second?
+Q2: How number of wickets and dot balls in the previous over affect the runs in the subsequent over across different overs of the innings?
 
 ```{r runs_per-over}
-cricket <-  read_csv("data-raw/deliveries_all.csv")
-matches <- read_csv("data-raw/matches_all.csv")
-
-cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id"))
+library(tsibble)
+
+cricket_all <-  read_csv("data-raw/deliveries_all.csv")
+matches_all <- read_csv("data-raw/matches_all.csv")
+
+cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
 
@@
   cricket_tsibble_all %>% 
-  prob_plot("over", "season",
+  prob_plot("over", 
+            "season",
             hierarchy_model,
@@
             plot_type = "quantile",
-                                quantile_prob = c(0.25, 0.5, 0.75), overlay = FALSE
-
-   )  +
+            quantile_prob = c(0.25, 0.5, 0.75), 
+            symmetric = FALSE)  +
      theme( # remove the vertical grid lines
@@
             # explicitly set the horizontal lines (or they will disappear too)
-            panel.grid.major.y = element_blank()
-     ) + scale_x_discrete(breaks = seq(2008, 2016, 3))
-```
-
-```{r violin, fig.cap= "Violin and box plots of run rate across innings of the match and matches won or lost for Mumbai Indians(top) and Chennai Super Kings(bottom).The violins for CSK has long right tails in 2nd innings for cases when they lost or won.
-MI has long right tail only when they won the match."}
-
-matches_select <- matches %>% select(id, toss_winner, winner, win_by_runs)
-
-cricket_winner <- cricket_tsibble_all %>% 
-  left_join(matches_select, by = c("match_id" = "id"))
-
-MI_winner <- cricket_winner %>% 
-             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians",
-                    season %in% c(2013, 2015)) %>% 
-             mutate(win = if_else(winner == "Mumbai Indians", 1, 0))
-
-MI_winner %>% 
-  gran_obs("win", "inning", hierarchy_model)
-
-CSK_winner <- cricket_winner %>% 
-             filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings",
-                    season %in% c(2013, 2015)) %>% 
-             mutate(win = if_else(winner == "Chennai Super Kings", 1, 0))
-
-CSK_winner %>% 
-  gran_obs("win", "inning", hierarchy_model)
-
- MI_winner %>% 
-            prob_plot( "win","inning",
+            panel.grid.major.y = element_blank()) + scale_x_discrete(breaks = seq(2008, 2016, 3))
+```
+
+From figure \autoref{cricex}, it can be observed that there is no clear upward shift in runs in the second innings as compared to the first innings.
+The variability of runs also increases as the teams approach towards the end of the innings, as observed through the longer and more distinct letter values.
+
+```{r cricex, fig.cap= "Letter value plot of runs per over across overs of the inning faceted by innings of the match.
+No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning."}
+
+library(tsibble)
+cricket_tsibble <- cricket %>%
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+
+cricket_tsibble %>%
+  filter(batting_team %in% c("Mumbai Indians",
+                             "Chennai Super Kings"))%>%
+  prob_plot("inning", 
+            "over",
+  response = "runs_per_over",
   hierarchy_model,
-                                response = "run_rate",
-                                plot_type = "violin",
-                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-                     overlay = FALSE) + ggtitle("") + 
-    geom_boxplot(width = 0.1, color = "blue")
-
- CSK_winner %>% filter(inning %in% c(1,2)) %>% 
-           prob_plot( "win","inning",
-                               hierarchy_model,
-                               response = "run_rate",
-                               plot_type = "violin",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-                    overlay = FALSE) + ggtitle("") + 
-   geom_boxplot(width = 0.1, color = "blue")
-```
-
-```{r}
-cricket_winner %>% as_tibble() %>% 
-             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians",
-                    winner!= "Mumbai Indians",
-                    win_by_runs!=0,
-                    season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary()
-
-cricket_winner %>% as_tibble() %>% 
-             filter(batting_team ==  "Chennai Super Kings",
-                    winner!= "Chennai Super Kings",
-                    win_by_runs!=0,
-                    inning ==2, 
-                    season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary()
-```
+  plot_type = "lv")
+```
+
+<!-- ```{r violin, fig.cap= "Violin and box plots of run rate across innings of the match and matches won or lost for Mumbai Indians(top) and Chennai Super Kings(bottom).The violins for CSK has long right tails in 2nd innings for cases when they lost or won.
+MI has long right tail only when they won the match."} -->
+
+<!-- matches_select <- matches %>% select(id, toss_winner, winner, win_by_runs) -->
+
+<!-- cricket_winner <- cricket_tsibble_all %>% -->
+<!--   left_join(matches_select, by = c("match_id" = "id")) -->
+
+<!-- MI_winner <- cricket_winner %>% -->
+<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
+<!--                     season %in% c(2013, 2015)) %>% -->
+<!--              mutate(win = if_else(winner == "Mumbai Indians", 1, 0)) -->
+
+<!-- MI_winner %>% -->
+<!--   gran_obs("win", "inning", hierarchy_model) -->
+
+<!-- CSK_winner <- cricket_winner %>% -->
+<!--              filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings", -->
+<!--                     season %in% c(2013, 2015)) %>% -->
+<!--              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0)) -->
+
+<!-- CSK_winner %>% -->
+<!--   gran_obs("win", "inning", hierarchy_model) -->
+
+<!--  MI_winner %>% -->
+<!--             prob_plot( "win","inning", -->
+<!--                                 hierarchy_model, -->
+<!--                                 response = "run_rate", -->
+<!--                                 plot_type = "violin", -->
+<!--                                 quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
+<!--                      overlay = FALSE) + ggtitle("") + -->
+<!--     geom_boxplot(width = 0.1, color = "blue") -->
+
+<!--  CSK_winner %>% filter(inning %in% c(1,2)) %>% -->
+<!--            prob_plot( "win","inning", -->
+<!--                                hierarchy_model, -->
+<!--                                response = "run_rate", -->
+<!--                                plot_type = "violin", -->
+<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
+<!--                     overlay = FALSE) + ggtitle("") + -->
+<!--    geom_boxplot(width = 0.1, color = "blue") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- cricket_winner %>% as_tibble() %>% -->
+<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
+<!--                     winner!= "Mumbai Indians", -->
+<!--                     win_by_runs!=0, -->
+<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
+
+<!-- cricket_winner %>% as_tibble() %>% -->
+<!--              filter(batting_team ==  "Chennai Super Kings", -->
+<!--                     winner!= "Chennai Super Kings", -->
+<!--                     win_by_runs!=0, -->
+<!--                     inning ==2, -->
+<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
+<!-- ``` -->
 
@@
 
-Figure \ref{fig:violin} shows that for CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
+<!-- Figure \ref{fig:violin} shows that for CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
 Run rate can go up to 6 runs per over depending on the situation.
@@
 However, we can delve deeper and try to see which overs are the most volatile for CSK and if it is different to that of MI?
-
-In Figure \ref{fig:lv}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
+-->
+
+<!-- In Figure \ref{fig:lv}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
 It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
 There is an upward shift of the distribution in the second part of the innings (after 10th over) and more variability (longer and distinct letter values) as well.
-CSK has been more uniform in their approach maintaining consistent scores per over.
+-->
+<!-- CSK has been more uniform in their approach maintaining consistent scores per over.
 Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}.
+-->
 
@@
 ```{r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."}
-
-MI <- cricket_tsibble_all %>% 
-  filter(season %in% c(2010:2015), batting_team=="Mumbai Indians")
-
-CSK <- cricket_tsibble_all %>% 
-  filter(season %in% c(2010:2015),
-         batting_team=="Kolkata Knight Riders",
-         inning %in% c(1,2))
-
- MI %>% 
-   gran_obs("inning", "over", hierarchy_model) %>%
-   knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
-
- CSK %>% 
-   gran_obs("inning", "over", hierarchy_model) %>%
-   knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
-
-MI %>% 
-  prob_plot("inning", "over",
-                               hierarchy_model,
-                               response = "run_rate",
-                               plot_type = "lv",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")
-
-CSK %>% 
-  prob_plot("inning", "over",
-                               hierarchy_model,
-                               response = "run_rate",
-                               plot_type = "lv",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
-
-cricket_tsibble_all %>% distinct(batting_team)
-
-```
-
-Q3: Is run rate set to reduce in overs where fielding is good?
+# 
+# 
+# 
+# MI <- cricket_tsibble_all %>%
+#   filter(season %in% c(2010:2015), batting_team=="Mumbai Indians")
+# 
+# 
+# CSK <- cricket_tsibble_all %>%
+#   filter(season %in% c(2010:2015),
+#          batting_team=="Kolkata Knight Riders",
+#          inning %in% c(1,2))
+# 
+# 
+#  MI %>%
+#    gran_obs("inning", "over", hierarchy_model) %>%
+#    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
+# 
+#  CSK %>%
+#    gran_obs("inning", "over", hierarchy_model) %>%
+#    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
+# 
+# MI %>%
+#   prob_plot("inning", "over",
+#                                hierarchy_model,
+#                                response = "run_rate",
+#                                plot_type = "lv",
+#                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")
+# 
+# 
+# CSK %>%
+#   prob_plot("inning", "over",
+#                                hierarchy_model,
+#                                response = "run_rate",
+#                                plot_type = "lv",
+#                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
+# 
+# cricket_tsibble_all %>% distinct(batting_team)
+
+```
+
+Q3: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalises runs in the subsequent over more?
 
@@
 Difference in run rates should be negative if fielding is good.
-Let us see if this fact is true.Thus, Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over.
+Let us see if this fact is true.Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over.
 
@@
 cricket_tsibble_fld %>% 
-  filter(fielding_wckts %in% c(0,1,2)) %>% 
+  filter(fielding_wckts %in% c(0,1)) %>% 
   prob_plot( "fielding_wckts","over",
@@
 
-```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."}
-
-cricket_dot <- cricket_season %>% 
-  mutate(wicket = if_else(is.na(dismissal_kind),0,1),
-         dot_balls = if_else(total_runs==0,1, 0)
-         ) %>% 
+```{r crickread}
+cricket_tsibble <- gravitas::cricket %>%
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+```
+
+```{r exdotplot, echo=FALSE}
+library(gravitas)
+library(tsibble)
+
+cricket_bowling <- cricket_season %>% 
+  mutate(dot_ball_proxy = if_else(total_runs==0,1,0)) %>% 
   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs),
-            wicket = sum(wicket),
-            dot_balls = sum(dot_balls))
-
-cricket_tsibble_dot <- cricket_dot %>% 
+  summarise(runs_per_over = sum(total_runs),
+            run_rate = sum(total_runs)/length(total_runs),
+            dot_balls = sum(dot_ball_proxy)) %>%
+            mutate(lagged_run_rate = c(0,diff(run_rate)))
+
+cricket_tsibble_bow <- cricket_bowling %>% 
   ungroup() %>% 
@@
 
-#cricket_tsibble_dot %>% gran_obs("over", "fielding_wckts")
-
-cricket_tsibble_dot  %>% filter(dot_balls %in% c(1,2,3,4)) %>% 
-  prob_plot("over","dot_balls",
+cricket_tsibble_bow %>% gran_obs("over", "dot_balls")
+
+cricket_tsibble_bow %>% 
+  filter(dot_balls %in% c(0, 1, 2))%>% 
+  prob_plot("dot_balls",
+            "over",
             hierarchy_model,
-                               response = "run_rate",
+            response = "lagged_run_rate",
             plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-           overlay = FALSE)  + ggtitle("")
-```
+            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
+
+```
+
+```{r exwicketplot}
+cricket_tsibble %>% 
+  filter(wicket %in% c(0, 1)) %>%
+  mutate(lag_wicket = lag(wicket)) %>% 
+  filter(lag_wicket %in% c(0, 1)) %>%
+  prob_plot("lag_wicket",
+            "over",
+            hierarchy_model,
+            response = "runs_per_over",
+            plot_type = "quantile",
+            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
+
+```
+
+<!-- # ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."} -->
+<!-- #  -->
+<!-- # cricket_dot <- cricket_season %>%  -->
+<!-- #   mutate(wicket = if_else(is.na(dismissal_kind),0,1), -->
+<!-- #          dot_balls = if_else(total_runs==0,1, 0) -->
+<!-- #          ) %>%  -->
+<!-- #   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
+<!-- #   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs), -->
+<!-- #             wicket = sum(wicket), -->
+<!-- #             dot_balls = sum(dot_balls))  -->
+<!-- #  -->
+<!-- #  -->
+<!-- #  -->
+<!-- # cricket_tsibble_dot <- cricket_dot %>%  -->
+<!-- #   ungroup() %>%  -->
+<!-- #   mutate(data_index = row_number()) %>%  -->
+<!-- #   as_tsibble(index = data_index) -->
+<!-- #  -->
+<!-- # #cricket_tsibble_dot %>% gran_obs("over", "fielding_wckts") -->
+<!-- #  -->
+<!-- # cricket_tsibble_dot  %>%  -->
+<!-- #   filter(dot_balls %in% c(1,2,3,4)) %>%  -->
+<!-- #   mutate(lag_dot_ball = lag(dot_balls)) -->
+<!-- #   prob_plot("over","dot_balls", -->
+<!-- #                                hierarchy_model, -->
+<!-- #                                response = "run_rate", -->
+<!-- #                                plot_type = "quantile", -->
+<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
+<!-- #            overlay = FALSE)  + ggtitle("") -->
+<!-- # ``` -->
+<!-- #  -->
+<!-- #  -->
 

### 2019-11-14 main.Rmd
changing structure

@@
 
-#### Relationships between two linear granularities and association to Periodicity
+### Relationships between two linear granularities and association to Periodicity
 
@@
 
+### Computation of time granularities
+
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf]. It is assumed that there exists a “bottom”
+granularity and Calendar algebra operations are designed to generate
+new granularities from the bottom one or recursively, from those already generated.
+Thus, the relationship between the operand(s) and the resulting granularities are encoded
+in the operations.
+
+The calendar algebra consists of two kinds of operations: grouping-oriented and
+granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity, while the
+granule-oriented operations do not change the granules of a granularity, but rather make
+choices of which granules should remain in the new granularity.
+
+- **grouping-oriented** : Let G be a full-integer labeled granularity, and m a positive integer. The grouping
+operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G
+into m-granule groups and making each group a granule of the resulting granularity.
+More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
+each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$. Granularity week is defined by $week = Group_7(day)$. Note that, we assume that the day labeled 1 starts a week.
+
+- **granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
+The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
+Since G2 partitions G1, each granule of G1 consists of some contiguous granules of G2.
+The granules of G1 can be partitioned into m-granule groups such that G1(1) to G1(m) are in one group, G1(m + 1) to G1(2m) are in the following group, and so on.
+The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
+For example, if G1 represents 30-day groups (i.e., G1 = Group_30(day)) and we want to add a day (i.e., k = 1) to the 5th one in each group of 12 months (i.e., l = 5 and m = 12), we may have $Alter^{12}_{5,1}(day, G1)$.
+Intuitively, this will make May have 31 days.
+
 ### Circular {#circular-gran-def}
 
-For mapping it in graphical elements, we need to consider all the periodic granules as similar.
-The formalism  is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+In a cyclic organization of time, the domain is composed of a set of recurring time values.
+The formalism of time granularities in cyclic arrangement is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
 Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
@@
 For example, if the time index is days, then a linear granularity might be weeks, months or years.
-Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain.
-We use modular arithmetic for manipulating the index of the periodic granularities so that they are treated as similar across the coarser granularity.
-
-Since the assumption in [@Bettini2000-vy] is that all granularities can be generated from the bottom granularity by calendar algebraic operations - we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time.
+Since time is arranged in a cyclic way typically to accomodate periodicity, we will try to formalize how periodic linear granularities can be represented in a cyclic way.
+Since the assumption in [@Bettini2000-vy] is that all granularities can be generated from the bottom granularity by calendar algebraic operations - we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time using modular arithmetic.
+
+<!-- Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain. . -->
 
@@
 
+The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
+
+Some examples of circular representation of finer granularities within coarse granularities in a  periodic and quasi-periodic set up are as follows:
+
 <!-- The formalism of circular granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
@@
 
-A linear granularity is a mapping of the index set to subsets of the time domain.
+<!-- A linear granularity is a mapping of the index set to subsets of the time domain.
 For example, if the time index is days, then a linear granularity might be weeks, months or years.
 Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain.
-We use modular arithmetic to define circular granularity and use the following definitions:
+We use modular arithmetic to define circular granularity and use the following definitions: -->
 
@@
 In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up.
-
-# Computation of time granularities
-
-An algebraic representation for time granularities, which
-we call the calendar algebra. It is assumed that there exists a “bottom”
-granularity known to the system. Calendar algebra operations are designed to generate
-new granularities from the bottom one or recursively, from those already generated.
-Thus, the relationship between the operand(s) and the resulting granularities are encoded
-in the operations.
-
-The calendar algebra consists of two kinds of operations: grouping-oriented and
-granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity, while the
-granule-oriented operations do not change the granules of a granularity, but rather make
-choices of which granules should remain in the new granularity.
-
-- **grouping-oriented** : Let G be a full-integer labeled granularity, and m a positive integer. The grouping
-operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G
-into m-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
-each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$. Granularity week is defined by $week = Group_7(day)$. Note that, we assume that the day labeled 1 starts a
-week.
-
-- **granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
-The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-Since G2 partitions G1, each granule of G1 consists of some contiguous granules of G2.
-The granules of G1 can be partitioned into m-granule groups such that G1(1) to G1(m) are in one group, G1(m + 1) to G1(2m) are in the following group, and so on.
-The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
-For example, if G1 represents 30-day groups (i.e., G1 = Group_30(day)) and we want to add a day (i.e., k = 1) to the 5th one in each group of 12 months (i.e., l = 5 and m = 12), we may have $Alter^{12}_{5,1}(day, G1)$.
-Intuitively, this will make May have 31 days.
 

### 2019-11-17 main.Rmd
cyclic time change

@@
 
+### Cyclic {#circular-gran-def}
+
+In a cyclic organization of time, the domain is composed of a set of recurring time values.
+Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday).
+Intutively, we need to manipulate the indices of periodic linear granularities so that it leads to repetitive patterns.
+The formalism of cyclic time granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the time index is days, then a linear granularity might be weeks, months or years.
+
+<!-- Since time is arranged in a cyclic way typically to accomodate periodicity, we will try to formalize how periodic linear granularities can be represented in a cyclic way. -->
+
+The assumption in [@Bettini2000-vy] is that all linear granularities can be generated from the bottom granularity by calendar algebraic operations.
+Hence we assume that a bottom granularity exists and is represented by the index of the tsibble.
+
+Suppose,
+z denotes the index of the tsibble. Also represents the bottom granularity.
+Let G be a periodic linear granularity, not equal to the bottom granularity, with each granule of G made up of $const(z, G)$ many granules of the bottom granularity.
+C be the cyclic representation of the linear time granularity in terms of the bottom granularity.
+
+Then, $$C(z, G) = z \mod c(z,G)$$
+
+Now, suppose we want a cyclic representation of two periodic granularities.
+
+$$C(G, H) = \lfloor z/const(z,G) \rfloor\mod const(G,H)$$
+
+C1 and C2 are two cyclic granularities in the hierarchy table with  $order(C1) <  order(C2)$. 
+Let $f(C_1, C_2)$ denotes the accessor function for representing the finer unit C1 in terms of a coarser unit C2 and $const(C1, C2)$ is a constant which relates C1 and C2.
+
+<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
+
+Then $f(x, y)$ can be  computed using modular arithmetic as follows:
+
+$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$
+
+The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
+
+Some examples of circular representation of finer granularities within coarse granularities in a  periodic and quasi-periodic set up are as follows:
+
+we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time using modular arithmetic.
+
+Extending from the idea of linear granulariites, where 
+$H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$ 
+can be re-written as,
+$H (j) = \bigcup_{i \in S_j \mod R}G((P/R) *(j - j \mod R) + i)$
+
+The index in the right hand side looks like an equivalence class with modulus P/R.  
+The value of $j\mod R$ is j for j< R and 1, 2, ...  for j>R.
+
+Here, i runs from 1 to no. of granules of G and j runs from 1 to number of granules of H.
+Hence, $j*\lfloor P/R \rfloor + i$.
+
+For j = 1, we have ($\lfloor P/R \rfloor + i$)
+For j = 2,  we have ($2*\lfloor P/R \rfloor + i$)
+
+$H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$
+
+<!-- Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain. . -->
+
 ### Computation of time granularities
@@
 Intuitively, this will make May have 31 days.
-
-### Circular {#circular-gran-def}
-
-In a cyclic organization of time, the domain is composed of a set of recurring time values.
-The formalism of time granularities in cyclic arrangement is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
-Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
-
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-Since time is arranged in a cyclic way typically to accomodate periodicity, we will try to formalize how periodic linear granularities can be represented in a cyclic way.
-Since the assumption in [@Bettini2000-vy] is that all granularities can be generated from the bottom granularity by calendar algebraic operations - we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time using modular arithmetic.
-
-<!-- Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain. . -->
-
-Suppose,
-z denotes the index of the tsibble,
-x,y are two units in the hierarchy table with  $order(x) <  order(y)$. 
-Let $f(x, y)$ denotes the accessor function for representing the finer unit x in terms of a coarser unit y and $c(x, y)$ is a constant which relates x and y.
-
-<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
-
-Then $f(x, y)$ can be  computed using modular arithmetic as follows:
-
-$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$
-
-The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
-
-Some examples of circular representation of finer granularities within coarse granularities in a  periodic and quasi-periodic set up are as follows:
 

### 2019-11-17 main.Rmd
harmony and clash section revised

@@
 
-Now, suppose we want a cyclic representation of two periodic granularities.
+Now, suppose we want a cyclic representation of two periodic granularities G and H, where H is periodic with respect to G.
 
@@
 
-# Synergy of time granularities
+# Synergy of cyclic time granularities
 
@@
 To define a general framework, we define harmony and clashes.
-Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in a non-compatible ways and hinder the process of exploration of the dependent variable if plotted together.
+
+<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
+-->
 
@@
 
-Suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps row numbers to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
-That is, let $S_{ij}$ be the set of row numbers such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
-The graphs that don't work are those where many of these 12 sets are empty.
-Now, many situations can lead to any of these sets being empty.
+Suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
+That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
+Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.Now, many situations can lead to any of these sets being empty.
 Let us discuss the following cases, where one or more of these sets can be empty.
@@
 For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
 These are structurally empty sets in that it is impossible for them to have any observations.
+
+<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
 
@@
 All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
-
-Therefore, combinations of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are called **harmonies** as they promote the exploratory analysis through visualization.
+
+Therefore, pair of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
+And the ones that do not lead to any missing combinations are called **harmonies**.
+
+<!-- as they promote the exploratory analysis through visualization. -->
 

### 2019-11-18 main.Rmd
distribution section review

@@
 
+<!-- Include why distribution plot, why 2D, why facetting in introduction -->
 <!-- Why distribution plot -->
 <!-- Huge data difficult to summaries using one statistic-->
-With huge amount of data being available, mean, median or any one summary statistic might not enough to understand a data set. Soon enough following questions become more interesting:
-
- - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?
-
- - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
+
+<!-- With huge amount of data being available, mean, median or any one summary statistic might not enough to understand a data set. Soon enough following questions become more interesting:   -->
+
+<!--  - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?  -->
+
+<!--  - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
 More generally, how the variation is the data set changes across different percentiles/deciles?
-
-  - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median?
-
-This is when displaying a probability distribution becomes a potentially useful approach.
-The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
+-->
+
+<!--   - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median? -->
+
+<!-- This is when displaying a probability distribution becomes a potentially useful approach. -->
+<!-- The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
 These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
+-->
 
@@
 <!-- Why 2D -->
-Graphical output devices are flat.
+
+<!-- Graphical output devices are flat.
 If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
@@
 Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
+-->
 
 <!-- Why facet-ing -->
-`ggplot2` (Wickham, 2016) is an excellent tool that facilitates the process of mapping different variables to a 2D frame through grammar of graphics.
+
+<!-- `ggplot2` (Wickham, 2016) is an excellent tool that facilitates the process of mapping different variables to a 2D frame through grammar of graphics.
 One aspect of ggplot2 is faceting, inspired by Trellis plots (Cleveland, 1993).
@@
 This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
+-->
 
 <!-- What are we visualizing -->
-We will see the distribution of the time series across bivariate temporal granularities through faceting approach with one temporal granularity plotted along the x-axis, the other one across facets and the dependent time series variable on the y-axis.
-
-Now, different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
+
+<!-- We will see the distribution of the time series across bivariate temporal granularities through faceting approach with one temporal granularity plotted along the x-axis, the other one across facets and the dependent time series variable on the y-axis.
+-->
+
+<!-- A well designed visual representation can replace cognitive calculations with simple perpetual inferences and improve speed of comprehension and decision making.
+However, creating a visualisation requires a number of nuanced judgements.
+The challenge being given a data set, the number of visual encodings and hence the space of possible visual design is extremely huge.
+-->
+
+The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk].
+The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, allows us to produce graphics using the same structured thinking that is used to design analysis.
+Briefly, any graphics are made up of **Layers**, **Facet**,**Scale** and a  **coordinate system**.**Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
+
+Creating a visualisation requires a number of nuanced judgements.
+However, we exploit some features of layers and faceting to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate temporal granularities.
+The general framework of visualizing the distribution invloves a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
+Different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
 We will look at each of these aspects separately and analyze the effect of each on visualization.
+While we accomplish that, we need to bear in mind that any customised species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
 
@@
 
-<!-- Lit review on statistical distribution plot-->
-
-<!-- Description and lit review of boxplot -->
+<!-- Common benefits and challenges of empirical and density based visualization -->
+
 There are several ways to plot statistical distributions.
-The displayed probabilities in these plots are either computed using kernel density estimation methods or emperical statistical methods.
-We will discuss the benefits and challenges of few conventional and recent ways to plot distributions using both of these methods.
-
-### Emperical methods
-
-Most commonly used techniques to display distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+The displayed probabilities in these plots are either computed using kernel density estimation methods or empirical statistical methods.
+The latter do not allow us to see the shape, skewness or nature of the tail or multi-modality with so much clarity as the former.
+However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
+The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
+Density plots which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution, unlike discretizing the distribution and showing only parts of it.
+They can be useful in spotting multimodality, however, they might become obscuring with too many categories and information on the entire distribution can add to cognitive load.
+Also, the probabilties in density plot are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
+As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
+The density based visualizations are subject to the same sample size restrictions and challenges that apply to any density estimation.
+In practice, the densities are estimated reasonably with at least 30 observations.
+Even with sample sizes of several hundred, however, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
+
+We will discuss few conventional and recent ways to plot distributions using both of these methods.
+
+### Empirical methods
+
+Most commonly used techniques to display a distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
 The box plot is a compact distributional summary, displaying less detail than a histogram.
@@
 Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
-However, we do not have an idea if the distribution is multimodal.
 Moreover, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
@@
 <!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.It shows only actual data values and no smoothed summaries.
-However, the letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles.
-These might lead to a lot of letter values being shown and leading to overload of information in one plot.
-Also, multimodality is difficult to spot.
+The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+“outliers” in letter value plots are those observations beyond the most extreme letter value.
+The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown and leading to overload of information in one plot.
 
@@
 Quantile plots visually portray the quantiles of the distribution of sample data.
-Much like the quartiles divide the data set equally into four equal parts, extensions might include dividing the data set even further.
-These plots are referred to as "quantile" plots with, where number of quantiles 9 for deciles and 99 for percentiles.
-No distributional assumptions are made about the data since empirical deciles are plotted.
-It avoids much clutter and just enable us to focus on specific probabilities.
-These plots do not allow us to see the shape, skewness or nature of the tail of the distribution with so much clarity as violin plots, ridge plots or letter value plots.
-Also, a large data set is required for the extreme percentiles to be estimated with any accuracy.
+Much like the quartiles divide the data set equally into four equal parts, extensions include dividing the data set even further.
+These plots are referred to as "quantile" plots.
+For example, number of quantiles would be 9 for a decile plot and 99 for percentile plot.
+It avoids much clutter and just enable us to focus on specific probabilities, typically representing typical and extreme behaviors.
+A large data set is required for the extreme percentiles to be estimated with any accuracy.
+These plots can display any quantiles instead of quartiles in a traditional boxplot.
+When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g: outside 1.5 times the interquartile range above the upper quartile and below the lower quartile).
+However, outliers are open to interpretation and not shown in an quantile plot.
 
@@
 
-A density plot which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution, unlike discretizing the distribution and showing only parts of it.
-
-Violin plots [@Hintze1998-zi, 1998] adds the information available from local density estimates to summary statistics provided by box plots.
+Traditional ways to visualize densities include violin plots [@Hintze1998-zi, 1998].
 The shape of the violin represents the density estimate of the variable.
 The more data points in a specific range, the larger the violin is for that range.
-Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories.
-Howevr, the density in violin are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
-As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
+Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
 
@@
 
-As a general rule, the quantile plots are really useful for picking patterns, whereas, other more involved methods of plotting are useful for studying anomalies or outlier behavior.
-It is good to be aware of the benefits and challenges while choosing any distribution plots, depending on the context.
+It is good to be conversant with the benefits and challenges while choosing any distribution plot, depending on the context.
+As a general rule, if we have too many categories, the quantile plots are useful for camparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
 

### 2019-11-18 main.Rmd
levels of granularities iumpact on visualisation

@@
 We will look at each of these aspects separately and analyze the effect of each on visualization.
+
 While we accomplish that, we need to bear in mind that any customised species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
+We are trying to address problems where even with good aesthetic qualities and good data, the graph might be confusing or misleading because of how people process the information.
 
@@
 
-Choice of plots depend on the levels(very high/high/medium/low) of the two granularities plotted.
-The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.Space and resolution become a problem if the number of levels of facet variables is too high.
+The levels (very high/high/medium/low) of the two granularities plotted might have an impact on the choice of plots.
+The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.
+Space and resolution become a problem if the number of levels of facet variables is too high.
 Hence, plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.
@@
 
-Space and cognitive power both are constrained if the number of levels of facet variables is very high.
-So as a general rule, we should avoid plotting a temporal granularity with very high levels across facets.
-
- - If levels of both the granularities are low, then any distribution plots might be chosen depending on which feature of the distribution who is interested to look at.
+ <!-- - If levels of both the granularities are low, then any statistical distribution plot might be chosen depending on which feature of the distribution one is interested to look at.  -->
+
+  - If level of the granularity plotted across x-axis is low or medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus, irrespective of the number of the levels of the facet variable.
 
@@
 
- - If level of the granularity plotted across x-axis is low, then any distribution plots might be chosen depending on which feature of the distribution who is interested to look at, irrespective of the number of the levels of the facet variable.
-
   - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.
+
+  - - If levels of granularities plotted are medium, empirical based methods of statistical distribution are preferred as they use less space by design than most density based methods.
 

### 2019-11-18 main.Rmd
synergy of time granularities

@@
 
-The levels (very high/high/medium/low) of the two granularities plotted might have an impact on the choice of plots.
+The levels of the two granularities plotted might have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
 The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.
-Space and resolution become a problem if the number of levels of facet variables is too high.
-Hence, plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.
+Plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.
 
@@
  <!-- - If levels of both the granularities are low, then any statistical distribution plot might be chosen depending on which feature of the distribution one is interested to look at.  -->
+
+The following principles are useful while choosing distribution plots given two temporal granularities.
 
@@
 ## Effect of synergy
+In Section \ref{sec:harmonyclash}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are build.
+In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.
 
@@
 
-We should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
-
-```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "h", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
+Hence we redefine harmony and clashes as follows : harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
+Hence, we should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
+
+```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
 (a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.

### 2019-11-18 main.Rmd
revision till smart meter

@@
 The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, allows us to produce graphics using the same structured thinking that is used to design analysis.
-Briefly, any graphics are made up of **Layers**, **Facet**,**Scale** and a  **coordinate system**.**Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
+Briefly, any graphics are made up of **Layers**, **Facet**,**Scale** and a  **coordinate system**.
+**Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
 
@@
 The displayed probabilities in these plots are either computed using kernel density estimation methods or empirical statistical methods.
-The latter do not allow us to see the shape, skewness or nature of the tail or multi-modality with so much clarity as the former.
+The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
 However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
@@
 
-It is good to be conversant with the benefits and challenges while choosing any distribution plot, depending on the context.
+Given a context, it is good to be conversant with the benefits and challenges while choosing a distribution plot.
 As a general rule, if we have too many categories, the quantile plots are useful for camparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
 
-### Levels of temporal granularities
-
-The levels of the two granularities plotted might have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
+### Effect of levels of temporal granularities
+
+The levels of the two granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
 The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.
@@
 
-  - If level of the granularity plotted across x-axis is low or medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus, irrespective of the number of the levels of the facet variable.
+  - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus.
 
@@
 
-  - - If levels of granularities plotted are medium, empirical based methods of statistical distribution are preferred as they use less space by design than most density based methods.
+  - If levels of any granularity plotted are more than medium, empirical based methods of distribution visualizing are preferred as they use less space by design than most density based methods.
 
@@
 
-## Effect of synergy
+## Effect of synergy of time granularities
+
 In Section \ref{sec:harmonyclash}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are build.
@@
 
-Hence we redefine harmony and clashes as follows : harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-Hence, we should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
+We redefine harmony and clashes as follows: harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
+As a result, we should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
 
@@
 
-## Effect of number of observation
+## Effect of number of observations
+
+Even with harmonies, visualizing probability distributions can be misleading either due to rarely occuring categories or unevenly distributed events.
 
@@
 
-Even with harmonies, problems might occur due to rarely occuring categories.
-
- Suppose we have $T$ hourly observations, and $C_1$ with $n_1$ categories and $C_2$ with $n_1$ categories.
-Each element of $C_1$ occurs approximately $T/n_1$ times while each element of $C_2$ occurs approximately $T/n_2$ times.
-There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(n_1n_2)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge n_1n_2k$, the visualization will be acceptable.
+Suppose we have $T$ observations, and two cyclic granularities -  $C_1$ with $n$ categories and $C_2$ with $m$ categories.
+Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
+There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
+If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
 The value of $k$ will depend on what type of visualization we are producing.
-For emperical estimates, even $k=10$ may be acceptable, but for density estimates, we probably need $k \ge 30$.
+For a decile plot, even $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
+Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
 
@@
 
-Let us take a specific example to illustrate this. Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
-
-### Unevenly distributed events
-
-Even when there are no rarely occuring events, when number of observations are not evenly distributed between different categories of the time granularities.Large differences between number of observations between and across facets, might have an impact on the estimation of dfistribution.
+### Unevenly distributed events (TODO)
+
+Even when there are no rarely occuring events, number of observations might vary hugely across within or across each panel.
+This might happen due to missing observations in the data or uneven locations of events in time domain.
 We use Gini's measure to compute if differences in number of observations across or within facets are significantly different.
@@
 
-<!-- <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions --> -->
+<!-- <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions --> 
 <!-- It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
@@
 
-In Figure \ref{fig:day_fortnight} and \ref{fig:day_fortnight_agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy beahvior of these two customers.
-In Figure \ref{fig:day_fortnight_agg} it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
-However, Figure \ref{fig:day_fortnight} adds to the beahvior by looking at the distribution.
+<!-- In Figure \ref{fig:day-fortnight} and \ref{fig:day-fortnight-agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy beahvior of these two customers.
+-->
+
+In Figure \ref{fig:day_fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy beahvior of these two customers.
+it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
+However, there is additional information which we can derive looking at the distribution.
 If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
@@
 
-```{r day_fortnight_agg}
+```{r day-fortnight_agg, echo = FALSE}
 library(lvplot)
@@
 
-While trying to explore the energy behavior of these customers systematically, the first thing we should have at our disposal for examining periodicities of energy behavior across time granularities is to know the number of time granularities we can look at exhaustively.
+While trying to explore the energy behavior of these customers systematically across time granularities, the first thing we should have at our disposal is to know the number of time granularities we can look at exhaustively.
 If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
@@
 
-The interval of this tsibble is 30 minutes, and hence the default for search_gran in this case, provides temporal granularities ranging from half-hour to year.
+The interval of this tsibble is 30 minutes, and hence the default in this case, provides temporal granularities ranging from half-hour to year.
 If these options are considered too many, the default options can be modified to limit the possibilities.
@@
 )
+
+smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
+
 ```
@@
 
-Now, we would like to see how these customers' behavior relate to the rest of the 
-50 households across these two measures -
-
-- if the regular behavior is regular and extreme behavior extreme or vice versa
-- if the weekend and weekday behavior are different for most of these households.
-
-Figure \ref{fig:wknd_wday50} shows the overlay quantile plot of energy consumption of 10 households.
+Now, we would like to see how these customers' behavior relate to the rest of the  50 households across these two measures -
+
+- if energy distribution is skewed towards extreme?
+
+- if the weekend and weekday behavior are different for most of these households?
+
+Figure \autoref{fig:wknd-day50} shows the quantile plot of energy consumption of 50 households.
+Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends.
+
+Figure \autoref{fig:my-hd} shows the area quantile plots of energy consumption across hours of the day faceted by months of the year.
 The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-It can be observed that the median is very closer to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-Weekend and weekday behavior is not very different for these ten customers, implying the first customer is more typical than the second customer when compared to these ten customers.
-
-```{r wknd_wday50}
-smart_meter10 %>% 
+It can be observed that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
+Moreover, only the pink band changes significantly in winter months (May - August).
+The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.
+Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
+It is to be noted here that offcourse the level of quantiles increased too in winter months, the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+
+```{r wknd-wday50}
+smart_meter <- 
+  read_rds("data/sm_cust50.rds")
+
+smart_meter %>% 
   prob_plot("wknd_wday",
@@
 
-```{r wknd_wday50_df}
-smart_meter10 %>% 
-  prob_plot("month_year",
-                            "day_fortnight",
-                            response = "general_supply_kwh",
-                            plot_type = "quantile",
-            symmetric = TRUE, 
-            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
-  scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3))
-```
+<!-- # ```{r wknd-wday50_df} -->
+<!-- # smart_meter10 %>%  -->
+<!-- #   prob_plot("month_year", -->
+<!-- #                             "day_fortnight", -->
+<!-- #                             response = "general_supply_kwh", -->
+<!-- #                             plot_type = "quantile", -->
+<!-- #             symmetric = TRUE,  -->
+<!-- #             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
+<!-- #   scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3)) -->
+<!-- # ``` -->
 
@@
 library(ggplot2)
-smart_meter10 %>%
+smart_meter %>%
   prob_plot("month_year", "hour_day",
@@
     symmetric = TRUE) +
-  ggtitle("") 
+  ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3))
 ```

### 2019-11-20 main.Rmd
cricket application update

@@
 
-## Single-order-up granularities
+### Single-order-up granularities
 
@@
 
-## Multiple order-up granularities
+### Multiple order-up granularities
 
@@
 
-# Case studies
+# Applications
 
@@
 
-The ball by ball data for IPL season 2008 to 2016 from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
+The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
 The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data cross overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016).
@@
 Q1: How run rates vary depending on if a team bats first or second?
+
 Q2: How number of wickets and dot balls in the previous over affect the runs in the subsequent over across different overs of the innings?
@@
 The number of dot balls is reflective of the quality of bowling in the game.
-Run rate of an over should ideally decrease if the number of dot balls increase or the batsman can likely to go for big shots and get dismissed.
+Run rate of an over should ideally decrease if the number of dot balls increase.
+However, what is the effect of dot balls in previous over on run rate in the subsequent over.
+Will players  batsman likely to go for big shots because they couldn't score good runs in the previous over?
+Or they should play consistently and avoid scoring high?
+
 Figure \ref{fig:dot} shows that for any over, increase in dot balls lead to decrease in run rate.
@@
 
-```{r exdotplot, echo=FALSE}
+```{r exdotplot, echo=FALSE, eval = FALSE}
 library(gravitas)
@@
             dot_balls = sum(dot_ball_proxy)) %>%
-            mutate(lagged_run_rate = c(0,diff(run_rate)))
+            mutate(diff_run_rate = c(0,diff(run_rate)))
 
@@
             hierarchy_model,
-            response = "lagged_run_rate",
+            response = "diff_run_rate",
             plot_type = "quantile",
@@
 ```{r exwicketplot}
+
+cricket_tsibble %>% 
+  filter(dot_balls %in% c(0, 1, 2)) %>%
+  mutate(lag_dot = lag(dot_balls)) %>% 
+  filter(lag_dot %in% c(0, 1, 2)) %>%
+  prob_plot("lag_dot",
+            "over",
+            hierarchy_model,
+            response = "runs_per_over",
+            plot_type = "quantile",
+            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
+
 cricket_tsibble %>% 
@@
 
-{r read_data, echo=FALSE}
+<!-- {r read_data, echo=FALSE} -->
 

### 2019-11-20 main.Rmd
definition of circular time granularities

@@
 Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday).
-Intutively, we need to manipulate the indices of periodic linear granularities so that it leads to repetitive patterns.
+Intutively, the indices of periodic linear granularities needs to be manipulated so that it leads to repetitive patterns.
+
 The formalism of cyclic time granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
@@
 z denotes the index of the tsibble. Also represents the bottom granularity.
-Let G be a periodic linear granularity, not equal to the bottom granularity, with each granule of G made up of $const(z, G)$ many granules of the bottom granularity.
-C be the cyclic representation of the linear time granularity in terms of the bottom granularity.
-
-Then, $$C(z, G) = z \mod c(z,G)$$
+Let G be a periodic linear granularity, not equal to the bottom granularity, with k(BG,G) as the period length of the bottom granularity with respect to linear granularity G.
+C(BG, G) is the circular granularitiy relating linear granularity G to the bottom granularity. Then,  C(BG, G) is given as follows: 
+Then, $$C_{BG, G}(z) = G(z\mod k(BG,G))$$
 

### 2019-11-20 main.Rmd
circular granularity definition update

@@
 Suppose,
-z denotes the index of the tsibble. Also represents the bottom granularity.
-Let G be a periodic linear granularity, not equal to the bottom granularity, with k(BG,G) as the period length of the bottom granularity with respect to linear granularity G.
+z denotes the index set.
+BG denotes the index of the tsibble. Also represents the bottom granularity.
+Let G be a linear granularity, periodic with respect to BG.
+k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.
 C(BG, G) is the circular granularitiy relating linear granularity G to the bottom granularity. Then,  C(BG, G) is given as follows: 
-Then, $$C_{BG, G}(z) = G(z\mod k(BG,G))$$
+Then, $$C_{BG, G}(z) = G(z\mod k(BG,G))$$.
+
+For example, suppose the bottom granularity day and circular granularity day-of-week are represented by D and $$C_{day, week}$$ respectively, then we will have:
+
+$$C_{day, week}(z) = D(z\mod 7)$$.
+
+Since, the period length of days with respect to weeks is 7.
+
+All circular granularities can be expressed in terms of the bottom linear granularity with equation.
 

### 2019-11-20 main.Rmd
Relationship between circular granularities

@@
 All circular granularities can be expressed in terms of the bottom linear granularity with equation.
+
+Relationship between circular granularities:
+
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as
+
+$$C_{(G, H)}(z) = G(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$$
+
+where, linear granularity H is periodic with respect to  linear granularity G and k(G, H) represents the period length of G within H.
+
+If linear granularities are not provided or bottom granularity is not known, any circular granularities of order m, provided the circular granularities till order (m-1) are provided.
 

### 2019-11-21 main.Rmd
circular granuarity discussion after definition update

--- 2019-11-20 main.Rmd
@@
 
+**Definition: Finer Than** : A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
+for each index i, there exists an index j such that $G(i) \subset H(j)$.
+
 **Definition: Groups Into** : A linear granularity G groups into a linear granularity H, denoted
@@
 
-Intuitively, the definition requires that all granules of H within the span of time
-between two exceptions have the same periodical behavior, characterized by R and P.
-
-Most practical problems seem to require only a granularity system containing a
-set of time granularities which are all periodical (or quasi-periodical) with respect to
-a basic granularity. The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
+Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by R and P.
+
+Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
+The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
 
@@
 Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday).
-Intutively, the indices of periodic linear granularities needs to be manipulated so that it leads to repetitive patterns.
+Intutively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorisation of time.
 
 The formalism of cyclic time granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
-Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
+A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble.
 A linear granularity is a mapping of the index set to subsets of the time domain.
 For example, if the time index is days, then a linear granularity might be weeks, months or years.
-
-<!-- Since time is arranged in a cyclic way typically to accomodate periodicity, we will try to formalize how periodic linear granularities can be represented in a cyclic way. -->
-
-The assumption in [@Bettini2000-vy] is that all linear granularities can be generated from the bottom granularity by calendar algebraic operations.
-Hence we assume that a bottom granularity exists and is represented by the index of the tsibble.
+Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
 
 Suppose,  
-z denotes the index of the tsibble. Also represents the bottom granularity.
-Let G be a periodic linear granularity, not equal to the bottom granularity, with k(BG,G) as the period length of the bottom granularity with respect to linear granularity G.
-C(BG, G) is the circular granularitiy relating linear granularity G to the bottom granularity. Then,  C(BG, G) is given as follows: 
-Then, $$C_{BG, G}(z) = G(z\mod k(BG,G))$$
-
-Now, suppose we want a cyclic representation of two periodic granularities G and H, where H is periodic with respect to G.
-
-$$C(G, H) = \lfloor z/const(z,G) \rfloor\mod const(G,H)$$
+z denotes the index set.  
+BG denotes the index of the tsibble (bottom granularity).
+G be a linear granularity, periodic with respect to BG.  
+k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.
+
+**Definition: Circular granularity** A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if
+
+\begin{equation} \label{eq2}
+$$C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+$$
+\end{equation}, such that 
+$BG(z) \neq \phi$.
+
+**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram given below.
+Each granule is represented by a box.
+The diagram also illustrate that the granules that overlap share elements from the underlying time domain.
+The first slot in the diagram shows the index set and the bottom granularity.
+The index of the tsibble considered is days.
+G and H are two linear granularities denoting days and weeks respectively.
+The period length of days with respect to weeks is 7, hence,
+$$C_{G, BG}(z)$$ represented by C(BG, G) is given by $$BG(z\mod 7)$$.
+The circular granularity C(BG, G) consists of repetitive pattern {BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)} which repeats itself after each period length.
+Similarly, C(BG, H) is given by $$BG(z\mod 14)$$ since the period length of days with respect to fortnight is 14.
+C(BG, H) also repeats the pattern ${BG(0), BG(1), \dots, BG(13)}$ in every period.
+Hence, circular granularities are represented as repetitive categorisation of time in the diagram.
+
+All circular granularities can be expressed in terms of the bottom linear granularity with equation \ref{eq2}
+
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as
+
+$$C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$$,
+where, linear granularity H is periodic with respect to  linear granularity G and k(G, H) represents the period length of G within H.
+
+### Computation of circular granularities
+
+The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
+We shall use the notion of a hierarchy table to define them.
+
+Definition:  **order** of a linear granularity in a hierarchy table H:$(G, l, m)$ is defined as the level of deconstruction of a linear granularity.
+In a hierarchy table, linear granularities are arranged from highest to lowest order of graininess.
+Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities provided in the hierarchy table.
+
+Let H: $(G, l, m)$ be a hierarchy model containing m linear granularities.
+$$G_{i}$$ represents the linear granularity of order i and $$l_{i}$$å represents the period length of $$G_{i}$$ with respect to $$G_{i+1}$$.
+
+If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
+
+\begin{equation} \label{eq1}
+\begin{split}
+C_{(x,y)}(z) & = C_{(x,x+1)}(z) + k(x, x+1)(C_{(x+1,y)}(z)-1) \\  
+  & =   C_{(x,x+1)}(z) + k(x, x+1)[ C_{(x+1,x+2)}(z) + c(x+1, x+2)( C_{(x+2, y)}(z) - 1) - 1]\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
+  &\vdots\\
+  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
+  +i, x+i+1) - 1)\\
+\end{split}
+\end{equation}
 
@@
 
-Extending from the idea of linear granulariites, where 
-$H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$ 
-can be re-written as,
-$H (j) = \bigcup_{i \in S_j \mod R}G((P/R) *(j - j \mod R) + i)$
-
-The index in the right hand side looks like an equivalence class with modulus P/R.  
-The value of $j\mod R$ is j for j< R and 1, 2, ...  for j>R.
-
-Here, i runs from 1 to no. of granules of G and j runs from 1 to number of granules of H.
-Hence, $j*\lfloor P/R \rfloor + i$.
-
-For j = 1, we have ($\lfloor P/R \rfloor + i$)
-For j = 2,  we have ($2*\lfloor P/R \rfloor + i$)
-
-$H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$
+<!-- Extending from the idea of linear granulariites, where  -->
+<!-- $H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$  -->
+<!-- can be re-written as, -->
+<!-- $H (j) = \bigcup_{i \in S_j \mod R}G((P/R) *(j - j \mod R) + i)$ -->
+
+<!-- The index in the right hand side looks like an equivalence class with modulus P/R.   -->
+<!-- The value of $j\mod R$ is j for j< R and 1, 2, ...  for j>R.   -->
+
+<!-- Here, i runs from 1 to no. of granules of G and j runs from 1 to number of granules of H. -->
+<!-- Hence, $j*\lfloor P/R \rfloor + i$. -->
+
+<!-- For j = 1, we have ($\lfloor P/R \rfloor + i$) -->
+<!-- For j = 2,  we have ($2*\lfloor P/R \rfloor + i$) -->
+
+<!-- $H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$  -->
 

### 2019-11-21 main.Rmd
circular granularity section done

@@
 
-### Relationships between two linear granularities and association to Periodicity
+#### Relationships between two linear granularities and association to Periodicity
 
@@
 
+**Definition: Bottom granularity** Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
+
+**Example:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $$\trianglelefteq $$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $$\trianglelefteq$$.
+
 Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
@@
 
-### Cyclic {#circular-gran-def}
+### Circular {#circular-gran-def}
 
@@
 Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday).
-Intutively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorisation of time.
+Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
 
@@
 
-Suppose,  
+#### Circular {#circular-gran-def}
+
+**Definition: Circular granularity** A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if
+
+\begin{equation} \label{eq2}
+C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
+BG(z) \neq \phi
+\end{equation}
+where,  
 z denotes the index set.  
@@
 
-**Definition: Circular granularity** A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if
-
-\begin{equation} \label{eq2}
-$$C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+$$
-\end{equation}, such that 
-$BG(z) \neq \phi$.
-
-**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram given below.
+**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram \autoref{fig:circular-gran}.
 Each granule is represented by a box.
@@
 G and H are two linear granularities denoting days and weeks respectively.
-The period length of days with respect to weeks is 7, hence,
-$$C_{G, BG}(z)$$ represented by C(BG, G) is given by $$BG(z\mod 7)$$.
+The period length of days with respect to weeks is 7, hence, $C_{G, BG}(z)$ represented by C(BG, G) is given by $BG(z\mod 7)$.
 The circular granularity C(BG, G) consists of repetitive pattern {BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)} which repeats itself after each period length.
-Similarly, C(BG, H) is given by $$BG(z\mod 14)$$ since the period length of days with respect to fortnight is 14.
+Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14.
 C(BG, H) also repeats the pattern ${BG(0), BG(1), \dots, BG(13)}$ in every period.
-Hence, circular granularities are represented as repetitive categorisation of time in the diagram.
-
-All circular granularities can be expressed in terms of the bottom linear granularity with equation \ref{eq2}
-
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as
-
-$$C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$$,
-where, linear granularity H is periodic with respect to  linear granularity G and k(G, H) represents the period length of G within H.
-
-### Computation of circular granularities
-
-The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
-We shall use the notion of a hierarchy table to define them.
-
-Definition:  **order** of a linear granularity in a hierarchy table H:$(G, l, m)$ is defined as the level of deconstruction of a linear granularity.
-In a hierarchy table, linear granularities are arranged from highest to lowest order of graininess.
-Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities provided in the hierarchy table.
-
-Let H: $(G, l, m)$ be a hierarchy model containing m linear granularities.
-$$G_{i}$$ represents the linear granularity of order i and $$l_{i}$$å represents the period length of $$G_{i}$$ with respect to $$G_{i+1}$$.
-
-If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
-
-\begin{equation} \label{eq1}
-\begin{split}
-C_{(x,y)}(z) & = C_{(x,x+1)}(z) + k(x, x+1)(C_{(x+1,y)}(z)-1) \\  
-  & =   C_{(x,x+1)}(z) + k(x, x+1)[ C_{(x+1,x+2)}(z) + c(x+1, x+2)( C_{(x+2, y)}(z) - 1) - 1]\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
-  &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
-  +i, x+i+1) - 1)\\
-\end{split}
-\end{equation}
+Hence, circular granularities are represented as repetitive categorization of time in the diagram.
+
+```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="\\label{fig:circular-gran} Relating circular granularities and bottom linear granularity "}
+knitr::include_graphics("Figs/circular-gran-diagram.png")
+```
+
+All circular granularities can be expressed in terms of the bottom linear granularity with equation \ref{eq2}.
+
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the period length of G within H.
+
+Table \ref{tab:definitions} shows representation of circular granularities relating two linear granularities.
+It is possible that none of these two linear granularities are bottom granularities.
+But the representation of the resultant circular granularity will be a function of the bottom granularity.
+For example, suppose the bottom granularity is minutes, and let $k_i$ is the period length of $C_i$.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lr@{~}lr@{~}r}
+\toprule
+Minute-of-Hour: & $C_1(z)$ & $= z \mod 60$ & $k_1 =$&$60$ \\
+Minute-of-Day: & $C_2(z)$ & $= z \mod 60*24$ & $k_2=$&$`r 60*24`$\\
+Hour-of-Day: & $C_4(z)$ & $= \lfloor z/60\rfloor\mod 24$  & $k_3 =$&$24$ \\
+Hour-of-Week: & $C_6(z)$ & $= \lfloor z/60\rfloor\mod 24*7$  & $k_4=$&$`r 24*7`$\\
+Day-of-Week: & $C_7(z)$ & $= \lfloor z/24*60\rfloor \mod 7$ & $k_5=$&$7$\\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Illustrative circular granularities with time index in minutes}
+\label{definitions}
+\end{table}
+
+### Aperiodic {#aperiodic-gran-def}
 
@@
 
-<!-- They also accomodate periodicity and hence the smaller subsets of bottom granularity are repeating themselves within the larger subsets of bottom granularity.  -->
+<!-- They also accommodate periodicity and hence the smaller subsets of bottom granularity are repeating themselves within the larger subsets of bottom granularity.  -->
 
@@
 The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
-We shall use the notion of a hierarchy table to define them.
-
-Consider a **hierarchy table** to be consisting of two columns:
-
- -  The first column represents the (temporal) units in ascending order of hierarchy.
-Any two units, which are essentially the linear granularities, can join together to form a circular/aperiodic granularity.
-
- - The second column represents the relationship between subsequent (temporal) units.
-For periodic granularities, this could be represented by a constant, whereas, for a aperiodic one, there is no single constant which can be used to represent their relationship.
-
-and, **order**, is defined as the position of the units in the hierarchical table.
-
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concering a single level as **single-order-up** granularities.
+We shall use the notion of a hierarchy table and order to define them.
+
+**Definition:  Order** of a linear granularity can be comprehended as the level graininess asscociated with a linear granularity.
+If we consider two linear granularities G and H, such that $G \preceq H$, then H is of higher order than G.
+
+**Notation:  Hierarchy table** Let H: $(G, C, k, n)$ be a hierarchy model containing m linear granularities.
+$G_{i}$ represents the linear granularity of order i and $k_{i}$å represents the period length of $$G_{i}$$ with respect to $$G_{i+1}$$ and $C_{l,m}$ represents the circular granularity that relates linear granularity of order l and m, with $\forall l,m \in n, l<m$.
+
+In a hierarchy table, linear granularities are arranged from highest to lowest order of linear granularities.
+
+<!-- Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities and the other granularities are subsequently more "coarse" than the earlier one.
+-->
+
+We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities.
 Let us look at few calendars to see examples of single and multiple order-up granularities.
+
+<!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
+-->
+
+### Computation of multiple-order from single-order
+
+\begin{equation} \label{eq1}
+\begin{split}
+C_{(x,y)}(z) & = C_{(x,x+1)}(z) + k(x, x+1)(C_{(x+1,y)}(z)-1) \\  
+  & =   C_{(x,x+1)}(z) + k(x, x+1)[ C_{(x+1,x+2)}(z) + c(x+1, x+2)( C_{(x+2, y)}(z) - 1) - 1]\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
+  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
+  &\vdots\\
+  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
+  +i, x+i+1) - 1)\\
+\end{split}
+\end{equation}
 
@@
 
-Now, let us dicuss the computation of circular granularities.
+Now, let us discuss the computation of circular granularities.
 
@@
 Computation of multiple-order-up granularities will differ if the units in the hierarchy table are periodic or aperiodic.
-We split the method of computation into two cases - one concerining all periodic single-order-up granularities and the second with mixed single-order-up granularities.
+We split the method of computation into two cases - one concerning all periodic single-order-up granularities and the second with mixed single-order-up granularities.
 
@@
 
-<!-- Include why distribution plot, why 2D, why facetting in introduction -->
+<!-- Include why distribution plot, why 2D, why faceting in introduction -->
 <!-- Why distribution plot -->
@@
 
-While we accomplish that, we need to bear in mind that any customised species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
+While we accomplish that, we need to bear in mind that any customized species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
 We are trying to address problems where even with good aesthetic qualities and good data, the graph might be confusing or misleading because of how people process the information.
@@
 They can be useful in spotting multimodality, however, they might become obscuring with too many categories and information on the entire distribution can add to cognitive load.
-Also, the probabilties in density plot are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
+Also, the probabilities in density plot are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
 As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
@@
 Given a context, it is good to be conversant with the benefits and challenges while choosing a distribution plot.
-As a general rule, if we have too many categories, the quantile plots are useful for camparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
+As a general rule, if we have too many categories, the quantile plots are useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
 
@@
 
-<!-- Facetting creates small multiples of data each representing a different subset of the data.
+<!-- Faceting creates small multiples of data each representing a different subset of the data.
 Small multiples are powerful for exploratory data analysis where we can rapidly compare patterns in different parts of the data and see whether they are the same or different.
@@
 
-<!-- In Figure \ref{fig:day-fortnight} and \ref{fig:day-fortnight-agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy beahvior of these two customers.
--->
-
-In Figure \ref{fig:day_fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy beahvior of these two customers.
+<!-- In Figure \ref{fig:day-fortnight} and \ref{fig:day-fortnight-agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
+-->
+
+In Figure \ref{fig:day_fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
@@
 Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that offcourse the level of quantiles increased too in winter months, the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+It is to be noted here that off course the level of quantiles increased too in winter months, the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
 
@@
 The application is not only restricted to temporal data.
-We provide an example of cricket to illustrate how this can be generalised in other applications.
+We provide an example of cricket to illustrate how this can be generalized in other applications.
 The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
@@
 
-<!-- How the scores per over vary across granularities/categorisations like innings of a match or matches of a season?
+<!-- How the scores per over vary across granularities/categorizations like innings of a match or matches of a season?
 Are these different for the winning teams in that season and the teams that couldn't qualify for the playoffs?
@@
 
-<!-- Figure \ref{fig:violin} shows that for CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
+<!-- Figure \ref{fig:violin} shows that for CSK, irrespective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
 Run rate can go up to 6 runs per over depending on the situation.
@@
 <!-- whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second.
-CSK's run rate distibution is extremely right tailed in the 2nd innings, but not in the first.
+CSK's run rate distribution is extremely right tailed in the 2nd innings, but not in the first.
 .
@@
 
-Q3: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalises runs in the subsequent over more?
+Q3: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalizes runs in the subsequent over more?
 

### 2019-11-21 main.Rmd
rough structure for circular done - trash not removed

@@
 
-**Notation:  Hierarchy table** Let H: $(G, C, k, n)$ be a hierarchy model containing m linear granularities.
-$G_{i}$ represents the linear granularity of order i and $k_{i}$å represents the period length of $$G_{i}$$ with respect to $$G_{i+1}$$ and $C_{l,m}$ represents the circular granularity that relates linear granularity of order l and m, with $\forall l,m \in n, l<m$.
+**Notation:  Hierarchy table** Let $H_n: (G, C, k)$ be a hierarchy model containing n linear granularities.
+$G_{l}$ represents the linear granularity of order l and $k_{l,m}$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
 
@@
 
-\begin{equation} \label{eq1}
+\begin{equation} \label{eq3}
 \begin{split}
-C_{(x,y)}(z) & = C_{(x,x+1)}(z) + k(x, x+1)(C_{(x+1,y)}(z)-1) \\  
-  & =   C_{(x,x+1)}(z) + k(x, x+1)[ C_{(x+1,x+2)}(z) + c(x+1, x+2)( C_{(x+2, y)}(z) - 1) - 1]\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
+C_{(G_l,G_m)}(z) & = C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1) \\  
+  & =   C_{(G_l,G_(l+1)}(z) + k(l, l+1)[ C_{G_(l+1),G_(l+2)}(z) + k(l+1, l+2)( C_{G_(l+2), G_m)}(z) - 1) - 1]\\
+  & =  C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{G_(l+1),G_(l+2)}(z) - 1) + k(l, l+1)k(l+1, l+2)(C(l+2, m) - 1)\\
+  & =  C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{G_(l+1),G_(l+2)}(z) - 1) + k(l, l+2)(C_{G_(l+2),G_(l+m)}(z) - 1)\\
   &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
-  +i, x+i+1) - 1)\\
+  & = \sum_{i=0}^{order(m) - order(l) - 1} k(l, l+i)(C_{G_(l+i),G_(l+i+1)}(z) - 1)\\
 \end{split}
@@
 <!-- some discussions on primary units of calendar -->
-So far we have used the Gregorian calendar as it is the most widely used calendar.
+**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar.
 But it is far from being the only one.
@@
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-Any civil day is divided into 24 hours and each hour into 60 minutes and each minute into 60 seconds.
-There are exceptions where in London, for example, length of each "hour" varies from about 39 minutes in December to 83 minutes in June.
 The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
@@
 
-| Units	      | Conversion factor
-|------------	|-------------------	|
-|  kin      	| 20             	    |
-|  uinal    	| 18            	    |
-|  tun      	| 20           	      |
-|  katun    	| 20            	    |
-|  baktun   	| 1          	        |
-
-Examples of multiple-order-up  granularities can be kin of the tun or kin of the baktun whereas examples of single-order-up granularities may include kin of the uinal, uinal of the tun etc.
+|  G	        |C                 |k
+|------------	|------------------|------------|
+|  kin      	|kin-of-uinal      |  20        |
+|  uinal    	|uinal-of-tun      |  18  	    |
+|  tun      	|tun-of-katun      |  20	      |
+|  katun    	|katun-of-baktun   |  20        |
+|  baktun   	| 1                |	1         |
+
+Examples of multiple-order-up  granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.
+
+Let us use the equation \ref{eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.
+
+\begin{equation} \label{eq4}
+\begin{split}
+C(uinal, baktun) & = C(uinal, tun) + kuinal, tun)C(tun,katun) + C(uinal, katun)C(katun, baktun) \\
+              & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\
+\end{split}
+\end{equation}
+
+### Computation of single-order from multiple-order
+
+For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
+
+\begin{equation} \label{eq5}
+C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
+\end{equation}
+
+**Example:** Considering the same example of Mayan Calendar,  it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq5}
+
+\begin{equation} \label{eq6}
+\begin{split}
+C(tun, katun) =  \lfloor C(uinal, baktun)(z)/18\rfloor \mod 20 \\
+\end{split}
+\end{equation}
 

### 2019-11-21 main.Rmd
clean circular gran section

@@
 
-C1 and C2 are two cyclic granularities in the hierarchy table with  $order(C1) <  order(C2)$. 
-Let $f(C_1, C_2)$ denotes the accessor function for representing the finer unit C1 in terms of a coarser unit C2 and $const(C1, C2)$ is a constant which relates C1 and C2.
-
-<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
-
-Then $f(x, y)$ can be  computed using modular arithmetic as follows:
-
-$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$
-
-The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
-
-Some examples of circular representation of finer granularities within coarse granularities in a  periodic and quasi-periodic set up are as follows:
-
-we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time using modular arithmetic.
-
-<!-- Extending from the idea of linear granulariites, where  -->
-<!-- $H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$  -->
-<!-- can be re-written as, -->
-<!-- $H (j) = \bigcup_{i \in S_j \mod R}G((P/R) *(j - j \mod R) + i)$ -->
-
-<!-- The index in the right hand side looks like an equivalence class with modulus P/R.   -->
-<!-- The value of $j\mod R$ is j for j< R and 1, 2, ...  for j>R.   -->
-
-<!-- Here, i runs from 1 to no. of granules of G and j runs from 1 to number of granules of H. -->
-<!-- Hence, $j*\lfloor P/R \rfloor + i$. -->
-
-<!-- For j = 1, we have ($\lfloor P/R \rfloor + i$) -->
-<!-- For j = 2,  we have ($2*\lfloor P/R \rfloor + i$) -->
-
-<!-- $H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$  -->
-
-<!-- Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain. . -->
-
-### Computation of time granularities
-
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf]. It is assumed that there exists a “bottom”
-granularity and Calendar algebra operations are designed to generate
-new granularities from the bottom one or recursively, from those already generated.
-Thus, the relationship between the operand(s) and the resulting granularities are encoded
-in the operations.
-
-The calendar algebra consists of two kinds of operations: grouping-oriented and
-granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity, while the
-granule-oriented operations do not change the granules of a granularity, but rather make
-choices of which granules should remain in the new granularity.
-
-- **grouping-oriented** : Let G be a full-integer labeled granularity, and m a positive integer. The grouping
-operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G
-into m-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
-each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$. Granularity week is defined by $week = Group_7(day)$. Note that, we assume that the day labeled 1 starts a week.
-
-- **granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
-The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-Since G2 partitions G1, each granule of G1 consists of some contiguous granules of G2.
-The granules of G1 can be partitioned into m-granule groups such that G1(1) to G1(m) are in one group, G1(m + 1) to G1(2m) are in the following group, and so on.
-The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
-For example, if G1 represents 30-day groups (i.e., G1 = Group_30(day)) and we want to add a day (i.e., k = 1) to the 5th one in each group of 12 months (i.e., l = 5 and m = 12), we may have $Alter^{12}_{5,1}(day, G1)$.
-Intuitively, this will make May have 31 days.
-
-<!-- The formalism of circular granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
-Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.
--->
-
-<!-- A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of time domain.
-We use modular arithmetic to define circular granularity and use the following definitions: -->
-
-<!-- A linear granularity, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to subsets of linear bottom granularity.
-Here, the domain is the index set and range contains subsets of linear bottom granularity.
-Circular granularity, joining two linear granularities, therefore is a mapping between these two subsets of linear bottom granularity.
-We use modular arithmetic to define circular granularity and use the following definitions: -->
-
-<!-- They also accommodate periodicity and hence the smaller subsets of bottom granularity are repeating themselves within the larger subsets of bottom granularity.  -->
-
-<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity. -->
-
-<!-- \begin{table}[ht] -->
-<!-- \begin{center} -->
-<!-- \begin{tabular}{lr@{~}lr@{~}r} -->
-<!-- \toprule -->
-<!-- MOH: & $C_1(s)$ & $= s \mod 60$ & $n_1 =$&$60$ \\ -->
-<!-- MOD: & $C_2(s)$ & $= s \mod `r 60*24`$ & $n_2=$&$`r 60*24`$\\ -->
-
-<!-- HOD: & $C_4(s)$ & $= \lfloor s/60\rfloor\mod 24$  & $n_4 =$&$24$ \\ -->
-<!-- HOW: & $C_6(s)$ & $= \lfloor s/60\rfloor\mod 24*7$  & $n_5=$&$`r 24*7`$\\ -->
-
-<!-- DOW: & $C_7(s)$ & $= \lfloor s/24*60\rfloor \mod 7$ & $n_6=$&$7$\\ -->
-<!-- \bottomrule -->
-<!-- \end{tabular} -->
-<!-- \end{center} -->
-<!-- \caption{Illustrative circular granularities with time index in minutes} -->
-<!-- \label{definitions} -->
-<!-- \end{table} -->
-
-<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
-Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
--->
-
-<!-- Formal definition of circular granularity -->
-
-**Definition: Equivalence class** Let $m \in N \backslash{0}$.
-For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
-
-The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$.
-Thus $Z_m = \{[0], [1], ..., [m-1]\}$.
-However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.
-
-**Definition:** A **circular granularity** $C$, joining two linear granularity with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that
-
-$C(s) = (s\mod m)$ for $s \in Z$.
-
-For example, suppose $C$ is a circular granularity denoting hour-of-day and we have hourly data for 100 hours.
-The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
-
-<!-- **Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once. -->
-
-<!-- **Definition:** A **circular granule** represents an equivalence class inside each cycle. -->
-
-### Aperiodic {#aperiodic-gran-def}
-<!-- Formal definition of aperiodic circular granularity -->
-
 **Definition:** An **Aperiodic circular granularity** can not be defined using modular arithmetic in a similar fashion.
@@
 
+C1 and C2 are two cyclic granularities in the hierarchy table with  $order(C1) <  order(C2)$. 
+Let $f(C_1, C_2)$ denotes the accessor function for representing the finer unit C1 in terms of a coarser unit C2 and $const(C1, C2)$ is a constant which relates C1 and C2.
+
+<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
+
+Then $f(x, y)$ can be  computed using modular arithmetic as follows:
+
+$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$
+
+The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
+
+<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
+Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
+-->
+
+<!-- Formal definition of circular granularity -->
+
+<!-- **Definition: Equivalence class** Let $m \in N \backslash{0}$.
+For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
+-->
+
+<!-- The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$.
+Thus $Z_m = \{[0], [1], ..., [m-1]\}$.
+However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.
+-->
+
+<!-- **Definition:** A **circular granularity** $C$, joining two linear granularity with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that -->
+
+<!-- $C(s) = (s\mod m)$ for $s \in Z$. -->
+
+<!-- For example, suppose $C$ is a circular granularity denoting hour-of-day and we have hourly data for 100 hours.
+The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
+-->
+
+<!-- **Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once. -->
+
+<!-- **Definition:** A **circular granule** represents an equivalence class inside each cycle. -->
+
+<!-- ### Aperiodic {#aperiodic-gran-def} -->
+<!-- Formal definition of aperiodic circular granularity -->
+
  <!-- Why we need single order up and multiple order up granularities -->
@@
 
-In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up.
+<!-- In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up. -->
+
+<!-- ### Computation of time granularities -->
+
+<!-- Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf].
+It is assumed that there exists a “bottom” -->
+<!-- granularity and Calendar algebra operations are designed to generate -->
+<!-- new granularities from the bottom one or recursively, from those already generated. -->
+<!-- Thus, the relationship between the operand(s) and the resulting granularities are encoded -->
+<!-- in the operations. -->
+
+<!-- The calendar algebra consists of two kinds of operations: grouping-oriented and -->
+<!-- granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity, while the -->
+<!-- granule-oriented operations do not change the granules of a granularity, but rather make -->
+<!-- choices of which granules should remain in the new granularity. -->
+
+<!-- - **grouping-oriented** : Let G be a full-integer labeled granularity, and m a positive integer. The grouping -->
+<!-- operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G -->
+<!-- into m-granule groups and making each group a granule of the resulting granularity. -->
+<!-- More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for -->
+<!-- each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$. Granularity week is defined by $week = Group_7(day)$. Note that, we assume that the day labeled 1 starts a week. -->
+
+<!-- - **granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
+The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
+Since G2 partitions G1, each granule of G1 consists of some contiguous granules of G2.
+The granules of G1 can be partitioned into m-granule groups such that G1(1) to G1(m) are in one group, G1(m + 1) to G1(2m) are in the following group, and so on.
+The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
+For example, if G1 represents 30-day groups (i.e., G1 = Group_30(day)) and we want to add a day (i.e., k = 1) to the 5th one in each group of 12 months (i.e., l = 5 and m = 12), we may have $Alter^{12}_{5,1}(day, G1)$.
+Intuitively, this will make May have 31 days.
+-->
 
@@
 
-Now, let us discuss the computation of circular granularities.
-
-### Single-order-up granularities
-
-Suppose, z denotes the index of the tsibble, x,y are two units in the hierarchy table  with  $order(x) <  order(y)$.
-Let $f(x, y)$ denotes the accessor function for computing the single-order-up granularity relating x and y and $c(x, y)$ is a constant which relates x and y.
-
-<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
-
-Then $f(x, y)$ can be  computed using modular arithmetic as follows:
-
-$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$ where $y = x+1$
-
-See table[] for an illustration of computing single order up granularities for Mayan calendar.
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-\textbf{Single-order-up granularities}
-
-$kin\_uinal:   = z \mod 20$  \\
-$uinal\_tun:   = \lfloor z/20\rfloor \mod 18 $\\
-$tun\_katun:    = \lfloor z/20*18\rfloor \mod 20 $\\
-$katun\_baktun:   = \lfloor z/20*18*20\rfloor \mod 20 $\\
-
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Illustrative single-order-up granularities for Mayan calendar with kin as the index}
-\label{single_order}
-\end{table}
-
-#### Aperiodic single-order-up granularities
-
-Aperiodic single-order-up granularities refers to the granularities which are formed with units that does not repeat its values in regular intervals or periods.
+<!-- Now, let us discuss the computation of circular granularities. -->
+
+<!-- #### Aperiodic single-order-up granularities -->
+
+<!-- Aperiodic single-order-up granularities refers to the granularities which are formed with units that does not repeat its values in regular intervals or periods.
 In Gregorian calendar, examples may include days of the month where each month may consists of 28, 29, 30 or 31 days.
@@
 We cannot compute the aperiodic single-order-up granularities using the index of the tsibble and modular arithmetic like in the periodic case using just the hierarchy table and relationships of units.
-
-### Multiple order-up granularities
-
-Computation of multiple-order-up granularities will differ if the units in the hierarchy table are periodic or aperiodic.
+-->
+
+<!-- ### Multiple order-up granularities  -->
+
+<!-- Computation of multiple-order-up granularities will differ if the units in the hierarchy table are periodic or aperiodic.
 We split the method of computation into two cases - one concerning all periodic single-order-up granularities and the second with mixed single-order-up granularities.
-
-### All single-order-up granularities are periodic
-
-z is the index set of the tsibble and x, y are two linear granularities with  $order(x) <  order(y)$.
+-->
+
+<!-- ### All single-order-up granularities are periodic -->
+
+<!-- z is the index set of the tsibble and x, y are two linear granularities with  $order(x) <  order(y)$.
 Also, $f(x, y)$ denotes the accessor function for computing circular granularity which relates x and y and $c(x, y)$ is a constant which relates x and y.
 It is easy to see that for $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
-
-Then, the accessor function f can be used recursively to obtain any multiple-order-up granularities as follows:
-
-\begin{equation} \label{eq1}
-\begin{split}
-f(x,y) & = f(x, x+1) + c(x, x+1)(f(x+1, y)-1) \\  
-  & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
-  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
-  &\vdots\\
-  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
-  +i, x+i+1) - 1)\\
-\end{split}x
-\end{equation}
-
-Let us use the equation to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.
-
-From Equation(1), we have
-\begin{equation} \label{eq1}
-\begin{split}
-f(uinal, baktun) & = f(uinal, tun) + c(uinal, tun)f(tun,katun) + c(uinal, katun)f(katun, baktun) \\
-              & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\
-\end{split}
-\end{equation}
-             (using equations \ref{single_order})
-
-### Mixed single-order-up granularities - circular or aperiodic
-
-Let us turn to Gregorian calendar for addressing this case.
-
-Suppose we have a hierarchy table for  Gregorian calendar as \ref{table_accessor2}.
+-->
+
+<!-- Then, the accessor function f can be used recursively to obtain any multiple-order-up granularities as follows: -->
+
+<!-- \begin{equation} \label{eq1} -->
+<!-- \begin{split} -->
+<!-- f(x,y) & = f(x, x+1) + c(x, x+1)(f(x+1, y)-1) \\   -->
+<!--   & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\ -->
+<!--   & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\ -->
+<!--   & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\ -->
+<!--   &\vdots\\ -->
+<!--   & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x -->
+<!--   +i, x+i+1) - 1)\\ -->
+<!-- \end{split}x -->
+<!-- \end{equation} -->
+
+<!-- Let us use the equation to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic. -->
+
+<!-- From Equation(1), we have -->
+<!-- \begin{equation} \label{eq1} -->
+<!-- \begin{split} -->
+<!-- f(uinal, baktun) & = f(uinal, tun) + c(uinal, tun)f(tun,katun) + c(uinal, katun)f(katun, baktun) \\ -->
+<!--               & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\ -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
+<!--              (using equations \ref{single_order}) -->
+
+<!-- ### Mixed single-order-up granularities - circular or aperiodic -->
+
+<!-- Let us turn to Gregorian calendar for addressing this case. -->
+
+<!-- Suppose we have a hierarchy table for  Gregorian calendar as \ref{table_accessor2}.
 Since months consists of unequal number of days, any temporal unit which is higher in order than months will also have unequal number of days.
@@
 Similarly, all single-order-up granularities which are formed using units whose orders are higher than months are also periodic.
-
-<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
+-->
+
+<!-- <!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
 However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
--->
-
-```{r table_accessor2}
-kable(tibble::tibble(
-  `linear granularities` = c("minute", "hour", "day", "month", "quarter", "year"),
-  `Conversion factor` = c(60,24, "aperiodic", 3, 4, 1), format = "markdown"))
-```
-
-There can be three scenarios for obtaining calendar categorization here:
-- granularities consisting of two units whose orders are less than day  
-- granularities consisting of two units whose orders are more than month  
-- granularities consisting of one unit with order at most day and another with order at least month
-
-The calendar categorization resulting from the first two cases are periodic and has been handled in the earlier section.
-
-The calendar categorization resulting from the last case are aperiodic.
+--> -->
+
+<!-- ```{r table_accessor2} -->
+<!-- kable(tibble::tibble( -->
+<!--   `linear granularities` = c("minute", "hour", "day", "month", "quarter", "year"), -->
+<!--   `Conversion factor` = c(60,24, "aperiodic", 3, 4, 1), format = "markdown")) -->
+<!-- ``` -->
+
+<!-- There can be three scenarios for obtaining calendar categorization here: -->
+<!-- - granularities consisting of two units whose orders are less than day   -->
+<!-- - granularities consisting of two units whose orders are more than month   -->
+<!-- - granularities consisting of one unit with order at most day and another with order at least month   -->
+
+<!-- The calendar categorization resulting from the first two cases are periodic and has been handled in the earlier section. -->
+
+<!-- The calendar categorization resulting from the last case are aperiodic.
 Examples might include day of the quarter or hour of the month.
 In this section, we will see how to obtain aperiodic circular granularities of these types.
-
-\begin{equation} \label{eq1}
-\begin{split}
-f_{(hour, month)}(z) = f_{(hour, day)}(z) + c(hour, day)f_{(day, month)}(z)\\
-\end{split}
-\end{equation}
-
-Here, the first part of the equation is a single-order-up granularity which can be obtained using last section.
+-->
+
+<!-- \begin{equation} \label{eq1} -->
+<!-- \begin{split} -->
+<!-- f_{(hour, month)}(z) = f_{(hour, day)}(z) + c(hour, day)f_{(day, month)}(z)\\ -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
+
+<!-- Here, the first part of the equation is a single-order-up granularity which can be obtained using last section.
 The second part is not single-order-up and can not be broken down further since each month consists of different number of days.
 In this case, it is important for us to know which month of the year and if the year is a leap year to obtain day of the month.
+-->
 

### 2019-11-21 main.Rmd
circular section updated - need to update aperiodic

@@
 **Definition: Groups Into** : A linear granularity G groups into a linear granularity H, denoted
-$$G \trianglelefteq H,$$
+$G \trianglelefteq H,$
 if for each index j there exists a (possibly infinite) subset S of the integers such
@@
 
-According to this definition, $$G \trianglelefteq H,$$ if each granule H (i) is the union of some
+According to this definition, $G \trianglelefteq H,$ if each granule H (i) is the union of some
 granules of G.
-For example, $$Days \trianglelefteq Months,$$.
+For example, $Days \trianglelefteq Months,$.
 This relationship, however, is not sufficient to fully describe a granularity in terms of another one.
-For example, it is clear that $$Days \trianglelefteq Months,$$, since each month is a grouping of a number of days, however it is also a periodical grouping.
+For example, it is clear that $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping.
 If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
@@
 
-**Example:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $$\trianglelefteq $$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $$\trianglelefteq$$.
-
-Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
+**Example:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+
+<!-- Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
 The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
-
-### Circular {#circular-gran-def}
+-->
+
+#### Computation of linear time granularities
+
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf]. It is assumed that there exists a “bottom”
+granularity and Calendar algebra operations are designed to generate
+new granularities from the bottom one or recursively, from those already generated.
+Thus, the relationship between the operand(s) and the resulting granularities are encoded
+in the operations.
+
+The calendar algebra consists of two kinds of operations: grouping-oriented and
+granule-oriented operations.
+The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
+Example can be to consider a calendar with only two linear granularities minute and hour and hour is generated by grouping every 60 minutes.
+The
+granule-oriented operations do not change the granules of a granularity, but rather make
+choices of which granules should remain in the new granularity.
+For example, one can choose to look at the granularity "Monday" and hence select only Mondays while looking at the linear granularity "day".
+
+Some relevant grouping oriented operations are discussed, which will be used in Section \ref{sec:cyclic-gran-def} to define circular and aperiodic granularities.
+
+- **The grouping operation** : Let G be a full-integer labeled granularity, and m a positive integer.
+The grouping operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G
+into m-granule groups and making each group a granule of the resulting granularity.
+More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
+each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
+
+**Example** :
+
+  - $minute = Group_{60}(second)$
+  - $hour = Group_{60}(minute)$,
+  - $day = Group_{24}(hour)$,
+  - $week = Group_{7}(day)$,
+
+where, second(1) would start a minute and likewise.
+
+- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
+The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
+The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
+**Example**
+
+  - pseudomonth = Alter^{12}_{(11,-1)}(day, Alter^{12}_{(9,-1)}(day, Alter^{12}_{(6,-1)}(day,
+                  Alter_^{12}_{(4,-1)}(day, Alter^{12}_{(2,-3)}(day, Group_31(day)))))),
+                  where the granularity pseudomonth is generated by grouping 31 days, 
+                  and then shrink April (4), June (6), September (9) and November (11)
+                  by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
+
+### Cyclic {#cyclic-gran-def}
 
@@
 
-The formalism of cyclic time granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+We propose a formalism of cyclic time granularities through the tsibble [@Wang2019-aw] framework of organizing temporal data.
 A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble.
@@
 Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
+
+The mapping of two linear granularities forming a cyclic granularity can be regular or irregular.
+It will be regular if, for example, they are formed by grouping operation.
+Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
+However, if they are formed by the altering-tick operation, granules of the resulting granularity is composed of different number of granules of bottom granularity.
+However, the relationship can still be periodic and hence they are also useful in addressing periodicities.
+The formalisms would however differ for circular and aperiodic granularities.
 
@@
 <!-- In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up. -->
-
-<!-- ### Computation of time granularities -->
-
-<!-- Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf].
-It is assumed that there exists a “bottom” -->
-<!-- granularity and Calendar algebra operations are designed to generate -->
-<!-- new granularities from the bottom one or recursively, from those already generated. -->
-<!-- Thus, the relationship between the operand(s) and the resulting granularities are encoded -->
-<!-- in the operations. -->
-
-<!-- The calendar algebra consists of two kinds of operations: grouping-oriented and -->
-<!-- granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity, while the -->
-<!-- granule-oriented operations do not change the granules of a granularity, but rather make -->
-<!-- choices of which granules should remain in the new granularity. -->
-
-<!-- - **grouping-oriented** : Let G be a full-integer labeled granularity, and m a positive integer. The grouping -->
-<!-- operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G -->
-<!-- into m-granule groups and making each group a granule of the resulting granularity. -->
-<!-- More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for -->
-<!-- each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$. Granularity week is defined by $week = Group_7(day)$. Note that, we assume that the day labeled 1 starts a week. -->
-
-<!-- - **granule-oriented** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
-The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-Since G2 partitions G1, each granule of G1 consists of some contiguous granules of G2.
-The granules of G1 can be partitioned into m-granule groups such that G1(1) to G1(m) are in one group, G1(m + 1) to G1(2m) are in the following group, and so on.
-The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
-For example, if G1 represents 30-day groups (i.e., G1 = Group_30(day)) and we want to add a day (i.e., k = 1) to the 5th one in each group of 12 months (i.e., l = 5 and m = 12), we may have $Alter^{12}_{5,1}(day, G1)$.
-Intuitively, this will make May have 31 days.
--->
 

### 2019-11-23 main.Rmd
introduction changed

@@
 <!-- problem of analyzing data with such finer scales and classical ways to deal with it -->
+
 Temporal data are available at various resolution depending on the context.
-Social and economic data like GDP are often collected and reported at coarser temporal scales like monthly, quarterly or annually.
-With recent advancement in technology, more and more data are recorded and stored at much finer temporal scales.
-Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be collected at every second.
+Social and economic data like GDP are often collected and reported at coarse temporal scales like monthly, quarterly or annually.
+With recent advancement in technology, more and more data are recorded at much finer temporal scales.
+Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be recorded every second.
 As the frequency of data increases, the number of questions about periodicity of the observed variable also increases.
 For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
-This approach requires deconstructing time in various possible ways.
-Calendar-based graphics[] can unpack the temporal variable, at different resolutions, to digest multiple seasonalities, and special events.
-They mainly pick out patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
-Any sub-daily resolution temporal data can also be displayed using this type of faceting [reference trellis plot] with days of the week, month of the year and another sub-daily deconstruction of time.
-
-<!-- # challenges of classical ways of displaying -->
+This approach requires deconstructing time in various possible ways called time granularities[@aigner2011visualization].
+It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on periodicity of the observed data.
+This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
+Analysts are expected to iteratively explore all possible choices of time granularities for comprehending possible periodicities in the data.
+But too many choices and a lack of a systematic approach to do so might become overwhelming.
+
+<!-- # existing ways and challenges -->
+
+Calendar-based graphics[@wang2019tsibble] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
+Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time.
 But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
-A temporal granularity which results from such a deconstruction may be intuitively described as a sequence of time granules, each one consisting of a set of time instants[].
-There can be several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
+There can be several classes of time deconstructions, viz.
+based on arrangement (linear vs.
+cyclic) or hierarchical order of calendar.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week.
-Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.
-Also, <!--details abt time gran -->
-the hierarchical structure of time creates a natural nested ordering.
+One of the first attempts of characterising these granularities occur in [@Bettini1998-ed].
+The definitions and rules defined are inadequate to reflect periodicities in time.
+Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior.
+Cyclic time granularities can be circular or aperiodic depending on if mappings between granularities are regular or irregular.
+Examples of circular can be hour of the day, day of the week and that of aperiodic granularities can be day of the month or public holidays.
+Time deconstructions can also be based on the hierarchical structure of time.
 For example, hours are nested within days, days within weeks, weeks within months, and so on.
-Hence, we can construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
-
-<!-- We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and are single-order-up.
--->
-
-<!-- TODO -->
-It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the observed data.
-This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
+lubridate [@G_Grolemund2011-vm] creates easy access and manipulation of common date-time objects.
+But most accessor functions are limited to single-order-up granularities.
 
@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-
-It is common to see aggregates of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates of usage [@2012-la] across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-
-We propose that the analysis of probability distributions of smart meter data at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA).
+We propose that the analysis of smart meter data can be benefitted from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
+However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analysed for different periodicities.
+
+<!-- at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA).
 EDA calls for utilizing visualization and transformation to explore data systematically.
 It is a process of generating hypothesis, testing them and consequently refining them through investigations.
+-->
 
 <!--existing closely related work and how ours relate --> 
- This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.
+ <!-- This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones. -->
 
 <!--actions on multiple order up - usage -->
-Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data.
+
+<!-- Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data.
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
@@
 But too many choices might leave him bewildered.
+-->
 
 <!-- challenges and how you are dealing wih it --> 
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow through proper study of what can be considered a prospective graphic for exploration.
-Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-Probability distributions can be examined using the range of graphics available in the [ggplot2](https://cran.r-project.org/package=ggplot2) package.
-
-In particular, this work provides the following tools.
-
-  * Functions to create multiple-order-up time granularities.
-This is an extension to the [lubridate]([https://cran.r-project.org/package=lubridate) package, which allows for the creation of some calendar categorizations, usually single-order-up.
-
-  * Checks on the feasibility of creating plots or drawing inferences from two granularities together.
-Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-
-<!-- The remainder of the paper is organized as follows.
-Section \@ref(sec:algorithm) details the construction of the calendar layout in depth.
-It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)).
-Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes.
-An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case).
-Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.
--->
-
-# Definitions of time granularities{#cirgran-def}
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow through the following:
+
+<!-- Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
+-->
+
+  * Formal characterization of cyclic granularities
+
+  * Facilitate manipulation of single and multiple order-up time granularities
+
+  * Checking feasibility of creating plots or drawing inferences from any two cyclic granularities
+
+  * Recommend prospective probability distributions for exploring distributions of univariate dependent variable across pair of granularities
+
+  <!-- Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
+-->
+
+The remainder of the paper is organized as follows.
+Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterisation of circular and aperiodic time granularities.
+Section \ref{sec:synergy} discusses the effect in data structure due to how pairs of cyclic time granularities are related.
+Section \ref{sec:visualization} discusses the role of plot choices, synergy of time granularities and number of observations to construct an informative and trustworthy visualization.
+Section \ref{sec:application} discusses how systematic exploration can be carried out for a temporal and non-temporal case.
+Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.
+
+# Definitions of time granularities{#sec:timegrandef}
 
@@
 
-### Linear {#linear-gran-def}
+### Linear {#sec:linear-gran-def}
 
@@
 The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
-**Example**
+
+**Example** :
 
@@
 
-### Cyclic {#cyclic-gran-def}
+### Cyclic {#sec:cyclic-gran-def}
 
@@
 Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
-However, if they are formed by the altering-tick operation, granules of the resulting granularity is composed of different number of granules of bottom granularity.
-However, the relationship can still be periodic and hence they are also useful in addressing periodicities.
-The formalisms would however differ for circular and aperiodic granularities.
-
-#### Circular {#circular-gran-def}
+However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
+However, the relationship can still be periodic, depending on the choice of periods and hence these are also useful in addressing periodicities.
+The formalisms would however differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
+
+#### Circular {#sec:circular-gran-def}
 
@@
 
-### Aperiodic {#aperiodic-gran-def}
-
-**Definition:** An **Aperiodic circular granularity** can not be defined using modular arithmetic in a similar fashion.
-The modulus for these type of calendar categorizations are not constant due to unequal length of some linear granularities.
-For example, please refer to the table below:
+### Aperiodic {#sec:aperiodic-gran-def}
+
+An **Aperiodic granularity** can not be defined using modular arithmetic in a similar fashion.
+The modulus for these type of calendar categorizations are not constant due to irregular mapping with bottom granularities.
+For example, table \ref{tab:} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
+The coarse granularity is assumed as the period within which the finer granularity will repeat its behavior aperiodically.
 
@@
 
-HOM: & $C_3(s)$ & $= s \mod `r 24*30`$ (approximately) & $n_3=$&$`r 24*31`$\\
-HOY: & $C_4(s)$ & $= s \mod `r 24*365`$ (except for leap years) & $n_4=$&$`r 24*366`$\\
-DOM: & $C_6(s)$ & $= \lfloor s/24\rfloor\mod 30$  (approximately) & $n_6=$&$31$\\
-DOY: & $C_7(s)$ & $= \lfloor s/24\rfloor \mod 365$ (except for leap years) & $n_7=$&$366$\\
-WOM: & $C_8(s)$ & $=\lfloor s/168 \rfloor \mod 4$ (approximately) & $n_8 =$&$ 5$\\
-WOY: & $C_9(s)$ & $=\lfloor s/168\rfloor \mod 52$ (approximately) & $n_9 =$&$ 53$\\
-MOY: & $C_{10}(s)$ & $= \lfloor s/720\rfloor \mod 12$ (approximately) & $n_{10}=$&$12$\\
+Day-of-Month: & $A_1(z)$ &  $k_1=31, 30, 29, 28$\\
+Hour-of-Month: & $A_2(z)$ & $k_2=24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $A_3(z)$ & $k_3=366, 365$\\
+Week-of-Month: & $A_4(z)$ & $k_4=5, 4$\\ 
 \bottomrule
@@
 \end{center}
-\caption{Illustrative aperiodic circular granularities with time index in hours}
+\caption{Illustrative aperiodic circular granularities with potential period lengths}
 \label{definitions}
@@
 
-C1 and C2 are two cyclic granularities in the hierarchy table with  $order(C1) <  order(C2)$. 
-Let $f(C_1, C_2)$ denotes the accessor function for representing the finer unit C1 in terms of a coarser unit C2 and $const(C1, C2)$ is a constant which relates C1 and C2.
-
-<!-- We consider a tsibble with an index (I), key(optional) and measured variables. Index is the variable with inherent ordering from past to future. -->
-
-Then $f(x, y)$ can be  computed using modular arithmetic as follows:
-
-$$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$
-
-The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
+**Definition: Aperiodic granularity** A aperiodic granularity A(BG, G) relates a linear granularity G to the bottom granularity, if
+
+$A_{BG,G}(z) = {0, 1, \dots ak_i}$, where ak_i denotes the number of granules of the bottom granularity present in the i^{th} granule of G.
+
+\begin{equation} \label{eq2}
+A_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
+BG(z) \neq \phi
+\end{equation}
+where,  
+z denotes the index set.  
+BG denotes the index of the tsibble (bottom granularity).  
+G be a linear granularity, periodic with respect to BG.    
+k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.
+
+<!-- Then $f(x, y)$ can be  computed using modular arithmetic as follows: -->
+
+<!-- $$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$ -->
+
+<!-- The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
+-->
 
@@
 
-# Synergy of cyclic time granularities
+# Synergy of cyclic time granularities {#sec:synergy}
 
@@
 
-# Visualization
+# Visualization {#sec:visualization}
 
@@
 
-In Section \ref{sec:harmonyclash}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are build.
+In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
 In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.
@@
 
-# Applications
-
-## Smart meter data of Australia {#smartmeter}
+# Applications {#sec:application}
+
+## Smart meter data of Australia {#sec:smartmeter}
 
@@
 
-In Figure \ref{fig:day_fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
+In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
@@
 
-smart_meter10 %>% 
-  filter(customer_id %in% c(10006704, 10017936)) %>% 
-  create_gran("day_fortnight") %>%
-  as_tibble() %>% 
-  group_by(customer_id, day_fortnight) %>% 
-  summarise(median_consumption= median(general_supply_kwh)) %>%
-  ggplot2::ggplot(aes(
-    x = as.factor(day_fortnight),
-    y = median_consumption,
-    color =  customer_id,
-    group = customer_id)) +
-  geom_line() +
-  xlab("day_fortnight") +
-  ylab("general_supply_kwh") + 
-  theme_bw() 
-```
-
-```{r day_fortnight}
+# smart_meter10 %>% 
+#   filter(customer_id %in% c(10006704, 10017936)) %>% 
+#   create_gran("day_fortnight") %>%
+#   as_tibble() %>% 
+#   group_by(customer_id, day_fortnight) %>% 
+#   summarise(median_consumption= median(general_supply_kwh)) %>%
+#   ggplot2::ggplot(aes(
+#     x = as.factor(day_fortnight),
+#     y = median_consumption,
+#     color =  customer_id,
+#     group = customer_id)) +
+#   geom_line() +
+#   xlab("day_fortnight") +
+#   ylab("general_supply_kwh") + 
+#   theme_bw() 
+```
+
+```{r day-fortnight}
 library(lvplot)
@@
 
-## T20 cricket data of Indian Premiere League {#cricket}
+## T20 cricket data of Indian Premiere League {#sec:cricket}
 

### 2019-11-23 main.Rmd
aperiodic granularity updated

@@
 **Definition: Groups periodically into** A granularity H is periodical with respect to a granularity G if
-(1) $$G \trianglelefteq H,$$, and
+(1) $G \trianglelefteq H,$, and
 (2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
@@
 Then, if ${S_0,...,S_(R-1)}$ are the sets of indexes of G describing
-$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$.
+$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
 
@@
 
-The mapping of two linear granularities forming a cyclic granularity can be regular or irregular.
+The mappings of two linear granularities forming a cyclic granularity can be regular or irregular.
 It will be regular if, for example, they are formed by grouping operation.
@@
 However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
-However, the relationship can still be periodic, depending on the choice of periods and hence these are also useful in addressing periodicities.
-The formalisms would however differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
+The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities.
+However, the formalisms would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
 
@@
 
-An **Aperiodic granularity** can not be defined using modular arithmetic in a similar fashion.
+An **Aperiodic granularity** can not be defined using modular arithmetic.
 The modulus for these type of calendar categorizations are not constant due to irregular mapping with bottom granularities.
-For example, table \ref{tab:} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
+For example, table \ref{tab:aperiodic} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
 The coarse granularity is assumed as the period within which the finer granularity will repeat its behavior aperiodically.
@@
 \caption{Illustrative aperiodic circular granularities with potential period lengths}
-\label{definitions}
+\label{aperiodic}
 \end{table}
 
-**Definition: Aperiodic granularity** A aperiodic granularity A(BG, G) relates a linear granularity G to the bottom granularity, if
-
-$A_{BG,G}(z) = {0, 1, \dots ak_i}$, where ak_i denotes the number of granules of the bottom granularity present in the i^{th} granule of G.
-
-\begin{equation} \label{eq2}
-A_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
-BG(z) \neq \phi
-\end{equation}
+**Definition: Aperiodic granularity** An aperiodic granularity A(G, BG) that relates linear granularities G and bottom granularity BG, where G is periodical with respect to BG, if
+
+$A_{BG, G}(z) = BG(z - \sum_{j=0}^{k-1}\vert S_{j \mod R} \vert)$ for $z \in S_k$,
+
 where,
-z denotes the index set.  
-BG denotes the index of the tsibble (bottom granularity).  
-G be a linear granularity, periodic with respect to BG.    
-k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.
+z denotes the index set,  
+P and R denotes the period of the granularity BG and number of granules of G in each period of BG,  
+$S_j$ are the sets of indices of BG describing $G_j$ such that $G(j) = \bigcup_{i \in S_j}BG(i)$,  
+$\vert S_j \vert$ is the cardinality of set $S_j$,
+
+Since G is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
+
+- $BG \trianglelefteq G,$
+- $G(j) = \bigcup_{i \in S_j \mod R}BG(P*\lfloor j/R \rfloor + i)$,
+
+**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram \autoref{fig:aperiodic-gran}.
+Each granule is represented by a box.
+Similar to Figure \ref:{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
+Two linear granularities G1 and G2 are considered, where $G1 = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $G2 = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$.
+This implies that G1 is made up of shrinking every 1st granule of Group_3(BG) by 1 and G2 is made up of shrinking every 1st granule of Group_3(BG) by 2.
+Number of granules of G1 and G2 in each period of BG is 2, but the number of granules of BG in each of those granules are different.
+A(G, G1) and A(G, G2) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G.
+
+```{r aperiodic-gran, echo=FALSE,out.width = "100%",fig.cap="\\label{fig:circular-gran} Relating circular granularities and bottom linear granularity "}
+knitr::include_graphics("Figs/aperiodic-gran.png")
+```
 

### 2019-11-23 main.Rmd
aperiodic gran concept included

@@
 
-**Definition: Aperiodic granularity** An aperiodic granularity A(G, BG) that relates linear granularities G and bottom granularity BG, where G is periodical with respect to BG, if
+**Definition: Aperiodic granularity** An aperiodic granularity A(G, BG) that relates linear granularities G and bottom granularity BG, with G being periodical with respect to BG, is given by
 
@@
 z denotes the index set,  
-P and R denotes the period of the granularity BG and number of granules of G in each period of BG,  
+R denotes the number of granules of G in each period of BG,  
 $S_j$ are the sets of indices of BG describing $G_j$ such that $G(j) = \bigcup_{i \in S_j}BG(i)$,  
@@
 - $BG \trianglelefteq G,$
-- $G(j) = \bigcup_{i \in S_j \mod R}BG(P*\lfloor j/R \rfloor + i)$,
+- $G(j) = \bigcup_{i \in S_j \mod R}BG(P*\lfloor j/R \rfloor + i)$,  where P denotes the period of the granularity BG.
 
@@
 Each granule is represented by a box.
-Similar to Figure \ref:{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
+Similar to Figure \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
 Two linear granularities G1 and G2 are considered, where $G1 = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $G2 = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$.
@@
 ### Computation of multiple-order from single-order
+
+#### All single order granularities are circular
 
@@
 
+#### Single order granularities are circular or aperiodic
+
+#### All single order granularities are circular
+
 ### Computation of single-order from multiple-order
@@
 <!-- In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up. -->
+
+#### Single order granularities are circular or aperiodic
 

### 2019-11-24 main.Rmd
all notations change in the aperiodic definition

@@
 
-**Definition: Aperiodic granularity** An aperiodic granularity A(G, BG) that relates linear granularities G and bottom granularity BG, with G being periodical with respect to BG, is given by
-
-$A_{BG, G}(z) = BG(z - \sum_{j=0}^{k-1}\vert S_{j \mod R} \vert)$ for $z \in S_k$,
+**Definition: Aperiodic granularity** An aperiodic granularity A(AG, BG) that relates linear granularities AG and bottom granularity BG, with AG being periodical with respect to BG, is given by
+
+$A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert)$ for $z \in T_k$,
 
 where,  
-z denotes the index set,  
-R denotes the number of granules of G in each period of BG,  
-$S_j$ are the sets of indices of BG describing $G_j$ such that $G(j) = \bigcup_{i \in S_j}BG(i)$,  
-$\vert S_j \vert$ is the cardinality of set $S_j$,
-
-Since G is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
-
-- $BG \trianglelefteq G,$
-- $G(j) = \bigcup_{i \in S_j \mod R}BG(P*\lfloor j/R \rfloor + i)$,  where P denotes the period of the granularity BG.
-
-**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram \autoref{fig:aperiodic-gran}.
+z denotes the index set (index of BG),
+R' denotes the number of granules of AG in each period of BG,  
+$T_w$ are the sets of indices of BG describing $AG_w$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$,  
+$\vert T_w \vert$ is the cardinality of set $T_w$,
+
+Since AG is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
+
+- $BG \trianglelefteq AG,$
+- $AG(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
+
+**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:aperiodic-gran}.
 Each granule is represented by a box.
-Similar to Figure \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
-Two linear granularities G1 and G2 are considered, where $G1 = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $G2 = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$.
-This implies that G1 is made up of shrinking every 1st granule of Group_3(BG) by 1 and G2 is made up of shrinking every 1st granule of Group_3(BG) by 2.
-Number of granules of G1 and G2 in each period of BG is 2, but the number of granules of BG in each of those granules are different.
-A(G, G1) and A(G, G2) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G.
+Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
+Two linear granularities AG and AH are considered, where $AG = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$.
+This implies that AG is made up of shrinking every 1st granule of Group_3(BG) by 1 and AH is made up of shrinking every 1st granule of Group_3(BG) by 2.
+Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different.
+A(BG, AG) and A(BG, AH) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH.
+For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}, suppose we want to compute A(BG,AG)(8), then since $8 \in T_{3}$
+
+\begin{equation} \label{eq4}
+\begin{split}
+A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) \\
+  & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
+  & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
+  & = BG(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
+  & = BG(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
+  & = BG(8 - 2.2 - 3) \\
+  & = BG(1) \\
+\end{split}
+\end{equation}
 
@@
 
-To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar.
+The linear granularities G, H, G1 and G2 considered in the definitions of both circular and aperiodic granularities are periodic with respect to the bottom granularity.
+The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
+
+<!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
 
@@
 Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+The behavioral patterns in households vary substantially, for example, some famailies use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
 -->

### 2019-11-24 main.Rmd
example for AH included

@@
 A(BG, AG) and A(BG, AH) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH.
-For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}, suppose we want to compute A(BG,AG)(8), then since $8 \in T_{3}$
+For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
+For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \ref{eq:eq4} and  \ref{eq:eq5}.
 
@@
 \begin{split}
-A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) \\
+A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , since 8 \in T_{3} \\
   & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
@@
 
+\begin{equation} \label{eq5}
+\begin{split}
+A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) \\ , 10 \in T_{1}
+  & = BG(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
+  & = BG(10 - \vert T_{0 \mod 2}\vert) \\
+  & = BG(10 - \vert T_{0}\vert) \\
+  & = BG(10 - 7) \\
+  & = BG(3) \\
+\end{split}
+\end{equation}
+
 ```{r aperiodic-gran, echo=FALSE,out.width = "100%",fig.cap="\\label{fig:circular-gran} Relating circular granularities and bottom linear granularity "}
-knitr::include_graphics("Figs/aperiodic-gran.png")
+knitr::include_graphics("Figs/aperiodic-granv1.png")
 ```
@@
 
-The linear granularities G, H, G1 and G2 considered in the definitions of both circular and aperiodic granularities are periodic with respect to the bottom granularity.
+The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity.
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.

### 2019-11-24 main.Rmd
quasi-periodic case added

@@
 
-$A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert)$ for $z \in T_k$,
-
+\begin{equation}\label{eq6}
+A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for \z \in T_k
+\end{equation}
 where,  
@@
 R' denotes the number of granules of AG in each period of BG,  
-$T_w$ are the sets of indices of BG describing $AG_w$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$,  
-$\vert T_w \vert$ is the cardinality of set $T_w$,
+$T_w$ are the sets of indices of BG describing $AG_w$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
+$\vert T_w \vert$ is the cardinality of set $T_w$.
 
@@
 \begin{split}
-A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , since 8 \in T_{3} \\
+A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , since \8 \in T_{3} \\
   & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
@@
 \begin{split}
-A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) \\ , 10 \in T_{1}
+A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) , since \10 \in T_{1}  \\
   & = BG(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
@@
 
+If we ignore leap years and all its exception, the periodicity of the grouping (days, months), P' = 365, R' = 12, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
+Considering leap years and all their exceptions, period is 400 years and hence $P' = (3*365 + 366)$, $R' = 12*400$.
+\autoref{eq:eq6} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
+
 <!-- Then $f(x, y)$ can be  computed using modular arithmetic as follows: -->
@@
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
+It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \ref{eq:eq6} can be modified as follows to account for exceptions $E = [e_begin, e_end]$ (\ref:{def:something})
+
+\begin{equation}\label{eq7}
+A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
+\end{equation}
+
+for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T(w)$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
+
+but the grouping is always periodic.
 

### 2019-11-24 main.Rmd
visualization intro delete

@@
 This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in [@Wickham2009-pk].
-bibliography: bibliography.bib
+bibliography: [bibliography.bib]
 output:
@@
 library(ggridges)
-library(viridis)
 library(tsibble)
@@
 \begin{equation}\label{eq6}
-A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for \z \in T_k
+A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for z \in T_k
 \end{equation}
@@
 \begin{split}
-A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , since \8 \in T_{3} \\
+A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , since 8 \in T_{3} \\
   & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
@@
 \begin{split}
-A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) , since \10 \in T_{1}  \\
+A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) , since 10 \in T_{1}  \\
   & = BG(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
@@
 
-### Computation of multiple-order from single-order
-
-#### All single order granularities are circular
+### Computation of multiple-order from single-order {#sec:multiplefromsingle}
 
@@
   &\vdots\\
-  & = \sum_{i=0}^{order(m) - order(l) - 1} k(l, l+i)(C_{G_(l+i),G_(l+i+1)}(z) - 1)\\
+  & = \sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_(l+i),G_(l+i+1)}(z) - 1)\\
 \end{split}
@@
 
-#### Single order granularities are circular or aperiodic
-
-#### All single order granularities are circular
-
-### Computation of single-order from multiple-order
+### Computation of single-order from multiple-order {#sec:singlefrommultiple}
 
@@
 
-<!-- In the next section, we discuss the computation of any single-order-up and multiple-order-up granularities in a periodic and aperiodic set up. -->
-
-#### Single order granularities are circular or aperiodic
-
-<!-- Computation of time granularities would differ in a periodic and non-periodic set up. -->
-
-<!-- Now, let us discuss the computation of circular granularities. -->
-
-<!-- #### Aperiodic single-order-up granularities -->
-
-<!-- Aperiodic single-order-up granularities refers to the granularities which are formed with units that does not repeat its values in regular intervals or periods.
-In Gregorian calendar, examples may include days of the month where each month may consists of 28, 29, 30 or 31 days.
-So there is no single number that can be used for converting days to months.
-We cannot compute the aperiodic single-order-up granularities using the index of the tsibble and modular arithmetic like in the periodic case using just the hierarchy table and relationships of units.
--->
-
-<!-- ### Multiple order-up granularities  -->
-
-<!-- Computation of multiple-order-up granularities will differ if the units in the hierarchy table are periodic or aperiodic.
-We split the method of computation into two cases - one concerning all periodic single-order-up granularities and the second with mixed single-order-up granularities.
--->
-
-<!-- ### All single-order-up granularities are periodic -->
-
-<!-- z is the index set of the tsibble and x, y are two linear granularities with  $order(x) <  order(y)$.
-Also, $f(x, y)$ denotes the accessor function for computing circular granularity which relates x and y and $c(x, y)$ is a constant which relates x and y.
-It is easy to see that for $order(x+1) =  order(y)$, the function is same as the single-order-up granularities.
--->
-
-<!-- Then, the accessor function f can be used recursively to obtain any multiple-order-up granularities as follows: -->
-
-<!-- \begin{equation} \label{eq1} -->
-<!-- \begin{split} -->
-<!-- f(x,y) & = f(x, x+1) + c(x, x+1)(f(x+1, y)-1) \\   -->
-<!--   & =  f(x, x+1) + c(x, x+1)[f(x+1, x+2) + c(x+1, x+2)(f(x+2, y) - 1) - 1]\\ -->
-<!--   & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\ -->
-<!--   & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\ -->
-<!--   &\vdots\\ -->
-<!--   & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x -->
-<!--   +i, x+i+1) - 1)\\ -->
-<!-- \end{split}x -->
-<!-- \end{equation} -->
-
-<!-- Let us use the equation to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic. -->
-
-<!-- From Equation(1), we have -->
-<!-- \begin{equation} \label{eq1} -->
-<!-- \begin{split} -->
-<!-- f(uinal, baktun) & = f(uinal, tun) + c(uinal, tun)f(tun,katun) + c(uinal, katun)f(katun, baktun) \\ -->
-<!--               & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-<!--              (using equations \ref{single_order}) -->
-
-<!-- ### Mixed single-order-up granularities - circular or aperiodic -->
-
-<!-- Let us turn to Gregorian calendar for addressing this case. -->
-
-<!-- Suppose we have a hierarchy table for  Gregorian calendar as \ref{table_accessor2}.
+In \ref{sec:singlefrommultiple} and \ref{sec:multiplefromsingle}, we assume that all single order granularities are circular.What happens when we have a mix of circular and aperiodic granularities.
+
+Let us turn to Gregorian calendar for addressing this case.
+Suppose we have a hierarchy table for  Gregorian calendar as \ref{table_accessor2}.
 Since months consists of unequal number of days, any temporal unit which is higher in order than months will also have unequal number of days.
-This is an example of a hierarchy table which has both periodic and aperiodic single-order-up granularities.
+This is an example of a hierarchy structure which has both circular and aperiodic single-order-up granularities.
 The single-order-up granularity day_month is aperiodic.
@@
 Similarly, all single-order-up granularities which are formed using units whose orders are higher than months are also periodic.
--->
-
-<!-- <!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
+
+<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
 However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
---> -->
-
-<!-- ```{r table_accessor2} -->
-<!-- kable(tibble::tibble( -->
-<!--   `linear granularities` = c("minute", "hour", "day", "month", "quarter", "year"), -->
-<!--   `Conversion factor` = c(60,24, "aperiodic", 3, 4, 1), format = "markdown")) -->
-<!-- ``` -->
-
-<!-- There can be three scenarios for obtaining calendar categorization here: -->
-<!-- - granularities consisting of two units whose orders are less than day   -->
-<!-- - granularities consisting of two units whose orders are more than month   -->
-<!-- - granularities consisting of one unit with order at most day and another with order at least month   -->
-
-<!-- The calendar categorization resulting from the first two cases are periodic and has been handled in the earlier section. -->
-
-<!-- The calendar categorization resulting from the last case are aperiodic.
+-->
+
+|  G	        |C/A               |k
+|------------	|------------------|------------|
+|  minute     |minute-of-hour    |  60        |
+|  hour    	  |hour-of-day       |  24  	    |
+|  day      	|day-of-month      | "aperiodic"|
+|  month    	|month-of-year     |  12        |
+|  year   	  | 1                |	1         |
+
+There can be three scenarios for obtaining multiple-order-granularities here:
+- granularities relating two linear granularities whose orders are less than day
+- granularities relating two linear granularities whose orders are more than month
+- granularities relating two linear granularities with order at most day and another with order at least month
+
+The single order-up granularites resulting from the first two cases are periodic and has been handled in \ref{sec:singlefrommultiple} and \ref{sec:multiplefromsingle}.
+The calendar categorization resulting from the last case are aperiodic.
 Examples might include day of the quarter or hour of the month.
-In this section, we will see how to obtain aperiodic circular granularities of these types.
--->
-
-<!-- \begin{equation} \label{eq1} -->
-<!-- \begin{split} -->
-<!-- f_{(hour, month)}(z) = f_{(hour, day)}(z) + c(hour, day)f_{(day, month)}(z)\\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-
-<!-- Here, the first part of the equation is a single-order-up granularity which can be obtained using last section.
-The second part is not single-order-up and can not be broken down further since each month consists of different number of days.
-In this case, it is important for us to know which month of the year and if the year is a leap year to obtain day of the month.
--->
+Let us consider the computation of $A_{hour, month}$ given single order-up granularities $A_{hour, day} = 12$, $A_{day, month} = 20$ and $C_{month, year} = 9$.
+
+\begin{equation} \label{eq1}
+\begin{split}
+A_{(hour, month)} & = (\sum_{w=0}^{9 - 1}(\vert T(w)\vert) +  20)/24 + 12 \\
+                  & = (\sum_{w=0}^{C_{month, year} - 1}(\vert T(w)\vert) +  A_{day, month})*k(hour_day) + A_{hour, day} \\
+                  & = (A{day_year} - 1)*k(hour_day) + A_{hour, day} \\
+\end{split}
+\end{equation}
+where, $month(w) = \bigcup_{z \in T_w}day(z)$
+
+Clearly, the computation of multiple order-up granularities become more involved with aperiodic mappings of linear granularities in the calendar structure.
+With one aperiodic granularity, we can write the following:
+
+A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included}
 
@@
 
-The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk].
+<!-- The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk].
 The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, allows us to produce graphics using the same structured thinking that is used to design analysis.
@@
 **Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
+-->
 
 Creating a visualisation requires a number of nuanced judgements.
-However, we exploit some features of layers and faceting to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate temporal granularities.
+We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate temporal granularities.
 The general framework of visualizing the distribution invloves a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
@@
 
+# Discussion {#sec:discussion}
+
+# Acknowledgements {-}
+
+# Bibliography
+
+```{r write-bib}
+write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib')
+```
+
 <!-- # ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."} -->
@@
 ```
-
-# Discussion {#sec:discussion}
-
-# Acknowledgements {-}
-
-# Bibliography
-
-<!-- ```{r write-bib} -->
-<!-- write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
-<!-- ``` -->

### 2019-11-24 main.Rmd
cricket story added

@@
 In a Twenty20 game the two teams have a single innings each, which is restricted to a maximum of 20 overs.
-Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls.
-A hierarchy like table \ref{tab:hierarchy} can be construed for this game format.
+Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls with some exceptions.
 
@@
 
-```{r hierarchy, echo=FALSE}
-hierarchy_model <- tibble::tibble(units = c("ball", "over", "inning", "match"), 
-                                  convert_fct = c(6, 20, 2, 1))
-knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket")
-```
-
-```{r readdata}
+<!-- ```{r hierarchy, echo=FALSE} -->
+<!-- hierarchy_model <- tibble::tibble(linear gran = c("ball", "over", "inning", "match"),  -->
+<!--                                   convert_fct = c(6, 20, 2, 1)) -->
+<!-- knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket") -->
+<!-- ``` -->
+
+Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, where each over, which represents an ordering from past to future, can form the index of the tsibble.
+The hierarchy table would look like the following:
+
+```{r hierarchy2}
 library(gravitas)
 library(tibble)
-glimpse(cricket)
-```
-
-Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, where each over, which represents an ordering from past to future, can form the index of the tsibble.
-The hierarchy table would look like the following:
-
-```{r hierarchy2}
-
 hierarchy_model <- tibble::tibble(
-  units = c("index", "over", "inning", "match"),
-  convert_fct = c(1, 20, 2, 1))
+  G = c("over", "inning", "match"),
+  k = c(20, 2, 1))
 
@@
 
-First, we look at the distribution of runs per over across over of the innings and seasons in \ref{fig:seas-over-inning}.
+First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas-over-inning}.
 The distribution of runs per over has not significantly changed from 2008 to 2016.
 There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
-Hence, we work with subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.
+Hence, we work with subsets of seasons to answer some of the questions:
+
+Q1: How run rates vary depending on if a team bats first or second?
 
 Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final 4 from 2010 to 2015.
-It would be interesting to observe their strategies throughout all matches in the two seasons.
-The following two questions might help us partially understand their strategies.
+It would be interesting to take their example in order to dive deeper into the first question.
 
@@
 <!-- - Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings? -->
-
-Q1: How run rates vary depending on if a team bats first or second?
-
-Q2: How number of wickets and dot balls in the previous over affect the runs in the subsequent over across different overs of the innings?
 
@@
 
-```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons."}
+```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."}
   cricket_tsibble_all %>% 
@@
   hierarchy_model,
-  plot_type = "lv")
-```
+  plot_type = "lv")  + ggtitle("")
+```
+
+Q2: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalizes runs in the subsequent over more?
+
+For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
+If a batsman is bowled out, it does not necessarily signify good fielding.
+So we only include catches and run out as a measure of fielding.
+Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.
+Let us see if this fact is true.
+Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
+
+```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and dismissals by catch and catch and bowled.
+This shows good fielding in one over leads to lower runs in the subsequent overs for at least 50 percentiles of times."}
+
+cricket_fielding <- cricket_season %>% 
+  mutate(fielding_proxy = if_else(dismissal_kind %in% c("caught", "caught and bowled"),1,0)) %>% 
+  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(runs_per_over = sum(total_runs),
+            run_rate = sum(total_runs)/length(total_runs),
+            fielding_wckts = sum(fielding_proxy)) %>%
+            mutate(diff_run_rate = c(0,diff(run_rate)))
+
+cricket_tsibble_fld <- cricket_fielding %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
+
+cricket_tsibble_fld %>% gran_obs("over", "fielding_wckts")
+
+cricket_tsibble_fld %>% 
+  filter(fielding_wckts %in% c(0,1)) %>% 
+  prob_plot( "fielding_wckts","over",
+                               hierarchy_model,
+                               response = "diff_run_rate",
+                               plot_type = "quantile",
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
+           overlay = TRUE)  + ggtitle("")
+
+```
+
+<!-- Q2: Amongst good fielding and bowling - which affect the runs of the subsequent overs more? -->
+
+Q3: Is runs set to reduce in the next over for dot balls in this over?
+
+A dot ball is a delivery bowled without any runs scored off it.
+The number of dot balls is reflective of the quality of bowling in the game.
+Run rate of an over should ideally decrease if the number of dot balls increase.
+However, what is the effect of dot balls on runs scored in the subsequent over.
+Will players  batsman likely to go for big shots because they couldn't score good runs in the previous over?
+Or they should play consistently and avoid scoring high?
+From figure \autoref{exdot} shows the quantile plot of runs across overs for none, one or two dot balls per over in the facets.
+Each of the quantiles decrease across facets.
+With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for 10th, 25th and 50th percentile quantiles.
+
+```{r crickread, ech0=FALSE}
+cricket_tsibble <- gravitas::cricket %>%
+  mutate(data_index = row_number()) %>%
+  as_tsibble(index = data_index)
+```
+
+```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
+For all three cases, difference in runs between subsequent overs decrease."}
+
+library(gravitas)
+library(tsibble)
+
+cricket_bowling <- cricket_season %>% 
+  mutate(dot_ball_proxy = if_else(total_runs==0,1,0),
+         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
+  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+  summarise(dot_balls = sum(dot_ball_proxy),
+            wickets = sum(wicket_proxy), 
+            total_runs = sum(total_runs)) %>%
+            mutate(diff_run_rate = c(0,diff(total_runs)))
+
+cricket_tsibble_bow <- cricket_bowling %>% 
+  ungroup() %>% 
+  mutate(data_index = row_number()) %>% 
+  as_tsibble(index = data_index)
+
+cricket_tsibble_bow %>%
+  filter(dot_balls %in% c(0, 1, 2)) %>%
+  prob_plot("dot_balls",
+            "over",
+            hierarchy_model,
+            response = "diff_run_rate",
+            plot_type = "quantile",
+            quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
+            symmetric = FALSE)  + ggtitle("")
+
+```
+
+# Discussion {#sec:discussion}
+
+# Acknowledgements {-}
+
+# Bibliography
+
+<!-- # ```{r write-bib} -->
+<!-- # write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
+<!-- # ``` -->
+
+<!-- library(gravitas) -->
+<!-- library(tsibble) -->
+
+<!--  cricket_bowling <- cricket_season %>% -->
+<!--    mutate(dot_ball_proxy = if_else(total_runs==0,1,0), -->
+<!--           wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% -->
+<!--   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
+<!--    summarise(dot_balls = if_else(sum(dot_ball_proxy)>0, "yes","no"), -->
+<!--              wickets = if_else(sum(wicket_proxy)>0, "yes","no"),  -->
+<!--             total_runs = sum(total_runs)) %>% -->
+<!--             mutate(diff_run_rate = c(0,diff(total_runs))) -->
+<!--   -->
+<!--   -->
+<!--  cricket_tsibble_bow <- cricket_bowling %>%  -->
+<!--    ungroup() %>%  -->
+<!--    mutate(data_index = row_number()) %>%  -->
+<!--    as_tsibble(index = data_index) -->
+
+<!-- cricket_tsibble_bow %>% gran_obs("over", "dot_balls") -->
+
+<!-- # cricket_tsibble_bow %>%  -->
+<!-- #   #filter(dot_balls %in% c(0, 1, 2))%>%  -->
+<!-- #   prob_plot("dot_balls", -->
+<!-- #             "over", -->
+<!-- #             hierarchy_model, -->
+<!-- #             response = "diff_run_rate", -->
+<!-- #             plot_type = "quantile", -->
+<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
+<!-- #             symmetric = FALSE) -->
+
+<!-- cricket_tsibble_bow %>% gran_obs("over", "wickets") -->
+
+<!-- cricket_tsibble_bow %>%  -->
+<!--   #filter(wickets %in% c(0, 1))%>%  -->
+<!--   gather("proxies", "yes_no", -c(season, -->
+<!--                                      match_id, -->
+<!--                                      batting_team, -->
+<!--                                      bowling_team, -->
+<!--                                      inning,     -->
+<!--                                      over, -->
+<!--                                      diff_run_rate, -->
+<!--                                      data_index)) %>%  -->
+<!--   filter(yes_no %in% c("no")) %>%  -->
+<!--   prob_plot("proxies", -->
+<!--             "over", -->
+<!--             hierarchy_model, -->
+<!--             response = "diff_run_rate", -->
+<!--             plot_type = "boxplot", -->
+<!--             symmetric = FALSE, -->
+<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
+<!-- ``` -->
+
+<!-- ```{r proxies-yes} -->
+<!-- cricket_tsibble_bow %>%  -->
+<!--   #filter(wickets %in% c(0, 1))%>%  -->
+<!--   gather("proxies", "yes_no", -c(season, -->
+<!--                                      match_id, -->
+<!--                                      batting_team, -->
+<!--                                      bowling_team, -->
+<!--                                      inning,     -->
+<!--                                      over, -->
+<!--                                      diff_run_rate, -->
+<!--                                      data_index)) %>%  -->
+<!--   filter(yes_no %in% c("yes")) %>%  -->
+<!--   prob_plot("proxies", -->
+<!--             "over", -->
+<!--             hierarchy_model, -->
+<!--             response = "diff_run_rate", -->
+<!--             plot_type = "lv", -->
+<!--             symmetric = FALSE, -->
+<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
+
+<!-- ``` -->
+
+<!-- ```{r exwicketplot, echo=FALSE} -->
+<!-- #  -->
+<!-- # cricket_tsibble %>%  -->
+<!-- #   filter(dot_balls %in% c(0, 1, 2)) %>% -->
+<!-- #   mutate(lag_dot = lag(dot_balls)) %>%  -->
+<!-- #   filter(lag_dot %in% c(0, 1, 2)) %>% -->
+<!-- #   prob_plot("lag_dot", -->
+<!-- #             "over", -->
+<!-- #             hierarchy_model, -->
+<!-- #             response = "runs_per_over", -->
+<!-- #             plot_type = "quantile", -->
+<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
+
+<!-- cricket_tsibble %>%  -->
+<!--   filter(wicket %in% c(0, 1)) %>% -->
+<!--   mutate(lag_wicket = lag(wicket)) %>%  -->
+<!--   filter(lag_wicket %in% c(0, 1)) %>% -->
+<!--   prob_plot("lag_wicket", -->
+<!--             "over", -->
+<!--             hierarchy_model, -->
+<!--             response = "runs_per_over", -->
+<!--             plot_type = "quantile", -->
+<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
+
+<!-- ``` -->
 
@@
 
-```{r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."}
+<!-- r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."} -->
 # 
@@
 
-```
-
-Q3: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalizes runs in the subsequent over more?
-
-For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
-If a batsman is bowled out, it does not necessarily signify good fielding.
-So we only include catches and run out as a measure of fielding.
-Difference in run rates should be negative if fielding is good.
-Let us see if this fact is true.Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over.
-
-```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and dismissals by catch and catch and bowled."}
-cricket_fielding <- cricket_season %>% 
-  mutate(fielding_proxy = if_else(dismissal_kind %in% c("caught", "caught and bowled"),1,0)) %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs),
-            run_rate = sum(total_runs)/length(total_runs),
-            fielding_wckts = sum(fielding_proxy)) %>%
-            mutate(lagged_run_rate = c(0,diff(run_rate)))
-
-cricket_tsibble_fld <- cricket_fielding %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-cricket_tsibble_fld %>% gran_obs("over", "fielding_wckts")
-
-cricket_tsibble_fld %>% 
-  filter(fielding_wckts %in% c(0,1)) %>% 
-  prob_plot( "fielding_wckts","over",
-                               hierarchy_model,
-                               response = "lagged_run_rate",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-           overlay = TRUE)  + ggtitle("")
-
-```
-
-Q4: Is run rate set to reduce in overs where number of dot balls are more/number of wickets are more?
-
-A dot ball is a delivery bowled without any runs scored off it.
-The number of dot balls is reflective of the quality of bowling in the game.
-Run rate of an over should ideally decrease if the number of dot balls increase.
-However, what is the effect of dot balls in previous over on run rate in the subsequent over.
-Will players  batsman likely to go for big shots because they couldn't score good runs in the previous over?
-Or they should play consistently and avoid scoring high?
-
-Figure \ref{fig:dot} shows that for any over, increase in dot balls lead to decrease in run rate.
-
-```{r crickread}
-cricket_tsibble <- gravitas::cricket %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-```
-
-```{r exdotplot, echo=FALSE, eval = FALSE}
-library(gravitas)
-library(tsibble)
-
-cricket_bowling <- cricket_season %>% 
-  mutate(dot_ball_proxy = if_else(total_runs==0,1,0)) %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs),
-            run_rate = sum(total_runs)/length(total_runs),
-            dot_balls = sum(dot_ball_proxy)) %>%
-            mutate(diff_run_rate = c(0,diff(run_rate)))
-
-cricket_tsibble_bow <- cricket_bowling %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-cricket_tsibble_bow %>% gran_obs("over", "dot_balls")
-
-cricket_tsibble_bow %>% 
-  filter(dot_balls %in% c(0, 1, 2))%>% 
-  prob_plot("dot_balls",
-            "over",
-            hierarchy_model,
-            response = "diff_run_rate",
-            plot_type = "quantile",
-            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-
-```
-
-```{r exwicketplot}
-
-cricket_tsibble %>% 
-  filter(dot_balls %in% c(0, 1, 2)) %>%
-  mutate(lag_dot = lag(dot_balls)) %>% 
-  filter(lag_dot %in% c(0, 1, 2)) %>%
-  prob_plot("lag_dot",
-            "over",
-            hierarchy_model,
-            response = "runs_per_over",
-            plot_type = "quantile",
-            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-
-cricket_tsibble %>% 
-  filter(wicket %in% c(0, 1)) %>%
-  mutate(lag_wicket = lag(wicket)) %>% 
-  filter(lag_wicket %in% c(0, 1)) %>%
-  prob_plot("lag_wicket",
-            "over",
-            hierarchy_model,
-            response = "runs_per_over",
-            plot_type = "quantile",
-            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))
-
-```
-
-# Discussion {#sec:discussion}
-
-# Acknowledgements {-}
-
-# Bibliography
-
-```{r write-bib}
-write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib')
-```
-
 <!-- # ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."} -->
@@
 <!--            panel.grid.major.y = element_line( size=.1, color="black" )  -->
-<!--     ) -->
+<!--     )
 
@@
 <!--   ) -->
-
-<!-- ``` -->
 
@@
 
-<!-- ``` -->
-
 <!-- ```{r read_data_all_seas, echo=FALSE} -->
@@
 <!--   ) -->
-<!-- ``` --
-```

### 2019-11-24 main.Rmd
smart meter improving

@@
 
-In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
+In \autoref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
@@
 
-The interval of this tsibble is 30 minutes, and hence the default in this case, provides temporal granularities ranging from half-hour to year.
-If these options are considered too many, the default options can be modified to limit the possibilities.
-For example, the most coarse temporal unit can be set to be a “month”.
+The interval of this tsibble is 30 minutes, and hence the temporal granularities range from half-hour to year.
+If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
 
@@
 
-This looks better. However, some intermediate temporal units might not be pertinent to the analysis and we might want to remove them from the list of granularities.
+Also, some intermediate temporal units that might not be pertinent to the analysis and we might want to remove them from the list of granularities.
 
@@
 From the search list, we found that we can look at six granularities, that amounts to analyzing six graphics.
-However, what happens if we want to see the distribution of energy across two granularities at a time?
+However, the distribution of energy needs to be visualized across two granularities at a time.
 This is equivalent to looking at the distribution of energy consumption across one granularity conditional on another one.
@@
 
-So, what is the number of pairs of granularities we can look at?
+What is the number of pairs of granularities we can look at?
 It is equivalent to taking 2 granularities from 6, which essentially means we need to examine 30 plots.
 The good news is, not all time granularities can be plotted together and we do not have to analyze so many plots!
-
 Harmony/clash can be identified to considerably reduce the number of visualizations that can aid exploratory analysis.
@@
 
-In \ref{fig:bothcust} We visualize the harmony pair (wknd_wday, hour_day) through a box plot.
+In \autoref{fig:bothcust} We visualize the harmony pair (wknd_wday, hour_day) through a box plot.
 Boxplot of energy consumption is shown across wknd_wday (facet) and hour-day (x-axis) for the same two households.
@@
 
-```{r}
+```{r bothcust}
 cust1 <- smart_meter10 %>% 

### 2019-11-24 main.Rmd
spell check

@@
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-One of the first attempts of characterising these granularities occur in [@Bettini1998-ed].
+One of the first attempts of characterizing these granularities occur in [@Bettini1998-ed].
 The definitions and rules defined are inadequate to reflect periodicities in time.
@@
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-We propose that the analysis of smart meter data can be benefitted from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
+We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
 However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analysed for different periodicities.
@@
 The remainder of the paper is organized as follows.
-Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterisation of circular and aperiodic time granularities.
+Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterization of circular and aperiodic time granularities.
 Section \ref{sec:synergy} discusses the effect in data structure due to how pairs of cyclic time granularities are related.
@@
 In a cyclic organization of time, the domain is composed of a set of recurring time values.
-Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday).
+Hence, any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
 Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
@@
 The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities.
-However, the formalisms would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
+However, the formalism would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
 
@@
 
-<!-- Q2: Amongst good fielding and bowling - which affect the runs of the subsequent overs more? -->
+<!-- Q2: Among good fielding and bowling - which affect the runs of the subsequent overs more? -->
 

### 2019-11-25 main.Rmd
cleaning

@@
 
-```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="\\label{fig:linear-time} The time domain distributed as linear granularities"}
+```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
 knitr::include_graphics("Figs/linear-time.png")
@@
 
-```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="\\label{fig:circular-gran} Relating circular granularities and bottom linear granularity "}
+```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
 knitr::include_graphics("Figs/circular-gran-diagram.png")
@@
 
-\begin{equation}\label{eq6}
+\begin{equation}\label{eq3}
 A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for z \in T_k
@@
 
-```{r aperiodic-gran, echo=FALSE,out.width = "100%",fig.cap="\\label{fig:circular-gran} Relating circular granularities and bottom linear granularity "}
+```{r aperiodic-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
 knitr::include_graphics("Figs/aperiodic-granv1.png")
@@
 
-\begin{equation}\label{eq7}
+\begin{equation}\label{eq6}
 A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
@@
 
-\begin{equation} \label{eq3}
+\begin{equation} \label{eq7}
 \begin{split}
@@
 
-\begin{equation} \label{eq4}
+\begin{equation} \label{eq8}
 \begin{split}
@@
 
-\begin{equation} \label{eq5}
+\begin{equation} \label{eq9}
 C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
@@
 
-\begin{equation} \label{eq6}
+\begin{equation} \label{eq10}
 \begin{split}
@@
 
-\begin{equation} \label{eq1}
+\begin{equation} \label{eq11}
 \begin{split}
@@
 
-Suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
+Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
 That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
@@
 
-### Unevenly distributed events (TODO)
+### Unevenly distributed events
 
@@
 This might happen due to missing observations in the data or uneven locations of events in time domain.
-We use Gini's measure to compute if differences in number of observations across or within facets are significantly different.
+In such cases, the visualizations that use kernel density estimates should be used with caution as sample size would directly affect both the variance of the estimator and the confidence intervals.
+Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels.
 
@@
 
-In \autoref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
+In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
@@
 
-```{r day-fortnight_agg, echo = FALSE}
-library(lvplot)
-library(ggplot2)
-library(dplyr)
-library(tibble)
-library(gravitas)
-library(tidyr)
-# smart_meter10 %>% 
-#   filter(customer_id %in% c(10006704, 10017936)) %>% 
-#   create_gran("day_fortnight") %>%
-#   as_tibble() %>% 
-#   group_by(customer_id, day_fortnight) %>% 
-#   summarise(median= log(median(general_supply_kwh)),
-#             total = log(sum(general_supply_kwh))) %>% 
-#   gather("daily_consumption", "value", - c(customer_id, day_fortnight)) %>% 
-#   ggplot2::ggplot(aes(
-#     x = as.factor(day_fortnight),
-#     y = value, 
-#     color = daily_consumption,
-#     group = daily_consumption)) +
-#   geom_line() +
-#   xlab("day_fortnight") +
-#   ylab("general_supply_kwh") + 
-#   facet_wrap(~customer_id) + 
-#   theme_bw()
-
-# smart_meter10 %>% 
-#   filter(customer_id %in% c(10006704, 10017936)) %>% 
-#   create_gran("day_fortnight") %>%
-#   as_tibble() %>% 
-#   group_by(customer_id, day_fortnight) %>% 
-#   summarise(median_consumption= median(general_supply_kwh)) %>%
-#   ggplot2::ggplot(aes(
-#     x = as.factor(day_fortnight),
-#     y = median_consumption,
-#     color =  customer_id,
-#     group = customer_id)) +
-#   geom_line() +
-#   xlab("day_fortnight") +
-#   ylab("general_supply_kwh") + 
-#   theme_bw() 
-```
-
-```{r day-fortnight}
+<!-- # ```{r day-fortnight_agg, echo = FALSE} -->
+<!-- # library(lvplot) -->
+<!-- # library(ggplot2) -->
+<!-- # library(dplyr) -->
+<!-- # library(tibble) -->
+<!-- # library(gravitas) -->
+<!-- # library(tidyr) -->
+<!-- # smart_meter10 %>%  -->
+<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
+<!-- #   create_gran("day_fortnight") %>% -->
+<!-- #   as_tibble() %>%  -->
+<!-- #   group_by(customer_id, day_fortnight) %>%  -->
+<!-- #   summarise(median= log(median(general_supply_kwh)), -->
+<!-- #             total = log(sum(general_supply_kwh))) %>%  -->
+<!-- #   gather("daily_consumption", "value", - c(customer_id, day_fortnight)) %>%  -->
+<!-- #   ggplot2::ggplot(aes( -->
+<!-- #     x = as.factor(day_fortnight), -->
+<!-- #     y = value,  -->
+<!-- #     color = daily_consumption, -->
+<!-- #     group = daily_consumption)) + -->
+<!-- #   geom_line() + -->
+<!-- #   xlab("day_fortnight") + -->
+<!-- #   ylab("general_supply_kwh") +  -->
+<!-- #   facet_wrap(~customer_id) +  -->
+<!-- #   theme_bw()  -->
+
+<!-- # smart_meter10 %>%  -->
+<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
+<!-- #   create_gran("day_fortnight") %>% -->
+<!-- #   as_tibble() %>%  -->
+<!-- #   group_by(customer_id, day_fortnight) %>%  -->
+<!-- #   summarise(median_consumption= median(general_supply_kwh)) %>% -->
+<!-- #   ggplot2::ggplot(aes( -->
+<!-- #     x = as.factor(day_fortnight), -->
+<!-- #     y = median_consumption, -->
+<!-- #     color =  customer_id, -->
+<!-- #     group = customer_id)) + -->
+<!-- #   geom_line() + -->
+<!-- #   xlab("day_fortnight") + -->
+<!-- #   ylab("general_supply_kwh") +  -->
+<!-- #   theme_bw()  -->
+<!-- ``` -->
+
+```{r day-fortnight, fig.cap="Letter value plot of two households across days of the fortnight.
+M, F, E, D and C represents the letter values of energy consumption.
+Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household."}
 library(lvplot)
@@
 
-While trying to explore the energy behavior of these customers systematically across time granularities, the first thing we should have at our disposal is to know the number of time granularities we can look at exhaustively.
+**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
 If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
@@
 
-Also, some intermediate temporal units that might not be pertinent to the analysis and we might want to remove them from the list of granularities.
+Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
 
@@
 
-Now that we have the list of granularities to look at, we can visualize the distribution.
-From the search list, we found that we can look at six granularities, that amounts to analyzing six graphics.
-However, the distribution of energy needs to be visualized across two granularities at a time.
-This is equivalent to looking at the distribution of energy consumption across one granularity conditional on another one.
-One way can be to plot one of the granularities on the x-axis and another on the facet.
-Different perspectives of the data can be derived depending on where the granularities are placed.
-
-What is the number of pairs of granularities we can look at?
-It is equivalent to taking 2 granularities from 6, which essentially means we need to examine 30 plots.
-The good news is, not all time granularities can be plotted together and we do not have to analyze so many plots!
-Harmony/clash can be identified to considerably reduce the number of visualizations that can aid exploratory analysis.
-
-```{r is_harmony, echo= TRUE}
-smart_meter10 %>% 
-  is_harmony(gran1 = "hour_day", 
-             gran2 = "day_week")
-
-smart_meter10 %>%
-  is_harmony(gran1 = "hour_day", 
-             gran2 = "day_week", 
-             facet_h = 14)
-
-smart_meter10 %>% 
-  is_harmony(gran1 = "day_month",
-             gran2 = "week_month")
-```
-
-Let us now look at all the harmonies that we can examine. Fortunately, we are left with only 13 out of 30 visualizations.
+Now that we have the list of cyclic granularities to look at, we can visualize the distribution.
+From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
+The distribution of energy needs to be visualized across two cyclic granularities at a time.
+It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.However, harmony/clash pairs can identified among those 30 pairs, to determine feasibility of plotting any pairs together.
+We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
+
+<!-- # ```{r is_harmony, echo= TRUE} -->
+<!-- # smart_meter10 %>%  -->
+<!-- #   is_harmony(gran1 = "hour_day",  -->
+<!-- #              gran2 = "day_week") -->
+<!-- #  -->
+<!-- # smart_meter10 %>% -->
+<!-- #   is_harmony(gran1 = "hour_day",  -->
+<!-- #              gran2 = "day_week",  -->
+<!-- #              facet_h = 14) -->
+<!-- #  -->
+<!-- # smart_meter10 %>%  -->
+<!-- #   is_harmony(gran1 = "day_month", -->
+<!-- #              gran2 = "week_month") -->
+<!-- # ``` -->
 
@@
 )
-
 smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
@@
 
-In \autoref{fig:bothcust} We visualize the harmony pair (wknd_wday, hour_day) through a box plot.
+In \ref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.
 Boxplot of energy consumption is shown across wknd_wday (facet) and hour-day (x-axis) for the same two households.
@@
 
-```{r bothcust}
+```{r bothcust_notrun}
 cust1 <- smart_meter10 %>% 
@@
 
-```{r bothcust}
+```{r bothcust, fig.cap = "Boxplot of energy consumption across hours of the day faceted by weekday/weekend for household 1 (customer id: 10006704) and household 2 (customer id: 10017936).
+Median behavior for the early morning hours of household 2 is extremely high for weekends compared to weekdays, not much differenc can be observed for household 1.
+"}
 cust1
@@
 
-Figure \autoref{fig:wknd-day50} shows the quantile plot of energy consumption of 50 households.
+Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of 50 households.
 Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends.
 
-Figure \autoref{fig:my-hd} shows the area quantile plots of energy consumption across hours of the day faceted by months of the year.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-It can be observed that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
+Figure \ref{fig:my-hd} shows that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
 Moreover, only the pink band changes significantly in winter months (May - August).
@@
 Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that off course the level of quantiles increased too in winter months, the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
-
-```{r wknd-wday50}
+It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather considtions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+
+```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
+The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."}
+
 smart_meter <- 
@@
 
-```{r my-hd, echo=TRUE}
+```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
+The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
+"}
 
@@
 <!-- ``` -->
+
+This case study shows systematic exploration of energy behavior starting with two households and comparing some of their features with all 50 households.
+First, it helps us to find the list of granularities to look at, then shrinks the number of possible visualizations by identifying harmonies and then explored harmony pairs to gain some insights on periodic behavior of the households.
 
@@
 
-From figure \autoref{cricex}, it can be observed that there is no clear upward shift in runs in the second innings as compared to the first innings.
+From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in runs in the second innings as compared to the first innings.
 The variability of runs also increases as the teams approach towards the end of the innings, as observed through the longer and more distinct letter values.
@@
 Or they should play consistently and avoid scoring high?
-From figure \autoref{exdot} shows the quantile plot of runs across overs for none, one or two dot balls per over in the facets.
+From figure \ref{fig:exdot} shows the quantile plot of runs across overs for none, one or two dot balls per over in the facets.
 Each of the quantiles decrease across facets.
@@
 ```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
-For all three cases, difference in runs between subsequent overs decrease."}
+For all three cases, difference in runs between subsequent overs decrease.", fig.pos="htb"}
 
@@
 <!-- r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."} -->
-# 
-# 
-# 
-# MI <- cricket_tsibble_all %>%
-#   filter(season %in% c(2010:2015), batting_team=="Mumbai Indians")
-# 
-# 
-# CSK <- cricket_tsibble_all %>%
-#   filter(season %in% c(2010:2015),
-#          batting_team=="Kolkata Knight Riders",
-#          inning %in% c(1,2))
-# 
-# 
-#  MI %>%
-#    gran_obs("inning", "over", hierarchy_model) %>%
-#    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
-# 
-#  CSK %>%
-#    gran_obs("inning", "over", hierarchy_model) %>%
-#    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015")
-# 
-# MI %>%
-#   prob_plot("inning", "over",
-#                                hierarchy_model,
-#                                response = "run_rate",
-#                                plot_type = "lv",
-#                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")
-# 
-# 
-# CSK %>%
-#   prob_plot("inning", "over",
-#                                hierarchy_model,
-#                                response = "run_rate",
-#                                plot_type = "lv",
-#                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")
-# 
-# cricket_tsibble_all %>% distinct(batting_team)
+<!-- #  -->
+<!-- #  -->
+<!-- #  -->
+<!-- # MI <- cricket_tsibble_all %>% -->
+<!-- #   filter(season %in% c(2010:2015), batting_team=="Mumbai Indians") -->
+<!-- #  -->
+<!-- #  -->
+<!-- # CSK <- cricket_tsibble_all %>% -->
+<!-- #   filter(season %in% c(2010:2015), -->
+<!-- #          batting_team=="Kolkata Knight Riders", -->
+<!-- #          inning %in% c(1,2)) -->
+<!-- #  -->
+<!-- #  -->
+<!-- #  MI %>% -->
+<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
+<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
+<!-- #  -->
+<!-- #  CSK %>% -->
+<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
+<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
+<!-- #  -->
+<!-- # MI %>% -->
+<!-- #   prob_plot("inning", "over", -->
+<!-- #                                hierarchy_model, -->
+<!-- #                                response = "run_rate", -->
+<!-- #                                plot_type = "lv", -->
+<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") -->
+<!-- #  -->
+<!-- #  -->
+<!-- # CSK %>% -->
+<!-- #   prob_plot("inning", "over", -->
+<!-- #                                hierarchy_model, -->
+<!-- #                                response = "run_rate", -->
+<!-- #                                plot_type = "lv", -->
+<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") -->
+<!-- #  -->
+<!-- # cricket_tsibble_all %>% distinct(batting_team) -->
 
@@
 <!--   ) -->
+
+```{r write-bib, eval = FALSE}
+write_bib(c("dplyr", "tidyr", "tsibble", "magrittr","rlang", "gravitas"), file = "rpkgs.bib")
+```

### 2019-11-25 main.Rmd
spell check rough draft

@@
 
-Calendar-based graphics[@wang2019tsibble] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
+Calendar-based graphics[@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
 Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time.
@@
 
+\newtheorem{mydef}{Definition}
+\begin{mydef}\label{def:quasi}
 **Definition: Quasi Periodical** A granularity H is quasi-periodical with respect to a granularity G if
@@
 between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
+\end{mydef}
 
@@
 
-We propose a formalism of cyclic time granularities through the tsibble [@Wang2019-aw] framework of organizing temporal data.
+We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
 A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble.
@@
 \caption{Illustrative circular granularities with time index in minutes}
-\label{definitions}
+\label{tab:definitions}
 \end{table}
@@
 \caption{Illustrative aperiodic circular granularities with potential period lengths}
-\label{aperiodic}
+\label{tab:aperiodic}
 \end{table}
@@
 
-\begin{equation}\label{eq3}
+\begin{equation}\label{eq:eq3}
 A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for z \in T_k
@@
 Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different.
-A(BG, AG) and A(BG, AH) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH.
+A(BG, AG) and A(BG, AH) are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH.
 For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
@@
 
-\begin{equation} \label{eq4}
+\begin{equation} \label{eq:eq4}
 \begin{split}
-A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , since 8 \in T_{3} \\
+A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
   & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
@@
   & = BG(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
-  & = BG(8 - 2.2 - 3) \\
+  & = BG(8 - 2*2 - 3) \\
   & = BG(1) \\
@@
 
-\begin{equation} \label{eq5}
+\begin{equation} \label{eq:eq5}
 \begin{split}
-A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) , since 10 \in T_{1}  \\
+A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
   & = BG(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
@@
 Considering leap years and all their exceptions, period is 400 years and hence $P' = (3*365 + 366)$, $R' = 12*400$.
-\autoref{eq:eq6} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
+\autoref{eq:eq3} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
 
@@
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
-It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \ref{eq:eq6} can be modified as follows to account for exceptions $E = [e_begin, e_end]$ (\ref:{def:something})
+It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \ref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (\autoref{def:quasi})
 
@@
 
-**Definition:  Order** of a linear granularity can be comprehended as the level graininess asscociated with a linear granularity.
+**Definition:  Order** of a linear granularity can be comprehended as the level graininess associated with a linear granularity.
 If we consider two linear granularities G and H, such that $G \preceq H$, then H is of higher order than G.
@@
 
-Let us use the equation \ref{eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.
+Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.
 
@@
 
-**Example:** Considering the same example of Mayan Calendar,  it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq5}
+**Example:** Considering the same example of Mayan Calendar,  it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:eq5}
 
@@
 Let us turn to Gregorian calendar for addressing this case.
-Suppose we have a hierarchy table for  Gregorian calendar as \ref{table_accessor2}.
+Suppose we have a hierarchy table using some linear granularties from Gregorian calendar.
 Since months consists of unequal number of days, any temporal unit which is higher in order than months will also have unequal number of days.
@@
 Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some famailies use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
 -->
@@
 Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather considtions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
 

### 2019-11-25 cricket_per_over.R

@@ -1,22 +1,0 @@
-cricket_per_over <- cricketdata %>% 
-  group_by(match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs)) 
-
-cricket_tsibble <- cricket_per_over %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-
-harmony(cricket_tsibble, lowest_unit = "over", highest_unit = "match", hierarchy_tbl =  hierarchy_model)
-cricket_per_over %>% 
-  
-  
-  cricket_tsibble %>% granplot("inning_match", "over_inning",
-                               hierarchy_model,
-                               response = "runs_per_over",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-                               
-                               
-  )

### 2019-11-25 main.Rmd
spell check rough draft

@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in [@Wickham2009-pk].
+This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2[@Wickham2009-pk].
 bibliography: [bibliography.bib]
@@
 
-Temporal data are available at various resolution depending on the context.
-Social and economic data like GDP are often collected and reported at coarse temporal scales like monthly, quarterly or annually.
+Temporal data are available at various resolutions depending on the context.
+Social and economic data like GDP is often collected and reported at coarse temporal scales like monthly, quarterly or annually.
 With recent advancement in technology, more and more data are recorded at much finer temporal scales.
 Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be recorded every second.
-As the frequency of data increases, the number of questions about periodicity of the observed variable also increases.
+As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases.
 For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
 This approach requires deconstructing time in various possible ways called time granularities[@aigner2011visualization].
-It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on periodicity of the observed data.
+It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
 This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
@@
 There can be several classes of time deconstructions, viz.
-based on arrangement (linear vs.
-cyclic) or hierarchical order of calendar.
+based on the arrangement (linear vs.
+cyclic) or hierarchical order of the calendar.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
@@
 
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world.
+The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
 Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates of usage [@2012-la] across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates of usage across households (for example, [@2012-la]), total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
 We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
-However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analysed for different periodicities.
+However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analyzed for different periodicities.
 
@@
 
-# Definitions of time granularities{#sec:timegrandef}
+# Time granularities{#sec:timegrandef}
 
@@
 
-Providing a formalism to some of these abstractions is important to model a time series across differently grained temporal domains.
+In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.
 
@@
 (iii) the indexes of first and last granules in H, if their value is not infinite. 
-Then, if ${S_0,...,S_(R-1)}$ are the sets of indexes of G describing
+Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
 $H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
@@
 
-\begin{equation} \label{eq2}
+\begin{equation} \label{eq:eq2}
 C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
@@
 
-All circular granularities can be expressed in terms of the bottom linear granularity with equation \ref{eq2}.
+All circular granularities can be expressed in terms of the bottom linear granularity with equation \autoref{eq:eq2}.
 
@@
 The modulus for these type of calendar categorizations are not constant due to irregular mapping with bottom granularities.
-For example, table \ref{tab:aperiodic} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
+For example,\autoref{tab:aperiodic} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
 The coarse granularity is assumed as the period within which the finer granularity will repeat its behavior aperiodically.
@@
 For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
-For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \ref{eq:eq4} and  \ref{eq:eq5}.
+For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
 
@@
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
-It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \ref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (\autoref{def:quasi})
+It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (\autoref{def:quasi})
 
@@
 for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T(w)$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
-
-but the grouping is always periodic.
 
@@
 
-In \ref{sec:singlefrommultiple} and \ref{sec:multiplefromsingle}, we assume that all single order granularities are circular.What happens when we have a mix of circular and aperiodic granularities.
+In \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}, we assume that all single order granularities are circular.What happens when we have a mix of circular and aperiodic granularities.
 
@@
 
-The single order-up granularites resulting from the first two cases are periodic and has been handled in \ref{sec:singlefrommultiple} and \ref{sec:multiplefromsingle}.
+The single order-up granularites resulting from the first two cases are periodic and has been handled in \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}.
 The calendar categorization resulting from the last case are aperiodic.
@@
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
-However, there is additional information which we can derive looking at the distribution.
+However, there is additional information that we can derive looking at the distribution.
 If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
@@
 
-In \ref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.
+In \autoref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.
 Boxplot of energy consumption is shown across wknd_wday (facet) and hour-day (x-axis) for the same two households.

### 2019-11-25 main.Rmd
definition reference changed

@@
 Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
-
-## Arrangement: Linear vs. Circular vs. aperiodic
-
-The arrangement of the time domain can result in deconstructing time in linear, circular or nearly circular ways.
+Classes of time abstractions may be based on arrangement or hierarchical order of time, discussed in further details in \autoref{sec:arrangement} and \autoref{sec:order}.
+
+## Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement}
+
+The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.
 Time granularities are **linear** if they respect the linear progression of time.
@@
 However, periodicity is very common in all kinds of data.
-Time granularities, which are constructed by using two linear granularities, can be utilized to explain such periodicities.
+Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
 The mappings between linear granularities can be regular or irregular.
@@
 
-There has been several attempts to provide the framework for formally characterizing time-granularities and identifying their structural properties, relationships and symbolic representations.
+There has been several attempts to provide the framework for formally characterizing time-granularities.
 One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
 
-**Definition:** A **time domain** is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+\newtheorem{timedom}{Definition}
+\begin{timedom}\label{def:timedom}
+A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+\end{timedom}
 
@@
 
-**Definition:** A linear **granularity** is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
+\begin{timedom}\label{def:linear}
+A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
 
@@
   (C2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
-
-**Definition:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
-
+\end{timedom}
+
+**Dicussion:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
 The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
@@
 
-**Definition: Finer Than** : A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
+\begin{timedom}\label{def:finerthan}
+A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
 for each index i, there exists an index j such that $G(i) \subset H(j)$.
-
-**Definition: Groups Into** : A linear granularity G groups into a linear granularity H, denoted
+\end{timedom}
+
+\begin{timedom}\label{def:groupsinto}
+A linear granularity G groups into a linear granularity H, denoted
 $G \trianglelefteq H,$
@@
 \end{equation}
+\end{timedom}
 
@@
 
+\begin{timedom}\label{def:linear}
 **Definition: Groups periodically into** A granularity H is periodical with respect to a granularity G if
@@
 $H (i + R) = \bigcup_{j \in S} G(j + P)$.
+\end{timedom}
 
@@
 
-\newtheorem{mydef}{Definition}
-\begin{mydef}\label{def:quasi}
+\begin{timedom}\label{def:quasi}
 **Definition: Quasi Periodical** A granularity H is quasi-periodical with respect to a granularity G if
@@
 between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
-\end{mydef}
+\end{timedom}
 
@@
 
+\begin{timedom}\label{def:BG}
 **Definition: Bottom granularity** Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
+\end{timedom}
 
@@
 
-## Order: Single vs. Multiple
+## Order: Single vs. Multiple {#sec:order}
 

### 2019-11-25 main.Rmd
fine tune linear gran section

@@
 \autoref{fig:linear-time} shows the implication of this condition.
-If we consider the bottom linear granularity [@aigner2011visualization] as hourly and the entire horizon has T hours then it will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+If we consider the chronon [@aigner2011visualization] as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
 
@@
 ```
-
-The definitions and rules for linear granularities are inadequate to reflect periodicities in time, like weekly, monthly or yearly seasonality.
-Hence, there is a need to define circular time granularities in a different approach.
 
@@
 
-According to this definition, $G \trianglelefteq H,$ if each granule H (i) is the union of some
+According to this definition, $G \trianglelefteq H,$ if each granule H (j) is the union of some
 granules of G.
 For example, $Days \trianglelefteq Months,$.
-This relationship, however, is not sufficient to fully describe a granularity in terms of another one.
-For example, it is clear that $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping.
+However, this  relationship is not fully described until we associate it with periodicity.
+For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping.
 If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
-Considering leap years and all their exceptions, the period becomes 400 years, but the grouping is always periodic.
-In order to define a granularity in terms of another granularity including the notion of periodicity, a particular case of the groups into relationship is considered.
-
+Considering leap years and all their exceptions, the period becomes 400 years.
 \begin{timedom}\label{def:linear}
-**Definition: Groups periodically into** A granularity H is periodical with respect to a granularity G if
+A granularity H is periodical with respect to a granularity G if
 (1) $G \trianglelefteq H,$, and
@@
 \begin{timedom}\label{def:quasi}
-**Definition: Quasi Periodical** A granularity H is quasi-periodical with respect to a granularity G if
+A granularity H is quasi-periodical with respect to a granularity G if
 (1) $$G \trianglelefteq H,$$, and

### 2019-11-25 main.Rmd
arrangment section fine tune

@@
 granules of G.
-For example, $Days \trianglelefteq Months,$.
+For example, $Days \trianglelefteq Months$.
 However, this  relationship is not fully described until we associate it with periodicity.
@@
 
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf]. It is assumed that there exists a “bottom”
-granularity and Calendar algebra operations are designed to generate
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf].
+It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
 new granularities from the bottom one or recursively, from those already generated.
@@
 Example can be to consider a calendar with only two linear granularities minute and hour and hour is generated by grouping every 60 minutes.
-The
-granule-oriented operations do not change the granules of a granularity, but rather make
-choices of which granules should remain in the new granularity.
+The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
 For example, one can choose to look at the granularity "Monday" and hence select only Mondays while looking at the linear granularity "day".
@@
 - **The grouping operation** : Let G be a full-integer labeled granularity, and m a positive integer.
-The grouping operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G
-into m-granule groups and making each group a granule of the resulting granularity.
+The grouping operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G into m-granule groups and making each group a granule of the resulting granularity.
 More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
@@
 
-**Definition: Circular granularity** A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if
+\begin{timedom}\label{def:circular}
+A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if
 
@@
 k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.  
-
-**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram \autoref{fig:circular-gran}.
+\end{timedom}
+
+**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
 Each granule is represented by a box.
-The diagram also illustrate that the granules that overlap share elements from the underlying time domain.
+The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
 The first slot in the diagram shows the index set and the bottom granularity.
@@
 \begin{equation}\label{eq:eq3}
-A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for z \in T_k
+A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
 \end{equation}
@@
 R' denotes the number of granules of AG in each period of BG,  
-$T_w$ are the sets of indices of BG describing $AG_w$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
+$T_w$ are the sets of indices of BG describing $AG(w)$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
 $\vert T_w \vert$ is the cardinality of set $T_w$.
@@
 Two linear granularities AG and AH are considered, where $AG = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$.
-This implies that AG is made up of shrinking every 1st granule of Group_3(BG) by 1 and AH is made up of shrinking every 1st granule of Group_3(BG) by 2.
+This implies that AG is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and AH is made up of shrinking every 1st granule of Group_3(BG) by 2 granules.
 Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different.
@@
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
-It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (\autoref{def:quasi})
+It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi})
 
@@
 
-for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T(w)$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
+for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
 

### 2019-11-25 main.Rmd
order fine tune till eq 7

@@
 
-**Definition:  Order** of a linear granularity can be comprehended as the level graininess associated with a linear granularity.
-If we consider two linear granularities G and H, such that $G \preceq H$, then H is of higher order than G.
+**Definition:  Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
+If we consider two linear granularities G and H, such that G is finer than or groups into H (by Definitions \ref{def:finerthan}, \ref{def:groupsinto}), then H is of higher order than G.
 
 **Notation:  Hierarchy table** Let $H_n: (G, C, k)$ be a hierarchy model containing n linear granularities.
-$G_{l}$ represents the linear granularity of order l and $k_{l,m}$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
-
-In a hierarchy table, linear granularities are arranged from highest to lowest order of linear granularities.
+$G_{l}$ represents the linear granularity of order l and $k(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
+
+In a hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
 
@@
 \begin{split}
-C_{(G_l,G_m)}(z) & = C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1) \\  
-  & =   C_{(G_l,G_(l+1)}(z) + k(l, l+1)[ C_{G_(l+1),G_(l+2)}(z) + k(l+1, l+2)( C_{G_(l+2), G_m)}(z) - 1) - 1]\\
-  & =  C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{G_(l+1),G_(l+2)}(z) - 1) + k(l, l+1)k(l+1, l+2)(C(l+2, m) - 1)\\
-  & =  C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{G_(l+1),G_(l+2)}(z) - 1) + k(l, l+2)(C_{G_(l+2),G_(l+m)}(z) - 1)\\
+C_{(G_l,G_m)}(z) & = C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1) \\  
+  & =   C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1] \\
+  & =  C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1)\\
+  & =  C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1)\\
   &\vdots\\
-  & = \sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_(l+i),G_(l+i+1)}(z) - 1)\\
+  & = \sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1)\\
 \end{split}
@@
 Nevertheless, for any calendar a hierarchy can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows (:
 
@@
 
-Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.
+Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic and regular.
 

### 2019-11-25 main.Rmd
checking notations for computation

@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates of usage across households (for example, [@2012-la]), total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates [@2012-la],  of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
@@
 Considering leap years and all their exceptions, the period becomes 400 years.
-\begin{timedom}\label{def:linear}
+\begin{timedom}\label{def:periodical}
 A granularity H is periodical with respect to a granularity G if
@@
 
-  - pseudomonth = Alter^{12}_{(11,-1)}(day, Alter^{12}_{(9,-1)}(day, Alter^{12}_{(6,-1)}(day,
-                  Alter_^{12}_{(4,-1)}(day, Alter^{12}_{(2,-3)}(day, Group_31(day)))))),
+  - pseudomonth = 
+  $Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, Alter^{12}_{6,-1}(day,   Alter^{12}_{4,-1)}(day, Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))$,
                   where the granularity pseudomonth is generated by grouping 31 days, 
@@
 Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
-Two linear granularities AG and AH are considered, where $AG = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$.
+Two linear granularities AG and AH are considered, where $AG = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$.
 This implies that AG is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and AH is made up of shrinking every 1st granule of Group_3(BG) by 2 granules.
@@
 Nevertheless, for any calendar a hierarchy can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows (:
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows [@Reingold2001-kf]:
 

### 2019-11-27 main.Rmd
split time granularity sections

@@
 
-# Time granularities{#sec:timegrandef}
+# Linear Time granularities {#sec:lineartime}
 
 <!-- general calendar categorization -->
-Often we partition time into months, weeks or days to relate it to data.
+<!-- Often we partition time into months, weeks or days to relate it to data.
 Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
@@
 Classes of time abstractions may be based on arrangement or hierarchical order of time, discussed in further details in \autoref{sec:arrangement} and \autoref{sec:order}.
-
-## Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement}
-
-The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.
+-->
+
+Often we partition time into months, weeks or days to relate it to data.
+Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
 Time granularities are **linear** if they respect the linear progression of time.
 Examples include hours, days, weeks and months.
-However, periodicity is very common in all kinds of data.
-Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
-The mappings between linear granularities can be regular or irregular.
-For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
-On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
-Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.
+
+## Definitions and Relationships
+
+<!-- Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement} -->
+
+<!-- The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.  -->
+
+ There has been several attempts to provide the framework for formally characterizing time-granularities.
+One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
 
@@
 
-In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.
-
-### Linear {#sec:linear-gran-def}
-
-There has been several attempts to provide the framework for formally characterizing time-granularities.
-One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
+<!-- In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.  -->
+
+<!-- ### Linear {#sec:linear-gran-def} -->
 
@@
 
-#### Relationships between two linear granularities and association to Periodicity
-
 [@Bettini1998-ed] talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
@@
 for each index i, there exists an index j such that $G(i) \subset H(j)$.
-\end{timedom}
+\end{timedom}.
+
+**Dicussion:**
 
@@
 
-According to this definition, $G \trianglelefteq H,$ if each granule H (j) is the union of some
+**Dicussion:** According to this definition, $G \trianglelefteq H,$ if each granule H (j) is the union of some
 granules of G.
@@
 Considering leap years and all their exceptions, the period becomes 400 years.
+
 \begin{timedom}\label{def:periodical}
@@
 
-This formula can be explained observing that H (j ) is either one of H (0), .
+**Dicussion:** This formula can be explained observing that H (j ) is either one of H (0), .
 .
@@
 
+**Dicussion:**
+
 Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by R and P.
@@
 \begin{timedom}\label{def:BG}
-**Definition: Bottom granularity** Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
+Bottom granularity - Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
 \end{timedom}
 
-**Example:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+**Dicussion:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
 
@@
 
-#### Computation of linear time granularities
+## Computation through Calendar Algebra
 
@@
 More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
-each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
+each integer $i$, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
 
@@
 
-- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
+- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and $l$, $k$, $m$ integers, where G2 partitions G1, and $1 \leq l \leq m$.
 The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
+The altering-tick operation modifies the granules of G1 so that the $l$th granule of each group has $|k|$ additional (or fewer when k < 0) granules of G2.
 
@@
 
-### Cyclic {#sec:cyclic-gran-def}
+# Cyclic Time granularities {#sec:circular-gran-def}
+
+Periodicity is very common in all kinds of data.
+Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
+The mappings between linear granularities can be regular or irregular.
+For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
+On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
+Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
+Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.
 
@@
 
-#### Circular {#sec:circular-gran-def}
+## Circular {#sec:circular-gran-def}
 
@@
 C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
-BG(z) \neq \phi
+BG(z) \neq \phi,
 \end{equation}
 where,  
-z denotes the index set.  
-BG denotes the index of the tsibble (bottom granularity).  
-G be a linear granularity, periodic with respect to BG.    
+z denotes the index set, 
+BG denotes the index of the tsibble (bottom granularity), 
+G be a linear granularity, periodic with respect to BG,     
 k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.  
@@
 
-### Aperiodic {#sec:aperiodic-gran-def}
+## Quasi-circular {#sec:circular-gran-def}
 
@@
 z denotes the index set (index of BG),
-R' denotes the number of granules of AG in each period of BG,  
+$R'$ denotes the number of granules of AG in each period of BG,  
 $T_w$ are the sets of indices of BG describing $AG(w)$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
@@
 
+# Acyclic time granularities {#sec:circular-gran-def}
+
 <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
 
-## Order: Single vs. Multiple {#sec:order}
+# Computing cyclic time granularities {#sec:order}
 
@@
 -->
-
-## Harmony and Clashes
 
@@
 
+# Data structure
+
 # Visualization {#sec:visualization}
@@
 Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
+Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
 Moreover, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
@@
 <!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
 “outliers” in letter value plots are those observations beyond the most extreme letter value.
@@
 
-Traditional ways to visualize densities include violin plots [@Hintze1998-zi, 1998].
+Traditional ways to visualize densities include violin plots [@Hintze1998-zi].
 The shape of the violin represents the density estimate of the variable.
@@
 
-Even when there are no rarely occuring events, number of observations might vary hugely across within or across each panel.
+Even when there are no rarely occuring events, number of observations might vary hugely within or across each panel.
 This might happen due to missing observations in the data or uneven locations of events in time domain.

### 2019-11-27 main.Rmd
linear and granule fixed

@@
 
-  (C1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
+  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
 than all elements of $G(j)$, and  
-  (C2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
+  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
 \end{timedom}
 
-**Dicussion:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
-The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+\begin{timedom}\label{def:granule}
+Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
+\end{granule}
+
+**Dicussion:** The first condition in \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
 \autoref{fig:linear-time} shows the implication of this condition.
+It shows the linear granularities hours, days, weeks, months and years, each of which are unidirectional in nature and arranged from past to future.
 If we consider the chronon [@aigner2011visualization] as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+Each granule of the linear granularities shown are represented through a box.
 
@@
 \end{timedom}.
-
-**Dicussion:**
 
@@
 
-**Dicussion:** According to this definition, $G \trianglelefteq H,$ if each granule H (j) is the union of some
-granules of G.
+**Dicussion:** According to this definition, $G \trianglelefteq H,$ if each granule H(j) is the union of some granules of G.
+
 For example, $Days \trianglelefteq Months$.

### 2019-11-27 main.Rmd
acyclic time granularities

@@
 \begin{timedom}\label{def:granule}
-Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
-\end{granule}
+Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
+\end{timedom}
 
@@
 ```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
-knitr::include_graphics("Figs/linear-time.png")
+knitr::include_graphics("Figs/linear-arr.png")
 ```
@@
 
-**Dicussion:** According to this definition, $G \trianglelefteq H,$ if each granule H(j) is the union of some granules of G.
-
-For example, $Days \trianglelefteq Months$.
-However, this  relationship is not fully described until we associate it with periodicity.
-For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping.
-If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
-Considering leap years and all their exceptions, the period becomes 400 years.
+**Dicussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.
+$G \preceq H$ if every granule in G is a subset of some granule in H.
+In \autoref{fig:groups-finer}, both $Days \trianglelefteq Weeks$ and $Days \preceq Weeks$ hold true.
+However, Weekends groups into Holidays but Weekends are not finer than Holidays.
+On the other hand, Weekends are finer than weeks but do not group into weeks.
+Consider another example where $Days \trianglelefteq Months$.
+This relationship is not fully described until we associate it with periodicity.
+Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
+The period becomes 400 years with the inclusion of leap years and all their exceptions, .
+
+```{r groups-finer,echo=FALSE,out.width = "100%",fig.cap="Illustrations of groups into and finer than relationships.
+Weekends groups into Holidays but Weekends are not finer than Holidays.
+On the other hand, Weekends are finer than weeks but do not group into weeks.
+Day groups into and are finer than weeks"}
+knitr::include_graphics("Figs/groupsandfiner.png")
+```
 
@@
 Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
-$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-
-**Dicussion:** This formula can be explained observing that H (j ) is either one of H (0), .
-.
-.
-, H (R - 1) or, for the periodicity of H, is one of these granules “shifted” ahead or behind on the time line of a finite number of periods.
-If $H (j ) \epsilon {H (0), .
-.
-.
-, H (R - 1)}$, then the formula ‘j mod R’ defines the index (among those in {0,...,R - 1}) of the granule that must be shifted to obtain H (j).The number of periods each granule of G composing H (j mod R) should be shifted is given by $\lfloor j/R \rfloor$.
+$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
+
+**Dicussion:**
 
@@
 # Acyclic time granularities {#sec:circular-gran-def}
+
+Re-occurances in time can also happen due to completely irregular mapping between linear granularities.
+Some public and school holidays repeat every year, but the time period on which they appear might vary from year to year.
+For example, U.S labour day is the first Monday in September and U.S.
+Memorial Day is the last day in May.
+
+A classic example can be that of Easter.
 

### 2019-11-29 main.Rmd
cyclic intro change

@@
 
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf].
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
 It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
@@
 
-**Example** :
+**Examples** :
 
@@
 
+For more variations of grouping operations and granule oriented operations, the readers are remcommended to see [@Ning2002-tf].s
+
 # Cyclic Time granularities {#sec:circular-gran-def}
+
+<!-- sub: regular and irregular mapping leads to circular and quasi-circular granularities -->
 
@@
 Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
-Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.
-
-In a cyclic organization of time, the domain is composed of a set of recurring time values.
-Hence, any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
-Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
+Mappings will be regular if, for example, they are formed by grouping operation.
+Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
+However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
+However, the formalism would differ for circular (regular mapping) and quasi-circular (irregular mapping) granularities.
+Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or public holidays.
+
+<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
 
@@
 Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
-
-The mappings of two linear granularities forming a cyclic granularity can be regular or irregular.
-It will be regular if, for example, they are formed by grouping operation.
-Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
-However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
-The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities.
-However, the formalism would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
+In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
+Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
 

### 2019-11-29 main.Rmd
changing notation of quasi-periodic

@@
 G be a linear granularity, periodic with respect to BG,     
-k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.  
+and k(BG,G) be the period length of the bottom granularity with respect to linear granularity G. Here, linear granularities BG and G has regular mapping.
 \end{timedom}
@@
 But the representation of the resultant circular granularity will be a function of the bottom granularity.
-For example, suppose the bottom granularity is minutes, and let $k_i$ is the period length of $C_i$.
+For example, suppose the bottom granularity is minutes, and let $k_i$ is the number of granules in each granule of $C_i$, also referred to as the period length.
 
@@
 
-An **Aperiodic granularity** can not be defined using modular arithmetic.
-The modulus for these type of calendar categorizations are not constant due to irregular mapping with bottom granularities.
-For example,\autoref{tab:aperiodic} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
-The coarse granularity is assumed as the period within which the finer granularity will repeat its behavior aperiodically.
+A **quasi-circular granularity** can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+For example, \autoref{tab:aperiodic} shows some example of quasi-circular granularities ($QC_i$) and plausible choices of number of granules of the finer granularity inside each granule of the coarser one($k_i$).
 
@@
 
-Day-of-Month: & $A_1(z)$ &  $k_1=31, 30, 29, 28$\\
-Hour-of-Month: & $A_2(z)$ & $k_2=24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $A_3(z)$ & $k_3=366, 365$\\
-Week-of-Month: & $A_4(z)$ & $k_4=5, 4$\\ 
+Day-of-Month: & $QC_1(z)$ &  $k_1=31, 30, 29, 28$\\
+Hour-of-Month: & $QC_2(z)$ & $k_2=24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $QC_3(z)$ & $k_3=366, 365$\\
+Week-of-Month: & $QC_4(z)$ & $k_4=5, 4$\\ 
 \bottomrule
@@
 
-**Definition: Aperiodic granularity** An aperiodic granularity A(AG, BG) that relates linear granularities AG and bottom granularity BG, with AG being periodical with respect to BG, is given by
+\begin{timedom}\label{def:circular}
+A quasi-circular granularity QC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
 
 \begin{equation}\label{eq:eq3}
-A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
+QC_{BG, G}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
 \end{equation}
@@
 z denotes the index set (index of BG),
-$R'$ denotes the number of granules of AG in each period of BG,  
-$T_w$ are the sets of indices of BG describing $AG(w)$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
-$\vert T_w \vert$ is the cardinality of set $T_w$.
-
-Since AG is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
-
-- $BG \trianglelefteq AG,$
-- $AG(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
+$R'$ denotes the number of granules of G in each period of the grouping (BG,G),   
+$T_w$ are the sets of indices of BG describing $G(w)$ such that $G(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+Here, linear granularities BG and G has irregular mapping.
+\end{timedom}
+
+Since G is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
+
+- $BG \trianglelefteq G,$
+- $G(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
 

### 2019-12-01 main.Rmd
acyclic definition

@@
 
-\begin{timedom}\label{def:circular}
+\begin{timedom}\label{def:quasi-circular}
 A quasi-circular granularity QC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
@@
 Re-occurances in time can also happen due to completely irregular mapping between linear granularities.
-Some public and school holidays repeat every year, but the time period on which they appear might vary from year to year.
+Some public and school holidays repeat every year, but there is no period within which their behavior remains constant.
 For example, U.S labour day is the first Monday in September and U.S.
 Memorial Day is the last day in May.
-
-A classic example can be that of Easter.
+A classic example can be that of Easter or street cleaning my old house occurs on the first and third Monday of the month between April and October, except for state holidays.As such all of these are still recurring events, but recurring with non-periodic patterns.
+
+\begin{timedom}\label{def:acyclic}
+An acyclic granularity AC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
+
+\begin{equation}\label{eq:eq3}
+QC_{BG, G}(z) = BG(z), z \in T_k
+\end{equation}
+where,  
+z denotes the index set (index of BG),
+$T_w$ are the sets of indices of BG describing $G(w)$ such that $G(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$. Here, linear granularities BG and G has irregular mapping.
+\end{timedom}
 
@@
 # Data structure
+
+Creating
 

### 2019-12-01 main.Rmd
data structure section

@@
 
-Creating
+<!-- How volumes of continuously collected data can be visualized to ease further decision making?  -->
+
+Creating good visualization requires good data structures.
+Commonly, simple sequences of time value pairs $<t_0, v_0> \dots, <t_n, v_n>$ are the basis of analysis and visualization.
+It is recognized that the initial approaches of just considering time as an ordinal dimension in a 2D or 3D visualisation are inadequate to capture characteristics of time-dependent information.
+For example, there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
+Moreover, one can use different calendars that define meaningful system of granularities for different application domain.
+Data can then be integrated at different levels of granularity enabling value aggregation or distribution along granularities.
+Since, we are essentially concerned with detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by anabling aggregation or distribution across cyclic time granularities.
+
+ Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
 

### 2019-12-02 main.Rmd
synergy includec in data structure and cyclic intro change

@@
 On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
-Mappings will be regular if, for example, they are formed by grouping operation.
-Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
-However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
-However, the formalism would differ for circular (regular mapping) and quasi-circular (irregular mapping) granularities.
+Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
 Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or public holidays.
+To elaborate more, mappings will be regular if, for example, linear granularities are formed by grouping operation.
+Grouping operations combine fixed granules of a linear granularity to form a new granule of the resulting linear granularity.
+However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
+Both circular and quasi-circular granularities accomodate periodicities in data.
+However, the formalisms would differ.
 
@@
 A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included}
-
-# Synergy of cyclic time granularities {#sec:synergy}
-
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-<!-- What  is interaction and some examples why it is important to study interaction -->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
-
-<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
--->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
--->
-
-<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
--->
-
-<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
--->
-
-<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
--->
-
-<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
--->
-
-<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
--->
-
-<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
--->
-
-Before exploring the behavior of a "dependent variable" across time granularities,
-it is important to know how these granularities interact with each other.
-If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
-To define a general framework, we define harmony and clashes.
-
-<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
--->
-
-<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
-It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
--->
-
-Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
-That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.Now, many situations can lead to any of these sets being empty.
-Let us discuss the following cases, where one or more of these sets can be empty.
-
-Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
-These are structurally empty sets in that it is impossible for them to have any observations.
-
-<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
-
-Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
-These are called "event-driven" empty combinations.
-Again, let us consider a specific example to illustrate this.
-Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
-Here $C_1$ can take 7 values while $C_2$ can take 2 values.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
-While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
-For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
-
-<!-- However, this is different from the first example because these are not structurally empty.
-Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
--->
-
-Thirdly, empty combinations can be a result of how granularities are constructed.
-Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
-Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
-This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
-Hence, they are referred to as "build-based" empty combinations.
-
-An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-
-Therefore, pair of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are called **harmonies**.
 
@@
 
-Creating good visualization requires good data structures.
+Effective exploration or good visualization require good data structures.
+Temporal data problems are typically grouped into two types of analysis, time series and longitudinal.
+A new tidy data structure to support exploration and modeling of temporal data is tsibble, where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
+This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.
 Commonly, simple sequences of time value pairs $<t_0, v_0> \dots, <t_n, v_n>$ are the basis of analysis and visualization.
@@
 Since, we are essentially concerned with detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by anabling aggregation or distribution across cyclic time granularities.
-
- Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
+We consider the following data structure for exploring distribution of a measured variable ((Figure \@ref(fig:data-structure)).
+These are based on conditioning the values taken on by one or more of the cyclic granularities in a data set.
+This means data subsets are formed for each level of the conditioned variable and the fixed variable.
+This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
+
+```{r data-structure, fig.cap = "The data structure for exploring periodicities along cyclic granularities.
+Two cyclic granularities serve as fixed and conditioned variable and the observational unit across combinations of these cyclic time granularities.", fig.pos="htb"}
+include_graphics("Figs/data-structure.png")
+```
+
+## Fixed and Conditioned variables {#sec:synergy}
+
+The cyclic time granularities which are essentially a function of the index of the tsibble are abstractions of time and hence are discrete variable, mapped as a factor.
+The relationships between the fixed and conditioned variable are of importance in this data structure due to the ways the cyclic granularities, which are mapped to fixed and conditional variable, relate to one another.
+To define a general framework, we define harmony and clashes.
+
+<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+These are structurally empty sets in that it is impossible for them to have any observations.
+-->
+
+<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
+-->
+
+<!-- What  is interaction and some examples why it is important to study interaction -->
+
+<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
+
+<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
+Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
+b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
+c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
+We can conclude that Month-of-Year and Day-of-Month are harmonies.
+-->
+
+<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
+The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
+All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
+Hence, Day-of-Month and Day-of-Year are clashes.
+-->
+
+<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+Violin plots are a combination of box plot and density plot.
+Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
+These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
+-->
+
+<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
+The y-axis represents week of the year and the x-axis represents electricity consumption.
+Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
+They are designed to bring out changes in distributions over time.
+The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
+So, Week-of-Month and Week-of-Year are clashes.
+-->
+
+<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
+Decile plots are useful in displaying distribution through deciles without much clutter.
+This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
+So Month-of-Year and Day-of-Year are clashes as well.
+-->
+
+<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
+Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
+Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
+-->
+
+<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
+Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
+There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
+-->
+
+<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
+Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
+-->
+
+<!-- Before exploring the behavior of a "dependent variable" across time granularities, -->
+<!-- it is important to know how these granularities interact with each other.
+If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
+-->
+
+<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
+-->
+
+<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
+It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
+-->
+
+Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
+That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
+Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
+Now, many situations can lead to any of these sets being empty.
+Let us discuss the following cases, where one or more of these sets can be empty.
+
+Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
+These are called "structurally" empty combinations.
+Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
+These are structurally empty sets in that it is impossible for them to have any observations.
+
+<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
+
+Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
+These are called "event-driven" empty combinations.
+Again, let us consider a specific example to illustrate this.
+Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
+Here $C_1$ can take 7 values while $C_2$ can take 2 values.
+So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
+While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
+For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
+
+<!-- However, this is different from the first example because these are not structurally empty.
+Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
+-->
+
+Thirdly, empty combinations can be a result of how granularities are constructed.
+Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
+Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
+This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
+Hence, they are referred to as "build-based" empty combinations.
+
+An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+
+Therefore, pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
+And the ones that do not lead to any missing combinations are called **harmonies**.
+
+## Measured variable
+
+The measured variable which is the univariate observational units can be dicrete or continuous.
+
+<!-- would invlove a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis. -->
+
+ <!-- Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space. -->
 

### 2019-12-02 main.Rmd
aperiodic granularity definition update

@@
 
-Periodicity is very common in all kinds of data.
-Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
+We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
+A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to the time index of a tsibble.
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the time index is days, then a linear granularity might be weeks, months or years.
+Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
+In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
+Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
+
+Cyclic time granularities which accomodate periodicities can be constructed by relating two linear granularities.
 The mappings between linear granularities can be regular or irregular.
@@
 On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
-Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or public holidays.
 To elaborate more, mappings will be regular if, for example, linear granularities are formed by grouping operation.
@@
 However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
-Both circular and quasi-circular granularities accomodate periodicities in data.
-However, the formalisms would differ.
+Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
+Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
+The formalisms would differ for circular and quasi-circular granularities although both accomodate periodicities in data.
 
 <!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
-
-We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
-A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble.
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
-In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
-Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
 
@@
 BG denotes the index of the tsibble (bottom granularity), 
-G be a linear granularity, periodic with respect to BG,     
-and k(BG,G) be the period length of the bottom granularity with respect to linear granularity G. Here, linear granularities BG and G has regular mapping.
+G is a linear granularity periodic with respect to BG with regular mapping,     
+and k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.
 \end{timedom}
@@
 But the representation of the resultant circular granularity will be a function of the bottom granularity.
-For example, suppose the bottom granularity is minutes, and let $k_i$ is the number of granules in each granule of $C_i$, also referred to as the period length.
+Suppose the bottom granularity is minutes, and $k_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$, also referred to as the period length.
 
@@
 
-A **quasi-circular granularity** can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
-For example, \autoref{tab:aperiodic} shows some example of quasi-circular granularities ($QC_i$) and plausible choices of number of granules of the finer granularity inside each granule of the coarser one($k_i$).
+A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+For example, \autoref{tab:quasi} shows some example of quasi-circular granularities ($QC_i$) and plausible choices of number of granules ($k_i$) of the finer granularity inside each granule of the coarser one in the cyclic granularity ($QC_i$).
 
@@
 
-Day-of-Month: & $QC_1(z)$ &  $k_1=31, 30, 29, 28$\\
-Hour-of-Month: & $QC_2(z)$ & $k_2=24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $QC_3(z)$ & $k_3=366, 365$\\
-Week-of-Month: & $QC_4(z)$ & $k_4=5, 4$\\ 
+Day-of-Month: & $QC_1(z)$ && $k_1=31, 30, 29, 28$\\
+Hour-of-Month: & $QC_2(z)$ && $k_2=24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $QC_3(z)$ && $k_3=366, 365$\\
+Week-of-Month: & $QC_4(z)$ && $k_4=5, 4$\\ 
 \bottomrule
@@
 \end{center}
-\caption{Illustrative aperiodic circular granularities with potential period lengths}
-\label{tab:aperiodic}
+\caption{Illustrative quasi-circular granularities with potential period lengths}
+\label{tab:quasi}
 \end{table}
 
-\begin{timedom}\label{def:quasi-circular}
-A quasi-circular granularity QC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
+\begin{timedom}\label{def:quasicircular}
+A quasi-circular granularity QC_(BG, G') that relates linear granularities G' and bottom granularity BG, if
 
 \begin{equation}\label{eq:eq3}
-QC_{BG, G}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
+QC_{BG, G'}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
 \end{equation}
+
 where,  
-z denotes the index set (index of BG),
-$R'$ denotes the number of granules of G in each period of the grouping (BG,G),   
-$T_w$ are the sets of indices of BG describing $G(w)$ such that $G(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-Here, linear granularities BG and G has irregular mapping.
+z denotes the index set,  
+BG denotes the index of the tsibble (bottom granularity),  
+G' is a linear granularity periodic with respect to BG with irregular mapping,   
+R' denotes the number of granules of G' in each period of the grouping (BG,G),   
+T_w are the sets of indices of BG describing G'(w) such that $G'(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{timedom}
 
-Since G is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
-
-- $BG \trianglelefteq G,$
-- $G(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
-
-**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:aperiodic-gran}.
+Since G' is periodical with respect to BG, from definition \ref{def:periodical} the following holds true:
+
+- $BG \trianglelefteq G'$
+- $G'(w) = \bigcup_{z \in T_w \mod R'}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the grouping $(BG, G')$.
+
+**Discussion**: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-gran}.
 Each granule is represented by a box.
 Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
-Two linear granularities AG and AH are considered, where $AG = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$.
-This implies that AG is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and AH is made up of shrinking every 1st granule of Group_3(BG) by 2 granules.
-Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different.
-A(BG, AG) and A(BG, AH) are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH.
-For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
-For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
+Two linear granularities AG and AH are considered, where $G' = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$.
+This implies that G' is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and H' is made up of shrinking every 1st granule of Group_3(BG) by 2 granules.
+Number of granules of G' and H' in each period of BG is 2, but the number of granules of BG in each of those granules are different.
+QC(BG, G') and QC(BG, H') are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G' or H'.
+For G', R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
+For H', R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
 
@@
 \begin{split}
-A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
+QC(BG,G')(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
   & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
@@
 \begin{split}
-A(BG,AH)(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
+QC(BG,H')(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
   & = BG(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
@@
 
-```{r aperiodic-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
+```{r quasi-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
 knitr::include_graphics("Figs/aperiodic-granv1.png")
@@
 
-# Acyclic time granularities {#sec:circular-gran-def}
-
-Re-occurances in time can also happen due to completely irregular mapping between linear granularities.
-Some public and school holidays repeat every year, but there is no period within which their behavior remains constant.
+## Aperiodic {#sec:circular-gran-def}
+
+Periodic and Quasi-periodic behavior can be defined by a repeating pattern, except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.
+Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
+Most public and school holidays repeat every year, but there is no period within which their behavior remains constant.
 For example, U.S labour day is the first Monday in September and U.S.
 Memorial Day is the last day in May.
-A classic example can be that of Easter or street cleaning my old house occurs on the first and third Monday of the month between April and October, except for state holidays.As such all of these are still recurring events, but recurring with non-periodic patterns.
+A classic example can be that of Easter.
+Examples of aperiodic granularity can also be an event that covers the first and third Monday of the month between June and October, except for state holidays.
+All of these are still recurring events, but recurring with non-periodic patterns.
+As such, plausible $k_i$ from \autoref{tab:aperiodic} can be infinite for aperiodic granularities.
 
 \begin{timedom}\label{def:acyclic}
-An acyclic granularity AC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
-
-\begin{equation}\label{eq:eq3}
-QC_{BG, G}(z) = BG(z), z \in T_k
+An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by
+
+\begin{equation}\label{eq:eq}
+\begin{split}
+A_{BG, G}(z) & = 1, z \in T_q\\
+     & = 0, z \not\in T_q\\
+\end{split}
 \end{equation}
@@
 z denotes the index set (index of BG),
-$T_w$ are the sets of indices of BG describing $G(w)$ such that $G(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$. Here, linear granularities BG and G has irregular mapping.
+$T_q$ are the sets of indices of BG describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
 \end{timedom}

### 2019-12-02 main.Rmd
acyclic time granularities def

@@
 Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
-Most public and school holidays repeat every year, but there is no period within which their behavior remains constant.
+Most public holidays repeat every year, but there is no period within which their behavior remains constant.
 For example, U.S labour day is the first Monday in September and U.S.
 Memorial Day is the last day in May.
-A classic example can be that of Easter.
-Examples of aperiodic granularity can also be an event that covers the first and third Monday of the month between June and October, except for state holidays.
+A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
+Examples of aperiodic granularity may also include school holidays or a scheduling event that covers the first and third Monday of the month between June and October, except for state holidays.
 All of these are still recurring events, but recurring with non-periodic patterns.

### 2019-12-02 gravitas/gravitas.Rmd
aperiodic completed

--- 2019-12-02 main.Rmd
@@
 This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2[@Wickham2009-pk].
-bibliography: [bibliography.bib]
-output:
-  bookdown::pdf_book:
-    #base_format: rticles::asa_article
-    fig_height: 5
-    fig_width: 8
-    fig_caption: yes
-    dev: "pdf"
+
+bibliography: bibliography.bib
+output: rticles::asa_article
 ---
@@
 
-```{r external, include = FALSE}
-# read_chunk('scripts/main.R')
-```
-
-```{r load}
-
-```
-
 # Introduction
@@
 
-# Linear Time granularities {#sec:lineartime}
+# Linear Time granularities{#sec:timegrandef}
 
 <!-- general calendar categorization -->
-<!-- Often we partition time into months, weeks or days to relate it to data.
+Often we partition time into months, weeks or days to relate it to data.
 Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
@@
 Classes of time abstractions may be based on arrangement or hierarchical order of time, discussed in further details in \autoref{sec:arrangement} and \autoref{sec:order}.
--->
-
-Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
+
+## Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement}
+
+The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.
 Time granularities are **linear** if they respect the linear progression of time.
 Examples include hours, days, weeks and months.
-
-## Definitions and Relationships
-
-<!-- Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement} -->
-
-<!-- The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.  -->
-
- There has been several attempts to provide the framework for formally characterizing time-granularities.
-One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
+However, periodicity is very common in all kinds of data.
+Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
+The mappings between linear granularities can be regular or irregular.
+For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
+On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
+Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
+Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.
 
@@
 
-<!-- In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.  -->
-
-<!-- ### Linear {#sec:linear-gran-def} -->
+In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.
+
+### Linear {#sec:linear-gran-def}
+
+There has been several attempts to provide the framework for formally characterizing time-granularities.
+One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
 
@@
 
-  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
+  (C1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
 than all elements of $G(j)$, and  
-  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
+  (C2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
 \end{timedom}
 
-\begin{timedom}\label{def:granule}
-Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
-\end{timedom}
-
-**Dicussion:** The first condition in \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+**Dicussion:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
+The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
 \autoref{fig:linear-time} shows the implication of this condition.
-It shows the linear granularities hours, days, weeks, months and years, each of which are unidirectional in nature and arranged from past to future.
 If we consider the chronon [@aigner2011visualization] as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
-Each granule of the linear granularities shown are represented through a box.
 
 ```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
-knitr::include_graphics("Figs/linear-arr.png")
+knitr::include_graphics("Figs/linear-time.png")
 ```
+
+#### Relationships between two linear granularities and association to Periodicity
 
@@
 for each index i, there exists an index j such that $G(i) \subset H(j)$.
-\end{timedom}.
+\end{timedom}
 
@@
 A linear granularity G groups into a linear granularity H, denoted
-$G \trianglelefteq H,$
+G \trianglelefteq H 
 if for each index j there exists a (possibly infinite) subset S of the integers such
@@
 
-**Dicussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.
-$G \preceq H$ if every granule in G is a subset of some granule in H.
-In \autoref{fig:groups-finer}, both $Days \trianglelefteq Weeks$ and $Days \preceq Weeks$ hold true.
-However, Weekends groups into Holidays but Weekends are not finer than Holidays.
-On the other hand, Weekends are finer than weeks but do not group into weeks.
-Consider another example where $Days \trianglelefteq Months$.
-This relationship is not fully described until we associate it with periodicity.
-Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions, .
-
-```{r groups-finer,echo=FALSE,out.width = "100%",fig.cap="Illustrations of groups into and finer than relationships.
-Weekends groups into Holidays but Weekends are not finer than Holidays.
-On the other hand, Weekends are finer than weeks but do not group into weeks.
-Day groups into and are finer than weeks"}
-knitr::include_graphics("Figs/groupsandfiner.png")
-```
-
+According to this definition, $G \trianglelefteq H$ if each granule H (j) is the union of some
+granules of G.
+For example, $Days \trianglelefteq Months$.
+However, this  relationship is not fully described until we associate it with periodicity.
+For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping.
+If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
+Considering leap years and all their exceptions, the period becomes 400 years.
 \begin{timedom}\label{def:periodical}
 A granularity H is periodical with respect to a granularity G if
-(1) $G \trianglelefteq H,$, and
-(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
-that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
-$H (i + R) = \bigcup_{j \in S} G(j + P)$.
+(1) G \trianglelefteq H, and
+(2) there exist R,P \epsilon Z+, where R is less than the number of granules of H, such
+that for all i \epsilon Z, if H(i) = \bigcup_{j \in S}G(j) and H (i + R) \neq \phi then
+H (i + R) = \bigcup_{j \in S} G(j + P).
 \end{timedom}
@@
 Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
-$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-
-**Dicussion:**
+$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
+
+This formula can be explained observing that H (j ) is either one of H (0), .
+.
+.
+, H (R - 1) or, for the periodicity of H, is one of these granules “shifted” ahead or behind on the time line of a finite number of periods.
+If $H (j ) \epsilon {H (0), .
+.
+.
+, H (R - 1)}$, then the formula ‘j mod R’ defines the index (among those in {0,...,R - 1}) of the granule that must be shifted to obtain H (j).The number of periods each granule of G composing H (j mod R) should be shifted is given by $\lfloor j/R \rfloor$.
 
@@
 A granularity H is quasi-periodical with respect to a granularity G if
-(1) $$G \trianglelefteq H,$$, and
+(1) G \trianglelefteq H, and
 (2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
 integers R,P, where R is less than the minimum of the number of granules of H
-between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
+between any 2 exceptions, such that for all i \epsilon Z, if H(i) = \bigcup_{k \in [0,k]} G(j_r)  and H (i + R) \neq \phi and i + R < min(E), where E is the closest existing exception after H (i)^2, then H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P ).
 \end{timedom}
 
-**Dicussion:**
-
 Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by R and P.
@@
 \begin{timedom}\label{def:BG}
-Bottom granularity - Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
+**Definition: Bottom granularity** Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
 \end{timedom}
 
-**Dicussion:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+**Example:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
 
@@
 
-## Computation through Calendar Algebra
-
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
+#### Computation of linear time granularities
+
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf].
 It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
@@
 More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
-each integer $i$, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
-
-**Examples** :
+each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
+
+**Example** :
 
@@
 
-- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and $l$, $k$, $m$ integers, where G2 partitions G1, and $1 \leq l \leq m$.
+- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
 The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-The altering-tick operation modifies the granules of G1 so that the $l$th granule of each group has $|k|$ additional (or fewer when k < 0) granules of G2.
+The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
 
@@
 
-For more variations of grouping operations and granule oriented operations, the readers are remcommended to see [@Ning2002-tf].s
-
-# Cyclic Time granularities {#sec:circular-gran-def}
-
-<!-- sub: regular and irregular mapping leads to circular and quasi-circular granularities -->
-
-Periodicity is very common in all kinds of data.
-Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
-The mappings between linear granularities can be regular or irregular.
-For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
-On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
-Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or public holidays.
-To elaborate more, mappings will be regular if, for example, linear granularities are formed by grouping operation.
-Grouping operations combine fixed granules of a linear granularity to form a new granule of the resulting linear granularity.
-However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
-Both circular and quasi-circular granularities accomodate periodicities in data.
-However, the formalisms would differ.
-
-<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
+### Cyclic {#sec:cyclic-gran-def}
+
+In a cyclic organization of time, the domain is composed of a set of recurring time values.
+Hence, any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
+Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
 
@@
 Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
-In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
-Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
-
-## Circular {#sec:circular-gran-def}
+
+The mappings of two linear granularities forming a cyclic granularity can be regular or irregular.
+It will be regular if, for example, they are formed by grouping operation.
+Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
+However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
+The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities.
+However, the formalism would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
+
+#### Circular {#sec:circular-gran-def}
 
@@
 C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
-BG(z) \neq \phi,
+BG(z) \neq \phi
 \end{equation}
 where,  
-z denotes the index set, 
-BG denotes the index of the tsibble (bottom granularity), 
-G be a linear granularity, periodic with respect to BG,     
-and k(BG,G) be the period length of the bottom granularity with respect to linear granularity G. Here, linear granularities BG and G has regular mapping.
+z denotes the index set.  
+BG denotes the index of the tsibble (bottom granularity).  
+G be a linear granularity, periodic with respect to BG.    
+k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.  
 \end{timedom}
@@
 But the representation of the resultant circular granularity will be a function of the bottom granularity.
-For example, suppose the bottom granularity is minutes, and let $k_i$ is the number of granules in each granule of $C_i$, also referred to as the period length.
+For example, suppose the bottom granularity is minutes, and let $k_i$ is the period length of $C_i$.
 
@@
 
-## Quasi-circular {#sec:circular-gran-def}
-
-A **quasi-circular granularity** can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
-For example, \autoref{tab:aperiodic} shows some example of quasi-circular granularities ($QC_i$) and plausible choices of number of granules of the finer granularity inside each granule of the coarser one($k_i$).
+### Aperiodic {#sec:aperiodic-gran-def}
+
+An **Aperiodic granularity** can not be defined using modular arithmetic.
+The modulus for these type of calendar categorizations are not constant due to irregular mapping with bottom granularities.
+For example,\autoref{tab:aperiodic} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
+The coarse granularity is assumed as the period within which the finer granularity will repeat its behavior aperiodically.
 
@@
 
-Day-of-Month: & $QC_1(z)$ &  $k_1=31, 30, 29, 28$\\
-Hour-of-Month: & $QC_2(z)$ & $k_2=24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $QC_3(z)$ & $k_3=366, 365$\\
-Week-of-Month: & $QC_4(z)$ & $k_4=5, 4$\\ 
+Day-of-Month: & $A_1(z)$ &  $k_1=31, 30, 29, 28$\\
+Hour-of-Month: & $A_2(z)$ & $k_2=24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $A_3(z)$ & $k_3=366, 365$\\
+Week-of-Month: & $A_4(z)$ & $k_4=5, 4$\\ 
 \bottomrule
@@
 
-\begin{timedom}\label{def:quasi-circular}
-A quasi-circular granularity QC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
+**Definition: Aperiodic granularity** An aperiodic granularity A(AG, BG) that relates linear granularities AG and bottom granularity BG, with AG being periodical with respect to BG, is given by
 
 \begin{equation}\label{eq:eq3}
-QC_{BG, G}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
+A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
 \end{equation}
@@
 z denotes the index set (index of BG),
-$R'$ denotes the number of granules of G in each period of the grouping (BG,G),   
-$T_w$ are the sets of indices of BG describing $G(w)$ such that $G(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-Here, linear granularities BG and G has irregular mapping.
-\end{timedom}
-
-Since G is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
-
-- $BG \trianglelefteq G,$
-- $G(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
+R' denotes the number of granules of AG in each period of BG,  
+$T_w$ are the sets of indices of BG describing $AG(w)$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
+$\vert T_w \vert$ is the cardinality of set $T_w$.
+
+Since AG is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
+
+- $BG \trianglelefteq AG$
+- $AG(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
 
@@
 
-# Acyclic time granularities {#sec:circular-gran-def}
-
-Re-occurances in time can also happen due to completely irregular mapping between linear granularities.
-Some public and school holidays repeat every year, but there is no period within which their behavior remains constant.
-For example, U.S labour day is the first Monday in September and U.S.
-Memorial Day is the last day in May.
-A classic example can be that of Easter or street cleaning my old house occurs on the first and third Monday of the month between April and October, except for state holidays.As such all of these are still recurring events, but recurring with non-periodic patterns.
-
-\begin{timedom}\label{def:acyclic}
-An acyclic granularity AC_(BG, G) that relates linear granularities G and bottom granularity BG, with AG being periodical with respect to BG, is given by
-
-\begin{equation}\label{eq:eq3}
-QC_{BG, G}(z) = BG(z), z \in T_k
-\end{equation}
-where,  
-z denotes the index set (index of BG),
-$T_w$ are the sets of indices of BG describing $G(w)$ such that $G(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$. Here, linear granularities BG and G has irregular mapping.
-\end{timedom}
-
 <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
 
-# Computing cyclic time granularities {#sec:order}
+## Order: Single vs. Multiple {#sec:order}
 
@@
 A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included}
+
+# Synergy of cyclic time granularities {#sec:synergy}
+
+<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+These are structurally empty sets in that it is impossible for them to have any observations.
+-->
+
+<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+So graphics involving $C_1$ and $C_2$ are potentially useful.
+-->
+
+<!-- What  is interaction and some examples why it is important to study interaction -->
+
+<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
+
+<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
+M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
+Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
+b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
+c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
+We can conclude that Month-of-Year and Day-of-Month are harmonies.
+-->
+
+<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
+The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
+All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
+Hence, Day-of-Month and Day-of-Year are clashes.
+-->
+
+<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
+Violin plots are a combination of box plot and density plot.
+Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
+These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
+-->
+
+<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
+The y-axis represents week of the year and the x-axis represents electricity consumption.
+Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
+They are designed to bring out changes in distributions over time.
+The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
+So, Week-of-Month and Week-of-Year are clashes.
+-->
+
+<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
+Decile plots are useful in displaying distribution through deciles without much clutter.
+This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
+So Month-of-Year and Day-of-Year are clashes as well.
+-->
+
+<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
+Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
+Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
+-->
+
+<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
+Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
+There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
+-->
+
+<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
+Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
+-->
+
+Before exploring the behavior of a "dependent variable" across time granularities,
+it is important to know how these granularities interact with each other.
+If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
+To define a general framework, we define harmony and clashes.
+
+<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
+-->
+
+## Harmony and Clashes
+
+<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
+It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
+-->
+
+Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
+That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
+Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.Now, many situations can lead to any of these sets being empty.
+Let us discuss the following cases, where one or more of these sets can be empty.
+
+Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
+These are called "structurally" empty combinations.
+Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
+Here $C_1$ can take 31 values while $C_2$ can take 5 values.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Many of these are empty.
+For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
+These are structurally empty sets in that it is impossible for them to have any observations.
+
+<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
+
+Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
+These are called "event-driven" empty combinations.
+Again, let us consider a specific example to illustrate this.
+Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
+Here $C_1$ can take 7 values while $C_2$ can take 2 values.
+So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
+While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
+For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
+
+<!-- However, this is different from the first example because these are not structurally empty.
+Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
+-->
+
+Thirdly, empty combinations can be a result of how granularities are constructed.
+Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
+Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
+This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
+Hence, they are referred to as "build-based" empty combinations.
+
+An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
+Here $C_1$ can take 7 values while $C_2$ can take 12 values.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
+All of these are non-empty because every DOW can occur in every month.
+
+Therefore, pair of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
+And the ones that do not lead to any missing combinations are called **harmonies**.
 
@@
 
-# Data structure
-
-<!-- How volumes of continuously collected data can be visualized to ease further decision making?  -->
-
-Effective exploration or good visualization require good data structures.
-Temporal data problems are typically grouped into two types of analysis, time series and longitudinal.
-A new tidy data structure to support exploration and modeling of temporal data is tsibble, where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
-This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.
-Commonly, simple sequences of time value pairs $<t_0, v_0> \dots, <t_n, v_n>$ are the basis of analysis and visualization.
-It is recognized that the initial approaches of just considering time as an ordinal dimension in a 2D or 3D visualisation are inadequate to capture characteristics of time-dependent information.
-For example, there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use different calendars that define meaningful system of granularities for different application domain.
-Data can then be integrated at different levels of granularity enabling value aggregation or distribution along granularities.
-Since, we are essentially concerned with detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by anabling aggregation or distribution across cyclic time granularities.
-We consider the following data structure for exploring distribution of a measured variable ((Figure \@ref(fig:data-structure)).
-These are based on conditioning the values taken on by one or more of the cyclic granularities in a data set.
-This means data subsets are formed for each level of the conditioned variable and the fixed variable.
-This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
-
-```{r data-structure, fig.cap = "The data structure for exploring periodicities along cyclic granularities.
-Two cyclic granularities serve as fixed and conditioned variable and the observational unit across combinations of these cyclic time granularities.", fig.pos="htb"}
-include_graphics("Figs/data-structure.png")
-```
-
-## Fixed and Conditioned variables {#sec:synergy}
-
-The cyclic time granularities which are essentially a function of the index of the tsibble are abstractions of time and hence are discrete variable, mapped as a factor.
-The relationships between the fixed and conditioned variable are of importance in this data structure due to the ways the cyclic granularities, which are mapped to fixed and conditional variable, relate to one another.
-To define a general framework, we define harmony and clashes.
-
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-<!-- What  is interaction and some examples why it is important to study interaction -->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
-
-<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
--->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
--->
-
-<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
--->
-
-<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
--->
-
-<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
--->
-
-<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
--->
-
-<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
--->
-
-<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
--->
-
-<!-- Before exploring the behavior of a "dependent variable" across time granularities, -->
-<!-- it is important to know how these granularities interact with each other.
-If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
--->
-
-<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
--->
-
-<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
-It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
--->
-
-Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
-That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
-Now, many situations can lead to any of these sets being empty.
-Let us discuss the following cases, where one or more of these sets can be empty.
-
-Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
-These are structurally empty sets in that it is impossible for them to have any observations.
-
-<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
-
-Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
-These are called "event-driven" empty combinations.
-Again, let us consider a specific example to illustrate this.
-Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
-Here $C_1$ can take 7 values while $C_2$ can take 2 values.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
-While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
-For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
-
-<!-- However, this is different from the first example because these are not structurally empty.
-Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
--->
-
-Thirdly, empty combinations can be a result of how granularities are constructed.
-Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
-Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
-This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
-Hence, they are referred to as "build-based" empty combinations.
-
-An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-
-Therefore, pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are called **harmonies**.
-
-## Measured variable
-
-The measured variable which is the univariate observational units can be dicrete or continuous.
-
-<!-- would invlove a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis. -->
-
- <!-- Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space. -->
-
 # Visualization {#sec:visualization}
@@
 Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
+Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
 Moreover, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
@@
 <!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
 “outliers” in letter value plots are those observations beyond the most extreme letter value.
@@
 
-Traditional ways to visualize densities include violin plots [@Hintze1998-zi].
+Traditional ways to visualize densities include violin plots [@Hintze1998-zi, 1998].
 The shape of the violin represents the density estimate of the variable.

### 2019-12-02 gravitas/gravitas.tex
aperiodic completed

--- 2019-12-02 gravitas/gravitas.Rmd
@@
----
-title: Exploring probability distributions for bivariate temporal granularities
-authors:
-- name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-  email: Sayani.Gupta@monash.edu
-
-- name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-keywords:
-- data visualization
-- statistical distributions
-- time granularities
-- periodicties
-- grammar of graphics
-- R
-
-abstract: |
- Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
-As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases.
-We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week.
-Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.
-
-  The hierarchical structure of linear granularities creates a natural nested ordering resulting in single-order-up and multiple-order-up granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
-
-   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2[@Wickham2009-pk].
-
-bibliography: bibliography.bib
-output: rticles::asa_article
----
-
-```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
-options("knitr.graphics.auto_pdf" = TRUE)
-library(knitr)
-library(tidyverse)
-library(lubridate)
-library(lvplot)
-library(ggridges)
-library(tsibble)
-library(gravitas)
-library(tsibble)
-opts_chunk$set(
-  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
-  fig.path = "figure/", fig.align = "center", fig.show = "hold",
-  cache = FALSE, cache.path = "cache/",
-  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
-)
-knitr::opts_knit$set(root.dir = here::here())
-```
-
-# Introduction
-
-<!--temporal granularities and why should we care -->
-
-<!-- problem of analyzing data with such finer scales and classical ways to deal with it -->
-
-Temporal data are available at various resolutions depending on the context.
-Social and economic data like GDP is often collected and reported at coarse temporal scales like monthly, quarterly or annually.
-With recent advancement in technology, more and more data are recorded at much finer temporal scales.
-Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be recorded every second.
-As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases.
-For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
-This approach requires deconstructing time in various possible ways called time granularities[@aigner2011visualization].
-It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
-This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
-Analysts are expected to iteratively explore all possible choices of time granularities for comprehending possible periodicities in the data.
-But too many choices and a lack of a systematic approach to do so might become overwhelming.
-
-<!-- # existing ways and challenges -->
-
-Calendar-based graphics[@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
-Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time.
-But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
-There can be several classes of time deconstructions, viz.
-based on the arrangement (linear vs.
-cyclic) or hierarchical order of the calendar.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-One of the first attempts of characterizing these granularities occur in [@Bettini1998-ed].
-The definitions and rules defined are inadequate to reflect periodicities in time.
-Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior.
-Cyclic time granularities can be circular or aperiodic depending on if mappings between granularities are regular or irregular.
-Examples of circular can be hour of the day, day of the week and that of aperiodic granularities can be day of the month or public holidays.
-Time deconstructions can also be based on the hierarchical structure of time.
-For example, hours are nested within days, days within weeks, weeks within months, and so on.
-Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
-lubridate [@G_Grolemund2011-vm] creates easy access and manipulation of common date-time objects.
-But most accessor functions are limited to single-order-up granularities.
-
-<!--motivation -->
-
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
-Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
-Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates [@2012-la],  of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
-However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analyzed for different periodicities.
-
-<!-- at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA).
-EDA calls for utilizing visualization and transformation to explore data systematically.
-It is a process of generating hypothesis, testing them and consequently refining them through investigations.
--->
-
-<!--existing closely related work and how ours relate --> 
- <!-- This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones. -->
-
-<!--actions on multiple order up - usage -->
-
-<!-- Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-However, this approach can lead to a myriad of choices all of which are not useful.
-Analysts are expected to iteratively visualize these choices for exploring possible patterns in the data.
-But too many choices might leave him bewildered.
--->
-
-<!-- challenges and how you are dealing wih it --> 
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow through the following:
-
-<!-- Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
--->
-
-  * Formal characterization of cyclic granularities
-
-  * Facilitate manipulation of single and multiple order-up time granularities
-
-  * Checking feasibility of creating plots or drawing inferences from any two cyclic granularities
-
-  * Recommend prospective probability distributions for exploring distributions of univariate dependent variable across pair of granularities
-
-  <!-- Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
--->
-
-The remainder of the paper is organized as follows.
-Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterization of circular and aperiodic time granularities.
-Section \ref{sec:synergy} discusses the effect in data structure due to how pairs of cyclic time granularities are related.
-Section \ref{sec:visualization} discusses the role of plot choices, synergy of time granularities and number of observations to construct an informative and trustworthy visualization.
-Section \ref{sec:application} discusses how systematic exploration can be carried out for a temporal and non-temporal case.
-Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.
-
-# Linear Time granularities{#sec:timegrandef}
-
-<!-- general calendar categorization -->
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+%% load any required packages here
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Exploring probability distributions for bivariate temporal granularities}
+
+  \author{
+        Sayani Gupta \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Rob J Hyndman \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Dianne Cook \\
+    Department of Econometrics and Business Statistics, Monash University\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Exploring probability distributions for bivariate temporal granularities}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+Recent advances in technology greatly facilitates recording and storing
+data at much finer temporal scales than was previously possible. As the
+frequency of time-oriented data increases, the number of questions about
+the observed variable that need to be addressed by visual representation
+also increases. We propose some new tools to explore this type of data,
+which deconstruct time in many different ways. There are several classes
+of time deconstructions including linear time granularities, circular
+time granularities and aperiodic calendar categorizations. Linear time
+granularities respect the linear progression of time such as hours,
+days, weeks and months. Circular time granularities accommodate
+periodicities in time such as hour of the day, and day of the week.
+Aperiodic calendar categorizations are neither linear nor circular, such
+as day of the month or public holidays.
+
+The hierarchical structure of linear granularities creates a natural
+nested ordering resulting in single-order-up and multiple-order-up
+granularities. For example, hour of the week and second of the hour are
+both multiple-order-up, while hour of the day and second of the minute
+are single-order-up.
+
+Visualizing data across granularities which are either single-order-up
+or multiple-order-up or periodic/aperiodic helps us to understand
+periodicities, pattern and anomalies in the data. Because of the large
+volume of data available, using displays of probability distributions
+conditional on one or more granularities is a potentially useful
+approach. This work provides tools for creating granularities and
+exploring the associated time series within the tidy workflow, so that
+probability distributions can be examined using the range of graphics
+available in ggplot2\citep{Wickham2009-pk}.
+\end{abstract}
+
+\noindent%
+{\it Keywords:} data visualization, statistical distributions, time granularities, periodicties, grammar of graphics, R
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Temporal data are available at various resolutions depending on the
+context. Social and economic data like GDP is often collected and
+reported at coarse temporal scales like monthly, quarterly or annually.
+With recent advancement in technology, more and more data are recorded
+at much finer temporal scales. Energy consumption is collected every
+half an hour, while energy supply is collected every minute and web
+search data might be recorded every second. As the frequency of data
+increases, the number of questions about the periodicity of the observed
+variable also increases. For example, data collected at an hourly scale
+can be analyzed using coarser temporal scales like days, months or
+quarters. This approach requires deconstructing time in various possible
+ways called time granularities\citep{aigner2011visualization}. It is
+important to be able to navigate through all of these temporal
+granularities to have multiple perspectives on the periodicity of the
+observed data. This idea aligns with the notion of EDA
+\citep{Tukey1977-jx} which emphasizes the use of multiple perspectives
+on data to help formulate hypotheses before proceeding to hypothesis
+testing. Visualizing probability distributions conditional on one or
+more granularities is a potentially useful approach for exploration.
+Analysts are expected to iteratively explore all possible choices of
+time granularities for comprehending possible periodicities in the data.
+But too many choices and a lack of a systematic approach to do so might
+become overwhelming.
+
+Calendar-based graphics\citep{wang2018calendar} are useful in
+visualizing patterns in the weekly and monthly structure well and are
+capable of checking the weekends or special days. Any sub-daily
+resolution temporal data can also be displayed using this type of
+faceting \citep{Wickham2009-pk} with days of the week, month of the year
+and another sub-daily deconstruction of time. But calendar effects are
+not restricted to conventional day-of-week, month-of-year ways of
+deconstructing time. There can be several classes of time
+deconstructions, viz. based on the arrangement (linear vs.~cyclic) or
+hierarchical order of the calendar. Linear time granularities respect
+the linear progression of time such as hours, days, weeks and months.
+One of the first attempts of characterizing these granularities occur in
+\citep{Bettini1998-ed}. The definitions and rules defined are inadequate
+to reflect periodicities in time. Hence, there is a need to define
+cyclic time granularities in a different approach, which can be useful
+in visualizing periodic behavior. Cyclic time granularities can be
+circular or aperiodic depending on if mappings between granularities are
+regular or irregular. Examples of circular can be hour of the day, day
+of the week and that of aperiodic granularities can be day of the month
+or public holidays. Time deconstructions can also be based on the
+hierarchical structure of time. For example, hours are nested within
+days, days within weeks, weeks within months, and so on. Hence, it is
+possible to construct single-order-up granularities like second of the
+minute or multiple-order-up granularities like second of the hour.
+lubridate \citep{G_Grolemund2011-vm} creates easy access and
+manipulation of common date-time objects. But most accessor functions
+are limited to single-order-up granularities.
+
+The motivation for this work comes from the desire to provide methods to
+better understand large quantities of measurements on energy usage
+reported by smart meters in households across Australia, and indeed many
+parts of the world. Smart meters currently provide half-hourly use in
+kWh for each household, from the time that they were installed, some as
+early as 2012. Households are distributed geographically and have
+different demographic properties such as the existence of solar panels,
+central heating or air conditioning. The behavioral patterns in
+households vary substantially, for example, some families use a dryer
+for their clothes while others hang them on a line, and some households
+might consist of night owls, while others are morning larks. It is
+common to see aggregates \citep{2012-la}, of usage across households,
+total kWh used each half hour by state, for example, because energy
+companies need to understand maximum loads that they will have to plan
+ahead to accommodate. But studying overall energy use hides the
+distributions of usage at finer scales, and making it more difficult to
+find solutions to improve energy efficiency. We propose that the
+analysis of smart meter data can be benefited from systematically
+exploring energy consumption by visualizing the probability
+distributions across different deconstructions of time to find regular
+patterns/anomalies. However, the motivation came through the smart meter
+example, this is a problem which relates to any data that needs to be
+analyzed for different periodicities.
+
+This work provides tools for systematically exploring bivariate
+granularities within the tidy workflow through the following:
+
+\begin{itemize}
+\item
+  Formal characterization of cyclic granularities
+\item
+  Facilitate manipulation of single and multiple order-up time
+  granularities
+\item
+  Checking feasibility of creating plots or drawing inferences from any
+  two cyclic granularities
+\item
+  Recommend prospective probability distributions for exploring
+  distributions of univariate dependent variable across pair of
+  granularities
+\end{itemize}
+
+The remainder of the paper is organized as follows. Section
+\ref{sec:timegrandef} details the background of linear granularities in
+depth and introduces the characterization of circular and aperiodic time
+granularities. Section \ref{sec:synergy} discusses the effect in data
+structure due to how pairs of cyclic time granularities are related.
+Section \ref{sec:visualization} discusses the role of plot choices,
+synergy of time granularities and number of observations to construct an
+informative and trustworthy visualization. Section \ref{sec:application}
+discusses how systematic exploration can be carried out for a temporal
+and non-temporal case. Section \ref{sec:discussion} summarizes this
+paper and discusses possible future direction.
+
+\hypertarget{sec:timegrandef}{%
+\section{Linear Time granularities}\label{sec:timegrandef}}
+
 Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
-Classes of time abstractions may be based on arrangement or hierarchical order of time, discussed in further details in \autoref{sec:arrangement} and \autoref{sec:order}.
-
-## Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement}
-
-The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.
-Time granularities are **linear** if they respect the linear progression of time.
-Examples include hours, days, weeks and months.
-However, periodicity is very common in all kinds of data.
-Time granularities which can accomodate periodicities can be constructed by relating two linear granularities.
-The mappings between linear granularities can be regular or irregular.
-For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
-On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
-Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular.
-Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.
-
-<!-- <!-- Representation with tsibble -->
-<!-- Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
--->
-
-<!-- <!-- What is calendar categorization --> 
-<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc.
-Many of these are circular, such as day-of-week, time-of-day.
-Some are nearly circular such as day-of-month.
-Some are irregular such as working day/non-working day.
-Let's call all of these "calendar categorizations".
-Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
--->
-
-<!-- What is circular granularity and some examples -->
-<!-- We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
-The number of categories is essentially the periodicity of a circular time granularity.
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
-Then the following categorizations can be computed.
--->
-
-In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.
-
-### Linear {#sec:linear-gran-def}
-
-There has been several attempts to provide the framework for formally characterizing time-granularities.
-One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
+Such discrete abstractions of time can be thought of as time
+granularities \citep{aigner2011visualization}. Examples of time
+abstractions may also include day-of-week, time-of-day, week-of-year,
+day-of-month, month-of-year, working day/non-working day, etc which are
+useful to represent different periodicities in the data. Classes of time
+abstractions may be based on arrangement or hierarchical order of time,
+discussed in further details in \autoref{sec:arrangement} and
+\autoref{sec:order}.
+
+\hypertarget{sec:arrangement}{%
+\subsection{Arrangement: Linear vs.~Circular
+vs.~aperiodic}\label{sec:arrangement}}
+
+The arrangement of the time domain can result in deconstructing time in
+linear, circular or aperiodic (nearly circular) ways. Time granularities
+are \textbf{linear} if they respect the linear progression of time.
+Examples include hours, days, weeks and months. However, periodicity is
+very common in all kinds of data. Time granularities which can
+accomodate periodicities can be constructed by relating two linear
+granularities. The mappings between linear granularities can be regular
+or irregular. For example, a regular mapping exists between minutes and
+hours, where 60 minutes always add up to 1 hour. On contrary, an
+irregular mapping exists between days and months, since one might can
+have days ranging from 28 to 31. Hence, time granularities can also be
+\textbf{circular} or \textbf{aperiodic} depending on if the mapping of
+the linear granularities is regular or irregular. Examples of circular
+time granularities include hour of the day, and day of the week, whereas
+examples for aperiodic time granularities can be day of the month or
+public holidays.
+
+In the following sections, we provide a formalism to all of these
+abstractions and identify their relationships and symbolic
+representations.
+
+\hypertarget{sec:linear-gran-def}{%
+\subsubsection{Linear}\label{sec:linear-gran-def}}
+
+There has been several attempts to provide the framework for formally
+characterizing time-granularities. One of the first attempts occur in
+\citep{Bettini1998-ed} with the help of the following definitions:
 
@@
 
-A time domain can be **discrete** (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be **dense** (if it is an infinite set).
-A time domain is assumed to be discrete for the purpose of our discussion.
+A time domain can be \textbf{discrete} (if there is unique predecessor
+and successor for every element except for the first and last one in the
+time domain), or it can be \textbf{dense} (if it is an infinite set). A
+time domain is assumed to be discrete for the purpose of our discussion.
 
@@
 
-**Dicussion:** Each non-empty subset $G(i)$ is called a **granule**, where $i$ is one of the indexes and $G$ is a linear granularity.
-The first condition implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
-\autoref{fig:linear-time} shows the implication of this condition.
-If we consider the chronon [@aigner2011visualization] as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
-
-```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
-knitr::include_graphics("Figs/linear-time.png")
-```
-
-#### Relationships between two linear granularities and association to Periodicity
-
-[@Bettini1998-ed] talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
+\textbf{Dicussion:} Each non-empty subset \(G(i)\) is called a
+\textbf{granule}, where \(i\) is one of the indexes and \(G\) is a
+linear granularity. The first condition implies that the granules in a
+linear granularity are non-overlapping and their index order is same as
+time order. \autoref{fig:linear-time} shows the implication of this
+condition. If we consider the chronon \citep{aigner2011visualization} as
+hourly, the time domain with T hours will have \(\lfloor T/24\rfloor\)
+days, \(\lfloor T/(24*7)\rfloor\) weeks and so on.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/linear-time}
+
+}
+
+\caption{The time domain distributed as linear granularities}\label{fig:linear-time}
+\end{figure}
+
+\hypertarget{relationships-between-two-linear-granularities-and-association-to-periodicity}{%
+\paragraph{Relationships between two linear granularities and
+association to
+Periodicity}\label{relationships-between-two-linear-granularities-and-association-to-periodicity}}
+
+\citep{Bettini1998-ed} talks about the relationships of linear time
+granularities and structure of a calendar and also relates them to the
+notion of periodicity in time.
 
@@
 
-According to this definition, $G \trianglelefteq H$ if each granule H (j) is the union of some
-granules of G.
-For example, $Days \trianglelefteq Months$.
-However, this  relationship is not fully described until we associate it with periodicity.
-For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping.
-If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
-Considering leap years and all their exceptions, the period becomes 400 years.
-\begin{timedom}\label{def:periodical}
-A granularity H is periodical with respect to a granularity G if
-(1) G \trianglelefteq H, and
-(2) there exist R,P \epsilon Z+, where R is less than the number of granules of H, such
-that for all i \epsilon Z, if H(i) = \bigcup_{j \in S}G(j) and H (i + R) \neq \phi then
-H (i + R) = \bigcup_{j \in S} G(j + P).
-\end{timedom}
+According to this definition, \(G \trianglelefteq H\) if each granule H
+(j) is the union of some granules of G. For example,
+\(Days \trianglelefteq Months\). However, this relationship is not fully
+described until we associate it with periodicity. For example,
+\(Days \trianglelefteq Months,\), since each month is a grouping of a
+number of days, however it is also a periodical grouping. If leap year
+is ignored, the periodicity of the grouping is 1 year, since each month
+would be defined as the grouping of the same number of days (31, 28, or
+30, depending on the month) every year. Considering leap years and all
+their exceptions, the period becomes 400 years.
+\textbackslash{}begin\{timedom\}\label{def:periodical} A granularity H
+is periodical with respect to a granularity G if (1) G
+\trianglelefteq H, and (2) there exist R,P \epsilon Z+, where R is less
+than the number of granules of H, such that for all i \epsilon Z, if
+H(i) = \bigcup\emph{\{j \in S\}G(j) and H (i + R) \neq \phi then H (i +
+R) = \bigcup}\{j \in S\} G(j + P). \textbackslash{}end\{timedom\}
 
 A granularity H which is periodical with respect to G is specified by:
-(i) the R sets of indexes of G, ${S_0,...,S_(R-1)}$ describing the granules of H within one period;
-(ii) the value of P; 
-(iii) the indexes of first and last granules in H, if their value is not infinite. 
-Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
-$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H (j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-
-This formula can be explained observing that H (j ) is either one of H (0), .
-.
-.
-, H (R - 1) or, for the periodicity of H, is one of these granules “shifted” ahead or behind on the time line of a finite number of periods.
-If $H (j ) \epsilon {H (0), .
-.
-.
-, H (R - 1)}$, then the formula ‘j mod R’ defines the index (among those in {0,...,R - 1}) of the granule that must be shifted to obtain H (j).The number of periods each granule of G composing H (j mod R) should be shifted is given by $\lfloor j/R \rfloor$.
-
-Granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
-
-\begin{timedom}\label{def:quasi}
-A granularity H is quasi-periodical with respect to a granularity G if
-(1) G \trianglelefteq H, and
-(2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
-integers R,P, where R is less than the minimum of the number of granules of H
-between any 2 exceptions, such that for all i \epsilon Z, if H(i) = \bigcup_{k \in [0,k]} G(j_r)  and H (i + R) \neq \phi and i + R < min(E), where E is the closest existing exception after H (i)^2, then H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P ).
-\end{timedom}
-
-Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by R and P.
+(i) the R sets of indexes of G, \({S_0,...,S_(R-1)}\) describing the
+granules of H within one period; (ii) the value of P; (iii) the indexes
+of first and last granules in H, if their value is not infinite. Then,
+if \({S_0,...,S_{R-1}}\) are the sets of indexes of G describing
+\(H (0), . . . , H (R - 1)\), respectively, then the description of an
+arbitrary granule H (j) is given by:
+\(\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)\).
+
+This formula can be explained observing that H (j ) is either one of H
+(0), . . . , H (R - 1) or, for the periodicity of H, is one of these
+granules ``shifted'' ahead or behind on the time line of a finite number
+of periods. If \(H (j ) \epsilon {H (0), . . . , H (R - 1)}\), then the
+formula `j mod R' defines the index (among those in \{0,\ldots{},R -
+1\}) of the granule that must be shifted to obtain H (j).The number of
+periods each granule of G composing H (j mod R) should be shifted is
+given by \(\lfloor j/R \rfloor\).
+
+Granularities can be periodical with respect to other granularities,
+except for a finite number of spans of time where they behave in an
+anomalous way \citep{Bettini2000-vy}.
+
+\textbackslash{}begin\{timedom\}\label{def:quasi} A granularity H is
+quasi-periodical with respect to a granularity G if (1) G
+\trianglelefteq H, and (2) there exist a set of intervals E1,\ldots{},Ez
+(the granularity exceptions) and positive integers R,P, where R is less
+than the minimum of the number of granules of H between any 2
+exceptions, such that for all i \epsilon Z, if H(i) = \bigcup\emph{\{k
+\in [0,k]\} G(j\_r) and H (i + R) \neq \phi and i + R \textless{}
+min(E), where E is the closest existing exception after H (i)\^{}2, then
+H (i + R) = \bigcup}\{k \in [0,k]\} G(j\_r + P ).
+\textbackslash{}end\{timedom\}
+
+Intuitively, the definition requires that all granules of H within the
+span of time between two exceptions have the same periodical behavior,
+characterized by R and P.
 
@@
 
-**Example:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
-
-<!-- Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
-The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
--->
-
-#### Computation of linear time granularities
-
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf].
-It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
-new granularities from the bottom one or recursively, from those already generated.
-Thus, the relationship between the operand(s) and the resulting granularities are encoded
-in the operations.
-
-The calendar algebra consists of two kinds of operations: grouping-oriented and
-granule-oriented operations.
-The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
-Example can be to consider a calendar with only two linear granularities minute and hour and hour is generated by grouping every 60 minutes.
-The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
-For example, one can choose to look at the granularity "Monday" and hence select only Mondays while looking at the linear granularity "day".
-
-Some relevant grouping oriented operations are discussed, which will be used in Section \ref{sec:cyclic-gran-def} to define circular and aperiodic granularities.
-
-- **The grouping operation** : Let G be a full-integer labeled granularity, and m a positive integer.
-The grouping operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G into m-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
-each integer i, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
-
-**Example** :
-
+\textbf{Example:} Given the set of all granularities defined over the
+time domain (Z; \textless{}), and the granularity relationship
+\(\trianglelefteq\) (groups into), the granularity mapping each index
+into the corresponding instant (same integer number as the index) is a
+bottom granularity with respect to \(\trianglelefteq\).
+
+\hypertarget{computation-of-linear-time-granularities}{%
+\paragraph{Computation of linear time
+granularities}\label{computation-of-linear-time-granularities}}
+
+Linear time granularities are computed through an algebraic
+representation for time granularities, which is referred to as calendar
+algebra\citep{Ning2002-tf}. It is assumed that there exists a ``bottom''
+granularity and Calendar algebra operations are designed to generate new
+granularities from the bottom one or recursively, from those already
+generated. Thus, the relationship between the operand(s) and the
+resulting granularities are encoded in the operations.
+
+The calendar algebra consists of two kinds of operations:
+grouping-oriented and granule-oriented operations. The grouping-oriented
+operations combine certain granules of a granularity together to form
+the granules of the new granularity. Example can be to consider a
+calendar with only two linear granularities minute and hour and hour is
+generated by grouping every 60 minutes. The granule-oriented operations
+do not change the granules of a granularity, but rather make choices of
+which granules should remain in the new granularity. For example, one
+can choose to look at the granularity ``Monday'' and hence select only
+Mondays while looking at the linear granularity ``day''.
+
+Some relevant grouping oriented operations are discussed, which will be
+used in Section \ref{sec:cyclic-gran-def} to define circular and
+aperiodic granularities.
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{The grouping operation} : Let G be a full-integer labeled
+  granularity, and m a positive integer. The grouping operation
+  \(Group_m(G)\) generates a new granularity G, by partitioning the
+  granules of G into m-granule groups and making each group a granule of
+  the resulting granularity. More precisely, \(G = Group_m(G)\) is the
+  full-integer labeled granularity such that for each integer i,
+  \(G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)\).
+\end{itemize}
+
+\textbf{Example} :
+
+\begin{verbatim}
 - $minute = Group_{60}(second)$
@@
 - $week = Group_{7}(day)$,
+\end{verbatim}
 
@@
 
-- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and l, k, m integers, where G2 partitions G1, and $1 \leq l \leq m$.
-The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-The altering-tick operation modifies the granules of G1 so that the lth granule of each group has |k| additional (or fewer when k < 0) granules of G2.
-
-**Example** :
-
-  - pseudomonth = 
-  $Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, Alter^{12}_{6,-1}(day,   Alter^{12}_{4,-1)}(day, Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))$,
+\begin{itemize}
+\tightlist
+\item
+  \textbf{The altering-tick operation} : Let G1, G2 be full-integer
+  labeled granularities, and l, k, m integers, where G2 partitions G1,
+  and \(1 \leq l \leq m\). The altering-tick operation
+  \(Alter^{m}_{l,k} (G2, G1)\) generates a new full-integer labeled
+  granularity by periodically expanding or shrinking granules of G1 in
+  terms of granules of G2. The altering-tick operation modifies the
+  granules of G1 so that the lth granule of each group has
+  \textbar{}k\textbar{} additional (or fewer when k \textless{} 0)
+  granules of G2.
+\end{itemize}
+
+\textbf{Example} :
+
+\begin{itemize}
+\tightlist
+\item
+  pseudomonth =
+  \(Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, Alter^{12}_{6,-1}(day, Alter^{12}_{4,-1)}(day, Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))\),
   where the granularity pseudomonth is generated by grouping 31 days,
   and then shrink April (4), June (6), September (9) and November (11)
-                  by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
-
-### Cyclic {#sec:cyclic-gran-def}
-
-In a cyclic organization of time, the domain is composed of a set of recurring time values.
-Hence, any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
-Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
-
-We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
-A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble.
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
-
-The mappings of two linear granularities forming a cyclic granularity can be regular or irregular.
-It will be regular if, for example, they are formed by grouping operation.
-Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity.
-However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity.
-The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities.
-However, the formalism would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.
-
-#### Circular {#sec:circular-gran-def}
+  by 1 day, and shrink February (2) by 3 days. \citep{Ning2002-tf}
+\end{itemize}
+
+\hypertarget{sec:cyclic-gran-def}{%
+\subsubsection{Cyclic}\label{sec:cyclic-gran-def}}
+
+In a cyclic organization of time, the domain is composed of a set of
+recurring time values. Hence, any time value can be proceeded and
+succeeded by another time value (for e.g Monday comes before Wednesday
+but Monday also succeeds Wednesday). Intuitively, cyclic granularities
+are additional abstractions of time which are not linear and hence the
+index order of linear granularities needs to be manipulated so that it
+leads to repetitive categorization of time.
+
+We propose a formalism of cyclic time granularities through the tsibble
+\citep{wang2019tsibble} framework of organizing temporal data. A time
+domain, as defined by Bettini, is essentially a mapping of the index set
+to the time index of a tsibble. A linear granularity is a mapping of the
+index set to subsets of the time domain. For example, if the time index
+is days, then a linear granularity might be weeks, months or years.
+Lining up with the the assumption in \citep{Bettini2000-vy} that all
+linear granularities can be generated from the bottom granularity by
+calendar algebraic operations, we assume that a bottom granularity
+exists and is represented by the index of the tsibble.
+
+The mappings of two linear granularities forming a cyclic granularity
+can be regular or irregular. It will be regular if, for example, they
+are formed by grouping operation. Grouping operation would ensure that
+each granule of the resulting granularity consists of same number of
+granules of bottom granularity. However, if they are formed by the
+altering-tick operation, for example, granules of the resulting
+granularity is composed of different number of granules of bottom
+granularity. The relationship can still be periodic, depending on the
+choice of periods and hence these are also useful in exploring
+periodicities. However, the formalism would differ for circular (regular
+mapping) and aperiodic (irregular mapping) granularities.
+
+\hypertarget{sec:circular-gran-def}{%
+\paragraph{Circular}\label{sec:circular-gran-def}}
 
@@
 
-**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
-Each granule is represented by a box.
-The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set and the bottom granularity.
-The index of the tsibble considered is days.
-G and H are two linear granularities denoting days and weeks respectively.
-The period length of days with respect to weeks is 7, hence, $C_{G, BG}(z)$ represented by C(BG, G) is given by $BG(z\mod 7)$.
-The circular granularity C(BG, G) consists of repetitive pattern {BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)} which repeats itself after each period length.
-Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14.
-C(BG, H) also repeats the pattern ${BG(0), BG(1), \dots, BG(13)}$ in every period.
-Hence, circular granularities are represented as repetitive categorization of time in the diagram.
-
-```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
-knitr::include_graphics("Figs/circular-gran-diagram.png")
-```
-
-All circular granularities can be expressed in terms of the bottom linear granularity with equation \autoref{eq:eq2}.
-
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the period length of G within H.
-
-Table \ref{tab:definitions} shows representation of circular granularities relating two linear granularities.
-It is possible that none of these two linear granularities are bottom granularities.
-But the representation of the resultant circular granularity will be a function of the bottom granularity.
-For example, suppose the bottom granularity is minutes, and let $k_i$ is the period length of $C_i$.
+\textbf{Discussion}: Example showing circular granularities relating two
+linear granularities each with bottom granularities are visually
+depicted in a series of slots in the \autoref{fig:circular-gran}. Each
+granule is represented by a box. The diagram also illustrates that the
+granules that overlap share elements from the underlying time domain.
+The first slot in the diagram shows the index set and the bottom
+granularity. The index of the tsibble considered is days. G and H are
+two linear granularities denoting days and weeks respectively. The
+period length of days with respect to weeks is 7, hence,
+\(C_{G, BG}(z)\) represented by C(BG, G) is given by \(BG(z\mod 7)\).
+The circular granularity C(BG, G) consists of repetitive pattern
+\{BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)\} which repeats itself
+after each period length. Similarly, C(BG, H) is given by
+\(BG(z\mod 14)\) since the period length of days with respect to
+fortnight is 14. C(BG, H) also repeats the pattern
+\({BG(0), BG(1), \dots, BG(13)}\) in every period. Hence, circular
+granularities are represented as repetitive categorization of time in
+the diagram.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/circular-gran-diagram}
+
+}
+
+\caption{Relating circular granularities and bottom linear granularity }\label{fig:circular-gran}
+\end{figure}
+
+All circular granularities can be expressed in terms of the bottom
+linear granularity with equation \autoref{eq:eq2}.
+
+In general, any circular granularity relating two linear granularity,
+none of which are bottom granularity can be expressed as
+\(C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))\), where
+linear granularity H is periodic with respect to linear granularity G
+and k(G, H) represents the period length of G within H.
+
+Table \ref{tab:definitions} shows representation of circular
+granularities relating two linear granularities. It is possible that
+none of these two linear granularities are bottom granularities. But the
+representation of the resultant circular granularity will be a function
+of the bottom granularity. For example, suppose the bottom granularity
+is minutes, and let \(k_i\) is the period length of \(C_i\).
 
@@
 Minute-of-Hour: & $C_1(z)$ & $= z \mod 60$ & $k_1 =$&$60$ \\
-Minute-of-Day: & $C_2(z)$ & $= z \mod 60*24$ & $k_2=$&$`r 60*24`$\\
+Minute-of-Day: & $C_2(z)$ & $= z \mod 60*24$ & $k_2=$&$1440$\\
 Hour-of-Day: & $C_4(z)$ & $= \lfloor z/60\rfloor\mod 24$  & $k_3 =$&$24$ \\
-Hour-of-Week: & $C_6(z)$ & $= \lfloor z/60\rfloor\mod 24*7$  & $k_4=$&$`r 24*7`$\\
+Hour-of-Week: & $C_6(z)$ & $= \lfloor z/60\rfloor\mod 24*7$  & $k_4=$&$168$\\
 Day-of-Week: & $C_7(z)$ & $= \lfloor z/24*60\rfloor \mod 7$ & $k_5=$&$7$\\
@@
 
-### Aperiodic {#sec:aperiodic-gran-def}
-
-An **Aperiodic granularity** can not be defined using modular arithmetic.
-The modulus for these type of calendar categorizations are not constant due to irregular mapping with bottom granularities.
-For example,\autoref{tab:aperiodic} shows some example of aperiodic granularities ($A_i$) and plausible choices of period lengths($k_i$).
-The coarse granularity is assumed as the period within which the finer granularity will repeat its behavior aperiodically.
+\hypertarget{sec:aperiodic-gran-def}{%
+\subsubsection{Aperiodic}\label{sec:aperiodic-gran-def}}
+
+An \textbf{Aperiodic granularity} can not be defined using modular
+arithmetic. The modulus for these type of calendar categorizations are
+not constant due to irregular mapping with bottom granularities. For
+example,\autoref{tab:aperiodic} shows some example of aperiodic
+granularities (\(A_i\)) and plausible choices of period
+lengths(\(k_i\)). The coarse granularity is assumed as the period within
+which the finer granularity will repeat its behavior aperiodically.
 
@@
 
-**Definition: Aperiodic granularity** An aperiodic granularity A(AG, BG) that relates linear granularities AG and bottom granularity BG, with AG being periodical with respect to BG, is given by
+\textbf{Definition: Aperiodic granularity} An aperiodic granularity
+A(AG, BG) that relates linear granularities AG and bottom granularity
+BG, with AG being periodical with respect to BG, is given by
 
@@
 A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
-\end{equation}
-where,  
-z denotes the index set (index of BG),
-R' denotes the number of granules of AG in each period of BG,  
-$T_w$ are the sets of indices of BG describing $AG(w)$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
-$\vert T_w \vert$ is the cardinality of set $T_w$.
-
-Since AG is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:
-
-- $BG \trianglelefteq AG$
-- $AG(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.
-
-**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:aperiodic-gran}.
-Each granule is represented by a box.
-Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
-Two linear granularities AG and AH are considered, where $AG = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$.
-This implies that AG is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and AH is made up of shrinking every 1st granule of Group_3(BG) by 2 granules.
-Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different.
-A(BG, AG) and A(BG, AH) are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH.
-For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
-For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
+\end{equation} where,\\
+z denotes the index set (index of BG), R' denotes the number of granules
+of AG in each period of BG,\\
+\(T_w\) are the sets of indices of BG describing \(AG(w)\) such that
+\(AG(w) = \bigcup_{z \in T_w}BG(z)\), and \(T_w = 0\) for
+\(w \epsilon Z_{<0}\)\\
+\(\vert T_w \vert\) is the cardinality of set \(T_w\).
+
+Since AG is periodical with respect to BG, from definition of ``Groups
+periodically into'' the following holds true:
+
+\begin{itemize}
+\tightlist
+\item
+  \(BG \trianglelefteq AG\)
+\item
+  \(AG(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)\),
+  where P' denotes the period of the granularity BG.
+\end{itemize}
+
+\textbf{Discussion}: Example showing aperiodic granularities relating
+two linear granularities each with bottom granularities are visually
+depicted in a series of slots in \autoref{fig:aperiodic-gran}. Each
+granule is represented by a box. Similar to \autoref{fig:circular-gran},
+the granules that overlap share elements from the underlying time domain
+and the first slot shows the index set and the bottom granularity. Two
+linear granularities AG and AH are considered, where
+\(AG = Alter^{2}_{(1,-1)}(BG, Group_3(BG))\) and
+\(AH = Alter^{2}_{(1,-2)}(BG, Group_7(BG))\). This implies that AG is
+made up by shrinking every 1st granule of Group\_3(BG) by 1 granule and
+AH is made up of shrinking every 1st granule of Group\_3(BG) by 2
+granules. Number of granules of AG and AH in each period of BG is 2, but
+the number of granules of BG in each of those granules are different.
+A(BG, AG) and A(BG, AH) are repetitive categorization of time, similar
+to circular granularities, except that the number of granules of BG is
+not necessarily the same across different granules of AG or AH. For AG,
+R' = 2, P'= 5, \(T_0\) = \{0,1\} and \(T_1\) = \{2, 3, 4\}. For AH, R' =
+2, P'= 12, \(T_0\) = \{0, 1, 2, 3, 4, 5, 6\} and \(T_1\) = \{7, 8, 9,
+10, 11\}, then we will have \autoref{eq:eq4} and \autoref{eq:eq5}.
 
@@
 
-```{r aperiodic-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
-knitr::include_graphics("Figs/aperiodic-granv1.png")
-```
-
-If we ignore leap years and all its exception, the periodicity of the grouping (days, months), P' = 365, R' = 12, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
-Considering leap years and all their exceptions, period is 400 years and hence $P' = (3*365 + 366)$, $R' = 12*400$.
-\autoref{eq:eq3} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
-
-<!-- Then $f(x, y)$ can be  computed using modular arithmetic as follows: -->
-
-<!-- $$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$ -->
-
-<!-- The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
--->
-
-<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
-Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
--->
-
-<!-- Formal definition of circular granularity -->
-
-<!-- **Definition: Equivalence class** Let $m \in N \backslash{0}$.
-For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
--->
-
-<!-- The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$.
-Thus $Z_m = \{[0], [1], ..., [m-1]\}$.
-However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.
--->
-
-<!-- **Definition:** A **circular granularity** $C$, joining two linear granularity with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that -->
-
-<!-- $C(s) = (s\mod m)$ for $s \in Z$. -->
-
-<!-- For example, suppose $C$ is a circular granularity denoting hour-of-day and we have hourly data for 100 hours.
-The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
--->
-
-<!-- **Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once. -->
-
-<!-- **Definition:** A **circular granule** represents an equivalence class inside each cycle. -->
-
-<!-- ### Aperiodic {#aperiodic-gran-def} -->
-<!-- Formal definition of aperiodic circular granularity -->
-
- <!-- Why we need single order up and multiple order up granularities -->
-<!-- A time series which is characterized by being composed of repeating cycles. -->
-
-<!-- Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
-Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
--->
-
-The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity.
-The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
-It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi})
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/aperiodic-granv1}
+
+}
+
+\caption{Relating circular granularities and bottom linear granularity }\label{fig:aperiodic-gran}
+\end{figure}
+
+If we ignore leap years and all its exception, the periodicity of the
+grouping (days, months), P' = 365, R' = 12, since each month would be
+defined as the grouping of the same number of days (31, 28, or 30,
+depending on the month) every year. Considering leap years and all their
+exceptions, period is 400 years and hence \(P' = (3*365 + 366)\),
+\(R' = 12*400\). \autoref{eq:eq3} can be useful to find the aperiodic
+granularity day-of-month in both of these cases, since the grouping is
+periodic is both cases.
+
+The linear granularities G, H, AG and AH considered in the definitions
+of circular and aperiodic granularities are all periodic with respect to
+the bottom granularity. The only difference is G, H have regular
+mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
+It can be noted here that if a linear granularity AG is quasi-periodic
+with respect to BG, then \autoref{eq:eq3} can be modified as follows to
+account for exceptions \(E = [e_{begin}, e_{end}]\) (by Definition
+\autoref{def:quasi})
 
@@
 
-for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
-
-<!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
-
-## Order: Single vs. Multiple {#sec:order}
-
-The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
-We shall use the notion of a hierarchy table and order to define them.
-
-**Definition:  Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
-If we consider two linear granularities G and H, such that G is finer than or groups into H (by Definitions \ref{def:finerthan}, \ref{def:groupsinto}), then H is of higher order than G.
-
-**Notation:  Hierarchy table** Let $H_n: (G, C, k)$ be a hierarchy model containing n linear granularities.
-$G_{l}$ represents the linear granularity of order l and $k(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
-
-In a hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
-
-<!-- Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities and the other granularities are subsequently more "coarse" than the earlier one.
--->
-
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities.
-Let us look at few calendars to see examples of single and multiple order-up granularities.
-
-<!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
--->
-
-### Computation of multiple-order from single-order {#sec:multiplefromsingle}
+for \(z \in T_k\) and \(e_k\) is the number of exceptions in
+\(\bigcup\limits_{w = 0}^{k} T_w\) and
+\(E_{u} = [e_{begin}(u), e_{end}(u)]\).
+
+\hypertarget{sec:order}{%
+\subsection{Order: Single vs.~Multiple}\label{sec:order}}
+
+The hierarchical structure of time creates a natural nested ordering
+which can produce \textbf{single-order-up} or \textbf{multiple-order-up}
+granularities. We shall use the notion of a hierarchy table and order to
+define them.
+
+\textbf{Definition: Order} of a linear granularity can be comprehended
+as the level of graininess associated with a linear granularity. If we
+consider two linear granularities G and H, such that G is finer than or
+groups into H (by Definitions \ref{def:finerthan},
+\ref{def:groupsinto}), then H is of higher order than G.
+
+\textbf{Notation: Hierarchy table} Let \(H_n: (G, C, k)\) be a hierarchy
+model containing n linear granularities. \(G_{l}\) represents the linear
+granularity of order l and \(k(l,m)\) represents the period length of
+\(G_{l}\) with respect to \(G_{m}\) and \(C_{G(l),G(m)}\) represents the
+circular granularity that relates linear granularity of order l and m,
+\(\forall l,m \in n, l<m\).
+
+In a hierarchy table, linear granularities are arranged from lowest to
+highest order of linear granularities.
+
+We refer to granularities which are nested within multiple levels as
+\textbf{multiple-order-up} granularities and those concerning a single
+level as \textbf{single-order-up} granularities. Let us look at few
+calendars to see examples of single and multiple order-up granularities.
+
+\hypertarget{sec:multiplefromsingle}{%
+\subsubsection{Computation of multiple-order from
+single-order}\label{sec:multiplefromsingle}}
 
@@
 
-<!-- some discussions on primary units of calendar -->
-**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar.
-But it is far from being the only one.
-All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars.
-Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
-Nevertheless, for any calendar a hierarchy can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows [@Reingold2001-kf]:
-
- - 1 kin = 1 day
- - 1 uinal = 20 kin
- - 1 tun  = 18 uinal
- - 1 katun = 20 tun
- - 1 baktun = 20 katun
-
-Thus, the hierarchy table for the Mayan calendar would look like the following:
-
-|  G	        |C                 |k
-|------------	|------------------|------------|
-|  kin      	|kin-of-uinal      |  20        |
-|  uinal    	|uinal-of-tun      |  18  	    |
-|  tun      	|tun-of-katun      |  20	      |
-|  katun    	|katun-of-baktun   |  20        |
-|  baktun   	| 1                |	1         |
-
-Examples of multiple-order-up  granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.
-
-Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic and regular.
+\textbf{Example:} So far we have used the Gregorian calendar as it is
+the most widely used calendar. But it is far from being the only one.
+All calendars fall under three types - solar, lunar or
+lunisolar/solilunar but the day is the basic unit of time underlying all
+calendars. Various calendars, however, use different conventions to
+structure days into larger units: weeks, months, years and cycle of
+years. The French revolutionary calendar divided each day into 10
+``hours'', each ``hour'' into 100 ``minutes'' and each ``minute'' into
+100 ``seconds''. Nevertheless, for any calendar a hierarchy can be
+defined. For example, in Mayan calendar, one day was referred to as 1
+kin and the calendar was structured as follows \citep{Reingold2001-kf}:
+
+\begin{itemize}
+\tightlist
+\item
+  1 kin = 1 day
+\item
+  1 uinal = 20 kin
+\item
+  1 tun = 18 uinal
+\item
+  1 katun = 20 tun
+\item
+  1 baktun = 20 katun
+\end{itemize}
+
+Thus, the hierarchy table for the Mayan calendar would look like the
+following:
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+G & C & k\tabularnewline
+\midrule
+\endhead
+kin & kin-of-uinal & 20\tabularnewline
+uinal & uinal-of-tun & 18\tabularnewline
+tun & tun-of-katun & 20\tabularnewline
+katun & katun-of-baktun & 20\tabularnewline
+baktun & 1 & 1\tabularnewline
+\bottomrule
+\end{longtable}
+
+Examples of multiple-order-up granularities can be kin-of-tun or
+kin-of-baktun whereas examples of single-order-up granularities may
+include kin-of-uinal, uinal-of-tun etc.
+
+Let us use the equation \ref{eq:eq3} to compute the multiple-order-up
+granularity uinal\_katun for Mayan calendar, which is periodic and
+regular.
 
@@
 
-### Computation of single-order from multiple-order {#sec:singlefrommultiple}
-
-For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
+\hypertarget{sec:singlefrommultiple}{%
+\subsubsection{Computation of single-order from
+multiple-order}\label{sec:singlefrommultiple}}
+
+For a hierarchy table \(H_n: (G, C, k)\) with
+\(l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}\) and \(l_2<l_1\) and
+\(m_2>m_1\), we have
 
@@
 
-**Example:** Considering the same example of Mayan Calendar,  it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:eq5}
+\textbf{Example:} Considering the same example of Mayan Calendar, it is
+possible to compute the single-order-up granularity tun-of-katun given
+the multiple-order-up granularity uinal-baktun using equation
+\ref{eq:eq5}
 
@@
 
-In \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}, we assume that all single order granularities are circular.What happens when we have a mix of circular and aperiodic granularities.
-
-Let us turn to Gregorian calendar for addressing this case.
-Suppose we have a hierarchy table using some linear granularties from Gregorian calendar.
-Since months consists of unequal number of days, any temporal unit which is higher in order than months will also have unequal number of days.
-This is an example of a hierarchy structure which has both circular and aperiodic single-order-up granularities.
-The single-order-up granularity day_month is aperiodic.
-Any single-order-up granularities which are formed by units below days are periodic.
-Similarly, all single-order-up granularities which are formed using units whose orders are higher than months are also periodic.
-
-<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
-However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
--->
-
-|  G	        |C/A               |k
-|------------	|------------------|------------|
-|  minute     |minute-of-hour    |  60        |
-|  hour    	  |hour-of-day       |  24  	    |
-|  day      	|day-of-month      | "aperiodic"|
-|  month    	|month-of-year     |  12        |
-|  year   	  | 1                |	1         |
-
-There can be three scenarios for obtaining multiple-order-granularities here:
-- granularities relating two linear granularities whose orders are less than day
-- granularities relating two linear granularities whose orders are more than month
-- granularities relating two linear granularities with order at most day and another with order at least month
-
-The single order-up granularites resulting from the first two cases are periodic and has been handled in \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}.
-The calendar categorization resulting from the last case are aperiodic.
-Examples might include day of the quarter or hour of the month.
-Let us consider the computation of $A_{hour, month}$ given single order-up granularities $A_{hour, day} = 12$, $A_{day, month} = 20$ and $C_{month, year} = 9$.
+In \autoref{sec:singlefrommultiple} and
+\autoref{sec:multiplefromsingle}, we assume that all single order
+granularities are circular.What happens when we have a mix of circular
+and aperiodic granularities.
+
+Let us turn to Gregorian calendar for addressing this case. Suppose we
+have a hierarchy table using some linear granularties from Gregorian
+calendar. Since months consists of unequal number of days, any temporal
+unit which is higher in order than months will also have unequal number
+of days. This is an example of a hierarchy structure which has both
+circular and aperiodic single-order-up granularities. The
+single-order-up granularity day\_month is aperiodic. Any single-order-up
+granularities which are formed by units below days are periodic.
+Similarly, all single-order-up granularities which are formed using
+units whose orders are higher than months are also periodic.
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+G & C/A & k\tabularnewline
+\midrule
+\endhead
+minute & minute-of-hour & 60\tabularnewline
+hour & hour-of-day & 24\tabularnewline
+day & day-of-month & ``aperiodic''\tabularnewline
+month & month-of-year & 12\tabularnewline
+year & 1 & 1\tabularnewline
+\bottomrule
+\end{longtable}
+
+There can be three scenarios for obtaining multiple-order-granularities
+here: - granularities relating two linear granularities whose orders are
+less than day - granularities relating two linear granularities whose
+orders are more than month - granularities relating two linear
+granularities with order at most day and another with order at least
+month
+
+The single order-up granularites resulting from the first two cases are
+periodic and has been handled in \autoref{sec:singlefrommultiple} and
+\autoref{sec:multiplefromsingle}. The calendar categorization resulting
+from the last case are aperiodic. Examples might include day of the
+quarter or hour of the month. Let us consider the computation of
+\(A_{hour, month}\) given single order-up granularities
+\(A_{hour, day} = 12\), \(A_{day, month} = 20\) and
+\(C_{month, year} = 9\).
 
@@
 \end{split}
-\end{equation}
-where, $month(w) = \bigcup_{z \in T_w}day(z)$
-
-Clearly, the computation of multiple order-up granularities become more involved with aperiodic mappings of linear granularities in the calendar structure.
-With one aperiodic granularity, we can write the following:
-
-A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included}
-
-# Synergy of cyclic time granularities {#sec:synergy}
-
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-<!-- What  is interaction and some examples why it is important to study interaction -->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
-
-<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
--->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
--->
-
-<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
--->
-
-<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
--->
-
-<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
--->
-
-<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
--->
-
-<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
--->
-
-<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
--->
-
-Before exploring the behavior of a "dependent variable" across time granularities,
-it is important to know how these granularities interact with each other.
-If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
-To define a general framework, we define harmony and clashes.
-
-<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
--->
-
-## Harmony and Clashes
-
-<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
-It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
--->
-
-Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
-That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.Now, many situations can lead to any of these sets being empty.
-Let us discuss the following cases, where one or more of these sets can be empty.
-
-Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
-These are structurally empty sets in that it is impossible for them to have any observations.
-
-<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
-
-Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
-These are called "event-driven" empty combinations.
-Again, let us consider a specific example to illustrate this.
-Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
-Here $C_1$ can take 7 values while $C_2$ can take 2 values.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
-While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
-For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
-
-<!-- However, this is different from the first example because these are not structurally empty.
-Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
--->
-
-Thirdly, empty combinations can be a result of how granularities are constructed.
-Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
-Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
-This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
-Hence, they are referred to as "build-based" empty combinations.
-
-An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-
-Therefore, pair of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are called **harmonies**.
-
-<!-- as they promote the exploratory analysis through visualization. -->
-
-```{r}
-# grid.arrange(p1,p2,p3,p4,nrow = 2)
-
-#
-# x <- gridExtra::arrangeGrob (p5, p6, p1, p2, p3, p4,newpage = TRUE)
-# grid::grid.draw(x)
-
-#
-# gridExtra::grid.arrange(
-#   p1,
-#   p2,
-#   p3,
-#   p4,
-#   p5,
-#   p6,
-#   nrow = 3,
-#   top = "Title of the page",
-#   bottom = grid::textGrob(
-#     "this footnote is right-justified",
-#     gp = grid::gpar(fontface = 3, fontsize = 9),
-#     hjust = 1,
-#     x = 1
-#   )
-# )
-# plot.new()
-# mtext('Various probability distribution plot of electricity consumption data of Victoria from 2012 to 2014 (top left) Letter Value plot by DoM and MoY (top right) Decile Plot by HoD and DoW (middle left) Box plot by DoY and DoM (middle right) Violin plot of DoM and WoM (bottom left) Ridge plot by WoM and WoY (bottom right) Decile plot by DoY and MoY', side = 3, line = - 10, outer = TRUE)
-```
-
-<!-- More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
-When C1 and C2 are used as aesthetics or facets, problems arise when we encounter empty combinations.
-Let $S_{i,j}$ be the set of combination of the levels of C1 and C2.
-In this example, we have 12 such sets $S_{i,j}$ because i can take 4 values and j can take 3 values.
--->
-
-<!-- Thus, the common link that differentiates harmonies from clashes are: -->
-
-<!--   a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis. -->
-
-<!--   b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets. -->
-
-# Visualization {#sec:visualization}
-
-<!-- Why visualize -->
-<!-- Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
-However, improper choice of models can lead to wrong predictions.
-One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
--->
-
-<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
-
-<!-- Include why distribution plot, why 2D, why faceting in introduction -->
-<!-- Why distribution plot -->
-<!-- Huge data difficult to summaries using one statistic-->
-
-<!-- With huge amount of data being available, mean, median or any one summary statistic might not enough to understand a data set. Soon enough following questions become more interesting:   -->
-
-<!--  - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?  -->
-
-<!--  - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
-More generally, how the variation is the data set changes across different percentiles/deciles?
--->
-
-<!--   - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median? -->
-
-<!-- This is when displaying a probability distribution becomes a potentially useful approach. -->
-<!-- The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
--->
-
-<!-- ## Distributions conditional on bivariate granularities -->
-
-<!-- Why 2D -->
-
-<!-- Graphical output devices are flat.
-If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
-As soon as the data move to three variables, we need to infer multi-dimensional structure by a two-dimensional medium[Cleveland, elements of graphing data].
-Any attempt to encode all these temporal granularities together to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
--->
-
-<!-- Why facet-ing -->
-
-<!-- `ggplot2` (Wickham, 2016) is an excellent tool that facilitates the process of mapping different variables to a 2D frame through grammar of graphics.
-One aspect of ggplot2 is faceting, inspired by Trellis plots (Cleveland, 1993).
-These are based on conditioning the values taken on by one or more of the variables in a data set.
-This means creating same plots using data subsets corresponding to each level of the conditioned variable.
-This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
--->
-
-<!-- What are we visualizing -->
-
-<!-- We will see the distribution of the time series across bivariate temporal granularities through faceting approach with one temporal granularity plotted along the x-axis, the other one across facets and the dependent time series variable on the y-axis.
--->
-
-<!-- A well designed visual representation can replace cognitive calculations with simple perpetual inferences and improve speed of comprehension and decision making.
-However, creating a visualisation requires a number of nuanced judgements.
-The challenge being given a data set, the number of visual encodings and hence the space of possible visual design is extremely huge.
--->
-
-<!-- The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, allows us to produce graphics using the same structured thinking that is used to design analysis.
-Briefly, any graphics are made up of **Layers**, **Facet**,**Scale** and a  **coordinate system**.
-**Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
--->
-
-Creating a visualisation requires a number of nuanced judgements.
-We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate temporal granularities.
-The general framework of visualizing the distribution invloves a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
-Different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
-We will look at each of these aspects separately and analyze the effect of each on visualization.
-
-While we accomplish that, we need to bear in mind that any customized species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
-We are trying to address problems where even with good aesthetic qualities and good data, the graph might be confusing or misleading because of how people process the information.
-
-<!-- Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
-The harmonies when plotted together can help exploration.
-But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
-This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
--->
-
-## Choice of Plots
-
-<!-- Common benefits and challenges of empirical and density based visualization -->
-
-There are several ways to plot statistical distributions.
-The displayed probabilities in these plots are either computed using kernel density estimation methods or empirical statistical methods.
-The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
-However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
-The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
-Density plots which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution, unlike discretizing the distribution and showing only parts of it.
-They can be useful in spotting multimodality, however, they might become obscuring with too many categories and information on the entire distribution can add to cognitive load.
-Also, the probabilities in density plot are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
-As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
-The density based visualizations are subject to the same sample size restrictions and challenges that apply to any density estimation.
-In practice, the densities are estimated reasonably with at least 30 observations.
-Even with sample sizes of several hundred, however, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
-
-We will discuss few conventional and recent ways to plot distributions using both of these methods.
-
-### Empirical methods
-
-Most commonly used techniques to display a distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-The box plot is a compact distributional summary, displaying less detail than a histogram.
-Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg, 1978] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
-Moreover, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
-Also, the number of outliers is large for larger data set since number of outliers is proportional to the number of observations.
-
-<!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg, 2006] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
-“outliers” in letter value plots are those observations beyond the most extreme letter value.
-The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown and leading to overload of information in one plot.
-
-<!-- Description and lit review of quantile plots -->
-
-Quantile plots visually portray the quantiles of the distribution of sample data.
-Much like the quartiles divide the data set equally into four equal parts, extensions include dividing the data set even further.
-These plots are referred to as "quantile" plots.
-For example, number of quantiles would be 9 for a decile plot and 99 for percentile plot.
-It avoids much clutter and just enable us to focus on specific probabilities, typically representing typical and extreme behaviors.
-A large data set is required for the extreme percentiles to be estimated with any accuracy.
-These plots can display any quantiles instead of quartiles in a traditional boxplot.
-When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g: outside 1.5 times the interquartile range above the upper quartile and below the lower quartile).
-However, outliers are open to interpretation and not shown in an quantile plot.
-
-<!-- Description and lit review of density plots -->
-
-<!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  -->
-
-### Kernel density estimation methods
-
-Traditional ways to visualize densities include violin plots [@Hintze1998-zi, 1998].
-The shape of the violin represents the density estimate of the variable.
-The more data points in a specific range, the larger the violin is for that range.
-Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
-
-The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
-This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
-
-<!-- Description and lit review of density plots -->
- A Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
-Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
-A clear advantage over boxplots is that these plots allow us to see multimodality in the distribution.
-However, these plots can be obscuring when there is overlap of distribution for two or more categories of the y-axis.
-Also, if there are lot of categories, it is difficult to compare the height of the densities across categories.
-
-<!-- Description and lit review of HDR plots -->
-The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
-The probabilities for which the summarization is required can be chosen based on the requirement.
-These regions do not need to be contiguous and help identify multi-modality.
-
-Given a context, it is good to be conversant with the benefits and challenges while choosing a distribution plot.
-As a general rule, if we have too many categories, the quantile plots are useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
-
-### Effect of levels of temporal granularities
-
-The levels of the two granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
-The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.
-Plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.
-
-Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are chosen based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-Any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
-
- <!-- - If levels of both the granularities are low, then any statistical distribution plot might be chosen depending on which feature of the distribution one is interested to look at.  -->
-
-The following principles are useful while choosing distribution plots given two temporal granularities.
-
-  - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus.
-
- - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories.
-
-  - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.
-
-  - If levels of any granularity plotted are more than medium, empirical based methods of distribution visualizing are preferred as they use less space by design than most density based methods.
-
-<!-- Let us consider case by case and see which plots are better suitable in which scenarios. -->
-
-<!--  Assumptions should be made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.  -->
-<!-- Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot. -->
-
-<!-- Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
-These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
-Let us consider case by case and see which plots are better suitable in which scenarios.
--->
-
-<!-- Faceting creates small multiples of data each representing a different subset of the data.
-Small multiples are powerful for exploratory data analysis where we can rapidly compare patterns in different parts of the data and see whether they are the same or different.
-But space and resolution quickly become a problem if the number of levels of facet variables is too high.
-Also, Cleveland suggested that a graphic should display as much information as it can, with the lowest possible cognitive strain to the viewer.
--->
-
-## Effect of synergy of time granularities
-
-In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
-In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.
-
-For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
-\autoref{fig:allFig}.
-\autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
-\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
-\autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig} (f) shows decile plots across  day of the year and month of the year.
-
-Clearly, in \autoref{fig:allFig}, we observe that some choices of time granularities work and others do not.
-In \autoref{fig:allFig} (c), there will be no observations for some combinations "Day-of-Month" and "Day-of-Year".
-In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every "Day-of-Week" can occur in any "Month-of-Year".
-Thus the graphs that don't work are those where many of the combination sets are empty.
-In other words, if there are levels of time granularities plotted across x-axis which are not spanned by levels of the time granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
-We hypothesize that the synergy of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
-
-We redefine harmony and clashes as follows: harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-As a result, we should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
-
-```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
+\end{equation} where, \(month(w) = \bigcup_{z \in T_w}day(z)\)
+
+Clearly, the computation of multiple order-up granularities become more
+involved with aperiodic mappings of linear granularities in the calendar
+structure. With one aperiodic granularity, we can write the following:
+
+A\_\{P, Q)\} \& = (A\{AG\_Q\} - 1)*k(P\_AG) + A\_\{P, AG\} \{TO REVISE:
+many aperiodic granularties not to be included\}
+
+\hypertarget{sec:synergy}{%
+\section{Synergy of cyclic time granularities}\label{sec:synergy}}
+
+Before exploring the behavior of a ``dependent variable'' across time
+granularities, it is important to know how these granularities interact
+with each other. If two time granulaties of interest interact, the
+relationship between each of the interacting variables and the dependent
+variable will depend on the value of the other interacting variable. To
+define a general framework, we define harmony and clashes.
+
+\hypertarget{harmony-and-clashes}{%
+\subsection{Harmony and Clashes}\label{harmony-and-clashes}}
+
+Suppose we have two cyclic granularities \(C_1\) and \(C_2\), such that
+\(C_1\) maps index set to a set \(\{A_1,A_2,A_3,\dots, A_n\)\}, and
+\(C_2\) maps index set to a set \(\{B_1,B_2,B_3,\dots, B_m\)\}. That is,
+let \(S_{ij}\) be the set of index set such that for all
+\(s \in S_{ij}\), \(C_1(s) = i\) and \(C_2(s) = j\). Since, \(C_1\) has
+n levels and \(C_2\) has m levels there will be \(nm\) such sets
+\(S_{ij}\).Now, many situations can lead to any of these sets being
+empty. Let us discuss the following cases, where one or more of these
+sets can be empty.
+
+Firstly, empty combinations can arise due to the structure of the
+calendar or hierarchy. These are called ``structurally'' empty
+combinations. Let us take a specific example, where \(C_1\) maps row
+numbers to Day-of-Month and \(C_2\) maps row numbers to Week-of-Month.
+Here \(C_1\) can take 31 values while \(C_2\) can take 5 values. There
+will be \(31\times 5=155\) sets \(S_{ij}\) corresponding to the possible
+combinations of WOM and DOM. Many of these are empty. For example
+\(S_{1,5}\), \(S_{21,2}\), etc.This is also intuitive since the first
+day of the month can never correspond to fifth week of the month. These
+are structurally empty sets in that it is impossible for them to have
+any observations.
+
+Secondly, empty combinations can turn up due to differences in event
+location or duration in a calendar. These are called ``event-driven''
+empty combinations. Again, let us consider a specific example to
+illustrate this. Let \(C_1\) be DOW and \(C_2\) be
+WorkingDay/NonWorkingDay. Here \(C_1\) can take 7 values while \(C_2\)
+can take 2 values. So there are 14 sets \(S_{ij}\) corresponding to the
+possible combinations of DOW and WD/NWD. While potentially all of these
+can be non-empty (it is possible to have a public holiday on any DOW),
+in practice many of these combinations will probably have very few
+observations. For example, there are few if any public holidays on
+Wednesdays or Thursdays in any given year in Melbourne.
+
+Thirdly, empty combinations can be a result of how granularities are
+constructed. Let \(C_1\) maps row numbers to ``Business days'', which
+are days from Monday to Friday except holidays and \(C_2\) is
+Day-of-Month. Then the weekends in Days-of-Month would not correspond to
+any Business days and would have missing observations due to the way the
+granularities are constructed. This is different from the structurally
+empty combinations because structure of the calendar does not lead to
+these missing combinations, but the construction of the granularity
+does. Hence, they are referred to as ``build-based'' empty combinations.
+
+An example when there will be no empty combinations could be where
+\(C_1\) maps row numbers to Day-of-Week and \(C_2\) maps row numbers to
+Month-of-Year. Here \(C_1\) can take 7 values while \(C_2\) can take 12
+values. So there are \(12\times7=84\) sets \(S_{ij}\) corresponding to
+the possible combinations of DOW and MOY. All of these are non-empty
+because every DOW can occur in every month.
+
+Therefore, pair of circular/aperiodic granularities which lead to
+structurally, event-driven or build-based empty-combinations are
+referred to as to as \textbf{clashes}. And the ones that do not lead to
+any missing combinations are called \textbf{harmonies}.
+
+\hypertarget{sec:visualization}{%
+\section{Visualization}\label{sec:visualization}}
+
+Creating a visualisation requires a number of nuanced judgements. We
+exploit some features of layers and faceting \citep{Wilkinson1999-nk}
+and \citep{Wickham2009-pk} to come up with a recommendation system while
+visualizing the distribution of the ``dependent'' variable across
+bivariate temporal granularities. The general framework of visualizing
+the distribution invloves a faceting approach with one temporal
+granularity plotted along the x-axis, the other one across facet and the
+dependent time series variable on the y-axis. Different distribution
+plots might be appropriate depending on which features of the
+distrbution we are interested to look at, the levels of time
+granularities being plotted, how granularities interact and number of
+observations available. We will look at each of these aspects separately
+and analyze the effect of each on visualization.
+
+While we accomplish that, we need to bear in mind that any customized
+species of visualization might be constructed by varying the set of
+mappings between data properties and visual attributes such as position,
+shape and color. We are trying to address problems where even with good
+aesthetic qualities and good data, the graph might be confusing or
+misleading because of how people process the information.
+
+\hypertarget{choice-of-plots}{%
+\subsection{Choice of Plots}\label{choice-of-plots}}
+
+There are several ways to plot statistical distributions. The displayed
+probabilities in these plots are either computed using kernel density
+estimation methods or empirical statistical methods. The latter do not
+allow us to see the shape, skewness, nature of the tail or
+multi-modality with so much clarity as the former. However, they avoid
+much clutter, where some specific probabilities representing typical and
+extreme behavior are on focus. The probabilities displayed are based on
+actual data, which is aligned to principles that governed Tukey's
+original boxplot. Density plots which uses a kernel density estimate to
+show the probability density function of the variable can show the
+entire distribution, unlike discretizing the distribution and showing
+only parts of it. They can be useful in spotting multimodality, however,
+they might become obscuring with too many categories and information on
+the entire distribution can add to cognitive load. Also, the
+probabilities in density plot are estimated through kernel density
+estimation and thus makes assumptions when selecting kernel or
+bandwidth. As a result, the plot shows smooth summaries based on the
+assumptions and not only on actual observations. The density based
+visualizations are subject to the same sample size restrictions and
+challenges that apply to any density estimation. In practice, the
+densities are estimated reasonably with at least 30 observations. Even
+with sample sizes of several hundred, however, choosing too large a
+value for bandwidth can cause the estimate to oversmooth the data.
+
+We will discuss few conventional and recent ways to plot distributions
+using both of these methods.
+
+\hypertarget{empirical-methods}{%
+\subsubsection{Empirical methods}\label{empirical-methods}}
+
+Most commonly used techniques to display a distribution of data include
+the histogram (Karl Pearson), which shows the prevalence of values
+grouped into bins and the box-and-whisker plots \citep{Tukey1977-jx}
+which convey statistical features such as the median, quartile
+boundaries, hinges, whiskers and extreme outliers. The box plot is a
+compact distributional summary, displaying less detail than a histogram.
+Due to wide spread popularity and simplicity in implementation, a number
+of variations are proposed to the original one which provides alternate
+definitions of quantiles, whiskers, fences and outliers. Notched box
+plots \citep[1978]{Mcgill1978-hg} has box widths proportional to the
+number of points in the group and display confidence interval around
+medians aims to overcome some drawbacks of box plots. The standard box
+plot and all of these variations are helpful to get an idea of the
+distribution at a glance. Moreover, for data less than 1000
+observations, detailed estimates of tail behavior beyond the quartiles
+are not trustworthy. Also, the number of outliers is large for larger
+data set since number of outliers is proportional to the number of
+observations.
+
+The letter-value box plot \citep[2006]{Hofmann2017-sg} was designed to
+adjust for number of outliers proportional to the data size and display
+more reliable estimates of tail. ``outliers'' in letter value plots are
+those observations beyond the most extreme letter value. The letter
+values are shown till the depths where the letter values are reliable
+estimates of their corresponding quantiles and hence might lead to a lot
+of letter values being shown and leading to overload of information in
+one plot.
+
+Quantile plots visually portray the quantiles of the distribution of
+sample data. Much like the quartiles divide the data set equally into
+four equal parts, extensions include dividing the data set even further.
+These plots are referred to as ``quantile'' plots. For example, number
+of quantiles would be 9 for a decile plot and 99 for percentile plot. It
+avoids much clutter and just enable us to focus on specific
+probabilities, typically representing typical and extreme behaviors. A
+large data set is required for the extreme percentiles to be estimated
+with any accuracy. These plots can display any quantiles instead of
+quartiles in a traditional boxplot. When reviewing a boxplot, an outlier
+is defined as a data point that is located outside the fences
+(``whiskers'') of the boxplot (e.g: outside 1.5 times the interquartile
+range above the upper quartile and below the lower quartile). However,
+outliers are open to interpretation and not shown in an quantile plot.
+
+\hypertarget{kernel-density-estimation-methods}{%
+\subsubsection{Kernel density estimation
+methods}\label{kernel-density-estimation-methods}}
+
+Traditional ways to visualize densities include violin plots
+\citep[1998]{Hintze1998-zi}. The shape of the violin represents the
+density estimate of the variable. The more data points in a specific
+range, the larger the violin is for that range. Adding two density plots
+gives a symmetric plot which makes it easier to see the magnitude of the
+density and compare across categories, enabling easier detection of
+clusters or bumps within a distribution.
+
+The summary plot \citep[2010]{Potter2010-qc} combines a minimal box plot
+with glyphs representing the first five moments (mean, standard
+deviation, skewness, kurtosis and tailings), and a sectioned density
+plot crossed with a violin plot (both color and width are mapped to
+estimated density), and an overlay of a reference distribution. This
+suffers from the same problem as boxplots or violin plot, as it is
+combination of those two.
+
+A Ridge line plot (sometimes called Joy plot) shows the distribution of
+a numeric value for several groups. Distribution can be represented
+using histograms or density plots, all aligned to the same horizontal
+scale and presented with a slight overlap. A clear advantage over
+boxplots is that these plots allow us to see multimodality in the
+distribution. However, these plots can be obscuring when there is
+overlap of distribution for two or more categories of the y-axis. Also,
+if there are lot of categories, it is difficult to compare the height of
+the densities across categories.
+
+The highest density region (HDR) box plot proposed by
+\citep{Hyndman1996-ft} displays a probability density region that
+contains points of relatively highest density. The probabilities for
+which the summarization is required can be chosen based on the
+requirement. These regions do not need to be contiguous and help
+identify multi-modality.
+
+Given a context, it is good to be conversant with the benefits and
+challenges while choosing a distribution plot. As a general rule, if we
+have too many categories, the quantile plots are useful for comparing
+patterns, whereas, other more involved methods of plotting are useful
+for studying anomalies, outlier or multimodal behavior.
+
+\hypertarget{effect-of-levels-of-temporal-granularities}{%
+\subsubsection{Effect of levels of temporal
+granularities}\label{effect-of-levels-of-temporal-granularities}}
+
+The levels of the two granularities plotted have an impact on the choice
+of plots since space and resolution might become a problem if the number
+of levels are too high. The criteria for different levels could be based
+on usual cognitive power while comparing across facets and display size
+available to us. Plot choices will also vary depending on which
+granularity is placed on the x-axis and which one across facets.
+
+Levels are categorized as very high/high/medium/low each for the facet
+variable and the x-axis variable. Default values for these levels are
+chosen based on levels of common temporal granularities like day of the
+month, day of a fortnight or day of a week. Any levels above 31 can be
+considered as very high, any levels between 14 to 31 can be taken as
+high and that between 7 to 14 can be taken as medium and below 7 as low.
+31, 14 and 7 are the levels of days-of-month, days-of-fortnight and
+days-of week respectively.
+
+The following principles are useful while choosing distribution plots
+given two temporal granularities.
+
+\begin{itemize}
+\item
+  If levels of both granularity plotted are low/medium, then any
+  distribution plots might be chosen depending on which feature of the
+  distribution needs focus.
+\item
+  If level of the granularity plotted across x-axis is more than medium,
+  ridge plots should be avoided to escape overlap of categories.
+\item
+  If level of the granularity plotted across x-axis is more than or
+  equal to high, quantile plots are preferred.
+\item
+  If levels of any granularity plotted are more than medium, empirical
+  based methods of distribution visualizing are preferred as they use
+  less space by design than most density based methods.
+\end{itemize}
+
+\hypertarget{effect-of-synergy-of-time-granularities}{%
+\subsection{Effect of synergy of time
+granularities}\label{effect-of-synergy-of-time-granularities}}
+
+In Section \ref{sec:synergy}, we discussed how pairs of granularities
+can have empty combinations either due to structure of calendar, event
+location or duration or due to the way they are built. In this section,
+we will see how these empty combinations affect the visualization when a
+dependent variable is plotted against these granularities.
+
+For illustration, distribution of half-hourly electricity consumption of
+Victoria is plotted across different time granularities in each of the
+panel in \autoref{fig:allFig}. \autoref{fig:allFig} (a) shows the letter
+value plot across days of the month faceted by months like January,
+March and December. \autoref{fig:allFig} (c) shows box plot across days
+of the year by the 1st, 15th, 29th and 31st days of the month.
+\autoref{fig:allFig} (d) showing violin plot across days of the month
+faceted by week of the month. \autoref{fig:allFig} (e), variations
+across week of the year conditional on week of the month can be observed
+through a ridge plot and \autoref{fig:allFig} (f) shows decile plots
+across day of the year and month of the year.
+
+Clearly, in \autoref{fig:allFig}, we observe that some choices of time
+granularities work and others do not. In \autoref{fig:allFig} (c), there
+will be no observations for some combinations ``Day-of-Month'' and
+``Day-of-Year''. In particular, the 1st day of the month can never
+correspond to 2nd, 3rd or 4th day of the year. On the contrary, for
+\autoref{fig:allFig} (a), we will not have any combinations with zero
+combinations because every ``Day-of-Week'' can occur in any
+``Month-of-Year''. Thus the graphs that don't work are those where many
+of the combination sets are empty. In other words, if there are levels
+of time granularities plotted across x-axis which are not spanned by
+levels of the time granularities plotted across facets or vice versa, we
+will have empty sets leading to potential ineffective graphs. We
+hypothesize that the synergy of these granularities are in play while
+deciding if the resulting plot would be a good candidate for exploratory
+analysis.
+
+We redefine harmony and clashes as follows: harmonies are pairs of
+granularities that aid exploratory data analysis, and clashes are pairs
+that are incompatible with each other for exploratory analysis. As a
+result, we should avoid plotting clashes as they hinder the exploratory
+process by having missing combinations of time granularities in each
+panel.
+
+\begin{figure}
+
+\includegraphics[width=0.5\linewidth]{figure/allFig-1} \includegraphics[width=0.5\linewidth]{figure/allFig-2} \includegraphics[width=0.5\linewidth]{figure/allFig-3} \includegraphics[width=0.5\linewidth]{figure/allFig-4} \includegraphics[width=0.5\linewidth]{figure/allFig-5} \includegraphics[width=0.5\linewidth]{figure/allFig-6} \hfill{}
+
+\caption{Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
 (a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
-Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
-
-# change all of these to gravitas function
-library(gravitas)
-library(tsibbledata)
-library(dplyr)
-library(ggplot2)
-
-par(mfrow = c(3, 2))
-
-VIC <- vic_elec
-
-VIC %>% 
-  create_gran("month_year") %>% 
-  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>% 
-  prob_plot("month_year","day_week",
-            response = "Demand",
-            plot_type = "lv") + 
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(a) Letter value plot")
-
-VIC %>% 
-  create_gran("day_week") %>% 
-  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>% 
-  prob_plot("day_week","hour_day",
-            response = "Demand",
-            plot_type = "quantile",
-            symmetric = FALSE) + 
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(b) Decile plot")
-
-VIC %>%
-  create_gran("day_month") %>% 
-  filter(day_month %in% c(1, 15, 29, 31)) %>%
-  prob_plot("day_month",
-                "day_year",
-                response = "Demand",
-                plot_type = "boxplot")+
-  ylab("Electricity Demand [KWh]") +
-  scale_x_discrete(breaks = seq(0, 366, 60)) + 
-  theme(legend.position = "bottom", 
-        strip.text = element_text(size = 7, margin = margin())) +
-  ggtitle("(c) Box plot")
-
-VIC %>%
-    create_gran("week_month") %>% 
-  filter(week_month %in% c(1, 2, 4)) %>%
-   prob_plot("week_month",
-                "day_month",
-                response = "Demand",
-                plot_type = "violin", alpha = 0.3) + ylab("") + 
-  xlab("Days of the Month") + 
-  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + 
-  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
-  ggtitle("(d) Violin plot")
-
-VIC %>%  create_gran("week_month") %>%
-  create_gran("week_year") %>%
-  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>% 
-  prob_plot("week_year", 
-            "week_month",
-            response = "Demand",
-            plot_type = "ridge")  + 
-  xlab("Electricity Demand [KWh]") +
-  ylab("Weeks of the Year") +
-  scale_x_continuous(breaks = seq(2000, 10000, 3000)) + 
-  theme(legend.position = "bottom", 
-        strip.text = element_text(size = 7, margin = margin())) +
-  ggtitle("(e) Ridge plot")
-
-VIC %>%
-  create_gran("month_year") %>% 
-  filter(month_year %in% c("Jan", "Jul", "Nov")) %>% 
-   prob_plot("month_year", 
-            "day_year",
-            response = "Demand",
-            plot_type = "quantile",
-            quantile_prob = seq(0.1, 0.9, 0.1),
-            symmetric = FALSE) +
-   #scale_x_discrete(breaks = seq(1, 336, 60)) + 
-   ylab("") + 
-   xlab("Day of the Year") + theme(legend.position = "none", 
-                                   strip.text = element_text(size = 7, 
-                                                             margin = margin())) +
-  ggtitle("(f) Decile plot")
-
-# VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
-```
-
-## Effect of number of observations
-
-Even with harmonies, visualizing probability distributions can be misleading either due to rarely occuring categories or unevenly distributed events.
-
-### Rarely occuring events
-
-Suppose we have $T$ observations, and two cyclic granularities -  $C_1$ with $n$ categories and $C_2$ with $m$ categories.
-Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
-There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
-The value of $k$ will depend on what type of visualization we are producing.
-For a decile plot, even $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
-Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
-
-<!-- The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations. -->
-
-### Unevenly distributed events
-
-Even when there are no rarely occuring events, number of observations might vary hugely within or across each panel.
-This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the visualizations that use kernel density estimates should be used with caution as sample size would directly affect both the variance of the estimator and the confidence intervals.
-Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels.
-
-# Applications {#sec:application}
-
-## Smart meter data of Australia {#sec:smartmeter}
-
-Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns and identify the extreme and regular households amongst these 50 households.
-
-<!-- typical and extreme behaviors of 50 households from that trial.  -->
-
-<!-- , and indeed many parts of the world.
-Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
--->
-
-<!-- <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions --> 
-<!-- It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
--->
-
-<!-- We want to  look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities. -->
-
-<!-- In Figure \ref{fig:day-fortnight} and \ref{fig:day-fortnight-agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
--->
-
-In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
-it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
-However, there is additional information that we can derive looking at the distribution.
-If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
-However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
-This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
-
-<!-- # ```{r day-fortnight_agg, echo = FALSE} -->
-<!-- # library(lvplot) -->
-<!-- # library(ggplot2) -->
-<!-- # library(dplyr) -->
-<!-- # library(tibble) -->
-<!-- # library(gravitas) -->
-<!-- # library(tidyr) -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!-- #   create_gran("day_fortnight") %>% -->
-<!-- #   as_tibble() %>%  -->
-<!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarise(median= log(median(general_supply_kwh)), -->
-<!-- #             total = log(sum(general_supply_kwh))) %>%  -->
-<!-- #   gather("daily_consumption", "value", - c(customer_id, day_fortnight)) %>%  -->
-<!-- #   ggplot2::ggplot(aes( -->
-<!-- #     x = as.factor(day_fortnight), -->
-<!-- #     y = value,  -->
-<!-- #     color = daily_consumption, -->
-<!-- #     group = daily_consumption)) + -->
-<!-- #   geom_line() + -->
-<!-- #   xlab("day_fortnight") + -->
-<!-- #   ylab("general_supply_kwh") +  -->
-<!-- #   facet_wrap(~customer_id) +  -->
-<!-- #   theme_bw()  -->
-
-<!-- # smart_meter10 %>%  -->
-<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!-- #   create_gran("day_fortnight") %>% -->
-<!-- #   as_tibble() %>%  -->
-<!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarise(median_consumption= median(general_supply_kwh)) %>% -->
-<!-- #   ggplot2::ggplot(aes( -->
-<!-- #     x = as.factor(day_fortnight), -->
-<!-- #     y = median_consumption, -->
-<!-- #     color =  customer_id, -->
-<!-- #     group = customer_id)) + -->
-<!-- #   geom_line() + -->
-<!-- #   xlab("day_fortnight") + -->
-<!-- #   ylab("general_supply_kwh") +  -->
-<!-- #   theme_bw()  -->
-<!-- ``` -->
-
-```{r day-fortnight, fig.cap="Letter value plot of two households across days of the fortnight.
+Only plots (a) and (b) show harmonised time variables.}\label{fig:allFig}
+\end{figure}
+
+\hypertarget{effect-of-number-of-observations}{%
+\subsection{Effect of number of
+observations}\label{effect-of-number-of-observations}}
+
+Even with harmonies, visualizing probability distributions can be
+misleading either due to rarely occuring categories or unevenly
+distributed events.
+
+\hypertarget{rarely-occuring-events}{%
+\subsubsection{Rarely occuring events}\label{rarely-occuring-events}}
+
+Suppose we have \(T\) observations, and two cyclic granularities -
+\(C_1\) with \(n\) categories and \(C_2\) with \(m\) categories. Each
+element of \(C_1\) occurs approximately \(T/n\) times while each element
+of \(C_2\) occurs approximately \(T/m\) times. There are no structurally
+empty combinations, and each combination will occur on average an equal
+number of times as \(T\rightarrow\infty\), so the average number of
+observations per combination is \(T/(mn)\). If we require at least \(k\)
+observations to create a meaningful panel, then provided \(T \ge mnk\),
+the visualization will be acceptable. The value of \(k\) will depend on
+what type of visualization we are producing. For a decile plot, even
+\(k=10\) may be acceptable, but for density estimates, we would need
+\(k \ge 30\). Rarely occurring categories such as the 366th day of the
+year, or the 31st day of the month can suffer from such problem.
+
+\hypertarget{unevenly-distributed-events}{%
+\subsubsection{Unevenly distributed
+events}\label{unevenly-distributed-events}}
+
+Even when there are no rarely occuring events, number of observations
+might vary hugely within or across each panel. This might happen due to
+missing observations in the data or uneven locations of events in time
+domain. In such cases, the visualizations that use kernel density
+estimates should be used with caution as sample size would directly
+affect both the variance of the estimator and the confidence intervals.
+Measures such as Gini's coefficient may be used to see if sample sizes
+vary widely across panels.
+
+\hypertarget{sec:application}{%
+\section{Applications}\label{sec:application}}
+
+\hypertarget{sec:smartmeter}{%
+\subsection{Smart meter data of Australia}\label{sec:smartmeter}}
+
+Smart meters provide large quantities of measurements on energy usage
+for households across Australia. One of the customer trial
+\citep{smart-meter} conducted as part of the Smart Grid Smart City
+(SGSC) project (2010-2014) in Newcastle, New South Wales and some parts
+of Sydney provides customer wise data on half-hourly energy usage and
+detailed information on appliance use, climate, retail and distributor
+product offers, and other related factors. It would be interesting to
+explore the energy consumption distribution for these customers and gain
+more insights on their energy behavior which are otherwise lost either
+due to aggregation or looking only at coarser temporal units. The idea
+here is to show how looking at the time across different granularities
+together can help identify different behavioral patterns and identify
+the extreme and regular households amongst these 50 households.
+
+In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for
+two customers to illustrate what kinds of insights can be drawn for the
+energy behavior of these two customers. it can be seen that for most
+days of the fortnight, the second household has much less consumption
+than the first one. However, there is additional information that we can
+derive looking at the distribution. If we consider letter value F as a
+regular behavior and letter values beyond F as not-so-regular behavior,
+we can conclude that the regular behavior of the first household is more
+stable than the second household. However, the distribution of tail of
+the first household is more variable, observed through distinct letter
+values, implying that their not-so-regular behavior is quite extreme.
+This shows, how looking at the distribution of the dependent variable
+can throw more light on the energy behavior of the customers, which are
+lost using aggregate or summary statistics.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/day-fortnight-1}
+
+}
+
+\caption{Letter value plot of two households across days of the fortnight.
 M, F, E, D and C represents the letter values of energy consumption.
-Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household."}
-library(lvplot)
-library(ggplot2)
-library(dplyr)
-library(tibble)
-library(gravitas)
-smart_meter10 %>%
-  filter(customer_id %in% c(10006704, 10017936)) %>% 
-  create_gran("day_fortnight") %>%
-  ggplot2::ggplot(aes(
-    x = as.factor(day_fortnight),
-    y = general_supply_kwh)) +
-  xlab("day_fortnight") +
-  geom_lv(
-    outlier.colour = "red",
-    aes(fill = ..LV..),
-    k = 5) +
-  facet_wrap(~customer_id) + 
-  scale_fill_lv() +
-  theme_bw() +
-  scale_y_sqrt() 
-```
-
-<!-- Figure \autoref{fig:my-hd} shows the overlay quantile plot of energy consumption of 50 households.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-Each facet represents a month and energy consumption across each hours of the day is shown inside each facet.
-It can be observed that the median is very close (and hence not distinctly visible) to the lower boundaries of all the other bands implying energy consumption for these households are extremely left skewed.75% of the households has distinct higher evening peaks compared to morning peaks, especially in the winter months.
-The top 5% households behave very differently than even the top 10% households in terms of total energy consumption across hours of the day.
--->
-
-**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
-
-```{r}
-library(tsibble)
-interval(smart_meter10)
-smart_meter10 %>% search_gran()
-```
-
-The interval of this tsibble is 30 minutes, and hence the temporal granularities range from half-hour to year.
-If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
-
-```{r}
-smart_meter10 %>% 
-  search_gran(highest_unit = "month")
-```
-
-Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
-
-```{r search_gran_limit2}
-
-smart_meter10 %>% search_gran(highest_unit = "month",
-                              filter_out = c("hhour", "fortnight")
-)
-```
-
-Now that we have the list of cyclic granularities to look at, we can visualize the distribution.
-From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-The distribution of energy needs to be visualized across two cyclic granularities at a time.
-It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.However, harmony/clash pairs can identified among those 30 pairs, to determine feasibility of plotting any pairs together.
-We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
-
-<!-- # ```{r is_harmony, echo= TRUE} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   is_harmony(gran1 = "hour_day",  -->
-<!-- #              gran2 = "day_week") -->
-<!-- #  -->
-<!-- # smart_meter10 %>% -->
-<!-- #   is_harmony(gran1 = "hour_day",  -->
-<!-- #              gran2 = "day_week",  -->
-<!-- #              facet_h = 14) -->
-<!-- #  -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   is_harmony(gran1 = "day_month", -->
-<!-- #              gran2 = "week_month") -->
-<!-- # ``` -->
-
-```{r harmony, echo=TRUE}
-smart_meter10 %>% harmony(
-  ugran = "month",
-  filter_out = c("hhour", "fortnight")
-)
-smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
-
-```
-
-In \autoref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.
-Boxplot of energy consumption is shown across wknd_wday (facet) and hour-day (x-axis) for the same two households.
-For the second household, outliers are less prominent implying their regular behavior is more stable.
-For the first household, energy behavior is not significantly different between weekdays and weekends.
-For the second household, median energy consumption for the early morning hours is extremely high for weekends compared to weekdays.
-
- For the second household, outliers are less prominent implying their regular behavior is more stable.
-For the first household, energy behavior is not significantly different between weekdays and weekends.
-For the second household, median energy consumption for the early morning hours is extremely high for weekends compared to weekdays.
-
-```{r bothcust_notrun}
-cust1 <- smart_meter10 %>% 
-  filter(customer_id %in% c(10006704)) %>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "boxplot") +
-  scale_y_sqrt() +
-  ggtitle("Energy consumption distribution for customer id: 10006704")
-
-cust2 <- smart_meter10 %>% 
-  filter(customer_id %in% c(10017936))%>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "boxplot") +
-  scale_y_sqrt() + 
-  ggtitle("Energy consumption distribution for customer id: 10017936")
-```
-
-```{r bothcust, fig.cap = "Boxplot of energy consumption across hours of the day faceted by weekday/weekend for household 1 (customer id: 10006704) and household 2 (customer id: 10017936).
+Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household.}\label{fig:day-fortnight}
+\end{figure}
+
+\textbf{gravitas} R package \citep{R-gravitas} is used to facilitate the
+systematic exploration here. While trying to explore the energy behavior
+of these customers systematically across cyclic time granularities, the
+first thing we should have at our disposal is to know which all cyclic
+time granularities we can look at exhaustively. If we consider
+conventional time deconstructions for a Gregorian calendar (second,
+minute, half-hour, hour, day, week, fortnight, month, quarter, semester,
+year), the following time granularities can be considered for this
+analysis.
+
+\begin{verbatim}
+#> 30m
+\end{verbatim}
+
+\begin{verbatim}
+#>  [1] "hhour_hour"         "hhour_day"          "hhour_week"        
+#>  [4] "hhour_fortnight"    "hhour_month"        "hhour_quarter"     
+#>  [7] "hhour_semester"     "hhour_year"         "hour_day"          
+#> [10] "hour_week"          "hour_fortnight"     "hour_month"        
+#> [13] "hour_quarter"       "hour_semester"      "hour_year"         
+#> [16] "day_week"           "day_fortnight"      "day_month"         
+#> [19] "day_quarter"        "day_semester"       "day_year"          
+#> [22] "week_fortnight"     "week_month"         "week_quarter"      
+#> [25] "week_semester"      "week_year"          "fortnight_month"   
+#> [28] "fortnight_quarter"  "fortnight_semester" "fortnight_year"    
+#> [31] "month_quarter"      "month_semester"     "month_year"        
+#> [34] "quarter_semester"   "quarter_year"       "semester_year"
+\end{verbatim}
+
+The interval of this tsibble is 30 minutes, and hence the temporal
+granularities range from half-hour to year. If these options are
+considered too many, the most coarse temporal unit can be set to be a
+``month''.
+
+\begin{verbatim}
+#>  [1] "hhour_hour"      "hhour_day"       "hhour_week"     
+#>  [4] "hhour_fortnight" "hhour_month"     "hour_day"       
+#>  [7] "hour_week"       "hour_fortnight"  "hour_month"     
+#> [10] "day_week"        "day_fortnight"   "day_month"      
+#> [13] "week_fortnight"  "week_month"      "fortnight_month"
+\end{verbatim}
+
+Also, some intermediate temporal units that might not be pertinent to
+the analysis can be removed from the list of cyclic granularities we
+want to look at.
+
+\begin{verbatim}
+#> [1] "hour_day"   "hour_week"  "hour_month" "day_week"   "day_month" 
+#> [6] "week_month"
+\end{verbatim}
+
+Now that we have the list of cyclic granularities to look at, we can
+visualize the distribution. From the search list, we found six cyclic
+granularities for which we would like to derive insights of energy
+behavior. The distribution of energy needs to be visualized across two
+cyclic granularities at a time. It is equivalent to taking 2
+granularities from 6, which essentially is equivalent to visualizing 30
+plots.However, harmony/clash pairs can identified among those 30 pairs,
+to determine feasibility of plotting any pairs together. We are left
+with 13 harmonies pair, each of which can be plotted together to look at
+the energy behavior from different perspectives.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }\KeywordTok{harmony}\NormalTok{(}
+  \DataTypeTok{ugran =} \StringTok{"month"}\NormalTok{,}
+  \DataTypeTok{filter_out =} \KeywordTok{c}\NormalTok{(}\StringTok{"hhour"}\NormalTok{, }\StringTok{"fortnight"}\NormalTok{)}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#> # A tibble: 13 x 4
+#>    facet_variable x_variable facet_levels x_levels
+#>    <chr>          <chr>             <int>    <int>
+#>  1 day_week       hour_day              7       24
+#>  2 day_month      hour_day             31       24
+#>  3 week_month     hour_day              5       24
+#>  4 day_month      hour_week            31      168
+#>  5 week_month     hour_week             5      168
+#>  6 day_week       hour_month            7      744
+#>  7 hour_day       day_week             24        7
+#>  8 day_month      day_week             31        7
+#>  9 week_month     day_week              5        7
+#> 10 hour_day       day_month            24       31
+#> 11 day_week       day_month             7       31
+#> 12 hour_day       week_month           24        5
+#> 13 day_week       week_month            7        5
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }\KeywordTok{gran_advice}\NormalTok{(}\StringTok{"wknd_wday"}\NormalTok{, }\StringTok{"hour_day"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#> The chosen granularities are harmonies 
+#>  
+#> Recommended plots are: violin lv quantile boxplot 
+#>  
+#> Number of observations are homogenous across facets 
+#>  
+#> Number of observations are homogenous within facets 
+#>  
+#> Cross tabulation of granularities : 
+#>  
+#> # A tibble: 24 x 3
+#>    hour_day Weekday Weekend
+#>    <fct>      <dbl>   <dbl>
+#>  1 0           7705    3097
+#>  2 1           7698    3100
+#>  3 2           7698    3101
+#>  4 3           7698    3102
+#>  5 4           7699    3099
+#>  6 5           7701    3098
+#>  7 6           7700    3099
+#>  8 7           7700    3098
+#>  9 8           7695    3098
+#> 10 9           7696    3098
+#> # ... with 14 more rows
+\end{verbatim}
+
+In \autoref{fig:bothcust} we visualize the harmony pair (wknd\_wday,
+hour\_day) through a box plot. Boxplot of energy consumption is shown
+across wknd\_wday (facet) and hour-day (x-axis) for the same two
+households. For the second household, outliers are less prominent
+implying their regular behavior is more stable. For the first household,
+energy behavior is not significantly different between weekdays and
+weekends. For the second household, median energy consumption for the
+early morning hours is extremely high for weekends compared to weekdays.
+
+For the second household, outliers are less prominent implying their
+regular behavior is more stable. For the first household, energy
+behavior is not significantly different between weekdays and weekends.
+For the second household, median energy consumption for the early
+morning hours is extremely high for weekends compared to weekdays.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/bothcust-1} \includegraphics[width=\textwidth]{figure/bothcust-2}
+
+}
+
+\caption{Boxplot of energy consumption across hours of the day faceted by weekday/weekend for household 1 (customer id: 10006704) and household 2 (customer id: 10017936).
 Median behavior for the early morning hours of household 2 is extremely high for weekends compared to weekdays, not much differenc can be observed for household 1.
-"}
-cust1
-cust2
-```
-
-Now, we would like to see how these customers' behavior relate to the rest of the  50 households across these two measures -
-
-- if energy distribution is skewed towards extreme?
-
-- if the weekend and weekday behavior are different for most of these households?
-
-Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of 50 households.
-Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends.
-
-Figure \ref{fig:my-hd} shows that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-Moreover, only the pink band changes significantly in winter months (May - August).
-The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.
-Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
-
-```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
-The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."}
-
-smart_meter <- 
-  read_rds("data/sm_cust50.rds")
-
-smart_meter %>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "quantile",
-            symmetric = FALSE, 
-            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
-  scale_y_sqrt()
-```
-
-<!-- # ```{r wknd-wday50_df} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   prob_plot("month_year", -->
-<!-- #                             "day_fortnight", -->
-<!-- #                             response = "general_supply_kwh", -->
-<!-- #                             plot_type = "quantile", -->
-<!-- #             symmetric = TRUE,  -->
-<!-- #             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
-<!-- #   scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3)) -->
-<!-- # ``` -->
-
-```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
+}\label{fig:bothcust}
+\end{figure}
+
+Now, we would like to see how these customers' behavior relate to the
+rest of the 50 households across these two measures -
+
+\begin{itemize}
+\item
+  if energy distribution is skewed towards extreme?
+\item
+  if the weekend and weekday behavior are different for most of these
+  households?
+\end{itemize}
+
+Figure \ref{fig:wknd-wday50} shows the quantile plot of energy
+consumption of 50 households. Similar quantiles for weekend and weekdays
+indicate that behaviors of most of these customers do not alter between
+weekdays and weekends.
+
+Figure \ref{fig:my-hd} shows that the median is very close to the lower
+boundaries of all the other bands implying energy consumption for these
+households are left skewed. Moreover, only the pink band changes
+significantly in winter months (May - August). The median consumption or
+width of other bands doesn't vary much across seasons, implying extreme
+behavior or extreme customers does not vary across seasons much. Most of
+the behavioral changes occur in the quartiles, that too in peak hours of
+the day. It is to be noted here that the the level of quantiles
+increased too in winter months (obvious because of weather conditions),
+the interesting part is to notice that the relationship between bands
+other than quartile stayed same across seasons.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/wknd-wday50-1}
+
+}
+
+\caption{Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
+The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends.}\label{fig:wknd-wday50}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/my-hd-1}
+
+}
+
+\caption{Area quantile plots of energy consumption across hours of the day faceted by months of the year.
 The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
 The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
-"}
-
-library(gravitas)
-library(tsibble)
-
-library(ggplot2)
-smart_meter %>%
-  prob_plot("month_year", "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99),
-    symmetric = TRUE) +
-  ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3))
-```
-
-<!-- Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
-This is true for both top 50% and 75% of the customers.
-Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
-All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
--->
-
-<!-- ```{r dw-hd} -->
-
-<!-- # Hourly usage across days of the week -->
-<!-- sm_cust50 %>% -->
-<!--   prob_plot("day_week",  -->
-<!--            "hour_day", -->
-<!--            "general_supply_kwh",  -->
-<!--            plot_type = "quantile",  -->
-<!--           quantile_prob = c(0.25, 0.5, 0.75), -->
-<!--            response = "general_supply_kwh", overlay = FALSE) + -->
-<!--   ggtitle("") -->
-
-<!-- ``` -->
-
-This case study shows systematic exploration of energy behavior starting with two households and comparing some of their features with all 50 households.
-First, it helps us to find the list of granularities to look at, then shrinks the number of possible visualizations by identifying harmonies and then explored harmony pairs to gain some insights on periodic behavior of the households.
-
-## T20 cricket data of Indian Premiere League {#sec:cricket}
-
-The application is not only restricted to temporal data.
-We provide an example of cricket to illustrate how this can be generalized in other applications.
-The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
-With eight teams, each team plays each other twice in a home-and-away round-robin format in the league phase.
-In a Twenty20 game the two teams have a single innings each, which is restricted to a maximum of 20 overs.
-Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls with some exceptions.
-
-The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data cross overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016).
-
-<!-- ```{r hierarchy, echo=FALSE} -->
-<!-- hierarchy_model <- tibble::tibble(linear gran = c("ball", "over", "inning", "match"),  -->
-<!--                                   convert_fct = c(6, 20, 2, 1)) -->
-<!-- knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket") -->
-<!-- ``` -->
-
-Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, where each over, which represents an ordering from past to future, can form the index of the tsibble.
-The hierarchy table would look like the following:
-
-```{r hierarchy2}
-library(gravitas)
-library(tibble)
-hierarchy_model <- tibble::tibble(
-  G = c("over", "inning", "match"),
-  k = c(20, 2, 1))
-
-knitr::kable(hierarchy_model, format = "markdown")
-
-```
-
-There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
-
-<!-- Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
--->
-
-<!-- Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that. -->
-
-<!-- a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions? -->
-
-<!-- How the scores per over vary across granularities/categorizations like innings of a match or matches of a season?
-Are these different for the winning teams in that season and the teams that couldn't qualify for the playoffs?
--->
-
-<!-- Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings? -->
-
-<!-- We will look at the ball by ball data for all batting teams.
-Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
-Also, we look at runs per over as that would have more variability compared to runs per ball and it be easier to observe the strategies of the winning team through that.
--->
-
-First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas-over-inning}.
-The distribution of runs per over has not significantly changed from 2008 to 2016.
-There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
-Hence, we work with subsets of seasons to answer some of the questions:
+}\label{fig:my-hd}
+\end{figure}
+
+This case study shows systematic exploration of energy behavior starting
+with two households and comparing some of their features with all 50
+households. First, it helps us to find the list of granularities to look
+at, then shrinks the number of possible visualizations by identifying
+harmonies and then explored harmony pairs to gain some insights on
+periodic behavior of the households.
+
+\hypertarget{sec:cricket}{%
+\subsection{T20 cricket data of Indian Premiere
+League}\label{sec:cricket}}
+
+The application is not only restricted to temporal data. We provide an
+example of cricket to illustrate how this can be generalized in other
+applications. The Indian Premier League (IPL) is a professional Twenty20
+cricket league in India contested by eight teams representing eight
+different cities in India. With eight teams, each team plays each other
+twice in a home-and-away round-robin format in the league phase. In a
+Twenty20 game the two teams have a single innings each, which is
+restricted to a maximum of 20 overs. Hence, in this format of cricket, a
+match will consist of 2 innings, an innings will consist of 20 overs, an
+over will consist of 6 balls with some exceptions.
+
+The ball by ball data for IPL season 2008 to 2016 is fetched from
+\href{https://www.kaggle.com/josephgpinto/ipl-data-analysis/data}{Kaggle}.
+The \texttt{cricket} data set in the \texttt{gravitas} package
+summarizes the ball-by-ball data cross overs and contains information
+for a sample of 214 matches spanning 9 seasons (2008 to 2016).
+
+Although there is no conventional time granularity in cricket, we can
+still represent the data set \texttt{cricket} through a
+\texttt{tsibble}, where each over, which represents an ordering from
+past to future, can form the index of the tsibble. The hierarchy table
+would look like the following:
+
+\begin{longtable}[]{@{}lr@{}}
+\toprule
+G & k\tabularnewline
+\midrule
+\endhead
+over & 20\tabularnewline
+inning & 2\tabularnewline
+match & 1\tabularnewline
+\bottomrule
+\end{longtable}
+
+There are many interesting questions that can possibly be answered with
+such a data set, however, we will explore a few and understand how the
+proposed approach in the paper can help answer some of the questions.
+
+First, we look at the distribution of runs per over across over of the
+innings and seasons in \autoref{fig:seas-over-inning}. The distribution
+of runs per over has not significantly changed from 2008 to 2016. There
+is no clear pattern/trend that runs per over is increasing or decreasing
+across seasons. Hence, we work with subsets of seasons to answer some of
+the questions:
 
@@
 
-Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final 4 from 2010 to 2015.
-It would be interesting to take their example in order to dive deeper into the first question.
-
-<!-- - Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first? -->
-
-<!-- - Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings? -->
-
-```{r runs_per-over}
-library(tsibble)
-
-cricket_all <-  read_csv("data-raw/deliveries_all.csv")
-matches_all <- read_csv("data-raw/matches_all.csv")
-
-cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
-
-cricket_per_over <- cricket_season %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))
-
-cricket_tsibble_all <- cricket_per_over %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-```
-
-```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."}
-  cricket_tsibble_all %>% 
-  prob_plot("over", 
-            "season",
-            hierarchy_model,
-            response = "runs_per_over",
-            plot_type = "quantile",
-            quantile_prob = c(0.25, 0.5, 0.75), 
-            symmetric = FALSE)  +
-     theme( # remove the vertical grid lines
-            panel.grid.major.x = element_blank() ,
-            # explicitly set the horizontal lines (or they will disappear too)
-            panel.grid.major.y = element_blank()) + scale_x_discrete(breaks = seq(2008, 2016, 3))
-```
-
-From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in runs in the second innings as compared to the first innings.
-The variability of runs also increases as the teams approach towards the end of the innings, as observed through the longer and more distinct letter values.
-
-```{r cricex, fig.cap= "Letter value plot of runs per over across overs of the inning faceted by innings of the match.
-No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning."}
-
-library(tsibble)
-cricket_tsibble <- cricket %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-cricket_tsibble %>%
-  filter(batting_team %in% c("Mumbai Indians",
-                             "Chennai Super Kings"))%>%
-  prob_plot("inning", 
-            "over",
-  response = "runs_per_over",
-  hierarchy_model,
-  plot_type = "lv")  + ggtitle("")
-```
-
-Q2: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalizes runs in the subsequent over more?
-
-For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
-If a batsman is bowled out, it does not necessarily signify good fielding.
-So we only include catches and run out as a measure of fielding.
-Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.
-Let us see if this fact is true.
-Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
-
-```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and dismissals by catch and catch and bowled.
-This shows good fielding in one over leads to lower runs in the subsequent overs for at least 50 percentiles of times."}
-
-cricket_fielding <- cricket_season %>% 
-  mutate(fielding_proxy = if_else(dismissal_kind %in% c("caught", "caught and bowled"),1,0)) %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs),
-            run_rate = sum(total_runs)/length(total_runs),
-            fielding_wckts = sum(fielding_proxy)) %>%
-            mutate(diff_run_rate = c(0,diff(run_rate)))
-
-cricket_tsibble_fld <- cricket_fielding %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-cricket_tsibble_fld %>% gran_obs("over", "fielding_wckts")
-
-cricket_tsibble_fld %>% 
-  filter(fielding_wckts %in% c(0,1)) %>% 
-  prob_plot( "fielding_wckts","over",
-                               hierarchy_model,
-                               response = "diff_run_rate",
-                               plot_type = "quantile",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-           overlay = TRUE)  + ggtitle("")
-
-```
-
-<!-- Q2: Among good fielding and bowling - which affect the runs of the subsequent overs more? -->
+Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of
+the best teams in IPL with multiple winning titles and always appearing
+in final 4 from 2010 to 2015. It would be interesting to take their
+example in order to dive deeper into the first question.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/seas-over-inning-1}
+
+}
+
+\caption{Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons.}\label{fig:seas-over-inning}
+\end{figure}
+
+From Figure \ref{fig:cricex}, it can be observed that there is no clear
+upward shift in runs in the second innings as compared to the first
+innings. The variability of runs also increases as the teams approach
+towards the end of the innings, as observed through the longer and more
+distinct letter values.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/cricex-1}
+
+}
+
+\caption{Letter value plot of runs per over across overs of the inning faceted by innings of the match.
+No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning.}\label{fig:cricex}
+\end{figure}
+
+Q2: Is run rate set to reduce in subsequent over if fielding/bowling is
+good in the previous over? Between fielding and bowling which penalizes
+runs in the subsequent over more?
+
+For establishing that the fielder fielded well in a particular over, we
+can see how many catches and run outs were made in that particular over.
+If a batsman is bowled out, it does not necessarily signify good
+fielding. So we only include catches and run out as a measure of
+fielding. Difference in runs across over should be negative if good
+fielding has an impact on the runs scored in the subsequent overs. Let
+us see if this fact is true. Figure \ref{fig:field} shows the difference
+between run rate between two subsequent overs are negative when good
+fielding leads to one or two dismissals in an over, implying good
+fielding in one over has indeed an impact on runs scored in the
+subsequent overs.
+
+\begin{verbatim}
+#> # A tibble: 4 x 21
+#>   fielding_wckts   `1`   `2`   `3`   `4`   `5`   `6`   `7`   `8`   `9`
+#>   <fct>          <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+#> 1 0               1039   997   972   978   962   957  1002  1012   993
+#> 2 1                120   145   171   165   180   181   138   124   140
+#> 3 2                  5    10     8     8     9    10     6    10     8
+#> 4 3                  0     0     0     0     0     0     0     0     0
+#> # ... with 11 more variables: `10` <dbl>, `11` <dbl>, `12` <dbl>,
+#> #   `13` <dbl>, `14` <dbl>, `15` <dbl>, `16` <dbl>, `17` <dbl>,
+#> #   `18` <dbl>, `19` <dbl>, `20` <dbl>
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/field-1}
+
+}
+
+\caption{Distribution of lagged run rate across overs of the innings and dismissals by catch and catch and bowled.
+This shows good fielding in one over leads to lower runs in the subsequent overs for at least 50 percentiles of times.}\label{fig:field}
+\end{figure}
 
@@
 
-A dot ball is a delivery bowled without any runs scored off it.
-The number of dot balls is reflective of the quality of bowling in the game.
-Run rate of an over should ideally decrease if the number of dot balls increase.
-However, what is the effect of dot balls on runs scored in the subsequent over.
-Will players  batsman likely to go for big shots because they couldn't score good runs in the previous over?
-Or they should play consistently and avoid scoring high?
-From figure \ref{fig:exdot} shows the quantile plot of runs across overs for none, one or two dot balls per over in the facets.
-Each of the quantiles decrease across facets.
-With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for 10th, 25th and 50th percentile quantiles.
-
-```{r crickread, ech0=FALSE}
-cricket_tsibble <- gravitas::cricket %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-```
-
-```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
-For all three cases, difference in runs between subsequent overs decrease.", fig.pos="htb"}
-
-library(gravitas)
-library(tsibble)
-
-cricket_bowling <- cricket_season %>% 
-  mutate(dot_ball_proxy = if_else(total_runs==0,1,0),
-         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(dot_balls = sum(dot_ball_proxy),
-            wickets = sum(wicket_proxy), 
-            total_runs = sum(total_runs)) %>%
-            mutate(diff_run_rate = c(0,diff(total_runs)))
-
-cricket_tsibble_bow <- cricket_bowling %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-cricket_tsibble_bow %>%
-  filter(dot_balls %in% c(0, 1, 2)) %>%
-  prob_plot("dot_balls",
-            "over",
-            hierarchy_model,
-            response = "diff_run_rate",
-            plot_type = "quantile",
-            quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
-            symmetric = FALSE)  + ggtitle("")
-
-```
-
-# Discussion {#sec:discussion}
-
-# Acknowledgements {-}
-
-# Bibliography
-
-<!-- # ```{r write-bib} -->
-<!-- # write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
-<!-- # ``` -->
-
-<!-- library(gravitas) -->
-<!-- library(tsibble) -->
-
-<!--  cricket_bowling <- cricket_season %>% -->
-<!--    mutate(dot_ball_proxy = if_else(total_runs==0,1,0), -->
-<!--           wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% -->
-<!--   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
-<!--    summarise(dot_balls = if_else(sum(dot_ball_proxy)>0, "yes","no"), -->
-<!--              wickets = if_else(sum(wicket_proxy)>0, "yes","no"),  -->
-<!--             total_runs = sum(total_runs)) %>% -->
-<!--             mutate(diff_run_rate = c(0,diff(total_runs))) -->
-<!--   -->
-<!--   -->
-<!--  cricket_tsibble_bow <- cricket_bowling %>%  -->
-<!--    ungroup() %>%  -->
-<!--    mutate(data_index = row_number()) %>%  -->
-<!--    as_tsibble(index = data_index) -->
-
-<!-- cricket_tsibble_bow %>% gran_obs("over", "dot_balls") -->
-
-<!-- # cricket_tsibble_bow %>%  -->
-<!-- #   #filter(dot_balls %in% c(0, 1, 2))%>%  -->
-<!-- #   prob_plot("dot_balls", -->
-<!-- #             "over", -->
-<!-- #             hierarchy_model, -->
-<!-- #             response = "diff_run_rate", -->
-<!-- #             plot_type = "quantile", -->
-<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!-- #             symmetric = FALSE) -->
-
-<!-- cricket_tsibble_bow %>% gran_obs("over", "wickets") -->
-
-<!-- cricket_tsibble_bow %>%  -->
-<!--   #filter(wickets %in% c(0, 1))%>%  -->
-<!--   gather("proxies", "yes_no", -c(season, -->
-<!--                                      match_id, -->
-<!--                                      batting_team, -->
-<!--                                      bowling_team, -->
-<!--                                      inning,     -->
-<!--                                      over, -->
-<!--                                      diff_run_rate, -->
-<!--                                      data_index)) %>%  -->
-<!--   filter(yes_no %in% c("no")) %>%  -->
-<!--   prob_plot("proxies", -->
-<!--             "over", -->
-<!--             hierarchy_model, -->
-<!--             response = "diff_run_rate", -->
-<!--             plot_type = "boxplot", -->
-<!--             symmetric = FALSE, -->
-<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
-<!-- ``` -->
-
-<!-- ```{r proxies-yes} -->
-<!-- cricket_tsibble_bow %>%  -->
-<!--   #filter(wickets %in% c(0, 1))%>%  -->
-<!--   gather("proxies", "yes_no", -c(season, -->
-<!--                                      match_id, -->
-<!--                                      batting_team, -->
-<!--                                      bowling_team, -->
-<!--                                      inning,     -->
-<!--                                      over, -->
-<!--                                      diff_run_rate, -->
-<!--                                      data_index)) %>%  -->
-<!--   filter(yes_no %in% c("yes")) %>%  -->
-<!--   prob_plot("proxies", -->
-<!--             "over", -->
-<!--             hierarchy_model, -->
-<!--             response = "diff_run_rate", -->
-<!--             plot_type = "lv", -->
-<!--             symmetric = FALSE, -->
-<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
-
-<!-- ``` -->
-
-<!-- ```{r exwicketplot, echo=FALSE} -->
-<!-- #  -->
-<!-- # cricket_tsibble %>%  -->
-<!-- #   filter(dot_balls %in% c(0, 1, 2)) %>% -->
-<!-- #   mutate(lag_dot = lag(dot_balls)) %>%  -->
-<!-- #   filter(lag_dot %in% c(0, 1, 2)) %>% -->
-<!-- #   prob_plot("lag_dot", -->
-<!-- #             "over", -->
-<!-- #             hierarchy_model, -->
-<!-- #             response = "runs_per_over", -->
-<!-- #             plot_type = "quantile", -->
-<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- cricket_tsibble %>%  -->
-<!--   filter(wicket %in% c(0, 1)) %>% -->
-<!--   mutate(lag_wicket = lag(wicket)) %>%  -->
-<!--   filter(lag_wicket %in% c(0, 1)) %>% -->
-<!--   prob_plot("lag_wicket", -->
-<!--             "over", -->
-<!--             hierarchy_model, -->
-<!--             response = "runs_per_over", -->
-<!--             plot_type = "quantile", -->
-<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- ``` -->
-
-<!-- ```{r violin, fig.cap= "Violin and box plots of run rate across innings of the match and matches won or lost for Mumbai Indians(top) and Chennai Super Kings(bottom).The violins for CSK has long right tails in 2nd innings for cases when they lost or won.
-MI has long right tail only when they won the match."} -->
-
-<!-- matches_select <- matches %>% select(id, toss_winner, winner, win_by_runs) -->
-
-<!-- cricket_winner <- cricket_tsibble_all %>% -->
-<!--   left_join(matches_select, by = c("match_id" = "id")) -->
-
-<!-- MI_winner <- cricket_winner %>% -->
-<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
-<!--                     season %in% c(2013, 2015)) %>% -->
-<!--              mutate(win = if_else(winner == "Mumbai Indians", 1, 0)) -->
-
-<!-- MI_winner %>% -->
-<!--   gran_obs("win", "inning", hierarchy_model) -->
-
-<!-- CSK_winner <- cricket_winner %>% -->
-<!--              filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings", -->
-<!--                     season %in% c(2013, 2015)) %>% -->
-<!--              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0)) -->
-
-<!-- CSK_winner %>% -->
-<!--   gran_obs("win", "inning", hierarchy_model) -->
-
-<!--  MI_winner %>% -->
-<!--             prob_plot( "win","inning", -->
-<!--                                 hierarchy_model, -->
-<!--                                 response = "run_rate", -->
-<!--                                 plot_type = "violin", -->
-<!--                                 quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!--                      overlay = FALSE) + ggtitle("") + -->
-<!--     geom_boxplot(width = 0.1, color = "blue") -->
-
-<!--  CSK_winner %>% filter(inning %in% c(1,2)) %>% -->
-<!--            prob_plot( "win","inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "violin", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!--                     overlay = FALSE) + ggtitle("") + -->
-<!--    geom_boxplot(width = 0.1, color = "blue") -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- cricket_winner %>% as_tibble() %>% -->
-<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
-<!--                     winner!= "Mumbai Indians", -->
-<!--                     win_by_runs!=0, -->
-<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
-
-<!-- cricket_winner %>% as_tibble() %>% -->
-<!--              filter(batting_team ==  "Chennai Super Kings", -->
-<!--                     winner!= "Chennai Super Kings", -->
-<!--                     win_by_runs!=0, -->
-<!--                     inning ==2, -->
-<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
-<!-- ``` -->
-
-<!-- For MI, cases when they lost, distribution of run rates not very different depending on if they bat first or second.
-However, cases when they won, distribution of run rate is very skewed when they bat in the 2nd innings implying winning by short margins.
--->
-
-<!-- Figure \ref{fig:violin} shows that for CSK, irrespective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
-Run rate can go up to 6 runs per over depending on the situation.
-For MI, this is not true.
-When they won, their run rates in the 2nd innings went up to 6 and led to their win, but stayed within 4 whenever they lost.
-This implies CSK is more of a fighter than MI when it comes to defending in the 2nd innings.
-However, we can delve deeper and try to see which overs are the most volatile for CSK and if it is different to that of MI?
--->
-
-<!-- In Figure \ref{fig:lv}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
-It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
-There is an upward shift of the distribution in the second part of the innings (after 10th over) and more variability (longer and distinct letter values) as well.
--->
-<!-- CSK has been more uniform in their approach maintaining consistent scores per over.
-Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}.
--->
-
-<!-- whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second.
-CSK's run rate distribution is extremely right tailed in the 2nd innings, but not in the first.
-.
---> -->
-
-<!-- ```{r 2013_15_MI} -->
-<!--  cricket_tsibble_all %>%  -->
-<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
-<!--            prob_plot("inning", "over", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
-<!--     ) -->
-<!--  cricket_tsibble_all %>%  -->
-<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
-<!--            prob_plot("inning", "over", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
-<!--     ) -->
-
-<!-- ``` -->
-
-<!-- r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."} -->
-<!-- #  -->
-<!-- #  -->
-<!-- #  -->
-<!-- # MI <- cricket_tsibble_all %>% -->
-<!-- #   filter(season %in% c(2010:2015), batting_team=="Mumbai Indians") -->
-<!-- #  -->
-<!-- #  -->
-<!-- # CSK <- cricket_tsibble_all %>% -->
-<!-- #   filter(season %in% c(2010:2015), -->
-<!-- #          batting_team=="Kolkata Knight Riders", -->
-<!-- #          inning %in% c(1,2)) -->
-<!-- #  -->
-<!-- #  -->
-<!-- #  MI %>% -->
-<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
-<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
-<!-- #  -->
-<!-- #  CSK %>% -->
-<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
-<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
-<!-- #  -->
-<!-- # MI %>% -->
-<!-- #   prob_plot("inning", "over", -->
-<!-- #                                hierarchy_model, -->
-<!-- #                                response = "run_rate", -->
-<!-- #                                plot_type = "lv", -->
-<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") -->
-<!-- #  -->
-<!-- #  -->
-<!-- # CSK %>% -->
-<!-- #   prob_plot("inning", "over", -->
-<!-- #                                hierarchy_model, -->
-<!-- #                                response = "run_rate", -->
-<!-- #                                plot_type = "lv", -->
-<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") -->
-<!-- #  -->
-<!-- # cricket_tsibble_all %>% distinct(batting_team) -->
-
-<!-- # ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."} -->
-<!-- #  -->
-<!-- # cricket_dot <- cricket_season %>%  -->
-<!-- #   mutate(wicket = if_else(is.na(dismissal_kind),0,1), -->
-<!-- #          dot_balls = if_else(total_runs==0,1, 0) -->
-<!-- #          ) %>%  -->
-<!-- #   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
-<!-- #   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs), -->
-<!-- #             wicket = sum(wicket), -->
-<!-- #             dot_balls = sum(dot_balls))  -->
-<!-- #  -->
-<!-- #  -->
-<!-- #  -->
-<!-- # cricket_tsibble_dot <- cricket_dot %>%  -->
-<!-- #   ungroup() %>%  -->
-<!-- #   mutate(data_index = row_number()) %>%  -->
-<!-- #   as_tsibble(index = data_index) -->
-<!-- #  -->
-<!-- # #cricket_tsibble_dot %>% gran_obs("over", "fielding_wckts") -->
-<!-- #  -->
-<!-- # cricket_tsibble_dot  %>%  -->
-<!-- #   filter(dot_balls %in% c(1,2,3,4)) %>%  -->
-<!-- #   mutate(lag_dot_ball = lag(dot_balls)) -->
-<!-- #   prob_plot("over","dot_balls", -->
-<!-- #                                hierarchy_model, -->
-<!-- #                                response = "run_rate", -->
-<!-- #                                plot_type = "quantile", -->
-<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!-- #            overlay = FALSE)  + ggtitle("") -->
-<!-- # ``` -->
-<!-- #  -->
-<!-- #  -->
-
-<!--   cricket_tsibble %>% prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="black" )  -->
-<!--     )
-
-<!--   cricket_tsibble %>% -->
-<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!--     cricket_tsibble %>% -->
-<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!-- ```{r runs_per-over} -->
-
-<!-- # all season per over runs -->
-<!-- cricket_per_over_all <- cricket_all_seas %>%  -->
-<!--   group_by(season, match_id, batting_team,inning, over) %>% -->
-<!--   summarise(runs_per_over = sum(total_runs))  -->
-
-<!-- # all season per over runs tsibble -->
-<!-- cricket_tsibble_all <- cricket_per_over_all %>%  -->
-<!--   ungroup() %>%  -->
-<!--   mutate(data_index = row_number()) %>%  -->
-<!--   as_tsibble(index = data_index) -->
-
-<!--   cricket_tsibble_all %>% prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   ) -->
-
-<!--   cricket_tsibble_all %>% prob_plot("inning_match", "season", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )   -->
-
-<!--   cricket_tsibble_all %>% prob_plot("season", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  +  -->
-
-<!--   cricket_tsibble %>% -->
-<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!--     cricket_tsibble %>% -->
-<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!-- {r read_data, echo=FALSE} -->
-
-<!-- cricket_tsibble <- cricketdata %>% -->
-<!--   mutate(data_index = row_number()) %>% -->
-<!--   as_tsibble(index = data_index) -->
-
-<!-- # cricket_tsibble %>% dynamic_create_gran( -->
-<!-- #   "ball_over", -->
-<!-- #   hierarchy_model -->
-<!-- # ) -->
-
-<!-- #over_inning    ball_over -->
-<!-- cricket_tsibble %>% granplot("over_inning", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!-- ) -->
-
-<!-- #inning_match   ball_over -->
-
-<!-- cricket_tsibble %>% granplot("inning_match", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!-- ) -->
-
-<!-- # inning_match   ball_inning -->
-<!-- #interesting plot -->
-<!-- cricket_tsibble %>% granplot("inning_match", "ball_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
-<!--   scale_x_discrete(breaks = seq(1,120,5)) -->
-
-<!-- # ball_over      over_inning -->
-
-<!-- cricket_tsibble %>% granplot("ball_over", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- # inning_match   over_inning -->
-
-<!-- cricket_tsibble %>% granplot("inning_match", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      over_match -->
-
-<!-- cricket_tsibble %>% granplot("ball_over", "over_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      inning_match -->
-
-<!-- cricket_tsibble %>% granplot("ball_over", "inning_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->
-
-<!-- cricket_tsibble %>% -->
-<!--   granplot( -->
-<!--     gran1 = "inning_match", -->
-<!--     gran2 = "over_inning", -->
-<!--     hierarchy_model, -->
-<!--     response = "total_runs", -->
-<!--     plot_type = "quantile" -->
-<!--   ) -->
-
-<!-- ```{r read_data_all_seas, echo=FALSE} -->
-
-<!-- load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda") -->
-
-<!-- cricket_tsibble_all <- cricket_all_seas %>% -->
-<!--   filter(season %in% c(2012:2015)) %>%  -->
-<!--   mutate(data_index = row_number()) %>% -->
-<!--   as_tsibble(index = data_index) -->
-
-<!-- # cricket_tsibble %>% dynamic_create_gran( -->
-<!-- #   "ball_over", -->
-<!-- #   hierarchy_model -->
-<!-- # ) -->
-
-<!-- #over_inning    ball_over  -->
-<!-- cricket_tsibble_all %>% granplot("over_inning", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!-- ) -->
-
-<!-- #inning_match   ball_over  -->
-
-<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!-- ) -->
-
-<!-- # inning_match   ball_inning  -->
-<!-- #interesting plot -->
-<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
-<!--   scale_x_discrete(breaks = seq(1,120,5)) -->
-
-<!-- # ball_over      over_inning   -->
-
-<!-- cricket_tsibble_all %>% granplot("ball_over", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- # inning_match   over_inning   -->
-
-<!-- cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      over_match   -->
-
-<!-- cricket_tsibble_all %>% granplot("ball_over", "over_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      inning_match   -->
-
-<!-- cricket_tsibble_all %>% granplot("ball_over", "inning_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->
-
-<!-- cricket_tsibble_all %>% -->
-<!--   granplot( -->
-<!--     gran1 = "inning_match", -->
-<!--     gran2 = "over_inning", -->
-<!--     hierarchy_model, -->
-<!--     response = "total_runs", -->
-<!--     plot_type = "quantile" -->
-<!--   ) -->
-
-```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr","rlang", "gravitas"), file = "rpkgs.bib")
-```
+A dot ball is a delivery bowled without any runs scored off it. The
+number of dot balls is reflective of the quality of bowling in the game.
+Run rate of an over should ideally decrease if the number of dot balls
+increase. However, what is the effect of dot balls on runs scored in the
+subsequent over. Will players batsman likely to go for big shots because
+they couldn't score good runs in the previous over? Or they should play
+consistently and avoid scoring high? From figure \ref{fig:exdot} shows
+the quantile plot of runs across overs for none, one or two dot balls
+per over in the facets. Each of the quantiles decrease across facets.
+With two dot balls per over, at least 50\% of the times, run rates
+decrease in the subsequent over as observed through the negative values
+for 10th, 25th and 50th percentile quantiles.
+
+\begin{figure}[htb]
+
+{\centering \includegraphics[width=\textwidth]{figure/exdot-1}
+
+}
+
+\caption{10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
+For all three cases, difference in runs between subsequent overs decrease.}\label{fig:exdot}
+\end{figure}
+
+\hypertarget{sec:discussion}{%
+\section{Discussion}\label{sec:discussion}}
+
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+\hypertarget{bibliography}{%
+\section{Bibliography}\label{bibliography}}
+
+--\textgreater{}
+
+\bibliographystyle{agsm}
+\bibliography{bibliography.bib}
+
+\end{document}

### 2019-12-02 main.Rmd
aperiodic completed

@@
 
-\begin{timedom}\label{def:quasicircular}
+\newtheorem{timedom1}{Definition}[subsection]
+\begin{timedom1}\label{def:quasicircular}
 A quasi-circular granularity QC_(BG, G') that relates linear granularities G' and bottom granularity BG, if
@@
 T_w are the sets of indices of BG describing G'(w) such that $G'(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-\end{timedom}
+\end{timedom1}
 
@@
 
-```{r quasi-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
+```{r quasi-gran, echo=FALSE,out.width = "100%",fig.cap="Relating quasi-circular granularities and bottom linear granularity "}
 knitr::include_graphics("Figs/aperiodic-granv1.png")
@@
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
-It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi})
+It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
 
 \begin{equation}\label{eq6}
-A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
+A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
 \end{equation}
@@
 Most public holidays repeat every year, but there is no period within which their behavior remains constant.
-For example, U.S labour day is the first Monday in September and U.S.
+A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
+U.S labour day is the first Monday in September and U.S.
 Memorial Day is the last day in May.
-A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
-Examples of aperiodic granularity may also include school holidays or a scheduling event that covers the first and third Monday of the month between June and October, except for state holidays.
-All of these are still recurring events, but recurring with non-periodic patterns.
-As such, plausible $k_i$ from \autoref{tab:aperiodic} can be infinite for aperiodic granularities.
+In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
+Examples of aperiodic granularity may also include school holidays or a scheduling event that might cover the first and third Monday of the month between June and October, except for state holidays.
+All of these are recurring events, but with non-periodic patterns.
+As such, plausible $k_i$ from \autoref{tab:quasi} can be infinite for aperiodic granularities.
 
@@
 \begin{split}
-A_{BG, G}(z) & = 1, z \in T_q\\
+A_{BG, G''}(z) & = 1, z \in T_q\\
      & = 0, z \not\in T_q\\
@@
 
+Discussion: Aperiodic time granularities map a time index to a categorical variable containing binary values.
+It takes the value 1 whenever the time index coincides with the events in context (e.g.
+public or school holiday) and 0 when it does not.
+Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector.
+An example of aperiodic granularity can be one that relates a bottom granularity ($BG$) and linear granularity ($G''$) such that $BG \trianglelefteq G''$, $BG \not\preceq G$ and $BG$ is not periodic or quasi-periodic with respect to $G''$, as shown in \autoref{fig:aperiodic}.
+
+```{r aperiodic, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
+knitr::include_graphics("Figs/aperiodic.png")
+```
+
 <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
@@
 
-### Computation of multiple-order from single-order {#sec:multiplefromsingle}
+## Multiple-order from single-order {#sec:multiplefromsingle}
 
@@
 
-### Computation of single-order from multiple-order {#sec:singlefrommultiple}
+## Single-order from multiple-order {#sec:singlefrommultiple}
 

### 2019-12-02 gravitas/data-raw/cricketdata.R

@@ -1,0 +1,59 @@
+library(readr)
+library(tsibble)
+library(dplyr)
+
+cricket <-  read_csv("data-raw/deliveries_all.csv")
+matches <- read_csv("data-raw/matches_all.csv")
+
+cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id")
+
+# each over needs to have 6 balls
+
+cricket_fltr_ball <- cricket_season %>%
+  filter(wide_runs + noball_runs== 0) %>%    # filtering out scores that are gained by no balls or wide
+  # still have to filter out overs for which number of balls per over less than 6
+  group_by(batting_team, match_id, inning, over) %>%
+  summarise(n_over = length(ball)) %>%
+  filter(n_over==6) %>%
+  mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
+
+
+cricket_over_crctd <- cricket_season %>%
+  mutate(cricket_key = paste(batting_team, match_id, inning, over, sep = "_")) %>%
+  filter(cricket_key %in% cricket_fltr_ball$key) %>%
+  filter(wide_runs + noball_runs== 0) %>%
+  mutate(ball_per_over = purrr::rep_along(match_id, 1:6))
+
+# each innings need to have 20 overs
+
+cricket_fltr_over <- cricket_over_crctd %>%
+  filter(inning %in% c(1,2)) %>%
+  group_by(inning, match_id) %>%
+  summarise(n = length(over)) %>% filter(n==120) %>%
+  mutate(key = paste(match_id, inning, sep = "_"))
+
+cricket_over_inning_crctd <- cricket_over_crctd %>%
+  mutate(over_key = paste(match_id, inning, sep = "_")) %>%
+  filter(over_key %in% cricket_fltr_over$key) %>%
+  filter(wide_runs + noball_runs== 0) %>%
+  mutate(ball_per_over = purrr::rep_along(match_id, 1:6)) %>%
+  filter(inning %in% c(1,2))
+
+
+# each match needs to have 2 innings
+
+cricket_filtr_inning <- cricket_over_inning_crctd %>%
+  group_by(match_id) %>% 
+  summarise(length(inning)) %>% 
+  filter(`length(inning)`==240)
+
+cricket_all_seas <- cricket_over_inning_crctd %>%
+  filter(match_id %in% cricket_filtr_inning$match_id) %>%
+  filter(wide_runs + noball_runs== 0) %>%
[…]

### 2019-12-02 main.Rmd
data structure

@@
 Effective exploration or good visualization require good data structures.
-Temporal data problems are typically grouped into two types of analysis, time series and longitudinal.
-A new tidy data structure to support exploration and modeling of temporal data is tsibble, where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
-This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.
 Commonly, simple sequences of time value pairs $<t_0, v_0> \dots, <t_n, v_n>$ are the basis of analysis and visualization.
 It is recognized that the initial approaches of just considering time as an ordinal dimension in a 2D or 3D visualisation are inadequate to capture characteristics of time-dependent information.
-For example, there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use different calendars that define meaningful system of granularities for different application domain.
-Data can then be integrated at different levels of granularity enabling value aggregation or distribution along granularities.
-Since, we are essentially concerned with detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by anabling aggregation or distribution across cyclic time granularities.
-We consider the following data structure for exploring distribution of a measured variable ((Figure \@ref(fig:data-structure)).
-These are based on conditioning the values taken on by one or more of the cyclic granularities in a data set.
-This means data subsets are formed for each level of the conditioned variable and the fixed variable.
+There is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
+Moreover, one can use calendars based on application domain that define meaningful system of granularities.
+Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities.
+Since, we are interested in detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.
+Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
+Instead, the big problem can be broken down into smaller pieces by focusing on two cyclic granularities at a time.
+
+A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
+Since all cyclic granularities can be expressed in terms of the bottom granularity, which is the index of the tsibble, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
+Data subsets are formed for each level of the conditioned variable and the fixed variable.
 This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
 
-```{r data-structure, fig.cap = "The data structure for exploring periodicities along cyclic granularities.
-Two cyclic granularities serve as fixed and conditioned variable and the observational unit across combinations of these cyclic time granularities.", fig.pos="htb"}
-include_graphics("Figs/data-structure.png")
+```{r data-structure, fig.cap = "The data structure for exploring periodicities in data.
+Two cyclic granularities serve as fixed and conditioned variable and the measurement variable is the observation across combinations of these cyclic time granularities.", fig.pos="htb"}
+include_graphics("Figs/data-structure-v1.png")
 ```

### 2019-12-02 main.Rmd
data structure

@@
 
-The cyclic time granularities which are essentially a function of the index of the tsibble are abstractions of time and hence are discrete variable, mapped as a factor.
-The relationships between the fixed and conditioned variable are of importance in this data structure due to the ways the cyclic granularities, which are mapped to fixed and conditional variable, relate to one another.
+The cyclic time granularities which are essentially categorisations of time are discrete variable, mapped as a factor since they have an inherent order.
+Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
 To define a general framework, we define harmony and clashes.

### 2019-12-02 main.Rmd
section headings changed

@@
 
-# Cyclic Time granularities {#sec:circular-gran-def}
+# Cyclic Time granularities {#sec:cyclic-gran-def}
 
@@
 
-## Quasi-circular {#sec:circular-gran-def}
+## Quasi-circular {#sec:quasi-circular-gran-def}
 
@@
 
-## Aperiodic {#sec:circular-gran-def}
+## Aperiodic {#sec:aperiodic-gran-def}
 
@@
 
-## Fixed and Conditioned variables {#sec:synergy}
+## Synergy of fixed and conditioned variables {#sec:synergy}
 
@@
 
-## Measured variable
-
-The measured variable which is the univariate observational units can be dicrete or continuous.
+## Density estimation of measured variable
+
+The measured variable can be dicrete or continuous.
+The probability distribution of these univariate measured variable could be studied by estimating the proability density function (pdf for continuous) or probability mass function (pmf for dicrete).
+Both pdf and pmf are concerned with point probabilities of the random variable.
+Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d.
+The densities can either be estimated or summarized using quantiles.Instead of assuming a parametric model for the distribution (e.g.
+a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g.
+differentiable).
+It is then possible to estimate the unknown density function f(·).
+
+The histogram is the oldest and most popular density estimator method.
+The form of the histogram depends very much on the tuning parameters class width and origin.
+The naive estimator was introduced by Fix and Hodges (1951) and first discussed by Rosenblatt(1956).
+Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962).
+The nearest neighbour estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977).
+There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.
+
+The quantile of a distribution is defined as $Q(p)= F^{-1}(p) = inf{{x: F(x) >p}}$, $0<p< 1$, where $F(x)$ is the distribution function.
+Let $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$, and let $Qi(p)$ denote the ith sample quantile definition, then
+
+\begin{equation}
+\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}
+\end{equation}
+
+where,  $(j - m)/n \leq p < (j - m + 1)/n$
+
+There have been several other nonparametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
+Also sample quantiles can be of the form that are not defined for all p including hinges and other letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
 
@@
 
-### Empirical methods
+### Quantile based methods
 
@@
 ```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
-For all three cases, difference in runs between subsequent overs decrease.", fig.pos="htb"}
+For all three cases, difference in runs between subsequent overs decrease.", fig.pos="h"}
 

### 2019-12-04 main.Rmd
changing data structure introduction

@@
 Effective exploration or good visualization require good data structures.
-Commonly, simple sequences of time value pairs $<t_0, v_0> \dots, <t_n, v_n>$ are the basis of analysis and visualization.
-It is recognized that the initial approaches of just considering time as an ordinal dimension in a 2D or 3D visualisation are inadequate to capture characteristics of time-dependent information.
+Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization.
+It is recognized that the initial approaches of just considering time as an ordinal dimension in visualisation are inadequate to capture characteristics of time-dependent information.
 There is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use calendars based on application domain that define meaningful system of granularities.
+Moreover, one can use calendars based on application domain that define contextual system of granularities.
 Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities.
@@
 Since all cyclic granularities can be expressed in terms of the bottom granularity, which is the index of the tsibble, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-Data subsets are formed for each level of the conditioned variable and the fixed variable.
-This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
+This is extending the columns of tsibble by including the cyclic granularities.Now suppose we want to explore the measurement variable $v$ across two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
+Also, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
+Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ such sets.
+These sets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities.
+
+<!-- This data representation can lead to comprehensive exploration by conditioning the levels taken by one or more of the cyclic granularities. -->
+
+<!-- This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data by creating same plots using data subsets corresponding to each level of the conditioned variable.
+-->
+
+<!-- These are based on conditioning the values taken on by one or more of the variables in a data set.
+This means creating same plots using data subsets corresponding to each level of the conditioned variable.
+This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
+-->
 
@@
 
-## Synergy of fixed and conditioned variables {#sec:synergy}
-
-The cyclic time granularities which are essentially categorisations of time are discrete variable, mapped as a factor since they have an inherent order.
+## Synergy of the cyclic granularities {#sec:synergy}
+
+The way cyclic granularities relate become important when we consider the data structure <$A_i$, $B_j(s)$, $v(s)$> for exploration.
+We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the levels of the cyclic granularities.
+To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
+
+<!-- The cyclic time granularities which are essentially categorisations of time are discrete variable, mapped as a factor since they have an inherent order.
 Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
 To define a general framework, we define harmony and clashes.
+-->
 
@@
 
-Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
+<!-- Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
 That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
 Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
-Now, many situations can lead to any of these sets being empty.
-Let us discuss the following cases, where one or more of these sets can be empty.
+-->
+
+<!-- The cyclic time granularities which are essentially categorisations of time are discrete variable, mapped as a factor since they have an inherent order.
+Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
+To define a general framework, we define harmony and clashes.
+-->
+
+<!-- Now, many situations can lead to some of these sets being empty. Let us discuss the following cases, where one or more of these sets can be empty. -->
 
@@
 There have been several other nonparametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
-Also sample quantiles can be of the form that are not defined for all p including hinges and other letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
+Also sample quantiles can be of the form that are not defined for all p including hinges, letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
 
@@
 ```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
-For all three cases, difference in runs between subsequent overs decrease.", fig.pos="h"}
+For all three cases, difference in runs between subsequent overs decrease.", fig.pos="ht"}
 

### 2019-12-04 main.Rmd
changing circular gran

@@
 \begin{timedom}\label{def:circular}
-A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if
+A circular granularity C(B, G) relates a linear granularity G to the bottom granularity B, if
 
 \begin{equation} \label{eq:eq2}
-C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
-BG(z) \neq \phi,
+C_{B, G}(z) = {L}(z\mod k(B,G)) \quad \forall z \in \mathbb{Z}^+
 \end{equation}
-where,  
-z denotes the index set, 
-BG denotes the index of the tsibble (bottom granularity), 
-G is a linear granularity periodic with respect to BG with regular mapping,     
-and k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.
+where,\\
+z denotes the index set, \\
+B denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity G with regular mapping, \\
+{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
+and k(B,G) is the number of granules of B in each granule of G.
 \end{timedom}
 
-**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
+**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorisation of the circular granularity.
+However, the labels can coincide with indexes in which case integers are directly used to refer to categorisations of the circular granularity.
+Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
 Each granule is represented by a box.
@@
 G and H are two linear granularities denoting days and weeks respectively.
-The period length of days with respect to weeks is 7, hence, $C_{G, BG}(z)$ represented by C(BG, G) is given by $BG(z\mod 7)$.
-The circular granularity C(BG, G) consists of repetitive pattern {BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)} which repeats itself after each period length.
+$k(day, week) =  7$ and hence, $C_{G, B}(z)$ represented by C(BG, G) is given by $B(z\mod 7)$.
+The circular granularity C(B, G) consists of repetitive pattern {B(0), B(1), B(2), B(3), B(4), B(5), B(6)} which repeats itself after each period length.
 Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14.
-C(BG, H) also repeats the pattern ${BG(0), BG(1), \dots, BG(13)}$ in every period.
+C(B, H) also repeats the pattern ${B(0), B(1), \dots, B(13)}$ in every period.
 Hence, circular granularities are represented as repetitive categorization of time in the diagram.

### 2019-12-04 CyclicGran.Rmd
demystifying definitions using day and month example

---
title: Formal definition of cyclic granularities
author: Sayani Gupta
branding: FALSE
fontfamily: mathpazo
output: pdf_document
---

# Quasi-circular

A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity. However, they are still formed with linear granularities, one of which "groups periodically into" the other. We will take the example of days and months to illustrate the notion of "groups periodically into" and "quasi-circular" granularities.

The grouping (day, month) has a period of 1 year. Ignoring leap years, this would mean that the behavior of days within months repeat every 365 days. That is, each month would consist of the same number of days in every year. We assume there are D linear granularity "day" and M linear granularity "month" in total.

\begin{equation}\label{eq:eq1}
\begin{split}
month(0) & = day(0)\bigcup day(1)\dots\bigcup day(30)\\
month(1) & = day(31)\bigcup day(32)\dots\bigcup day(58)\\
month(2) & = day(59)\bigcup day(60)\dots\bigcup day(89)\\
\dots\\
month(11) & = day(334)\dots day(363)\bigcup day(364)\\
\end{split}
\end{equation}

If we know the composition of each of the months in terms of days for one year, we can find the composition of any month beyond 1 year since the "pattern" repeats itself along the time domain due to the periodic property.

In other words,  if 
\begin{equation}\label{eq:eq2}
month(j) = day(a_1)\bigcup day(a_2)\dots\bigcup day(a_k) \quad for \quad j \in {0,1, 2, \dots, 11}
\end{equation}
then,  
\begin{equation}
month(j + R) = day(a_1 + P)\bigcup day(a_2 + P)\dots\bigcup day(a_k + P) \quad for \quad j + R \leq M
\end{equation}

Here, P = 365 and R = 12 will have the meaning of the period of the grouping (day, month) and the number of months in each of these periods.

Generalizing it to any two linear granularities G and H , the formal way of defining the property "groups periodically into" would like the following:

\newtheorem{definition}{Definition}
\begin{definition}\label{def:periodical}
A granularity H is periodical with respect to a granularity G if
(1) $G \trianglelefteq H$, and
(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
that for all $j \epsilon Z$, if $H(j) = \bigcup_{i \in S}G(i)$ and $H (j + R) \neq \phi$ then
$H (j + R) = \bigcup_{i \in S} G(i + P)$.
\end{definition}

Another way to represent \autoref{eq:eq1}

\begin{equation}\label{eq:eq3}
\begin{split}
month(0) & = \bigcup_{i \in S_0}day(i), \quad S_0 = {0, 1, 2, \dots, 30}\\
month(1) & = \bigcup_{i \in S_1}day(i), \quad S_1 = {31, 32, \dots, 58}\\
month(2) & = \bigcup_{i \in S_2}day(i), \quad S_2 = {59, 60, \dots, 89}\\
\vdots\\
month(11) &  = \bigcup_{i \in S_{11}}day(i), \quad S_{11} = {334, 335, \dots, 364}\\
\end{split}
\end{equation}

Here, ${S_0,...,S_{11}}$ are the sets of indexes of G describing $month(0), . . . , month(11)$. Then from Definition \autoref{def:periodical}, it also follows that if $H$ is periodical with respect to $G$, then

\begin{equation}\label{eq:eq4}
H(j) = \bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)
\end{equation},

where ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing $H (0), . . . , H (R - 1)$ respectively. Here $j \mod R$ represents the index among those in ${0, 1, \dots, R-1}$ that must be shifted to obtain $H(j)$. The number of periods each granule of G composing $H(j \mod R)$ should be shifted is given by $\lfloor j/R\rfloor$.

Thus,  to obtain $month(13)$ in terms of days we can either use \autoref{def:periodical} or \autoref{eq:eq4}
\begin{equation}
\begin{split}
month(13) & =  month (1 + 12)\\
 & = \bigcup_{i \in S_1} day(i + 365) \quad since \quad month(1) =  \bigcup_{i \in S_1}day(i), \quad S_1 = {31, 32, \dots, 58}\\
 & = day(31 + 365)\bigcup day(32 + 365)\bigcup day(58 + 365)\\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
month(13) & = \bigcup_{i \in S_{13 \mod 12}}day(365*\lfloor 13/12 \rfloor + i)\\
 & = \bigcup_{i \in S_1}day(365*1 + i)\\
 & = day(31 + 365)\bigcup day(32 + 365)\bigcup day(58 + 365)\\
\end{split}
\end{equation}

Till now, we saw how the formal definitions of "groups periodically into" can be seen through the example of (day, month). Now, we shall discuss how some quasi-circular granularity can be obtained. \autoref{fig:} shows that index set 34 should correspond to L(3) which is "4th-day-of-month", that of 60 should correspond to L(1), which is "2nd-day-of-month" and 334 should correspond to L(0), which is "1st-day-of-month". Now, quasi-circular granularities can be defined as follows to obtain these labels:

```{r quasi-circular-explain,echo=FALSE,out.width = "100%",fig.cap="Quasi-circular granularity day-of-month"}
knitr::include_graphics("Figs/quasi-circular-explain.png")
```

<!-- quad number \quad of \quad days \quad till \quad month(11-1) -->

<!-- Now, $35 \in month(1)$ and hence $N_k = 31$.  -->

<!-- Thus, $Q_{day, month}(35) = 35 - 31 =  L(4)$  -->

<!-- Similarly, $Q_{day, month}(60) = 60 - (31+28) =  L(1)$, $60 \in month(2)$ and N_k is number of days in month(0) and month(1). -->

\begin{definition}\label{quasicircular}
A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if

\begin{equation}\label{eq:eq3}
Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
\end{equation}

where,  
$z$ denotes the index set,
$B$ denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity $G'$ with irregular mapping, {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$, $T_w$ are the sets of indices of B describing $G'(w)$ such that 
\begin{equation}
G'(w) = \bigcup_{z \in T_w}B(z),\\
\end{equation}
$T_w = 0 \quad for \quad w\epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
\end{definition}

Here, $\vert T_w \vert$ = $\vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R)^{th}$ granule will have the same number of granules of B. $T_w$ are the sets of indexes of $B$ describing $G'(0), . . . , G'(R'- 1)$. The term $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$.

\begin{equation}
\begin{split}
Q_{day, month}(34) & = L(35 - 31) \quad since \quad 35 \in \quad month(1)\\
Q_{day, month}(60) & = L(60 - 31 - 28), \quad since\quad 60 \in \quad month(2)\\
Q_{day, month}(334) & = L(334 - 334) \quad since \quad 334 \in \quad month(11)\\
\end{split}
\end{equation}.

\begin{definition}\label{aperiodic}
An aperiodic granularity $A_(B, G'')$ that relates linear granularities G and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by

\begin{equation}\label{eq:eq4}
\begin{split}
A_{B, G''}(z) & = L(1), z \in T_q\\
     & = L(0), z \not\in T_q\\
\end{split}
\end{equation}
where,  
$z$ denotes the index set (index of $B$),
$T_q$ are the sets of indices of $B$ describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
\end{definition}

Discussion: Aperiodic time granularities map a time index to a categorical variable containing binary values. It takes the value 1 whenever the time index coincides with the events in context (e.g. public or school holiday) and 0 when it does not. Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector. An example of aperiodic granularity can be one that relates a bottom granularity ($BG$) and linear granularity ($G''$) such that $BG \trianglelefteq G''$, $BG \not\preceq G$ and $BG$ is not periodic or quasi-periodic with respect to $G''$, as shown in \autoref{fig:aperiodic}.

```{r aperiodic, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
knitr::include_graphics("Figs/aperiodic.png")

### 2019-12-04 main.Rmd
demystifying definitions using day and month example

@@
 
-\newtheorem{timedom}{Definition}
+\newtheorem{timedom}{Definition}[section]
+
 \begin{timedom}\label{def:timedom}
@@
 However, the labels can coincide with indexes in which case integers are directly used to refer to categorisations of the circular granularity.
+Note that each
+circular granularity can use a different set of labels.
 Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
@@
 $k(day, week) =  7$ and hence, $C_{G, B}(z)$ represented by C(BG, G) is given by $B(z\mod 7)$.
-The circular granularity C(B, G) consists of repetitive pattern {B(0), B(1), B(2), B(3), B(4), B(5), B(6)} which repeats itself after each period length.
-Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14.
-C(B, H) also repeats the pattern ${B(0), B(1), \dots, B(13)}$ in every period.
+The circular granularity C(B, G) consists of repetitive pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
+$L$ can be defined such that the labels corresponding to ${0,1,\dots,6}$ correspond to different days of the week ${Sun, Mon ,\dots, Sat}$ .
+Similarly, C(B, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14.
+C(B, H) also repeats the pattern ${L'(0), \dots, L'(13)}$ in every period.
 Hence, circular granularities are represented as repetitive categorization of time in the diagram.
@@
 
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = BG(\lfloor z/k(BG,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the period length of G within H.
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/k(B,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the number of granules of G in one granule of H.
 
@@
 It is possible that none of these two linear granularities are bottom granularities.
-But the representation of the resultant circular granularity will be a function of the bottom granularity.
-Suppose the bottom granularity is minutes, and $k_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$, also referred to as the period length.
+But the representation of the resultant circular granularity will be a function of the index set.
+Suppose the bottom granularity is minutes, $k_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$ and $L_i$ is the associated label mapping.
 
@@
 \toprule
-Minute-of-Hour: & $C_1(z)$ & $= z \mod 60$ & $k_1 =$&$60$ \\
-Minute-of-Day: & $C_2(z)$ & $= z \mod 60*24$ & $k_2=$&$`r 60*24`$\\
-Hour-of-Day: & $C_4(z)$ & $= \lfloor z/60\rfloor\mod 24$  & $k_3 =$&$24$ \\
-Hour-of-Week: & $C_6(z)$ & $= \lfloor z/60\rfloor\mod 24*7$  & $k_4=$&$`r 24*7`$\\
-Day-of-Week: & $C_7(z)$ & $= \lfloor z/24*60\rfloor \mod 7$ & $k_5=$&$7$\\
+Minute-of-Hour: & $C_1(z)$ & $= L_1(z \mod 60)$ & $k_1 =$&$60$ \\
+Minute-of-Day: & $C_2(z)$ & $= L_2(z \mod 60*24)$ & $k_2=$&$`r 60*24`$\\
+Hour-of-Day: & $C_4(z)$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $k_3 =$&$24$ \\
+Hour-of-Week: & $C_6(z)$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $k_4=$&$`r 24*7`$\\
+Day-of-Week: & $C_7(z)$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $k_5=$&$7$\\
 \bottomrule
@@
 A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
-For example, \autoref{tab:quasi} shows some example of quasi-circular granularities ($QC_i$) and plausible choices of number of granules ($k_i$) of the finer granularity inside each granule of the coarser one in the cyclic granularity ($QC_i$).
+For example, \autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) and plausible choices of number of granules ($k_i$) of the finer granularity inside each granule of the coarser one in the cyclic granularity ($Q_i$).
 
@@
 
-Day-of-Month: & $QC_1(z)$ && $k_1=31, 30, 29, 28$\\
-Hour-of-Month: & $QC_2(z)$ && $k_2=24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $QC_3(z)$ && $k_3=366, 365$\\
-Week-of-Month: & $QC_4(z)$ && $k_4=5, 4$\\ 
+Day-of-Month: & $Q_1(z)$ && $k_1=31, 30, 29, 28$\\
+Hour-of-Month: & $Q_2(z)$ && $k_2=24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $Q_3(z)$ && $k_3=366, 365$\\
+Week-of-Month: & $Q_4(z)$ && $k_4=5, 4$\\ 
 \bottomrule
@@
 
-\newtheorem{timedom1}{Definition}[subsection]
-\begin{timedom1}\label{def:quasicircular}
-A quasi-circular granularity QC_(BG, G') that relates linear granularities G' and bottom granularity BG, if
+\begin{timedom}\label{quasicircular}
+A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if
 
 \begin{equation}\label{eq:eq3}
-QC_{BG, G'}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
+Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
 \end{equation}
@@
 where,  
-z denotes the index set,  
-BG denotes the index of the tsibble (bottom granularity),  
-G' is a linear granularity periodic with respect to BG with irregular mapping,   
-R' denotes the number of granules of G' in each period of the grouping (BG,G),   
-T_w are the sets of indices of BG describing G'(w) such that $G'(w) = \bigcup_{z \in T_w}BG(z)$,$T_w = 0$ for $w \epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-\end{timedom1}
+$z$ denotes the index set,\\
+$B$ denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity $G'$ with irregular mapping,\\
+{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
+$T_w$ are the sets of indices of B describing $G'(w)$ such that 
+\begin{equation}
+G'(w) = \bigcup_{z \in T_w}B(z),\\
+\end{equation}
+$T_w = 0$ for $w \epsilon Z_{<0}$,\\
+$\vert T_w \vert$ is the cardinality of set $T_w$.
+\end{timedom}
 
@@
 
-**Discussion**: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-gran}.
+**Discussion**: According to the first condition, assume, for instance, that G'(w) is the union of the granules G(a1), G(a2), .
+.
+.
+, G(ak).
+The periodicity property (condition (2) in the definition) ensures that the Rth granule after H(i), i.e., H(i + R), if non-empty, is the union of G(a1 + P ), G(a2 + P), .
+.
+.
+, G(ak + P).
+
+Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-gran}.
 Each granule is represented by a box.
@@
 
-\begin{timedom}\label{def:acyclic}
+\begin{timedom}\label{aperiodic}
 An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by

### 2019-12-04 CyclicGran.Rmd
spell check done

@@
 $z$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity $G'$ with irregular mapping, {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$, $T_w$ are the sets of indices of B describing $G'(w)$ such that
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$, $T_w$ are the sets of indices of B describing $G'(w)$ such that
 \begin{equation}

### 2019-12-04 main.Rmd
spell check done

@@
 
-**Dicussion:** The first condition in \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+**Discussion:** The first condition in \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
 \autoref{fig:linear-time} shows the implication of this condition.
@@
 
-**Dicussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.
+**Discussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.
 $G \preceq H$ if every granule in G is a subset of some granule in H.
@@
 
-**Dicussion:**
+**Discussion:**
 
@@
 
-**Dicussion:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+**Discussion:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
 
@@
 
-For more variations of grouping operations and granule oriented operations, the readers are remcommended to see [@Ning2002-tf].s
+For more variations of grouping operations and granule oriented operations, the readers are recommended to see [@Ning2002-tf].s
 
@@
 Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
-In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
+In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
 Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
 
-Cyclic time granularities which accomodate periodicities can be constructed by relating two linear granularities.
+Cyclic time granularities which accommodate periodicities can be constructed by relating two linear granularities.
 The mappings between linear granularities can be regular or irregular.
@@
 Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
-The formalisms would differ for circular and quasi-circular granularities although both accomodate periodicities in data.
+The formalism would differ for circular and quasi-circular granularities although both accommodate periodicities in data.
 
@@
 z denotes the index set, \\
-B denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity G with regular mapping, \\
+B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity G with regular mapping, \\
 {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
@@
 
-**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorisation of the circular granularity.
-However, the labels can coincide with indexes in which case integers are directly used to refer to categorisations of the circular granularity.
+**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity.
+However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
 Note that each
@@
 $z$ denotes the index set,\\
-$B$ denotes a full-integer labelled bottom granularity which groups periodicially into linear granularity $G'$ with irregular mapping,\\
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping,\\
 {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
@@
 
-<!-- The cyclic time granularities which are essentially categorisations of time are discrete variable, mapped as a factor since they have an inherent order.
+<!-- The cyclic time granularities which are essentially categorizations of time are discrete variable, mapped as a factor since they have an inherent order.
 Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
@@
 The measured variable can be dicrete or continuous.
-The probability distribution of these univariate measured variable could be studied by estimating the proability density function (pdf for continuous) or probability mass function (pmf for dicrete).
+The probability distribution of these univariate measured variable could be studied by estimating the probability density function (pdf for continuous) or probability mass function (pmf for dicrete).
 Both pdf and pmf are concerned with point probabilities of the random variable.
@@
 Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962).
-The nearest neighbour estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977).
+The nearest neighbor estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977).
 There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.
@@
 
-There have been several other nonparametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
+There have been several other non parametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
 Also sample quantiles can be of the form that are not defined for all p including hinges, letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
 
-<!-- would invlove a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis. -->
+<!-- would involve a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis. -->
 

### 2019-12-05 CyclicGran.Rmd
demystifying definitions using day and month example

@@
 ---
-title: Formal definition of cyclic granularities
+title: Revisiting some definitions with examples from Gregorian calendar
 author: Sayani Gupta
@@
 
-# Quasi-circular
+# Groups-periodically
 
-A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
-However, they are still formed with linear granularities, one of which "groups periodically into" the other.
-We will take the example of days and months to illustrate the notion of "groups periodically into" and "quasi-circular" granularities.
+The concept already existed in the literature. This is an attempt to demistify the notations using the linear granularities day and month.
 
@@
 
+# Circular (Periodic linear granularities + regular mapping)
+
+A **circular** granularity can be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+They are formed with linear granularities, one of which "groups periodically into" the other.
+
+\begin{definition}\label{def:circular}
+A circular granularity C(B, G) relates a linear granularity G to the bottom granularity B, if
+
+\begin{equation} \label{eq:eq2}
+C_{B, G}(z) = {L}(z\mod k(B,G)) \quad \forall z \in \mathbb{Z}^+
+\end{equation}
+where,
+z denotes the index set,
+B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity G with regular mapping, 
+{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,
+and k(B,G) is the number of granules of B in each granule of G.
+\end{definition}
+
+In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity.
+However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
+Note that each
+circular granularity can use different label mappings.
+In \autoref{fig:day-of-week}, the label mapping L can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sunday, Monday, \dots, Saturday)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({0, 1, \dots, 6)}$
+depending on the context.
+
+```{r day-of-week,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
+knitr::include_graphics("Figs/day-of-week.png")
+```
+
+# Quasi-circular (Periodic linear granularities + irregular mapping)
+
+A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+However, they are still formed with linear granularities, one of which "groups periodically into" the other.
+We will take the example of days and months to illustrate the notion of "groups periodically into" and "quasi-circular" granularities.
+
 Till now, we saw how the formal definitions of "groups periodically into" can be seen through the example of (day, month).
 Now, we shall discuss how some quasi-circular granularity can be obtained.
-\autoref{fig:} shows that index set 34 should correspond to L(3) which is "4th-day-of-month", that of 60 should correspond to L(1), which is "2nd-day-of-month" and 334 should correspond to L(0), which is "1st-day-of-month".
+\autoref{fig:quasi-circular-explain} shows that index set 34 should correspond to L(3) which is "4th-day-of-month", that of 60 should correspond to L(1), which is "2nd-day-of-month" and 334 should correspond to L(0), which is "1st-day-of-month".
 Now, quasi-circular granularities can be defined as follows to obtain these labels:
@@
 \begin{definition}\label{quasicircular}
-A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if
+A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if
 
@@
 
+# Aperiodic (Aperiodic linear granularities)
+
 \begin{definition}\label{aperiodic}
-An aperiodic granularity $A_(B, G'')$ that relates linear granularities G and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by
+An aperiodic granularity $A_{B, G''}$ that relates linear granularities G and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by
 

### 2019-12-05 CyclicGran.Rmd
demystifying definitions using day and month example

@@
 ---
-title: Revisiting some definitions with examples from Gregorian calendar
+title: Revisiting some definitions with common examples from Gregorian calendar
 author: Sayani Gupta
@@
 
-Thus,  to obtain $month(13)$ in terms of days we can either use \autoref{def:periodical} or \autoref{eq:eq4}
+Thus,  to obtain $month(13)$ in terms of days we can either use Definition \autoref{def:periodical} or \autoref{eq:eq4}
 \begin{equation}
@@
 
-A **circular** granularity can be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+A **circular** granularity can be defined using modular arithmetic due to its regular mapping with the bottom granularity.
 They are formed with linear granularities, one of which "groups periodically into" the other.
+In the following definition, we would assume that bottom granularity $B$ groups periodically into $G$ and number of granules of $B$ in each granule of $G$ is constant.
+Thus, if $P$ and $R$  have the meaning of the period of the grouping $(B, G)$ and the number of granules of $G$ in each of these periods, then for circular granularity, $P/R \in \mathbb{Z}^+$.
+Also, since each granule of $G$ has same number of granules of $B$, $R$ can be taken to be $1$ and $P$ can be interpreted as the number of granules of B in any granule of G.
 
@@
 \begin{equation} \label{eq:eq2}
-C_{B, G}(z) = {L}(z\mod k(B,G)) \quad \forall z \in \mathbb{Z}^+
+\begin{split}
+C_{B, G}(z) & = {L}(z\mod P) \quad \forall z \in \mathbb{Z}^+ \\
+\end{split}
 \end{equation}
@@
 B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity G with regular mapping, 
-{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,
-and k(B,G) is the number of granules of B in each granule of G.
+{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (P-1)}$,
+and P is the number of granules of B in each granule of G.
 \end{definition}
@@
 A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+Here $P/R \not \in \mathbb{Z}^+$.
 However, they are still formed with linear granularities, one of which "groups periodically into" the other.
-We will take the example of days and months to illustrate the notion of "groups periodically into" and "quasi-circular" granularities.
-
-Till now, we saw how the formal definitions of "groups periodically into" can be seen through the example of (day, month).
-Now, we shall discuss how some quasi-circular granularity can be obtained.
-\autoref{fig:quasi-circular-explain} shows that index set 34 should correspond to L(3) which is "4th-day-of-month", that of 60 should correspond to L(1), which is "2nd-day-of-month" and 334 should correspond to L(0), which is "1st-day-of-month".
+We shall discuss an example of  quasi-circular granularity day-of-month in \autoref{fig:quasi-circular-explain}.
+The index 34 should correspond to L(3) which is "4th-day-of-month", that of 60 should correspond to L(1), which is "2nd-day-of-month" and 334 should correspond to L(0), which is "1st-day-of-month".
 Now, quasi-circular granularities can be defined as follows to obtain these labels:
@@
 
+In \autoref{fig:quasi-circular-explain}, the label mapping L can be defined as $L: ({0,1,2, \dots, 31}) \longmapsto\ ({DOM_1, DOM_2, \dots, DOM_{31})}$ or $L: ({0,1,2, \dots, 31}) \longmapsto\ ({1, 2, \dots, 31)}$
+depending on the context.
+
 \begin{equation}
@@
 \begin{split}
-A_{B, G''}(z) & = L(1), z \in T_q\\
-     & = L(0), z \not\in T_q\\
+A_{B, G''}(z) & = L(1), \quad z \in T_q\\
+     & = L(0), \quad z \not\in T_q\\
 \end{split}
 \end{equation}
-where,  
-$z$ denotes the index set (index of $B$),
-$T_q$ are the sets of indices of $B$ describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
+where  
+$z$ denotes the index set,
+$T_q$ are the sets of indices of $B$ describing $G''(q)$ such that $$G''(q) = \bigcup_{z \in T_q}B(z)$$
 \end{definition}
 
-Discussion: Aperiodic time granularities map a time index to a categorical variable containing binary values.
+ Aperiodic time granularities map a time index to binary categories.
 It takes the value 1 whenever the time index coincides with the events in context (e.g.
 public or school holiday) and 0 when it does not.
-Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector.
-An example of aperiodic granularity can be one that relates a bottom granularity ($BG$) and linear granularity ($G''$) such that $BG \trianglelefteq G''$, $BG \not\preceq G$ and $BG$ is not periodic or quasi-periodic with respect to $G''$, as shown in \autoref{fig:aperiodic}.
+\autoref{fig:aperiodic} shows the linear granularity academic semester and days.
+Academic semesters are not periodic with respect to days.
+The label mapping L can be defined as $L: ({0,1}) \longmapsto\ ({On, Off)}$ or can be customised based on context.
 
-```{r aperiodic, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
-knitr::include_graphics("Figs/aperiodic.png")
+```{r aperiodic, echo=FALSE,out.width = "100%",fig.cap=" Aperiodic granularity day-semester"}
+knitr::include_graphics("Figs/academ-semester.png")

### 2019-12-05 CyclicGran.Rmd
demystifying defintions

@@
 fontfamily: mathpazo
-output: pdf_document
+output:
+  MonashEBSTemplates::memo:
+    keep_tex: true
+header-includes:
+  - \usepackage[lf,t]{carlito}
+  - \titleformat{\section}[block]{\fontsize{12}{14}\bfseries\sffamily}{\thesection}{0.4em}{}
+  - \def\mod{\text{~mod~}}
 ---
@@
 
-The concept already existed in the literature. This is an attempt to demistify the notations using the linear granularities day and month.
+The concept already existed in the literature. This is an attempt to demystify the notations using the linear granularities day and month.
 
@@
 \dots\\
-month(11) & = day(334)\dots day(363)\bigcup day(364)\\
+month(11) & = day(334)\dots day(363)\bigcup day(364)\\a
 \end{split}
@@
 
-Generalizing it to any two linear granularities G and H , the formal way of defining the property "groups periodically into" would like the following:
+Generalizing it to any two linear granularities $G$ and $H$ , the formal way of defining the property "groups periodically into" would like the following:
 
@@
 \begin{definition}\label{def:periodical}
-A granularity H is periodical with respect to a granularity G if
-(1) $G \trianglelefteq H$, and
+A granularity H is periodical with respect to a granularity $G$ if
+(1) $G \trianglelefteq H$, and \\
 (2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
@@
 
-Here, ${S_0,...,S_{11}}$ are the sets of indexes of G describing $month(0), .
+Here, ${S_0,...,S_{11}}$ are the sets of indexes of $G$ describing $month(0), .
 .
@@
 \begin{equation}\label{eq:eq4}
-H(j) = \bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)
+H(j) = \bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i), 
 \end{equation},
 
-where ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing $H (0), .
+where ${S_0,...,S_{R-1}}$ are the sets of indexes of $G$ describing $H (0), .
 .
@@
 Here $j \mod R$ represents the index among those in ${0, 1, \dots, R-1}$ that must be shifted to obtain $H(j)$.
-The number of periods each granule of G composing $H(j \mod R)$ should be shifted is given by $\lfloor j/R\rfloor$.
+The number of periods each granule of $G$ composing $H(j \mod R)$ should be shifted is given by $\lfloor j/R\rfloor$.
 
 Thus,  to obtain $month(13)$ in terms of days we can either use Definition \autoref{def:periodical} or \autoref{eq:eq4}
+
 \begin{equation}
@@
 \begin{definition}\label{def:circular}
-A circular granularity C(B, G) relates a linear granularity G to the bottom granularity B, if
+A circular granularity C(B, G) relates a linear granularity $G$ to the bottom granularity B, if
 
-\begin{equation} \label{eq:eq2}
+\begin{equation} \label{eq:eq5}
 \begin{split}
@@
 \end{equation}
-where,
+where
 z denotes the index set,
-B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity G with regular mapping, 
+B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, 
 {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (P-1)}$,
@@
 A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
-Here $P/R \not \in \mathbb{Z}^+$.
 However, they are still formed with linear granularities, one of which "groups periodically into" the other.
@@
 The index 34 should correspond to L(3) which is "4th-day-of-month", that of 60 should correspond to L(1), which is "2nd-day-of-month" and 334 should correspond to L(0), which is "1st-day-of-month".
-Now, quasi-circular granularities can be defined as follows to obtain these labels:
+Now, quasi-circular granularities can be defined as follows to manipulate the indices to our need:
 
@@
 
-\begin{equation}\label{eq:eq3}
+\begin{equation}\label{eq:eq6}
 Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
@@
 
-where,  
+where
 $z$ denotes the index set,
@@
 \begin{split}
-Q_{day, month}(34) & = L(35 - 31) \quad since \quad 35 \in \quad month(1)\\
-Q_{day, month}(60) & = L(60 - 31 - 28), \quad since\quad 60 \in \quad month(2)\\
-Q_{day, month}(334) & = L(334 - 334) \quad since \quad 334 \in \quad month(11)\\
+Q_{day, month}(34) & = L(34 - 31)       = L(3) \quad since \quad 35 \in \quad month(1)\\
+Q_{day, month}(60) & = L(60 - 31 - 28)  = L(2) , \quad since\quad 60 \in \quad month(2)\\
+Q_{day, month}(334) & = L(334 - 334)    = L(0) \quad since \quad 334 \in \quad month(11)
 \end{split}
@@
 \begin{definition}\label{aperiodic}
-An aperiodic granularity $A_{B, G''}$ that relates linear granularities G and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by
+An aperiodic granularity $A_{B, G''}$ that relates linear granularities $G''$ and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by
 
-\begin{equation}\label{eq:eq4}
+\begin{equation}\label{eq:eq7}
 \begin{split}
@@
  Aperiodic time granularities map a time index to binary categories.
-It takes the value 1 whenever the time index coincides with the events in context (e.g.
-public or school holiday) and 0 when it does not.
+It takes a value (eg.
+1) whenever the time index coincides with the events in context (e.g.
+public or school holiday) and another value 0 (e.g.
+0) when it does not.
 \autoref{fig:aperiodic} shows the linear granularity academic semester and days.
 Academic semesters are not periodic with respect to days.
-The label mapping L can be defined as $L: ({0,1}) \longmapsto\ ({On, Off)}$ or can be customised based on context.
+The label mapping L can be defined as $L: ({0,1}) \longmapsto\ ({On, Off)}$ or can be customized based on context.
 

### 2019-12-05 CyclicGran.tex
demystifying defintions

\documentclass[10pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{csquotes}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Revisiting some definitions with common examples from Gregorian calendar},
            pdfauthor={Sayani Gupta},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2cm,right=2cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,backend=biber, natbib=true,]{biblatex}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Revisiting some definitions with common examples from Gregorian calendar}

\author{Sayani Gupta}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\usepackage{bm,url}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\lfoot{}\cfoot{}\rfoot{}
\lhead{\textsf{Revisiting some definitions with common examples from Gregorian calendar}}
\rhead{\textsf{\thepage}}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\RequirePackage{float}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*3}{*1}
\titlespacing{\subsection}{0pt}{*1}{*0.5}

%% LINE AND PAGE BREAKING
\sloppy
\raggedbottom
\usepackage[bottom]{footmisc}
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

\usepackage[showonlyrefs]{mathtools}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}
\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[article]{url}{}
%\DeclareFieldFormat[book]{url}{}
%\DeclareFieldFormat[inbook]{url}{}
%\DeclareFieldFormat[incollection]{url}{}
%\DeclareFieldFormat[inproceedings]{url}{}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
%\DeclareFieldFormat{extrayear}{}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}
\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother
\renewcommand*{\finalnamedelim}{%
  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
  \addspace\&\space}

%%% Change title format
\usepackage{fancybox,color}
\usepackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}
\usepackage{titling}

\pretitle{%

\vspace*{-1cm}

\LARGE\bfseries}
\posttitle{\vspace*{0.3cm}\par}
\preauthor{\large}
\postauthor{\hfill}
\predate{\small}
\postdate{\vspace*{0.1cm}}

\newenvironment{fminipage}%
{\begin{Sbox}\begin{minipage}{\textwidth}}%
{\end{minipage}\end{Sbox}\colorbox[gray]{0.8}{\TheSbox}}
\raggedbottom

\usepackage[australian]{babel}
\date{\today}

\usepackage[lf,t]{carlito}
\titleformat{\section}[block]{\fontsize{12}{14}\bfseries\sffamily}{\thesection}{0.4em}{}
\def\mod{\text{~mod~}}

\begin{document}
\vspace*{-2cm}
\begin{fminipage}\sffamily
\maketitle
\end{fminipage}\vspace*{0.5cm}
\setstretch{1.2}

\hypertarget{groups-periodically}{%
\section{Groups-periodically}\label{groups-periodically}}

The concept already existed in the literature. This is an attempt to demystify the notations using the linear granularities day and month.

The grouping (day, month) has a period of 1 year. Ignoring leap years, this would mean that the behavior of days within months repeat every 365 days. That is, each month would consist of the same number of days in every year. We assume there are D linear granularity \enquote{day} and M linear granularity \enquote{month} in total.

\begin{equation}\label{eq:eq1}
\begin{split}
month(0) & = day(0)\bigcup day(1)\dots\bigcup day(30)\\
month(1) & = day(31)\bigcup day(32)\dots\bigcup day(58)\\
month(2) & = day(59)\bigcup day(60)\dots\bigcup day(89)\\
\dots\\
month(11) & = day(334)\dots day(363)\bigcup day(364)\\a
\end{split}
\end{equation}

If we know the composition of each of the months in terms of days for one year, we can find the composition of any month beyond 1 year since the \enquote{pattern} repeats itself along the time domain due to the periodic property.

In other words, if
\begin{equation}\label{eq:eq2}
month(j) = day(a_1)\bigcup day(a_2)\dots\bigcup day(a_k) \quad for \quad j \in {0,1, 2, \dots, 11}
\end{equation}
then,\\
\begin{equation}
month(j + R) = day(a_1 + P)\bigcup day(a_2 + P)\dots\bigcup day(a_k + P) \quad for \quad j + R \leq M
\end{equation}

Here, P = 365 and R = 12 will have the meaning of the period of the grouping (day, month) and the number of months in each of these periods.

Generalizing it to any two linear granularities \(G\) and \(H\) , the formal way of defining the property \enquote{groups periodically into} would like the following:

\newtheorem{definition}{Definition}
\begin{definition}\label{def:periodical}
A granularity H is periodical with respect to a granularity $G$ if
(1) $G \trianglelefteq H$, and \\
(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
that for all $j \epsilon Z$, if $H(j) = \bigcup_{i \in S}G(i)$ and $H (j + R) \neq \phi$ then
$H (j + R) = \bigcup_{i \in S} G(i + P)$.
\end{definition}

Another way to represent \autoref{eq:eq1}

\begin{equation}\label{eq:eq3}
\begin{split}
month(0) & = \bigcup_{i \in S_0}day(i), \quad S_0 = {0, 1, 2, \dots, 30}\\
month(1) & = \bigcup_{i \in S_1}day(i), \quad S_1 = {31, 32, \dots, 58}\\
month(2) & = \bigcup_{i \in S_2}day(i), \quad S_2 = {59, 60, \dots, 89}\\
\vdots\\
month(11) &  = \bigcup_{i \in S_{11}}day(i), \quad S_{11} = {334, 335, \dots, 364}\\
\end{split}
\end{equation}

Here, \({S_0,...,S_{11}}\) are the sets of indexes of \(G\) describing \(month(0), . . . , month(11)\). Then from Definition \autoref{def:periodical}, it also follows that if \(H\) is periodical with respect to \(G\), then

\begin{equation}\label{eq:eq4}
H(j) = \bigcup_{i \in S_j \text{~mod~}R}G(P*\lfloor j/R \rfloor + i), 
\end{equation},

where \({S_0,...,S_{R-1}}\) are the sets of indexes of \(G\) describing \(H (0), . . . , H (R - 1)\) respectively. Here \(j \text{~mod~}R\) represents the index among those in \({0, 1, \dots, R-1}\) that must be shifted to obtain \(H(j)\). The number of periods each granule of \(G\) composing \(H(j \text{~mod~}R)\) should be shifted is given by \(\lfloor j/R\rfloor\).

Thus, to obtain \(month(13)\) in terms of days we can either use Definition \autoref{def:periodical} or \autoref{eq:eq4}

\begin{equation}
\begin{split}
month(13) & =  month (1 + 12)\\
 & = \bigcup_{i \in S_1} day(i + 365) \quad since \quad month(1) =  \bigcup_{i \in S_1}day(i), \quad S_1 = {31, 32, \dots, 58}\\
 & = day(31 + 365)\bigcup day(32 + 365)\bigcup day(58 + 365)\\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
month(13) & = \bigcup_{i \in S_{13 \text{~mod~}12}}day(365*\lfloor 13/12 \rfloor + i)\\
 & = \bigcup_{i \in S_1}day(365*1 + i)\\
 & = day(31 + 365)\bigcup day(32 + 365)\bigcup day(58 + 365)\\
\end{split}
\end{equation}

\hypertarget{circular-periodic-linear-granularities-regular-mapping}{%
\section{Circular (Periodic linear granularities + regular mapping)}\label{circular-periodic-linear-granularities-regular-mapping}}

A \textbf{circular} granularity can be defined using modular arithmetic due to its regular mapping with the bottom granularity. They are formed with linear granularities, one of which \enquote{groups periodically into} the other. In the following definition, we would assume that bottom granularity \(B\) groups periodically into \(G\) and number of granules of \(B\) in each granule of \(G\) is constant. Thus, if \(P\) and \(R\) have the meaning of the period of the grouping \((B, G)\) and the number of granules of \(G\) in each of these periods, then for circular granularity, \(P/R \in \mathbb{Z}^+\). Also, since each granule of \(G\) has same number of granules of \(B\), \(R\) can be taken to be \(1\) and \(P\) can be interpreted as the number of granules of B in any granule of G.

\begin{definition}\label{def:circular}
A circular granularity C(B, G) relates a linear granularity $G$ to the bottom granularity B, if

\begin{equation} \label{eq:eq5}
\begin{split}
C_{B, G}(z) & = {L}(z\text{~mod~}P) \quad \forall z \in \mathbb{Z}^+ \\
\end{split}
\end{equation}
where
z denotes the index set,
B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, 
{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (P-1)}$,
and P is the number of granules of B in each granule of G.
\end{definition}

In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity. However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity. Note that each
circular granularity can use different label mappings. In \autoref{fig:day-of-week}, the label mapping L can be defined as \(L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}\) or \(L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sunday, Monday, \dots, Saturday)}\) or \(L: ({0,1,2, \dots, 6}) \longmapsto\ ({0, 1, \dots, 6)}\)
depending on the context.

\begin{figure}
\includegraphics[width=1\linewidth]{Figs/day-of-week} \caption{Circular granularity day-of-week}\label{fig:day-of-week}
\end{figure}

\hypertarget{quasi-circular-periodic-linear-granularities-irregular-mapping}{%
\section{Quasi-circular (Periodic linear granularities + irregular mapping)}\label{quasi-circular-periodic-linear-granularities-irregular-mapping}}

A \textbf{quasi-circular} granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity. However, they are still formed with linear granularities, one of which \enquote{groups periodically into} the other. We shall discuss an example of quasi-circular granularity day-of-month in \autoref{fig:quasi-circular-explain}. The index 34 should correspond to L(3) which is \enquote{4th-day-of-month}, that of 60 should correspond to L(1), which is \enquote{2nd-day-of-month} and 334 should correspond to L(0), which is \enquote{1st-day-of-month}. Now, quasi-circular granularities can be defined as follows to manipulate the indices to our need:

\begin{figure}
\includegraphics[width=1\linewidth]{Figs/quasi-circular-explain} \caption{Quasi-circular granularity day-of-month}\label{fig:quasi-circular-explain}
\end{figure}

\begin{definition}\label{quasicircular}
A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if

\begin{equation}\label{eq:eq6}
Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \text{~mod~}R'} \vert), \quad z \in T_k
\end{equation}

where
$z$ denotes the index set,
$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, {L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$, $T_w$ are the sets of indices of B describing $G'(w)$ such that 
\begin{equation}
G'(w) = \bigcup_{z \in T_w}B(z),\\
\end{equation}
$T_w = 0 \quad for \quad w\epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
\end{definition}

Here, \(\vert T_w \vert\) = \(\vert T_{w\text{~mod~}R'} \vert\) since every \(w^{th}\) and \((w+R)^{th}\) granule will have the same number of granules of B. \(T_w\) are the sets of indexes of \(B\) describing \(G'(0), . . . , G'(R'- 1)\). The term \(\sum_{w=0}^{k-1}\vert T_{w \text{~mod~}R'}\vert\) denotes the number of granules of \(B\) till the \((k-1)^{th}\) granule of \(G'\).

In \autoref{fig:quasi-circular-explain}, the label mapping L can be defined as \(L: ({0,1,2, \dots, 31}) \longmapsto\ ({DOM_1, DOM_2, \dots, DOM_{31})}\) or \(L: ({0,1,2, \dots, 31}) \longmapsto\ ({1, 2, \dots, 31)}\)
depending on the context.

\begin{equation}
\begin{split}
Q_{day, month}(34) & = L(34 - 31)       = L(3) \quad since \quad 35 \in \quad month(1)\\
Q_{day, month}(60) & = L(60 - 31 - 28)  = L(2) , \quad since\quad 60 \in \quad month(2)\\
Q_{day, month}(334) & = L(334 - 334)    = L(0) \quad since \quad 334 \in \quad month(11)
\end{split}
\end{equation}.

\hypertarget{aperiodic-aperiodic-linear-granularities}{%
\section{Aperiodic (Aperiodic linear granularities)}\label{aperiodic-aperiodic-linear-granularities}}

\begin{definition}\label{aperiodic}
An aperiodic granularity $A_{B, G''}$ that relates linear granularities $G''$ and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by

\begin{equation}\label{eq:eq7}
\begin{split}
A_{B, G''}(z) & = L(1), \quad z \in T_q\\
     & = L(0), \quad z \not\in T_q\\
\end{split}
\end{equation}
where  
$z$ denotes the index set,
$T_q$ are the sets of indices of $B$ describing $G''(q)$ such that $$G''(q) = \bigcup_{z \in T_q}B(z)$$
\end{definition}

Aperiodic time granularities map a time index to binary categories. It takes a value (eg. 1) whenever the time index coincides with the events in context (e.g.~public or school holiday) and another value 0 (e.g.~0) when it does not. \autoref{fig:aperiodic} shows the linear granularity academic semester and days. Academic semesters are not periodic with respect to days. The label mapping L can be defined as \(L: ({0,1}) \longmapsto\ ({On, Off)}\) or can be customized based on context.

\begin{figure}
\includegraphics[width=1\linewidth]{Figs/academ-semester} \caption{ Aperiodic granularity day-semester}\label{fig:aperiodic}
\end{figure}

\printbibliography

\end{document}

### 2019-12-05 CyclicGran.Rmd
demystifying defintions

@@
 \dots\\
-month(11) & = day(334)\dots day(363)\bigcup day(364)\\a
+month(11) & = day(334)\dots day(363)\bigcup day(364)\\
 \end{split}
@@
 Q_{day, month}(60) & = L(60 - 31 - 28)  = L(2) , \quad since\quad 60 \in \quad month(2)\\
-Q_{day, month}(334) & = L(334 - 334)    = L(0) \quad since \quad 334 \in \quad month(11)
+Q_{day, month}(334) & = L(334 - 334)    = L(0) \quad since \quad 334 \in \quad month(11)\\
+Q_{day, month}(439) & = L(439 - \sum_{w=0}^{14-1}\vert T_{w \mod 12} \vert)\quad since \quad 439 \in \quad month(14)\\
+& = L(439 - \sum_{w=0}^{13}\vert T_{w \mod R'} \vert)\\
+&  = L(439 - (2\vert T_{0} \vert + 2\vert T_{1}\vert + \vert T_{3}\vert + \vert T_{4}\vert + \dots + \vert T_{11}\vert))\\
+&  = L(439 - 2\times 31 - 2\times 28 - 31 - 30 - \dots - 30)\\
+&  = L(439 - 423)\\
+&  = L(6)
 \end{split}
 \end{equation}.
+
+The idea is we can find any day of the month beyond 1 year provided we have the composition of days across months in one year due to the periodic property.
 

### 2019-12-05 CyclicGran.tex
demystifying defintions

@@
 Q_{day, month}(60) & = L(60 - 31 - 28)  = L(2) , \quad since\quad 60 \in \quad month(2)\\
-Q_{day, month}(334) & = L(334 - 334)    = L(0) \quad since \quad 334 \in \quad month(11)
+Q_{day, month}(334) & = L(334 - 334)    = L(0) \quad since \quad 334 \in \quad month(11)\\
+Q_{day, month}(439) & = L(439 - \sum_{w=0}^{14-1}\vert T_{w \text{~mod~}12} \vert)\quad since \quad 439 \in \quad month(14)\\
+& = L(439 - \sum_{w=0}^{13}\vert T_{w \text{~mod~}R'} \vert)\\
+&  = L(439 - (2\vert T_{0} \vert + 2\vert T_{1}\vert + \vert T_{3}\vert + \vert T_{4}\vert + \dots + \vert T_{11}\vert))\\
+&  = L(439 - 2\times 31 - 2\times 28 - 31 - 30 - \dots - 30)\\
+&  = L(439 - 423)\\
+&  = L(6)
 \end{split}
 \end{equation}.
+
+The idea is we can find any day of the month beyond 1 year provided we have the composition of days across months in one year due to the periodic property.
 

### 2019-12-05 main.Rmd
density estimation section

@@
 
+Density estimates are ueful in investigation of the properties of the measured variable.
+They can give valuable indication of features like skewness, multimodality or tail behavior.
 The measured variable can be dicrete or continuous.
-The probability distribution of these univariate measured variable could be studied by estimating the probability density function (pdf for continuous) or probability mass function (pmf for dicrete).
-Both pdf and pmf are concerned with point probabilities of the random variable.
+The probability distribution of these univariate measured variable could be studied by estimating the probability density function (continuous) or probability mass function (dicrete).
+The densities can either be estimated or summarized using empirical quantiles.
+The approach can be parametric or non-parametric.
+A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself.
+
+The most commonly used density estimation method is the kernel density estimator. The Kernel density estimator with Kernel K is defined by:
+
+$$\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h$$ , where $h$ is the smoothing parameter or the bandwidth.
+where $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from a continuous univariate distribution with probabilitty density function f(·).
+
+ <!-- Both pdf and pmf are concerned with point probabilities of the random variable.
 Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d.
-The densities can either be estimated or summarized using quantiles.Instead of assuming a parametric model for the distribution (e.g.
+-->
+
+<!-- Instead of assuming a parametric model for the distribution (e.g.
 a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g.
 differentiable).
-It is then possible to estimate the unknown density function f(·).
+It is then possible to estimate the u2nknown density function f(·).
+-->
 

### 2019-12-07 main.Rmd
the measured variable estimation section

@@
 We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
-However, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analyzed for different periodicities.
+Although, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analyzed for different periodicities.
 
@@
 
-## Density estimation of measured variable
-
-Density estimates are ueful in investigation of the properties of the measured variable.
+## Estimation of the measured variable
+
+Density estimates are useful in exploring the properties of the measured variable.
 They can give valuable indication of features like skewness, multimodality or tail behavior.
-The measured variable can be dicrete or continuous.
-The probability distribution of these univariate measured variable could be studied by estimating the probability density function (continuous) or probability mass function (dicrete).
-The densities can either be estimated or summarized using empirical quantiles.
+The measured variable in the considered data structure can be dicrete or continuous.
+The probability distribution of them could be studied by estimating the probability density function (continuous) or probability mass function (dicrete).
 The approach can be parametric or non-parametric.
 A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself.
-
-The most commonly used density estimation method is the kernel density estimator. The Kernel density estimator with Kernel K is defined by:
-
-$$\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h$$ , where $h$ is the smoothing parameter or the bandwidth.
-where $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from a continuous univariate distribution with probabilitty density function f(·).
+These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
+The entire distribution could be estimated using different density estimation methods or summarized using empirical quantiles.
+Several non-parametric density estimators and quantile estimators are proposed in the literature, the most commonly used form are as follows:
+
+<!-- The most commonly used density estimation method is the kernel density estimator.  -->
+
+If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$ and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$, the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by:
+
+\begin{equation}
+\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h
+\end{equation}
+
+where $h$ is the smoothing parameter or the bandwidth. The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by:
+
+<!-- $Q$, $(q \equiv  Q')$, which is the quantile density function -->
+<!-- $ Q_{n} (u) = \sum_{i=1}^ n X_{(i)} I((i-1)/n <u \leq i/n)$\\ -->
+<!--             $= X_(1) + \sum_{i=1}^ n (X_{(i)} - X_{(i-1)}) I((i-1)/n < u)$ -->
+
+<!-- and most widely used definition of sample quantile is given in  by: -->
+
+<!-- $$\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}$$ where,  $(j - m)/n \leq p < (j - m + 1)/n$,  -->
+<!-- for some $m \in R$ and $0 < \gamma < 1$.  $\hat{Qi(p)}$ is the ith sample quantile  -->
+
+\begin{equation}
+\hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u)
+\end{equation}
+
+where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$.
 
@@
 
-The histogram is the oldest and most popular density estimator method.
+<!-- The histogram is the oldest and most popular density estimator method.
 The form of the histogram depends very much on the tuning parameters class width and origin.
@@
 There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.
-
-The quantile of a distribution is defined as $Q(p)= F^{-1}(p) = inf{{x: F(x) >p}}$, $0<p< 1$, where $F(x)$ is the distribution function.
-Let $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$, and let $Qi(p)$ denote the ith sample quantile definition, then
+-->
+
+<!-- The inverse $Q = F^{-1}$, which is the quantile function can be defined as $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$.  -->
+
+The MSE of the estimators are given as follows:
 
 \begin{equation}
-\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}
+\begin{split}
+MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
+MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
+\end{split}
 \end{equation}
 
-where,  $(j - m)/n \leq p < (j - m + 1)/n$
-
-There have been several other non parametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
+where $\kappa = \int k^2(y)dy$.
+
+<!-- other non parametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
 Also sample quantiles can be of the form that are not defined for all p including hinges, letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
+-->
 
@@
 Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels.
+the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$, unless $h$ is chosen as a function of $n$.
+However, the variance of the $f(x)$ is $(1/nh)\int K(t)^2dt$ is a function of n.
 

### 2019-12-07 main.Rmd
estimation

@@
 
-\newtheorem{timedom}{Definition}[section]
-
-\begin{timedom}\label{def:timedom}
+\newtheorem{definition}{Definition}[section]
+
+\begin{definition}\label{def:definition}
 A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{def:linear}
+\begin{definition}\label{def:linear}
 A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
@@
   (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
-\end{timedom}
-
-\begin{timedom}\label{def:granule}
+\end{definition}
+
+\begin{definition}\label{def:granule}
 Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{def:finerthan}
+\begin{definition}\label{def:finerthan}
 A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
 for each index i, there exists an index j such that $G(i) \subset H(j)$.
-\end{timedom}.
-
-\begin{timedom}\label{def:groupsinto}
+\end{definition}.
+
+\begin{definition}\label{def:groupsinto}
 A linear granularity G groups into a linear granularity H, denoted
@@
 \end{equation}
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{def:periodical}
+\begin{definition}\label{def:periodical}
 A granularity H is periodical with respect to a granularity G if
@@
 $H (i + R) = \bigcup_{j \in S} G(j + P)$.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{def:quasi}
+\begin{definition}\label{def:quasi}
 A granularity H is quasi-periodical with respect to a granularity G if
@@
 between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{def:BG}
+\begin{definition}\label{def:BG}
 Bottom granularity - Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{def:circular}
-A circular granularity C(B, G) relates a linear granularity G to the bottom granularity B, if
-
-\begin{equation} \label{eq:eq2}
-C_{B, G}(z) = {L}(z\mod k(B,G)) \quad \forall z \in \mathbb{Z}^+
+\begin{definition}\label{def:circular}
+A circular granularity C(B, G) relates a linear granularity $G$ to the bottom granularity B, if
+
+\begin{equation} \label{eq:eq5}
+\begin{split}
+C_{B, G}(z) & = {L}(z\mod P(B, G)) \quad \forall z \in \mathbb{Z}^+ \\
+\end{split}
 \end{equation}
-where,\\
-z denotes the index set, \\
-B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity G with regular mapping, \\
-{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
-and k(B,G) is the number of granules of B in each granule of G.
-\end{timedom}
+where
+z denotes the index set,
+B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, 
+$P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$,
+and ${L}$ is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (P-1)}$,
+
+\end{definition}
 
@@
 However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
-Note that each
-circular granularity can use a different set of labels.
-Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
+Note that each circular granularity can use a different set of labels.
+Example showing circular granularity relating two linear granularities with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
 Each granule is represented by a box.
@@
 The index of the tsibble considered is days.
-G and H are two linear granularities denoting days and weeks respectively.
-$k(day, week) =  7$ and hence, $C_{G, B}(z)$ represented by C(BG, G) is given by $B(z\mod 7)$.
-The circular granularity C(B, G) consists of repetitive pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
-$L$ can be defined such that the labels corresponding to ${0,1,\dots,6}$ correspond to different days of the week ${Sun, Mon ,\dots, Sat}$ .
-Similarly, C(B, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14.
-C(B, H) also repeats the pattern ${L'(0), \dots, L'(13)}$ in every period.
-Hence, circular granularities are represented as repetitive categorization of time in the diagram.
-
-```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
-knitr::include_graphics("Figs/circular-gran-diagram.png")
+Note that each
+circular granularity can use different label mappings.
+In \autoref{fig:day-of-week}, Day and Week are two linear granularities and $k(Day, Week) =  7$.
+The circular granularity C(Day, Week) representing Day-of-Week will consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
+The label mapping $L$ can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sunday, Monday, \dots, Saturday)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({0, 1, \dots, 6)}$ depending on the context.
+
+<!-- # ```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "} -->
+<!-- # knitr::include_graphics("Figs/circular-gran-diagram.png") -->
+<!-- # ``` -->
+
+```{r day-of-week,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
+knitr::include_graphics("Figs/day-of-week.png")
 ```
@@
 
-\begin{timedom}\label{quasicircular}
+\begin{definition}\label{quasicircular}
 A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if
@@
 $\vert T_w \vert$ is the cardinality of set $T_w$.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{timedom}\label{aperiodic}
+\begin{definition}\label{aperiodic}
 An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by
@@
 $T_q$ are the sets of indices of BG describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
-\end{timedom}
+\end{definition}
 
@@
 
-\begin{equation}
+\begin{equation}\label{eq:MSE}
 \begin{split}
@@
 This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the visualizations that use kernel density estimates should be used with caution as sample size would directly affect both the variance of the estimator and the confidence intervals.
-Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels.
-the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$, unless $h$ is chosen as a function of $n$.
-However, the variance of the $f(x)$ is $(1/nh)\int K(t)^2dt$ is a function of n.
+In such cases, the visualizations that use density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators.
+\autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.
+
+<!-- the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$,  -->
+
+<!-- Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels. -->
 

### 2019-12-07 main.Rmd
circular definition typo correct

@@
 \begin{definition}\label{def:circular}
-A circular granularity C(B, G) relates a linear granularity $G$ to the bottom granularity B, if
+A circular granularity $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
 
@@
 $P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$,
-and ${L}$ is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (P-1)}$,
+and ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$.
 
@@
 Note that each circular granularity can use a different set of labels.
-Example showing circular granularity relating two linear granularities with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}.
+Example showing circular granularity relating two linear granularities with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-dow}.
 Each granule is represented by a box.
@@
 The index of the tsibble considered is days.
-Note that each
-circular granularity can use different label mappings.
-In \autoref{fig:day-of-week}, Day and Week are two linear granularities and $k(Day, Week) =  7$.
-The circular granularity C(Day, Week) representing Day-of-Week will consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
-The label mapping $L$ can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sunday, Monday, \dots, Saturday)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({0, 1, \dots, 6)}$ depending on the context.
+Note that each circular granularity can use different label mappings.
+Day and Week are two linear granularities with $P =  7$.
+The circular granularity C(Day, Week) representing Day-of-Week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
+The label mapping $L$ can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{Sunday, Monday, \dots, Saturday\}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{0, 1, \dots, 6\}$ depending on the context.
 
@@
 
-```{r day-of-week,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
-knitr::include_graphics("Figs/day-of-week.png")
-```
-
-All circular granularities can be expressed in terms of the bottom linear granularity with equation \autoref{eq:eq2}.
-
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/k(B,G) \rfloor\mod k(G,H))$, where linear granularity H is periodic with respect to linear granularity G and k(G, H) represents the number of granules of G in one granule of H.
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
+knitr::include_graphics("Figs/circular-dow.png")
+```
+
+<!-- All circular granularities can be expressed in terms of the bottom linear granularity with equation \autoref{eq:eq2}. -->
+
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity H is periodic with respect to linear granularity G with regular mapping such that the number of granules of G in each granule of H is $P(G,H)$.
 
@@
 But the representation of the resultant circular granularity will be a function of the index set.
-Suppose the bottom granularity is minutes, $k_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$ and $L_i$ is the associated label mapping.
+Suppose the bottom granularity is minutes, $P_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$ and $L_i$ is the associated label mapping.
 
@@
 \toprule
-Minute-of-Hour: & $C_1(z)$ & $= L_1(z \mod 60)$ & $k_1 =$&$60$ \\
-Minute-of-Day: & $C_2(z)$ & $= L_2(z \mod 60*24)$ & $k_2=$&$`r 60*24`$\\
-Hour-of-Day: & $C_4(z)$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $k_3 =$&$24$ \\
-Hour-of-Week: & $C_6(z)$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $k_4=$&$`r 24*7`$\\
-Day-of-Week: & $C_7(z)$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $k_5=$&$7$\\
+Minute-of-Hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
+Minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
+Hour-of-Day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
+Hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
+Day-of-Week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
 \bottomrule
@@
 A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
-For example, \autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) and plausible choices of number of granules ($k_i$) of the finer granularity inside each granule of the coarser one in the cyclic granularity ($Q_i$).
+However, they are still formed with linear granularities, one of which "groups periodically into" the other.
+\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) and plausible choices of number of granules of the finer granularity inside each granule of the coarser one in the cyclic granularity ($Q_i$) is ($P_i$).
 
@@
 
-Day-of-Month: & $Q_1(z)$ && $k_1=31, 30, 29, 28$\\
-Hour-of-Month: & $Q_2(z)$ && $k_2=24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $Q_3(z)$ && $k_3=366, 365$\\
-Week-of-Month: & $Q_4(z)$ && $k_4=5, 4$\\ 
+Day-of-Month: & $Q_1(z)$ && $P_1=31, 30, 29, 28$\\
+Hour-of-Month: & $Q_2(z)$ && $P_2=24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $Q_3(z)$ && $P_3=366, 365$\\
+Week-of-Month: & $Q_4(z)$ && $P_4=5, 4$\\ 
 \bottomrule
@@
 \begin{definition}\label{quasicircular}
-A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if
-
-\begin{equation}\label{eq:eq3}
+A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if
+
+\begin{equation}\label{eq:eq6}
 Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
@@
 
-where,  
-$z$ denotes the index set,\\
-$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping,\\
-{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
-$T_w$ are the sets of indices of B describing $G'(w)$ such that 
+where
+$z$ denotes the index set,
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, ${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (k-1)\}$, $T_w$ are the sets of indices of B describing $G'(w)$ such that
 \begin{equation}
@@
 \end{equation}
-$T_w = 0$ for $w \epsilon Z_{<0}$,\\
-$\vert T_w \vert$ is the cardinality of set $T_w$.
+$T_w = 0 \quad for \quad w\epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{definition}
 
-Since G' is periodical with respect to BG, from definition \ref{def:periodical} the following holds true:
-
-- $BG \trianglelefteq G'$
-- $G'(w) = \bigcup_{z \in T_w \mod R'}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the grouping $(BG, G')$.
+Since G' is periodical with respect to BG, from definition \ref{def:periodical} the following still holds true:
+
+- $B \trianglelefteq G'$
+- there exist $R', P' \epsilon Z+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ and $G(w + R') \neq \phi$ then $G'(w + R') = \bigcup_{z \in T_w}B(z + P')$ and hence $G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the grouping $(B, G')$.
 

### 2019-12-07 main.Rmd
quasi circular done

@@
 \begin{split}
-C_{B, G}(z) & = {L}(z\mod P(B, G)) \quad \forall z \in \mathbb{Z}^+ \\
+C_{B, G}(z) & = {L}(z\mod P(B, G)) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
 \end{split}
@@
 
-**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity.
-However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
-Note that each circular granularity can use a different set of labels.
-Example showing circular granularity relating two linear granularities with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-dow}.
+**Example**:  Example showing circular granularity relating two linear granularities "Day" and "Week" is visually depicted in a series of slots in the \autoref{fig:circular-dow}.
 Each granule is represented by a box.
 The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set and the bottom granularity.
+The first slot in the diagram shows the index set.
 The index of the tsibble considered is days.
-Note that each circular granularity can use different label mappings.
 Day and Week are two linear granularities with $P =  7$.
-The circular granularity C(Day, Week) representing Day-of-Week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
-The label mapping $L$ can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{Sunday, Monday, \dots, Saturday\}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{0, 1, \dots, 6\}$ depending on the context.
+The circular granularity $C_{Day, Week}$ representing Day-of-Week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
+
+**Discussion**: Note that each circular granularity can use different label mappings.
+In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity.
+However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
+Hence, the label mapping $L$ in \autoref{fig:circular-dow} can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{Sunday, Monday, \dots, Saturday\}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{0, 1, \dots, 6\}$ depending on the context.
 
@@
 In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity H is periodic with respect to linear granularity G with regular mapping such that the number of granules of G in each granule of H is $P(G,H)$.
-
-Table \ref{tab:definitions} shows representation of circular granularities relating two linear granularities.
-It is possible that none of these two linear granularities are bottom granularities.
+Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
+It is possible that none of the two linear granularities are bottom granularities.
 But the representation of the resultant circular granularity will be a function of the index set.
-Suppose the bottom granularity is minutes, $P_i$ is the number of granules of the finer granularity in each granule of the coarser granularity in the circular granularity $C_i$ and $L_i$ is the associated label mapping.
+It is assumed that the bottom granularity is minutes.
 
@@
 However, they are still formed with linear granularities, one of which "groups periodically into" the other.
-\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) and plausible choices of number of granules of the finer granularity inside each granule of the coarser one in the cyclic granularity ($Q_i$) is ($P_i$).
+\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) with ($P_i$) denoting the plausible choices of number of granules of the finer granularity inside each granule of the coarser one.
 
@@
 
-Day-of-Month: & $Q_1(z)$ && $P_1=31, 30, 29, 28$\\
-Hour-of-Month: & $Q_2(z)$ && $P_2=24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $Q_3(z)$ && $P_3=366, 365$\\
-Week-of-Month: & $Q_4(z)$ && $P_4=5, 4$\\ 
+Day-of-Month: & $Q_1$ && $P_1 = 31, 30, 29, 28$\\
+Hour-of-Month: & $Q_2$ && $P_2 = 24*31, 24*30, 24*29, 24*28$\\
+Day-of-Year: & $Q_3$ && $P_3 = 366, 365$\\
+Week-of-Month: & $Q_4$ && $P_4 = 5, 4$\\ 
 \bottomrule
@@
 
-\begin{definition}\label{quasicircular}
+\begin{definition}\label{def:quasicircular}
 A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if
 
-\begin{equation}\label{eq:eq6}
+\begin{equation}\label{eq:quasi}
 Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
@@
 where
-$z$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, ${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (k-1)\}$, $T_w$ are the sets of indices of B describing $G'(w)$ such that
-\begin{equation}
-G'(w) = \bigcup_{z \in T_w}B(z),\\
-\end{equation}
-$T_w = 0 \quad for \quad w\epsilon Z_{<0}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ denote the period of the grouping $(B, G')$ and the number of granules of $G'$ in each of these periods, ${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$, $T_w$ are the sets of indices of $B$s describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{definition}
 
-Since G' is periodical with respect to BG, from definition \ref{def:periodical} the following still holds true:
-
-- $B \trianglelefteq G'$
-- there exist $R', P' \epsilon Z+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ and $G(w + R') \neq \phi$ then $G'(w + R') = \bigcup_{z \in T_w}B(z + P')$ and hence $G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the grouping $(B, G')$.
-
-**Discussion**: According to the first condition, assume, for instance, that G'(w) is the union of the granules G(a1), G(a2), .
-.
-.
-, G(ak).
-The periodicity property (condition (2) in the definition) ensures that the Rth granule after H(i), i.e., H(i + R), if non-empty, is the union of G(a1 + P ), G(a2 + P), .
-.
-.
-, G(ak + P).
-
-Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-gran}.
+**Example**: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
 Each granule is represented by a box.
-Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity.
-Two linear granularities AG and AH are considered, where $G' = Alter^{2}_{(1,-1)}(BG, Group_3(BG))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(BG))$.
-This implies that G' is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and H' is made up of shrinking every 1st granule of Group_3(BG) by 2 granules.
-Number of granules of G' and H' in each period of BG is 2, but the number of granules of BG in each of those granules are different.
-QC(BG, G') and QC(BG, H') are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G' or H'.
-For G', R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}.
-For H', R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
+Two linear granularities $G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))$ are considered.
+This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
+Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.
+$Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
+For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
+For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
 
@@
 \begin{split}
-QC(BG,G')(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
-  & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
-  & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
-  & = BG(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
-  & = BG(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
-  & = BG(8 - 2*2 - 3) \\
-  & = BG(1) \\
+Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
+  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
+  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
+  & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
+  & = L(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
+  & = L(8 - 2*2 - 3) \\
+  & = L(1) \\
 \end{split}
@@
 \begin{split}
-QC(BG,H')(10) & = BG(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
-  & = BG(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
-  & = BG(10 - \vert T_{0 \mod 2}\vert) \\
-  & = BG(10 - \vert T_{0}\vert) \\
-  & = BG(10 - 7) \\
-  & = BG(3) \\
+Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
+  & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
+  & = L(10 - \vert T_{0 \mod 2}\vert) \\
+  & = L(10 - \vert T_{0}\vert) \\
+  & = L(10 - 7) \\
+  & = L(3) \\
 \end{split}
@@
 
-```{r quasi-gran, echo=FALSE,out.width = "100%",fig.cap="Relating quasi-circular granularities and bottom linear granularity "}
-knitr::include_graphics("Figs/aperiodic-granv1.png")
-```
-
-If we ignore leap years and all its exception, the periodicity of the grouping (days, months), P' = 365, R' = 12, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
+**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, from definition \ref{def:periodical} the following holds true:
+
+- $B \trianglelefteq G'$
+- there exist $R', P' \epsilon Z_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
+- number of granules of $B$ in each granule of $G'$ is not a constant
+
+Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
+The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
+The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R)^{th}$ granule of $G'$ will have the same number of granules of $B$.
+The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$.
+Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in in \autoref{def:quasicircular}.
+
+```{r quasi-circular-example, echo=FALSE,out.width = "100%",fig.cap="Relating quasi-circular granularities and bottom linear granularity "}
+knitr::include_graphics("Figs/quasi-circular-example.png")
+```
+
+<!-- If we ignore leap years and all its exception, the periodicity of the grouping (days, months), P' = 365, R' = 12, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
 Considering leap years and all their exceptions, period is 400 years and hence $P' = (3*365 + 366)$, $R' = 12*400$.
 \autoref{eq:eq3} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
+-->
 
@@
 
-The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity.
+<!-- The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity.
 The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
-It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
+-->
+
+It can be noted here that if a linear granularity $G'$ is quasi-periodic with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
 
 \begin{equation}\label{eq6}
-A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
+A_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
 \end{equation}
@@
 All of these are recurring events, but with non-periodic patterns.
-As such, plausible $k_i$ from \autoref{tab:quasi} can be infinite for aperiodic granularities.
+As such, plausible $P_i$ from \autoref{tab:quasi} can be infinite for aperiodic granularities.
+
+Aperiodic cyclic granularities are defined using aperiodic linear granularities.
+Consider n aperiodic linear granularity $A_i$ such that $B \trianglelefteq A_i, \forall i\in{1, 2, \dots, n}$.
+Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i, j\}}$ of the integers such
+that
+\begin{equation*}
+A_(j) = \bigcup_{z \in T_{\{i, j\}}}B(z)
+\end{equation*}
 
 \begin{definition}\label{aperiodic}
-An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by
-
-\begin{equation}\label{eq:eq}
+An aperiodic granularity $AX_{B, G''}$ that relates linear granularities $G''$ and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by
+
+\begin{equation}\label{eq:eq7}
 \begin{split}
-A_{BG, G''}(z) & = 1, z \in T_q\\
-     & = 0, z \not\in T_q\\
+A_{B, G''}(z) & = L(i) \quad z \in T_q(i) \quad and \quad \forall i \in E \\
+     & = L(0) otherwise
 \end{split}     
@@
 where, 
-z denotes the index set (index of BG),
-$T_q$ are the sets of indices of BG describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
+$z$ denotes the index set,
+$T_q$ are the sets of indices of $B$ describing $G''(q)$ such that $$G''(q) = \bigcup_{z \in T_q}B(z)$$ and $T_q(z) = \bigcup_{i \in E} T_q(i)$, E denotes the set of linear aperiodic granularities.
 \end{definition}
 
-Discussion: Aperiodic time granularities map a time index to a categorical variable containing binary values.
+Discussion: Aperiodic cyclic time granularities relate many aperiodic linear granularities to the bottom granularity.
+The number of categories obtained coincides with the number of aperiodic linear granularities considered.
+
+map a time index to a categorical variable containing binary values.
 It takes the value 1 whenever the time index coincides with the events in context (e.g.
@@
 Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector.
-An example of aperiodic granularity can be one that relates a bottom granularity ($BG$) and linear granularity ($G''$) such that $BG \trianglelefteq G''$, $BG \not\preceq G$ and $BG$ is not periodic or quasi-periodic with respect to $G''$, as shown in \autoref{fig:aperiodic}.
+An example of aperiodic granularity can be one that relates a bottom granularity ($B$) and linear granularity ($G''$) such that $BG \trianglelefteq G''$, $BG \not\preceq G$ and $B$ is not periodic or quasi-periodic with respect to $G''$, as shown in \autoref{fig:aperiodic}.
 

### 2019-12-07 main.Rmd
aperiodic granularity done

@@
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularity $A_i$ such that $B \trianglelefteq A_i, \forall i\in{1, 2, \dots, n}$.
-Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i, j\}}$ of the integers such
-that
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ none of which can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
+Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
+Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that
 \begin{equation*}
-A_(j) = \bigcup_{z \in T_{\{i, j\}}}B(z)
+M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)
 \end{equation*}
+and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
 
 \begin{definition}\label{aperiodic}
-An aperiodic granularity $AX_{B, G''}$ that relates linear granularities $G''$ and bottom granularity $B$, with $G''$ being aperiodical with respect to $B$, is given by
-
-\begin{equation}\label{eq:eq7}
+An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
+
+\begin{equation}\label{eq:aperiodic}
 \begin{split}
-A_{B, G''}(z) & = L(i) \quad z \in T_q(i) \quad and \quad \forall i \in E \\
-     & = L(0) otherwise
+A_{B, M}(z) & = L(i), \quad z \in T_{i_j} \quad and \quad \forall i \in {1, 2, \dots, n}\\
+     & = L(0), \quad otherwise
 \end{split}     
@@
 where, 
-$z$ denotes the index set,
-$T_q$ are the sets of indices of $B$ describing $G''(q)$ such that $$G''(q) = \bigcup_{z \in T_q}B(z)$$ and $T_q(z) = \bigcup_{i \in E} T_q(i)$, E denotes the set of linear aperiodic granularities.
+$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
+$T_{i_j}$ are the sets of indices of $B$ describing $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and $M = \bigcup_{j=1}^{n}M_{j}$.
 \end{definition}
 
-Discussion: Aperiodic cyclic time granularities relate many aperiodic linear granularities to the bottom granularity.
-The number of categories obtained coincides with the number of aperiodic linear granularities considered.
-
-map a time index to a categorical variable containing binary values.
+**Example**: \autoref{fig:aperiodic-example} shows two aperiodic linear granularity $M_1$ and $M_2$.
+Each granule is represented by a box.
+The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
+The first slot in the diagram shows the index set and A_{B, M} consists of three categories, one corresponding to the first aperiodic event $M_1$, the second corresponding to aperiodic event $M_2$ and the third corresponding to no event/aperiodic linear granularity.
+
+**Discussion**: The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
+The index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
+
+<!-- map a time index to a categorical variable containing binary values.
 It takes the value 1 whenever the time index coincides with the events in context (e.g.
@@
 Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector.
-An example of aperiodic granularity can be one that relates a bottom granularity ($B$) and linear granularity ($G''$) such that $BG \trianglelefteq G''$, $BG \not\preceq G$ and $B$ is not periodic or quasi-periodic with respect to $G''$, as shown in \autoref{fig:aperiodic}.
-
-```{r aperiodic, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
-knitr::include_graphics("Figs/aperiodic.png")
+-->
+
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
+knitr::include_graphics("Figs/aperiodic-example.png")
 ```

### 2019-12-07 main.Rmd
cyclic gran section full update

@@
 The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set and A_{B, M} consists of three categories, one corresponding to the first aperiodic event $M_1$, the second corresponding to aperiodic event $M_2$ and the third corresponding to no event/aperiodic linear granularity.
+The first slot in the diagram shows the index set and $A_{B, M}$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
 

### 2019-12-08 main.Rmd
cyclic calendar algebra update

@@
 
-# Computing cyclic time granularities {#sec:order}
-
+# Cyclic calendar algebra {#sec:cyclic-calendar}
+
+In \autoref{sec:}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed.
+In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
+This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
+Since, our method caters to computation of cyclic granularities, we shall refer to it as "cyclic calendar algebra".
+
+The cyclic calendar algebra consists of two kinds of operations: one that requires computing multiple order-up from single order-up cyclic granularities and the other that require the reverse.
 The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
+We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities.
 We shall use the notion of a hierarchy table and order to define them.
 
-**Definition:  Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
-If we consider two linear granularities G and H, such that G is finer than or groups into H (by Definitions \ref{def:finerthan}, \ref{def:groupsinto}), then H is of higher order than G.
-
-**Notation:  Hierarchy table** Let $H_n: (G, C, k)$ be a hierarchy model containing n linear granularities.
-$G_{l}$ represents the linear granularity of order l and $k(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
-
-In a hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
+**Notation:  Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
+$G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
+**Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
+For example, if we consider two linear granularities $G$ and $H$, such that $G$ is finer than or groups into $H$, then $H$ is of higher order than $G$.
+In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
+Let us look at some examples of single and multiple order-up granularities.
 
@@
 
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities.
-Let us look at few calendars to see examples of single and multiple order-up granularities.
-
 <!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
 -->
+
+<!-- some discussions on primary units of calendar -->
+**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar.
+But it is far from being the only one.
+All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars.
+Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
+The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
+Nevertheless, for any calendar a hierarchy can be defined.
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows [@Reingold2001-kf]:
+
+ - 1 kin = 1 day
+ - 1 uinal = 20 kin
+ - 1 tun  = 18 uinal
+ - 1 katun = 20 tun
+ - 1 baktun = 20 katun
+
+Thus, the hierarchy table for the Mayan calendar would look like the following:
+
+|  G	        |C                 |P
+|------------	|------------------|------------|
+|  kin      	|kin-of-uinal      |  20        |
+|  uinal    	|uinal-of-tun      |  18  	    |
+|  tun      	|tun-of-katun      |  20	      |
+|  katun    	|katun-of-baktun   |  20        |
+|  baktun   	| 1                |	1         |
+
+Examples of multiple-order-up  granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.
 
@@
 \end{equation}
-
-<!-- some discussions on primary units of calendar -->
-**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar.
-But it is far from being the only one.
-All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars.
-Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
-Nevertheless, for any calendar a hierarchy can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows [@Reingold2001-kf]:
-
- - 1 kin = 1 day
- - 1 uinal = 20 kin
- - 1 tun  = 18 uinal
- - 1 katun = 20 tun
- - 1 baktun = 20 katun
-
-Thus, the hierarchy table for the Mayan calendar would look like the following:
-
-|  G	        |C                 |k
-|------------	|------------------|------------|
-|  kin      	|kin-of-uinal      |  20        |
-|  uinal    	|uinal-of-tun      |  18  	    |
-|  tun      	|tun-of-katun      |  20	      |
-|  katun    	|katun-of-baktun   |  20        |
-|  baktun   	| 1                |	1         |
-
-Examples of multiple-order-up  granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.
 

### 2019-12-08 main.Rmd
single to multiple all circular

@@
 
-# Linear Time granularities {#sec:lineartime}
+# Linear time granularities {#sec:lineartime}
 
@@
 
-# Cyclic Time granularities {#sec:cyclic-gran-def}
+# Cyclic time granularities {#sec:cyclic-gran-def}
 
@@
 
-The cyclic calendar algebra consists of two kinds of operations: one that requires computing multiple order-up from single order-up cyclic granularities and the other that require the reverse.
+The cyclic calendar algebra consists of two kinds of operations:  
+(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities  
+(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities
+
 The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
-We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities.
 We shall use the notion of a hierarchy table and order to define them.
 
-**Notation:  Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
+**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
 $G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
+
 **Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
@@
 In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
-Let us look at some examples of single and multiple order-up granularities.
+
+We refer to granularities which are nested within multiple levels of the hierarchy table as multiple-order-up granularities and those concerning a single level as single-order-up granularities.
 
@@
 <!-- some discussions on primary units of calendar -->
-**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar.
+**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
 But it is far from being the only one.
-All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars.
+All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
@@
 Nevertheless, for any calendar a hierarchy can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows [@Reingold2001-kf]:
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
 
@@
 
-## Multiple-order from single-order {#sec:multiplefromsingle}
+## Single-to-multiple {#sec:single-to-multiple}
+
+### All circular single order-up granularities
+
+Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity.
+Since, the operation requires the use of modular arithmetic, it is important that the label mapping of each circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
+The label mapping of the resultant multiple-order-up granularity can however be chosen arbitrarily, depending on the context.
 
@@
 \begin{split}
-C_{(G_l,G_m)}(z) & = C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1) \\  
-  & =   C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1] \\
-  & =  C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1)\\
-  & =  C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1)\\
+C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1)) \\  
+  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1]) \\
+  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1))\\
+  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1))\\
   &\vdots\\
-  & = \sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1)\\
+  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1))\\
 \end{split}
@@
 
+*Examplle*
 Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic and regular.
@@
 
-## Single-order from multiple-order {#sec:singlefrommultiple}
+### Circular or aperiodic single order-up granularities
+
+## multiple-to-single {#sec:multiple-to-single}
 

### 2019-12-08 main.Rmd
computation section single-to-multiple update

@@
 
-### All circular single order-up granularities
+### All circular single order-up granularities {#sec:all-circular}
 
@@
 
-*Examplle*
-Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic and regular.
+*Example:*Let us use the equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar..
 
@@
 \begin{split}
-C(uinal, baktun) & = C(uinal, tun) + kuinal, tun)C(tun,katun) + C(uinal, katun)C(katun, baktun) \\
-              & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\
+C_{uinal, baktun}(z) & = L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
+              & = L(\lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
 \end{split}
@@
 
-### Circular or aperiodic single order-up granularities
-
-## multiple-to-single {#sec:multiple-to-single}
-
-For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
-
-\begin{equation} \label{eq9}
-C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
-\end{equation}
-
-**Example:** Considering the same example of Mayan Calendar,  it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:eq5}
-
-\begin{equation} \label{eq10}
+### Circular or quasi-circular single order-up granularities
+
+Let us revisit Gregorian calendar for addressing this case.
+Suppose we have a hierarchy table using some linear granularties from Gregorian calendar.
+Since months consists of unequal number of days, any linear granularity unit with higher order than months will also have unequal number of days.
+This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities.
+The single-order-up granularity day_month is quasi-circular.
+Any single-order-up granularities which are formed by linear granularities below days are circular.
+Similarly, all single-order-up granularities which are formed using linear granularities with orders higher than months are also circular.
+
+<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
+However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
+-->
+
+|  G	        |C                 |P
+|------------	|------------------|----------------|
+|  minute     |minute-of-hour    |  60            |
+|  hour    	  |hour-of-day       |  24  	        |
+|  day      	|day-of-month      |"quasi-circular"|
+|  month    	|month-of-year     |  12            |
+|  year   	  | 1                |	1             |
+
+There can be three scenarios for obtaining multiple-order-granularities here:
+- cyclic granularities relating two linear granularities whose orders are less than day  
+- cyclic granularities relating two linear granularities whose orders are more than month  
+- granularities relating two linear granularities with order at most day and another with order at least month
+
+The multiple order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:singlefrommultiple} and \autoref{sec:all-circular}.
+The calendar categorization resulting from the last case are quasi-circular.
+Examples might include day of the quarter or hour of the month.
+Let us consider the computation of $A_{hour, month}$ given single order-up granularities $A_{hour, day} = 12$, $A_{day, month} = 20$ and $C_{month, year} = 9$.
+
+\begin{equation} \label{eq:hour-month}
 \begin{split}
-C(tun, katun) =  \lfloor C(uinal, baktun)(z)/18\rfloor \mod 20 \\
+C_{hour, month}(z) & = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))
 \end{split}
@@
 
-In \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}, we assume that all single order granularities are circular.What happens when we have a mix of circular and aperiodic granularities.
-
-Let us turn to Gregorian calendar for addressing this case.
-Suppose we have a hierarchy table using some linear granularties from Gregorian calendar.
-Since months consists of unequal number of days, any temporal unit which is higher in order than months will also have unequal number of days.
-This is an example of a hierarchy structure which has both circular and aperiodic single-order-up granularities.
-The single-order-up granularity day_month is aperiodic.
-Any single-order-up granularities which are formed by units below days are periodic.
-Similarly, all single-order-up granularities which are formed using units whose orders are higher than months are also periodic.
-
-<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
-However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
--->
-
-|  G	        |C/A               |k
-|------------	|------------------|------------|
-|  minute     |minute-of-hour    |  60        |
-|  hour    	  |hour-of-day       |  24  	    |
-|  day      	|day-of-month      | "aperiodic"|
-|  month    	|month-of-year     |  12        |
-|  year   	  | 1                |	1         |
-
-There can be three scenarios for obtaining multiple-order-granularities here:
-- granularities relating two linear granularities whose orders are less than day
-- granularities relating two linear granularities whose orders are more than month
-- granularities relating two linear granularities with order at most day and another with order at least month
-
-The single order-up granularites resulting from the first two cases are periodic and has been handled in \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}.
-The calendar categorization resulting from the last case are aperiodic.
-Examples might include day of the quarter or hour of the month.
-Let us consider the computation of $A_{hour, month}$ given single order-up granularities $A_{hour, day} = 12$, $A_{day, month} = 20$ and $C_{month, year} = 9$.
+\begin{equation} \label{eq:day-year}
+\begin{split}
+C_{day, year}(z) & = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}-1}(\vert T_{w} \vert))
+\end{split}
+\end{equation}
+
+where, $T_w$ are the sets of indices of linear granularity day such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 
@@
 A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included}
+
+## multiple-to-single {#sec:multiple-to-single}
+
+For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
+
+\begin{equation} \label{eq9}
+C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
+\end{equation}
+
+**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:eq5}
+
+\begin{equation} \label{eq10}
+\begin{split}
+C_{tun, katun}(z) =  L(\lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20) 
+\end{split}
+\end{equation}
+
+In \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}, we assume that all single order granularities are circular.
+What happens when we have a mix of circular and aperiodic granularities.
 
@@
 # Discussion {#sec:discussion}
+
+Exploratory data analysis is iterative, finding and summarizing patterns and then probing more deeply.
+Some guidelines and best practices to explore without wandering aimlessly can save
 

### 2019-12-08 main.Rmd
intro edit for section summary

@@
 For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
-This approach requires deconstructing time in various possible ways called time granularities[@aigner2011visualization].
+This approach requires deconstructing time in various possible ways called time granularities [@aigner2011visualization].
 It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
@@
 
-Calendar-based graphics[@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
+Calendar-based graphics [@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
 Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time.
@@
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-One of the first attempts of characterizing these granularities occur in [@Bettini1998-ed].
+One of the first attempts of characterizing these granularities occur in @Bettini1998-ed.
 The definitions and rules defined are inadequate to reflect periodicities in time.
 Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior.
-Cyclic time granularities can be circular or aperiodic depending on if mappings between granularities are regular or irregular.
-Examples of circular can be hour of the day, day of the week and that of aperiodic granularities can be day of the month or public holidays.
+Cyclic time granularities can be circular, quasi-circular or aperiodic.
+Examples of circular can be hour of the day, day of the week, that of quasi-circular can be day of the month and examples of aperiodic granularities can be public or school holidays.
 Time deconstructions can also be based on the hierarchical structure of time.
@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates [@2012-la],  of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates (see @2012-la) of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
@@
 
-  * Facilitate manipulation of single and multiple order-up time granularities
-
-  * Checking feasibility of creating plots or drawing inferences from any two cyclic granularities
+  * Facilitate manipulation of single and multiple order-up time granularities through cyclic calendar algebra
+
+  * Checking feasibility of creating plots or drawing inferences for any two cyclic granularities
 
@@
 The remainder of the paper is organized as follows.
-Section \ref{sec:timegrandef} details the background of linear granularities in depth and introduces the characterization of circular and aperiodic time granularities.
-Section \ref{sec:synergy} discusses the effect in data structure due to how pairs of cyclic time granularities are related.
-Section \ref{sec:visualization} discusses the role of plot choices, synergy of time granularities and number of observations to construct an informative and trustworthy visualization.
-Section \ref{sec:application} discusses how systematic exploration can be carried out for a temporal and non-temporal case.
+\autoref{sec:lineartime} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
+\autoref{sec:cyclic-gran-def} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
+Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
+\autoref{sec:data-structure} discusses the data structure for exploring distribution of an observed variable across pair of cyclic time granularities.
+\autoref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
+Section \ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application.
 Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.
@@
 
-In \autoref{sec:}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed.
+In \autoref{sec:cyclic-gran-def}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
@@
 
-The multiple order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:singlefrommultiple} and \autoref{sec:all-circular}.
-The calendar categorization resulting from the last case are quasi-circular.
-Examples might include day of the quarter or hour of the month.
-Let us consider the computation of $A_{hour, month}$ given single order-up granularities $A_{hour, day} = 12$, $A_{day, month} = 20$ and $C_{month, year} = 9$.
+The multiple order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:all-circular}.
+Any cyclic granularity resulting from the last case are quasi-circular.
+Examples might include hour-of-month or day-of-year.
+Let us consider the computation of $C_{hour, month}(z)$ for $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$.
 
@@
 \begin{split}
-C_{hour, month}(z) & = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))
+C_{hour, month}(z) & = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))\\
+& = L(12 + 24*20)\\
 \end{split}
@@
 \begin{split}
-C_{day, year}(z) & = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}-1}(\vert T_{w} \vert))
+C_{day, year}(z) & = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert))\\
 \end{split}
@@
 
-\begin{equation} \label{eq11}
+Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
+$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
+However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
+It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$.
+To keep it simple, we include only the case where there is just one single order-up quasi-circular granularity in the hierarchy table and any multiple order-up quasi-circular granularity need to be expressed with single order-up circular granularities and a quasi-circular granularity, that is not necessarily single order-up.
+
+\begin{equation} \label{eq:multifromsingle-quasi}
 \begin{split}
-A_{(hour, month)} & = (\sum_{w=0}^{9 - 1}(\vert T(w)\vert) +  20)/24 + 12 \\
-                  & = (\sum_{w=0}^{C_{month, year} - 1}(\vert T(w)\vert) +  A_{day, month})*k(hour_day) + A_{hour, day} \\
-                  & = (A{day_year} - 1)*k(hour_day) + A_{hour, day} \\
+C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)) \\ 
+& = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1)) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)\\
 \end{split}
 \end{equation}
-where, $month(w) = \bigcup_{z \in T_w}day(z)$
-
-Clearly, the computation of multiple order-up granularities become more involved with aperiodic mappings of linear granularities in the calendar structure.
-With one aperiodic granularity, we can write the following:
-
-A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included}
-
-## multiple-to-single {#sec:multiple-to-single}
+
+where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping.
+
+<!-- The aperiodic granularity $C_{day, quarter}(z))$ in \autoref{eq:hour-month} can be broken further. -->
+
+<!-- \begin{equation} \label{eq11} -->
+<!-- \begin{split} -->
+<!-- A_{(hour, month)} & = (\sum_{w=0}^{9 - 1}(\vert T(w)\vert) +  20)/24 + 12 \\ -->
+<!--                   & = (\sum_{w=0}^{C_{month, year} - 1}(\vert T(w)\vert) +  A_{day, month})*k(hour_day) + A_{hour, day} \\ -->
+<!--                   & = (A{day_year} - 1)*k(hour_day) + A_{hour, day} \\ -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
+<!-- where, $month(w) = \bigcup_{z \in T_w}day(z)$ -->
+
+<!--  With one aperiodic granularity, we can write the following:  -->
+<!-- A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included} -->
+
+## Multiple-to-single {#sec:multiple-to-single}
+
+### Muliple order-up circular granularities {#sec:all-circular-multiple}
 
@@
 
-\begin{equation} \label{eq9}
+\begin{equation} \label{eq:all-circular-multiple}
 C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
@@
 
-**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:eq5}
+**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}
 
@@
 
-In \autoref{sec:singlefrommultiple} and \autoref{sec:multiplefromsingle}, we assume that all single order granularities are circular.
-What happens when we have a mix of circular and aperiodic granularities.
+### Muliple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
+
+<!-- \begin{equation} \label{eq:quasi-circular-multiple} -->
+<!-- C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod P(m_1, m_2)) -->
+<!-- \end{equation} -->
+
+The representation of single order-up quasi-circular granularities using multiple order-up quasi-circular granularities is not discussed in this paper.
 
@@
 
-# Data structure
+# Data structure{#sec:data-structure}
 
@@
 # Acknowledgements {-}
+
+The authors would like to thank Prof. Antony Unwin for his immense inpu
 

### 2019-12-08 main.Rmd
Discussion section updated

@@
 
-The representation of single order-up quasi-circular granularities using multiple order-up quasi-circular granularities is not discussed in this paper.
+The representation of single order-up quasi-circular granularities using multiple order-up quasi-circular granularities is not discussed in this paper and is left for future work.
 
@@
 
-Exploratory data analysis is iterative, finding and summarizing patterns and then probing more deeply.
+Exploratory data analysis is iteratively finding and summarizing patterns.
+With temporal data available at more finer scales, exploring periodicity become overwhelming with so many possible cyclic granularities that could be generated for a context.
+This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by choosing a list of harmonies that need to be plotted together given a hierarchy or calendar structure and recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+
+The infrastructure of systematic exploration of observed variable across pair of cyclic time granularities has been implemented in the
+R package `gravitas`.
+A missing piece of the gravitas package is to enable user-defined temporal calendars and compute multiple order-up quasi-circular granularities from the hierarchy table.
+Also, computation of cyclic aperiodic granularities would require computing aperiodic linear granularities first.
+A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
+These functionalities can be imported in the gravitas package to accomodate for aperiodic cyclic granularities.
+
 Some guidelines and best practices to explore without wandering aimlessly can save

### 2019-12-08 main.Rmd
discussion change

@@
   affiliation: Department of Econometrics and Business Statistics, Monash University
+
+- name: Antony Unwin
+  affiliation: University of Augsburg
 
@@
 We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations.
+There are several classes of time deconstructions including linear and cyclic time granularities.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week.
-Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.
+Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month and aperiodic such as public or school holidays.
 
@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2[@Wickham2009-pk].
+This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009-pk].
 bibliography: [bibliography.bib]
@@
 Exploratory data analysis is iteratively finding and summarizing patterns.
-With temporal data available at more finer scales, exploring periodicity become overwhelming with so many possible cyclic granularities that could be generated for a context.
-This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by choosing a list of harmonies that need to be plotted together given a hierarchy or calendar structure and recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context.
+This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and therby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
 
@@
 
-Some guidelines and best practices to explore without wandering aimlessly can save
-
 # Acknowledgements {-}
 
-The authors would like to thank Prof. Antony Unwin for his immense inpu
+The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages.
+This manuscipt was created with knitr [@knitr] and R Markdown [@rmarkdown].
+The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all required materials required to reproduce this article.
 
 # Bibliography
-
-<!-- # ```{r write-bib} -->
-<!-- # write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
-<!-- # ``` -->
-
-<!-- library(gravitas) -->
-<!-- library(tsibble) -->
-
-<!--  cricket_bowling <- cricket_season %>% -->
-<!--    mutate(dot_ball_proxy = if_else(total_runs==0,1,0), -->
-<!--           wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% -->
-<!--   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
-<!--    summarise(dot_balls = if_else(sum(dot_ball_proxy)>0, "yes","no"), -->
-<!--              wickets = if_else(sum(wicket_proxy)>0, "yes","no"),  -->
-<!--             total_runs = sum(total_runs)) %>% -->
-<!--             mutate(diff_run_rate = c(0,diff(total_runs))) -->
-<!--   -->
-<!--   -->
-<!--  cricket_tsibble_bow <- cricket_bowling %>%  -->
-<!--    ungroup() %>%  -->
-<!--    mutate(data_index = row_number()) %>%  -->
-<!--    as_tsibble(index = data_index) -->
-
-<!-- cricket_tsibble_bow %>% gran_obs("over", "dot_balls") -->
-
-<!-- # cricket_tsibble_bow %>%  -->
-<!-- #   #filter(dot_balls %in% c(0, 1, 2))%>%  -->
-<!-- #   prob_plot("dot_balls", -->
-<!-- #             "over", -->
-<!-- #             hierarchy_model, -->
-<!-- #             response = "diff_run_rate", -->
-<!-- #             plot_type = "quantile", -->
-<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!-- #             symmetric = FALSE) -->
-
-<!-- cricket_tsibble_bow %>% gran_obs("over", "wickets") -->
-
-<!-- cricket_tsibble_bow %>%  -->
-<!--   #filter(wickets %in% c(0, 1))%>%  -->
-<!--   gather("proxies", "yes_no", -c(season, -->
-<!--                                      match_id, -->
-<!--                                      batting_team, -->
-<!--                                      bowling_team, -->
-<!--                                      inning,     -->
-<!--                                      over, -->
-<!--                                      diff_run_rate, -->
-<!--                                      data_index)) %>%  -->
-<!--   filter(yes_no %in% c("no")) %>%  -->
-<!--   prob_plot("proxies", -->
-<!--             "over", -->
-<!--             hierarchy_model, -->
-<!--             response = "diff_run_rate", -->
-<!--             plot_type = "boxplot", -->
-<!--             symmetric = FALSE, -->
-<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
-<!-- ``` -->
-
-<!-- ```{r proxies-yes} -->
-<!-- cricket_tsibble_bow %>%  -->
-<!--   #filter(wickets %in% c(0, 1))%>%  -->
-<!--   gather("proxies", "yes_no", -c(season, -->
-<!--                                      match_id, -->
-<!--                                      batting_team, -->
-<!--                                      bowling_team, -->
-<!--                                      inning,     -->
-<!--                                      over, -->
-<!--                                      diff_run_rate, -->
-<!--                                      data_index)) %>%  -->
-<!--   filter(yes_no %in% c("yes")) %>%  -->
-<!--   prob_plot("proxies", -->
-<!--             "over", -->
-<!--             hierarchy_model, -->
-<!--             response = "diff_run_rate", -->
-<!--             plot_type = "lv", -->
-<!--             symmetric = FALSE, -->
-<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
-
-<!-- ``` -->
-
-<!-- ```{r exwicketplot, echo=FALSE} -->
-<!-- #  -->
-<!-- # cricket_tsibble %>%  -->
-<!-- #   filter(dot_balls %in% c(0, 1, 2)) %>% -->
-<!-- #   mutate(lag_dot = lag(dot_balls)) %>%  -->
-<!-- #   filter(lag_dot %in% c(0, 1, 2)) %>% -->
-<!-- #   prob_plot("lag_dot", -->
-<!-- #             "over", -->
-<!-- #             hierarchy_model, -->
-<!-- #             response = "runs_per_over", -->
-<!-- #             plot_type = "quantile", -->
-<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- cricket_tsibble %>%  -->
-<!--   filter(wicket %in% c(0, 1)) %>% -->
-<!--   mutate(lag_wicket = lag(wicket)) %>%  -->
-<!--   filter(lag_wicket %in% c(0, 1)) %>% -->
-<!--   prob_plot("lag_wicket", -->
-<!--             "over", -->
-<!--             hierarchy_model, -->
-<!--             response = "runs_per_over", -->
-<!--             plot_type = "quantile", -->
-<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- ``` -->
-
-<!-- ```{r violin, fig.cap= "Violin and box plots of run rate across innings of the match and matches won or lost for Mumbai Indians(top) and Chennai Super Kings(bottom).The violins for CSK has long right tails in 2nd innings for cases when they lost or won.
-MI has long right tail only when they won the match."} -->
-
-<!-- matches_select <- matches %>% select(id, toss_winner, winner, win_by_runs) -->
-
-<!-- cricket_winner <- cricket_tsibble_all %>% -->
-<!--   left_join(matches_select, by = c("match_id" = "id")) -->
-
-<!-- MI_winner <- cricket_winner %>% -->
-<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
-<!--                     season %in% c(2013, 2015)) %>% -->
-<!--              mutate(win = if_else(winner == "Mumbai Indians", 1, 0)) -->
-
-<!-- MI_winner %>% -->
-<!--   gran_obs("win", "inning", hierarchy_model) -->
-
-<!-- CSK_winner <- cricket_winner %>% -->
-<!--              filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings", -->
-<!--                     season %in% c(2013, 2015)) %>% -->
-<!--              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0)) -->
-
-<!-- CSK_winner %>% -->
-<!--   gran_obs("win", "inning", hierarchy_model) -->
-
-<!--  MI_winner %>% -->
-<!--             prob_plot( "win","inning", -->
-<!--                                 hierarchy_model, -->
-<!--                                 response = "run_rate", -->
-<!--                                 plot_type = "violin", -->
-<!--                                 quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!--                      overlay = FALSE) + ggtitle("") + -->
-<!--     geom_boxplot(width = 0.1, color = "blue") -->
-
-<!--  CSK_winner %>% filter(inning %in% c(1,2)) %>% -->
-<!--            prob_plot( "win","inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "violin", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!--                     overlay = FALSE) + ggtitle("") + -->
-<!--    geom_boxplot(width = 0.1, color = "blue") -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- cricket_winner %>% as_tibble() %>% -->
-<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
-<!--                     winner!= "Mumbai Indians", -->
-<!--                     win_by_runs!=0, -->
-<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
-
-<!-- cricket_winner %>% as_tibble() %>% -->
-<!--              filter(batting_team ==  "Chennai Super Kings", -->
-<!--                     winner!= "Chennai Super Kings", -->
-<!--                     win_by_runs!=0, -->
-<!--                     inning ==2, -->
-<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
-<!-- ``` -->
-
-<!-- For MI, cases when they lost, distribution of run rates not very different depending on if they bat first or second.
-However, cases when they won, distribution of run rate is very skewed when they bat in the 2nd innings implying winning by short margins.
--->
-
-<!-- Figure \ref{fig:violin} shows that for CSK, irrespective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing.
-Run rate can go up to 6 runs per over depending on the situation.
-For MI, this is not true.
-When they won, their run rates in the 2nd innings went up to 6 and led to their win, but stayed within 4 whenever they lost.
-This implies CSK is more of a fighter than MI when it comes to defending in the 2nd innings.
-However, we can delve deeper and try to see which overs are the most volatile for CSK and if it is different to that of MI?
--->
-
-<!-- In Figure \ref{fig:lv}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match.
-It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings.
-There is an upward shift of the distribution in the second part of the innings (after 10th over) and more variability (longer and distinct letter values) as well.
--->
-<!-- CSK has been more uniform in their approach maintaining consistent scores per over.
-Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}.
--->
-
-<!-- whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second.
-CSK's run rate distribution is extremely right tailed in the 2nd innings, but not in the first.
-.
---> -->
-
-<!-- ```{r 2013_15_MI} -->
-<!--  cricket_tsibble_all %>%  -->
-<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
-<!--            prob_plot("inning", "over", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
-<!--     ) -->
-<!--  cricket_tsibble_all %>%  -->
-<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
-<!--            prob_plot("inning", "over", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "run_rate", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
-<!--     ) -->
-
-<!-- ``` -->
-
-<!-- r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."} -->
-<!-- #  -->
-<!-- #  -->
-<!-- #  -->
-<!-- # MI <- cricket_tsibble_all %>% -->
-<!-- #   filter(season %in% c(2010:2015), batting_team=="Mumbai Indians") -->
-<!-- #  -->
-<!-- #  -->
-<!-- # CSK <- cricket_tsibble_all %>% -->
-<!-- #   filter(season %in% c(2010:2015), -->
-<!-- #          batting_team=="Kolkata Knight Riders", -->
-<!-- #          inning %in% c(1,2)) -->
-<!-- #  -->
-<!-- #  -->
-<!-- #  MI %>% -->
-<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
-<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
-<!-- #  -->
-<!-- #  CSK %>% -->
-<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
-<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
-<!-- #  -->
-<!-- # MI %>% -->
-<!-- #   prob_plot("inning", "over", -->
-<!-- #                                hierarchy_model, -->
-<!-- #                                response = "run_rate", -->
-<!-- #                                plot_type = "lv", -->
-<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") -->
-<!-- #  -->
-<!-- #  -->
-<!-- # CSK %>% -->
-<!-- #   prob_plot("inning", "over", -->
-<!-- #                                hierarchy_model, -->
-<!-- #                                response = "run_rate", -->
-<!-- #                                plot_type = "lv", -->
-<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") -->
-<!-- #  -->
-<!-- # cricket_tsibble_all %>% distinct(batting_team) -->
-
-<!-- # ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."} -->
-<!-- #  -->
-<!-- # cricket_dot <- cricket_season %>%  -->
-<!-- #   mutate(wicket = if_else(is.na(dismissal_kind),0,1), -->
-<!-- #          dot_balls = if_else(total_runs==0,1, 0) -->
-<!-- #          ) %>%  -->
-<!-- #   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
-<!-- #   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs), -->
-<!-- #             wicket = sum(wicket), -->
-<!-- #             dot_balls = sum(dot_balls))  -->
-<!-- #  -->
-<!-- #  -->
-<!-- #  -->
-<!-- # cricket_tsibble_dot <- cricket_dot %>%  -->
-<!-- #   ungroup() %>%  -->
-<!-- #   mutate(data_index = row_number()) %>%  -->
-<!-- #   as_tsibble(index = data_index) -->
-<!-- #  -->
-<!-- # #cricket_tsibble_dot %>% gran_obs("over", "fielding_wckts") -->
-<!-- #  -->
-<!-- # cricket_tsibble_dot  %>%  -->
-<!-- #   filter(dot_balls %in% c(1,2,3,4)) %>%  -->
-<!-- #   mutate(lag_dot_ball = lag(dot_balls)) -->
-<!-- #   prob_plot("over","dot_balls", -->
-<!-- #                                hierarchy_model, -->
-<!-- #                                response = "run_rate", -->
-<!-- #                                plot_type = "quantile", -->
-<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
-<!-- #            overlay = FALSE)  + ggtitle("") -->
-<!-- # ``` -->
-<!-- #  -->
-<!-- #  -->
-
-<!--   cricket_tsibble %>% prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + -->
-<!--     theme( # remove the vertical grid lines -->
-<!--            panel.grid.major.x = element_blank() , -->
-<!--            # explicitly set the horizontal lines (or they will disappear too) -->
-<!--            panel.grid.major.y = element_line( size=.1, color="black" )  -->
-<!--     )
-
-<!--   cricket_tsibble %>% -->
-<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!--     cricket_tsibble %>% -->
-<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!-- ```{r runs_per-over} -->
-
-<!-- # all season per over runs -->
-<!-- cricket_per_over_all <- cricket_all_seas %>%  -->
-<!--   group_by(season, match_id, batting_team,inning, over) %>% -->
-<!--   summarise(runs_per_over = sum(total_runs))  -->
-
-<!-- # all season per over runs tsibble -->
-<!-- cricket_tsibble_all <- cricket_per_over_all %>%  -->
-<!--   ungroup() %>%  -->
-<!--   mutate(data_index = row_number()) %>%  -->
-<!--   as_tsibble(index = data_index) -->
-
-<!--   cricket_tsibble_all %>% prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   ) -->
-
-<!--   cricket_tsibble_all %>% prob_plot("inning_match", "season", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )   -->
-
-<!--   cricket_tsibble_all %>% prob_plot("season", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!--   )  +  -->
-
-<!--   cricket_tsibble %>% -->
-<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!--     cricket_tsibble %>% -->
-<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
-<!--     prob_plot("inning_match", "over_inning", -->
-<!--                                hierarchy_model, -->
-<!--                                response = "runs_per_over", -->
-<!--                                plot_type = "quantile", -->
-<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!--   ) -->
-
-<!-- {r read_data, echo=FALSE} -->
-
-<!-- cricket_tsibble <- cricketdata %>% -->
-<!--   mutate(data_index = row_number()) %>% -->
-<!--   as_tsibble(index = data_index) -->
-
-<!-- # cricket_tsibble %>% dynamic_create_gran( -->
-<!-- #   "ball_over", -->
-<!-- #   hierarchy_model -->
-<!-- # ) -->
-
-<!-- #over_inning    ball_over -->
-<!-- cricket_tsibble %>% granplot("over_inning", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!-- ) -->
-
-<!-- #inning_match   ball_over -->
-
-<!-- cricket_tsibble %>% granplot("inning_match", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!-- ) -->
-
-<!-- # inning_match   ball_inning -->
-<!-- #interesting plot -->
-<!-- cricket_tsibble %>% granplot("inning_match", "ball_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
-<!--   scale_x_discrete(breaks = seq(1,120,5)) -->
-
-<!-- # ball_over      over_inning -->
-
-<!-- cricket_tsibble %>% granplot("ball_over", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- # inning_match   over_inning -->
-
-<!-- cricket_tsibble %>% granplot("inning_match", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      over_match -->
-
-<!-- cricket_tsibble %>% granplot("ball_over", "over_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      inning_match -->
-
-<!-- cricket_tsibble %>% granplot("ball_over", "inning_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->
-
-<!-- cricket_tsibble %>% -->
-<!--   granplot( -->
-<!--     gran1 = "inning_match", -->
-<!--     gran2 = "over_inning", -->
-<!--     hierarchy_model, -->
-<!--     response = "total_runs", -->
-<!--     plot_type = "quantile" -->
-<!--   ) -->
-
-<!-- ```{r read_data_all_seas, echo=FALSE} -->
-
-<!-- load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda") -->
-
-<!-- cricket_tsibble_all <- cricket_all_seas %>% -->
-<!--   filter(season %in% c(2012:2015)) %>%  -->
-<!--   mutate(data_index = row_number()) %>% -->
-<!--   as_tsibble(index = data_index) -->
-
-<!-- # cricket_tsibble %>% dynamic_create_gran( -->
-<!-- #   "ball_over", -->
-<!-- #   hierarchy_model -->
-<!-- # ) -->
-
-<!-- #over_inning    ball_over  -->
-<!-- cricket_tsibble_all %>% granplot("over_inning", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-
-<!-- ) -->
-
-<!-- #inning_match   ball_over  -->
-
-<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_over", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
-<!-- ) -->
-
-<!-- # inning_match   ball_inning  -->
-<!-- #interesting plot -->
-<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--    plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
-<!--   scale_x_discrete(breaks = seq(1,120,5)) -->
-
-<!-- # ball_over      over_inning   -->
-
-<!-- cricket_tsibble_all %>% granplot("ball_over", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs", -->
-<!--   plot_type = "quantile", -->
-<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->
-
-<!-- # inning_match   over_inning   -->
-
-<!-- cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      over_match   -->
-
-<!-- cricket_tsibble_all %>% granplot("ball_over", "over_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) -->
-
-<!-- # ball_over      inning_match   -->
-
-<!-- cricket_tsibble_all %>% granplot("ball_over", "inning_match", -->
-<!--   hierarchy_model, -->
-<!--   response = "total_runs" -->
-<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->
-
-<!-- cricket_tsibble_all %>% -->
-<!--   granplot( -->
-<!--     gran1 = "inning_match", -->
-<!--     gran2 = "over_inning", -->
-<!--     hierarchy_model, -->
-<!--     response = "total_runs", -->
-<!--     plot_type = "quantile" -->
-<!--   ) -->
 

### 2019-12-08 main.Rmd
spell check done

@@
 
-\begin{equation} \label{eq:eq5}
+\begin{equation} \label{eq:circular-gran}
 \begin{split}
@@
 For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
-For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
+For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:circular-gran}.
 
@@
 
-\begin{equation} \label{eq7}
+\begin{equation} \label{eq:eq7}
 \begin{split}
@@
 With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context.
-This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and therby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and thereby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
 
@@
 A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
-These functionalities can be imported in the gravitas package to accomodate for aperiodic cyclic granularities.
+These functionalities can be imported in the gravitas package to accommodate for aperiodic cyclic granularities.
 
@@
 
-The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages.
-This manuscipt was created with knitr [@knitr] and R Markdown [@rmarkdown].
+The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
+Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs for improving the analysis of smart meter application.
+The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/) This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
 The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all required materials required to reproduce this article.

### 2019-12-09 main.Rmd
data structure change

@@
 Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization.
-It is recognized that the initial approaches of just considering time as an ordinal dimension in visualisation are inadequate to capture characteristics of time-dependent information.
-There is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
+It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
 Moreover, one can use calendars based on application domain that define contextual system of granularities.
 Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities.
-Since, we are interested in detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.
+Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.
 Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem can be broken down into smaller pieces by focusing on two cyclic granularities at a time.
+Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
 
 A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
-Since all cyclic granularities can be expressed in terms of the bottom granularity, which is the index of the tsibble, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
+Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
 This is extending the columns of tsibble by including the cyclic granularities.Now suppose we want to explore the measurement variable $v$ across two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.

### 2019-12-09 main.Rmd
synergy of time granularities

@@
 
-```{r data-structure, fig.cap = "The data structure for exploring periodicities in data.
-Two cyclic granularities serve as fixed and conditioned variable and the measurement variable is the observation across combinations of these cyclic time granularities.", fig.pos="htb"}
+```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure", fig.pos="htb"}
 include_graphics("Figs/data-structure-v1.png")
@@
 
-The way cyclic granularities relate become important when we consider the data structure <$A_i$, $B_j(s)$, $v(s)$> for exploration.
-We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the levels of the cyclic granularities.
+The way cyclic granularities relate become important when we consider the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ for exploration.
+We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the categories/levels of the cyclic granularities.
 To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
@@
 These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Let us take a specific example, where $C_1$ maps index set to day-of-month and $C_2$ maps index set to week-of-month.
+Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of day-of-month and week-of-month.
 Many of these are empty.
@@
 Again, let us consider a specific example to illustrate this.
-Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay.
-Here $C_1$ can take 7 values while $C_2$ can take 2 values.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD.
-While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations.
-For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.
+Let $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
+Here $C_1$ can have 7 levels while $C_2$ can have 2 levels.
+So there are 14 sets $S_{ij}$ corresponding to the possible combinations of day-of-week and WorkingDay/NonWorkingDay.
+While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations.
+For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
 
@@
 Thirdly, empty combinations can be a result of how granularities are constructed.
-Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month.
-Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
-This is different from the structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
+Let $C_1$ maps index set to Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
+Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
+This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
 Hence, they are referred to as "build-based" empty combinations.
 
-An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-
-Therefore, pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are called **harmonies**.
+An example when there will be no empty combinations could be where $C_1$ maps index set to day-of-week and $C_2$ maps index set to month-of-year.
+Here $C_1$ can have 7 levels while $C_2$ can have 12 levels.
+So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of day-of-week and month-of-year.
+All of these are non-empty because every day-of-week can occur in every month.
+
+A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
+And the ones that do not lead to any missing combinations are referrednto as **harmonies**.
 

### 2019-12-09 main.Rmd
visualization intro update

@@
 
-Density estimates are useful in exploring the properties of the measured variable.
-They can give valuable indication of features like skewness, multimodality or tail behavior.
+Estimating the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
 The measured variable in the considered data structure can be dicrete or continuous.
@@
 The entire distribution could be estimated using different density estimation methods or summarized using empirical quantiles.
-Several non-parametric density estimators and quantile estimators are proposed in the literature, the most commonly used form are as follows:
+Severalforms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
 
@@
 Creating a visualisation requires a number of nuanced judgements.
-We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate temporal granularities.
-The general framework of visualizing the distribution invloves a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
+We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
+The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
 Different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
@@
 There are several ways to plot statistical distributions.
-The displayed probabilities in these plots are either computed using kernel density estimation methods or empirical statistical methods.
+The displayed probabilities in these plots are either computed using kernel density estimation methods or quantile estimation methods.
 The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
 However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
-The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
 Density plots which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution, unlike discretizing the distribution and showing only parts of it.
+The methods are subject to the same sample size restrictions and challenges that apply to any density estimation.
+In practice, the densities are estimated reasonably with at least 30 observations.
+Even with sample sizes of several hundred, however, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
+
+We will discuss few conventional and recent ways to plot distributions using both of these methods.
+
+<!-- The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
 They can be useful in spotting multimodality, however, they might become obscuring with too many categories and information on the entire distribution can add to cognitive load.
@@
 As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
-The density based visualizations are subject to the same sample size restrictions and challenges that apply to any density estimation.
-In practice, the densities are estimated reasonably with at least 30 observations.
-Even with sample sizes of several hundred, however, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
-
-We will discuss few conventional and recent ways to plot distributions using both of these methods.
+-->
 
@@
 
-Most commonly used techniques to display a distribution of data include the histogram (Karl Pearson), which shows the prevalence of values grouped into bins and the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+Most commonly used techniques to display a distribution of data include the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
 The box plot is a compact distributional summary, displaying less detail than a histogram.

### 2019-12-09 main.Rmd
visualization sextion update

@@
 There are several ways to plot statistical distributions.
-The displayed probabilities in these plots are either computed using kernel density estimation methods or quantile estimation methods.
+The entire distribution can be displayed or some quantiles can be visualized with probabilities in these plots either computed using kernel density estimation methods or quantile estimation methods.
 The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
 However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
-Density plots which uses a kernel density estimate to show the probability density function of the variable can show the entire distribution, unlike discretizing the distribution and showing only parts of it.
+Density plots can show the entire distribution, unlike discretizing and showing only parts of it.
 The methods are subject to the same sample size restrictions and challenges that apply to any density estimation.
@@
 Most commonly used techniques to display a distribution of data include the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-The box plot is a compact distributional summary, displaying less detail than a histogram.
+The box plot is a compact distributional summary.
 Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
 Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
-Moreover, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
+However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
 Also, the number of outliers is large for larger data set since number of outliers is proportional to the number of observations.
@@
 “outliers” in letter value plots are those observations beyond the most extreme letter value.
-The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown and leading to overload of information in one plot.
+The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
 
@@
 Quantile plots visually portray the quantiles of the distribution of sample data.
-Much like the quartiles divide the data set equally into four equal parts, extensions include dividing the data set even further.
-These plots are referred to as "quantile" plots.
-For example, number of quantiles would be 9 for a decile plot and 99 for percentile plot.
-It avoids much clutter and just enable us to focus on specific probabilities, typically representing typical and extreme behaviors.
-A large data set is required for the extreme percentiles to be estimated with any accuracy.
 These plots can display any quantiles instead of quartiles in a traditional boxplot.
-When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g: outside 1.5 times the interquartile range above the upper quartile and below the lower quartile).
+This gives more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
+When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
+outside 1.5 times the interquartile range above the upper quartile and below the lower quartile).
 However, outliers are open to interpretation and not shown in an quantile plot.
@@
 
-### Kernel density estimation methods
+### Kernel density based methods
 
@@
 
-The summary plot [@Potter2010-qc, 2010] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot (both color and width are mapped to estimated density), and an overlay of a reference distribution.
+The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
 This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
@@
 Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
-A clear advantage over boxplots is that these plots allow us to see multimodality in the distribution.
+Like other density based displays, these plots allow us to see multimodality in the distribution.
 However, these plots can be obscuring when there is overlap of distribution for two or more categories of the y-axis.
-Also, if there are lot of categories, it is difficult to compare the height of the densities across categories.
+Also, with lot of categories, it is difficult to compare the height of the densities across categories.
 
@@
 
-### Effect of levels of temporal granularities
-
-The levels of the two granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
+### Effect of levels of cyclic granularities
+
+The levels of the two cyclic granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
 The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.
@@
 
-The following principles are useful while choosing distribution plots given two temporal granularities.
+The following principles are useful while choosing distribution plots given two temporal granularities:
 
@@
 
-  - If levels of any granularity plotted are more than medium, empirical based methods of distribution visualizing are preferred as they use less space by design than most density based methods.
+  - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.
 
@@
 
-## Effect of synergy of time granularities
+## Effect of synergy of cyclic granularities
 
@@
 
-Clearly, in \autoref{fig:allFig}, we observe that some choices of time granularities work and others do not.
-In \autoref{fig:allFig} (c), there will be no observations for some combinations "Day-of-Month" and "Day-of-Year".
+Clearly, in \autoref{fig:allFig}, we observe that some choices of cyclic time granularities work and others do not.
+In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year.
 In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every "Day-of-Week" can occur in any "Month-of-Year".
+On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every day-of-week can occur in any month-of-year.
 Thus the graphs that don't work are those where many of the combination sets are empty.
-In other words, if there are levels of time granularities plotted across x-axis which are not spanned by levels of the time granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
-We hypothesize that the synergy of these  granularities are in play while deciding if the resulting plot would be a good candidate for exploratory analysis.
+In other words, if there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
+We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
 
@@
 
-Suppose we have $T$ observations, and two cyclic granularities -  $C_1$ with $n$ categories and $C_2$ with $m$ categories.
+Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
 Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
-There are no structurally empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
+There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
 If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.

### 2019-12-09 main.Rmd
edit till introduction

@@
 
-The remainder of the paper is organized as follows.
-\autoref{sec:lineartime} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
-\autoref{sec:cyclic-gran-def} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
+The remainder of the paper is organized as follows: \ref{sec:linear-time} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
+\ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
 Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
-\autoref{sec:data-structure} discusses the data structure for exploring distribution of an observed variable across pair of cyclic time granularities.
-\autoref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
+\ref{sec:data-structure} discusses the data structure for exploring distribution of the associated time series across pair of cyclic time granularities.
+\ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
 Section \ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application.
@@
 
-# Linear time granularities {#sec:lineartime}
+# Linear time granularities {#sec:linear-time}
 
@@
 
-## Definitions and Relationships
+## Definitions and relationships
 
@@
 
-## Computation through Calendar Algebra
+## Computation through calendar algebra
 
@@
 
-Some relevant grouping oriented operations are discussed, which will be used in Section \ref{sec:cyclic-gran-def} to define circular and aperiodic granularities.
+Some relevant grouping oriented operations are discussed, which will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
 
@@
 
-# Cyclic time granularities {#sec:cyclic-gran-def}
+# Cyclic time granularities {#sec:cyclic-gran}
 
@@
 
-In \autoref{sec:cyclic-gran-def}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed.
+In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
@@
 
-## Choice of Plots
+## Choice of statistical distribution plots
 

### 2019-12-09 main.Rmd
changes till linear granularity

@@
 
-The remainder of the paper is organized as follows: \ref{sec:linear-time} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
-\ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
+The remainder of the paper is organized as follows: Section \ref{sec:linear-time} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
+Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
 Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
-\ref{sec:data-structure} discusses the data structure for exploring distribution of the associated time series across pair of cyclic time granularities.
-\ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
+Section \ref{sec:data-structure} discusses the data structure for exploring distribution of the associated time series across pair of cyclic time granularities.
+Section \ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
 Section \ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application.
@@
 
- There has been several attempts to provide the framework for formally characterizing time-granularities.
-One of the first attempts occur in [@Bettini1998-ed] with the help of the following definitions:
+ There has been several attempts to provide the framework for formally characterizing time-granularities. One of the first attempts occur in @Bettini1998-ed with the help of the following definitions:
 
@@
 
-\newtheorem{definition}{Definition}[section]
+\newtheorem{definition}{Definition}
 
 \begin{definition}\label{def:definition}
-A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+(time domain): A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
 \end{definition}
@@
 \begin{definition}\label{def:linear}
-A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
+(linear granularity): A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
 
@@
 \begin{definition}\label{def:granule}
-Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
+(granule) Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
 \end{definition}
 
-**Discussion:** The first condition in \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+**Discussion:** The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
 \autoref{fig:linear-time} shows the implication of this condition.
 It shows the linear granularities hours, days, weeks, months and years, each of which are unidirectional in nature and arranged from past to future.
-If we consider the chronon [@aigner2011visualization] as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+If we consider the chronon (see @aigner2011visualization) as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
 Each granule of the linear granularities shown are represented through a box.
@@
 \begin{definition}\label{def:finerthan}
-A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
+(finer than): A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
 for each index i, there exists an index j such that $G(i) \subset H(j)$.
@@
 \begin{definition}\label{def:groupsinto}
-A linear granularity G groups into a linear granularity H, denoted
+(groups into): A linear granularity G groups into a linear granularity H, denoted
 $G \trianglelefteq H,$
@@
 $G \preceq H$ if every granule in G is a subset of some granule in H.
-In \autoref{fig:groups-finer}, both $Days \trianglelefteq Weeks$ and $Days \preceq Weeks$ hold true.
-However, Weekends groups into Holidays but Weekends are not finer than Holidays.
-On the other hand, Weekends are finer than weeks but do not group into weeks.
-Consider another example where $Days \trianglelefteq Months$.
+In \autoref{fig:groups-finer}, both $day \trianglelefteq week$ and $day \preceq week$ hold true.
+However, weekends are finer than weeks but do not group into weeks.
+And weekends neither groups into or finer than public holidays.
+Consider another example where $day \trianglelefteq month$.
 This relationship is not fully described until we associate it with periodicity.
@@
 ```{r groups-finer,echo=FALSE,out.width = "100%",fig.cap="Illustrations of groups into and finer than relationships.
-Weekends groups into Holidays but Weekends are not finer than Holidays.
-On the other hand, Weekends are finer than weeks but do not group into weeks.
-Day groups into and are finer than weeks"}
+Weekends are finer than weeks but do not group into weeks.
+And weekends neither groups into or finer than public holidays"}
 knitr::include_graphics("Figs/groupsandfiner.png")
@@
 \begin{definition}\label{def:periodical}
-A granularity H is periodical with respect to a granularity G if
-(1) $G \trianglelefteq H,$, and
+(groups periodically): A granularity H is periodical with respect to a granularity G if
+(1) $G \trianglelefteq H,$ and
 (2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
@@
 
-A granularity H which is periodical with respect to G is specified by:
-(i) the R sets of indexes of G, ${S_0,...,S_(R-1)}$ describing the granules of H within one period;
+**Discussion:** A granularity H which is periodical with respect to G is specified by:
+(i) the R sets of indexes of G, ${S_0,...,S_{R-1}}$ describing the granules of H within one period;
 (ii) the value of P; 
@@
 Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
-$H (0), . . . , H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-
-**Discussion:**
-
-Granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
+$H (0), .
+.
+.
+, H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
+Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
 
 \begin{definition}\label{def:quasi}
-A granularity H is quasi-periodical with respect to a granularity G if
+(groups quasi-periodically): A granularity H is quasi-periodical with respect to a granularity G if
 (1) $$G \trianglelefteq H,$$, and
@@
 
-**Discussion:**
-
-Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by R and P.
-
-\begin{definition}\label{def:BG}
-Bottom granularity - Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set.
+**Discussion:** Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
+
+\begin{definition}\label{def:B}
+(bottom granularity): Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
 \end{definition}
 
-**Discussion:** Given the set of all granularities defined over the time domain (Z; <), and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+**Discussion:** Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
 
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
+It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate
 new granularities from the bottom one or recursively, from those already generated.
-Thus, the relationship between the operand(s) and the resulting granularities are encoded
-in the operations.
-
 The calendar algebra consists of two kinds of operations: grouping-oriented and
@@
 The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
-Example can be to consider a calendar with only two linear granularities minute and hour and hour is generated by grouping every 60 minutes.
+Example can be to consider a calendar with only two linear granularities "minute" and "hour" and "hour" is generated by grouping every $60$ "minutes".
 The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
-For example, one can choose to look at the granularity "Monday" and hence select only Mondays while looking at the linear granularity "day".
+For example, one can choose to look at the granularity "Monday" and hence select only "Monday" from linear granularity "day".
 
@@
 
-- **The grouping operation** : Let G be a full-integer labeled granularity, and m a positive integer.
-The grouping operation $Group_m(G)$ generates a new granularity G, by partitioning the granules of G into m-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G = Group_m(G)$ is the full-integer labeled granularity such that for
-each integer $i$, $G(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G(j)$.
+- **The grouping operation** : Let $G1$ be a full-integer labeled granularity, and $m$ a positive integer.
+The grouping operation $Group_m(G)$ generates a new granularity $G2$, by partitioning the granules of $G$ into m-granule groups and making each group a granule of the resulting granularity.
+More precisely, $G2 = Group_m(G1)$ is the full-integer labeled granularity such that for
+each integer $i$, $G2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G1(j)$.
 
@@
 
-  - pseudomonth = 
-  $Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, Alter^{12}_{6,-1}(day,   Alter^{12}_{4,-1)}(day, Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))$,
-                  where the granularity pseudomonth is generated by grouping 31 days, 
-                  and then shrink April (4), June (6), September (9) and November (11)
-                  by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
-
-For more variations of grouping operations and granule oriented operations, the readers are recommended to see [@Ning2002-tf].s
+begin{equation}
+pseudomonth & = Alter^{12}_{11,-1}(day, \\
+  Alter^{12}_{9,-1}(day, \\
+  Alter^{12}_{6,-1}(day,  \\
+  Alter^{12}_{4,-1)}(day, \\
+  Alter^{12}_{(2,-3)}(day, \\
+  Group_{31}(day))))))
+end{equation}
+
+where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
+
+For more variations of grouping operations and granule oriented operations, the readers are recommended to see @Ning2002-tf.
 

### 2019-12-09 main.Rmd
changes till quasi-circular

@@
 
-begin{equation}
+\begin{equation}
+\begin{split}
 pseudomonth & = Alter^{12}_{11,-1}(day, \\
@@
   Group_{31}(day))))))
-end{equation}
+\end{split}
+\end{equation}
 
@@
 We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of the index set to the time index of a tsibble.
+A tsibble consists of an index, key and measured variables.
+An index is a variable with inherent ordering from past to present and key is a set of variables that define observational units over time.
+In a tsibble each observation should be uniquely identified by index and key.
+
+A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the time index of a tsibble.
 A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
-Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.
+For example, if the time index of a tsibble is days, then a linear granularity might be weeks, months or years.
+A bottom granularity is represented by the index of the tsibble.
+Lining up with the the assumption in @Bettini2000-vy, that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists.
 In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
@@
 
-Cyclic time granularities which accommodate periodicities can be constructed by relating two linear granularities.
-The mappings between linear granularities can be regular or irregular.
+Cyclic time granularities which accommodate for periodicities can be constructed by relating two linear granularities.
+The mappings between these linear granularities can be regular or irregular.
 For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
-On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31.
+On contrary, an irregular mapping exists between days and months, since one might have days ranging from 28 to 31.
 To elaborate more, mappings will be regular if, for example, linear granularities are formed by grouping operation.
@@
 Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
-Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
-The formalism would differ for circular and quasi-circular granularities although both accommodate periodicities in data.
+Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
+The formalism would differ for circular and quasi-circular granularities although both accommodate for periodicities in data.
 
@@
 \begin{definition}\label{def:circular}
-A circular granularity $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
+(circular granularity): A circular granularity $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
 
@@
 
-**Example**:  Example showing circular granularity relating two linear granularities "Day" and "Week" is visually depicted in a series of slots in the \autoref{fig:circular-dow}.
+**Example**:  Example showing circular granularity relating two linear granularities "day" and "week" is visually depicted in a series of slots in the \autoref{fig:circular-dow}.
 Each granule is represented by a box.
@@
 The first slot in the diagram shows the index set.
-The index of the tsibble considered is days.
-Day and Week are two linear granularities with $P =  7$.
-The circular granularity $C_{Day, Week}$ representing Day-of-Week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
+The bottom granularity, represented by the index of the tsibble considered is "day".
+"day" and "week" are two linear granularities with $P =  7$.
+The circular granularity $C_{day, week}$ representing day-of-week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
 
@@
 \toprule
-Minute-of-Hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
-Minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
-Hour-of-Day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
-Hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
-Day-of-Week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
+minute-of-Hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
+minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
+hour-of-Day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
+hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
+day-of-Week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
 \bottomrule
@@
 
-A **quasi-circular** granularity can not be defined using modular arithmetic due to its irregular mapping with the bottom granularity.
+A **quasi-circular** granularity can not be defined using modular arithmetic since they are formed using two linear granularities with irregular mapping.
 However, they are still formed with linear granularities, one of which "groups periodically into" the other.
@@
 \begin{definition}\label{def:quasicircular}
-A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if
+(quasi-circular granularity): A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if
 

### 2019-12-09 main.Rmd
till quasi-circular

@@
 \begin{split}
-Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
+Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\
   & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
@@
 \begin{split}
-Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
+Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,\quad since \quad  10 \in T_{1}  \\
   & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
@@
 The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$.
-Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in in \autoref{def:quasicircular}.
+Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.
 
@@
 
-It can be noted here that if a linear granularity $G'$ is quasi-periodic with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
+It can be noted here that if a linear granularity $G'$ is quasi-periodic (Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
 
@@
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ none of which can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ (that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity).
 Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
@@
 \begin{definition}\label{aperiodic}
-An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
+(aperiodic cyclic granularity): An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
 
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and $M = \bigcup_{j=1}^{n}M_{j}$.
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{j=1}^{n}M_{j}$ and index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
 \end{definition}
@@
 **Discussion**: The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
-The index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
 
@@
 
-In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed.
+In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.

### 2019-12-09 main.Rmd
minor changes till cyclic granularities fixed

@@
 
-The cyclic calendar algebra consists of two kinds of operations:  
+The cyclic calendar algebra consists of broadly two kinds of operations:  
 (1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities  
@@
 \begin{split}
-C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1)) \\  
-  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1]) \\
-  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1))\\
-  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1))\\
+C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z))) \\  
+  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z))) \\
+  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
+  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
   &\vdots\\
-  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1))\\
+  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z)))\\
 \end{split}
@@
 
-*Example:*Let us use the equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar..
+*Example:* Let us use the equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
 
@@
 Suppose we have a hierarchy table using some linear granularties from Gregorian calendar.
-Since months consists of unequal number of days, any linear granularity unit with higher order than months will also have unequal number of days.
+Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.
 This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities.
-The single-order-up granularity day_month is quasi-circular.
+The single-order-up granularity day-of-month is quasi-circular.
 Any single-order-up granularities which are formed by linear granularities below days are circular.
@@
 |  hour    	  |hour-of-day       |  24  	        |
-|  day      	|day-of-month      |"quasi-circular"|
+|  day      	|day-of-month      |quasi-circular  |
 |  month    	|month-of-year     |  12            |
@@
 
-where, $T_w$ are the sets of indices of linear granularity day such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 
@@
 C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)) \\ 
-& = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1)) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)\\
+& = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))) + P(l, m')(C_{G_{m'},G_{m}}(z))\\
 \end{split}
@@
 A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are referrednto as **harmonies**.
+And the ones that do not lead to any missing combinations are referred as **harmonies**.
 
@@
 The entire distribution could be estimated using different density estimation methods or summarized using empirical quantiles.
-Severalforms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
+Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
 

### 2019-12-09 main.Rmd
minor changes till cyclic granularities fixed

@@
 
-\begin{equation}
+\begin{equation*}
 \begin{split}
-pseudomonth & = Alter^{12}_{11,-1}(day, \\
-  Alter^{12}_{9,-1}(day, \\
-  Alter^{12}_{6,-1}(day,  \\
-  Alter^{12}_{4,-1)}(day, \\
-  Alter^{12}_{(2,-3)}(day, \\
-  Group_{31}(day))))))
+pseudomonth = Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, \\
+  Alter^{12}_{6,-1}(day, Alter^{12}_{4,-1)}(day, \\
+  Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))
 \end{split}
-\end{equation}
-
-where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
+\end{equation*}
+
+, where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
 
@@
 Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs for improving the analysis of smart meter application.
-The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/) This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
+The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
+This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
 The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all required materials required to reproduce this article.

### 2019-12-09 main.Rmd
visualization section 1 naming and corresponding changes

@@
 
-## Choice of statistical distribution plots
+## Choice of exploratory visualization tool
 
@@
 
-There are several ways to plot statistical distributions.
-The entire distribution can be displayed or some quantiles can be visualized with probabilities in these plots either computed using kernel density estimation methods or quantile estimation methods.
+There are several ways to visualize statistical distributions.
+The entire distribution can be displayed through density estimates using non-paramteric density estimation methods or only summarized values can be displayed using descriptive statistics.
 The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
 However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
-Density plots can show the entire distribution, unlike discretizing and showing only parts of it.
-The methods are subject to the same sample size restrictions and challenges that apply to any density estimation.
+Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it.
+These displays depend on the specific estimation procedure (e.g., kernel density estimate) as well as on additional smoothing parameters and hence can vary for the same data set.
+Furthermore, these methods are subject to the same sample size restrictions and challenges that apply to any density estimation.
 In practice, the densities are estimated reasonably with at least 30 observations.
-Even with sample sizes of several hundred, however, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
+However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
+The visualization tools based on descriptive statistics display only actual observations from the sample, and remain free of any tuning parameters or Kernels.
 
@@
 
-### Quantile based methods
-
-Most commonly used techniques to display a distribution of data include the box-and-whisker plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-The box plot is a compact distributional summary.
+### Descriptive statistics based methods
+
+Most commonly used techniques include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{th}$ and $(n + 1 - ([n/2] + 1)/2)^{th}$ order statistics are taken as estimates of the quartiles.
 Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of these variations are helpful to get an idea of the distribution at a glance.
+Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
 However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
@@
 The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+These are particularly useful for large data sets.
 “outliers” in letter value plots are those observations beyond the most extreme letter value.
@@
 These plots can display any quantiles instead of quartiles in a traditional boxplot.
-This gives more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
+Some of the most implemented definitions of sample quantiles based on one or two order statistics can be found in @Hyndman1996-ty.
+These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
 When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
@@
 
-### Kernel density based methods
+### Non-parametric density based methods
 

### 2019-12-09 main.Rmd
typos correct

@@
 Let us see if this fact is true.
-Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one or two dismissals in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
+Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
 
@@
 Each of the quantiles decrease across facets.
-With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for 10th, 25th and 50th percentile quantiles.
+With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for different quantiles.
 

### 2019-12-09 main.Rmd
cricket example making clear

@@
   G = c("over", "inning", "match"),
-  k = c(20, 2, 1))
-
+  C = c("over_inning", "inning_match", 1),
+  P = c(20, 2, 1))
 knitr::kable(hierarchy_model, format = "markdown")
@@
 
+In Q2 and Q3s we considered dots balls per over or catches per over as aperiodic cyclic granularities, which not necessarily appear in the hierarchy table since we can not compute these granularities as a function of circular or quasi-circular granularities.
+Also, their order in the hierarchy table is not clear as they are essentially obtained from some aperiodic linear granularity.
+
 # Discussion {#sec:discussion}

### 2019-12-10 main.Rmd
language change in cricket analysis

--- 2019-12-09 main.Rmd
@@
 The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
-With eight teams, each team plays each other twice in a home-and-away round-robin format in the league phase.
-In a Twenty20 game the two teams have a single innings each, which is restricted to a maximum of 20 overs.
+In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season can consist of several matches.
+The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
+The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016), each of which has 6 balls in an over and 2 innings in a match.
+s
+
+<!-- the two teams have a single innings each, which is restricted to a maximum of 20 overs.
 Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls with some exceptions.
-
-The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data cross overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016).
+-->
 
@@
 
-Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, where each over, which represents an ordering from past to future, can form the index of the tsibble.
+Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, with each over representing an ordering from past to future, can form the index of the tsibble.
 The hierarchy table would look like the following:
@@
 hierarchy_model <- tibble::tibble(
-  G = c("over", "inning", "match"),
-  k = c(20, 2, 1))
-
+  G = c("over", "inning", "match", "season"),
+  C = c("over-of-inning", "inning-of-match", "match-of-season",1),
+  P = c(20, 2, "aperiodic", 1))
 knitr::kable(hierarchy_model, format = "markdown")
@@
 
-First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas-over-inning}.
+First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperidic cyclic granularity) in \autoref{fig:seas-over-inning}.
 The distribution of runs per over has not significantly changed from 2008 to 2016.
@@
 
-Mumbai Indians(MI) and Chennai Super kings(CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final 4 from 2010 to 2015.
+Mumbai Indians (MI) and Chennai Super kings (CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final playoffs from 2010 to 2015.
 It would be interesting to take their example in order to dive deeper into the first question.
@@
 
-Q2: Is run rate set to reduce in subsequent over if fielding/bowling is good in the previous over? Between fielding and bowling which penalizes runs in the subsequent over more?
+Q2: Is runs per over set to reduce in subsequent over if fielding is good in the previous over?
 
@@
 If a batsman is bowled out, it does not necessarily signify good fielding.
-So we only include catches and run out as a measure of fielding.
+So we only include number of catches and run out in an over as a measure of good fielding.
 Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.
-Let us see if this fact is true.
-Figure \ref{fig:field} shows the difference between run rate between two subsequent overs are negative when good fielding leads to one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
+We can verify this fact through Figure \ref{fig:field} which shows that the difference in runs between two subsequent overs are negative when good fielding leads to one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
 
@@
 
-Q3: Is runs set to reduce in the next over for dot balls in this over?
+Q3: Are runs set to reduce in the next over for dot balls in the previous over?
 
@@
 
+In Q2 and Q3, we considered dots balls per over or catches per over as aperiodic cyclic granularities.
+They do not appear in the hierarchy table since we can not compute these granularities as a function of circular or quasi-circular granularities.
+Also, their order in the hierarchy table is not clear as they are essentially obtained from some aperiodic linear granularity.
+
 # Discussion {#sec:discussion}

### 2019-12-14 main.Rmd
changing section visualization

@@
 
-Creating a visualisation requires a number of nuanced judgements.
+The three key components of every visualization is: data, aesthetic mappings between variables in the data and visual attributes, and at least one layer which describes how to render each observation.
+Layers are usually created with a geom function.
+Considering the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by:
+- Data: <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$
+- Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position.
+- Layer: any geom that summarizes distribution
+
+To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or facetting. Facetting creates tables of graphics by displaying the same graph for each subset.
+- Facet:  $B_j(s)$
+
 We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.

### 2019-12-15 main.Rmd
grammar of graphics section update

--- 2019-12-10 main.Rmd
@@
 
-<!-- The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, allows us to produce graphics using the same structured thinking that is used to design analysis.
-Briefly, any graphics are made up of **Layers**, **Facet**,**Scale** and a  **coordinate system**.
+The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk]. 
+The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, suggests that graphics are made up of distinct layers of grammatical elements and meaningful plots are built around appropriate aesthetic mappings.
+<!-- seven grammatical elements in total, out of them three are essential.  -->
+
+Briefly, the layered grammar defines the components of plot as:
+
+- **Layers**:  made up of data, mapping, statistical transformation, geometric objects and position adjustments
+
+  - **Data**: observations made on a number of variables, typically thought of as a sample from a population distribution
+
+  - **Mappings** describing how variables are mapped to elements of the plot, such as horizontal axis or color
+
+  - **Statistical transformation** can summarize the data in many useful ways. The geometric object determines the physical representation of the data in the graph.
+
+  - **Geometric objects** control the type of plot that is created and are classified by their dimensions, for example points (0 dimension), line (1 dimension), polygons(2 dimensions) and so on.
+
+  - **Position adjustment** of elements on the plot are needed for dense data, otherwise data points might obscure one another.
+
+- **Facet**ting is a mechanism for splitting the data into subsets, then plotting each subset in a different panel.
+This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
+
+- **Scale** controls how data is mapped to aesthetic attributes. For every aesthetic property employed in a layer, we need one scale.
+
+- **A coordinate system** maps the position of objects onto the plane of the plot, and controls how the axes and grid lines are drawn.
+
+Drawing from the grammar of graphics, if <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ serves as the basis of visualizing the probability distribution of the measured variable, the following components of the layered grammar can be specified:
+
+- Data: <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$
+- Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position
+- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution
+- Geometric objects: e.g. boxplot, letter value, violin, ridge, highest density regions
+
+After the essential layers, the next layer being used is the facet which dictates how to split up our plot.
+
+- Facet:  $B_j(s)$
+
+<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or facetting.
+Facetting creates tables of graphics by displaying the same graph for each subset.
+-->
+
+<!-- The three essential component layers are data, aesthetic mapping and the geometric objects. -->
+<!-- Data are observations made on a number of variables, typically thought of as a sample from a population distribution.
+Mappings describe how variables are mapped to elements of the plot, such as horizontal axis or color -->
+
+<!-- the second layer is the aesthetic mapping, which refers to the scales on to which we want to map our data -->
+<!-- The third essential layer is the geometry which decides how the plot will look. After the essential layers, the next layer being used is the facet which dictates how to split up our plot -->
+
+<!-- Aesthetics refers to the scales on to which we want to map our data. we create plots by layering grammarical elements on top of one another and use aesthetic maps -->
+
+<!-- The grammar of graphics establish the building blocks for creating meaningful visualizations. It -->
+<!-- This means we have a framework that allows us to communicate in a way.  -->
+
+<!-- .
+allows us to produce graphics using the same structured thinking that is used to design analysis.
+Briefly, any graphics are made up of **Layers**, **Facet**, **Scale** and a  **coordinate system**.
 **Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
@@
 
-Creating a visualisation requires a number of nuanced judgements.
-We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
+<!-- The three key components of every visualization is: data, aesthetic mappings between variables in the data and visual attributes, and at least one layer which describes how to render each observation.
+Layers are usually created with a geom function.
+Considering the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
+<!-- - Data: <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
+<!-- - Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position. -->
+<!-- - Layer: any geom that summarizes distribution -->
+
+<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or facetting.
+Facetting creates tables of graphics by displaying the same graph for each subset.
+-->
+<!-- - Facet:  $B_j(s)$ -->
+
+<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
 The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
+-->
+
 Different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
 We will look at each of these aspects separately and analyze the effect of each on visualization.
-
 While we accomplish that, we need to bear in mind that any customized species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.

### 2019-12-15 main.Rmd
update name of subsections

@@
 
-## Estimation of the measured variable
+## Summarizing properties of the distribution
 
@@
 The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk]. 
-The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, suggests that graphics are made up of distinct layers of grammatical elements and meaningful plots are built around appropriate aesthetic mappings.
-<!-- seven grammatical elements in total, out of them three are essential.  -->
-
-Briefly, the layered grammar defines the components of plot as:
+The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, suggests that graphics are made up of distinct layers of grammatical elements.
+Briefly, the grammatical elements of any plot are as follows:
 
@@
 
-- **Facet**ting is a mechanism for splitting the data into subsets, then plotting each subset in a different panel.
-This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
+- **Facet** dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
 
@@
 - Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution
-- Geometric objects: e.g. boxplot, letter value, violin, ridge, highest density regions
-
-After the essential layers, the next layer being used is the facet which dictates how to split up our plot.
-
+- Geometric objects: examples can include boxplot, letter value, violin, ridge, highest density region plots
 - Facet:  $B_j(s)$
@@
 
-Different distribution plots might be appropriate depending on which features of the distrbution we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
+Hence, the ingredients for the plots are subsets of numbers (or categorical measurements), and the choice of faceting or axis alignment for the subset strata.
+The basic plot choice is display of density.
+Different plots might be appropriate depending on which features of the density we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
 We will look at each of these aspects separately and analyze the effect of each on visualization.
@@
 
-## Choice of exploratory visualization tool
+## Choice of
 

### 2019-12-15 main.Rmd
measured variable section update

@@
 
-## Summarizing properties of the distribution
-
-Estimating the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
-The measured variable in the considered data structure can be dicrete or continuous.
-The probability distribution of them could be studied by estimating the probability density function (continuous) or probability mass function (dicrete).
+## Measured variable
+
+<!-- Moreover, the measured variable in the considered data structure can be dicrete or continuous. -->
+
+Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
+With huge amount of data being available, mean, median or any one summary statistic might not be enough to understand the different properties of the measured variable.
+Instead the distribution could be summarized using some summary statistics together, each of which can highlight different properties of the data.
+Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries.
+Another way of summarizing could be estimating the probability density function (continuous) or probability mass function (dicrete).
 The approach can be parametric or non-parametric.
@@
 These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
-The entire distribution could be estimated using different density estimation methods or summarized using empirical quantiles.
 Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
@@
 
-Hence, the ingredients for the plots are subsets of numbers (or categorical measurements), and the choice of faceting or axis alignment for the subset strata.
+<!-- Hence, the ingredients for the plots are subsets of numbers (or categorical measurements), and the choice of faceting or axis alignment for the subset strata.
 The basic plot choice is display of density.
@@
 We are trying to address problems where even with good aesthetic qualities and good data, the graph might be confusing or misleading because of how people process the information.
+-->
 
@@
 
-## Choice of
+## Choice of statistical transformations and geometric objects
 
@@
 
-There are several ways to visualize statistical distributions.
-The entire distribution can be displayed through density estimates using non-paramteric density estimation methods or only summarized values can be displayed using descriptive statistics.
+There are several ways to summarize a statistical distributions.
+The entire distribution can be estimated through density estimates using non-paramteric density estimation methods or only summarized using descriptive statistics.
 The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
@@
 These displays depend on the specific estimation procedure (e.g., kernel density estimate) as well as on additional smoothing parameters and hence can vary for the same data set.
-Furthermore, these methods are subject to the same sample size restrictions and challenges that apply to any density estimation.
 In practice, the densities are estimated reasonably with at least 30 observations.
 However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
-The visualization tools based on descriptive statistics display only actual observations from the sample, and remain free of any tuning parameters or Kernels.
+The summaries based on descriptive statistics remain free of any tuning parameters or Kernels.
 
@@
 
+<!--  Furthermore, these methods are subject to the same sample size restrictions and challenges that apply to any density estimation.  -->
 <!-- The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
@@
 
-Given a context, it is good to be conversant with the benefits and challenges while choosing a distribution plot.
-As a general rule, if we have too many categories, the quantile plots are useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
+Each type of density display has different parameters.
+These are quantities that need to be estimated given the data.
+Each is equipped with some benefits and challenges.
+For example, if we have too many categories in the variables mapped to facet or x-axis, the quantile plots might be useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
 

### 2019-12-15 main.Rmd
choice of plots section update

@@
 
-## Measured variable
+## Measured variable {#sec:measured-variable}
 
@@
 
-Drawing from the grammar of graphics, if <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ serves as the basis of visualizing the probability distribution of the measured variable, the following components of the layered grammar can be specified:
+Drawing from the grammar of graphics, if <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ serves as the basis of visualizing the distribution of the measured variable, the following components of the layered grammar can be specified:
 
@@
 
-## Choice of statistical transformations and geometric objects
+## Choice of plots
 
@@
 
-There are several ways to summarize a statistical distributions.
-The entire distribution can be estimated through density estimates using non-paramteric density estimation methods or only summarized using descriptive statistics.
+Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization.
+The basic plot choice is display of density.
+In Section \ref{sec:measured-variable}, it is dicussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
 The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
-However, they avoid much clutter, where some specific probabilities representing typical and extreme behavior are on focus.
-Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it.
-These displays depend on the specific estimation procedure (e.g., kernel density estimate) as well as on additional smoothing parameters and hence can vary for the same data set.
+However, they avoid much clutter and help us to focus on some specific properties of the data.
+as well as on additional smoothing parameters and hence can vary for the same data set.
 In practice, the densities are estimated reasonably with at least 30 observations.
 However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
-The summaries based on descriptive statistics remain free of any tuning parameters or Kernels.
+The summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.
 
 We will discuss few conventional and recent ways to plot distributions using both of these methods.
+
+<!-- Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it.
+These displays depend on the specific estimation procedure (e.g., kernel density estimate)  -->
 
@@
 
-### Descriptive statistics based methods
-
-Most commonly used techniques include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+### Descriptive statistics based displays
+
+Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
 These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{th}$ and $(n + 1 - ([n/2] + 1)/2)^{th}$ order statistics are taken as estimates of the quartiles.
@@
 
-### Non-parametric density based methods
-
-Traditional ways to visualize densities include violin plots [@Hintze1998-zi].
+### Kernel density based displays
+
+Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi].
 The shape of the violin represents the density estimate of the variable.

### 2019-12-15 main.Rmd
choice of plots updated

@@
 However, they avoid much clutter and help us to focus on some specific properties of the data.
-as well as on additional smoothing parameters and hence can vary for the same data set.
+Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
 In practice, the densities are estimated reasonably with at least 30 observations.
@@
 
-Quantile plots visually portray the quantiles of the distribution of sample data.
-These plots can display any quantiles instead of quartiles in a traditional boxplot.
-Some of the most implemented definitions of sample quantiles based on one or two order statistics can be found in @Hyndman1996-ty.
+Quantile plots can display any quantiles instead of quartiles in a traditional boxplot.
 These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
@@
 <!-- Description and lit review of density plots -->
- A Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups.
-Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap.
+ A Ridge line plot (sometimes called Joy plot) shows the distribution for several groups.
+Distribution can be represented using density plots, all aligned to the same horizontal scale and presented with a slight overlap.
 Like other density based displays, these plots allow us to see multimodality in the distribution.

### 2019-12-15 main.Rmd
minor
visualization section update

@@
 The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, suggests that graphics are made up of distinct layers of grammatical elements.
-Briefly, the grammatical elements of any plot are as follows:
-
-- **Layers**:  made up of data, mapping, statistical transformation, geometric objects and position adjustments
-
-  - **Data**: observations made on a number of variables, typically thought of as a sample from a population distribution
-
-  - **Mappings** describing how variables are mapped to elements of the plot, such as horizontal axis or color
-
-  - **Statistical transformation** can summarize the data in many useful ways. The geometric object determines the physical representation of the data in the graph.
-
-  - **Geometric objects** control the type of plot that is created and are classified by their dimensions, for example points (0 dimension), line (1 dimension), polygons(2 dimensions) and so on.
-
-  - **Position adjustment** of elements on the plot are needed for dense data, otherwise data points might obscure one another.
-
-- **Facet** dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
-
-- **Scale** controls how data is mapped to aesthetic attributes. For every aesthetic property employed in a layer, we need one scale.
-
-- **A coordinate system** maps the position of objects onto the plane of the plot, and controls how the axes and grid lines are drawn.
-
-Drawing from the grammar of graphics, if <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ serves as the basis of visualizing the distribution of the measured variable, the following components of the layered grammar can be specified:
-
-- Data: <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$
+There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
+Data are observations made on a number of variables, typically thought of as a sample from a population distribution.
+Aesthetic mapping describes how variables are mapped to elements of the plot, such as horizontal axis or color.
+Geometries determine the physical representation of the data in the plot.
+The remaining grammatical elements of a graphic namely facets, statistics, coordinates and themes are optional and contains detail of the plot.
+Facets dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
+
+<!-- Briefly, the grammatical elements of any plot are as follows: -->
+
+<!-- - **Layers**:  made up of data, mapping, statistical transformation, geometric objects and position adjustments -->
+
+<!--   - **Data**: observations made on a number of variables, typically thought of as a sample from a population distribution -->
+
+<!--   - **Mappings** describing how variables are mapped to elements of the plot, such as horizontal axis or color -->
+
+<!--   - **Statistical transformation** can summarize the data in many useful ways. The geometric object determines the physical representation of the data in the graph. -->
+
+<!--   - **Geometric objects** control the type of plot that is created and are classified by their dimensions, for example points (0 dimension), line (1 dimension), polygons(2 dimensions) and so on.
+-->
+
+<!--   - **Position adjustment** of elements on the plot are needed for dense data, otherwise data points might obscure one another. -->
+
+<!-- - **Facet** dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
+-->
+
+<!-- - **Scale** controls how data is mapped to aesthetic attributes. For every aesthetic property employed in a layer, we need one scale.  -->
+
+<!-- - **A coordinate system** maps the position of objects onto the plane of the plot, and controls how the axes and grid lines are drawn.  -->
+
+Drawing from the grammar of graphics, if <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
+
+- Data: <$A_i$, $B_j$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ 
 - Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position
-- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution
-- Geometric objects: examples can include boxplot, letter value, violin, ridge, highest density region plots
-- Facet:  $B_j(s)$
+- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v(s)$
+- Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
+- Facet:  $B_j$
 
@@
 
-## Choice of plots
+## Choice of statistical transformations and geometric objects
 
@@
 Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization.
-The basic plot choice is display of density.
+The basic plot choice for our data structure is the one that can display densities.
 In Section \ref{sec:measured-variable}, it is dicussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
@@
 
-Each type of density display has different parameters.
-These are quantities that need to be estimated given the data.
+Each type of density display has different parameters, that need to be estimated given the data.
 Each is equipped with some benefits and challenges.
@@
 
-### Effect of levels of cyclic granularities
+## Levels within facet and aesthetic variables
 
@@
 
-## Effect of synergy of cyclic granularities
+## Interaction of facet and aesthetic variables
 
@@
 
-## Effect of number of observations
+## Statistical transformation computation
 

### 2019-12-15 main.Rmd
interchangeability section update

@@
 
-## Levels within facet and aesthetic variables
+## Facet and aesthetic variables
+
+### Levels
 
@@
 
-## Interaction of facet and aesthetic variables
+### Interaction
 
@@
 
-## Statistical transformation computation
-
-Even with harmonies, visualizing probability distributions can be misleading either due to rarely occuring categories or unevenly distributed events.
+### Interchangeability
+
+When we consider the layers of grammars considered for our data structure <$A_i$, $B_j$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, mapping $A_i$ mapped to $x$ position and $B_j$ to facets
+suggests that levels of $A_i$'s which are placed in close proximity are to be compared against each other and each level of $B_j$ represent a group.
+When $A_i$ mapped to facets and $B_j$ is mapped to $x$ axis, distributions are compared across levels of $B_j$ and levels if $A_i$ start acting as a group.
+However, the variables used in the layers are the same, the difference in mapping emphasizes different behavior of the variables.
+
+## Number of observations and statistical transformations
+
+Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occuring categories or unevenly distributed events.
 
@@
 If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
-The value of $k$ will depend on what type of visualization we are producing.
-For a decile plot, even $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
+The value of $k$ will depend on the statistical transformation we are producing.
+For computing deciles, $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
 Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
@@
 
-Even when there are no rarely occuring events, number of observations might vary hugely within or across each panel.
+Even when there are no rarely occuring events, number of observations might vary hugely within or across each facet.
 This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the visualizations that use density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators.
+In such cases, the statistical transformations based on density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators.
 \autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.

### 2019-12-16 main.Rmd
data set changed from Ai, Bj to C1, C2

@@
 Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-This is extending the columns of tsibble by including the cyclic granularities.Now suppose we want to explore the measurement variable $v$ across two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-Also, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
-Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ such sets.
-These sets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities.
+This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities.
+Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
+Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+
+<!--  These subsets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities. -->
 
@@
 
-The way cyclic granularities relate become important when we consider the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ for exploration.
+The way cyclic granularities relate become important when we consider the data structure in \@autoref{fig:data-structure}.
+While considering data sets of the form $<C_1, C_2, v>$, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
+Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
 We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the categories/levels of the cyclic granularities.
@@
 
-Drawing from the grammar of graphics, if <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
-
-- Data: <$A_i$, $B_j$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ 
-- Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position
+Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
+
+- Data: <$C_1$, $C_2$, $v$>
+- Aesthetic mapping: $C_1$ mapped to $x$ position and $v$ to $y$ position
 - Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v(s)$
 - Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
-- Facet:  $B_j$
-
-<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or facetting.
-Facetting creates tables of graphics by displaying the same graph for each subset.
--->
+- Facet:  $C_2$
+
+<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
 
@@
 
-<!-- Aesthetics refers to the scales on to which we want to map our data. we create plots by layering grammarical elements on top of one another and use aesthetic maps -->
+<!-- Aesthetics refers to the scales on to which we want to map our data. we create plots by layering grammatical elements on top of one another and use aesthetic maps -->
 
@@
 
-<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or facetting.
-Facetting creates tables of graphics by displaying the same graph for each subset.
--->
+<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
 <!-- - Facet:  $B_j(s)$ -->
@@
 The basic plot choice for our data structure is the one that can display densities.
-In Section \ref{sec:measured-variable}, it is dicussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
+In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
 The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
@@
 When we consider the layers of grammars considered for our data structure <$A_i$, $B_j$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, mapping $A_i$ mapped to $x$ position and $B_j$ to facets
-suggests that levels of $A_i$'s which are placed in close proximity are to be compared against each other and each level of $B_j$ represent a group.
+suggests that levels of $C_1$, that is $A_i$'s which are placed in close proximity are to be compared against each other and each level of $B_j$ represent a group.
 When $A_i$ mapped to facets and $B_j$ is mapped to $x$ axis, distributions are compared across levels of $B_j$ and levels if $A_i$ start acting as a group.

### 2019-12-16 main.Rmd
minor corrections

@@
 - Aesthetic mapping: $C_1$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v(s)$
+- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v$
 - Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
@@
 The basic plot choice is display of density.
-Different plots might be appropriate depending on which features of the density we are interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
+Different plots might be appropriate depending on which features of the density we arae interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
 We will look at each of these aspects separately and analyze the effect of each on visualization.
@@
 
-When we consider the layers of grammars considered for our data structure <$A_i$, $B_j$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, mapping $A_i$ mapped to $x$ position and $B_j$ to facets
-suggests that levels of $C_1$, that is $A_i$'s which are placed in close proximity are to be compared against each other and each level of $B_j$ represent a group.
-When $A_i$ mapped to facets and $B_j$ is mapped to $x$ axis, distributions are compared across levels of $B_j$ and levels if $A_i$ start acting as a group.
-However, the variables used in the layers are the same, the difference in mapping emphasizes different behavior of the variables.
+When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets
+$A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
+Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
+Hence, in this case $A_i$'s are compared against each other within each group.
+When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$.
+Although the variables used in the layers are the same, the difference in mapping leads to emphasis of different behavior of the variables.
 

### 2019-12-17 main.Rmd
section heading changed to mapping
changing structure of smart meter application by bringing all households first

@@
 
+We will discuss the effect of the levels of cyclic granularities in this section.
 The levels of the two cyclic granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
@@
 
+We will discuss the effect of the interaction of cyclic granularities in this section.
 In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
@@
 
-### Interchangeability
-
-When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets
-$A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
+### Mapping
+
+We will discuss the effect of the mapping of cyclic granularities in this section.
+When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
@@
 
+<!-- # how data looks -->
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.
-It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns and identify the extreme and regular households amongst these 50 households.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption in `general_supply_kwh` in every half hour from February 2012 to March 2014.
+\ref{fig:smart-meter} shows the data structure of the available data.
+It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
+The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns and identify the extreme and regular households.
+We take a random sample of 50 households to see how we can achieve that.
+
+```{r smart-meter, fig.cap = "Smart meter data structure"}
+
+smart_meter <- 
+  read_rds("data/sm_cust50.rds")
+```
+
+<!-- how gravitas is used to deal with systematic exploration -->
+
+**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
+If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
+
+```{r}
+library(tsibble)
+interval(smart_meter10)
+smart_meter10 %>% search_gran()
+```
+
+The interval of this tsibble is 30 minutes, and hence the temporal granularities range from half-hour to year.
+If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
+
+```{r}
+smart_meter10 %>% 
+  search_gran(highest_unit = "month")
+```
+
+Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
+
+```{r search_gran_limit2}
+
+smart_meter10 %>% search_gran(highest_unit = "month",
+                              filter_out = c("hhour", "fortnight")
+)
+```
+
+Now that we have the list of cyclic granularities to look at, we can visualize the distribution.
+From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
+The distribution of energy needs to be visualized across two cyclic granularities at a time.
+It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.However, harmony/clash pairs can identified among those 30 pairs, to determine feasibility of plotting any pairs together.
+We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
+
+<!-- on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.  -->
 
@@
 
-**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
-
-```{r}
-library(tsibble)
-interval(smart_meter10)
-smart_meter10 %>% search_gran()
-```
-
-The interval of this tsibble is 30 minutes, and hence the temporal granularities range from half-hour to year.
-If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
-
-```{r}
-smart_meter10 %>% 
-  search_gran(highest_unit = "month")
-```
-
-Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
-
-```{r search_gran_limit2}
-
-smart_meter10 %>% search_gran(highest_unit = "month",
-                              filter_out = c("hhour", "fortnight")
-)
-```
-
-Now that we have the list of cyclic granularities to look at, we can visualize the distribution.
-From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-The distribution of energy needs to be visualized across two cyclic granularities at a time.
-It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.However, harmony/clash pairs can identified among those 30 pairs, to determine feasibility of plotting any pairs together.
-We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
-
 <!-- # ```{r is_harmony, echo= TRUE} -->
@@
 The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
-Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs for improving the analysis of smart meter application.
+Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
 This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
-The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all required materials required to reproduce this article.
+The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article.
 

### 2019-12-17 main.Rmd
structure change of smart meter section - talking about paradox

@@
 
-smart_meter <- 
+sm <- 
   read_rds("data/sm_cust50.rds")
+
+smart_meter <- sm %>% as_tsibble()
+
+smart_meter
 ```
@@
 If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
-
-```{r}
-library(tsibble)
-interval(smart_meter10)
-smart_meter10 %>% search_gran()
-```
-
 The interval of this tsibble is 30 minutes, and hence the temporal granularities range from half-hour to year.
 If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
+
+```{r}
+library(gravitas)
+smart_meter %>% search_gran()
+```
 
@@
 
-Now that we have the list of cyclic granularities to look at, we can visualize the distribution.
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using \autoref{}.
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-The distribution of energy needs to be visualized across two cyclic granularities at a time.
-It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.However, harmony/clash pairs can identified among those 30 pairs, to determine feasibility of plotting any pairs together.
+But the distribution of energy needs to be visualized across two cyclic granularities at a time.
+It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.
+However, harmony/clash pairs can be identified among those 30 pairs, to determine feasibility of plotting any pairs together.
 We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
+We will look at few harmony pairs and see how the proposed workflow can be utilized.
+
+Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of 50 households.
+Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the fact
+
+Figure \ref{fig:my-hd} shows that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
+Moreover, only the pink band changes significantly in winter months (May - August).
+The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.
+Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
+It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+
+```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
+The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."}
+
+smart_meter <- 
+  read_rds("data/sm_cust50.rds")
+
+smart_meter %>% 
+  prob_plot("wknd_wday",
+                            "hour_day",
+                            response = "general_supply_kwh",
+                            plot_type = "quantile",
+            symmetric = FALSE, 
+            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
+  scale_y_sqrt()
+```
+
+<!-- # ```{r wknd-wday50_df} -->
+<!-- # smart_meter10 %>%  -->
+<!-- #   prob_plot("month_year", -->
+<!-- #                             "day_fortnight", -->
+<!-- #                             response = "general_supply_kwh", -->
+<!-- #                             plot_type = "quantile", -->
+<!-- #             symmetric = TRUE,  -->
+<!-- #             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
+<!-- #   scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3)) -->
+<!-- # ``` -->
+
+```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
+The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
+"}
+
+library(gravitas)
+library(tsibble)
+
+library(ggplot2)
+smart_meter %>%
+  prob_plot("month_year", "hour_day",
+    response = "general_supply_kwh",
+    plot_type = "quantile",
+    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99),
+    symmetric = TRUE) +
+  ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3))
+```
+
+<!-- Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
+This is true for both top 50% and 75% of the customers.
+Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
+All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
+-->
+
+<!-- ```{r dw-hd} -->
+
+<!-- # Hourly usage across days of the week -->
+<!-- sm_cust50 %>% -->
+<!--   prob_plot("day_week",  -->
+<!--            "hour_day", -->
+<!--            "general_supply_kwh",  -->
+<!--            plot_type = "quantile",  -->
+<!--           quantile_prob = c(0.25, 0.5, 0.75), -->
+<!--            response = "general_supply_kwh", overlay = FALSE) + -->
+<!--   ggtitle("") -->
+
+<!-- ``` -->
 
@@
 
-Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of 50 households.
-Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends.
-
-Figure \ref{fig:my-hd} shows that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-Moreover, only the pink band changes significantly in winter months (May - August).
-The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.
-Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
-
-```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
-The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."}
-
-smart_meter <- 
-  read_rds("data/sm_cust50.rds")
-
-smart_meter %>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "quantile",
-            symmetric = FALSE, 
-            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
-  scale_y_sqrt()
-```
-
-<!-- # ```{r wknd-wday50_df} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   prob_plot("month_year", -->
-<!-- #                             "day_fortnight", -->
-<!-- #                             response = "general_supply_kwh", -->
-<!-- #                             plot_type = "quantile", -->
-<!-- #             symmetric = TRUE,  -->
-<!-- #             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
-<!-- #   scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3)) -->
-<!-- # ``` -->
-
-```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
-"}
-
-library(gravitas)
-library(tsibble)
-
-library(ggplot2)
-smart_meter %>%
-  prob_plot("month_year", "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99),
-    symmetric = TRUE) +
-  ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3))
-```
-
-<!-- Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
-This is true for both top 50% and 75% of the customers.
-Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
-All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
--->
-
-<!-- ```{r dw-hd} -->
-
-<!-- # Hourly usage across days of the week -->
-<!-- sm_cust50 %>% -->
-<!--   prob_plot("day_week",  -->
-<!--            "hour_day", -->
-<!--            "general_supply_kwh",  -->
-<!--            plot_type = "quantile",  -->
-<!--           quantile_prob = c(0.25, 0.5, 0.75), -->
-<!--            response = "general_supply_kwh", overlay = FALSE) + -->
-<!--   ggtitle("") -->
-
-<!-- ``` -->
-
 This case study shows systematic exploration of energy behavior starting with two households and comparing some of their features with all 50 households.

### 2020-01-16 main.Rmd
application smart meter data structure

@@
 
+Data structure for visualization:
+
+The variable `reading_datetime` is the index, which represents the time variable.
+The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed.
+Data sets of the form <C1, C2, `general_supply_kwh`> will form the basis of exploration for each key in the data set, the cyclic granularity C1 mapped to x axis and `general_supply_kwh` to y-axis with the cyclic granularity C2 as the facet.
+Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
+ridge or highest density region plots can then be used to explore the data set.
+
+Cyclic granularities search and computation:
+
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
+
 ```{r smart-meter, fig.cap = "Smart meter data structure"}
@@
 **gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
+
+The variable 'reading_datetime` is the index, which represents the time variable.
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following time granularities can be considered for this analysis.
-The interval of this tsibble is 30 minutes, and hence the temporal granularities range from half-hour to year.
-If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
+
+These cyclic granularities are a function of the index variable and are computed based on if they are cicular (@ref:Sec), pseudo-circular (@ref:Sec) or aperiodic (@ref:Sec).
+If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following cyclic time granularities can be considered for this analysis.
+The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.
 
@@
 
+If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
+
 ```{r}
@@
 Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using \autoref{}.
+
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.

### 2020-01-27 main.Rmd
smart meter section update - aligning it with how paper flows

@@
 The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns and identify the extreme and regular households.
-We take a random sample of 50 households to see how we can achieve that.
-
-Data structure for visualization:
-
-The variable `reading_datetime` is the index, which represents the time variable.
+
+<!-- We take a random sample of 50 households to see how we can achieve that.  -->
+
+### Data structure for visualization:
+
+The variable `reading_datetime` is the index of the tsibble and represents the time variable.
 The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed.
-Data sets of the form <C1, C2, `general_supply_kwh`> will form the basis of exploration for each key in the data set, the cyclic granularity C1 mapped to x axis and `general_supply_kwh` to y-axis with the cyclic granularity C2 as the facet.
+Data sets of the form <C1, C2, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity C1 would be mapped to x axis and `general_supply_kwh` to y-axis with the cyclic granularity C2 mapped to facet.
 Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
@@
 
-Cyclic granularities search and computation:
-
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-
 ```{r smart-meter, fig.cap = "Smart meter data structure"}
@@
 
+### Cyclic granularities search and computation:
+
+**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
+
 <!-- how gravitas is used to deal with systematic exploration -->
 
-**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
-
-The variable 'reading_datetime` is the index, which represents the time variable.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-
-These cyclic granularities are a function of the index variable and are computed based on if they are cicular (@ref:Sec), pseudo-circular (@ref:Sec) or aperiodic (@ref:Sec).
+These cyclic granularities are a function of the index variable and are computed based on if they are cicular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}).
 If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following cyclic time granularities can be considered for this analysis.
@@
 
-```{r}
+```{r, echo=TRUE}
 library(gravitas)
@@
 
-```{r}
+```{r, echo=TRUE}
 smart_meter10 %>% 
@@
 
-```{r search_gran_limit2}
+```{r search_gran_limit2, echo=TRUE}
 
@@
 
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using \autoref{}.
+### Screening and visualising harmonies:
+
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:asperiodic-gran-def}.
 
@@
 
-Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of 50 households.
-Similar quantiles for weekend and weekdays indicate that behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the fact
-
-Figure \ref{fig:my-hd} shows that the median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-Moreover, only the pink band changes significantly in winter months (May - August).
-The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.
-Most of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of quantiles increased too in winter months (obvious because of weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+```{r harmony, echo=TRUE}
+smart_meter10 %>% harmony(
+  ugran = "month",
+  filter_out = c("hhour", "fortnight")
+)
+smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
+
+```
+
+Each of these 13 pairs need to be visualised to have multiple perspectives on the data.
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
+This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households.
+Quantiles for weekend and weekdays for these customers look exactly the same.
+It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors.
+Consider another example in Figure \ref{fig:my-hd} showing the area quantile plots of 50 customers.
+The median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
+The pink band changes significantly in winter months (May - August) implying that much of the behavioral changes occur in the quartiles, that too in peak hours of the day.
+It is to be noted here that the the level of all quantiles increased too in winter months (increased energy usage in weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+
+<!-- The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.  -->
 
@@
 
-In Figure \ref{fig:day-fortnight}, the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
+Since we established that looking at random samples with different demographics might confound our analysis, next the smart meter data is filtered for two customers in Figure \ref{fig:day-fortnight},to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
@@
 
-```{r harmony, echo=TRUE}
-smart_meter10 %>% harmony(
-  ugran = "month",
-  filter_out = c("hhour", "fortnight")
-)
-smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
-
-```
-
 In \autoref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.

### 2020-01-27 main.Rmd
smart meter section broken up as how paper is organized

@@
 
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:asperiodic-gran-def}.
+
 ### Screening and visualising harmonies:
 
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:asperiodic-gran-def}.
-
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-But the distribution of energy needs to be visualized across two cyclic granularities at a time.
-It is equivalent to taking 2 granularities from 6, which essentially is equivalent to visualizing 30 plots.
+Given the data structure <$C1$ $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
+Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to visualizing 30 plots.
 However, harmony/clash pairs can be identified among those 30 pairs, to determine feasibility of plotting any pairs together.
 We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
-We will look at few harmony pairs and see how the proposed workflow can be utilized.
+
+<!-- We will look at few harmony pairs and see how the proposed workflow can be utilized.  -->
+
+<!-- From the Gestalt theory with the same data  <$C1$ $C2$, `general_supply_kwh`>, mapping $C1$ on the x-axis vs facet might lead to different insights. -->
 
@@
 
-Each of these 13 pairs need to be visualised to have multiple perspectives on the data.
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
+However, if the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
 This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
@@
 
-Now, we would like to see how these customers' behavior relate to the rest of the  50 households across these two measures -
-
-- if energy distribution is skewed towards extreme?
-
-- if the weekend and weekday behavior are different for most of these households?
-
-This case study shows systematic exploration of energy behavior starting with two households and comparing some of their features with all 50 households.
-First, it helps us to find the list of granularities to look at, then shrinks the number of possible visualizations by identifying harmonies and then explored harmony pairs to gain some insights on periodic behavior of the households.
+<!-- Now, we would like to see how these customers' behavior relate to the rest of the  50 households across these two measures -   -->
+
+<!-- - if energy distribution is skewed towards extreme? -->
+
+<!-- - if the weekend and weekday behavior are different for most of these households? -->
+
+This case study shows systematic exploration of energy behavior for two random households comparing some of their features with all 50 households.
+First, it helps us to find the list of granularities to look at, then shrinks the number of possible visualizations by identifying harmonies and lastly visualise harmony pairs to gain some insights on periodic behavior of the households.
 

### 2020-01-27 main.Rmd
application smart meter data structure

@@
 
-Since we established that looking at random samples with different demographics might confound our analysis, next the smart meter data is filtered for two customers in Figure \ref{fig:day-fortnight},to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
+Since we established that looking at random samples with different demographics might confound our analysis, next the smart meter data is filtered for two customers, to illustrate kinds of insights that can be drawn for the energy behavior of these two customers
+
+<!-- can be added later -->
+
+<!-- In Figure \ref{fig:day-fortnight},to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
 it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
@@
 This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
+-->
 
@@
 
-```{r day-fortnight, fig.cap="Letter value plot of two households across days of the fortnight.
+<!-- ```{r day-fortnight, fig.cap="Letter value plot of two households across days of the fortnight.
 M, F, E, D and C represents the letter values of energy consumption.
-Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household."}
-library(lvplot)
-library(ggplot2)
-library(dplyr)
-library(tibble)
-library(gravitas)
-smart_meter10 %>%
-  filter(customer_id %in% c(10006704, 10017936)) %>% 
-  create_gran("day_fortnight") %>%
-  ggplot2::ggplot(aes(
-    x = as.factor(day_fortnight),
-    y = general_supply_kwh)) +
-  xlab("day_fortnight") +
-  geom_lv(
-    outlier.colour = "red",
-    aes(fill = ..LV..),
-    k = 5) +
-  facet_wrap(~customer_id) + 
-  scale_fill_lv() +
-  theme_bw() +
-  scale_y_sqrt() 
-```
+Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household."} -->
+<!-- library(lvplot) -->
+<!-- library(ggplot2) -->
+<!-- library(dplyr) -->
+<!-- library(tibble) -->
+<!-- library(gravitas) -->
+<!-- smart_meter10 %>% -->
+<!--   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
+<!--   create_gran("day_fortnight") %>% -->
+<!--   ggplot2::ggplot(aes( -->
+<!--     x = as.factor(day_fortnight), -->
+<!--     y = general_supply_kwh)) + -->
+<!--   xlab("day_fortnight") + -->
+<!--   geom_lv( -->
+<!--     outlier.colour = "red", -->
+<!--     aes(fill = ..LV..), -->
+<!--     k = 5) + -->
+<!--   facet_wrap(~customer_id) +  -->
+<!--   scale_fill_lv() + -->
+<!--   theme_bw() + -->
+<!--   scale_y_sqrt()  -->
+<!-- ``` -->
 
@@
 In \autoref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.
-Boxplot of energy consumption is shown across wknd_wday (facet) and hour-day (x-axis) for the same two households.
-For the second household, outliers are less prominent implying their regular behavior is more stable.
-For the first household, energy behavior is not significantly different between weekdays and weekends.
-For the second household, median energy consumption for the early morning hours is extremely high for weekends compared to weekdays.
-
- For the second household, outliers are less prominent implying their regular behavior is more stable.
-For the first household, energy behavior is not significantly different between weekdays and weekends.
-For the second household, median energy consumption for the early morning hours is extremely high for weekends compared to weekdays.
+Boxplot of energy consumption is shown across `wknd_wday` (facet) and `hour_day` (x-axis) for the same two households.
+For the first household, the shape of the distribution does not change significantly for weekdays and weekends.
+The distribution is roughly smooth with a morning and an evening peak.
+For the second household, the quantiles behave differently in peak hours.
+For the first few hours on weekdays, energy consumption starts at a much higher level than weekends.
+Consider looking at letter value plots instead of quantile plots to look at the same data in \autoref{fig:bothcust-lv}.
+There is additional information that we can derive looking at the distribution.
+There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day.
+Also hours from 7 to 13 seems most volatile given the shape of the violin.
+If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
+For example in \autoref{fig:cust-lv}, it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
+Also, outliers are more prominent in the latter part of the day.
+All of these indicates that looking at different distribution plots or changing the mapping might draw lights on different aspect of the energy behavior for the same customer.
+
+<!-- If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
+However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
+This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
+-->
 
@@
                             response = "general_supply_kwh",
-                            plot_type = "boxplot") +
+                            plot_type = "quantile",
+            symmetric = FALSE,
+            quantile_prob = c(0.25, 0.5, 0.75, 0.9)) +
   scale_y_sqrt() +
@@
                             response = "general_supply_kwh",
-                            plot_type = "boxplot") +
+                            plot_type = "quantile",
+            symmetric = FALSE,
+             quantile_prob = c(0.25, 0.5, 0.75, 0.9)) +
   scale_y_sqrt() + 
@@
 cust2
+```
+
+```{r bothcust-lv, fig.cap ="towrite."}
+# 
+# smart_meter10 %>% 
+#   filter(customer_id %in% c(10006704)) %>% 
+#   prob_plot("wknd_wday",
+#                             "hour_day",
+#                             response = "general_supply_kwh",
+#                             plot_type = "violin") +
+#   scale_y_sqrt() +
+#   ggtitle("Energy consumption distribution for customer id: 10006704")
+
+smart_meter10 %>% 
+  filter(customer_id %in% c(10017936))%>% 
+  prob_plot("wknd_wday",
+                            "hour_day",
+                            response = "general_supply_kwh",
+                            plot_type = "violin") +
+  scale_y_sqrt() + ggtitle("Energy consumption distribution for customer id: 10017936")
+```
+
+```{r cust-ridge, fig.cap ="towrite something"}
+# 
+# smart_meter10 %>% 
+#   filter(customer_id %in% c(10006704)) %>% 
+#   prob_plot("wknd_wday",
+#                             "hour_day",
+#                             response = "general_supply_kwh",
+#                             plot_type = "violin") +
+#   scale_y_sqrt() +
+#   ggtitle("Energy consumption distribution for customer id: 10006704")
+
+smart_meter10 %>% 
+  filter(customer_id %in% c(10017936))%>% 
+  prob_plot("hour_day", 
+            "wknd_wday",
+            response = "general_supply_kwh",
+            plot_type = "boxplot") +
+            scale_y_sqrt() +
+            # scale_fill_brewer(palette = "Dark2")
+  ggtitle("Energy consumption distribution for customer id: 10017936")
 ```

### 2020-01-28 main.Rmd
visulaisation section minor changes

@@
 
-The grammar of graphics introduced a framework to incorporate underlying features for statistical graphics to construct them by relating data space to the graphic space [@Wilkinson1999-nk]. 
-The layered grammar of graphics proposed by [@Wickham2009-pk], which is a modification over the former, suggests that graphics are made up of distinct layers of grammatical elements.
+The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk].
+The layered grammar of graphics proposed by [@Wickham2009-pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
 There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
@@
 In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
-The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
-However, they avoid much clutter and help us to focus on some specific properties of the data.
 Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
-In practice, the densities are estimated reasonably with at least 30 observations.
-However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data.
-The summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.
+The descriptive statistics based methods do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
+However, they avoid clutter and help us to focus on some specific properties of the data.
+Also, the summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.
+
+<!--  However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. -->
 

### 2020-01-28 main.Rmd
section levels update

@@
 
-Quantile plots can display any quantiles instead of quartiles in a traditional boxplot.
+Quantile plots display quantiles instead of quartiles in a traditional boxplot.
 These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
-When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
-outside 1.5 times the interquartile range above the upper quartile and below the lower quartile).
-However, outliers are open to interpretation and not shown in an quantile plot.
+While in a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
+outside 1.5 times the interquartile range above the upper quartile and below the lower quartile), in quantile plots outliers are open to interpretation and not shown.
 
@@
 We will discuss the effect of the levels of cyclic granularities in this section.
-The levels of the two cyclic granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
-The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.
-Plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.
-
-Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are chosen based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-Any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
+Recall that cyclic granularities $C_1$ and $C_2$ are such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
+$A_i$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.
+The levels have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
+Plot choices will also depend on the aesthetic mapping of the cyclic granularities.
+A potential approach can be to categorize the levels as very high/high/medium/low for each cyclic granularities and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
+Default values for these categorisations can be chosen based on levels of common temporal granularities like days-of-month, days-of-fortnight and days-of-week.
+For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.
+
+<!-- The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.  -->
 

### 2020-01-30 main.Rmd
header change
minor changes and names of sub headers in smsrt meter

@@
 
-### Mapping
+### Interchangeability of mappings
 
@@
 One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption in `general_supply_kwh` in every half hour from February 2012 to March 2014.
-\ref{fig:smart-meter} shows the data structure of the available data.
+\ref{tab:smart-meter} shows the data structure of the available data.
 It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
@@
 
-smart_meter
+smart_meter %>% kable()
 ```
@@
 
-**gravitas** R package [@R-gravitas] is used to facilitate the systematic exploration here.
+R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here.
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.

### 2020-01-30 main.Rmd
commit after long delays of Rstudio

@@
 The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed.
-Data sets of the form <C1, C2, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity C1 would be mapped to x axis and `general_supply_kwh` to y-axis with the cyclic granularity C2 mapped to facet.
+Data sets of the form <$C1$, $C2$, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity $C1$ would be mapped to $x$ axis and `general_supply_kwh` to $y$ axis with the cyclic granularity $C2$ mapped to facet.
 Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
@@
 
-smart_meter <- sm %>% as_tsibble()
-
-smart_meter %>% kable()
+smart_meter <- sm %>% as_tsibble() %>% dplyr::select(customer_id, 
+                              reading_datetime, 
+                              general_supply_kwh, 
+                              everything()) %>% head()
+
+knitr::kable(smart_meter, format = "markdown")
 ```
@@
 
-```{r, echo=TRUE}
+```{r search, echo=TRUE}
 library(gravitas)
-smart_meter %>% search_gran()
-```
-
-If these options are considered too many, the most coarse temporal unit can be set to be a “month”.
-
-```{r, echo=TRUE}
-smart_meter10 %>% 
+smart_smart_search <- smart_meter %>% search_gran()
+knitr::kable(smart_smart_search, format = "markdown")
+```
+
+If these options are considered too many, the most coarse temporal unit can be set to a “month”.
+
+```{r search_limit1, echo=TRUE}
+smart_meter_limit1 <- smart_meter10 %>% 
   search_gran(highest_unit = "month")
+knitr::kable(smart_meter_limit1, format = "markdown")
 ```
@@
 
-smart_meter10 %>% search_gran(highest_unit = "month",
+smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month",
                               filter_out = c("hhour", "fortnight")
 )
+knitr::kable(smart_meter_limit2, format = "markdown")
 ```
@@
 Given the data structure <$C1$ $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
-Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to visualizing 30 plots.
-However, harmony/clash pairs can be identified among those 30 pairs, to determine feasibility of plotting any pairs together.
+Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualisation.
+However, harmony/clash pairs can be identified among those 30 possibilties to determine feasibility of plotting any pairs together.
 We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
@@
 )
-smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
-
-```
-
-However, if the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
+# smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
+
+```
+
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
 This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
 For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households.
-Quantiles for weekend and weekdays for these customers look exactly the same.
+Quantiles for weekends and weekdays for these customers look exactly the same.
 It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors.
@@
 
-Since we established that looking at random samples with different demographics might confound our analysis, next the smart meter data is filtered for two customers, to illustrate kinds of insights that can be drawn for the energy behavior of these two customers
+Looking at random samples with different demographics might confound our analysis and a strategy for analyzing multiple keys together could be to first screen some basis time series or demographic features and then look at their energy behavior.
+This is beyond the scope of the current work.
+For the purpose of this paper, the smart meter data is filtered for two customers to illustrate kinds of insights that can be drawn after proceeding with the systematic exploration.
 
@@
 There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day.
-Also hours from 7 to 13 seems most volatile given the shape of the violin.
+Also hours from 7 to 13 look most volatile.
 If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
@@
 
-This case study shows systematic exploration of energy behavior for two random households comparing some of their features with all 50 households.
-First, it helps us to find the list of granularities to look at, then shrinks the number of possible visualizations by identifying harmonies and lastly visualise harmony pairs to gain some insights on periodic behavior of the households.
+This case study shows systematic exploration of energy behavior for two random households to gain some insights on periodic behavior of the households.
+First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualise a harmony pair and shows the effect of different distribution plots or reverse mapping.
 

### 2020-01-30 main.Rmd
all kables removed in smart meter section

@@
                               everything()) %>% head()
-
-knitr::kable(smart_meter, format = "markdown")
+smart_meter
+# knitr::kable(smart_meter, format = "markdown")
 ```
@@
 library(gravitas)
-smart_smart_search <- smart_meter %>% search_gran()
-knitr::kable(smart_smart_search, format = "markdown")
+smart_meter %>% search_gran()
+#smart_smart_search <- smart_meter %>% search_gran()
+#knitr::kable(smart_smart_search, format = "markdown")
 ```
@@
 ```{r search_limit1, echo=TRUE}
-smart_meter_limit1 <- smart_meter10 %>% 
+smart_meter10 %>% 
   search_gran(highest_unit = "month")
-knitr::kable(smart_meter_limit1, format = "markdown")
+# smart_meter_limit1 <- smart_meter10 %>% 
+#   search_gran(highest_unit = "month")
+# knitr::kable(smart_meter_limit1, format = "markdown")
 ```
@@
 ```{r search_gran_limit2, echo=TRUE}
-
-smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month",
+smart_meter10 %>% search_gran(highest_unit = "month",
                               filter_out = c("hhour", "fortnight")
                               )
-knitr::kable(smart_meter_limit2, format = "markdown")
+# smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month",
+#                               filter_out = c("hhour", "fortnight")
+# )
+# knitr::kable(smart_meter_limit2, format = "markdown")
 ```

### 2020-01-30 main.Rmd
cricket intro change

@@
 
-The application is not only restricted to temporal data.
+The method is not only restricted to temporal data.
 We provide an example of cricket to illustrate how this can be generalized in other applications.
+Although there is no conventional time component in cricket, each ball can be thought to represent a basic unit of time as the game progresses forward with each ball.
+In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season can consist of several matches.
+Thus similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
+An idea of cyclic granularities can be conceived with this hierarchy in mind.
+Examples include ball of the over, over of the innings, ball of the innings.
+Although these cyclic granularities are circular by design of the hierarchy, there might be cases where these are aperiodic.
+As an example, in most cases an over will consist of 6 balls with some exceptions of in case of a wide ball or if the innings finish before the over finishes.
+Irrespective of the type of cyclic granularities, it can be interesting to visualize the periodic behavior of scores/performance under similar data structure as temporal datasets.
+
 The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
-In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season can consist of several matches.
 The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
 The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016), each of which has 6 balls in an over and 2 innings in a match.
-s
 

### 2020-01-30 main.Rmd
cricket section making concise - removing the pattern across season

--- 2020-01-30 main.Rmd
@@
 
-The application is not only restricted to temporal data.
-We provide an example of cricket to illustrate how this can be generalized in other applications.
+The method is not only restricted to temporal data.
+We provide an example of cricket to illustrate how this can be generalized to other applications.
+Although there is no conventional time component in cricket, each ball can be thought to represent an ordering from past to future with the game progressing forward with each ball.
+In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
+Thus similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
+An idea of cyclic granularities can be conceived with this hierarchy in mind.
+Examples may include ball of the over, over of the innings or ball of the innings.
+Although most of these cyclic granularities are circular in theory by design of the hierarchy, in reality these may be aperiodic.
+For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when a inning finishes before the over finishes.
+Thus, the cyclic granularity ball-of-the-over will be circular in most cases and aperiodic in others.
+However, irrespective of the type of cyclic granularities, it can be interesting to visualize the periodicity of  scores/performance under similar data structure as temporal datasets.
+
 The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
-In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season can consist of several matches.
 The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016), each of which has 6 balls in an over and 2 innings in a match.
-s
+The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 5 overs and each match has 2 innings.
 
@@
 
-Although there is no conventional time granularity in cricket, we can still represent the data set `cricket` through a `tsibble`, with each over representing an ordering from past to future, can form the index of the tsibble.
-The hierarchy table would look like the following:
-
-```{r hierarchy2}
-library(gravitas)
-library(tibble)
-hierarchy_model <- tibble::tibble(
-  G = c("over", "inning", "match", "season"),
-  C = c("over-of-inning", "inning-of-match", "match-of-season",1),
-  P = c(20, 2, "aperiodic", 1))
-knitr::kable(hierarchy_model, format = "markdown")
-
-```
+<!-- The hierarchy table would look like the following: -->
+
+<!-- # ```{r hierarchy2} -->
+<!-- # library(gravitas) -->
+<!-- # library(tibble) -->
+<!-- # hierarchy_model <- tibble::tibble( -->
+<!-- #   G = c("over", "inning", "match", "season"), -->
+<!-- #   C = c("over-of-inning", "inning-of-match", "match-of-season",1), -->
+<!-- #   P = c(20, 2, "aperiodic", 1)) -->
+<!-- # knitr::kable(hierarchy_model, format = "markdown") -->
+<!-- #  -->
+<!-- # ``` -->
 
@@
 
-First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperidic cyclic granularity) in \autoref{fig:seas-over-inning}.
+<!-- Making it short -->
+
+<!-- First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperidic cyclic granularity) in \autoref{fig:seas-over-inning}.
 The distribution of runs per over has not significantly changed from 2008 to 2016.
 There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
-Hence, we work with subsets of seasons to answer some of the questions:
+Hence, we work with subsets of seasons to answer some of the questions: -->
 
@@
 
-```{r runs_per-over}
-library(tsibble)
-
-cricket_all <-  read_csv("data-raw/deliveries_all.csv")
-matches_all <- read_csv("data-raw/matches_all.csv")
-
-cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
-
-cricket_per_over <- cricket_season %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))
-
-cricket_tsibble_all <- cricket_per_over %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-```
-
-```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."}
-  cricket_tsibble_all %>% 
-  prob_plot("over", 
-            "season",
-            hierarchy_model,
-            response = "runs_per_over",
-            plot_type = "quantile",
-            quantile_prob = c(0.25, 0.5, 0.75), 
-            symmetric = FALSE)  +
-     theme( # remove the vertical grid lines
-            panel.grid.major.x = element_blank() ,
-            # explicitly set the horizontal lines (or they will disappear too)
-            panel.grid.major.y = element_blank()) + scale_x_discrete(breaks = seq(2008, 2016, 3))
-```
+<!-- # ```{r runs_per-over} -->
+<!-- # library(tsibble) -->
+<!-- #  -->
+<!-- # cricket_all <-  read_csv("data-raw/deliveries_all.csv") -->
+<!-- # matches_all <- read_csv("data-raw/matches_all.csv") -->
+<!-- #  -->
+<!-- # cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id")) -->
+<!-- #                                          -->
+<!-- # cricket_per_over <- cricket_season %>%  -->
+<!-- #   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
+<!-- #   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))  -->
+<!-- #  -->
+<!-- # cricket_tsibble_all <- cricket_per_over %>%  -->
+<!-- #   ungroup() %>%  -->
+<!-- #   mutate(data_index = row_number()) %>%  -->
+<!-- #   as_tsibble(index = data_index) -->
+<!-- # ``` -->
+<!-- #  -->
+<!-- #  -->
+<!-- # ```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."} -->
+<!-- #   cricket_tsibble_all %>%  -->
+<!-- #   prob_plot("over",  -->
+<!-- #             "season", -->
+<!-- #             hierarchy_model, -->
+<!-- #             response = "runs_per_over", -->
+<!-- #             plot_type = "quantile", -->
+<!-- #             quantile_prob = c(0.25, 0.5, 0.75),  -->
+<!-- #             symmetric = FALSE)  + -->
+<!-- #      theme( # remove the vertical grid lines -->
+<!-- #             panel.grid.major.x = element_blank() , -->
+<!-- #             # explicitly set the horizontal lines (or they will disappear too) -->
+<!-- #             panel.grid.major.y = element_blank()) + scale_x_discrete(breaks = seq(2008, 2016, 3)) -->
+<!-- # ``` -->
 

### 2020-01-30 main.Rmd
section on cricket updated

@@
 Mumbai Indians (MI) and Chennai Super kings (CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final playoffs from 2010 to 2015.
-It would be interesting to take their example in order to dive deeper into the first question.
+It would be interesting to take their example in order to dive deeper into this question.
 
@@
 
-<!-- # ```{r runs_per-over} -->
-<!-- # library(tsibble) -->
-<!-- #  -->
-<!-- # cricket_all <-  read_csv("data-raw/deliveries_all.csv") -->
-<!-- # matches_all <- read_csv("data-raw/matches_all.csv") -->
-<!-- #  -->
-<!-- # cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id")) -->
-<!-- #                                          -->
-<!-- # cricket_per_over <- cricket_season %>%  -->
-<!-- #   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
-<!-- #   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))  -->
-<!-- #  -->
-<!-- # cricket_tsibble_all <- cricket_per_over %>%  -->
-<!-- #   ungroup() %>%  -->
-<!-- #   mutate(data_index = row_number()) %>%  -->
-<!-- #   as_tsibble(index = data_index) -->
-<!-- # ``` -->
-<!-- #  -->
-<!-- #  -->
+```{r runs_per-over}
+ library(tsibble)
+
+ cricket_all <-  read_csv("data-raw/deliveries_all.csv")
+ matches_all <- read_csv("data-raw/matches_all.csv")
+
+ cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
+
+ cricket_per_over <- cricket_season %>%
+   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))
+
+ cricket_tsibble_all <- cricket_per_over %>%
+   ungroup() %>%
+   mutate(data_index = row_number()) %>%
+   as_tsibble(index = data_index)
+```
+#
+#
 <!-- # ```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."} -->
@@
 
-From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in runs in the second innings as compared to the first innings.
-The variability of runs also increases as the teams approach towards the end of the innings, as observed through the longer and more distinct letter values.
+From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
+This might indicate that palyers are vulnerable to score through each over and make more runs in the first innings, whereas if they in the second innings they are less vulnberable as they have a target in mind.
+Also longer and more distinct letter values in the second innings suggests that the variability is more in the second innings.
+To answer the question, winning teams like CSK and MI have indeed different startegies when it comes to batting in the first or second innings.
 
@@
   hierarchy_model,
-  plot_type = "lv")  + ggtitle("")
+  plot_type = "lv")  + ggtitle("") +
+  scale_fill_brewer(palette = "Dark2")
+
 ```
@@
 cricket_tsibble_fld %>% 
-  filter(fielding_wckts %in% c(0,1)) %>% 
-  prob_plot( "fielding_wckts","over",
+  mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets")) %>% 
+  #filter(fielding_wckts %in% c(0,1)) %>% 
+  prob_plot( "field","over",
                                hierarchy_model,
                                response = "diff_run_rate",
-                               plot_type = "quantile",
+                               plot_type = "lv",
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
@@
   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(dot_balls = sum(dot_ball_proxy),
-            wickets = sum(wicket_proxy), 
-            total_runs = sum(total_runs)) %>%
-            mutate(diff_run_rate = c(0,diff(total_runs)))
+  summarise(runs_per_over = sum(total_runs),
+            run_rate = sum(total_runs)/length(total_runs),
+            dot_balls = sum(dot_ball_proxy)) %>% 
+            mutate(diff_run_rate = c(0,diff(run_rate)))
 
@@
 cricket_tsibble_bow %>%
-  filter(dot_balls %in% c(0, 1, 2)) %>%
-  prob_plot("dot_balls",
+  mutate(dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) %>% 
+  #filter(dot_balls %in% c(0, 1, 2)) %>%
+  prob_plot("dot",
             "over",

### 2020-01-30 main.Rmd
changing cricket intro and figures - need updation of caption

@@
                                response = "diff_run_rate",
-                               plot_type = "lv",
+                               plot_type = "boxplot",
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),

### 2020-01-30 main.Rmd
caption changed - cricket analysis

@@
 Although most of these cyclic granularities are circular in theory by design of the hierarchy, in reality these may be aperiodic.
-For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when a inning finishes before the over finishes.
+For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
 Thus, the cyclic granularity ball-of-the-over will be circular in most cases and aperiodic in others.
@@
 The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 5 overs and each match has 2 innings.
+The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
+There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
 
@@
 <!-- # ``` -->
-
-There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
 
@@
 From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
-This might indicate that palyers are vulnerable to score through each over and make more runs in the first innings, whereas if they in the second innings they are less vulnberable as they have a target in mind.
+This might indicate that palyers are vulnerable to score through each over and make more runs in the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
 Also longer and more distinct letter values in the second innings suggests that the variability is more in the second innings.
-To answer the question, winning teams like CSK and MI have indeed different startegies when it comes to batting in the first or second innings.
+To answer Q1, winning teams like CSK and MI have indeed different startegies when it comes to batting in the first or second innings.
 
@@
 Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.
-We can verify this fact through Figure \ref{fig:field} which shows that the difference in runs between two subsequent overs are negative when good fielding leads to one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
-
-```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and dismissals by catch and catch and bowled.
-This shows good fielding in one over leads to lower runs in the subsequent overs for at least 50 percentiles of times."}
+We can verify this fact through Figure \ref{fig:field} which shows that the difference in run rates between two subsequent overs are negative when good fielding leads to at least one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
+
+```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+This shows good fielding in previous over leads to lower run rate in the subsequent over.
+Also run rates are much less volatile in case there are at least one wicket in the previous over."}
 
@@
 
-```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over.
-For all three cases, difference in runs between subsequent overs decrease.", fig.pos="ht"}
+```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), more than zero (facet 2) dot balls per over.
+For all quantiles, difference in runs between subsequent overs decrease with at least one dot ball in the previous over.", fig.pos="ht"}
 
@@
 In Q2 and Q3, we considered dots balls per over or catches per over as aperiodic cyclic granularities.
-They do not appear in the hierarchy table since we can not compute these granularities as a function of circular or quasi-circular granularities.
+They do not appear in the hierarchy since we can not compute these granularities as a function of circular or quasi-circular granularities.
 Also, their order in the hierarchy table is not clear as they are essentially obtained from some aperiodic linear granularity.
+These are similar to holidays in temporal data which also can not be placed at a fixed order in a hierarchy table in case they are not finer than or groups into other temporal units.
 

### 2020-01-31 main.Rmd
caption changed - smart meter

@@
 
-```{r bothcust, fig.cap = "Boxplot of energy consumption across hours of the day faceted by weekday/weekend for household 1 (customer id: 10006704) and household 2 (customer id: 10017936).
+```{r bothcust, fig.cap = "Quantile plots of energy consumption across hours of the day faceted by weekday/weekend for household 1 (customer id: 10006704) and household 2 (customer id: 10017936).
 Median behavior for the early morning hours of household 2 is extremely high for weekends compared to weekdays, not much differenc can be observed for household 1.
@@
 
-```{r bothcust-lv, fig.cap ="towrite."}
+```{r bothcust-lv, fig.cap ="Violin plots of energy consumption across hours of the day faceted by weekday/weekend for household 2 (customer id: 10017936).
+Early morning hours are bimodal implying both high and low consumption are probable.
+Volatility is high from 7 to 13 hours as evident from the narrow violins."}
 # 
@@
 
-```{r cust-ridge, fig.cap ="towrite something"}
+```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for household 2 (customer id: 10017936).
+Median behavior for the early morning hours of household 2 is extremely high for weekends compared to weekdays."}
 # 

### 2020-02-03 main.Rmd
cricket data analysis - code optimised

@@
 Thus, the cyclic granularity ball-of-the-over will be circular in most cases and aperiodic in others.
-However, irrespective of the type of cyclic granularities, it can be interesting to visualize the periodicity of  scores/performance under similar data structure as temporal datasets.
+However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal dataset similar to any temporal dataset.
 
@@
 The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
+The hierarchy table would look like the following:
+
+```{r hierarchy2}
+ library(gravitas)
+ library(tibble)
+ hierarchy_model <- tibble::tibble(
+   G = c("over", "inning", "match", "season"),
+   C = c("over-of-inning", "inning-of-match", "match-of-season",1),
+   P = c(20, 2, "aperiodic", 1))
+ knitr::kable(hierarchy_model, format = "markdown")
+```
+
 There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
@@
 
-<!-- The hierarchy table would look like the following: -->
-
-<!-- # ```{r hierarchy2} -->
-<!-- # library(gravitas) -->
-<!-- # library(tibble) -->
-<!-- # hierarchy_model <- tibble::tibble( -->
-<!-- #   G = c("over", "inning", "match", "season"), -->
-<!-- #   C = c("over-of-inning", "inning-of-match", "match-of-season",1), -->
-<!-- #   P = c(20, 2, "aperiodic", 1)) -->
-<!-- # knitr::kable(hierarchy_model, format = "markdown") -->
-<!-- #  -->
-<!-- # ``` -->
-
 <!-- Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
@@
 
-Q1: How run rates vary depending on if a team bats first or second?
+Q1: How total runs vary depending on if a team bats first or second?
 
@@
 It would be interesting to take their example in order to dive deeper into this question.
-
+Circular cyclic granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circulargran} with over as index of the tsibble.
+From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
+This might indicate that players are vulnerable to score through each over and make more runs in the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
+To answer Q1, winning teams like CSK and MI have indeed different startegies when it comes to batting in the first or second innings.
+
+<!-- Also longer and more distinct letter values in the second innings suggests that the variability is more in the second innings. -->
 <!-- - Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first? -->
@@
  cricket_per_over <- cricket_season %>%
-   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))
+   group_by(season,
+            match_id,
+            batting_team,
+            bowling_team,
+            inning,
+            over) %>%
+   summarise(runs_per_over = sum(total_runs),
+             run_rate = sum(total_runs)/length(total_runs))
 
@@
 
-From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
-This might indicate that palyers are vulnerable to score through each over and make more runs in the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
-Also longer and more distinct letter values in the second innings suggests that the variability is more in the second innings.
-To answer Q1, winning teams like CSK and MI have indeed different startegies when it comes to batting in the first or second innings.
-
 ```{r cricex, fig.cap= "Letter value plot of runs per over across overs of the inning faceted by innings of the match.
@@
   as_tsibble(index = data_index)
+
+hierarchy_model <- tibble::tibble(
+  units = c("index", "over", "inning", "match"),
+  convert_fct = c(1, 20, 2, 1)
+)
+
+# cricket_tsibble %>%
+#   filter(batting_team %in% c("Mumbai Indians",
+#                              "Chennai Super Kings"))%>%
+#   create_gran("inning_match", hierarchy_model) %>% create_gran("over_inning", hierarchy_model)
+
+cricket_tsibble %>%
+  filter(batting_team %in% c("Mumbai Indians",
+                             "Chennai Super Kings"))%>%
+  gran_advice("inning", 
+            "over")
 
@@
 
+cricket_tsibble %>%
+  filter(batting_team %in% c("Mumbai Indians",
+                             "Chennai Super Kings"))%>%
+  gran_advice("inning", 
+            "over")
+
 ```
@@
 
-cricket_fielding <- cricket_season %>% 
-  mutate(fielding_proxy = if_else(dismissal_kind %in% c("caught", "caught and bowled"),1,0)) %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+cricket_dot_field <- cricket_season %>% 
+  mutate(fielding_proxy = if_else(dismissal_kind %in% 
+                                    c("caught", "caught and bowled"),1,0),
+         dot_ball_proxy = if_else(total_runs==0,1,0),
+         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
+  group_by(season,
+           match_id,
+           batting_team,
+           bowling_team, 
+           inning,
+           over) %>%
   summarise(runs_per_over = sum(total_runs),
             run_rate = sum(total_runs)/length(total_runs),
-            fielding_wckts = sum(fielding_proxy)) %>%
+            fielding_wckts = sum(fielding_proxy),
+            dot_balls = sum(dot_ball_proxy)) %>%
             mutate(diff_run_rate = c(0,diff(run_rate)))
 
-cricket_tsibble_fld <- cricket_fielding %>% 
+cricket_tsibble <- cricket_dot_field %>% 
   ungroup() %>% 
@@
 
-cricket_tsibble_fld %>% gran_obs("over", "fielding_wckts")
-
-cricket_tsibble_fld %>% 
-  mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets")) %>% 
+cricket_tsibble %>% gran_obs("over", "fielding_wckts")
+
+cricket_data_ts <- cricket_tsibble %>% 
+  mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets"),
+         dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls"))
+
   #filter(fielding_wckts %in% c(0,1)) %>% 
-  prob_plot( "field","over",
+  cricket_data_ts %>% prob_plot( "over","field",
                                hierarchy_model,
                                response = "diff_run_rate",
-                               plot_type = "boxplot",
+                               plot_type = "violin",
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
            overlay = TRUE)  + ggtitle("")
-
 ```
@@
 
-```{r crickread, ech0=FALSE}
-cricket_tsibble <- gravitas::cricket %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-```
+<!-- # ```{r crickread, ech0=FALSE} -->
+<!-- # cricket_tsibble <- gravitas::cricket %>% -->
+<!-- #   mutate(data_index = row_number()) %>% -->
+<!-- #   as_tsibble(index = data_index) -->
+<!-- # ``` -->
 
@@
 
-library(gravitas)
-library(tsibble)
-
-cricket_bowling <- cricket_season %>% 
-  mutate(dot_ball_proxy = if_else(total_runs==0,1,0),
-         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
-  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-  summarise(runs_per_over = sum(total_runs),
-            run_rate = sum(total_runs)/length(total_runs),
-            dot_balls = sum(dot_ball_proxy)) %>% 
-            mutate(diff_run_rate = c(0,diff(run_rate)))
-
-cricket_tsibble_bow <- cricket_bowling %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-cricket_tsibble_bow %>%
-  mutate(dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) %>% 
+# library(gravitas)
+# library(tsibble)
+# 
+# 
+# cricket_bowling <- cricket_season %>% 
+#   mutate(dot_ball_proxy = if_else(total_runs==0,1,0),
+#          wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
+#   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
+#   summarise(runs_per_over = sum(total_runs),
+#             run_rate = sum(total_runs)/length(total_runs),
+#             dot_balls = sum(dot_ball_proxy)) %>% 
+#             mutate(diff_run_rate = c(0,diff(run_rate)))
+# 
+# 
+# cricket_tsibble_bow <- cricket_bowling %>% 
+#   ungroup() %>% 
+#   mutate(data_index = row_number()) %>% 
+#   as_tsibble(index = data_index)
+#
+
+# cricket_tsibble_bow %>%
+#   mutate(dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) %>% 
+#   #filter(dot_balls %in% c(0, 1, 2)) %>%
+#   prob_plot("dot",
+#             "over",
+#             hierarchy_model,
+#             response = "diff_run_rate",
+#             plot_type = "quantile",
+#             quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
+#             symmetric = FALSE)  + ggtitle("")
+
+cricket_data_ts %>% 
   #filter(dot_balls %in% c(0, 1, 2)) %>%
@@
             symmetric = FALSE)  + ggtitle("")
-
 ```

### 2020-02-03 main.Rmd
smart meter application - all captions and headers fixed.

@@
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption in `general_supply_kwh` in every half hour from February 2012 to March 2014.
-\ref{tab:smart-meter} shows the data structure of the available data.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
 It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns and identify the extreme and regular households.
-
+The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns.
+<!-- and identify the extreme and regular households. -->
+
+### Data structure for visualization:
+
+The data structure for a sample of 50 households from that trial is as follows:
 <!-- We take a random sample of 50 households to see how we can achieve that.  -->
 
-### Data structure for visualization:
+```{r smart-meter, fig.cap = "Smart meter data structure"}
+sm <- 
+  read_rds("data/sm_cust50.rds")
+
+smart_meter <- sm %>% as_tsibble() %>% dplyr::select(customer_id, 
+                              reading_datetime, 
+                              general_supply_kwh)
+smart_meter
+```
 
@@
 
-```{r smart-meter, fig.cap = "Smart meter data structure"}
-
-sm <- 
-  read_rds("data/sm_cust50.rds")
-
-smart_meter <- sm %>% as_tsibble() %>% dplyr::select(customer_id, 
-                              reading_datetime, 
-                              general_supply_kwh, 
-                              everything()) %>% head()
-smart_meter
-# knitr::kable(smart_meter, format = "markdown")
-```
-
 ### Cyclic granularities search and computation:
@@
 smart_meter %>% search_gran()
-#smart_smart_search <- smart_meter %>% search_gran()
-#knitr::kable(smart_smart_search, format = "markdown")
-```
+```
+
+<!-- #smart_smart_search <- smart_meter %>% search_gran() -->
+<!-- #knitr::kable(smart_smart_search, format = "markdown") -->
 
@@
   search_gran(highest_unit = "month")
-# smart_meter_limit1 <- smart_meter10 %>% 
-#   search_gran(highest_unit = "month")
-# knitr::kable(smart_meter_limit1, format = "markdown")
-```
+```
+
+<!-- # smart_meter_limit1 <- smart_meter10 %>%  -->
+<!-- #   search_gran(highest_unit = "month") -->
+<!-- # knitr::kable(smart_meter_limit1, format = "markdown") -->
 
@@
                               )
-# smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month",
-#                               filter_out = c("hhour", "fortnight")
-# )
-# knitr::kable(smart_meter_limit2, format = "markdown")
-```
+```
+
+<!-- # smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month", -->
+<!-- #                               filter_out = c("hhour", "fortnight") -->
+<!-- # ) -->
+<!-- # knitr::kable(smart_meter_limit2, format = "markdown") -->
 
@@
 <!-- From the Gestalt theory with the same data  <$C1$ $C2$, `general_supply_kwh`>, mapping $C1$ on the x-axis vs facet might lead to different insights. -->
+s
 
@@
 )
-# smart_meter10 %>% gran_advice("wknd_wday", "hour_day")
-
-```
-
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
-This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
-For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households.
-Quantiles for weekends and weekdays for these customers look exactly the same.
-It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors.
-Consider another example in Figure \ref{fig:my-hd} showing the area quantile plots of 50 customers.
-The median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-The pink band changes significantly in winter months (May - August) implying that much of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of all quantiles increased too in winter months (increased energy usage in weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+```
+
+<!-- # smart_meter10 %>% gran_advice("wknd_wday", "hour_day") -->
 
@@
 
-```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
-The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."}
-
-smart_meter <- 
-  read_rds("data/sm_cust50.rds")
-
-smart_meter %>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "quantile",
-            symmetric = FALSE, 
-            quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
-  scale_y_sqrt()
-```
+<!-- No need to share the same for 50 customers - cut down text and anyway nothing much interesting is coming -->
+
+<!-- ```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
+The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."} -->
+
+<!-- smart_meter <-  -->
+<!--   read_rds("data/sm_cust50.rds") -->
+
+<!-- smart_meter %>%  -->
+<!--   prob_plot("wknd_wday", -->
+<!--                             "hour_day", -->
+<!--                             response = "general_supply_kwh", -->
+<!--                             plot_type = "quantile", -->
+<!--             symmetric = FALSE,  -->
+<!--             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
+<!--   scale_y_sqrt() -->
+<!-- ``` -->
 
@@
 
-```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
+<!-- ```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
 The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
 The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
-"}
-
-library(gravitas)
-library(tsibble)
-
-library(ggplot2)
-smart_meter %>%
-  prob_plot("month_year", "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99),
-    symmetric = TRUE) +
-  ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3))
-```
+"} -->
+
+<!-- library(gravitas) -->
+<!-- library(tsibble) -->
+
+<!-- library(ggplot2) -->
+<!-- smart_meter %>% -->
+<!--   prob_plot("month_year", "hour_day", -->
+<!--     response = "general_supply_kwh", -->
+<!--     plot_type = "quantile", -->
+<!--     quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99), -->
+<!--     symmetric = TRUE) + -->
+<!--   ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3)) -->
+<!-- ``` -->
 
@@
 -->
-
-Looking at random samples with different demographics might confound our analysis and a strategy for analyzing multiple keys together could be to first screen some basis time series or demographic features and then look at their energy behavior.
-This is beyond the scope of the current work.
-For the purpose of this paper, the smart meter data is filtered for two customers to illustrate kinds of insights that can be drawn after proceeding with the systematic exploration.
 
@@
 
-In \autoref{fig:bothcust} we visualize the harmony pair (wknd_wday, hour_day) through a box plot.
-Boxplot of energy consumption is shown across `wknd_wday` (facet) and `hour_day` (x-axis) for the same two households.
+<!-- Boxplot of energy consumption is shown across `wknd_wday` (facet) and `hour_day` (x-axis) for the same two households.
 For the first household, the shape of the distribution does not change significantly for weekdays and weekends.
 The distribution is roughly smooth with a morning and an evening peak.
-For the second household, the quantiles behave differently in peak hours.
-For the first few hours on weekdays, energy consumption starts at a much higher level than weekends.
-Consider looking at letter value plots instead of quantile plots to look at the same data in \autoref{fig:bothcust-lv}.
-There is additional information that we can derive looking at the distribution.
-There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day.
+-->
+
+In \autoref{fig:bothcust}, the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot. 
+The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
+The energy consumption is extremely skewed with 1st, 10th and 25th percentile lying very close whereas 75th, 90th and 99th lying further away from each other.This is common across both weekdays and weekends.
+For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer as compared to weekends.
+
+Consider looking at letter value plots instead of quantile plots to look at the same data in \autoref{fig:bothcust-lv}. There is additional information that we can derive looking at the distribution.
+There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
 Also hours from 7 to 13 look most volatile.
@@
 Also, outliers are more prominent in the latter part of the day.
-All of these indicates that looking at different distribution plots or changing the mapping might draw lights on different aspect of the energy behavior for the same customer.
+All of these indicates that looking at different distribution plots or changing the mapping might shed lights on different aspect of the energy behavior for the same customer.
 
@@
 
-```{r bothcust_notrun}
-cust1 <- smart_meter10 %>% 
-  filter(customer_id %in% c(10006704)) %>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "quantile",
-            symmetric = FALSE,
-            quantile_prob = c(0.25, 0.5, 0.75, 0.9)) +
-  scale_y_sqrt() +
-  ggtitle("Energy consumption distribution for customer id: 10006704")
-
+```{r granadvice, echo = TRUE}
+smart_meter10 %>% 
+  filter(customer_id %in% c(10017936))%>% 
+  gran_advice("wknd_wday",
+                            "hour_day")
+```
+
+```{r bothcust, fig.cap = "Area quantile plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936 displays the smoothed energy consumption curve for the entire day.
+The customer typically consumes more in the morning hours following which their consumption is low till late afternoon hours.
+Median consumption for the early morning hours starts higher and continue to remain high for longer during weekdays compared to weekdend."}
+
+# plot for only customer is shown
+# cust1 <- smart_meter10 %>% 
+#filter(customer_id %in% c(10006704)) %>% 
+#   prob_plot("wknd_wday",
+#                             "hour_day",
+#                             response = "general_supply_kwh",
+#                             plot_type = "quantile",
+#             symmetric = FALSE,
+#             quantile_prob = c(0.25, 0.5, 0.75, 0.9)) 
+#   scale_y_sqrt() +
+#   ggtitle("Energy consumption distribution for customer id: 10006704")
 cust2 <- smart_meter10 %>% 
@@
                             plot_type = "quantile",
-            symmetric = FALSE,
-             quantile_prob = c(0.25, 0.5, 0.75, 0.9)) +
-  scale_y_sqrt() + 
-  ggtitle("Energy consumption distribution for customer id: 10017936")
-```
-
-```{r bothcust, fig.cap = "Quantile plots of energy consumption across hours of the day faceted by weekday/weekend for household 1 (customer id: 10006704) and household 2 (customer id: 10017936).
-Median behavior for the early morning hours of household 2 is extremely high for weekends compared to weekdays, not much differenc can be observed for household 1.
-"}
-cust1
+            symmetric = TRUE,
+             quantile_prob = c(0.01,0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
+  scale_y_sqrt() 
 cust2
@@
 
-```{r bothcust-lv, fig.cap ="Violin plots of energy consumption across hours of the day faceted by weekday/weekend for household 2 (customer id: 10017936).
-Early morning hours are bimodal implying both high and low consumption are probable.
-Volatility is high from 7 to 13 hours as evident from the narrow violins."}
+```{r bothcust-lv, fig.cap ="Violin plots of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
+Early morning hours are bimodal implying both high and low consumptions are probable.
+Bimodality can be caused by difference in behavior across different seasons.
+Volatility is highest from 7 to 13 hours as evident from the narrow violins."}
 # 
@@
                             plot_type = "violin") +
-  scale_y_sqrt() + ggtitle("Energy consumption distribution for customer id: 10017936")
-```
-
-```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for household 2 (customer id: 10017936).
-Median behavior for the early morning hours of household 2 is extremely high for weekends compared to weekdays."}
+  scale_y_sqrt() 
+```
+
+```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
+The change in mapping of the cyclic granularities leads to easier comparison within weekdays and weekends for each hour.
+Median consumption till 6am is higher for weekends compared to weekdays although the quartiles look similar.
+Outliers appear more during the end of the day implying more uncertain behavior in latter hours of the day compared to early morning hours."}
 # 
@@
             plot_type = "boxplot") +
-            scale_y_sqrt() +
+            scale_y_sqrt()
             # scale_fill_brewer(palette = "Dark2")
-  ggtitle("Energy consumption distribution for customer id: 10017936")
+
 ```
@@
 
-This case study shows systematic exploration of energy behavior for two random households to gain some insights on periodic behavior of the households.
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
+This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+A strategy for analyzing multiple keys together could be to first screen some basis time series or demographic features and then look at their energy behavior.
+This is beyond the scope of the current work.
+
+<!-- For the purpose of this paper, the smart meter data is filtered for two customers to illustrate kinds of insights that can be drawn after proceeding with the systematic exploration. -->
+
+<!-- For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households.
+Quantiles for weekends and weekdays for these customers look exactly the same.
+It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors.
+Consider another example in Figure \ref{fig:my-hd} showing the area quantile plots of 50 customers.
+The median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
+The pink band changes significantly in winter months (May - August) implying that much of the behavioral changes occur in the quartiles, that too in peak hours of the day.
+It is to be noted here that the the level of all quantiles increased too in winter months (increased energy usage in weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
+-->
+
+This case study shows systematic exploration of energy behavior for a random household to gain some insights on periodic behavior of the households.
 First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualise a harmony pair and shows the effect of different distribution plots or reverse mapping.

### 2020-02-03 main.Rmd
minor changes in cricket and smart meter

@@
 
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:asperiodic-gran-def}.
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:aperiodic-gran-def}.
 
@@
 <!-- From the Gestalt theory with the same data  <$C1$ $C2$, `general_supply_kwh`>, mapping $C1$ on the x-axis vs facet might lead to different insights. -->
-s
 
@@
 It would be interesting to take their example in order to dive deeper into this question.
-Circular cyclic granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circulargran} with over as index of the tsibble.
+Circular cyclic granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
 From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
@@
 ```
-#
-#
+
 <!-- # ```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."} -->
@@
 
-cricket_tsibble %>%
-  filter(batting_team %in% c("Mumbai Indians",
-                             "Chennai Super Kings"))%>%
-  gran_advice("inning", 
-            "over")
+# cricket_tsibble %>%
+#   filter(batting_team %in% c("Mumbai Indians",
+#                              "Chennai Super Kings"))%>%
+#   gran_advice("inning", 
+#             "over")
 

### 2020-02-05 main.Rmd
revise with minor changes in circular granularity
first equation revise minor
aperiodic cylic granularity

@@
 But the representation of the resultant circular granularity will be a function of the index set.
-It is assumed that the bottom granularity is minutes.
+The bottom granularity is assumed to be minutes.
 
@@
   & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
-  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
   & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
-  & = L(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
+  & = L(8 - 2*\vert T_{0}\vert - \vert T_{1}\vert) \\
   & = L(8 - 2*2 - 3) \\
@@
 
-**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, from definition \ref{def:periodical} the following holds true:
+**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then the following holds true:
 
@@
 - there exist $R', P' \epsilon Z_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
-- number of granules of $B$ in each granule of $G'$ is not a constant
+
+from definition \ref{def:periodical}
+
+and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping)
 
@@
 The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
-The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R)^{th}$ granule of $G'$ will have the same number of granules of $B$.
+The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R')^{th}$ granule of $G'$ will have the same number of granules of $B$.
 The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$.
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{j=1}^{n}M_{j}$ and index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{i=1}^{n}M_{i}$ and index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
 \end{definition}
@@
 
-**Discussion**: The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
+**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labour day and M can be conceived as the aperiodic linear granularity denoting any holidays.
+The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
 

### 2020-02-06 main.Rmd
-1 deleted from cyclic calenar algra - single to multiple conversion
changes in single to multiple

@@
 Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity.
-Since, the operation requires the use of modular arithmetic, it is important that the label mapping of each circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
+Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
 The label mapping of the resultant multiple-order-up granularity can however be chosen arbitrarily, depending on the context.
@@
 \begin{split}
-C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z))) \\  
+C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_{l+1},G_m)}(z))) \\  
   & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z))) \\
-  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
-  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
+  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
+  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
   &\vdots\\
@@
 
-*Example:* Let us use the equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
+*Example:* Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
 
@@
 Examples might include hour-of-month or day-of-year.
-Let us consider the computation of $C_{hour, month}(z)$ for $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$.
+Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$.
 
@@
 However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
-It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$.
-To keep it simple, we include only the case where there is just one single order-up quasi-circular granularity in the hierarchy table and any multiple order-up quasi-circular granularity need to be expressed with single order-up circular granularities and a quasi-circular granularity, that is not necessarily single order-up.
-
-\begin{equation} \label{eq:multifromsingle-quasi}
+It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
+
+Generally speaking, to keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy.
+Any multiple order-up quasi-circular granularity can then be considered to be a function of some circular (single or multiple order-up) and a quasi-circular single order-up granularity.
+
+<!-- Both of them can be circular or quasi-circular or it can be each of both. In case they are both quasi-circular or one circular and the other quasi-circular, we need to adjust the formula -  -->
+
+For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, two different approaches need to be employed for computing multiple order-up granularities from single order-up ones.
+
+- Case when $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+
+\begin{equation} \label{eq:multifromsingle-quasi1}
 \begin{split}
-C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)) \\ 
-& = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))) + P(l, m')(C_{G_{m'},G_{m}}(z))\\
+C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\ 
 \end{split}
@@
 
-where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping.
+- Case when  $C_{l,m'}(z)$  is quasi-circular and $C_{m',m}(z)$ is circular
+
+\begin{equation} \label{eq:multifromsingle-quasi2}
+\begin{split}
+C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert))\\
+\end{split}
+\end{equation}
+
+where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+
+<!-- \begin{equation} \label{eq:multifromsingle-quasi} -->
+<!-- \begin{split} -->
+<!-- C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\  -->
+<!-- & = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))) + P(l, m')(C_{G_{m'},G_{m}}(z))\\ -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
+
+<!-- where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping. -->
+
+If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
 

### 2020-02-07 main.Rmd
adding latex package after asa article output

@@
 bibliography: [bibliography.bib]
+preamble: >
+  \usepackage{mathtools}
+  \usepackage{amssymb} 
+  \usepackage{Booktabs}
 output:
   bookdown::pdf_book:
-    #base_format: rticles::asa_article
+    base_format: rticles::asa_article
     fig_height: 5
@@
 (groups into): A linear granularity G groups into a linear granularity H, denoted
-$G \trianglelefteq H,$
+$G \trianglelefteq H$,
 if for each index j there exists a (possibly infinite) subset S of the integers such
@@
 (groups periodically): A granularity H is periodical with respect to a granularity G if
-(1) $G \trianglelefteq H,$ and
+(1) $G \trianglelefteq H$, and
 (2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
@@
 (groups quasi-periodically): A granularity H is quasi-periodical with respect to a granularity G if
-(1) $$G \trianglelefteq H,$$, and
+(1) $G \trianglelefteq H$, and
 (2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
@@
 
-If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
+If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z) are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
 

### 2020-02-07 main.Rmd
commits to include all latex packages for including jcgs format

@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009-pk].
+This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
 bibliography: [bibliography.bib]
@@
   \usepackage{amssymb} 
-  \usepackage{Booktabs}
+  \usepackage{booktabs}
+  \usepackage{longtable}
 output:
@@
 Calendar-based graphics [@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
-Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009-pk] with days of the week, month of the year and another sub-daily deconstruction of time.
+Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year and another sub-daily deconstruction of time.
 But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
@@
 The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by [@Wickham2009-pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
+The layered grammar of graphics proposed by [@Wickham2009pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
 There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
@@
 
-<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009-pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
+<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
 The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
@@
 If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
-For example in \autoref{fig:cust-lv}, it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
+For example in \autoref{fig:bothcust-lv}, it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
 Also, outliers are more prominent in the latter part of the day.
@@
 
-```{r granadvice, echo = TRUE}
+```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE}
 smart_meter10 %>% 
@@
 ```{r cricex, fig.cap= "Letter value plot of runs per over across overs of the inning faceted by innings of the match.
-No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning."}
+No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning.", warning=FALSE, message = FALSE}
 
@@
 This shows good fielding in previous over leads to lower run rate in the subsequent over.
-Also run rates are much less volatile in case there are at least one wicket in the previous over."}
+Also run rates are much less volatile in case there are at least one wicket in the previous over.", warning=FALSE}
 

### 2020-02-07 gravitas/data-raw/cricketdata.R
cleaning up folder

@@ -1,59 +1,0 @@
-library(readr)
-library(tsibble)
-library(dplyr)
-
-cricket <-  read_csv("data-raw/deliveries_all.csv")
-matches <- read_csv("data-raw/matches_all.csv")
-
-cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id")
-
-# each over needs to have 6 balls
-
-cricket_fltr_ball <- cricket_season %>%
-  filter(wide_runs + noball_runs== 0) %>%    # filtering out scores that are gained by no balls or wide
-  # still have to filter out overs for which number of balls per over less than 6
-  group_by(batting_team, match_id, inning, over) %>%
-  summarise(n_over = length(ball)) %>%
-  filter(n_over==6) %>%
-  mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
-
-
-cricket_over_crctd <- cricket_season %>%
-  mutate(cricket_key = paste(batting_team, match_id, inning, over, sep = "_")) %>%
-  filter(cricket_key %in% cricket_fltr_ball$key) %>%
-  filter(wide_runs + noball_runs== 0) %>%
-  mutate(ball_per_over = purrr::rep_along(match_id, 1:6))
-
-# each innings need to have 20 overs
-
-cricket_fltr_over <- cricket_over_crctd %>%
-  filter(inning %in% c(1,2)) %>%
-  group_by(inning, match_id) %>%
-  summarise(n = length(over)) %>% filter(n==120) %>%
-  mutate(key = paste(match_id, inning, sep = "_"))
-
-cricket_over_inning_crctd <- cricket_over_crctd %>%
-  mutate(over_key = paste(match_id, inning, sep = "_")) %>%
-  filter(over_key %in% cricket_fltr_over$key) %>%
-  filter(wide_runs + noball_runs== 0) %>%
-  mutate(ball_per_over = purrr::rep_along(match_id, 1:6)) %>%
-  filter(inning %in% c(1,2))
-
-
-# each match needs to have 2 innings
-
-cricket_filtr_inning <- cricket_over_inning_crctd %>%
-  group_by(match_id) %>% 
-  summarise(length(inning)) %>% 
-  filter(`length(inning)`==240)
-
-cricket_all_seas <- cricket_over_inning_crctd %>%
-  filter(match_id %in% cricket_filtr_inning$match_id) %>%
-  filter(wide_runs + noball_runs== 0) %>%
[…]

### 2020-02-07 main.tex
changes for asa journal rticle format

--- 2020-02-07 main.Rmd
@@
----
-title: Exploring probability distributions for bivariate temporal granularities
-authors:
-- name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-  email: Sayani.Gupta@monash.edu
-
-- name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Antony Unwin
-  affiliation: University of Augsburg
-
-keywords:
-- data visualization
-- statistical distributions
-- time granularities
-- periodicties
-- grammar of graphics
-- R
-
-abstract: |
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+%% load any required packages here
+\usepackage{mathtools} \usepackage{amssymb}  \usepackage{booktabs} \usepackage{longtable}
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Exploring probability distributions for bivariate temporal granularities}
+
+  \author{
+        Sayani Gupta \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Rob J Hyndman \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Dianne Cook \\
+    Department of Econometrics and Business Statistics, Monash University\\
+     and \\     Antony Unwin \\
+    University of Augsburg\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Exploring probability distributions for bivariate temporal granularities}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
 Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
-bibliography: [bibliography.bib]
-preamble: >
-  \usepackage{mathtools}
-  \usepackage{amssymb} 
-  \usepackage{booktabs}
-  \usepackage{longtable}
-output:
-  bookdown::pdf_book:
-    base_format: rticles::asa_article
-    fig_height: 5
-    fig_width: 8
-    fig_caption: yes
-    dev: "pdf"
----
-
-```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
-options("knitr.graphics.auto_pdf" = TRUE)
-library(knitr)
-library(tidyverse)
-library(lubridate)
-library(lvplot)
-library(ggridges)
-library(tsibble)
-library(gravitas)
-library(tsibble)
-opts_chunk$set(
-  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
-  fig.path = "figure/", fig.align = "center", fig.show = "hold",
-  cache = FALSE, cache.path = "cache/",
-  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
-)
-knitr::opts_knit$set(root.dir = here::here())
-```
-
-```{r external, include = FALSE}
-# read_chunk('scripts/main.R')
-```
-
-```{r load}
-
-```
-
-# Introduction
-
-<!--temporal granularities and why should we care -->
-
-<!-- problem of analyzing data with such finer scales and classical ways to deal with it -->
+This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 \citep{Wickham2009pk}.
+\end{abstract}
+
+\noindent%
+{\it Keywords:} data visualization, statistical distributions, time granularities, periodicties, grammar of graphics, R
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
-This approach requires deconstructing time in various possible ways called time granularities [@aigner2011visualization].
+This approach requires deconstructing time in various possible ways called time granularities \citep{aigner2011visualization}.
 It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
-This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+This idea aligns with the notion of EDA \citep{Tukey1977-jx} which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
 Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
@@
 
-<!-- # existing ways and challenges -->
-
-Calendar-based graphics [@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
-Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year and another sub-daily deconstruction of time.
+Calendar-based graphics \citep{wang2018calendar} are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
+Any sub-daily resolution temporal data can also be displayed using this type of faceting \citep{Wickham2009pk} with days of the week, month of the year and another sub-daily deconstruction of time.
 But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
 There can be several classes of time deconstructions, viz.
-based on the arrangement (linear vs.
-cyclic) or hierarchical order of the calendar.
+based on the arrangement (linear vs.~cyclic) or hierarchical order of the calendar.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-One of the first attempts of characterizing these granularities occur in @Bettini1998-ed.
+One of the first attempts of characterizing these granularities occur in \citet{Bettini1998-ed}.
 The definitions and rules defined are inadequate to reflect periodicities in time.
@@
 Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
-lubridate [@G_Grolemund2011-vm] creates easy access and manipulation of common date-time objects.
+lubridate \citep{G_Grolemund2011-vm} creates easy access and manipulation of common date-time objects.
 But most accessor functions are limited to single-order-up granularities.
-
-<!--motivation -->
 
@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates (see @2012-la) of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+It is common to see aggregates (see \citet{2012-la}) of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
 But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
@@
 
-<!-- at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA).
-EDA calls for utilizing visualization and transformation to explore data systematically.
-It is a process of generating hypothesis, testing them and consequently refining them through investigations.
--->
-
-<!--existing closely related work and how ours relate --> 
- <!-- This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones. -->
-
-<!--actions on multiple order up - usage -->
-
-<!-- Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-However, this approach can lead to a myriad of choices all of which are not useful.
-Analysts are expected to iteratively visualize these choices for exploring possible patterns in the data.
-But too many choices might leave him bewildered.
--->
-
-<!-- challenges and how you are dealing wih it --> 
 This work provides tools for systematically exploring bivariate granularities within the tidy workflow through the following:
 
-<!-- Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
--->
-
-  * Formal characterization of cyclic granularities
-
-  * Facilitate manipulation of single and multiple order-up time granularities through cyclic calendar algebra
-
-  * Checking feasibility of creating plots or drawing inferences for any two cyclic granularities
-
-  * Recommend prospective probability distributions for exploring distributions of univariate dependent variable across pair of granularities
-
-  <!-- Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
--->
+\begin{itemize}
+\item
+  Formal characterization of cyclic granularities
+\item
+  Facilitate manipulation of single and multiple order-up time granularities through cyclic calendar algebra
+\item
+  Checking feasibility of creating plots or drawing inferences for any two cyclic granularities
+\item
+  Recommend prospective probability distributions for exploring distributions of univariate dependent variable across pair of granularities
+\end{itemize}
 
@@
 
-# Linear time granularities {#sec:linear-time}
-
-<!-- general calendar categorization -->
-<!-- Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
-Classes of time abstractions may be based on arrangement or hierarchical order of time, discussed in further details in \autoref{sec:arrangement} and \autoref{sec:order}.
--->
+\hypertarget{sec:linear-time}{%
+\section{Linear time granularities}\label{sec:linear-time}}
 
 Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Time granularities are **linear** if they respect the linear progression of time.
+Such discrete abstractions of time can be thought of as time granularities \citep{aigner2011visualization}.
+Time granularities are \textbf{linear} if they respect the linear progression of time.
 Examples include hours, days, weeks and months.
 
-## Definitions and relationships
-
-<!-- Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement} -->
-
-<!-- The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.  -->
-
- There has been several attempts to provide the framework for formally characterizing time-granularities. One of the first attempts occur in @Bettini1998-ed with the help of the following definitions:
-
-<!-- <!-- Representation with tsibble -->
-<!-- Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
--->
-
-<!-- <!-- What is calendar categorization --> 
-<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc.
-Many of these are circular, such as day-of-week, time-of-day.
-Some are nearly circular such as day-of-month.
-Some are irregular such as working day/non-working day.
-Let's call all of these "calendar categorizations".
-Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
--->
-
-<!-- What is circular granularity and some examples -->
-<!-- We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
-The number of categories is essentially the periodicity of a circular time granularity.
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
-Then the following categorizations can be computed.
--->
-
-<!-- In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.  -->
-
-<!-- ### Linear {#sec:linear-gran-def} -->
+\hypertarget{definitions-and-relationships}{%
+\subsection{Definitions and relationships}\label{definitions-and-relationships}}
+
+There has been several attempts to provide the framework for formally characterizing time-granularities.
+One of the first attempts occur in \citet{Bettini1998-ed} with the help of the following definitions:
 
@@
 
-A time domain can be **discrete** (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be **dense** (if it is an infinite set).
+A time domain can be \textbf{discrete} (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be \textbf{dense} (if it is an infinite set).
 A time domain is assumed to be discrete for the purpose of our discussion.
@@
 
-**Discussion:** The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+\textbf{Discussion:} The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
 \autoref{fig:linear-time} shows the implication of this condition.
 It shows the linear granularities hours, days, weeks, months and years, each of which are unidirectional in nature and arranged from past to future.
-If we consider the chronon (see @aigner2011visualization) as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+If we consider the chronon (see \citet{aigner2011visualization}) as hourly, the time domain with T hours will have \(\lfloor T/24\rfloor\) days, \(\lfloor T/(24*7)\rfloor\) weeks and so on.
 Each granule of the linear granularities shown are represented through a box.
 
-```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
-knitr::include_graphics("Figs/linear-arr.png")
-```
-
-[@Bettini1998-ed] talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/linear-arr}
+
+}
+
+\caption{The time domain distributed as linear granularities}\label{fig:linear-time}
+\end{figure}
+
+\citep{Bettini1998-ed} talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
 
@@
 for each index i, there exists an index j such that $G(i) \subset H(j)$.
-\end{definition}.
+\end{definition}
+
+.
 
@@
 
-**Discussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.
-$G \preceq H$ if every granule in G is a subset of some granule in H.
-In \autoref{fig:groups-finer}, both $day \trianglelefteq week$ and $day \preceq week$ hold true.
+\textbf{Discussion:} \(G \trianglelefteq H\) if each granule H(j) is the union of some granules of G.
+\(G \preceq H\) if every granule in G is a subset of some granule in H.
+In \autoref{fig:groups-finer}, both \(day \trianglelefteq week\) and \(day \preceq week\) hold true.
 However, weekends are finer than weeks but do not group into weeks.
 And weekends neither groups into or finer than public holidays.
-Consider another example where $day \trianglelefteq month$.
+Consider another example where \(day \trianglelefteq month\).
 This relationship is not fully described until we associate it with periodicity.
@@
 
-```{r groups-finer,echo=FALSE,out.width = "100%",fig.cap="Illustrations of groups into and finer than relationships.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/groupsandfiner}
+
+}
+
+\caption{Illustrations of groups into and finer than relationships.
 Weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or finer than public holidays"}
-knitr::include_graphics("Figs/groupsandfiner.png")
-```
+And weekends neither groups into or finer than public holidays}\label{fig:groups-finer}
+\end{figure}
 
@@
 
-**Discussion:** A granularity H which is periodical with respect to G is specified by:
-(i) the R sets of indexes of G, ${S_0,...,S_{R-1}}$ describing the granules of H within one period;
+\textbf{Discussion:} A granularity H which is periodical with respect to G is specified by:
+(i) the R sets of indexes of G, \({S_0,...,S_{R-1}}\) describing the granules of H within one period;
 (ii) the value of P;
 (iii) the indexes of first and last granules in H, if their value is not infinite.
-Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
-$H (0), .
+Then, if \({S_0,...,S_{R-1}}\) are the sets of indexes of G describing
+\(H (0), .
 .
 .
-, H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
+, H (R - 1)\), respectively, then the description of an arbitrary granule H(j) is given by: \(\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)\).
+Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way \citep{Bettini2000-vy}.
 
@@
 
-**Discussion:** Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
+\textbf{Discussion:} Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by \(R\) and \(P\).
 
@@
 
-**Discussion:** Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
-
-<!-- Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
-The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
--->
-
-## Computation through calendar algebra
-
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate
+\textbf{Discussion:} Given the set of all granularities defined over the time domain \((Z; <)\), and the granularity relationship \(\trianglelefteq\) (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to \(\trianglelefteq\).
+
+\hypertarget{computation-through-calendar-algebra}{%
+\subsection{Computation through calendar algebra}\label{computation-through-calendar-algebra}}
+
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra \citep{Ning2002-tf}.
+It is assumed that there exists a ``bottom'' granularity and calendar algebra operations are designed to generate
 new granularities from the bottom one or recursively, from those already generated.
@@
 The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
-Example can be to consider a calendar with only two linear granularities "minute" and "hour" and "hour" is generated by grouping every $60$ "minutes".
+Example can be to consider a calendar with only two linear granularities ``minute'' and ``hour'' and ``hour'' is generated by grouping every \(60\) ``minutes''.
 The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
-For example, one can choose to look at the granularity "Monday" and hence select only "Monday" from linear granularity "day".
+For example, one can choose to look at the granularity ``Monday'' and hence select only ``Monday'' from linear granularity ``day''.
 
@@
 
-- **The grouping operation** : Let $G1$ be a full-integer labeled granularity, and $m$ a positive integer.
-The grouping operation $Group_m(G)$ generates a new granularity $G2$, by partitioning the granules of $G$ into m-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G2 = Group_m(G1)$ is the full-integer labeled granularity such that for
-each integer $i$, $G2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G1(j)$.
-
-**Examples** :
-
-  - $minute = Group_{60}(second)$
-  - $hour = Group_{60}(minute)$,
-  - $day = Group_{24}(hour)$,
-  - $week = Group_{7}(day)$,
+\begin{itemize}
+\tightlist
+\item
+  \textbf{The grouping operation} : Let \(G1\) be a full-integer labeled granularity, and \(m\) a positive integer.
+The grouping operation \(Group_m(G)\) generates a new granularity \(G2\), by partitioning the granules of \(G\) into m-granule groups and making each group a granule of the resulting granularity.
+  More precisely, \(G2 = Group_m(G1)\) is the full-integer labeled granularity such that for
+  each integer \(i\), \(G2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G1(j)\).
+\end{itemize}
+
+\textbf{Examples} :
+
+\begin{itemize}
+\tightlist
+\item
+  \(minute = Group_{60}(second)\)
+\item
+  \(hour = Group_{60}(minute)\),
+\item
+  \(day = Group_{24}(hour)\),
+\item
+  \(week = Group_{7}(day)\),
+\end{itemize}
 
@@
 
-- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and $l$, $k$, $m$ integers, where G2 partitions G1, and $1 \leq l \leq m$.
-The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-The altering-tick operation modifies the granules of G1 so that the $l$th granule of each group has $|k|$ additional (or fewer when k < 0) granules of G2.
-
-**Example** :
+\begin{itemize}
+\tightlist
+\item
+  \textbf{The altering-tick operation} : Let G1, G2 be full-integer labeled granularities, and \(l\), \(k\), \(m\) integers, where G2 partitions G1, and \(1 \leq l \leq m\).
+The altering-tick operation \(Alter^{m}_{l,k} (G2, G1)\) generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
+The altering-tick operation modifies the granules of G1 so that the \(l\)th granule of each group has \(|k|\) additional (or fewer when k \textless{} 0) granules of G2.
+\end{itemize}
+
+\textbf{Example} :
 
@@
 
-, where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
-
-For more variations of grouping operations and granule oriented operations, the readers are recommended to see @Ning2002-tf.
-
-# Cyclic time granularities {#sec:cyclic-gran}
-
-<!-- sub: regular and irregular mapping leads to circular and quasi-circular granularities -->
-
-We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
+, where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days.
+\citep{Ning2002-tf}
+
+For more variations of grouping operations and granule oriented operations, the readers are recommended to see \citet{Ning2002-tf}.
+
+\hypertarget{sec:cyclic-gran}{%
+\section{Cyclic time granularities}\label{sec:cyclic-gran}}
+
+We propose a formalism of cyclic time granularities through the tsibble \citep{wang2019tsibble} framework of organizing temporal data.
 A tsibble consists of an index, key and measured variables.
@@
 
-A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the time index of a tsibble.
+A time domain, as defined by \citet{Bettini1998-ed}, is essentially a mapping of the index set to the time index of a tsibble.
 A linear granularity is a mapping of the index set to subsets of the time domain.
@@
 A bottom granularity is represented by the index of the tsibble.
-Lining up with the the assumption in @Bettini2000-vy, that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists.
+Lining up with the the assumption in \citet{Bettini2000-vy}, that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists.
 In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
@@
 However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
-Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
+Cyclic time granularities are referred to as \textbf{circular} if mappings between linear time granularities are regular and \textbf{quasi-circular} if the same is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 
-<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
-
-## Circular {#sec:circular-gran-def}
+\hypertarget{sec:circular-gran-def}{%
+\subsection{Circular}\label{sec:circular-gran-def}}
 
@@
 
-**Example**:  Example showing circular granularity relating two linear granularities "day" and "week" is visually depicted in a series of slots in the \autoref{fig:circular-dow}.
+\textbf{Example}: Example showing circular granularity relating two linear granularities ``day'' and ``week'' is visually depicted in a series of slots in the \autoref{fig:circular-dow}.
 Each granule is represented by a box.
@@
 The first slot in the diagram shows the index set.
-The bottom granularity, represented by the index of the tsibble considered is "day".
-"day" and "week" are two linear granularities with $P =  7$.
-The circular granularity $C_{day, week}$ representing day-of-week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
-
-**Discussion**: Note that each circular granularity can use different label mappings.
+The bottom granularity, represented by the index of the tsibble considered is ``day''.
+``day'' and ``week'' are two linear granularities with \(P = 7\).
+The circular granularity \(C_{day, week}\) representing day-of-week will thus consist of pattern \{L(0), L(1), L(2), L(3), L(4), L(5), L(6)\} which repeats itself after each period length.
+
+\textbf{Discussion}: Note that each circular granularity can use different label mappings.
 In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity.
 However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
-Hence, the label mapping $L$ in \autoref{fig:circular-dow} can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{Sunday, Monday, \dots, Saturday\}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{0, 1, \dots, 6\}$ depending on the context.
-
-<!-- # ```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "} -->
-<!-- # knitr::include_graphics("Figs/circular-gran-diagram.png") -->
-<!-- # ``` -->
-
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
-knitr::include_graphics("Figs/circular-dow.png")
-```
-
-<!-- All circular granularities can be expressed in terms of the bottom linear granularity with equation \autoref{eq:eq2}. -->
-
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity H is periodic with respect to linear granularity G with regular mapping such that the number of granules of G in each granule of H is $P(G,H)$.
-Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
+Hence, the label mapping \(L\) in \autoref{fig:circular-dow} can be defined as \(L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}\) or \(L: \{0,1,2, \dots, 6\} \longmapsto\ \{Sunday, Monday, \dots, Saturday\}\) or \(L: \{0,1,2, \dots, 6\} \longmapsto\ \{0, 1, \dots, 6\}\) depending on the context.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/circular-dow}
+
+}
+
+\caption{Circular granularity day-of-week}\label{fig:circular-dow}
+\end{figure}
+
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as \(C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))\), where linear granularity H is periodic with respect to linear granularity G with regular mapping such that the number of granules of G in each granule of H is \(P(G,H)\).
+Table \ref{tab:definitions} shows representation of circular granularities \(C_i\) relating two linear granularities with \(P_i\) being the number of granules of the finer granularity in each granule of the coarser granularity and \(L_i\) is the associated label mapping.
 It is possible that none of the two linear granularities are bottom granularities.
@@
 minute-of-Hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
-minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
+minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$1440$\\
 hour-of-Day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
-hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
+hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$168$\\
 day-of-Week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
@@
 
-## Quasi-circular {#sec:quasi-circular-gran-def}
-
-A **quasi-circular** granularity can not be defined using modular arithmetic since they are formed using two linear granularities with irregular mapping.
-However, they are still formed with linear granularities, one of which "groups periodically into" the other.
-\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) with ($P_i$) denoting the plausible choices of number of granules of the finer granularity inside each granule of the coarser one.
+\hypertarget{sec:quasi-circular-gran-def}{%
+\subsection{Quasi-circular}\label{sec:quasi-circular-gran-def}}
+
+A \textbf{quasi-circular} granularity can not be defined using modular arithmetic since they are formed using two linear granularities with irregular mapping.
+However, they are still formed with linear granularities, one of which ``groups periodically into'' the other.
+\autoref{tab:quasi} shows some example of quasi-circular granularities (\(Q_i\)) with (\(P_i\)) denoting the plausible choices of number of granules of the finer granularity inside each granule of the coarser one.
 
@@
 
-**Example**: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
+\textbf{Example}: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
 Each granule is represented by a box.
-Two linear granularities $G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))$ are considered.
-This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
-Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.
-$Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
-For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
-For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:circular-gran}.
+Two linear granularities \(G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))\) and \(H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))\) are considered.
+This implies that \(G'\) is made up by shrinking every 1st granule of \(Group_3(B)\) by \(1\) granule and \(H'\) is made up of shrinking every 1st granule of \(Group_3(B)\) by \(2\) granules.
+Number of granules of \(G'\) and \(H'\) in each period of \(B\) is 2 but the number of granules of \(B\) in each of those granules are different.
+\(Q_{B, G'}\) and \(Q_{B, H'}\) are repetitive categorization of time, similar to circular granularities, except that the number of granules of \(B\) is not necessarily the same across different granules of \(G'\) or \(H'\).
+For \(G'\), \(T_0 = \{0,1\}\) and \(T_1 = \{2, 3, 4\}\).
+For \(H'\), \(T_0 = \{0, 1, 2, 3, 4, 5, 6\}\) and \(T_1 = \{7, 8, 9, 10, 11\}\), then we will have \autoref{eq:eq4} and \autoref{eq:circular-gran}.
 
@@
 
-**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then the following holds true:
-
-- $B \trianglelefteq G'$
-- there exist $R', P' \epsilon Z_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
+\textbf{Discussion}: If linear granularity \(G'\) is periodical with respect to \(B\) with irregular mapping, then the following holds true:
+
+\begin{itemize}
+\tightlist
+\item
+  \(B \trianglelefteq G'\)
+\item
+  there exist \(R', P' \epsilon Z_+\) such that if \(G'(w) = \bigcup_{z \in T_w}B(z)\) then \[G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)\]
+\end{itemize}
 
@@
 
-and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping)
-
-Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
-The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
-The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R')^{th}$ granule of $G'$ will have the same number of granules of $B$.
-The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$.
-Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.
-
-```{r quasi-circular-example, echo=FALSE,out.width = "100%",fig.cap="Relating quasi-circular granularities and bottom linear granularity "}
-knitr::include_graphics("Figs/quasi-circular-example.png")
-```
-
-<!-- If we ignore leap years and all its exception, the periodicity of the grouping (days, months), P' = 365, R' = 12, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
-Considering leap years and all their exceptions, period is 400 years and hence $P' = (3*365 + 366)$, $R' = 12*400$.
-\autoref{eq:eq3} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
--->
-
-<!-- Then $f(x, y)$ can be  computed using modular arithmetic as follows: -->
-
-<!-- $$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$ -->
-
-<!-- The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
--->
-
-<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
-Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
--->
-
-<!-- Formal definition of circular granularity -->
-
-<!-- **Definition: Equivalence class** Let $m \in N \backslash{0}$.
-For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
--->
-
-<!-- The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$.
-Thus $Z_m = \{[0], [1], ..., [m-1]\}$.
-However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.
--->
-
-<!-- **Definition:** A **circular granularity** $C$, joining two linear granularity with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that -->
-
-<!-- $C(s) = (s\mod m)$ for $s \in Z$. -->
-
-<!-- For example, suppose $C$ is a circular granularity denoting hour-of-day and we have hourly data for 100 hours.
-The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
--->
-
-<!-- **Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once. -->
-
-<!-- **Definition:** A **circular granule** represents an equivalence class inside each cycle. -->
-
-<!-- ### Aperiodic {#aperiodic-gran-def} -->
-<!-- Formal definition of aperiodic circular granularity -->
-
- <!-- Why we need single order up and multiple order up granularities -->
-<!-- A time series which is characterized by being composed of repeating cycles. -->
-
-<!-- Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
-Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
--->
-
-<!-- The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity.
-The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
--->
-
-It can be noted here that if a linear granularity $G'$ is quasi-periodic (Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
+and, number of granules of \(B\) in each granule of \(G'\) is not a constant (irregular mapping)
+
+Here \(w \mod R'\) represents the index that must be shifted to obtain \(G'(w)\).
+The idea here is if we know the composition of each of the granules of \(G'\) in terms of granules of \(B\) for one period, we can find the composition of any granule of \(G'\) beyond a period since the ``pattern'' repeats itself along the time domain due to the periodic property.
+The periodic property also ensures that \(\vert T_w \vert = \vert T_{w\mod R'} \vert\) since every \(w^{th}\) and \((w+R')^{th}\) granule of \(G'\) will have the same number of granules of \(B\).
+The term \(\sum_{w=0}^{k-1}\vert T_{w}\vert\) denotes the number of granules of \(B\) till the \((k-1)^{th}\) granule of \(G'\).
+Since \(\vert T_w \vert = \vert T_{w\mod R'} \vert\), the number of granules of \(B\) till the \((k-1)^{th}\) granule of \(G'\) becomes \(\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert\) in Definition \autoref{def:quasicircular}.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/quasi-circular-example}
+
+}
+
+\caption{Relating quasi-circular granularities and bottom linear granularity }\label{fig:quasi-circular-example}
+\end{figure}
+
+It can be noted here that if a linear granularity \(G'\) is quasi-periodic (Definition \autoref{def:quasi}) with respect to \(B\), then \autoref{eq:quasi} can be modified as follows to account for exceptions \(E = [e_{begin}, e_{end}]\) (by Definition \autoref{def:quasi}).
 
@@
 
-for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
-
-## Aperiodic {#sec:aperiodic-gran-def}
+for \(z \in T_k\) and \(e_k\) is the number of exceptions in \(\bigcup\limits_{w = 0}^{k} T_w\) and \(E_{u} = [e_{begin}(u), e_{end}(u)]\).
+
+\hypertarget{sec:aperiodic-gran-def}{%
+\subsection{Aperiodic}\label{sec:aperiodic-gran-def}}
 
@@
 All of these are recurring events, but with non-periodic patterns.
-As such, plausible $P_i$ from \autoref{tab:quasi} can be infinite for aperiodic granularities.
+As such, plausible \(P_i\) from \autoref{tab:quasi} can be infinite for aperiodic granularities.
 
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ (that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity).
-Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
-Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that
+Consider n aperiodic linear granularities \(M_i\) \(\forall \{i\in{1, 2, \dots, n\}}\) (that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity).
+Further assume that \(B \trianglelefteq M_i\) \(\forall \{i\in{1, 2, \dots, n\}}\).
+Then according to \autoref{def:groupsinto}, for each index \(j\) there exists a (possibly infinite) subset \(T_{\{i_j\}}\) of the integers such that
 \begin{equation*}
@@
 \end{equation*}
-and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
+and suppose \(M = \bigcup_{i=1}^{n}M_{i}\) is formed by collecting the granules of \(\{M_1, M_2, \dots, M_n\}\).
 
@@
 
-**Example**: \autoref{fig:aperiodic-example} shows two aperiodic linear granularity $M_1$ and $M_2$.
+\textbf{Example}: \autoref{fig:aperiodic-example} shows two aperiodic linear granularity \(M_1\) and \(M_2\).
 Each granule is represented by a box.
 The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set and $A_{B, M}$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
-
-**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labour day and M can be conceived as the aperiodic linear granularity denoting any holidays.
+The first slot in the diagram shows the index set and \(A_{B, M}\) consists of three categories, \(L(1)\) corresponding to the first aperiodic event \(M_1\), \(L(2)\) corresponding to aperiodic event \(M_2\) and the \(L(0)\) corresponding to no event/aperiodic linear granularity.
+
+\textbf{Discussion}: Here, \(M_1\) can be an aperiodic linear granularity denoting block leaves during Christmas, \(M_2\) can be the one denoting Easter, \(M_n\) can be Labour day and M can be conceived as the aperiodic linear granularity denoting any holidays.
 The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
 
-<!-- map a time index to a categorical variable containing binary values.
-It takes the value 1 whenever the time index coincides with the events in context (e.g.
-public or school holiday) and 0 when it does not.
-Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector.
--->
-
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
-knitr::include_graphics("Figs/aperiodic-example.png")
-```
-
-<!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
-
-# Cyclic calendar algebra {#sec:cyclic-calendar}
-
-In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/aperiodic-example}
+
+}
+
+\caption{Relating aperiodic granularity and bottom linear granularity }\label{fig:aperiodic-example}
+\end{figure}
+
+\hypertarget{sec:cyclic-calendar}{%
+\section{Cyclic calendar algebra}\label{sec:cyclic-calendar}}
+
+In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in \citet{Bettini1998-ed}.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
-This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
-Since, our method caters to computation of cyclic granularities, we shall refer to it as "cyclic calendar algebra".
-
-The cyclic calendar algebra consists of broadly two kinds of operations:  
-(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities  
-(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities
-
-The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
+This is similar to calendar algebra in \citet{Ning2002-tf}, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
+Since, our method caters to computation of cyclic granularities, we shall refer to it as ``cyclic calendar algebra''.
+
+The cyclic calendar algebra consists of broadly two kinds of operations:\\
+(1) \textbf{single-to-multiple:} representation of multiple-order-up cyclic granularities from single order-up cyclic granularities\\
+(2) \textbf{multiple-to-single:} representation of single-order-up cyclic granularities from multiple order-up cyclic granularities
+
+The hierarchical structure of time creates a natural nested ordering which can produce \textbf{single-order-up} or \textbf{multiple-order-up} granularities.
 We shall use the notion of a hierarchy table and order to define them.
 
-**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
-$G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
-
-**Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
-For example, if we consider two linear granularities $G$ and $H$, such that $G$ is finer than or groups into $H$, then $H$ is of higher order than $G$.
+\textbf{Hierarchy table} Let \(H_n: (G, C, P)\) be a hierarchy model containing n linear granularities.
+\(G_{l}\) represents the linear granularity of order l and \(P(l,m)\) represents the period length of \(G_{l}\) with respect to \(G_{m}\) and \(C_{G(l),G(m)}\) represents the cyclic granularity that relates linear granularity of order l and m, \(\forall l,m \in n, l<m\).
+
+\textbf{Order} of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
+For example, if we consider two linear granularities \(G\) and \(H\), such that \(G\) is finer than or groups into \(H\), then \(H\) is of higher order than \(G\).
 In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
@@
 
-<!-- Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities and the other granularities are subsequently more "coarse" than the earlier one.
--->
-
-<!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
--->
-
-<!-- some discussions on primary units of calendar -->
-**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
+\textbf{Example:} So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
 But it is far from being the only one.
-All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars [@Reingold2001-kf].
+All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars \citep{Reingold2001-kf}.
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
+The French revolutionary calendar divided each day into 10 ``hours'', each ``hour'' into 100 ``minutes'' and each ``minute'' into 100 ``seconds''.
 Nevertheless, for any calendar a hierarchy can be defined.
@@
 
- - 1 kin = 1 day
- - 1 uinal = 20 kin
- - 1 tun  = 18 uinal
- - 1 katun = 20 tun
- - 1 baktun = 20 katun
+\begin{itemize}
+\tightlist
+\item
+  1 kin = 1 day
+\item
+  1 uinal = 20 kin
+\item
+  1 tun = 18 uinal
+\item
+  1 katun = 20 tun
+\item
+  1 baktun = 20 katun
+\end{itemize}
 
@@
 
-|  G	        |C                 |P
-|------------	|------------------|------------|
-|  kin      	|kin-of-uinal      |  20        |
-|  uinal    	|uinal-of-tun      |  18  	    |
-|  tun      	|tun-of-katun      |  20	      |
-|  katun    	|katun-of-baktun   |  20        |
-|  baktun   	| 1                |	1         |
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+G & C & P\tabularnewline
+\midrule
+\endhead
+kin & kin-of-uinal & 20\tabularnewline
+uinal & uinal-of-tun & 18\tabularnewline
+tun & tun-of-katun & 20\tabularnewline
+katun & katun-of-baktun & 20\tabularnewline
+baktun & 1 & 1\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-## Single-to-multiple {#sec:single-to-multiple}
-
-### All circular single order-up granularities {#sec:all-circular}
+\hypertarget{sec:single-to-multiple}{%
+\subsection{Single-to-multiple}\label{sec:single-to-multiple}}
+
+\hypertarget{sec:all-circular}{%
+\subsubsection{All circular single order-up granularities}\label{sec:all-circular}}
 
 Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity.
-Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
+Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, \(L(x) = x\quad \forall x\).
 The label mapping of the resultant multiple-order-up granularity can however be chosen arbitrarily, depending on the context.
@@
 
-*Example:* Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
+\emph{Example:} Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal\_katun for Mayan calendar.
 
@@
 
-### Circular or quasi-circular single order-up granularities
+\hypertarget{circular-or-quasi-circular-single-order-up-granularities}{%
+\subsubsection{Circular or quasi-circular single order-up granularities}\label{circular-or-quasi-circular-single-order-up-granularities}}
 
@@
 
-<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
-However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
--->
-
-|  G	        |C                 |P
-|------------	|------------------|----------------|
-|  minute     |minute-of-hour    |  60            |
-|  hour    	  |hour-of-day       |  24  	        |
-|  day      	|day-of-month      |quasi-circular  |
-|  month    	|month-of-year     |  12            |
-|  year   	  | 1                |	1             |
-
-There can be three scenarios for obtaining multiple-order-granularities here:  
-- cyclic granularities relating two linear granularities whose orders are less than day  
-- cyclic granularities relating two linear granularities whose orders are more than month  
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+G & C & P\tabularnewline
+\midrule
+\endhead
+minute & minute-of-hour & 60\tabularnewline
+hour & hour-of-day & 24\tabularnewline
+day & day-of-month & quasi-circular\tabularnewline
+month & month-of-year & 12\tabularnewline
+year & 1 & 1\tabularnewline
+\bottomrule
+\end{longtable}
+
+There can be three scenarios for obtaining multiple-order-granularities here:\\
+- cyclic granularities relating two linear granularities whose orders are less than day\\
+- cyclic granularities relating two linear granularities whose orders are more than month\\
 - granularities relating two linear granularities with order at most day and another with order at least month
@@
 Examples might include hour-of-month or day-of-year.
-Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$.
+Let us consider the computation of \(C_{hour, month}(z)\) for a \(z\) such that \(C_{hour, day}(z) = 12\), \(C_{day, month}(z) = 20\) and \(C_{month, year}(z) = 9\).
 
@@
 
-where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+where, \(T_w\) is such that \(month(w) = \bigcup_{z \in T_w}day(z)\) and \(\vert T_w \vert\) is the cardinality of set \(T_w\).
 
 Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
-$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
-However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
-It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
+\(C_{hour, month}(z)\) can be represented as function of circular granularity \(C_{hour, day}(z)\) and quasi-circular granularity \(C_{day, month}(z)\), both of which are single order-up.
+However, \(C_{day, year}(z)\) can not be computed only with single order-up granularities \(C_{day, month}(z)\) and \(C_{month, year}(z)\).
+It also requires the knowledge of the composition of the linear granularity \(days\) within \(months\) in an \(year\) for computation of \(\vert T_w \vert\).
 
@@
 
-<!-- Both of them can be circular or quasi-circular or it can be each of both. In case they are both quasi-circular or one circular and the other quasi-circular, we need to adjust the formula -  -->
-
-For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, two different approaches need to be employed for computing multiple order-up granularities from single order-up ones.
-
-- Case when $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+For example, with \(C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))\), two different approaches need to be employed for computing multiple order-up granularities from single order-up ones.
+
+\begin{itemize}
+\tightlist
+\item
+  Case when \(C_{l,m'}(z)\) is circular and \(C_{m',m}(z)\) is quasi-circular
+\end{itemize}
 
@@
 
-- Case when  $C_{l,m'}(z)$  is quasi-circular and $C_{m',m}(z)$ is circular
+\begin{itemize}
+\tightlist
+\item
+  Case when \(C_{l,m'}(z)\) is quasi-circular and \(C_{m',m}(z)\) is circular
+\end{itemize}
 
@@
 
-where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-
-<!-- \begin{equation} \label{eq:multifromsingle-quasi} -->
-<!-- \begin{split} -->
-<!-- C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\  -->
-<!-- & = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))) + P(l, m')(C_{G_{m'},G_{m}}(z))\\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-
-<!-- where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping. -->
-
-If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z) are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
-
-<!-- The aperiodic granularity $C_{day, quarter}(z))$ in \autoref{eq:hour-month} can be broken further. -->
-
-<!-- \begin{equation} \label{eq11} -->
-<!-- \begin{split} -->
-<!-- A_{(hour, month)} & = (\sum_{w=0}^{9 - 1}(\vert T(w)\vert) +  20)/24 + 12 \\ -->
-<!--                   & = (\sum_{w=0}^{C_{month, year} - 1}(\vert T(w)\vert) +  A_{day, month})*k(hour_day) + A_{hour, day} \\ -->
-<!--                   & = (A{day_year} - 1)*k(hour_day) + A_{hour, day} \\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-<!-- where, $month(w) = \bigcup_{z \in T_w}day(z)$ -->
-
-<!--  With one aperiodic granularity, we can write the following:  -->
-<!-- A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included} -->
-
-## Multiple-to-single {#sec:multiple-to-single}
-
-### Muliple order-up circular granularities {#sec:all-circular-multiple}
-
-For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
+where, \(T_w\) is such that \(G_{m'}(w) = \bigcup_{z \in T_w}G_{l}\) and \(\vert T_w \vert\) is the cardinality of set \(T_w\).
+
+If \(C_{G_{l},G_{m'}}(z)\) or \$C\_\{G\_\{l\},G\_\{m'\}\}(z) are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
+
+\hypertarget{sec:multiple-to-single}{%
+\subsection{Multiple-to-single}\label{sec:multiple-to-single}}
+
+\hypertarget{sec:all-circular-multiple}{%
+\subsubsection{Muliple order-up circular granularities}\label{sec:all-circular-multiple}}
+
+For a hierarchy table \(H_n: (G, C, k)\) with \(l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}\) and \(l_2<l_1\) and \(m_2>m_1\), we have
 
@@
 
-**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}
+\textbf{Example:} Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}
 
@@
 
-### Muliple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
-
-<!-- \begin{equation} \label{eq:quasi-circular-multiple} -->
-<!-- C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod P(m_1, m_2)) -->
-<!-- \end{equation} -->
+\hypertarget{sec:quasi-circular-multiple}{%
+\subsubsection{Muliple order-up quasi-circular granularities}\label{sec:quasi-circular-multiple}}
 
@@
 
-<!-- as they promote the exploratory analysis through visualization. -->
-
-```{r}
-# grid.arrange(p1,p2,p3,p4,nrow = 2)
-
-#
-# x <- gridExtra::arrangeGrob (p5, p6, p1, p2, p3, p4,newpage = TRUE)
-# grid::grid.draw(x)
-
-#
-# gridExtra::grid.arrange(
-#   p1,
-#   p2,
-#   p3,
-#   p4,
-#   p5,
-#   p6,
-#   nrow = 3,
-#   top = "Title of the page",
-#   bottom = grid::textGrob(
-#     "this footnote is right-justified",
-#     gp = grid::gpar(fontface = 3, fontsize = 9),
-#     hjust = 1,
-#     x = 1
-#   )
-# )
-# plot.new()
-# mtext('Various probability distribution plot of electricity consumption data of Victoria from 2012 to 2014 (top left) Letter Value plot by DoM and MoY (top right) Decile Plot by HoD and DoW (middle left) Box plot by DoY and DoM (middle right) Violin plot of DoM and WoM (bottom left) Ridge plot by WoM and WoY (bottom right) Decile plot by DoY and MoY', side = 3, line = - 10, outer = TRUE)
-```
-
-<!-- More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
-When C1 and C2 are used as aesthetics or facets, problems arise when we encounter empty combinations.
-Let $S_{i,j}$ be the set of combination of the levels of C1 and C2.
-In this example, we have 12 such sets $S_{i,j}$ because i can take 4 values and j can take 3 values.
--->
-
-<!-- Thus, the common link that differentiates harmonies from clashes are: -->
-
-<!--   a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis. -->
-
-<!--   b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets. -->
-
-# Data structure{#sec:data-structure}
-
-<!-- How volumes of continuously collected data can be visualized to ease further decision making?  -->
+\hypertarget{sec:data-structure}{%
+\section{Data structure}\label{sec:data-structure}}
 
 Effective exploration or good visualization require good data structures.
-Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization.
+Commonly, simple sequences of time value pairs \((<t_0, v_0> \dots, <t_n, v_n>)\) are the basis of analysis and visualization.
 It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
@@
 
-A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
-Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
+A recent tidy data structure to support exploration and modeling of temporal data is tsibble \citep{wang2019tsibble}, where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
+Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in (Figure \ref{fig:data-structure} for exploration of temporal data of this kind.
 This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities.
-Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
-Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
-
-<!--  These subsets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities. -->
-
-<!-- This data representation can lead to comprehensive exploration by conditioning the levels taken by one or more of the cyclic granularities. -->
-
-<!-- This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data by creating same plots using data subsets corresponding to each level of the conditioned variable.
--->
-
-<!-- These are based on conditioning the values taken on by one or more of the variables in a data set.
-This means creating same plots using data subsets corresponding to each level of the conditioned variable.
-This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
--->
-
-```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure", fig.pos="htb"}
-include_graphics("Figs/data-structure-v1.png")
-```
-
-## Synergy of the cyclic granularities {#sec:synergy}
-
-The way cyclic granularities relate become important when we consider the data structure in \@autoref{fig:data-structure}.
-While considering data sets of the form $<C_1, C_2, v>$, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
-Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
-We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the categories/levels of the cyclic granularities.
+Let us consider two cyclic granularities \(C_1\) and \(C_2\), such that \(C_1\) maps index set to a set \(\{A_1, A_2, A_3, \dots, A_n\)\}, and \(C_2\) maps index set to a set \(\{B_1, B_2, B_3, \dots, B_m\)\} and \(v\) denotes the measurement variable.
+Data sets of the form \textless{}\(C_1\), \(C_2\), \(v\)\textgreater{} can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+
+\begin{figure}[htb]
+
+{\centering \includegraphics[width=\textwidth]{Figs/data-structure-v1}
+
+}
+
+\caption{The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure}\label{fig:data-structure}
+\end{figure}
+
+\hypertarget{sec:synergy}{%
+\subsection{Synergy of the cyclic granularities}\label{sec:synergy}}
+
+The way cyclic granularities relate become important when we consider the data structure in @autoref\{fig:data-structure\}.
+While considering data sets of the form \(<C_1, C_2, v>\), let \(S_{ij}\) be the set of index set such that for all \(s \in S_{ij}\), \(C_1(s) = A_i\) and \(C_2(s) = B_j\).
+Data subsets like \textless{}\(A_i\), \(B_j\), \(v(s)\)\textgreater{} can be obtained for all \(i \in {1, 2, \dots, n}\) and \(j \in {1, 2, \dots, m}\) which will lead to \(nm\) data subsets.
+We will discuss few cases, where one or more of these \(nm\) sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the categories/levels of the cyclic granularities.
 To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
 
-<!-- The cyclic time granularities which are essentially categorizations of time are discrete variable, mapped as a factor since they have an inherent order.
-Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
-To define a general framework, we define harmony and clashes.
--->
-
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
+Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
+These are called ``structurally'' empty combinations.
+Let us take a specific example, where \(C_1\) maps index set to day-of-month and \(C_2\) maps index set to week-of-month.
+Here \(C_1\) can have 31 levels/categories while \(C_2\) can have 5 categories.
+There will be \(31\times 5=155\) sets \(S_{ij}\) corresponding to the possible combinations of day-of-month and week-of-month.
 Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
+For example \(S_{1,5}\), \(S_{21,2}\), etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
 These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-<!-- What  is interaction and some examples why it is important to study interaction -->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
-
-<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
--->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
--->
-
-<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
--->
-
-<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
--->
-
-<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
--->
-
-<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
--->
-
-<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
--->
-
-<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
--->
-
-<!-- Before exploring the behavior of a "dependent variable" across time granularities, -->
-<!-- it is important to know how these granularities interact with each other.
-If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
--->
-
-<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
--->
-
-<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
-It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
--->
-
-<!-- Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
-That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
--->
-
-<!-- The cyclic time granularities which are essentially categorizations of time are discrete variable, mapped as a factor since they have an inherent order.
-Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
-To define a general framework, we define harmony and clashes.
--->
-
-<!-- Now, many situations can lead to some of these sets being empty. Let us discuss the following cases, where one or more of these sets can be empty. -->
-
-Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ maps index set to day-of-month and $C_2$ maps index set to week-of-month.
-Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of day-of-month and week-of-month.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
-These are structurally empty sets in that it is impossible for them to have any observations.
-
-<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
 
 Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
-These are called "event-driven" empty combinations.
+These are called ``event-driven'' empty combinations.
 Again, let us consider a specific example to illustrate this.
-Let $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
-Here $C_1$ can have 7 levels while $C_2$ can have 2 levels.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of day-of-week and WorkingDay/NonWorkingDay.
+Let \(C_1\) be day-of-week and \(C_2\) be WorkingDay/NonWorkingDay.
+Here \(C_1\) can have 7 levels while \(C_2\) can have 2 levels.
+So there are 14 sets \(S_{ij}\) corresponding to the possible combinations of day-of-week and WorkingDay/NonWorkingDay.
 While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations.
@@
 
-<!-- However, this is different from the first example because these are not structurally empty.
-Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
--->
-
 Thirdly, empty combinations can be a result of how granularities are constructed.
-Let $C_1$ maps index set to Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
+Let \(C_1\) maps index set to Business-days, which are days from Monday to Friday except holidays and \(C_2\) is day-of-month.
 Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
 This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
-Hence, they are referred to as "build-based" empty combinations.
-
-An example when there will be no empty combinations could be where $C_1$ maps index set to day-of-week and $C_2$ maps index set to month-of-year.
-Here $C_1$ can have 7 levels while $C_2$ can have 12 levels.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of day-of-week and month-of-year.
+Hence, they are referred to as ``build-based'' empty combinations.
+
+An example when there will be no empty combinations could be where \(C_1\) maps index set to day-of-week and \(C_2\) maps index set to month-of-year.
+Here \(C_1\) can have 7 levels while \(C_2\) can have 12 levels.
+So there are \(12\times7=84\) sets \(S_{ij}\) corresponding to the possible combinations of day-of-week and month-of-year.
 All of these are non-empty because every day-of-week can occur in every month.
 
-A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are referred as **harmonies**.
-
-## Measured variable {#sec:measured-variable}
-
-<!-- Moreover, the measured variable in the considered data structure can be dicrete or continuous. -->
+A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as \textbf{clashes}.
+And the ones that do not lead to any missing combinations are referred as \textbf{harmonies}.
+
+\hypertarget{sec:measured-variable}{%
+\subsection{Measured variable}\label{sec:measured-variable}}
 
@@
 These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
-Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
-
-<!-- The most commonly used density estimation method is the kernel density estimator.  -->
-
-If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$ and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$, the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by:
+Several forms of non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( \citet{Silverman1986-cl}, \citet{Jones1992-pj}) are as follows:
+
+If we assume that \(X_1, X_2, \dots, X_n \sim F\) are independent and identically distributed observations from an univariate distribution with probability density/mass function \(f(.)\) and \(X_{(1)},\dots X_{(n)}\) denote the order statistics of \(X_1, X_2, \dots, X_n\), the Kernel density estimator \(\hat{f(x)}\) with Kernel \(K\) is defined by:
 
@@
 
-where $h$ is the smoothing parameter or the bandwidth. The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by:
-
-<!-- $Q$, $(q \equiv  Q')$, which is the quantile density function -->
-<!-- $ Q_{n} (u) = \sum_{i=1}^ n X_{(i)} I((i-1)/n <u \leq i/n)$\\ -->
-<!--             $= X_(1) + \sum_{i=1}^ n (X_{(i)} - X_{(i-1)}) I((i-1)/n < u)$ -->
-
-<!-- and most widely used definition of sample quantile is given in  by: -->
-
-<!-- $$\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}$$ where,  $(j - m)/n \leq p < (j - m + 1)/n$,  -->
-<!-- for some $m \in R$ and $0 < \gamma < 1$.  $\hat{Qi(p)}$ is the ith sample quantile  -->
+where \(h\) is the smoothing parameter or the bandwidth. The sample quantile estimator \(\hat{q(u)}\) with Kernel \(K\) is given by:
 
@@
 
-where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$.
-
- <!-- Both pdf and pmf are concerned with point probabilities of the random variable.
-Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d.
--->
-
-<!-- Instead of assuming a parametric model for the distribution (e.g.
-a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g.
-differentiable).
-It is then possible to estimate the u2nknown density function f(·).
--->
-
-<!-- The histogram is the oldest and most popular density estimator method.
-The form of the histogram depends very much on the tuning parameters class width and origin.
-The naive estimator was introduced by Fix and Hodges (1951) and first discussed by Rosenblatt(1956).
-Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962).
-The nearest neighbor estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977).
-There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.
--->
-
-<!-- The inverse $Q = F^{-1}$, which is the quantile function can be defined as $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$.  -->
+where \(Q(u)= F^{-1}(u) = inf({x: F(x) >u)}\), \(0<u< 1\) and \(q \equiv Q'\).
 
@@
 
-where $\kappa = \int k^2(y)dy$.
-
-<!-- other non parametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
-Also sample quantiles can be of the form that are not defined for all p including hinges, letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
--->
-
-<!-- would involve a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis. -->
-
- <!-- Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space. -->
-
-# Visualization {#sec:visualization}
-
-<!-- Why visualize -->
-<!-- Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
-However, improper choice of models can lead to wrong predictions.
-One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
--->
-
-<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
-
-<!-- Include why distribution plot, why 2D, why faceting in introduction -->
-<!-- Why distribution plot -->
-<!-- Huge data difficult to summaries using one statistic-->
-
-<!-- With huge amount of data being available, mean, median or any one summary statistic might not enough to understand a data set. Soon enough following questions become more interesting:   -->
-
-<!--  - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?  -->
-
-<!--  - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
-More generally, how the variation is the data set changes across different percentiles/deciles?
--->
-
-<!--   - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median? -->
-
-<!-- This is when displaying a probability distribution becomes a potentially useful approach. -->
-<!-- The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
--->
-
-<!-- ## Distributions conditional on bivariate granularities -->
-
-<!-- Why 2D -->
-
-<!-- Graphical output devices are flat.
-If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
-As soon as the data move to three variables, we need to infer multi-dimensional structure by a two-dimensional medium[Cleveland, elements of graphing data].
-Any attempt to encode all these temporal granularities together to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
--->
-
-<!-- Why facet-ing -->
-
-<!-- `ggplot2` (Wickham, 2016) is an excellent tool that facilitates the process of mapping different variables to a 2D frame through grammar of graphics.
-One aspect of ggplot2 is faceting, inspired by Trellis plots (Cleveland, 1993).
-These are based on conditioning the values taken on by one or more of the variables in a data set.
-This means creating same plots using data subsets corresponding to each level of the conditioned variable.
-This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
--->
-
-<!-- What are we visualizing -->
-
-<!-- We will see the distribution of the time series across bivariate temporal granularities through faceting approach with one temporal granularity plotted along the x-axis, the other one across facets and the dependent time series variable on the y-axis.
--->
-
-<!-- A well designed visual representation can replace cognitive calculations with simple perpetual inferences and improve speed of comprehension and decision making.
-However, creating a visualisation requires a number of nuanced judgements.
-The challenge being given a data set, the number of visual encodings and hence the space of possible visual design is extremely huge.
--->
-
-The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by [@Wickham2009pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
+where \(\kappa = \int k^2(y)dy\).
+
+\hypertarget{sec:visualization}{%
+\section{Visualization}\label{sec:visualization}}
+
+The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space \citep{Wilkinson1999-nk}.
+The layered grammar of graphics proposed by \citep{Wickham2009pk}, which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
 There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
@@
 
-<!-- Briefly, the grammatical elements of any plot are as follows: -->
-
-<!-- - **Layers**:  made up of data, mapping, statistical transformation, geometric objects and position adjustments -->
-
-<!--   - **Data**: observations made on a number of variables, typically thought of as a sample from a population distribution -->
-
-<!--   - **Mappings** describing how variables are mapped to elements of the plot, such as horizontal axis or color -->
-
-<!--   - **Statistical transformation** can summarize the data in many useful ways. The geometric object determines the physical representation of the data in the graph. -->
-
-<!--   - **Geometric objects** control the type of plot that is created and are classified by their dimensions, for example points (0 dimension), line (1 dimension), polygons(2 dimensions) and so on.
--->
-
-<!--   - **Position adjustment** of elements on the plot are needed for dense data, otherwise data points might obscure one another. -->
-
-<!-- - **Facet** dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
--->
-
-<!-- - **Scale** controls how data is mapped to aesthetic attributes. For every aesthetic property employed in a layer, we need one scale.  -->
-
-<!-- - **A coordinate system** maps the position of objects onto the plane of the plot, and controls how the axes and grid lines are drawn.  -->
-
-Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
-
-- Data: <$C_1$, $C_2$, $v$>
-- Aesthetic mapping: $C_1$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v$
-- Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
-- Facet:  $C_2$
-
-<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
-
-<!-- The three essential component layers are data, aesthetic mapping and the geometric objects. -->
-<!-- Data are observations made on a number of variables, typically thought of as a sample from a population distribution.
-Mappings describe how variables are mapped to elements of the plot, such as horizontal axis or color -->
-
-<!-- the second layer is the aesthetic mapping, which refers to the scales on to which we want to map our data -->
-<!-- The third essential layer is the geometry which decides how the plot will look. After the essential layers, the next layer being used is the facet which dictates how to split up our plot -->
-
-<!-- Aesthetics refers to the scales on to which we want to map our data. we create plots by layering grammatical elements on top of one another and use aesthetic maps -->
-
-<!-- The grammar of graphics establish the building blocks for creating meaningful visualizations. It -->
-<!-- This means we have a framework that allows us to communicate in a way.  -->
-
-<!-- .
-allows us to produce graphics using the same structured thinking that is used to design analysis.
-Briefly, any graphics are made up of **Layers**, **Facet**, **Scale** and a  **coordinate system**.
-**Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
--->
-
-<!-- The three key components of every visualization is: data, aesthetic mappings between variables in the data and visual attributes, and at least one layer which describes how to render each observation.
-Layers are usually created with a geom function.
-Considering the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
-<!-- - Data: <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
-<!-- - Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position. -->
-<!-- - Layer: any geom that summarizes distribution -->
-
-<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
-<!-- - Facet:  $B_j(s)$ -->
-
-<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
-The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
--->
-
-<!-- Hence, the ingredients for the plots are subsets of numbers (or categorical measurements), and the choice of faceting or axis alignment for the subset strata.
-The basic plot choice is display of density.
-Different plots might be appropriate depending on which features of the density we arae interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
-We will look at each of these aspects separately and analyze the effect of each on visualization.
-While we accomplish that, we need to bear in mind that any customized species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
-We are trying to address problems where even with good aesthetic qualities and good data, the graph might be confusing or misleading because of how people process the information.
--->
-
-<!-- Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
-The harmonies when plotted together can help exploration.
-But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
-This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
--->
-
-## Choice of statistical transformations and geometric objects
-
-<!-- Common benefits and challenges of empirical and density based visualization -->
+Drawing from the grammar of graphics, if \textless{}\(C_1\), \(C_2\), \(v\)\textgreater{} serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
+
+\begin{itemize}
+\tightlist
+\item
+  Data: \textless{}\(C_1\), \(C_2\), \(v\)\textgreater{}
+\item
+  Aesthetic mapping: \(C_1\) mapped to \(x\) position and \(v\) to \(y\) position
+\item
+  Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of \(v\)
+\item
+  Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
+\item
+  Facet: \(C_2\)
+\end{itemize}
+
+\hypertarget{choice-of-statistical-transformations-and-geometric-objects}{%
+\subsection{Choice of statistical transformations and geometric objects}\label{choice-of-statistical-transformations-and-geometric-objects}}
 
@@
 
-<!--  However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. -->
-
 We will discuss few conventional and recent ways to plot distributions using both of these methods.
 
-<!-- Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it.
-These displays depend on the specific estimation procedure (e.g., kernel density estimate)  -->
-
-<!--  Furthermore, these methods are subject to the same sample size restrictions and challenges that apply to any density estimation.  -->
-<!-- The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
-They can be useful in spotting multimodality, however, they might become obscuring with too many categories and information on the entire distribution can add to cognitive load.
-Also, the probabilities in density plot are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
-As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
--->
-
-### Descriptive statistics based displays
-
-Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{th}$ and $(n + 1 - ([n/2] + 1)/2)^{th}$ order statistics are taken as estimates of the quartiles.
+\hypertarget{descriptive-statistics-based-displays}{%
+\subsubsection{Descriptive statistics based displays}\label{descriptive-statistics-based-displays}}
+
+Most commonly used displays include the box plots \citep{Tukey1977-jx} which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+These displays are based on descriptive statistics, for example, \(([n/2] + 1)/2^{th}\) and \((n + 1 - ([n/2] + 1)/2)^{th}\) order statistics are taken as estimates of the quartiles.
 Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
+Notched box plots \citep{Mcgill1978-hg} has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
+The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
 However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
@@
 
-<!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+The letter-value box plot \citep{Hofmann2017-sg} was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
 These are particularly useful for large data sets.
-“outliers” in letter value plots are those observations beyond the most extreme letter value.
+``outliers'' in letter value plots are those observations beyond the most extreme letter value.
 The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
-
-<!-- Description and lit review of quantile plots -->
 
@@
 These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
-While in a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
-outside 1.5 times the interquartile range above the upper quartile and below the lower quartile), in quantile plots outliers are open to interpretation and not shown.
-
-<!-- Description and lit review of density plots -->
-
-<!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  -->
-
-### Kernel density based displays
-
-Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi].
+While in a boxplot, an outlier is defined as a data point that is located outside the fences (``whiskers'') of the boxplot (e.g.~outside 1.5 times the interquartile range above the upper quartile and below the lower quartile), in quantile plots outliers are open to interpretation and not shown.
+
+\hypertarget{kernel-density-based-displays}{%
+\subsubsection{Kernel density based displays}\label{kernel-density-based-displays}}
+
+Traditional ways to visualize kernel densities include violin plots \citep{Hintze1998-zi}.
 The shape of the violin represents the density estimate of the variable.
@@
 
-The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
+The summary plot \citep{Potter2010-qc} combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
 This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
 
-<!-- Description and lit review of density plots -->
 A Ridge line plot (sometimes called Joy plot) shows the distribution for several groups.
@@
 
-<!-- Description and lit review of HDR plots -->
-The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
+The highest density region (HDR) box plot proposed by \citep{Hyndman1996-ft} displays a probability density region that contains points of relatively highest density.
 The probabilities for which the summarization is required can be chosen based on the requirement.
@@
 
-## Facet and aesthetic variables
-
-### Levels
+\hypertarget{facet-and-aesthetic-variables}{%
+\subsection{Facet and aesthetic variables}\label{facet-and-aesthetic-variables}}
+
+\hypertarget{levels}{%
+\subsubsection{Levels}\label{levels}}
 
 We will discuss the effect of the levels of cyclic granularities in this section.
-Recall that cyclic granularities $C_1$ and $C_2$ are such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-$A_i$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.
+Recall that cyclic granularities \(C_1\) and \(C_2\) are such that \(C_1\) maps index set to a set \(\{A_1, A_2, A_3, \dots, A_n\)\}, and \(C_2\) maps index set to a set \(\{B_1, B_2, B_3, \dots, B_m\)\}.
+\(A_i\)'s and \(B_j\)'s are referred to as the levels of \(C_1\) and \(C_2\) for all \(i \in {1, 2, \dots, n}\) and \(j \in {1, 2, \dots, m}\) respectively.
 The levels have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
@@
 
-<!-- The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.  -->
-
- <!-- - If levels of both the granularities are low, then any statistical distribution plot might be chosen depending on which feature of the distribution one is interested to look at.  -->
-
 The following principles are useful while choosing distribution plots given two temporal granularities:
 
-  - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus.
-
- - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories.
-
-  - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.
-
-  - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.
-
-<!-- Let us consider case by case and see which plots are better suitable in which scenarios. -->
-
-<!--  Assumptions should be made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.  -->
-<!-- Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot. -->
-
-<!-- Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
-These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
-Let us consider case by case and see which plots are better suitable in which scenarios.
--->
-
-<!-- Faceting creates small multiples of data each representing a different subset of the data.
-Small multiples are powerful for exploratory data analysis where we can rapidly compare patterns in different parts of the data and see whether they are the same or different.
-But space and resolution quickly become a problem if the number of levels of facet variables is too high.
-Also, Cleveland suggested that a graphic should display as much information as it can, with the lowest possible cognitive strain to the viewer.
--->
-
-### Interaction
+\begin{itemize}
+\item
+  If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus.
+\item
+  If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories.
+\item
+  If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.
+\item
+  If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.
+\end{itemize}
+
+\hypertarget{interaction}{%
+\subsubsection{Interaction}\label{interaction}}
 
@@
 
-```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
+\begin{figure}
+
+\includegraphics[width=0.5\linewidth]{figure/allFig-1} \includegraphics[width=0.5\linewidth]{figure/allFig-2} \includegraphics[width=0.5\linewidth]{figure/allFig-3} \includegraphics[width=0.5\linewidth]{figure/allFig-4} \includegraphics[width=0.5\linewidth]{figure/allFig-5} \includegraphics[width=0.5\linewidth]{figure/allFig-6} \hfill{}
+
+\caption{Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
 (a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
-Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
-
-# change all of these to gravitas function
-library(gravitas)
-library(tsibbledata)
-library(dplyr)
-library(ggplot2)
-
-par(mfrow = c(3, 2))
-
-VIC <- vic_elec
-
-VIC %>% 
-  create_gran("month_year") %>% 
-  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>% 
-  prob_plot("month_year","day_week",
-            response = "Demand",
-            plot_type = "lv") + 
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(a) Letter value plot")
-
-VIC %>% 
-  create_gran("day_week") %>% 
-  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>% 
-  prob_plot("day_week","hour_day",
-            response = "Demand",
-            plot_type = "quantile",
-            symmetric = FALSE) + 
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(b) Decile plot")
-
-VIC %>%
-  create_gran("day_month") %>% 
-  filter(day_month %in% c(1, 15, 29, 31)) %>%
-  prob_plot("day_month",
-                "day_year",
-                response = "Demand",
-                plot_type = "boxplot")+
-  ylab("Electricity Demand [KWh]") +
-  scale_x_discrete(breaks = seq(0, 366, 60)) + 
-  theme(legend.position = "bottom", 
-        strip.text = element_text(size = 7, margin = margin())) +
-  ggtitle("(c) Box plot")
-
-VIC %>%
-    create_gran("week_month") %>% 
-  filter(week_month %in% c(1, 2, 4)) %>%
-   prob_plot("week_month",
-                "day_month",
-                response = "Demand",
-                plot_type = "violin", alpha = 0.3) + ylab("") + 
-  xlab("Days of the Month") + 
-  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) + 
-  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
-  ggtitle("(d) Violin plot")
-
-VIC %>%  create_gran("week_month") %>%
-  create_gran("week_year") %>%
-  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>% 
-  prob_plot("week_year", 
-            "week_month",
-            response = "Demand",
-            plot_type = "ridge")  + 
-  xlab("Electricity Demand [KWh]") +
-  ylab("Weeks of the Year") +
-  scale_x_continuous(breaks = seq(2000, 10000, 3000)) + 
-  theme(legend.position = "bottom", 
-        strip.text = element_text(size = 7, margin = margin())) +
-  ggtitle("(e) Ridge plot")
-
-VIC %>%
-  create_gran("month_year") %>% 
-  filter(month_year %in% c("Jan", "Jul", "Nov")) %>% 
-   prob_plot("month_year", 
-            "day_year",
-            response = "Demand",
-            plot_type = "quantile",
-            quantile_prob = seq(0.1, 0.9, 0.1),
-            symmetric = FALSE) +
-   #scale_x_discrete(breaks = seq(1, 336, 60)) + 
-   ylab("") + 
-   xlab("Day of the Year") + theme(legend.position = "none", 
-                                   strip.text = element_text(size = 7, 
-                                                             margin = margin())) +
-  ggtitle("(f) Decile plot")
-
-# VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
-```
-
-### Interchangeability of mappings
+Only plots (a) and (b) show harmonised time variables.}\label{fig:allFig}
+\end{figure}
+
+\hypertarget{interchangeability-of-mappings}{%
+\subsubsection{Interchangeability of mappings}\label{interchangeability-of-mappings}}
 
 We will discuss the effect of the mapping of cyclic granularities in this section.
-When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
+When we consider data sets of the form \textless{}\(C_1\), \(C_2\), \(v\)\textgreater{} with \(C_1\) mapped to \(x\) position and \(C_2\) to facets \(A_i\)'s are placed in close proximity and each \(B_j\) represent a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case $A_i$'s are compared against each other within each group.
-When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$.
+Hence, in this case \(A_i\)'s are compared against each other within each group.
+When \(C_2\) is mapped to facets and \(C_1\) is mapped to \(x\) axis, comparisons would ideally happen across all \(B_j\) within each \(A_i\).
 Although the variables used in the layers are the same, the difference in mapping leads to emphasis of different behavior of the variables.
 
-## Number of observations and statistical transformations
+\hypertarget{number-of-observations-and-statistical-transformations}{%
+\subsection{Number of observations and statistical transformations}\label{number-of-observations-and-statistical-transformations}}
 
@@
 
-### Rarely occuring events
-
-Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
-Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
-There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
-The value of $k$ will depend on the statistical transformation we are producing.
-For computing deciles, $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
+\hypertarget{rarely-occuring-events}{%
+\subsubsection{Rarely occuring events}\label{rarely-occuring-events}}
+
+Suppose we have \(T\) observations, and two cyclic granularities \(C_1\) with \(n\) categories and \(C_2\) with \(m\) categories.
+Each element of \(C_1\) occurs approximately \(T/n\) times while each element of \(C_2\) occurs approximately \(T/m\) times.
+There are no empty combinations, and each combination will occur on average an equal number of times as \(T\rightarrow\infty\), so the average number of observations per combination is \(T/(mn)\).
+If we require at least \(k\) observations to create a meaningful panel, then provided \(T \ge mnk\), the visualization will be acceptable.
+The value of \(k\) will depend on the statistical transformation we are producing.
+For computing deciles, \(k=10\) may be acceptable, but for density estimates, we would need \(k \ge 30\).
 Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
 
-<!-- The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations. -->
-
-### Unevenly distributed events
+\hypertarget{unevenly-distributed-events}{%
+\subsubsection{Unevenly distributed events}\label{unevenly-distributed-events}}
 
@@
 In such cases, the statistical transformations based on density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators.
-\autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.
-
-<!-- the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$,  -->
-
-<!-- Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels. -->
-
-# Applications {#sec:application}
-
-## Smart meter data of Australia {#sec:smartmeter}
-
-<!-- # how data looks -->
+\autoref{eq:MSE} shows that the variance of both \(\hat{f(x)}\) and \(\hat{q(u)}\) is a function of the sample size \(n\).
+
+\hypertarget{sec:application}{%
+\section{Applications}\label{sec:application}}
+
+\hypertarget{sec:smartmeter}{%
+\subsection{Smart meter data of Australia}\label{sec:smartmeter}}
+
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
+One of the customer trial \citep{smart-meter} conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
 It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
 The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns.
-<!-- and identify the extreme and regular households. -->
-
-### Data structure for visualization:
+
+\hypertarget{data-structure-for-visualization}{%
+\subsubsection{Data structure for visualization:}\label{data-structure-for-visualization}}
 
 The data structure for a sample of 50 households from that trial is as follows:
-<!-- We take a random sample of 50 households to see how we can achieve that.  -->
-
-```{r smart-meter, fig.cap = "Smart meter data structure"}
-sm <- 
-  read_rds("data/sm_cust50.rds")
-
-smart_meter <- sm %>% as_tsibble() %>% dplyr::select(customer_id, 
-                              reading_datetime, 
-                              general_supply_kwh)
-smart_meter
-```
-
-The variable `reading_datetime` is the index of the tsibble and represents the time variable.
-The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed.
-Data sets of the form <$C1$, $C2$, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity $C1$ would be mapped to $x$ axis and `general_supply_kwh` to $y$ axis with the cyclic granularity $C2$ mapped to facet.
-Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
+
+\begin{verbatim}
+#> # A tsibble: 1,450,232 x 3 [30m] <UTC>
+#> # Key:       customer_id [50]
+#>    customer_id reading_datetime    general_supply_kwh
+#>    <chr>       <dttm>                           <dbl>
+#>  1 10006414    2012-02-10 08:00:00              0.141
+#>  2 10006414    2012-02-10 08:30:00              0.088
+#>  3 10006414    2012-02-10 09:00:00              0.078
+#>  4 10006414    2012-02-10 09:30:00              0.151
+#>  5 10006414    2012-02-10 10:00:00              0.146
+#>  6 10006414    2012-02-10 10:30:00              0.077
+#>  7 10006414    2012-02-10 11:00:00              0.052
+#>  8 10006414    2012-02-10 11:30:00              0.055
+#>  9 10006414    2012-02-10 12:00:00              0.055
+#> 10 10006414    2012-02-10 12:30:00              0.252
+#> # … with 1,450,222 more rows
+\end{verbatim}
+
+The variable \texttt{reading\_datetime} is the index of the tsibble and represents the time variable.
+The variable \texttt{customer\_id} denoting different households represent the key of the tsibble and \texttt{general\_supply\_kwh} is the time series variable which needs to be analyzed.
+Data sets of the form \textless{}\(C1\), \(C2\), \texttt{general\_supply\_kwh}\textgreater{} will form the basis of exploration for each key in the data set, where the cyclic granularity \(C1\) would be mapped to \(x\) axis and \texttt{general\_supply\_kwh} to \(y\) axis with the cyclic granularity \(C2\) mapped to facet.
+Any geometry displaying the probability distribution of \texttt{general\_supply\_kwh} like boxplot, letter value, violin,
 ridge or highest density region plots can then be used to explore the data set.
 
-### Cyclic granularities search and computation:
-
-R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here.
+\hypertarget{cyclic-granularities-search-and-computation}{%
+\subsubsection{Cyclic granularities search and computation:}\label{cyclic-granularities-search-and-computation}}
+
+R package \textbf{gravitas} \citep{R-gravitas} is used to facilitate the systematic exploration here.
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-
-<!-- how gravitas is used to deal with systematic exploration -->
 
@@
 
-```{r search, echo=TRUE}
-library(gravitas)
-smart_meter %>% search_gran()
-```
-
-<!-- #smart_smart_search <- smart_meter %>% search_gran() -->
-<!-- #knitr::kable(smart_smart_search, format = "markdown") -->
-
-If these options are considered too many, the most coarse temporal unit can be set to a “month”.
-
-```{r search_limit1, echo=TRUE}
-smart_meter10 %>% 
-  search_gran(highest_unit = "month")
-```
-
-<!-- # smart_meter_limit1 <- smart_meter10 %>%  -->
-<!-- #   search_gran(highest_unit = "month") -->
-<!-- # knitr::kable(smart_meter_limit1, format = "markdown") -->
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(gravitas)}
+\NormalTok{smart_meter }\OperatorTok{%>%}\StringTok{ }\KeywordTok{search_gran}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#>  [1] "hhour_hour"         "hhour_day"          "hhour_week"        
+#>  [4] "hhour_fortnight"    "hhour_month"        "hhour_quarter"     
+#>  [7] "hhour_semester"     "hhour_year"         "hour_day"          
+#> [10] "hour_week"          "hour_fortnight"     "hour_month"        
+#> [13] "hour_quarter"       "hour_semester"      "hour_year"         
+#> [16] "day_week"           "day_fortnight"      "day_month"         
+#> [19] "day_quarter"        "day_semester"       "day_year"          
+#> [22] "week_fortnight"     "week_month"         "week_quarter"      
+#> [25] "week_semester"      "week_year"          "fortnight_month"   
+#> [28] "fortnight_quarter"  "fortnight_semester" "fortnight_year"    
+#> [31] "month_quarter"      "month_semester"     "month_year"        
+#> [34] "quarter_semester"   "quarter_year"       "semester_year"
+\end{verbatim}
+
+If these options are considered too many, the most coarse temporal unit can be set to a ``month''.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{search_gran}\NormalTok{(}\DataTypeTok{highest_unit =} \StringTok{"month"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#>  [1] "hhour_hour"      "hhour_day"       "hhour_week"      "hhour_fortnight"
+#>  [5] "hhour_month"     "hour_day"        "hour_week"       "hour_fortnight" 
+#>  [9] "hour_month"      "day_week"        "day_fortnight"   "day_month"      
+#> [13] "week_fortnight"  "week_month"      "fortnight_month"
+\end{verbatim}
 
@@
 
-```{r search_gran_limit2, echo=TRUE}
-smart_meter10 %>% search_gran(highest_unit = "month",
-                              filter_out = c("hhour", "fortnight")
-                              )
-```
-
-<!-- # smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month", -->
-<!-- #                               filter_out = c("hhour", "fortnight") -->
-<!-- # ) -->
-<!-- # knitr::kable(smart_meter_limit2, format = "markdown") -->
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }\KeywordTok{search_gran}\NormalTok{(}\DataTypeTok{highest_unit =} \StringTok{"month"}\NormalTok{,}
+                              \DataTypeTok{filter_out =} \KeywordTok{c}\NormalTok{(}\StringTok{"hhour"}\NormalTok{, }\StringTok{"fortnight"}\NormalTok{)}
+\NormalTok{                              )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#> [1] "hour_day"   "hour_week"  "hour_month" "day_week"   "day_month" 
+#> [6] "week_month"
+\end{verbatim}
 
@@
 
-### Screening and visualising harmonies:
+\hypertarget{screening-and-visualising-harmonies}{%
+\subsubsection{Screening and visualising harmonies:}\label{screening-and-visualising-harmonies}}
 
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-Given the data structure <$C1$ $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
+Given the data structure \textless{}\(C1\) \(C2\), \texttt{general\_supply\_kwh}\textgreater{}, each of those six cyclic granularities can either be mapped to x-axis or to facet.
 Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualisation.
@@
 
-<!-- We will look at few harmony pairs and see how the proposed workflow can be utilized.  -->
-
-<!-- From the Gestalt theory with the same data  <$C1$ $C2$, `general_supply_kwh`>, mapping $C1$ on the x-axis vs facet might lead to different insights. -->
-
-```{r harmony, echo=TRUE}
-smart_meter10 %>% harmony(
-  ugran = "month",
-  filter_out = c("hhour", "fortnight")
-)
-```
-
-<!-- # smart_meter10 %>% gran_advice("wknd_wday", "hour_day") -->
-
-<!-- The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.  -->
-
-<!-- No need to share the same for 50 customers - cut down text and anyway nothing much interesting is coming -->
-
-<!-- ```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
-The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."} -->
-
-<!-- smart_meter <-  -->
-<!--   read_rds("data/sm_cust50.rds") -->
-
-<!-- smart_meter %>%  -->
-<!--   prob_plot("wknd_wday", -->
-<!--                             "hour_day", -->
-<!--                             response = "general_supply_kwh", -->
-<!--                             plot_type = "quantile", -->
-<!--             symmetric = FALSE,  -->
-<!--             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
-<!--   scale_y_sqrt() -->
-<!-- ``` -->
-
-<!-- # ```{r wknd-wday50_df} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   prob_plot("month_year", -->
-<!-- #                             "day_fortnight", -->
-<!-- #                             response = "general_supply_kwh", -->
-<!-- #                             plot_type = "quantile", -->
-<!-- #             symmetric = TRUE,  -->
-<!-- #             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
-<!-- #   scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3)) -->
-<!-- # ``` -->
-
-<!-- ```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
-"} -->
-
-<!-- library(gravitas) -->
-<!-- library(tsibble) -->
-
-<!-- library(ggplot2) -->
-<!-- smart_meter %>% -->
-<!--   prob_plot("month_year", "hour_day", -->
-<!--     response = "general_supply_kwh", -->
-<!--     plot_type = "quantile", -->
-<!--     quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99), -->
-<!--     symmetric = TRUE) + -->
-<!--   ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3)) -->
-<!-- ``` -->
-
-<!-- Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
-This is true for both top 50% and 75% of the customers.
-Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
-All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
--->
-
-<!-- ```{r dw-hd} -->
-
-<!-- # Hourly usage across days of the week -->
-<!-- sm_cust50 %>% -->
-<!--   prob_plot("day_week",  -->
-<!--            "hour_day", -->
-<!--            "general_supply_kwh",  -->
-<!--            plot_type = "quantile",  -->
-<!--           quantile_prob = c(0.25, 0.5, 0.75), -->
-<!--            response = "general_supply_kwh", overlay = FALSE) + -->
-<!--   ggtitle("") -->
-
-<!-- ``` -->
-
-<!-- on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.  -->
-
-<!-- typical and extreme behaviors of 50 households from that trial.  -->
-
-<!-- , and indeed many parts of the world.
-Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
--->
-
-<!-- <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions --> 
-<!-- It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
--->
-
-<!-- We want to  look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities. -->
-
-<!-- In Figure \ref{fig:day-fortnight} and \ref{fig:day-fortnight-agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
--->
-
-<!-- can be added later -->
-
-<!-- In Figure \ref{fig:day-fortnight},to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
-it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
-However, there is additional information that we can derive looking at the distribution.
-If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
-However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
-This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
--->
-
-<!-- # ```{r day-fortnight_agg, echo = FALSE} -->
-<!-- # library(lvplot) -->
-<!-- # library(ggplot2) -->
-<!-- # library(dplyr) -->
-<!-- # library(tibble) -->
-<!-- # library(gravitas) -->
-<!-- # library(tidyr) -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!-- #   create_gran("day_fortnight") %>% -->
-<!-- #   as_tibble() %>%  -->
-<!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarise(median= log(median(general_supply_kwh)), -->
-<!-- #             total = log(sum(general_supply_kwh))) %>%  -->
-<!-- #   gather("daily_consumption", "value", - c(customer_id, day_fortnight)) %>%  -->
-<!-- #   ggplot2::ggplot(aes( -->
-<!-- #     x = as.factor(day_fortnight), -->
-<!-- #     y = value,  -->
-<!-- #     color = daily_consumption, -->
-<!-- #     group = daily_consumption)) + -->
-<!-- #   geom_line() + -->
-<!-- #   xlab("day_fortnight") + -->
-<!-- #   ylab("general_supply_kwh") +  -->
-<!-- #   facet_wrap(~customer_id) +  -->
-<!-- #   theme_bw()  -->
-
-<!-- # smart_meter10 %>%  -->
-<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!-- #   create_gran("day_fortnight") %>% -->
-<!-- #   as_tibble() %>%  -->
-<!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarise(median_consumption= median(general_supply_kwh)) %>% -->
-<!-- #   ggplot2::ggplot(aes( -->
-<!-- #     x = as.factor(day_fortnight), -->
-<!-- #     y = median_consumption, -->
-<!-- #     color =  customer_id, -->
-<!-- #     group = customer_id)) + -->
-<!-- #   geom_line() + -->
-<!-- #   xlab("day_fortnight") + -->
-<!-- #   ylab("general_supply_kwh") +  -->
-<!-- #   theme_bw()  -->
-<!-- ``` -->
-
-<!-- ```{r day-fortnight, fig.cap="Letter value plot of two households across days of the fortnight.
-M, F, E, D and C represents the letter values of energy consumption.
-Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household."} -->
-<!-- library(lvplot) -->
-<!-- library(ggplot2) -->
-<!-- library(dplyr) -->
-<!-- library(tibble) -->
-<!-- library(gravitas) -->
-<!-- smart_meter10 %>% -->
-<!--   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!--   create_gran("day_fortnight") %>% -->
-<!--   ggplot2::ggplot(aes( -->
-<!--     x = as.factor(day_fortnight), -->
-<!--     y = general_supply_kwh)) + -->
-<!--   xlab("day_fortnight") + -->
-<!--   geom_lv( -->
-<!--     outlier.colour = "red", -->
-<!--     aes(fill = ..LV..), -->
-<!--     k = 5) + -->
-<!--   facet_wrap(~customer_id) +  -->
-<!--   scale_fill_lv() + -->
-<!--   theme_bw() + -->
-<!--   scale_y_sqrt()  -->
-<!-- ``` -->
-
-<!-- Figure \autoref{fig:my-hd} shows the overlay quantile plot of energy consumption of 50 households.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-Each facet represents a month and energy consumption across each hours of the day is shown inside each facet.
-It can be observed that the median is very close (and hence not distinctly visible) to the lower boundaries of all the other bands implying energy consumption for these households are extremely left skewed.75% of the households has distinct higher evening peaks compared to morning peaks, especially in the winter months.
-The top 5% households behave very differently than even the top 10% households in terms of total energy consumption across hours of the day.
--->
-
-<!-- # ```{r is_harmony, echo= TRUE} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   is_harmony(gran1 = "hour_day",  -->
-<!-- #              gran2 = "day_week") -->
-<!-- #  -->
-<!-- # smart_meter10 %>% -->
-<!-- #   is_harmony(gran1 = "hour_day",  -->
-<!-- #              gran2 = "day_week",  -->
-<!-- #              facet_h = 14) -->
-<!-- #  -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   is_harmony(gran1 = "day_month", -->
-<!-- #              gran2 = "week_month") -->
-<!-- # ``` -->
-
-<!-- Boxplot of energy consumption is shown across `wknd_wday` (facet) and `hour_day` (x-axis) for the same two households.
-For the first household, the shape of the distribution does not change significantly for weekdays and weekends.
-The distribution is roughly smooth with a morning and an evening peak.
--->
-
-In \autoref{fig:bothcust}, the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot. 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }\KeywordTok{harmony}\NormalTok{(}
+  \DataTypeTok{ugran =} \StringTok{"month"}\NormalTok{,}
+  \DataTypeTok{filter_out =} \KeywordTok{c}\NormalTok{(}\StringTok{"hhour"}\NormalTok{, }\StringTok{"fortnight"}\NormalTok{)}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#> # A tibble: 13 x 4
+#>    facet_variable x_variable facet_levels x_levels
+#>    <chr>          <chr>             <int>    <int>
+#>  1 day_week       hour_day              7       24
+#>  2 day_month      hour_day             31       24
+#>  3 week_month     hour_day              5       24
+#>  4 day_month      hour_week            31      168
+#>  5 week_month     hour_week             5      168
+#>  6 day_week       hour_month            7      744
+#>  7 hour_day       day_week             24        7
+#>  8 day_month      day_week             31        7
+#>  9 week_month     day_week              5        7
+#> 10 hour_day       day_month            24       31
+#> 11 day_week       day_month             7       31
+#> 12 hour_day       week_month           24        5
+#> 13 day_week       week_month            7        5
+\end{verbatim}
+
+In \autoref{fig:bothcust}, the distribution of energy consumption is plotted across the harmony pair (wknd\_wday, hour\_day) through an area quantile plot.
 The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
@@
 
-<!-- If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
-However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
-This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
--->
-
-```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE}
-smart_meter10 %>% 
-  filter(customer_id %in% c(10017936))%>% 
-  gran_advice("wknd_wday",
-                            "hour_day")
-```
-
-```{r bothcust, fig.cap = "Area quantile plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936 displays the smoothed energy consumption curve for the entire day.
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{filter}\NormalTok{(customer_id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{10017936}\NormalTok{))}\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{gran_advice}\NormalTok{(}\StringTok{"wknd_wday"}\NormalTok{,}
+                            \StringTok{"hour_day"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#> The chosen granularities are harmonies 
+#>  
+#> Recommended plots are: violin lv quantile boxplot 
+#>  
+#> Number of observations are homogenous across facets 
+#>  
+#> Number of observations are homogenous within facets 
+#>  
+#> Cross tabulation of granularities : 
+#>  
+#> # A tibble: 24 x 3
+#>    hour_day Weekday Weekend
+#>    <fct>      <dbl>   <dbl>
+#>  1 0            910     366
+#>  2 1            908     366
+#>  3 2            909     366
+#>  4 3            910     366
+#>  5 4            910     366
+#>  6 5            910     366
+#>  7 6            909     366
+#>  8 7            908     366
+#>  9 8            908     366
+#> 10 9            908     366
+#> # … with 14 more rows
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/bothcust-1}
+
+}
+
+\caption{Area quantile plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936 displays the smoothed energy consumption curve for the entire day.
 The customer typically consumes more in the morning hours following which their consumption is low till late afternoon hours.
-Median consumption for the early morning hours starts higher and continue to remain high for longer during weekdays compared to weekdend."}
-
-# plot for only customer is shown
-# cust1 <- smart_meter10 %>% 
-#filter(customer_id %in% c(10006704)) %>% 
-#   prob_plot("wknd_wday",
-#                             "hour_day",
-#                             response = "general_supply_kwh",
-#                             plot_type = "quantile",
-#             symmetric = FALSE,
-#             quantile_prob = c(0.25, 0.5, 0.75, 0.9)) 
-#   scale_y_sqrt() +
-#   ggtitle("Energy consumption distribution for customer id: 10006704")
-cust2 <- smart_meter10 %>% 
-  filter(customer_id %in% c(10017936))%>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "quantile",
-            symmetric = TRUE,
-             quantile_prob = c(0.01,0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
-  scale_y_sqrt() 
-cust2
-```
-
-```{r bothcust-lv, fig.cap ="Violin plots of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
+Median consumption for the early morning hours starts higher and continue to remain high for longer during weekdays compared to weekdend.}\label{fig:bothcust}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/bothcust-lv-1}
+
+}
+
+\caption{Violin plots of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
 Early morning hours are bimodal implying both high and low consumptions are probable.
 Bimodality can be caused by difference in behavior across different seasons.
-Volatility is highest from 7 to 13 hours as evident from the narrow violins."}
-# 
-# smart_meter10 %>% 
-#   filter(customer_id %in% c(10006704)) %>% 
-#   prob_plot("wknd_wday",
-#                             "hour_day",
-#                             response = "general_supply_kwh",
-#                             plot_type = "violin") +
-#   scale_y_sqrt() +
-#   ggtitle("Energy consumption distribution for customer id: 10006704")
-
-smart_meter10 %>% 
-  filter(customer_id %in% c(10017936))%>% 
-  prob_plot("wknd_wday",
-                            "hour_day",
-                            response = "general_supply_kwh",
-                            plot_type = "violin") +
-  scale_y_sqrt() 
-```
-
-```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
+Volatility is highest from 7 to 13 hours as evident from the narrow violins.}\label{fig:bothcust-lv}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/cust-ridge-1}
+
+}
+
+\caption{Boxplot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
 The change in mapping of the cyclic granularities leads to easier comparison within weekdays and weekends for each hour.
 Median consumption till 6am is higher for weekends compared to weekdays although the quartiles look similar.
-Outliers appear more during the end of the day implying more uncertain behavior in latter hours of the day compared to early morning hours."}
-# 
-# smart_meter10 %>% 
-#   filter(customer_id %in% c(10006704)) %>% 
-#   prob_plot("wknd_wday",
-#                             "hour_day",
-#                             response = "general_supply_kwh",
-#                             plot_type = "violin") +
-#   scale_y_sqrt() +
-#   ggtitle("Energy consumption distribution for customer id: 10006704")
-
-smart_meter10 %>% 
-  filter(customer_id %in% c(10017936))%>% 
-  prob_plot("hour_day", 
-            "wknd_wday",
-            response = "general_supply_kwh",
-            plot_type = "boxplot") +
-            scale_y_sqrt()
-            # scale_fill_brewer(palette = "Dark2")
-
-```
-
-<!-- Now, we would like to see how these customers' behavior relate to the rest of the  50 households across these two measures -   -->
-
-<!-- - if energy distribution is skewed towards extreme? -->
-
-<!-- - if the weekend and weekday behavior are different for most of these households? -->
-
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
-This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+Outliers appear more during the end of the day implying more uncertain behavior in latter hours of the day compared to early morning hours.}\label{fig:cust-ridge}
+\end{figure}
+
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one \texttt{customer\_id} shows a particular behavior, but this behavior is reversed when all keys are combined together.
+This is also intuitive because \texttt{customer\_id}'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
 A strategy for analyzing multiple keys together could be to first screen some basis time series or demographic features and then look at their energy behavior.
@@
 
-<!-- For the purpose of this paper, the smart meter data is filtered for two customers to illustrate kinds of insights that can be drawn after proceeding with the systematic exploration. -->
-
-<!-- For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households.
-Quantiles for weekends and weekdays for these customers look exactly the same.
-It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors.
-Consider another example in Figure \ref{fig:my-hd} showing the area quantile plots of 50 customers.
-The median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-The pink band changes significantly in winter months (May - August) implying that much of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of all quantiles increased too in winter months (increased energy usage in weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
--->
-
 This case study shows systematic exploration of energy behavior for a random household to gain some insights on periodic behavior of the households.
@@
 
-## T20 cricket data of Indian Premiere League {#sec:cricket}
+\hypertarget{sec:cricket}{%
+\subsection{T20 cricket data of Indian Premiere League}\label{sec:cricket}}
 
@@
 The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
-The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
+The ball by ball data for IPL season 2008 to 2016 is fetched from \href{https://www.kaggle.com/josephgpinto/ipl-data-analysis/data}{Kaggle}.
+The \texttt{cricket} data set in the \texttt{gravitas} package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
 The hierarchy table would look like the following:
 
-```{r hierarchy2}
- library(gravitas)
- library(tibble)
- hierarchy_model <- tibble::tibble(
-   G = c("over", "inning", "match", "season"),
-   C = c("over-of-inning", "inning-of-match", "match-of-season",1),
-   P = c(20, 2, "aperiodic", 1))
- knitr::kable(hierarchy_model, format = "markdown")
-```
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+G & C & P\tabularnewline
+\midrule
+\endhead
+over & over-of-inning & 20\tabularnewline
+inning & inning-of-match & 2\tabularnewline
+match & match-of-season & aperiodic\tabularnewline
+season & 1 & 1\tabularnewline
+\bottomrule
+\end{longtable}
 
 There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
-
-<!-- the two teams have a single innings each, which is restricted to a maximum of 20 overs.
-Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls with some exceptions.
--->
-
-<!-- ```{r hierarchy, echo=FALSE} -->
-<!-- hierarchy_model <- tibble::tibble(linear gran = c("ball", "over", "inning", "match"),  -->
-<!--                                   convert_fct = c(6, 20, 2, 1)) -->
-<!-- knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket") -->
-<!-- ``` -->
-
-<!-- Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
--->
-
-<!-- Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that. -->
-
-<!-- a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions? -->
-
-<!-- How the scores per over vary across granularities/categorizations like innings of a match or matches of a season?
-Are these different for the winning teams in that season and the teams that couldn't qualify for the playoffs?
--->
-
-<!-- Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings? -->
-
-<!-- We will look at the ball by ball data for all batting teams.
-Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
-Also, we look at runs per over as that would have more variability compared to runs per ball and it be easier to observe the strategies of the winning team through that.
--->
-
-<!-- Making it short -->
-
-<!-- First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperidic cyclic granularity) in \autoref{fig:seas-over-inning}.
-The distribution of runs per over has not significantly changed from 2008 to 2016.
-There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
-Hence, we work with subsets of seasons to answer some of the questions: -->
 
@@
 It would be interesting to take their example in order to dive deeper into this question.
-Circular cyclic granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
+Circular cyclic granularities ``over-of-inning'' and ``inning-of-match'' can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
 From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
@@
 
-<!-- Also longer and more distinct letter values in the second innings suggests that the variability is more in the second innings. -->
-<!-- - Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first? -->
-
-<!-- - Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings? -->
-
-```{r runs_per-over}
- library(tsibble)
-
- cricket_all <-  read_csv("data-raw/deliveries_all.csv")
- matches_all <- read_csv("data-raw/matches_all.csv")
-
- cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
-
- cricket_per_over <- cricket_season %>%
-   group_by(season,
-            match_id,
-            batting_team,
-            bowling_team,
-            inning,
-            over) %>%
-   summarise(runs_per_over = sum(total_runs),
-             run_rate = sum(total_runs)/length(total_runs))
-
- cricket_tsibble_all <- cricket_per_over %>%
-   ungroup() %>%
-   mutate(data_index = row_number()) %>%
-   as_tsibble(index = data_index)
-```
-
-<!-- # ```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."} -->
-<!-- #   cricket_tsibble_all %>%  -->
-<!-- #   prob_plot("over",  -->
-<!-- #             "season", -->
-<!-- #             hierarchy_model, -->
-<!-- #             response = "runs_per_over", -->
-<!-- #             plot_type = "quantile", -->
-<!-- #             quantile_prob = c(0.25, 0.5, 0.75),  -->
-<!-- #             symmetric = FALSE)  + -->
-<!-- #      theme( # remove the vertical grid lines -->
-<!-- #             panel.grid.major.x = element_blank() , -->
-<!-- #             # explicitly set the horizontal lines (or they will disappear too) -->
-<!-- #             panel.grid.major.y = element_blank()) + scale_x_discrete(breaks = seq(2008, 2016, 3)) -->
-<!-- # ``` -->
-
-```{r cricex, fig.cap= "Letter value plot of runs per over across overs of the inning faceted by innings of the match.
-No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning.", warning=FALSE, message = FALSE}
-
-library(tsibble)
-cricket_tsibble <- cricket %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-hierarchy_model <- tibble::tibble(
-  units = c("index", "over", "inning", "match"),
-  convert_fct = c(1, 20, 2, 1)
-)
-
-# cricket_tsibble %>%
-#   filter(batting_team %in% c("Mumbai Indians",
-#                              "Chennai Super Kings"))%>%
-#   create_gran("inning_match", hierarchy_model) %>% create_gran("over_inning", hierarchy_model)
-
-cricket_tsibble %>%
-  filter(batting_team %in% c("Mumbai Indians",
-                             "Chennai Super Kings"))%>%
-  gran_advice("inning", 
-            "over")
-
-cricket_tsibble %>%
-  filter(batting_team %in% c("Mumbai Indians",
-                             "Chennai Super Kings"))%>%
-  prob_plot("inning", 
-            "over",
-  response = "runs_per_over",
-  hierarchy_model,
-  plot_type = "lv")  + ggtitle("") +
-  scale_fill_brewer(palette = "Dark2")
-
-# cricket_tsibble %>%
-#   filter(batting_team %in% c("Mumbai Indians",
-#                              "Chennai Super Kings"))%>%
-#   gran_advice("inning", 
-#             "over")
-
-```
+\begin{verbatim}
+#> The chosen granularities are harmonies 
+#>  
+#> Recommended plots are: violin lv quantile boxplot 
+#>  
+#> Number of observations are homogenous across facets 
+#>  
+#> Number of observations are homogenous within facets 
+#>  
+#> Cross tabulation of granularities : 
+#>  
+#> # A tibble: 20 x 3
+#>    over    `1`   `2`
+#>    <fct> <dbl> <dbl>
+#>  1 1        65    44
+#>  2 2        65    44
+#>  3 3        65    44
+#>  4 4        65    44
+#>  5 5        65    44
+#>  6 6        65    44
+#>  7 7        65    44
+#>  8 8        65    44
+#>  9 9        65    44
+#> 10 10       65    44
+#> 11 11       65    44
+#> 12 12       65    44
+#> 13 13       65    44
+#> 14 14       65    44
+#> 15 15       65    44
+#> 16 16       65    44
+#> 17 17       65    44
+#> 18 18       65    44
+#> 19 19       65    44
+#> 20 20       65    44
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/cricex-1}
+
+}
+
+\caption{Letter value plot of runs per over across overs of the inning faceted by innings of the match.
+No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning.}\label{fig:cricex}
+\end{figure}
 
@@
 
-```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+\begin{verbatim}
+#> # A tibble: 4 x 21
+#>   fielding_wckts   `1`   `2`   `3`   `4`   `5`   `6`   `7`   `8`   `9`  `10`
+#>   <fct>          <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+#> 1 0               1039   997   972   978   962   957  1002  1012   993   984
+#> 2 1                120   145   171   165   180   181   138   124   140   153
+#> 3 2                  5    10     8     8     9    10     6    10     8     4
+#> 4 3                  0     0     0     0     0     0     0     0     0     0
+#> # … with 10 more variables: `11` <dbl>, `12` <dbl>, `13` <dbl>, `14` <dbl>,
+#> #   `15` <dbl>, `16` <dbl>, `17` <dbl>, `18` <dbl>, `19` <dbl>, `20` <dbl>
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figure/field-1}
+
+}
+
+\caption{Distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
 This shows good fielding in previous over leads to lower run rate in the subsequent over.
-Also run rates are much less volatile in case there are at least one wicket in the previous over.", warning=FALSE}
-
-cricket_dot_field <- cricket_season %>% 
-  mutate(fielding_proxy = if_else(dismissal_kind %in% 
-                                    c("caught", "caught and bowled"),1,0),
-         dot_ball_proxy = if_else(total_runs==0,1,0),
-         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
-  group_by(season,
-           match_id,
-           batting_team,
-           bowling_team, 
-           inning,
-           over) %>%
-  summarise(runs_per_over = sum(total_runs),
-            run_rate = sum(total_runs)/length(total_runs),
-            fielding_wckts = sum(fielding_proxy),
-            dot_balls = sum(dot_ball_proxy)) %>%
-            mutate(diff_run_rate = c(0,diff(run_rate)))
-
-cricket_tsibble <- cricket_dot_field %>% 
-  ungroup() %>% 
-  mutate(data_index = row_number()) %>% 
-  as_tsibble(index = data_index)
-
-cricket_tsibble %>% gran_obs("over", "fielding_wckts")
-
-cricket_data_ts <- cricket_tsibble %>% 
-  mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets"),
-         dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls"))
-
-  #filter(fielding_wckts %in% c(0,1)) %>% 
-  cricket_data_ts %>% prob_plot( "over","field",
-                               hierarchy_model,
-                               response = "diff_run_rate",
-                               plot_type = "violin",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-           overlay = TRUE)  + ggtitle("")
-```
-
-<!-- Q2: Among good fielding and bowling - which affect the runs of the subsequent overs more? -->
+Also run rates are much less volatile in case there are at least one wicket in the previous over.}\label{fig:field}
+\end{figure}
 
@@
 Each of the quantiles decrease across facets.
-With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for different quantiles.
-
-<!-- # ```{r crickread, ech0=FALSE} -->
-<!-- # cricket_tsibble <- gravitas::cricket %>% -->
-<!-- #   mutate(data_index = row_number()) %>% -->
-<!-- #   as_tsibble(index = data_index) -->
-<!-- # ``` -->
-
-```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), more than zero (facet 2) dot balls per over.
-For all quantiles, difference in runs between subsequent overs decrease with at least one dot ball in the previous over.", fig.pos="ht"}
-
-# library(gravitas)
-# library(tsibble)
-# 
-# 
-# cricket_bowling <- cricket_season %>% 
-#   mutate(dot_ball_proxy = if_else(total_runs==0,1,0),
-#          wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% 
-#   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-#   summarise(runs_per_over = sum(total_runs),
-#             run_rate = sum(total_runs)/length(total_runs),
-#             dot_balls = sum(dot_ball_proxy)) %>% 
-#             mutate(diff_run_rate = c(0,diff(run_rate)))
-# 
-# 
-# cricket_tsibble_bow <- cricket_bowling %>% 
-#   ungroup() %>% 
-#   mutate(data_index = row_number()) %>% 
-#   as_tsibble(index = data_index)
-#
-
-# cricket_tsibble_bow %>%
-#   mutate(dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) %>% 
-#   #filter(dot_balls %in% c(0, 1, 2)) %>%
-#   prob_plot("dot",
-#             "over",
-#             hierarchy_model,
-#             response = "diff_run_rate",
-#             plot_type = "quantile",
-#             quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
-#             symmetric = FALSE)  + ggtitle("")
-
-cricket_data_ts %>% 
-  #filter(dot_balls %in% c(0, 1, 2)) %>%
-  prob_plot("dot",
-            "over",
-            hierarchy_model,
-            response = "diff_run_rate",
-            plot_type = "quantile",
-            quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
-            symmetric = FALSE)  + ggtitle("")
-```
+With two dot balls per over, at least 50\% of the times, run rates decrease in the subsequent over as observed through the negative values for different quantiles.
+
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=\textwidth]{figure/exdot-1}
+
+}
+
+\caption{10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), more than zero (facet 2) dot balls per over.
+For all quantiles, difference in runs between subsequent overs decrease with at least one dot ball in the previous over.}\label{fig:exdot}
+\end{figure}
 
@@
 
-# Discussion {#sec:discussion}
+\hypertarget{sec:discussion}{%
+\section{Discussion}\label{sec:discussion}}
 
@@
 The infrastructure of systematic exploration of observed variable across pair of cyclic time granularities has been implemented in the
-R package `gravitas`.
+R package \texttt{gravitas}.
 A missing piece of the gravitas package is to enable user-defined temporal calendars and compute multiple order-up quasi-circular granularities from the hierarchy table.
 Also, computation of cyclic aperiodic granularities would require computing aperiodic linear granularities first.
-A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
+A few R packages like \texttt{almanac} and \texttt{gs} provide functionality to create recurring events that are not periodic.
 These functionalities can be imported in the gravitas package to accommodate for aperiodic cyclic granularities.
 
-# Acknowledgements {-}
-
-The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
-Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
-The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
-The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article.
-
-# Bibliography
-
-```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr","rlang", "gravitas"), file = "rpkgs.bib")
-```
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+The authors would like to thank the cohort at \href{https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team}{NUMBATS}, Monash University for sharing their wisdom and experience of developing R packages, Dr.~Peter Toscas \href{https://data61.csiro.au/}{from Data61 CSIRO} for providing useful inputs on improving the analysis of smart meter application.
+The package \texttt{gravitas} was built under \href{https://summerofcode.withgoogle.com/archive/}{Google Summer of Code, 2019}.
+This article was created with knitr \citep{knitr} and R Markdown \citep{rmarkdown}.
+The project's Github repository \url{https://github.com/Sayani07/paper-gravitas} contains all materials required to reproduce this article.
+
+\bibliographystyle{agsm}
+\bibliography{bibliography.bib}
+
+\end{document}

### 2020-02-08 main.Rmd
consistent naming of granularities

@@
 \toprule
-minute-of-Hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
-minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
-hour-of-Day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
-hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
-day-of-Week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
+minute-of-hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
+minute-of-day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
+hour-of-day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
+hour-of-week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
+day-of-week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
 \bottomrule
@@
 
-Day-of-Month: & $Q_1$ && $P_1 = 31, 30, 29, 28$\\
-Hour-of-Month: & $Q_2$ && $P_2 = 24*31, 24*30, 24*29, 24*28$\\
-Day-of-Year: & $Q_3$ && $P_3 = 366, 365$\\
-Week-of-Month: & $Q_4$ && $P_4 = 5, 4$\\ 
+day-of-month: & $Q_1$ && $P_1 = 31, 30, 29, 28$\\
+hour-of-month: & $Q_2$ && $P_2 = 24*31, 24*30, 24*29, 24*28$\\
+day-of-year: & $Q_3$ && $P_3 = 366, 365$\\
+week-of-month: & $Q_4$ && $P_4 = 5, 4$\\ 
 \bottomrule
@@
 The levels have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
-Plot choices will also depend on the aesthetic mapping of the cyclic granularities.
 A potential approach can be to categorize the levels as very high/high/medium/low for each cyclic granularities and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
@@
 ```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
+(a) Letter value plot by day-of-month and month-of-year (b) Decile plot by hour-of-day and day-of-week (c) Box plot by day-of-year and day-of-month, (d) Violin plot of day-of-month and week-of-month, (e) Ridge plot by week-of-month and week-of-year, (f) Decile plot by day-of-year and month-of-year.
 Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}

### 2020-02-08 main.Rmd
caption and observations manipulate for graphs in cricket data

@@
 It would be interesting to take their example in order to dive deeper into this question.
-Circular cyclic granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
+Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
 From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
@@
 So we only include number of catches and run out in an over as a measure of good fielding.
-Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.
-We can verify this fact through Figure \ref{fig:field} which shows that the difference in run rates between two subsequent overs are negative when good fielding leads to at least one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
-
-```{r field, fig.cap = "Distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
+Figure \ref{fig:field} shows that at least one dismissal in an over leads to a lower median and less volatility in the distribution of difference in run rates.
+
+<!-- Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.  -->
+
+```{r field, fig.cap = "Violin plots with an overlaid boxplot showing distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
 This shows good fielding in previous over leads to lower run rate in the subsequent over.
@@
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-           overlay = TRUE)  + ggtitle("")
+           overlay = TRUE)  + ggtitle("") + geom_boxplot(width = 0.5)
 ```
@@
 Or they should play consistently and avoid scoring high?
-From figure \ref{fig:exdot} shows the quantile plot of runs across overs for none, one or two dot balls per over in the facets.
-Each of the quantiles decrease across facets.
+Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
+With at least one dot balls per over, the distribution of difference in run rates as observed through the quantiles plotted have shifted downwards.
+This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
+
+<!-- Each of the quantiles decrease across facets.
 With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for different quantiles.
+-->
 
@@
 
-In Q2 and Q3, we considered dots balls per over or catches per over as aperiodic cyclic granularities.
-They do not appear in the hierarchy since we can not compute these granularities as a function of circular or quasi-circular granularities.
-Also, their order in the hierarchy table is not clear as they are essentially obtained from some aperiodic linear granularity.
-These are similar to holidays in temporal data which also can not be placed at a fixed order in a hierarchy table in case they are not finer than or groups into other temporal units.
+In Q2 and Q3, dots balls per over or catches per over are considered as aperiodic cyclic granularities with dot balls or catches as aperiodic linear granularities.
+These aperiodic linear granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
+These are similar to holidays or special events in temporal data.
+While any special event that corresponds to a time domain can be treated as an aperiodic linear granularity in a temporal case, dot balls or wickets that corresponds to certain balls (index) could be treated as aperiodic events in cricket.
+
+<!-- For example, in the temporal case, aperiodic cyclic granularity "Easter" is computed based on which all time indices correspond to the event "Easter".
+Similarly, for aperiodic cyclic granularity "dot balls per over" is computed based on which all balls correspond to the event "dot balls".
+-->
 

### 2020-02-08 main.tex
cricket analysis graph captions update

@@
 \toprule
-minute-of-Hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
-minute-of-Day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$1440$\\
-hour-of-Day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
-hour-of-Week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$168$\\
-day-of-Week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
+minute-of-hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
+minute-of-day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$1440$\\
+hour-of-day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
+hour-of-week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$168$\\
+day-of-week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
 \bottomrule
@@
 \caption{Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY.
+(a) Letter value plot by day-of-month and month-of-year (b) Decile plot by hour-of-day and day-of-week (c) Box plot by day-of-year and day-of-month, (d) Violin plot of day-of-month and week-of-month, (e) Ridge plot by week-of-month and week-of-year, (f) Decile plot by day-of-year and month-of-year.
 Only plots (a) and (b) show harmonised time variables.}\label{fig:allFig}
@@
 It would be interesting to take their example in order to dive deeper into this question.
-Circular cyclic granularities ``over-of-inning'' and ``inning-of-match'' can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
+Circular granularities ``over-of-inning'' and ``inning-of-match'' can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
 From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
@@
 So we only include number of catches and run out in an over as a measure of good fielding.
-Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.
-We can verify this fact through Figure \ref{fig:field} which shows that the difference in run rates between two subsequent overs are negative when good fielding leads to at least one dismissal in an over, implying good fielding in one over has indeed an impact on runs scored in the subsequent overs.
+Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
+Figure \ref{fig:field} shows that at least one dismissal in an over leads to a lower median and less volatility in the distribution of difference in run rates.
 
@@
 
-\caption{Distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+\caption{Violin plots with an overlaid boxplot showing distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
 This shows good fielding in previous over leads to lower run rate in the subsequent over.
@@
 Or they should play consistently and avoid scoring high?
-From figure \ref{fig:exdot} shows the quantile plot of runs across overs for none, one or two dot balls per over in the facets.
-Each of the quantiles decrease across facets.
-With two dot balls per over, at least 50\% of the times, run rates decrease in the subsequent over as observed through the negative values for different quantiles.
+Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
+With at least one dot balls per over, the distribution of difference in run rates as observed through the quantiles plotted have shifted downwards.
+This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
 
@@
 
-In Q2 and Q3, we considered dots balls per over or catches per over as aperiodic cyclic granularities.
-They do not appear in the hierarchy since we can not compute these granularities as a function of circular or quasi-circular granularities.
-Also, their order in the hierarchy table is not clear as they are essentially obtained from some aperiodic linear granularity.
-These are similar to holidays in temporal data which also can not be placed at a fixed order in a hierarchy table in case they are not finer than or groups into other temporal units.
+In Q2 and Q3, dots balls per over or catches per over are considered as aperiodic cyclic granularities with dot balls or catches as aperiodic linear granularities.
+These aperiodic linear granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
+These are similar to holidays or special events in temporal data.
+While any special event that corresponds to a time domain can be treated as an aperiodic linear granularity in a temporal case, dot balls or wickets that corresponds to certain balls (index) could be treated as aperiodic events in cricket.
 

### 2020-02-08 main.Rmd
spell check

@@
 A potential approach can be to categorize the levels as very high/high/medium/low for each cyclic granularities and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
-Default values for these categorisations can be chosen based on levels of common temporal granularities like days-of-month, days-of-fortnight and days-of-week.
+Default values for these categorizations can be chosen based on levels of common temporal granularities like days-of-month, days-of-fortnight and days-of-week.
 For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.
@@
 
-These cyclic granularities are a function of the index variable and are computed based on if they are cicular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}).
+These cyclic granularities are a function of the index variable and are computed based on if they are circular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}).
 If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following cyclic time granularities can be considered for this analysis.
@@
 
-### Screening and visualising harmonies:
+### Screening and visualizing harmonies:
 
@@
 Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualisation.
-However, harmony/clash pairs can be identified among those 30 possibilties to determine feasibility of plotting any pairs together.
+However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together.
 We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
@@
 This case study shows systematic exploration of energy behavior for a random household to gain some insights on periodic behavior of the households.
-First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualise a harmony pair and shows the effect of different distribution plots or reverse mapping.
+First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualize a harmony pair and shows the effect of different distribution plots or reverse mapping.
 
@@
 This might indicate that players are vulnerable to score through each over and make more runs in the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
-To answer Q1, winning teams like CSK and MI have indeed different startegies when it comes to batting in the first or second innings.
+To answer Q1, winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.
 

### 2020-02-08 main.tex
spell check

@@
 
-\hypertarget{screening-and-visualising-harmonies}{%
-\subsubsection{Screening and visualising harmonies:}\label{screening-and-visualising-harmonies}}
+\hypertarget{screening-and-visualizing-harmonies}{%
+\subsubsection{Screening and visualizing harmonies:}\label{screening-and-visualizing-harmonies}}
 

### 2020-02-08 main.Rmd
title change

@@
 ---
-title: Exploring probability distributions for bivariate temporal granularities
+title: Visualizing probability distributions across bivariate cyclic temporal granularities
+
+# to produce blinded version set to 1
+blinded: 0
+
 authors:
@@
 - time granularities
+- calendar algebra
 - periodicties

### 2020-02-08 main.tex
title change

@@
 % NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{}
+\newcommand{\blind}{0}
 
@@
 {
-  \title{\bf Exploring probability distributions for bivariate temporal granularities}
+  \title{\bf Visualizing probability distributions across bivariate cyclic temporal granularities}
 
@@
   \begin{center}
-    {\LARGE\bf Exploring probability distributions for bivariate temporal granularities}
+    {\LARGE\bf Visualizing probability distributions across bivariate cyclic temporal granularities}
   \end{center}
@@
 \noindent%
-{\it Keywords:} data visualization, statistical distributions, time granularities, periodicties, grammar of graphics, R
+{\it Keywords:} data visualization, statistical distributions, time granularities, calendar algebra, periodicties, grammar of graphics, R
 \vfill

### 2020-02-10 main.Rmd
till cyclic granularity section

@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ denote the period of the grouping $(B, G')$ and the number of granules of $G'$ in each of these periods, ${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$, $T_w$ are the sets of indices of $B$s describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ denote the period of the grouping $(B, G')$ and the number of granules of $G'$ in each of these periods, ${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$, $T_w$ are the sets of indices of $B$ describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{definition}
@@
 
-```{r quasi-circular-example, echo=FALSE,out.width = "100%",fig.cap="Relating quasi-circular granularities and bottom linear granularity "}
+```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="Quasi-circular granularities $Q(B, G')$ and $Q(B, H')$ relating bottom granularity $B$ and linear granularities $G'$ and $H'$ respectively."}
 knitr::include_graphics("Figs/quasi-circular-example.png")
@@
 Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
-Most public holidays repeat every year, but there is no period within which their behavior remains constant.
+Most public holidays repeat every year, but there is no finite (or reasonably small) period within which their behavior remains constant.
 A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
@@
 All of these are recurring events, but with non-periodic patterns.
-As such, plausible $P_i$ from \autoref{tab:quasi} can be infinite for aperiodic granularities.
+As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
 
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ (that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity).
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$, that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
 Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
@@
 and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
+Here, index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity.
 
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{i=1}^{n}M_{i}$ and index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{i=1}^{n}M_{i}$.
 \end{definition}
@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Relating aperiodic granularity and bottom linear granularity "}
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Aperiodic cyclic granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$"}
 knitr::include_graphics("Figs/aperiodic-example.png")

### 2020-02-10 main.tex
till cyclic granularity section

@@
 
-\caption{Relating quasi-circular granularities and bottom linear granularity }\label{fig:quasi-circular-example}
+\caption{Quasi-circular granularities $Q(B, G')$ and $Q(B, H')$ relating bottom granularity $B$ and linear granularities $G'$ and $H'$ respectively.}\label{fig:quasi-circular-example}
 \end{figure}
@@
 All of these are recurring events, but with non-periodic patterns.
-As such, plausible \(P_i\) from \autoref{tab:quasi} can be infinite for aperiodic granularities.
+As such, plausible \(P_i\) from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
 
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities \(M_i\) \(\forall \{i\in{1, 2, \dots, n\}}\) (that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity).
+Consider n aperiodic linear granularities \(M_i\) \(\forall \{i\in{1, 2, \dots, n\}}\), that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
 Further assume that \(B \trianglelefteq M_i\) \(\forall \{i\in{1, 2, \dots, n\}}\).
@@
 and suppose \(M = \bigcup_{i=1}^{n}M_{i}\) is formed by collecting the granules of \(\{M_1, M_2, \dots, M_n\}\).
+Here, index \({\{i_j\}}\) stands for the \(j^{th}\) granule of the \(i^{th}\) linear aperiodic granularity.
 
@@
 
-\caption{Relating aperiodic granularity and bottom linear granularity }\label{fig:aperiodic-example}
+\caption{Aperiodic cyclic granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$}\label{fig:aperiodic-example}
 \end{figure}

### 2020-02-10 main.Rmd
minor edits in cyclic calendar algebra

@@
 **Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
-$G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.
+$G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in {1, 2, \dots,n}$ and $l<m$.
 
@@
 \begin{split}
-C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_{l+1},G_m)}(z))) \\  
-  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z))) \\
-  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
-  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
+C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{(G_{l+1},G_m)}(z))) \\  
+  & =   L(C_{(G_l,G_{l+1}}(z) + P(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)( C_{G_{l+2}, G_m}(z))) \\
+  & =  L(C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+1)P(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
+  & =  L(C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
   &\vdots\\
-  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z)))\\
+  & = L(\sum_{i=0}^{m - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z)))\\
 \end{split}
@@
 C_{uinal, baktun}(z) & = L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
-              & = L(\lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
+              & = L(\lfloor z/20\rfloor \mod 18  +
+              20*\lfloor z/20*18\rfloor \mod 20 +
+              20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
 \end{split}
@@
 
-If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z) are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
+If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
 

### 2020-02-10 main.Rmd
cleaning cricket and cyclic calendar algrabra

@@
 
-The way cyclic granularities relate become important when we consider the data structure in \@autoref{fig:data-structure}.
+The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
 While considering data sets of the form $<C_1, C_2, v>$, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
@@
 Thus, the cyclic granularity ball-of-the-over will be circular in most cases and aperiodic in others.
-However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal dataset similar to any temporal dataset.
+
+<!-- However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal dataset similar to any temporal dataset.
+-->
 
@@
 The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
-The hierarchy table would look like the following:
+This could be useful in a periodic world when we wish to compute any circular/quasi-circular granularity based on a hierarchy table which look like the following:
 
@@
 
-There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal dataset similar to any temporal dataset.
+There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
+We will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
 
@@
 <!-- - Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings? -->
-
-```{r runs_per-over}
- library(tsibble)
-
- cricket_all <-  read_csv("data-raw/deliveries_all.csv")
- matches_all <- read_csv("data-raw/matches_all.csv")
-
- cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
-
- cricket_per_over <- cricket_season %>%
-   group_by(season,
-            match_id,
-            batting_team,
-            bowling_team,
-            inning,
-            over) %>%
-   summarise(runs_per_over = sum(total_runs),
-             run_rate = sum(total_runs)/length(total_runs))
-
- cricket_tsibble_all <- cricket_per_over %>%
-   ungroup() %>%
-   mutate(data_index = row_number()) %>%
-   as_tsibble(index = data_index)
-```
 
@@
 
-# cricket_tsibble %>%
-#   filter(batting_team %in% c("Mumbai Indians",
-#                              "Chennai Super Kings"))%>%
-#   create_gran("inning_match", hierarchy_model) %>% create_gran("over_inning", hierarchy_model)
-
-cricket_tsibble %>%
-  filter(batting_team %in% c("Mumbai Indians",
-                             "Chennai Super Kings"))%>%
-  gran_advice("inning", 
-            "over")
-
 cricket_tsibble %>%
@@
 
-# cricket_tsibble %>%
-#   filter(batting_team %in% c("Mumbai Indians",
-#                              "Chennai Super Kings"))%>%
-#   gran_advice("inning", 
-#             "over")
-
 ```
@@
 
-```{r field, fig.cap = "Violin plots with an overlaid boxplot showing distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
-This shows good fielding in previous over leads to lower run rate in the subsequent over.
-Also run rates are much less volatile in case there are at least one wicket in the previous over.", warning=FALSE}
+<!-- fetching raw data since dot and fielding information not available in dismissal type not available in cricket data gravitas -->
+
+```{r runs_per-over}
+ library(tsibble)
+
+ cricket_all <-  read_csv("data-raw/deliveries_all.csv")
+ matches_all <- read_csv("data-raw/matches_all.csv")
+
+ cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
+
+ # cricket_per_over <- cricket_season %>%
+ #   group_by(season,
+ #            match_id,
+ #            batting_team,
+ #            bowling_team,
+ #            inning,
+ #            over) %>%
+ #   summarise(runs_per_over = sum(total_runs),
+ #             run_rate = sum(total_runs)/length(total_runs))
+ # 
+ # cricket_tsibble_all <- cricket_per_over %>%
+ #   ungroup() %>%
+ #   mutate(data_index = row_number()) %>%
+ #   as_tsibble(index = data_index)
+```
+
+```{r}
 
@@
 
-cricket_tsibble %>% gran_obs("over", "fielding_wckts")
-
-cricket_data_ts <- cricket_tsibble %>% 
+cricket_data<- cricket_tsibble %>% 
   mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets"),
          dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) 
+```
+
+```{r echo = TRUE}
+cricket_data %>% gran_obs("field", "over")
+cricket_data %>% gran_obs("dot", "over")
+```
+
+```{r field,  fig.cap = "Violin plots with an overlaid boxplot showing distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+This shows good fielding in previous over leads to lower run rate in the subsequent over.
+Also run rates are much less volatile in case there are at least one wicket in the previous over.", warning=FALSE}
 
   #filter(fielding_wckts %in% c(0,1)) %>% 
-  cricket_data_ts %>% prob_plot( "over","field",
+  cricket_data %>% prob_plot("over","field",
                                hierarchy_model,
@@
 
-cricket_data_ts %>% 
+cricket_data %>% 
   #filter(dot_balls %in% c(0, 1, 2)) %>%

### 2020-02-10 main.tex
cleaning cricket and cyclic calendar algrabra

@@
 \textbf{Hierarchy table} Let \(H_n: (G, C, P)\) be a hierarchy model containing n linear granularities.
-\(G_{l}\) represents the linear granularity of order l and \(P(l,m)\) represents the period length of \(G_{l}\) with respect to \(G_{m}\) and \(C_{G(l),G(m)}\) represents the cyclic granularity that relates linear granularity of order l and m, \(\forall l,m \in n, l<m\).
+\(G_{l}\) represents the linear granularity of order l and \(P(l,m)\) represents the period length of \(G_{l}\) with respect to \(G_{m}\) and \(C_{G(l),G(m)}\) represents the cyclic granularity that relates linear granularity of order l and m, \(\forall l,m \in {1, 2, \dots,n}\) and \(l<m\).
 
@@
 
-If \(C_{G_{l},G_{m'}}(z)\) or \$C\_\{G\_\{l\},G\_\{m'\}\}(z) are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
+If \(C_{G_{l},G_{m'}}(z)\) or \(C_{G_{l},G_{m'}}(z)\) are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
 
@@
 
-The way cyclic granularities relate become important when we consider the data structure in @autoref\{fig:data-structure\}.
+The way cyclic granularities relate become important when we consider the data structure in Figure \ref{fig:data-structure}.
 While considering data sets of the form \(<C_1, C_2, v>\), let \(S_{ij}\) be the set of index set such that for all \(s \in S_{ij}\), \(C_1(s) = A_i\) and \(C_2(s) = B_j\).
@@
 Thus, the cyclic granularity ball-of-the-over will be circular in most cases and aperiodic in others.
-However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a non-temporal dataset similar to any temporal dataset.
 
@@
 The \texttt{cricket} data set in the \texttt{gravitas} package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
-The hierarchy table would look like the following:
+This could be useful in a periodic world when we wish to compute any circular/quasi-circular granularity based on a hierarchy table which look like the following:
 
@@
 
-There are many interesting questions that can possibly be answered with such a data set, however, we will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a non-temporal dataset similar to any temporal dataset.
+There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
+We will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
 
@@
 
-\begin{verbatim}
-#> The chosen granularities are harmonies 
-#>  
-#> Recommended plots are: violin lv quantile boxplot 
-#>  
-#> Number of observations are homogenous across facets 
-#>  
-#> Number of observations are homogenous within facets 
-#>  
-#> Cross tabulation of granularities : 
-#>  
-#> # A tibble: 20 x 3
-#>    over    `1`   `2`
-#>    <fct> <dbl> <dbl>
-#>  1 1        65    44
-#>  2 2        65    44
-#>  3 3        65    44
-#>  4 4        65    44
-#>  5 5        65    44
-#>  6 6        65    44
-#>  7 7        65    44
-#>  8 8        65    44
-#>  9 9        65    44
-#> 10 10       65    44
-#> 11 11       65    44
-#> 12 12       65    44
-#> 13 13       65    44
-#> 14 14       65    44
-#> 15 15       65    44
-#> 16 16       65    44
-#> 17 17       65    44
-#> 18 18       65    44
-#> 19 19       65    44
-#> 20 20       65    44
-\end{verbatim}
-
 \begin{figure}
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cricket_data }\OperatorTok{%>%}\StringTok{ }\KeywordTok{gran_obs}\NormalTok{(}\StringTok{"field"}\NormalTok{, }\StringTok{"over"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{verbatim}
-#> # A tibble: 4 x 21
-#>   fielding_wckts   `1`   `2`   `3`   `4`   `5`   `6`   `7`   `8`   `9`  `10`
-#>   <fct>          <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
-#> 1 0               1039   997   972   978   962   957  1002  1012   993   984
-#> 2 1                120   145   171   165   180   181   138   124   140   153
-#> 3 2                  5    10     8     8     9    10     6    10     8     4
-#> 4 3                  0     0     0     0     0     0     0     0     0     0
-#> # … with 10 more variables: `11` <dbl>, `12` <dbl>, `13` <dbl>, `14` <dbl>,
-#> #   `15` <dbl>, `16` <dbl>, `17` <dbl>, `18` <dbl>, `19` <dbl>, `20` <dbl>
+#> # A tibble: 20 x 3
+#>    over  `>0 wickets` `no wickets`
+#>    <fct>        <dbl>        <dbl>
+#>  1 1              125         1039
+#>  2 2              155          997
+#>  3 3              179          972
+#>  4 4              173          978
+#>  5 5              189          962
+#>  6 6              191          957
+#>  7 7              144         1002
+#>  8 8              134         1012
+#>  9 9              148          993
+#> 10 10             157          984
+#> 11 11             173          962
+#> 12 12             177          951
+#> 13 13             169          954
+#> 14 14             205          913
+#> 15 15             208          892
+#> 16 16             219          865
+#> 17 17             231          833
+#> 18 18             291          743
+#> 19 19             287          695
+#> 20 20             306          572
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cricket_data }\OperatorTok{%>%}\StringTok{ }\KeywordTok{gran_obs}\NormalTok{(}\StringTok{"dot"}\NormalTok{, }\StringTok{"over"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+#> # A tibble: 20 x 3
+#>    over  `>0 dot balls` `no dot balls`
+#>    <fct>          <dbl>          <dbl>
+#>  1 1               1146             18
+#>  2 2               1140             12
+#>  3 3               1111             40
+#>  4 4               1112             39
+#>  5 5               1099             52
+#>  6 6               1099             49
+#>  7 7               1057             89
+#>  8 8               1026            120
+#>  9 9               1027            114
+#> 10 10              1010            131
+#> 11 11               992            143
+#> 12 12               971            157
+#> 13 13               970            153
+#> 14 14               957            161
+#> 15 15               935            165
+#> 16 16               904            180
+#> 17 17               881            183
+#> 18 18               852            182
+#> 19 19               788            194
+#> 20 20               683            195
 \end{verbatim}

### 2020-02-10 main.Rmd
cricketdata change -  with lag dot

@@
 
-<!-- fetching raw data since dot and fielding information not available in dismissal type not available in cricket data gravitas -->
+<!-- fetching raw data since dot and fielding information not available in dismissal type not available in cricket data gravitas. Also `cricket` is aggregated across overs -->
 
@@
   mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets"),
-         dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) 
+         dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls"),
+         lag_field =  lag(field),
+         lag_dot = lag(dot)) %>% filter(lag_field!=0, lag_dot!=0)
 ```
@@
   #filter(fielding_wckts %in% c(0,1)) %>% 
-  cricket_data %>% prob_plot("over","field",
+  cricket_data %>% prob_plot("over","lag_field",
                                hierarchy_model,
-                               response = "diff_run_rate",
+                               response = "run_rate",
                                plot_type = "violin",
@@
 Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
-With at least one dot balls per over, the distribution of difference in run rates as observed through the quantiles plotted have shifted downwards.
+With at least one dot balls per over, the distribution of run rates in facet 1 increase slower compared to that in facet 2.
 This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
@@
 
-```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), more than zero (facet 2) dot balls per over.
-For all quantiles, difference in runs between subsequent overs decrease with at least one dot ball in the previous over.", fig.pos="ht"}
+```{r exdot, fig.cap="25th, 50th, 75th quantiles of runs per over are drawn  across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
+For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.", fig.pos="ht"}
 
@@
   #filter(dot_balls %in% c(0, 1, 2)) %>%
-  prob_plot("dot",
+  prob_plot("lag_dot",
             "over",
             hierarchy_model,
-            response = "diff_run_rate",
+            response = "run_rate",
             plot_type = "quantile",
-            quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
+            quantile_prob = c(0.25, 0.5, 0.75),
             symmetric = FALSE)  + ggtitle("")

### 2020-02-10 main.tex
cricketdata change -  with lag dot

@@
 #>    <fct>        <dbl>        <dbl>
-#>  1 1              125         1039
+#>  1 1              125         1038
 #>  2 2              155          997
@@
 #>    <fct>          <dbl>          <dbl>
-#>  1 1               1146             18
+#>  1 1               1145             18
 #>  2 2               1140             12
@@
 
-\caption{10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), more than zero (facet 2) dot balls per over.
-For all quantiles, difference in runs between subsequent overs decrease with at least one dot ball in the previous over.}\label{fig:exdot}
+\caption{25th, 50th, 75th quantiles of runs per over are drawn  across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
+For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.}\label{fig:exdot}
 \end{figure}

### 2020-02-10 main.Rmd
cricketdata change -  with lag dot and wicket

@@
 Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
-Figure \ref{fig:field} shows that at least one dismissal in an over leads to a lower median and less volatility in the distribution of difference in run rates.
+Figure \ref{fig:field} shows that at least one dismissal in a previous over leads to a lower median and less spread in the distribution of run rates of the subsequent over.
 
@@
 
-```{r field,  fig.cap = "Violin plots with an overlaid boxplot showing distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+```{r field,  fig.cap = "Violin plots with an overlaid boxplot showing distribution of run rate across overs of the innings and (lagged) wickets dismissed due to fielding.
 This shows good fielding in previous over leads to lower run rate in the subsequent over.
-Also run rates are much less volatile in case there are at least one wicket in the previous over.", warning=FALSE}
+Also distribution of run rates has less spread in case there is at least one wicket in the previous over.", warning=FALSE}
 

### 2020-02-10 main.tex
cricketdata change -  with lag dot and wicket

@@
 
-\caption{Violin plots with an overlaid boxplot showing distribution of lagged run rate across overs of the innings and wickets dismissed due to fielding.
+\caption{Violin plots with an overlaid boxplot showing distribution of run rate across overs of the innings and (lagged) wickets dismissed due to fielding.
 This shows good fielding in previous over leads to lower run rate in the subsequent over.
-Also run rates are much less volatile in case there are at least one wicket in the previous over.}\label{fig:field}
+Also distribution of run rates has less spread in case there is at least one wicket in the previous over.}\label{fig:field}
 \end{figure}

### 2020-02-10 main.Rmd
cricketdata change -  with lag dot and wicket

@@
                                plot_type = "violin",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-           overlay = TRUE)  + ggtitle("") + geom_boxplot(width = 0.5)
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") + geom_boxplot(width = 0.5)
 ```

### 2020-02-13 main.Rmd
Updated title and abstract

@@
 ---
-title: Visualizing probability distributions across bivariate cyclic temporal granularities
+title: Visualizing probability distributions across bivariate nonlinear temporal granularities
 
@@
  Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
-As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases.
+As the frequency of time-oriented data increases, the number of questions about the observed variable that can be addressed by visual representation also increases.
 We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear and cyclic time granularities.
+There are several classes of time deconstructions including linear and nonlinear time granularities.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month and aperiodic such as public or school holidays.
+Nonlinear time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
 
@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
+This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
 bibliography: [bibliography.bib]

### 2020-02-13 main.Rmd
Cleaning up blank lines
Start of intro

@@
 With recent advancement in technology, more and more data are recorded at much finer temporal scales.
-Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be recorded every second.
+Energy consumption may be collected every half an hour, energy supply may be collected every minute, and web search data might be recorded every second.
 As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases.
@@
 This approach requires deconstructing time in various possible ways called time granularities [@aigner2011visualization].
+
 It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
@@
 Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
-Analysts are expected to iteratively explore all possible choices of time granularities for comprehending possible periodicities in the data.
+Analysts are expected to iteratively explore many possible choices of time granularities for comprehending possible periodicities in the data.
 But too many choices and a lack of a systematic approach to do so might become overwhelming.

### 2020-02-13 main.Rmd
Improved wording in Section 1

@@
 Temporal data are available at various resolutions depending on the context.
-Social and economic data like GDP is often collected and reported at coarse temporal scales like monthly, quarterly or annually.
+Social and economic data like GDP is often collected and reported at coarse temporal scales such as monthly, quarterly or annually.
 With recent advancement in technology, more and more data are recorded at much finer temporal scales.
@@
 As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases.
-For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
+For example, data collected at an hourly scale can be analyzed using coarser temporal scales such as days, months or quarters.
 This approach requires deconstructing time in various possible ways called time granularities [@aigner2011visualization].
@@
 
-Calendar-based graphics [@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
-Any sub-daily resolution temporal data can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year and another sub-daily deconstruction of time.
-But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
-There can be several classes of time deconstructions, viz.
-based on the arrangement (linear vs.
-cyclic) or hierarchical order of the calendar.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-One of the first attempts of characterizing these granularities occur in @Bettini1998-ed.
-The definitions and rules defined are inadequate to reflect periodicities in time.
-Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior.
-Cyclic time granularities can be circular, quasi-circular or aperiodic.
-Examples of circular can be hour of the day, day of the week, that of quasi-circular can be day of the month and examples of aperiodic granularities can be public or school holidays.
+Calendar-based graphics [@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure, and are helpful when checking for the effects of weekends or special days.
+Any temporal data at sub-daily resolution can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year, or another sub-daily deconstruction of time.
+But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions.
+There can be many different time deconstructions, based on the calendar or on categorizations of time granularities.
+
+Linear time granularities respect the linear progression of time such as hours, days, weeks and months, and non-repeating.
+One of the first attempts to characterize these granularities is due to @Bettini1998-ed.
+However, the definitions and rules defined are inadequate for describing nonlinear or repeating granularities.
+Hence, there is a need to define some new nonlinear time granularities, that can be useful in visualizations.
+Nonlinear time granularities can be circular, quasi-circular or aperiodic.
+Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularty is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+
 Time deconstructions can also be based on the hierarchical structure of time.
 For example, hours are nested within days, days within weeks, weeks within months, and so on.
-Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
-lubridate [@G_Grolemund2011-vm] creates easy access and manipulation of common date-time objects.
-But most accessor functions are limited to single-order-up granularities.
+Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour.
+The lubridate package [@G_Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
+But most of its accessor functions are limited to single-order-up granularities.
 
@@
 
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
+The motivation for this work stems from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
 Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates (see @2012-la) of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
-Although, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analyzed for different periodicities.
+It is common to see aggregates [see @2012-la] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
+But studying overall energy use hides the distributions of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.
+We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
+Although, the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.
 
@@
 <!-- challenges and how you are dealing wih it -->
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow through the following:
-
-<!-- Pairs of granularities are categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
--->
-
-  * Formal characterization of cyclic granularities
-
-  * Facilitate manipulation of single and multiple order-up time granularities through cyclic calendar algebra
-
-  * Checking feasibility of creating plots or drawing inferences for any two cyclic granularities
-
-  * Recommend prospective probability distributions for exploring distributions of univariate dependent variable across pair of granularities
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow. In particular, we
+
+  * provide a formal characterization of nonlinear granularities;
+  * facilitate manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra;
+  * develop an approach to check the feasibility of creating plots or drawing inferences for any two nonlinear granularities;
+  * recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pair of granularities.
 
@@
 
-The remainder of the paper is organized as follows: Section \ref{sec:linear-time} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
-Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
-Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
-Section \ref{sec:data-structure} discusses the data structure for exploring distribution of the associated time series across pair of cyclic time granularities.
+The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provide some background material on linear granularities and introduces calendar algebra for computing different linear granularities.
+Section \ref{sec:cyclic-gran} formally characterizes different nonlinear time granularities by extending the framework of linear time granularities.
+Section \ref{sec:cyclic-calendar} introduces calendar algebra for computing nonlinear time granularities.
+Section \ref{sec:data-structure} discusses the data structure for exploring the conditional distributions of the associated time series across pairs of nonlinear time granularities.
 Section \ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.

### 2020-02-13 main.Rmd
Added email address

@@
   affiliation: Department of Econometrics and Business Statistics, Monash University
-  email: Sayani.Gupta@monash.edu
+  thanks: "Email: Sayani.Gupta@monash.edu"
 
@@
     dev: "pdf"
+    keep_tex: yes
 ---

### 2020-02-13 main.Rmd
Improvements to Section 2

@@
 
-Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
+Discrete abstraction of time such as hours, days, weeks and months can be thought of as linear "time granularities" [@aigner2011visualization].
 Time granularities are **linear** if they respect the linear progression of time.
-Examples include hours, days, weeks and months.
 
@@
 
- There has been several attempts to provide the framework for formally characterizing time-granularities. One of the first attempts occur in @Bettini1998-ed with the help of the following definitions:
+There have been several attempts to provide a framework for formally characterizing time granularities. One of the first attempts was provided by @Bettini1998-ed who proposed following definitions:
 
@@
 
-\begin{definition}\label{def:definition}
-(time domain): A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+\begin{definition}[Time domain]\label{def:definition}
+A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
 \end{definition}
 
-A time domain can be **discrete** (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be **dense** (if it is an infinite set).
-A time domain is assumed to be discrete for the purpose of our discussion.
-
-\begin{definition}\label{def:linear}
-(linear granularity): A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
+We will assume the time domain is *discrete*; i.e., there is unique predecessor and successor for every element in the time domain except for the first and last.
+
+\begin{definition}[Linear granularity]\label{def:linear}
+A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
 
   (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
-than all elements of $G(j)$, and
+than all elements of $G(j)$; and
   (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
@@
 
-\begin{definition}\label{def:granule}
-(granule) Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
+\begin{definition}[Granule]\label{def:granule}
+Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
 \end{definition}
 
-**Discussion:** The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
-\autoref{fig:linear-time} shows the implication of this condition.
-It shows the linear granularities hours, days, weeks, months and years, each of which are unidirectional in nature and arranged from past to future.
-If we consider the chronon (see @aigner2011visualization) as hourly, the time domain with T hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
-Each granule of the linear granularities shown are represented through a box.
-
-```{r linear-time,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities"}
+The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+\autoref{fig:linear-time} illustrates the linear time granularites hours, days, weeks, months and years, each of which are non-repeating in nature and arranged from past to future.
+If we consider the chronon [see @aigner2011visualization] as hourly, the time domain with $T$ hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+
+```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="The time domain distributed as linear granularities."}
 knitr::include_graphics("Figs/linear-arr.png")
@@
 
-[@Bettini1998-ed] talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
-
-\begin{definition}\label{def:finerthan}
-(finer than): A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
-for each index i, there exists an index j such that $G(i) \subset H(j)$.
+@Bettini1998-ed discusses the relationships between linear time granularities, the structure of a calendar, and periodicities in time.
+
+\begin{definition}[Finer than]\label{def:finerthan}
+A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$.
 \end{definition}.
 
-\begin{definition}\label{def:groupsinto}
-(groups into): A linear granularity G groups into a linear granularity H, denoted
-$G \trianglelefteq H$,
-if for each index j there exists a (possibly infinite) subset S of the integers such
-that
-\begin{equation}
-H(j) = \bigcup_{i \in S}G(i)
-\end{equation}
+\begin{definition}[Groups into]\label{def:groupsinto}
+A linear granularity $G$ groups into a linear granularity $H$, denoted
+$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that
+\[
+  H(j) = \bigcup_{i \in S}G(i).
+\]
 \end{definition}
 
-**Discussion:** $G \trianglelefteq H$ if each granule H(j) is the union of some granules of G.
-$G \preceq H$ if every granule in G is a subset of some granule in H.
-In \autoref{fig:groups-finer}, both $day \trianglelefteq week$ and $day \preceq week$ hold true.
+These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true.
 However, weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or finer than public holidays.
-Consider another example where $day \trianglelefteq month$.
+And weekends neither groups into or is finer than public holidays.
+Consider another example where day $\trianglelefteq$ month.
 This relationship is not fully described until we associate it with periodicity.
 Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions, .
-
-```{r groups-finer,echo=FALSE,out.width = "100%",fig.cap="Illustrations of groups into and finer than relationships.
+The period becomes 400 years with the inclusion of leap years and all their exceptions.
+
+```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships.
 Weekends are finer than weeks but do not group into weeks.
@@
 
-\begin{definition}\label{def:periodical}
-(groups periodically): A granularity H is periodical with respect to a granularity G if
-(1) $G \trianglelefteq H$, and
-(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
-that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
-$H (i + R) = \bigcup_{j \in S} G(j + P)$.
+\begin{definition}[Groups periodically]\label{def:periodical}
+A granularity $H$ is periodical with respect to a granularity $G$ if:
+(1) $G \trianglelefteq H$; and
+(2) there exist $R$, $P \in Z+$, where $R$ is less than the number of granules of $H$, such that for all $i \in Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
 \end{definition}
 
-**Discussion:** A granularity H which is periodical with respect to G is specified by:
-(i) the R sets of indexes of G, ${S_0,...,S_{R-1}}$ describing the granules of H within one period;
-(ii) the value of P;
-(iii) the indexes of first and last granules in H, if their value is not infinite.
-Then, if ${S_0,...,S_{R-1}}$ are the sets of indexes of G describing
-$H (0), .
-.
-.
-, H (R - 1)$, respectively, then the description of an arbitrary granule H(j) is given by: $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
+A granularity $H$ which is periodical with respect to $G$ is specified by:
+(i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period;
+(ii) the value of $P$;
+(iii) the indexes of first and last granules in $H$, if their value is not infinite.
+Then, if ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
+$H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
 Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
 
-\begin{definition}\label{def:quasi}
-(groups quasi-periodically): A granularity H is quasi-periodical with respect to a granularity G if
-(1) $G \trianglelefteq H$, and
-(2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
-integers R,P, where R is less than the minimum of the number of granules of H
-between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
+\begin{definition}[Groups quasi-periodically]\label{def:quasi}
+A granularity $H$ is quasi-periodical with respect to a granularity $G$ if:
+(1) $G \trianglelefteq H$; and
+(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
 \end{definition}
 
-**Discussion:** Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
-
-\begin{definition}\label{def:B}
-(bottom granularity): Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
+Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
+
+\begin{definition}[Bottom granularity]\label{def:B}
+Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
 \end{definition}
 
-**Discussion:** Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
 
@@
 - $B \trianglelefteq G'$
-- there exist $R', P' \epsilon Z_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
+- there exist $R'$, $P' \in Z_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
 

### 2020-02-13 main.Rmd
Simplified preamble

@@
 preamble: >
-  \usepackage{mathtools}
-  \usepackage{amssymb}
-  \usepackage{booktabs}
-  \usepackage{longtable}
+  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
 output:

### 2020-02-13 main.Rmd
Questions regarding section 2

@@
 
+\todo[inline]{Give an example}
+
 \begin{definition}[Groups quasi-periodically]\label{def:quasi}
@@
 (1) $G \trianglelefteq H$; and
-(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
+(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
 \end{definition}
@@
 Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
+
+\todo[inline]{Give an example}
 
@@
 
+\todo[inline]{What is g-rel?}
+
 Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate
-new granularities from the bottom one or recursively, from those already generated.
-The calendar algebra consists of two kinds of operations: grouping-oriented and
-granule-oriented operations.
+It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
 The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
-Example can be to consider a calendar with only two linear granularities "minute" and "hour" and "hour" is generated by grouping every $60$ "minutes".
+For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes".
 The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
@@
 
-Some relevant grouping oriented operations are discussed, which will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
-
-- **The grouping operation** : Let $G1$ be a full-integer labeled granularity, and $m$ a positive integer.
-The grouping operation $Group_m(G)$ generates a new granularity $G2$, by partitioning the granules of $G$ into m-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G2 = Group_m(G1)$ is the full-integer labeled granularity such that for
-each integer $i$, $G2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G1(j)$.
-
-**Examples** :
-
-  - $minute = Group_{60}(second)$
-  - $hour = Group_{60}(minute)$,
-  - $day = Group_{24}(hour)$,
-  - $week = Group_{7}(day)$,
-
-where, second(1) would start a minute and likewise.
-
-- **The altering-tick operation** : Let G1, G2 be full-integer labeled granularities, and $l$, $k$, $m$ integers, where G2 partitions G1, and $1 \leq l \leq m$.
-The altering-tick operation $Alter^{m}_{l,k} (G2, G1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-The altering-tick operation modifies the granules of G1 so that the $l$th granule of each group has $|k|$ additional (or fewer when k < 0) granules of G2.
-
-**Example** :
-
-\begin{equation*}
-\begin{split}
-pseudomonth = Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, \\
-  Alter^{12}_{6,-1}(day, Alter^{12}_{4,-1)}(day, \\
-  Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))
-\end{split}
-\end{equation*}
-
-, where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days. [@Ning2002-tf]
-
-For more variations of grouping operations and granule oriented operations, the readers are recommended to see @Ning2002-tf.
-
-# Cyclic time granularities {#sec:cyclic-gran}
+Some relevant grouping-oriented operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
+
+- **The grouping operation**: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
+The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
+More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
+$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
+
+\todo[inline]{What does "full-integer" mean?}
+
+For example:
+\begin{align*}
+  \text{minute} &= \text{Group}_{60}(\text{second});\\
+  \text{hour}   &= \text{Group}_{60}(\text{minute});\\
+  \text{day}    &= \text{Group}_{24}(\text{hour});\\
+  \text{week}   &= \text{Group}_{7}(\text{day});
+\end{align*}
+
+where $\text{second}(1)$ would start a minute and likewise.
+
+\todo[inline]{What does this last phrase mean?}
+
+- **The altering-tick operation** : Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
+The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$.
+The altering-tick operation modifies the granules of $G_1$ so that the $l$th granule of each group has $|k|$ additional (or fewer when $k < 0$) granules of $G_2$.
+
+For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
+\begin{align*}
+\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \\
+  &~~\text{Alter}^{12}_{9,-1}(\text{day}, \\
+  &~~~~\text{Alter}^{12}_{6,-1}(\text{day}, \\
+  &~~~~~~\text{Alter}^{12}_{4,-1)}(\text{day}, \\
+  &~~~~~~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \\
+  &~~~~~~~~~~\text{Group}_{31}(\text{day})))))),
+\end{align*}
+where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
+
+\todo[inline]{Why is this called a ``pseudomonth''?}
+
+For more variations of grouping operations and granule oriented operations, see @Ning2002-tf.
+
+# Nonlinear time granularities {#sec:cyclic-gran}
 
@@
 
-We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
+We propose a formalism of nonlinear time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
 A tsibble consists of an index, key and measured variables.
@@
 In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
-Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
-
-Cyclic time granularities which accommodate for periodicities can be constructed by relating two linear granularities.
+Hence intuitively, nonlinear granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
+
+Nonlinear time granularities which accommodate for periodicities can be constructed by relating two linear granularities.
 The mappings between these linear granularities can be regular or irregular.
@@
 However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
-Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
+Nonlinear time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 
-Aperiodic cyclic granularities are defined using aperiodic linear granularities.
+Aperiodic nonlinear granularities are defined using aperiodic linear granularities.
 Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$, that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
@@
 
-\begin{definition}\label{aperiodic}
-(aperiodic cyclic granularity): An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
-
+\begin{definition}[Aperiodic nonlinear granularity]\label{aperiodic}
+An aperiodic nonlinear granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
 \begin{equation}\label{eq:aperiodic}
@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Aperiodic cyclic granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$"}
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Aperiodic nonlinear granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$"}
 knitr::include_graphics("Figs/aperiodic-example.png")
@@
 
-In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
+In \autoref{sec:cyclic-gran}, we discussed how we can define nonlinear granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.

### 2020-02-13 main.tex
Recompiled

@@
 %% load any required packages here
-\usepackage{mathtools} \usepackage{amssymb}  \usepackage{booktabs} \usepackage{longtable}
+\usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
 
@@
 {
-  \title{\bf Visualizing probability distributions across bivariate cyclic temporal granularities}
+  \title{\bf Visualizing probability distributions across bivariate nonlinear temporal granularities}
 
   \author{
-        Sayani Gupta \\
+        Sayani Gupta \thanks{Email: \href{mailto:Sayani.Gupta@monash.edu}{\nolinkurl{Sayani.Gupta@monash.edu}}} \\
     Department of Econometrics and Business Statistics, Monash University\\
@@
   \begin{center}
-    {\LARGE\bf Visualizing probability distributions across bivariate cyclic temporal granularities}
+    {\LARGE\bf Visualizing probability distributions across bivariate nonlinear temporal granularities}
   \end{center}
@@
 Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
-As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases.
+As the frequency of time-oriented data increases, the number of questions about the observed variable that can be addressed by visual representation also increases.
 We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear and cyclic time granularities.
+There are several classes of time deconstructions including linear and nonlinear time granularities.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month and aperiodic such as public or school holidays.
+Nonlinear time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
 
@@
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 \citep{Wickham2009pk}.
+This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 \citep{Wickham2009pk}.
 \end{abstract}
@@
 Temporal data are available at various resolutions depending on the context.
-Social and economic data like GDP is often collected and reported at coarse temporal scales like monthly, quarterly or annually.
+Social and economic data like GDP is often collected and reported at coarse temporal scales such as monthly, quarterly or annually.
 With recent advancement in technology, more and more data are recorded at much finer temporal scales.
-Energy consumption is collected every half an hour, while energy supply is collected every minute and web search data might be recorded every second.
+Energy consumption may be collected every half an hour, energy supply may be collected every minute, and web search data might be recorded every second.
 As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases.
-For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters.
+For example, data collected at an hourly scale can be analyzed using coarser temporal scales such as days, months or quarters.
 This approach requires deconstructing time in various possible ways called time granularities \citep{aigner2011visualization}.
+
 It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
@@
 Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
-Analysts are expected to iteratively explore all possible choices of time granularities for comprehending possible periodicities in the data.
+Analysts are expected to iteratively explore many possible choices of time granularities for comprehending possible periodicities in the data.
 But too many choices and a lack of a systematic approach to do so might become overwhelming.
 
-Calendar-based graphics \citep{wang2018calendar} are useful in visualizing patterns in the weekly and monthly structure well and are capable of checking the weekends or special days.
-Any sub-daily resolution temporal data can also be displayed using this type of faceting \citep{Wickham2009pk} with days of the week, month of the year and another sub-daily deconstruction of time.
-But calendar effects are not restricted to conventional day-of-week, month-of-year ways of deconstructing time.
-There can be several classes of time deconstructions, viz.
-based on the arrangement (linear vs.~cyclic) or hierarchical order of the calendar.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-One of the first attempts of characterizing these granularities occur in \citet{Bettini1998-ed}.
-The definitions and rules defined are inadequate to reflect periodicities in time.
-Hence, there is a need to define cyclic time granularities in a different approach, which can be useful in visualizing periodic behavior.
-Cyclic time granularities can be circular, quasi-circular or aperiodic.
-Examples of circular can be hour of the day, day of the week, that of quasi-circular can be day of the month and examples of aperiodic granularities can be public or school holidays.
+Calendar-based graphics \citep{wang2018calendar} are useful in visualizing patterns in the weekly and monthly structure, and are helpful when checking for the effects of weekends or special days.
+Any temporal data at sub-daily resolution can also be displayed using this type of faceting \citep{Wickham2009pk} with days of the week, month of the year, or another sub-daily deconstruction of time.
+But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions.
+There can be many different time deconstructions, based on the calendar or on categorizations of time granularities.
+
+Linear time granularities respect the linear progression of time such as hours, days, weeks and months, and non-repeating.
+One of the first attempts to characterize these granularities is due to \citet{Bettini1998-ed}.
+However, the definitions and rules defined are inadequate for describing nonlinear or repeating granularities.
+Hence, there is a need to define some new nonlinear time granularities, that can be useful in visualizations.
+Nonlinear time granularities can be circular, quasi-circular or aperiodic.
+Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularty is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+
 Time deconstructions can also be based on the hierarchical structure of time.
 For example, hours are nested within days, days within weeks, weeks within months, and so on.
-Hence, it is possible to construct single-order-up granularities like second of the minute or multiple-order-up granularities like second of the hour.
-lubridate \citep{G_Grolemund2011-vm} creates easy access and manipulation of common date-time objects.
-But most accessor functions are limited to single-order-up granularities.
-
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
+Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour.
+The lubridate package \citep{G_Grolemund2011-vm} provides tools to access and manipulate common date-time objects.
+But most of its accessor functions are limited to single-order-up granularities.
+
+The motivation for this work stems from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
 Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates (see \citet{2012-la}) of usage across households, total kWh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-We propose that the analysis of smart meter data can be benefited from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
-Although, the motivation came through the smart meter example, this is a problem which relates to any data that needs to be analyzed for different periodicities.
-
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow through the following:
+It is common to see aggregates \citep[see][]{2012-la} of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
+But studying overall energy use hides the distributions of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.
+We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
+Although, the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.
+
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow. In particular, we
 
 \begin{itemize}
-\item
-  Formal characterization of cyclic granularities
-\item
-  Facilitate manipulation of single and multiple order-up time granularities through cyclic calendar algebra
-\item
-  Checking feasibility of creating plots or drawing inferences for any two cyclic granularities
-\item
-  Recommend prospective probability distributions for exploring distributions of univariate dependent variable across pair of granularities
+\tightlist
+\item
+  provide a formal characterization of nonlinear granularities;
+\item
+  facilitate manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra;
+\item
+  develop an approach to check the feasibility of creating plots or drawing inferences for any two nonlinear granularities;
+\item
+  recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pair of granularities.
 \end{itemize}
 
-The remainder of the paper is organized as follows: Section \ref{sec:linear-time} details the background of linear granularities in depth and briefly explain calendar algebra for computing different linear granularities.
-Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
-Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
-Section \ref{sec:data-structure} discusses the data structure for exploring distribution of the associated time series across pair of cyclic time granularities.
+The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provide some background material on linear granularities and introduces calendar algebra for computing different linear granularities.
+Section \ref{sec:cyclic-gran} formally characterizes different nonlinear time granularities by extending the framework of linear time granularities.
+Section \ref{sec:cyclic-calendar} introduces calendar algebra for computing nonlinear time granularities.
+Section \ref{sec:data-structure} discusses the data structure for exploring the conditional distributions of the associated time series across pairs of nonlinear time granularities.
 Section \ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
@@
 
-Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities \citep{aigner2011visualization}.
+Discrete abstraction of time such as hours, days, weeks and months can be thought of as linear ``time granularities'' \citep{aigner2011visualization}.
 Time granularities are \textbf{linear} if they respect the linear progression of time.
-Examples include hours, days, weeks and months.
 
@@
 
-There has been several attempts to provide the framework for formally characterizing time-granularities.
-One of the first attempts occur in \citet{Bettini1998-ed} with the help of the following definitions:
+There have been several attempts to provide a framework for formally characterizing time granularities.
+One of the first attempts was provided by \citet{Bettini1998-ed} who proposed following definitions:
 
@@
 
-\begin{definition}\label{def:definition}
-(time domain): A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+\begin{definition}[Time domain]\label{def:definition}
+A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
 \end{definition}
 
-A time domain can be \textbf{discrete} (if there is unique predecessor and successor for every element except for the first and last one in the time domain), or it can be \textbf{dense} (if it is an infinite set).
-A time domain is assumed to be discrete for the purpose of our discussion.
-
-\begin{definition}\label{def:linear}
-(linear granularity): A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
+We will assume the time domain is \emph{discrete}; i.e., there is unique predecessor and successor for every element in the time domain except for the first and last.
+
+\begin{definition}[Linear granularity]\label{def:linear}
+A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
 
   (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
-than all elements of $G(j)$, and  
+than all elements of $G(j)$; and
   (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
@@
 
-\begin{definition}\label{def:granule}
-(granule) Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
+\begin{definition}[Granule]\label{def:granule}
+Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
 \end{definition}
 
-\textbf{Discussion:} The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
-\autoref{fig:linear-time} shows the implication of this condition.
-It shows the linear granularities hours, days, weeks, months and years, each of which are unidirectional in nature and arranged from past to future.
-If we consider the chronon (see \citet{aigner2011visualization}) as hourly, the time domain with T hours will have \(\lfloor T/24\rfloor\) days, \(\lfloor T/(24*7)\rfloor\) weeks and so on.
-Each granule of the linear granularities shown are represented through a box.
+The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
+\autoref{fig:linear-time} illustrates the linear time granularites hours, days, weeks, months and years, each of which are non-repeating in nature and arranged from past to future.
+If we consider the chronon \citep[see][]{aigner2011visualization} as hourly, the time domain with \(T\) hours will have \(\lfloor T/24\rfloor\) days, \(\lfloor T/(24*7)\rfloor\) weeks and so on.
 
@@
 
-\caption{The time domain distributed as linear granularities}\label{fig:linear-time}
+\caption{The time domain distributed as linear granularities.}\label{fig:linear-time}
 \end{figure}
 
-\citep{Bettini1998-ed} talks about the relationships of linear time granularities and structure of a calendar and also relates them to the notion of periodicity in time.
-
-\begin{definition}\label{def:finerthan}
-(finer than): A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
-for each index i, there exists an index j such that $G(i) \subset H(j)$.
+\citet{Bettini1998-ed} discusses the relationships between linear time granularities, the structure of a calendar, and periodicities in time.
+
+\begin{definition}[Finer than]\label{def:finerthan}
+A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$.
 \end{definition}
@@
 
-\begin{definition}\label{def:groupsinto}
-(groups into): A linear granularity G groups into a linear granularity H, denoted
-$G \trianglelefteq H$,
-if for each index j there exists a (possibly infinite) subset S of the integers such
-that
-\begin{equation}
-H(j) = \bigcup_{i \in S}G(i)
-\end{equation}
+\begin{definition}[Groups into]\label{def:groupsinto}
+A linear granularity $G$ groups into a linear granularity $H$, denoted
+$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that
+\[
+  H(j) = \bigcup_{i \in S}G(i).
+\]
 \end{definition}
 
-\textbf{Discussion:} \(G \trianglelefteq H\) if each granule H(j) is the union of some granules of G.
-\(G \preceq H\) if every granule in G is a subset of some granule in H.
-In \autoref{fig:groups-finer}, both \(day \trianglelefteq week\) and \(day \preceq week\) hold true.
+These definitions are illustrated in \autoref{fig:groups-finer}, where both day \(\trianglelefteq\) week, and day \(\preceq\) week, hold true.
 However, weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or finer than public holidays.
-Consider another example where \(day \trianglelefteq month\).
+And weekends neither groups into or is finer than public holidays.
+Consider another example where day \(\trianglelefteq\) month.
 This relationship is not fully described until we associate it with periodicity.
 Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions, .
+The period becomes 400 years with the inclusion of leap years and all their exceptions.
 
@@
 
-\begin{definition}\label{def:periodical}
-(groups periodically): A granularity H is periodical with respect to a granularity G if
-(1) $G \trianglelefteq H$, and
-(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
-that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
-$H (i + R) = \bigcup_{j \in S} G(j + P)$.
+\begin{definition}[Groups periodically]\label{def:periodical}
+A granularity $H$ is periodical with respect to a granularity $G$ if:
+(1) $G \trianglelefteq H$; and
+(2) there exist $R$, $P \in Z+$, where $R$ is less than the number of granules of $H$, such that for all $i \in Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
 \end{definition}
 
-\textbf{Discussion:} A granularity H which is periodical with respect to G is specified by:
-(i) the R sets of indexes of G, \({S_0,...,S_{R-1}}\) describing the granules of H within one period;
-(ii) the value of P;
-(iii) the indexes of first and last granules in H, if their value is not infinite.
-Then, if \({S_0,...,S_{R-1}}\) are the sets of indexes of G describing
-\(H (0), .
-.
-.
-, H (R - 1)\), respectively, then the description of an arbitrary granule H(j) is given by: \(\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)\).
+A granularity \(H\) which is periodical with respect to \(G\) is specified by:
+(i) the \(R\) sets of indexes of \(G\), \({S_0,\dots,S_{R-1}}\) describing the granules of \(H\) within one period;
+(ii) the value of \(P\);
+(iii) the indexes of first and last granules in \(H\), if their value is not infinite.
+Then, if \({S_0,\dots,S_{R-1}}\) are the sets of indexes of \(G\) describing
+\(H(0), \dots, H(R - 1)\), respectively, then the description of an arbitrary granule \(H(j)\) is given by \(\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)\).
 Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way \citep{Bettini2000-vy}.
 
-\begin{definition}\label{def:quasi}
-(groups quasi-periodically): A granularity H is quasi-periodical with respect to a granularity G if
-(1) $G \trianglelefteq H$, and
-(2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
-integers R,P, where R is less than the minimum of the number of granules of H
-between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
+\todo[inline]{Give an example}
+
+\begin{definition}[Groups quasi-periodically]\label{def:quasi}
+A granularity $H$ is quasi-periodical with respect to a granularity $G$ if:
+(1) $G \trianglelefteq H$; and
+(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
 \end{definition}
 
-\textbf{Discussion:} Intuitively, the definition requires that all granules of H within the span of time between two exceptions have the same periodical behavior, characterized by \(R\) and \(P\).
-
-\begin{definition}\label{def:B}
-(bottom granularity): Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
+Intuitively, the definition requires that all granules of \(H\) within the span of time between two exceptions have the same periodical behavior, characterized by \(R\) and \(P\).
+
+\todo[inline]{Give an example}
+
+\begin{definition}[Bottom granularity]\label{def:B}
+Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
 \end{definition}
 
-\textbf{Discussion:} Given the set of all granularities defined over the time domain \((Z; <)\), and the granularity relationship \(\trianglelefteq\) (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to \(\trianglelefteq\).
+\todo[inline]{What is g-rel?}
+
+Given the set of all granularities defined over the time domain \((Z; <)\), and the granularity relationship \(\trianglelefteq\) (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to \(\trianglelefteq\).
 
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra \citep{Ning2002-tf}.
-It is assumed that there exists a ``bottom'' granularity and calendar algebra operations are designed to generate
-new granularities from the bottom one or recursively, from those already generated.
-The calendar algebra consists of two kinds of operations: grouping-oriented and
-granule-oriented operations.
+It is assumed that there exists a ``bottom'' granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
 The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
-Example can be to consider a calendar with only two linear granularities ``minute'' and ``hour'' and ``hour'' is generated by grouping every \(60\) ``minutes''.
+For example, consider a calendar with only two linear granularities: ``minute'' and ``hour'', where ``hour'' is generated by grouping every \(60\) ``minutes''.
 The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
@@
 
-Some relevant grouping oriented operations are discussed, which will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
+Some relevant grouping-oriented operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
 
@@
 \item
-  \textbf{The grouping operation} : Let \(G1\) be a full-integer labeled granularity, and \(m\) a positive integer.
-The grouping operation \(Group_m(G)\) generates a new granularity \(G2\), by partitioning the granules of \(G\) into m-granule groups and making each group a granule of the resulting granularity.
-  More precisely, \(G2 = Group_m(G1)\) is the full-integer labeled granularity such that for
-  each integer \(i\), \(G2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G1(j)\).
+  \textbf{The grouping operation}: Let \(G_1\) be a full-integer labeled granularity, and \(m\) a positive integer.
+The grouping operation Group\(_m(G)\) generates a new granularity \(G_2\), by partitioning the granules of \(G_1\) into \(m\)-granule groups and making each group a granule of the resulting granularity.
+More precisely, \(G_2 = \text{Group}_m(G_1)\) is the full-integer labeled granularity such that for each integer \(i\),
+  \[G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).\]
 \end{itemize}
 
-\textbf{Examples} :
+\todo[inline]{What does "full-integer" mean?}
+
+For example:
+\begin{align*}
+  \text{minute} &= \text{Group}_{60}(\text{second});\\
+  \text{hour}   &= \text{Group}_{60}(\text{minute});\\
+  \text{day}    &= \text{Group}_{24}(\text{hour});\\
+  \text{week}   &= \text{Group}_{7}(\text{day});
+\end{align*}
+
+where \(\text{second}(1)\) would start a minute and likewise.
+
+\todo[inline]{What does this last phrase mean?}
 
@@
 \item
-  \(minute = Group_{60}(second)\)
-\item
-  \(hour = Group_{60}(minute)\),
-\item
-  \(day = Group_{24}(hour)\),
-\item
-  \(week = Group_{7}(day)\),
+  \textbf{The altering-tick operation} : Let \(G_1\), \(G_2\) be full-integer labeled granularities, and \(l\), \(k\), \(m\) integers, where \(G_2\) partitions \(G_1\), and \(1 \leq l \leq m\).
+The altering-tick operation \(\text{Alter}^{m}_{l,k} (G_2, G_1)\) generates a new full-integer labeled granularity by periodically expanding or shrinking granules of \(G_1\) in terms of granules of \(G_2\).
+The altering-tick operation modifies the granules of \(G_1\) so that the \(l\)th granule of each group has \(|k|\) additional (or fewer when \(k < 0\)) granules of \(G_2\).
 \end{itemize}
 
-where, second(1) would start a minute and likewise.
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{The altering-tick operation} : Let G1, G2 be full-integer labeled granularities, and \(l\), \(k\), \(m\) integers, where G2 partitions G1, and \(1 \leq l \leq m\).
-The altering-tick operation \(Alter^{m}_{l,k} (G2, G1)\) generates a new full-integer labeled granularity by periodically expanding or shrinking granules of G1 in terms of granules of G2.
-The altering-tick operation modifies the granules of G1 so that the \(l\)th granule of each group has \(|k|\) additional (or fewer when k \textless{} 0) granules of G2.
-\end{itemize}
-
-\textbf{Example} :
-
-\begin{equation*}
-\begin{split}
-pseudomonth = Alter^{12}_{11,-1}(day, Alter^{12}_{9,-1}(day, \\
-  Alter^{12}_{6,-1}(day, Alter^{12}_{4,-1)}(day, \\
-  Alter^{12}_{(2,-3)}(day, Group_{31}(day))))))
-\end{split}
-\end{equation*}
-
-, where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days.
-\citep{Ning2002-tf}
-
-For more variations of grouping operations and granule oriented operations, the readers are recommended to see \citet{Ning2002-tf}.
+For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
+\begin{align*}
+\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \\
+  &~~\text{Alter}^{12}_{9,-1}(\text{day}, \\
+  &~~~~\text{Alter}^{12}_{6,-1}(\text{day}, \\
+  &~~~~~~\text{Alter}^{12}_{4,-1)}(\text{day}, \\
+  &~~~~~~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \\
+  &~~~~~~~~~~\text{Group}_{31}(\text{day})))))),
+\end{align*}
+where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days \citep{Ning2002-tf}.
+
+\todo[inline]{Why is this called a ``pseudomonth''?}
+
+For more variations of grouping operations and granule oriented operations, see \citet{Ning2002-tf}.
 
 \hypertarget{sec:cyclic-gran}{%
-\section{Cyclic time granularities}\label{sec:cyclic-gran}}
-
-We propose a formalism of cyclic time granularities through the tsibble \citep{wang2019tsibble} framework of organizing temporal data.
+\section{Nonlinear time granularities}\label{sec:cyclic-gran}}
+
+We propose a formalism of nonlinear time granularities through the tsibble \citep{wang2019tsibble} framework of organizing temporal data.
 A tsibble consists of an index, key and measured variables.
@@
 However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
-Cyclic time granularities are referred to as \textbf{circular} if mappings between linear time granularities are regular and \textbf{quasi-circular} if the same is irregular.
+Nonlinear time granularities are referred to as \textbf{circular} if mappings between linear time granularities are regular and \textbf{quasi-circular} if the same is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 \item
-  there exist \(R', P' \epsilon Z_+\) such that if \(G'(w) = \bigcup_{z \in T_w}B(z)\) then \[G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)\]
+  there exist \(R'\), \(P' \in Z_+\) such that if \(G'(w) = \bigcup_{z \in T_w}B(z)\) then \[G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)\]
 \end{itemize}
@@
 
-Aperiodic cyclic granularities are defined using aperiodic linear granularities.
+Aperiodic nonlinear granularities are defined using aperiodic linear granularities.
 Consider n aperiodic linear granularities \(M_i\) \(\forall \{i\in{1, 2, \dots, n\}}\), that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
@@
 
-\caption{Aperiodic cyclic granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$}\label{fig:aperiodic-example}
+\caption{Aperiodic nonlinear granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$}\label{fig:aperiodic-example}
 \end{figure}
@@
 
-In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in \citet{Bettini1998-ed}.
+In \autoref{sec:cyclic-gran}, we discussed how we can define nonlinear granularities by extending the framework of linear granularities defined in \citet{Bettini1998-ed}.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
@@
 
-The cyclic calendar algebra consists of broadly two kinds of operations:\\
-(1) \textbf{single-to-multiple:} representation of multiple-order-up cyclic granularities from single order-up cyclic granularities\\
+The cyclic calendar algebra consists of broadly two kinds of operations:
+(1) \textbf{single-to-multiple:} representation of multiple-order-up cyclic granularities from single order-up cyclic granularities
 (2) \textbf{multiple-to-single:} representation of single-order-up cyclic granularities from multiple order-up cyclic granularities
@@
 
-There can be three scenarios for obtaining multiple-order-granularities here:\\
-- cyclic granularities relating two linear granularities whose orders are less than day\\
-- cyclic granularities relating two linear granularities whose orders are more than month\\
+There can be three scenarios for obtaining multiple-order-granularities here:
+- cyclic granularities relating two linear granularities whose orders are less than day
+- cyclic granularities relating two linear granularities whose orders are more than month
 - granularities relating two linear granularities with order at most day and another with order at least month
@@
 #> 10 10006414    2012-02-10 12:30:00              0.252
-#> # … with 1,450,222 more rows
+#> # ... with 1,450,222 more rows
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{smart_meter10 }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{search_gran}\NormalTok{(}\DataTypeTok{highest_unit =} \StringTok{"month"}\NormalTok{)}
@@
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-Given the data structure \textless{}\(C1\) \(C2\), \texttt{general\_supply\_kwh}\textgreater{}, each of those six cyclic granularities can either be mapped to x-axis or to facet.
+Given the data structure \textless{}\(C1\) \(C2\), \texttt{general\_supply\_kwh}\textgreater, each of those six cyclic granularities can either be mapped to x-axis or to facet.
 Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualisation.
@@
 \begin{Highlighting}[]
-\NormalTok{smart_meter10 }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(customer_id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{10017936}\NormalTok{))}\OperatorTok{%>%}\StringTok{ }
+\NormalTok{smart_meter10 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(customer_id }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{10017936}\NormalTok{))}\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{gran_advice}\NormalTok{(}\StringTok{"wknd_wday"}\NormalTok{,}
@@
 #> 10 9            908     366
-#> # … with 14 more rows
+#> # ... with 14 more rows
 \end{verbatim}

### 2020-02-13 main.Rmd
include examples of periodic and quasi-periodic

@@
 
-\todo[inline]{Give an example}
+```{r group-periodicial, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$."}
+knitr::include_graphics("Figs/groups_periodically.png")
+```
 
@@
 
-\todo[inline]{Give an example}
+```{r quasi-group-periodicial, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."}
+knitr::include_graphics("Figs/quasi_groups_periodically.png")
+```
 

### 2020-02-13 main.Rmd
example illustration of groups periodically and quasi-periodically

--- 2020-02-13 main.Rmd
@@
 
-\todo[inline]{Give an example}
+```{r quasi-group-periodicial, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."}
+knitr::include_graphics("Figs/quasi_groups_periodically.png")
+```
+
+**Example**:  Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodicial} and \autoref{fig:quasi-group-periodicial} respectively.
+Each granule is represented by a box.
+In these figures, the granules that overlap share elements from the underlying time domain.
+In both of these cases, each granule of $G$ is made up of two and three granules of $B$ alternately.
+So the period is $P = 5$ and number of granules of $G$ in each period is given by $R=2$.
+The only difference in \autoref{fig:quasi-group-periodicial} is that there is an exception after the first granule of $G$ which consists of one granule of $B$.
+Except for this exception $E1$, $B$ groups periodically into $G$.
 

### 2020-02-13 main.Rmd
all todo questions answered

@@
 
-\todo[inline]{What is g-rel?}
+\todo[inline]{What is g-rel? - g-rel can be any granularity relationships like -   
+groups into, finer than, groups periodically (the granularity relationships defined in this paper) or sub-granular, shift equivalent (not discussed in this paper but other relationships that the author discussed in @Bettini1998-ed)}
 
@@
 
-\todo[inline]{What does "full-integer" mean?}
+\todo[inline]{What does "full-integer" mean? It might be the case that we don not always
+have integers as the label sets of granularities. A full-integer labeled granularities are defined in definition 2 since labels coincide with indices. These labels can be an
+arbitrary subset of (possibly noncontiguous) integers and could be used to identify granules.
+The author has introduced the definition of "labeled granularity" in @Ning2002-tf which could be added after definition of linear granularity to clarify this.
+}
 
@@
 
-\todo[inline]{What does this last phrase mean?}
+\todo[inline]{What does this last phrase mean?
+- if  the time horizon is such that it starts from the 15th second of a day, then doing repeated operations of grouping might result in a granule of "resultant" granularity day or week, which is different from a time-stamped day or week.
+Could be phrased better or deleted.}
+
+When the granules for second and minutes are arranged, the first granule of seconds and minutes granules should that overlap share elements from the underlying time domain}
 
@@
 
-\todo[inline]{Why is this called a ``pseudomonth''?}
+\todo[inline]{Why is this called a ``pseudomonth''? - Here,  granularity pseudomonth is generated by grouping 31 days, and then shrink April, June, September and
+November by 1 day, and shrink February (2) by 3 days.
+It doesn't incorporate adjustments for leap year or seconds in which case the period for the altering operation should be 400 instead of 12.
+``pseudomonth`` computes months only in non-leap years.}
 

### 2020-02-13 main.Rmd
spell check

@@
 Nonlinear time granularities can be circular, quasi-circular or aperiodic.
-Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularty is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
 
@@
 
-```{r group-periodicial, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$."}
+```{r group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$."}
 knitr::include_graphics("Figs/groups_periodically.png")
@@
 
-```{r quasi-group-periodicial, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."}
+```{r quasi-group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."}
 knitr::include_graphics("Figs/quasi_groups_periodically.png")
@@
 
-**Example**:  Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodicial} and \autoref{fig:quasi-group-periodicial} respectively.
+**Example**:  Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodical} and \autoref{fig:quasi-group-periodical} respectively.
 Each granule is represented by a box.
@@
 So the period is $P = 5$ and number of granules of $G$ in each period is given by $R=2$.
-The only difference in \autoref{fig:quasi-group-periodicial} is that there is an exception after the first granule of $G$ which consists of one granule of $B$.
+The only difference in \autoref{fig:quasi-group-periodical} is that there is an exception after the first granule of $G$ which consists of one granule of $B$.
 Except for this exception $E1$, $B$ groups periodically into $G$.
@@
 
-\todo[inline]{What does "full-integer" mean? It might be the case that we don not always
+\todo[inline]{What does "full-integer" mean? It might be the case that we do not always
 have integers as the label sets of granularities. A full-integer labeled granularities are defined in definition 2 since labels coincide with indices. These labels can be an
-arbitrary subset of (possibly noncontiguous) integers and could be used to identify granules.
+arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
 The author has introduced the definition of "labeled granularity" in @Ning2002-tf which could be added after definition of linear granularity to clarify this.

### 2020-02-13 main.tex
re-compile after addressing Rob's comments till Section 2

@@
 
-\todo[inline]{Give an example}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/groups_periodically}
+
+}
+
+\caption{Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$.}\label{fig:group-periodical}
+\end{figure}
 
@@
 
-\todo[inline]{Give an example}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figs/quasi_groups_periodically}
+
+}
+
+\caption{Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception.}\label{fig:quasi-group-periodical}
+\end{figure}
+
+\textbf{Example}: Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodical} and \autoref{fig:quasi-group-periodical} respectively.
+Each granule is represented by a box.
+In these figures, the granules that overlap share elements from the underlying time domain.
+In both of these cases, each granule of \(G\) is made up of two and three granules of \(B\) alternately.
+So the period is \(P = 5\) and number of granules of \(G\) in each period is given by \(R=2\).
+The only difference in \autoref{fig:quasi-group-periodical} is that there is an exception after the first granule of \(G\) which consists of one granule of \(B\).
+Except for this exception \(E1\), \(B\) groups periodically into \(G\).
 
@@
 
-\todo[inline]{What does "full-integer" mean?}
+\todo[inline]{What does "full-integer" mean? It might be the case that we do not always
+have integers as the label sets of granularities. A full-integer labeled granularities are defined in definition 2 since labels coincide with indices. These labels can be an
+arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
+The author has introduced the definition of "labeled granularity" in @Ning2002-tf which could be added after definition of linear granularity to clarify this.
+}
 
@@
 
-\todo[inline]{What does this last phrase mean?}
+\todo[inline]{What does this last phrase mean?
+- if  the time horizon is such that it starts from the 15th second of a day, then doing repeated operations of grouping might result in a granule of "resultant" granularity day or week, which is different from a time-stamped day or week.
+Could be phrased better or deleted.}
+
+When the granules for second and minutes are arranged, the first granule of seconds and minutes granules should that overlap share elements from the underlying time domain\}
 
@@
 #> 10 10006414    2012-02-10 12:30:00              0.252
-#> # ... with 1,450,222 more rows
+#> # … with 1,450,222 more rows
 \end{verbatim}
@@
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-Given the data structure \textless{}\(C1\) \(C2\), \texttt{general\_supply\_kwh}\textgreater, each of those six cyclic granularities can either be mapped to x-axis or to facet.
+Given the data structure \textless{}\(C1\) \(C2\), \texttt{general\_supply\_kwh}\textgreater{}, each of those six cyclic granularities can either be mapped to x-axis or to facet.
 Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualisation.
@@
 #> 10 9            908     366
-#> # ... with 14 more rows
+#> # … with 14 more rows
 \end{verbatim}

### 2020-02-14 main.Rmd
todo additions made

@@
 
-\todo[inline]{What is g-rel? - g-rel can be any granularity relationships like -   
-groups into, finer than, groups periodically (the granularity relationships defined in this paper) or sub-granular, shift equivalent (not discussed in this paper but other relationships that the author discussed in @Bettini1998-ed)}
-
+Here, g-rel could be any relationship defined between granularities like "groups into", "finer than" or "groups periodically".
 Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
@@
 
+<!-- It might be the case that we do not always have integers as the label sets of granularities.  -->
+
+<!-- \begin{definition}[Labelled granularity]\label{def:labelgran} -->
+
+The notion of labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
+A labeled granularity is a pair $(\L, G)$, where $L$ is a subset of the integers, and $G$ is a mapping from $L$ to the subsets of the time domain such that (1) if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$, and (2) if $i<k<j$ , where $i$, $j$ , and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
+$L$ is called the label set of $G$.
+When $L$ is exactly the integers, the granularities are called full-integer labeled granularity like the one defined in Definition \ref{def:linear}.
+However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
+Some calendar operations will only work on full-integer labeled granularities, while others could be implemented using more flexible labelings.
+
 Some relevant grouping-oriented operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-
-\todo[inline]{What does "full-integer" mean? It might be the case that we do not always
-have integers as the label sets of granularities. A full-integer labeled granularities are defined in definition 2 since labels coincide with indices. These labels can be an
-arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
-The author has introduced the definition of "labeled granularity" in @Ning2002-tf which could be added after definition of linear granularity to clarify this.
-}
 
@@
 
-where $\text{second}(1)$ would start a minute and likewise.
-
-\todo[inline]{What does this last phrase mean?
-- if  the time horizon is such that it starts from the 15th second of a day, then doing repeated operations of grouping might result in a granule of "resultant" granularity day or week, which is different from a time-stamped day or week.
-Could be phrased better or deleted.}
-
-When the granules for second and minutes are arranged, the first granule of seconds and minutes granules should that overlap share elements from the underlying time domain}
+where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
+Otherwise, repeated operations of grouping might result in a granule of "resultant" granularity "day" or "week", which is different from a time-stamped "day" or "week".
 
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
+Let us consider a granularity "pseudomonth" which computes months only for non-leap years.
 \begin{align*}
@@
 where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
-
-\todo[inline]{Why is this called a ``pseudomonth''? - Here,  granularity pseudomonth is generated by grouping 31 days, and then shrink April, June, September and
-November by 1 day, and shrink February (2) by 3 days.
-It doesn't incorporate adjustments for leap year or seconds in which case the period for the altering operation should be 400 instead of 12.
-``pseudomonth`` computes months only in non-leap years.}
 

### 2020-02-14 main.Rmd
text example included (no diagram for periodic and quasi-periodic)

@@
 
-```{r group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$."}
-knitr::include_graphics("Figs/groups_periodically.png")
-```
+<!-- # ```{r group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$."} -->
+<!-- # knitr::include_graphics("Figs/groups_periodically.png") -->
+<!-- # ``` -->
 
@@
 
-```{r quasi-group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."}
-knitr::include_graphics("Figs/quasi_groups_periodically.png")
-```
-
-**Example**:  Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodical} and \autoref{fig:quasi-group-periodical} respectively.
+<!-- # ```{r quasi-group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."} -->
+<!-- # knitr::include_graphics("Figs/quasi_groups_periodically.png") -->
+<!-- # ``` -->
+
+<!-- **Example**:  Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodical} and \autoref{fig:quasi-group-periodical} respectively.
 Each granule is represented by a box.
@@
 Except for this exception $E1$, $B$ groups periodically into $G$.
+-->
+
+**Example**: Consider the Gregorian calendar without leap year and all their exceptions.
+Then each month is a grouping of the same number of days over years.
+So the period is $P = 365$ and number of granules of "months" in each period is given by $R=12$.
+Thus, we could say days groups periodically into months.
+Now consider the Gregorian calendar with leap years but none of their exceptions.
+In such a case, if we consider every February $29^{th}$ as exceptions, we will still have each month as a grouping of the same number of days over years.
+In that case, we may say days groups quasi-periodically into months.
+The value of $P$ and $R$ would be 365 and 12 respectively while $E_i$'s would be the time domain corresponding to $29^{th}$ February of any year.
 

### 2020-02-17 main.Rmd
recompile after todo
Further improvements to Section 2

@@
 (1) $G \trianglelefteq H$; and
-(2) there exist $R$, $P \in Z+$, where $R$ is less than the number of granules of $H$, such that for all $i \in Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
+(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
 \end{definition}
@@
 (1) $G \trianglelefteq H$; and
-(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
+(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
 \end{definition}
@@
 
-**Example**: Consider the Gregorian calendar without leap year and all their exceptions.
-Then each month is a grouping of the same number of days over years.
-So the period is $P = 365$ and number of granules of "months" in each period is given by $R=12$.
+Consider the Gregorian calendar without leap years.
+Then each month is a grouping of the same number of days each year.
+The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$.
 Thus, we could say days groups periodically into months.
-Now consider the Gregorian calendar with leap years but none of their exceptions.
-In such a case, if we consider every February $29^{th}$ as exceptions, we will still have each month as a grouping of the same number of days over years.
-In that case, we may say days groups quasi-periodically into months.
-The value of $P$ and $R$ would be 365 and 12 respectively while $E_i$'s would be the time domain corresponding to $29^{th}$ February of any year.
+Now consider the Gregorian calendar with leap years, so that days group quasi-periodically into months.
+The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s would be the time domain corresponding to $29^{\text{th}}$ February of any year.
 
@@
 Here, g-rel could be any relationship defined between granularities like "groups into", "finer than" or "groups periodically".
-Given the set of all granularities defined over the time domain $(Z; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
+Given the set of all granularities defined over the time domain $(\mathbb{Z}; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
 
@@
 For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes".
-The granule-oriented operations do not change the granules of a granularity, but rather make choices of which granules should remain in the new granularity.
-For example, one can choose to look at the granularity "Monday" and hence select only "Monday" from linear granularity "day".
+The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity.
+For example, one can define the granularity "Monday" and hence select only Mondays from linear granularity "day".
 
@@
 
-The notion of labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
-A labeled granularity is a pair $(\L, G)$, where $L$ is a subset of the integers, and $G$ is a mapping from $L$ to the subsets of the time domain such that (1) if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$, and (2) if $i<k<j$ , where $i$, $j$ , and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
+The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
+A labeled granularity is a pair $(L, G)$, where $L$ is a subset of the integers, and $G$ is a mapping from $L$ to the subsets of the time domain such that:
+
+  1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and
+  2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
+
 $L$ is called the label set of $G$.
-When $L$ is exactly the integers, the granularities are called full-integer labeled granularity like the one defined in Definition \ref{def:linear}.
+When $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
 However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
-Some calendar operations will only work on full-integer labeled granularities, while others could be implemented using more flexible labelings.
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.
 
@@
 
-- **The grouping operation**: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
+The grouping operation:
+: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
 The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-
 For example:
@@
 \end{align*}
-
 where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
@@
 
-- **The altering-tick operation** : Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
+The altering-tick operation:
+: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
 The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$.
-The altering-tick operation modifies the granules of $G_1$ so that the $l$th granule of each group has $|k|$ additional (or fewer when $k < 0$) granules of $G_2$.
+The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$.
 

### 2020-02-17 main.Rmd
Work on sec 3

@@
   \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
+  \def\mod{~\text{mod}~}
 output:
@@
 A tsibble consists of an index, key and measured variables.
-An index is a variable with inherent ordering from past to present and key is a set of variables that define observational units over time.
-In a tsibble each observation should be uniquely identified by index and key.
+An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
+In a tsibble, each observation (row) should be uniquely identified by index and key.
 
@@
 A bottom granularity is represented by the index of the tsibble.
-Lining up with the the assumption in @Bettini2000-vy, that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists.
-In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be preceded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday).
-Hence intuitively, nonlinear granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
-
-Nonlinear time granularities which accommodate for periodicities can be constructed by relating two linear granularities.
+Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.
+
+Nonlinear time granularities with periodicities can be constructed by relating two linear granularities.
 The mappings between these linear granularities can be regular or irregular.
-For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour.
-On contrary, an irregular mapping exists between days and months, since one might have days ranging from 28 to 31.
-To elaborate more, mappings will be regular if, for example, linear granularities are formed by grouping operation.
-Grouping operations combine fixed granules of a linear granularity to form a new granule of the resulting linear granularity.
-However, if granularities are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of the root granularity.
-Nonlinear time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the same is irregular.
+For example, a regular mapping exists between minutes and hours (where 60 minutes always add up to 1 hour), while an irregular mapping exists between days and months (as the number of days per month ranges from 28 to 31).
+Mappings will be regular if linear granularities are formed by a grouping operation, but granularities formed by the altering-tick operation will be irregular.
+Nonlinear time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the mapping is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
-The formalism would differ for circular and quasi-circular granularities although both accommodate for periodicities in data.
 
@@
 
-\begin{definition}\label{def:circular}
-(circular granularity): A circular granularity $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
-
+\begin{definition}[Circular granularity]\label{def:circular}
+A circular granularity $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
 \begin{equation} \label{eq:circular-gran}
@@
 and ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$.
-
 \end{definition}
 
-**Example**:  Example showing circular granularity relating two linear granularities "day" and "week" is visually depicted in a series of slots in the \autoref{fig:circular-dow}.
-Each granule is represented by a box.
-The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set.
-The bottom granularity, represented by the index of the tsibble considered is "day".
-"day" and "week" are two linear granularities with $P =  7$.
-The circular granularity $C_{day, week}$ representing day-of-week will thus consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length.
-
-**Discussion**: Note that each circular granularity can use different label mappings.
-In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity.
-However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity.
-Hence, the label mapping $L$ in \autoref{fig:circular-dow} can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{Sunday, Monday, \dots, Saturday\}$ or $L: \{0,1,2, \dots, 6\} \longmapsto\ \{0, 1, \dots, 6\}$ depending on the context.
+\autoref{fig:circular-dow} depicts the circular granularity "day of the week", constructed from two linear granularities "day" (the bottom granularity) and "week".
+The granules that overlap share elements from the underlying time domain.
+In this example, $P=7$, and the circular granularity $C_{\text{day}, \text{week}}$ representing day-of-week consists of the repeated pattern $\{L(0), L(1), L(2),$ $L(3), L(4), L(5), L(6)\}$.
+
+Each circular granularity can use descriptive label mappings, or be simply labelled using the index. The label mapping $L$ in \autoref{fig:circular-dow} can be defined as
+$$
+  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sun}, \text{Mon}, \dots, \text{Sat}\}
+$$
+or
+$$
+  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}
+$$
+for example.
 
@@
 
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity H is periodic with respect to linear granularity G with regular mapping such that the number of granules of G in each granule of H is $P(G,H)$.
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
 Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
@@
 - $B \trianglelefteq G'$
-- there exist $R'$, $P' \in Z_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
+- there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
 
@@
 The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
-The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R')^{th}$ granule of $G'$ will have the same number of granules of $B$.
-The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$.
-Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.
+The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$.
+The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$.
+Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.
 
@@
 and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
-Here, index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity.
+Here, index ${\{i_j\}}$ stands for the $j^{\text{th}}$ granule of the $i^{\text{th}}$ linear aperiodic granularity.
 
@@
 Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{th}$ and $(n + 1 - ([n/2] + 1)/2)^{th}$ order statistics are taken as estimates of the quartiles.
+These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.
 Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.

### 2020-02-17 main.Rmd
Query on Section 3

@@
 
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
+In general, any circular granularity relating two linear granularity, none of which are bottom granularity\todo{Why not?} can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
 Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.

### 2020-02-17 data-raw/sm_cust50.R
cosde for creating sm_cust50

@@ -1,0 +1,35 @@
+library(tidyverse)
+library(tsibble)
+library(gravitas)
+library(readr)
+
+
+# Data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in ["https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter"].
+
+download.file("http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z", destfile ="data-raw/sm_raw.csv")
+
+smart_meter_data_raw <- read_csv("data-raw/sm_raw.csv", n_max = 3e6)
+
+sm_cust <- smart_meter_data_raw %>% 
+  distinct(CUSTOMER_ID) %>%
+  .$CUSTOMER_ID %>% 
+  sample(size= 50)
+
+smart_meter_data <- smart_meter_data_raw %>%
+  dplyr::filter(CUSTOMER_ID %in% sm_cust) %>%
+  dplyr::rename_all(tolower) %>%
+  dplyr::arrange(customer_id, reading_datetime) %>%
+  dplyr::group_by(customer_id) %>%
+  dplyr::mutate(reading_datetime = case_when(
+    duplicated(reading_datetime) ~ reading_datetime + lubridate::hours(1),
+    TRUE ~ reading_datetime
+  ))
+
+smart_meter_data$customer_id <- as.character(smart_meter_data$customer_id)
+
+sm_cust50 <- smart_meter_data %>%
+  as_tsibble(index = reading_datetime, key = customer_id) %>%
+  ungroup() %>%
+  dplyr::select(-calendar_key)
+
+save(sm_cust50, file = "data/sm_cust50.RData", compress = "xz")

### 2020-02-17 data-raw/sm_cust50.R
rds format

@@ -32,4 +32,4 @@ sm_cust50 <- smart_meter_data %>%
   ungroup() %>%
   dplyr::select(-calendar_key)
 
-save(sm_cust50, file = "data/sm_cust50.RData", compress = "xz")
+save(sm_cust50, file = "data/sm_cust50.rds", compress = "xz")

### 2020-02-17 main.Rmd
uinal, baktun spacing corrected

@@
 \begin{split}
-C_{uinal, baktun}(z) & = L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
-              & = L(\lfloor z/20\rfloor \mod 18  +
-              20*\lfloor z/20*18\rfloor \mod 20 +
-              20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
+C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
+               = & L(\lfloor z/20\rfloor \mod 18  \\
+               &~~~ + 20*\lfloor z/20*18\rfloor \mod 20  \\
+               &~~~ + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
 \end{split}
@@
 ```{r smart-meter, fig.cap = "Smart meter data structure"}
-sm <-
-  read_rds("data/sm_cust50.rds")
+sm <- read_rds("data/sm_cust50.rds")
 

### 2020-02-18 data-raw/sm_cust50.R
set seed for reproducibility

@@ -3,13 +3,16 @@ library(tsibble)
 library(gravitas)
 library(readr)
 
+# steps for obtaining raw data
+# Raw data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter 
+# file name:CD_INTERVAL_READING_ALL_NO_QUOTES.csv
+# Download the file using the link: http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z
+# save it in data-raw folder and run this code to obtain sm_cust50.rds
 
-# Data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in ["https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter"].
+smart_meter_data_raw <- read_csv("data-raw/CD_INTERVAL_READING_ALL_NO_QUOTES.csv", n_max = 3e6)
 
-download.file("http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z", destfile ="data-raw/sm_raw.csv")
 
-smart_meter_data_raw <- read_csv("data-raw/sm_raw.csv", n_max = 3e6)
-
+set.seed(12345)
 sm_cust <- smart_meter_data_raw %>% 
   distinct(CUSTOMER_ID) %>%
   .$CUSTOMER_ID %>% 

### 2020-02-18 data-raw/sm_cust50.R
Scripted download of smart metre data

@@ -1,38 +1,41 @@
 library(tidyverse)
 library(tsibble)
-library(gravitas)
-library(readr)
+#devtools::install_github("jimhester/archive")
+library(archive)
 
-# steps for obtaining raw data
-# Raw data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter 
+# Raw data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in
+# https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter
 # file name:CD_INTERVAL_READING_ALL_NO_QUOTES.csv
-# Download the file using the link: http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z
-# save it in data-raw folder and run this code to obtain sm_cust50.rds
 
-smart_meter_data_raw <- read_csv("data-raw/CD_INTERVAL_READING_ALL_NO_QUOTES.csv", n_max = 3e6)
+download.file(
+  "http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z",
+  "data-raw/smart-metre.7z",
+  mode = "wb"
+)
+smart_meter_data_raw <- archive_read("data-raw/smart-metre.7z") %>%
+  read_csv(n_max = 3e6)
 
+#smart_meter_data_raw <- read_csv("data-raw/CD_INTERVAL_READING_ALL_NO_QUOTES.csv", n_max = 3e6)
 
 set.seed(12345)
-sm_cust <- smart_meter_data_raw %>% 
+sm_cust <- smart_meter_data_raw %>%
   distinct(CUSTOMER_ID) %>%
-  .$CUSTOMER_ID %>% 
+  pull(CUSTOMER_ID) %>%
   sample(size= 50)
 
-smart_meter_data <- smart_meter_data_raw %>%
-  dplyr::filter(CUSTOMER_ID %in% sm_cust) %>%
-  dplyr::rename_all(tolower) %>%
[…]

### 2020-02-24 main.Rmd
content cut iteration 2
 - def bottom gran
 + def calendar

@@
 
-\begin{definition}[Finer than]\label{def:finerthan}
-A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$.
-\end{definition}.
+<!-- CONTENT CUT -->
+<!-- \begin{definition}[Finer than]\label{def:finerthan} -->
+<!-- A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$. -->
+<!-- \end{definition}. -->
 
@@
 
-These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true.
+<!-- CONTENT CUT -->
+<!-- These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true.
 However, weekends are finer than weeks but do not group into weeks.
@@
 The period becomes 400 years with the inclusion of leap years and all their exceptions.
-
-```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships.
+-->
+
+For example, day $\trianglelefteq$ week since every granule of "week" is the union of some set of granules of "day".
+Consider another example where day $\trianglelefteq$ month.
+This relationship however is incomplete without its association to periodicity.
+Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
+The period becomes 400 years with the inclusion of leap years and all their exceptions.
+
+<!-- CONTENT CUT -->
+<!-- # ```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships.
 Weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or finer than public holidays"}
-knitr::include_graphics("Figs/groupsandfiner.png")
-```
+And weekends neither groups into or finer than public holidays"} -->
+<!-- knitr::include_graphics("Figs/groupsandfiner.png") -->
+<!-- ``` -->
 
@@
 
-A granularity $H$ which is periodical with respect to $G$ is specified by:
-(i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period;
-(ii) the value of $P$;
-(iii) the indexes of first and last granules in $H$, if their value is not infinite.
-Then, if ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
+<!-- CONTENT CUT -->
+<!-- A granularity $H$ which is periodical with respect to $G$ is specified by: -->
+<!-- (i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period; -->
+<!-- (ii) the value of $P$; -->
+<!-- (iii) the indexes of first and last granules in $H$, if their value is not infinite. -->
+
+If ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
 $H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
 Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
+These are called quasi-periodic relationships.
+Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
 
@@
 
-\begin{definition}[Groups quasi-periodically]\label{def:quasi}
-A granularity $H$ is quasi-periodical with respect to a granularity $G$ if:
-(1) $G \trianglelefteq H$; and
-(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
-\end{definition}
-
-Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
+<!-- CONTENT CUT -->
+<!-- \begin{definition}[Groups quasi-periodically]\label{def:quasi} -->
+<!-- A granularity $H$ is quasi-periodical with respect to a granularity $G$ if: -->
+<!-- (1) $G \trianglelefteq H$; and -->
+<!-- (2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
+-->
+<!-- \end{definition} -->
 
@@
 
-Consider the Gregorian calendar without leap years.
-Then each month is a grouping of the same number of days each year.
+For example, in a Gregorian calendar without leap years each month is a grouping of the same number of days each year.
 The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$.
-Thus, we could say days groups periodically into months.
-Now consider the Gregorian calendar with leap years, so that days group quasi-periodically into months.
-The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s would be the time domain corresponding to $29^{\text{th}}$ February of any year.
-
-\begin{definition}[Bottom granularity]\label{def:B}
-Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
-\end{definition}
-
-Here, g-rel could be any relationship defined between granularities like "groups into", "finer than" or "groups periodically".
+Thus, we could say days group periodically into months.
+In a Gregorian calendar with leap years, days group quasi-periodically into months.
+The value of $P$ and $R$ would still be 365 and 12 respectively, while the <!--$E_i$s--> exceptions would be the time domain corresponding to $29^{\text{th}}$ February of any year.
+
+<!-- CONTENT CUT -->
+<!-- \begin{definition}[Bottom granularity]\label{def:B} -->
+<!-- Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
+-->
+<!-- \end{definition} -->
+
+<!-- Here, g-rel could be any relationship defined between granularities as defined in @Bettini1998-ed.
 Given the set of all granularities defined over the time domain $(\mathbb{Z}; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
-
-<!-- Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
-The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
--->
+-->
+
+A discrete time model often uses a fixed smallest granularity also known as _*bottom granularity_* (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _*chronons*_ (@Jensen1998-qn).
+Chronons may be grouped into granules and a linear granularity is a non-overlapping mapping of
+granules to subsets of the time domain (see @Dyreson2000-mr). A system of multiple granularities in lattice structures is referred to as a calendar.
 
@@
 It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
-The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
+
+<!-- CONTENT CUT -->
+<!-- The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
 The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
 For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes".
-The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity.
+-->
+<!-- The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity.
 For example, one can define the granularity "Monday" and hence select only Mondays from linear granularity "day".
-
-<!-- It might be the case that we do not always have integers as the label sets of granularities.  -->
+-->
 
@@
 
-The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
-A labeled granularity is a pair $(L, G)$, where $L$ is a subset of the integers, and $G$ is a mapping from $L$ to the subsets of the time domain such that:
-
-  1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and
-  2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
-
-$L$ is called the label set of $G$.
-When $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
+In addition to the integer index, each granule could be associated with a textual representation which is called the label.
+The label can be mapped to index for input and the index can later be mapped to this label for output.
+When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
 Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.
 
-Some relevant grouping-oriented operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
+<!-- CONTENT CUT -->
+<!-- The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
+A labeled granularity is a pair $(L, G)$, where the label set $L$ is a subset of the integers instead of the set of integers and -->
+<!-- follows the conditions (1) and (2) in \ref{def:linear} where $i$, $j$, and $k$ are in $L$.
+When the label set $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->  -->
+
+<!-- $G$ is a mapping from $L$ to the subsets of the time domain such that: -->
+
+<!--   1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and -->
+<!--   2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty. -->
+
+Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
 
@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-For example:
-\begin{align*}
-  \text{minute} &= \text{Group}_{60}(\text{second});\\
-  \text{hour}   &= \text{Group}_{60}(\text{minute});\\
-  \text{day}    &= \text{Group}_{24}(\text{hour});\\
-  \text{week}   &= \text{Group}_{7}(\text{day});
-\end{align*}
-where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
+Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$
+
+<!-- CONTENT CUT -->
+<!-- where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
 Otherwise, repeated operations of grouping might result in a granule of "resultant" granularity "day" or "week", which is different from a time-stamped "day" or "week".
+-->
 
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-Let us consider a granularity "pseudomonth" which computes months only for non-leap years.
+Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
+
 \begin{align*}
-\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \\
-  &~~\text{Alter}^{12}_{9,-1}(\text{day}, \\
-  &~~~~\text{Alter}^{12}_{6,-1}(\text{day}, \\
-  &~~~~~~\text{Alter}^{12}_{4,-1)}(\text{day}, \\
-  &~~~~~~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \\
-  &~~~~~~~~~~\text{Group}_{31}(\text{day})))))),
+\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \text{Alter}^{12}_{9,-1}(\text{day}, \\
+  &~~\text{Alter}^{12}_{6,-1}(\text{day}, \text{Alter}^{12}_{4,-1)}(\text{day}, \\
+  &~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \text{Group}_{31}(\text{day})))))),
 \end{align*}
-where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
 
@@
 
-In general, any circular granularity relating two linear granularity, none of which are bottom granularity\todo{Why not?} can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
+In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
 Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
@@
 These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ maps index set to day-of-month and $C_2$ maps index set to week-of-month.
+Let us take a specific example, where $C_1$ be day-of-month and $C_2$ be week-of-month.
 Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories.
@@
 Here $C_1$ can have 7 levels while $C_2$ can have 2 levels.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of day-of-week and WorkingDay/NonWorkingDay.
+So there are 14 sets $S_{ij}$ corresponding to the possible combinations of $C_1$ and $C_2$.
 While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations.
-For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
+For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
 
@@
 Thirdly, empty combinations can be a result of how granularities are constructed.
-Let $C_1$ maps index set to Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
+Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
 Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
@@
 These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
-Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
+Several forms of  non-parametric density estimators <!--and quantile estimators--> exist in the literature, the most commonly used form in @Silverman1986-cl<!--,  @Jones1992-pj)--> is as follows:
 
@@
 
-where $h$ is the smoothing parameter or the bandwidth. The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by:
-
-<!-- $Q$, $(q \equiv  Q')$, which is the quantile density function -->
-<!-- $ Q_{n} (u) = \sum_{i=1}^ n X_{(i)} I((i-1)/n <u \leq i/n)$\\ -->
-<!--             $= X_(1) + \sum_{i=1}^ n (X_{(i)} - X_{(i-1)}) I((i-1)/n < u)$ -->
-
-<!-- and most widely used definition of sample quantile is given in  by: -->
-
-<!-- $$\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}$$ where,  $(j - m)/n \leq p < (j - m + 1)/n$,  -->
-<!-- for some $m \in R$ and $0 < \gamma < 1$.  $\hat{Qi(p)}$ is the ith sample quantile  -->
-
-\begin{equation}
-\hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u)
-\end{equation}
-
-where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$.
-
- <!-- Both pdf and pmf are concerned with point probabilities of the random variable.
-Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d.
--->
-
-<!-- Instead of assuming a parametric model for the distribution (e.g.
-a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g.
-differentiable).
-It is then possible to estimate the u2nknown density function f(·).
--->
-
-<!-- The histogram is the oldest and most popular density estimator method.
-The form of the histogram depends very much on the tuning parameters class width and origin.
-The naive estimator was introduced by Fix and Hodges (1951) and first discussed by Rosenblatt(1956).
-Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962).
-The nearest neighbor estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977).
-There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.
--->
-
-<!-- The inverse $Q = F^{-1}$, which is the quantile function can be defined as $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$.  -->
-
-The MSE of the estimators are given as follows:
+where $h$ is the smoothing parameter or the bandwidth. The MSE of the estimator is given by:
 
@@
 MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
-MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
 \end{split}
@@
 where $\kappa = \int k^2(y)dy$.
+
+<!-- CONTENT CUT -->
+<!-- The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by: -->
+
+<!-- \begin{equation} -->
+<!-- \hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u) -->
+<!-- \end{equation} -->
+
+<!-- where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$. -->
+<!-- MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa -->
 
@@
 In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
-Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
+We will discuss few conventional and recent ways to plot distributions using both of these methods.
+
+<!-- CONTENT CUT -->
+<!-- Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
 The descriptive statistics based methods do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
@@
 Also, the summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.
-
-<!--  However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. -->
-
-We will discuss few conventional and recent ways to plot distributions using both of these methods.
-
-<!-- Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it.
-These displays depend on the specific estimation procedure (e.g., kernel density estimate)  -->
+-->
 
@@
 
-Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
-These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.
-Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
+Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.--> Due to wide spread popularity and simplicity in implementation, a number of variations are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.
 Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
 However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
-Also, the number of outliers is large for larger data set since number of outliers is proportional to the number of observations.
+Also, the number of outliers is large for larger data sets.
 
 <!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
-These are particularly useful for large data sets.
+The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.
 “outliers” in letter value plots are those observations beyond the most extreme letter value.
@@
 Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
+It becomes less useful when there are too many levels or categories to display.
 
@@
   filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>%
-  prob_plot("month_year","day_week",
+  prob_plot("month_year", "day_week",
     response = "Demand",
-            plot_type = "lv") +
+    plot_type = "lv"
+  ) +
   ylab("Electricity Demand [KWh]") +
@@
   filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
-  prob_plot("day_week","hour_day",
+  prob_plot("day_week", "hour_day",
     response = "Demand",
     plot_type = "quantile",
-            symmetric = FALSE) +
+    symmetric = FALSE
+  ) +
   ylab("Electricity Demand [KWh]") +
@@
     response = "Demand",
-                plot_type = "boxplot")+
+    plot_type = "boxplot"
+  ) +
   ylab("Electricity Demand [KWh]") +
   scale_x_discrete(breaks = seq(0, 366, 60)) +
-  theme(legend.position = "bottom",
-        strip.text = element_text(size = 7, margin = margin())) +
+  theme(
+    legend.position = "bottom",
+    strip.text = element_text(size = 7, margin = margin())
+  ) +
   ggtitle("(c) Box plot")
@@
     response = "Demand",
-                plot_type = "violin", alpha = 0.3) + ylab("") +
+    plot_type = "violin", alpha = 0.3
+  ) + ylab("") +
   xlab("Days of the Month") +
@@
 
-VIC %>%  create_gran("week_month") %>%
+VIC %>%
+  create_gran("week_month") %>%
   create_gran("week_year") %>%
-  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>%
+  filter(week_month %in% c(1, 2, 5), week_year %in% c(1:20)) %>%
   prob_plot("week_year",
@@
     response = "Demand",
-            plot_type = "ridge")  +
+    plot_type = "ridge"
+  ) +
   xlab("Electricity Demand [KWh]") +
@@
   scale_x_continuous(breaks = seq(2000, 10000, 3000)) +
-  theme(legend.position = "bottom",
-        strip.text = element_text(size = 7, margin = margin())) +
+  theme(
+    legend.position = "bottom",
+    strip.text = element_text(size = 7, margin = margin())
+  ) +
   ggtitle("(e) Ridge plot")
@@
     quantile_prob = seq(0.1, 0.9, 0.1),
-            symmetric = FALSE) +
-   #scale_x_discrete(breaks = seq(1, 336, 60)) +
+    symmetric = FALSE
+  ) +
+  # scale_x_discrete(breaks = seq(1, 336, 60)) +
   ylab("") +
-   xlab("Day of the Year") + theme(legend.position = "none",
-                                   strip.text = element_text(size = 7,
-                                                             margin = margin())) +
+  xlab("Day of the Year") + theme(
+    legend.position = "none",
+    strip.text = element_text(
+      size = 7,
+      margin = margin()
+    )
+  ) +
   ggtitle("(f) Decile plot")
@@
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
 It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
@@
 ```{r smart-meter, fig.cap = "Smart meter data structure"}
-sm <- read_rds("data/sm_cust50.rds")
-
-smart_meter <- sm %>% as_tsibble() %>% dplyr::select(customer_id,
+load("data/sm_cust50.Rdata")
+
+smart_meter <- sm_cust50 %>%
+  as_tsibble() %>%
+  dplyr::select(
+    customer_id,
     reading_datetime,
-                              general_supply_kwh)
-smart_meter
+    general_supply_kwh
+  )
+smart_meter %>% head(n = 5)
 ```
@@
 ```{r search_gran_limit2, echo=TRUE}
-smart_meter10 %>% search_gran(highest_unit = "month",
+smart_meter10 %>% search_gran(
+  highest_unit = "month",
   filter_out = c("hhour", "fortnight")
@@
 smart_meter10 %>%
-  filter(customer_id %in% c(10017936))%>%
-  gran_advice("wknd_wday",
-                            "hour_day")
+  filter(customer_id %in% c(10017936)) %>%
+  gran_advice(
+    "wknd_wday",
+    "hour_day"
+  )
 ```
@@
 # cust1 <- smart_meter10 %>%
-#filter(customer_id %in% c(10006704)) %>%
+# filter(customer_id %in% c(10006704)) %>%
 #   prob_plot("wknd_wday",
@@
 cust2 <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936))%>%
+  filter(customer_id %in% c(10017936)) %>%
   prob_plot("wknd_wday",
@@
     symmetric = TRUE,
-             quantile_prob = c(0.01,0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
+    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
   scale_y_sqrt()
@@
 smart_meter10 %>%
-  filter(customer_id %in% c(10017936))%>%
+  filter(customer_id %in% c(10017936)) %>%
   prob_plot("wknd_wday",
@@
     response = "general_supply_kwh",
-                            plot_type = "violin") +
+    plot_type = "violin"
+  ) +
   scale_y_sqrt()
@@
 smart_meter10 %>%
-  filter(customer_id %in% c(10017936))%>%
+  filter(customer_id %in% c(10017936)) %>%
   prob_plot("hour_day",
@@
     response = "general_supply_kwh",
-            plot_type = "boxplot") +
+    plot_type = "boxplot"
+  ) +
   scale_y_sqrt()
 # scale_fill_brewer(palette = "Dark2")
-
 ```
@@
   G = c("over", "inning", "match", "season"),
-   C = c("over-of-inning", "inning-of-match", "match-of-season",1),
-   P = c(20, 2, "aperiodic", 1))
+  C = c("over-of-inning", "inning-of-match", "match-of-season", 1),
+  P = c(20, 2, "aperiodic", 1)
+)
 knitr::kable(hierarchy_model, format = "markdown")
@@
 cricket_tsibble %>%
-  filter(batting_team %in% c("Mumbai Indians",
-                             "Chennai Super Kings"))%>%
+  filter(batting_team %in% c(
+    "Mumbai Indians",
+    "Chennai Super Kings"
+  )) %>%
   prob_plot("inning",
@@
     hierarchy_model,
-  plot_type = "lv")  + ggtitle("") +
+    plot_type = "lv"
+  ) + ggtitle("") +
   scale_fill_brewer(palette = "Dark2")
-
 ```
@@
 
- cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
+cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" = "id"))
 
@@
 cricket_dot_field <- cricket_season %>%
-  mutate(fielding_proxy = if_else(dismissal_kind %in%
-                                    c("caught", "caught and bowled"),1,0),
-         dot_ball_proxy = if_else(total_runs==0,1,0),
-         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>%
-  group_by(season,
+  mutate(
+    fielding_proxy = if_else(dismissal_kind %in%
+      c("caught", "caught and bowled"), 1, 0),
+    dot_ball_proxy = if_else(total_runs == 0, 1, 0),
+    wicket_proxy = if_else(is.na(dismissal_kind), 0, 1)
+  ) %>%
+  group_by(
+    season,
     match_id,
@@
     inning,
-           over) %>%
-  summarise(runs_per_over = sum(total_runs),
-            run_rate = sum(total_runs)/length(total_runs),
+    over
+  ) %>%
+  summarise(
+    runs_per_over = sum(total_runs),
+    run_rate = sum(total_runs) / length(total_runs),
     fielding_wckts = sum(fielding_proxy),
-            dot_balls = sum(dot_ball_proxy)) %>%
-            mutate(diff_run_rate = c(0,diff(run_rate)))
+    dot_balls = sum(dot_ball_proxy)
+  ) %>%
+  mutate(diff_run_rate = c(0, diff(run_rate)))
 
@@
 
-cricket_data<- cricket_tsibble %>%
-  mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets"),
-         dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls"),
+cricket_data <- cricket_tsibble %>%
+  mutate(
+    field = if_else(fielding_wckts == 0, "no wickets", ">0 wickets"),
+    dot = if_else(dot_balls == 0, "no dot balls", ">0 dot balls"),
     lag_field = lag(field),
-         lag_dot = lag(dot)) %>% filter(lag_field!=0, lag_dot!=0)
-```
-
-```{r echo = TRUE}
+    lag_dot = lag(dot)
+  ) %>%
+  filter(lag_field != 0, lag_dot != 0)
+```
+
+```{r echo = FALSE, eval = FALSE}
 cricket_data %>% gran_obs("field", "over")
@@
 
-  #filter(fielding_wckts %in% c(0,1)) %>%
-  cricket_data %>% prob_plot("over","lag_field",
+# filter(fielding_wckts %in% c(0,1)) %>%
+cricket_data %>% prob_plot("over", "lag_field",
   hierarchy_model,
@@
   plot_type = "violin",
-                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") + geom_boxplot(width = 0.5)
+  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
+) + ggtitle("") + geom_boxplot(width = 0.5)
 ```
@@
 cricket_data %>%
-  #filter(dot_balls %in% c(0, 1, 2)) %>%
+  # filter(dot_balls %in% c(0, 1, 2)) %>%
   prob_plot("lag_dot",
@@
     quantile_prob = c(0.25, 0.5, 0.75),
-            symmetric = FALSE)  + ggtitle("")
+    symmetric = FALSE
+  ) + ggtitle("")
 ```
@@
 ```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr","rlang", "gravitas"), file = "rpkgs.bib")
-```
+write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas"), file = "rpkgs.bib")
+```

### 2020-02-24 main.Rmd
minor changes in Section 2 for content cut

@@
 ```
+In addition to the integer index, each granule could be associated with a textual representation which is called the label.
+The label can be mapped to index for input and the index can later be mapped to this label for output.
+When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
 
@@
 
-A discrete time model often uses a fixed smallest granularity also known as _*bottom granularity_* (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _*chronons*_ (@Jensen1998-qn).
-Chronons may be grouped into granules and a linear granularity is a non-overlapping mapping of
-granules to subsets of the time domain (see @Dyreson2000-mr). A system of multiple granularities in lattice structures is referred to as a calendar.
+A discrete time model often uses a fixed smallest granularity also known as *_bottom granularity_* (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called *_chronons_* (@Jensen1998-qn).
+Chronons may be grouped into granules and a non-overlapping mapping of granules to subsets of the time domain is a linear granularity.
+A system of multiple granularities in lattice structures is referred to as a *_calendar_* (see @Dyreson2000-mr).
 
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+It is assumed that there exists a *_bottom granularity_* and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic granularities.
 
@@
 
-In addition to the integer index, each granule could be associated with a textual representation which is called the label.
-The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.
-
 <!-- CONTENT CUT -->
@@
 <!-- follows the conditions (1) and (2) in \ref{def:linear} where $i$, $j$, and $k$ are in $L$.
-When the label set $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->  -->
+When the label set $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->
 
@@
 <!--   2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty. -->
-
-Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
 
@@
 
-For more variations of grouping operations and granule oriented operations, see @Ning2002-tf.
+For more variations of calendar algrabra operations, see @Ning2002-tf.
 

### 2020-02-24 main.Rmd
content cut - section 2 - iteration 2

@@
 
-@Bettini1998-ed discusses the relationships between linear time granularities, the structure of a calendar, and periodicities in time.
+@Bettini1998-ed discusses the relationships between linear time granularities, periodicities in time, and the structure of a calendar.
 
@@
 
-A discrete time model often uses a fixed smallest granularity also known as *_bottom granularity_* (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called *_chronons_* (@Jensen1998-qn).
+A discrete time model often uses a fixed smallest granularity also known as _bottom granularity_ (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
 Chronons may be grouped into granules and a non-overlapping mapping of granules to subsets of the time domain is a linear granularity.
-A system of multiple granularities in lattice structures is referred to as a *_calendar_* (see @Dyreson2000-mr).
+A system of multiple granularities in lattice structures is referred to as a _calendar_ (see @Dyreson2000-mr).
 

### 2020-02-24 main.Rmd
content cut - Section 3 - iteration 1

@@
 preamble: >
+  \setlength {\marginparwidth }{2cm}
   \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
@@
 An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-In a tsibble, each observation (row) should be uniquely identified by index and key.
+In a tsibble, each observation (row) is uniquely identified by index and key.
 
@@
 Nonlinear time granularities with periodicities can be constructed by relating two linear granularities.
-The mappings between these linear granularities can be regular or irregular.
-For example, a regular mapping exists between minutes and hours (where 60 minutes always add up to 1 hour), while an irregular mapping exists between days and months (as the number of days per month ranges from 28 to 31).
-Mappings will be regular if linear granularities are formed by a grouping operation, but granularities formed by the altering-tick operation will be irregular.
-Nonlinear time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the mapping is irregular.
+The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months).
+<!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).--> A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.
+Nonlinear time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 
-\autoref{fig:circular-dow} depicts the circular granularity "day of the week", constructed from two linear granularities "day" (the bottom granularity) and "week".
+\autoref{fig:circular-dow} depicts the circular granularity "Day-of-Week", constructed from linear granularities "day" (the bottom granularity) and "week".
 The granules that overlap share elements from the underlying time domain.
@@
 Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
-It is possible that none of the two linear granularities are bottom granularities.
-But the representation of the resultant circular granularity will be a function of the index set.
-The bottom granularity is assumed to be minutes.
+<!--The representation of the resultant circular granularity will be a function of the index set.--> The bottom granularity is assumed to be minutes in this case.
 
@@
 Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\
-  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
-  & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
-  & = L(8 - 2*\vert T_{0}\vert - \vert T_{1}\vert) \\
-  & = L(8 - 2*2 - 3) \\
   & = L(1) \\
@@
 
+<!-- CONTENT CUT -->
+  <!-- & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\ -->
+  <!-- & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\ -->
+  <!-- & = L(8 - 2*\vert T_{0}\vert - \vert T_{1}\vert) \\ -->
+  <!-- & = L(8 - 2*2 - 3) \\ -->
+
+  <!-- & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\ -->
+  <!-- & = L(10 - \vert T_{0 \mod 2}\vert) \\ -->
+  <!-- & = L(10 - \vert T_{0}\vert) \\ -->
+  <!-- & = L(10 - 7) \\ -->
+
 \begin{equation} \label{eq:eq5}
@@
 Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,\quad since \quad  10 \in T_{1}  \\
-  & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
-  & = L(10 - \vert T_{0 \mod 2}\vert) \\
-  & = L(10 - \vert T_{0}\vert) \\
-  & = L(10 - 7) \\
   & = L(3) \\
@@
 
-**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then the following holds true:
-
-- $B \trianglelefteq G'$
-- there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
-
-from definition \ref{def:periodical}
-
-and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping)
+**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})
+
+<!-- - $B \trianglelefteq G'$ -->
+
+-
+
+<!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->
 
@@
 
-It can be noted here that if a linear granularity $G'$ is quasi-periodic (Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
+It can be noted here that if a linear granularity $G'$ is quasi-periodic (Bettini2000-vy) <!--(Definition \autoref{def:quasi})--> with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$.<!--(by Definition \autoref{def:quasi}).-->
 
@@
 for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
-
 ## Aperiodic {#sec:aperiodic-gran-def}

### 2020-02-24 main.Rmd
content cut - cyclic calendar algebra - iteration 1

@@
 
-**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})
-
-<!-- - $B \trianglelefteq G'$ -->
-
--
-
-<!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->
-
+**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
 Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
@@
 for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
+
 ## Aperiodic {#sec:aperiodic-gran-def}
@@
 Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
-Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that
-\begin{equation*}
-M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)
-\end{equation*}
-and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
+Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
 Here, index ${\{i_j\}}$ stands for the $j^{\text{th}}$ granule of the $i^{\text{th}}$ linear aperiodic granularity.
@@
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities
-(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities
-
-The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
+(1) **single-to-multiple:** and (2) **multiple-to-single:** which entails the representation of multiple-order-up cyclic granularities from single order-up cyclic granularities and vice versa.
+The hierarchical structure of time creates a natural nested ordering which can produce _single-order-up_ or _multiple-order-up_ granularities.
 We shall use the notion of a hierarchy table and order to define them.
 
-**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
+**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy table containing $n$ linear granularities.
 $G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in {1, 2, \dots,n}$ and $l<m$.
@@
 
-We refer to granularities which are nested within multiple levels of the hierarchy table as multiple-order-up granularities and those concerning a single level as single-order-up granularities.
+We refer to granularities which are nested within multiple levels of the hierarchy table as _multiple-order-up_ granularities and those concerning a single level as _single-order-up_ granularities.
 
@@
 But it is far from being the only one.
-All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars [@Reingold2001-kf].
+<!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->.
+The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
 The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
-Nevertheless, for any calendar a hierarchy can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
-
- - 1 kin = 1 day
- - 1 uinal = 20 kin
- - 1 tun  = 18 uinal
- - 1 katun = 20 tun
- - 1 baktun = 20 katun
+Nevertheless, for any calendar a hierarchy table can be defined.
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun
 
@@
 
-*Example:* Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
+**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
 
@@
 C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
-               = & L(\lfloor z/20\rfloor \mod 18  \\
-               &~~~ + 20*\lfloor z/20*18\rfloor \mod 20  \\
+               = & L(\lfloor z/20\rfloor \mod 18 + 20*\lfloor z/20*18\rfloor \mod 20  \\
                &~~~ + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
@@
 Examples might include hour-of-month or day-of-year.
-Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$.
-
-\begin{equation} \label{eq:hour-month}
-\begin{split}
-C_{hour, month}(z) & = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))\\
-& = L(12 + 24*20)\\
-\end{split}
-\end{equation}
-
-\begin{equation} \label{eq:day-year}
-\begin{split}
-C_{day, year}(z) & = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert))\\
-\end{split}
-\end{equation}
-
-where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-
-Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
-$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
-However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
-It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
-
-Generally speaking, to keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy.
-Any multiple order-up quasi-circular granularity can then be considered to be a function of some circular (single or multiple order-up) and a quasi-circular single order-up granularity.
-
-<!-- Both of them can be circular or quasi-circular or it can be each of both. In case they are both quasi-circular or one circular and the other quasi-circular, we need to adjust the formula -  -->
-
-For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, two different approaches need to be employed for computing multiple order-up granularities from single order-up ones.
-
-- Case when $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+
+Generally speaking, to keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
+Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.
+Two different approaches need to be employed for the follwing cases:
+
+<!-- For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, -->
+- $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
 
@@
 
-- Case when  $C_{l,m'}(z)$  is quasi-circular and $C_{m',m}(z)$ is circular
+- $C_{l,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
 
@@
 where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+
+<!-- Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$. -->
+
+**Example:** Let us use Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2} to compute the multiple-order-up granularity hour_month and day_year for the Gregorian calendar considered.
+
+ - $C_{hour, month}(z) = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))$  
+ - $C_{day, year}(z) = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert))$, where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$.
+
+<!-- CONTENT CUT -->
+<!-- Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
+$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
+However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
+It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
+-->
+<!-- If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}. -->
 
@@
 <!-- where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping. -->
-
-If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
 

### 2020-02-24 main.Rmd
data structure introduction changed

@@
 Effective exploration or good visualization require good data structures.
-Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization.
 It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use calendars based on application domain that define contextual system of granularities.
-Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities.
-Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.
-Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
+Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.
+However, restructuring time in this manner leads to re-organisation of the data structure, where each level of a cyclic granularity correspond to multiple values of the measured variable, unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
+It is common to see consolidation of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity to correspond to an unique value of the measured variable.
+But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
+Summarizing the probability distribution of these multiple observations using some summary statistics together, each of which can highlight different properties of the data could be a potential way to compehend periocities in the data.
+Hence, a possible data structure would be to consider different summarisation of the probability distribution of the measured variable across two cyclic granularities.
+Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might become clumsy.
 Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
+
+<!--Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.-->
+<!-- Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
+Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
+-->
 
@@
 Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
-We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the categories/levels of the cyclic granularities.
-To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
+We will discuss few cases, where one or more of these $nm$ sets will be empty, the implication of which we will discuss in Section \ref{sec:vis-interaction}.
+
+<!--To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
 
@@
 Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called "structurally" empty combinations.
+These are "structurally" empty combinations in that it is impossible for them to have any value of the measured variable.
 Let us take a specific example, where $C_1$ be day-of-month and $C_2$ be week-of-month.
@@
 Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
-These are structurally empty sets in that it is impossible for them to have any observations.
+For example $S_{1,5}$, $S_{21,2}$, etc.
+This is also intuitive since the first day of the month can never correspond to fifth week of the month.
 
@@
 These are called "event-driven" empty combinations.
-Again, let us consider a specific example to illustrate this.
-Let $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
+Let us consider $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
 Here $C_1$ can have 7 levels while $C_2$ can have 2 levels.
@@
 Thirdly, empty combinations can be a result of how granularities are constructed.
+These are called "build-based" empty combinations.
 Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
@@
 This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
-Hence, they are referred to as "build-based" empty combinations.
-
-An example when there will be no empty combinations could be where $C_1$ maps index set to day-of-week and $C_2$ maps index set to month-of-year.
+
+An example when there will be no empty combinations could be where $C_1$ and $C_2$ maps index set to day-of-week and month-of-year respectively.
 Here $C_1$ can have 7 levels while $C_2$ can have 12 levels.
@@
 
-## Measured variable {#sec:measured-variable}
+## Summarizing the distribution of a measured variable {#sec:measured-variable}
 
@@
 Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
-With huge amount of data being available, mean, median or any one summary statistic might not be enough to understand the different properties of the measured variable.
-Instead the distribution could be summarized using some summary statistics together, each of which can highlight different properties of the data.
 Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries.
@@
 
-### Interaction
+### Interaction {#sec:vis-interaction}
 

### 2020-02-25 main.Rmd
CONTENT CUT - DATA STRUCTURE - iteration 1

@@
 
-For more variations of calendar algrabra operations, see @Ning2002-tf.
+For more variations of calendar algebra operations, see @Ning2002-tf.
 
@@
 A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
-U.S labour day is the first Monday in September and U.S.
+U.S labor day is the first Monday in September and U.S.
 Memorial Day is the last day in May.
@@
 
-**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labour day and M can be conceived as the aperiodic linear granularity denoting any holidays.
+**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Lab our day and M can be conceived as the aperiodic linear granularity denoting any holidays.
 The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
@@
 Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.
-Two different approaches need to be employed for the follwing cases:
+Two different approaches need to be employed for the following cases:
 
@@
 However, restructuring time in this manner leads to re-organisation of the data structure, where each level of a cyclic granularity correspond to multiple values of the measured variable, unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
-It is common to see consolidation of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity to correspond to an unique value of the measured variable.
+It is common to see consolidation of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
 But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the probability distribution of these multiple observations using some summary statistics together, each of which can highlight different properties of the data could be a potential way to compehend periocities in the data.
-Hence, a possible data structure would be to consider different summarisation of the probability distribution of the measured variable across two cyclic granularities.
-Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might become clumsy.
-Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
+Summarizing the probability distribution of these multiple observations across cyclic granularities using some summary statistics together, each of which can highlight different features of the data could be a potential way to comprehend periocities in the data.
+Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
+Instead, the big problem is broken down by focusing on two cyclic granularities at a time.
 
@@
 While considering data sets of the form $<C_1, C_2, v>$, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
-Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
+Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
 We will discuss few cases, where one or more of these $nm$ sets will be empty, the implication of which we will discuss in Section \ref{sec:vis-interaction}.
@@
 These are "structurally" empty combinations in that it is impossible for them to have any value of the measured variable.
-Let us take a specific example, where $C_1$ be day-of-month and $C_2$ be week-of-month.
-Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories.
+Let us take a specific example, where $C_1$ be day-of-month with 31 levels and $C_2$ be week-of-month with 5 levels.
 There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of day-of-month and week-of-month.
@@
 These are called "event-driven" empty combinations.
-Let us consider $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
-Here $C_1$ can have 7 levels while $C_2$ can have 2 levels.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of $C_1$ and $C_2$.
-While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations.
+Let us consider $C_1$ be day-of-week with 7 levels and $C_2$ be WorkingDay/NonWorkingDay with 2 levels.
+While potentially all of these 14 sets $S_{ij}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
 For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
@@
 Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
-Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
+Then the days denoting weekends in a month would not correspond to any Business days.
 This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
@@
 
-If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$ and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$, the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by:
-
-\begin{equation}
-\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h
-\end{equation}
-
-where $h$ is the smoothing parameter or the bandwidth. The MSE of the estimator is given by:
-
+If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$, <!--and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$,--> the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by $\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h$, where $h$ is the smoothing parameter or the bandwidth.
+The MSE of the estimator is given by:
 \begin{equation}\label{eq:MSE}
 \begin{split}
-MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
+MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
 \end{split}
@@
 <!-- where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$. -->
-<!-- MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa -->
+<!--  -->
+<!-- MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\ -->
 
@@
 <!-- A well designed visual representation can replace cognitive calculations with simple perpetual inferences and improve speed of comprehension and decision making.
-However, creating a visualisation requires a number of nuanced judgements.
+However, creating a visualization requires a number of nuanced judgements.
 The challenge being given a data set, the number of visual encodings and hence the space of possible visual design is extremely huge.
@@
 Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization.
-The basic plot choice for our data structure is the one that can display densities.
-In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
-We will discuss few conventional and recent ways to plot distributions using both of these methods.
+The basic plot choice for our data structure is the one that can display distributions<!--In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated.-->.
+We will discuss few conventional and recent ways to plot distributions using both Kernel density estimates and descriptive statistics.
 
@@
 
-Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.--> Due to wide spread popularity and simplicity in implementation, a number of variations are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.
-Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
+Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers at a glance.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.--> Due to wide spread popularity and simplicity in implementation, a number of variations like Notched box plots [@Mcgill1978-hg] are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.<!--has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot.
+The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
+-->.
 However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
@@
 <!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.
-“outliers” in letter value plots are those observations beyond the most extreme letter value.
-The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
+The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.<!--“outliers” in letter value plots are those observations beyond the most extreme letter value.--> The letter values are displayed till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
 
@@
 These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
-While in a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
-outside 1.5 times the interquartile range above the upper quartile and below the lower quartile), in quantile plots outliers are open to interpretation and not shown.
+While in a boxplot and letter value plots are well defined <!--an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
+outside 1.5 times the interquartile range above the upper quartile and below the lower quartile)-->, in quantile plots outliers are open to interpretation and not shown.
 
@@
 The shape of the violin represents the density estimate of the variable.
-The more data points in a specific range, the larger the violin is for that range.
-Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
+<!--The more data points in a specific range, the larger the violin is for that range.
+-->Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
 It becomes less useful when there are too many levels or categories to display.
 
-The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments (mean, standard deviation, skewness, kurtosis and tailings), and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
+The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
 This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
@@
 <!-- Description and lit review of density plots -->
- A Ridge line plot (sometimes called Joy plot) shows the distribution for several groups.
+ A Ridge line plot <!--(sometimes called Joy plot)--> shows the distribution for several groups.
 Distribution can be represented using density plots, all aligned to the same horizontal scale and presented with a slight overlap.
@@
 Each type of density display has different parameters, that need to be estimated given the data.
-Each is equipped with some benefits and challenges.
-For example, if we have too many categories in the variables mapped to facet or x-axis, the quantile plots might be useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior.
+Each is equipped with some benefits and challenges which should be borne in mind while using them for exploration.
+<!--For example, if we have too many categories in the variables mapped to facet or x-axis, the quantile plots might be useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior-->
 
@@
 
-We will discuss the effect of the levels of cyclic granularities in this section.
-Recall that cyclic granularities $C_1$ and $C_2$ are such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-$A_i$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.
-The levels have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high.
-A potential approach can be to categorize the levels as very high/high/medium/low for each cyclic granularities and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
-Default values for these categorizations can be chosen based on levels of common temporal granularities like days-of-month, days-of-fortnight and days-of-week.
-For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.
+We will discuss the effect of the levels of cyclic granularities on the choice of plots <!--Recall that for cyclic granularities $C_1$ and $C_2$ such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
+$A_i$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.The levels have an impact on the--> since space and resolution might become a problem if the number of levels are too high.
+A potential approach could be to categorize the levels as very high/high/medium/low for each cyclic granularity and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
+Default values for these categorizations can be chosen based on levels of common temporal granularities like days-of-month, days-of-fortnight and days-of-week.<!--For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.>
 
@@
 
-We will discuss the effect of the interaction of cyclic granularities in this section.
-In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
-In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.
-
-For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
+We will discuss the effect of the interaction of cyclic granularities in this section.<!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
+In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
 \autoref{fig:allFig}.
@@
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case $A_i$'s are compared against each other within each group.
-When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$.
-Although the variables used in the layers are the same, the difference in mapping leads to emphasis of different behavior of the variables.
+Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$-->.
+With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
 
@@
 This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the statistical transformations based on density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators.
-\autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.
+In such cases, the statistical transformations based on density should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators (see \autoref{eq:MSE})
+
+<!--  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$--> is a function of the sample size $n$. -->
 
@@
 Given the data structure <$C1$ $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
-Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualisation.
+Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization.
 However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together.

### 2020-02-25 main.Rmd
nonlinear to cyclic terminology - consistent with mid_can review document

@@
 ---
-title: Visualizing probability distributions across bivariate nonlinear temporal granularities
+title: Visualizing probability distributions across bivariate cyclic temporal granularities
 
@@
 We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear and nonlinear time granularities.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Nonlinear time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
+There are several classes of time deconstructions including linear and cyclic time granularities.
+Linear time granularities respect the linear progression of time such as hours, days, weeks and monthsCyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
 
@@
 One of the first attempts to characterize these granularities is due to @Bettini1998-ed.
-However, the definitions and rules defined are inadequate for describing nonlinear or repeating granularities.
-Hence, there is a need to define some new nonlinear time granularities, that can be useful in visualizations.
-Nonlinear time granularities can be circular, quasi-circular or aperiodic.
+However, the definitions and rules defined are inadequate for describing cyclic or repeating granularities.
+Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizationsCyclic time granularities can be circular, quasi-circular or aperiodic.
 Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
@@
 
-  * provide a formal characterization of nonlinear granularities;
+  * provide a formal characterization of cyclic granularities;
   * facilitate manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra;
-  * develop an approach to check the feasibility of creating plots or drawing inferences for any two nonlinear granularities;
+  * develop an approach to check the feasibility of creating plots or drawing inferences for any two cyclic granularities;
   * recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pair of granularities.
@@
 The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provide some background material on linear granularities and introduces calendar algebra for computing different linear granularities.
-Section \ref{sec:cyclic-gran} formally characterizes different nonlinear time granularities by extending the framework of linear time granularities.
-Section \ref{sec:cyclic-calendar} introduces calendar algebra for computing nonlinear time granularities.
-Section \ref{sec:data-structure} discusses the data structure for exploring the conditional distributions of the associated time series across pairs of nonlinear time granularities.
+Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
+Section \ref{sec:cyclic-calendar} introduces calendar algebra for computing cyclic time granularities.
+Section \ref{sec:data-structure} discusses the data structure for exploring the conditional distributions of the associated time series across pairs of cyclic time granularities.
 Section \ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
@@
 
-# Nonlinear time granularities {#sec:cyclic-gran}
+# cyclic time granularities {#sec:cyclic-gran}
 
@@
 
-We propose a formalism of nonlinear time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
+We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
 A tsibble consists of an index, key and measured variables.
@@
 
-Nonlinear time granularities with periodicities can be constructed by relating two linear granularities.
+cyclic time granularities with periodicities can be constructed by relating two linear granularities.
 The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months).
 <!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).--> A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.
-Nonlinear time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular.
+Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 
-Aperiodic nonlinear granularities are defined using aperiodic linear granularities.
+Aperiodic cyclic granularities are defined using aperiodic linear granularities.
 Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$, that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
@@
 
-\begin{definition}[Aperiodic nonlinear granularity]\label{aperiodic}
-An aperiodic nonlinear granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
+\begin{definition}[Aperiodic cyclic granularity]\label{aperiodic}
+An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
 \begin{equation}\label{eq:aperiodic}
@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Aperiodic nonlinear granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$"}
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Aperiodic cyclic granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$"}
 knitr::include_graphics("Figs/aperiodic-example.png")
@@
 
-In \autoref{sec:cyclic-gran}, we discussed how we can define nonlinear granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
+In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.

### 2020-02-25 main.Rmd
nonlinear to cyclic

--- 2020-02-24 main.Rmd
@@
 We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear and nonlinear time granularities.
+There are several classes of time deconstructions including linear and cyclic time granularities.
 Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Nonlinear time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
+Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
 
@@
 preamble: >
-  \setlength {\marginparwidth }{2cm}
   \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
@@
 One of the first attempts to characterize these granularities is due to @Bettini1998-ed.
-However, the definitions and rules defined are inadequate for describing nonlinear or repeating granularities.
-Hence, there is a need to define some new nonlinear time granularities, that can be useful in visualizations.
-Nonlinear time granularities can be circular, quasi-circular or aperiodic.
+However, the definitions and rules defined are inadequate for describing cyclic or repeating granularities.
+Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizations.
+Cyclic time granularities can be circular, quasi-circular or aperiodic.
 Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
@@
 ```
-In addition to the integer index, each granule could be associated with a textual representation which is called the label.
-The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
-
-@Bettini1998-ed discusses the relationships between linear time granularities, periodicities in time, and the structure of a calendar.
-
-<!-- CONTENT CUT -->
-<!-- \begin{definition}[Finer than]\label{def:finerthan} -->
-<!-- A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$. -->
-<!-- \end{definition}. -->
+
+@Bettini1998-ed discusses the relationships between linear time granularities, the structure of a calendar, and periodicities in time.
+
+\begin{definition}[Finer than]\label{def:finerthan}
+A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$.
+\end{definition}.
 
@@
 
-<!-- CONTENT CUT -->
-<!-- These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true.
+These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true.
 However, weekends are finer than weeks but do not group into weeks.
@@
 The period becomes 400 years with the inclusion of leap years and all their exceptions.
--->
-
-For example, day $\trianglelefteq$ week since every granule of "week" is the union of some set of granules of "day".
-Consider another example where day $\trianglelefteq$ month.
-This relationship however is incomplete without its association to periodicity.
-Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions.
-
-<!-- CONTENT CUT -->
-<!-- # ```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships.
+
+```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships.
 Weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or finer than public holidays"} -->
-<!-- knitr::include_graphics("Figs/groupsandfiner.png") -->
-<!-- ``` -->
+And weekends neither groups into or finer than public holidays"}
+knitr::include_graphics("Figs/groupsandfiner.png")
+```
 
@@
 
-<!-- CONTENT CUT -->
-<!-- A granularity $H$ which is periodical with respect to $G$ is specified by: -->
-<!-- (i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period; -->
-<!-- (ii) the value of $P$; -->
-<!-- (iii) the indexes of first and last granules in $H$, if their value is not infinite. -->
-
-If ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
+A granularity $H$ which is periodical with respect to $G$ is specified by:
+(i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period;
+(ii) the value of $P$;
+(iii) the indexes of first and last granules in $H$, if their value is not infinite.
+Then, if ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
 $H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
 Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
-These are called quasi-periodic relationships.
-Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
 
@@
 
-<!-- CONTENT CUT -->
-<!-- \begin{definition}[Groups quasi-periodically]\label{def:quasi} -->
-<!-- A granularity $H$ is quasi-periodical with respect to a granularity $G$ if: -->
-<!-- (1) $G \trianglelefteq H$; and -->
-<!-- (2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
--->
-<!-- \end{definition} -->
+\begin{definition}[Groups quasi-periodically]\label{def:quasi}
+A granularity $H$ is quasi-periodical with respect to a granularity $G$ if:
+(1) $G \trianglelefteq H$; and
+(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
+\end{definition}
+
+Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
 
@@
 
-For example, in a Gregorian calendar without leap years each month is a grouping of the same number of days each year.
+Consider the Gregorian calendar without leap years.
+Then each month is a grouping of the same number of days each year.
 The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$.
-Thus, we could say days group periodically into months.
-In a Gregorian calendar with leap years, days group quasi-periodically into months.
-The value of $P$ and $R$ would still be 365 and 12 respectively, while the <!--$E_i$s--> exceptions would be the time domain corresponding to $29^{\text{th}}$ February of any year.
-
-<!-- CONTENT CUT -->
-<!-- \begin{definition}[Bottom granularity]\label{def:B} -->
-<!-- Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
--->
-<!-- \end{definition} -->
-
-<!-- Here, g-rel could be any relationship defined between granularities as defined in @Bettini1998-ed.
+Thus, we could say days groups periodically into months.
+Now consider the Gregorian calendar with leap years, so that days group quasi-periodically into months.
+The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s would be the time domain corresponding to $29^{\text{th}}$ February of any year.
+
+\begin{definition}[Bottom granularity]\label{def:B}
+Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
+\end{definition}
+
+Here, g-rel could be any relationship defined between granularities like "groups into", "finer than" or "groups periodically".
 Given the set of all granularities defined over the time domain $(\mathbb{Z}; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
--->
-
-A discrete time model often uses a fixed smallest granularity also known as _bottom granularity_ (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
-Chronons may be grouped into granules and a non-overlapping mapping of granules to subsets of the time domain is a linear granularity.
-A system of multiple granularities in lattice structures is referred to as a _calendar_ (see @Dyreson2000-mr).
+
+<!-- Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity.
+The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations.
+-->
 
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a *_bottom granularity_* and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
-Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic granularities.
-
-<!-- CONTENT CUT -->
-<!-- The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
+It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
 The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
 For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes".
--->
-<!-- The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity.
+The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity.
 For example, one can define the granularity "Monday" and hence select only Mondays from linear granularity "day".
--->
+
+<!-- It might be the case that we do not always have integers as the label sets of granularities.  -->
 
@@
 
-<!-- CONTENT CUT -->
-<!-- The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
-A labeled granularity is a pair $(L, G)$, where the label set $L$ is a subset of the integers instead of the set of integers and -->
-<!-- follows the conditions (1) and (2) in \ref{def:linear} where $i$, $j$, and $k$ are in $L$.
-When the label set $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->
-
-<!-- $G$ is a mapping from $L$ to the subsets of the time domain such that: -->
-
-<!--   1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and -->
-<!--   2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty. -->
+The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
+A labeled granularity is a pair $(L, G)$, where $L$ is a subset of the integers, and $G$ is a mapping from $L$ to the subsets of the time domain such that:
+
+  1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and
+  2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
+
+$L$ is called the label set of $G$.
+When $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
+However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.
+
+Some relevant grouping-oriented operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.
 
@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$
-
-<!-- CONTENT CUT -->
-<!-- where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
+For example:
+\begin{align*}
+  \text{minute} &= \text{Group}_{60}(\text{second});\\
+  \text{hour}   &= \text{Group}_{60}(\text{minute});\\
+  \text{day}    &= \text{Group}_{24}(\text{hour});\\
+  \text{week}   &= \text{Group}_{7}(\text{day});
+\end{align*}
+where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
 Otherwise, repeated operations of grouping might result in a granule of "resultant" granularity "day" or "week", which is different from a time-stamped "day" or "week".
--->
 
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
-
+Let us consider a granularity "pseudomonth" which computes months only for non-leap years.
 \begin{align*}
-\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \text{Alter}^{12}_{9,-1}(\text{day}, \\
-  &~~\text{Alter}^{12}_{6,-1}(\text{day}, \text{Alter}^{12}_{4,-1)}(\text{day}, \\
-  &~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \text{Group}_{31}(\text{day})))))),
+\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \\
+  &~~\text{Alter}^{12}_{9,-1}(\text{day}, \\
+  &~~~~\text{Alter}^{12}_{6,-1}(\text{day}, \\
+  &~~~~~~\text{Alter}^{12}_{4,-1)}(\text{day}, \\
+  &~~~~~~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \\
+  &~~~~~~~~~~\text{Group}_{31}(\text{day})))))),
 \end{align*}
-
-For more variations of calendar algrabra operations, see @Ning2002-tf.
-
-# Nonlinear time granularities {#sec:cyclic-gran}
+where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
+
+For more variations of grouping operations and granule oriented operations, see @Ning2002-tf.
+
+# cyclic time granularities {#sec:cyclic-gran}
 
@@
 
-We propose a formalism of nonlinear time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
+We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
 A tsibble consists of an index, key and measured variables.
 An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-In a tsibble, each observation (row) is uniquely identified by index and key.
+In a tsibble, each observation (row) should be uniquely identified by index and key.
 
@@
 
-Nonlinear time granularities with periodicities can be constructed by relating two linear granularities.
-The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months).
-<!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).--> A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.
-Nonlinear time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular.
+cyclic time granularities with periodicities can be constructed by relating two linear granularities.
+The mappings between these linear granularities can be regular or irregular.
+For example, a regular mapping exists between minutes and hours (where 60 minutes always add up to 1 hour), while an irregular mapping exists between days and months (as the number of days per month ranges from 28 to 31).
+Mappings will be regular if linear granularities are formed by a grouping operation, but granularities formed by the altering-tick operation will be irregular.
+Cyclic time granularities are referred to as **circular** if mappings between linear time granularities are regular and **quasi-circular** if the mapping is irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 
-\autoref{fig:circular-dow} depicts the circular granularity "Day-of-Week", constructed from linear granularities "day" (the bottom granularity) and "week".
+\autoref{fig:circular-dow} depicts the circular granularity "day of the week", constructed from two linear granularities "day" (the bottom granularity) and "week".
 The granules that overlap share elements from the underlying time domain.
@@
 Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
-<!--The representation of the resultant circular granularity will be a function of the index set.--> The bottom granularity is assumed to be minutes in this case.
+It is possible that none of the two linear granularities are bottom granularities.
+But the representation of the resultant circular granularity will be a function of the index set.
+The bottom granularity is assumed to be minutes.
 
@@
 Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\
+  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
+  & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
+  & = L(8 - 2*\vert T_{0}\vert - \vert T_{1}\vert) \\
+  & = L(8 - 2*2 - 3) \\
   & = L(1) \\
@@
 
-<!-- CONTENT CUT -->
-  <!-- & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\ -->
-  <!-- & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\ -->
-  <!-- & = L(8 - 2*\vert T_{0}\vert - \vert T_{1}\vert) \\ -->
-  <!-- & = L(8 - 2*2 - 3) \\ -->
-
-  <!-- & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\ -->
-  <!-- & = L(10 - \vert T_{0 \mod 2}\vert) \\ -->
-  <!-- & = L(10 - \vert T_{0}\vert) \\ -->
-  <!-- & = L(10 - 7) \\ -->
-
 \begin{equation} \label{eq:eq5}
@@
 Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,\quad since \quad  10 \in T_{1}  \\
+  & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
+  & = L(10 - \vert T_{0 \mod 2}\vert) \\
+  & = L(10 - \vert T_{0}\vert) \\
+  & = L(10 - 7) \\
   & = L(3) \\
@@
 
-**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
+**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then the following holds true:
+
+- $B \trianglelefteq G'$
+- there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$
+
+from definition \ref{def:periodical}
+
+and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping)
+
 Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
@@
 
-It can be noted here that if a linear granularity $G'$ is quasi-periodic (Bettini2000-vy) <!--(Definition \autoref{def:quasi})--> with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$.<!--(by Definition \autoref{def:quasi}).-->
+It can be noted here that if a linear granularity $G'$ is quasi-periodic (Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).
 
@@
 
-Aperiodic nonlinear granularities are defined using aperiodic linear granularities.
+Aperiodic cyclic granularities are defined using aperiodic linear granularities.
 Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$, that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
 Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
-Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
+Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that
+\begin{equation*}
+M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)
+\end{equation*}
+and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
 Here, index ${\{i_j\}}$ stands for the $j^{\text{th}}$ granule of the $i^{\text{th}}$ linear aperiodic granularity.
 
-\begin{definition}[Aperiodic nonlinear granularity]\label{aperiodic}
-An aperiodic nonlinear granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
+\begin{definition}[Aperiodic cyclic granularity]\label{aperiodic}
+An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
 \begin{equation}\label{eq:aperiodic}
@@
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple:** and (2) **multiple-to-single:** which entails the representation of multiple-order-up cyclic granularities from single order-up cyclic granularities and vice versa.
-The hierarchical structure of time creates a natural nested ordering which can produce _single-order-up_ or _multiple-order-up_ granularities.
+(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities
+(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities
+
+The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities.
 We shall use the notion of a hierarchy table and order to define them.
 
-**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy table containing $n$ linear granularities.
+**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities.
 $G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in {1, 2, \dots,n}$ and $l<m$.
@@
 
-We refer to granularities which are nested within multiple levels of the hierarchy table as _multiple-order-up_ granularities and those concerning a single level as _single-order-up_ granularities.
+We refer to granularities which are nested within multiple levels of the hierarchy table as multiple-order-up granularities and those concerning a single level as single-order-up granularities.
 
@@
 But it is far from being the only one.
-<!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->.
-The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
+All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
 The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
-Nevertheless, for any calendar a hierarchy table can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun
+Nevertheless, for any calendar a hierarchy can be defined.
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:
+
+ - 1 kin = 1 day
+ - 1 uinal = 20 kin
+ - 1 tun  = 18 uinal
+ - 1 katun = 20 tun
+ - 1 baktun = 20 katun
 
@@
 
-**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
+*Example:* Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
 
@@
 C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
-               = & L(\lfloor z/20\rfloor \mod 18 + 20*\lfloor z/20*18\rfloor \mod 20  \\
+               = & L(\lfloor z/20\rfloor \mod 18  \\
+               &~~~ + 20*\lfloor z/20*18\rfloor \mod 20  \\
                &~~~ + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
@@
 Examples might include hour-of-month or day-of-year.
-
-Generally speaking, to keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
-Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.
-Two different approaches need to be employed for the follwing cases:
-
-<!-- For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, -->
-- $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$.
+
+\begin{equation} \label{eq:hour-month}
+\begin{split}
+C_{hour, month}(z) & = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))\\
+& = L(12 + 24*20)\\
+\end{split}
+\end{equation}
+
+\begin{equation} \label{eq:day-year}
+\begin{split}
+C_{day, year}(z) & = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert))\\
+\end{split}
+\end{equation}
+
+where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+
+Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
+$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
+However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
+It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
+
+Generally speaking, to keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy.
+Any multiple order-up quasi-circular granularity can then be considered to be a function of some circular (single or multiple order-up) and a quasi-circular single order-up granularity.
+
+<!-- Both of them can be circular or quasi-circular or it can be each of both. In case they are both quasi-circular or one circular and the other quasi-circular, we need to adjust the formula -  -->
+
+For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, two different approaches need to be employed for computing multiple order-up granularities from single order-up ones.
+
+- Case when $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
 
@@
 
-- $C_{l,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
+- Case when  $C_{l,m'}(z)$  is quasi-circular and $C_{m',m}(z)$ is circular
 
@@
 where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-
-<!-- Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$. -->
-
-**Example:** Let us use Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2} to compute the multiple-order-up granularity hour_month and day_year for the Gregorian calendar considered.
-
- - $C_{hour, month}(z) = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))$  
- - $C_{day, year}(z) = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert))$, where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$.
-
-<!-- CONTENT CUT -->
-<!-- Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
-$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
-However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
-It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
--->
-<!-- If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}. -->
 
@@
 <!-- where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping. -->
+
+If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}.
 
@@
 Effective exploration or good visualization require good data structures.
+Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization.
 It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.
-However, restructuring time in this manner leads to re-organisation of the data structure, where each level of a cyclic granularity correspond to multiple values of the measured variable, unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
-It is common to see consolidation of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity to correspond to an unique value of the measured variable.
-But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the probability distribution of these multiple observations using some summary statistics together, each of which can highlight different properties of the data could be a potential way to compehend periocities in the data.
-Hence, a possible data structure would be to consider different summarisation of the probability distribution of the measured variable across two cyclic granularities.
-Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might become clumsy.
+Moreover, one can use calendars based on application domain that define contextual system of granularities.
+Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities.
+Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.
+Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
 Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
-
-<!--Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.-->
-<!-- Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
--->
 
@@
 Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
-We will discuss few cases, where one or more of these $nm$ sets will be empty, the implication of which we will discuss in Section \ref{sec:vis-interaction}.
-
-<!--To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
+We will discuss few cases, where one or more of these $nm$ sets will be empty, which will essentially mean that the measured variable can not be explored for all combinations of the categories/levels of the cyclic granularities.
+To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
 
@@
 Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are "structurally" empty combinations in that it is impossible for them to have any value of the measured variable.
-Let us take a specific example, where $C_1$ be day-of-month and $C_2$ be week-of-month.
+These are called "structurally" empty combinations.
+Let us take a specific example, where $C_1$ maps index set to day-of-month and $C_2$ maps index set to week-of-month.
 Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories.
@@
 Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-This is also intuitive since the first day of the month can never correspond to fifth week of the month.
+For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month.
+These are structurally empty sets in that it is impossible for them to have any observations.
 
@@
 These are called "event-driven" empty combinations.
-Let us consider $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
+Again, let us consider a specific example to illustrate this.
+Let $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay.
 Here $C_1$ can have 7 levels while $C_2$ can have 2 levels.
-So there are 14 sets $S_{ij}$ corresponding to the possible combinations of $C_1$ and $C_2$.
+So there are 14 sets $S_{ij}$ corresponding to the possible combinations of day-of-week and WorkingDay/NonWorkingDay.
 While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations.
-For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
+For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
 
@@
 Thirdly, empty combinations can be a result of how granularities are constructed.
-These are called "build-based" empty combinations.
-Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
+Let $C_1$ maps index set to Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
 Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed.
 This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
-
-An example when there will be no empty combinations could be where $C_1$ and $C_2$ maps index set to day-of-week and month-of-year respectively.
+Hence, they are referred to as "build-based" empty combinations.
+
+An example when there will be no empty combinations could be where $C_1$ maps index set to day-of-week and $C_2$ maps index set to month-of-year.
 Here $C_1$ can have 7 levels while $C_2$ can have 12 levels.
@@
 
-## Summarizing the distribution of a measured variable {#sec:measured-variable}
+## Measured variable {#sec:measured-variable}
 
@@
 Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
+With huge amount of data being available, mean, median or any one summary statistic might not be enough to understand the different properties of the measured variable.
+Instead the distribution could be summarized using some summary statistics together, each of which can highlight different properties of the data.
 Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries.
@@
 These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
-Several forms of non-parametric density estimators <!--and quantile estimators--> exist in the literature, the most commonly used form in @Silverman1986-cl<!--,  @Jones1992-pj)--> is as follows:
+Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
 
@@
 
-where $h$ is the smoothing parameter or the bandwidth. The MSE of the estimator is given by:
+where $h$ is the smoothing parameter or the bandwidth. The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by:
+
+<!-- $Q$, $(q \equiv  Q')$, which is the quantile density function -->
+<!-- $ Q_{n} (u) = \sum_{i=1}^ n X_{(i)} I((i-1)/n <u \leq i/n)$\\ -->
+<!--             $= X_(1) + \sum_{i=1}^ n (X_{(i)} - X_{(i-1)}) I((i-1)/n < u)$ -->
+
+<!-- and most widely used definition of sample quantile is given in  by: -->
+
+<!-- $$\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}$$ where,  $(j - m)/n \leq p < (j - m + 1)/n$,  -->
+<!-- for some $m \in R$ and $0 < \gamma < 1$.  $\hat{Qi(p)}$ is the ith sample quantile  -->
+
+\begin{equation}
+\hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u)
+\end{equation}
+
+where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$.
+
+ <!-- Both pdf and pmf are concerned with point probabilities of the random variable.
+Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d.
+-->
+
+<!-- Instead of assuming a parametric model for the distribution (e.g.
+a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g.
+differentiable).
+It is then possible to estimate the u2nknown density function f(·).
+-->
+
+<!-- The histogram is the oldest and most popular density estimator method.
+The form of the histogram depends very much on the tuning parameters class width and origin.
+The naive estimator was introduced by Fix and Hodges (1951) and first discussed by Rosenblatt(1956).
+Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962).
+The nearest neighbor estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977).
+There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.
+-->
+
+<!-- The inverse $Q = F^{-1}$, which is the quantile function can be defined as $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$.  -->
+
+The MSE of the estimators are given as follows:
 
@@
 MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
+MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
 \end{split}
@@
 where $\kappa = \int k^2(y)dy$.
-
-<!-- CONTENT CUT -->
-<!-- The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by: -->
-
-<!-- \begin{equation} -->
-<!-- \hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u) -->
-<!-- \end{equation} -->
-
-<!-- where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$. -->
-<!-- MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa -->
 
@@
 In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
-We will discuss few conventional and recent ways to plot distributions using both of these methods.
-
-<!-- CONTENT CUT -->
-<!-- Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
+Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
 The descriptive statistics based methods do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
@@
 Also, the summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.
--->
+
+<!--  However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. -->
+
+We will discuss few conventional and recent ways to plot distributions using both of these methods.
+
+<!-- Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it.
+These displays depend on the specific estimation procedure (e.g., kernel density estimate)  -->
 
@@
 
-Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.--> Due to wide spread popularity and simplicity in implementation, a number of variations are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.
+Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.
+These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.
+Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers.
 Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
 However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
-Also, the number of outliers is large for larger data sets.
+Also, the number of outliers is large for larger data set since number of outliers is proportional to the number of observations.
 
 <!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.
+The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail.
+These are particularly useful for large data sets.
 “outliers” in letter value plots are those observations beyond the most extreme letter value.
@@
 Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
-It becomes less useful when there are too many levels or categories to display.
 
@@
 
-### Interaction {#sec:vis-interaction}
+### Interaction
 
@@
   filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>%
-  prob_plot("month_year", "day_week",
+  prob_plot("month_year","day_week",
             response = "Demand",
-    plot_type = "lv"
-  ) +
+            plot_type = "lv") +
   ylab("Electricity Demand [KWh]") +
@@
   filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
-  prob_plot("day_week", "hour_day",
+  prob_plot("day_week","hour_day",
             response = "Demand",
             plot_type = "quantile",
-    symmetric = FALSE
-  ) +
+            symmetric = FALSE) +
   ylab("Electricity Demand [KWh]") +
@@
                 response = "Demand",
-    plot_type = "boxplot"
-  ) +
+                plot_type = "boxplot")+
   ylab("Electricity Demand [KWh]") +
   scale_x_discrete(breaks = seq(0, 366, 60)) +
-  theme(
-    legend.position = "bottom",
-    strip.text = element_text(size = 7, margin = margin())
-  ) +
+  theme(legend.position = "bottom",
+        strip.text = element_text(size = 7, margin = margin())) +
   ggtitle("(c) Box plot")
@@
                 response = "Demand",
-    plot_type = "violin", alpha = 0.3
-  ) + ylab("") +
+                plot_type = "violin", alpha = 0.3) + ylab("") +
   xlab("Days of the Month") +
@@
 
-VIC %>%
-  create_gran("week_month") %>%
+VIC %>%  create_gran("week_month") %>%
   create_gran("week_year") %>%
-  filter(week_month %in% c(1, 2, 5), week_year %in% c(1:20)) %>%
+  filter(week_month%in% c(1, 2, 5), week_year %in% c(1:20)) %>%
   prob_plot("week_year",
@@
             response = "Demand",
-    plot_type = "ridge"
-  ) +
+            plot_type = "ridge")  +
   xlab("Electricity Demand [KWh]") +
@@
   scale_x_continuous(breaks = seq(2000, 10000, 3000)) +
-  theme(
-    legend.position = "bottom",
-    strip.text = element_text(size = 7, margin = margin())
-  ) +
+  theme(legend.position = "bottom",
+        strip.text = element_text(size = 7, margin = margin())) +
   ggtitle("(e) Ridge plot")
@@
             quantile_prob = seq(0.1, 0.9, 0.1),
-    symmetric = FALSE
-  ) +
-  # scale_x_discrete(breaks = seq(1, 336, 60)) +
+            symmetric = FALSE) +
+   #scale_x_discrete(breaks = seq(1, 336, 60)) +
    ylab("") +
-  xlab("Day of the Year") + theme(
-    legend.position = "none",
-    strip.text = element_text(
-      size = 7,
-      margin = margin()
-    )
-  ) +
+   xlab("Day of the Year") + theme(legend.position = "none",
+                                   strip.text = element_text(size = 7,
+                                                             margin = margin())) +
   ggtitle("(f) Decile plot")
@@
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
+One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
 It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
@@
 ```{r smart-meter, fig.cap = "Smart meter data structure"}
-load("data/sm_cust50.Rdata")
-
-smart_meter <- sm_cust50 %>%
-  as_tsibble() %>%
-  dplyr::select(
-    customer_id,
+sm <- read_rds("data/sm_cust50.rds")
+
+smart_meter <- sm %>% as_tsibble() %>% dplyr::select(customer_id,
                               reading_datetime,
-    general_supply_kwh
-  )
-smart_meter %>% head(n = 5)
+                              general_supply_kwh)
+smart_meter
 ```
@@
 ```{r search_gran_limit2, echo=TRUE}
-smart_meter10 %>% search_gran(
-  highest_unit = "month",
+smart_meter10 %>% search_gran(highest_unit = "month",
                               filter_out = c("hhour", "fortnight")
@@
 smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  gran_advice(
-    "wknd_wday",
-    "hour_day"
-  )
+  filter(customer_id %in% c(10017936))%>%
+  gran_advice("wknd_wday",
+                            "hour_day")
 ```
@@
 # cust1 <- smart_meter10 %>%
-# filter(customer_id %in% c(10006704)) %>%
+#filter(customer_id %in% c(10006704)) %>%
 #   prob_plot("wknd_wday",
@@
 cust2 <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
+  filter(customer_id %in% c(10017936))%>%
   prob_plot("wknd_wday",
@@
             symmetric = TRUE,
-    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
+             quantile_prob = c(0.01,0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) +
   scale_y_sqrt()
@@
 smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
+  filter(customer_id %in% c(10017936))%>%
   prob_plot("wknd_wday",
@@
                             response = "general_supply_kwh",
-    plot_type = "violin"
-  ) +
+                            plot_type = "violin") +
   scale_y_sqrt()
@@
 smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
+  filter(customer_id %in% c(10017936))%>%
   prob_plot("hour_day",
@@
             response = "general_supply_kwh",
-    plot_type = "boxplot"
-  ) +
+            plot_type = "boxplot") +
             scale_y_sqrt()
             # scale_fill_brewer(palette = "Dark2")
+
 ```
@@
    G = c("over", "inning", "match", "season"),
-  C = c("over-of-inning", "inning-of-match", "match-of-season", 1),
-  P = c(20, 2, "aperiodic", 1)
-)
+   C = c("over-of-inning", "inning-of-match", "match-of-season",1),
+   P = c(20, 2, "aperiodic", 1))
  knitr::kable(hierarchy_model, format = "markdown")
@@
 cricket_tsibble %>%
-  filter(batting_team %in% c(
-    "Mumbai Indians",
-    "Chennai Super Kings"
-  )) %>%
+  filter(batting_team %in% c("Mumbai Indians",
+                             "Chennai Super Kings"))%>%
   prob_plot("inning",
@@
   hierarchy_model,
-    plot_type = "lv"
-  ) + ggtitle("") +
+  plot_type = "lv")  + ggtitle("") +
   scale_fill_brewer(palette = "Dark2")
+
 ```
@@
 
-cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" = "id"))
+ cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" ="id"))
 
@@
 cricket_dot_field <- cricket_season %>%
-  mutate(
-    fielding_proxy = if_else(dismissal_kind %in%
-      c("caught", "caught and bowled"), 1, 0),
-    dot_ball_proxy = if_else(total_runs == 0, 1, 0),
-    wicket_proxy = if_else(is.na(dismissal_kind), 0, 1)
-  ) %>%
-  group_by(
-    season,
+  mutate(fielding_proxy = if_else(dismissal_kind %in%
+                                    c("caught", "caught and bowled"),1,0),
+         dot_ball_proxy = if_else(total_runs==0,1,0),
+         wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>%
+  group_by(season,
            match_id,
@@
            inning,
-    over
-  ) %>%
-  summarise(
-    runs_per_over = sum(total_runs),
-    run_rate = sum(total_runs) / length(total_runs),
+           over) %>%
+  summarise(runs_per_over = sum(total_runs),
+            run_rate = sum(total_runs)/length(total_runs),
             fielding_wckts = sum(fielding_proxy),
-    dot_balls = sum(dot_ball_proxy)
-  ) %>%
-  mutate(diff_run_rate = c(0, diff(run_rate)))
+            dot_balls = sum(dot_ball_proxy)) %>%
+            mutate(diff_run_rate = c(0,diff(run_rate)))
 
@@
 
-cricket_data <- cricket_tsibble %>%
-  mutate(
-    field = if_else(fielding_wckts == 0, "no wickets", ">0 wickets"),
-    dot = if_else(dot_balls == 0, "no dot balls", ">0 dot balls"),
+cricket_data<- cricket_tsibble %>%
+  mutate(field = if_else(fielding_wckts==0, "no wickets", ">0 wickets"),
+         dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls"),
          lag_field =  lag(field),
-    lag_dot = lag(dot)
-  ) %>%
-  filter(lag_field != 0, lag_dot != 0)
-```
-
-```{r echo = FALSE, eval = FALSE}
+         lag_dot = lag(dot)) %>% filter(lag_field!=0, lag_dot!=0)
+```
+
+```{r echo = TRUE}
 cricket_data %>% gran_obs("field", "over")
@@
 
-# filter(fielding_wckts %in% c(0,1)) %>%
-cricket_data %>% prob_plot("over", "lag_field",
+  #filter(fielding_wckts %in% c(0,1)) %>%
+  cricket_data %>% prob_plot("over","lag_field",
                                hierarchy_model,
@@
                                plot_type = "violin",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-) + ggtitle("") + geom_boxplot(width = 0.5)
+                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") + geom_boxplot(width = 0.5)
 ```
@@
 cricket_data %>%
-  # filter(dot_balls %in% c(0, 1, 2)) %>%
+  #filter(dot_balls %in% c(0, 1, 2)) %>%
   prob_plot("lag_dot",
@@
             quantile_prob = c(0.25, 0.5, 0.75),
-    symmetric = FALSE
-  ) + ggtitle("")
+            symmetric = FALSE)  + ggtitle("")
 ```
@@
 ```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas"), file = "rpkgs.bib")
-```
+write_bib(c("dplyr", "tidyr", "tsibble", "magrittr","rlang", "gravitas"), file = "rpkgs.bib")
+```

### 2020-02-25 main.Rmd
Monash working paper template

@@
 
-authors:
-- name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-  thanks: "Email: Sayani.Gupta@monash.edu"
-
-- name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
-
-- name: Antony Unwin
-  affiliation: University of Augsburg
-
-keywords:
-- data visualization
-- statistical distributions
-- time granularities
-- calendar algebra
-- periodicties
-- grammar of graphics
-- R
+author:
+  - familyname: Gupta
+    othernames: Sayani
+    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+\newline \newline CSIRO Data61, \newline Private Bag 10, South Clayton, VIC 3169
+    email : Sayani.Gupta@monash.edu
+  - familyname: Hyndman
+    othernames: Rob J
+    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+    email : rob.hyndman@monash.edu
+  - familyname: Cook
+    othernames: Dianne
+    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+    email : dicook@monash.edu
+  - familyname: Unwin
+    othernames: Antony
+    address: University of Augsburg
+    email : au50au@me.com
+
+keywords: "data visualization, statistical distributions, time granularities, calendar algebra, periodicties, grammar of graphics, R"
 
@@
 This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
+toc: false
 bibliography: [bibliography.bib]
@@
   \def\mod{~\text{mod}~}
+biblio-style: authoryear-comp
 output:
-  bookdown::pdf_book:
-    base_format: rticles::asa_article
+  #bookdown::pdf_book:
+    MonashEBSTemplates::workingpaper:
+    #base_format: rticles::asa_article
     fig_height: 5
@@
     keep_tex: yes
+
 ---
@@
 knitr::opts_knit$set(root.dir = here::here())
+knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed = TRUE)})
 ```
@@
 
-# cyclic time granularities {#sec:cyclic-gran}
+# Cyclic time granularities {#sec:cyclic-gran}
 

### 2020-03-05 main.Rmd
spell check

--- 2020-02-25 main.Rmd
@@
 There are several classes of time deconstructions including linear and cyclic time granularities.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and monthsCyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
+Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
+Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
 
@@
 However, the definitions and rules defined are inadequate for describing cyclic or repeating granularities.
-Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizationsCyclic time granularities can be circular, quasi-circular or aperiodic.
+Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizations.
+Cyclic time granularities can be circular, quasi-circular or aperiodic.
 Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
@@
 
-cyclic time granularities with periodicities can be constructed by relating two linear granularities.
+Cyclic time granularities with periodicities can be constructed by relating two linear granularities.
 The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months).
@@
   plot_type = "violin",
-  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
-) + ggtitle("") + geom_boxplot(width = 0.5)
+) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0,4.5) +
+      theme(
+      axis.text = element_text(size = 24),
+      strip.text = element_text(size = 24),
+      axis.title = element_text(size = 24),
+      legend.title = element_text(size = 24),
+      legend.text = element_text(size = 24),
+      legend.position = "bottom",
+      plot.title =  element_text(size = 24)
+    )
 ```

### 2020-03-05 main.Rmd
content cut short - clashes and harmonies

@@
 Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are "structurally" empty combinations in that it is impossible for them to have any value of the measured variable.
+These are called "structurally" empty combinations.
 Let us take a specific example, where $C_1$ be day-of-month with 31 levels and $C_2$ be week-of-month with 5 levels.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of day-of-month and week-of-month.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
+There will be $31\times 5=155$ sets $S_{ij}$ corresponding to possible combinations of $C_1$ and $C_2$.
+Many of these like $S_{1,5}$, $S_{21,2}$ are empty.
 This is also intuitive since the first day of the month can never correspond to fifth week of the month.
@@
 These are called "build-based" empty combinations.
-Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month.
+Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ be day-of-month.
 Then the days denoting weekends in a month would not correspond to any Business days.
@@
 Here $C_1$ can have 7 levels while $C_2$ can have 12 levels.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of day-of-week and month-of-year.
+So there are $12\times7=84$ sets $S_{ij}$.
 All of these are non-empty because every day-of-week can occur in every month.
@@
 
-```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  gran_advice(
-    "wknd_wday",
-    "hour_day"
-  )
-```
-
 ```{r bothcust, fig.cap = "Area quantile plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936 displays the smoothed energy consumption curve for the entire day.
@@
   plot_type = "violin",
-) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0,4.5) +
-      theme(
-      axis.text = element_text(size = 24),
-      strip.text = element_text(size = 24),
-      axis.title = element_text(size = 24),
-      legend.title = element_text(size = 24),
-      legend.text = element_text(size = 24),
-      legend.position = "bottom",
-      plot.title =  element_text(size = 24)
-    )
+) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0,4.5)
 ```
@@
 
-Q3: Are runs set to reduce in the next over for dot balls in the previous over?
-
-A dot ball is a delivery bowled without any runs scored off it.
+<!-- Q3: Are runs set to reduce in the next over for dot balls in the previous over? -->
+
+<!-- A dot ball is a delivery bowled without any runs scored off it.
 The number of dot balls is reflective of the quality of bowling in the game.
@@
 Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
-With at least one dot balls per over, the distribution of run rates in facet 1 increase slower compared to that in facet 2.
+-->
+<!-- With at least one dot balls per over, the distribution of run rates in facet 1 increase slower compared to that in facet 2.
 This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
-
-<!-- Each of the quantiles decrease across facets.
-With two dot balls per over, at least 50% of the times, run rates decrease in the subsequent over as observed through the negative values for different quantiles.
--->
-
-<!-- # ```{r crickread, ech0=FALSE} -->
-<!-- # cricket_tsibble <- gravitas::cricket %>% -->
-<!-- #   mutate(data_index = row_number()) %>% -->
-<!-- #   as_tsibble(index = data_index) -->
-<!-- # ``` -->
-
-```{r exdot, fig.cap="25th, 50th, 75th quantiles of runs per over are drawn  across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
-For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.", fig.pos="ht"}
-
-# library(gravitas)
-# library(tsibble)
-#
-#
-# cricket_bowling <- cricket_season %>%
-#   mutate(dot_ball_proxy = if_else(total_runs==0,1,0),
-#          wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>%
-#   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
-#   summarise(runs_per_over = sum(total_runs),
-#             run_rate = sum(total_runs)/length(total_runs),
-#             dot_balls = sum(dot_ball_proxy)) %>%
-#             mutate(diff_run_rate = c(0,diff(run_rate)))
-#
-#
-# cricket_tsibble_bow <- cricket_bowling %>%
-#   ungroup() %>%
-#   mutate(data_index = row_number()) %>%
-#   as_tsibble(index = data_index)
-#
-
-# cricket_tsibble_bow %>%
-#   mutate(dot = if_else(dot_balls==0,"no dot balls", ">0 dot balls")) %>%
-#   #filter(dot_balls %in% c(0, 1, 2)) %>%
-#   prob_plot("dot",
-#             "over",
-#             hierarchy_model,
-#             response = "diff_run_rate",
-#             plot_type = "quantile",
-#             quantile_prob = c(0.1, 0.2, 0.3, 0.4, 0.5),
-#             symmetric = FALSE)  + ggtitle("")
-
-cricket_data %>%
-  # filter(dot_balls %in% c(0, 1, 2)) %>%
-  prob_plot("lag_dot",
-    "over",
-    hierarchy_model,
-    response = "run_rate",
-    plot_type = "quantile",
-    quantile_prob = c(0.25, 0.5, 0.75),
-    symmetric = FALSE
-  ) + ggtitle("")
-```
-
-In Q2 and Q3, dots balls per over or catches per over are considered as aperiodic cyclic granularities with dot balls or catches as aperiodic linear granularities.
+-->
+
+<!-- ```{r exdot, fig.cap="25th, 50th, 75th quantiles of runs per over are drawn  across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
+For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.", fig.pos="ht"} -->
+
+<!-- cricket_data %>% -->
+<!--   # filter(dot_balls %in% c(0, 1, 2)) %>% -->
+<!--   prob_plot("lag_dot", -->
+<!--     "over", -->
+<!--     hierarchy_model, -->
+<!--     response = "run_rate", -->
+<!--     plot_type = "quantile", -->
+<!--     quantile_prob = c(0.25, 0.5, 0.75), -->
+<!--     symmetric = FALSE -->
+<!--   ) + ggtitle("") -->
+<!-- ``` -->
+
+In Q2, wickets per over are considered as aperiodic cyclic granularities with dot balls or catches as aperiodic linear granularities.
 These aperiodic linear granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
@@
 While any special event that corresponds to a time domain can be treated as an aperiodic linear granularity in a temporal case, dot balls or wickets that corresponds to certain balls (index) could be treated as aperiodic events in cricket.
-
-<!-- For example, in the temporal case, aperiodic cyclic granularity "Easter" is computed based on which all time indices correspond to the event "Easter".
-Similarly, for aperiodic cyclic granularity "dot balls per over" is computed based on which all balls correspond to the event "dot balls".
--->
 
@@
 
+# Appendix
+
+```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE}
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  gran_advice(
+    "wknd_wday",
+    "hour_day"
+  )
+```
+
 # Bibliography

### 2020-03-05 main.Rmd
figure a) and f) kept only for showing harmony and clashes

@@
 \autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
-\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
+<!---\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
 \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig} (f) shows decile plots across  day of the year and month of the year.
-
-Clearly, in \autoref{fig:allFig}, we observe that some choices of cyclic time granularities work and others do not.
-In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year.
+\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig}---> (b) shows decile plots across day of the year and month of the year.
+
+Clearly, in \autoref{fig:allFig}, we observe that choice (a) of cyclic time granularities work but (b) does not work.
+In (b) we can not compare the distributions across different facets becuase every day-of-year cannot occur for every month.
+For example, 21st day of the year can never correspond to the month March or December of any year.
+<!---In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year.
 In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-On the contrary, for \autoref{fig:allFig} (a), we will not have any combinations with zero combinations because every day-of-week can occur in any month-of-year.
-Thus the graphs that don't work are those where many of the combination sets are empty.
+On the contrary, for \autoref{fig:allFig}---> In (a) we will not have any combinations because every day-of-week can occur in every month-of-year.
+<!---Thus the graphs that don't work are those where many of the combination sets are empty.--->
+
 In other words, if there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
@@
 
-We redefine harmony and clashes as follows: harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
-As a result, we should avoid plotting clashes as they hinder the exploratory process by having missing combinations of time granularities in each panel.
+We redefine harmony and clashes as follows: harmonies are pairs of granularities that do not contain emptry combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations.
+As a result, we should avoid plotting clashes.<!---as they hinder the exploratory process by having missing combinations of time granularities in each panel.--->
 
 ```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by day-of-month and month-of-year (b) Decile plot by hour-of-day and day-of-week (c) Box plot by day-of-year and day-of-month, (d) Violin plot of day-of-month and week-of-month, (e) Ridge plot by week-of-month and week-of-year, (f) Decile plot by day-of-year and month-of-year.
-Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}
+(a) Letter value plot by day-of-month and month-of-year (b) Decile plot by day-of-year and month-of-year.
+Plot (a) and (b) show harmony and clashes respectively.", fig.show = 'hold'}
 
@@
 
-par(mfrow = c(3, 2))
+par(mfrow = c(1, 2))
 
@@
 
-VIC %>%
-  create_gran("day_week") %>%
-  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
-  prob_plot("day_week", "hour_day",
-    response = "Demand",
-    plot_type = "quantile",
-    symmetric = FALSE
-  ) +
-  ylab("Electricity Demand [KWh]") +
-  ggtitle("(b) Decile plot")
-
-VIC %>%
-  create_gran("day_month") %>%
-  filter(day_month %in% c(1, 15, 29, 31)) %>%
-  prob_plot("day_month",
-    "day_year",
-    response = "Demand",
-    plot_type = "boxplot"
-  ) +
-  ylab("Electricity Demand [KWh]") +
-  scale_x_discrete(breaks = seq(0, 366, 60)) +
-  theme(
-    legend.position = "bottom",
-    strip.text = element_text(size = 7, margin = margin())
-  ) +
-  ggtitle("(c) Box plot")
-
-VIC %>%
-  create_gran("week_month") %>%
-  filter(week_month %in% c(1, 2, 4)) %>%
-  prob_plot("week_month",
-    "day_month",
-    response = "Demand",
-    plot_type = "violin", alpha = 0.3
-  ) + ylab("") +
-  xlab("Days of the Month") +
-  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) +
-  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
-  ggtitle("(d) Violin plot")
-
-VIC %>%
-  create_gran("week_month") %>%
-  create_gran("week_year") %>%
-  filter(week_month %in% c(1, 2, 5), week_year %in% c(1:20)) %>%
-  prob_plot("week_year",
-    "week_month",
-    response = "Demand",
-    plot_type = "ridge"
-  ) +
-  xlab("Electricity Demand [KWh]") +
-  ylab("Weeks of the Year") +
-  scale_x_continuous(breaks = seq(2000, 10000, 3000)) +
-  theme(
-    legend.position = "bottom",
-    strip.text = element_text(size = 7, margin = margin())
-  ) +
-  ggtitle("(e) Ridge plot")
+# VIC %>%
+#   create_gran("day_week") %>%
+#   filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
+#   prob_plot("day_week", "hour_day",
+#     response = "Demand",
+#     plot_type = "quantile",
+#     symmetric = FALSE
+#   ) +
+#   ylab("Electricity Demand [KWh]") +
+#   ggtitle("(b) Decile plot")
+# 
+# VIC %>%
+#   create_gran("day_month") %>%
+#   filter(day_month %in% c(1, 15, 29, 31)) %>%
+#   prob_plot("day_month",
+#     "day_year",
+#     response = "Demand",
+#     plot_type = "boxplot"
+#   ) +
+#   ylab("Electricity Demand [KWh]") +
+#   scale_x_discrete(breaks = seq(0, 366, 60)) +
+#   theme(
+#     legend.position = "bottom",
+#     strip.text = element_text(size = 7, margin = margin())
+#   ) +
+#   ggtitle("(c) Box plot")
+# 
+# VIC %>%
+#   create_gran("week_month") %>%
+#   filter(week_month %in% c(1, 2, 4)) %>%
+#   prob_plot("week_month",
+#     "day_month",
+#     response = "Demand",
+#     plot_type = "violin", alpha = 0.3
+#   ) + ylab("") +
+#   xlab("Days of the Month") +
+#   theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) +
+#   scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
+#   ggtitle("(d) Violin plot")
+# 
+# VIC %>%
+#   create_gran("week_month") %>%
+#   create_gran("week_year") %>%
+#   filter(week_month %in% c(1, 2, 5), week_year %in% c(1:20)) %>%
+#   prob_plot("week_year",
+#     "week_month",
+#     response = "Demand",
+#     plot_type = "ridge"
+#   ) +
+#   xlab("Electricity Demand [KWh]") +
+#   ylab("Weeks of the Year") +
+#   scale_x_continuous(breaks = seq(2000, 10000, 3000)) +
+#   theme(
+#     legend.position = "bottom",
+#     strip.text = element_text(size = 7, margin = margin())
+#   ) +
+#   ggtitle("(e) Ridge plot")
 
@@
     plot_type = "quantile",
-    quantile_prob = seq(0.1, 0.9, 0.1),
+    quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
     symmetric = FALSE
@@
   xlab("Day of the Year") + theme(
-    legend.position = "none",
+    legend.position = "bottom",
     strip.text = element_text(
@@
   ) +
-  ggtitle("(f) Decile plot")
+  ggtitle("(b) Decile plot")
 

### 2020-03-05 main.Rmd
content cut from applications

@@
 
-**Example:** Let us use Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2} to compute the multiple-order-up granularity hour_month and day_year for the Gregorian calendar considered.
+**Example:** Let us use Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2} to compute the multiple-order-up granularity hour-of-month and day-of-year for the Gregorian calendar considered.
 
@@
 This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the statistical transformations based on density should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators (see \autoref{eq:MSE})
-
-<!--  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$--> is a function of the sample size $n$. -->
+In such cases, the statistical transformations based on density should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators (see \autoref{eq:MSE}) <!--  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$--> is a function of the sample size $n$.
 
@@
 These cyclic granularities are a function of the index variable and are computed based on if they are circular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}).
-If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, quarter, semester, year), the following cyclic time granularities can be considered for this analysis.
+If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, year), the following cyclic time granularities can be considered for this analysis.
 The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.
 
-```{r search, echo=TRUE}
+```{r search, echo=FALSE}
 library(gravitas)
-smart_meter %>% search_gran()
+smart_meter %>% search_gran(filter_out = c("semester", "quarter"))
 ```
@@
 If these options are considered too many, the most coarse temporal unit can be set to a “month”.
-
-```{r search_limit1, echo=TRUE}
-smart_meter10 %>%
-  search_gran(highest_unit = "month")
-```
+Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
+
+<!-- CONTENT CUT -->
+<!-- # ```{r search_limit1, echo=TRUE} -->
+<!-- # smart_meter10 %>% -->
+<!-- #   search_gran(highest_unit = "month") -->
+<!-- # ``` -->
 
@@
 <!-- # knitr::kable(smart_meter_limit1, format = "markdown") -->
-
-Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
 
@@
   plot_type = "violin",
-) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0,4.5)
+) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0, 4.5)
 ```
@@
 
-# Appendix
-
-```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  gran_advice(
-    "wknd_wday",
-    "hour_day"
-  )
-```
+<!-- # Appendix -->
+
+<!-- ```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE} -->
+<!-- smart_meter10 %>% -->
+<!--   filter(customer_id %in% c(10017936)) %>% -->
+<!--   gran_advice( -->
+<!--     "wknd_wday", -->
+<!--     "hour_day" -->
+<!--   ) -->
+<!-- ``` -->
+
+<!-- Q3: Are runs set to reduce in the next over for dot balls in the previous over? -->
+
+<!-- A dot ball is a delivery bowled without any runs scored off it.
+The number of dot balls is reflective of the quality of bowling in the game.
+Run rate of an over should ideally decrease if the number of dot balls increase.
+However, what is the effect of dot balls on runs scored in the subsequent over.
+Will players  batsman likely to go for big shots because they couldn't score good runs in the previous over?
+Or they should play consistently and avoid scoring high?
+Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
+-->
+<!-- With at least one dot balls per over, the distribution of run rates in facet 1 increase slower compared to that in facet 2.
+This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
+-->
+
+<!-- ```{r exdot, fig.cap="25th, 50th, 75th quantiles of runs per over are drawn  across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
+For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.", fig.pos="ht"} -->
+
+<!-- cricket_data %>% -->
+<!--   # filter(dot_balls %in% c(0, 1, 2)) %>% -->
+<!--   prob_plot("lag_dot", -->
+<!--     "over", -->
+<!--     hierarchy_model, -->
+<!--     response = "run_rate", -->
+<!--     plot_type = "quantile", -->
+<!--     quantile_prob = c(0.25, 0.5, 0.75), -->
+<!--     symmetric = FALSE -->
+<!--   ) + ggtitle("") -->
+<!-- ``` -->
 

### 2020-03-06 main.Rmd
mesasured variable section delete

@@
 library(tsibble)
+library(ggpubr)
 opts_chunk$set(
@@
 
-## Summarizing the distribution of a measured variable {#sec:measured-variable}
-
-<!-- Moreover, the measured variable in the considered data structure can be dicrete or continuous. -->
-
 Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
+However, we should have the consider the effect of number of observations on the summarization.
+Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
+Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
+There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
+If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
+The value of $k$ will depend on the statistical transformation we are producing.
+For computing deciles, $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
+Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
+
+<!-- ## Summarizing the distribution of a measured variable {#sec:measured-variable} -->
+
+<!-- <!-- Moreover, the measured variable in the considered data structure can be dicrete or continuous. -->
+
+<!-- Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
 Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries.
@@
 These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
-Several forms of non-parametric density estimators <!--and quantile estimators--> exist in the literature, the most commonly used form in @Silverman1986-cl<!--,  @Jones1992-pj)--> is as follows:
-
-<!-- The most commonly used density estimation method is the kernel density estimator.  -->
-
-If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$, <!--and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$,--> the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by $\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h$, where $h$ is the smoothing parameter or the bandwidth.
+Several forms of non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form in @Silverman1986-cl<!--,  @Jones1992-pj)is as follows:-->
+
+<!-- <!-- The most commonly used density estimation method is the kernel density estimator.  -->
+
+<!-- If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$, <!--and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$,the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by $\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h$, where $h$ is the smoothing parameter or the bandwidth.
 The MSE of the estimator is given by:
-\begin{equation}\label{eq:MSE}
-\begin{split}
-MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
-\end{split}
-\end{equation}
-
-where $\kappa = \int k^2(y)dy$.
+<!-- \begin{equation}\label{eq:MSE} -->
+<!-- \begin{split} -->
+<!-- MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
+
+<!-- where $\kappa = \int k^2(y)dy$. -->
 
@@
 
-```{r bothcust, fig.cap = "Area quantile plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936 displays the smoothed energy consumption curve for the entire day.
-The customer typically consumes more in the morning hours following which their consumption is low till late afternoon hours.
-Median consumption for the early morning hours starts higher and continue to remain high for longer during weekdays compared to weekdend."}
+```{r bothcust, fig.cap = "(a) Area quantile and (b) violin plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
+Plot (a) shows that the customer typically consumes more in the morning hours following which their consumption is low till late afternoon hours.
+Median consumption for the early morning hours starts higher and continue to remain high for longer during weekdays compared to weekdend.
+Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable.
+Bimodality can be caused by difference in behavior across different seasons.
+Volatility is highest from 7 to 13 hours as evident from the narrow violins."}
 
@@
 #   ggtitle("Energy consumption distribution for customer id: 10006704")
-cust2 <- smart_meter10 %>%
+cust2_quantile <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@
   scale_y_sqrt()
-cust2
-```
-
-```{r bothcust-lv, fig.cap ="Violin plots of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
-Early morning hours are bimodal implying both high and low consumptions are probable.
-Bimodality can be caused by difference in behavior across different seasons.
-Volatility is highest from 7 to 13 hours as evident from the narrow violins."}
 #
@@
 
-smart_meter10 %>%
+cust2_violin <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@
   scale_y_sqrt()
+
+ggarrange(cust2_quantile, cust2_violin, nrow = 2)
 ```

### 2020-03-06 main.Rmd
data structure section content cut

@@
 
+## Data structure for exploring two cyclic granularities
+
 A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
@@
 
-## Synergy of the cyclic granularities {#sec:synergy}
+## Harmonies and Clashes {#sec:synergy}
 
@@
 Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
-We will discuss few cases, where one or more of these $nm$ sets will be empty, the implication of which we will discuss in Section \ref{sec:vis-interaction}.
+Now, some situations can lead to some of these sets being empty.
+We will discuss few cases, where one or more of these $nm$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
+<!--The implication of these empty sets on visualization will be discussed in Section \ref{sec:vis-interaction}.
+-->
+
+\begin{definition}[Clash]\label{def:definition}
+A pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
+\end{definition}
+
+\begin{definition}[Harmony]\label{def:definition}
+A pair of cyclic granularities that do not contain any missing combinations of its categories.
+\end{definition}
 
@@
 
-<!-- Let's take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month.
-Here $C_1$ can take 31 values while $C_2$ can take 5 values.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM.
-Many of these are empty.
-For example $S_{1,5}$, $S_{21,2}$, etc.
-In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful.
-These are structurally empty sets in that it is impossible for them to have any observations.
--->
-
-<!-- Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year.
-Here $C_1$ can take 7 values while $C_2$ can take 12 values.
-So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY.
-All of these are non-empty because every DOW can occur in every month.
-So graphics involving $C_1$ and $C_2$ are potentially useful.
--->
-
-<!-- What  is interaction and some examples why it is important to study interaction -->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month. -->
-
-<!-- Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations.
-M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.
-Some observations that can be made from these letter-value plots include a) Right tails for most days in January lying above the median is more extended than the left tails contrary to days in April which has longer left tails.
-b) Days in mid January and February are characterized by high variation in consumption level, but days in January have longer right tails.
-c) Last five days in December (Christmas holidays) shows low variation in consumption with longer left tails implying people typically consume less electricity.
-We can conclude that Month-of-Year and Day-of-Month are harmonies.
--->
-
-<!-- \autoref{fig:allFig} (c) shows box plot of electricity consumption of Victoria from 2012 to 2014 across days of the year by the 1st, 15th, 29th and 31st days of the month.
-The box plot is a a very compact distributional summary, displaying median, quartile boundaries, hinges, whiskers and outliers.
-All facets do not contain data across same x-axis levels leading to difficulty in comparison across facets.
-Hence, Day-of-Month and Day-of-Year are clashes.
--->
-
-<!-- Take another example \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-Violin plots are a combination of box plot and density plot.
-Here, the first week of the month correspond to certain days of the month which are different for different week of the month.
-These kinds of graphics can hinder our comprehension across different weeks of the month and can be categorized as a incompatible combination to plot.
--->
-
-<!-- In \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot.
-The y-axis represents week of the year and the x-axis represents electricity consumption.
-Ridge plots are density plots all aligned to the same horizontal scale and presented with a slight overlap.
-They are designed to bring out changes in distributions over time.
-The data is distributed unequally for different levels of weeks of the year for different facets, which hinders comparison across facets.
-So, Week-of-Month and Week-of-Year are clashes.
--->
-
-<!-- \autoref{fig:allFig} (f) shows decile plots of consumption across  day of the year and month of the year.
-Decile plots are useful in displaying distribution through deciles without much clutter.
-This plot, however, seem ineffective for comparison as levels of x-axis for which observed data is plotted are completely disjoint across facets.
-So Month-of-Year and Day-of-Year are clashes as well.
--->
-
-<!-- In \autoref{fig:allFig} (c), we have empty combinations when we plot observations with "Day-of-Month" and "Day-of-Year".
-Here, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-Hence,  "Day-of-Month" and "Day-of-Year" are clashes due to the way they map to calendar.
--->
-
-<!-- In \autoref{fig:allFig} (a), we will not have any empty combinations because every DoM can occur in all MoY.
-Here, mapping from days to months is irregular in the sense that one month can consist of 28, 29, 30 or 31 days.
-There is no denying that the 29th, 30th or 31st day of the month needs to be analysed with caution due to the irregular mapping, however, the first 28 days of the month will occur in all months of the year.
--->
-
 <!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
@@
 
-<!-- Before exploring the behavior of a "dependent variable" across time granularities, -->
-<!-- it is important to know how these granularities interact with each other.
-If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable.
--->
-
 <!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
 -->
-
-<!-- When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations.
-It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.
--->
-
-<!-- Suppose we have two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.
-That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$.
-Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.
--->
-
-<!-- The cyclic time granularities which are essentially categorizations of time are discrete variable, mapped as a factor since they have an inherent order.
-Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
-To define a general framework, we define harmony and clashes.
--->
-
-<!-- Now, many situations can lead to some of these sets being empty. Let us discuss the following cases, where one or more of these sets can be empty. -->
 
@@
 This is also intuitive since the first day of the month can never correspond to fifth week of the month.
+Hence the pair (day-of-month, week-of-month) is a clash.
 
@@
 All of these are non-empty because every day-of-week can occur in every month.
-
-A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
-And the ones that do not lead to any missing combinations are referred as **harmonies**.
+Hence, the pair (day-of-week, month-of-year) is a harmony.
 
 Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
-However, we should have the consider the effect of number of observations on the summarization.
+However, we should have the consider the effect of number of observations on the summarization even for harmonies.
 Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.

### 2020-03-06 main.Rmd
cut literature review of different plot types

@@
 We will discuss few conventional and recent ways to plot distributions using both Kernel density estimates and descriptive statistics.
+Descriptive statistics based displays include box plots [@Tukey1977-jx] or different variations of it like Notched box plots [@Mcgill1978-hg].
+More recent ways are the letter-value box plot [@Hofmann2017-sg] or Quantile plots which display quantiles instead of quartiles in a traditional boxplot.
+Kernel density based plots for displaying a distribution include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, highest density regions (HDR) box [@Hyndman1996-ft].
+Each type of density display has different parameters, that need to be estimated given the data.
+Each is equipped with some benefits and challenges which should be borne in mind while using them for exploration.
 
@@
 
-### Descriptive statistics based displays
-
-Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers at a glance.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.--> Due to wide spread popularity and simplicity in implementation, a number of variations like Notched box plots [@Mcgill1978-hg] are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.<!--has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot.
+<!-- CONTENT CUT -->
+<!-- ### Descriptive statistics based displays -->
+
+<!-- Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers at a glance.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.
+Due to wide spread popularity and simplicity in implementation, a number of variations like Notched box plots [@Mcgill1978-hg] are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.<!--has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
+Challenge of boxplot.
 The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
--->.
 However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
 Also, the number of outliers is large for larger data sets.
-
-<!-- Description and lit review of letter-value plots -->
-The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.<!--“outliers” in letter value plots are those observations beyond the most extreme letter value.--> The letter values are displayed till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
-
-<!-- Description and lit review of quantile plots -->
-
-Quantile plots display quantiles instead of quartiles in a traditional boxplot.
+-->
+
+<!-- <!-- Description and lit review of letter-value plots --> 
+<!-- The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.<!--“outliers” in letter value plots are those observations beyond the most extreme letter value.
+The letter values are displayed till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
+-->
+
+<!-- <!-- Description and lit review of quantile plots -->
+
+<!-- Quantile plots display quantiles instead of quartiles in a traditional boxplot.
 These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
 While in a boxplot and letter value plots are well defined <!--an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
-outside 1.5 times the interquartile range above the upper quartile and below the lower quartile)-->, in quantile plots outliers are open to interpretation and not shown.
-
-<!-- Description and lit review of density plots -->
-
-<!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  -->
-
-### Kernel density based displays
-
-Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi].
+outside 1.5 times the interquartile range above the upper quartile and below the lower quartile), in quantile plots outliers are open to interpretation and not shown.
+-->
+
+<!-- <!-- Description and lit review of density plots -->
+
+<!-- <!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  --> -->
+
+<!-- ### Kernel density based displays -->
+
+<!-- Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi].
 The shape of the violin represents the density estimate of the variable.
 <!--The more data points in a specific range, the larger the violin is for that range.
--->Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
+Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
 It becomes less useful when there are too many levels or categories to display.
-
-The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
+-->
+
+<!-- The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
 This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
-
-<!-- Description and lit review of density plots -->
- A Ridge line plot <!--(sometimes called Joy plot)--> shows the distribution for several groups.
+-->
+
+<!-- <!-- Description and lit review of density plots -->
+<!--  A Ridge line plot <!--(sometimes called Joy plot) shows the distribution for several groups.
 Distribution can be represented using density plots, all aligned to the same horizontal scale and presented with a slight overlap.
@@
 Also, with lot of categories, it is difficult to compare the height of the densities across categories.
-
-<!-- Description and lit review of HDR plots -->
-The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
+-->
+
+<!-- <!-- Description and lit review of HDR plots -->
+<!-- The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
 The probabilities for which the summarization is required can be chosen based on the requirement.
 These regions do not need to be contiguous and help identify multi-modality.
-
-Each type of density display has different parameters, that need to be estimated given the data.
+-->
+
+<!-- Each type of density display has different parameters, that need to be estimated given the data.
 Each is equipped with some benefits and challenges which should be borne in mind while using them for exploration.
@@
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
+This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
 The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article.

### 2020-03-11 main.Rmd
minor changes in quasi-periodic linear granularities and linear calendar algebra

@@
 These are called quasi-periodic relationships.
-Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.
+
+<!-- CONTENT CUT -->
+<!-- Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$. -->
 
@@
 
-For example, in a Gregorian calendar without leap years each month is a grouping of the same number of days each year.
-The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$.
-Thus, we could say days group periodically into months.
-In a Gregorian calendar with leap years, days group quasi-periodically into months.
-The value of $P$ and $R$ would still be 365 and 12 respectively, while the <!--$E_i$s--> exceptions would be the time domain corresponding to $29^{\text{th}}$ February of any year.
+For example, in a Gregorian calendar without leap years we could say days group periodically into months with the period $P = 365$ and the number of granules of "months" in each period given by $R=12$.
+In a Gregorian calendar with leap years, days group quasi-periodically into months with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
 
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a *_bottom granularity_* and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
-Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic granularities.
+It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.
 
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
+Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
 

### 2020-03-11 main.Rmd
content cut till cyclic calendar algebra

@@
 \autoref{fig:linear-time} illustrates the linear time granularites hours, days, weeks, months and years, each of which are non-repeating in nature and arranged from past to future.
-If we consider the chronon [see @aigner2011visualization] as hourly, the time domain with $T$ hours will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
 
@@
 ```
-In addition to the integer index, each granule could be associated with a textual representation which is called the label.
-The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
+In addition to the integer index, each granule could be associated with a textual representation which is called the label.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.--> When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
 
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
-
-\begin{align*}
-\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \text{Alter}^{12}_{9,-1}(\text{day}, \\
-  &~~\text{Alter}^{12}_{6,-1}(\text{day}, \text{Alter}^{12}_{4,-1)}(\text{day}, \\
-  &~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \text{Group}_{31}(\text{day})))))),
-\end{align*}
-
+We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
 For more variations of calendar algebra operations, see @Ning2002-tf.
+
+<!-- CONTENT CUT -->
+<!-- \begin{align*} -->
+<!-- \text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \text{Alter}^{12}_{9,-1}(\text{day}, \\ -->
+<!--   &~~\text{Alter}^{12}_{6,-1}(\text{day}, \text{Alter}^{12}_{4,-1)}(\text{day}, \\ -->
+<!--   &~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \text{Group}_{31}(\text{day})))))), -->
+<!-- \end{align*} -->
 
@@
 An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-In a tsibble, each observation (row) is uniquely identified by index and key.
-
-A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the time index of a tsibble.
+<!--In a tsibble, each observation (row) is uniquely identified by index and key.--> A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the time index of a tsibble.
 A linear granularity is a mapping of the index set to subsets of the time domain.
 For example, if the time index of a tsibble is days, then a linear granularity might be weeks, months or years.
-A bottom granularity is represented by the index of the tsibble.
-Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.
-
-Cyclic time granularities with periodicities can be constructed by relating two linear granularities.
-The mappings between these linear granularities can be regular (minutes and hours) or irregular (days and months).
-<!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).--> A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.
-Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mapping is irregular.
+A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.-->
+
+Cyclic granularities can be constructed by relating two linear granularities.
+The mappings between these linear granularities could be _regular_ with every granule of coarser linear granularity being made up of same number of granules of finer linear granularity or _irregular_.
+<!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).
+A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.--> Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mappings are irregular.
 Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
@@
 Two linear granularities $G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))$ are considered.
-This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
-Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.
-$Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
+<!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
+Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> $Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
 For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
@@
 
-It can be noted here that if a linear granularity $G'$ is quasi-periodic (Bettini2000-vy) <!--(Definition \autoref{def:quasi})--> with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$.<!--(by Definition \autoref{def:quasi}).-->
-
-\begin{equation}\label{eq6}
-A_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert)
-\end{equation}
-
-for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
+<!-- CONTENT CUT -->
+<!-- It can be noted here that if a linear granularity $G'$ is quasi-periodic (Bettini2000-vy) <!--(Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$.<!--(by Definition \autoref{def:quasi}).-->
+
+<!-- \begin{equation}\label{eq6} -->
+<!-- A_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert) -->
+<!-- \end{equation} -->
+
+<!-- for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$. -->
 
@@
 
-Periodic and Quasi-periodic behavior can be defined by a repeating pattern, except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.
-Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
+<!--Periodic and Quasi-periodic behavior can be defined by a repeating pattern except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.--> Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
 Most public holidays repeat every year, but there is no finite (or reasonably small) period within which their behavior remains constant.
-A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
-U.S labor day is the first Monday in September and U.S.
+A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.<!--U.S labor day is the first Monday in September and U.S.
 Memorial Day is the last day in May.
-In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
-Examples of aperiodic granularity may also include school holidays or a scheduling event that might cover the first and third Monday of the month between June and October, except for state holidays.
-All of these are recurring events, but with non-periodic patterns.
+--> In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
+Examples of aperiodic granularity may also include school holidays or a scheduling event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns.
 As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
@@
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$, that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.
-Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$.
+<!--that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.--> Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
 Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
@@
 **Example**: \autoref{fig:aperiodic-example} shows two aperiodic linear granularity $M_1$ and $M_2$.
-Each granule is represented by a box.
+<!--Each granule is represented by a box.
 The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set and $A_{B, M}$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
-
-**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Lab our day and M can be conceived as the aperiodic linear granularity denoting any holidays.
-The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.
+-->The first slot in the diagram shows the index set and $A_{B, M}$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
+
+**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labor day and M can be conceived as the aperiodic linear granularity denoting any holidays.
+<!--The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.-->
 
@@
 In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
-In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
-This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
-Since, our method caters to computation of cyclic granularities, we shall refer to it as "cyclic calendar algebra".
+In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities <!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
+Since, our method caters to computation of cyclic granularities, we shall refer to it as--> through cyclic calendar algebra operations.
 
@@
 But it is far from being the only one.
-<!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->.
-The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
+<!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
@@
 The single-order-up granularity day-of-month is quasi-circular.
-Any single-order-up granularities which are formed by linear granularities below days are circular.
-Similarly, all single-order-up granularities which are formed using linear granularities with orders higher than months are also circular.
+<!--Any single-order-up granularities which are formed by linear granularities below days are circular.
+Similarly, all single-order-up granularities which are formed using linear granularities with orders higher than months are also circular.-->
 
@@
 
-Generally speaking, to keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
+To keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
 Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.

### 2020-03-12 main.Rmd
delete redundant statements application- smaet meter

@@
 output: 
-  bookdown::pdf_book:
-    base_format: rticles::asa_article
+  rticles::asa_article:
     fig_height: 5
@@
 - $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
-
 \begin{equation} \label{eq:multifromsingle-quasi1}
@@
 - $C_{l,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
-
 \begin{equation} \label{eq:multifromsingle-quasi2}
@@
 One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
-It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.
-The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns.
+<!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to look at the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
 <!-- and identify the extreme and regular households. -->
@@
 
-The data structure for a sample of 50 households from that trial is as follows:
+The data structure for a sample of 50 households from that trial is depicted in \ where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
 <!-- We take a random sample of 50 households to see how we can achieve that.  -->
@@
   )
-smart_meter %>% head(n = 5)
+smart_meter %>% head(n = 5) %>% knitr::kable()
 ```
 
-The variable `reading_datetime` is the index of the tsibble and represents the time variable.
-The variable `customer_id` denoting different households represent the key of the tsibble and `general_supply_kwh` is the time series variable which needs to be analyzed.
-Data sets of the form <$C1$, $C2$, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity $C1$ would be mapped to $x$ axis and `general_supply_kwh` to $y$ axis with the cyclic granularity $C2$ mapped to facet.
+<!--of the tsibble and represents the time variable.
+The variable  denoting different households represent the key of the tsibble and  is the time series variable which needs to be analyzed.Data sets of the form <$C1$, $C2$, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity $C1$ would be mapped to $x$ axis and `general_supply_kwh` to $y$ axis with the cyclic granularity $C2$ mapped to facet.
 Any geometry displaying the probability distribution of `general_supply_kwh` like  boxplot, letter value, violin,
-ridge or highest density region plots can then be used to explore the data set.
+ridge or highest density region plots can then be used to explore the data set.-->
 
@@
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-
-<!-- how gravitas is used to deal with systematic exploration -->
-
 These cyclic granularities are a function of the index variable and are computed based on if they are circular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}).
@@
 The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.
-
-```{r search, echo=FALSE}
+The cyclic granularities which could be considered are "hhour_hour", "hhour_day", "hhour_week",  "hhour_fortnight",  "hhour_month", "hhour_year", "hour_day", "hour_week", "hour_fortnight", "hour_month", "hour_year", "day_week", "day_fortnight",  "day_month" , "day_year", "week_fortnight",  "week_month", "week_year" , "fortnight_month", "fortnight_month" and  "month_year".
+These options are too many and hence the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour or fortnight could be dropped.
+Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:aperiodic-gran-def}.
+
+```{r search, echo=FALSE, eval=FALSE}
 library(gravitas)
@@
 <!-- #knitr::kable(smart_smart_search, format = "markdown") -->
-
-If these options are considered too many, the most coarse temporal unit can be set to a “month”.
-Also, some intermediate temporal units that might not be pertinent to the analysis can be removed from the list of cyclic granularities we want to look at.
 
@@
 
-```{r search_gran_limit2, echo=TRUE}
+```{r search_gran_limit2, echo=FALSE, eval = FALSE}
 smart_meter10 %>% search_gran(
@@
   filter_out = c("hhour", "fortnight")
-)
+)%>% knitr::kable(format = "markdown")
 ```
@@
 
-```{r harmony, echo=TRUE}
+```{r harmony, echo=FALSE}
 smart_meter10 %>% harmony(
@@
   filter_out = c("hhour", "fortnight")
-)
+) %>% knitr::kable(format = "markdown")
 ```
@@
 
-Consider looking at letter value plots instead of quantile plots to look at the same data in \autoref{fig:bothcust-lv}. There is additional information that we can derive looking at the distribution.
+Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust-lv}.
+There is additional information that we can derive looking at the distribution.
 There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
@@
 If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
-This is also intuitive because `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+This is also intuitive because heterogenous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
 A strategy for analyzing multiple keys together could be to first screen some basis time series or demographic features and then look at their energy behavior.

### 2020-03-12 main.Rmd
software implementation added

@@
 
-It is important to be able to navigate through all of these temporal granularities to have multiple perspectives on the periodicity of the observed data.
+It is important to be able to navigate through all of these time granularities to have multiple perspectives on the periodicity of the observed data.
 This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
@@
 Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.
+
+The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN. More details about the package 
+can be found on the package website https://sayani07.github.io/gravitas/.
+This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
+The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article and the code is also available online
+in the supplemental materials.
 
@@
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-These cyclic granularities are a function of the index variable and are computed based on if they are circular (Section \ref{sec:circular-gran-def}), pseudo-circular (Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def}).
-If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, fortnight, month, year), the following cyclic time granularities can be considered for this analysis.
+These cyclic granularities are a function of the index variable and are computed based on if they are circular <!--(Section \ref{sec:circular-gran-def})-->, pseudo-circular <!--(Section \ref{sec:quasi-circular-gran-def})--> or aperiodic <!--(Section \ref{sec:aperiodic-gran-def})-->.
+If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year), the following cyclic time granularities can be considered for this analysis.
 The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.
-The cyclic granularities which could be considered are "hhour_hour", "hhour_day", "hhour_week",  "hhour_fortnight",  "hhour_month", "hhour_year", "hour_day", "hour_week", "hour_fortnight", "hour_month", "hour_year", "day_week", "day_fortnight",  "day_month" , "day_year", "week_fortnight",  "week_month", "week_year" , "fortnight_month", "fortnight_month" and  "month_year".
-These options are too many and hence the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour or fortnight could be dropped.
+The cyclic granularities which could be considered are "hhour_hour", "hhour_day", "hhour_week",  "hhour_month", "hhour_year", "hour_day", "hour_week", "hour_month", "hour_year", "day_week", "day_month" , "day_year",  "week_month", "week_year" and  "month_year".
+These options are too many and hence the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped.
 Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
@@
 
-```{r search, echo=FALSE, eval=FALSE}
+```{r search, echo=FALSE, eval = FALSE}
 library(gravitas)
-smart_meter %>% search_gran(filter_out = c("semester", "quarter"))
+smart_meter %>% search_gran(filter_out = c("semester", "quarter", "fortnight"))
 ```
@@
 
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:aperiodic-gran-def}.
+<!-- Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:aperiodic-gran-def}.
+-->
 
@@
 
-Mumbai Indians (MI) and Chennai Super kings (CSK) are considered one of the best teams in IPL with multiple winning titles and always appearing in final playoffs from 2010 to 2015.
+Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
 It would be interesting to take their example in order to dive deeper into this question.
-Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.
-From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
+<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
 This might indicate that players are vulnerable to score through each over and make more runs in the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
@@
 For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
-If a batsman is bowled out, it does not necessarily signify good fielding.
+<!--If a batsman is bowled out, it does not necessarily signify good fielding.
 So we only include number of catches and run out in an over as a measure of good fielding.
 Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
-Figure \ref{fig:field} shows that at least one dismissal in a previous over leads to a lower median and less spread in the distribution of run rates of the subsequent over.
+--> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable.
+Figure \ref{fig:field} shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
 
@@
 
-```{r field,  fig.cap = "Violin plots with an overlaid boxplot showing distribution of run rate across overs of the innings and (lagged) wickets dismissed due to fielding.
-This shows good fielding in previous over leads to lower run rate in the subsequent over.
-Also distribution of run rates has less spread in case there is at least one wicket in the previous over.", warning=FALSE}
+```{r field,  fig.cap = "Box plots showing distribution of run rate across overs of the innings and (lagged) wickets dismissed due to fielding.
+This shows at least one wicket in the previous over leads to downward shift of median run rate and quartile spread in the subsequent over for most overs.", warning = FALSE}
 
 # filter(fielding_wckts %in% c(0,1)) %>%
-cricket_data %>% prob_plot("over", "lag_field",
+cricket_data %>%
+  filter(over!=1) %>% 
+  prob_plot("over", "lag_field",
   hierarchy_model,
   response = "run_rate",
-  plot_type = "violin",
-) + ggtitle("") + geom_boxplot(width = 0.5, aes(colour = lag_field)) + ylim(0, 4.5)
+  plot_type = "boxplot",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)
+) + ggtitle("") + scale_y_log10()
+#+ geom_boxplot(width = 0.5, aes(colour = lag_field)) 
 ```
@@
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
-The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article.
 

### 2020-03-12 main.Rmd
changed to pdf book and base_format: asa article + linear calendar algrabra section shrinked further

@@
 output:
-  rticles::asa_article:
+  bookdown::pdf_book:
+    base_format: rticles::asa_article
     fig_height: 5
@@
 Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.
+For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$.
+For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$.
+The _altering-tick operation_ generates new full-integer labeled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$.
+For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
+We can --> consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
+For more variations of calendar algebra operations, see @Ning2002-tf.
 
@@
 
-The grouping operation:
-: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
+<!-- CONTENT CUT -->
+<!-- The grouping operation: -->
+<!-- : Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
 The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
-$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$
+More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$, -->
+<!-- $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$ -->
+<!-- Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$ -->
 
@@
 
-The altering-tick operation:
-: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
+<!-- CONTENT CUT -->
+<!-- The altering-tick operation: -->
+<!-- : Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
 The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$.
 The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$.
-
-For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
+-->
+
+<!-- For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
 We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
 For more variations of calendar algebra operations, see @Ning2002-tf.
+-->
 

### 2020-03-12 main.Rmd
elements of grammar explaination cut short, levels cut short

@@
 The layered grammar of graphics proposed by [@Wickham2009pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
-There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
+Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
+
+<!-- CONTENT CUT -->
+<!-- There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
 Data are observations made on a number of variables, typically thought of as a sample from a population distribution.
@@
 Facets dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
+-->
 
@@
 
-Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
-
 - Data: <$C_1$, $C_2$, $v$>
-- Aesthetic mapping: $C_1$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v$
-- Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
-- Facet:  $C_2$
+- Aesthetic mapping (mapping of variables to elements of the plot): $C_1$ mapped to $x$ position and $v$ to $y$ position
+- Statistical transformation (summarize data): Any descriptive or smoothing statistics that summarizes distribution of $v$
+- Geometric objects (physical representation of the data): Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
+- Facet (split plots):  $C_2$
 
@@
 
-The following principles are useful while choosing distribution plots given two temporal granularities:
-
-  - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus.
-
- - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories.
-
-  - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.
-
-  - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.
+<!-- CONTENT CUT -->
+<!-- The following principles are useful while choosing distribution plots given two temporal granularities: -->
+
+<!--   - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus. -->
+
+<!--  - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories. -->
+
+<!--   - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred. -->
+
+<!--   - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.
+-->
 

### 2020-03-12 main.Rmd
one quasi-circular granularity content cut

@@
 
-Let us revisit Gregorian calendar for addressing this case.
-Suppose we have a hierarchy table using some linear granularties from Gregorian calendar.
+<!-- CONTENT CUT -->
+<!-- There can be three scenarios for obtaining multiple-order-granularities here:   -->
+<!--  - cyclic granularities relating two linear granularities whose orders are less than day   -->
+<!--  - cyclic granularities relating two linear granularities whose orders are more than month   -->
+<!--  - granularities relating two linear granularities with order at most day and another with order at least month -->
+
+<!-- The multiple order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:all-circular}.
+Any cyclic granularity resulting from the last case are quasi-circular.
+Examples might include hour-of-month or day-of-year.
+-->
+
+We consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
+Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.
+Two different approaches need to be employed for the following cases:
+
+<!-- For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, -->
+- $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+\begin{equation} \label{eq:multifromsingle-quasi1}
+\begin{split}
+C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\
+\end{split}
+\end{equation}
+
+- $C_{l,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
+\begin{equation} \label{eq:multifromsingle-quasi2}
+\begin{split}
+C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert))\\
+\end{split}
+\end{equation}
+
+where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+
+<!-- Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$. -->
+
+**Example:** Let us revisit Gregorian calendar for addressing this case.
+Suppose we have a hierarchy using linear granularties from Gregorian calendar.
 Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.
-This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities.
-The single-order-up granularity day-of-month is quasi-circular.
-<!--Any single-order-up granularities which are formed by linear granularities below days are circular.
-Similarly, all single-order-up granularities which are formed using linear granularities with orders higher than months are also circular.-->
-
-<!-- Hence, any calendar categorization which includes one linear granularity with orders at most as days and another one with orders at least as months will be aperiodic in nature.
-However, it is very much possible for a calendar categorization consisting of two linear granularities higher in order than month to be periodic.For example, the categorization month_quarter would be periodic since each quarter always consists of three months.
--->
+This is an example of a hierarchy structure which has circular and quasi-circular single-order-up granularities.
+Only one single-order-up granularity day-of-month is quasi-circular.
 
@@
 
-There can be three scenarios for obtaining multiple-order-granularities here:  
- - cyclic granularities relating two linear granularities whose orders are less than day  
- - cyclic granularities relating two linear granularities whose orders are more than month  
- - granularities relating two linear granularities with order at most day and another with order at least month
-
-The multiple order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:all-circular}.
-Any cyclic granularity resulting from the last case are quasi-circular.
-Examples might include hour-of-month or day-of-year.
-
-To keep it simple, we consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
-Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.
-Two different approaches need to be employed for the following cases:
-
-<!-- For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, -->
-- $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
-\begin{equation} \label{eq:multifromsingle-quasi1}
-\begin{split}
-C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\
-\end{split}
-\end{equation}
-
-- $C_{l,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
-\begin{equation} \label{eq:multifromsingle-quasi2}
-\begin{split}
-C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert))\\
-\end{split}
-\end{equation}
-
-where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-
-<!-- Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$. -->
-
-**Example:** Let us use Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2} to compute the multiple-order-up granularity hour-of-month and day-of-year for the Gregorian calendar considered.
+Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we have:
+
+<!-- to compute the multiple-order-up granularity hour-of-month and day-of-year for the Gregorian calendar considered. -->
 

### 2020-03-12 main.Rmd
data structure and rarely occuring events

@@
 
-## Data structure for exploring two cyclic granularities
+## Data structure for exploring two cyclic granularities {#sec:twocylcic}
 
@@
 
-## Harmonies and Clashes {#sec:synergy}
+### Harmonies and Clashes {#sec:synergy}
 
@@
 Hence, the pair (day-of-week, month-of-year) is a harmony.
+
+### Summarize distribution of measurements {#sec:summarize-measured}
 
@@
 
-<!-- <!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  --> -->
+<!-- <!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  -->
 
@@
 
-Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occuring categories or unevenly distributed events.
-
-### Rarely occuring events
-
-Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
+Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occuring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
+
+<!-- CONTENT CUT -->
+<!-- ### Rarely occuring events -->
+
+<!-- Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
 Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
@@
 Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
+-->
 
@@
 This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the statistical transformations based on density should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators (see \autoref{eq:MSE}) <!--  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$--> is a function of the sample size $n$.
+In such cases, the statistical transformations based on density should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators.
+<!-- (see \autoref{eq:MSE})  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$is a function of the sample size $n$.-->
 
@@
 This is also intuitive because heterogenous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
-A strategy for analyzing multiple keys together could be to first screen some basis time series or demographic features and then look at their energy behavior.
+A strategy for analyzing multiple keys together could be to first group them basis time series or demographic features and then look at their energy behavior.
 This is beyond the scope of the current work.
@@
 There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
-We will explore a few and understand how the proposed approach in the paper can help answer some of the questions.
+<!--We will explore a few interesting questions and understand how the proposed approach in the paper can help answer some of the questions.-->
 
@@
 <!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
-This might indicate that players are vulnerable to score through each over and make more runs in the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
+This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
 To answer Q1, winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.
@@
 
-In Q2, wickets per over are considered as aperiodic cyclic granularities with dot balls or catches as aperiodic linear granularities.
-These aperiodic linear granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
+In Q2, wickets per over are considered as aperiodic cyclic granularities with wickets as aperiodic linear granularities.
+These granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
 These are similar to holidays or special events in temporal data.
-While any special event that corresponds to a time domain can be treated as an aperiodic linear granularity in a temporal case, dot balls or wickets that corresponds to certain balls (index) could be treated as aperiodic events in cricket.
+<!-- CONTENT CUT -->
+<!--While any special event that corresponds to a time domain can be treated as an aperiodic linear granularity in a temporal case, dot balls or wickets that corresponds to certain balls (index) could be treated as aperiodic events in cricket.-->
 
@@
 
-The infrastructure of systematic exploration of observed variable across pair of cyclic time granularities has been implemented in the
+<!--The infrastructure of systematic exploration of observed variable across pair of cyclic time granularities has been implemented in the
 R package `gravitas`.
-A missing piece of the gravitas package is to enable user-defined temporal calendars and compute multiple order-up quasi-circular granularities from the hierarchy table.
+--> A missing piece is to enable user-defined temporal calendars and compute multiple order-up quasi-circular granularities from the hierarchy table.
 Also, computation of cyclic aperiodic granularities would require computing aperiodic linear granularities first.

### 2020-03-12 main.Rmd
include interchangeability mapping

@@
 
-### Interaction {#sec:vis-interaction}
-
-We will discuss the effect of the interaction of cyclic granularities in this section.<!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
-In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
-\autoref{fig:allFig}.
+### Synergy of cyclic granularities {#sec:vis-interaction}
+
+If there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
+We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
+Harmonies are pairs of granularities that do not contain emptry combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations.
+As a result, we should avoid plotting clashes.
+
+<!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
+In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in \autoref{fig:allFig}.
 \autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
+This plot works because every day-of-week can occur in every month-of-year.
+
 <!---\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month.
 \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig}---> (b) shows decile plots across day of the year and month of the year.
-
-Clearly, in \autoref{fig:allFig}, we observe that choice (a) of cyclic time granularities work but (b) does not work.
-In (b) we can not compare the distributions across different facets becuase every day-of-year cannot occur for every month.
-For example, 21st day of the year can never correspond to the month March or December of any year.
+\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig}(b) shows decile plots across day of the year and month of the year.--->
+
+<!-- Clearly, in \autoref{fig:allFig}, we observe that  -->
+
+<!--In (b) we can not compare the distributions across different facets becuase every day-of-year cannot occur for every month.
+For example, 21st day of the year can never correspond to the month March or December of any year.-->
 <!---In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year.
 In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-On the contrary, for \autoref{fig:allFig}---> In (a) we will not have any combinations because every day-of-week can occur in every month-of-year.
-<!---Thus the graphs that don't work are those where many of the combination sets are empty.--->
-
-In other words, if there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
-We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
-
-We redefine harmony and clashes as follows: harmonies are pairs of granularities that do not contain emptry combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations.
-As a result, we should avoid plotting clashes.<!---as they hinder the exploratory process by having missing combinations of time granularities in each panel.--->
+On the contrary, for \autoref{fig:allFig}--->  <!---Thus the graphs that don't work are those where many of the combination sets are empty.--->
+
+<!---as they hinder the exploratory process by having missing combinations of time granularities in each panel.--->
 
@@
 
-# change all of these to gravitas function
 library(gravitas)
@@
   create_gran("month_year") %>%
-  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>%
+  filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  create_gran("day_week") %>%
+  filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
   prob_plot("month_year", "day_week",
@@
   ylab("Electricity Demand [KWh]") +
-  ggtitle("(a) Letter value plot")
+  ggtitle("(a) Letter value plot with month-of-year facet")
 
@@
 
+# VIC %>%
+#   create_gran("month_year") %>%
+#   filter(month_year %in% c("Jan", "Jul", "Nov")) %>%
+#   prob_plot("month_year",
+#     "day_year",
+#     response = "Demand",
+#     plot_type = "quantile",
+#     quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
+#     symmetric = FALSE
+#   ) +
+#   # scale_x_discrete(breaks = seq(1, 336, 60)) +
+#   ylab("") +
+#   xlab("Day of the Year") + theme(
+#     legend.position = "bottom",
+#     strip.text = element_text(
+#       size = 7,
+#       margin = margin()
+#     )
+#   ) +
+#   ggtitle("(b) Decile plot")
+
 VIC %>%
   create_gran("month_year") %>%
-  filter(month_year %in% c("Jan", "Jul", "Nov")) %>%
-  prob_plot("month_year",
-    "day_year",
+  filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  create_gran("day_week") %>%
+  filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("day_week",
+    "month_year",
     response = "Demand",
-    plot_type = "quantile",
-    quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-    symmetric = FALSE
+    plot_type = "lv"
   ) +
-  # scale_x_discrete(breaks = seq(1, 336, 60)) +
   ylab("") +
-  xlab("Day of the Year") + theme(
-    legend.position = "bottom",
-    strip.text = element_text(
-      size = 7,
-      margin = margin()
-    )
-  ) +
-  ggtitle("(b) Decile plot")
+  xlab("Day of the Year") +
+  ggtitle("(b) Letter value plot with day-of-week facet")
 
@@
 With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
+\autoref{fig:allFig} (a) shows the letter value plot across days of the week faceted by months like January, March and December and \autoref{fig:allFig} (b) shows letter value of across months faceted by days.
+The first one helps us to compare different days of the weeks within a month and the second one helps to compare different months for each day of the week.
 

### 2020-03-13 main.Rmd
acknowledgment part shifted to end

@@
 
-The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN. More details about the package 
-can be found on the package website https://sayani07.github.io/gravitas/.
-This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
-The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article and the code is also available online
-in the supplemental materials.
-
 # Linear time granularities {#sec:linear-time}
@@
 Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.
-For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$.
+
+<!-- CONTENT CUT -->
+<!-- For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$.
 For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$.
@@
 For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-We can --> consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
+We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
 For more variations of calendar algebra operations, see @Ning2002-tf.
+-->
 
@@
 
-<!-- CONTENT CUT -->
-<!-- The grouping operation: -->
-<!-- : Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
+The grouping operation:
+: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
 The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$, -->
-<!-- $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$ -->
-<!-- Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$ -->
+More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
+$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
+Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$
 
@@
 
-<!-- CONTENT CUT -->
-<!-- The altering-tick operation: -->
-<!-- : Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
+The altering-tick operation:
+: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
 The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$.
 The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$.
--->
-
-<!-- For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
+
+For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
 We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
 For more variations of calendar algebra operations, see @Ning2002-tf.
--->
 
@@
     runs_per_over = sum(total_runs),
-    run_rate = sum(total_runs) / length(total_runs),
+    run_rate = sum(total_runs)*6 / length(total_runs),
     fielding_wckts = sum(fielding_proxy),
@@
  quantile_prob = c(0.25, 0.5, 0.75)
-) + ggtitle("") + scale_y_log10()
+) + ggtitle("") 
 #+ geom_boxplot(width = 0.5, aes(colour = lag_field)) 
@@
 
+# Software implementation {#sec:software}
+
+The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN.
+
 # Acknowledgements {-}
@@
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
+More details about the package
+can be found on the package website https://sayani07.github.io/gravitas/.
+This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
+The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article and the code is also available online
+in the supplemental materials.
 

### 2020-03-15 main.Rmd
finer, coarser and bottom granularity defined informally

@@
 
-Discrete abstraction of time such as hours, days, weeks and months can be thought of as linear "time granularities" [@aigner2011visualization].
-Time granularities are **linear** if they respect the linear progression of time.
+Discrete abstraction of time such as weeks, months or holidays can be thought of as "time granularities". Time granularities are **linear** if they respect the linear progression of time.
 
@@
 
-A discrete time model often uses a fixed smallest granularity also known as _bottom granularity_ (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
-Chronons may be grouped into granules and a non-overlapping mapping of granules to subsets of the time domain is a linear granularity.
+A discrete time model often uses a fixed smallest linear granularity also known as _bottom granularity_ (@Bettini1998-ed) <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
+Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
 A system of multiple granularities in lattice structures is referred to as a _calendar_ (see @Dyreson2000-mr).
@@
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
-It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom up.
+It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one.
 Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.
@@
 For more variations of calendar algebra operations, see @Ning2002-tf.
--->
 
@@
 : Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
-The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$.
-The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$.
+The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
 

### 2020-03-15 main.Rmd
revisions till cyclic time granularities

@@
 
-```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="The time domain distributed as linear granularities."}
+```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="The time domain distributed as linear granularities hours, days, weeks and months.
+These are ordered with ordering guided by integers and hence is unidirectional and non-repeating."}
 knitr::include_graphics("Figs/linear-arr.png")
@@
 
-We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data.
+We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] data structure.
 A tsibble consists of an index, key and measured variables.
@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
-knitr::include_graphics("Figs/circular-dow.png")
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities Day and Week and circular granularity Day-of-Week with $P=7$ is constructed using the two linear granularities.
+Day-of-Week has a repeating pattern unlike Day and Week and has an ordering only within a period represented by hashed arrows."}
+knitr::include_graphics("Figs/circular-dow-arrow.png")
 ```
@@
 \end{center}
-\caption{Illustrative circular granularities with time index in minutes}
+\caption{Illustrative circular granularities with bottom granularity "minutes".
+These circular granularities relate two linear granularities one of which groups periodically into the other with regular mapping.}
 \label{tab:definitions}
@@
 \end{center}
-\caption{Illustrative quasi-circular granularities with potential period lengths}
+\caption{Illustrative quasi-circular granularities with potential period lengths.
+These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping.
+Illustrative quasi-circular granularities }
 \label{tab:quasi}
@@
 
-```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="Quasi-circular granularities $Q(B, G')$ and $Q(B, H')$ relating bottom granularity $B$ and linear granularities $G'$ and $H'$ respectively."}
-knitr::include_graphics("Figs/quasi-circular-example.png")
+```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="The time domain distributed as linear granularities $B$, $G'$, $H'$ and quasi-circular granularities $Q(B, G')$ and $Q(B, H')$.
+Both $Q(B, G')$ and $Q(B, H')$ have repeating patterns unlike $G'$ or $H'$ and the patterns are guided by the distribution of $B$ within $G'$ and $H'$ within a period shown by the hashed arrow."}
+knitr::include_graphics("Figs/quasi-circular-ex.png")
 ```
@@
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$.
-<!--that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.--> Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ such that \ref{def:periodical} does not hold true for them with respect to $B$ (the bottom granularity).
+<!--that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.--> However, $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
 Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{i=1}^{n}M_{i}$.
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{i=1}^{n}M_{i}$.
 \end{definition}
@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Aperiodic cyclic granularity $A(B, M1)$ relating aperiodic linear granularity $M1$ and bottom granularity $B$"}
-knitr::include_graphics("Figs/aperiodic-example.png")
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear aperiodic granularities $B$, $M1$, $M2$, $M$ and aperiodic cyclic granularity $A(B, M)$.
+$A(B, M)$ have repeating patterns with three categories, two of which correspond to the aperiodic linear granularities $M1$ and $M2$ and other one correponding to no aperiodic granularity."}
+knitr::include_graphics("Figs/aperiodic-ex.png")
 ```

### 2020-03-15 harmony_select.R

@@ -1,0 +1,27 @@
+harmony_data <- smart_meter10 %>% harmony(
+  ugran = "month",
+  filter_out = c("hhour", "fortnight")
+) 
+
+for(i in 1:nrow(harmony_data))
+{
+  
+ data_new <- create_gran(smart_meter10, harmony_data[1,1]$facet_variable)%>% create_gran(harmony_data[1,2]$x_variable)
+ 
+data_new$hour_day = as.numeric(data_new$hour_day)
+density(data_new$hour_day)
+data_new$day_week = as.numeric(data_new$day_week)
+density(data_new$day_week)
+
+data_new[-c(1:2)]  %>% 
+  pivot_wider(names_from = day_week,
+              values_from = general_supply_kwh,
+              values_fn = list(general_supply_kwh = list)) %>% 
+  map(unlist) 
+
+
+%>% map_dfr(unlist)
+  
+    
+}
+    

### 2020-03-19 main.Rmd
hierarchy table and order re-write

@@
 
-In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
+In Section \ref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities <!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
@@
 
-**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy table containing $n$ linear granularities.
-$G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in {1, 2, \dots,n}$ and $l<m$.
-
-**Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity.
-For example, if we consider two linear granularities $G$ and $H$, such that $G$ is finer than or groups into $H$, then $H$ is of higher order than $G$.
-In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.
-
-We refer to granularities which are nested within multiple levels of the hierarchy table as _multiple-order-up_ granularities and those concerning a single level as _single-order-up_ granularities.
+**Order** of a linear granularity can be comprehended as the level of coarseness associated with a linear granularity.
+For two linear granularities $G$ and $H$, if $G$ groups into $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+In any hierarchy table, linear granularities are arranged from lowest to highest order.
+
+**Hierarchy table** Let $H_n: (G, C, K)$ be a hierarchy table containing $n$ linear granularities.
+$G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
+$K \equiv P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ for circular granularity $C_{G(l),G(m)}$ and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
+We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
+It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.
+
+We refer to granularities which are nested within multiple levels of the hierarchy table as _multiple-order-up_ and those concerning a single level as _single-order-up_ granularities.
 

### 2020-03-19 main.Rmd
till cyclic calendar algebra

@@
 Nevertheless, for any calendar a hierarchy table can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun
-
-Thus, the hierarchy table for the Mayan calendar would look like the following:
-
-|  G	        |C                 |P
-|------------	|------------------|------------|
-|  kin      	|kin-of-uinal      |  20        |
-|  uinal    	|uinal-of-tun      |  18  	    |
-|  tun      	|tun-of-katun      |  20	      |
-|  katun    	|katun-of-baktun   |  20        |
-|  baktun   	| 1                |	1         |
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
+\autoref{tab:tab-mayan} represents the the hierarchy table for the Mayan calendar.
+
+```{r tab-mayan}
+table_mayan <- tibble(G = c("kin", "uinal", "tun", "katun", "baktun"),
+                      C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
+                      K = c(20, 18, 20, 20, 1)) 
+kable(table_mayan, 
+      caption = "Hierarchy table for Mayan calendar.
+$G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
+All cyclic granularities are circular with $K$ as the period length.")
+```
 
@@
 Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
-The label mapping of the resultant multiple-order-up granularity can however be chosen arbitrarily, depending on the context.
+The label mapping of the resultant multiple-order-up granularity can however be chosen depending on the context.
 
@@
 
+, for $l < m + 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$
+
 **Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
@@
 \begin{split}
-C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + C_{uinal, katun}C_{katun, baktun}(z)) \\
-               = & L(\lfloor z/20\rfloor \mod 18 + 20*\lfloor z/20*18\rfloor \mod 20  \\
-               &~~~ + 20*18*20\lfloor z/20*18*20\rfloor \mod 20) \\
+C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + P(uinal, katun)C_{katun, baktun}(z)) \\
+               = & L(\lfloor z/20\rfloor \mod 18 + 18*\lfloor z/20*18\rfloor \mod 20  \\
+               &~~~ + 18*20\lfloor z/20*18*20\rfloor \mod 20) \\
 \end{split}
@@
 We consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
-Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be considered to be a function of two single order-up granularities $C_{l,m'}(z)$ and $C_{m',m}(z)$.
+Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
 Two different approaches need to be employed for the following cases:
@@
 
-**Example:** Let us revisit Gregorian calendar for addressing this case.
-Suppose we have a hierarchy using linear granularties from Gregorian calendar.
+**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from Gregorian calendar.
 Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.
-This is an example of a hierarchy structure which has circular and quasi-circular single-order-up granularities.
+This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities.
 Only one single-order-up granularity day-of-month is quasi-circular.
 
-|  G	        |C                 |P
+|  G	        |C                 |K
 |------------	|------------------|----------------|
@@
 |  hour    	  |hour-of-day       |  24  	        |
-|  day      	|day-of-month      |quasi-circular  |
+|  day      	|day-of-month      | k(day, month)  |
 |  month    	|month-of-year     |  12            |
@@
 
-For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
+For a hierarchy table $H_n: (G, C, K)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
 
 \begin{equation} \label{eq:all-circular-multiple}
-C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
+C_{G_{l_1}, G_{m_1}}(z) = C_{G_{l_2}, G_{m_2}}(\lfloor z/k(l_2,l_1) \rfloor\mod P(m_1, m_2))
 \end{equation}
 
-**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}
+**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
 

### 2020-03-19 main.Rmd
data structure section update

@@
 Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.
-However, restructuring time in this manner leads to re-organisation of the data structure, where each level of a cyclic granularity correspond to multiple values of the measured variable, unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
-It is common to see consolidation of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
-But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the probability distribution of these multiple observations across cyclic granularities using some summary statistics together, each of which can highlight different features of the data could be a potential way to comprehend periocities in the data.
-Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
-Instead, the big problem is broken down by focusing on two cyclic granularities at a time.
-
-<!--Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.-->
-<!-- Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
--->
-
-## Data structure for exploring two cyclic granularities {#sec:twocylcic}
-
 A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
@@
 Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+
+<!--Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.-->
+<!-- Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
+Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
+-->
 
@@
 
-### Summarize distribution of measurements {#sec:summarize-measured}
-
-Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
-However, we should have the consider the effect of number of observations on the summarization even for harmonies.
+### Summarizing the measured variable {#sec:summarize-measured}
+
+Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity correspond to multiple values of the measured variable.
+This is unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
+It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
+But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
+Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities could be a potential way to compare and contrast periocities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
+Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
+
+However, we need to consider the effect of number of observations on the summarization even for harmonies.
 Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
@@
 There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
+If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge mnk$ will be acceptable.
 The value of $k$ will depend on the statistical transformation we are producing.
-For computing deciles, $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
+For computing deciles, $k=10$ may be acceptable, but for density estimates, $k \ge 30$ is suggested.
 Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.

### 2020-03-19 main.Rmd
update before application

@@
 Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
-Now, some situations can lead to some of these sets being empty.
+Now, some situations can lead to few or many of these sets being empty.
 We will discuss few cases, where one or more of these $nm$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
@@
 - Aesthetic mapping (mapping of variables to elements of the plot): $C_1$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation (summarize data): Any descriptive or smoothing statistics that summarizes distribution of $v$
+- Statistical transformation (data summarization): Any descriptive or smoothing statistics that summarizes distribution of $v$
 - Geometric objects (physical representation of the data): Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
@@
 We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
-Harmonies are pairs of granularities that do not contain emptry combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations.
+Harmonies are pairs of granularities that do not contain empty combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations.
 As a result, we should avoid plotting clashes.
@@
 
-<!--In (b) we can not compare the distributions across different facets becuase every day-of-year cannot occur for every month.
+<!--In (b) we can not compare the distributions across different facets because every day-of-year cannot occur for every month.
 For example, 21st day of the year can never correspond to the month March or December of any year.-->
@@
 
-```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014.
-(a) Letter value plot by day-of-month and month-of-year (b) Decile plot by day-of-year and month-of-year.
-Plot (a) and (b) show harmony and clashes respectively.", fig.show = 'hold'}
+```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Letter value plot of electricity consumption data of Victoria from 2012 to 2014 (a) across day-of-week faceted by month-of-year and (b) across month-of-year faceted by day-of-week.
+Plots (a) and (b) both show harmonies since each month consists of all days of the week and every days of the week can occur in every month.
+However, Plot (a) helps comparison of different days of the week within each month and Plot (b) facilitates comparing months of the year for different days of the week.", fig.show = 'hold'}
 
@@
   ylab("Electricity Demand [KWh]") +
+  xlab("day-of-week") + 
   ggtitle("(a) Letter value plot with month-of-year facet")
@@
   ylab("") +
-  xlab("Day of the Year") +
+  xlab("month-of-year") +
   ggtitle("(b) Letter value plot with day-of-week facet")
@@
 If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
-This is also intuitive because heterogenous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+This is also intuitive because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
 A strategy for analyzing multiple keys together could be to first group them basis time series or demographic features and then look at their energy behavior.

### 2020-03-19 main.Rmd
supplementary materials section added

@@
 This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
+The ideas presented in this article have been implemented in the open source R package gravitas.
+R code and data are available in the supplementary materials.
+
 bibliography: [bibliography.bib]
@@
 
-The data structure for a sample of 50 households from that trial is depicted in \ where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
+The data structure for a sample of 50 households from that trial is depicted in \autoref{tab:smart-data} where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
 <!-- We take a random sample of 50 households to see how we can achieve that.  -->
 
-```{r smart-meter, fig.cap = "Smart meter data structure"}
+```{r smart-data}
 load("data/sm_cust50.Rdata")
 
-smart_meter <- sm_cust50 %>%
+smart_data <- smart_meter <- sm_cust50 %>%
   as_tsibble() %>%
@@
     general_supply_kwh
-  )
-smart_meter %>% head(n = 5) %>% knitr::kable()
+  ) %>% head(n = 5)
+
+  knitr::kable(smart_data,
+               caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
 ```
@@
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-These cyclic granularities are a function of the index variable and are computed based on if they are circular <!--(Section \ref{sec:circular-gran-def})-->, pseudo-circular <!--(Section \ref{sec:quasi-circular-gran-def})--> or aperiodic <!--(Section \ref{sec:aperiodic-gran-def})-->.
-If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year), the following cyclic time granularities can be considered for this analysis.
+These cyclic granularities are a function of the index variable and are computed based on if they are circular <!--(Section \ref{sec:circular-gran-def})-->, quasi-circular.
+<!--(Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def})--> If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year), the following cyclic time granularities can be considered for this analysis.
 The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.
-The cyclic granularities which could be considered are "hhour_hour", "hhour_day", "hhour_week",  "hhour_month", "hhour_year", "hour_day", "hour_week", "hour_month", "hour_year", "day_week", "day_month" , "day_year",  "week_month", "week_year" and  "month_year".
+The cyclic granularities which could be considered are hhour_hour, hhour_day, hhour_week, hhour_month, hhour_year, hour_day, hour_week, hour_month, hour_year, day_week, day_month, day_year, week_month, week_year and month_year.
 These options are too many and hence the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped.
 Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:aperiodic-gran-def}.
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} or \ref{sec:cyclic-calendar}.
 
@@
 library(gravitas)
-smart_meter %>% search_gran(filter_out = c("semester", "quarter", "fortnight"))
+smart_meter %>%
+  search_gran(filter_out = c("semester", "quarter", "fortnight"))
 ```
@@
 
-### Screening and visualizing harmonies:
+### Screening and visualizing harmonies
 
@@
 However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together.
-We are left with 13 harmonies pair, each of which can be plotted together to look at the energy behavior from different perspectives.
+\autoref{tab:harmony-tab} shows 13 harmonies pairs, each of which can be plotted together to look at the energy behavior from different perspectives.
 
@@
 
-```{r harmony, echo=FALSE}
-smart_meter10 %>% harmony(
-  ugran = "month",
-  filter_out = c("hhour", "fortnight")
-) %>% knitr::kable(format = "markdown")
+```{r harmony-tab, echo=FALSE}
+smart_harmony <- smart_meter10 %>% 
+  harmony(ugran = "month",
+          filter_out = c("hhour", "fortnight"))%>%
+  rename(`facet variable` = facet_variable,
+         `x-axis variable` = x_variable,
+         `Levels of facet variable` = facet_levels,
+         `Levels of x-axis variable` = x_levels)
+
+  knitr::kable(smart_harmony,
+                caption = "Harmonies with two cyclic granularity one placed on the facet and the other on x-axis with the corresponding levels.
+Out of 30 possible combinations of cyclic granularities that could be plotted only 13 are harmonies.")
 ```
@@
 
-# Software implementation {#sec:software}
-
-The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN.
-
 # Acknowledgements {-}
@@
 
+# Supplemental Materials
+
+**Data and scripts:** Datasets and R code to reproduce all figures in this article (main.Rmd).
+
+**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN.
+The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, check feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommend prospective probability distributions through factors described in the article.
+Version 0.1.2 of the package was used for the results presented in the article and is available from Github (https://github.com/Sayani07/gravitas).
+
+**R-packages:** Each of the R packages used in this article are available online (URLs are provided in the bibliography).
+
 # Bibliography

### 2020-03-19 main.Rmd
corrections done till end.

@@
 
-\begin{definition}[Clash]\label{def:definition}
+\begin{definition}[Clash]\label{def:harmony}
 A pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
@@
 
-\begin{definition}[Harmony]\label{def:definition}
+\begin{definition}[Harmony]\label{def:clash}
 A pair of cyclic granularities that do not contain any missing combinations of its categories.
@@
 
-In \autoref{fig:bothcust}, the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot.
+In \autoref{fig:bothcust} (b), the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot.
 The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
@@
 
-Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust-lv}.
+Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust}(c).
 There is additional information that we can derive looking at the distribution.
@@
 If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
-For example in \autoref{fig:bothcust-lv}, it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
+For example in \autoref{fig:bothcust}(a), it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
 Also, outliers are more prominent in the latter part of the day.
@@
 
-```{r bothcust, fig.cap = "(a) Area quantile and (b) violin plot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
-Plot (a) shows that the customer typically consumes more in the morning hours following which their consumption is low till late afternoon hours.
-Median consumption for the early morning hours starts higher and continue to remain high for longer during weekdays compared to weekdend.
-Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable.
-Bimodality can be caused by difference in behavior across different seasons.
-Volatility is highest from 7 to 13 hours as evident from the narrow violins."}
+```{r bothcust, fig.cap = "(a) Area quantile and (b) violin plot and (c) box plot of energy consumption for customer id: 10017936.
+Plot (a) helps to compare different hours of the day within weekdays and weekends.
+Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable due to different seasons.
+Plot (c) shows the effect of change in mapping which leads to easier comparison between weekdays and weekends for each hour."}
 
@@
   ) +
-  scale_y_sqrt()
+  scale_y_sqrt() +
+  ggtitle(" (b) Area quantile plot faceted by weekend-weekday")  + ylab("Electricity Demand [KWh]")
 #
@@
   ) +
-  scale_y_sqrt()
-
-ggarrange(cust2_quantile, cust2_violin, nrow = 2)
-```
-
-```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
-The change in mapping of the cyclic granularities leads to easier comparison within weekdays and weekends for each hour.
-Median consumption till 6am is higher for weekends compared to weekdays although the quartiles look similar.
-Outliers appear more during the end of the day implying more uncertain behavior in latter hours of the day compared to early morning hours."}
-#
-# smart_meter10 %>%
-#   filter(customer_id %in% c(10006704)) %>%
-#   prob_plot("wknd_wday",
-#                             "hour_day",
-#                             response = "general_supply_kwh",
-#                             plot_type = "violin") +
-#   scale_y_sqrt() +
-#   ggtitle("Energy consumption distribution for customer id: 10006704")
+  scale_y_sqrt()+
+  ggtitle(" (c) Violin plot faceted by weekend-weekday") + ylab("")
 
@@
   ) +
-  scale_y_sqrt()
-# scale_fill_brewer(palette = "Dark2")
+  scale_y_sqrt()+
+  ggtitle(" (a) Box plot faceted by hour-of-day")  + ylab("Electricity Demand [KWh]")
+
+ggarrange(cust2_quantile, cust2_violin, nrow = 2)
 ```
+
+<!-- # ```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
+The change in mapping of the cyclic granularities leads to easier comparison within weekdays and weekends for each hour.
+Median consumption till 6am is higher for weekends compared to weekdays although the quartiles look similar.
+Outliers appear more during the end of the day implying more uncertain behavior in latter hours of the day compared to early morning hours."} -->
+<!-- #  -->
+<!-- # smart_meter10 %>% -->
+<!-- #   filter(customer_id %in% c(10017936)) %>% -->
+<!-- #   prob_plot("hour_day", -->
+<!-- #     "wknd_wday", -->
+<!-- #     response = "general_supply_kwh", -->
+<!-- #     plot_type = "boxplot" -->
+<!-- #   ) + -->
+<!-- #   scale_y_sqrt() -->
+<!-- # ``` -->
 
@@
 The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
-This could be useful in a periodic world when we wish to compute any circular/quasi-circular granularity based on a hierarchy table which look like the following:
-
-```{r hierarchy2}
+This could be useful in a periodic world when we wish to compute any circular/quasi-circular granularity based on a hierarchy table which look like \autoref{tab:hierarchy-cric}.
+
+```{r hierarchy-cric}
 library(gravitas)
@@
   C = c("over-of-inning", "inning-of-match", "match-of-season", 1),
-  P = c(20, 2, "aperiodic", 1)
+  K = c(20, 2, "k(match, season)", 1)
 )
-knitr::kable(hierarchy_model, format = "markdown")
+knitr::kable(hierarchy_model,
+             caption = "Hierarchy table for cricket where each over is nested within innings, innings is nested within match and so on.
+Cyclic granularity over-of-inning, inning-of-match are assumed to be circular which means each inning will have 20 overs and each match has 2 innings and  match-of-season is assumed to be quasi-periodic.
+")
 ```
@@
 
-Q1: How total runs vary depending on if a team bats first or second?
-
+First, it would be interesting to see if the distribution of total runs vary depending on if a team bats first or second?
 Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
@@
 
-```{r cricex, fig.cap= "Letter value plot of runs per over across overs of the inning faceted by innings of the match.
-No upward shift in runs in the second innings like that in the first implying teams are more vulnerable to score more in the first innings as they approach the end of the inning.", warning=FALSE, message = FALSE}
+Another interesting question could be if runs per over set to reduce in subsequent over if fielding is good in the previous over?
+For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
+<!--If a batsman is bowled out, it does not necessarily signify good fielding.
+So we only include number of catches and run out in an over as a measure of good fielding.
+Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
+--> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable.
+Figure \ref{fig:field} shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
+
+<!-- Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.  -->
+
+<!-- fetching raw data since dot and fielding information not available in dismissal type not available in cricket data gravitas. Also `cricket` is aggregated across overs -->
+
+```{r cricex, fig.cap= "(a) Letter value plot shows that there is no upward shift in runs in the second innings like that in the first implying teams are vulnerable to score more in the first innings as they approach the end of the inning.
+(b) Box plot shows at least one wicket in the previous over leads to downward shift of median run rate and quartile spread in the subsequent over for most overs.", warning=FALSE, message = FALSE, out.width = "90%"}
 
@@
     plot_type = "lv"
-  ) + ggtitle("") +
-  scale_fill_brewer(palette = "Dark2")
-```
-
-Q2: Is runs per over set to reduce in subsequent over if fielding is good in the previous over?
-
-For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
-<!--If a batsman is bowled out, it does not necessarily signify good fielding.
-So we only include number of catches and run out in an over as a measure of good fielding.
-Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
---> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable.
-Figure \ref{fig:field} shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
-
-<!-- Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.  -->
-
-<!-- fetching raw data since dot and fielding information not available in dismissal type not available in cricket data gravitas. Also `cricket` is aggregated across overs -->
-
-```{r runs_per-over}
+  ) + 
+  scale_fill_brewer(palette = "Dark2") + 
+  ggtitle("(a) Runs per over across over faceted by inning") +
+  theme(legend.position = "right")
+
 library(tsibble)
@@
 #   as_tsibble(index = data_index)
-```
-
-```{r}
 
@@
   mutate(
-    field = if_else(fielding_wckts == 0, "no wickets", ">0 wickets"),
+    field = if_else(fielding_wckts == 0, "0 wicket", ">0 wickets"),
     dot = if_else(dot_balls == 0, "no dot balls", ">0 dot balls"),
@@
   filter(lag_field != 0, lag_dot != 0)
-```
-
-```{r echo = FALSE, eval = FALSE}
-cricket_data %>% gran_obs("field", "over")
-cricket_data %>% gran_obs("dot", "over")
-```
-
-```{r field,  fig.cap = "Box plots showing distribution of run rate across overs of the innings and (lagged) wickets dismissed due to fielding.
-This shows at least one wicket in the previous over leads to downward shift of median run rate and quartile spread in the subsequent over for most overs.", warning = FALSE}
+
+cricket_data$lag_field <- factor(cricket_data$field, levels = c("0 wicket", ">0 wickets"))
 
 # filter(fielding_wckts %in% c(0,1)) %>%
+# 
  cricket_data %>%
@@
  quantile_prob = c(0.25, 0.5, 0.75)
-) + ggtitle("") 
-#+ geom_boxplot(width = 0.5, aes(colour = lag_field)) 
+) + ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") + 
+  #ylab("runs per over")  +
+  xlab("Number of wickets in previous over")
+# + geom_boxplot(width = 0.5, aes(colour = lag_field)) 
 ```
@@
 
-In Q2, wickets per over are considered as aperiodic cyclic granularities with wickets as aperiodic linear granularities.
+Wickets per over are considered as aperiodic cyclic granularities with wickets as aperiodic linear granularities.
 These granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
@@
 
-**Data and scripts:** Datasets and R code to reproduce all figures in this article (main.Rmd).
+**Data and scripts:** Data sets and R codes to reproduce all figures in this article (main.Rmd).
 

### 2020-03-19 main.Rmd
minor corrections

@@
 
+<!-- # ```{r tab-mayan} -->
+<!-- # table_mayan <- tibble(`Circular granularity` = c("minute-of-hour","minute-of-day", -->
+<!-- #                                                  "hour-of-day","hour-of-week", "day-of-week"), -->
+<!-- #                       Expression = c($L_1(z \mod 60)$, L_2(z \mod 60*24), L_3(\lfloor z/60\rfloor\mod 24)$,  L_4(\lfloor z/60\rfloor\mod 24*7)$, L_5(\lfloor z/24*60\rfloor \mod 7)$)) -->
+<!-- #                                                   -->
+<!-- #                                                   -->
+<!-- #                       C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1), -->
+<!-- #                       K = c(20, 18, 20, 20, 1))  -->
+<!-- # kable(table_mayan,  -->
+<!-- #       format = "latex", -->
+<!-- #       caption = "Hierarchy table for Mayan calendar.
+$G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
+All cyclic granularities are circular with $K$ as the period length.") -->
+<!-- # ``` -->
+
 ## Quasi-circular {#sec:quasi-circular-gran-def}
@@
 kable(table_mayan, 
+      format = "latex",
       caption = "Hierarchy table for Mayan calendar.
@@
 
-**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from Gregorian calendar.
+**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
 Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.
@@
 
-|  G	        |C                 |K
-|------------	|------------------|----------------|
-|  minute     |minute-of-hour    |  60            |
-|  hour    	  |hour-of-day       |  24  	        |
-|  day      	|day-of-month      | k(day, month)  |
-|  month    	|month-of-year     |  12            |
-|  year   	  | 1                |	1             |
+```{r tab-gregorian}
+table_greg <- tibble(G = c("minute", "hour", "day", "month", "year"),
+                      C = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
+                     K = c(20, 18,"k(day, month)", 20, 1))
+
+kable(table_greg, 
+      caption = "Illustrative hierarchy table for Gregorian calendar.
+$G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
+All these cyclic granularities are circular with $K$ as the period length except for day-of-month which is quasi-circular.")
+```
 
@@
 
-```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure", fig.pos="htb"}
+```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
 include_graphics("Figs/data-structure-v1.png")
@@
   filter_out = c("hhour", "fortnight")
-)%>% knitr::kable(format = "markdown")
+)%>% knitr::kable()
 ```
@@
 Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable due to different seasons.
-Plot (c) shows the effect of change in mapping which leads to easier comparison between weekdays and weekends for each hour."}
+Plot (c) shows the effect of change in mapping which leads to easier comparison between weekdays and weekends for each hour.", out.width="90%"}
 
@@
 It would be interesting to take their example in order to dive deeper into this question.
-<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}, it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
+<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
 This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
@@
 --> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable.
-Figure \ref{fig:field} shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
+Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
 

### 2020-03-23 main.Rmd
till cyclic calendar algebra re-check and maintain consistency

@@
 
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months, and non-repeating.
+Linear time granularities respect the linear progression of time and are non-repeating such as hours, days, weeks and months.
 One of the first attempts to characterize these granularities is due to @Bettini1998-ed.
@@
 Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
-Section \ref{sec:cyclic-calendar} introduces calendar algebra for computing cyclic time granularities.
+Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
 Section \ref{sec:data-structure} discusses the data structure for exploring the conditional distributions of the associated time series across pairs of cyclic time granularities.
@@
 
-<!-- CONTENT CUT -->
-<!-- \begin{definition}[Finer than]\label{def:finerthan} -->
-<!-- A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$. -->
-<!-- \end{definition}. -->
+\begin{definition}[Finer than]\label{def:finerthan}
+A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
+\[
+G(i) \subset H(j).
+\]
+\end{definition}.
 
@@
 
-For example, day $\trianglelefteq$ week since every granule of "week" is the union of some set of granules of "day".
-Consider another example where day $\trianglelefteq$ month.
+For example, $day \trianglelefteq week$ and $day \preceq week$ since every granule of "week" is the union of some set of granules of "day" and each "day" is a subset of a "week".
+Consider another example where $day \trianglelefteq month$.
 This relationship however is incomplete without its association to periodicity.
@@
 It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one.
-Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.
+Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and quasi-circular granularities.
 
@@
 The altering-tick operation: 
-: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$.
+: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
 The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].
+We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
 For more variations of calendar algebra operations, see @Ning2002-tf.
@@
 An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-<!--In a tsibble, each observation (row) is uniquely identified by index and key.--> A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the time index of a tsibble.
+<!--In a tsibble, each observation (row) is uniquely identified by index and key.--> A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the index of a tsibble.
 A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the time index of a tsibble is days, then a linear granularity might be weeks, months or years.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
 A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.-->
@@
 Cyclic granularities can be constructed by relating two linear granularities.
-The mappings between these linear granularities could be _regular_ with every granule of coarser linear granularity being made up of same number of granules of finer linear granularity or _irregular_.
+The mappings between these linear granularities could be either _regular_ or _irregular_ with every granule of coarser linear granularity being made up of same (regular) or different (irregular) number of granules of finer linear granularity.
 <!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).
@@
 The granules that overlap share elements from the underlying time domain.
-In this example, $P=7$, and the circular granularity $C_{\text{day}, \text{week}}$ representing day-of-week consists of the repeated pattern $\{L(0), L(1), L(2),$ $L(3), L(4), L(5), L(6)\}$.
+In this example, $P=7$, and the circular granularity $C_{\text{day}, \text{week}}$ representing Day-of-Week consists of the repeated pattern $\{L(0), L(1), L(2),$ $L(3), L(4), L(5), L(6)\}$.
 
@@
 \begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-minute-of-hour: & $C_1$ & $= L_1(z \mod 60)$ & $P_1 =$&$60$ \\
-minute-of-day: & $C_2$ & $= L_2(z \mod 60*24)$ & $P_2=$&$`r 60*24`$\\
-hour-of-day: & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
-hour-of-week: & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
-day-of-week: & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$ & $P_5=$&$7$\\
+\hline
+circular granularity && expression  && period length\\
+\hline
+minute-of-hour & $C_1$ & $= L_1(z \mod 60)$  & $P_1 = $&$60$ \\
+minute-of-day & $C_2$ & $= L_2(z \mod 60*24)$  & $P_2=$&$`r 60*24`$\\
+hour-of-day & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
+hour-of-week & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
+day-of-week & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$  & $P_5=$&$7$\\
 \bottomrule
@@
 \caption{Illustrative circular granularities with bottom granularity "minutes".
-These circular granularities relate two linear granularities one of which groups periodically into the other with regular mapping.}
+Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and $P_i$ represents the period length.}
 \label{tab:definitions}
@@
 
-<!-- # ```{r tab-mayan} -->
-<!-- # table_mayan <- tibble(`Circular granularity` = c("minute-of-hour","minute-of-day", -->
-<!-- #                                                  "hour-of-day","hour-of-week", "day-of-week"), -->
-<!-- #                       Expression = c($L_1(z \mod 60)$, L_2(z \mod 60*24), L_3(\lfloor z/60\rfloor\mod 24)$,  L_4(\lfloor z/60\rfloor\mod 24*7)$, L_5(\lfloor z/24*60\rfloor \mod 7)$)) -->
-<!-- #                                                   -->
-<!-- #                                                   -->
-<!-- #                       C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1), -->
-<!-- #                       K = c(20, 18, 20, 20, 1))  -->
-<!-- # kable(table_mayan,  -->
-<!-- #       format = "latex", -->
-<!-- #       caption = "Hierarchy table for Mayan calendar.
+<!-- ```{r tab-circular} -->
+<!--  table_mayan <-  -->
+<!--   tibble(C = c("minute-of-hour", -->
+<!--                                     "minute-of-day",  -->
+<!--                                     "hour-of-day", -->
+<!--                                     "hour-of-week", -->
+<!--                                     "day-of-week"), -->
+<!--                        Expression = c("\$L\_1(z \mod 60)\$", -->
+<!--                                         "\$L\_2(z \mod 60*24)\$", -->
+<!--                                         "\$L\_3(\lfloor z/60\rfloor\mod 24)\$", -->
+<!--                                         "\$L\_4(\lfloor z/60\rfloor\mod 24*7)\$", -->
+<!--                                         "\$L\_5(\lfloor z/24*60\rfloor \mod 7)\$")) -->
+
+<!--  kable(table_mayan, -->
+<!--        format = "latex", -->
+<!--        caption = "Hierarchy table for Mayan calendar.
 $G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
 All cyclic granularities are circular with $K$ as the period length.") -->
-<!-- # ``` -->
+<!-- ``` -->
 
@@
 \begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-
-day-of-month: & $Q_1$ && $P_1 = 31, 30, 29, 28$\\
-hour-of-month: & $Q_2$ && $P_2 = 24*31, 24*30, 24*29, 24*28$\\
-day-of-year: & $Q_3$ && $P_3 = 366, 365$\\
-week-of-month: & $Q_4$ && $P_4 = 5, 4$\\
+\hline
+quasi-circular granularity && potential period lengths\\
+\hline
+$Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\
+$Q_2 =$ hour-of-month && $P_2 = 24*31, 24*30, 24*29, 24*28$\\
+$Q_3 =$ day-of-year && $P_3 = 366, 365$\\
+$Q_4 =$ week-of-month && $P_4 = 5, 4$\\
 \bottomrule
@@
 \end{center}
-\caption{Illustrative quasi-circular granularities with potential period lengths.
-These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping.
-Illustrative quasi-circular granularities }
+\caption{Illustrative quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
+These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping and hence can not be expressed through modular arithmetic.}
 \label{tab:quasi}
@@
 ```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="The time domain distributed as linear granularities $B$, $G'$, $H'$ and quasi-circular granularities $Q(B, G')$ and $Q(B, H')$.
-Both $Q(B, G')$ and $Q(B, H')$ have repeating patterns unlike $G'$ or $H'$ and the patterns are guided by the distribution of $B$ within $G'$ and $H'$ within a period shown by the hashed arrow."}
+Both $Q(B, G')$ and $Q(B, H')$ have repeating patterns unlike $G'$ or $H'$ and the patterns are guided by the distribution of $B$ within $G'$ and $H'$ within a period shown by the hashed arrows."}
 knitr::include_graphics("Figs/quasi-circular-ex.png")
@@
 Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ such that \ref{def:periodical} does not hold true for them with respect to $B$ (the bottom granularity).
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ such that Definition \ref{def:periodical} does not hold true for them with respect to $B$ (the bottom granularity).
 <!--that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.--> However, $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
-Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
+Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$.
+Suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
 Here, index ${\{i_j\}}$ stands for the $j^{\text{th}}$ granule of the $i^{\text{th}}$ linear aperiodic granularity.
@@
 The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
--->The first slot in the diagram shows the index set and $A_{B, M}$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
+-->The first slot in the diagram shows the index set and $A(B, M)$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
 
@@
 **Order** of a linear granularity can be comprehended as the level of coarseness associated with a linear granularity.
-For two linear granularities $G$ and $H$, if $G$ groups into $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+For example, for two linear granularities $G$ and $H$, if $G$ groups into $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
 In any hierarchy table, linear granularities are arranged from lowest to highest order.
@@
       format = "latex",
+      booktabs = TRUE,
       caption = "Hierarchy table for Mayan calendar.
@@
 
-**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.
+**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar.
 
@@
 kable(table_greg, 
+      format = "latex",
+      booktabs = TRUE,
       caption = "Illustrative hierarchy table for Gregorian calendar.
@@
   knitr::kable(smart_data,
+               format = "latex",
+               booktabs = TRUE,
                caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
@@
   highest_unit = "month",
-  filter_out = c("hhour", "fortnight")
-)%>% knitr::kable()
+  filter_out = c("hhour", "fortnight"))%>%
+  knitr::kable(format = "latex",
+      booktabs = TRUE)
 ```
@@
   knitr::kable(smart_harmony,
+               format = "latex",
+               booktabs = TRUE,
                 caption = "Harmonies with two cyclic granularity one placed on the facet and the other on x-axis with the corresponding levels.
@@
 knitr::kable(hierarchy_model,
+             format = "latex",
+             booktabs = TRUE,
              caption = "Hierarchy table for cricket where each over is nested within innings, innings is nested within match and so on.

### 2020-03-23 main.Rmd
reorganised and changed language  of abstract

@@
 abstract: |
- Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible.
-As the frequency of time-oriented data increases, the number of questions about the observed variable that can be addressed by visual representation also increases.
-We propose some new tools to explore this type of data, which deconstruct time in many different ways.
-There are several classes of time deconstructions including linear and cyclic time granularities.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month, and aperiodic such as public or school holidays.
-
-  The hierarchical structure of linear granularities creates a natural nested ordering resulting in single-order-up and multiple-order-up granularities.
-For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.
-
-   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-This work provides tools for creating granularities and exploring the associated time series within a tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
-The ideas presented in this article have been implemented in the open source R package gravitas.
-R code and data are available in the supplementary materials.
-
+ Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets.
+This paper describes several classes of time deconstructions using linear and cyclic time granularities.
+Linear time granularities respect the linear progression of time such as hours, days, weeks and months with respect to a baseline.
+Cyclic time granularities can be circular such as hour of the day, quasi-circular such as day of the month, and aperiodic such as public  holidays.
+The hierarchical structure of granularities creates a nested ordering.
+Hour of the day and second of the minute are single-order-up.
+Hour of the week is multiple-order-up, because it passes over day  of the week.
+Methods are provided for creating all possible granularities for a time index.
+A recommendation algorithm provides an indication whether a pair of granularities could be interesting to examine together, called a harmony or clash.
+
+ The time granularities can used to create visualizations of the data to explore for periodicities, associations and anomalies.
+The granularities can be considered to be categorical variables (ordered and unordered) which induces a grouping of the observations.
+Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
+A recommendation of appropriate distribution display is provided.
+
+ The methods are implemented in the open source R package gravitas.
+The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow( FIXME REF), and the probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
 bibliography: [bibliography.bib]

### 2020-03-23 main.Rmd
figure 7 caption

@@
 
-```{r bothcust, fig.cap = "(a) Area quantile and (b) violin plot and (c) box plot of energy consumption for customer id: 10017936.
-Plot (a) helps to compare different hours of the day within weekdays and weekends.
-Plot (b) shows early morning hours are bimodal implying both high and low consumptions are probable due to different seasons.
-Plot (c) shows the effect of change in mapping which leads to easier comparison between weekdays and weekends for each hour.", out.width="90%"}
+```{r bothcust, fig.cap = "Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
+Two granularities are used: hour of the day (I) and  weekday/weekend (II).
+Plot (a) shows granularity I facetted by granularity II, and plots (b), (c) shows the converse mapping.
+Plot (a) makes a comparison of usage by workday within each hour of the day using side-by-side boxplots.
+Generally, on a work day there is more consumption early in the day.
+Plots (b) and (c) examine the temporal trend of consumption over the course of a day, separately for the type of day.
+Plot (b) uses an area quantile to put the emphasis on the time series, for example, the median consumption over time shows prolonged usage in the morning on weekdays.
+Plot (c) uses a violin plot to place emphasis on distributional differences across hours.
+It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.
+(FIXME: make nice axis labels)", out.width="90%"}
 

### 2020-03-24 main.Rmd
linear  granularities revised

@@
 
-Discrete abstraction of time such as weeks, months or holidays can be thought of as "time granularities". Time granularities are **linear** if they respect the linear progression of time.
-
-## Definitions and relationships
+Discrete abstraction of time such as weeks, months or holidays can be thought of as "time granularities".
+Time granularities are **linear** if they respect the linear progression of time.
+There have been several attempts to provide a framework for formally characterizing time granularities, including @Bettini1998-ed which forms the basis of the work described here.
+
+## Definitions
 
@@
 
-There have been several attempts to provide a framework for formally characterizing time granularities. One of the first attempts was provided by @Bettini1998-ed who proposed following definitions:
+<!-- There have been several attempts to provide a framework for formally characterizing time granularities.
+One of the first attempts was provided by @Bettini1998-ed who proposed following definitions:-->
 
@@
 
-\begin{definition}[Time domain]\label{def:definition}
-A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+\begin{definition}\label{def:definition}
+A {\bf time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$. 
 \end{definition}
 
-We will assume the time domain is *discrete*; i.e., there is unique predecessor and successor for every element in the time domain except for the first and last.
-
-\begin{definition}[Linear granularity]\label{def:linear}
-A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:
-
+The time domain is assumed to be *discrete*, and there is unique predecessor and successor for every element in the time domain except for the first and last.
+
+\begin{definition}\label{def:index set}
+There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of positive integers. Often, this is thought of as $t=1, \dots, T$.
+\end{definition}
+
+\begin{definition}\label{def:linear}
+A {\bf linear granularity} is a mapping $G$ from the integers (the index set, $Z$) to subsets of the time domain such that:
   (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
@@
   (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
+Each non-empty subset $G(i)$ is called a {\bf granule}.
 \end{definition}
 
-\begin{definition}[Granule]\label{def:granule}
-Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
-\end{definition}
-
-The first condition in Definition \ref{def:linear} implies that the granules in a linear granularity are non-overlapping and their index order is same as time order.
-\autoref{fig:linear-time} illustrates the linear time granularites hours, days, weeks, months and years, each of which are non-repeating in nature and arranged from past to future.
-If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.
+\noindent This implies that the granules in a linear granularity are non-overlapping, continuous and ordering is maintained.
+The indexing for each granule can also be associated with textual representation, called the label.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.
+When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
+
+\autoref{fig:linear-time} illustrates the linear time granularities.<!--, each of which are non-repeating in nature and arranged from past to future.
+If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.-->
 
@@
 ```
-In addition to the integer index, each granule could be associated with a textual representation which is called the label.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.--> When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
-
-@Bettini1998-ed discusses the relationships between linear time granularities, periodicities in time, and the structure of a calendar.
-
-\begin{definition}[Finer than]\label{def:finerthan}
-A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
-\[
-G(i) \subset H(j).
-\]
+
+## Relativities
+
+<!--@Bettini1998-ed discusses the relationships between linear time granularities, periodicities in time, and the structure of a calendar.-->
+
+Properties of pairs of granularities fall into various categories.
+
+\begin{definition}\label{def:finerthan}
+A linear granularity $G$ is {\bf finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
+$G(i) \subset H(j).$
 \end{definition}.
 
-\begin{definition}[Groups into]\label{def:groupsinto}
-A linear granularity $G$ groups into a linear granularity $H$, denoted
-$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that
-\[
-  H(j) = \bigcup_{i \in S}G(i).
-\]
+\begin{definition}\label{def:groupsinto}
+A linear granularity $G$ {\bf groups into} a linear granularity $H$, denoted
+$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
 \end{definition}
@@
 
-\begin{definition}[Groups periodically]\label{def:periodical}
-A granularity $H$ is periodical with respect to a granularity $G$ if:
+\begin{definition}\label{def:periodical}
+A granularity $H$ is {\bf periodical} with respect to a granularity $G$ if:
 (1) $G \trianglelefteq H$; and
@@
 
-A discrete time model often uses a fixed smallest linear granularity also known as _bottom granularity_ (@Bettini1998-ed) <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
+A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed _bottom granularity_.
+<!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
 Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
-A system of multiple granularities in lattice structures is referred to as a _calendar_ (see @Dyreson2000-mr).
+A system of multiple granularities in lattice structures is referred to as a _calendar_ by @Dyreson2000-mr.
 

### 2020-03-24 main.Rmd
linear  granularities relativities revised

@@
 $H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].
-These are called quasi-periodic relationships.
+Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way, called **quasi-periodic** relationships by @Bettini2000-vy.
 
@@
 
-For example, in a Gregorian calendar without leap years we could say days group periodically into months with the period $P = 365$ and the number of granules of "months" in each period given by $R=12$.
+\noindent For example, in a Gregorian calendar without leap years we could say days group periodically into months with the period $P = 365$ and the number of granules of "months" in each period given by $R=12$.
 In a Gregorian calendar with leap years, days group quasi-periodically into months with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
@@
 
-A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed _bottom granularity_.
+A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
 <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
 Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
-A system of multiple granularities in lattice structures is referred to as a _calendar_ by @Dyreson2000-mr.
+A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson2000-mr.
 

### 2020-03-24 main.Rmd
reference update

--- 2020-03-23 main.Rmd
@@
 
-Calendar-based graphics [@wang2018calendar] are useful in visualizing patterns in the weekly and monthly structure, and are helpful when checking for the effects of weekends or special days.
+Calendar-based graphics [@wang2020calendar] are useful in visualizing patterns in the weekly and monthly structure, and are helpful when checking for the effects of weekends or special days.
 Any temporal data at sub-daily resolution can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year, or another sub-daily deconstruction of time.
@@
 Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour.
-The lubridate package [@G_Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
+The lubridate package [@Grolemund_2011] provides tools to access and manipulate common date-time objects.
 But most of its accessor functions are limited to single-order-up granularities.
@@
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates [see @2012-la] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
+It is common to see aggregates [see @Goodwin_2012] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
 But studying overall energy use hides the distributions of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.

### 2020-03-24 main.Rmd
add DOIs to references

--- 2020-03-24 main.Rmd
@@
 Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour.
-The lubridate package [@G_Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
+The lubridate package [@Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
 But most of its accessor functions are limited to single-order-up granularities.
@@
 Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
-A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson2000-mr.
+A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson_2000.
 
@@
 
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf].
+Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning_2002].
 It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one.
@@
 For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
-For more variations of calendar algebra operations, see @Ning2002-tf.
+We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning_2002].
+For more variations of calendar algebra operations, see @Ning_2002.
 
@@
 ```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib")
 ```

### 2020-03-25 main.Rmd
Addressed issue 10: fig 6 doesn't really work for the purpose

@@
 As a result, we should avoid plotting clashes.
+For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarter-of-year.
+This plot does not work because every quarter does not correspond to all months of the year for example, the first quarter would not correspond to the month December of an year.
 
 <!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
-In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in \autoref{fig:allFig}.
-\autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
-This plot works because every day-of-week can occur in every month-of-year.
+In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.-->
 
@@
 
-```{r allFig, fig.height=5.5, out.width="50%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Letter value plot of electricity consumption data of Victoria from 2012 to 2014 (a) across day-of-week faceted by month-of-year and (b) across month-of-year faceted by day-of-week.
-Plots (a) and (b) both show harmonies since each month consists of all days of the week and every days of the week can occur in every month.
-However, Plot (a) helps comparison of different days of the week within each month and Plot (b) facilitates comparing months of the year for different days of the week.", fig.show = 'hold'}
-
-library(gravitas)
-library(tsibbledata)
-library(dplyr)
-library(ggplot2)
-
-par(mfrow = c(1, 2))
-
-VIC <- vic_elec
-
-VIC %>%
-  create_gran("month_year") %>%
-  filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
-  create_gran("day_week") %>%
-  filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
-  prob_plot("month_year", "day_week",
-    response = "Demand",
-    plot_type = "lv"
+```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Letter value plot of electricity consumption data (a) across month-of-year faceted by quarter-of-year, (b) across weekday/weekend faceted by quarter-of-year and (c) across quarter-of-year faceted by weekday/weekend.
+Plots (b) and (c) both show harmonies since each quarter consists of both weekdays and weekends.
+Conversely, weekend and weekday can occur in every quarter.
+Plot (b) helps to compare weekends and weekdays for each quarter.
+It can be seen that for every quarter weeekend and weekday consumption are fairly similar except for the second quarter where the letter value D and E behave differently, whereas, Plot (c) helps to compare quarters within weekdays and weekends.
+For example, the quartile spread of consumption shrinks/lowers from first to fourth quarter, whereas this pattern is not true for weekends.
+Plot (a) shows a clash since all quarters do not correspond to all months of the year.
+", fig.show = 'hold'}
+
+load("data/sm_cust50.rds")
+cust50 <- sm_cust50 %>%  distinct(customer_id)
+
+VIC <- sm_cust50 %>% 
+  filter(customer_id==cust50$customer_id[2])
+
+scene1 <-  VIC %>%
+  #create_gran("quarter_year") %>%
+  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  #create_gran("day_week") %>%
+  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("quarter_year", "wknd_wday",
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            symmetric = FALSE
   ) +
-  ylab("Electricity Demand [KWh]") +
-  xlab("day-of-week") + 
-  ggtitle("(a) Letter value plot with month-of-year facet")
+  ylab("") +
+  xlab("weekend/weekday") +
+  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10()
+
+scene2 <- VIC %>%
+  #create_gran("month_year") %>%
+  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  #create_gran("day_week") %>%
+  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("wknd_wday",
+            "quarter_year",
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            symmetric = FALSE
+  ) +
+  ylab("") +
+    xlab("quarter-of-year")  +
+  scale_fill_brewer(palette = "Paired") + theme(legend.position = "right") + ggtitle("") + scale_y_log10()
 
 # VIC %>%
-#   create_gran("day_week") %>%
-#   filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
-#   prob_plot("day_week", "hour_day",
-#     response = "Demand",
-#     plot_type = "quantile",
+#   #create_gran("month_year") %>%
+#   #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+#   #create_gran("day_week") %>%
+#   #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+#   prob_plot("month_year",
+#             "day_month",
+#             response = "general_supply_kwh",
+#             plot_type = "lv",
 #             symmetric = FALSE
 #   ) +
-#   ylab("Electricity Demand [KWh]") +
-#   ggtitle("(b) Decile plot")
-#
-# VIC %>%
-#   create_gran("day_month") %>%
-#   filter(day_month %in% c(1, 15, 29, 31)) %>%
-#   prob_plot("day_month",
-#     "day_year",
-#     response = "Demand",
-#     plot_type = "boxplot"
-#   ) +
-#   ylab("Electricity Demand [KWh]") +
-#   scale_x_discrete(breaks = seq(0, 366, 60)) +
-#   theme(
-#     legend.position = "bottom",
-#     strip.text = element_text(size = 7, margin = margin())
-#   ) +
-#   ggtitle("(c) Box plot")
-#
-# VIC %>%
-#   create_gran("week_month") %>%
-#   filter(week_month %in% c(1, 2, 4)) %>%
-#   prob_plot("week_month",
-#     "day_month",
-#     response = "Demand",
-#     plot_type = "violin", alpha = 0.3
-#   ) + ylab("") +
-#   xlab("Days of the Month") +
-#   theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) +
-#   scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
-#   ggtitle("(d) Violin plot")
-#
-# VIC %>%
-#   create_gran("week_month") %>%
-#   create_gran("week_year") %>%
-#   filter(week_month %in% c(1, 2, 5), week_year %in% c(1:20)) %>%
-#   prob_plot("week_year",
-#     "week_month",
-#     response = "Demand",
-#     plot_type = "ridge"
-#   ) +
-#   xlab("Electricity Demand [KWh]") +
-#   ylab("Weeks of the Year") +
-#   scale_x_continuous(breaks = seq(2000, 10000, 3000)) +
-#   theme(
-#     legend.position = "bottom",
-#     strip.text = element_text(size = 7, margin = margin())
-#   ) +
-#   ggtitle("(e) Ridge plot")
-
-# VIC %>%
-#   create_gran("month_year") %>%
-#   filter(month_year %in% c("Jan", "Jul", "Nov")) %>%
-#   prob_plot("month_year",
-#     "day_year",
-#     response = "Demand",
-#     plot_type = "quantile",
-#     quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
-#     symmetric = FALSE
-#   ) +
-#   # scale_x_discrete(breaks = seq(1, 336, 60)) +
 #   ylab("") +
-#   xlab("Day of the Year") + theme(
-#     legend.position = "bottom",
-#     strip.text = element_text(
-#       size = 7,
-#       margin = margin()
-#     )
-#   ) +
-#   ggtitle("(b) Decile plot")
-
-VIC %>%
-  create_gran("month_year") %>%
-  filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
-  create_gran("day_week") %>%
-  filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
-  prob_plot("day_week",
+#   xlab("day_month")  + 
+#   ggtitle("(d) Letter value plot across days of months & months")
+
+scene3 <- VIC %>%
+  #create_gran("month_year") %>%
+  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  #create_gran("day_week") %>%
+  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("quarter_year",
             "month_year",
-    response = "Demand",
-    plot_type = "lv"
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            symmetric = FALSE
   ) +
   ylab("") +
-  xlab("month-of-year") +
-  ggtitle("(b) Letter value plot with day-of-week facet")
-
-# VIC_moy_doy %>% ggplot(aes(x = yday, y = Value, col = as.factor(Quantile), group = yday)) + geom_line() + facet_wrap(~Indx_Month) + scale_x_continuous(breaks = seq(1, 336, 60)) + ylab("") + xlab("Day of the Year") + theme(legend.position = "none", strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot")
+  xlab("months") + theme(legend.position = "right") + scale_fill_brewer(palette = "Set2") + ggtitle("") + scale_y_log10()
+
+gg_fig <- ggarrange(scene3, 
+          ggarrange(scene1, scene2, ncol = 2, labels = c("(b)", "(c)")),
+          nrow = 2, labels = "(a)")
+          #label.y = "Electricity Demand [KWh]"\
+
+ggpubr::annotate_figure(gg_fig,
+                left = text_grob("Electricity Demand [KWh]",  rot = 90))
+
 ```
-
 ### Interchangeability of mappings
@@
 With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
-\autoref{fig:allFig} (a) shows the letter value plot across days of the week faceted by months like January, March and December and \autoref{fig:allFig} (b) shows letter value of across months faceted by days.
-The first one helps us to compare different days of the weeks within a month and the second one helps to compare different months for each day of the week.
+\autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarter-of-year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
+\autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
 

### 2020-03-25 main.Rmd
Addressed issue 10: fig 6 doesn't really work for the purpose

@@
 As a result, we should avoid plotting clashes.
-For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarter-of-year.
-This plot does not work because every quarter does not correspond to all months of the year for example, the first quarter would not correspond to the month December of an year.
 
 <!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
-In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.-->
+In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption is plotted across different time granularities in each of the panel in \autoref{fig:allFig}.
+\autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December.
+This plot works because every day-of-week can occur in every month-of-year.
 
@@
 With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
-\autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarter-of-year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
-\autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
+\autoref{fig:allFig} (a) shows the letter value plot across days of the week faceted by months like January, March and December and \autoref{fig:allFig} (b) shows letter value of across months faceted by days.
+The first one helps us to compare different days of the weeks within a month and the second one helps to compare different months for each day of the week.
 

### 2020-03-25 main.Rmd
Issue 11: tables need cleaning up

--- 2020-03-25 main.Rmd
@@
 library(ggpubr)
+library(xtable)
+library(kableExtra)
 opts_chunk$set(
@@
 For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
-\autoref{tab:tab-mayan} represents the the hierarchy table for the Mayan calendar.
+\autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.
 
 ```{r tab-mayan}
-table_mayan <- tibble(G = c("kin", "uinal", "tun", "katun", "baktun"),
-                      C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
-                      K = c(20, 18, 20, 20, 1)) 
+table_mayan <- tibble(`linear (G)` = c("kin", "uinal", "tun", "katun", "baktun"),
+                      `single-order-up cyclic (C)` = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
+                       `period length/conversion operator (K)`= c(20, 18, 20, 20, 1)) 
 kable(table_mayan, 
@@
       booktabs = TRUE,
-      caption = "Hierarchy table for Mayan calendar.
-$G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
-All cyclic granularities are circular with $K$ as the period length.")
+      align = "c",
+      caption = "Hierarchy table for Mayan calendar with all circular single-order-up granularities.")
+  #row_spec(0, bold = TRUE) %>% 
+  #row_spec(1:(nrow(table_mayan)-1), hline_after = T) %>% kable_styling(latex_options = "striped")
 ```
@@
 ```{r tab-gregorian}
-table_greg <- tibble(G = c("minute", "hour", "day", "month", "year"),
-                      C = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
-                     K = c(20, 18,"k(day, month)", 20, 1))
+table_greg <- tibble(`linear (G)` = c("minute", "hour", "day", "month", "year"),
+                      `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
+                     `period length/conversion operator (K)` = c(20, 18,"k(day, month)", 20, 1))
 
@@
       booktabs = TRUE,
-      caption = "Illustrative hierarchy table for Gregorian calendar.
-$G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
-All these cyclic granularities are circular with $K$ as the period length except for day-of-month which is quasi-circular.")
+      caption = "Illustrative hierarchy table for Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>% 
+  #row_spec(0, bold = TRUE) %>%
+   kable_styling()
 ```
@@
 <!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to look at the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
+
 <!-- and identify the extreme and regular households. -->
 
-### Data structure for visualization:
-
-The data structure for a sample of 50 households from that trial is depicted in \autoref{tab:smart-data} where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
-<!-- We take a random sample of 50 households to see how we can achieve that.  -->
+<!-- ### Data structure for visualization: -->
+
+<!-- The data structure for a sample of 50 households from that trial is depicted in \autoref{tab:smart-data} where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
+-->
+<!-- <!-- We take a random sample of 50 households to see how we can achieve that.  -->
 
 ```{r smart-data}
-load("data/sm_cust50.Rdata")
-
-smart_data <- smart_meter <- sm_cust50 %>%
-  as_tsibble() %>%
-  dplyr::select(
-    customer_id,
-    reading_datetime,
-    general_supply_kwh
-  ) %>% head(n = 5)
-
-  knitr::kable(smart_data,
-               format = "latex",
-               booktabs = TRUE,
-               caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
+# load("data/sm_cust50.Rdata")
+# 
+# smart_data <- smart_meter <- sm_cust50 %>%
+#   as_tsibble() %>%
+#   dplyr::select(
+#     customer_id,
+#     reading_datetime,
+#     general_supply_kwh
+#   ) %>% head(n = 5)
+# 
+#   knitr::kable(smart_data,
+#                format = "latex",
+#                booktabs = TRUE,
+#                caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
 ```
@@
   knitr::kable(format = "latex",
-      booktabs = TRUE)
+      booktabs = TRUE) %>%
+  #row_spec(0, bold = TRUE)%>%
+  kable_styling()
 ```
@@
                booktabs = TRUE,
-                caption = "Harmonies with two cyclic granularity one placed on the facet and the other on x-axis with the corresponding levels.
-Out of 30 possible combinations of cyclic granularities that could be plotted only 13 are harmonies.")
+                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis.") %>% 
+    #row_spec(0, bold = TRUE) %>% 
+    kable_styling() %>% 
+     add_footnote("NOTE: Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.",
+                  notation="none")
 ```
@@
 hierarchy_model <- tibble::tibble(
-  G = c("over", "inning", "match", "season"),
-  C = c("over-of-inning", "inning-of-match", "match-of-season", 1),
-  K = c(20, 2, "k(match, season)", 1)
+  `linear (G)` = c("over", "inning", "match", "season"),
+  `single-order-up cyclic (C)` = c("over-of-inning", "inning-of-match", "match-of-season", 1),
+  `period length/conversion operator (K)` = c(20, 2, "k(match, season)", 1)
 )
@@
              booktabs = TRUE,
-             caption = "Hierarchy table for cricket where each over is nested within innings, innings is nested within match and so on.
-Cyclic granularity over-of-inning, inning-of-match are assumed to be circular which means each inning will have 20 overs and each match has 2 innings and  match-of-season is assumed to be quasi-periodic.
-")
+             caption = "Hierarchy table for cricket with each over is nested within innings, innings nested within match and match within season.") %>% 
+  #row_spec(0, bold = TRUE)%>%
+  kable_styling()
 ```

### 2020-03-26 main.Rmd
section identifier
illutrative removed, justification corrected and note included in caption for tables

@@
 \end{center}
-\caption{Illustrative circular granularities with bottom granularity "minutes".
+\caption{Examples of circular granularities with bottom granularity minutes.
 Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and $P_i$ represents the period length.}
@@
 \end{center}
-\caption{Illustrative quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
+\caption{Examples of quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
 These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping and hence can not be expressed through modular arithmetic.}
@@
       booktabs = TRUE,
-      align = "c",
-      caption = "Hierarchy table for Mayan calendar with all circular single-order-up granularities.")
+      caption = "Hierarchy table for Mayan calendar with all circular single-order-up granularities.") %>% 
+  kable_styling()
   #row_spec(0, bold = TRUE) %>% 
-  #row_spec(1:(nrow(table_mayan)-1), hline_after = T) %>% kable_styling(latex_options = "striped")
+  #row_spec(1:(nrow(table_mayan)-1), hline_after = T)   
 ```
@@
       booktabs = TRUE,
-      caption = "Illustrative hierarchy table for Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>% 
+      caption = "Hierarchy table for Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>% 
   #row_spec(0, bold = TRUE) %>%
@@
 
-### Harmonies and Clashes {#sec:synergy}
+## Harmonies and Clashes {#sec:synergy}
 
@@
 
-### Summarizing the measured variable {#sec:summarize-measured}
+## Summarizing the measured variable {#sec:summarize-measured}
 
@@
                booktabs = TRUE,
-                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis.") %>% 
+                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis.
+Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.") %>%
     #row_spec(0, bold = TRUE) %>% 
-    kable_styling() %>% 
-     add_footnote("NOTE: Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.",
-                  notation="none")
+    kable_styling()
 ```

### 2020-03-26 main.Rmd
index starts from 0 for figure 1

@@
 These are ordered with ordering guided by integers and hence is unidirectional and non-repeating."}
-knitr::include_graphics("Figs/linear-arr.png")
+knitr::include_graphics("Figs/linear-arr1.png")
 ```

### 2020-03-26 main.Rmd
axis labels fix, caption first line fixed and Di FIXME labels

@@
  The methods are implemented in the open source R package gravitas.
-The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow( FIXME REF), and the probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
+The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
 bibliography: [bibliography.bib]
@@
 \caption{Examples of circular granularities with bottom granularity minutes.
-Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and $P_i$ represents the period length.}
+Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and $P_i$ represents the period length.
+Circular granularities could be expressed using modular arithmetic due to their regular mapping.
+}
 \label{tab:definitions}
@@
 \caption{Examples of quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
-These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping and hence can not be expressed through modular arithmetic.}
+These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping and hence can not be expressed through modular arithmetic.
+}
 \label{tab:quasi}
@@
 
-```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="The time domain distributed as linear granularities $B$, $G'$, $H'$ and quasi-circular granularities $Q(B, G')$ and $Q(B, H')$.
-Both $Q(B, G')$ and $Q(B, H')$ have repeating patterns unlike $G'$ or $H'$ and the patterns are guided by the distribution of $B$ within $G'$ and $H'$ within a period shown by the hashed arrows."}
+```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="The time domain distributed as linear granularities $B$ (bottom granularity), $G'$, $H'$ and quasi-circular granularities $Q(B, G')$ and $Q(B, H')$.
+Both $Q(B, G')$ and $Q(B, H')$ have repeating patterns unlike $G'$ or $H'$ and the patterns are guided by the distribution of $B$ within $G'$ and $H'$ and remains constant within a period shown by the hashed arrows."}
 knitr::include_graphics("Figs/quasi-circular-ex.png")
@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear aperiodic granularities $B$, $M1$, $M2$, $M$ and aperiodic cyclic granularity $A(B, M)$.
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear aperiodic granularities $M1$, $M2$, $M$ and aperiodic cyclic granularity $A(B, M)$ relating them with bottom granularity $B$.
 $A(B, M)$ have repeating patterns with three categories, two of which correspond to the aperiodic linear granularities $M1$ and $M2$ and other one correponding to no aperiodic granularity."}
@@
 
-```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Letter value plot of electricity consumption data (a) across month-of-year faceted by quarter-of-year, (b) across weekday/weekend faceted by quarter-of-year and (c) across quarter-of-year faceted by weekday/weekend.
+```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Distribution of energy consumption displayed through letter value plots.
+Plot (a) displays consumption across month-of-year faceted by quarter-of-year, (b) across weekday/weekend faceted by quarter-of-year and (c) across quarter-of-year faceted by weekday/weekend.
 Plots (b) and (c) both show harmonies since each quarter consists of both weekdays and weekends.
@@
 Plot (b) helps to compare weekends and weekdays for each quarter.
-It can be seen that for every quarter weeekend and weekday consumption are fairly similar except for the second quarter where the letter value D and E behave differently, whereas, Plot (c) helps to compare quarters within weekdays and weekends.
+It can be seen that for every quarter weeekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently, whereas, Plot (c) helps to compare quarters within weekdays and weekends.
 For example, the quartile spread of consumption shrinks/lowers from first to fourth quarter, whereas this pattern is not true for weekends.
@@
                booktabs = TRUE,
-                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis.
+                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis along with their number of categories/levels.
 Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.") %>%
@@
 Plot (c) uses a violin plot to place emphasis on distributional differences across hours.
-It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.
-(FIXME: make nice axis labels)", out.width="90%"}
+It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.", out.width="90%"}
 
@@
   scale_y_sqrt() +
-  ggtitle(" (b) Area quantile plot faceted by weekend-weekday")  + ylab("Electricity Demand [KWh]")
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("Electricity Demand [KWh]") + xlab("hour-of-day") + ggtitle("") + ylab("")
 #
@@
   scale_y_sqrt()+
-  ggtitle(" (c) Violin plot faceted by weekend-weekday") + ylab("")
-
-smart_meter10 %>%
+  #ggtitle(" (c) Violin plot faceted by weekend-weekday") +
+  ylab("") + xlab("hours of the day") + ggtitle("") + 
+  scale_x_discrete( breaks = seq(0, 23, 5))
+
+cust2_box <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@
   ) +
-  scale_y_sqrt()+
-  ggtitle(" (a) Box plot faceted by hour-of-day")  + ylab("Electricity Demand [KWh]")
-
-ggarrange(cust2_quantile, cust2_violin, nrow = 2)
+  scale_y_sqrt() +
+  #ggtitle(" (a) Box plot faceted by hour-of-day") +
+  xlab("weekday(wdy)/ weekend(wnd)") + ggtitle("") + ylab("") + 
+  scale_x_discrete(labels = c('wdy','wnd'))
+
+#ggarrange(cust2_quantile, cust2_violin, nrow = 2)
+
+gg_fig <- ggarrange(cust2_box, 
+          ggarrange(cust2_quantile, cust2_violin, nrow = 2, labels = c("(b)", "(c)")),
+          ncol = 2, labels = "(a)")
+          #label.y = "Electricity Demand [KWh]"\
+
+ggpubr::annotate_figure(gg_fig,
+                left = text_grob("Electricity Demand [KWh]",  rot = 90))
 ```
@@
 
-```{r cricex, fig.cap= "(a) Letter value plot shows that there is no upward shift in runs in the second innings like that in the first implying teams are vulnerable to score more in the first innings as they approach the end of the inning.
-(b) Box plot shows at least one wicket in the previous over leads to downward shift of median run rate and quartile spread in the subsequent over for most overs.", warning=FALSE, message = FALSE, out.width = "90%"}
+```{r cricex, fig.cap= "Runs per over shown with different distribution displays, and granularity arrangements.
+Plot (a) shows letter value plot across overs faceted by innings.
+There is no upward shift in runs in the second innings like that in the first implying teams are vulnerable to score more in the first innings as they approach the end of the inning.
+Plot (b) shows box plot across number of wickets in previous over faceted by current over.
+This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning=FALSE, message = FALSE, out.width = "90%"}
 
@@
   )) %>%
+  mutate(inning = paste0("inning:", inning)) %>% 
   prob_plot("inning",
@@
   ggtitle("(a) Runs per over across over faceted by inning") +
-  theme(legend.position = "right")
+  theme(legend.position = "right") +
+  ggtitle("(a)") +
+  ylab("runs per hour") +
+  xlab("over-of-inning") 
+  # ggplot2::theme(
+  #     strip.text = ggplot2::element_text(
+  #       size = 7,
+  #       margin = ggplot2::margin()
+  #     )
+  #   ) 
+  #
 
@@
 ) + ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") + 
-  #ylab("runs per over")  +
-  xlab("Number of wickets in previous over")
+  ylab("runs per over")  +
+  xlab("Number of wickets in previous over") +
+  ggtitle("(b)")
 # + geom_boxplot(width = 0.5, aes(colour = lag_field)) 

### 2020-03-26 main.Rmd
Redundant package loading

@@
 library(gravitas)
-library(tsibble)
 library(ggpubr)

### 2020-03-26 main.Rmd
Fixed table justification

@@
 \begin{center}
-\begin{tabular}{lr@{~}lr@{~}r}
+\begin{tabular}{lll}
 \hline
-circular granularity && expression  && period length\\
+circular granularity & expression & period length \\
 \hline
-minute-of-hour & $C_1$ & $= L_1(z \mod 60)$  & $P_1 = $&$60$ \\
-minute-of-day & $C_2$ & $= L_2(z \mod 60*24)$  & $P_2=$&$`r 60*24`$\\
-hour-of-day & $C_3$ & $= L_3(\lfloor z/60\rfloor\mod 24)$  & $P_3 =$&$24$ \\
-hour-of-week & $C_4$ & $= L_4(\lfloor z/60\rfloor\mod 24*7)$  & $P_4=$&$`r 24*7`$\\
-day-of-week & $C_5$ & $= L_5(\lfloor z/24*60\rfloor \mod 7)$  & $P_5=$&$7$\\
+minute-of-hour                               &
+  $C_1 = L_1(z \mod 60)$                     &
+  $P_1 = \phantom{99}60$ \\
+minute-of-day                                &
+  $C_2 = L_2(z \mod 60*24)$                  &
+  $P_2= `r 60*24`$\\
+hour-of-day                                  &
+  $C_3 = L_3(\lfloor z/60\rfloor\mod 24)$    &
+  $P_3 = \phantom{99}24$ \\
+hour-of-week                                 &
+  $C_4 = L_4(\lfloor z/60\rfloor\mod 24*7)$  &
+  $P_4= \phantom{9}`r 24*7`$\\
+day-of-week                                  &
+  $C_5 = L_5(\lfloor z/24*60\rfloor \mod 7)$ &
+  $P_5= \phantom{999}7$\\
 \bottomrule

### 2020-03-27 main.Rmd
all plot labels lowercase + for weekends/weekdays no xlabels needed

@@
       booktabs = TRUE,
+      align = "l",
       caption = "Hierarchy table for Mayan calendar with all circular single-order-up granularities.") %>% 
@@
   ylab("") +
-  xlab("weekend/weekday") +
-  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10()
+  #xlab("weekend/weekday") +
+  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + xlab("")
 
@@
   ylab("") +
-    xlab("quarter-of-year")  +
+    xlab("quarters of the year")  +
   scale_fill_brewer(palette = "Paired") + theme(legend.position = "right") + ggtitle("") + scale_y_log10()
@@
   ylab("") +
-  xlab("months") + theme(legend.position = "right") + scale_fill_brewer(palette = "Set2") + ggtitle("") + scale_y_log10()
+  xlab("months of the year") + theme(legend.position = "right") + scale_fill_brewer(palette = "Set2") + ggtitle("") + scale_y_log10()
 
@@
           nrow = 2, labels = "(a)")
-          #label.y = "Electricity Demand [KWh]"\
+          #label.y = "electricity demand [KWh]"\
 
 ggpubr::annotate_figure(gg_fig,
-                left = text_grob("Electricity Demand [KWh]",  rot = 90))
+                left = text_grob("electricity demand [KWh]",  rot = 90))
 
@@
   #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("Electricity Demand [KWh]") + xlab("hour-of-day") + ggtitle("") + ylab("")
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")
 #
@@
   #ggtitle(" (a) Box plot faceted by hour-of-day") +
-  xlab("weekday(wdy)/ weekend(wnd)") + ggtitle("") + ylab("") + 
+  xlab("") +
+  ggtitle("") + ylab("") + 
   scale_x_discrete(labels = c('wdy','wnd'))
@@
           ncol = 2, labels = "(a)")
-          #label.y = "Electricity Demand [KWh]"\
+          #label.y = "electricity demand [KWh]"\
 
 ggpubr::annotate_figure(gg_fig,
-                left = text_grob("Electricity Demand [KWh]",  rot = 90))
+                left = text_grob("electricity demand [KWh]",  rot = 90))
 ```
@@
   ggtitle("(a)") +
-  ylab("runs per hour") +
-  xlab("over-of-inning") 
+  ylab("runs per over") +
+  xlab("overs of the innings") 
   # ggplot2::theme(
@@
   ylab("runs per over")  +
-  xlab("Number of wickets in previous over") +
+  xlab("number of wickets in previous over") +
   ggtitle("(b)")

### 2020-03-27 main.Rmd
labels a, b and c in plots no parenthesis

@@
 gg_fig <- ggarrange(scene3, 
-          ggarrange(scene1, scene2, ncol = 2, labels = c("(b)", "(c)")),
-          nrow = 2, labels = "(a)")
+          ggarrange(scene1, scene2, ncol = 2, labels = c("b", "c")),
+          nrow = 2, labels = "a")
           #label.y = "electricity demand [KWh]"\
@@
 gg_fig <- ggarrange(cust2_box, 
-          ggarrange(cust2_quantile, cust2_violin, nrow = 2, labels = c("(b)", "(c)")),
-          ncol = 2, labels = "(a)")
+          ggarrange(cust2_quantile, cust2_violin, nrow = 2, labels = c("b", "c")),
+          ncol = 2, labels = "a")
           #label.y = "electricity demand [KWh]"\
@@
   scale_fill_brewer(palette = "Dark2") + 
-  ggtitle("(a) Runs per over across over faceted by inning") +
+  #ggtitle("(a) Runs per over across over faceted by inning") +
   theme(legend.position = "right") +
-  ggtitle("(a)") +
+  ggtitle("a") +
   ylab("runs per over") +
-  xlab("overs of the innings") 
+  xlab("overs of the innings") +
+  theme(plot.title = element_text(face = "bold"))
   # ggplot2::theme(
@@
   symmetric = FALSE,
- quantile_prob = c(0.25, 0.5, 0.75)
-) + ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") + 
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") + 
   ylab("runs per over")  +
   xlab("number of wickets in previous over") +
-  ggtitle("(b)")
+  ggtitle("b") +
+   theme(plot.title = element_text(face = "bold"))
 # + geom_boxplot(width = 0.5, aes(colour = lag_field)) 

### 2020-03-31 main.Rmd
data structure image update: include all cyclic granularities

@@
 A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
-Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities.
-Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
+Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
+This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$ such that $C_i$ maps index set to $n_i$ levels or categories and $v$ denotes the measurement variable.
+The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities.
+Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities each relating one finer linear granularity with a coarser one could be constructed.
+Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
+Instead, this big problem is broken down by focusing on two cyclic granularities at a time.
+Hence data sets of the form <$C_i$, $C_j$, $v$> forms the basis for exploration and analysis of the measured variable.
+
+<!-- CONTENT SHORT -->
+<!-- across bivariate cyclic granularities.  -->
+<!-- Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
 Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+-->
 
@@
 
-```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including two cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
-include_graphics("Figs/data-structure-v1.png")
+```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
+include_graphics("Figs/data-structure-v2.png")
 ```
@@
 
+<!-- Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
+Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+-->
+
 The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
-While considering data sets of the form $<C_1, C_2, v>$, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
-Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ data subsets.
+Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
+Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, l}$ and $j \in {1, 2, \dots, m}$  which will lead to $lm$ data subsets.
 Now, some situations can lead to few or many of these sets being empty.
-We will discuss few cases, where one or more of these $nm$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
+We will discuss few cases, where one or more of these $lm$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
 <!--The implication of these empty sets on visualization will be discussed in Section \ref{sec:vis-interaction}.

### 2020-04-01 main.Rmd
add notations in application

@@
 B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping,
-$P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$,
-and ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$.
+$P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$.
 \end{definition}
 
-\autoref{fig:circular-dow} depicts the circular granularity "Day-of-Week", constructed from linear granularities "day" (the bottom granularity) and "week".
+${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$.
+\autoref{fig:circular-dow} depicts the circular granularity "Day-of-Week", constructed from linear granularities "Day" (the bottom granularity) and "Week".
 The granules that overlap share elements from the underlying time domain.
@@
 
-### Circular or quasi-circular single order-up granularities
+### Circular or quasi-circular single order-up granularities {#sec:notall-circular}
 
@@
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-These cyclic granularities are a function of the index variable and are computed based on if they are circular <!--(Section \ref{sec:circular-gran-def})-->, quasi-circular.
-<!--(Section \ref{sec:quasi-circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def})--> If we consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year), the following cyclic time granularities can be considered for this analysis.
-The interval of this tsibble is 30 minutes, and hence the temporal granularities may range from half-hour to year.
-The cyclic granularities which could be considered are hhour_hour, hhour_day, hhour_week, hhour_month, hhour_year, hour_day, hour_week, hour_month, hour_year, day_week, day_month, day_year, week_month, week_year and month_year.
-These options are too many and hence the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped.
+Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
+Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
+Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed.
+If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped.
 Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} or \ref{sec:cyclic-calendar}.
+
+## Computation
+
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:cyclic-calendar}, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} or (for hour_month, day_month, week_month).
+
+The index set maps the half-hourly time instants to a set of positive integers.
+$$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular})
+$$ hour\_week(z) =  day\_week(z) + 24*hour\_day(z)$$ (Section \ref{sec:all-circular})
+<!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) -->
+$$ hour\_month(z) =  hour\_day(z) + 24*day\_month(z)$$ (Section \ref{sec:notall-circular})
+
+<!-- These cyclic granularities are a function of the index variable and are computed based on if they are circular <!--(Section \ref{sec:circular-gran-def}), quasi-circular.
+<!--(Section \ref{sec:quasi- circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def})-->
 
@@
 
-```{r search_gran_limit2, echo=FALSE, eval = FALSE}
+```{r search_gran_limit2, echo = FALSE, eval = FALSE}
 smart_meter10 %>% search_gran(

### 2020-04-01 main.Rmd
L removed from definition of circular granularity

@@
 \begin{split}
-C_{B, G}(z) & = {L}(z\mod P(B, G)) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
+C_{B, G}(z) & = z\mod P(B, G) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
 \end{split}
@@
 
-${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$.
+<!-- The label can be mapped to index for input and the index can later be mapped to this label for output.
+When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.
+-->
+
 \autoref{fig:circular-dow} depicts the circular granularity "Day-of-Week", constructed from linear granularities "Day" (the bottom granularity) and "Week".
 The granules that overlap share elements from the underlying time domain.
-In this example, $P=7$, and the circular granularity $C_{\text{day}, \text{week}}$ representing Day-of-Week consists of the repeated pattern $\{L(0), L(1), L(2),$ $L(3), L(4), L(5), L(6)\}$.
-
-Each circular granularity can use descriptive label mappings, or be simply labelled using the index. The label mapping $L$ in \autoref{fig:circular-dow} can be defined as
+In this example, $P=7$, and the circular granularity $C_{\text{day}, \text{week}}$ representing Day-of-Week consists of the repeated pattern $\{0, 1, 2, 3, 4, 5, 6\}$.
+Each circular granularity can use descriptive label mappings, or be simply labelled using the index.
+<!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for \autoref{fig:circular-dow} can be defined as
 $$
@@
 
-In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
-Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping.
+In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H)$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
+Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity.
 <!--The representation of the resultant circular granularity will be a function of the index set.--> The bottom granularity is assumed to be minutes in this case.
@@
 minute-of-hour                               &
-  $C_1 = L_1(z \mod 60)$                     &
+  $C_1 = z \mod 60$                     &
   $P_1 = \phantom{99}60$ \\
 minute-of-day                                &
-  $C_2 = L_2(z \mod 60*24)$                  &
+  $C_2 = z \mod 60*24$                  &
   $P_2= `r 60*24`$\\
 hour-of-day                                  &
-  $C_3 = L_3(\lfloor z/60\rfloor\mod 24)$    &
+  $C_3 = \lfloor z/60\rfloor\mod 24$    &
   $P_3 = \phantom{99}24$ \\
 hour-of-week                                 &
-  $C_4 = L_4(\lfloor z/60\rfloor\mod 24*7)$  &
+  $C_4 = \lfloor z/60\rfloor\mod 24*7$  &
   $P_4= \phantom{9}`r 24*7`$\\
 day-of-week                                  &
-  $C_5 = L_5(\lfloor z/24*60\rfloor \mod 7)$ &
+  $C_5 = \lfloor z/24*60\rfloor \mod 7$ &
   $P_5= \phantom{999}7$\\

### 2020-04-02 main.Rmd
circular diagram and caption changed

@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear granularities Day and Week and circular granularity Day-of-Week with $P=7$ is constructed using the two linear granularities.
-Day-of-Week has a repeating pattern unlike Day and Week and has an ordering only within a period represented by hashed arrows."}
-knitr::include_graphics("Figs/circular-dow-arrow.png")
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Terminology and intuition of circular granularities are shown.
+Panel a shows the time domain distributed as linear granularities Day and Week and circular granularity Day-of-Week with $P=7$ is constructed using the two linear granularities.
+Day-of-Week has a repeating pattern unlike Day and Week and has an ordering only within a period represented by hashed arrows.
+Panel b and c depicts the intuition through a measuring tape analogy where the linear measuring tape has been folded to obtain circular granularities from linear ones to accomodate for repeating patterns."}
+knitr::include_graphics("Figs/circular-illustration.png")
 ```

### 2020-04-02 main.Rmd
added Nick in acknowledgement
aperiodic picture and caption update

@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="The time domain distributed as linear aperiodic granularities $M1$, $M2$, $M$ and aperiodic cyclic granularity $A(B, M)$ relating them with bottom granularity $B$.
-$A(B, M)$ have repeating patterns with three categories, two of which correspond to the aperiodic linear granularities $M1$ and $M2$ and other one correponding to no aperiodic granularity."}
-knitr::include_graphics("Figs/aperiodic-ex.png")
+```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Terminology and intuition of aperiodic cyclic granularities are shown.
+Panel a shows the time domain distributed as linear granularities Day, Semester, Semester Week, within semester breaks, in-session and aperiodic cyclic granularity Week-of-semester.
+$A(B, M)$ consists of three categories, 1, 2 and 0 corresponding to within semester breaks, in-session and non-semester period respectively.
+Panel b depicts the intuition through a measuring tape analogy which needs to be folded at different points every year since $M1$ and $M2$ will fall in different weeks."}
+knitr::include_graphics("Figs/aperiodic-ex2.png")
 ```
@@
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
+We would also like to thank Nicholas Spyrison for many useful discussions and feedback on the manuscipt.
 More details about the package

### 2020-04-02 main.Rmd
quasi-circular diagram and caption update

@@
 
-```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="The time domain distributed as linear granularities $B$ (bottom granularity), $G'$, $H'$ and quasi-circular granularities $Q(B, G')$ and $Q(B, H')$.
-Both $Q(B, G')$ and $Q(B, H')$ have repeating patterns unlike $G'$ or $H'$ and the patterns are guided by the distribution of $B$ within $G'$ and $H'$ and remains constant within a period shown by the hashed arrows."}
-knitr::include_graphics("Figs/quasi-circular-ex.png")
+```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="
+Terminology and intuition of quasi-circular granularities are shown.
+Panel a shows the time domain distributed as linear granularities $B$ (bottom granularity), $G$, $H'$ and quasi-circular granularities $Q(B, G)$ and $Q(B, H')$.
+Both $Q(B, G)$ and $Q(B, H')$ have repeating patterns unlike $G$ or $H'$ and the patterns are guided by the distribution of $B$ within $G$ and $H'$ and remains constant within a period shown by the hashed arrows.
+Panel b depicts the intuition through a measuring tape analogy where the folds are not of same length but has same length over a period."}
+knitr::include_graphics("Figs/quasi-circular-ex2.png")
 ```
@@
 ```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Terminology and intuition of aperiodic cyclic granularities are shown.
-Panel a shows the time domain distributed as linear granularities Day, Semester, Semester Week, within semester breaks, in-session and aperiodic cyclic granularity Week-of-semester.
-$A(B, M)$ consists of three categories, 1, 2 and 0 corresponding to within semester breaks, in-session and non-semester period respectively.
+Panel a shows the time domain distributed as linear granularities Day, Semester, within semester breaks, in-session and aperiodic cyclic granularity Week-of-Semester.
+$A(B, M)$ consists of three categories 0, 1, 2 non-semester period, within-semester breaks and in-session semester period respectively.
 Panel b depicts the intuition through a measuring tape analogy which needs to be folded at different points every year since $M1$ and $M2$ will fall in different weeks."}

### 2020-04-02 main.Rmd
wickets labels and facet label "innings" changed

@@
 
-```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="
-Terminology and intuition of quasi-circular granularities are shown.
+```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="Terminology and intuition of quasi-circular granularities are shown.
 Panel a shows the time domain distributed as linear granularities $B$ (bottom granularity), $G$, $H'$ and quasi-circular granularities $Q(B, G)$ and $Q(B, H')$.
@@
 Panel b depicts the intuition through a measuring tape analogy where the folds are not of same length but has same length over a period."}
+
 knitr::include_graphics("Figs/quasi-circular-ex2.png")
@@
   ggtitle("") + ylab("") + 
-  scale_x_discrete(labels = c('wdy','wnd'))
+  scale_x_discrete(labels = c('wday','wend')) +
+  ggplot2::theme(axis.text.x = element_text(size = 7))
 
@@
   )) %>%
-  mutate(inning = paste0("inning:", inning)) %>% 
+  mutate(inning = paste0("innings: ", inning)) %>% 
   prob_plot("inning",
@@
   xlab("overs of the innings") +
-  theme(plot.title = element_text(face = "bold"))
-  # ggplot2::theme(
-  #     strip.text = ggplot2::element_text(
-  #       size = 7,
-  #       margin = ggplot2::margin()
-  #     )
-  #   ) 
-  #
+  theme(plot.title = element_text(face = "bold")) +
+  ggplot2::theme(
+       strip.text = ggplot2::element_text(
+        size = 10,
+         margin = ggplot2::margin(b=0, t=0)
+      )
+     )
 
@@
   mutate(
-    field = if_else(fielding_wckts == 0, "0 wicket", ">0 wickets"),
+    field = if_else(fielding_wckts == 0, "0", "1+"),
     dot = if_else(dot_balls == 0, "no dot balls", ">0 dot balls"),
@@
 
-cricket_data$lag_field <- factor(cricket_data$field, levels = c("0 wicket", ">0 wickets"))
+cricket_data$lag_field <- factor(cricket_data$field, levels = c("0", "1+"))
 

### 2020-04-02 hamrony_select.Rmd
rank harmonies and ppt diagram update

---
title: "Ranking harmonies"
author: "Sayani Gupta"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gravitas)
library(philentropy)
```

# Ranking harmonies based on maximum paiwise distance between distributions of levels

Step 1:

For each harmony pair ($C_1, C_2$) with $m$ and $n$ levels respectively , assign a tibble with $C_1$ denoting rows and $C_2$ denoting columns. In the raw data structure, each cell of this $m*n$ order tibble would be a vector of multiple observations.

Step 2:

Since the length of each of these vectors inside each cell is different, the estimated density values of this data is obtained for n = 512(default?) for each cell. We still have a  $m*n$ order tibble with length of vector in each cell equal to 512 now.

Step 3:

For each of these $m$ levels of $C_1$, we need to compute a matrix which provides the pairwise distances of all the $n$ levels of $C_2$. Each of these matrix would be $n*n$ (if symmetric distance obtained just a traingular matric needs to be computed. Thus, in step 3 we have a list of $m$ elements, each of which is a matrix of order $n*n$.

Step 4:

For each of the  $m$ elements, maximum pairwise distance is obtained for each column of the $n*n$ matrix.  Thus, in step 4, maximum pairwise distance for each of the levels of $C_2$ are recorded. Again we have a (m*n) data frame with each cell as a scalar representing the maximum paiwise distances for each level of $C_2$.

Step 5:

Take maximum value of this matrix or take sum of the elements of this matrix to obtain a measure of variation for this harmony pair. The harmony pair with the maximum variation amongst these will be most interesting.

For example, let us take the harmony pair (wknd_wday, hour_day):

```{r}
library(tidyverse)
library(gravitas)
```

# all harmony pairs
```{r}
harmony_data <- smart_meter10 %>% harmony(
  ugran = "month",
  filter_out = c("hhour", "fortnight")
) 
```

# if they have same pairs but one placed on facet and other on x-axis
# then computational load can be minimised by

```{r create_gran}
 data_new <- smart_meter10 %>% 
   create_gran(harmony_data$facet_variable[7]) %>%
   create_gran(harmony_data$x_variable[7]) 
```

```{r step1}
step1 <- data_new[-c(1:2)]  %>% 
  pivot_wider(names_from = harmony_data$x_variable[1],
              values_from = general_supply_kwh,
              values_fn = list(general_supply_kwh = list))

step1
```

```{r step2}
density_extractx <- function(x)
{
  density(x)$y
}

colNms <- colnames(step1)[2:8]

step2 <- NULL

for (i in 1:length(colNms)) {
  step2[[i]] <- lapply(step1[[colNms[i]]], density_extractx)
}
str(step2)
```

```{r step3-4}
compute_JSD <- function(x, y)
{
  mat <- rbind(x, y)
  return(JSD(mat))
}

rowTibb <- step1[, 1]
step3 <- rep(list(diag(nrow(rowTibb))), length(colNms))
#step4 <- matrix(NA, ncol = nrow(rowTibb), nrow = length(colNms))
step4 <- array(NA, dim = nrow(rowTibb))
## Logic
# for each of the list 7 DOW
#__ find the stepped sum difference of density vector elements
for (k in 1:length(colNms)){

  dist <- matrix(NA, 
                 nrow = nrow(rowTibb),
                 ncol = nrow(rowTibb)) ## Matrix
  row_of_col_max <- NULL
  for(i in 1:nrow(rowTibb))
  {
    for (j in 1:nrow(rowTibb))
    {
      m1 <- step2[[k]][[i]]
      m2 <- step2[[k]][[j]]
      message(paste0("K:",k," I:",i," J:", j))
      dist[i, j] <- compute_JSD(m1, m2)
      # row_of_col_max[j] <- max(dist[, j]) 
       row_of_col_max <- max(dist)  
      # maximum of the entire matrix
    }
  }
  step3[[k]] <- dist
  step4[k] <- row_of_col_max
  step5 <- mean(step4)
}

step3
step4
step5
```

### 2020-04-02 main.Rmd
rank harmonies and ppt diagram update

@@
 
-### Unevenly distributed events
-
 Even when there are no rarely occuring events, number of observations might vary hugely within or across each facet.
@@
 <!-- Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels. -->
+
+## Ranking harmonies based on total variability
+
+Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
+Ranking could be based on how well a harmony pair displays the maximum variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+For each facet level, the most important display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
+And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data.
+Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leiber divergence measure.
+Rpobability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
 
@@
 Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
-
-## Computation
-
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:cyclic-calendar}, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} or (for hour_month, day_month, week_month).
+Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:cyclic-calendar}, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
 
@@
 ```{r harmony-tab, echo=FALSE}
+harmonies <- smart_meter10 %>% 
+  harmony(ugran = "month",
+          filter_out = c("hhour", "fortnight"))
+
 smart_harmony <- smart_meter10 %>%
-  harmony(ugran = "month",
-          filter_out = c("hhour", "fortnight"))%>%
+  filter(customer_id %in% c(10017936)) %>% 
+  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh")%>%
   rename(`facet variable` = facet_variable,
          `x-axis variable` = x_variable,
-         `Levels of facet variable` = facet_levels,
-         `Levels of x-axis variable` = x_levels)
+         `facet levels` = facet_levels,
+         `x-axis levels` = x_levels,
+         `average maximum variation` = mean_max_variation)
 
@@
                 caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis along with their number of categories/levels.
-Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.") %>%
+Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.
+The harmony pairs are ranked basis the average maximum variation which captures maximum pairwise distance between x-axis levels for each facets and then average them over facets.") %>%
     #row_spec(0, bold = TRUE) %>% 

### 2020-04-02 harmony_select.R

@@ -1,27 +1,103 @@
+library(tidyverse)
+library(gravitas)
+library(philentropy)
+
 harmony_data <- smart_meter10 %>% harmony(
   ugran = "month",
   filter_out = c("hhour", "fortnight")
 ) 
 
-for(i in 1:nrow(harmony_data))
-{
-  
- data_new <- create_gran(smart_meter10, harmony_data[1,1]$facet_variable)%>% create_gran(harmony_data[1,2]$x_variable)
- 
-data_new$hour_day = as.numeric(data_new$hour_day)
-density(data_new$hour_day)
+ data_new <- smart_meter10 %>% 
+   create_gran(harmony_data[1,1]$facet_variable) %>%
+   create_gran(harmony_data[1,2]$x_variable)
+ 0
+.data_new$hour_day = as.numeric(data_new$hour_day)
+#density(data_new$hour_day)
 data_new$day_week = as.numeric(data_new$day_week)
-density(data_new$day_week)
+#density(data_new$day_week)
+# 
+# data_new[-c(1:2)]  %>% 
+#   pivot_wider(names_from = day_week,
+#               values_from = general_supply_kwh,
+#               values_fn = list(general_supply_kwh = list)) %>%
+#   map(function(x) density(unlist(x)))
 
-data_new[-c(1:2)]  %>% 
+
+step1 <- data_new[-c(1:2)]  %>% 
   pivot_wider(names_from = day_week,
               values_from = general_supply_kwh,
-              values_fn = list(general_supply_kwh = list)) %>% 
-  map(unlist) 
+              values_fn = list(general_supply_kwh = list))
 
 
-%>% map_dfr(unlist)
-  
-    
+density_extractx <- function(x)
+{
+  density(x)$y
 }
-    
+
+colNms <- colnames(step1)[2:8]
+
+step2 <- NULL
+
+for (i in 1:length(colNms)) {
+  step2[[i]] <- lapply(step1[[colNms[i]]], density_extractx)
+}
+#names(step2) <- d$hour_day
+str(step2)
+
+
+# mat1 <- matrix(1:6, nrow=2)
+# mat2 <- matrix(2:7, nrow=2)
+sumDiff_mat <- function (mat1, mat2) {
+  sum(mat2 - mat1) 
+}
+
+compute_JSD <- function(x, y)
+{
+  mat <- rbind(x, y)
+  return(JSD(mat))
+}
+
+
+rowTibb <- step1[, 1]
[…]

### 2020-04-02 simulation.R

@@ -1,0 +1,64 @@
+
+library(gravitas)
+library(tsibbledata)
+library(dplyr)
+library(ggplot2)
+
+
+cust50 <- sm_cust50 %>%  distinct(customer_id)
+
+VIC <- sm_cust50 %>% 
+  filter(customer_id==cust50$customer_id[2])
+
+VIC %>%
+  #create_gran("quarter_year") %>%
+  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  #create_gran("day_week") %>%
+  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("quarter_year", "wknd_wday",
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            symmetric = FALSE
+  ) +
+  ylab("Electricity Demand [KWh]") +
+  xlab("weekend/weekday") + 
+  ggtitle("(a) Letter value plot with quarter-of-year facet")+
+  scale_fill_brewer(palette = "Dark2") 
+
+
+
+VIC %>%
+  #create_gran("month_year") %>%
+  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  #create_gran("day_week") %>%
+  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("wknd_wday",
+            "quarter_year",
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            symmetric = FALSE
+  ) +
+  ylab("") +
+    xlab("quarter-of-year")  + 
+  ggtitle("(b) Letter value plot with weekend/weekday facet") +
+  scale_fill_brewer(palette = "Paired") 
+
+
+
+
+VIC %>%
+  #create_gran("month_year") %>%
+  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
+  #create_gran("day_week") %>%
+  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
+  prob_plot("quarter_year",
+            "month_year",
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            k = 3, 
+            symmetric = FALSE
+  ) +
+  ylab("") +
+  xlab("day_month")  + 
+  ggtitle("(c) Letter value plot across quarters & months") 
+

### 2020-04-02 harmony_select.R
rank harmonies code deleted and included on package

@@ -1,27 +1,0 @@
-harmony_data <- smart_meter10 %>% harmony(
-  ugran = "month",
-  filter_out = c("hhour", "fortnight")
-) 
-
-for(i in 1:nrow(harmony_data))
-{
-  
- data_new <- create_gran(smart_meter10, harmony_data[1,1]$facet_variable)%>% create_gran(harmony_data[1,2]$x_variable)
- 
-data_new$hour_day = as.numeric(data_new$hour_day)
-density(data_new$hour_day)
-data_new$day_week = as.numeric(data_new$day_week)
-density(data_new$day_week)
-
-data_new[-c(1:2)]  %>% 
-  pivot_wider(names_from = day_week,
-              values_from = general_supply_kwh,
-              values_fn = list(general_supply_kwh = list)) %>% 
-  map(unlist) 
-
-
-%>% map_dfr(unlist)
-  
-    
-}
-    

### 2020-04-02 main.R

@@ -1,0 +1,1861 @@
+### Supplemental Materials (Data and scripts)
+
+
+
+
+
+
+
+## ---- load
+library(knitr)
+library(tidyverse)
+library(lubridate)
+library(lvplot)
+library(ggridges)
+library(tsibble)
+library(gravitas)
+library(tsibble)
+library(ggpubr)
+# install.packages("devtools")
+# devtools::install_github("Sayani07/gravitas")
+library(gravitas)
+
+## ---- linear-time
+knitr::include_graphics("Figs/linear-arr.png")
+
+## ---- mvtsplot1
+# Figure 1 : Multiple parallel time series plot of a real world data set
+# obtained using a fiber optic cable
+t <- nrow(anomalous_stream_1)
+f <- ncol(anomalous_stream_1)
+g <- as_tibble(anomalous_stream_1) %>%
+  gather() %>%
+  mutate(key = rep((1:f), each = t)) %>%
+  mutate(Time = rep(1:t, f))
+colnames(g) <- c("Cable", "Value", "Time")
+p1 <- ggplot(g, aes(x = Time, y = Cable, fill = Value)) +
+  geom_tile() +
+  scale_fill_gradientn(
+    colours = c("#F0E442", "#000000", "#000000"),
+    values = c(0, 0.1, max(anomalous_stream_1))
+  ) +
+  ylab("Cable (Sensor ID / Time Series ID)") +
+  scale_y_continuous(breaks = seq(0, 600, 100)) +
+  scale_x_continuous(breaks = seq(0, 1400, 200)) +
+  xlab("Time") +
+  theme(legend.position = "none", axis.title = element_text(size = 15))
+print(p1)
+
+
+
+
+
+
+
+## ---- outtype
+# Figure 2: Different types of anomalies in temporal data
+set.seed(1234)
+ts1 <- as.ts(rnorm(500, 10, 3))
+ts1[300] <- 35
+ts1[150] <- 30
+ts1 <- as_tibble(ts1) %>% mutate(Time = 1:500)
+colnames(ts1) <- c("Value", "Time")
+point <- tibble(Time = c(150, 300), Value = c(30, 35))
+p1 <- ggplot(ts1) +
+  geom_line(aes(x = Time, y = Value), size = 0.5, alpha = 0.7) +
[…]

### 2020-04-02 harmony_select.R
delete redundant codes

@@ -1,103 +1,0 @@
-library(tidyverse)
-library(gravitas)
-library(philentropy)
-
-harmony_data <- smart_meter10 %>% harmony(
-  ugran = "month",
-  filter_out = c("hhour", "fortnight")
-) 
-
- data_new <- smart_meter10 %>% 
-   create_gran(harmony_data[1,1]$facet_variable) %>%
-   create_gran(harmony_data[1,2]$x_variable)
- 0
-.data_new$hour_day = as.numeric(data_new$hour_day)
-#density(data_new$hour_day)
-data_new$day_week = as.numeric(data_new$day_week)
-#density(data_new$day_week)
-# 
-# data_new[-c(1:2)]  %>% 
-#   pivot_wider(names_from = day_week,
-#               values_from = general_supply_kwh,
-#               values_fn = list(general_supply_kwh = list)) %>%
-#   map(function(x) density(unlist(x)))
-
-
-step1 <- data_new[-c(1:2)]  %>% 
-  pivot_wider(names_from = day_week,
-              values_from = general_supply_kwh,
-              values_fn = list(general_supply_kwh = list))
-
-
-density_extractx <- function(x)
-{
-  density(x)$y
-}
-
-colNms <- colnames(step1)[2:8]
-
-step2 <- NULL
-
-for (i in 1:length(colNms)) {
-  step2[[i]] <- lapply(step1[[colNms[i]]], density_extractx)
-}
-#names(step2) <- d$hour_day
-str(step2)
-
-
-# mat1 <- matrix(1:6, nrow=2)
-# mat2 <- matrix(2:7, nrow=2)
-sumDiff_mat <- function (mat1, mat2) {
-  sum(mat2 - mat1) 
-}
-
-compute_JSD <- function(x, y)
-{
-  mat <- rbind(x, y)
-  return(JSD(mat))
-}
-
-
-rowTibb <- step1[, 1]
-step3 <- rep(list(diag(nrow(rowTibb))), length(colNms))
-step4 <- matrix(NA, ncol = nrow(rowTibb), nrow = length(colNms))
-## Logic
-# for each of the list 7 DOW
-#__ find the stepped sum difference of density vector elements
-for (k in 1:length(colNms)){
-  
-  dist <- matrix(NA, 
-                 nrow = nrow(rowTibb),
-                 ncol = nrow(rowTibb)) ## Matrix
-  row_of_col_max <- NULL
[…]

### 2020-04-06 main.Rmd
consistent with Di's revised way of writing definitions

@@
 
-\begin{definition}[Circular granularity]\label{def:circular}
-A circular granularity $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
+\begin{definition}\label{def:circular}
+A {\bf circular granularity} $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
 \begin{equation} \label{eq:circular-gran}
@@
 z denotes the index set,
-B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping,
+B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, and 
 $P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$.
@@
 \begin{definition}\label{def:quasicircular}
-(quasi-circular granularity): A quasi-circular granularity $Q_{B, G'}$ that relates linear granularities $G'$ and bottom granularity $B$, if
-
+A {\bf quasi-circular granularity} $Q_{B, G'}$ relates linear granularities $G'$ and bottom granularity $B$, if
 \begin{equation}\label{eq:quasi}
-Q_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), \quad z \in T_k
+Q_{B, G'}(z) = z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert,for\quad z \in T_k
 \end{equation}
-
-where
+where,
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ denote the period of the grouping $(B, G')$ and the number of granules of $G'$ in each of these periods, ${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$, $T_w$ are the sets of indices of $B$ describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ are the period of the grouping $(B, G')$ and the number of granules of $G'$ in each of these periods, $T_w$ are the sets of indices of $B$ describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{definition}
+
+<!--${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$,-->
 
@@
 
-\begin{definition}[Aperiodic cyclic granularity]\label{aperiodic}
-An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
+\begin{definition}\label{aperiodic}
+An {\bf aperiodic cyclic granularity} $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
 \begin{equation}\label{eq:aperiodic}
 \begin{split}
-A_{B, M}(z) & = L(i), \quad z \in T_{i_j} \quad and \quad \forall i \in {1, 2, \dots, n}\\
-     & = L(0), \quad otherwise
+A_{B, M}(z) & =
+\begin{cases}
+i, & \text{for}\quad z \in T_{i_j} \\
+0  & \text{otherwise} 
+\end{cases}
 \end{split}
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{i=1}^{n}M_{i}$.
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
 \end{definition}
+
+<!-- \forall i \in {1, 2, \dots, n}\\ -->
 
@@
 
-\begin{definition}[Clash]\label{def:harmony}
-A pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
+\begin{definition}\label{def:harmony}
+A {\bf clash} is a pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
 \end{definition}
 
-\begin{definition}[Harmony]\label{def:clash}
-A pair of cyclic granularities that do not contain any missing combinations of its categories.
+\begin{definition}\label{def:clash}
+A {\bf harmony} is a pair of cyclic granularities that do not contain any missing combinations of its categories.
 \end{definition}
@@
 Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
-Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:cyclic-calendar}, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
-
-The index set maps the half-hourly time instants to a set of positive integers.
-$$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular})
-$$ hour\_week(z) =  day\_week(z) + 24*hour\_day(z)$$ (Section \ref{sec:all-circular})
+Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Sections \ref{sec:cyclic-calendar}.
+<!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
+-->
+Here, the index set $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ maps the half-hourly time instants to a set of positive integers.
+Then, multiple-order-up circular granularity $hour\_week$ and quasi-circular granularity $hour\_month$ could be computed from single-order-up granularities as follows:
+<!-- $$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular}) -->
+\begin{align}
+  hour\_week(z) =  day\_week(z) + 24*hour\_day(z)
+\end{align}
+
 <!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) -->
-$$ hour\_month(z) =  hour\_day(z) + 24*day\_month(z)$$ (Section \ref{sec:notall-circular})
+\begin{align}
+hour\_month(z) =  hour\_day(z) + 24*day\_month(z)
+\end{align}
 

### 2020-04-11 main.Rmd
choosing harmonies section fix

@@
 Panel b and c depicts the intuition through a measuring tape analogy where the linear measuring tape has been folded to obtain circular granularities from linear ones to accomodate for repeating patterns."}
-knitr::include_graphics("Figs/circular-illustration.png")
+knitr::include_graphics("Figs/circular-example.png")
 ```
@@
 -->
+
+## Choosing harmonies to show
+
+Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
+Ranking could be based on how well a harmony pair displays the maximum variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+For each facet level, the most important display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
+And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data.
+Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leiber divergence measure.
+Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+<!-- Suppose we have ${x_1, x_2, \dots, x_n} which are random sample from some underlying population. --> <!-- Let $p(x)$ and $q(x)$ be two probability distributions of a random variable x. -->
+<!-- We compute Kullback-Liebler distance based on quantiles $q_1$ and $q_2$ of $v$ at probabilities $p$ corresponding to $C_1$ and $C_2$ respectively. -->
+
+<!-- \begin{equation} -->
+<!-- D(q_1,q_2) = \int^{\infty}_{-\infty}q_1(x)logq_1(x)/q_2(x)\,dx -->
+<!-- \end{equation} -->
+
+<!-- But since this is not symmetric, Jensen-Shanon distance is calculated which is based on Kullback-Liebler divergence. -->
+
+<!-- $$JS(q_1, q_2) = [D(q_1, r) + D(q_2, r)]/2 where r = (q_1 + q_2)/2$$ -->
+
+1.
+Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
+Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
+
+2.
+There will be $m$ time series variable corresponding to each level $A_i$ of $C_1$.
+We compute the pairwise differences in distributions of these $m$ time series variables.
+Thus there will be  ${m \choose 2}$ pairwise distances correponding to each $A_i$'s.
+
+3.
+Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+suppose $q$ is the sample quantile vector computed for probability vector $p$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by:
+
+$$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$
+where,
+ $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)logq_j(x)/q_k(x)\,dx$$
+is the Kullback-Leibler divergence of $q_j$ with respect to $q_k$.
+
+4.
+For each $\forall \i = {1, 2, \dots, l}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{j} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+Here, maximum is being adjusted against the range of the data to accomodate for comparison of maximum distances for different ranges.
+
+5. The mean maximum pairwise distance (MMPD) over l = {1, 2, \dots, l}$ is computed by 
+$MMPD = (1/l)\sum_{1}^{l}{D_{j}}$.
+
+6.
+Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD since any harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+
+<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
 

### 2020-04-12 main.Rmd
include new diagrams

@@
 Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
-Analysts are expected to iteratively explore many possible choices of time granularities for comprehending possible periodicities in the data.
+Analysts are expected to alliteratively explore many possible choices of time granularities for comprehending possible periodicities in the data.
 But too many choices and a lack of a systematic approach to do so might become overwhelming.
@@
 However, this approach can lead to a myriad of choices all of which are not useful.
-Analysts are expected to iteratively visualize these choices for exploring possible patterns in the data.
+Analysts are expected to alliteratively visualize these choices for exploring possible patterns in the data.
 But too many choices might leave him bewildered.
@@
 Panel b and c depicts the intuition through a measuring tape analogy where the linear measuring tape has been folded to obtain circular granularities from linear ones to accomodate for repeating patterns."}
-knitr::include_graphics("Figs/circular-example.png")
+knitr::include_graphics("Figs/circular-example-1.png")
 ```
@@
 
-knitr::include_graphics("Figs/quasi-circular-ex2.png")
+# knitr::include_graphics("Figs/quasi-circular-ex2.png")
 ```
@@
 Panel b depicts the intuition through a measuring tape analogy which needs to be folded at different points every year since $M1$ and $M2$ will fall in different weeks."}
-knitr::include_graphics("Figs/aperiodic-ex2.png")
+knitr::include_graphics("Figs/quasi-periodic-example.png")
 ```
@@
 But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities could be a potential way to compare and contrast periocities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
+Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities could be a potential way to compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
 Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
@@
 <!-- Graphical output devices are flat.
-If there are only two variables, a cartesian graph of one variable against the other can show the configuration of points.
+If there are only two variables, a Cartesian graph of one variable against the other can show the configuration of points.
 As soon as the data move to three variables, we need to infer multi-dimensional structure by a two-dimensional medium[Cleveland, elements of graphing data].
@@
 We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  ${m \choose 2}$ pairwise distances correponding to each $A_i$'s.
+Thus there will be  ${m \choose 2}$ pairwise distances corresponding to each $A_i$'s.
 
@@
 4.
-For each $\forall \i = {1, 2, \dots, l}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{j} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
-Here, maximum is being adjusted against the range of the data to accomodate for comparison of maximum distances for different ranges.
+For each $\forall i = {1, 2, \dots, l}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{j} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
 
@@
 
-Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occuring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
+Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
 
 <!-- CONTENT CUT -->
-<!-- ### Rarely occuring events -->
+<!-- ### Rarely occurring events -->
 
@@
 
-Even when there are no rarely occuring events, number of observations might vary hugely within or across each facet.
+Even when there are no rarely occurring events, number of observations might vary hugely within or across each facet.
 This might happen due to missing observations in the data or uneven locations of events in time domain.
@@
 Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leiber divergence measure.
-Rpobability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
 
@@
 <!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarise(median= log(median(general_supply_kwh)), -->
+<!-- #   summarize(median= log(median(general_supply_kwh)), -->
 <!-- #             total = log(sum(general_supply_kwh))) %>%  -->
@@
 <!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarise(median_consumption= median(general_supply_kwh)) %>% -->
+<!-- #   summarize(median_consumption= median(general_supply_kwh)) %>% -->
 <!-- #   ggplot2::ggplot(aes( -->
@@
 
-<!-- However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal dataset similar to any temporal dataset.
+<!-- However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal data set similar to any temporal data set.
 -->
@@
 
-However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal dataset similar to any temporal dataset.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal data set similar to any temporal data set.
 There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
@@
 
-<!-- First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperidic cyclic granularity) in \autoref{fig:seas-over-inning}.
+<!-- First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperiodic cyclic granularity) in \autoref{fig:seas-over-inning}.
 The distribution of runs per over has not significantly changed from 2008 to 2016.
@@
 
-Exploratory data analysis is iteratively finding and summarizing patterns.
+Exploratory data analysis is alliteratively finding and summarizing patterns.
 With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context.
@@
 A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
-These functionalities can be imported in the gravitas package to accommodate for aperiodic cyclic granularities.
+These functionalists can be imported in the gravitas package to accommodate for aperiodic cyclic granularities.
 
@@
 
-The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
-Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
+The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Mon ash University for sharing their wisdom and experience of developing R packages, Dr.
+Peter Tosca [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-We would also like to thank Nicholas Spyrison for many useful discussions and feedback on the manuscipt.
+We would also like to thank Nicholas Prison for many useful discussions and feedback on the manuscript.
 More details about the package

### 2020-04-13 main.Rmd
minor modifications on spelling and grammar

@@
 
-The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provide some background material on linear granularities and introduces calendar algebra for computing different linear granularities.
+The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provides some background material on linear granularities and introduces calendar algebra for computing different linear granularities.
 Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
@@
 
-For example, $day \trianglelefteq week$ and $day \preceq week$ since every granule of "week" is the union of some set of granules of "day" and each "day" is a subset of a "week".
+For example, $day \trianglelefteq week$ and $day \preceq week$ since every granule of $week$ is the union of some set of granules of $day$ and each $day$ is a subset of a $week$.
 Consider another example where $day \trianglelefteq month$.
 This relationship however is incomplete without its association to periodicity.
-Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
+Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
 The period becomes 400 years with the inclusion of leap years and all their exceptions.
@@
 
-\noindent For example, in a Gregorian calendar without leap years we could say days group periodically into months with the period $P = 365$ and the number of granules of "months" in each period given by $R=12$.
-In a Gregorian calendar with leap years, days group quasi-periodically into months with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
+\noindent For example, in a Gregorian calendar without leap years we could say $day$ groups periodically into $month$ with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
+In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
 
@@
 
-**Example**: Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
+**Example**: 
+<!-- Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
 Each granule is represented by a box.
-Two linear granularities $G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))$ are considered.
-<!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
-Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> $Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
+-->
+
+Consider two linear granularities $G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))$.<!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
+Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
 For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
-For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:circular-gran}.
+For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:eq:eq5}.
 
@@
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple:** and (2) **multiple-to-single:** which entails the representation of multiple-order-up cyclic granularities from single order-up cyclic granularities and vice versa.
+(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of multiple-order-up cyclic granularities from single order-up cyclic granularities and vice versa.
 The hierarchical structure of time creates a natural nested ordering which can produce _single-order-up_ or _multiple-order-up_ granularities.
@@
 The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities.
-Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities each relating one finer linear granularity with a coarser one could be constructed.
+Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
 Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
@@
 Then the days denoting weekends in a month would not correspond to any Business days.
-This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does.
+This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the choice of the granularities does.
 
@@
 
-Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity correspond to multiple values of the measured variable.
+Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
 This is unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
@@
 
-## Choosing harmonies to show
-
-Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
-Ranking could be based on how well a harmony pair displays the maximum variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+<!-- ## Choosing harmonies to show -->
+
+<!-- Even after screening harmonies from the list of all possible pairs of cyclic time granularities, the options could be too many and it might be useful to rank these harmonies.
+Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most important display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
 And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data.
-Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leiber divergence measure.
+Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
 Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+-->
 <!-- Suppose we have ${x_1, x_2, \dots, x_n} which are random sample from some underlying population. --> <!-- Let $p(x)$ and $q(x)$ be two probability distributions of a random variable x. -->
@@
 <!-- $$JS(q_1, q_2) = [D(q_1, r) + D(q_2, r)]/2 where r = (q_1 + q_2)/2$$ -->
-
-1.
-Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
-
-2.
-There will be $m$ time series variable corresponding to each level $A_i$ of $C_1$.
-We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  ${m \choose 2}$ pairwise distances corresponding to each $A_i$'s.
-
-3.
-Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
-suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by:
-
-$$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$
-where,
- $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)logq_j(x)/q_k(x)\,dx$$
-is the Kullback-Leibler divergence of $q_j$ with respect to $q_k$.
-
-4.
-For each $\forall i = {1, 2, \dots, l}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{j} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
-Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
-
-5. The mean maximum pairwise distance (MMPD) over l = {1, 2, \dots, l}$ is computed by 
-$MMPD = (1/l)\sum_{1}^{l}{D_{j}}$.
-
-6.
-Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD since any harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
-
-<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
 
@@
 
-## Ranking harmonies based on total variability
+## Choosing harmonies to show
 
@@
 And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data.
-Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leiber divergence measure.
+Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
 Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+
+1.
+Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
+Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
+
+2.
+There will be $m$ time series variable corresponding to each level $A_i$ of $C_1$.
+We compute the pairwise differences in distributions of these $m$ time series variables.
+Thus there will be  ${m \choose 2}$ pairwise distances corresponding to each $A_i$'s.
+
+3.
+Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+Suppose $q$ is the sample quantile vector computed for probability vector $p$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by:
+
+$$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$
+where,
+ $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$
+is the Kullback-Leibler divergence of $q_j$ with respect to $q_k$.
+
+4.
+For each $\forall i = {1, 2, \dots, l}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
+
+5. The mean maximum pairwise distance (MMPD) over $l = {1, 2, \dots, l}$ is computed by 
+$MMPD = (1/l)\sum_{i=1}^{l}{D_{j}}$.
+
+6.
+Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD since any harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+
+<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
 
@@
 The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-We would also like to thank Nicholas Prison for many useful discussions and feedback on the manuscript.
+We would also like to thank Nicholas Spyrison for many useful discussions and feedback on the manuscript.
 More details about the package

### 2020-04-13 main.Rmd
circular granularities section - diagram caption and period terminology made more clear

@@
 z denotes the index set,
-B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, and 
-$P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$.
+B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, and $P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$, also called the period of the grouping $(B, G)$.
 \end{definition}
@@
 
-\autoref{fig:circular-dow} depicts the circular granularity "Day-of-Week", constructed from linear granularities "Day" (the bottom granularity) and "Week".
-The granules that overlap share elements from the underlying time domain.
-In this example, $P=7$, and the circular granularity $C_{\text{day}, \text{week}}$ representing Day-of-Week consists of the repeated pattern $\{0, 1, 2, 3, 4, 5, 6\}$.
+\autoref{fig:circular-dow} depicts the linear (a and b) and the corresponding stack (c and d) display of time.
+The circular granularity day-of-week and day-of-fortnight are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$).
+The granules that overlap share elements from the underlying time domain in the linear display.
+The circular granularity $C_{B , G}$ and $C_{B, H}$ <!--= C_{\text{day}, \text{week}}$--> representing day-of-week and day-of-fortnight respectively consist of repeated pattern of categories $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
+The granules that overlap represents the same category in the stack display.
 Each circular granularity can use descriptive label mappings, or be simply labelled using the index.
-<!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for \autoref{fig:circular-dow} can be defined as
+<!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
 $$
@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Terminology and intuition of circular granularities are shown.
-Panel a shows the time domain distributed as linear granularities Day and Week and circular granularity Day-of-Week with $P=7$ is constructed using the two linear granularities.
-Day-of-Week has a repeating pattern unlike Day and Week and has an ordering only within a period represented by hashed arrows.
-Panel b and c depicts the intuition through a measuring tape analogy where the linear measuring tape has been folded to obtain circular granularities from linear ones to accomodate for repeating patterns."}
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="  Circular granularities illustrated through linear (a and b) and stack (c and d) display of time.
+The linear display is similar to a measuring tape with the measurement marking distributed as linear granularities day, week and fortnight.
+The stack display gives the intuition of circular granularities by cutting the measuring tape and stacking on top of one another to represent repetitive pattern of categories.
+The granules that overlap represent same time domain in the linear display and same category in the stack display.
+Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time."}
 knitr::include_graphics("Figs/circular-example-1.png")
@@
 
-In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H)$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$.
-Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity.
-<!--The representation of the resultant circular granularity will be a function of the index set.--> The bottom granularity is assumed to be minutes in this case.
+In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H)$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping and $P(G,H)$ is the period of the grouping $(G, H)$  <!--such that the number of granules of $G$ in each granule of $H$ is-->.
+Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with period $P_i$ <!--being the number of granules of the finer granularity in each granule of the coarser granularity.The representation of the resultant circular granularity will be a function of the index set.--> and minutes as the bottom granularity.
 
@@
 \hline
-circular granularity & expression & period length \\
+circular granularity & expression & period \\
 \hline
@@
 \caption{Examples of circular granularities with bottom granularity minutes.
-Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and $P_i$ represents the period length.
+Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
 Circular granularities could be expressed using modular arithmetic due to their regular mapping.
@@
 For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
-For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:eq:eq5}.
+For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
 
@@
 We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  ${m \choose 2}$ pairwise distances corresponding to each $A_i$'s.
+Thus there will be  ${m \choose 2}$ pairwise distances corresponding to each $A_i$.
 

### 2020-04-13 main.Rmd
quasi-circular section: diagram and examples update

@@
 \caption{Examples of quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
-These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping and hence can not be expressed through modular arithmetic.
+These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping leading many potential period lengths and hence can not be expressed through modular arithmetic.
 }
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ are the period of the grouping $(B, G')$ and the number of granules of $G'$ in each of these periods, $T_w$ are the sets of indices of $B$ describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ are the period of the grouping $(B, G')$ and the number of granules of $G'$ in $P'$, $T_w$ are the sets of indices of $B$ describing $G'(w)$ such that  $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{definition}
@@
 
-**Example**: 
+<!-- **Example**:  -->
 <!-- Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
@@
 
-Consider two linear granularities $G' = Alter^{2}_{(1,-1)}(BG, Group_3(B))$ and $H' = Alter^{2}_{(1,-2)}(BG, Group_7(B))$.<!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
-Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ and $Q_{B, H'}$ are repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$ or $H'$.
-For $G'$, $T_0 = \{0,1\}$ and $T_1 = \{2, 3, 4\}$.
-For $H'$, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.
-
-\begin{equation} \label{eq:eq4}
-\begin{split}
-Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\
-  & = L(1) \\
-\end{split}
-\end{equation}
+Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
+Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
+Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$.
+Hence using \ref:{def:quasicircular}, then we will have:
+
+<!-- \begin{equation} \label{eq:eq4} -->
+<!-- \begin{split} -->
+<!-- Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\ -->
+<!--   & = L(1) \\ -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
 
@@
 
-\begin{equation} \label{eq:eq5}
+\begin{align}
 \begin{split}
-Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,\quad since \quad  10 \in T_{1}  \\
-  & = L(3) \\
+Q_{B,H'}(10) & = 10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert ,\quad since \quad  10 \in T_{1}  \\
+  & = 3 \\
 \end{split}
-\end{equation}
-
-**Discussion**: If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
+\end{align}
+
+<!-- **Discussion**:  -->
+If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
 Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
@@
 The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$.
-The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$.
+The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ in Definition \autoref{def:quasicircular} denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$.
 Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.
 
-```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap="Terminology and intuition of quasi-circular granularities are shown.
+```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
+The linear display shows linear granularities days, weeks, semester weeks, semester Terminology and intuition of quasi-circular granularities are shown.
 Panel a shows the time domain distributed as linear granularities $B$ (bottom granularity), $G$, $H'$ and quasi-circular granularities $Q(B, G)$ and $Q(B, H')$.
@@
 ```
+
+<!-- Circular granularities illustrated through linear (a and b) and stack (c and d) display of time.
+The linear display is similar to a measuring tape with the measurement marking distributed as linear granularities day, week and fortnight.
+The stack display gives the intuition of circular granularities by cutting the measuring tape and stacking on top of one another to represent repetitive pattern of categories.
+The granules that overlap represent same time domain in the linear display and same category in the stack display.
+Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time.
+-->
 
@@
 
-```{r aperiodic-example, echo=FALSE,out.width = "100%",fig.cap="Terminology and intuition of aperiodic cyclic granularities are shown.
-Panel a shows the time domain distributed as linear granularities Day, Semester, within semester breaks, in-session and aperiodic cyclic granularity Week-of-Semester.
-$A(B, M)$ consists of three categories 0, 1, 2 non-semester period, within-semester breaks and in-session semester period respectively.
-Panel b depicts the intuition through a measuring tape analogy which needs to be folded at different points every year since $M1$ and $M2$ will fall in different weeks."}
-knitr::include_graphics("Figs/quasi-periodic-example.png")
+```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
+The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future.
+Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks after which there is a study break of 1 week before final exams which continue for 1.3 weeks.
+This pattern remains same for all semester and hence $Q_{H, M}$ with $P = 18.3 weeks$ denoting week of the semester week type will be a quasi-circular granularity with repeating patterns.
+$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years (shown through c)."}
+
+knitr::include_graphics("Figs/quasi.png")
+
 ```
-
-<!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
+<!-- #  -->
+<!-- # <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. --> -->
 

### 2020-04-13 main.Rmd
aperiodic section update - diagram and caption

@@
 Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$.
-Hence using \ref:{def:quasicircular}, then we will have:
+Hence using \ref{def:quasicircular}, then we will have:
 
@@
 
-\begin{definition}\label{aperiodic}
-An {\bf aperiodic cyclic granularity} $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
+\begin{definition}\label{def:aperiodic}
+An {\bf aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if
 \begin{equation}\label{eq:aperiodic}
@@
 $z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
+$B$ denotes a full-integer labelled bottom granularity which groups into linear granularity $M$ but not periodically, $T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
 \end{definition}
@@
 
-**Example**: \autoref{fig:aperiodic-example} shows two aperiodic linear granularity $M_1$ and $M_2$.
+Consider a school semester which lasts for 18.3 weeks, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks after which there is a study break of 1 week before final exams which continue for 1.3 weeks.
+Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
+Both $M_1$ and $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$).
+Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity.
+This is because the placement of the semester within an year would vary across years (\autoref{fig:aperiodic-example} c).
+\autoref{fig:aperiodic-example} b shows the stack display of time with granules representing same categories (semester break/in-session) stacked on top of one another.
+\autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}.
+It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester remains constant over years.
+Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.
+
 <!--Each granule is represented by a box.
 The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
--->The first slot in the diagram shows the index set and $A(B, M)$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.
-
-**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labor day and M can be conceived as the aperiodic linear granularity denoting any holidays.
+The first slot in the diagram shows the index set and $A(B, M)$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.-->
+
+<!-- Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labor day and M can be conceived as the aperiodic linear granularity denoting any holidays.
 <!--The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.-->
@@
 <!-- #  -->
-<!-- # <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. --> -->
+<!-- # <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
 

### 2020-04-15 main.Rmd
calendar algebra minor changes

@@
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of multiple-order-up cyclic granularities from single order-up cyclic granularities and vice versa.
-The hierarchical structure of time creates a natural nested ordering which can produce _single-order-up_ or _multiple-order-up_ granularities.
+(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities and vice versa.
+The hierarchical structure of time creates a natural nested ordering which can produce single-order-up or multiple-order-up granularities.
 We shall use the notion of a hierarchy table and order to define them.
@@
 **Order** of a linear granularity can be comprehended as the level of coarseness associated with a linear granularity.
-For example, for two linear granularities $G$ and $H$, if $G$ groups into $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+For example, for two linear granularities $G$ and $H$, if $G$ groups into or finer than $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
 In any hierarchy table, linear granularities are arranged from lowest to highest order.
@@
 $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
-$K \equiv P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ for circular granularity $C_{G(l),G(m)}$ and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
+$K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
 We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
@@
 <!-- some discussions on primary units of calendar -->
-**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
+<!-- **Example:**  -->
+
+So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
 But it is far from being the only one.
@@
 \begin{split}
-C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{(G_{l+1},G_m)}(z))) \\
-  & =   L(C_{(G_l,G_{l+1}}(z) + P(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)( C_{G_{l+2}, G_m}(z))) \\
-  & =  L(C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+1)P(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
-  & =  L(C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
+C_{G_l,G_m}(z) & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_m}(z)) \\
+  & = C_{(G_l,G_{l+1}}(z) + P(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)]( C_{G_{l+2}, G_m}(z)) \\
+  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+1)P(l+1, l+2)(C_{G_{l+2}, G_{m}})\\
+  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+2)(C_{G_{l+2},G_{l+m}}(z))\\
   &\vdots\\
-  & = L(\sum_{i=0}^{m - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z)))\\
+  & = \sum_{i=0}^{m - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))\\
 \end{split}
@@
 
-, for $l < m + 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$
-
-**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar.
+, for $l < m - 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$
+
+**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar with the index set  $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ mapping "kins" to a set of positive integers.
 
@@
 \begin{split}
-C_{uinal, baktun}(z)  = & L(C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + P(uinal, katun)C_{katun, baktun}(z)) \\
-               = & L(\lfloor z/20\rfloor \mod 18 + 18*\lfloor z/20*18\rfloor \mod 20  \\
-               &~~~ + 18*20\lfloor z/20*18*20\rfloor \mod 20) \\
+C_{uinal, baktun}(z)  = & C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + P(uinal, katun)C_{katun, baktun}(z) \\
+               = & \lfloor z/20\rfloor \mod 18 + 18*\lfloor z/20*18\rfloor \mod 20  \\
+               &~~~ + 18*20\lfloor z/20*18*20\rfloor \mod 20 \\
 \end{split}
@@
 \begin{split}
-C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\
+C_{G_l,G_{m}}(z) & = C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)) \\
 \end{split}
@@
 \begin{split}
-C_{G_l,G_{l+2}}(z) & = L(C_{G_{l},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert))\\
+C_{G_l,G_{m}}(z) & = C_{G_{l},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
 \end{split}
@@
 
- - $C_{hour, month}(z) = L(C_{hour, day}(z) + P(hour, day)*C_{day, month}(z))$  
- - $C_{day, year}(z) = L(C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert))$, where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$.
+ - $C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)$  
+ - $C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert)$, where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$
 
@@
 \begin{equation} \label{eq:all-circular-multiple}
-C_{G_{l_1}, G_{m_1}}(z) = C_{G_{l_2}, G_{m_2}}(\lfloor z/k(l_2,l_1) \rfloor\mod P(m_1, m_2))
+C_{G_{l_1}, G_{m_1}}(z) = \lfloor C_{G_{l_2}, G_{m_2}}(z)/P(l_2,l_1) \rfloor \mod P(l_1, m_1)
 \end{equation}
 
-**Example:** Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
+<!-- **Example:**  -->
+Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
 
@@
 \begin{split}
-C_{tun, katun}(z) =  L(\lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20)
+C_{tun, katun}(z) =  \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20
 \end{split}
@@
 The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
-Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
+Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}, let $S_{ij}$ be a subset of the index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
 Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, l}$ and $j \in {1, 2, \dots, m}$  which will lead to $lm$ data subsets.

### 2020-04-15 main.Rmd
till visuyalization minor

@@
 Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$ such that $C_i$ maps index set to $n_i$ levels or categories and $v$ denotes the measurement variable.
+This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$ and $v$ denotes the measurement variable.
 The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities.
@@
 The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
-Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}, let $S_{ij}$ be a subset of the index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
+Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
+Here, $A_i$'s or $B_j$'s are the levels/categories corresponding to $C_1$ and $C_2$ respectively.
+Let $S_{ij}$ be a subset of the index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
 Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, l}$ and $j \in {1, 2, \dots, m}$  which will lead to $lm$ data subsets.
@@
 Then the days denoting weekends in a month would not correspond to any Business days.
-This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the choice of the granularities does.
+This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularities does.
 
@@
 However, we need to consider the effect of number of observations on the summarization even for harmonies.
-Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
-Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
-There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
-If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge mnk$ will be acceptable.
+Suppose we have $T$ observations, and two cyclic granularities $C_1$ and $C_2$ with $l$ and $m$ categories respectively.
+Each element of $C_1$ occurs approximately $T/l$ times while each element of $C_2$ occurs approximately $T/m$ times.
+There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(lm)$.
+If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge lmk$ will be acceptable.
 The value of $k$ will depend on the statistical transformation we are producing.
@@
 For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarter-of-year.
-This plot does not work because every quarter does not correspond to all months of the year for example, the first quarter would not correspond to the month December of an year.
+This plot does not work because every quarter does not correspond to all months of the year, for example, the first quarter would not correspond to the month December of an year.
 
@@
 Plot (b) helps to compare weekends and weekdays for each quarter.
-It can be seen that for every quarter weeekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently, whereas, Plot (c) helps to compare quarters within weekdays and weekends.
-For example, the quartile spread of consumption shrinks/lowers from first to fourth quarter, whereas this pattern is not true for weekends.
+It can be seen that for every quarter weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently, whereas, Plot (c) helps to compare quarters within weekdays and weekends.
+For example, the quartile spread of consumption shrinks/lowers from first to fourth quarter for weekdays, whereas this pattern is not true for weekends.
 Plot (a) shows a clash since all quarters do not correspond to all months of the year.
@@
 
-Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
+Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
 

### 2020-04-15 main.Rmd
choosing harmonies

@@
 Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
-Ranking could be based on how well a harmony pair displays the maximum variation in distributions of the measured variable within and across each levels of the cyclic granularities.
-For each facet level, the most important display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data.
-Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
+Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
+And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying variation in distributions.
 Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+Moreover, distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
 
@@
 Suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by:
-
-$$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$
-where,
- $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$
-is the Kullback-Leibler divergence of $q_j$ with respect to $q_k$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
+ $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
 
 4.
-For each $\forall i = {1, 2, \dots, l}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
 Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
 
-5. The mean maximum pairwise distance (MMPD) over $l = {1, 2, \dots, l}$ is computed by 
-$MMPD = (1/l)\sum_{i=1}^{l}{D_{j}}$.
+5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by 
+$MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$.
 
 6.
-Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD since any harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
+Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
 

### 2020-04-15 main.Rmd
cricket and acknowledgement check

@@
 For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
-Thus, the cyclic granularity ball-of-the-over will be circular in most cases and aperiodic in others.
+Thus, the cyclic granularity ball-of-over will be circular in most cases and aperiodic in others.
 
@@
 
-However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a  non-temporal data set similar to any temporal data set.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to throw light on the periodic behavior of a  non-temporal data set similar to any temporal data set.
 There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
@@
 Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
-It would be interesting to take their example in order to dive deeper into this question.
+We take their example in order to dive deeper into this question.
 <!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
 This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
-To answer Q1, winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.
+Thus winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.
 
@@
 Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
---> Dismissals in previous over can lead to a batsman being either defensive or play more vulnerable.
-Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissals.
+--> Dismissals in previous over can lead to a batsman being either defensive or be more resilient.
+Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissal.
 
@@
 
-Exploratory data analysis is alliteratively finding and summarizing patterns.
+Exploratory data analysis is iteratively finding and summarizing patterns.
 With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context.
-This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and thereby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+This work provides a framework to systematically explore distribution of an univariate measured variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and thereby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
 
@@
 A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
-These functionalists can be imported in the gravitas package to accommodate for aperiodic cyclic granularities.
+These functions can be imported in the `gravitas` package to accommodate for aperiodic cyclic granularities.
 
@@
 
-The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Mon ash University for sharing their wisdom and experience of developing R packages, Dr.
+The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
 Peter Tosca [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.

### 2020-04-15 main.Rmd
cyclic calendar algebra multiple-to-single for quasi updated

@@
 
-The representation of single order-up quasi-circular granularities using multiple order-up quasi-circular granularities is not discussed in this paper and is left for future work.
-
+<!-- CONTENT CHANGE -->
+<!-- The representation of single order-up quasi-circular granularities using multiple order-up quasi-circular granularities is not discussed in this paper and is left for future work. -->
+
+This case is similar to \ref{eq:multifromsingle-quasi2}.
+Single-order-up quasi-circular granularity ($C_{G_{l},G_{m'}}$) could be obtained from multiple-order-up quasi-circular granularity ($C_{G_l,G_{m}}$) and single/multiple order-up circular granularity $C_{m',m}$ using \ref{eq:multifromsingle-quasi2}.
+
+\begin{equation} \label{eq:multifromsingle-quasi3}
+\begin{split}
+C_{G_{l},G_{m'}}(z) & = C_{G_l,G_{m}}(z) - \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
+\end{split}
+\end{equation}
+
+where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$, $\vert T_w \vert$ is the cardinality of set $T_w$.
 <!-- as they promote the exploratory analysis through visualization. -->
@@
 R package `gravitas`.
---> A missing piece is to enable user-defined temporal calendars and compute multiple order-up quasi-circular granularities from the hierarchy table.
+--> A missing piece is to enable user-defined temporal calendars.
 Also, computation of cyclic aperiodic granularities would require computing aperiodic linear granularities first.

### 2020-04-15 main.Rmd
smart meter application update

@@
 
-### Levels
+### Levels {#sec:levels}
 
@@
 We will discuss the effect of the mapping of cyclic granularities in this section.
-When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
+When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets, then $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$-->.
-With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
+Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$--> With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
 \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarter-of-year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
@@
 4.
-For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
 Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
@@
 Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
+Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
+If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units like hhour could be dropped.
+Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
+Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Section \ref{sec:cyclic-calendar}.
+<!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
+-->
+with the index set mapping half-hourly time instants to a set of positive integers.
+
+<!-- R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
+Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
+Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
 Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed.
@@
 -->
-Here, the index set $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ maps the half-hourly time instants to a set of positive integers.
-Then, multiple-order-up circular granularity $hour\_week$ and quasi-circular granularity $hour\_month$ could be computed from single-order-up granularities as follows:
-<!-- $$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular}) -->
-\begin{align}
-  hour\_week(z) =  day\_week(z) + 24*hour\_day(z)
-\end{align}
-
-<!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) -->
-\begin{align}
-hour\_month(z) =  hour\_day(z) + 24*day\_month(z)
-\end{align}
+<!-- Here, the index set $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ maps the half-hourly time instants to a set of positive integers.
+Then, multiple-order-up circular granularity $hour\_week$ and quasi-circular granularity $hour\_month$ could be computed from single-order-up granularities as follows: -->
+<!-- <!-- $$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular}) --> 
+<!-- \begin{align} -->
+<!--   hour\_week(z) =  day\_week(z) + 24*hour\_day(z) -->
+<!-- \end{align} -->
+
+<!-- <!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) --> 
+<!-- \begin{align} -->
+<!-- hour\_month(z) =  hour\_day(z) + 24*day\_month(z) -->
+<!-- \end{align} -->
 
@@
 From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-Given the data structure <$C1$ $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
-Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization.
+Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet.
+Thus the problem is equivalent to choosing 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization.
+However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting pairs together.
+\autoref{tab:harmony-tab} shows 13 harmonies pairs with any cyclic granularities with levels more than 31 removed as effective exploration becomes difficult with many levels (Section \ref{sec:levels}).
+For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day.
+The geometry used for displaying the distribution is chosen as area-quantiles \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
+\autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.
+
+<!-- From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
+Given the data structure <$C1$, $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
+Thus the problem is equivalent to taking 2 granularities at a time from six, which essentially is equivalent to having 30 data subsets for visualization.
 However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together.
 \autoref{tab:harmony-tab} shows 13 harmonies pairs, each of which can be plotted together to look at the energy behavior from different perspectives.
+-->
 

### 2020-04-16 main.Rmd
added install gravitas code

@@
 library(tsibble)
+# remotes::install_github("Sayani07/gravitas")
 library(gravitas)

### 2020-04-16 main.Rmd
cache = TRUE for the rank_harmony code
update paper after updating gravitas package

--- 2020-04-15 main.Rmd
@@
 library(tsibble)
+remotes::install_github("Sayani07/gravitas")
 library(gravitas)
@@
 
-```{r harmony-tab, echo=FALSE}
+```{r harmony-tab, echo=FALSE, cache=TRUE}
 harmonies <- smart_meter10 %>% 

### 2020-04-16 main.Rmd
data set of rank_harmony added in cache

@@
 ```{r harmony-tab, echo=FALSE, cache=TRUE}
+
+if (!file.exists("cache/smart_harmony.rds")) { 
   harmonies <- smart_meter10 %>% 
@@
          `average maximum variation` = mean_max_variation)
+
+  write_rds("cache/smart_harmony.rds") 
+}
+smart_harmony <- read_rds("cache/smart_harmony.rds")
 

### 2020-04-16 main.Rmd
cache ignored except for smart_harmony - file.exists and dir.exists

--- 2020-04-16 main.Rmd
@@
 ```{r harmony-tab, echo=FALSE, cache=TRUE}
+if(!dir.exists("cache/"))
+{
+  dir.create("cache")
+}
+if (!file.exists("cache/smart_harmony.rds")) { 
   harmonies <- smart_meter10 %>% 
@@
          `average maximum variation` = mean_max_variation)
+
+  write_rds(smart_harmony, "cache/smart_harmony.rds")
+}
+smart_harmony <- read_rds("cache/smart_harmony.rds")
 

### 2020-04-16 main.Rmd
review till cyclic calendar algebra

@@
 
-<!-- Finally, visualizing data across single/multiple order-up granularities help us to understand periodicities, pattern and anomalies in the data.
+<!-- Finally, visualizing data across single/multiple-order-up granularities help us to understand periodicities, pattern and anomalies in the data.
 Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
@@
 \autoref{fig:circular-dow} depicts the linear (a and b) and the corresponding stack (c and d) display of time.
-The circular granularity day-of-week and day-of-fortnight are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$).
+Circular granularities day-of-week $C_{B , G}$ and day-of-fortnight $C_{B, H}$ are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$).
 The granules that overlap share elements from the underlying time domain in the linear display.
-The circular granularity $C_{B , G}$ and $C_{B, H}$ <!--= C_{\text{day}, \text{week}}$--> representing day-of-week and day-of-fortnight respectively consist of repeated pattern of categories $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
+The circular granularities $C_{B , G}$ and $C_{B , H}$  <!--= C_{\text{day}, \text{week}}$ representing day-of-week and day-of-fortnight respectively --> consist of repeated pattern of categories $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
 The granules that overlap represents the same category in the stack display.
@@
 However, they are still formed with linear granularities, one of which "groups periodically into" the other.
-\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) with ($P_i$) denoting the plausible choices of number of granules of the finer granularity inside each granule of the coarser one.
+\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) with ($P_i$) denoting the plausible choices of period of the grouping of two linear granularities.
+<!-- number of granules of the finer granularity inside each granule of the coarser one. -->
 
@@
 \caption{Examples of quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
-These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping leading many potential period lengths and hence can not be expressed through modular arithmetic.
+These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping leading to many potential period lengths and hence can not be expressed through modular arithmetic.
 }
@@
 \begin{equation}\label{eq:quasi}
-Q_{B, G'}(z) = z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert,for\quad z \in T_k
+\begin{split}
+Q_{B, G'}(z) =
+z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, & \quad \text{for}\quad z \in T_{k} 
+\end{split}
 \end{equation}
@@
 
+<!-- \begin{split} -->
+<!-- A_{B, M}(z) & = -->
+<!-- \begin{cases} -->
+<!-- i, & \text{for}\quad z \in T_{i_j} \\ -->
+<!-- 0  & \text{otherwise}  -->
+<!-- \end{cases} -->
+<!-- \end{split} -->
+
 <!--${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$,-->
@@
 
-Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
+Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
 Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
@@
 
-Consider a school semester which lasts for 18.3 weeks, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks after which there is a study break of 1 week before final exams which continue for 1.3 weeks.
+Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 9 days.
 Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
-Both $M_1$ and $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$).
+Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$).
 Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity.
@@
 \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}.
-It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester remains constant over years.
+It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
 Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.
@@
 The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future.
-Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks after which there is a study break of 1 week before final exams which continue for 1.3 weeks.
-This pattern remains same for all semester and hence $Q_{H, M}$ with $P = 18.3 weeks$ denoting week of the semester week type will be a quasi-circular granularity with repeating patterns.
+Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks.
+This pattern remains same for all semester and hence $Q_{H, M}$ with $P = 18.3 weeks$ denoting semester week of the semester week type will be a quasi-circular granularity with repeating patterns.
 $A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years (shown through c)."}
@@
 $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
-$K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
+$K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
 We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
@@
 
-<!-- The multiple order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:all-circular}.
+<!-- The multiple-order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:all-circular}.
 Any cyclic granularity resulting from the last case are quasi-circular.
@@
 We consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
-Any multiple order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
-Two different approaches need to be employed for the following cases:
+Any multiple-order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
+Two different approaches need to be employed leading to the following cases:
 
@@
 \begin{split}
-C_{G_l,G_{m}}(z) & = C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)) \\
+C_{G_l,G_{m}}(z) & = C_{G_{l},G_{m'}}(z) + P(l, m')C_{G_{m'},G_{m}}(z) \\
 \end{split}
@@
 <!-- CONTENT CUT -->
-<!-- Clearly, the computation of multiple order-up granularities become more involved with quasi-circular granularities.
+<!-- Clearly, the computation of multiple-order-up granularities become more involved with quasi-circular granularities.
 $C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
@@
 -->
-<!-- If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}. -->
+<!-- If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple-order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}. -->
 
@@
 <!-- CONTENT CHANGE -->
-<!-- The representation of single order-up quasi-circular granularities using multiple order-up quasi-circular granularities is not discussed in this paper and is left for future work. -->
-
-This case is similar to \ref{eq:multifromsingle-quasi2}.
-Single-order-up quasi-circular granularity ($C_{G_{l},G_{m'}}$) could be obtained from multiple-order-up quasi-circular granularity ($C_{G_l,G_{m}}$) and single/multiple order-up circular granularity $C_{m',m}$ using \ref{eq:multifromsingle-quasi2}.
+<!-- The representation of single order-up quasi-circular granularities using multiple-order-up quasi-circular granularities is not discussed in this paper and is left for future work. -->
+
+This case is similar to Equation \ref{eq:multifromsingle-quasi2}.
+Single-order-up quasi-circular granularity ($C_{G_{l},G_{m'}}$) could be obtained from multiple-order-up quasi-circular granularity ($C_{G_l,G_{m}}$) and single/multiple-order-up circular granularity $C_{m',m}$ using \ref{eq:multifromsingle-quasi2}.
 

### 2020-04-17 main.Rmd
Fixed latex warning from m\choose 2

@@
 We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  ${m \choose 2}$ pairwise distances corresponding to each $A_i$.
+Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_i$.
 
@@
 4.
-For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
 Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.

### 2020-04-18 main.Rmd
renv lock updated with R 3.6.3 rticles and bookdown

@@
 options("knitr.graphics.auto_pdf" = TRUE)
+library(bookdown)
+library(rticles)
 library(knitr)

### 2020-04-21 main.Rmd
diagram changed

@@
 
-\autoref{fig:circular-dow} depicts the linear (a and b) and the corresponding stack (c and d) display of time.
+\autoref{fig:circular-dow} depicts the linear (a) and the corresponding stack (b) display of time.
 Circular granularities day-of-week $C_{B , G}$ and day-of-fortnight $C_{B, H}$ are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$).
@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap=" Circular granularities illustrated through linear (a and b) and stack (c and d) display of time.
-The linear display is similar to a measuring tape with the measurement marking distributed as linear granularities day, week and fortnight.
-The stack display gives the intuition of circular granularities by cutting the measuring tape and stacking on top of one another to represent repetitive pattern of categories.
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap=" Circular granularities illustrated through linear (a) and stack (b) display of time.
 The granules that overlap represent same time domain in the linear display and same category in the stack display.
 Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time."}
-knitr::include_graphics("Figs/circular-example-1.png")
+knitr::include_graphics("Figs/circularex.png")
 ```
@@
 Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks.
-This pattern remains same for all semester and hence $Q_{H, M}$ with $P = 18.3 weeks$ denoting semester week of the semester week type will be a quasi-circular granularity with repeating patterns.
-$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years (shown through c)."}
-
-knitr::include_graphics("Figs/quasi.png")
+This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 18.3 weeks will be a quasi-circular granularity with repeating patterns.
+$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years."}
+
+knitr::include_graphics("Figs/quasiex.png")
 

### 2020-04-23 main.Rmd
proof reading

@@
 Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$.
-Hence using \ref{def:quasicircular}, then we will have:
+Hence using Defintion \autoref{def:quasicircular} we will have:
 
@@
 <!-- **Discussion**:  -->
-If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
+If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
 Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
@@
 <!--Periodic and Quasi-periodic behavior can be defined by a repeating pattern except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.--> Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
-Most public holidays repeat every year, but there is no finite (or reasonably small) period within which their behavior remains constant.
+Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant.
 A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.<!--U.S labor day is the first Monday in September and U.S.
@@
 Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity.
-This is because the placement of the semester within an year would vary across years (\autoref{fig:aperiodic-example} c).
+This is because the placement of the semester within an year would vary across years.
 \autoref{fig:aperiodic-example} b shows the stack display of time with granules representing same categories (semester break/in-session) stacked on top of one another.
@@
 \begin{split}
-C_{G_l,G_m}(z) & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_m}(z)) \\
-  & = C_{(G_l,G_{l+1}}(z) + P(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)]( C_{G_{l+2}, G_m}(z)) \\
-  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+1)P(l+1, l+2)(C_{G_{l+2}, G_{m}})\\
-  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+2)(C_{G_{l+2},G_{l+m}}(z))\\
+C_{G_l,G_m}(z) & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_m}(z)\\
+  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)( C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)C_{G_{l+2}, G_m}(z)) \\
+  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_{l+2}}(z) + P(l, l+1)P(l+1, l+2)C_{G_{l+2}, G_{m}}\\
+  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_{l+2}}(z) + P(l, l+2)C_{G_{l+2},G_{l+m}}(z)\\
   &\vdots\\
-  & = \sum_{i=0}^{m - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))\\
+  & = \sum_{i=0}^{m - l - 1} P(l, l+i)C_{G_{l+i},G_{l+i+1}}(z)\\
 \end{split}
@@
 
-This case is similar to Equation \ref{eq:multifromsingle-quasi2}.
-Single-order-up quasi-circular granularity ($C_{G_{l},G_{m'}}$) could be obtained from multiple-order-up quasi-circular granularity ($C_{G_l,G_{m}}$) and single/multiple-order-up circular granularity $C_{m',m}$ using \ref{eq:multifromsingle-quasi2}.
-
-\begin{equation} \label{eq:multifromsingle-quasi3}
-\begin{split}
-C_{G_{l},G_{m'}}(z) & = C_{G_l,G_{m}}(z) - \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
-\end{split}
-\end{equation}
-
-where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$, $\vert T_w \vert$ is the cardinality of set $T_w$.
+Single-order-up quasi-circular granularity could be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}.
+
+<!-- \begin{equation} \label{eq:multifromsingle-quasi3} -->
+<!-- \begin{split} -->
+<!-- C_{G_{l},G_{m'}}(z) & = C_{G_l,G_{m}}(z) - \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\ -->
+<!-- \end{split} -->
+<!-- \end{equation} -->
+
+<!-- where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$, $\vert T_w \vert$ is the cardinality of set $T_w$. -->
 <!-- as they promote the exploratory analysis through visualization. -->

### 2020-04-23 main.Rmd
proof read changes

@@
 Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$.
-Hence using Defintion \autoref{def:quasicircular} we will have:
+Hence using Definition \autoref{def:quasicircular} we will have:
 
@@
 For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day.
-The geometry used for displaying the distribution is chosen as area-quantiles \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
+The geometry used for displaying the distribution is chosen as area-quantiles.
+\autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
 \autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.

### 2020-04-23 main.Rmd
up to 3.2

--- 2020-04-23 main.Rmd
@@
   \setlength {\marginparwidth }{2cm}
-  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes}
+  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
   \def\mod{~\text{mod}~}
@@
 
-\autoref{fig:linear-time} illustrates the linear time granularities.<!--, each of which are non-repeating in nature and arranged from past to future.
+\autoref{fig:linear-time} illustrates the linear time granularities.
+\todo[inline]{Expand  into  a paragraph - how does  it illustrate the linear granularities?
+How does  hour, day,...
+relate to G...} <!--, each of which are non-repeating in nature and arranged from past to future.
 If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.-->
@@
 -->
-
-For example, $day \trianglelefteq week$ and $day \preceq week$ since every granule of $week$ is the union of some set of granules of $day$ and each $day$ is a subset of a $week$.
+\newtheorem*{example}{Example}
+
+\begin{example}
+{\rm Let $day \trianglelefteq week$ and $day \preceq week$ since every granule of $week$ is the union of some set of granules of $day$ and each $day$ is a subset of a $week$.
 Consider another example where $day \trianglelefteq month$.
@@
 Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions.
+The period becomes 400 years with the inclusion of leap years and all their exceptions.}
+\end{example}
 
@@
 
-\noindent For example, in a Gregorian calendar without leap years we could say $day$ groups periodically into $month$ with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
-In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
+\begin{example}
+{\rm In a Gregorian calendar without leap years we could say $day$ groups periodically into $month$ with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
+In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.}
+\end{example}
 
@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$
+
+\begin{example} 
+{\rm minute = G$_{60}(\text{second})$, hour = G$_{60}(\text{minute})$, day = G$_{24}(\text{hour})$.}\todo[inline]{Can you use the same units shown in Fig 1?}
+\end{example}
 
@@
 
-We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] data structure.
+Cyclic granularities represeent cyclical components in time.
+Weeks is an example of a cyclical granularity because every 7 days there's another Monday.
+Cyclic granularities can be constructed by conducting modular arithmetic on two linear granularities.
+Cycles can be either _regular_, called **circular**, or _irregular_, **quasi-circular**.
+<!-- with every granule of coarser linear granularity being made up of same (regular) or different (irregular) number of granules of finer linear granularity.
+For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).
+A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.
+Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mappings are irregular.
+Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.-->
+
+<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
+
+Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] data structure.
 A tsibble consists of an index, key and measured variables.
@@
 For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
-A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.-->
-
-Cyclic granularities can be constructed by relating two linear granularities.
-The mappings between these linear granularities could be either _regular_ or _irregular_ with every granule of coarser linear granularity being made up of same (regular) or different (irregular) number of granules of finer linear granularity.
-<!--For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).
-A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.--> Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mappings are irregular.
-Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.
-
-<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
+A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.-->\todo[inline]{Would it be worth to start this discussion about  data structures earlier, with the linear granularity, and follow it through when appropriate; or put all discussion on  implementation in the data structures section.}
 
@@
 
-\autoref{fig:circular-dow} depicts the linear (a) and the corresponding stack (b) display of time.
-Circular granularities day-of-week $C_{B , G}$ and day-of-fortnight $C_{B, H}$ are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$).
-The granules that overlap share elements from the underlying time domain in the linear display.
-The circular granularities $C_{B , G}$ and $C_{B , H}$  <!--= C_{\text{day}, \text{week}}$ representing day-of-week and day-of-fortnight respectively --> consist of repeated pattern of categories $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
-The granules that overlap represents the same category in the stack display.
-Each circular granularity can use descriptive label mappings, or be simply labelled using the index.
+\autoref{fig:circular-dow} \todo[inline]{C(B,G) should be above C(B,H) in figure, and yuo could add the integers to the stacked sets.} illustrates the linear and the corresponding cyclical granularities.
+Cyclical granularities can  be considered to be cutting  the  linear  granularity into pieces, and stacking them to match the cycles (as shown in b).
+$B, G, H$ (day, week, fortnight, respectively) are linear granularities.
+The circular granularity $C_{B , G}$ (day-of-week)is constructed from $B$ and  $G$.
+The circular granularity $C_{B, H}$ (day-of-fortnight) is constructed from $B$ and $H$.
+These overlapping cyclical granularities share elements from the linear granularity.
+Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns  $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
+
+Each circular granularity can use descriptive label mappings.
 <!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap=" Circular granularities illustrated through linear (a) and stack (b) display of time.
-The granules that overlap represent same time domain in the linear display and same category in the stack display.
-Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time."}
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Illustration of circular relative to linear granularities (a).
+Circular granularities can be considered to be cutting the linear granularity into pieces and stacking them (b).
+The circular  granularity creates repeated integer sequences."}
 knitr::include_graphics("Figs/circularex.png")
@@
 
-### Muliple order-up circular granularities {#sec:all-circular-multiple}
+### Multiple order-up circular granularities {#sec:all-circular-multiple}
 
@@
 
+## Choosing harmonies to show
+
+Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
+Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
+And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying variation in distributions.
+Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+Moreover, distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
+
+1.
+Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
+Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
+
+2.
+There will be $m$ time series variable corresponding to each level $A_i$ of $C_1$.
+We compute the pairwise differences in distributions of these $m$ time series variables.
+Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_i$.
+
+3.
+Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+Suppose $q$ is the sample quantile vector computed for probability vector $p$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
+ $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
+
+4.
+For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
+
+5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by
+$MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$.
+
+6.
+Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
+Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+
+<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
+
 ## Summarizing the measured variable {#sec:summarize-measured}
@@
 <!-- Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels. -->
-
-## Choosing harmonies to show
-
-Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
-Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
-For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying variation in distributions.
-Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-Moreover, distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
-
-1.
-Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
-
-2.
-There will be $m$ time series variable corresponding to each level $A_i$ of $C_1$.
-We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_i$.
-
-3.
-Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
-Suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
- $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
-
-4.
-For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
-Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
-
-5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by
-$MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$.
-
-6.
-Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
-Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
-
-<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
 

### 2020-04-23 main.Rmd
proof read change

--- 2020-04-23 main.Rmd
@@
 For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day.
-The geometry used for displaying the distribution is chosen as area-quantiles.
-\autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
+The geometry used for displaying the distribution is chosen as area-quantiles in \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
 \autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.
@@
 The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
-The energy consumption is extremely skewed with 1st, 10th and 25th percentile lying very close whereas 75th, 90th and 99th lying further away from each other.This is common across both weekdays and weekends.
+The energy consumption is extremely skewed with 1st, 10th and 25th percentile lying very close whereas 75th, 90th and 99th lying further away from each other.
+This is common across both weekdays and weekends.
 For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer as compared to weekends.

### 2020-04-23 main.Rmd
circular diagram updated with added integers in stack display

--- 2020-04-23 main.Rmd
@@
 For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day.
-The geometry used for displaying the distribution is chosen as area-quantiles \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
+The geometry used for displaying the distribution is chosen as area-quantiles in \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
 \autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.

### 2020-04-23 main.Rmd
linear diagram update

@@
 These are ordered with ordering guided by integers and hence is unidirectional and non-repeating."}
-knitr::include_graphics("Figs/linear-arr1.png")
+knitr::include_graphics("Figs/linear-arr2.png")
 ```

### 2020-04-23 main.Rmd
\todo[inline]{Can you use the same units shown in Fig 1?}
{Expand  into  a paragraph - how does  it illustrate the linear granularities? How does  hour, day,... relate to G...}

@@
 When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.-->
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
 
 \autoref{fig:linear-time} illustrates the linear time granularities.
+Here, "hour" is a linear granularity which is a mapping from index set to hourly time domain.
+Similarly, "day", "week", "month" and "year" are also linear granularities which maps from index set to subsets of the hourly time domain.
+For example, day($0$) =  {hour($0$), hour($1$), hour($2$), \dots, hour($23$)}, day($1$) =  {hour($24$), hour($25$), hour($26$), \dots, hour($47$)} and so on.
+Thus, if we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
+These linear granularities are ordered with ordering guided by the index set which is a set of integers.
+Hence, they are uni-directional and non-repeating.
+
 \todo[inline]{Expand  into  a paragraph - how does  it illustrate the linear granularities?
@@
 
-```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="The time domain distributed as linear granularities hours, days, weeks and months.
+```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="The time domain distributed as linear granularities hour, day, week, month and year.
 These are ordered with ordering guided by integers and hence is unidirectional and non-repeating."}
-knitr::include_graphics("Figs/linear-arr2.png")
+knitr::include_graphics("Figs/linear-arr3.png")
 ```
@@
 \begin{example} 
-{\rm minute = G$_{60}(\text{second})$, hour = G$_{60}(\text{minute})$, day = G$_{24}(\text{hour})$.}\todo[inline]{Can you use the same units shown in Fig 1?}
+{\rm day = G$_{24}(\text{hour})$, week = G$_{7}(\text{day})$.}
 \end{example}

### 2020-04-24 main.Rmd
english and 'more objective'-type edits from Nick.

@@
 Methods are provided for creating all possible granularities for a time index.
-A recommendation algorithm provides an indication whether a pair of granularities could be interesting to examine together, called a harmony or clash.
-
- The time granularities can used to create visualizations of the data to explore for periodicities, associations and anomalies.
-The granularities can be considered to be categorical variables (ordered and unordered) which induces a grouping of the observations.
+A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together, called a harmony or when they cannot, called a clash.
+
+ The time granularities can be used to create visualizations of the data to explore for periodicities, associations and anomalies.
+The granularities can be considered to be categorical variables (ordered or unordered) which induces a grouping of the observations.
 Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
@@
 
- The methods are implemented in the open source R package gravitas.
-The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
+ The methods are implemented in the open source R package `gravitas`.
+The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
 bibliography: [bibliography.bib]
@@
 This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
-Analysts are expected to alliteratively explore many possible choices of time granularities for comprehending possible periodicities in the data.
-But too many choices and a lack of a systematic approach to do so might become overwhelming.
+Visualizing probability distributions conditional on one or more granularities is an indispensible tool for exploration.
+Analysts are expected to comprehesively explore the many ways to view and consider temporal data.
+However, the plethora of choices and the lack of a systematic approach to do so quickly becomes overwhelming.
 
@@
 Cyclic time granularities can be circular, quasi-circular or aperiodic.
-Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+Examples of circular granularities are hour of the day and day of the week; an example of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
 
@@
 The lubridate package [@Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
-But most of its accessor functions are limited to single-order-up granularities.
+However, most of its accessor functions are limited to single-order-up granularities.
 
@@
 \begin{definition}\label{def:index set}
-There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of positive integers. Often, this is thought of as $t=1, \dots, T$.
+There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of positive integers. Often, this is thought of as $t=0, \dots, T$.
 \end{definition}
@@
 When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
-
-\autoref{fig:linear-time} illustrates the linear time granularities.
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed as the **bottom granularity**.
+
+\autoref{fig:linear-time} illustrates a number of common linear time granularities.
 Here, "hour" is a linear granularity which is a mapping from index set to hourly time domain.
-Similarly, "day", "week", "month" and "year" are also linear granularities which maps from index set to subsets of the hourly time domain.
+Similarly, "day", "week", "month" and "year" are also linear granularities which maps from index set to subsets of theeir respective time domains.
 For example, day($0$) =  {hour($0$), hour($1$), hour($2$), \dots, hour($23$)}, day($1$) =  {hour($24$), hour($25$), hour($26$), \dots, hour($47$)} and so on.
@@
 Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions.}
+This grouping period becomes 4 and 400 years with the inclusion of leap years and leap centries respectively.}
 \end{example}
@@
 Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mappings are irregular.
-Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.-->
+Examples of common circular granularities include hour of the day and day of the week, whereas examples for quasi-circular granularities are day of the month and week of the month.-->
 
@@
 
-Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] data structure.
+Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
 A tsibble consists of an index, key and measured variables.
@@
 Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
-Circular granularities could be expressed using modular arithmetic due to their regular mapping.
+Circular granularities can be expressed using modular arithmetic due to their regular mapping.
 }
@@
 --> In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
-Examples of aperiodic granularity may also include school holidays or a scheduling event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns.
+Examples of aperiodic granularity may also include school holidays or a scheduled event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns.
 As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
@@
 
-Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 9 days.
+Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
 Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
@@
 So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
-But it is far from being the only one.
+However, it is far from being the only one.
 <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
+The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
 Nevertheless, for any calendar a hierarchy table can be defined.
-For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
+For example, in the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
 \autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.
@@
 Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
-The label mapping of the resultant multiple-order-up granularity can however be chosen depending on the context.
+The label mapping of the resultant multiple-order-up granularity can be chosen depending on the context.
 
@@
                       `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
-                     `period length/conversion operator (K)` = c(20, 18,"k(day, month)", 20, 1))
+                     `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1))
 
@@
       booktabs = TRUE,
-      caption = "Hierarchy table for Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
+      caption = "Hierarchy table for the Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
   #row_spec(0, bold = TRUE) %>%
@@
 
-Effective exploration or good visualization require good data structures.
+Effective exploration and visualization require proper data structures.
 It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
@@
 Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
-Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
+Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
 Instead, this big problem is broken down by focusing on two cyclic granularities at a time.
@@
 
-## Harmonies and Clashes {#sec:synergy}
+## Harmonies and clashes {#sec:synergy}
 
@@
 
-## Choosing harmonies to show
-
-Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
+## Choosing harmonies to show {#sec:chooseharmonies}
+
+Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the number of combinations quickly becomes too large and thus, it is beneficial useful to rank these harmonies.
 Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying variation in distributions.
+The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
 Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-Moreover, distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
+Distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
 
@@
 It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
-But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities could be a potential way to compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
+However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
+Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities is a valid way to compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
 Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
@@
 - Aesthetic mapping (mapping of variables to elements of the plot): $C_1$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation (data summarization): Any descriptive or smoothing statistics that summarizes distribution of $v$
-- Geometric objects (physical representation of the data): Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
+- Statistical transformation (data summarization): any descriptive or smoothing statistics that summarizes distribution of $v$
+- Geometric objects (physical representation of the data): any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
 - Facet (split plots):  $C_2$
@@
 We will discuss few conventional and recent ways to plot distributions using both Kernel density estimates and descriptive statistics.
-Descriptive statistics based displays include box plots [@Tukey1977-jx] or different variations of it like Notched box plots [@Mcgill1978-hg].
-More recent ways are the letter-value box plot [@Hofmann2017-sg] or Quantile plots which display quantiles instead of quartiles in a traditional boxplot.
+Descriptive statistics based displays include box plots [@Tukey1977-jx] or different variations of it like notched box plots [@Mcgill1978-hg].
+More recent ways are the letter-value box plot [@Hofmann2017-sg] or quantile plots which display quantiles instead of quartiles in a traditional boxplot.
 Kernel density based plots for displaying a distribution include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, highest density regions (HDR) box [@Hyndman1996-ft].
@@
 
-If there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
+If there are levels of cyclic granularities plotted across the  x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
 We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
@@
 Conversely, weekend and weekday can occur in every quarter.
+Analysts should avoid plotting such clashes.
 Plot (b) helps to compare weekends and weekdays for each quarter.
@@
 
-R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
+The R package `gravitas` [@R-gravitas] is used to facilitate the systematic exploration here.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to cosider is which cyclic time granularities we can look at exhaustively.
 Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
 Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
-Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
-If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units like hhour could be dropped.
-Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
+Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ (via \ref{sec:chooseharmonies} item #4) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
+If these options are considered too numerous, the smallest and largest granularities are good candidates to disconsider.
+Hense, we removed half hour and year.
+Then we will be left $(4*3/2) = 6$ cyclic granularites, namely: "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
 Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Section \ref{sec:cyclic-calendar}.
@@
 -->
-with the index set mapping half-hourly time instants to a set of positive integers.
+with the index set mapping half-hourly time periods to a set of positive integers.
 
@@
 Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet.
-Thus the problem is equivalent to choosing 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization.
-However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting pairs together.
-\autoref{tab:harmony-tab} shows 13 harmonies pairs with any cyclic granularities with levels more than 31 removed as effective exploration becomes difficult with many levels (Section \ref{sec:levels}).
-For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day.
-The geometry used for displaying the distribution is chosen as area-quantiles in \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
-\autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.
+Choosing 2 of the possible 6 granularities, which is equivalent to having ${2 \choose 6} = 30$ canidates for visualization.
+Fortunately, harmony granularities can be identified among those 30 possibilities to narrow the search.
+\autoref{tab:harmony-tab} shows 13 harmonies pairs, after removing clashes and any cyclic granularities with levels more than 31, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
+For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/weekend) and $C2$ is hour of the day.
+The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust} (b and c respectively).
+\autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type with distribution geometrically displayed as boxplots.
 
@@
 In \autoref{fig:bothcust} (b), the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+The black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
 The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
-The energy consumption is extremely skewed with 1st, 10th and 25th percentile lying very close whereas 75th, 90th and 99th lying further away from each other.
+The energy consumption is extremely (positive- or right-) skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
 This is common across both weekdays and weekends.
@@
 There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
-Also hours from 7 to 13 look most volatile.
+Also the hours from 7 to 13 look most volatile.
 If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
@@
 Also, outliers are more prominent in the latter part of the day.
-All of these indicates that looking at different distribution plots or changing the mapping might shed lights on different aspect of the energy behavior for the same customer.
+All of these indicate that looking at different distribution geometry or changing the mapping might shed lights on different aspect of the energy behavior for the same sample population.
 
@@
 
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
-This is also intuitive because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one observation shows a particular behavior, but this behavior paradoxically decomes obsured by aggregation.
+For example in a particular neighborhood one household may have the least daily power consumption for a full week, yet still not be the household with the minimum weekly power consumption.
+This is an intuitive possibility, because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
 A strategy for analyzing multiple keys together could be to first group them basis time series or demographic features and then look at their energy behavior.
@@
 
-The method is not only restricted to temporal data.
-We provide an example of cricket to illustrate how this can be generalized to other applications.
+The method is not only restricted to temporal data, and can be generalised to many hierarchical granularies (with  continuous and uni-directional nature).
+We illustrate this with an application to the sport cricket.
 Although there is no conventional time component in cricket, each ball can be thought to represent an ordering from past to future with the game progressing forward with each ball.
-In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
-Thus similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
-An idea of cyclic granularities can be conceived with this hierarchy in mind.
-Examples may include ball of the over, over of the innings or ball of the innings.
-Although most of these cyclic granularities are circular in theory by design of the hierarchy, in reality these may be aperiodic.
+In the Twenty20 format, an over will consist of 6 balls (with some exceptions), an inning is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
+Thus, similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
+The idea of cyclic granularities can be likewise mapped to this hierarchy.
+Example granularites then include ball of the over, over of the inning and ball of the inning.
+Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularies will are aperiodic.
 For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
@@
              booktabs = TRUE,
-             caption = "Hierarchy table for cricket with each over is nested within innings, innings nested within match and match within season.") %>%
+             caption = "Hierarchy table for cricket where overs are nested within an inning, innings nested within a match and matches within a season.") %>%
   #row_spec(0, bold = TRUE)%>%
@@
 
-However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to throw light on the periodic behavior of a  non-temporal data set similar to any temporal data set.
-There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to shed light on the periodic behavior of a non-temporal data set similar to any temporal data set.
+There are many interesting questions that could possibly be answered with such a data set irrespective of the type of cyclic granularities.
 <!--We will explore a few interesting questions and understand how the proposed approach in the paper can help answer some of the questions.-->
@@
 
-First, it would be interesting to see if the distribution of total runs vary depending on if a team bats first or second?
-Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
+First, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second inning.
+The Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
 We take their example in order to dive deeper into this question.
-<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
-This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
-Thus winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.
+<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
+This seem indicate that players feel mounting presure to score more runs as they approach towards the end of the first inning.
+Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting press and may adopt a more conservative strategy, to score runs.
+Thus winning teams like CSK and MI seem to emply different inning strategies when it comes to their batting order.
 
@@
 
-Another interesting question could be if runs per over set to reduce in subsequent over if fielding is good in the previous over?
-For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
+Another interesting question could is: do runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
+For establishing that the fielding quality, we apply an indicator function on dismissals (1 if there was at least one catch in the previous over, 0 otherwise).
+Runs in the current over is then the observation variable.
 <!--If a batsman is bowled out, it does not necessarily signify good fielding.
@@
 Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
---> Dismissals in previous over can lead to a batsman being either defensive or be more resilient.
-Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissal.
+--> Dismissals in the previous over can lead to a batsman adopting a more defensive play style.
+Figure \ref{fig:cricex}(b) shows that when there are no dismissals in the previous over, leads to a higher median and quartile spread of run rate per over as compared to the case when there has been at least one dismissal in the previous over.
 
@@
 Plot (a) shows letter value plot across overs faceted by innings.
-There is no upward shift in runs in the second innings like that in the first implying teams are vulnerable to score more in the first innings as they approach the end of the inning.
-Plot (b) shows box plot across number of wickets in previous over faceted by current over.
-This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning=FALSE, message = FALSE, out.width = "90%"}
+The team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
+Plot (b) shows box plot of runs per over across an indicator of wickets in previous over faceted by current over.
+This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning = FALSE, message = FALSE, out.width = "90%"}
 
@@
 
-Exploratory data analysis is iteratively finding and summarizing patterns.
-With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context.
-This work provides a framework to systematically explore distribution of an univariate measured variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and thereby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+Exploratory data analysis and data analysis in general involve many iterations of finding and summarizing patterns.
+With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible  granularities to explore.
+This work provides a framework to systematically explore distribution of an univariate measured variable across two cyclic time granularities by creating any cyclic granularity, ranking a list of harmonies and thereby identifying possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
 
@@
 
-The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
+The authors would like to thank the cohort [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages and Dr.
 Peter Tosca [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
-The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-We would also like to thank Nicholas Spyrison for many useful discussions and feedback on the manuscript.
-More details about the package
-can be found on the package website https://sayani07.github.io/gravitas/.
-This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
-The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article and the code is also available online
-in the supplemental materials.
+The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
+We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
+More details about the package can be found on the package website [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
+This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+This paper's Github repository, [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
 
@@
 
-**Data and scripts:** Data sets and R codes to reproduce all figures in this article (main.Rmd).
-
-**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN.
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.Rmd).
+
+**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
 The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, check feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommend prospective probability distributions through factors described in the article.
-Version 0.1.2 of the package was used for the results presented in the article and is available from Github (https://github.com/Sayani07/gravitas).
-
-**R-packages:** Each of the R packages used in this article are available online (URLs are provided in the bibliography).
+Version 0.1.2 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).
+
+**R-packages:** Each of the R packages used in this article are available online with URLs provided in the bibliography.
 

### 2020-04-25 main.Rmd
small edits, and subjective wording/comments etc.

@@
 library(tsibble)
+if (requireNamespace("gravitas"))
   remotes::install_github("Sayani07/gravitas")
@@
 )
-knitr::opts_knit$set(root.dir = here::here())
+#knitr::opts_knit$set(root.dir = here::here())
 ```
@@
       align = "l",
-      caption = "Hierarchy table for Mayan calendar with all circular single-order-up granularities.") %>%
+      caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities.") %>%
   kable_styling()
@@
 \begin{definition}\label{def:clash}
-A {\bf harmony} is a pair of cyclic granularities that do not contain any missing combinations of its categories.
+A {\bf harmony} is a pair of cyclic granularities that does not contain any missing combinations of its categories.
 \end{definition}

### 2020-04-25 main.Rmd
small changes to install/require gravita, st it knits for me.

@@
 library(tsibble)
-if (requireNamespace("gravitas"))
+if (requireNamespace("gravitas") == F)
   remotes::install_github("Sayani07/gravitas")

### 2020-04-26 main.Rmd
linear granularoty caption change

--- 2020-04-24 main.Rmd
@@
 Methods are provided for creating all possible granularities for a time index.
-A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together, called a harmony or when they cannot, called a clash.
-
- The time granularities can be used to create visualizations of the data to explore for periodicities, associations and anomalies.
-The granularities can be considered to be categorical variables (ordered or unordered) which induces a grouping of the observations.
+A recommendation algorithm provides an indication whether a pair of granularities could be interesting to examine together, called a harmony or clash.
+
+ The time granularities can used to create visualizations of the data to explore for periodicities, associations and anomalies.
+The granularities can be considered to be categorical variables (ordered and unordered) which induces a grouping of the observations.
 Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
@@
 
- The methods are implemented in the open source R package `gravitas`.
-The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
+ The methods are implemented in the open source R package gravitas.
+The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009pk].
 bibliography: [bibliography.bib]
@@
 This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Visualizing probability distributions conditional on one or more granularities is an indispensible tool for exploration.
-Analysts are expected to comprehesively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly becomes overwhelming.
+Visualizing probability distributions conditional on one or more granularities is a potentially useful approach for exploration.
+Analysts are expected to alliteratively explore many possible choices of time granularities for comprehending possible periodicities in the data.
+But too many choices and a lack of a systematic approach to do so might become overwhelming.
 
@@
 Cyclic time granularities can be circular, quasi-circular or aperiodic.
-Examples of circular granularities are hour of the day and day of the week; an example of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+Examples of circular granularities are hour of the day and day of the week; an examples of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
 
@@
 The lubridate package [@Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
-However, most of its accessor functions are limited to single-order-up granularities.
+But most of its accessor functions are limited to single-order-up granularities.
 
@@
 \begin{definition}\label{def:index set}
-There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of positive integers. Often, this is thought of as $t=0, \dots, T$.
+There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of positive integers. Often, this is thought of as $t=1, \dots, T$.
 \end{definition}
@@
 When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed as the **bottom granularity**.
-
-\autoref{fig:linear-time} illustrates a number of common linear time granularities.
+Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> \autoref{fig:linear-time} illustrates the linear time granularities.
 Here, "hour" is a linear granularity which is a mapping from index set to hourly time domain.
-Similarly, "day", "week", "month" and "year" are also linear granularities which maps from index set to subsets of theeir respective time domains.
-For example, day($0$) =  {hour($0$), hour($1$), hour($2$), \dots, hour($23$)}, day($1$) =  {hour($24$), hour($25$), hour($26$), \dots, hour($47$)} and so on.
-Thus, if we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
+Similarly, "day", "week", "month" and "year" are also linear granularities which maps from index set to subsets of the hourly time domain.
+<!--For example, day($0$) =  {hour($0$), hour($1$), hour($2$), \dots, hour($23$)}, day($1$) =  {hour($24$), hour($25$), hour($26$), \dots, hour($47$)} and so on.--> If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
 These linear granularities are ordered with ordering guided by the index set which is a set of integers.
@@
 Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
-This grouping period becomes 4 and 400 years with the inclusion of leap years and leap centries respectively.}
+The period becomes 400 years with the inclusion of leap years and all their exceptions.}
 \end{example}
@@
 Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mappings are irregular.
-Examples of common circular granularities include hour of the day and day of the week, whereas examples for quasi-circular granularities are day of the month and week of the month.-->
+Examples of common circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or week of the month.-->
 
@@
 
-Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
+Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] data structure.
 A tsibble consists of an index, key and measured variables.
@@
 Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
-Circular granularities can be expressed using modular arithmetic due to their regular mapping.
+Circular granularities could be expressed using modular arithmetic due to their regular mapping.
 }
@@
 --> In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
-Examples of aperiodic granularity may also include school holidays or a scheduled event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns.
+Examples of aperiodic granularity may also include school holidays or a scheduling event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns.
 As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
@@
 
-Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
+Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 9 days.
 Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
@@
 So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
-However, it is far from being the only one.
+But it is far from being the only one.
 <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
+The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds".
 Nevertheless, for any calendar a hierarchy table can be defined.
-For example, in the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
+For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
 \autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.
@@
 Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
-The label mapping of the resultant multiple-order-up granularity can be chosen depending on the context.
+The label mapping of the resultant multiple-order-up granularity can however be chosen depending on the context.
 
@@
                       `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
-                     `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1))
+                     `period length/conversion operator (K)` = c(20, 18,"k(day, month)", 20, 1))
 
@@
       booktabs = TRUE,
-      caption = "Hierarchy table for the Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
+      caption = "Hierarchy table for Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
   #row_spec(0, bold = TRUE) %>%
@@
 
-Effective exploration and visualization require proper data structures.
+Effective exploration or good visualization require good data structures.
 It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
@@
 Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
-Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
+Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
 Instead, this big problem is broken down by focusing on two cyclic granularities at a time.
@@
 
-## Harmonies and clashes {#sec:synergy}
+## Harmonies and Clashes {#sec:synergy}
 
@@
 
-## Choosing harmonies to show {#sec:chooseharmonies}
-
-Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the number of combinations quickly becomes too large and thus, it is beneficial useful to rank these harmonies.
+## Choosing harmonies to show
+
+Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies.
 Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
+And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying variation in distributions.
 Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-Distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
+Moreover, distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
 
@@
 It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
-However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities is a valid way to compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
+But that approach  hides the distributions of the measured variable induced by the re-organised data structure.
+Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities could be a potential way to compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
 Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
@@
 - Aesthetic mapping (mapping of variables to elements of the plot): $C_1$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation (data summarization): any descriptive or smoothing statistics that summarizes distribution of $v$
-- Geometric objects (physical representation of the data): any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
+- Statistical transformation (data summarization): Any descriptive or smoothing statistics that summarizes distribution of $v$
+- Geometric objects (physical representation of the data): Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
 - Facet (split plots):  $C_2$
@@
 We will discuss few conventional and recent ways to plot distributions using both Kernel density estimates and descriptive statistics.
-Descriptive statistics based displays include box plots [@Tukey1977-jx] or different variations of it like notched box plots [@Mcgill1978-hg].
-More recent ways are the letter-value box plot [@Hofmann2017-sg] or quantile plots which display quantiles instead of quartiles in a traditional boxplot.
+Descriptive statistics based displays include box plots [@Tukey1977-jx] or different variations of it like Notched box plots [@Mcgill1978-hg].
+More recent ways are the letter-value box plot [@Hofmann2017-sg] or Quantile plots which display quantiles instead of quartiles in a traditional boxplot.
 Kernel density based plots for displaying a distribution include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, highest density regions (HDR) box [@Hyndman1996-ft].
@@
 
-If there are levels of cyclic granularities plotted across the  x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
+If there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
 We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
@@
 Conversely, weekend and weekday can occur in every quarter.
-Analysts should avoid plotting such clashes.
 Plot (b) helps to compare weekends and weekdays for each quarter.
@@
 
-The R package `gravitas` [@R-gravitas] is used to facilitate the systematic exploration here.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to cosider is which cyclic time granularities we can look at exhaustively.
+R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
 Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
 Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
-Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ (via \ref{sec:chooseharmonies} item #4) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
-If these options are considered too numerous, the smallest and largest granularities are good candidates to disconsider.
-Hense, we removed half hour and year.
-Then we will be left $(4*3/2) = 6$ cyclic granularites, namely: "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
+Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
+If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units like hhour could be dropped.
+Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
 Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Section \ref{sec:cyclic-calendar}.
@@
 -->
-with the index set mapping half-hourly time periods to a set of positive integers.
+with the index set mapping half-hourly time instants to a set of positive integers.
 
@@
 Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet.
-Choosing 2 of the possible 6 granularities, which is equivalent to having ${2 \choose 6} = 30$ canidates for visualization.
-Fortunately, harmony granularities can be identified among those 30 possibilities to narrow the search.
-\autoref{tab:harmony-tab} shows 13 harmonies pairs, after removing clashes and any cyclic granularities with levels more than 31, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
-For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/weekend) and $C2$ is hour of the day.
-The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust} (b and c respectively).
-\autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type with distribution geometrically displayed as boxplots.
+Thus the problem is equivalent to choosing 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization.
+However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting pairs together.
+\autoref{tab:harmony-tab} shows 13 harmonies pairs with any cyclic granularities with levels more than 31 removed as effective exploration becomes difficult with many levels (Section \ref{sec:levels}).
+For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day.
+The geometry used for displaying the distribution is chosen as area-quantiles in \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b).
+\autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.
 
@@
 In \autoref{fig:bothcust} (b), the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot.
-The black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
 The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
-The energy consumption is extremely (positive- or right-) skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
+The energy consumption is extremely skewed with 1st, 10th and 25th percentile lying very close whereas 75th, 90th and 99th lying further away from each other.
 This is common across both weekdays and weekends.
@@
 There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
-Also the hours from 7 to 13 look most volatile.
+Also hours from 7 to 13 look most volatile.
 If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
@@
 Also, outliers are more prominent in the latter part of the day.
-All of these indicate that looking at different distribution geometry or changing the mapping might shed lights on different aspect of the energy behavior for the same sample population.
+All of these indicates that looking at different distribution plots or changing the mapping might shed lights on different aspect of the energy behavior for the same customer.
 
@@
 
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one observation shows a particular behavior, but this behavior paradoxically decomes obsured by aggregation.
-For example in a particular neighborhood one household may have the least daily power consumption for a full week, yet still not be the household with the minimum weekly power consumption.
-This is an intuitive possibility, because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one `customer_id` shows a particular behavior, but this behavior is reversed when all keys are combined together.
+This is also intuitive because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
 A strategy for analyzing multiple keys together could be to first group them basis time series or demographic features and then look at their energy behavior.
@@
 
-The method is not only restricted to temporal data, and can be generalised to many hierarchical granularies (with  continuous and uni-directional nature).
-We illustrate this with an application to the sport cricket.
+The method is not only restricted to temporal data.
+We provide an example of cricket to illustrate how this can be generalized to other applications.
 Although there is no conventional time component in cricket, each ball can be thought to represent an ordering from past to future with the game progressing forward with each ball.
-In the Twenty20 format, an over will consist of 6 balls (with some exceptions), an inning is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
-Thus, similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
-The idea of cyclic granularities can be likewise mapped to this hierarchy.
-Example granularites then include ball of the over, over of the inning and ball of the inning.
-Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularies will are aperiodic.
+In a Twenty20 game, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
+Thus similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
+An idea of cyclic granularities can be conceived with this hierarchy in mind.
+Examples may include ball of the over, over of the innings or ball of the innings.
+Although most of these cyclic granularities are circular in theory by design of the hierarchy, in reality these may be aperiodic.
 For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
@@
              booktabs = TRUE,
-             caption = "Hierarchy table for cricket where overs are nested within an inning, innings nested within a match and matches within a season.") %>%
+             caption = "Hierarchy table for cricket with each over is nested within innings, innings nested within match and match within season.") %>%
   #row_spec(0, bold = TRUE)%>%
@@
 
-However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to shed light on the periodic behavior of a non-temporal data set similar to any temporal data set.
-There are many interesting questions that could possibly be answered with such a data set irrespective of the type of cyclic granularities.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to throw light on the periodic behavior of a  non-temporal data set similar to any temporal data set.
+There are many interesting questions that can possibly be answered with such a data set irrespective of the type of cyclic granularities.
 <!--We will explore a few interesting questions and understand how the proposed approach in the paper can help answer some of the questions.-->
@@
 
-First, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second inning.
-The Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
+First, it would be interesting to see if the distribution of total runs vary depending on if a team bats first or second?
+Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
 We take their example in order to dive deeper into this question.
-<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
-This seem indicate that players feel mounting presure to score more runs as they approach towards the end of the first inning.
-Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting press and may adopt a more conservative strategy, to score runs.
-Thus winning teams like CSK and MI seem to emply different inning strategies when it comes to their batting order.
+<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that there is no clear upward shift in median and quartile deviation of runs in the second innings as compared to the first innings.
+This might indicate that players are vulnerable to score through each over and make more runs as they approach towards the end of the first innings, whereas if they bat in the second innings they have a target in mind and are not as vulnerable.
+Thus winning teams like CSK and MI have indeed different strategies when it comes to batting in the first or second innings.
 
@@
 
-Another interesting question could is: do runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
-For establishing that the fielding quality, we apply an indicator function on dismissals (1 if there was at least one catch in the previous over, 0 otherwise).
-Runs in the current over is then the observation variable.
+Another interesting question could be if runs per over set to reduce in subsequent over if fielding is good in the previous over?
+For establishing that the fielder fielded well in a particular over, we can see how many catches and run outs were made in that particular over.
 <!--If a batsman is bowled out, it does not necessarily signify good fielding.
@@
 Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
---> Dismissals in the previous over can lead to a batsman adopting a more defensive play style.
-Figure \ref{fig:cricex}(b) shows that when there are no dismissals in the previous over, leads to a higher median and quartile spread of run rate per over as compared to the case when there has been at least one dismissal in the previous over.
+--> Dismissals in previous over can lead to a batsman being either defensive or be more resilient.
+Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over will lead to a higher median and quartile spread of run rates compared to the case when there has been at least one dismissal.
 
@@
 Plot (a) shows letter value plot across overs faceted by innings.
-The team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
-Plot (b) shows box plot of runs per over across an indicator of wickets in previous over faceted by current over.
-This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning = FALSE, message = FALSE, out.width = "90%"}
+There is no upward shift in runs in the second innings like that in the first implying teams are vulnerable to score more in the first innings as they approach the end of the inning.
+Plot (b) shows box plot across number of wickets in previous over faceted by current over.
+This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning=FALSE, message = FALSE, out.width = "90%"}
 
@@
 
-Exploratory data analysis and data analysis in general involve many iterations of finding and summarizing patterns.
-With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible  granularities to explore.
-This work provides a framework to systematically explore distribution of an univariate measured variable across two cyclic time granularities by creating any cyclic granularity, ranking a list of harmonies and thereby identifying possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+Exploratory data analysis is iteratively finding and summarizing patterns.
+With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context.
+This work provides a framework to systematically explore distribution of an univariate measured variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and thereby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
 
@@
 
-The authors would like to thank the cohort [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages and Dr.
+The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr.
 Peter Tosca [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
-The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
-More details about the package can be found on the package website [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
-This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
-This paper's Github repository, [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
+We would also like to thank Nicholas Spyrison for many useful discussions and feedback on the manuscript.
+More details about the package
+can be found on the package website https://sayani07.github.io/gravitas/.
+This article was created with knitr [@knitr, @R-knitr] and R Markdown [@rmarkdown, @R-rmarkdown].
+The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article and the code is also available online
+in the supplemental materials.
 
@@
 
-**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.Rmd).
-
-**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
+**Data and scripts:** Data sets and R codes to reproduce all figures in this article (main.Rmd).
+
+**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package gravitas [@R-gravitas], available from CRAN.
 The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, check feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommend prospective probability distributions through factors described in the article.
-Version 0.1.2 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).
-
-**R-packages:** Each of the R packages used in this article are available online with URLs provided in the bibliography.
+Version 0.1.2 of the package was used for the results presented in the article and is available from Github (https://github.com/Sayani07/gravitas).
+
+**R-packages:** Each of the R packages used in this article are available online (URLs are provided in the bibliography).
 

### 2020-04-27 main.Rmd
put all discussion on implementation in the data structures section.

--- 2020-04-25 main.Rmd
@@
 These are ordered with ordering guided by integers and hence is unidirectional and non-repeating."}
-knitr::include_graphics("Figs/linear-arr3.png")
+knitr::include_graphics("Figs/linear-ex.png")
 ```
@@
 
-Cyclic granularities represeent cyclical components in time.
+\todo[inline]{Check with Di to revise the first two statements}
+Cyclic granularities represent cyclical repetitions in time.
 Weeks is an example of a cyclical granularity because every 7 days there's another Monday.
@@
 <!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
-
-Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
-A tsibble consists of an index, key and measured variables.
-An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-<!--In a tsibble, each observation (row) is uniquely identified by index and key.--> A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the index of a tsibble.
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
-A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.-->\todo[inline]{Would it be worth to start this discussion about data structures earlier, with the linear granularity, and follow it through when appropriate; or put all discussion on implementation in the data structures section.}
 
@@
 The circular granularity creates repeated integer sequences."}
-knitr::include_graphics("Figs/circularex.png")
+knitr::include_graphics("Figs/circular-ex.png")
 ```
@@
 
-knitr::include_graphics("Figs/quasiex.png")
+knitr::include_graphics("Figs/quasi-ex.png")
 
@@
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities and vice versa.
+(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
 The hierarchical structure of time creates a natural nested ordering which can produce single-order-up or multiple-order-up granularities.
@@
 **Order** of a linear granularity can be comprehended as the level of coarseness associated with a linear granularity.
-For example, for two linear granularities $G$ and $H$, if $G$ groups into or finer than $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+For example, for two linear granularities $G$ and $H$, if $G$ _groups into_ or _finer than_ $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
 In any hierarchy table, linear granularities are arranged from lowest to highest order.
@@
 
-We consider the case of only one quasi-circular single order-up granularity in the hierarchy table.
-Any multiple-order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
-Two different approaches need to be employed leading to the following cases:
+We consider the case of only one quasi-circular single order-up granularity in the hierarchy table. Any multiple-order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a
+dicrete combination of circular and quasi-circular granularities.
+<!--function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
+--> Depending on the order of combination, two different approaches need to be employed leading to the following cases:
 
@@
 Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.
-A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
+A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
+A tsibble consists of an index, key and measured variables.
+An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
+A bottom granularity is represented by the index of the tsibble.
 Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
@@
 Hence data sets of the form <$C_i$, $C_j$, $v$> forms the basis for exploration and analysis of the measured variable.
+
+<!-- CONTENT SHORT -->
+<!-- Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
+A tsibble consists of an index, key and measured variables.
+An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
+<!--In a tsibble, each observation (row) is uniquely identified by index and key.A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the index of a tsibble.
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
+A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.\todo[inline]{Would it be worth to start this discussion about data structures earlier, with the linear granularity, and follow it through when appropriate; or put all discussion on implementation in the data structures section.} -->
 
@@
 \begin{definition}\label{def:clash}
-A {\bf harmony} is a pair of cyclic granularities that does not contain any missing combinations of its categories.
+A {\bf harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
 \end{definition}
@@
   #xlab("weekend/weekday") +
-  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + xlab("")
+  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + xlab("")+ 
+  theme_minimal()
 
@@
     xlab("quarters of the year")  +
-  scale_fill_brewer(palette = "Paired") + theme(legend.position = "right") + ggtitle("") + scale_y_log10()
+  scale_fill_brewer(palette = "Paired") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + theme_minimal()
 
@@
   ylab("") +
-  xlab("months of the year") + theme(legend.position = "right") + scale_fill_brewer(palette = "Set2") + ggtitle("") + scale_y_log10()
+  xlab("months of the year") + theme(legend.position = "right") + scale_fill_brewer(palette = "Set2") + ggtitle("") + scale_y_log10()+ theme_minimal()
 
@@
 ggpubr::annotate_figure(gg_fig,
-                left = text_grob("electricity demand [KWh]",  rot = 90))
+                left = text_grob("electricity demand [KWh]",  rot = 90)) + theme_minimal()
 
@@
   #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
+  theme_minimal()
 #
@@
   ylab("") + xlab("hours of the day") + ggtitle("") +
-  scale_x_discrete( breaks = seq(0, 23, 5))
+  scale_x_discrete( breaks = seq(0, 23, 5))+ 
+  theme_minimal()
 
@@
   scale_x_discrete(labels = c('wday','wend')) +
-  ggplot2::theme(axis.text.x = element_text(size = 7))
+  ggplot2::theme(axis.text.x = element_text(size = 7))+ 
+  theme_minimal()
 
@@
 <!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
-This seem indicate that players feel mounting presure to score more runs as they approach towards the end of the first inning.
+This seem to indicate that players feel mounting presure to score more runs as they approach towards the end of the first inning.
 Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting press and may adopt a more conservative strategy, to score runs.
-Thus winning teams like CSK and MI seem to emply different inning strategies when it comes to their batting order.
+Thus winning teams like CSK and MI seem to employ different inning strategies when it comes to their batting order.
 
@@
 
-Another interesting question could is: do runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
-For establishing that the fielding quality, we apply an indicator function on dismissals (1 if there was at least one catch in the previous over, 0 otherwise).
+Another interesting question could be: do runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
+For establishing the fielding quality, we apply an indicator function on dismissals (1 if there was at least one wicket in the previous over due to run out or catch, 0 otherwise).
 Runs in the current over is then the observation variable.
@@
 --> Dismissals in the previous over can lead to a batsman adopting a more defensive play style.
-Figure \ref{fig:cricex}(b) shows that when there are no dismissals in the previous over, leads to a higher median and quartile spread of run rate per over as compared to the case when there has been at least one dismissal in the previous over.
+Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over as compared to the case when there has been at least one dismissal in the previous over.
 
@@
       )
-     )
+     ) +  theme_minimal() 
+#geom_smooth(aes( x = over,
+#                               y=runs_per_over), method = lm, #formula = y ~ splines::bs(x, 3), se = FALSE)
 
@@
   ggtitle("b") +
-   theme(plot.title = element_text(face = "bold"))
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
 # + geom_boxplot(width = 0.5, aes(colour = lag_field))

### 2020-04-27 main.Rmd
Expand into a paragraph - how does it illustrate the linear granularities? How does hour, day,... relate to G...}

@@
 
-\todo[inline]{Expand into a paragraph - how does it illustrate the linear granularities?
-How does hour, day,...
-relate to G...} <!--, each of which are non-repeating in nature and arranged from past to future.
-If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.-->
+ <!--, each of which are non-repeating in nature and arranged from past to future. If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.-->
 
@@
 
-\autoref{fig:circular-dow} \todo[inline]{C(B,G) should be above C(B,H) in figure, and yuo could add the integers to the stacked sets.} illustrates the linear and the corresponding cyclical granularities.
+\autoref{fig:circular-dow} illustrates the linear and the corresponding cyclical granularities.
 Cyclical granularities can be considered to be cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).

### 2020-04-30 algo_rank_harmony.Rmd
algo updated with noramlity assumption

---
title: "rank_harmony"
author: "Sayani Gupta"
date: "24/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_M$\} and $v$ is the measured variable. Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$

2. There will be $J$ groups corresponding to each level $A_i$ of $C_1$. We compute the pairwise differences in distributions of these $J$groups. Thus there will be  $k = \binom{J}{2}$ pairwise distances corresponding to each $A_i$. Let $D_1$,...,$D_k$ be these distances and let their maximum be $M_k$. $D_{k} \sim N(\mu, \sigma^2)$. Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the Gumbel distribution.

3. So we get normalized variables $N_i = \frac{M_k - a_k}{b_k}$ \forall $A_i$, $\{i \in 1, 2, \dots, I\}$, each of which follows a standard Gumbel distribution.

4. We have been able to normalize for the number of levels of x-axis variable till Step 3. Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role. For this we can take the mean, median or any quantile value. But the median (and any quantiles) of the Gumbel distribution grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$. Hence, $median(N_i)/log(I)$ can serve as the maximum pairwise distance (MMPD).

5.  Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD. Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.

3. Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance. Suppose $q$ is the sample quantile vector computed for probability vector $p$. Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
 $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.

4. For each $i \in \{1, 2, \dots, l\}$, $\binom{m}{2}$ paiwise distances between distributions could be formed. Let $X_{i,p} \sim N(\mu, \sigma^2)$ $\forall i = {1, 2, \dots, l}$ and $\forall p = {1, 2, \dots, \binom{m}{2}}$ and $Y_p =  max\{X_{i,p}\}_{i=1}^{l}$. We will look at the distribution of the sample maximum $Y_p$.
Let $\overline Y$ and $S_{y}^2$ be the sample mean and standard deviation of the $\{Y_j\}$'s. One of the normalised limit distribution when talking about order statistics could be Gumbel distribution (or can be any generalized extreme value distribution).

$E(Y_{j}) = \mu + \sigma k_n$
$V(Y_{j}) = \sigma^2 c_n$

where $c_n$ and $k_n$ are the normalizing and centering constants.

```{r}
library(readr)
read_rds("cache/smart_harmony_n.rds")
read_rds("cache/smart_harmony_ak.rds")
```

### 2020-04-30 algo_rank_harmony.Rmd
algorithms updated and results conciliates for all versions

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(readr)
 ```
@@
 Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_M$\} and $v$ is the measured variable.
-Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$
+Hence for each combination ($A_i$, $B_j$), we have s subset of the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$
 
@@
 There will be $J$ groups corresponding to each level $A_i$ of $C_1$.
-We compute the pairwise differences in distributions of these $J$groups.
+We compute the pairwise differences in distributions of these $J$ groups.
 Thus there will be  $k = \binom{J}{2}$ pairwise distances corresponding to each $A_i$.
 Let $D_1$,...,$D_k$ be these distances and let their maximum be $M_k$.
-$D_{k} \sim N(\mu, \sigma^2)$.
-Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the Gumbel distribution.
 
-3. So we get normalized variables $N_i = \frac{M_k - a_k}{b_k}$ \forall $A_i$, $\{i \in 1, 2, \dots, I\}$, each of which follows a standard Gumbel distribution.
+## Version 1 (suggested by Rob)
 
-4.
-We have been able to normalize for the number of levels of x-axis variable till Step 3.
+ 3.1.
+If the $D_i$ distribution has heavy tails, then under EVT the maximum converges to Frechet and lim $P(M_k / a_k)$ converges to a limiting distribution where $a_k$ = the $(1-\frac{1}{k})$ quantile of the $D_i$ distribution.
+So normalization could be done  using $M_k / a_k$.
+
+ 4.1.
+For each facet, we have a normalised variable using Step 3.1.
+We can take the maximum across all facets as the final mesasure.
+But this maximum is not normalized against the number of facet levels.
+Or we can take the mean (to be discussed).
+
+```{r}
+read_rds("cache/smart_harmony_ak.rds")
+```
+
+## Version 2 (suggested by Sayani - new)
+
+ 3.2.
+Let $D_{k} \sim N(\mu, \sigma^2)$.
+Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the standard Gumbel distribution.
+So normalization could be done  using $\frac{M_k - a_k}{b_k}$, where $\Phi^{-1}(1/k)$ could be approximated by $(1 - \frac{1}{k})$ sample quantile and $\mu$ and $\sigma$ by sample mean and standard deviation.
+
+ 4.2. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now 
+$\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have $N_i = \frac{M_k - a_k}{b_k}$ each of which follows a standard Gumbel distribution.
 Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role.
-For this we can take the mean, median or any quantile value.
-But the median (and any quantiles) of the Gumbel distribution grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$.
-Hence, $median(N_i)/log(I)$ can serve as the maximum pairwise distance (MMPD).
+For this we can take the mean, median or any quantiles of these normalised variables.
+But the median (and any quantiles) of these Gumbel distributions grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$.
+Hence, median($N_i$)/log($I$) could serve as the final measure.
 
-5.
-Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
-Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+```{r}
+library(readr)
+read_rds("cache/smart_harmony_musigma.rds")
+```
 
-3.
-Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
-Suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
- $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
+## Version 3 (suggested by Sayani - old benchmark)
 
-4.
-For each $i \in \{1, 2, \dots, l\}$, $\binom{m}{2}$ paiwise distances between distributions could be formed.
-Let $X_{i,p} \sim N(\mu, \sigma^2)$ $\forall i = {1, 2, \dots, l}$ and $\forall p = {1, 2, \dots, \binom{m}{2}}$ and $Y_p =  max\{X_{i,p}\}_{i=1}^{l}$.
-We will look at the distribution of the sample maximum $Y_p$.
-Let $\overline Y$ and $S_{y}^2$ be the sample mean and standard deviation of the $\{Y_j\}$'s.
-One of the normalised limit distribution when talking about order statistics could be Gumbel distribution (or can be any generalized extreme value distribution).
+ 3.3.
+For each $i \in \{1, 2, \dots, I\}$, the maximum pairwise distance could be chosen from  $\binom{I}{2}$ distances using $D_{i}$ = max($d_{jl}$)/$J$(max($d_{jl}$) - min($d_{jl}$)) $\forall j, l = {1, 2, \dots, J}$.
 
-$E(Y_{j}) = \mu + \sigma k_n$
-$V(Y_{j}) = \sigma^2 c_n$
-
-where $c_n$ and $k_n$ are the normalizing and centering constants.
+ 4.3. Mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, I}$ is computed by
+$MMPD = (1/I)\sum_{i=1}^{l}{D_{i}}$.
 
@@
 read_rds("cache/smart_harmony_n.rds")
-read_rds("cache/smart_harmony_ak.rds")
 ```
+
+ 5.
+Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD (final measure).
+Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.

### 2020-04-30 algo_rank_harmony.Rmd
algo rank file with data update

@@
  1.
-Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_M$\} and $v$ is the measured variable.
+Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_J$\} and $v$ is the measured variable.
 Hence for each combination ($A_i$, $B_j$), we have s subset of the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$
@@
 ```{r}
-read_rds("cache/smart_harmony_ak.rds")
+read_rds("data/smart_harmony_ak.rds")
 ```
@@
  4.2. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now 
-$\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have $N_i = \frac{M_k - a_k}{b_k}$ each of which follows a standard Gumbel distribution.
+$\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have a normalised variable each of which follows a standard Gumbel distribution.
 Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role.
@@
 library(readr)
-read_rds("cache/smart_harmony_musigma.rds")
+read_rds("data/smart_harmony_musigma.rds")
 ```
 
-## Version 3 (suggested by Sayani - old benchmark)
+## Version 3 (old benchmark)
 
@@
 library(readr)
-read_rds("cache/smart_harmony_n.rds")
+read_rds("data/smart_harmony_n.rds")
 ```

### 2020-04-30 main.Rmd
algo rank file with data update

@@
 if (!file.exists("cache/smart_harmony.rds")) {
-  harmonies <- smart_meter10 %>%
+  sm <- smart_meter10 %>%
+  filter(customer_id %in% c(10017936))
+
+  harmonies <- sm %>%
   harmony(ugran = "month",
+          filter_in = "wknd_wday", 
           filter_out = c("hhour", "fortnight"))
 
-smart_harmony <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
+smart_harmony <- sm %>%
   rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh")%>%
@@
 
-  write_rds(smart_harmony, "cache/smart_harmony.rds")
+  write_rds(smart_harmony,"cache/smart_harmony_ak.rds")
 }

### 2020-04-30 main.Rmd
added examples for each definitions

@@
 This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Visualizing probability distributions conditional on one or more granularities is an indispensible tool for exploration.
-Analysts are expected to comprehesively explore the many ways to view and consider temporal data.
+Visualizing probability distributions conditional on one or more granularities is an indispensable tool for exploration.
+Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
 However, the plethora of choices and the lack of a systematic approach to do so quickly becomes overwhelming.
@@
 
-The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provides some background material on linear granularities and introduces calendar algebra for computing different linear granularities.
+The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provides some background material on linear granularities and calendar algebra for computing different linear granularities.
 Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
@@
 
-\newtheorem{definition}{Definition}
+\newtheorem{definition}{Definition}[section]
 
@@
 $G(i) \subset H(j).$
-\end{definition}.
+\end{definition}
 
@@
 \begin{example}
-{\rm Let $day \trianglelefteq week$ and $day \preceq week$ since every granule of $week$ is the union of some set of granules of $day$ and each $day$ is a subset of a $week$.
+{\rm Both $day \trianglelefteq week$ and $day \preceq week$ holds, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
 Consider another example where $day \trianglelefteq month$.
@@
 Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
-This grouping period becomes 4 and 400 years with the inclusion of leap years and leap centries respectively.}
+This grouping period becomes 4 and 400 years with the inclusion of leap years and leap seconds respectively.}
 \end{example}
@@
 \begin{definition}\label{def:periodical}
-A granularity $H$ is {\bf periodical} with respect to a granularity $G$ if:
+A granularity $G$ is {\bf periodical} with respect to a granularity $H$ if:
 (1) $G \trianglelefteq H$; and
@@
 \begin{example}
-{\rm In a Gregorian calendar without leap years we could say $day$ groups periodically into $month$ with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
-In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.}
+{\rm In a Gregorian calendar without leap years we could say $day$ groups periodically into month with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
+In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.}
 \end{example}
@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-
 \begin{example} 
-{\rm day = G$_{24}(\text{hour})$, week = G$_{7}(\text{day})$.}
+{\rm Due to even length of $day$ and $week$, we can derive them from $hour$ using the grouping operation as follows: day = G$_{24}(\text{hour})$, week = G$_{24*7}(\text{hour})$.}
 \end{example}
-
-<!-- CONTENT CUT -->
-<!-- where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise.
-Otherwise, repeated operations of grouping might result in a granule of "resultant" granularity "day" or "week", which is different from a time-stamped "day" or "week".
--->
 
@@
 
-For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning_2002].
+\begin{example} 
+{\rm Due to the uneven length of $month$, we need to use the altering-tick operation to compute $month$ from $day$.
+We can consider a granularity $pseudomonth$ which computes $month$ only for non-leap years by grouping 31 granules of $day$, and then shrink April, June, September and November by 1 granule of $day$, and shrink February by 3 granules of $day$ [@Ning_2002].
 For more variations of calendar algebra operations, see @Ning_2002.
+}
+\end{example}
 
@@
 
-\todo[inline]{Check with Di to revise the first two statements}
 Cyclic granularities represent cyclical repetitions in time.
-Weeks is an example of a cyclical granularity because every 7 days there's another Monday.
-Cyclic granularities can be constructed by conducting modular arithmetic on two linear granularities.
-Cycles can be either _regular_, called **circular**, or _irregular_, **quasi-circular**.
+They can be thought of as additional categorizations of time that are not linear.<!--Weeks is an example of a cyclical granularity because every 7 days there's another Monday.--> Cyclic granularities can be constructed by operating on two linear granularities.
+Cycles can be either _regular_, called **circular**, or _irregular_, **quasi-circular** when these two linear granularities relate periodically.
 <!-- with every granule of coarser linear granularity being made up of same (regular) or different (irregular) number of granules of finer linear granularity.
@@
 
-\autoref{fig:circular-dow} illustrates the linear and the corresponding cyclical granularities.
+\begin{example}
+{\rm \autoref{fig:circular-dow} illustrates the linear and the corresponding cyclical granularities.
 Cyclical granularities can be considered to be cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
@@
 Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns  $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
-
 Each circular granularity can use descriptive label mappings.
-<!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
+Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
 $$
@@
 $$
-for example.
+for example.}
+\end{example}
 
@@
 
-Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.
-Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
+\begin{example}
+{\rm Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.
+Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
 Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$.
@@
 
-<!-- \begin{equation} \label{eq:eq4} -->
-<!-- \begin{split} -->
-<!-- Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\ -->
-<!--   & = L(1) \\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-
-<!-- CONTENT CUT -->
-  <!-- & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\ -->
-  <!-- & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\ -->
-  <!-- & = L(8 - 2*\vert T_{0}\vert - \vert T_{1}\vert) \\ -->
-  <!-- & = L(8 - 2*2 - 3) \\ -->
-
-  <!-- & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\ -->
-  <!-- & = L(10 - \vert T_{0 \mod 2}\vert) \\ -->
-  <!-- & = L(10 - \vert T_{0}\vert) \\ -->
-  <!-- & = L(10 - 7) \\ -->
-
-\begin{align}
+\begin{align*}
 \begin{split}
@@
 \end{split}
-\end{align}
+\end{align*}
+.}
+\end{example}
 
@@
 
-Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
+\begin{example}
+{\rm Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
 Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
@@
 It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
-Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.
+Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.}
+\end{example}
 
@@
 knitr::include_graphics("Figs/quasi-ex.png")
-
 ```
+
 <!-- #  -->
@@
 
-**Example:** Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar with the index set  $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ mapping "kins" to a set of positive integers.
+\begin{example} 
+{\rm Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar with the index set  $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ mapping "kins" to a set of positive integers.
 
@@
 \end{equation}
+}
+\end{example}
 
@@
 
-**Example:** <!--Let us revisit Gregorian calendar for addressing this case.--> Suppose we have a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
+\begin{example} 
+{\rm Suppose we have a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
 Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.
@@
 Only one single-order-up granularity day-of-month is quasi-circular.
+
+Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we have:
+
+ $$C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)$$
+ $$C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert)$$, where $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$
+
+}
+\end{example}
 
@@
 ```
-
-Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we have:
-
-<!-- to compute the multiple-order-up granularity hour-of-month and day-of-year for the Gregorian calendar considered. -->
-
- - $C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)$
- - $C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert)$, where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$
 
@@
 
-<!-- **Example:**  -->
-Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
+\begin{example}
+{\rm Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
 
@@
 The R package `gravitas` [@R-gravitas] is used to facilitate the systematic exploration here.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to cosider is which cyclic time granularities we can look at exhaustively.
+While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to consider is which cyclic time granularities we can look at exhaustively.
 Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
@@
 Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ (via \ref{sec:chooseharmonies} item #4) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
-If these options are considered too numerous, the smallest and largest granularities are good candidates to disconsider.
-Hense, we removed half hour and year.
+If these options are considered too numerous, the smallest and largest granularities are good candidates to consider.
+Hence, we removed half hour and year.
 Then we will be left $(4*3/2) = 6$ cyclic granularites, namely: "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
@@
 Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet.
-Choosing 2 of the possible 6 granularities, which is equivalent to having ${2 \choose 6} = 30$ canidates for visualization.
+Choosing 2 of the possible 6 granularities, which is equivalent to having ${2 \choose 6} = 30$ candidates for visualization.
 Fortunately, harmony granularities can be identified among those 30 possibilities to narrow the search.
@@
 
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one observation shows a particular behavior, but this behavior paradoxically decomes obsured by aggregation.
+If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one observation shows a particular behavior, but this behavior paradoxically becomes obscured by aggregation.
 For example in a particular neighborhood one household may have the least daily power consumption for a full week, yet still not be the household with the minimum weekly power consumption.
@@
 
-The method is not only restricted to temporal data, and can be generalised to many hierarchical granularies (with  continuous and uni-directional nature).
+The method is not only restricted to temporal data, and can be generalized to many hierarchical granularities (with  continuous and uni-directional nature).
 We illustrate this with an application to the sport cricket.
@@
 Example granularites then include ball of the over, over of the inning and ball of the inning.
-Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularies will are aperiodic.
+Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularities will are aperiodic.
 For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
@@
 <!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
-This seem to indicate that players feel mounting presure to score more runs as they approach towards the end of the first inning.
+This seem to indicate that players feel mounting pressure to score more runs as they approach towards the end of the first inning.
 Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting press and may adopt a more conservative strategy, to score runs.

### 2020-04-30 main.Rmd
minor edits - from section visualization figs

@@
 \end{align*}
-.}
+}
 \end{example}
@@
 The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by [@Wickham2009pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
+The layered grammar of graphics proposed by @Wickham2009pk, which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
 Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
@@
 A potential approach could be to categorize the levels as very high/high/medium/low for each cyclic granularity and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
-Default values for these categorizations can be chosen based on levels of common temporal granularities like days-of-month, days-of-fortnight and days-of-week.<!--For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.>
+Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight or days of the week.<!--For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.>
 
@@
 As a result, we should avoid plotting clashes.
-For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarter-of-year.
-This plot does not work because every quarter does not correspond to all months of the year, for example, the first quarter would not correspond to the month December of an year.
+For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarters of the year.
+This plot does not work because every quarter does not correspond to all months of the year, for example, the first quarter of the year would never correspond to the month December of an year.
 
@@
 ```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Distribution of energy consumption displayed through letter value plots.
-Plot (a) displays consumption across month-of-year faceted by quarter-of-year, (b) across weekday/weekend faceted by quarter-of-year and (c) across quarter-of-year faceted by weekday/weekend.
+Plot (a) displays consumption across month-of-year faceted by quarters of the year, (b) across weekday/weekend faceted by quarters of the year and (c) across quarters of the year faceted by weekday/weekend.
 Plots (b) and (c) both show harmonies since each quarter consists of both weekdays and weekends.
@@
 Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$--> With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
-\autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarter-of-year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
+\autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
 \autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
@@
 
-The R package `gravitas` [@R-gravitas] is used to facilitate the systematic exploration here.
+The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] is used to facilitate the systematic exploration here.
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to consider is which cyclic time granularities we can look at exhaustively.
@@
 Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
-Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ (via \ref{sec:chooseharmonies} item #4) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
+Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ (via \ref{sec:data-structure}) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
 If these options are considered too numerous, the smallest and largest granularities are good candidates to consider.
@@
 Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet.
-Choosing 2 of the possible 6 granularities, which is equivalent to having ${2 \choose 6} = 30$ candidates for visualization.
+Choosing 2 of the possible 6 granularities, which is equivalent to having ${6 \choose 2} = 30$ candidates for visualization.
 Fortunately, harmony granularities can be identified among those 30 possibilities to narrow the search.
@@
 
-In \autoref{fig:bothcust} (b), the distribution of energy consumption is plotted across the harmony pair (wknd_wday, hour_day) through an area quantile plot.
+In \autoref{fig:bothcust} (b), the distribution of energy consumption is plotted across the harmony pair (weekday/weekend, hour of the day) through an area quantile plot.
 The black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
@@
 There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
-Also the hours from 7 to 13 look most volatile.
-If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
+<!--Also the hours from 7 to 13 look most volatile.--> If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
 For example in \autoref{fig:bothcust}(a), it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
@@
 
-However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to shed light on the periodic behavior of a non-temporal data set similar to any temporal data set.
+However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to shed light on the aperiodic behavior of a non-temporal data set similar to aperiodic events like formal meetings, workshops, conferences, school semesters in a temporal set up.
 There are many interesting questions that could possibly be answered with such a data set irrespective of the type of cyclic granularities.
@@
 This seem to indicate that players feel mounting pressure to score more runs as they approach towards the end of the first inning.
-Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting press and may adopt a more conservative strategy, to score runs.
+Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting pressure and may adopt a more conservative strategy, to score runs.
 Thus winning teams like CSK and MI seem to employ different inning strategies when it comes to their batting order.
@@
 
-```{r cricex, fig.cap= "Runs per over shown with different distribution displays, and granularity arrangements.
+```{r cricex, fig.cap= "Runs per over shown with different distribution displays, and granularities.
 Plot (a) shows letter value plot across overs faceted by innings.
-The team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
+For the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
 Plot (b) shows box plot of runs per over across an indicator of wickets in previous over faceted by current over.

### 2020-04-30 main.Rmd
quasi periodic diagram changed again to stack 0 weeks on top of each other
fixed section 2.3 and caption for figure 1

--- 2020-04-30 main.Rmd
@@
 
-```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="The time domain distributed as linear granularities hour, day, week, month and year.
-These are ordered with ordering guided by integers and hence is unidirectional and non-repeating."}
+```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="Illustration of time domain, linear granularities and index set.
+Hour, day, week, month and year are linear granularities and can also be considered to be time domains.
+These are ordered with ordering guided by integers and hence is unidirectional and non-repeating.
+Hours could also be considered the index set, and a bottom granularity.
+"}
 knitr::include_graphics("Figs/linear-ex.png")
@@
 
-The grouping operation:
-: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
-The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
+\begin{definition}\label{def:grouping_operation}
+Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
+{\bf The grouping operation} Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
 More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
+\end{definition}
+
 \begin{example} 
@@
 
-The altering-tick operation:
-: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
-The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
+\begin{definition}\label{altering-tick}
+Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
+{\bf The altering-tick operation} $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
+\end{definition}
 
 \begin{example}
-{\rm Due to the uneven length of $month$, we need to use the altering-tick operation to compute $month$ from $day$.
-We can consider a granularity $pseudomonth$ which computes $month$ only for non-leap years by grouping 31 granules of $day$, and then shrink April, June, September and November by 1 granule of $day$, and shrink February by 3 granules of $day$ [@Ning_2002].
-For more variations of calendar algebra operations, see @Ning_2002.
-}
+The altering-tick operation can be used to compute the granularity $pseudomonth$, for the uneven lengths of $month$, by grouping 31 granules of $day$, followed by shrinking April, June, September and November by 1 granule of $day$, and shrinking February by 3 granules of $day$ for non-leap years.
 \end{example}
+
+\noindent For more variations of calendar algebra operations, see @Ning_2002.
 

### 2020-05-01 main.Rmd

--- 2020-04-30 main.Rmd
@@
 
-knitr::include_graphics("Figs/quasi-ex.png")
+knitr::include_graphics("Figs/aperiodic-ex3.png")
 ```

### 2020-05-01 algo_rank_harmony.Rmd

@@
 library(readr)
+library(gravitas)
+library(tidyverse)
 ```
@@
 
+```{r}
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  prob_plot("day_week",
+    "day_month",
+    response = "general_supply_kwh",
+    plot_type = "quantile",
+    symmetric = TRUE,
+    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  scale_y_sqrt() +
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
+  theme_minimal()
+```
+
 ## Version 2 (suggested by Sayani - new)
@@
 ```
+```{r}
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  prob_plot("hour_day",
+    "wknd_wday",
+    response = "general_supply_kwh",
+    plot_type = "quantile",
+    symmetric = TRUE,
+    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  scale_y_sqrt() +
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
+  theme_minimal()
+```
 
@@
 
+```{r}
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  prob_plot("week_month",
+    "wknd_wday",
+    response = "general_supply_kwh",
+    plot_type = "quantile",
+    symmetric = TRUE,
+    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  scale_y_sqrt() +
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
+  theme_minimal()
+```
+
  5.

### 2020-05-06 main.Rmd
grouping operation section - integer arithmetic intro

@@
 
-\begin{definition}\label{altering-tick}
-Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
+If we assume that linear granularities are integer sets with each integer representing a granule, we would be able to express the grouping operation through integer arithmetic which would pair better with the modular arithmetic used in the next section.
+The grouping operation Group$_m(G)$ essentially leads to a new granularity whose granules are the integer part of the quotient when the integer set spanned by linear granularity $G$ is divided by $m$ (the divisor), such that, $G_2(i) = \lfloor G_1(j)/m \rfloor$.
+While new granularities are generated with ordering guided by quotients, we shall see how for some cyclic granularities ordering is guided by the remainder with the assumption of linear granularities being represented by integer sets.
+
+<!-- Alternatively, given any integer $G_1(j)$ and a positive integer $m$, if there exists unique integers $G_2(i)$ and $R$ such that $G1(j) = m * G2(i) + R$ where $0 ≤ R < m$, then $G_2$ is a granularity that is formed -->
+
+<!-- CONTENT CUT altering-tick -->
+<!-- \begin{definition}\label{altering-tick} -->
+<!-- Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
 {\bf The altering-tick operation} $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
-\end{definition}
-
-\begin{example}
-The altering-tick operation can be used to compute the granularity $pseudomonth$, for the uneven lengths of $month$, by grouping 31 granules of $day$, followed by shrinking April, June, September and November by 1 granule of $day$, and shrinking February by 3 granules of $day$ for non-leap years.
-\end{example}
+-->
+<!-- \end{definition} -->
+
+<!-- \begin{example} -->
+<!-- The altering-tick operation can be used to compute the granularity $pseudomonth$, for the uneven lengths of $month$, by grouping 31 granules of $day$, followed by shrinking April, June, September and November by 1 granule of $day$, and shrinking February by 3 granules of $day$ for non-leap years.
+-->
+<!-- \end{example} -->
 

### 2020-05-06 algo_rank_harmony.Rmd
algo update

@@
 
+```{r}
+read_rds("data/smart_harmony_gamma.rds")
+```
+
+## Version 1 (Distances follow Gamma)
+
+ 3.1.
+Suppose the reponse variable follows Normal distribution.
+Then any distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution.
+Hence if we proceed with the assumption that the $D_i$ distribution has Gamma distribution, then under EVT the maximum converges to Gumbel distribution and lim $P(M_k - b_k)/a_k)$ converges to a limiting Gumbel distribution where $b_k = (\frac{1}{\beta} (ln n + (\alpha - 1)ln ln n - ln \gamma(\alpha))$ and $a_k = \frac{1}{\beta}$.
+Here $\alpha$ and $\beta$ could be estimated using $(\mu/s)^2$ and $(s^2/\mu)$ respectively.
+
+ 4.1.
+For each facet, we have a normalised variable using Step 3.1.
+We can take the maximum across all facets as the final mesasure.
+But this maximum is not normalized against the number of facet levels.
+Or we can take the mean (to be discussed).
+
+```{r}
+read_rds("data/smart_harmony_ak.rds")
+```
+
 ## Version 1 (suggested by Rob)
 
- 3.1.
+ 3.2.
 If the $D_i$ distribution has heavy tails, then under EVT the maximum converges to Frechet and lim $P(M_k / a_k)$ converges to a limiting distribution where $a_k$ = the $(1-\frac{1}{k})$ quantile of the $D_i$ distribution.
@@
 
- 4.1.
-For each facet, we have a normalised variable using Step 3.1.
+ 4.2.
+For each facet, we have a normalised variable using Step 3.2.
 We can take the maximum across all facets as the final mesasure.
@@
 
-## Version 2 (suggested by Sayani - new)
+## Version 3 (Distances follow Normal)
 
- 3.2.
+ 3.3.
 Let $D_{k} \sim N(\mu, \sigma^2)$.
@@
 
- 4.2. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now 
+ 4.3. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now 
 $\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have a normalised variable each of which follows a standard Gumbel distribution.
@@
 
-## Version 3 (old benchmark)
+## Version 4 (old benchmark)
 
- 3.3.
+ 3.4.
 For each $i \in \{1, 2, \dots, I\}$, the maximum pairwise distance could be chosen from  $\binom{I}{2}$ distances using $D_{i}$ = max($d_{jl}$)/$J$(max($d_{jl}$) - min($d_{jl}$)) $\forall j, l = {1, 2, \dots, J}$.
 
- 4.3. Mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, I}$ is computed by
+ 4.4. Mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, I}$ is computed by
 $MMPD = (1/I)\sum_{i=1}^{l}{D_{i}}$.
@@
 Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+
+## results on cricket data
+
+Gamma
+
+```{r}
+library(readr)
+read_rds("data/smart_harmony_normal_cric.rds")
+```
+Normal 
+```{r}
+read_rds("data/smart_harmony_gamma_cric.rds")
+```

### 2020-05-06 main.Rmd
minor

@@
 
-<!-- Pairs of granularities can be categorized as either a *harmony* or *clash*, where harmonies are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
+<!-- Pairs of granularities can be categorized as either a *harmony* or *clash*, where = are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
 -->
@@
 smart_harmony <- sm %>%
-  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh")%>%
+  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh") %>%
   rename(`facet variable` = facet_variable,

### 2020-05-07 main.Rmd
revised definition 7

@@
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
+The grouping operation Group$_m(G)$, provides a new granularity whose granules are the integer part of the quotient when the integer set spanned by linear granularity $G$ is divided by $m$, such that, $G_2(i) = \lfloor G_1(j)/m \rfloor$.
 \end{definition}
@@
 \end{example}
-
-If we assume that linear granularities are integer sets with each integer representing a granule, we would be able to express the grouping operation through integer arithmetic which would pair better with the modular arithmetic used in the next section.
-The grouping operation Group$_m(G)$ essentially leads to a new granularity whose granules are the integer part of the quotient when the integer set spanned by linear granularity $G$ is divided by $m$ (the divisor), such that, $G_2(i) = \lfloor G_1(j)/m \rfloor$.
-While new granularities are generated with ordering guided by quotients, we shall see how for some cyclic granularities ordering is guided by the remainder with the assumption of linear granularities being represented by integer sets.
 

### 2020-05-07 main.Rmd
removed mention of altering-tick operation from paper

@@
 
-\newtheorem{definition}{Definition}[section]
+\newtheorem{definition}{Definition}
 
@@
 \begin{example}
-{\rm Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.
+{\rm Consider $G'$ such that every two consecutive granules of $G'$ are made up of $7$ and $5$ granules of $B$ respectively within each period of the grouping ($B$, $G'$).
 Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
@@
 \end{example}
+
+<!--such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.-->
 

### 2020-05-07 main.Rmd
working  on section 4

@@
 
-# Cyclic calendar algebra {#sec:cyclic-calendar}
-
-In Section \ref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
-In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities <!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
-Since, our method caters to computation of cyclic granularities, we shall refer to it as--> through cyclic calendar algebra operations.
+# Hierarchy and ordering {#sec:cyclic-calendar}
+
+<!-- In Section \ref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
+In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities --><!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
+Since, our method caters to computation of cyclic granularities, we shall refer to it as--> <!-- through cyclic calendar algebra operations.-->
 
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
+(1) **single-to-multiple** and (2) **multiple-to-single** \todo[inline]{these don't make sense to me, they are just the reverse of each other, so why are both needed.
+its not  clear from the  next block of text what each operation does} which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
 The hierarchical structure of time creates a natural nested ordering which can produce single-order-up or multiple-order-up granularities.
@@
 
-**Order** of a linear granularity can be comprehended as the level of coarseness associated with a linear granularity.
-For example, for two linear granularities $G$ and $H$, if $G$ _groups into_ or _finer than_ $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+\begin{definition}\label{def:order}
+The {\bf order} of a linear granularity is the level of coarseness associated with a linear granularity.
+For example, for two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
 In any hierarchy table, linear granularities are arranged from lowest to highest order.
-
-**Hierarchy table** Let $H_n: (G, C, K)$ be a hierarchy table containing $n$ linear granularities.
+\end{definition}
+
+\begin{definition}\label{def:hierarchy} 
+Let $H_n: (G, C, K)$ be a {\bf hierarchy table} containing $n$ linear granularities.
 $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
@@
 It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.
+\end{definition}
 

### 2020-05-07 algo_rank_harmony.Rmd
added gamma and weibull norming constants and paramter estimation

@@
 ---
-title: "rank_harmony"
+title: "Choosing harmonies based on average maximum pairwise distance of distributions"
 author: "Sayani Gupta"
-date: "24/04/2020"
-output: html_document
+output:
+  html_document:
+    theme: united
+    highlight: tango
+    toc: true
 ---
@@
 ```
+
+# Pointers
+
+- A major result in the field of EVT is the probability density of the maximum or minimum can only be one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+But the normalising constants vary depending on the underlying distribution of the variables.
+Hence to normalize appropriately, it is important to assume a distribution of distances in our case.
+
+- Under broad conditions, the similarities that arise from any distribution $L_p$ norms - a class of commonly applied distance matrix from one feature vector to other vectors could be
+
+  a) Normal: Pekalska and Duin [23] assumed that Lp-distances between
+feature vectors are normally distributed, if the feature values are independent, identically distributed, and have limited variance.
+We can not assume that feature values are independent.
+
+  b) Gamma: Ferencz et al. [7] have considered the Gamma distribution to model the L2-distances.
+This is intuitive since if the reponse variable follows Normal distribution, then $L_2$ norm distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution.
+
+  c) Weibull: Weibull distributed, if feature vectors are correlated and non-identically distributed. [The Distribution Family of Similarity Distances]
+
+- For each of the cases, the normalising constants are a function of the paramters which have been estimated using method of moments.
+
+  a) **Normal**  
+  $D_{k} \sim N(\mu, \sigma^2)$  
+    *Normalising constants*:  
+    $a_k = \mu − \sigma\Phi^{-1}(1/k) \quad \quad b_k = -\sigma/\Phi^{-1}(1/k)$  
+    *Estimated parameters*:  
+    $\hat{\mu}=\dfrac{1}{n}\sum\limits_{i=1}^n X_i \quad \quad \hat{\sigma^2}=\dfrac{1}{n}\sum\limits_{i=1}^n (X_i-\bar{X})^2$  
+    $\hat{\Phi^{-1}(1/k)} =  (1 - \frac{1}{k})_{th}$ sample quantile
+
+  b) **Gamma**  
+  $D_{k} \sim \gamma(\alpha, \beta)$  
+    *Normalising constants:*  
+    $b_k = (\frac{1}{\beta} (ln n + (\alpha - 1)ln ln n - ln       \gamma(\alpha))  \quad \quad a_k = \frac{1}{\beta}$  
+    *Estimated parameters:*  
+    $\alpha = (\mu/s)^2 \quad \quad \beta = (s^2/\mu)$
+
+  c)  **Weibull**
+
+\begin{equation}
+
+f_D(d; \lambda, k) = \left\{ \begin{array}{cl}
+    \frac{k}{\lambda}(\frac{d}{\lambda})^{k-1} e^{-(d/\lambda)^k} & \ ; \ d \geq 0 \\
+    0 & \ ; \ d < 0 \end{array} \right. \
+\end{equation}  
+    *Normalising constants:*  
+  $b_k = F^{-1}(1-1/n)  \quad \quad a_k = A(b_k)$, where $A(x)$ is an auxilliary function which can be chosen as $(1-F(x)/f(x))$.
+
+  \begin{equation}
+F(x) = 1 - e^{{-(\frac{d}{\lambda})}^k}
+\end{equation}
+    *Estimated parameters:*  
+  $k = (\frac{\sigma}{\mu})^{-1.086} \quad \quad \lambda = \frac{\mu}{\Gamma(1+1/k)} \quad \quad \hat{\mu}=\dfrac{1}{n}\sum\limits_{i=1}^n D_i \quad \quad \hat{\sigma^2}=\dfrac{1}{n-1}\sum\limits_{i=1}^n (X_i-\bar{D})^2$
+
+# Algorithm
 
@@
 Let $D_1$,...,$D_k$ be these distances and let their maximum be $M_k$.
-
-```{r}
-read_rds("data/smart_harmony_gamma.rds")
-```
 
@@
 ```{r}
-read_rds("data/smart_harmony_ak.rds")
+read_rds("data/smart_harmony_gamma.rds")
 ```

### 2020-05-07 algo_rank_harmony.Rmd
update algorithm

@@
 title: "Choosing harmonies based on average maximum pairwise distance of distributions"
-author: "Sayani Gupta"
 output:
@@
 feature vectors are normally distributed, if the feature values are independent, identically distributed, and have limited variance.
+(“The Dissimilarity Representation for Pattern Recognition | Series in Machine Perception and Artificial Intelligence.” n.d.
+Accessed May 8, 2020.
+https://www.worldscientific.com/worldscibooks/10.1142/5965.)
 We can not assume that feature values are independent.
 
-  b) Gamma: Ferencz et al. [7] have considered the Gamma distribution to model the L2-distances.
+  b) Gamma: (Ferencz, A., E.
+G.
+Learned-Miller, and J.
+Malik.
+2005.
+“Building a Classification Cascade for Visual Identification from One Example.” In Tenth IEEE International Conference on Computer Vision (ICCV’05) Volume 1, 1:286–93 Vol.
+1.) have considered the Gamma distribution to model the L2-distances.
 This is intuitive since if the reponse variable follows Normal distribution, then $L_2$ norm distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution.
@@
 
-# Algorithm
+<!-- # Algorithm -->
 
- 1.
+<!--  1.
 Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_J$\} and $v$ is the measured variable.
-Hence for each combination ($A_i$, $B_j$), we have s subset of the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$
+Hence for each combination ($A_i$, $B_j$), we have s subset of the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$ -->
 
- 2.
+<!--  2.
 There will be $J$ groups corresponding to each level $A_i$ of $C_1$.
@@
 Let $D_1$,...,$D_k$ be these distances and let their maximum be $M_k$.
+-->
 
-## Version 1 (Distances follow Gamma)
+# Smart meter data
 
- 3.1.
+## Distances follow Weibull
+
+```{r}
+library(readr)
+read_rds("data/smart_harmony_weibull.rds") %>% knitr::kable()
+```
+
+## Distances follow Gamma
+
+ <!-- 3.1.
 Suppose the reponse variable follows Normal distribution.
@@
 Here $\alpha$ and $\beta$ could be estimated using $(\mu/s)^2$ and $(s^2/\mu)$ respectively.
+-->
 
- 4.1.
+ <!-- 4.1.
 For each facet, we have a normalised variable using Step 3.1.
@@
 Or we can take the mean (to be discussed).
+-->
 
 ```{r}
-read_rds("data/smart_harmony_gamma.rds")
+read_rds("data/smart_harmony_gamma.rds") %>% knitr::kable()
 ```
+
+## Distances follow Normal
+
+<!--  3.3.
+Let $D_{k} \sim N(\mu, \sigma^2)$.
+Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the standard Gumbel distribution.
+So normalization could be done  using $\frac{M_k - a_k}{b_k}$, where $\Phi^{-1}(1/k)$ could be approximated by $(1 - \frac{1}{k})$ sample quantile and $\mu$ and $\sigma$ by sample mean and standard deviation.
+-->
+
+<!--  4.3. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now  -->
+<!-- $\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have a normalised variable each of which follows a standard Gumbel distribution.
+Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role.
+For this we can take the mean, median or any quantiles of these normalised variables.
+But the median (and any quantiles) of these Gumbel distributions grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$.
+Hence, median($N_i$)/log($I$) could serve as the final measure.
+-->
+
+```{r}
+library(readr)
+read_rds("data/smart_harmony_normal.rds") %>% knitr::kable()
+```
+
+```{r, echo = FALSE, eval = FALSE}
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  prob_plot("hour_day",
+    "wknd_wday",
+    response = "general_supply_kwh",
+    plot_type = "quantile",
+    symmetric = TRUE,
+    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  scale_y_sqrt() +
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
+  theme_minimal()
+```
+
+<!--  5.
+Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD (final measure).
+Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+-->
+
+# Cricket data
+
+## Distances follow Weibull
+```{r}
+read_rds("data/smart_cric_weibul.rds")%>% knitr::kable()
+```
+
+## Distances follow Gamma
+
+```{r}
+library(readr)
+read_rds("data/smart_cric_gamma.rds") %>% knitr::kable()
+```
+
+## Distances follow Normal 
+```{r}
+read_rds("data/smart_cric_normal.rds")%>% knitr::kable()
+```
+
+# Old benchmarks
 
@@
 
-```{r}
+```{r, eval=FALSE, echo=FALSE}
 read_rds("data/smart_harmony_ak.rds")
@@
 
-```{r}
+```{r, eval = FALSE}
 smart_meter10 %>%
@@
     "day_month",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    symmetric = TRUE,
-    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
-```
-
-## Version 3 (Distances follow Normal)
-
- 3.3.
-Let $D_{k} \sim N(\mu, \sigma^2)$.
-Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the standard Gumbel distribution.
-So normalization could be done  using $\frac{M_k - a_k}{b_k}$, where $\Phi^{-1}(1/k)$ could be approximated by $(1 - \frac{1}{k})$ sample quantile and $\mu$ and $\sigma$ by sample mean and standard deviation.
-
- 4.3. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now 
-$\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have a normalised variable each of which follows a standard Gumbel distribution.
-Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role.
-For this we can take the mean, median or any quantiles of these normalised variables.
-But the median (and any quantiles) of these Gumbel distributions grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$.
-Hence, median($N_i$)/log($I$) could serve as the final measure.
-
-```{r}
-library(readr)
-read_rds("data/smart_harmony_musigma.rds")
-```
-```{r}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("hour_day",
-    "wknd_wday",
     response = "general_supply_kwh",
@@
 
-```{r}
+```{r, eval = FALSE, echo=FALSE}
 library(readr)
-read_rds("data/smart_harmony_n.rds")
+read_rds("data/smart_harmony_n.rds") %>% knitr::kable()
 ```
 
-```{r}
+```{r, echo = FALSE, eval = FALSE}
 smart_meter10 %>%
@@
 ```
-
- 5.
-Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD (final measure).
-Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
-
-## results on cricket data
-
-Gamma
-
-```{r}
-library(readr)
-read_rds("data/smart_harmony_normal_cric.rds")
-```
-Normal 
-```{r}
-read_rds("data/smart_harmony_gamma_cric.rds")
-```

### 2020-05-07 algo_rank_harmony.Rmd
reference updated

@@
 
-- Under broad conditions, the similarities that arise from any distribution $L_p$ norms - a class of commonly applied distance matrix from one feature vector to other vectors could be
+- Under broad conditions, the similarities that arise from any distribution $L^p$ norms - a class of commonly applied distance matrix from one feature vector to other vectors could be
 
-  a) Normal: Pekalska and Duin [23] assumed that Lp-distances between
+  a) Normal: Pekalska and Duin assumed that Lp-distances between
 feature vectors are normally distributed, if the feature values are independent, identically distributed, and have limited variance.
-(“The Dissimilarity Representation for Pattern Recognition | Series in Machine Perception and Artificial Intelligence.” n.d.
+We can not assume that feature values are independent.(“The Dissimilarity Representation for Pattern Recognition | Series in Machine Perception and Artificial Intelligence.”   n.d.
 Accessed May 8, 2020.
 https://www.worldscientific.com/worldscibooks/10.1142/5965.)
-We can not assume that feature values are independent.
 
-  b) Gamma: (Ferencz, A., E.
+  b) Gamma: Ferencz, Miller and Malik have considered the
+Gamma distribution to model the L2-distances.
+This is intuitive since if the reponse variable follows
+Normal distribution, then $L_2$ norm distance measure which
+is similar to taking squares of differences of Normal
+variables would follow a Chi-Squared distribution, a special
+case of Gamma distribution.(Ferencz, A., E.
 G.
@@
 “Building a Classification Cascade for Visual Identification from One Example.” In Tenth IEEE International Conference on Computer Vision (ICCV’05) Volume 1, 1:286–93 Vol.
-1.) have considered the Gamma distribution to model the L2-distances.
-This is intuitive since if the reponse variable follows Normal distribution, then $L_2$ norm distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution.
+1.)
 
-  c) Weibull: Weibull distributed, if feature vectors are correlated and non-identically distributed. [The Distribution Family of Similarity Distances]
+  c) Weibull: Weibull distributed, if feature vectors are
+correlated and non-identically distributed.
+(Burghouts, Gertjan, Arnold Smeulders, and Jan-Mark
+Geusebroek. 2008. “The Distribution Family of Similarity
+Distances.” In Advances in Neural Information Processing
+Systems 20, edited by J. C. Platt, D. Koller, Y. Singer, and
+S. T. Roweis, 201–8. Curran Associates, Inc.)
 
-- For each of the cases, the normalising constants are a function of the paramters which have been estimated using method of moments.
+- For each of the cases, the norming constants are a function
+of the paramters which have been estimated using method of moments.
+
+ Norming constants are taken from Embrechts, Paul, Claudia   Klüppelberg, and Thomas Mikosch. 2013. Modelling Extremal Events: For Insurance and Finance. Springer Science & Business Media.
 
@@
     *Estimated parameters:*  
-  $k = (\frac{\sigma}{\mu})^{-1.086} \quad \quad \lambda = \frac{\mu}{\Gamma(1+1/k)} \quad \quad \hat{\mu}=\dfrac{1}{n}\sum\limits_{i=1}^n D_i \quad \quad \hat{\sigma^2}=\dfrac{1}{n-1}\sum\limits_{i=1}^n (X_i-\bar{D})^2$
+  $k = (\frac{\sigma}{\mu})^{-1.086} \quad \quad \lambda = \frac{\mu}{\Gamma(1+1/k)} \quad \quad \hat{\mu}=\dfrac{1}{n}\sum\limits_{i=1}^n D_i=\bar{D} \quad \quad \hat{\sigma^2}=\dfrac{1}{n-1}\sum\limits_{i=1}^n (X_i-\bar{D})^2$
 

### 2020-05-08 main.Rmd
non-negative integers, not positive integers

@@
 \begin{definition}\label{def:index set}
-There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of positive integers. Often, this is thought of as $t=0, \dots, T$.
+There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of non-negative integers. Often, this is thought of as $t=0, \dots, T$.
 \end{definition}
@@
 
-# Hierarchy and ordering {#sec:cyclic-calendar}
+# Relativities {#sec:cyclic-calendar}
 

### 2020-05-08 main.Rmd
structural changes to Section 3. Merging Section 4 to Section 3

@@
 -->
+
+\begin{definition}\label{def:order}
+The {\bf order} of a linear granularity is the level of coarseness associated with a linear granularity.
+For example, for two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+\end{definition}
 
@@
 
-# Relativities {#sec:cyclic-calendar}
+## Relativities {#sec:cyclic-calendar}
 
@@
 
+The hierarchical structure of time creates a natural nested ordering which can produce multiple-order-up granularities. We shall use the notion of a hierarchy table to define them.
+
+\begin{definition}\label{def:hierarchy} 
+A {\bf hierarchy table} helps to define cyclic granularities relative to linear granularities.
+The hierarchy table $H_n: (G, C, K)$ arranges $n$ linear granularities from lowest to higher oder and lays down the way they relate through $K$.
+Let $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
+Then $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
+\end{definition}
+
+\noindent We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
+It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.
+
+\begin{definition}\label{def:norderup} 
+Cyclic granularities which are nested within multiple levels of the hierarchy table are called {\bf multiple-order-up} granularities.
+\end{definition}
+
+<!-- \begin{definition}\label{def:norderup}  -->
+
+<!-- \end{definition} -->
+
+A _single-order-up_ is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
+
+## Computation through cyclic calendar algebra
+
 The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple** and (2) **multiple-to-single** \todo[inline]{these don't make sense to me, they are just the reverse of each other, so why are both needed.
-its not  clear from the  next block of text what each operation does} which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
-The hierarchical structure of time creates a natural nested ordering which can produce single-order-up or multiple-order-up granularities.
-We shall use the notion of a hierarchy table and order to define them.
-
-\begin{definition}\label{def:order}
-The {\bf order} of a linear granularity is the level of coarseness associated with a linear granularity.
-For example, for two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
-In any hierarchy table, linear granularities are arranged from lowest to highest order.
-\end{definition}
-
-\begin{definition}\label{def:hierarchy} 
-Let $H_n: (G, C, K)$ be a {\bf hierarchy table} containing $n$ linear granularities.
-$G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
-$K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
-We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
-It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.
-\end{definition}
-
-We refer to granularities which are nested within multiple levels of the hierarchy table as _multiple-order-up_ and those concerning a single level as _single-order-up_ granularities.
+(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
 
@@
 
-## Single-to-multiple {#sec:single-to-multiple}
-
-### All circular single order-up granularities {#sec:all-circular}
+### Single-to-multiple {#sec:single-to-multiple}
+
+#### All circular single order-up granularities {#sec:all-circular}
 
@@
 
-### Circular or quasi-circular single order-up granularities {#sec:notall-circular}
+#### Circular or quasi-circular single order-up granularities {#sec:notall-circular}
 
@@
 
-## Multiple-to-single {#sec:multiple-to-single}
-
-### Multiple order-up circular granularities {#sec:all-circular-multiple}
+### Multiple-to-single {#sec:multiple-to-single}
+
+#### Multiple order-up circular granularities {#sec:all-circular-multiple}
 
@@
 
-### Muliple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
+#### Muliple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
 

### 2020-05-12 main.Rmd
order definition and example fine tuned

@@
 \noindent This implies that the granules in a linear granularity are non-overlapping, continuous and ordering is maintained.
-The indexing for each granule can also be associated with textual representation, called the label.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.
+The indexing for each granule can also be associated with textual representation, called the label.
+A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.
 When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
 Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> \autoref{fig:linear-time} illustrates the linear time granularities.
-Here, "hour" is a linear granularity which is a mapping from index set to hourly time domain.
-Similarly, "day", "week", "month" and "year" are also linear granularities which maps from index set to subsets of the hourly time domain.
+Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities which maps from index set to subsets of the hourly time domain.
 <!--For example, day($0$) =  {hour($0$), hour($1$), hour($2$), \dots, hour($23$)}, day($1$) =  {hour($24$), hour($25$), hour($26$), \dots, hour($47$)} and so on.--> If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
@@
 The {\bf order} of a linear granularity is the level of coarseness associated with a linear granularity.
-For example, for two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$.
+A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
 \end{definition}
 
-A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
+\begin{example}
+{\rm For two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ then $G$ is of lower order than $H$.
+Moreover, if we consider bottom granularity as day, linear granularity week will have lower order than month since each week consist of less number of days than each month.
+Here neither  week {\em groups into} month or  week {\em finer than} month, but their relative order could be determined.}
+\end{example}
+
 <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).

### 2020-05-12 main.Rmd
hierarchy table fixed

@@
 
+<!-- that the aperiodic linear granularities cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.  -->
+
 <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
-Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
+Chronons may be grouped into granules of a bottom granularity and -->
+
+## Computation through calendar algebra
+
+Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
 A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson_2000.
-
-## Computation through calendar algebra
-
 Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning_2002].
@@
 \begin{definition}\label{def:hierarchy} 
-A {\bf hierarchy table} helps to define cyclic granularities relative to linear granularities.
-The hierarchy table $H_n: (G, C, K)$ arranges $n$ linear granularities from lowest to higher oder and lays down the way they relate through $K$.
+The {\bf hierarchy table} $H_n: (G, C, K)$ arranges $n$ periodic linear granularities from lowest to higher order and lays down the way they relate through $K$ so that cyclic granularities $C$ could be defined relative to these linear granularities.
 Let $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
-Then $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.
+Then $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$, if $C_{G_{l}, G_{m}}$ is a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$, if $C_{G_{l}, G_{m}}$ is quasi-circular.
 \end{definition}
 
-\noindent We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
-It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.
-
-\begin{definition}\label{def:norderup} 
-Cyclic granularities which are nested within multiple levels of the hierarchy table are called {\bf multiple-order-up} granularities.
-\end{definition}
-
-<!-- \begin{definition}\label{def:norderup}  -->
-
-<!-- \end{definition} -->
-
-A _single-order-up_ is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
-
-## Computation through cyclic calendar algebra
-
-The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
-
-<!-- Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities and the other granularities are subsequently more "coarse" than the earlier one.
--->
-
-<!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
--->
-
-<!-- some discussions on primary units of calendar -->
-<!-- **Example:**  -->
-
-So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar.
+\begin{example}
+{\rm So far we have used examples of linear granularities from Gregorian calendar as it is the most widely used calendar.
 However, it is far from being the only one.
-<!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
+The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
 Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
@@
 For example, in the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
-\autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.
+\autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.}
+\end{example}
 
@@
   #row_spec(1:(nrow(table_mayan)-1), hline_after = T)
+  #.
 ```
 
-Examples of multiple-order-up granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.
+<!-- for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.  -->
+<!-- so that cyclic granularities $C$ could be defined relative to these linear granularities. -->
+
+\noindent It might not be always possible to position an aperiodic linear granularity in a hierarchy table.
+Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
+
+\begin{definition}\label{def:norderup} 
+Cyclic granularities which are nested within multiple levels of the hierarchy table are called {\bf multiple-order-up} granularities.
+\end{definition}
+
+A _single-order-up_ is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
+ <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->
+
+\begin{example}
+{\rm For \autoref{tab:tab-mayan}, examples of multiple-order-up granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.}
+\end{example}
+
+<!-- \begin{definition}\label{def:norderup}  -->
+
+<!-- \end{definition} -->
+
+## Computation through cyclic calendar algebra
+
+The cyclic calendar algebra consists of broadly two kinds of operations:
+(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
+
+<!-- Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities and the other granularities are subsequently more "coarse" than the earlier one.
+-->
+
+<!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
+-->
+
+<!-- some discussions on primary units of calendar -->
+<!-- **Example:**  -->
 

### 2020-05-12 main.Rmd
revised the circular relativies section

@@
 
+```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
+The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future.
+Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks.
+This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 18.3 weeks will be a quasi-circular granularity with repeating patterns.
+$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years."}
+
+knitr::include_graphics("Figs/aperiodic-ex3.png")
+```
+
 \begin{definition}\label{def:aperiodic}
@@
 
-```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
-The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future.
-Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks.
-This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 18.3 weeks will be a quasi-circular granularity with repeating patterns.
-$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years."}
-
-knitr::include_graphics("Figs/aperiodic-ex3.png")
-```
-
 <!-- #  -->
@@
 
-The hierarchical structure of time creates a natural nested ordering which can produce multiple-order-up granularities. We shall use the notion of a hierarchy table to define them.
+The hierarchical structure of time creates a natural nested ordering  which can be used in the computation of relative pairs of granularities.
+This is illustrated using two examples, a Mayan calendar, which is simple, in addition to the more complex Gregorian calendar.
 
 \begin{definition}\label{def:hierarchy} 
-The {\bf hierarchy table} $H_n: (G, C, K)$ arranges $n$ periodic linear granularities from lowest to higher order and lays down the way they relate through $K$ so that cyclic granularities $C$ could be defined relative to these linear granularities.
+The nested ordering of linear granularities can be organized into a {\bf hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order.
+It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$.
+can be defined relative to the linear granularities.
 Let $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
@@
 \begin{example}
-{\rm So far we have used examples of linear granularities from Gregorian calendar as it is the most widely used calendar.
-However, it is far from being the only one.
-The day is the basic unit of time underlying most calendars [@Reingold2001-kf].
-Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years.
-The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
-Nevertheless, for any calendar a hierarchy table can be defined.
-For example, in the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun.
-\autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.}
+{\rm  \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar.
+In the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun.
+}
 \end{example}
@@
 
+ <!--So far we have used examples of linear granularities from Gregorian calendar as it is the most widely used calendar. However, it is far from being the only one.-->
+\noindent It is interesting to note, that in the Mayan calendar, like most calendars, that day is the basic unit of time [@Reingold2001-kf].
+The structuring of larger units,  weeks, months, years and cycle of years, though, varies substantially.
+An exception is the French revolutionary calendar, which divides each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
+Nevertheless, for any calendar a hierarchy table can be defined.
+
 <!-- for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.  -->
@@
 
-\noindent It might not be always possible to position an aperiodic linear granularity in a hierarchy table.
+Note that, it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
 Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
@@
 \begin{definition}\label{def:norderup} 
-Cyclic granularities which are nested within multiple levels of the hierarchy table are called {\bf multiple-order-up} granularities.
+The hierarchy table contains {\bf multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
+A {\bf single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
 \end{definition}
 
-A _single-order-up_ is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
  <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->
@@
 \begin{example}
-{\rm For \autoref{tab:tab-mayan}, examples of multiple-order-up granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.}
+{\rm In \autoref{tab:tab-mayan}, kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.}
 \end{example}

### 2020-05-13 main.Rmd
Section 3.5 changes made - no definition-example framework and examples for every operation need not be given

@@
 
-#### All circular single order-up granularities {#sec:all-circular}
-
-Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity.
-Since, the operation requires the use of modular arithmetic, it is important that the label mapping of the individual circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
+<!-- #### All circular single order-up granularities {#sec:all-circular}   -->
+
+<!-- Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity. -->
+
+\noindent Methods to obtain multiple-order-up granularity will be different based on if the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
+The methods require the use of integer arithmetic, and hence it is important that the label mapping of the individual single-order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
 The label mapping of the resultant multiple-order-up granularity can be chosen depending on the context.
+Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using Equation \ref{eq:eq7} for $l < m - 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$.
 
@@
 
-, for $l < m - 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$
-
-\begin{example} 
-{\rm Let us use Equation \ref{eq:eq7} to compute the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar with the index set  $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ mapping "kins" to a set of positive integers.
-
-\begin{equation} \label{eq8}
+\noindent For example, the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar <!--with the index set  $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ mapping "kins" to a set of positive integers--> could be obtained using Equation \ref{eq:eq8}.
+
+\begin{equation} \label{eq:eq8}
 \begin{split}
@@
 \end{equation}
-}
-\end{example}
-
-#### Circular or quasi-circular single order-up granularities {#sec:notall-circular}
+
+<!-- #### Circular or quasi-circular single order-up granularities {#sec:notall-circular}   -->
 
@@
 
-We consider the case of only one quasi-circular single order-up granularity in the hierarchy table. Any multiple-order-up quasi-circular granularity $C_{l, m}(z)$ can then be obtained as a
-dicrete combination of circular and quasi-circular granularities.
+\noindent We consider the case of only one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
+Any multiple-order-up quasi-circular granularity $C_{l, m}(z)$ could then be obtained as a dicrete combination of circular and quasi-circular granularities.
 <!--function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
---> Depending on the order of combination, two different approaches need to be employed leading to the following cases:
+--> Depending on the order of the combination, again two different approaches need to be employed leading to the following cases:
 
@@
 
-where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+\noindent where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 
@@
 
-\begin{example} 
-{\rm Suppose we have a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
-Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.
-This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities.
-Only one single-order-up granularity day-of-month is quasi-circular.
-
-Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we have:
+\noindent Consider a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
+<!--Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.--> This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities with day-of-month as the only single-order-up quasi-circular granularity.
+
+\noindent Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we then have:
 
@@
  $$C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert)$$, where $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$
-
-}
-\end{example}
 
@@
 
-#### Multiple order-up circular granularities {#sec:all-circular-multiple}
-
-For a hierarchy table $H_n: (G, C, K)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have
+<!-- #### Multiple order-up circular granularities {#sec:all-circular-multiple}   -->
+
+\noindent Similar to single-to-multiple operations, multiple-to-single operations also involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
+For a hierarchy table $H_n: (G, C, K)$ with only  circular single-order-up granularities in the hierarchy and $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, multiple-order-up granularities could be obtained using \ref{eq:all-circular-multiple}.
 
@@
 
-\begin{example}
-{\rm Considering the same example of Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun given the multiple-order-up granularity uinal-baktun using equation \ref{eq:all-circular-multiple}.
-
-\begin{equation} \label{eq10}
-\begin{split}
-C_{tun, katun}(z) =  \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20
-\end{split}
-\end{equation}
-}
-\end{example}
+\noindent For example, in Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) =  \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.
 

### 2020-05-14 main.Rmd
spell check

@@
 A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.--> \autoref{fig:linear-time} illustrates the linear time granularities.
+When the label set is exactly the integers, the granularities are called "full-integer" labelled granularity, such as in Definition \ref{def:linear}.
+Some calendar operations will only work on full-integer labelled granularities, while others can be implemented using more flexible labelings.--> \autoref{fig:linear-time} illustrates the linear time granularities.
 Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities which maps from index set to subsets of the hourly time domain.
@@
 For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$.
-The _altering-tick operation_ generates new full-integer labeled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$.
+The _altering-tick operation_ generates new full-integer labelled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$.
 For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
@@
 <!-- CONTENT CUT -->
-<!-- The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
-A labeled granularity is a pair $(L, G)$, where the label set $L$ is a subset of the integers instead of the set of integers and -->
+<!-- The notion of a labelled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
+A labelled granularity is a pair $(L, G)$, where the label set $L$ is a subset of the integers instead of the set of integers and -->
 <!-- follows the conditions (1) and (2) in \ref{def:linear} where $i$, $j$, and $k$ are in $L$.
-When the label set $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->
+When the label set $L$ is exactly the integers, the granularities are called "full-integer" labelled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->
 
@@
 \begin{definition}\label{def:grouping_operation}
-Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer.
+Let $G_1$ be a full-integer labelled granularity, and $m$ a positive integer.
 {\bf The grouping operation} Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
+More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labelled granularity such that for each integer $i$,
 $$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
@@
 <!-- \begin{definition}\label{altering-tick} -->
-<!-- Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
+<!-- Let $G_1$, $G_2$ be full-integer labelled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
 {\bf The altering-tick operation} $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
@@
 <!-- The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.
+When the label set is exactly the integers, the granularities are called "full-integer" labelled granularity, such as in Definition \ref{def:linear}.
+Some calendar operations will only work on full-integer labelled granularities, while others can be implemented using more flexible labelings.
 -->
@@
 
-#### Muliple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
+#### Multiple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
 

### 2020-05-14 algo_rank_harmony.Rmd
data structure intro changed

@@
 library(readr)
-read_rds("data/smart_harmony_weibull.rds") %>% knitr::kable()
+read_rds("data/smart_harmony_wb.rds") %>% knitr::kable()
 ```
@@
 ```{r}
-read_rds("data/smart_harmony_gamma.rds") %>% knitr::kable()
+read_rds("data/smart_harmony_g.rds") %>% knitr::kable()
 ```
@@
 library(readr)
-read_rds("data/smart_harmony_normal.rds") %>% knitr::kable()
+read_rds("data/smart_harmony_norm.rds") %>% knitr::kable()
 ```
@@
 
-# Cricket data
-
-## Distances follow Weibull
-```{r}
-read_rds("data/smart_cric_weibul.rds")%>% knitr::kable()
-```
-
-## Distances follow Gamma
-
-```{r}
-library(readr)
-read_rds("data/smart_cric_gamma.rds") %>% knitr::kable()
-```
-
-## Distances follow Normal 
-```{r}
-read_rds("data/smart_cric_normal.rds")%>% knitr::kable()
-```
-
-# Old benchmarks
-
-## Version 1 (suggested by Rob)
+## Version 1 (General - suggested by Rob)
 
@@
 
-```{r, eval=FALSE, echo=FALSE}
-read_rds("data/smart_harmony_ak.rds")
+```{r}
+read_rds("data/smart_harmony_g.rds") %>% knitr::kable()
 ```
 
-```{r, eval = FALSE}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("day_week",
-    "day_month",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    symmetric = TRUE,
-    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
+<!-- ```{r, eval=FALSE, echo=FALSE} -->
+<!-- read_rds("data/smart_harmony_ak.rds") -->
+<!-- ``` -->
+
+<!-- ```{r, eval = FALSE} -->
+<!-- smart_meter10 %>% -->
+<!--   filter(customer_id %in% c(10017936)) %>% -->
+<!--   prob_plot("day_week", -->
+<!--     "day_month", -->
+<!--     response = "general_supply_kwh", -->
+<!--     plot_type = "quantile", -->
+<!--     symmetric = TRUE, -->
+<!--     quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99) -->
+<!--   ) + -->
+<!--   scale_y_sqrt() + -->
+<!--   #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") + -->
+<!--   ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+  -->
+<!--   theme_minimal() -->
+<!-- ``` -->
+
+<!-- ## Version 4 (old benchmark) -->
+
+<!--  3.4.
+For each $i \in \{1, 2, \dots, I\}$, the maximum pairwise distance could be chosen from  $\binom{I}{2}$ distances using $D_{i}$ = max($d_{jl}$)/$J$(max($d_{jl}$) - min($d_{jl}$)) $\forall j, l = {1, 2, \dots, J}$.
+-->
+
+<!--  4.4. Mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, I}$ is computed by -->
+<!-- $MMPD = (1/I)\sum_{i=1}^{l}{D_{i}}$. -->
+
+<!-- ```{r, eval = FALSE, echo=FALSE} -->
+<!-- library(readr) -->
+<!-- read_rds("data/smart_harmony_n.rds") %>% knitr::kable() -->
+<!-- ``` -->
+
+<!-- ```{r, echo = FALSE, eval = FALSE} -->
+<!-- smart_meter10 %>% -->
+<!--   filter(customer_id %in% c(10017936)) %>% -->
+<!--   prob_plot("week_month", -->
+<!--     "wknd_wday", -->
+<!--     response = "general_supply_kwh", -->
+<!--     plot_type = "quantile", -->
+<!--     symmetric = TRUE, -->
+<!--     quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99) -->
+<!--   ) + -->
+<!--   scale_y_sqrt() + -->
+<!--   #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") + -->
+<!--   ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+  -->
+<!--   theme_minimal() -->
+<!-- ``` -->
+
+# Cricket data
+
+## Distances follow Weibull
+```{r}
+read_rds("data/ smart_cric_weibull.rds")%>% knitr::kable()
 ```
 
-## Version 4 (old benchmark)
+## Distances follow Gamma
 
- 3.4.
-For each $i \in \{1, 2, \dots, I\}$, the maximum pairwise distance could be chosen from  $\binom{I}{2}$ distances using $D_{i}$ = max($d_{jl}$)/$J$(max($d_{jl}$) - min($d_{jl}$)) $\forall j, l = {1, 2, \dots, J}$.
-
- 4.4. Mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, I}$ is computed by
-$MMPD = (1/I)\sum_{i=1}^{l}{D_{i}}$.
-
-```{r, eval = FALSE, echo=FALSE}
+```{r}
 library(readr)
-read_rds("data/smart_harmony_n.rds") %>% knitr::kable()
+read_rds("data/ smart_cric_gamma.rds") %>% knitr::kable()
 ```
 
-```{r, echo = FALSE, eval = FALSE}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("week_month",
-    "wknd_wday",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    symmetric = TRUE,
-    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
+## Distances follow Normal 
+```{r}
+read_rds("data/ smart_cric_normal.rds")%>% knitr::kable()
 ```
+
+## Distances follow general 
+```{r}
+read_rds("data/ smart_cric_general.rds")%>% knitr::kable()
+```

### 2020-05-14 main.Rmd
data structure intro changed

@@
 
-Effective exploration and visualization require proper data structures.
+<!-- # Why data structure for cyclic granularities exploration -->
+
+\noindent Effective exploration and visualization require proper data structures.
 It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
 Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.
-A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
+
+<!-- #recent data structures for temporal exploration -->
+
+\noindent A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble] data structure.
 A tsibble consists of an index, key and measured variables.
@@
 A bottom granularity is represented by the index of the tsibble.
-Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
+
+<!--# our proposal of extension of tsibble data strcuture and -->
+<!--# why looking at pair of granularities instead of all -->
+
+\noindent All cyclic granularities can be expressed in terms of the index set, and hence, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
 This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$ and $v$ denotes the measurement variable.
@@
 Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
-Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
-Instead, this big problem is broken down by focusing on two cyclic granularities at a time.
+Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic granularities.
+Any attempt to encode all or many of these cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
+Instead, this big problem is broken down by focusing on pair of cyclic granularities at a time.
 Hence data sets of the form <$C_i$, $C_j$, $v$> forms the basis for exploration and analysis of the measured variable.
@@
 
-Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the number of combinations quickly becomes too large and thus, it is beneficial useful to rank these harmonies.
+<!-- Why do we need ranking system -->
+<!-- set of pairs - building the ranks for N_C and N_L -->
+
+Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the number of combinations quickly becomes too large and thus, it is beneficial to rank these harmonies.
 Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
@@
 Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+
+<!-- Calculating distances and lit review -->
+
+<!-- Normalizing -->
+
  Distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
@@
 ## Summarizing the measured variable {#sec:summarize-measured}
+
+<!-- Why we looking at probability measure -->
+
+<!-- T+1 observation not T -->
 

### 2020-05-14 algo_validation.Rmd
- compared norming constants for similar levels
- checked with other papers regarding norming constants

---
title: "Untitled"
author: "Sayani Gupta"
date: "14/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(gravitas)
library(tidyverse)
```

```{r}
cricket_data <- read_rds("data/cricket_data.rds")
cricket_data %>% hist_distance(gran1 = "lag_field", gran2 = "over", response = "runs_per_over", hierarchy_tbl = hierarchy_model)
```

```{r}

  hierarchy_model <- tibble::tibble(
      units = c("index", "over", "inning", "match"),
      convert_fct = c(1, 20, 2, 1))

.data <- cricket_data

hierarchy_tbl <- hierarchy_model

response <- "runs_per_over"

harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)

smart_cric_normal <- cricket_data %>% 
  rank_harmony(harmony_tbl = harmonies_cric, 
               response = response,
               hierarchy_tbl = hierarchy_model,
               dist_distribution = "normal")

smart_cric_gamma <- cricket_data %>% 
  rank_harmony(harmony_tbl = harmonies_cric, 
               response = response,
               hierarchy_tbl = hierarchy_model,
               dist_distribution = "gamma")

smart_cric_weibull <- cricket_data %>% 
  rank_harmony(harmony_tbl = harmonies_cric, 
               response = response,
               hierarchy_tbl = hierarchy_model,
               dist_distribution = "weibull")
```

Now the pair (over, lag_field) and (over_inning_match) has the same levels (20,2). Hence we can see if the norming constants are working properly using these as example.

```{r}
cricket_data %>% 
  prob_plot("over",
            "lag_field",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

```{r}
cricket_data %>% 
  prob_plot("over",
            "inning_match",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

```{r}
cricket_data %>% 
  prob_plot("lag_field",
            "over_match",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

```{r}
cricket_data %>% 
  prob_plot("inning_match",
            "over_inning",
            plot_type = "boxplot",
            response = response,
            hierarchy_tbl = hierarchy_model)
```

### 2020-05-19 main.Rmd
subheader changed to computation only

@@
 
-## Computation through calendar algebra
+## Computation
 
@@
 
-## Computation through cyclic calendar algebra
-
+## Computation
+
+In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
+This is similar to calendar algebra in @Ning2002-tf for linear granularities but caters to computation of cyclic granularities, we shall refer to it as through cyclic calendar algebra operations.
 The cyclic calendar algebra consists of broadly two kinds of operations:
@@
 \noindent All cyclic granularities can be expressed in terms of the index set, and hence, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$ and $v$ denotes the measurement variable.
+This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$.
 The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities.
-Suppose we have $n$ linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
-Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic granularities.
+Suppose we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
+Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.
 Any attempt to encode all or many of these cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
-Instead, this big problem is broken down by focusing on pair of cyclic granularities at a time.
-Hence data sets of the form <$C_i$, $C_j$, $v$> forms the basis for exploration and analysis of the measured variable.
+Instead, this big problem is broken down by focusing on pair of cyclic granularities ($C_i$, $C_j$) at a time and data sets of the form <$C_i$, $C_j$, $v$> forms the basis for exploration and analysis of the measured variable $v$.
 
@@
 ```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
-include_graphics("Figs/data-structure-v2.png")
+include_graphics("Figs/data-struc-rev.png")
 ```
@@
 The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
-Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-Here, $A_i$'s or $B_j$'s are the levels/categories corresponding to $C_1$ and $C_2$ respectively.
-Let $S_{ij}$ be a subset of the index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$.
-Data subsets for each combination of levels ($A_i$, $B_j$) like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, l}$ and $j \in {1, 2, \dots, m}$  which will lead to $lm$ data subsets.
+Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_J$\}.
+Here, $A_k$'s or $B_l$'s are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
+Let $S_{kl}$ be a subset of the index set such that for all $s \in S_{kl}$, $C_i(s) = A_k$ and $C_j(s) = B_l$.
+Data subsets for each combination of levels ($A_k$, $B_l$) like <$A_k$, $B_l$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, l}$ and $j \in {1, 2, \dots, m}$  which will lead to $lm$ data subsets.
 Now, some situations can lead to few or many of these sets being empty.
@@
 These are called "structurally" empty combinations.
-Let us take a specific example, where $C_1$ be day-of-month with 31 levels and $C_2$ be week-of-month with 5 levels.
-There will be $31\times 5=155$ sets $S_{ij}$ corresponding to possible combinations of $C_1$ and $C_2$.
+Let us take a specific example, where $C_i$ be day-of-month with 31 levels and $C_j$ be week-of-month with 5 levels.
+There will be $31\times 5=155$ sets $S_{kl}$ corresponding to possible combinations of $C_i$ and $C_j$.
 Many of these like $S_{1,5}$, $S_{21,2}$ are empty.
@@
 
-<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->
+<!--  In fact, most of these 155 sets will be empty, making the combination of $C_i$ and $C_j$ in a graph unhelpful. -->
 
@@
 These are called "event-driven" empty combinations.
-Let us consider $C_1$ be day-of-week with 7 levels and $C_2$ be WorkingDay/NonWorkingDay with 2 levels.
-While potentially all of these 14 sets $S_{ij}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
+Let us consider $C_i$ be day-of-week with 7 levels and $C_j$ be WorkingDay/NonWorkingDay with 2 levels.
+While potentially all of these 14 sets $S_{kl}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
 For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
@@
 <!-- However, this is different from the first example because these are not structurally empty.
-Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting.
+Also, the graphic may still be useful if we want to look at whether the interaction of $C_i$ and $C_j$ affects the distribution we are plotting.
 -->
@@
 These are called "build-based" empty combinations.
-Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ be day-of-month.
+Let $C_i$ be Business-days, which are days from Monday to Friday except holidays and $C_j$ be day-of-month.
 Then the days denoting weekends in a month would not correspond to any Business days.
@@
 
-An example when there will be no empty combinations could be where $C_1$ and $C_2$ maps index set to day-of-week and month-of-year respectively.
-Here $C_1$ can have 7 levels while $C_2$ can have 12 levels.
-So there are $12\times7=84$ sets $S_{ij}$.
+An example when there will be no empty combinations could be where $C_i$ and $C_j$ maps index set to day-of-week and month-of-year respectively.
+Here $C_i$ can have 7 levels while $C_j$ can have 12 levels.
+So there are $12\times7=84$ sets $S_{kl}$.
 All of these are non-empty because every day-of-week can occur in every month.
@@
 
-Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the number of combinations quickly becomes too large and thus, it is beneficial to rank these harmonies.
-Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granulairites.
+Hence, all possible pairs of cyclic time granularities that could be analyzed together is $N_C \choose 2$.
+Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption.
+Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.
+The harmony pair $(C_i, C_j)$ would be most interesting if the average variation in the measured variable between different levels of $C_i$ and $C_j$ is maximum.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
 The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
-Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-
-<!-- Calculating distances and lit review -->
-
-<!-- Normalizing -->
-
- Distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
+-->
+
+<!-- Calculating distances between distributions -->
+
+\noindent One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable across different levels of $C_j$ and integrating it for all levels of $C_i$.
+Distances or similarity measures have been extensively studied and applied to fields like clustering, classification and anomaly detection.
+We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
+One of the common ways to measure distances between distributions is through Jenson-Shanon divergence which is based on Kullback-Leibler divergence.
+In our work, probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+Suppose $q$ is a sample quantile vector of the measured variable computed for probability vector $p$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
+ $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
+
+<!-- Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth. -->
+
+<!-- # distribution of distances -->
+<!-- # Normalizing and lit review -->
+
+\noindent Since we are looking to rank the harmony pairs based on if it captures the maximum variation across levels of $C_i$ and $C_j$, we need to make sure that this maximum value is independent of the number of levels of the two cyclic granularities.
+A major result in the field of EVT is the probability distribution of the maximum can only be one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+But the normalising constants vary depending on the underlying distribution of the measured variables.
+Hence to normalize appropriately, it is important to assume a distribution of distances in our case.
+
+<!-- For ranking the harmony pairs, these distances need to be normalised by the number of levels of . -->
+
+<!-- Distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
+-->
 
@@
 Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_i$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
+Hence for each combination ($A_k$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
 
 2.
-There will be $m$ time series variable corresponding to each level $A_i$ of $C_1$.
+There will be $m$ time series variable corresponding to each level $A_k$ of $C_1$.
 We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_i$.
-
-3.
-Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
-Suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
- $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
+Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_k$.
+
+3. Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+
+<!-- Suppose $q$ is the sample quantile vector computed for probability vector $p$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where, -->
+<!--  $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$. -->
 
@@
 
-6.
-Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
-Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.
+6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
+
+<!-- with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$. -->
 
@@
 Layers are usually created with a geom function.
-Considering the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
-<!-- - Data: <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
-<!-- - Aesthetic mapping: $A_i$ mapped to $x$ position and $v(s)$ to $y$ position. -->
+Considering the data structure <$A_k$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
+<!-- - Data: <$A_k$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
+<!-- - Aesthetic mapping: $A_k$ mapped to $x$ position and $v(s)$ to $y$ position. -->
 <!-- - Layer: any geom that summarizes distribution -->
@@
 
-<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_i$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
+<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_k$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
 The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
@@
 We will discuss the effect of the levels of cyclic granularities on the choice of plots <!--Recall that for cyclic granularities $C_1$ and $C_2$ such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-$A_i$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.The levels have an impact on the--> since space and resolution might become a problem if the number of levels are too high.
+$A_k$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.The levels have an impact on the--> since space and resolution might become a problem if the number of levels are too high.
 A potential approach could be to categorize the levels as very high/high/medium/low for each cyclic granularity and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.

### 2020-05-19 main.Rmd
harmony clash notation changed

@@
 \noindent All cyclic granularities can be expressed in terms of the index set, and hence, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities $C_i$.
-The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities.
-Suppose we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
-Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.
+This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities.
+The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
+Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.
 Any attempt to encode all or many of these cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
-Instead, this big problem is broken down by focusing on pair of cyclic granularities ($C_i$, $C_j$) at a time and data sets of the form <$C_i$, $C_j$, $v$> forms the basis for exploration and analysis of the measured variable $v$.
+Instead, this big problem is broken down by focusing on pair of cyclic granularities ($C_i$, $C_j$) at a time for $i, j \in \{1, 2, \dots, N_C\}$.
+Data sets of the form <$C_i$, $C_j$, $v$> then forms the basis for exploration and analysis of the measured variable $v$.
 
@@
 The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
-Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_J$\}.
+Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k | k \in \mathbb{N}, k \leq K\}$ and $C_j$ maps index set to a set $\{B_l | l \in \mathbb{N}, l \leq L\}$.
 Here, $A_k$'s or $B_l$'s are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
 Let $S_{kl}$ be a subset of the index set such that for all $s \in S_{kl}$, $C_i(s) = A_k$ and $C_j(s) = B_l$.
-Data subsets for each combination of levels ($A_k$, $B_l$) like <$A_k$, $B_l$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, l}$ and $j \in {1, 2, \dots, m}$  which will lead to $lm$ data subsets.
+Data subsets for each combination of levels ($A_k$, $B_l$) like <$A_k$, $B_l$, $v(s)$> can be obtained for all $k \in {1, 2, \dots, K}$ and $l \in {1, 2, \dots, L}$  which will lead to $KL$ data subsets.
 Now, some situations can lead to few or many of these sets being empty.
-We will discuss few cases, where one or more of these $lm$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
+We will discuss few cases, where one or more of these $KL$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
 <!--The implication of these empty sets on visualization will be discussed in Section \ref{sec:vis-interaction}.
@@
 
-Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
+\noindent Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
 These are called "structurally" empty combinations.
@@
 
-Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
+\noindent Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
 These are called "event-driven" empty combinations.
@@
 
-Thirdly, empty combinations can be a result of how granularities are constructed.
+\noindent Thirdly, empty combinations can be a result of how granularities are constructed.
 These are called "build-based" empty combinations.
@@
 
-An example when there will be no empty combinations could be where $C_i$ and $C_j$ maps index set to day-of-week and month-of-year respectively.
+\noindent An example when there will be no empty combinations could be where $C_i$ and $C_j$ maps index set to day-of-week and month-of-year respectively.
 Here $C_i$ can have 7 levels while $C_j$ can have 12 levels.

### 2020-05-21 main.Rmd
data structure algorithm made short and include normalization section

@@
 
-<!--# our proposal of extension of tsibble data strcuture and -->
+<!--# our proposal of extension of tsibble data structure and -->
 <!--# why looking at pair of granularities instead of all -->
@@
 
-\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granulairites.
+\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities.
 Hence, all possible pairs of cyclic time granularities that could be analyzed together is $N_C \choose 2$.
@@
 Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.
-The harmony pair $(C_i, C_j)$ would be most interesting if the average variation in the measured variable between different levels of $C_i$ and $C_j$ is maximum.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
@@
 
-\noindent One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable across different levels of $C_j$ and integrating it for all levels of $C_i$.
+\noindent One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable.
+<!--across different levels of $C_j$ and integrating it for all levels of $C_i$.
 Distances or similarity measures have been extensively studied and applied to fields like clustering, classification and anomaly detection.
-We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
+---> We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
 One of the common ways to measure distances between distributions is through Jenson-Shanon divergence which is based on Kullback-Leibler divergence.
-In our work, probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-Suppose $q$ is a sample quantile vector of the measured variable computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where,
- $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$.
+Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+The Jensen-Shanon distance between two sample quantile vectors $q_u$ and $q_v$ computed for probability vector $p$ will be given by $$d_{uv} = [D(q_u, r) + D(q_v, r)]/2 \quad where \quad r = (q_u + q_v)/2$$ where,
+ $$D(q_u,q_v) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_u$ and $q_v$.
 
@@
 
-<!-- # distribution of distances -->
-<!-- # Normalizing and lit review -->
-
-\noindent Since we are looking to rank the harmony pairs based on if it captures the maximum variation across levels of $C_i$ and $C_j$, we need to make sure that this maximum value is independent of the number of levels of the two cyclic granularities.
-A major result in the field of EVT is the probability distribution of the maximum can only be one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
-But the normalising constants vary depending on the underlying distribution of the measured variables.
+<!-- # why Normalizing required --->
+
+\noindent The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the cyclic granularities.
+But maximum is not robust to the number of levels and seems to be higher for harmonies with higher levels.
+Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
+\noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+Suppose $X_{1},X_{2}\ldots ,X_{n}$  be a sequence of independent and identically-distributed random variables, and $M_{n}=\max\{X_{1},\ldots ,X_{n}\}$.
+If a sequence of pairs of real numbers $a_{n}>0$ and $\lim _{{n\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
+The normalizing constants, however, vary depending on the underlying distribution of $X_i$'s.
 Hence to normalize appropriately, it is important to assume a distribution of distances in our case.
 
-<!-- For ranking the harmony pairs, these distances need to be normalised by the number of levels of . -->
-
-<!-- Distances between these probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
--->
+<!-- # distribution of distances and lit review -->
+
+\noindent Menendez1997-in and Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
+However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.
+
+\noindent The algorithm employed addressing all the issues decried above is summarized as follows:
+<!-- Algorithm -->
+
+<!-- 1.
+Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
+Hence for each combination ($A_k$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$ -->
+
+\noindent Algorithm
+**Input** $(C_i, C_j, v) \quad \forall i, j \in N_C$  
+**Output** MMPD (Median Maximum Pairwise Distances) measuring the average variation across different levels of $C_i$ and $C_j$ $\forall i, j \in N_C$
 
 1.
-Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_k$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$
+For each pair of cyclic granularity $(C_i, C_j)$, there will be $L$ groups corresponding to each level $A_k$ of $C_i$.
+Thus there will be  $m = \binom{L}{2}$ pairwise distances corresponding to each $A_k$.
 
 2.
+Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+Let $d_1, d_2, d_3, \dots d_m$ be the pairwise Jensen-Shannon distances.
+Suppose the maximum pairwise distance for $A_k$ maximum be $M_k$.
+
+3.
+The distribution of distances is heavy tailed.
+But if they are gamma or weibull could be judged by looking at the QQ-plots of the empirical distribution.
+The estimates of the parameters are obtained using Maximum Likelihood Estimation.
+
+4.
+The norming constants $(a_k, b_k)$ are obtained from @Embrechts2013-pd and each $M_k$ is normalized using normalized maximum $NM_k = \frac{M_k - b_k}{a_k}$ for each $A_k$, $k \in {1, 2, \dots, K}$.
+We have normalized the number of levels of $C_j$ using Steps 1-4.
+
+<!-- the norming constants, which are a function of the parameters are obtained using method of moments. -->
+
+<!-- Since the distances follow a Chi-squared distribution with , the norming constants are taken as $a_n = $ and $b_n = $ -->
+
+5.
+We also need to aggregate these normalized maximum across different levels of $C_i$.
+For this we can take the mean, median or any quantiles of these normalized maximums.
+The median (and any quantiles) of these maximum distributions (which converges to a Gumbel distribution) grow at the rate of (some power of) $log K$ with $K \xrightarrow{} \infty$.
+Hence, M/log($K$) could serve as the final measure (MMPD), where $M = median\{NM_1, NM_2, \dots, NM_K\}$.
+
+6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
+
+<!-- 2.
 There will be $m$ time series variable corresponding to each level $A_k$ of $C_1$.
@@
 Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_k$.
-
-3. Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+-->
 
@@
 
-4.
+<!-- 4.
 For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
 Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
-
-5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by
-$MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$.
-
-6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
+-->
+
+<!-- 5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by -->
+<!-- $MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$. -->
+
+<!-- 6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD. -->
 

### 2020-05-21 algo_rank_harmony.Rmd
updated united table

@@
 library(readr)
-read_rds("data/smart_harmony_wb.rds") %>% knitr::kable()
+smart_weibull <- read_rds("data/smart_harmony_wb.rds")  %>% mutate(rank_weibull = row_number())
 ```
@@
 ```{r}
-read_rds("data/smart_harmony_g.rds") %>% knitr::kable()
+smart_gamma <- read_rds("data/smart_gamma.rds")  %>% mutate(rank_gamma = row_number())
 ```
@@
 library(readr)
-read_rds("data/smart_harmony_norm.rds") %>% knitr::kable()
+smart_normal <- read_rds("data/smart_harmony_norm.rds") %>% mutate(rank_normal = row_number())
 ```
@@
 ```{r}
-read_rds("data/smart_harmony_g.rds") %>% knitr::kable()
+smart_general <- read_rds("data/smart_general.rds") %>% mutate(rank_general = row_number())
+
+data_unite <- smart_normal %>% 
+  left_join(smart_gamma, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%
+  rename("MMPD_normal" = "mean_max_variation.x", "MMPD_gamma" = "mean_max_variation.y") %>% 
+    left_join(smart_weibull,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  rename("MMPD_weibull" = "mean_max_variation") %>% 
+  left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  rename("MMPD_general" = "mean_max_variation") %>% 
+  select(-c(facet_levels, x_levels))
+
+data_unite %>% select(-c(MMPD_normal, MMPD_gamma, MMPD_weibull, MMPD_general))
+data_unite %>% select(-c(rank_normal, rank_gamma, rank_weibull, rank_general))
+
 ```

### 2020-05-27 tikz_diagram.R
clean folder

@@ -1,76 +1,0 @@
-\documentclass{minimal}
-\usepackage{tikz}
-\usetikzlibrary{calc,trees,positioning,arrows,chains,shapes.geometric,%
-  decorations.pathreplacing,decorations.pathmorphing,shapes,%
-  matrix,shapes.symbols}
-
-\tikzset{
-  >=stealth',
-  punktchain/.style={
-    rectangle, 
-    rounded corners, 
-    % fill=black!10,
-    draw=black, very thick,
-    text width=10em, 
-    minimum height=3em, 
-    text centered, 
-    on chain},
-  line/.style={draw, thick, <-},
-  element/.style={
-    tape,
-    top color=white,
-    bottom color=blue!50!black!60!,
-    minimum width=8em,
-    draw=blue!40!black!90, very thick,
-    text width=10em, 
-    minimum height=3.5em, 
-    text centered, 
-    on chain},
-  every join/.style={->, thick,shorten >=1pt},
-  decoration={brace},
-  tuborg/.style={decorate},
-  tubnode/.style={midway, right=2pt},
-}
-\begin{document}
-\begin{tikzpicture}
-  [node distance=.8cm,
-  start chain=going below,]
-     \node[punktchain, join] (intro) {Introduktion};
-     \node[punktchain, join] (probf)      {Problemformulering};
-     \node[punktchain, join] (investeringer)      {Investeringsteori};
-     \node[punktchain, join] (perfekt) {Det perfekte kapitalmarked};
-     \node[punktchain, join, ] (emperi) {Emperi};
-      \node (asym) [punktchain ]  {Asymmetrisk information};
-      \begin{scope}[start branch=venstre,
-        %We need to redefine the join-style to have the -> turn out right
-        every join/.style={->, thick, shorten <=1pt}, ]
-        \node[punktchain, on chain=going left, join=by {<-}]
-            (risiko) {Risiko og gamble};
-      \end{scope}
-      \begin{scope}[start branch=hoejre,]
-      \node (finans) [punktchain, on chain=going right] {Det finansielle system};
-    \end{scope}
-  \node[punktchain, join,] (disk) {Det imperfekte finansielle marked};
-  \node[punktchain, join,] (makro) {Investeringsmæssige konsekvenser};
-  \node[punktchain, join] (konk) {Konklusion};
[…]

### 2020-05-27 choosing_harmonies.Rmd
choosing harmonies assimilated doc

---
title: Screening harmonies
authors:
- name: Sayani Gupta
  affiliation: Department of Econometrics and Business Statistics, Monash University
  email: Sayani.Gupta@monash.edu

bibliography: bibliography.bib
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(viridis)
library(tsibble)
library(gravitas)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = FALSE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

```{r external, include = FALSE}
# read_chunk('scripts/main.R')
```

```{r load}

```

# Idea

Even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption. Hence, there is a need to rank the  harmonies basis how well they capture the variation in the measured variable and additionally reduce the number of harmonies for further exploration/visualization. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data. Thus the idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.

# Computing distances

One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable. We do this through Jensen-Shannon divergence which is based on Kullback-Leibler divergence. Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.

 We shall call this measure of variation as  Median Maximum Pairwise Distances (MMPD)

# Normalize distances

The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the harmonies. But maximum is not robust to the number of levels and is higher for harmonies with higher levels. Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels. The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process. The normalizing constants, however, vary depending on the underlying distribution and hence it is important to assume a distribution of distances in our case.

## Theoretical evidence

@Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution. Let $p^{(1)} \equiv (p_1^{(1)}, p_2^{(1)}, \dots, p_k^{(1)})$ and $p^{(2)} \equiv (p_1^{(2)}, p_1^{(2)}, \dots, p_k^{(2)})$ denote two probability distributions, where $k$ is the number of components of the probability vector p. Then 2N(ln2)D is known to converge to {$\chi^2$} distribution with degrees of freedom = (k - 1) and

$D[p^{(1)}, p^{(2)}] = H((p^{(1)} + p^{(2)})/2)  - H(p^{(1)}) - H(p^{(2)})$, $H_{(p)} = \sum_{i = 1}^{k} p_ilog_2p_i$ and N is the total number of events.

## Empirical evidence

However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles. The QQ plots of four harmony pairs are plotted below. It could be seen that Chi-square distribution serves as a pretty good fit to the data (specially in the extreme right tail).

```{r}
library(tsibbledata)
library(ggplot2)
library(tsibble)
library(lvplot)
library(dplyr)
library(gravitas)
library(purrr)
library(magrittr)
sm <- smart_meter10 %>%
filter(customer_id %in% c(10017936))
.data = sm
gran1 = "wknd_wday"
gran2 = "hour_day"
response  = "general_supply_kwh"
```

### Distribution fitting of distances for the harmony pair (weekend/weekday, hour-of-day)

```{r}
.data %>% qqplot_distance(gran1 = "wknd_wday", gran2 = "hour_day", response = "general_supply_kwh")
```

### Distribution fitting of distances for the harmony pair (day-of-week, hour-of-day)

```{r}
.data %>% qqplot_distance(gran1 = "day_week", gran2 = "hour_day", response = "general_supply_kwh")
```

### Distribution fitting of distances for the harmony pair (hour-of-day, week-of-month)

```{r}
.data %>% qqplot_distance(gran1 = "hour_day", gran2 = "week_month", response = "general_supply_kwh")
```

### Distribution fitting of distances for the harmony pair (day-of-month, day-of-week)

```{r}
.data %>% qqplot_distance(gran1 = "day_month", gran2 = "day_week", response = "general_supply_kwh")
```

# Estimating parameters

Currently, MME is used. But MASS::fitdistr() and package fitdistrplus also provide methods to estimate paramters through MLE. WIP.

# Choose thresholds for harmonies

WIP

# Results

##  Smart meter data

### Maximum distance between levels of x-axis variable and median across levels of facet variable

```{r}
library(readr)
smart_weibull <- read_rds("data/reverse_smart_weibull.rds")  %>% mutate(rank_weibull = row_number())

smart_gamma <- read_rds("data/reverse_smart_gamma.rds")  %>% mutate(rank_gamma = row_number())

smart_normal <- read_rds("data/reverse_smart_normal.rds") %>% mutate(rank_normal = row_number())

smart_chi <-read_rds("data/reverse_smart_chisq.rds") %>% mutate(rank_chi = row_number())
smart_general <- read_rds("data/reverse_smart_general.rds") %>% mutate(rank_general = row_number())

data_unite <- smart_chi %>% 
  left_join(smart_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
    left_join(smart_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
  rename("gamma" = "mean_max_variation") %>% 
  left_join(smart_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("normal" = "mean_max_variation") %>% 
   left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("general" = "mean_max_variation") 
```

```{r}
data_unite %>% 
  select(-c(normal, gamma, weibull, general, chisq)) %>%
  dplyr::rename("normal"  = "rank_normal",
         "gamma"  = "rank_gamma",
         "weibull"  = "rank_weibull",
         "general"  = "rank_general",
         "chi"  = "rank_chi") %>% 
  knitr::kable(format = "latex")

# data_unite %>%
#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
#   knitr::kable()
```

### Maximum distance between levels of facet variable and median across levels of x-axis variable

```{r}
library(readr)
smart_weibull <- read_rds("data/smart_harmony_wb.rds")  %>% mutate(rank_weibull = row_number())

smart_gamma <- read_rds("data/smart_gamma.rds")  %>% mutate(rank_gamma = row_number())

smart_normal <- read_rds("data/smart_harmony_norm.rds") %>% mutate(rank_normal = row_number())

smart_chi <-read_rds("data/smart_harmony_chisq.rds") %>% mutate(rank_chi = row_number())
smart_general <- read_rds("data/smart_general.rds") %>% mutate(rank_general = row_number())

data_unite <- smart_chi %>% 
  left_join(smart_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
    left_join(smart_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
  rename("gamma" = "mean_max_variation") %>% 
  left_join(smart_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("normal" = "mean_max_variation") %>% 
   left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("general" = "mean_max_variation") 
```

```{r}
data_unite %>% 
  select(-c(normal, gamma, weibull, general, chisq)) %>% 
  dplyr::rename("normal"  = "rank_normal",
         "gamma"  = "rank_gamma",
         "weibull"  = "rank_weibull",
         "general"  = "rank_general",
         "chi"  = "rank_chi") %>% 
  knitr::kable(format = "latex")

# data_unite %>%
#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
#   knitr::kable()
```

## Graphical evidence

```{r, echo = FALSE, eval = FALSE}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("wknd_wday",
    "hour_day",
    response = "general_supply_kwh",
    plot_type = "boxplot",
    #symmetric = TRUE,
    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

```{r, echo = FALSE, eval = FALSE}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("day_week",
    "week_month",
    response = "general_supply_kwh",
    plot_type = "boxplot",
    #symmetric = TRUE,
    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()

smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("week_month",
    "day_week",
    response = "general_supply_kwh",
    plot_type = "boxplot",
    #symmetric = TRUE,
    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

## cricket data

```{r}
cricket_data <- read_rds("data/cricket_data.rds")

  hierarchy_model <- tibble::tibble(
      units = c("index", "over", "inning", "match"),
      convert_fct = c(1, 20, 2, 1))

hierarchy_tbl <- hierarchy_model

response <- "runs_per_over"

harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)

```

### Distribution fitting of distances for the harmony pair (lag_field, over)

```{r}
cricket_data  %>% qqplot_distance(gran1 = "lag_field", gran2 = "over", response = "runs_per_over", hierarchy_tbl = hierarchy_model)
```

```{r}
cric_weibull <- read_rds("data/cric_weibull.rds") %>% mutate(rank_weibull = row_number())

cric_gamma <-read_rds("data/cric_gamma.rds") %>% mutate(rank_gamma = row_number())

cric_normal <-read_rds("data/cric_normal.rds")%>% mutate(rank_normal = row_number())

cric_general <- read_rds("data/cric_general.rds")%>% mutate(rank_general = row_number())

cric_chisq <- read_rds("data/cric_chisq.rds")%>% mutate(rank_chisq = row_number())

data_unite <- cric_chisq %>% 
  left_join(cric_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
    left_join(cric_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
  rename("gamma" = "mean_max_variation") %>% 
  left_join(cric_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("normal" = "mean_max_variation") %>% 
   left_join(cric_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("general" = "mean_max_variation") 
```

```{r}
data_unite %>% 
  select(-c(normal, gamma, weibull, general, chisq)) %>%
  dplyr::rename("normal"  = "rank_normal",
         "gamma"  = "rank_gamma",
         "weibull"  = "rank_weibull",
         "general"  = "rank_general",
         "chi"  = "rank_chisq") %>% 
  knitr::kable(format = "latex")

# data_unite %>%
#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
#   knitr::kable()
```

### Maximum distance between levels of facet variable and median across levels of x-axis variable

```{r}
cric_weibull <- read_rds("data/ smart_cric_weibull.rds") %>% mutate(rank_weibull = row_number())

cric_gamma <-read_rds("data/ smart_cric_gamma.rds") %>% mutate(rank_gamma = row_number())

cric_normal <-read_rds("data/ smart_cric_normal.rds")%>% mutate(rank_normal = row_number())

cric_general <- read_rds("data/ smart_cric_general.rds")%>% mutate(rank_general = row_number())

cric_chisq <- read_rds("data/smart_cric_chisq.rds")%>% mutate(rank_chisq = row_number())

data_unite <- cric_chisq %>% 
  left_join(cric_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
    left_join(cric_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
  rename("gamma" = "mean_max_variation") %>% 
  left_join(cric_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("normal" = "mean_max_variation") %>% 
   left_join(cric_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
  rename("general" = "mean_max_variation") 
```

```{r}
data_unite %>% 
  select(-c(normal, gamma, weibull, general, chisq)) %>%
  dplyr::rename("normal"  = "rank_normal",
         "gamma"  = "rank_gamma",
         "weibull"  = "rank_weibull",
         "general"  = "rank_general",
         "chi"  = "rank_chisq") %>% 
  knitr::kable(format = "latex")

# data_unite %>%
#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
#   knitr::kable()
```

# Bibliography

```{r write-bib, eval = FALSE}
write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib")
```

### 2020-05-27 algo_rank_harmony.Rmd
added choosing_harmony document

@@
 smart_normal <- read_rds("data/smart_harmony_norm.rds") %>% mutate(rank_normal = row_number())
+
+library(readr)
+smart_chi <-read_rds("data/smart_harmony_chisq.rds") %>% mutate(rank_chi = row_number())
 ```
@@
 
-data_unite <- smart_normal %>% 
+data_unite <- smart_chi %>% 
+  left_join(smart_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
+   rename("MMPD_chisq" = "mean_max_variation.x", "MMPD_weibull" = "mean_max_variation.y") %>% 
     left_join(smart_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
-  rename("MMPD_normal" = "mean_max_variation.x", "MMPD_gamma" = "mean_max_variation.y") %>% 
-    left_join(smart_weibull,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  rename("MMPD_weibull" = "mean_max_variation") %>% 
-  left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  rename("MMPD_general" = "mean_max_variation") %>% 
-  select(-c(facet_levels, x_levels))
+  rename("MMPD_gamma" = "mean_max_variation") %>% 
+  left_join(smart_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
+  rename("MMPD_normal" = "mean_max_variation") %>% 
+   left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
+  rename("MMPD_general" = "mean_max_variation") %>%  select(-c(facet_levels, x_levels))
 
-data_unite %>% select(-c(MMPD_normal, MMPD_gamma, MMPD_weibull, MMPD_general))
-data_unite %>% select(-c(rank_normal, rank_gamma, rank_weibull, rank_general))
-
+data_unite %>% select(-c(MMPD_normal, MMPD_gamma, MMPD_weibull, MMPD_general, -MMPD_chisq)) %>% knitr::kable()
+data_unite %>% select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% knitr::kable()
 ```
@@
 ```{r}
-read_rds("data/ smart_cric_weibull.rds")%>% knitr::kable()
+cric_weibull <- read_rds("data/ smart_cric_weibull.rds")%>% mutate(rank_weibull = row_number())
 ```
@@
 library(readr)
-read_rds("data/ smart_cric_gamma.rds") %>% knitr::kable()
+cric_gamma <-read_rds("data/ smart_cric_gamma.rds") %>% mutate(rank_gamma = row_number())
 ```
@@
 ```{r}
-read_rds("data/ smart_cric_normal.rds")%>% knitr::kable()
+cric_normal <-read_rds("data/ smart_cric_normal.rds")%>% mutate(rank_normal = row_number())
 ```
@@
 ```{r}
-read_rds("data/ smart_cric_general.rds")%>% knitr::kable()
+cric_general <- read_rds("data/ smart_cric_general.rds")%>% mutate(rank_general = row_number())
+
+data_unite <- smart_normal %>% 
+  left_join(smart_gamma, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%
+  rename("MMPD_normal" = "mean_max_variation.x", "MMPD_gamma" = "mean_max_variation.y") %>% 
+    left_join(smart_weibull,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  rename("MMPD_weibull" = "mean_max_variation") %>% 
+  left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  rename("MMPD_general" = "mean_max_variation") %>% 
+  select(-c(facet_levels, x_levels))
+
+data_unite %>% select(-c(MMPD_normal, MMPD_gamma, MMPD_weibull, MMPD_general))
+data_unite %>% select(-c(rank_normal, rank_gamma, rank_weibull, rank_general))
 ```

### 2020-05-27 algo_validation.Rmd
added choosing_harmony document

@@
 ```{r}
-cricket_data <- read_rds("data/cricket_data.rds")
-cricket_data %>% hist_distance(gran1 = "lag_field", gran2 = "over", response = "runs_per_over", hierarchy_tbl = hierarchy_model)
+
+# cricket_data %>% histogram_distance(gran1 = "over", gran2 = "inning_match", response = "runs_per_over", hierarchy_tbl = hierarchy_model)
 ```
@@
 ```{r}
+cricket_data <- read_rds("data/cricket_data.rds")
 
@@
 ```
+
+```{r, eval = FALSE}
+sm <- smart_meter10 %>%
+  filter(customer_id %in% c(10017936))
+ .data = sm
+ response  = "general_supply_kwh"
+ .data %>% hist_distance(gran1 = "hour_day", gran2 = "wknd_wday", response = "general_supply_kwh")
+```

### 2020-05-27 choosing_harmonies.Rmd
choosing harmonies update

@@
 library(gravitas)
+library(ggpubr)
 opts_chunk$set(
@@
 
-```{r, echo = FALSE, eval = FALSE}
-smart_meter10 %>%
+```{r, echo = FALSE}
+p1 <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@
   ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
+  theme_minimal() + 
+  ggtitle("Rank 1 Section 6.1.1")  + 
+  scale_x_discrete(breaks = seq(0, 23, 5))
+
+p2 <- smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  prob_plot("hour_day",
+    "wknd_wday",
+    response = "general_supply_kwh",
+    plot_type = "boxplot",
+    #symmetric = TRUE,
+    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  scale_y_sqrt() +
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("electricity demand [KWh]") + xlab("") + ggtitle("") + ylab("")+ 
+  theme_minimal() + 
+  ggtitle("Rank 1 Section 6.1.2") + scale_x_discrete(labels = c('wday','wend'))
+
 ```
 
-```{r, echo = FALSE, eval = FALSE}
-smart_meter10 %>%
+```{r, echo = FALSE}
+p3 <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
-  prob_plot("day_week",
+  prob_plot("week_month",
+    "hour_day",
+    response = "general_supply_kwh",
+    plot_type = "boxplot",
+    #symmetric = TRUE,
+    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
+  ) +
+  scale_y_sqrt() +
+  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
+  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
+  theme_minimal() + 
+  ggtitle("Rank 2 Section 6.1.2") + 
+  scale_x_discrete(breaks = seq(0, 23, 5))
+
+p4 <- smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>%
+  prob_plot("hour_day",
     "week_month",
@@
   #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
+  ylab("electricity demand [KWh]") + xlab("week of the month") + ggtitle("") + ylab("")+ 
+  theme_minimal()  + 
+  ggtitle("Rank 2 Section 6.1.2")
 
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("week_month",
-    "day_week",
-    response = "general_supply_kwh",
-    plot_type = "boxplot",
-    #symmetric = TRUE,
-    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
+ggpubr::ggarrange(p1, p2,  ncol = 2)
+
+ggpubr::ggarrange(p3, p4,  ncol = 2)
 ```

### 2020-05-28 main.Rmd
number of observation write-up done

@@
 Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption.
-Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.
-The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
@@
 
-\noindent One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable.
+\noindent One of the potential ways to evaluate this variation is by computing the maximum pairwise distances between the distributions of the measured variable across different levels of the cyclic granularities.
 <!--across different levels of $C_j$ and integrating it for all levels of $C_i$.
@@
 Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
-\noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+\noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
 Suppose $X_{1},X_{2}\ldots ,X_{n}$  be a sequence of independent and identically-distributed random variables, and $M_{n}=\max\{X_{1},\ldots ,X_{n}\}$.
@@
 
-\noindent The algorithm employed addressing all the issues decried above is summarized as follows:
+\noindent The algorithm employed addressing all the issues described above is summarized as follows:
 <!-- Algorithm -->
@@
 
+<!-- For a tsibble with index ranging from ${1, 2, \dots, (T + 1)}$, that is (T+1) observations -->
+
+<!-- In linear granularities, if we consider a data set of size T where each time point corresponds to an unique value of the measured variable.  -->
+
+Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
+<!--This is unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.--> It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
+However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
+Summarizing the distribution of the measured variables through these multiple observations across cyclic granularities could be a potential way to explore and bring forward different features of the  data.
+<!-- compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
+Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
+
+<!-- using some summary statistics, each of which can highlight different features of the data or through probability densities -->
+
 <!-- T+1 observation not T -->
-
-Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
-This is unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.
-It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
-However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the distribution of these multiple observations across cyclic granularities using some summary statistics, each of which can highlight different features of the data or through probability densities is a valid way to compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
-Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
-
 However, we need to consider the effect of number of observations on the summarization even for harmonies.
-Suppose we have $T$ observations, and two cyclic granularities $C_1$ and $C_2$ with $l$ and $m$ categories respectively.
-Each element of $C_1$ occurs approximately $T/l$ times while each element of $C_2$ occurs approximately $T/m$ times.
-There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(lm)$.
-If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge lmk$ will be acceptable.
-The value of $k$ will depend on the statistical transformation we are producing.
-For computing deciles, $k=10$ may be acceptable, but for density estimates, $k \ge 30$ is suggested.
+Consider a data set with $(T + 1)$ observations, and two cyclic granularities $C_i$ and $C_j$ with $L$ and $M$ categories respectively <!--Each element of $C_i$ occurs approximately $(T+1)/L$ times while each element of $C_j$ occurs approximately $(T+1)/M$ times.
+and there are no or very few empty combinations since we only consider harmonies.--> Each combination might occur on average an equal number of times as $T\rightarrow\infty$ when the number of observations per combination is $nobs = (T + 1)/(LM)$ or they can vary anywhere between $nobs = 0$ and $nobs = (T+1)$.
+Now the statistical transformation used for summarizing the measured variable  for a combination should be chosen in line with this $nobs$.
+For example, it might be useful to compute deciles only when $nobs>=10$.
 Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
@@
 Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
+For Kernel density based plots $nobs>30$ is suggested, whereas for summary statistics based plots like boxplots $nobs = 10$ is acceptable and for quantile plots $nobs$ should at least be equal to the number of quantiles plotted.
+
+<!-- If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge lmk$ will be acceptable.
+The value of $k$ will depend on the statistical transformation we are producing.
+For computing deciles, $k=10$ may be acceptable, but for density estimates, $k \ge 30$ is suggested.
+-->
 
@@
   response = "run_rate",
-  plot_type = "boxplot",
+  plot_type = "quantile",
   symmetric = FALSE,

### 2020-05-28 main.Rmd
minor changes to number of observations

@@
 It is common to see aggregates [see @Goodwin_2012] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
-But studying overall energy use hides the distributions of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.
+But studying overall energy use hides the distribution of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.
 We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
@@
 
-## Choosing harmonies to show {#sec:chooseharmonies}
-
-<!-- Why do we need ranking system -->
-<!-- set of pairs - building the ranks for N_C and N_L -->
-
-\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities.
-Hence, all possible pairs of cyclic time granularities that could be analyzed together is $N_C \choose 2$.
-Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption.
-Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
-For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
--->
-
-<!-- Calculating distances between distributions -->
-
-\noindent One of the potential ways to evaluate this variation is by computing the maximum pairwise distances between the distributions of the measured variable across different levels of the cyclic granularities.
-<!--across different levels of $C_j$ and integrating it for all levels of $C_i$.
-Distances or similarity measures have been extensively studied and applied to fields like clustering, classification and anomaly detection.
----> We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
-One of the common ways to measure distances between distributions is through Jenson-Shanon divergence which is based on Kullback-Leibler divergence.
-Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-The Jensen-Shanon distance between two sample quantile vectors $q_u$ and $q_v$ computed for probability vector $p$ will be given by $$d_{uv} = [D(q_u, r) + D(q_v, r)]/2 \quad where \quad r = (q_u + q_v)/2$$ where,
- $$D(q_u,q_v) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_u$ and $q_v$.
-
-<!-- Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth. -->
-
-<!-- # why Normalizing required --->
-
-\noindent The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the cyclic granularities.
-But maximum is not robust to the number of levels and seems to be higher for harmonies with higher levels.
-Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
-\noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
-Suppose $X_{1},X_{2}\ldots ,X_{n}$  be a sequence of independent and identically-distributed random variables, and $M_{n}=\max\{X_{1},\ldots ,X_{n}\}$.
-If a sequence of pairs of real numbers $a_{n}>0$ and $\lim _{{n\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
-The normalizing constants, however, vary depending on the underlying distribution of $X_i$'s.
-Hence to normalize appropriately, it is important to assume a distribution of distances in our case.
-
-<!-- # distribution of distances and lit review -->
-
-\noindent Menendez1997-in and Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
-However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.
-
-\noindent The algorithm employed addressing all the issues described above is summarized as follows:
-<!-- Algorithm -->
-
-<!-- 1.
-Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_k$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$ -->
-
-\noindent Algorithm
-**Input** $(C_i, C_j, v) \quad \forall i, j \in N_C$  
-**Output** MMPD (Median Maximum Pairwise Distances) measuring the average variation across different levels of $C_i$ and $C_j$ $\forall i, j \in N_C$
-
-1.
-For each pair of cyclic granularity $(C_i, C_j)$, there will be $L$ groups corresponding to each level $A_k$ of $C_i$.
-Thus there will be  $m = \binom{L}{2}$ pairwise distances corresponding to each $A_k$.
-
-2.
-Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
-Let $d_1, d_2, d_3, \dots d_m$ be the pairwise Jensen-Shannon distances.
-Suppose the maximum pairwise distance for $A_k$ maximum be $M_k$.
-
-3.
-The distribution of distances is heavy tailed.
-But if they are gamma or weibull could be judged by looking at the QQ-plots of the empirical distribution.
-The estimates of the parameters are obtained using Maximum Likelihood Estimation.
-
-4.
-The norming constants $(a_k, b_k)$ are obtained from @Embrechts2013-pd and each $M_k$ is normalized using normalized maximum $NM_k = \frac{M_k - b_k}{a_k}$ for each $A_k$, $k \in {1, 2, \dots, K}$.
-We have normalized the number of levels of $C_j$ using Steps 1-4.
-
-<!-- the norming constants, which are a function of the parameters are obtained using method of moments. -->
-
-<!-- Since the distances follow a Chi-squared distribution with , the norming constants are taken as $a_n = $ and $b_n = $ -->
-
-5.
-We also need to aggregate these normalized maximum across different levels of $C_i$.
-For this we can take the mean, median or any quantiles of these normalized maximums.
-The median (and any quantiles) of these maximum distributions (which converges to a Gumbel distribution) grow at the rate of (some power of) $log K$ with $K \xrightarrow{} \infty$.
-Hence, M/log($K$) could serve as the final measure (MMPD), where $M = median\{NM_1, NM_2, \dots, NM_K\}$.
-
-6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
-
-<!-- 2.
-There will be $m$ time series variable corresponding to each level $A_k$ of $C_1$.
-We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_k$.
--->
-
-<!-- Suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where, -->
-<!--  $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$. -->
-
-<!-- 4.
-For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
-Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
--->
-
-<!-- 5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by -->
-<!-- $MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$. -->
-
-<!-- 6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD. -->
-
-<!-- with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$. -->
-
-<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
-
 ## Summarizing the measured variable {#sec:summarize-measured}
@@
 However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the distribution of the measured variables through these multiple observations across cyclic granularities could be a potential way to explore and bring forward different features of the  data.
+Summarizing the distribution of the measured variable through these multiple observations could be a potential way to explore and bring forward different features of the data.
 <!-- compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
@@
 However, we need to consider the effect of number of observations on the summarization even for harmonies.
-Consider a data set with $(T + 1)$ observations, and two cyclic granularities $C_i$ and $C_j$ with $L$ and $M$ categories respectively <!--Each element of $C_i$ occurs approximately $(T+1)/L$ times while each element of $C_j$ occurs approximately $(T+1)/M$ times.
-and there are no or very few empty combinations since we only consider harmonies.--> Each combination might occur on average an equal number of times as $T\rightarrow\infty$ when the number of observations per combination is $nobs = (T + 1)/(LM)$ or they can vary anywhere between $nobs = 0$ and $nobs = (T+1)$.
-Now the statistical transformation used for summarizing the measured variable  for a combination should be chosen in line with this $nobs$.
-For example, it might be useful to compute deciles only when $nobs>=10$.
-Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
+Consider a data set with $(T + 1)$ observations, and two cyclic granularities $C_i$ and $C_j$ with $L$ and $M$ categories respectively.
+<!--Each element of $C_i$ occurs approximately $(T+1)/L$ times while each element of $C_j$ occurs approximately $(T+1)/M$ times.
+and there are no or very few empty combinations since we only consider harmonies.--> Let $nobs$ be the number of observations for a combination of levels from $(C_i, C_j)$.
+Now, $nobs$ might be equal for all combinations on an average <!--as $T\rightarrow\infty$ when the number of observations per combination is-->  so that $nobs = (T + 1)/(LM)$ as $T\rightarrow\infty$ or ideally take any value between $\{0, 1, 2, \dots, T\}$.
+The statistical transformation used for summarizing the measured variable for a combination should be chosen in line with  $nobs$.
+For example, it might be useful to compute deciles only when $nobs \ge 10$.
+Rarely occurring categories such as the 366th day of the year, or the 31st day of the month are more likely to suffer from this problem.
 
@@
 ```
+
 ### Interchangeability of mappings
@@
 
+## Choosing harmonies to show {#sec:chooseharmonies}
+
+<!-- Why do we need ranking system -->
+<!-- set of pairs - building the ranks for N_C and N_L -->
+
+\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities.
+Hence, all possible pairs of cyclic time granularities that could be analyzed together is $N_C \choose 2$.
+Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption.
+Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
+The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
+-->
+
+<!-- Calculating distances between distributions -->
+
+\noindent One of the potential ways to evaluate this variation is by computing the maximum pairwise distances between the distributions of the measured variable across different levels of the cyclic granularities.
+<!--across different levels of $C_j$ and integrating it for all levels of $C_i$.
+Distances or similarity measures have been extensively studied and applied to fields like clustering, classification and anomaly detection.
+---> We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
+One of the common ways to measure distances between distributions is through Jenson-Shanon divergence which is based on Kullback-Leibler divergence.
+Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
+The Jensen-Shanon distance between two sample quantile vectors $q_u$ and $q_v$ computed for probability vector $p$ will be given by $$d_{uv} = [D(q_u, r) + D(q_v, r)]/2 \quad where \quad r = (q_u + q_v)/2$$ where,
+ $$D(q_u,q_v) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_u$ and $q_v$.
+
+<!-- Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth. -->
+
+<!-- # why Normalizing required --->
+
+\noindent The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the cyclic granularities.
+But maximum is not robust to the number of levels and seems to be higher for harmonies with higher levels.
+Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
+\noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+Suppose $X_{1},X_{2}\ldots ,X_{n}$  be a sequence of independent and identically-distributed random variables, and $M_{n}=\max\{X_{1},\ldots ,X_{n}\}$.
+If a sequence of pairs of real numbers $a_{n}>0$ and $\lim _{{n\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
+The normalizing constants, however, vary depending on the underlying distribution of $X_i$'s.
+Hence to normalize appropriately, it is important to assume a distribution of distances in our case.
+
+<!-- # distribution of distances and lit review -->
+
+\noindent Menendez1997-in and Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
+However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.
+
+\noindent The algorithm employed addressing all the issues described above is summarized as follows:
+<!-- Algorithm -->
+
+<!-- 1.
+Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
+Hence for each combination ($A_k$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$ -->
+
+\noindent Algorithm
+**Input** $(C_i, C_j, v) \quad \forall i, j \in N_C$  
+**Output** MMPD (Median Maximum Pairwise Distances) measuring the average variation across different levels of $C_i$ and $C_j$ $\forall i, j \in N_C$
+
+1.
+For each pair of cyclic granularity $(C_i, C_j)$, there will be $L$ groups corresponding to each level $A_k$ of $C_i$.
+Thus there will be  $m = \binom{L}{2}$ pairwise distances corresponding to each $A_k$.
+
+2.
+Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
+Let $d_1, d_2, d_3, \dots d_m$ be the pairwise Jensen-Shannon distances.
+Suppose the maximum pairwise distance for $A_k$ maximum be $M_k$.
+
+3.
+The distribution of distances is heavy tailed.
+But if they are gamma or weibull could be judged by looking at the QQ-plots of the empirical distribution.
+The estimates of the parameters are obtained using Maximum Likelihood Estimation.
+
+4.
+The norming constants $(a_k, b_k)$ are obtained from @Embrechts2013-pd and each $M_k$ is normalized using normalized maximum $NM_k = \frac{M_k - b_k}{a_k}$ for each $A_k$, $k \in {1, 2, \dots, K}$.
+We have normalized the number of levels of $C_j$ using Steps 1-4.
+
+<!-- the norming constants, which are a function of the parameters are obtained using method of moments. -->
+
+<!-- Since the distances follow a Chi-squared distribution with , the norming constants are taken as $a_n = $ and $b_n = $ -->
+
+5.
+We also need to aggregate these normalized maximum across different levels of $C_i$.
+For this we can take the mean, median or any quantiles of these normalized maximums.
+The median (and any quantiles) of these maximum distributions (which converges to a Gumbel distribution) grow at the rate of (some power of) $log K$ with $K \xrightarrow{} \infty$.
+Hence, M/log($K$) could serve as the final measure (MMPD), where $M = median\{NM_1, NM_2, \dots, NM_K\}$.
+
+6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
+
+<!-- 2.
+There will be $m$ time series variable corresponding to each level $A_k$ of $C_1$.
+We compute the pairwise differences in distributions of these $m$ time series variables.
+Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_k$.
+-->
+
+<!-- Suppose $q$ is the sample quantile vector computed for probability vector $p$.
+Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where, -->
+<!--  $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$. -->
+
+<!-- 4.
+For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
+Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
+-->
+
+<!-- 5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by -->
+<!-- $MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$. -->
+
+<!-- 6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD. -->
+
+<!-- with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$. -->
+
+<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
+
 ## Number of observations and statistical transformations
 
-Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (\ref{sec:summarize-measured}) or unevenly distributed events.
-For Kernel density based plots $nobs>30$ is suggested, whereas for summary statistics based plots like boxplots $nobs = 10$ is acceptable and for quantile plots $nobs$ should at least be equal to the number of quantiles plotted.
+Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section \ref{sec:summarize-measured}) or unevenly distributed events.
+For Kernel density based plots $nobs>30$ is suggested, for summary statistics based plots like boxplots $nobs = 10$ is acceptable and for quantile plots $nobs$ should at least be equal to the number of quantiles plotted.
 

### 2020-05-28 choosing_harmonies.Rmd
after adding conditional ggplot to gravitas
chhosing harmony figure update

@@
 
-Currently, MME is used. But MASS::fitdistr() and package fitdistrplus also provide methods to estimate paramters through MLE. WIP.
+Currently, MME is used. But MASS::fitdistr() and package fitdistrplus also provide methods to estimate parameters through MLE. WIP.
 
@@
 
-WIP
+Threholds could be chosen for distances (chi-squared) or maximum distance (Gumbel distribution).
+
+Critical values of the Chi-squared statistic could be obtained for appropriate degrees of freedom.
+
+Chi-squared statistic: And test values less than critical value would imply that the distances are not significantly different from zero , implying distributions are similar.
+If distributions are similar most times, then the plot is not interesting.
+Hence, all pairs for which most of the pairwise distances are significantly different from zero would only be included in the harmony rank table.
+
+Gumbel statistic: The test decides if the sample of maximum distance is from Gumbel or not.
+If it is from Gumbel, the value of the test statistic should be ideally zero.
+And we want to take all pairs for which the statistic is significantly different from zero.
+So we choose test values greater than the Gumbel critical value at 5% significance.
 
@@
          "general"  = "rank_general",
-         "chi"  = "rank_chi") %>% 
+         "chi"  = "rank_chi",
+         "f_L" = "facet_levels",
+         "x_L" = "x_levels") %>% 
+  knitr::kable(format = "latex")
+
+  threshold <- read_rds("data/smart_harmony_threshold.rds") %>% 
+               mutate(thresval = "select") %>% 
+    select(facet_variable, x_variable, thresval)
+
+data_unite %>% 
+  select(-c(normal, gamma, weibull, general, chisq)) %>%
+  left_join(threshold, 
+            by = c("facet_variable", "x_variable"))%>%
+  select(-facet_levels, -x_levels) %>% 
     knitr::kable(format = "latex")
@@
 #   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>%
+#   filter()
 #   knitr::kable()

### 2020-05-29 main.Rmd
algo made really short

@@
 
-\noindent Menendez1997-in and Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
+\noindent @Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
 However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.
@@
 
-\noindent Algorithm
 **Input** $(C_i, C_j, v) \quad \forall i, j \in N_C$  
@@
 
-1.
+1. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$ for the pair $(C_i, C_j)$.
+
+2. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories.
+
+3. Compute maximum within the $m$ distances.
+
+4. Compute normalised maximum distance ($NM$) using appropriate norming constants so that the distribution of the normalised maximum converges to a Gumbel distribution as $m\rightarrow\infty$.
+
+5. Use Step 1-4 to compute normalised maximum distance for each $k$ levels ($NM_k$).
+
+6. Compute median maximum pairwise distance $$MMPD = median\{NM_1, NM_2, \dots, NM_K\}$$.
+
+7. Repeat Step 1 to 6 for all harmony pairs.
+
+<!-- 1.
 For each pair of cyclic granularity $(C_i, C_j)$, there will be $L$ groups corresponding to each level $A_k$ of $C_i$.
 Thus there will be  $m = \binom{L}{2}$ pairwise distances corresponding to each $A_k$.
-
-2.
+-->
+
+<!-- 2.
 Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
@@
 Suppose the maximum pairwise distance for $A_k$ maximum be $M_k$.
-
-3.
+-->
+
+<!-- 3.
 The distribution of distances is heavy tailed.
@@
 The estimates of the parameters are obtained using Maximum Likelihood Estimation.
-
-4.
+-->
+
+<!-- 4.
 The norming constants $(a_k, b_k)$ are obtained from @Embrechts2013-pd and each $M_k$ is normalized using normalized maximum $NM_k = \frac{M_k - b_k}{a_k}$ for each $A_k$, $k \in {1, 2, \dots, K}$.
 We have normalized the number of levels of $C_j$ using Steps 1-4.
-
-<!-- the norming constants, which are a function of the parameters are obtained using method of moments. -->
-
-<!-- Since the distances follow a Chi-squared distribution with , the norming constants are taken as $a_n = $ and $b_n = $ -->
-
-5.
+-->
+
+<!-- 5.
 We also need to aggregate these normalized maximum across different levels of $C_i$.
@@
 Hence, M/log($K$) could serve as the final measure (MMPD), where $M = median\{NM_1, NM_2, \dots, NM_K\}$.
-
-6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD.
-
-<!-- 2.
-There will be $m$ time series variable corresponding to each level $A_k$ of $C_1$.
-We compute the pairwise differences in distributions of these $m$ time series variables.
-Thus there will be  $\binom{m}{2}$ pairwise distances corresponding to each $A_k$.
--->
-
-<!-- Suppose $q$ is the sample quantile vector computed for probability vector $p$.
-Then Jensen-Shanon distance between two sample quantiles $q_j$ and $q_k$ will be given by $$d_{jk} = [D(q_j, r) + D(q_k, r)]/2 \quad where \quad r = (q_j + q_k)/2$$ where, -->
-<!--  $$D(q_j,q_k) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_j$ and $q_k$. -->
-
-<!-- 4.
-For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  $\binom{m}{2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$.
-Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.
--->
-
-<!-- 5. The mean maximum pairwise distance (MMPD) over $i = {1, 2, \dots, l}$ is computed by -->
-<!-- $MMPD = (1/l)\sum_{i=1}^{l}{D_{i}}$. -->
+-->
 
 <!-- 6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD. -->
-
-<!-- with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$. -->
-
-<!-- for each level of $C_1$, maximum variation in distribution has been captured against different levels of $C_2$. -->
 

### 2020-05-30 main.Rmd
panel.border changed for cricket example

@@
   ggtitle("b") +
-   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
-# + geom_boxplot(width = 0.5, aes(colour = lag_field))
+   theme(plot.title = element_text(face = "bold")) +
+        theme(axis.ticks = element_blank(), legend.background = element_blank(),
+            legend.key = element_blank(), panel.background = element_blank(), strip.background = element_blank(),
+            plot.background = element_blank(), complete = TRUE, panel.grid.major = element_line(colour = "#E0E0E0"),
+            panel.border = element_rect(colour = "#E0E0E0", fill = NA))
 ```

### 2020-06-03 main.Rmd
choosing harmony section update

@@
 \noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
-Suppose $X_{1},X_{2}\ldots ,X_{n}$  be a sequence of independent and identically-distributed random variables, and $M_{n}=\max\{X_{1},\ldots ,X_{n}\}$.
-If a sequence of pairs of real numbers $a_{n}>0$ and $\lim _{{n\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
-The normalizing constants, however, vary depending on the underlying distribution of $X_i$'s.
-Hence to normalize appropriately, it is important to assume a distribution of distances in our case.
+Suppose $d_{1},d_{2}\ldots ,d_{m}$  be a sequence of independent and identically-distributed pairwise distances, and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
+If a sequence of pairs of real numbers $a_{m}>0$ and $\lim _{{n\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
+The normalizing constants, however, vary depending on the underlying distribution of $d_i$'s.
+Hence to normalize appropriately, it is important to assume a distribution of $d_i$'s in our case.
 
@@
 
-**Input** $(C_i, C_j, v) \quad \forall i, j \in N_C$  
-**Output** MMPD (Median Maximum Pairwise Distances) measuring the average variation across different levels of $C_i$ and $C_j$ $\forall i, j \in N_C$
-
-1. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$ for the pair $(C_i, C_j)$.
+- **Input:** Data corresponding to all harmony pairs, i.e., data sets of the form $(C_i, C_j, v)$ 
+$\forall i, j \in N_C$  
+- **Output:** MMPD (Median Maximum Pairwise Distances) measuring the average variation across different levels of $C_i$ and $C_j$ $\forall i, j \in N_C$
+
+1. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$.
 
@@
 
-5. Use Step 1-4 to compute normalised maximum distance for each $k$ levels ($NM_k$).
-
-6. Compute median maximum pairwise distance $$MMPD = median\{NM_1, NM_2, \dots, NM_K\}$$.
-
-7. Repeat Step 1 to 6 for all harmony pairs.
+5. Use Steps 1-4 to compute normalised maximum distance for each $k$ levels ($NM_k$).
+
+6. Compute $MMPD = median(NM_1, NM_2, \dots, NM_K)/log(K)$. Median normalised distance grows at the rate of log(K) as $K\rightarrow\infty$ and hence is normalised for the value of K in $MMPD$.
+
+7. Repeat Steps 1 to 6 for all harmony pairs.
 

### 2020-06-04 main.Rmd
reference of JSD added and definition of Kullback Leibler improved

@@
 ---> We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
-One of the common ways to measure distances between distributions is through Jenson-Shanon divergence which is based on Kullback-Leibler divergence.
-Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-The Jensen-Shanon distance between two sample quantile vectors $q_u$ and $q_v$ computed for probability vector $p$ will be given by $$d_{uv} = [D(q_u, r) + D(q_v, r)]/2 \quad where \quad r = (q_u + q_v)/2$$ where,
- $$D(q_u,q_v) = \int^{\infty}_{-\infty}q_j(x)log\frac{q_j(x)}{q_k(x)}\,dx$$ is the Kullback-Leibler divergence between $q_u$ and $q_v$.
-
-<!-- Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth. -->
+One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence.
+The Jensen-Shanon distance between twon probability distribution p1 and p2 is given by:  $q_u$ and $q_v$ computed for probability vector $p$ will be given by $$d_{p_1, p2} = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
+ $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.
+
+Probability distributions are estimated through quantiles instead of kernel density so that there is minimal dependency on selecting kernel or bandwidth.
 

### 2020-06-04 main.Rmd
gestalt theory in stating why the idea is of median maximum pairwise distances

@@
 minute-of-day                                &
-  $C_2 = z \mod 60*24$                  &
+  $C_j = z \mod 60*24$                  &
   $P_2= `r 60*24`$\\
@@
 <!-- across bivariate cyclic granularities.  -->
-<!-- Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
-Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+<!-- Let us consider two cyclic granularities $C_1$ and $C_j$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
+Data sets of the form <$C_1$, $C_j$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
 -->
@@
 
-<!-- Let us consider two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
-Data sets of the form <$C_1$, $C_2$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
+<!-- Let us consider two cyclic granularities $C_1$ and $C_j$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
+Data sets of the form <$C_1$, $C_j$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
 -->
@@
 The layered grammar of graphics proposed by @Wickham2009pk, which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
-Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
+Drawing from the grammar of graphics, if <$C_i$, $C_j$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
 
@@
 
-- Data: <$C_1$, $C_2$, $v$>
-- Aesthetic mapping (mapping of variables to elements of the plot): $C_1$ mapped to $x$ position and $v$ to $y$ position
+- Data: <$C_i$, $C_j$, $v$>
+- Aesthetic mapping (mapping of variables to elements of the plot): $C_i$ mapped to $x$ position and $v$ to $y$ position
 - Statistical transformation (data summarization): any descriptive or smoothing statistics that summarizes distribution of $v$
 - Geometric objects (physical representation of the data): any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
-- Facet (split plots):  $C_2$
+- Facet (split plots):  $C_j$
 
@@
 Layers are usually created with a geom function.
-Considering the data structure <$A_k$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
-<!-- - Data: <$A_k$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
+Considering the data structure <$A_k$, $B_l(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
+<!-- - Data: <$A_k$, $B_l(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
 <!-- - Aesthetic mapping: $A_k$ mapped to $x$ position and $v(s)$ to $y$ position. -->
@@
 <!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
-<!-- - Facet:  $B_j(s)$ -->
-
-<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_k$, $B_j(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
+<!-- - Facet:  $B_l(s)$ -->
+
+<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_k$, $B_l(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
 The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
@@
 
-We will discuss the effect of the levels of cyclic granularities on the choice of plots <!--Recall that for cyclic granularities $C_1$ and $C_2$ such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-$A_k$'s and $B_j$'s are referred to as the levels of $C_1$ and $C_2$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.The levels have an impact on the--> since space and resolution might become a problem if the number of levels are too high.
+The levels of cyclic granularities has a role to play on the choice of plots <!--Recall that for cyclic granularities $C_i$ and $C_j$ such that $C_i$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
+$A_k$'s and $B_l$'s are referred to as the levels of $C_i$ and $C_j$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.The levels have an impact on the--> since space and resolution might become a problem with too many levels.
 A potential approach could be to categorize the levels as very high/high/medium/low for each cyclic granularity and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
@@
 <!-- Suppose we have ${x_1, x_2, \dots, x_n} which are random sample from some underlying population. --> <!-- Let $p(x)$ and $q(x)$ be two probability distributions of a random variable x. -->
-<!-- We compute Kullback-Liebler distance based on quantiles $q_1$ and $q_2$ of $v$ at probabilities $p$ corresponding to $C_1$ and $C_2$ respectively. -->
+<!-- We compute Kullback-Liebler distance based on quantiles $q_1$ and $q_2$ of $v$ at probabilities $p$ corresponding to $C_i$ and $C_j$ respectively. -->
 
@@
 
-If there are levels of cyclic granularities plotted across the  x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs.
-We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
-Harmonies are pairs of granularities that do not contain empty combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations.
-As a result, we should avoid plotting clashes.
+For data sets of the form ($C_i$, $C_j$, $v$) $\forall i, j \in N_C$, if there are levels of $C_i$ (or $C_j$) not spanned by levels of $C_j$ (or $C_i$), empty sets are formed leading to potential ineffective graphs.
+The conjecture is that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
+Harmonies are pairs of granularities that do not contain empty combinations and thus aid exploratory data analysis.
+Plotting Clashes should be avoided since they contain empty combinations.
 For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarters of the year.
@@
 We will discuss the effect of the mapping of cyclic granularities in this section.
-When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets, then $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet.
+When we consider data sets of the form <$C_i$, $C_j$, $v$> with $C_i$ mapped to $x$ position and $C_j$ to facets, then $A_k$'s are placed in close proximity and each $B_l$ represent a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$--> With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.
+Hence, in this case $A_k$'s are compared against each other within each group.<!-- When $C_j$ is mapped to facets and $C_i$ is mapped to $x$ axis, comparisons would ideally happen across all $B_l$ within each $A_k$--> With the mapping of $C_i$ and $C_j$ reversed, emphasis will shift to different behavior of the variables.
 \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
@@
 One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence.
-The Jensen-Shanon distance between twon probability distribution p1 and p2 is given by:  $q_u$ and $q_v$ computed for probability vector $p$ will be given by $$d_{p_1, p2} = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
+The Jensen-Shanon distance between two probability distribution p1 and p2 is given by $$d_{p_1, p2} = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
  $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.
-
 Probability distributions are estimated through quantiles instead of kernel density so that there is minimal dependency on selecting kernel or bandwidth.
 
+ <!-- $q_u$ and $q_v$ computed for probability vector $p$ will be given by -->
+
 <!-- # why Normalizing required --->
 
-\noindent The harmony pairs could be arranged from highest to lowest average maximum pairwise distances across different levels of the cyclic granularities.
+\noindent The choice of aesthetic mapping and facets for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$ representing a group/facet.
+A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet.
+A probable solution for ranking interesting harmony pairs could be to arrange from highest to lowest average maximum pairwise distances.
 But maximum is not robust to the number of levels and seems to be higher for harmonies with higher levels.
-Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
+Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.
 \noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
@@
 <!-- 1.
-Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_k$, $B_j$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$ -->
+Suppose $C_i$ and $C_2$ are two cyclic granularities such that $C_i$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
+Hence for each combination ($A_k$, $B_l$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$ -->
 
@@
 
-<!-- Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories.
-Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
+<!-- Suppose we have $T$ observations, and two cyclic granularities $C_i$ with $n$ categories and $C_2$ with $m$ categories.
+Each element of $C_i$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
 There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.

### 2020-06-04 choosing_harmonies.Rmd
choosing harmonies section further update

@@
 
+```{r}
+cricket_data <- read_rds("data/cricket_data.rds")
+cricket_data %>%
+  filter(over!=1) %>%
+  prob_plot("over", "lag_field",
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "violin",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("b") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
+```
+
 # Bibliography

### 2020-06-04 main.Rmd
choosing harmonies section further update

@@
 One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence.
-The Jensen-Shanon distance between two probability distribution p1 and p2 is given by $$d_{p_1, p2} = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
+The Jensen-Shanon distance between two probability distribution p1 and p2 is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
  $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.
@@
 But maximum is not robust to the number of levels and seems to be higher for harmonies with higher levels.
-Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.
-\noindent The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
-Suppose $d_{1},d_{2}\ldots ,d_{m}$  be a sequence of independent and identically-distributed pairwise distances, and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
-If a sequence of pairs of real numbers $a_{m}>0$ and $\lim _{{n\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
-The normalizing constants, however, vary depending on the underlying distribution of $d_i$'s.
-Hence to normalize appropriately, it is important to assume a distribution of $d_i$'s in our case.
+Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.<!--- \noindent The Fisher–Tippett–Gnedenko theorem states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+--> Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances, and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
+If a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
+The normalizing constants $(a_{m}, b_{m})$ vary depending on the underlying distribution of the pairwise distances.
+Hence to normalize appropriately, it is important to assume a distribution of these distances.
 
@@
 
-\noindent @Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
-However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.
-
-\noindent The algorithm employed addressing all the issues described above is summarized as follows:
+\noindent @Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a chi-squared distribution with degrees of freedom equal to the number of discretization used to estimate the continuous distribution.
+With large degrees of freedom, chi-squared is asymptotically normal by the CLT.
+Hence, the norming constants for the distribution of maximum distance are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of a standard normal distribution.
+
+<!--However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.-->
+
+\noindent The algorithm employed is summarized as follows:
 <!-- Algorithm -->
@@
 
-1. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$.
-
-2. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories.
-
-3. Compute maximum within the $m$ distances.
-
-4. Compute normalised maximum distance ($NM$) using appropriate norming constants so that the distribution of the normalised maximum converges to a Gumbel distribution as $m\rightarrow\infty$.
-
-5. Use Steps 1-4 to compute normalised maximum distance for each $k$ levels ($NM_k$).
-
-6. Compute $MMPD = median(NM_1, NM_2, \dots, NM_K)/log(K)$. Median normalised distance grows at the rate of log(K) as $K\rightarrow\infty$ and hence is normalised for the value of K in $MMPD$.
-
-7. Repeat Steps 1 to 6 for all harmony pairs.
+1. Fix harmony pair $(C_i, C_j)$.
+
+2. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$.
+
+3. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories.
+
+4. Compute maximum within the $m$ computed distances.
+
+5. Compute normalized maximum distance ($NM$) using appropriate norming constants so that the distribution of the normalized maximum converges to a Gumbel distribution as $m\rightarrow\infty$.
+
+6. Use Steps 1-4 to compute normalized maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.
+
+7. Compute MMPD = median $(NM_1, NM_2, \dots, NM_K)$/log($K$). Median normalized distance grows at the rate of log($K$) as $K\rightarrow\infty$ and hence is normalized for the value of $K$ in MMPD.
+
+8. Repeat Steps 1 to 7 for all harmony pairs.
 

### 2020-06-04 main.Rmd
till smart meter section

@@
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
-<!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to look at the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
+One of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
+<!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to visualize the distribution of the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
 
@@
 Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
-Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ (via \ref{sec:data-structure}) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.
-If these options are considered too numerous, the smallest and largest granularities are good candidates to consider.
-Hence, we removed half hour and year.
-Then we will be left $(4*3/2) = 6$ cyclic granularites, namely: "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
+Considering $6$ linear granularities half-hour, hour, day, week, month and year in the hierarchy table, $N_C = (6*5/2) = 15$.
+<!--(via Section \ref{sec:data-structure}) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.--> If $N_C$ seem too large, the smallest and largest linear granularities could be considered to be removed from the hierarchy table.
+We remove half-year and year to have $N_C = (4*3/2) = 6$ and obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
+Further, we add cyclic granularity day-type( "wknd_wday") to capture weekend and weekday behavior.
 Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Section \ref{sec:cyclic-calendar}.
 <!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
--->
-with the index set mapping half-hourly time periods to a set of positive integers.
+with the index set mapping half-hourly time periods to a set of positive integers.-->
 
@@
 smart_meter %>%
-  search_gran(filter_out = c("semester", "quarter", "fortnight"))
+  search_gran(filter_out = c("semester", 
+                             "quarter",
+                             "fortnight"))
 ```
@@
 
-From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet.
-Choosing 2 of the possible 6 granularities, which is equivalent to having ${6 \choose 2} = 30$ candidates for visualization.
-Fortunately, harmony granularities can be identified among those 30 possibilities to narrow the search.
-\autoref{tab:harmony-tab} shows 13 harmonies pairs, after removing clashes and any cyclic granularities with levels more than 31, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
-For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/weekend) and $C2$ is hour of the day.
+From the search list, $N_C = 7$ cyclic granularities are chosen for which we would like to derive insights of energy behavior.
+Recalling the data structure <$C_i$, $C_j$, `general_supply_kwh`> for exploration $\forall i, j \in \{1, 2, ldots, 7\}$, each of these $7$ cyclic granularities can either be mapped to x-axis or to facet.
+Choosing $2$ of the possible $7$ granularities, which is equivalent to having ${7 \choose 2} = 21$ candidates for visualization.
+Fortunately, harmony granularities can be identified among those $21$ possibilities to narrow the search.
+\autoref{tab:harmony-tab} shows $10$ harmony pairs after removing clashes and any cyclic granularities with levels more than 31, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
+For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
 The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust} (b and c respectively).
-\autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type with distribution geometrically displayed as boxplots.
+\autoref{fig:bothcust} (a) displays reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour of the day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.
 

### 2020-06-04 main.Rmd
added citation for KL divergence

@@
 ---> We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
-One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence.
-The Jensen-Shanon distance between two probability distribution p1 and p2 is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
+One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence (@Kullback1951-jy).
+The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
  $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.
@@
 
-\noindent The choice of aesthetic mapping and facets for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$ representing a group/facet.
+\noindent The choice of aesthetic mapping and facets for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet.
 A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet.
 A probable solution for ranking interesting harmony pairs could be to arrange from highest to lowest average maximum pairwise distances.
-But maximum is not robust to the number of levels and seems to be higher for harmonies with higher levels.
+But maximum is not robust to the number of levels and seems to be higher for harmonies with higher number of levels.
 Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.<!--- \noindent The Fisher–Tippett–Gnedenko theorem states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
---> Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances, and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
-If a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, the Fréchet or the Weibull family.
+--> Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
+Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family.
 The normalizing constants $(a_{m}, b_{m})$ vary depending on the underlying distribution of the pairwise distances.
@@
 \noindent @Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a chi-squared distribution with degrees of freedom equal to the number of discretization used to estimate the continuous distribution.
-With large degrees of freedom, chi-squared is asymptotically normal by the CLT.
-Hence, the norming constants for the distribution of maximum distance are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of a standard normal distribution.
+With large degrees of freedom, chi-squared is asymptotically normal by the Central Limit Theorem.
+Hence, the norming constants for the distribution of $M_m$ are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of a standard normal distribution.
+Typically normal distributions will have Gumbel limiting distribution for the maxima, which have medians that grow as $m\rightarrow\infty$ at the rate of log $m$.
+
+<!-- Median normalized distance grows at the rate of log($K$) as $K\rightarrow\infty$ and hence is normalized for the value of $K$ in MMPD.  -->
 
@@
 
-4. Compute maximum within the $m$ computed distances.
-
-5. Compute normalized maximum distance ($NM$) using appropriate norming constants so that the distribution of the normalized maximum converges to a Gumbel distribution as $m\rightarrow\infty$.
+4. Identify maximum within the $m$ computed distances.
+
+5. Compute normalized maximum distance ($NM$) using appropriate norming constants.
+<!--so that the distribution of the normalized maximum converges to a Gumbel distribution as $m\rightarrow\infty$.-->
 
@@
 
-7. Compute MMPD = median $(NM_1, NM_2, \dots, NM_K)$/log($K$). Median normalized distance grows at the rate of log($K$) as $K\rightarrow\infty$ and hence is normalized for the value of $K$ in MMPD.
+7. Compute MMPD = median $(NM_1, NM_2, \dots, NM_K)$/log($K$).
 
@@
 Example granularites then include ball of the over, over of the inning and ball of the inning.
-Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularities will are aperiodic.
+Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularities are aperiodic.
 For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
@@
 
-First, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second inning.
+First, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second innings.
 The Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
@@
 
-Wickets per over are considered as aperiodic cyclic granularities with wickets as aperiodic linear granularities.
+Wickets per over are considered as an aperiodic cyclic granularity with wickets as an aperiodic linear granularity.
 These granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.

### 2020-06-04 choosing_harmonies_Normal.Rmd
choosing harmonies pdf update

--- 2020-05-27 choosing_harmonies.Rmd
@@
 
-## Theoretical evidence
-
-@Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a Chi-squared distribution.
-Let $p^{(1)} \equiv (p_1^{(1)}, p_2^{(1)}, \dots, p_k^{(1)})$ and $p^{(2)} \equiv (p_1^{(2)}, p_1^{(2)}, \dots, p_k^{(2)})$ denote two probability distributions, where $k$ is the number of components of the probability vector p.
-Then 2N(ln2)D is known to converge to {$\chi^2$} distribution with degrees of freedom = (k - 1) and
-
-$D[p^{(1)}, p^{(2)}] = H((p^{(1)} + p^{(2)})/2)  - H(p^{(1)}) - H(p^{(2)})$, $H_{(p)} = \sum_{i = 1}^{k} p_ilog_2p_i$ and N is the total number of events.
-
-## Empirical evidence
-
-However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.
-The QQ plots of four harmony pairs are plotted below.
-It could be seen that Chi-square distribution serves as a pretty good fit to the data (specially in the extreme right tail).
-
-```{r}
-library(tsibbledata)
-library(ggplot2)
-library(tsibble)
-library(lvplot)
-library(dplyr)
-library(gravitas)
-library(purrr)
-library(magrittr)
-sm <- smart_meter10 %>%
-filter(customer_id %in% c(10017936))
-.data = sm
-gran1 = "wknd_wday"
-gran2 = "hour_day"
-response  = "general_supply_kwh"
-```
-
-### Distribution fitting of distances for the harmony pair (weekend/weekday, hour-of-day)
-
-```{r}
-.data %>% qqplot_distance(gran1 = "wknd_wday", gran2 = "hour_day", response = "general_supply_kwh")
-```
-
-### Distribution fitting of distances for the harmony pair (day-of-week, hour-of-day)
-
-```{r}
-.data %>% qqplot_distance(gran1 = "day_week", gran2 = "hour_day", response = "general_supply_kwh")
-```
-
-### Distribution fitting of distances for the harmony pair (hour-of-day, week-of-month)
-
-```{r}
-.data %>% qqplot_distance(gran1 = "hour_day", gran2 = "week_month", response = "general_supply_kwh")
-```
-
-### Distribution fitting of distances for the harmony pair (day-of-month, day-of-week)
-
-```{r}
-.data %>% qqplot_distance(gran1 = "day_month", gran2 = "day_week", response = "general_supply_kwh")
-```
-
-# Estimating parameters
-
-Currently, MME is used. But MASS::fitdistr() and package fitdistrplus also provide methods to estimate paramters through MLE. WIP.
-
 # Choose thresholds for harmonies
 
-WIP
+| facets/x-axis     	| A_1  	| A_2  	| A_3  	| .. 	| .. 	| A_K  	|
+|-----	|------	|------	|------	|----	|----	|------	|
+| B_1 	| p_11 	| p_12 	| p_13 	| .. 	| .. 	| p_1K 	|
+| B_2 	|      	|      	|      	|    	|    	|      	|
+| B_3 	|      	|      	|      	|    	|    	|      	|
+| ..  	|      	|      	|      	|    	|    	|      	|
+| ..  	|      	|      	|      	|    	|    	|      	|
+| B_L 	| p_L1 	| p_L2 	| p_L3 	| .. 	| .. 	| p_LK 	|
+
+$H_{01}: p_{11} = p_{21} = \ldots = p_{L1}$  
+$H_{02}: p_{12} = p_{22} = \ldots = p_{L2}$  
+\vdots
+$H_{0K}: p_{1K} = p_{2K} = \ldots = p_{LK}$
+
+$m$ = $L \choose 2$ (unordered)  
+$m$ = $L-1$ (ordered)
+
+|  facets/distances	| A_1  	| A_2  	| A_3  	| .. 	| .. 	| A_K  	|
+|------------------	|------	|------	|------	|----	|----	|------	|
+| d_1              	| d_11 	| d_12 	| d_13 	| .. 	| .. 	| d_1K 	|
+| d_2              	|      	|      	|      	|    	|    	|      	|
+| d_3              	|      	|      	|      	|    	|    	|      	|
+| ..               	|      	|      	|      	|    	|    	|      	|
+| ..               	|      	|      	|      	|    	|    	|      	|
+| d_m              	| d_m1 	| d_m2 	| d_m3 	| .. 	| .. 	| d_mL 	|
+
+$H_{01}: d_{11} = d_{21} = \ldots = d_{m1} = 0$  
+$H_{02}: d_{12} = d_{22} = \ldots = d_{L2} = 0$  
+\vdots
+$H_{0m}: d_{1m} = d_{2m} = \ldots = d_{Lm} = 0$
+
+ - can do ANOVA at this stage
+ - interpretation of results (if interaction of levels significant when testing if means of distributions of distances are equal to zero )
+
+| facets/max-dist 	| A_1       	| A_2       | .. 	| .. 	| A_K       	|
+|-----------------	|-----------	|-----------	|-----------	|----	|----	|-----------	|
+| max-dist        	| max(d_11, .., dm1) 	| max(d_12, ...d_m2) 	| .. 	| .. 	| max(d_1K,...d_mK) 	|
+
+$H_{01}: max(d_{11}, .., d_{m1}) = 0$  
+$H_{02}: max(d_{12}, ...d_{m2})  = 0$  
+\vdots
+$H_{0K}: max(d_{1K},...d_{mK}) = 0$
+
+ -  normalised maximum distribution follows standardised Gumbel distribution
+ -  multiple hypothesis testing problem where p-values needs to be adjusted with Fisher's combination test (preferred) or Bonferroni's correction
+ - What is the test statistic for multiple hypothesis problem?
 
@@
 
-### Maximum distance between levels of x-axis variable and median across levels of facet variable
+normal: standard normal ordered distances  
+normal_nonstd: non-standard normal ordered distances  
+normal_un: standard normal unordered distances  
+normal_nonstd_un: non-standard normal unordered   distances
 
@@
 library(readr)
-smart_weibull <- read_rds("data/reverse_smart_weibull.rds")  %>% mutate(rank_weibull = row_number())
+smart_normal <- read_rds("data/ smart_harmony_normal.rds")  %>% mutate(rank_normal = row_number())
 
-smart_gamma <- read_rds("data/reverse_smart_gamma.rds")  %>% mutate(rank_gamma = row_number())
+smart_normal_nonstd <- read_rds("data/smart_harmony_nonstd.rds") %>% mutate(rank_normal_nonstnd = row_number())
 
-smart_normal <- read_rds("data/reverse_smart_normal.rds") %>% mutate(rank_normal = row_number())
+smart_normal_un <- read_rds("data/smart_harmony_normal_unordered.rds")  %>% mutate(rank_normal_un = row_number())
 
-smart_chi <-read_rds("data/reverse_smart_chisq.rds") %>% mutate(rank_chi = row_number())
-smart_general <- read_rds("data/reverse_smart_general.rds") %>% mutate(rank_general = row_number())
+smart_normal_nonstd_un <- read_rds("data/smart_harmony_nonstd_unordered.rds") %>% mutate(rank_normal_nonstnd_un = row_number())
 
-data_unite <- smart_chi %>% 
-  left_join(smart_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
-    left_join(smart_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
-  rename("gamma" = "mean_max_variation") %>% 
-  left_join(smart_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("normal" = "mean_max_variation") %>% 
-   left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("general" = "mean_max_variation") 
+smart_normal %>% 
+  left_join(smart_normal_nonstd, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
+   rename("normal" = "mean_max_variation.x", "normal_nonstd" = "mean_max_variation.y") %>%
+
+  left_join(smart_normal_un, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
+  rename("normal_un" = "mean_max_variation") %>%
+
+  left_join(smart_normal_nonstd_un, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
+  rename("normal_nonstd_un" = "mean_max_variation")%>% select(-c(normal, normal_nonstd, normal_un, normal_nonstd_un))%>%
+  dplyr::rename("normal"  = "rank_normal",
+         "normal_nonstd"  = "rank_normal_nonstnd",
+         "normal_un"  = "rank_normal_un",
+         "normal_nonstd_un"  = "rank_normal_nonstnd_un")%>%   knitr::kable(format = "latex")
 ```
 
-```{r}
-data_unite %>% 
-  select(-c(normal, gamma, weibull, general, chisq)) %>%
-  dplyr::rename("normal"  = "rank_normal",
-         "gamma"  = "rank_gamma",
-         "weibull"  = "rank_weibull",
-         "general"  = "rank_general",
-         "chi"  = "rank_chi") %>% 
-  knitr::kable(format = "latex")
+<!-- # unordered -->
 
-# data_unite %>%
-#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
-#   knitr::kable()
-```
+<!-- ```{r} -->
+<!-- library(readr) -->
+<!-- smart_normal <- read_rds("data/smart_harmony_normal_unordered.rds")  %>% mutate(rank_normal = row_number()) -->
 
-### Maximum distance between levels of facet variable and median across levels of x-axis variable
+<!-- smart_normal_nonstd <- read_rds("data/smart_harmony_nonstd_unordered.rds") %>% mutate(rank_normal_nonstnd = row_number()) -->
 
-```{r}
-library(readr)
-smart_weibull <- read_rds("data/smart_harmony_wb.rds")  %>% mutate(rank_weibull = row_number())
-
-smart_gamma <- read_rds("data/smart_gamma.rds")  %>% mutate(rank_gamma = row_number())
-
-smart_normal <- read_rds("data/smart_harmony_norm.rds") %>% mutate(rank_normal = row_number())
-
-smart_chi <-read_rds("data/smart_harmony_chisq.rds") %>% mutate(rank_chi = row_number())
-smart_general <- read_rds("data/smart_general.rds") %>% mutate(rank_general = row_number())
-
-data_unite <- smart_chi %>% 
-  left_join(smart_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
-    left_join(smart_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
-  rename("gamma" = "mean_max_variation") %>% 
-  left_join(smart_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("normal" = "mean_max_variation") %>% 
-   left_join(smart_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("general" = "mean_max_variation") 
-```
-
-```{r}
-data_unite %>% 
-  select(-c(normal, gamma, weibull, general, chisq)) %>% 
-  dplyr::rename("normal"  = "rank_normal",
-         "gamma"  = "rank_gamma",
-         "weibull"  = "rank_weibull",
-         "general"  = "rank_general",
-         "chi"  = "rank_chi") %>% 
-  knitr::kable(format = "latex")
-
-# data_unite %>%
-#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
-#   knitr::kable()
-```
+<!-- smart_normal %>%  -->
+<!--   left_join(smart_normal_nonstd, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  -->
+<!--    rename("normal" = "mean_max_variation.x", "normal_nonstd" = "mean_max_variation.y") %>%  -->
+<!--   select(-c(normal, normal_nonstd))%>% -->
+<!--   dplyr::rename("normal_unordered"  = "rank_normal", -->
+<!--          "normal_nonstd_unordered"  = "rank_normal_nonstnd")%>%   knitr::kable(format = "latex") -->
+<!-- ``` -->
 
@@
 
+```{r}
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>% prob_plot( gran1 = "day_week", gran2 = "day_month", plot_type = "boxplot")
+
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>% prob_plot(gran1 = "wknd_wday", gran2 = "day_month", plot_type = "boxplot")
+
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>% 
+  prob_plot(gran1 = "wknd_wday", gran2 = "hour_day", plot_type = "boxplot")
+
+```
+
 ## cricket data
@@
 harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)
-
-```
-
-### Distribution fitting of distances for the harmony pair (lag_field, over)
-
-```{r}
-cricket_data  %>% qqplot_distance(gran1 = "lag_field", gran2 = "over", response = "runs_per_over", hierarchy_tbl = hierarchy_model)
 ```
@@
 ```{r}
-cric_weibull <- read_rds("data/cric_weibull.rds") %>% mutate(rank_weibull = row_number())
-
-cric_gamma <-read_rds("data/cric_gamma.rds") %>% mutate(rank_gamma = row_number())
-
-cric_normal <-read_rds("data/cric_normal.rds")%>% mutate(rank_normal = row_number())
-
-cric_general <- read_rds("data/cric_general.rds")%>% mutate(rank_general = row_number())
-
-cric_chisq <- read_rds("data/cric_chisq.rds")%>% mutate(rank_chisq = row_number())
-
-data_unite <- cric_chisq %>% 
-  left_join(cric_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
-    left_join(cric_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
-  rename("gamma" = "mean_max_variation") %>% 
-  left_join(cric_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("normal" = "mean_max_variation") %>% 
-   left_join(cric_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("general" = "mean_max_variation") 
-```
-
-```{r}
-data_unite %>% 
-  select(-c(normal, gamma, weibull, general, chisq)) %>%
-  dplyr::rename("normal"  = "rank_normal",
-         "gamma"  = "rank_gamma",
-         "weibull"  = "rank_weibull",
-         "general"  = "rank_general",
-         "chi"  = "rank_chisq") %>% 
-  knitr::kable(format = "latex")
-
-# data_unite %>%
-#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
-#   knitr::kable()
-```
-
-### Maximum distance between levels of facet variable and median across levels of x-axis variable
-
-```{r}
-cric_weibull <- read_rds("data/ smart_cric_weibull.rds") %>% mutate(rank_weibull = row_number())
-
-cric_gamma <-read_rds("data/ smart_cric_gamma.rds") %>% mutate(rank_gamma = row_number())
-
-cric_normal <-read_rds("data/ smart_cric_normal.rds")%>% mutate(rank_normal = row_number())
-
-cric_general <- read_rds("data/ smart_cric_general.rds")%>% mutate(rank_general = row_number())
-
-cric_chisq <- read_rds("data/smart_cric_chisq.rds")%>% mutate(rank_chisq = row_number())
-
-data_unite <- cric_chisq %>% 
-  left_join(cric_weibull, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-   rename("chisq" = "mean_max_variation.x", "weibull" = "mean_max_variation.y") %>% 
-    left_join(cric_gamma,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  
-  rename("gamma" = "mean_max_variation") %>% 
-  left_join(cric_normal,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("normal" = "mean_max_variation") %>% 
-   left_join(cric_general,  by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("general" = "mean_max_variation") 
-```
-
-```{r}
-data_unite %>% 
-  select(-c(normal, gamma, weibull, general, chisq)) %>%
-  dplyr::rename("normal"  = "rank_normal",
-         "gamma"  = "rank_gamma",
-         "weibull"  = "rank_weibull",
-         "general"  = "rank_general",
-         "chi"  = "rank_chisq") %>% 
-  knitr::kable(format = "latex")
-
-# data_unite %>%
-#   select(-c(rank_normal, rank_gamma, rank_weibull, rank_general, rank_chi)) %>% 
-#   knitr::kable()
+cricket_data <- read_rds("data/cricket_data.rds")
+cricket_data %>%
+  filter(over!=1) %>%
+  prob_plot("over", "lag_field",
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "violin",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("b") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
 ```

### 2020-06-04 choosing_harmonies_Normal.Rmd
choosing threshold update

@@
 | ..               	|      	|      	|      	|    	|    	|      	|
-| d_m              	| d_m1 	| d_m2 	| d_m3 	| .. 	| .. 	| d_mL 	|
+| d_m              	| d_m1 	| d_m2 	| d_m3 	| .. 	| .. 	| d_mK 	|
 
 $H_{01}: d_{11} = d_{21} = \ldots = d_{m1} = 0$  
-$H_{02}: d_{12} = d_{22} = \ldots = d_{L2} = 0$  
+$H_{02}: d_{12} = d_{22} = \ldots = d_{m2} = 0$  
 \vdots
-$H_{0m}: d_{1m} = d_{2m} = \ldots = d_{Lm} = 0$
+$H_{0K}: d_{1K} = d_{2K} = \ldots = d_{mK} = 0$
 
@@
 
-```{r}
+```{r allgraph}
 library(readr)
@@
 ```{r, echo = FALSE}
-p1 <- smart_meter10 %>%
+smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@
   theme_minimal() + 
-  ggtitle("Rank 1 Section 6.1.1")  + 
   scale_x_discrete(breaks = seq(0, 23, 5))
 
-p2 <- smart_meter10 %>%
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>% prob_plot( gran1 = "day_week", gran2 = "day_month", plot_type = "boxplot") + scale_x_discrete(breaks = seq(0, 31, 5))
+
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>% prob_plot(gran1 = "wknd_wday", gran2 = "day_month", plot_type = "boxplot") + scale_x_discrete(breaks = seq(0, 31, 5))
+
+smart_meter10 %>%
+  filter(customer_id %in% c(10017936)) %>% 
+  prob_plot(gran1 = "wknd_wday", gran2 = "hour_day", plot_type = "boxplot")
+
+smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@
   ylab("electricity demand [KWh]") + xlab("") + ggtitle("") + ylab("")+ 
-  theme_minimal() + 
-  ggtitle("Rank 1 Section 6.1.2") + scale_x_discrete(labels = c('wday','wend'))
-
-```
-
-```{r, echo = FALSE}
-p3 <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("week_month",
-    "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "boxplot",
-    #symmetric = TRUE,
-    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal() + 
-  ggtitle("Rank 2 Section 6.1.2") + 
-  scale_x_discrete(breaks = seq(0, 23, 5))
-
-p4 <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("hour_day",
-    "week_month",
-    response = "general_supply_kwh",
-    plot_type = "boxplot",
-    #symmetric = TRUE,
-    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("week of the month") + ggtitle("") + ylab("")+ 
-  theme_minimal()  + 
-  ggtitle("Rank 2 Section 6.1.2")
-
-ggpubr::ggarrange(p1, p2,  ncol = 2)
-
-ggpubr::ggarrange(p3, p4,  ncol = 2)
-```
-
-```{r}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>% prob_plot( gran1 = "day_week", gran2 = "day_month", plot_type = "boxplot")
-
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>% prob_plot(gran1 = "wknd_wday", gran2 = "day_month", plot_type = "boxplot")
-
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>% 
-  prob_plot(gran1 = "wknd_wday", gran2 = "hour_day", plot_type = "boxplot")
+  theme_minimal() + scale_x_discrete(labels = c('wday','wend'))
 
@@
 
-```{r}
+```{r cricket}
 cricket_data <- read_rds("data/cricket_data.rds")
@@
 
-```{r}
+```{r cricket_lag}
 cricket_data <- read_rds("data/cricket_data.rds")

### 2020-06-12 main.Rmd
opening paragraph of section 4

@@
 
-<!-- # Why data structure for cyclic granularities exploration -->
-
-\noindent Effective exploration and visualization require proper data structures.
-It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.
-
-<!-- #recent data structures for temporal exploration -->
-
-\noindent A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble] data structure.
+<!-- # Why data structure and basis for cyclic granularities exploration -->
+
+\noindent Effective exploration and visualization benefits from well-organized data structures.
+@wang2019tsibble introduced a tidy data structure, tsibble, to support exploration and modeling of temporal data.
+This forms the basis of the structure for cyclic granularities.
 A tsibble consists of an index, key and measured variables.
@@
 A bottom granularity is represented by the index of the tsibble.
+<!-- It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
+Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.-->
+
+<!-- #recent data structures for temporal exploration -->
 
@@
 
-\noindent All cyclic granularities can be expressed in terms of the index set, and hence, we consider the data structure in Figure \@ref(fig:data-structure) for exploration of temporal data of this kind.
-This is a two-dimensional array extending the columns of tsibble by including the cyclic granularities.
-The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
+\noindent All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure).
+This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.\todo[inline]{think your diagram should color what's tsibble differently from what's augmented} The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
 Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.

### 2020-06-17 main.Rmd
change color for tsibble and gravitas data structure

@@
 \noindent All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure).
-This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.\todo[inline]{think your diagram should color what's tsibble differently from what's augmented} The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
+This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
 Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.
@@
 ```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
-include_graphics("Figs/data-struc-rev.png")
+include_graphics("Figs/data-struc-diffcol.png")
 ```

### 2020-07-15 choosing_harmonies_Normal.Rmd
all before updating gravitas package and makeing rank harmony efficient

@@
 library(ggpubr)
+library(readr)
 opts_chunk$set(
@@
 | ..               	|      	|      	|      	|    	|    	|      	|
-| d_m              	| d_m1 	| d_m2 	| d_m3 	| .. 	| .. 	| d_mK 	|
+| d_m              	| d_m1 	| d_m2 	| d_m3 	 .. 	| .. 	| d_mK 	|
 
@@
  - What is the test statistic for multiple hypothesis problem?
+
+Permutation test:
+
+<!-- paiwise percentile tests for distances.
+Each facet is a category with m ordered/unordered pairwise distances.
+We are testing if the 95th percentile of these facets are equal.
+If at least two of them are different, we keep the harmony pair, else remove them from the list.
+-->
+
+<!-- 1st screening criterion: 95th percentile different for different facet labels -->
+
+<!-- 2nd screening/ranking criterion: within the harmonies foir which facet levels are different, MMPD is used to rank them -->
+
+**Assumption:** random permutation without considering ordering 
+(Local)
+
+1. Given the original sequence for $\{C_i, C_j\}$; $\{v_t: t=0, 1, 2, \dots, T-1\}$, the MMPD is computed and is represented by $MMPD_{obs}$.
+
+2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
+
+3. MMPD is computed for all random permutation and is represented by $MMPD_{sample}$.
+
+4.  Steps (2) and (3) are repeated a large number
+of times M (e.g. 1000).
+
+5. For each permutation, one $MMPD_{sample}$ value is obtained.
+
+6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
+
+7. If  $MMPD_{obs}> MMPD_{threshold}$, then it implies the observed MMPD is 95% of the times higher than the any other random permutation of the original sequence.
+
+Pros: Considering thresholds locally for each harmony pairs would imply that data is reshuffled only within that harmony pair and generated MMPDs are only reflection of how MMPDs are when measured between that harmony pairs only.
+
+Cons: The threshold value is considered locally for each harmony pairs and hence it does not align the original MMPD values we get.
+That is, a harmony pair with lower MMPD might get selected over that with a higher MMPD.
+Difficult to understand but it means that although MMPD value is smaller, it is significantly different from zero.
+
+**Assumption:** random permutation without considering ordering 
+(Local)
+
+1. Given the data; $\{v_t: t=0, 1, 2, \dots, T-1\}$, the MMPD is computed and is represented by $MMPD_{obs}$.
+
+2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
+
+3. MMPD is computed for all random permutation of the data and is represented by $MMPD_{sample}$.
+
+4.  Steps (2) and (3) are repeated a large number
+of times M (e.g. 1000).
+
+5. For each permutation, one $MMPD_{sample}$ value is obtained.
+
+6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
+
+7. If  $MMPD_{obs}> MMPD_{threshold}$, harmony pairs are accepted. Onlyone threshold for all harmony pairs.
+
+Pros: Considering thresholds global for all harmony pairs would imply less computation time.
+
+Cons: Only one threshold for all harmony pairs might not be an appropriate measure but a good benchmark.
+
+```{r}
+library(readr)
+signi <- read_rds("data/significant_harmonies.rds")
+signi
+local_thres <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE, FALSE, FALSE)
+
+signi %>% select(-facet_levels, -x_levels) %>% 
+  mutate(local_threshold = local_thres) %>% knitr::kable(format = "latex")
+
+```
+
+<!-- **Assumption:** random permutation considering ordering -->
+
+<!-- Intuition? -->
+
+<!-- Original: -->
+
+<!-- | facets/observation     	| 1  	| 2  	| 3  	| .. 	| .. 	| n  	| -->
+<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
+<!-- | B_1 	| x_1 	| x_2 	| x_3 	| .. 	| .. 	| x_n 	| -->
+<!-- | B_2 	| y_1 	| y_2 	| y_3 	| .. 	| .. 	| y_n 	| -->
+<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
+<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
+<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
+<!-- | B_L 	| z_1 	| z_2 	| z_3 	| .. 	| .. 	| z_n 	| -->
+
+<!-- Unordered: Permuting across levels -->
+
+<!-- | facets/observation     	| 1  	| 2  	| 3  	| .. 	| .. 	| n  	| -->
+<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
+<!-- | B_1 	| y_3 	| x_2 	| z_2 	| .. 	| .. 	| y_1 	| -->
+<!-- | B_2 	| x_n 	| y_2 	| x_1 	| .. 	| .. 	| z_3 	| -->
+<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
+<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
+<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
+<!-- | B_L 	| x_10 	| x_3 	| y_n 	| .. 	| .. 	| x_5 	| -->
+
+<!-- Ordered: Permuting from the same levels -->
+
+<!-- | facets/observation     	| 1  	| 2  	| 3  	| .. 	| .. 	| n  	| -->
+<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
+<!-- | B_1 	| x_3 	| x_10 	| x_100 	| .. 	| .. 	| x_1 	| -->
+<!-- | B_2 	| y_n 	| y_2 	| y_10 	| .. 	| .. 	| y_3 	| -->
+<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
+<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
+<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
+<!-- | B_L 	| z_10 	| z_200 	| z_10 	| .. 	| .. 	| z_5 	| -->
+
+```{r}
+sm <- smart_meter10 %>%
+filter(customer_id %in% c(10017936))
+.data = sm
+sm %>% prob_plot("week_month", "day_week", plot_type = "boxplot")
+
+sm %>% prob_plot("day_month", "hour_day", plot_type = "quantile", symmetric = FALSE, quantile_prob = c(0.05, 0.25, 0.5, 0.75, 0.95))
+
+```
 
@@
 
-```{r cricket}
-cricket_data <- read_rds("data/cricket_data.rds")
+# error after updating package and tsibble
 
-  hierarchy_model <- tibble::tibble(
-      units = c("index", "over", "inning", "match"),
-      convert_fct = c(1, 20, 2, 1))
+<!-- ```{r cricket} -->
+<!-- cricket_data <- read_rds("data/cricket_data.rds") -->
 
-hierarchy_tbl <- hierarchy_model
+<!--   hierarchy_model <- tibble::tibble( -->
+<!--       units = c("index", "over", "inning", "match"), -->
+<!--       convert_fct = c(1, 20, 2, 1)) -->
 
-response <- "runs_per_over"
+<!-- hierarchy_tbl <- hierarchy_model -->
 
-harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)
-```
+<!-- response <- "runs_per_over" -->
 
-```{r cricket_lag}
-cricket_data <- read_rds("data/cricket_data.rds")
-cricket_data %>%
-  filter(over!=1) %>%
-  prob_plot("over", "lag_field",
-  hierarchy_model,
-  response = "run_rate",
-  plot_type = "violin",
-  symmetric = FALSE,
- quantile_prob = c(0.25, 0.5, 0.75)) +
-   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
-  ylab("runs per over")  +
-  xlab("number of wickets in previous over") +
-  ggtitle("b") +
-   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
-```
+<!-- harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model) -->
+<!-- ``` -->
 
-# Bibliography
+<!-- ```{r cricket_lag} -->
+<!-- cricket_data <- read_rds("data/cricket_data.rds") -->
+<!-- cricket_data %>% -->
+<!--   filter(over!=1) %>% -->
+<!--   prob_plot("over", "lag_field", -->
+<!--   hierarchy_model, -->
+<!--   response = "run_rate", -->
+<!--   plot_type = "violin", -->
+<!--   symmetric = FALSE, -->
+<!--  quantile_prob = c(0.25, 0.5, 0.75)) + -->
+<!--    #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") + -->
+<!--   ylab("runs per over")  + -->
+<!--   xlab("number of wickets in previous over") + -->
+<!--   ggtitle("b") + -->
+<!--    theme(plot.title = element_text(face = "bold")) +  theme_minimal() -->
+<!-- ``` -->
 
-```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib")
-```
+<!-- # Bibliography -->
+
+<!-- ```{r write-bib, eval = FALSE} -->
+<!-- write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib") -->
+<!-- ``` -->
+
+# Does normalisation work?

### 2020-07-16 choosing_harmonies.Rmd
choosing harmonies pdf update
add all ranking check

@@
   rename("general" = "mean_max_variation")
+
 ```
@@
 ggpubr::ggarrange(p3, p4,  ncol = 2)
+```
+
+```{r}
+prob_plot(.data,  gran1 = "day_week", gran2 = "day_month", plot_type = "boxplot")
+
+prob_plot(.data,  gran1 = "wknd_wday", gran2 = "day_month", plot_type = "boxplot")
+
+prob_plot(.data,  gran1 = "wknd_wday", gran2 = "hour_day", plot_type = "boxplot")
+
 ```

### 2020-07-16 choosing_harmonies_Normal.Rmd
add all ranking check

@@
 One of the potential ways to evaluate this variation is by computing the pairwise distances between the distributions of the measured variable.
-We do this through Jensen-Shannon divergence which is based on Kullback-Leibler divergence.
+We do this through Jensen-Shannon distance which is based on Kullback-Leibler divergence.
 Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
@@
 The normalizing constants, however, vary depending on the underlying distribution and hence it is important to assume a distribution of distances in our case.
+
+```{r rank harmony}
+library(gravitas)
+library(ggplot2)
+library(dplyr)
+sm <- smart_meter10 %>%
+filter(customer_id %in% c("10017936"))
+harmonies <- sm %>%
+harmony(ugran = "month",
+filter_in = "wknd_wday",
+filter_out = c("hhour", "fortnight"))
+.data = sm
+response  = "general_supply_kwh"
+harmony_tbl =  harmonies
+smart_harmony <- .data %>% rank_harmony(harmony_tbl = harmonies,
+response = "general_supply_kwh", dist_ordered = FALSE)
+```
 
@@
 
-# error after updating package and tsibble
+```{r cricket}
+cricket_data <- read_rds("data/cricket_data.rds")
 
-<!-- ```{r cricket} -->
-<!-- cricket_data <- read_rds("data/cricket_data.rds") -->
+  hierarchy_model <- tibble::tibble(
+      units = c("index", "over", "inning", "match"),
+      convert_fct = c(1, 20, 2, 1))
 
-<!--   hierarchy_model <- tibble::tibble( -->
-<!--       units = c("index", "over", "inning", "match"), -->
-<!--       convert_fct = c(1, 20, 2, 1)) -->
+hierarchy_tbl <- hierarchy_model
 
-<!-- hierarchy_tbl <- hierarchy_model -->
+response <- "runs_per_over"
 
-<!-- response <- "runs_per_over" -->
+harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)
+```
 
-<!-- harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model) -->
-<!-- ``` -->
-
-<!-- ```{r cricket_lag} -->
-<!-- cricket_data <- read_rds("data/cricket_data.rds") -->
-<!-- cricket_data %>% -->
-<!--   filter(over!=1) %>% -->
-<!--   prob_plot("over", "lag_field", -->
-<!--   hierarchy_model, -->
-<!--   response = "run_rate", -->
-<!--   plot_type = "violin", -->
-<!--   symmetric = FALSE, -->
-<!--  quantile_prob = c(0.25, 0.5, 0.75)) + -->
-<!--    #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") + -->
-<!--   ylab("runs per over")  + -->
-<!--   xlab("number of wickets in previous over") + -->
-<!--   ggtitle("b") + -->
-<!--    theme(plot.title = element_text(face = "bold")) +  theme_minimal() -->
-<!-- ``` -->
+```{r cricket_lag}
+cricket_data <- read_rds("data/cricket_data.rds")
+cricket_data %>%
+  filter(over!=1) %>%
+  prob_plot("over", "lag_field",
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "violin",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("b") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
+```
 
@@
 # Does normalisation work?
+
+```{r}
+library(ggplot2)
+library(gravitas)
+sm <- smart_meter10 %>% dplyr::filter(customer_id %in% c("10017936"))
+
+harmonies <- sm %>% 
+  harmony(ugran = "month",
+          filter_in = "wknd_wday",
+          filter_out = c("hhour", "fortnight"))
+
+.data = sm
+response  = "general_supply_kwh"
+harmony_tbl =  harmonies
+smart_harmony <- .data %>% gravitas::rank_harmony(harmony_tbl = harmonies,
+response = "general_supply_kwh", dist_ordered = FALSE)
+
+smart_harmony %>% ggplot(aes(x = x_levels, y =  r)) + geom_point()
+```

### 2020-07-16 threshold.Rmd
add all ranking check

---
title: "Threshold computation"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 1. Given the original sequence for all harmony pairs; $\{v_t: t=0, 1, 2, \dots, T-1\}$, the $MMPD$ is computed for all harmony pairs and is represented by the vector $MMPD_{obs}$, each element of the vector corresponding to the $MMPD$ of each harmony pair.
 2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
 3. MMPD is computed for all harmony pairs for this permutation and is represented by the vector $MMPD_{sample}$.
 4. Steps (2) and (3) are repeated a large number of times M (e.g. 1000).
 5. For each permutation, one $MMPD_{sample}$ vector is obtained.
 6. $95^{th}$ percentile of all elements of the $MMPD_{sample}$ from different permutations is computed and stored in $MMPD_{threshold}$.
 7. Harmony pairs for which $MMPD_{obs} > MMPD_{threshold}$ are chosen.

### 2020-07-16 choosing_harmonies_Normal4.Rmd
cleaning and arranging files

--- 2020-07-16 choosing_harmonies_Normal.Rmd
@@
 The normalizing constants, however, vary depending on the underlying distribution and hence it is important to assume a distribution of distances in our case.
+
+# Does normalisation work?
+(Show with by comparing with maximum and 
+showing for similar levels)
+# Does it match with the threshold?
+# Should they tally with each other?
 
@@
 
-<!-- # Bibliography -->
-
-<!-- ```{r write-bib, eval = FALSE} -->
-<!-- write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib") -->
-<!-- ``` -->
-
-# Does normalisation work?
-
-```{r}
+```{r smart_harmony}
 library(ggplot2)
@@
 
-.data = sm
-response  = "general_supply_kwh"
 harmony_tbl =  harmonies
-smart_harmony <- .data %>% gravitas::rank_harmony(harmony_tbl = harmonies,
-response = "general_supply_kwh", dist_ordered = FALSE)
+
+smart_harmony <- sm %>% 
+  rank_harmony(harmony_tbl = harmonies,
+               response = "general_supply_kwh", 
+               dist_ordered = FALSE)
+
+smart_harmony %>% kable()
 
 smart_harmony %>% ggplot(aes(x = x_levels, y =  r)) + geom_point()
+
+smart_harmony %>% ggplot(aes(x = facet_levels, y =  r)) + geom_point()
 ```

### 2020-07-16 data-raw/sm_cust50.R
cleaning folder

@@ -1,41 +1,0 @@
-library(tidyverse)
-library(tsibble)
-#devtools::install_github("jimhester/archive")
-library(archive)
-
-# Raw data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in
-# https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter
-# file name:CD_INTERVAL_READING_ALL_NO_QUOTES.csv
-
-download.file(
-  "http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z",
-  "data-raw/smart-metre.7z",
-  mode = "wb"
-)
-smart_meter_data_raw <- archive_read("data-raw/smart-metre.7z") %>%
-  read_csv(n_max = 3e6)
-
-#smart_meter_data_raw <- read_csv("data-raw/CD_INTERVAL_READING_ALL_NO_QUOTES.csv", n_max = 3e6)
-
-set.seed(12345)
-sm_cust <- smart_meter_data_raw %>%
-  distinct(CUSTOMER_ID) %>%
-  pull(CUSTOMER_ID) %>%
-  sample(size= 50)
-
-sm_cust50 <- smart_meter_data_raw %>%
-  filter(CUSTOMER_ID %in% sm_cust) %>%
-  rename_all(tolower) %>%
-  arrange(customer_id, reading_datetime) %>%
-  group_by(customer_id) %>%
-  mutate(
-    reading_datetime = case_when(
-    duplicated(reading_datetime) ~ reading_datetime + lubridate::hours(1),
-    TRUE ~ reading_datetime
-  )) %>% 
-  ungroup() %>%
-  mutate(customer_id = as.character(customer_id)) %>%
-  as_tsibble(index = reading_datetime, key = customer_id) %>%
-  select(-calendar_key)
-
-save(sm_cust50, file = "data/sm_cust50.rds", compress = "xz")

### 2020-07-16 norm_threshold4.Rmd
still cleaning folder

--- 2020-07-16 choosing_harmonies_Normal4.Rmd
@@
 ---
-title: Screening harmonies
+title: Finalised versions of normalisation and threshold
 authors:
@@
 # Does normalisation work?
-(Show with by comparing with maximum and 
-showing for similar levels)
-# Does it match with the threshold?
-# Should they tally with each other?
+(Show with by comparing with maximum and showing for similar levels)
+
+```{r smart_harmony}
+library(ggplot2)
+library(gravitas)
+sm <- smart_meter10 %>% dplyr::filter(customer_id %in% c("10017936"))
+
+harmonies <- sm %>% 
+  harmony(ugran = "month",
+          filter_in = "wknd_wday",
+          filter_out = c("hhour", "fortnight"))
+
+harmony_tbl =  harmonies
+
+smart_harmony <- sm %>% 
+  rank_harmony(harmony_tbl = harmonies,
+               response = "general_supply_kwh", 
+               dist_ordered = FALSE)
+
+smart_harmony %>% kable()
+
+smart_harmony %>% ggplot(aes(x = x_levels, y =  r)) + geom_point()
+
+smart_harmony %>% ggplot(aes(x = facet_levels, y =  r)) + geom_point()
+```
 
@@
 
-| facets/x-axis     	| A_1  	| A_2  	| A_3  	| .. 	| .. 	| A_K  	|
-|-----	|------	|------	|------	|----	|----	|------	|
-| B_1 	| p_11 	| p_12 	| p_13 	| .. 	| .. 	| p_1K 	|
-| B_2 	|      	|      	|      	|    	|    	|      	|
-| B_3 	|      	|      	|      	|    	|    	|      	|
-| ..  	|      	|      	|      	|    	|    	|      	|
-| ..  	|      	|      	|      	|    	|    	|      	|
-| B_L 	| p_L1 	| p_L2 	| p_L3 	| .. 	| .. 	| p_LK 	|
-
-$H_{01}: p_{11} = p_{21} = \ldots = p_{L1}$  
-$H_{02}: p_{12} = p_{22} = \ldots = p_{L2}$  
-\vdots
-$H_{0K}: p_{1K} = p_{2K} = \ldots = p_{LK}$
-
-$m$ = $L \choose 2$ (unordered)  
-$m$ = $L-1$ (ordered)
-
-|  facets/distances	| A_1  	| A_2  	| A_3  	| .. 	| .. 	| A_K  	|
-|------------------	|------	|------	|------	|----	|----	|------	|
-| d_1              	| d_11 	| d_12 	| d_13 	| .. 	| .. 	| d_1K 	|
-| d_2              	|      	|      	|      	|    	|    	|      	|
-| d_3              	|      	|      	|      	|    	|    	|      	|
-| ..               	|      	|      	|      	|    	|    	|      	|
-| ..               	|      	|      	|      	|    	|    	|      	|
-| d_m              	| d_m1 	| d_m2 	| d_m3 	 .. 	| .. 	| d_mK 	|
-
-$H_{01}: d_{11} = d_{21} = \ldots = d_{m1} = 0$  
-$H_{02}: d_{12} = d_{22} = \ldots = d_{m2} = 0$  
-\vdots
-$H_{0K}: d_{1K} = d_{2K} = \ldots = d_{mK} = 0$
-
- - can do ANOVA at this stage
- - interpretation of results (if interaction of levels significant when testing if means of distributions of distances are equal to zero )
-
-| facets/max-dist 	| A_1       	| A_2       | .. 	| .. 	| A_K       	|
-|-----------------	|-----------	|-----------	|-----------	|----	|----	|-----------	|
-| max-dist        	| max(d_11, .., dm1) 	| max(d_12, ...d_m2) 	| .. 	| .. 	| max(d_1K,...d_mK) 	|
-
-$H_{01}: max(d_{11}, .., d_{m1}) = 0$  
-$H_{02}: max(d_{12}, ...d_{m2})  = 0$  
-\vdots
-$H_{0K}: max(d_{1K},...d_{mK}) = 0$
-
- -  normalised maximum distribution follows standardised Gumbel distribution
- -  multiple hypothesis testing problem where p-values needs to be adjusted with Fisher's combination test (preferred) or Bonferroni's correction
- - What is the test statistic for multiple hypothesis problem?
-
 Permutation test:
 
-<!-- paiwise percentile tests for distances.
-Each facet is a category with m ordered/unordered pairwise distances.
-We are testing if the 95th percentile of these facets are equal.
-If at least two of them are different, we keep the harmony pair, else remove them from the list.
--->
-
-<!-- 1st screening criterion: 95th percentile different for different facet labels -->
-
-<!-- 2nd screening/ranking criterion: within the harmonies foir which facet levels are different, MMPD is used to rank them -->
-
 **Assumption:** random permutation without considering ordering 
-(Local)
-
-1. Given the original sequence for $\{C_i, C_j\}$; $\{v_t: t=0, 1, 2, \dots, T-1\}$, the MMPD is computed and is represented by $MMPD_{obs}$.
-
-2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
-
-3. MMPD is computed for all random permutation and is represented by $MMPD_{sample}$.
-
-4.  Steps (2) and (3) are repeated a large number
-of times M (e.g. 1000).
-
-5. For each permutation, one $MMPD_{sample}$ value is obtained.
-
-6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
-
-7. If  $MMPD_{obs}> MMPD_{threshold}$, then it implies the observed MMPD is 95% of the times higher than the any other random permutation of the original sequence.
-
-Pros: Considering thresholds locally for each harmony pairs would imply that data is reshuffled only within that harmony pair and generated MMPDs are only reflection of how MMPDs are when measured between that harmony pairs only.
-
-Cons: The threshold value is considered locally for each harmony pairs and hence it does not align the original MMPD values we get.
-That is, a harmony pair with lower MMPD might get selected over that with a higher MMPD.
-Difficult to understand but it means that although MMPD value is smaller, it is significantly different from zero.
-
-**Assumption:** random permutation without considering ordering 
-(Local)
+(global)
 
@@
 
-```{r}
-library(readr)
-signi <- read_rds("data/significant_harmonies.rds")
-signi
-local_thres <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE, FALSE, FALSE)
-
-signi %>% select(-facet_levels, -x_levels) %>% 
-  mutate(local_threshold = local_thres) %>% knitr::kable(format = "latex")
-
-```
-
-<!-- **Assumption:** random permutation considering ordering -->
-
-<!-- Intuition? -->
-
-<!-- Original: -->
-
-<!-- | facets/observation     	| 1  	| 2  	| 3  	| .. 	| .. 	| n  	| -->
-<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
-<!-- | B_1 	| x_1 	| x_2 	| x_3 	| .. 	| .. 	| x_n 	| -->
-<!-- | B_2 	| y_1 	| y_2 	| y_3 	| .. 	| .. 	| y_n 	| -->
-<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
-<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
-<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
-<!-- | B_L 	| z_1 	| z_2 	| z_3 	| .. 	| .. 	| z_n 	| -->
-
-<!-- Unordered: Permuting across levels -->
-
-<!-- | facets/observation     	| 1  	| 2  	| 3  	| .. 	| .. 	| n  	| -->
-<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
-<!-- | B_1 	| y_3 	| x_2 	| z_2 	| .. 	| .. 	| y_1 	| -->
-<!-- | B_2 	| x_n 	| y_2 	| x_1 	| .. 	| .. 	| z_3 	| -->
-<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
-<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
-<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
-<!-- | B_L 	| x_10 	| x_3 	| y_n 	| .. 	| .. 	| x_5 	| -->
-
-<!-- Ordered: Permuting from the same levels -->
-
-<!-- | facets/observation     	| 1  	| 2  	| 3  	| .. 	| .. 	| n  	| -->
-<!-- |-----	|------	|------	|------	|----	|----	|------	| -->
-<!-- | B_1 	| x_3 	| x_10 	| x_100 	| .. 	| .. 	| x_1 	| -->
-<!-- | B_2 	| y_n 	| y_2 	| y_10 	| .. 	| .. 	| y_3 	| -->
-<!-- | B_3 	|      	|      	|      	|    	|    	|      	| -->
-<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
-<!-- | ..  	|      	|      	|      	|    	|    	|      	| -->
-<!-- | B_L 	| z_10 	| z_200 	| z_10 	| .. 	| .. 	| z_5 	| -->
-
-```{r}
-sm <- smart_meter10 %>%
-filter(customer_id %in% c(10017936))
-.data = sm
-sm %>% prob_plot("week_month", "day_week", plot_type = "boxplot")
-
-sm %>% prob_plot("day_month", "hour_day", plot_type = "quantile", symmetric = FALSE, quantile_prob = c(0.05, 0.25, 0.5, 0.75, 0.95))
-
-```
-
-# Results
-
-##  Smart meter data
-
-normal: standard normal ordered distances  
-normal_nonstd: non-standard normal ordered distances  
-normal_un: standard normal unordered distances  
-normal_nonstd_un: non-standard normal unordered   distances
-
-```{r allgraph}
-library(readr)
-smart_normal <- read_rds("data/ smart_harmony_normal.rds")  %>% mutate(rank_normal = row_number())
-
-smart_normal_nonstd <- read_rds("data/smart_harmony_nonstd.rds") %>% mutate(rank_normal_nonstnd = row_number())
-
-smart_normal_un <- read_rds("data/smart_harmony_normal_unordered.rds")  %>% mutate(rank_normal_un = row_number())
-
-smart_normal_nonstd_un <- read_rds("data/smart_harmony_nonstd_unordered.rds") %>% mutate(rank_normal_nonstnd_un = row_number())
-
-smart_normal %>% 
-  left_join(smart_normal_nonstd, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-   rename("normal" = "mean_max_variation.x", "normal_nonstd" = "mean_max_variation.y") %>%
-
-  left_join(smart_normal_un, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("normal_un" = "mean_max_variation") %>%
-
-  left_join(smart_normal_nonstd_un, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>% 
-  rename("normal_nonstd_un" = "mean_max_variation")%>% select(-c(normal, normal_nonstd, normal_un, normal_nonstd_un))%>%
-  dplyr::rename("normal"  = "rank_normal",
-         "normal_nonstd"  = "rank_normal_nonstnd",
-         "normal_un"  = "rank_normal_un",
-         "normal_nonstd_un"  = "rank_normal_nonstnd_un")%>%   knitr::kable(format = "latex")
-```
-
-<!-- # unordered -->
-
-<!-- ```{r} -->
-<!-- library(readr) -->
-<!-- smart_normal <- read_rds("data/smart_harmony_normal_unordered.rds")  %>% mutate(rank_normal = row_number()) -->
-
-<!-- smart_normal_nonstd <- read_rds("data/smart_harmony_nonstd_unordered.rds") %>% mutate(rank_normal_nonstnd = row_number()) -->
-
-<!-- smart_normal %>%  -->
-<!--   left_join(smart_normal_nonstd, by = c("facet_variable", "x_variable", "facet_levels", "x_levels")) %>%  -->
-<!--    rename("normal" = "mean_max_variation.x", "normal_nonstd" = "mean_max_variation.y") %>%  -->
-<!--   select(-c(normal, normal_nonstd))%>% -->
-<!--   dplyr::rename("normal_unordered"  = "rank_normal", -->
-<!--          "normal_nonstd_unordered"  = "rank_normal_nonstnd")%>%   knitr::kable(format = "latex") -->
-<!-- ``` -->
-
-## Graphical evidence
-
-```{r, echo = FALSE}
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("wknd_wday",
-    "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "boxplot",
-    #symmetric = TRUE,
-    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal() + 
-  scale_x_discrete(breaks = seq(0, 23, 5))
-
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>% prob_plot( gran1 = "day_week", gran2 = "day_month", plot_type = "boxplot") + scale_x_discrete(breaks = seq(0, 31, 5))
-
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>% prob_plot(gran1 = "wknd_wday", gran2 = "day_month", plot_type = "boxplot") + scale_x_discrete(breaks = seq(0, 31, 5))
-
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>% 
-  prob_plot(gran1 = "wknd_wday", gran2 = "hour_day", plot_type = "boxplot")
-
-smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("hour_day",
-    "wknd_wday",
-    response = "general_supply_kwh",
-    plot_type = "boxplot",
-    #symmetric = TRUE,
-    #quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("") + ggtitle("") + ylab("")+ 
-  theme_minimal() + scale_x_discrete(labels = c('wday','wend'))
-
-```
+# Does it match with the threshold?
+# Should they tally with each other?
 
@@
 ```
-
-```{r smart_harmony}
-library(ggplot2)
-library(gravitas)
-sm <- smart_meter10 %>% dplyr::filter(customer_id %in% c("10017936"))
-
-harmonies <- sm %>% 
-  harmony(ugran = "month",
-          filter_in = "wknd_wday",
-          filter_out = c("hhour", "fortnight"))
-
-harmony_tbl =  harmonies
-
-smart_harmony <- sm %>% 
-  rank_harmony(harmony_tbl = harmonies,
-               response = "general_supply_kwh", 
-               dist_ordered = FALSE)
-
-smart_harmony %>% kable()
-
-smart_harmony %>% ggplot(aes(x = x_levels, y =  r)) + geom_point()
-
-smart_harmony %>% ggplot(aes(x = facet_levels, y =  r)) + geom_point()
-```

### 2020-07-23 norm_threshold5.Rmd
adding recent simulation files

--- 2020-07-16 norm_threshold4.Rmd
@@
 
-# Does normalisation work?
-(Show with by comparing with maximum and showing for similar levels)
+```{r smart_harmony}
 
-```{r smart_harmony}
-library(ggplot2)
-library(gravitas)
 sm <- smart_meter10 %>% dplyr::filter(customer_id %in% c("10017936"))
@@
                response = "general_supply_kwh", 
-               dist_ordered = FALSE)
+               dist_ordered = TRUE)
 
-smart_harmony %>% kable()
+smart_harmony %>% 
+  mutate(MMPD = round(MMPD, 3), max_pd = round(max_pd, 3)) %>% 
+  mutate(rankn = row_number()) %>%
+  rename("rankun" = "r") %>% kable()
+```
 
-smart_harmony %>% ggplot(aes(x = x_levels, y =  r)) + geom_point()
+```{r cricket_harmony}
 
-smart_harmony %>% ggplot(aes(x = facet_levels, y =  r)) + geom_point()
+cricket_data <- read_rds("data/cricket_data.rds")
+
+  hierarchy_model <- tibble::tibble(
+      units = c("index", "over", "inning", "match"),
+      convert_fct = c(1, 20, 2, 1))
+
+.data <- cricket_data
+
+hierarchy_tbl <- hierarchy_model
+
+response <- "runs_per_over"
+
+harmonies_cric <- cricket_data %>% 
+  harmony(lgran = "over",
+          ugran = "match", 
+          filter_in = c("lag_field", "over"),
+          hierarchy_tbl = hierarchy_model)
+
+cric_harmony <- cricket_data %>% 
+  rank_harmony(harmony_tbl = harmonies_cric, 
+               response = response,
+               hierarchy_tbl = hierarchy_model,
+               dist_ordered = TRUE) %>% 
+  dplyr::filter(!(facet_variable=="over_inning"|x_variable =="over_inning"))
+
+cric_harmony %>% 
+  mutate(rankn = row_number()) %>%
+  rename("rankun" = "r") %>% 
+  mutate(MMPD = round(MMPD, 3), max_pd = round(max_pd, 3)) %>% 
+  kable()
+```
+
+# Does normalisation work?
+
+# Comparing similar levels
+
+check (inning_match, over) and (lag_field, over) with similar levels.
+(inning_match, over) is ranked above (lag_field, over). Let us see if it makes sense graphically.
+
+## indicating difference in variation better than max (2 by 2)
+
+```{r cric_graph_evidence}
+p1 <- cricket_data %>%
+  dplyr::filter(over!=1) %>%
+  prob_plot("inning_match", "lag_field",
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "quantile",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
+
+p2 <- cricket_data %>%
+  dplyr::filter(over!=1) %>%
+  prob_plot("lag_field", "inning_match",
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "quantile",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
+
+ggpubr::ggarrange(p1, p2, nrow = 2, labels = c("MMPD = 0.08,  max_pd = 0.068", "MMPD = 0.06,  max_pd = 0.064")) + ggtitle("Normalised max better indicating difference")
+```
+
+## indicating difference in variation for harmonies with same levels (20 by 2)
+
+```{r cric_samelev}
+
+p1 <- cricket_data %>%
+  dplyr::filter(over!=1) %>%
+  prob_plot("over","lag_field",
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "boxplot",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("MMPD = .058,  max_pd = 0.368") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
+
+p2 <- cricket_data %>%
+  dplyr::filter(over!=1) %>%
+  prob_plot("over", "inning_match", 
+  hierarchy_model,
+  response = "run_rate",
+  plot_type = "boxplot",
+  symmetric = FALSE,
+ quantile_prob = c(0.25, 0.5, 0.75)) +
+   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
+  ylab("runs per over")  +
+  xlab("number of wickets in previous over") +
+  ggtitle("MMPD = 0.039,  max_pd = 0.201") +
+   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
+p1
+p2
+
+#ggpubr::ggarrange(p1, p2, ncol = 2, labels = c()) + ggtitle("Normalised max better indicating difference")
+```
+
+## How does it compare with taking unnormalised maximums?
+
+```{r normalisation_work}
+smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) +geom_point()
+smart_harmony %>% mutate(lev = facet_levels*x_levels, r_MMPD = row_number()) %>% ggplot(aes(x = lev, y = r_MMPD)) +geom_line()
 ```
@@
 response = "general_supply_kwh", dist_ordered = FALSE)
+```
+
+# How does the normalization work for the x-levels?
+
+Let us fix the facet-level and see for different x-levels (31, 2) and (31, 7). Contrary to normalised max, non-normalised max ranks (31, 7) higher than (31, 2). Let us see which has higher variation.
+
+```{r same-facet-level}
+sm <- smart_meter10 %>%
+filter(customer_id %in% c(10017936))
+.data = sm
+p1 <- sm %>% prob_plot("day_month", "wknd_wday", plot_type = "boxplot", symmetric = FALSE)  + 
+  scale_y_log10()
+
+p2 <- sm %>% prob_plot("day_month", "day_week", plot_type = "boxplot", symmetric = FALSE, quantile_prob = c(0.05, 0.25, 0.5, 0.75, 0.95)) + 
+  scale_y_log10()
+
+p1
+p2
+```
+
+# How does the normalization work for the facet-levels?
+
+Let us fix the x-level and see for different facet-levels.
+(2, 24) and (7, 24). Contrary to normalised max, non-normalised max ranks (7, 24)  higher than (2, 24). Let us see which has higher variation.
+
+```{r same-x-level}
+sm <- smart_meter10 %>%
+filter(customer_id %in% c(10017936))
+.data = sm
+p1 <- sm %>% prob_plot("wknd_wday", "hour_day", plot_type = "boxplot", symmetric = FALSE)  + 
+  scale_y_log10()
+
+p2 <- sm %>% prob_plot("week_month", "hour_day", plot_type = "boxplot", symmetric = FALSE, quantile_prob = c(0.05, 0.25, 0.5, 0.75, 0.95)) + 
+  scale_y_log10()
+
+p1
+p2
 ```
@@
 # Does it match with the threshold?
+
+No
+
 # Should they tally with each other?
 
-## cricket data
-
-```{r cricket}
-cricket_data <- read_rds("data/cricket_data.rds")
-
-  hierarchy_model <- tibble::tibble(
-      units = c("index", "over", "inning", "match"),
-      convert_fct = c(1, 20, 2, 1))
-
-hierarchy_tbl <- hierarchy_model
-
-response <- "runs_per_over"
-
-harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)
-```
-
-```{r cricket_lag}
-cricket_data <- read_rds("data/cricket_data.rds")
-cricket_data %>%
-  filter(over!=1) %>%
-  prob_plot("over", "lag_field",
-  hierarchy_model,
-  response = "run_rate",
-  plot_type = "violin",
-  symmetric = FALSE,
- quantile_prob = c(0.25, 0.5, 0.75)) +
-   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
-  ylab("runs per over")  +
-  xlab("number of wickets in previous over") +
-  ggtitle("b") +
-   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
-```
+Yes ideally.
+But given the number of permutations they are not matching.
+Tried with increasing the number of permutations - didn't help.
+Also, permutations taking very long since for each permutation MMPD is calcuated for every harmony pair at each step.

### 2020-07-23 simulations_MMPD6.Rmd
adding recent simulation files

--- 2020-07-16 norm_threshold4.Rmd
@@
                response = "general_supply_kwh", 
-               dist_ordered = FALSE)
+               dist_ordered = TRUE)
 
-smart_harmony %>% kable()
+smart_harmony %>% 
+  mutate(MMPD = round(MMPD, 3), max_pd = round(max_pd, 3)) %>% 
+  mutate(rankn = row_number()) %>%
+  rename("rankun" = "r") %>% kable()
+```
 
-smart_harmony %>% ggplot(aes(x = x_levels, y =  r)) + geom_point()
+```{r cricket_harmony}
 
-smart_harmony %>% ggplot(aes(x = facet_levels, y =  r)) + geom_point()
+cricket_data <- read_rds("data/cricket_data.rds")
+
+  hierarchy_model <- tibble::tibble(
+      units = c("index", "over", "inning", "match"),
+      convert_fct = c(1, 20, 2, 1))
+
+.data <- cricket_data
+
+hierarchy_tbl <- hierarchy_model
+
+response <- "runs_per_over"
+
+harmonies_cric <- cricket_data %>% 
+  harmony(lgran = "over",
+          ugran = "match", 
+          filter_in = c("lag_field", "over"),
+          hierarchy_tbl = hierarchy_model)
+
+cric_harmony <- cricket_data %>% 
+  rank_harmony(harmony_tbl = harmonies_cric, 
+               response = response,
+               hierarchy_tbl = hierarchy_model,
+               dist_ordered = TRUE) %>% 
+  dplyr::filter(!(facet_variable=="over_inning"|x_variable =="over_inning"))
+
+cric_harmony %>% 
+  mutate(rankn = row_number()) %>%
+  rename("rankun" = "r") %>% 
+  mutate(MMPD = round(MMPD, 3), max_pd = round(max_pd, 3)) %>% 
+  kable()
 ```
+
+# Does normalisation work?
+
+## Number of levels and ranking
+
+```{r normalisation_work}
+smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) +geom_point()
+smart_harmony %>% mutate(lev = facet_levels*x_levels, r_MMPD = row_number()) %>% ggplot(aes(x = lev, y = r_MMPD)) +geom_line()
+```
+
+## Comparing levels using simulated data
+
+- MMPD should help choose the significantly different distributions only
+- The ranking should be from most different to least different
+
+Both of these should be checked again taking unnormalised statistic
+
+## Scenario 1: Simulated same normal distributions for all combinations of the pair of cyclic granularities
+
+## Scenario 2: Simulated same normal distributions for all x-levels within a facet, but different distributions across facets
+
+## Scenario 3: Simulated different normal distributions for different x-levels but they do not vary across facets
+
+# Scenario 4: Mixture of secenario 2 and 3
 
@@
 Cons: Only one threshold for all harmony pairs might not be an appropriate measure but a good benchmark.
-
-# Does it match with the threshold?
-# Should they tally with each other?
-
-## cricket data
-
-```{r cricket}
-cricket_data <- read_rds("data/cricket_data.rds")
-
-  hierarchy_model <- tibble::tibble(
-      units = c("index", "over", "inning", "match"),
-      convert_fct = c(1, 20, 2, 1))
-
-hierarchy_tbl <- hierarchy_model
-
-response <- "runs_per_over"
-
-harmonies_cric <- cricket_data %>% harmony(lgran = "over", ugran = "match", filter_in = c("lag_field", "over"), hierarchy_tbl = hierarchy_model)
-```
-
-```{r cricket_lag}
-cricket_data <- read_rds("data/cricket_data.rds")
-cricket_data %>%
-  filter(over!=1) %>%
-  prob_plot("over", "lag_field",
-  hierarchy_model,
-  response = "run_rate",
-  plot_type = "violin",
-  symmetric = FALSE,
- quantile_prob = c(0.25, 0.5, 0.75)) +
-   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
-  ylab("runs per over")  +
-  xlab("number of wickets in previous over") +
-  ggtitle("b") +
-   theme(plot.title = element_text(face = "bold")) +  theme_minimal()
-```

### 2020-07-23 simulations_MMPD6.Rmd
simulation graph

@@
 library(readr)
+
 opts_chunk$set(
@@
 ```{r normalisation_work}
-smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) +geom_point()
+smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) + geom_line()
 smart_harmony %>% mutate(lev = facet_levels*x_levels, r_MMPD = row_number()) %>% ggplot(aes(x = lev, y = r_MMPD)) +geom_line()
@@
 
+```{r scene1}
+library(readr)
+library(ggplot2)
+sim_data1 <- read_rds("sim_data1.rds")
+
+sim_data1 %>%
+  dplyr::filter(name==1) %>%
+  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
+
+global_harmony_iter1 <- read_rds("data/global_harmony_iter1.rds")
+
+global_harmony_iter1 %>% kable()
+```
+
 ## Scenario 2: Simulated same normal distributions for all x-levels within a facet, but different distributions across facets
 
+```{r scene2}
+library(readr)
+library(ggplot2)
+sim_data2 <- read_rds("sim_data2.rds")
+
+sim_data2 %>%
+  dplyr::filter(name==1) %>%
+  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
+
+global_harmony_iter2 <- read_rds("data/global_harmony_iter2.rds")
+
+global_harmony_iter2 %>% kable()
+```
+
 ## Scenario 3: Simulated different normal distributions for different x-levels but they do not vary across facets
+
+```{r scene3}
+library(readr)
+library(ggplot2)
+
+sim_data3 <- read_rds("sim_data3.rds")
+
+sim_data3 %>%
+  dplyr::filter(name==1) %>%
+  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
+
+sim_data3 %>%
+  dplyr::filter(name==4) %>%
+  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
+
+global_harmony_iter3 <- read_rds("data/global_harmony_iter3.rds")
+
+global_harmony_iter3 %>% kable()
+```
 

### 2020-07-27 simulations_MMPD6.Rmd
fixing MMPD document

@@
 Hence, there is a need to rank the harmonies basis how well they capture the variation in the measured variable and additionally reduce the number of harmonies for further exploration/visualization.
+Assuming a numeric response variable, our graphics are displays of distributions compared across combinations of categorical variables, one placed at x-axis and the other on the facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
 Hence, displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data.
-Thus the idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
+Thus, the idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
 
@@
 Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-
-We shall call this measure of variation as  Median Maximum Pairwise Distances (MMPD)
+We call this measure of variation as  Median Maximum Pairwise Distances (MMPD)
 
@@
 The normalizing constants, however, vary depending on the underlying distribution and hence it is important to assume a distribution of distances in our case.
+
+## Distribution of distances
+
+Theoretical: JS distances are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values.
+Taking sample percentiles to approximate the integral would mean taking $m = 99$.
+With large $m$, chi-squared is asymptotically normal by the CLT.
+Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution.
+
+Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made.
+
+Empirical: Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution.
+
+```{r}
+smart_harmony <-read_rds("data/smart_harmony_nonst.rds")
+smart_harmony
+```
 
@@
 
-```{r cricket_harmony}
-
-cricket_data <- read_rds("data/cricket_data.rds")
-
-  hierarchy_model <- tibble::tibble(
-      units = c("index", "over", "inning", "match"),
-      convert_fct = c(1, 20, 2, 1))
-
-.data <- cricket_data
-
-hierarchy_tbl <- hierarchy_model
-
-response <- "runs_per_over"
-
-harmonies_cric <- cricket_data %>% 
-  harmony(lgran = "over",
-          ugran = "match", 
-          filter_in = c("lag_field", "over"),
-          hierarchy_tbl = hierarchy_model)
-
-cric_harmony <- cricket_data %>% 
-  rank_harmony(harmony_tbl = harmonies_cric, 
-               response = response,
-               hierarchy_tbl = hierarchy_model,
-               dist_ordered = TRUE) %>% 
-  dplyr::filter(!(facet_variable=="over_inning"|x_variable =="over_inning"))
-
-cric_harmony %>% 
-  mutate(rankn = row_number()) %>%
-  rename("rankun" = "r") %>% 
-  mutate(MMPD = round(MMPD, 3), max_pd = round(max_pd, 3)) %>% 
-  kable()
-```
-
 # Does normalisation work?
@@
 ```{r normalisation_work}
-smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) + geom_line()
-smart_harmony %>% mutate(lev = facet_levels*x_levels, r_MMPD = row_number()) %>% ggplot(aes(x = lev, y = r_MMPD)) +geom_line()
+smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) + geom_line() + geom_point(colour = "blue")
+smart_harmony %>% mutate(lev = facet_levels*x_levels, r_MMPD = row_number()) %>% ggplot(aes(x = lev, y = r_MMPD)) +geom_line() + geom_point(colour = "blue")
 ```

### 2020-07-27 simulations_MMPD6.Rmd
theorem  added

@@
 Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
-The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
-The normalizing constants, however, vary depending on the underlying distribution and hence it is important to assume a distribution of distances in our case.
+The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.Suppose $d_{1},d_{2}\ldots ,d_{n}$ be a sequence of independent and identically-distributed pairwise distances and $M_{n}=\max\{d_{1},\ldots ,d_{n}\}$.
+Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{n}, b_{n})$ exists such that each $a_{n}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family.
+The normalizing constants $(a_{n}, b_{n})$ vary depending on the underlying distribution of the pairwise distances.
+Hence to normalize appropriately, it is important to assume a distribution of these distances.
 
@@
 Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution.
-
-Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made.
+Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function.
+$n$ is the number of pairwise comparisons being made.
 

### 2020-07-27 simulations_MMPD6.Rmd
updating documentation of MMPD

@@
 Hence, displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data.
-Thus, the idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
+Here, we have two main objectives:
+
+- To choose harmonies for which distributions of categories are significantly different 
+- To rank the selected harmonies from highest to lowest variation in the distribution of their categories.
+The idea here is to rate a harmony pair higher if this variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
 
@@
 We do this through Jensen-Shannon distance which is based on Kullback-Leibler divergence.
-Probability distributions are represented through sample quantiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
-We call this measure of variation as  Median Maximum Pairwise Distances (MMPD)
 
-# Normalize distances
+The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
+$$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.
+Probability distributions are estimated through quantiles instead of kernel density so that there is minimal dependency on selecting kernel or bandwidth.
+
+We call this measure of variation as  Median Maximum Pairwise Distances (MMPD).
+
+## Normalize distances
 
@@
 Thus these maximum pairwise distances need to be normalized for different harmonies in a way that eliminates the effect of different levels.
-The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.Suppose $d_{1},d_{2}\ldots ,d_{n}$ be a sequence of independent and identically-distributed pairwise distances and $M_{n}=\max\{d_{1},\ldots ,d_{n}\}$.
+The Fisher–Tippett–Gnedenko theorem in the field of Extreme Value Theory states that the maximum of a sample of iid random variables after proper re- normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
+
+More formally, $d_{1},d_{2}\ldots ,d_{n}$ be a sequence of independent and identically-distributed pairwise distances and $M_{n}=\max\{d_{1},\ldots ,d_{n}\}$.
 Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{n}, b_{n})$ exists such that each $a_{n}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{n}-b_{n}}{a_{n}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family.
@@
 Empirical: Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution.
+
+# Choose thresholds for harmonies through permutation test
+
+**Assumption:** random permutation without considering ordering 
+(global)
+
+1. Given the data; $\{v_t: t=0, 1, 2, \dots, T-1\}$, the MMPD is computed and is represented by $MMPD_{obs}$.
+
+2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
+
+3. MMPD is computed for all random permutation of the data and is represented by $MMPD_{sample}$.
+
+4.  Steps (2) and (3) are repeated a large number
+of times M (e.g. 1000).
+
+5. For each permutation, one $MMPD_{sample}$ value is obtained.
+
+6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
+
+7. If  $MMPD_{obs}> MMPD_{threshold}$, harmony pairs are accepted. Onlyone threshold for all harmony pairs.
+
+Pros: Considering thresholds global for all harmony pairs would imply less computation time.
+
+Cons: Only one threshold for all harmony pairs means we are assuming distribution of all harmonies pairs are similar, which might not be the case.But nevertheless, it is a good benchmark.
 
@@
 # Does normalisation work?
+
+## Minimal example
+
+Consider three cyclic granularities $A$, $B$ and $C$ with $2$, $3$ and $4$ categories.
+Thus, the harmony table consisting of all possible harmony pairs (assuming all pairs are harmonies), would look like the following:
+
+```{r harmony_min}
+# harmonies <- tibble::tibble(facet_variable = c("A", "B", "A", "C", "B", "C"),
+#                             x_variable  = c("B","A", "C", "A", "C", "B"),
+#                             facet_levels = c(2, 3, 2, 4, 3, 4),
+#                             x_levels = c(3, 2, 4, 2, 4, 3))
+
+harmonies <- tibble::tibble(facet_variable = c("A", "B"),
+                            x_variable  = c("B","A"),
+                            facet_levels = c(2, 3),
+                            x_levels = c(3, 2))
+
+harmonies %>% knitr::kable()
+```
+
+### Scenario 1: Simulated same normal distributions for all combinations of categories for all rows in the harmony table.
 
@@
 ```
-
-# Choose thresholds for harmonies
-
-Permutation test:
-
-**Assumption:** random permutation without considering ordering 
-(global)
-
-1. Given the data; $\{v_t: t=0, 1, 2, \dots, T-1\}$, the MMPD is computed and is represented by $MMPD_{obs}$.
-
-2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
-
-3. MMPD is computed for all random permutation of the data and is represented by $MMPD_{sample}$.
-
-4.  Steps (2) and (3) are repeated a large number
-of times M (e.g. 1000).
-
-5. For each permutation, one $MMPD_{sample}$ value is obtained.
-
-6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
-
-7. If  $MMPD_{obs}> MMPD_{threshold}$, harmony pairs are accepted. Onlyone threshold for all harmony pairs.
-
-Pros: Considering thresholds global for all harmony pairs would imply less computation time.
-
-Cons: Only one threshold for all harmony pairs might not be an appropriate measure but a good benchmark.

### 2020-07-27 main.Rmd
combination changed to permutation

@@
 \noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities.
-Hence, all possible pairs of cyclic time granularities that could be analyzed together is $N_C \choose 2$.
-Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left could be large and overwhelming for human consumption.
+Hence, all possible pairs of cyclic time granularities that could be analyzed together is $^{N_C} P_2$, with each of the circular granularity either mapped to facets or x-axis.
+Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption.
 Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.

### 2020-07-27 main.Rmd
fixing MMPD section

@@
 \noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities.
-Hence, all possible pairs of cyclic time granularities that could be analyzed together is $^{N_C} P_2$, with each of the circular granularity either mapped to facets or x-axis.
+Hence, all possible pairs of cyclic time granularities that could be analyzed together is $^{N_C} P_2$, with each of the circular granularities either mapped to facets or x-axis.
 Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption.
-Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+Hence, there is a need to refine our search and only look at harmonies for which variation in the categories of the cyclic granularities are significant.
+Moreover, there is a need to rank these selected harmonies from highest to lowest variation of their categories.
+
+<!-- rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
 The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
--->
+--> -->
 
@@
 
-\noindent The choice of aesthetic mapping and facets for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet.
+\noindent The choice of data structure for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet.
 A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet.
@@
 ```{r harmony-tab, echo=FALSE, cache=TRUE}
-if(!dir.exists("cache/"))
-{
-  dir.create("cache")
-}
-if (!file.exists("cache/smart_harmony.rds")) {
+# if(!dir.exists("cache/"))
+# {
+#   dir.create("cache")
+# }
+# if (!file.exists("cache/smart_harmony.rds")) {
   sm <- smart_meter10 %>%
@@
          `x-axis levels` = x_levels,
-         `average maximum variation` = mean_max_variation)
-
-  write_rds(smart_harmony,"cache/smart_harmony_ak.rds")
-}
-smart_harmony <- read_rds("cache/smart_harmony.rds")
+         `average maximum variation` = MMPD)
+
+#   write_rds(smart_harmony,"cache/smart_harmony_ak.rds")
+# }
+#smart_harmony <- read_rds("cache/smart_harmony.rds")
 

### 2020-07-29 simulations_MMPD6.Rmd
MMPD document

@@
 
+Type I Error. Reject the null hypothesis when there is in fact no significant effect.
+
+```{r, echo = FALSE}
+library(ggplot2)
+library(gravitas)
+library(purrr)
+library(distributional)
+library(magrittr)
+library(tidyr)
+library(dplyr)
+harmonies <- tibble::tibble(facet_variable = c("A", "B"),x_variable  = c("B","A"), facet_levels = c(2, 3),x_levels = c(3, 2))
+data <- gravitas:::sim_distharmony1(.data = harmonies[1,], sim_dist =  dist_normal(mu = 2, sigma = 5))
+
+data %>% select(-dist) %>% unnest(sim_dist) %>%
+ggplot(aes(x = Var2, y = sim_dist)) +
+facet_wrap(~Var1) + geom_boxplot()
+
+data %>% select(-dist) %>% unnest(sim_dist) %>%
+ggplot(aes(x = Var1, y = sim_dist)) +
+facet_wrap(~Var2) + geom_boxplot()
+
+dat1 <- data %>% select(-dist) %>%
+unnest(sim_dist) %>%
+pivot_wider(names_from = Var1, values_from  = sim_dist)
+dat2 <- data %>% select(-dist) %>%
+unnest(sim_dist) %>%
+pivot_wider(names_from = Var2, values_from  = sim_dist)
+
+step1_data <- list(dat1, dat2)
+response = "general_supply_kwh"
+# global_harmony_iter1 <-  smart_meter10 %>%
+#   global_threshold(harmony_tbl = harmonies,
+#   response, step1_data = step1_data,
+#    response, hierarchy_tbl = hierarchy_model)
+
+```
+
+```{r, echo = FALSE}
+library(ggplot2)
+library(gravitas)
+library(purrr)
+library(distributional)
+library(magrittr)
+library(tidyr)
+harmonies <- tibble::tibble(facet_variable = c("A", "B"),x_variable  = c("B","A"), facet_levels = c(2, 3),x_levels = c(3, 2))
+data <- sim_distharmony1(.data = harmonies[1,], sim_dist =  c(rep(dist_normal(mu = 2, sigma = 5),3), rep(dist_normal(mu = 5, sigma = 10), 3)))
+data %>% select(-dist) %>% unnest(sim_dist) %>%
+ggplot(aes(x = Var2, y = sim_dist)) +
+facet_wrap(~Var1) + geom_boxplot()
+data %>% select(-dist) %>% unnest(sim_dist) %>%
+ggplot(aes(x = Var1, y = sim_dist)) +
+facet_wrap(~Var2) + geom_boxplot()
+ dat1 <- data %>% select(-dist) %>%
+ unnest(sim_dist) %>%
+ pivot_wider(names_from = Var1, values_from  = sim_dist)
+ dat2 <- data %>% select(-dist) %>%
+  unnest(sim_dist) %>%
+  pivot_wider(names_from = Var2, values_from  = sim_dist)
+  step1_data <- list(dat1, dat2)
+  response = "general_supply_kwh"
+  # global_harmony_iter1 <-  smart_meter10 %>%
+  # global_threshold(harmony_tbl = harmonies,
+  # response, step1_data = step1_data,
+  #  response, hierarchy_tbl = hierarchy_model)
+
+```
+
 ## Number of levels and ranking

### 2020-07-29 main.Rmd
intro of which harmonies to show changed

@@
 Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption.
-Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+Hence, there is a need to select only those harmonies which have significant variation in the measured variable and hence might appear to be interesting when displayed.
+Moreover, there is a need to rank these harmonies basis how interesting they are, which could be measured through the captured amount of variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.

### 2020-07-29 main.Rmd
intro changed again of choosing harmonies

--- 2020-07-28 main.Rmd
@@
 Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption.
-Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
+Hence, there is a need to rank the harmonies basis how well they capture the variation in the measured variable and additionally reduce the number of harmonies for further exploration/visualization.
+Displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data.
+Hence, the idea  would be to select only those harmonies which have significant variation in the measured variable and to rank the selected harmonies, such that a harmony pair is rated more important if the variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
+This twofold objective is achieved by defining a statistic that essentially captures the variation in a harmony pair and a threshold below which the variation is not considered significant.
+
+<!-- Hence, there is a need to select only those harmonies which have significant variation in the measured variable and hence might appear to be interesting when displayed.
+Moreover, there is a need to rank these harmonies basis how interesting they are, which could be measured through the captured amount of variation in the measured variable.
+-->
+
+<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
 For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
@@
 
-\noindent One of the potential ways to evaluate this variation is by computing the maximum pairwise distances between the distributions of the measured variable across different levels of the cyclic granularities.
-<!--across different levels of $C_j$ and integrating it for all levels of $C_i$.
-Distances or similarity measures have been extensively studied and applied to fields like clustering, classification and anomaly detection.
----> We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.
+\noindent One of the potential ways to evaluate this variation is by computing the maximum pairwise distances between the distributions of the measured variable.
+<!--We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.-->
 One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence (@Kullback1951-jy).

### 2020-07-29 main.Rmd
cleaning extra text

@@
 
- <!-- $q_u$ and $q_v$ computed for probability vector $p$ will be given by -->
-
 <!-- # why Normalizing required --->
-
-\noindent The choice of aesthetic mapping and facets for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet.
-A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet.
-A probable solution for ranking interesting harmony pairs could be to arrange from highest to lowest average maximum pairwise distances.
-But maximum is not robust to the number of levels and seems to be higher for harmonies with higher number of levels.
-Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.<!--- \noindent The Fisher–Tippett–Gnedenko theorem states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process.
---> Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
+<!-- too much details - not required -->
+
+\noindent The choice of mapping considered for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet.
+A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet.<!-- But maximum is not robust to the number of levels and seems to be higher for harmonies with higher number of levels.--> However, these maximum pairwise distances need to be normalized so that the effect of different levels of $C_i$ and $C_j$ are eliminated.
+Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
 Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family.
@@
 With large degrees of freedom, chi-squared is asymptotically normal by the Central Limit Theorem.
-Hence, the norming constants for the distribution of $M_m$ are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of a standard normal distribution.
+Hence, the norming constants for the distribution of $M_m$ are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of the assymptotic normal distribution.
 Typically normal distributions will have Gumbel limiting distribution for the maxima, which have medians that grow as $m\rightarrow\infty$ at the rate of log $m$.
@@
 
-\noindent The algorithm employed is summarized as follows:
+\noindent The algorithm employed for computing MMPD is summarized as follows:
 <!-- Algorithm -->
@@
 8. Repeat Steps 1 to 7 for all harmony pairs.
-
-<!-- 1.
-For each pair of cyclic granularity $(C_i, C_j)$, there will be $L$ groups corresponding to each level $A_k$ of $C_i$.
-Thus there will be  $m = \binom{L}{2}$ pairwise distances corresponding to each $A_k$.
--->
-
-<!-- 2.
-Pairwise distances are computed between different quantiles of the time series variables using Jensen-Shanon distance.
-Let $d_1, d_2, d_3, \dots d_m$ be the pairwise Jensen-Shannon distances.
-Suppose the maximum pairwise distance for $A_k$ maximum be $M_k$.
--->
-
-<!-- 3.
-The distribution of distances is heavy tailed.
-But if they are gamma or weibull could be judged by looking at the QQ-plots of the empirical distribution.
-The estimates of the parameters are obtained using Maximum Likelihood Estimation.
--->
-
-<!-- 4.
-The norming constants $(a_k, b_k)$ are obtained from @Embrechts2013-pd and each $M_k$ is normalized using normalized maximum $NM_k = \frac{M_k - b_k}{a_k}$ for each $A_k$, $k \in {1, 2, \dots, K}$.
-We have normalized the number of levels of $C_j$ using Steps 1-4.
--->
-
-<!-- 5.
-We also need to aggregate these normalized maximum across different levels of $C_i$.
-For this we can take the mean, median or any quantiles of these normalized maximums.
-The median (and any quantiles) of these maximum distributions (which converges to a Gumbel distribution) grow at the rate of (some power of) $log K$ with $K \xrightarrow{} \infty$.
-Hence, M/log($K$) could serve as the final measure (MMPD), where $M = median\{NM_1, NM_2, \dots, NM_K\}$.
--->
-
-<!-- 6. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD. -->
 

### 2020-07-29 main.Rmd
include algo for threshold
algo for threshold updated

@@
 
-6. Use Steps 1-4 to compute normalized maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.
+6. Use Steps 1-5 to compute normalized maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.
 
@@
 8. Repeat Steps 1 to 7 for all harmony pairs.
+
+\noindent The algorithm employed for selecting harmonies with significant variation is summarized as follows:
+
+- **Input:** Harmony pairs $(C_i, C_j)$ $\forall i, j \in N_C$  
+- **Output:** Harmony pairs $(C_i, C_j)$ for which MMPD is significant.
+
+1. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, MMPD is computed for all harmony pairs and is represented by $MMPD_{obs}$.
+
+2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
+
+3. MMPD is computed for all harmony pairs for this random permutation of the data and is represented by $MMPD_{sample}$.
+
+4.  Steps (2) and (3) are repeated a large number
+of times M (e.g. 100).
+
+5. For each permutation, one $MMPD_{sample}$ vector is obtained.
+
+6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
+
+7. Harmony pairs for which $MMPD_{obs}> MMPD_{threshold}$ are accepted.
 

### 2020-07-29 main.Rmd
changes to choosing harmonies section done

@@
 With large degrees of freedom, chi-squared is asymptotically normal by the Central Limit Theorem.
-Hence, the norming constants for the distribution of $M_m$ are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of the assymptotic normal distribution.
-Typically normal distributions will have Gumbel limiting distribution for the maxima, which have medians that grow as $m\rightarrow\infty$ at the rate of log $m$.
+Hence, the norming constants for the distribution of $M_m$ are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of the limiting normal distribution.
+Typically normal distributions will have Gumbel limiting distribution for the maxima, which have medians that grow at the rate of log $m$ as $m\rightarrow\infty$.
+This fact is used to normalize the statistic for different facet levels.
 

### 2020-07-29 main.Rmd
smart meter application updated

@@
 From the search list, $N_C = 7$ cyclic granularities are chosen for which we would like to derive insights of energy behavior.
-Recalling the data structure <$C_i$, $C_j$, `general_supply_kwh`> for exploration $\forall i, j \in \{1, 2, ldots, 7\}$, each of these $7$ cyclic granularities can either be mapped to x-axis or to facet.
-Choosing $2$ of the possible $7$ granularities, which is equivalent to having ${7 \choose 2} = 21$ candidates for visualization.
-Fortunately, harmony granularities can be identified among those $21$ possibilities to narrow the search.
+Recalling the data structure <$C_i$, $C_j$, `general_supply_kwh`> for exploration $\forall i, j \in \{1, 2, \ldots, 7\}$, each of these $7$ cyclic granularities can either be mapped to x-axis or to facet.
+Choosing $2$ of the possible $7$ granularities, which is equivalent to having $^{7}P_2 = 42$ candidates for visualization.
+Fortunately, harmonies can be identified among those $42$ possibilities to narrow the search.
 \autoref{tab:harmony-tab} shows $10$ harmony pairs after removing clashes and any cyclic granularities with levels more than 31, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
-For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
+The MMPD is also shown along with a threshold indicator taking the value TRUE only when variation is significant.
+Starting from $42$ possible pairs of cyclic granularities to visualize, we are finally left with only $6$, which is a very sizable number of displays for exploration.
+
+Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the siginificance of MMPD, threshold and the impact of different distribution plots or reverse mapping.
+\noindent For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
 The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust} (b and c respectively).
@@
 
-smart_harmony <- sm %>%
-  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh") %>%
+global_harmony <- sm %>%
+  global_threshold(harmony_tbl = harmonies,
+                   response = "general_supply_kwh")%>%
+  # select(facet_variable,x_variable,
+  #        facet_levels,x_levels,MMPD) %>% 
   rename(`facet variable` = facet_variable,
@@
                 caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis along with their number of categories/levels.
-Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.
-The harmony pairs are ranked basis the average maximum variation which captures maximum pairwise distance between x-axis levels for each facets and then average them over facets.") %>%
+Out of 42 possible combinations of cyclic granularities, only these 10 are harmony pairs.
+The harmony pairs are ranked basis highest to lowest MMPD, only the first 6 of which are significant.") %>%
     #row_spec(0, bold = TRUE) %>%
@@
 This case study shows systematic exploration of energy behavior for a random household to gain some insights on periodic behavior of the households. 
-First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualize a harmony pair and shows the effect of different distribution plots or reverse mapping.
+<!--First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualize a harmony pair and shows the effect of different distribution plots or reverse mapping.-->
 

### 2020-07-29 main.Rmd
global threshold code update

@@
           filter_out = c("hhour", "fortnight"))
-
+set.seed(12345)
 global_harmony <- sm %>%
   global_threshold(harmony_tbl = harmonies,
-                   response = "general_supply_kwh")%>%
-  # select(facet_variable,x_variable,
-  #        facet_levels,x_levels,MMPD) %>% 
+                   response = "general_supply_kwh", nsamp = 5)%>%
+  select(facet_variable,x_variable,
+         facet_levels,x_levels,MMPD, gt_MMPD) %>%
   rename(`facet variable` = facet_variable,
@@
          `x-axis levels` = x_levels,
-         `average maximum variation` = MMPD)
+         `above threshold` = gt_MMPD)
 
@@
 
-  knitr::kable(smart_harmony,
+  knitr::kable(global_harmony,
                format = "latex",

### 2020-07-30 simulations_MMPD6.Rmd
little changes in simulation

@@
 
-step1_data <- list(dat1, dat2)
+harmony_data <- list(dat1, dat2)
 response = "general_supply_kwh"
+
+global_harmony <- harmony_data %>%
+  global_threshold(harmony_tbl = harmonies,
+               response = "sim_dist",
+               dist_distribution = "normal",
+               dist_ordered = TRUE,
+               create_gran_data = FALSE)
+
 # global_harmony_iter1 <-  smart_meter10 %>%

### 2020-07-30 main.Rmd
polishing the MMPD section

--- 2020-07-29 main.Rmd
@@
 Hence, the idea  would be to select only those harmonies which have significant variation in the measured variable and to rank the selected harmonies, such that a harmony pair is rated more important if the variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
-This twofold objective is achieved by defining a statistic that essentially captures the variation in a harmony pair and a threshold below which the variation is not considered significant.
+This twofold objective is achieved by defining a statistic that could essentially capture the variation of the measured variable in a harmony pair and constructing a test to check if the variation is statistically significant.
 
@@
 <!--We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.-->
-One of the common ways to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence (@Kullback1951-jy).
+A common way to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence (@Kullback1951-jy).
 The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
@@
 
-The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] is used to facilitate the systematic exploration here.
+The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] consisting of `reading_datetime`, `customer_id` and `general_supply_kwh` denoting the index, key and measured variable of the tsibble is used to facilitate the systematic exploration.
 While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to consider is which cyclic time granularities we can look at exhaustively.
@@
 Fortunately, harmonies can be identified among those $42$ possibilities to narrow the search.
-\autoref{tab:harmony-tab} shows $10$ harmony pairs after removing clashes and any cyclic granularities with levels more than 31, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
-The MMPD is also shown along with a threshold indicator taking the value TRUE only when variation is significant.
+\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
+The MMPD is also shown along with indicator (*) only when variation of measured variable across the harmony pair significant.
 Starting from $42$ possible pairs of cyclic granularities to visualize, we are finally left with only $6$, which is a very sizable number of displays for exploration.
 
-Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the siginificance of MMPD, threshold and the impact of different distribution plots or reverse mapping.
+Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the siginificance of MMPD, threshold and the impact of different distribution plots and reverse mapping.
 \noindent For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
@@
           filter_out = c("hhour", "fortnight"))
+set.seed(12345)
 
@@
   global_threshold(harmony_tbl = harmonies,
-                   response = "general_supply_kwh")%>%
-  # select(facet_variable,x_variable,
-  #        facet_levels,x_levels,MMPD) %>% 
+                   response = "general_supply_kwh", nsamp = 5) %>%
   rename(`facet variable` = facet_variable,
@@
          `facet levels` = facet_levels,
-         `x-axis levels` = x_levels,
-         `average maximum variation` = MMPD)
+         `x-axis levels` = x_levels) %>% 
+  mutate(MMPD = if_else(gt_MMPD == TRUE,
+                        paste0(MMPD, "*"), paste0(MMPD))) %>%
+  select(`facet variable`,`x-axis variable`,
+         `facet levels`,`x-axis levels` ,MMPD)
 
@@
 
-  knitr::kable(smart_harmony,
+  knitr::kable(global_harmony,
                format = "latex",
                booktabs = TRUE,
-                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis along with their number of categories/levels.
-Out of 42 possible combinations of cyclic granularities, only these 10 are harmony pairs.
-The harmony pairs are ranked basis highest to lowest MMPD, only the first 6 of which are significant.") %>%
+                caption = "Harmonies with a pair of cyclic granularity one placed on facet and the other on x-axis.
+Out of 42 possible combinations of cyclic granularities, only 16 are harmony pairs.
+The harmony pairs are ranked basis highest to lowest MMPD and only the first 6 show significant variation in the measured variable.") %>%
     #row_spec(0, bold = TRUE) %>%
@@
 
-In \autoref{fig:bothcust} (b), the distribution of energy consumption is plotted across the harmony pair (weekday/weekend, hour of the day) through an area quantile plot.
-The black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+In \autoref{fig:bothcust} (b), <!--the distribution of energy consumption is plotted across the harmony pair (weekday/weekend, hour of the day) through an area quantile plot-->the black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
 The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
@@
 
-This case study shows systematic exploration of energy behavior for a random household to gain some insights on periodic behavior of the households. 
+This case study shows systematic exploration of energy behavior for a household to gain exhaustive insights on periodic behavior of the households. 
 <!--First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualize a harmony pair and shows the effect of different distribution plots or reverse mapping.-->
@@
 Plot (a) shows letter value plot across overs faceted by innings.
-For the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
-Plot (b) shows box plot of runs per over across an indicator of wickets in previous over faceted by current over.
+For the team batting in the first innings there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
+Plot (b) shows quantile plot of runs per over across an indicator of wickets in previous over faceted by current over.
 This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning = FALSE, message = FALSE, out.width = "90%"}

### 2020-07-30 check_gt_mnylvls.R
simulations in a different file

@@ -1,0 +1,555 @@
+##----load
+library(ggplot2)
+library(gravitas)
+library(purrr)
+library(distributional)
+library(magrittr)
+library(tidyr)
+library(dplyr)
+
+##--rank-harmony
+#' Plotting probability distributions across granularities
+#'
+#' Plot probability distribution of univariate series across bivariate temporal granularities.
+#'
+#' @param .data a tsibble, if cyclic granularity needs to be constructed or a list consisting of tibbles for each pair of cyclic granularity in the harmony table
+#' @param harmony_tbl A tibble of harmonies and their levels obtained from the function().
+#' @param response response variable.
+#' @param prob numeric vector of probabilities with values in [0,1].
+#' @param hierarchy_tbl A hierarchy table specifying the hierarchy of units
+#' @param dist_distribution Underlying distribution of distances. Look at hist_distance()
+#' @param dist_ordered if levels of the time granularity is ordered.
+#' @param alpha significance level
+#' @param create_gran_data if data corresponding to a pair of cyclic granularity needs to be created
+#' @return  A tibble of harmonies and their levels ranked in descending order of average maximum pairwise distance of the harmony pairs.
+#
+#' @examples
+#' library(tsibbledata)
+#' library(ggplot2)
+#' library(tsibble)
+#' library(lvplot)
+#' library(dplyr)
+#' library(gravitas)
+#' library(purrr)
+#' library(magrittr)
+#' sm <- smart_meter10 %>%
+#' filter(customer_id %in% c("10017936"))
+#' harmonies <- sm %>%
+#' harmony(ugran = "month",
+#'        filter_in = "wknd_wday",
+#'        filter_out = c("hhour", "fortnight"))
+#' .data = sm
+#' response  = "general_supply_kwh"
+#' harmony_tbl =  harmonies
+#' smart_harmony <- .data %>% rank_harmony(harmony_tbl = harmonies,
+#' response = "general_supply_kwh", dist_ordered = TRUE)
+#' harmony_tbl <- PBS %>% harmony(ugran = "year")
+#' rank_harmony(PBS, harmony_tbl = harmony_tbl, response = "Cost")
+#' @export rank_harmony
+#' @export create_harmony_data
+
[…]

### 2020-07-30 simulations_MMPD6.Rmd
simulations in a different file
simulation results updated

--- 2020-07-27 simulations_MMPD6.Rmd
@@
 knitr::opts_knit$set(root.dir = here::here())
+read_chunk('check_gt_mnylvls.R')
 ```
@@
 ```{r external, include = FALSE}
-# read_chunk('scripts/main.R')
+
 ```
@@
 ```{r load}
+
+```
+
+```{r rank_harmony}
+
+```
+
+```{r global_threshold}
 
@@
 
-7. If  $MMPD_{obs}> MMPD_{threshold}$, harmony pairs are accepted. Onlyone threshold for all harmony pairs.
+7. If  $MMPD_{obs}> MMPD_{threshold}$, harmony pairs are accepted. Only one threshold for all harmony pairs.
 
@@
 
-```{r}
+```{r, eval = FALSE, echo = FALSE}
 smart_harmony <-read_rds("data/smart_harmony_nonst.rds")
@@
 
-```{r smart_harmony}
+```{r smart_harmony,eval = FALSE, echo = FALSE}
 
@@
 
-# Does normalisation work?
+# Does normalization work? -  Minimal examples
 
-## Minimal example
-
-Consider three cyclic granularities $A$, $B$ and $C$ with $2$, $3$ and $4$ categories.
+Consider two cyclic granularities $A$ and $B$ with $2$ and $3$categories.
 Thus, the harmony table consisting of all possible harmony pairs (assuming all pairs are harmonies), would look like the following:
@@
 
-### Scenario 1: Simulated same normal distributions for all combinations of categories for all rows in the harmony table.
+The output table has the value of MMPD (normalized median maximum pairwise distances), max_pd (un-normalized maximum pairwise distances), r(Rank of max_pd), gt_MMPD(global threshold of MMPD indicator) and gt_maxpd(global threshold of max_pd indicator).
 
-## Number of levels and ranking
+## Scenario 1: Simulated same normal distributions for all combinations of categories for all harmony pairs.
 
-```{r normalisation_work}
-smart_harmony %>% mutate(lev = facet_levels*x_levels) %>% ggplot(aes(x = lev, y = r)) + geom_line() + geom_point(colour = "blue")
-smart_harmony %>% mutate(lev = facet_levels*x_levels, r_MMPD = row_number()) %>% ggplot(aes(x = lev, y = r_MMPD)) +geom_line() + geom_point(colour = "blue")
+Failure to reject the null hypothesis when there is in fact no significant effect.
+
+```{r samenull_2by4}
+
 ```
 
-## Comparing levels using simulated data
+*Conclusion*: The test does not select harmony pairs when distributions of categories are same
 
-- MMPD should help choose the significantly different distributions only
-- The ranking should be from most different to least different
+## Scenario 2: Simulated different distributions for all combinations of categories for harmony pairs for few levels.
 
-Both of these should be checked again taking unnormalised statistic
+```{r diffnull_2by4}
 
-## Scenario 1: Simulated same normal distributions for all combinations of the pair of cyclic granularities
-
-```{r scene1}
-library(readr)
-library(ggplot2)
-sim_data1 <- read_rds("sim_data1.rds")
-
-sim_data1 %>%
-  dplyr::filter(name==1) %>%
-  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
-
-global_harmony_iter1 <- read_rds("data/global_harmony_iter1.rds")
-
-global_harmony_iter1 %>% kable()
 ```
 
-## Scenario 2: Simulated same normal distributions for all x-levels within a facet, but different distributions across facets
+*Conclusion*: The test select the harmony pair for which distribution of x-axis categories are significantly different
 
-```{r scene2}
-library(readr)
-library(ggplot2)
-sim_data2 <- read_rds("sim_data2.rds")
+## Scenario 3: Simulated different distributions for all combinations of categories for all harmony pairs with many levels.
 
-sim_data2 %>%
-  dplyr::filter(name==1) %>%
-  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
+```{r diffnull_7by11}
 
-global_harmony_iter2 <- read_rds("data/global_harmony_iter2.rds")
-
-global_harmony_iter2 %>% kable()
 ```
 
-## Scenario 3: Simulated different normal distributions for different x-levels but they do not vary across facets
+*Conclusion*: The test with MMPD rejects the harmony pair even for which distribution of x-axis categories are significantly different.
+But the test with only maximum selects the pair with varying x-axis categories.
 
-```{r scene3}
-library(readr)
-library(ggplot2)
+## Scenario 4: Cumulative 3 levels with 2, 7 and 11 and testing level and power
 
-sim_data3 <- read_rds("sim_data3.rds")
+```{r samenull_3levels}
 
-sim_data3 %>%
-  dplyr::filter(name==1) %>%
-  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
-
-sim_data3 %>%
-  dplyr::filter(name==4) %>%
-  ggplot(aes(x = Var2, y = sim_data)) + geom_boxplot() + facet_wrap(~Var1)
-
-global_harmony_iter3 <- read_rds("data/global_harmony_iter3.rds")
-
-global_harmony_iter3 %>% kable()
 ```
 
-# Scenario 4: Mixture of secenario 2 and 3
+*Conclusion*: With 3 levels the test incorrectly chooses 1 harmony pair with similar distribution. The harmony pair which is displayed.
 
-```{r rank harmony}
-library(gravitas)
-library(ggplot2)
-library(dplyr)
-sm <- smart_meter10 %>%
-filter(customer_id %in% c("10017936"))
-harmonies <- sm %>%
-harmony(ugran = "month",
-filter_in = "wknd_wday",
-filter_out = c("hhour", "fortnight"))
-.data = sm
-response  = "general_supply_kwh"
-harmony_tbl =  harmonies
-smart_harmony <- .data %>% rank_harmony(harmony_tbl = harmonies,
-response = "general_supply_kwh", dist_ordered = FALSE)
+```{r diffnull_3levels}
+
 ```
+
+*Conclusion*: The test with MMPD selects just one pair, as opposed to the test with maximum.
+This needs to be checked against what we expect from the test.
+The harmony pairs which are selected (either through MMPD or maximum) are displayed.

### 2020-07-30 check_gt_mnylvls.R

@@ -6,48 +6,10 @@ library(distributional)
 library(magrittr)
 library(tidyr)
 library(dplyr)
+library(ggpubr)
 
-##--rank-harmony
-#' Plotting probability distributions across granularities
-#'
-#' Plot probability distribution of univariate series across bivariate temporal granularities.
-#'
-#' @param .data a tsibble, if cyclic granularity needs to be constructed or a list consisting of tibbles for each pair of cyclic granularity in the harmony table
-#' @param harmony_tbl A tibble of harmonies and their levels obtained from the function().
-#' @param response response variable.
-#' @param prob numeric vector of probabilities with values in [0,1].
-#' @param hierarchy_tbl A hierarchy table specifying the hierarchy of units
-#' @param dist_distribution Underlying distribution of distances. Look at hist_distance()
-#' @param dist_ordered if levels of the time granularity is ordered.
-#' @param alpha significance level
-#' @param create_gran_data if data corresponding to a pair of cyclic granularity needs to be created
-#' @return  A tibble of harmonies and their levels ranked in descending order of average maximum pairwise distance of the harmony pairs.
-#
-#' @examples
-#' library(tsibbledata)
-#' library(ggplot2)
-#' library(tsibble)
-#' library(lvplot)
-#' library(dplyr)
-#' library(gravitas)
-#' library(purrr)
-#' library(magrittr)
-#' sm <- smart_meter10 %>%
-#' filter(customer_id %in% c("10017936"))
-#' harmonies <- sm %>%
-#' harmony(ugran = "month",
-#'        filter_in = "wknd_wday",
-#'        filter_out = c("hhour", "fortnight"))
-#' .data = sm
-#' response  = "general_supply_kwh"
-#' harmony_tbl =  harmonies
-#' smart_harmony <- .data %>% rank_harmony(harmony_tbl = harmonies,
-#' response = "general_supply_kwh", dist_ordered = TRUE)
-#' harmony_tbl <- PBS %>% harmony(ugran = "year")
-#' rank_harmony(PBS, harmony_tbl = harmony_tbl, response = "Cost")
-#' @export rank_harmony
-#' @export create_harmony_data
 
+##----rank_harmony
 rank_harmony <- function(.data = NULL,
[…]

### 2020-08-05 main.Rmd
heading changed for main with choose harmonies
making minor changes in data structure and visualization

@@
 \noindent All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure).
-This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
+This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.
+The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
 Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.
@@
 However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
-Summarizing the distribution of the measured variable through these multiple observations could be a potential way to explore and bring forward different features of the data.
+Summarizing the distribution of the measured variable using these multiple observations could be a potential way to explore and bring forward different features of the data.
 <!-- compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
@@
 However, we need to consider the effect of number of observations on the summarization even for harmonies.
-Consider a data set with $(T + 1)$ observations, and two cyclic granularities $C_i$ and $C_j$ with $L$ and $M$ categories respectively.
-<!--Each element of $C_i$ occurs approximately $(T+1)/L$ times while each element of $C_j$ occurs approximately $(T+1)/M$ times.
-and there are no or very few empty combinations since we only consider harmonies.--> Let $nobs$ be the number of observations for a combination of levels from $(C_i, C_j)$.
-Now, $nobs$ might be equal for all combinations on an average <!--as $T\rightarrow\infty$ when the number of observations per combination is-->  so that $nobs = (T + 1)/(LM)$ as $T\rightarrow\infty$ or ideally take any value between $\{0, 1, 2, \dots, T\}$.
-The statistical transformation used for summarizing the measured variable for a combination should be chosen in line with  $nobs$.
-For example, it might be useful to compute deciles only when $nobs \ge 10$.
-Rarely occurring categories such as the 366th day of the year, or the 31st day of the month are more likely to suffer from this problem.
+Consider a data set with $(T + 1)$ observations, and two cyclic granularities $C_i$ and $C_j$ with $K$ and $L$ categories respectively.
+Let $nobs$ be the number of observations for a combination of levels from $(C_i, C_j)$.
+Now, $nobs$ might be equal for all combinations on an average so that $nobs = (T + 1)/(KL)$ as $T\rightarrow\infty$ or ideally take any value between $\{0, 1, 2, \dots, T\}$.
+The statistical transformation used for summarizing the measured variable should be chosen in line with  $nobs$.
+For example, it might be useful to compute deciles for summarization only when $nobs \ge 10$.
+Rarely occurring categories such as the 366th day of the year, or the 31st day of the month are likely to suffer from this problem.
 
@@
 
-VIC <- sm_cust50 %>%
+VIC <- sm_cust50 %>% as_tsibble(regular = FALSE) %>% 
   filter(customer_id==cust50$customer_id[2])
@@
 scene2 <- VIC %>%
-  #create_gran("month_year") %>%
-  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>%
-  #create_gran("day_week") %>%
-  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
   prob_plot("wknd_wday",
@@
 
-# VIC %>%
-#   #create_gran("month_year") %>%
-#   #filter(month_year %in% c("Jan", "Mar", "Dec")) %>%
-#   #create_gran("day_week") %>%
-#   #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
-#   prob_plot("month_year",
-#             "day_month",
-#             response = "general_supply_kwh",
-#             plot_type = "lv",
-#             symmetric = FALSE
-#   ) +
-#   ylab("") +
-#   xlab("day_month")  +
-#   ggtitle("(d) Letter value plot across days of months & months")
-
 scene3 <- VIC %>%
-  #create_gran("month_year") %>%
-  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>%
-  #create_gran("day_week") %>%
-  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
   prob_plot("quarter_year",
@@
 \autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
-
-## Choosing harmonies to show {#sec:chooseharmonies}
-
-<!-- Why do we need ranking system -->
-<!-- set of pairs - building the ranks for N_C and N_L -->
-
-\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities.
-Hence, all possible pairs of cyclic time granularities that could be analyzed together is $^{N_C} P_2$, with each of the circular granularity either mapped to facets or x-axis.
-Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption.
-Hence, there is a need to rank the harmonies basis how well they capture the variation in the measured variable and additionally reduce the number of harmonies for further exploration/visualization.
-Displays that capture more variation within different categories in the same group would be important to bring out different patterns of the data.
-Hence, the idea  would be to select only those harmonies which have significant variation in the measured variable and to rank the selected harmonies, such that a harmony pair is rated more important if the variation between different levels of the x-axis variable is higher on an average across all levels of facet variables.
-This twofold objective is achieved by defining a statistic that could essentially capture the variation of the measured variable in a harmony pair and constructing a test to check if the variation is statistically significant.
-
-<!-- Hence, there is a need to select only those harmonies which have significant variation in the measured variable and hence might appear to be interesting when displayed.
-Moreover, there is a need to rank these harmonies basis how interesting they are, which could be measured through the captured amount of variation in the measured variable.
--->
-
-<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
-For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions.
--->
-
-<!-- Calculating distances between distributions -->
-
-\noindent One of the potential ways to evaluate this variation is by computing the maximum pairwise distances between the distributions of the measured variable.
-<!--We consider distances between probability distributions which lines up with our pivotal idea of being able to visualize probability distributions across pairs of cyclic granularities.-->
-A common way to measure distances between distributions is through Jenson-Shanon divergence (@Lin1991-fj) which is based on Kullback-Leibler divergence (@Kullback1951-jy).
-The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where,
- $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.
-Probability distributions are estimated through quantiles instead of kernel density so that there is minimal dependency on selecting kernel or bandwidth.
-
-<!-- # why Normalizing required --->
-<!-- too much details - not required -->
-
-\noindent The choice of mapping considered for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet.
-A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet.<!-- But maximum is not robust to the number of levels and seems to be higher for harmonies with higher number of levels.--> However, these maximum pairwise distances need to be normalized so that the effect of different levels of $C_i$ and $C_j$ are eliminated.
-Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$.
-Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family.
-The normalizing constants $(a_{m}, b_{m})$ vary depending on the underlying distribution of the pairwise distances.
-Hence to normalize appropriately, it is important to assume a distribution of these distances.
-
-<!-- # distribution of distances and lit review -->
-
-\noindent @Menendez1997-in and @Grosse2002-ex provide studies of the statistical properties of the Jensen-Shannon divergences and suggest that the theoretical asymptotic distribution of Jensen-Shannon divergence converges to a chi-squared distribution with degrees of freedom equal to the number of discretization used to estimate the continuous distribution.
-With large degrees of freedom, chi-squared is asymptotically normal by the Central Limit Theorem.
-Hence, the norming constants for the distribution of $M_m$ are chosen to be $b_m = \Phi^{-1}(1- 1/m)$ and $a_m = 1/m\phi(b_m)$, where $\Phi^{-1}$ and $\phi$ are the quantile and density function of the limiting normal distribution.
-Typically normal distributions will have Gumbel limiting distribution for the maxima, which have medians that grow at the rate of log $m$ as $m\rightarrow\infty$.
-This fact is used to normalize the statistic for different facet levels.
-
-<!-- Median normalized distance grows at the rate of log($K$) as $K\rightarrow\infty$ and hence is normalized for the value of $K$ in MMPD.  -->
-
-<!--However, since we are dealing with small finite samples, a more appropriate approach would be to look at the distribution of the samples through histogram, density plots or using QQ-plot to see how well the empirical quantiles match the theoretical quantiles.-->
-
-\noindent The algorithm employed for computing MMPD is summarized as follows:
-<!-- Algorithm -->
-
-<!-- 1.
-Suppose $C_i$ and $C_2$ are two cyclic granularities such that $C_i$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_l$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ is the measured variable.
-Hence for each combination ($A_k$, $B_l$), we have the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, l}$ and $\forall j = {1, 2, \dots, m}$ -->
-
-- **Input:** Data corresponding to all harmony pairs, i.e., data sets of the form $(C_i, C_j, v)$ 
-$\forall i, j \in N_C$  
-- **Output:** MMPD (Median Maximum Pairwise Distances) measuring the average variation across different levels of $C_i$ and $C_j$ $\forall i, j \in N_C$
-
-1. Fix harmony pair $(C_i, C_j)$.
-
-2. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$.
-
-3. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories.
-
-4. Identify maximum within the $m$ computed distances.
-
-5. Compute normalized maximum distance ($NM$) using appropriate norming constants.
-<!--so that the distribution of the normalized maximum converges to a Gumbel distribution as $m\rightarrow\infty$.-->
-
-6. Use Steps 1-5 to compute normalized maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.
-
-7. Compute MMPD = median $(NM_1, NM_2, \dots, NM_K)$/log($K$).
-
-8. Repeat Steps 1 to 7 for all harmony pairs.
-
-\noindent The algorithm employed for selecting harmonies with significant variation is summarized as follows:
-
-- **Input:** Harmony pairs $(C_i, C_j)$ $\forall i, j \in N_C$  
-- **Output:** Harmony pairs $(C_i, C_j)$ for which MMPD is significant.
-
-1. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, MMPD is computed for all harmony pairs and is represented by $MMPD_{obs}$.
-
-2. From the original sequence a random permutation is obtained: $\{v_t^*: t=0, 1, 2, \dots, T-1\}$.
-
-3. MMPD is computed for all harmony pairs for this random permutation of the data and is represented by $MMPD_{sample}$.
-
-4.  Steps (2) and (3) are repeated a large number
-of times M (e.g. 100).
-
-5. For each permutation, one $MMPD_{sample}$ vector is obtained.
-
-6. $95^{th}$ percentile of this $MMPD_{sample}$ distribution is computed and stored in $MMPD_{threshold}$.
-
-7. Harmony pairs for which $MMPD_{obs}> MMPD_{threshold}$ are accepted.
 

### 2020-08-05 main.Rmd
smart meter data minor changes proof-read

@@
 The conjecture is that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
-Harmonies are pairs of granularities that do not contain empty combinations and thus aid exploratory data analysis.
-Plotting Clashes should be avoided since they contain empty combinations.
+Harmonies are pairs of granularities that do not contain empty combinations and could possibly be useful for exploring patterns.
+However, plotting clashes should be avoided since they might contain combinations which are empty or have too few observations.
 For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarters of the year.
 This plot does not work because every quarter does not correspond to all months of the year, for example, the first quarter of the year would never correspond to the month December of an year.
-
-<!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built.
-In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.-->
-
-<!---\autoref{fig:allFig} (c) shows box plot across days of the year by the 1st, 15th, 29th and 31st days of the month.
-\autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month.
-\autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig}(b) shows decile plots across day of the year and month of the year.--->
-
-<!-- Clearly, in \autoref{fig:allFig}, we observe that  -->
-
-<!--In (b) we can not compare the distributions across different facets because every day-of-year cannot occur for every month.
-For example, 21st day of the year can never correspond to the month March or December of any year.-->
-<!---In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year.
-In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year.
-On the contrary, for \autoref{fig:allFig}--->  <!---Thus the graphs that don't work are those where many of the combination sets are empty.--->
-
-<!---as they hinder the exploratory process by having missing combinations of time granularities in each panel.--->
 
@@
 
-Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section \ref{sec:summarize-measured}) or unevenly distributed events.
-For Kernel density based plots $nobs>30$ is suggested, for summary statistics based plots like boxplots $nobs = 10$ is acceptable and for quantile plots $nobs$ should at least be equal to the number of quantiles plotted.
-
-<!-- If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge lmk$ will be acceptable.
+Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section \ref{sec:summarize-measured}).
+Even when there are no rarely occurring events, number of observations might vary hugely within or across each facet.
+This might happen due to missing observations in the data or uneven locations of events in time domain.
+In such cases, the statistical transformations should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators.
+
+<!-- For Kernel density based plots $nobs>30$ is suggested, for summary statistics based plots like boxplots $nobs = 10$ is acceptable and for quantile plots $nobs$ should at least be equal to the number of quantiles plotted.
+If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge lmk$ will be acceptable.
 The value of $k$ will depend on the statistical transformation we are producing.
@@
 
-Even when there are no rarely occurring events, number of observations might vary hugely within or across each facet.
-This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the statistical transformations based on density should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators.
 <!-- (see \autoref{eq:MSE})  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$is a function of the sample size $n$.-->
@@
 Considering $6$ linear granularities half-hour, hour, day, week, month and year in the hierarchy table, $N_C = (6*5/2) = 15$.
-<!--(via Section \ref{sec:data-structure}) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.--> If $N_C$ seem too large, the smallest and largest linear granularities could be considered to be removed from the hierarchy table.
-We remove half-year and year to have $N_C = (4*3/2) = 6$ and obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
+<!--(via Section \ref{sec:data-structure}) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.--> If $N_C$ seem too large, the smallest and largest linear granularities could be considered to be removed for analysis.
+Granularities half-year and year are removed to have $N_C = (4*3/2) = 6$ and obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
 Further, we add cyclic granularity day-type( "wknd_wday") to capture weekend and weekday behavior.
-Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Section \ref{sec:cyclic-calendar}.
+Now that we have a list of cyclic granularities to look at, we should be able to compute them using Section \ref{sec:cyclic-calendar}.
 <!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
@@
 Choosing $2$ of the possible $7$ granularities, which is equivalent to having $^{7}P_2 = 42$ candidates for visualization.
-Fortunately, harmonies can be identified among those $42$ possibilities to narrow the search.
+Harmonies can be identified among those $42$ possibilities to narrow the search.
 \autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective  exploration becomes difficult with many levels (Section \ref{sec:levels}).
-The MMPD is also shown along with indicator (*) only when variation of measured variable across the harmony pair significant.
-Starting from $42$ possible pairs of cyclic granularities to visualize, we are finally left with only $6$, which is a very sizable number of displays for exploration.
-
-Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the siginificance of MMPD, threshold and the impact of different distribution plots and reverse mapping.
-\noindent For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
+
+<!--The MMPD is also shown along with indicator (*) only when variation of measured variable across the harmony pair significant.
+Starting from $42$ possible pairs of cyclic granularities to visualize, we are finally left with only $6$, which is a very sizable number of displays for exploration.-->
+
+\noindent Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the <!--siginificance of MMPD, threshold and the -->impact of different distribution plots and reverse mapping.
+For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
 The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust} (b and c respectively).
 \autoref{fig:bothcust} (a) displays reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour of the day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.
+In \autoref{fig:bothcust} (b), <!--the distribution of energy consumption is plotted across the harmony pair (weekday/weekend, hour of the day) through an area quantile plot-->the black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
+The energy consumption is extremely (positive- or right-) skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
+This is common across both weekdays and weekends.
+For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer as compared to weekends.
+
+\noindent Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust}(c) for some more insights.
+There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
+<!--Also the hours from 7 to 13 look most volatile.--> If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
+For example in \autoref{fig:bothcust}(a), it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends.
+Also, outliers are more prominent in the latter hours of the day.
+All of these indicate that looking at different distribution geometry or changing the mapping might shed lights on different aspect of the energy behavior for the same sample population.
 
@@
           filter_in = "wknd_wday", 
-          filter_out = c("hhour", "fortnight"))
-set.seed(12345)
-
-global_harmony <- sm %>%
-  global_threshold(harmony_tbl = harmonies,
-                   response = "general_supply_kwh", nsamp = 5) %>%
+          filter_out = c("hhour", "fortnight")) %>%
     rename(`facet variable` = facet_variable,
@@
          `facet levels` = facet_levels,
-         `x-axis levels` = x_levels) %>% 
-  mutate(MMPD = if_else(gt_MMPD == TRUE,
-                        paste0(MMPD, "*"), paste0(MMPD))) %>%
-  select(`facet variable`,`x-axis variable`,
-         `facet levels`,`x-axis levels` ,MMPD)
-
-#   write_rds(smart_harmony,"cache/smart_harmony_ak.rds")
-# }
-#smart_harmony <- read_rds("cache/smart_harmony.rds")
-
-  knitr::kable(global_harmony,
+         `x-axis levels` = x_levels)
+
+# set.seed(12345)
+# 
+# global_harmony <- sm %>%
+#   global_threshold(harmony_tbl = harmonies,
+#                    response = "general_supply_kwh", nsamp = 5) %>%
+#   rename(`facet variable` = facet_variable,
+#          `x-axis variable` = x_variable,
+#          `facet levels` = facet_levels,
+#          `x-axis levels` = x_levels) %>% 
+#   mutate(MMPD = if_else(gt_MMPD == TRUE,
+#                         paste0(MMPD, "*"), paste0(MMPD))) %>%
+#   select(`facet variable`,`x-axis variable`,
+#          `facet levels`,`x-axis levels`)
+
+  knitr::kable(harmonies,
                format = "latex",
@@
                 caption = "Harmonies with a pair of cyclic granularity one placed on facet and the other on x-axis.
-Out of 42 possible combinations of cyclic granularities, only 16 are harmony pairs.
-The harmony pairs are ranked basis highest to lowest MMPD and only the first 6 show significant variation in the measured variable.") %>%
+Out of 42 possible combinations of cyclic granularities, only 16 are harmony pairs.") %>%
     #row_spec(0, bold = TRUE) %>%
@@
 -->
-
-In \autoref{fig:bothcust} (b), <!--the distribution of energy consumption is plotted across the harmony pair (weekday/weekend, hour of the day) through an area quantile plot-->the black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
-The energy consumption is extremely (positive- or right-) skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
-This is common across both weekdays and weekends.
-For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer as compared to weekends.
-
-Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust}(c).
-There is additional information that we can derive looking at the distribution.
-There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
-<!--Also the hours from 7 to 13 look most volatile.--> If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
-For example in \autoref{fig:bothcust}(a), it can be seen that median energy consumption for the early morning hours is extremely high for weekdays compared to weekends.
-Also, outliers are more prominent in the latter part of the day.
-All of these indicate that looking at different distribution geometry or changing the mapping might shed lights on different aspect of the energy behavior for the same sample population.
 
@@
 This is an intuitive possibility, because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
-A strategy for analyzing multiple keys together could be to first group them basis time series or demographic features and then look at their energy behavior.
+A strategy for analyzing periodicities of multiple keys together could be to group the keys basis their energy behavior across different cyclic granularities.
 This is beyond the scope of the current work.
@@
 
-This case study shows systematic exploration of energy behavior for a household to gain exhaustive insights on periodic behavior of the households. 
+<!--This case study shows systematic exploration of energy behavior for a household to gain exhaustive insights on periodic behavior of the households. -->
+
 <!--First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualize a harmony pair and shows the effect of different distribution plots or reverse mapping.-->

### 2020-08-05 main.Rmd
cricket application minor change

@@
 
-\noindent Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the <!--siginificance of MMPD, threshold and the -->impact of different distribution plots and reverse mapping.
+\noindent Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the <!--significance of MMPD, threshold and the -->impact of different distribution plots and reverse mapping.
 For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
@@
 Although there is no conventional time component in cricket, each ball can be thought to represent an ordering from past to future with the game progressing forward with each ball.
-In the Twenty20 format, an over will consist of 6 balls (with some exceptions), an inning is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
+In the Twenty20 format, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
 Thus, similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
 The idea of cyclic granularities can be likewise mapped to this hierarchy.
-Example granularites then include ball of the over, over of the inning and ball of the inning.
+Example granularites then include ball of the over, over of the innings and ball of the innings.
 Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularities are aperiodic.
-For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an inning finishes before the over finishes.
+For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an innings finishes before the over finishes.
 Thus, the cyclic granularity ball-of-over will be circular in most cases and aperiodic in others.
@@
 
-The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
+\noindent The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
 The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each inning has 20 overs and each match has 2 innings.
+The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each innings has 20 overs and each match has 2 innings.
 This could be useful in a periodic world when we wish to compute any circular/quasi-circular granularity based on a hierarchy table which look like \autoref{tab:hierarchy-cric}.
+However, even if the situation is not periodic, it can be interesting to visualize the distribution of a measured variable to shed light on the aperiodic behavior of a non-temporal data set similar to aperiodic events like formal meetings, workshops, conferences, school semesters in a temporal set up.
+<!--There are many interesting questions that could possibly be answered with such a data set irrespective of the type of cyclic granularities.-->
 
@@
 
-However, even if the situation is not periodic and a similar hierarchy can not be formed, it can be interesting to visualize the distribution of a measured variable across relevant cyclic granularities to shed light on the aperiodic behavior of a non-temporal data set similar to aperiodic events like formal meetings, workshops, conferences, school semesters in a temporal set up.
-There are many interesting questions that could possibly be answered with such a data set irrespective of the type of cyclic granularities.
  <!--We will explore a few interesting questions and understand how the proposed approach in the paper can help answer some of the questions.-->
@@
 
-First, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second innings.
+\noindent There are many interesting questions that could possibly be answered with the `cricket` data set.
+For example, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second innings.
 The Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
 We take their example in order to dive deeper into this question.
-<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first inning there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
+<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
 This seem to indicate that players feel mounting pressure to score more runs as they approach towards the end of the first inning.
-Whereas teams batting in the second inning have a set target in mind and are not subjected to such mounting pressure and may adopt a more conservative strategy, to score runs.
+Whereas teams batting in the second innings have a set target in mind and are not subjected to such mounting pressure and may adopt a more conservative strategy, to score runs.
 Thus winning teams like CSK and MI seem to employ different inning strategies when it comes to their batting order.
@@
 
-Another interesting question could be: do runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
+\noindent Another fascinating exploration would be to examine if runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
 For establishing the fielding quality, we apply an indicator function on dismissals (1 if there was at least one wicket in the previous over due to run out or catch, 0 otherwise).
@@
 Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over as compared to the case when there has been at least one dismissal in the previous over.
+Wickets per over are considered as an aperiodic cyclic granularity with wickets as an aperiodic linear granularity.
+These granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
+These are similar to holidays or special events in temporal data.
 
@@
 
-Wickets per over are considered as an aperiodic cyclic granularity with wickets as an aperiodic linear granularity.
-These granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
-These are similar to holidays or special events in temporal data.
 <!-- CONTENT CUT -->

### 2020-08-05 main.Rmd
minor edits

@@
 
-Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ such that Definition \ref{def:periodical} does not hold true for them with respect to $B$ (the bottom granularity).
+\noindent Aperiodic cyclic granularities are defined using aperiodic linear granularities.
+Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ such that Definition \ref{def:periodical} does not hold true with respect to $B$.
 <!--that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.--> However, $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
@@
 \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}.
-It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
-Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.}
+It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.}
 \end{example}
+
+ <!-- Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks. -->
 
@@
 Nevertheless, for any calendar a hierarchy table can be defined.
-
-<!-- for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$.  -->
-<!-- so that cyclic granularities $C$ could be defined relative to these linear granularities. -->
-
-Note that, it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
+Note that it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
 Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
@@
 In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
-This is similar to calendar algebra in @Ning2002-tf for linear granularities but caters to computation of cyclic granularities, we shall refer to it as through cyclic calendar algebra operations.
+This is similar to calendar algebra in @Ning_2002 for linear granularities but caters to computation of cyclic granularities.
+We shall refer to it as through cyclic calendar algebra operations.
 The cyclic calendar algebra consists of broadly two kinds of operations:
@@
 
-\noindent Similar to single-to-multiple operations, multiple-to-single operations also involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
+\noindent Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
 For a hierarchy table $H_n: (G, C, K)$ with only  circular single-order-up granularities in the hierarchy and $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, multiple-order-up granularities could be obtained using \ref{eq:all-circular-multiple}.
@@
 These functions can be imported in the `gravitas` package to accommodate for aperiodic cyclic granularities.
+Another future direction of work could be to further refine the search of harmonies by only selecting pairs of cyclic granularities for which variation in measured variable is significant and rating these selected harmony pairs in order of importance for exploration.
 
@@
 The authors would like to thank the cohort [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages and Dr.
-Peter Tosca [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
+Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
 The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).

### 2020-08-05 main.Rmd
shifting code to script folder

@@
 options("knitr.graphics.auto_pdf" = TRUE)
-library(bookdown)
-library(rticles)
 library(knitr)
-library(tidyverse)
-library(lubridate)
-library(lvplot)
-library(ggridges)
-library(tsibble)
-if (requireNamespace("gravitas") == F)
-  remotes::install_github("Sayani07/gravitas")
-library(gravitas)
-library(ggpubr)
-library(xtable)
-library(kableExtra)
 opts_chunk$set(
@@
 
-```{r external, include = FALSE}
-# read_chunk('scripts/main.R')
+```{r external, include = FALSE, cache = FALSE}
+read_chunk('scripts/main.R')
 ```
@@
 "}
-knitr::include_graphics("Figs/linear-ex.png")
 ```
@@
 The circular granularity creates repeated integer sequences."}
-knitr::include_graphics("Figs/circular-ex.png")
+
 ```
@@
 ```{r search, echo=FALSE, eval = FALSE}
-library(gravitas)
 smart_meter %>%
@@
 ```{r hierarchy-cric}
-library(gravitas)
-library(tibble)
 hierarchy_model <- tibble::tibble(
@@
 
-library(tsibble)
-
 cricket_all <- read_csv("data-raw/deliveries_all.csv")
@@
 R package `gravitas`.
---> A missing piece is to enable user-defined temporal calendars.
-Also, computation of cyclic aperiodic granularities would require computing aperiodic linear granularities first.
+--> A missing piece in the package `gravitas` is the  computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
 A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.

### 2020-08-05 main.R

@@ -1,1861 +1,0 @@
-### Supplemental Materials (Data and scripts)
-
-
-
-
-
-
-
-## ---- load
-library(knitr)
-library(tidyverse)
-library(lubridate)
-library(lvplot)
-library(ggridges)
-library(tsibble)
-library(gravitas)
-library(tsibble)
-library(ggpubr)
-# install.packages("devtools")
-# devtools::install_github("Sayani07/gravitas")
-library(gravitas)
-
-## ---- linear-time
-knitr::include_graphics("Figs/linear-arr.png")
-
-## ---- mvtsplot1
-# Figure 1 : Multiple parallel time series plot of a real world data set
-# obtained using a fiber optic cable
-t <- nrow(anomalous_stream_1)
-f <- ncol(anomalous_stream_1)
-g <- as_tibble(anomalous_stream_1) %>%
-  gather() %>%
-  mutate(key = rep((1:f), each = t)) %>%
-  mutate(Time = rep(1:t, f))
-colnames(g) <- c("Cable", "Value", "Time")
-p1 <- ggplot(g, aes(x = Time, y = Cable, fill = Value)) +
-  geom_tile() +
-  scale_fill_gradientn(
-    colours = c("#F0E442", "#000000", "#000000"),
-    values = c(0, 0.1, max(anomalous_stream_1))
-  ) +
-  ylab("Cable (Sensor ID / Time Series ID)") +
-  scale_y_continuous(breaks = seq(0, 600, 100)) +
-  scale_x_continuous(breaks = seq(0, 1400, 200)) +
-  xlab("Time") +
-  theme(legend.position = "none", axis.title = element_text(size = 15))
-print(p1)
-
-
-
-
-
-
-
-## ---- outtype
-# Figure 2: Different types of anomalies in temporal data
-set.seed(1234)
-ts1 <- as.ts(rnorm(500, 10, 3))
-ts1[300] <- 35
-ts1[150] <- 30
-ts1 <- as_tibble(ts1) %>% mutate(Time = 1:500)
-colnames(ts1) <- c("Value", "Time")
-point <- tibble(Time = c(150, 300), Value = c(30, 35))
-p1 <- ggplot(ts1) +
-  geom_line(aes(x = Time, y = Value), size = 0.5, alpha = 0.7) +
[…]

### 2020-08-05 scripts/main.R

@@ -1,0 +1,22 @@
+## ---- load
+library(bookdown)
+library(rticles)
+library(knitr)
+library(tidyverse)
+library(lubridate)
+library(lvplot)
+library(ggridges)
+library(tsibble)
+if (requireNamespace("gravitas") == F)
+  remotes::install_github("Sayani07/gravitas")
+library(gravitas)
+library(ggpubr)
+library(xtable)
+library(kableExtra)
+library(tibble)
+
+## ---- linear-time
+knitr::include_graphics("Figs/linear-ex.png")
+
+## ---- circular-dow
+knitr::include_graphics("Figs/circular-ex.png")

### 2020-08-05 main.Rmd
all R codes shifted to scripts folder

@@
 
-knitr::include_graphics("Figs/aperiodic-ex3.png")
 ```
@@
 ```{r tab-mayan}
-table_mayan <- tibble(`linear (G)` = c("kin", "uinal", "tun", "katun", "baktun"),
-                      `single-order-up cyclic (C)` = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
-                       `period length/conversion operator (K)`= c(20, 18, 20, 20, 1))
-kable(table_mayan,
-      format = "latex",
-      booktabs = TRUE,
-      align = "l",
-      caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities.") %>%
-  kable_styling()
-  #row_spec(0, bold = TRUE) %>%
-  #row_spec(1:(nrow(table_mayan)-1), hline_after = T)
-  #.
 ```
@@
 ```{r tab-gregorian}
-table_greg <- tibble(`linear (G)` = c("minute", "hour", "day", "month", "year"),
-                      `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
-                     `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1))
-
-kable(table_greg,
-      format = "latex",
-      booktabs = TRUE,
-      caption = "Hierarchy table for the Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
-  #row_spec(0, bold = TRUE) %>%
-   kable_styling()
 ```
@@
 ```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
-include_graphics("Figs/data-struc-diffcol.png")
 ```
@@
 ", fig.show = 'hold'}
-
-load("data/sm_cust50.rds")
-cust50 <- sm_cust50 %>%  distinct(customer_id)
-
-VIC <- sm_cust50 %>% as_tsibble(regular = FALSE) %>% 
-  filter(customer_id==cust50$customer_id[2])
-
-scene1 <-  VIC %>%
-  #create_gran("quarter_year") %>%
-  #filter(month_year %in% c("Jan", "Mar", "Dec")) %>%
-  #create_gran("day_week") %>%
-  #filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
-  prob_plot("quarter_year", "wknd_wday",
-            response = "general_supply_kwh",
-            plot_type = "lv",
-            symmetric = FALSE
-  ) +
-  ylab("") +
-  #xlab("weekend/weekday") +
-  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + xlab("")+ 
-  theme_minimal()
-
-scene2 <- VIC %>%
-  prob_plot("wknd_wday",
-            "quarter_year",
-            response = "general_supply_kwh",
-            plot_type = "lv",
-            symmetric = FALSE
-  ) +
-  ylab("") +
-    xlab("quarters of the year")  +
-  scale_fill_brewer(palette = "Paired") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + theme_minimal()
-
-scene3 <- VIC %>%
-  prob_plot("quarter_year",
-            "month_year",
-            response = "general_supply_kwh",
-            plot_type = "lv",
-            symmetric = FALSE
-  ) +
-  ylab("") +
-  xlab("months of the year") + theme(legend.position = "right") + scale_fill_brewer(palette = "Set2") + ggtitle("") + scale_y_log10()+ theme_minimal()
-
-gg_fig <- ggarrange(scene3,
-          ggarrange(scene1, scene2, ncol = 2, labels = c("b", "c")),
-          nrow = 2, labels = "a")
-          #label.y = "electricity demand [KWh]"\
-
-ggpubr::annotate_figure(gg_fig,
-                left = text_grob("electricity demand [KWh]",  rot = 90)) + theme_minimal()
 
@@
 ```{r search, echo=FALSE, eval = FALSE}
-smart_meter %>%
-  search_gran(filter_out = c("semester", 
-                             "quarter",
-                             "fortnight"))
 ```
@@
 ```{r search_gran_limit2, echo = FALSE, eval = FALSE}
-smart_meter10 %>% search_gran(
-  highest_unit = "month",
-  filter_out = c("hhour", "fortnight"))%>%
-  knitr::kable(format = "latex",
-      booktabs = TRUE) %>%
-  #row_spec(0, bold = TRUE)%>%
-  kable_styling()
 ```
@@
 ```{r harmony-tab, echo=FALSE, cache=TRUE}
-# if(!dir.exists("cache/"))
-# {
-#   dir.create("cache")
-# }
-# if (!file.exists("cache/smart_harmony.rds")) {
-  sm <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936))
-
-  harmonies <- sm %>%
-  harmony(ugran = "month",
-          filter_in = "wknd_wday", 
-          filter_out = c("hhour", "fortnight")) %>%
-    rename(`facet variable` = facet_variable,
-         `x-axis variable` = x_variable,
-         `facet levels` = facet_levels,
-         `x-axis levels` = x_levels)
-
-# set.seed(12345)
-# 
-# global_harmony <- sm %>%
-#   global_threshold(harmony_tbl = harmonies,
-#                    response = "general_supply_kwh", nsamp = 5) %>%
-#   rename(`facet variable` = facet_variable,
-#          `x-axis variable` = x_variable,
-#          `facet levels` = facet_levels,
-#          `x-axis levels` = x_levels) %>% 
-#   mutate(MMPD = if_else(gt_MMPD == TRUE,
-#                         paste0(MMPD, "*"), paste0(MMPD))) %>%
-#   select(`facet variable`,`x-axis variable`,
-#          `facet levels`,`x-axis levels`)
-
-  knitr::kable(harmonies,
-               format = "latex",
-               booktabs = TRUE,
-                caption = "Harmonies with a pair of cyclic granularity one placed on facet and the other on x-axis.
-Out of 42 possible combinations of cyclic granularities, only 16 are harmony pairs.") %>%
-    #row_spec(0, bold = TRUE) %>%
-    kable_styling()
 ```
@@
 It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.", out.width="90%"}
-
-# plot for only customer is shown
-# cust1 <- smart_meter10 %>%
-# filter(customer_id %in% c(10006704)) %>%
-#   prob_plot("wknd_wday",
-#                             "hour_day",
-#                             response = "general_supply_kwh",
-#                             plot_type = "quantile",
-#             symmetric = FALSE,
-#             quantile_prob = c(0.25, 0.5, 0.75, 0.9))
-#   scale_y_sqrt() +
-#   ggtitle("Energy consumption distribution for customer id: 10006704")
-cust2_quantile <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("wknd_wday",
-    "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "quantile",
-    symmetric = TRUE,
-    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
-  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
-  theme_minimal()
-#
-# smart_meter10 %>%
-#   filter(customer_id %in% c(10006704)) %>%
-#   prob_plot("wknd_wday",
-#                             "hour_day",
-#                             response = "general_supply_kwh",
-#                             plot_type = "violin") +
-#   scale_y_sqrt() +
-#   ggtitle("Energy consumption distribution for customer id: 10006704")
-
-cust2_violin <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("wknd_wday",
-    "hour_day",
-    response = "general_supply_kwh",
-    plot_type = "violin"
-  ) +
-  scale_y_sqrt()+
-  #ggtitle(" (c) Violin plot faceted by weekend-weekday") +
-  ylab("") + xlab("hours of the day") + ggtitle("") +
-  scale_x_discrete( breaks = seq(0, 23, 5))+ 
-  theme_minimal()
-
-cust2_box <- smart_meter10 %>%
-  filter(customer_id %in% c(10017936)) %>%
-  prob_plot("hour_day",
-    "wknd_wday",
-    response = "general_supply_kwh",
-    plot_type = "boxplot"
-  ) +
-  scale_y_sqrt() +
-  #ggtitle(" (a) Box plot faceted by hour-of-day") +
-  xlab("") +
-  ggtitle("") + ylab("") +
-  scale_x_discrete(labels = c('wday','wend')) +
-  ggplot2::theme(axis.text.x = element_text(size = 7))+ 
-  theme_minimal()
-
-#ggarrange(cust2_quantile, cust2_violin, nrow = 2)
-
-gg_fig <- ggarrange(cust2_box,
-          ggarrange(cust2_quantile, cust2_violin, nrow = 2, labels = c("b", "c")),
-          ncol = 2, labels = "a")
-          #label.y = "electricity demand [KWh]"\
-
-ggpubr::annotate_figure(gg_fig,
-                left = text_grob("electricity demand [KWh]",  rot = 90))
 ```
@@
 ```{r hierarchy-cric}
-hierarchy_model <- tibble::tibble(
-  `linear (G)` = c("over", "inning", "match", "season"),
-  `single-order-up cyclic (C)` = c("over-of-inning", "inning-of-match", "match-of-season", 1),
-  `period length/conversion operator (K)` = c(20, 2, "k(match, season)", 1)
-)
+
 knitr::kable(hierarchy_model,
@@
 
-library(tsibble)
-cricket_tsibble <- cricket %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-hierarchy_model <- tibble::tibble(
-  units = c("index", "over", "inning", "match"),
-  convert_fct = c(1, 20, 2, 1)
-)
-
-cricket_tsibble %>%
-  filter(batting_team %in% c(
-    "Mumbai Indians",
-    "Chennai Super Kings"
-  )) %>%
-  mutate(inning = paste0("innings: ", inning)) %>%
-  prob_plot("inning",
-    "over",
-    response = "runs_per_over",
-    hierarchy_model,
-    plot_type = "lv"
-  ) +
-  scale_fill_brewer(palette = "Dark2") +
-  #ggtitle("(a) Runs per over across over faceted by inning") +
-  theme(legend.position = "right") +
-  ggtitle("a") +
-  ylab("runs per over") +
-  xlab("overs of the innings") +
-  theme(plot.title = element_text(face = "bold")) +
-  ggplot2::theme(
-       strip.text = ggplot2::element_text(
-        size = 10,
-         margin = ggplot2::margin(b=0, t=0)
-      )
-     ) +  theme_minimal() 
-#geom_smooth(aes( x = over,
-#                               y=runs_per_over), method = lm, #formula = y ~ splines::bs(x, 3), se = FALSE)
-
-cricket_all <- read_csv("data-raw/deliveries_all.csv")
-matches_all <- read_csv("data-raw/matches_all.csv")
-
-cricket_season <- cricket_all %>% left_join(matches_all, by = c("match_id" = "id"))
-
-# cricket_per_over <- cricket_season %>%
-#   group_by(season,
-#            match_id,
-#            batting_team,
-#            bowling_team,
-#            inning,
-#            over) %>%
-#   summarise(runs_per_over = sum(total_runs),
-#             run_rate = sum(total_runs)/length(total_runs))
-#
-# cricket_tsibble_all <- cricket_per_over %>%
-#   ungroup() %>%
-#   mutate(data_index = row_number()) %>%
-#   as_tsibble(index = data_index)
-
-cricket_dot_field <- cricket_season %>%
-  mutate(
-    fielding_proxy = if_else(dismissal_kind %in%
-      c("caught", "caught and bowled"), 1, 0),
-    dot_ball_proxy = if_else(total_runs == 0, 1, 0),
-    wicket_proxy = if_else(is.na(dismissal_kind), 0, 1)
-  ) %>%
-  group_by(
-    season,
-    match_id,
-    batting_team,
-    bowling_team,
-    inning,
-    over
-  ) %>%
-  summarise(
-    runs_per_over = sum(total_runs),
-    run_rate = sum(total_runs)*6 / length(total_runs),
-    fielding_wckts = sum(fielding_proxy),
-    dot_balls = sum(dot_ball_proxy)
-  ) %>%
-  mutate(diff_run_rate = c(0, diff(run_rate)))
-
-cricket_tsibble <- cricket_dot_field %>%
-  ungroup() %>%
-  mutate(data_index = row_number()) %>%
-  as_tsibble(index = data_index)
-
-cricket_data <- cricket_tsibble %>%
-  mutate(
-    field = if_else(fielding_wckts == 0, "0", "1+"),
-    dot = if_else(dot_balls == 0, "no dot balls", ">0 dot balls"),
-    lag_field = lag(field),
-    lag_dot = lag(dot)
-  ) %>%
-  filter(lag_field != 0, lag_dot != 0)
-
-cricket_data$lag_field <- factor(cricket_data$field, levels = c("0", "1+"))
-
-# filter(fielding_wckts %in% c(0,1)) %>%
-#
- cricket_data %>%
-  filter(over!=1) %>%
-  prob_plot("over", "lag_field",
-  hierarchy_model,
-  response = "run_rate",
-  plot_type = "quantile",
-  symmetric = FALSE,
- quantile_prob = c(0.25, 0.5, 0.75)) +
-   #ggtitle("(b) Runs per over across overs faceted by number of wickets in previous over") +
-  ylab("runs per over")  +
-  xlab("number of wickets in previous over") +
-  ggtitle("b") +
-   theme(plot.title = element_text(face = "bold")) +
-        theme(axis.ticks = element_blank(), legend.background = element_blank(),
-            legend.key = element_blank(), panel.background = element_blank(), strip.background = element_blank(),
-            plot.background = element_blank(), complete = TRUE, panel.grid.major = element_line(colour = "#E0E0E0"),
-            panel.border = element_rect(colour = "#E0E0E0", fill = NA))
 ```

### 2020-08-05 scripts/main.R

@@ -20,3 +20,293 @@ knitr::include_graphics("Figs/linear-ex.png")
 
 ## ---- circular-dow
 knitr::include_graphics("Figs/circular-ex.png")
+
+## ----aperiodic-example
+knitr::include_graphics("Figs/aperiodic-ex3.png")
+
+## ----tab-mayan
+table_mayan <- tibble(`linear (G)` = c("kin", "uinal", "tun", "katun", "baktun"),
+                      `single-order-up cyclic (C)` = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
+                      `period length/conversion operator (K)`= c(20, 18, 20, 20, 1))
+kable(table_mayan,
+      format = "latex",
+      booktabs = TRUE,
+      align = "l",
+      caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities.") %>%
+  kable_styling()
+
+##----tab-gregorian
+table_greg <- tibble(`linear (G)` = c("minute", "hour", "day", "month", "year"),
+                     `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
+                     `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1))
+
+kable(table_greg,
+      format = "latex",
+      booktabs = TRUE,
+      caption = "Hierarchy table for the Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
+  #row_spec(0, bold = TRUE) %>%
+  kable_styling()
+
+## ----data-structure
+include_graphics("Figs/data-struc-diffcol.png")
+
+## ----allFig
+load("data/sm_cust50.rds")
+cust50 <- sm_cust50 %>%  distinct(customer_id)
+
+VIC <- sm_cust50 %>% as_tsibble(regular = FALSE) %>% 
+  filter(customer_id==cust50$customer_id[2])
+
+scene1 <-  VIC %>%
+  prob_plot("quarter_year", "wknd_wday",
+            response = "general_supply_kwh",
+            plot_type = "lv",
+            symmetric = FALSE
+  ) +
+  ylab("") +
+  #xlab("weekend/weekday") +
+  scale_fill_brewer(palette = "Dark2") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + xlab("")+ 
+  theme_minimal()
+
+
+scene2 <- VIC %>%
+  prob_plot("wknd_wday",
[…]

### 2020-08-05 main.Rmd
all figure caption clean

@@
 
-```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="Illustration of time domain, linear granularities and index set.
+```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="(ref:linear-time)"}
+```
+
+(ref:linear-time) Illustration of time domain, linear granularities and index set.
 Hour, day, week, month and year are linear granularities and can also be considered to be time domains.
@@
 Hours could also be considered the index set, and a bottom granularity.
-"}
-```
 
@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap="Illustration of circular relative to linear granularities (a).
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap= "(ref:circular-dow)"}
+
+```
+
+(ref:circular-dow) Illustration of circular relative to linear granularities (a).
 Circular granularities can be considered to be cutting the linear granularity into pieces and stacking them (b).
-The circular granularity creates repeated integer sequences."}
-
-```
+The circular granularity creates repeated integer sequences.
 
@@
 
-```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
+```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap= "(ref:aperiodic-example)"}
+
+```
+(ref:aperiodic-example)  Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
 The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future.
@@
 This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 18.3 weeks will be a quasi-circular granularity with repeating patterns.
-$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years."}
-
-```
+$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years.
 
@@
 
-```{r data-structure, fig.cap = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.", fig.pos="htb"}
+```{r data-structure, fig.cap = "(ref:data-structure)", fig.pos="htb"}
 ```
+
+(ref:data-structure) The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.
 
@@
 
-```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap="Distribution of energy consumption displayed through letter value plots.
+```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
+
+```
+
+(ref:allFig) Distribution of energy consumption displayed through letter value plots.
 Plot (a) displays consumption across month-of-year faceted by quarters of the year, (b) across weekday/weekend faceted by quarters of the year and (c) across quarters of the year faceted by weekday/weekend.
@@
 Plot (a) shows a clash since all quarters do not correspond to all months of the year.
-", fig.show = 'hold'}
-
-```
 
@@
 
-```{r bothcust, fig.cap = "Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
+```{r bothcust, fig.cap = "(ref:bothcust)", out.width="90%"}
+```
+
+(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
 Two granularities are used: hour of the day (I) and weekday/weekend (II).
@@
 Plot (c) uses a violin plot to place emphasis on distributional differences across hours.
-It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.", out.width="90%"}
-```
+It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.
 
@@
 
-```{r cricex, fig.cap= "Runs per over shown with different distribution displays, and granularities.
+```{r cricex, fig.cap= "(ref:cricex)", warning = FALSE, message = FALSE, out.width = "90%"}
+
+```
+
+(ref:cricex) Runs per over shown with different distribution displays, and granularities.
 Plot (a) shows letter value plot across overs faceted by innings.
@@
 Plot (b) shows quantile plot of runs per over across an indicator of wickets in previous over faceted by current over.
-This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.", warning = FALSE, message = FALSE, out.width = "90%"}
-
-```
+This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.
 

### 2020-08-05 main.Rmd
cleaning some extra text

@@
 
- <!-- Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks. -->
-
-<!--Each granule is represented by a box.
-The diagram also illustrates that the granules that overlap share elements from the underlying time domain.
-The first slot in the diagram shows the index set and $A(B, M)$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.-->
-
-<!-- Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labor day and M can be conceived as the aperiodic linear granularity denoting any holidays.
-<!--The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.-->
-
-<!-- map a time index to a categorical variable containing binary values.
-It takes the value 1 whenever the time index coincides with the events in context (e.g.
-public or school holiday) and 0 when it does not.
-Here, 1 and 0 are proxies to denote the binary categorization and can be replaced with a logical vector.
--->
-
-<!-- #  -->
-<!-- # <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
-
 ## Relativities {#sec:cyclic-calendar}
@@
 
-<!-- CONTENT CUT -->
-<!-- Clearly, the computation of multiple-order-up granularities become more involved with quasi-circular granularities.
-$C_{hour, month}(z)$ can be represented as function of circular granularity $C_{hour, day}(z)$ and quasi-circular granularity $C_{day, month}(z)$, both of which are single order-up.
-However, $C_{day, year}(z)$ can not be computed only with single order-up granularities $C_{day, month}(z)$ and $C_{month, year}(z)$.
-It also requires the knowledge of the composition of the linear granularity $days$ within $months$ in an $year$ for computation of $\vert T_w \vert$.
--->
-<!-- If $C_{G_{l},G_{m'}}(z)$ or $C_{G_{l},G_{m'}}(z)$ are multiple-order-up circular granularities, they could be computed using similar approach in Section \ref{sec:all-circular}. -->
-
-<!-- \begin{equation} \label{eq:multifromsingle-quasi} -->
-<!-- \begin{split} -->
-<!-- C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z))) \\  -->
-<!-- & = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))) + P(l, m')(C_{G_{m'},G_{m}}(z))\\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-
-<!-- where $G_{m'}$ is the first linear granularity in the hierarchy table that groups periodically into the bottom granularity with irregular mapping. -->
-
-<!-- The aperiodic granularity $C_{day, quarter}(z))$ in \autoref{eq:hour-month} can be broken further. -->
-
-<!-- \begin{equation} \label{eq11} -->
-<!-- \begin{split} -->
-<!-- A_{(hour, month)} & = (\sum_{w=0}^{9 - 1}(\vert T(w)\vert) +  20)/24 + 12 \\ -->
-<!--                   & = (\sum_{w=0}^{C_{month, year} - 1}(\vert T(w)\vert) +  A_{day, month})*k(hour_day) + A_{hour, day} \\ -->
-<!--                   & = (A{day_year} - 1)*k(hour_day) + A_{hour, day} \\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-<!-- where, $month(w) = \bigcup_{z \in T_w}day(z)$ -->
-
-<!--  With one aperiodic granularity, we can write the following:  -->
-<!-- A_{P, Q)} & = (A{AG_Q} - 1)*k(P_AG) + A_{P, AG} {TO REVISE: many aperiodic granularties not to be included} -->
-
 ### Multiple-to-single {#sec:multiple-to-single}
@@
 Single-order-up quasi-circular granularity could be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}.
-
-<!-- \begin{equation} \label{eq:multifromsingle-quasi3} -->
-<!-- \begin{split} -->
-<!-- C_{G_{l},G_{m'}}(z) & = C_{G_l,G_{m}}(z) - \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\ -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-
-<!-- where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$, $\vert T_w \vert$ is the cardinality of set $T_w$. -->
-<!-- as they promote the exploratory analysis through visualization. -->
-
-```{r}
-# grid.arrange(p1,p2,p3,p4,nrow = 2)
-
-#
-# x <- gridExtra::arrangeGrob (p5, p6, p1, p2, p3, p4,newpage = TRUE)
-# grid::grid.draw(x)
-
-#
-# gridExtra::grid.arrange(
-#   p1,
-#   p2,
-#   p3,
-#   p4,
-#   p5,
-#   p6,
-#   nrow = 3,
-#   top = "Title of the page",
-#   bottom = grid::textGrob(
-#     "this footnote is right-justified",
-#     gp = grid::gpar(fontface = 3, fontsize = 9),
-#     hjust = 1,
-#     x = 1
-#   )
-# )
-# plot.new()
-# mtext('Various probability distribution plot of electricity consumption data of Victoria from 2012 to 2014 (top left) Letter Value plot by DoM and MoY (top right) Decile Plot by HoD and DoW (middle left) Box plot by DoY and DoM (middle right) Violin plot of DoM and WoM (bottom left) Ridge plot by WoM and WoY (bottom right) Decile plot by DoY and MoY', side = 3, line = - 10, outer = TRUE)
-```
-
-<!-- More generally, if we have two circular granularities C1 and C2 which has [A, B, C, D] and [X, Y, Z] categories/levels.
-When C1 and C2 are used as aesthetics or facets, problems arise when we encounter empty combinations.
-Let $S_{i,j}$ be the set of combination of the levels of C1 and C2.
-In this example, we have 12 such sets $S_{i,j}$ because i can take 4 values and j can take 3 values.
--->
-
-<!-- Thus, the common link that differentiates harmonies from clashes are: -->
-
-<!--   a) There should not be any levels of the faceted variable which is empty for one or more levels of the factor plotted across x-axis. -->
-
-<!--   b) There should not be any level of the factor plotted across x-axis which doesn't have values for all levels of factors plotted across facets. -->
 

### 2020-08-06 main.Rmd
minor
fixing R citations

@@
 ```{r hierarchy-cric}
-
-knitr::kable(hierarchy_model,
-             format = "latex",
-             booktabs = TRUE,
-             caption = "Hierarchy table for cricket where overs are nested within an inning, innings nested within a match and matches within a season.") %>%
-  #row_spec(0, bold = TRUE)%>%
-  kable_styling()
 ```
@@
 More details about the package can be found on the package website [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
-This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+This article was created with `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
 This paper's Github repository, [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
@@
 ```{r write-bib, eval = FALSE}
-write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib")
+write_bib(c("rmarkdown", "bookdown", "rticles", "knitr", "ggplot2", "dplyr", "tsibble", "gravitas", "ggpubr", "kableExtra", "magrittr"), file = "rpkgs.bib")
 ```

### 2020-08-06 scripts/main.R

@@ -3,15 +3,9 @@ library(bookdown)
 library(rticles)
 library(knitr)
 library(tidyverse)
-library(lubridate)
-library(lvplot)
-library(ggridges)
 library(tsibble)
-if (requireNamespace("gravitas") == F)
-  remotes::install_github("Sayani07/gravitas")
 library(gravitas)
 library(ggpubr)
-library(xtable)
 library(kableExtra)
 library(tibble)
 
@@ -194,6 +188,13 @@ hierarchy_model <- tibble::tibble(
   `period length/conversion operator (K)` = c(20, 2, "k(match, season)", 1)
 )
 
+knitr::kable(hierarchy_model,
+             format = "latex",
+             booktabs = TRUE,
+             caption = "Hierarchy table for cricket where overs are nested within an innings, innings nested within a match and matches within a season.") %>%
+  #row_spec(0, bold = TRUE)%>%
+  kable_styling()
+
 ## ----cricex
 cricket_tsibble <- cricket %>%
   mutate(data_index = row_number()) %>%
@@ -310,3 +311,4 @@ cricket_data %>%
         legend.key = element_blank(), panel.background = element_blank(), strip.background = element_blank(),
         plot.background = element_blank(), complete = TRUE, panel.grid.major = element_line(colour = "#E0E0E0"),
         panel.border = element_rect(colour = "#E0E0E0", fill = NA))
+

### 2020-08-06 main.Rmd
fix Earo's papers from to appear to appeared version

@@
 \noindent Effective exploration and visualization benefits from well-organized data structures.
-@wang2019tsibble introduced a tidy data structure, tsibble, to support exploration and modeling of temporal data.
+@wang2020tsibble introduced a tidy data structure, tsibble, to support exploration and modeling of temporal data.
 This forms the basis of the structure for cyclic granularities.
@@
 Two granularities are used: hour of the day (I) and weekday/weekend (II).
-Plot (a) shows granularity I facetted by granularity II, and plots (b), (c) shows the converse mapping.
+Plot (a) shows granularity I faceted by granularity II, and plots (b), (c) shows the converse mapping.
 Plot (a) makes a comparison of usage by workday within each hour of the day using side-by-side boxplots.

### 2020-08-06 main.Rmd
fixing renv package dependencies

@@
 ```{r write-bib, eval = FALSE}
-write_bib(c("rmarkdown", "bookdown", "rticles", "knitr", "ggplot2", "dplyr", "tsibble", "gravitas", "ggpubr", "kableExtra", "magrittr"), file = "rpkgs.bib")
+write_bib(c("rmarkdown", "bookdown", "rticles", "knitr", "ggplot2", "dplyr", "tsibble", "gravitas", "ggpubr", "kableExtra", "magrittr", "readr", "lubridate", "renv","archive"), file = "rpkgs.bib")
 ```

### 2020-08-06 data-raw/cricketdata.R

@@ -5,12 +5,13 @@ library(dplyr)
 cricket <-  read_csv("data-raw/deliveries_all.csv")
 matches <- read_csv("data-raw/matches_all.csv")
 
-cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id")
+cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id"))
 
 # each over needs to have 6 balls
 
 cricket_fltr_ball <- cricket_season %>%
-  filter(wide_runs + noball_runs== 0) %>%    # filtering out scores that are gained by no balls or wide
+  filter(wide_runs + noball_runs== 0) %>% 
+  # filtering out scores that are gained by no balls or wide
   # still have to filter out overs for which number of balls per over less than 6
   group_by(batting_team, match_id, inning, over) %>%
   summarise(n_over = length(ball)) %>%

### 2020-08-06 data-raw/sm_cust50.R

@@ -1,4 +1,4 @@
-library(tidyverse)
+library(dplyr)
 library(tsibble)
 #devtools::install_github("jimhester/archive")
 library(archive)

### 2020-08-06 scripts/main.R

@@ -2,12 +2,14 @@
 library(bookdown)
 library(rticles)
 library(knitr)
-library(tidyverse)
+library(ggplot2)
+library(dplyr)
 library(tsibble)
 library(gravitas)
 library(ggpubr)
 library(kableExtra)
-library(tibble)
+library(magrittr)
+library(readr)
 
 ## ---- linear-time
 knitr::include_graphics("Figs/linear-ex.png")

### 2020-08-06 scripts/main.R
cache removed from the code harmony-tab

@@ -48,9 +48,11 @@ include_graphics("Figs/data-struc-diffcol.png")
 
 ## ----allFig
 load("data/sm_cust50.rds")
+sm_cust50 <- sm_cust50 %>% as_tsibble(regular = FALSE)
+
 cust50 <- sm_cust50 %>%  distinct(customer_id)
 
-VIC <- sm_cust50 %>% as_tsibble(regular = FALSE) %>% 
+VIC <- sm_cust50 %>% 
   filter(customer_id==cust50$customer_id[2])
 
 scene1 <-  VIC %>%

### 2020-08-26 main.Rmd
cache removed from the code harmony-tab
Code styling

--- 2020-08-06 main.Rmd
@@
 )
-#knitr::opts_knit$set(root.dir = here::here())
+# knitr::opts_knit$set(root.dir = here::here())
 ```
@@
 ```{r external, include = FALSE, cache = FALSE}
-read_chunk('scripts/main.R')
+read_chunk("scripts/main.R")
 ```
@@
 
-```{r harmony-tab, echo=FALSE, cache=TRUE}
+```{r harmony-tab, echo=FALSE}
 ```
@@
 ```{r write-bib, eval = FALSE}
-write_bib(c("rmarkdown", "bookdown", "rticles", "knitr", "ggplot2", "dplyr", "tsibble", "gravitas", "ggpubr", "kableExtra", "magrittr"), file = "rpkgs.bib")
+write_bib(c("rmarkdown", "bookdown", "rticles", "knitr", "ggplot2", "dplyr", "tsibble", "gravitas", "ggpubr", "kableExtra", "magrittr", "readr", "lubridate", "renv", "archive"), file = "rpkgs.bib")
 ```

### 2020-08-26 data-raw/cricketdata.R

@@ -1,58 +1,58 @@
 library(readr)
-library(tsibble)
 library(dplyr)
 
-cricket <-  read_csv("data-raw/deliveries_all.csv")
+cricket <- read_csv("data-raw/deliveries_all.csv")
 matches <- read_csv("data-raw/matches_all.csv")
 
-cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id"))
+cricket_season <- cricket %>%
+  left_join(matches, by = c("match_id" = "id"))
 
 # each over needs to have 6 balls
 
 cricket_fltr_ball <- cricket_season %>%
-  filter(wide_runs + noball_runs== 0) %>% 
+  filter(wide_runs + noball_runs == 0) %>%
   # filtering out scores that are gained by no balls or wide
   # still have to filter out overs for which number of balls per over less than 6
   group_by(batting_team, match_id, inning, over) %>%
   summarise(n_over = length(ball)) %>%
-  filter(n_over==6) %>%
+  filter(n_over == 6) %>%
   mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
 
 
 cricket_over_crctd <- cricket_season %>%
   mutate(cricket_key = paste(batting_team, match_id, inning, over, sep = "_")) %>%
   filter(cricket_key %in% cricket_fltr_ball$key) %>%
-  filter(wide_runs + noball_runs== 0) %>%
+  filter(wide_runs + noball_runs == 0) %>%
   mutate(ball_per_over = purrr::rep_along(match_id, 1:6))
 
 # each innings need to have 20 overs
 
 cricket_fltr_over <- cricket_over_crctd %>%
-  filter(inning %in% c(1,2)) %>%
+  filter(inning %in% c(1, 2)) %>%
   group_by(inning, match_id) %>%
-  summarise(n = length(over)) %>% filter(n==120) %>%
+  summarise(n = length(over)) %>%
+  filter(n == 120) %>%
   mutate(key = paste(match_id, inning, sep = "_"))
 
 cricket_over_inning_crctd <- cricket_over_crctd %>%
   mutate(over_key = paste(match_id, inning, sep = "_")) %>%
   filter(over_key %in% cricket_fltr_over$key) %>%
-  filter(wide_runs + noball_runs== 0) %>%
+  filter(wide_runs + noball_runs == 0) %>%
   mutate(ball_per_over = purrr::rep_along(match_id, 1:6)) %>%
[…]

### 2020-08-26 scripts/main.R

@@ -21,26 +21,32 @@ knitr::include_graphics("Figs/circular-ex.png")
 knitr::include_graphics("Figs/aperiodic-ex3.png")
 
 ## ----tab-mayan
-table_mayan <- tibble(`linear (G)` = c("kin", "uinal", "tun", "katun", "baktun"),
-                      `single-order-up cyclic (C)` = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
-                      `period length/conversion operator (K)`= c(20, 18, 20, 20, 1))
+table_mayan <- tibble(
+  `linear (G)` = c("kin", "uinal", "tun", "katun", "baktun"),
+  `single-order-up cyclic (C)` = c("kin-of-uinal", "uinal-of-tun", " tun-of-katun", "katun-of-baktun", 1),
+  `period length/conversion operator (K)` = c(20, 18, 20, 20, 1)
+)
 kable(table_mayan,
-      format = "latex",
-      booktabs = TRUE,
-      align = "l",
-      caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities.") %>%
+  format = "latex",
+  booktabs = TRUE,
+  align = "l",
+  caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities."
+) %>%
   kable_styling()
 
-##----tab-gregorian
-table_greg <- tibble(`linear (G)` = c("minute", "hour", "day", "month", "year"),
-                     `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
-                     `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1))
+## ----tab-gregorian
+table_greg <- tibble(
+  `linear (G)` = c("minute", "hour", "day", "month", "year"),
+  `single-order-up cyclic (C)` = c("minute-of-hour", "hour-of-day", "day-of-month", "month-of-year", 1),
+  `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1)
+)
 
 kable(table_greg,
-      format = "latex",
-      booktabs = TRUE,
-      caption = "Hierarchy table for the Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>%
-  #row_spec(0, bold = TRUE) %>%
+  format = "latex",
+  booktabs = TRUE,
[…]

### 2020-08-26 data-raw/sm_cust50.R
delete redundant pictures

@@ -1,44 +1,0 @@
-library(readr)
-library(dplyr)
-library(tsibble)
-# devtools::install_github("jimhester/archive")
-library(archive)
-
-# Raw data obtained from Electricity Use Interval Reading of Smart-Grid Smart-City Customer Trial Data available in
-# https://data.gov.au/dataset/ds-dga-4e21dea3-9b87-4610-94c7-15a8a77907ef/distribution/dist-dga-b71eb954-196a-4901-82fd-69b17f88521e/details?q=smart%20meter
-# file name:CD_INTERVAL_READING_ALL_NO_QUOTES.csv
-
-if (!file.exists("data-raw/smart-metre.7z")) {
-  download.file(
-    "http://datagovau.s3-ap-southeast-2.amazonaws.com/CDINTERVALREADINGALLNOQUOTES.csv.7z",
-    "data-raw/smart-metre.7z",
-    mode = "wb"
-  )
-}
-
-smart_meter_data_raw <- archive_read("data-raw/smart-metre.7z") %>%
-  read_csv(n_max = 3e6)
-
-set.seed(12345)
-sm_cust <- smart_meter_data_raw %>%
-  distinct(CUSTOMER_ID) %>%
-  pull(CUSTOMER_ID) %>%
-  sample(size = 50)
-
-sm_cust50 <- smart_meter_data_raw %>%
-  filter(CUSTOMER_ID %in% sm_cust) %>%
-  rename_all(tolower) %>%
-  arrange(customer_id, reading_datetime) %>%
-  group_by(customer_id) %>%
-  mutate(
-    reading_datetime = case_when(
-      duplicated(reading_datetime) ~ reading_datetime + lubridate::hours(1),
-      TRUE ~ reading_datetime
-    )
-  ) %>%
-  ungroup() %>%
-  mutate(customer_id = as.character(customer_id)) %>%
-  as_tsibble(index = reading_datetime, key = customer_id) %>%
-  select(-calendar_key)
-
-save(sm_cust50, file = "data/sm_cust50.rds", compress = "xz")

### 2020-08-26 scripts/main.R

@@ -55,13 +55,13 @@ kable(table_greg,
 include_graphics("Figs/data-struc-diffcol.png")
 
 ## ----allFig
-load("data/sm_cust50.rds")
-sm_cust50 <- sm_cust50 %>% as_tsibble(regular = FALSE)
+# load("data/sm_cust50.rds")
+# sm_cust50 <- sm_cust50 %>% as_tsibble(regular = FALSE)
+# 
+# cust50 <- sm_cust50 %>% distinct(customer_id)
 
-cust50 <- sm_cust50 %>% distinct(customer_id)
-
-VIC <- sm_cust50 %>%
-  filter(customer_id == cust50$customer_id[2])
+VIC <- gravitas::smart_meter10 %>%
+  filter(customer_id == "10006486")
 
 scene1 <- VIC %>%
   prob_plot("quarter_year", "wknd_wday",
@@ -85,8 +85,6 @@ scene2 <- VIC %>%
   ylab("") +
   xlab("quarters of the year") +
   scale_fill_brewer(palette = "Paired") + theme(legend.position = "right") + ggtitle("") + scale_y_log10() + theme_minimal()
-
-
 
 
 scene3 <- VIC %>%

### 2020-08-27 data-raw/cricketdata.R
Removed redundant data files and cleaned up script

@@ -1,60 +1,0 @@
-library(readr)
-library(dplyr)
-
-cricket <- read_csv("data-raw/deliveries_all.csv")
-matches <- read_csv("data-raw/matches_all.csv")
-
-cricket_season <- cricket %>%
-  left_join(matches, by = c("match_id" = "id"))
-
-# each over needs to have 6 balls
-
-cricket_fltr_ball <- cricket_season %>%
-  filter(wide_runs + noball_runs == 0) %>%
-  # filtering out scores that are gained by no balls or wide
-  # still have to filter out overs for which number of balls per over less than 6
-  group_by(batting_team, match_id, inning, over) %>%
-  summarise(n_over = length(ball)) %>%
-  filter(n_over == 6) %>%
-  mutate(key = paste(batting_team, match_id, inning, over, sep = "_"))
-
-
-cricket_over_crctd <- cricket_season %>%
-  mutate(cricket_key = paste(batting_team, match_id, inning, over, sep = "_")) %>%
-  filter(cricket_key %in% cricket_fltr_ball$key) %>%
-  filter(wide_runs + noball_runs == 0) %>%
-  mutate(ball_per_over = purrr::rep_along(match_id, 1:6))
-
-# each innings need to have 20 overs
-
-cricket_fltr_over <- cricket_over_crctd %>%
-  filter(inning %in% c(1, 2)) %>%
-  group_by(inning, match_id) %>%
-  summarise(n = length(over)) %>%
-  filter(n == 120) %>%
-  mutate(key = paste(match_id, inning, sep = "_"))
-
-cricket_over_inning_crctd <- cricket_over_crctd %>%
-  mutate(over_key = paste(match_id, inning, sep = "_")) %>%
-  filter(over_key %in% cricket_fltr_over$key) %>%
-  filter(wide_runs + noball_runs == 0) %>%
-  mutate(ball_per_over = purrr::rep_along(match_id, 1:6)) %>%
-  filter(inning %in% c(1, 2))
-
-# each match needs to have 2 innings
-
-cricket_filtr_inning <- cricket_over_inning_crctd %>%
-  group_by(match_id) %>%
-  summarise(length(inning)) %>%
-  filter(`length(inning)` == 240)
-
-cricket_all_seas <- cricket_over_inning_crctd %>%
-  filter(match_id %in% cricket_filtr_inning$match_id) %>%
-  filter(wide_runs + noball_runs == 0) %>%
[…]

### 2020-08-27 scripts/main.R

@@ -1,26 +1,24 @@
 ## ---- load
-library(bookdown)
-library(rticles)
 library(knitr)
-library(ggplot2)
+library(readr)
 library(dplyr)
+library(ggplot2)
 library(tsibble)
-if (requireNamespace("gravitas") == F)
+if (requireNamespace("gravitas") == FALSE) {
   remotes::install_github("Sayani07/gravitas")
+}
 library(gravitas)
 library(ggpubr)
 library(kableExtra)
-library(magrittr)
-library(readr)
 
 ## ---- linear-time
-knitr::include_graphics("Figs/linear-ex.png")
+include_graphics("Figs/linear-ex.png")
 
 ## ---- circular-dow
-knitr::include_graphics("Figs/circular-ex.png")
+include_graphics("Figs/circular-ex.png")
 
 ## ----aperiodic-example
-knitr::include_graphics("Figs/aperiodic-ex3.png")
+include_graphics("Figs/aperiodic-ex3.png")
 
 ## ----tab-mayan
 table_mayan <- tibble(
@@ -28,12 +26,13 @@ table_mayan <- tibble(
   `single-order-up cyclic (C)` = c("kin-of-uinal", "uinal-of-tun", " tun-of-katun", "katun-of-baktun", 1),
   `period length/conversion operator (K)` = c(20, 18, 20, 20, 1)
 )
-kable(table_mayan,
-  format = "latex",
-  booktabs = TRUE,
-  align = "l",
-  caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities."
-) %>%
+table_mayan %>%
+  kable(
+    format = "latex",
+    booktabs = TRUE,
+    align = "l",
+    caption = "Hierarchy table for Mayan calendar with circular single-order-up granularities."
+  ) %>%
   kable_styling()
 
 ## ----tab-gregorian
@@ -43,11 +42,12 @@ table_greg <- tibble(
   `period length/conversion operator (K)` = c(60, 24, "k(day, month)", 12, 1)
 )
 
-kable(table_greg,
-  format = "latex",
-  booktabs = TRUE,
-  caption = "Hierarchy table for the Gregorian calendar with both circular and quasi-circular single-order-up granularities."
-) %>%
+table_greg %>%
+  kable(
+    format = "latex",
+    booktabs = TRUE,
[…]

### 2020-08-27 main.Rmd
Clean up of Rmd file

@@
 - name: Sayani Gupta
-  affiliation: Department of Econometrics and Business Statistics, Monash University
   thanks: "Email: Sayani.Gupta@monash.edu"
+  affiliation: Department of Econometrics and Business Statistics, Monash University, Australia
 
 - name: Rob J Hyndman
-  affiliation: Department of Econometrics and Business Statistics, Monash University
+  affiliation: Department of Econometrics and Business Statistics, Monash University, Australia
 
 - name: Dianne Cook
-  affiliation: Department of Econometrics and Business Statistics, Monash University
+  affiliation: Department of Econometrics and Business Statistics, Monash University, Australia
 
 - name: Antony Unwin
-  affiliation: University of Augsburg
+  affiliation: University of Augsburg, Germany
 
@@
  Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets.
-This paper describes several classes of time deconstructions using linear and cyclic time granularities.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months with respect to a baseline.
-Cyclic time granularities can be circular such as hour of the day, quasi-circular such as day of the month, and aperiodic such as public holidays.
-The hierarchical structure of granularities creates a nested ordering.
-Hour of the day and second of the minute are single-order-up.
-Hour of the week is multiple-order-up, because it passes over day of the week.
+This paper describes classes of time deconstructions using linear and cyclic time granularities.
+Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
+Cyclic time granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
+The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up.
+Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week.
 Methods are provided for creating all possible granularities for a time index.
-A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together, called a harmony or when they cannot, called a clash.
-
- The time granularities can be used to create visualizations of the data to explore for periodicities, associations and anomalies.
-The granularities can be considered to be categorical variables (ordered or unordered) which induces a grouping of the observations.
+A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+
+ Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies.
+The granularities form categorical variables (ordered or unordered) which induce groupings of the observations.
 Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
-A recommendation of appropriate distribution display is provided.
-
- The methods are implemented in the open source R package `gravitas`.
-The functions for creating granularities and exploring the associated time series are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
+A recommendation of appropriate distribution displays is provided.
+
+ The methods are implemented in the open source R package `gravitas`, providing functions for creating granularities and exploring the associated time series which are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
 bibliography: [bibliography.bib]
 preamble: >
-  \setlength {\marginparwidth }{2cm}
-  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
+  \usepackage{mathtools,amssymb,booktabs,amsthm}
   \def\mod{~\text{mod}~}
+  \newtheorem{definition}{Definition}
+  \newtheorem*{example}{Example}
+  \usepackage{mathptmx}
+  \usepackage{caption}
+  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
+  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
+  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
 output:
@@
 opts_chunk$set(
-  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
-  fig.path = "figure/", fig.align = "center", fig.show = "hold",
-  cache = TRUE, cache.path = "cache/",
+  echo = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  comment = "#>",
+  fig.path = "figure/",
+  fig.align = "center",
+  fig.show = "hold",
+  cache = TRUE,
+  cache.path = "cache/",
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
 )
-# knitr::opts_knit$set(root.dir = here::here())
 ```
@@
 ```{r load}
-
 ```
@@
 # Introduction
-
-<!--temporal granularities and why should we care -->
-
-<!-- problem of analyzing data with such finer scales and classical ways to deal with it -->
 
@@
 
-<!-- # existing ways and challenges -->
-
 Calendar-based graphics [@wang2020calendar] are useful in visualizing patterns in the weekly and monthly structure, and are helpful when checking for the effects of weekends or special days.
@@
 However, most of its accessor functions are limited to single-order-up granularities.
-
-<!--motivation -->
 
@@
 
-<!-- at finer or coarser scales can be benefited from the approach of Exploratory Data Analysis (EDA).
-EDA calls for utilizing visualization and transformation to explore data systematically.
-It is a process of generating hypothesis, testing them and consequently refining them through investigations.
--->
-
-<!--existing closely related work and how ours relate -->
- <!-- This paper utilizes the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones. -->
-
-<!--actions on multiple order up - usage -->
-
-<!-- Finally, visualizing data across single/multiple-order-up granularities help us to understand periodicities, pattern and anomalies in the data.
-Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach.
-However, this approach can lead to a myriad of choices all of which are not useful.
-Analysts are expected to alliteratively visualize these choices for exploring possible patterns in the data.
-But too many choices might leave him bewildered.
--->
-
-<!-- challenges and how you are dealing wih it -->
 This work provides tools for systematically exploring bivariate granularities within the tidy workflow. In particular, we
@@
   * recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pair of granularities.
-
-<!-- Pairs of granularities can be categorized as either a *harmony* or *clash*, where = are pairs of granularities that aid exploratory data analysis, and clashes are pairs that are incompatible with each other for exploratory analysis.
--->
 
@@
 
-<!-- general calendar categorization -->
-<!-- Often we partition time into months, weeks or days to relate it to data.
-Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization].
-Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc which are useful to represent different periodicities in the data.
-Classes of time abstractions may be based on arrangement or hierarchical order of time, discussed in further details in \autoref{sec:arrangement} and \autoref{sec:order}.
--->
-
 Discrete abstraction of time such as weeks, months or holidays can be thought of as "time granularities".
@@
 
-<!-- Arrangement: Linear vs. Circular vs. aperiodic {#sec:arrangement} -->
-
-<!-- The arrangement of the time domain can result in deconstructing time in linear, circular or aperiodic (nearly circular) ways.  -->
-
-<!-- There have been several attempts to provide a framework for formally characterizing time granularities.
-One of the first attempts was provided by @Bettini1998-ed who proposed following definitions:-->
-
-<!-- <!-- Representation with tsibble -->
-<!-- Suppose we have a tsibble with a time index in one column and keys and variables in other columns.
-A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index.
-A linear granularity is a mapping of row numbers to subsets of the time domain.
-For example, if the time index is days, then a linear granularity might be weeks, months or years.
--->
-
-<!-- <!-- What is calendar categorization -->
-<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity.
-Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, working day/non-working day, etc.
-Many of these are circular, such as day-of-week, time-of-day.
-Some are nearly circular such as day-of-month.
-Some are irregular such as working day/non-working day.
-Let's call all of these "calendar categorizations".
-Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.
--->
-
-<!-- What is circular granularity and some examples -->
-<!-- We specify the circular categorizations using modular arithmetic and call these **circular granularities**.
-The number of categories is essentially the periodicity of a circular time granularity.
-For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$.
-Then the following categorizations can be computed.
--->
-
-<!-- In the following sections, we provide a formalism to all of these abstractions and identify their relationships and symbolic representations.  -->
-
-<!-- ### Linear {#sec:linear-gran-def} -->
-
-\newtheorem{definition}{Definition}
-
 \begin{definition}\label{def:definition}
-A {\bf time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
-\end{definition}
-
-The time domain is assumed to be *discrete*, and there is unique predecessor and successor for every element in the time domain except for the first and last.
+A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+\end{definition}
+
+\noindent The time domain is assumed to be *discrete*, and there is unique predecessor and successor for every element in the time domain except for the first and last.
 
 \begin{definition}\label{def:index set}
-There is a unique {\bf index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of non-negative integers. Often, this is thought of as $t=0, \dots, T$.
+There is a unique \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of non-negative integers. Often, this is thought of as $t=0, \dots, T$.
 \end{definition}
@@
 \begin{definition}\label{def:linear}
-A {\bf linear granularity} is a mapping $G$ from the integers (the index set, $Z$) to subsets of the time domain such that:
-  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
-than all elements of $G(j)$; and
+A \textbf{linear granularity} is a mapping $G$ from the integers (the index set, $Z$) to subsets of the time domain such that:
+  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less than all elements of $G(j)$; and
   (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
-Each non-empty subset $G(i)$ is called a {\bf granule}.
+Each non-empty subset $G(i)$ is called a \textbf{granule}.
 \end{definition}
@@
 The indexing for each granule can also be associated with textual representation, called the label.
-A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.<!--The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labelled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labelled granularities, while others can be implemented using more flexible labelings.--> \autoref{fig:linear-time} illustrates the linear time granularities.
+A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
+\autoref{fig:linear-time} illustrates the linear time granularities.
 Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities which maps from index set to subsets of the hourly time domain.
-<!--For example, day($0$) =  {hour($0$), hour($1$), hour($2$), \dots, hour($23$)}, day($1$) =  {hour($24$), hour($25$), hour($26$), \dots, hour($47$)} and so on.--> If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
+If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
 These linear granularities are ordered with ordering guided by the index set which is a set of integers.
@@
 
- <!--, each of which are non-repeating in nature and arranged from past to future. If we consider $T$ hours, we will have $\lfloor T/24\rfloor$ days, $\lfloor T/(24*7)\rfloor$ weeks and so on.-->
-
-```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="(ref:linear-time)"}
+```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="(ref:linear-time)", fig.pos="!htb"}
 ```
@@
 
-<!--@Bettini1998-ed discusses the relationships between linear time granularities, periodicities in time, and the structure of a calendar.-->
-
 Properties of pairs of granularities fall into various categories.
@@
 \begin{definition}\label{def:finerthan}
-A linear granularity $G$ is {\bf finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
+A linear granularity $G$ is \textbf{finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
 $G(i) \subset H(j).$
@@
 \begin{definition}\label{def:groupsinto}
-A linear granularity $G$ {\bf groups into} a linear granularity $H$, denoted
+A linear granularity $G$ \textbf{groups into} a linear granularity $H$, denoted
 $G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
@@
 
-<!-- CONTENT CUT -->
-<!-- These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true.
-However, weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or is finer than public holidays.
-Consider another example where day $\trianglelefteq$ month.
-This relationship is not fully described until we associate it with periodicity.
-Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored.
-The period becomes 400 years with the inclusion of leap years and all their exceptions.
--->
-\newtheorem*{example}{Example}
-
-\begin{example}
-{\rm Both $day \trianglelefteq week$ and $day \preceq week$ holds, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
+\noindent For example, both $day \trianglelefteq week$ and $day \preceq week$ hold, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
 Consider another example where $day \trianglelefteq month$.
@@
 Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
-This grouping period becomes 4 and 400 years with the inclusion of leap years and leap seconds respectively.}
-\end{example}
-
-<!-- CONTENT CUT -->
-<!-- # ```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships.
-Weekends are finer than weeks but do not group into weeks.
-And weekends neither groups into or finer than public holidays"} -->
-<!-- knitr::include_graphics("Figs/groupsandfiner.png") -->
-<!-- ``` -->
+This grouping period becomes 4 and 400 years with the inclusion of leap years and leap seconds respectively.
 
 \begin{definition}\label{def:periodical}
-A granularity $G$ is {\bf periodical} with respect to a granularity $H$ if:
+A granularity $G$ is \textbf{periodical} with respect to a granularity $H$ if:
 (1) $G \trianglelefteq H$; and
@@
 \end{definition}
-
-<!-- CONTENT CUT -->
-<!-- A granularity $H$ which is periodical with respect to $G$ is specified by: -->
-<!-- (i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period; -->
-<!-- (ii) the value of $P$; -->
-<!-- (iii) the indexes of first and last granules in $H$, if their value is not infinite. -->
 
@@
 
-<!-- CONTENT CUT -->
-<!-- Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$. -->
-
-<!-- # ```{r group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$."} -->
-<!-- # knitr::include_graphics("Figs/groups_periodically.png") -->
-<!-- # ``` -->
-
-<!-- CONTENT CUT -->
-<!-- \begin{definition}[Groups quasi-periodically]\label{def:quasi} -->
-<!-- A granularity $H$ is quasi-periodical with respect to a granularity $G$ if: -->
-<!-- (1) $G \trianglelefteq H$; and -->
-<!-- (2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
--->
-<!-- \end{definition} -->
-
-<!-- # ```{r quasi-group-periodical, echo=FALSE, out.width = "100%", fig.cap="Linear granularity $B$ groups quasi-periodically into linear granularity $G$ with $P = 5$ and $R = 2$, alternating between combining 2 and 3 granules of $B$ with $E1$ as the exception."} -->
-<!-- # knitr::include_graphics("Figs/quasi_groups_periodically.png") -->
-<!-- # ``` -->
-
-<!-- **Example**:  Example showing periodic and quasi-periodic linear granularities are displayed in \autoref{fig:group-periodical} and \autoref{fig:quasi-group-periodical} respectively.
-Each granule is represented by a box.
-In these figures, the granules that overlap share elements from the underlying time domain.
-In both of these cases, each granule of $G$ is made up of two and three granules of $B$ alternately.
-So the period is $P = 5$ and number of granules of $G$ in each period is given by $R=2$.
-The only difference in \autoref{fig:quasi-group-periodical} is that there is an exception after the first granule of $G$ which consists of one granule of $B$.
-Except for this exception $E1$, $B$ groups periodically into $G$.
--->
-
-\begin{example}
-{\rm In a Gregorian calendar without leap years we could say $day$ groups periodically into month with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
-In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.}
-\end{example}
-
-<!-- CONTENT CUT -->
-<!-- \begin{definition}[Bottom granularity]\label{def:B} -->
-<!-- Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
--->
-<!-- \end{definition} -->
-
-<!-- Here, g-rel could be any relationship defined between granularities as defined in @Bettini1998-ed.
-Given the set of all granularities defined over the time domain $(\mathbb{Z}; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.
--->
+For example, in a Gregorian calendar without leap years we could say $day$ groups periodically into month with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
+In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
 
 \begin{definition}\label{def:order}
-The {\bf order} of a linear granularity is the level of coarseness associated with a linear granularity.
+The \textbf{order} of a linear granularity is the level of coarseness associated with a linear granularity.
 A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
@@
 
-\begin{example}
-{\rm For two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ then $G$ is of lower order than $H$.
-Moreover, if we consider bottom granularity as day, linear granularity week will have lower order than month since each week consist of less number of days than each month.
-Here neither  week {\em groups into} month or  week {\em finer than} month, but their relative order could be determined.}
-\end{example}
-
-<!-- that the aperiodic linear granularities cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.  -->
-
-<!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn).
-Chronons may be grouped into granules of a bottom granularity and -->
+With two linear granularities $G$ and $H$, if $G$ *groups into* or *finer than* $H$ then $G$ is of lower order than $H$.
+For example, if the bottom granularity is day, then granularity week will have lower order than month since each week consist of fewer days than each month.
+Here neither week *groups into* month or week *finer than* month, but their relative order can be determined.
 
@@
 
-<!-- CONTENT CUT -->
-<!-- For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$.
-For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$.
-The _altering-tick operation_ generates new full-integer labelled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$.
-For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days.
-We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
-For more variations of calendar algebra operations, see @Ning2002-tf.
-
-<!-- CONTENT CUT -->
-<!-- The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations.
-The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity.
-For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes".
--->
-<!-- The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity.
-For example, one can define the granularity "Monday" and hence select only Mondays from linear granularity "day".
--->
-
-<!-- \begin{definition}[Labelled granularity]\label{def:labelgran} -->
-
-<!-- CONTENT CUT -->
-<!-- The notion of a labelled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra.
-A labelled granularity is a pair $(L, G)$, where the label set $L$ is a subset of the integers instead of the set of integers and -->
-<!-- follows the conditions (1) and (2) in \ref{def:linear} where $i$, $j$, and $k$ are in $L$.
-When the label set $L$ is exactly the integers, the granularities are called "full-integer" labelled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->
-
-<!-- $G$ is a mapping from $L$ to the subsets of the time domain such that: -->
-
-<!--   1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and -->
-<!--   2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty. -->
-
 \begin{definition}\label{def:grouping_operation}
 Let $G_1$ be a full-integer labelled granularity, and $m$ a positive integer.
-{\bf The grouping operation} Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
+\textbf{The grouping operation} Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
 More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labelled granularity such that for each integer $i$,
@@
 
-<!-- Alternatively, given any integer $G_1(j)$ and a positive integer $m$, if there exists unique integers $G_2(i)$ and $R$ such that $G1(j) = m * G2(i) + R$ where $0 ≤ R < m$, then $G_2$ is a granularity that is formed -->
-
-<!-- CONTENT CUT altering-tick -->
-<!-- \begin{definition}\label{altering-tick} -->
-<!-- Let $G_1$, $G_2$ be full-integer labelled granularities, and $l$, $k$, $m$ integers, with $G_1 \trianglelefteq G_2$, $G_1\preceq G_2$, and $1 \leq l \leq m$.
-{\bf The altering-tick operation} $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
--->
-<!-- \end{definition} -->
-
-<!-- \begin{example} -->
-<!-- The altering-tick operation can be used to compute the granularity $pseudomonth$, for the uneven lengths of $month$, by grouping 31 granules of $day$, followed by shrinking April, June, September and November by 1 granule of $day$, and shrinking February by 3 granules of $day$ for non-leap years.
--->
-<!-- \end{example} -->
-
 \noindent For more variations of calendar algebra operations, see @Ning_2002.
 
-<!-- CONTENT CUT -->
-<!-- \begin{align*} -->
-<!-- \text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \text{Alter}^{12}_{9,-1}(\text{day}, \\ -->
-<!--   &~~\text{Alter}^{12}_{6,-1}(\text{day}, \text{Alter}^{12}_{4,-1)}(\text{day}, \\ -->
-<!--   &~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \text{Group}_{31}(\text{day})))))), -->
-<!-- \end{align*} -->
-
 # Cyclic time granularities {#sec:cyclic-gran}
 
-<!-- sub: regular and irregular mapping leads to circular and quasi-circular granularities -->
-
 Cyclic granularities represent cyclical repetitions in time.
-They can be thought of as additional categorizations of time that are not linear.<!--Weeks is an example of a cyclical granularity because every 7 days there's another Monday.--> Cyclic granularities can be constructed by operating on two linear granularities.
+They can be thought of as additional categorizations of time that are not linear.
+Cyclic granularities can be constructed by operating on two linear granularities.
 Cycles can be either _regular_, called **circular**, or _irregular_, **quasi-circular** when these two linear granularities relate periodically.
-<!-- with every granule of coarser linear granularity being made up of same (regular) or different (irregular) number of granules of finer linear granularity.
-For example, a regular mapping exists between  (where 60 minutes always add up to 1 hour), while an irregular mapping exists between  (as the number of days per month ranges from 28 to 31).
-A grouping operation in Calendar algebra could lead to a regular mapping, whereas a altering-tick operation could result in an irregular mapping.
-Cyclic time granularities are referred to as **circular** if mappings are regular and **quasi-circular** if the mappings are irregular.
-Examples of common circular granularities include hour of the day and day of the week, whereas examples for quasi-circular granularities are day of the month and week of the month.-->
-
-<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->
 
@@
 \begin{definition}\label{def:circular}
-A {\bf circular granularity} $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
+A \textbf{circular granularity} $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
 \begin{equation} \label{eq:circular-gran}
@@
 
-<!-- The label can be mapped to index for input and the index can later be mapped to this label for output.
-When the label set is exactly the integers, the granularities are called "full-integer" labelled granularity, such as in Definition \ref{def:linear}.
-Some calendar operations will only work on full-integer labelled granularities, while others can be implemented using more flexible labelings.
--->
-
-\begin{example}
-{\rm \autoref{fig:circular-dow} illustrates the linear and the corresponding cyclical granularities.
+For example, \autoref{fig:circular-dow} illustrates the linear and the corresponding cyclical granularities.
 Cyclical granularities can be considered to be cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
@@
 These overlapping cyclical granularities share elements from the linear granularity.
-Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns  $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
+Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns $\{0, 1, \dots, 6\}$ and $\{0, 1, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
 Each circular granularity can use descriptive label mappings.
-Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
-$$
-  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sun}, \text{Mon}, \dots, \text{Sat}\}
-$$
-or
-$$
-  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}
-$$
-for example.}
-\end{example}
-
-<!-- # ```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "} -->
-<!-- # knitr::include_graphics("Figs/circular-gran-diagram.png") -->
-<!-- # ``` -->
-
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap= "(ref:circular-dow)"}
-
+Suppose ${L}$ is a label mapping that defines an unique label for each index $\ell \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined (for example) as
+\begin{align*}
+  L: \{0,1, \dots, 6\} & \longmapsto\ \{\text{Sun}, \text{Mon}, \dots, \text{Sat}\}\\
+\text{or}\qquad
+  L: \{0,1, \dots, 6\} & \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}.
+\end{align*}
+
+```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap= "(ref:circular-dow)",fig.pos="!htb"}
 ```
@@
 
-<!-- All circular granularities can be expressed in terms of the bottom linear granularity with equation \autoref{eq:eq2}. -->
-
-In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H)$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping and $P(G,H)$ is the period of the grouping $(G, H)$  <!--such that the number of granules of $G$ in each granule of $H$ is-->.
-Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with period $P_i$ <!--being the number of granules of the finer granularity in each granule of the coarser granularity.The representation of the resultant circular granularity will be a function of the index set.--> and minutes as the bottom granularity.
+In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H)$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping and $P(G,H)$ is the period of the grouping $(G, H)$.
+Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with period $P_i$ and minutes as the bottom granularity.
 
@@
 \begin{tabular}{lll}
-\hline
+\toprule
 circular granularity & expression & period \\
-\hline
+\midrule
 minute-of-hour                               &
@@
 
-<!-- ```{r tab-circular} -->
-<!--  table_mayan <-  -->
-<!--   tibble(C = c("minute-of-hour", -->
-<!--                                     "minute-of-day",  -->
-<!--                                     "hour-of-day", -->
-<!--                                     "hour-of-week", -->
-<!--                                     "day-of-week"), -->
-<!--                        Expression = c("\$L\_1(z \mod 60)\$", -->
-<!--                                         "\$L\_2(z \mod 60*24)\$", -->
-<!--                                         "\$L\_3(\lfloor z/60\rfloor\mod 24)\$", -->
-<!--                                         "\$L\_4(\lfloor z/60\rfloor\mod 24*7)\$", -->
-<!--                                         "\$L\_5(\lfloor z/24*60\rfloor \mod 7)\$")) -->
-
-<!--  kable(table_mayan, -->
-<!--        format = "latex", -->
-<!--        caption = "Hierarchy table for Mayan calendar.
-$G$ and $C$ represent the linear and single order-up cyclic granularities respectively.
-All cyclic granularities are circular with $K$ as the period length.") -->
-<!-- ``` -->
-
 ## Quasi-circular {#sec:quasi-circular-gran-def}
@@
 \autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) with ($P_i$) denoting the plausible choices of period of the grouping of two linear granularities.
-<!-- number of granules of the finer granularity inside each granule of the coarser one. -->
 
@@
 \begin{tabular}{lr@{~}lr@{~}r}
-\hline
+\toprule
 quasi-circular granularity && potential period lengths\\
-\hline
+\midrule
 $Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\
@@
 
-<!-- \begin{split} -->
-<!-- A_{B, M}(z) & = -->
-<!-- \begin{cases} -->
-<!-- i, & \text{for}\quad z \in T_{i_j} \\ -->
-<!-- 0  & \text{otherwise}  -->
-<!-- \end{cases} -->
-<!-- \end{split} -->
-
-<!--${L}$ is a label mapping that defines an unique label for each index $l \in \{0,1,\dots, (\lceil P'/R' \rceil-1)\}$,-->
-
-<!-- **Example**:  -->
-<!-- Example showing quasi-circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:quasi-circular-example}.
-Each granule is represented by a box.
--->
-
-\begin{example}
-{\rm Consider $G'$ such that every two consecutive granules of $G'$ are made up of $7$ and $5$ granules of $B$ respectively within each period of the grouping ($B$, $G'$).
+Consider $G'$ such that every two consecutive granules of $G'$ are made up of $7$ and $5$ granules of $B$ respectively within each period of the grouping ($B$, $G'$).
 Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
-Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$.
+Here, $T_0 = \{0, 1, \dots, 6\}$ and $T_1 = \{7, \dots 11\}$.
 Hence using Definition \autoref{def:quasicircular} we will have:
-
-\begin{align*}
-\begin{split}
-Q_{B,H'}(10) & = 10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert ,\quad since \quad 10 \in T_{1}  \\
-  & = 3 \\
-\end{split}
-\end{align*}
-}
-\end{example}
-
-<!--such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.-->
-
-<!-- **Discussion**:  -->
-If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->.
-Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
+$$
+  Q_{B,H'}(10)  = 10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert =3
+$$
+since $10 \in T_{1}$.
+
+If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical}) Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
 The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
@@
 Both $Q(B, G)$ and $Q(B, H')$ have repeating patterns unlike $G$ or $H'$ and the patterns are guided by the distribution of $B$ within $G$ and $H'$ and remains constant within a period shown by the hashed arrows.
-Panel b depicts the intuition through a measuring tape analogy where the folds are not of same length but has same length over a period."}
-
-# knitr::include_graphics("Figs/quasi-circular-ex2.png")
-```
-
-<!-- Circular granularities illustrated through linear (a and b) and stack (c and d) display of time.
-The linear display is similar to a measuring tape with the measurement marking distributed as linear granularities day, week and fortnight.
-The stack display gives the intuition of circular granularities by cutting the measuring tape and stacking on top of one another to represent repetitive pattern of categories.
-The granules that overlap represent same time domain in the linear display and same category in the stack display.
-Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time.
--->
-
-<!-- If we ignore leap years and all its exception, the periodicity of the grouping (days, months), P' = 365, R' = 12, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year.
-Considering leap years and all their exceptions, period is 400 years and hence $P' = (3*365 + 366)$, $R' = 12*400$.
-\autoref{eq:eq3} can be useful to find the aperiodic granularity day-of-month in both of these cases, since the grouping is periodic is both cases.
--->
-
-<!-- Then $f(x, y)$ can be computed using modular arithmetic as follows: -->
-
-<!-- $$f(x, y) = \lfloor z/c(z,x) \rfloor\mod c(x,y)$$ -->
-
-<!-- The constant c(x,y) is for granularities for groups into periodically, for granualrities which group into each other in a quasi-periodic way, no such constants exist and no such close form solution can be obtained.
--->
-
-<!-- It is easy to note here that due to unequal length of some linear granularities like months or years, these formulae can not be used for computing week-of-month or day-of-year.
-Thus, we start with the definition of circular granularity and then move on to computing aperiodic granularities.
--->
-
-<!-- Formal definition of circular granularity -->
-
-<!-- **Definition: Equivalence class** Let $m \in N \backslash{0}$.
-For any $a \in Z$ (set of integers), $[a]$ is defined as the equivalence class to which a belongs if $[a]$ = {$b \in Z | a \equiv (b \mod m)$}.
--->
-
-<!-- The set of all equivalence classes of the integers for a modulus $m$ is called the ring of integers modulo $m$, denoted by $Z_m$.
-Thus $Z_m = \{[0], [1], ..., [m-1]\}$.
-However, we often write $Z_m = \{0, 1, ..., (m-1)\}$, which is the set of integers modulo $m$.
--->
-
-<!-- **Definition:** A **circular granularity** $C$, joining two linear granularity with a modular period m is defined to be a mapping from the integers $Z$ (Index Set) to $Z_m$, such that -->
-
-<!-- $C(s) = (s\mod m)$ for $s \in Z$. -->
-
-<!-- For example, suppose $C$ is a circular granularity denoting hour-of-day and we have hourly data for 100 hours.
-The modular period $m = 24$, since each day consists of 24 hours and $C$ is a mapping from ${1,2,\dots, 100}$ to ${0,1,2, \dots, 23}$ such that  $C(s)= s \mod 24$ for $s \in {1,2,\dots, 100}$.
--->
-
-<!-- **Definition:** A **cycle** is defined as the progression of each circular granularity with modular period m through {$1,2,\dots,(m-1),0$} once. -->
-
-<!-- **Definition:** A **circular granule** represents an equivalence class inside each cycle. -->
-
-<!-- ### Aperiodic {#aperiodic-gran-def} -->
-<!-- Formal definition of aperiodic circular granularity -->
-
- <!-- Why we need single order up and multiple order up granularities -->
-<!-- A time series which is characterized by being composed of repeating cycles. -->
-
-<!-- Identifying repeating (periodic/aperiodic) patterns are necessary in revealing patterns and future trends of a temporal data.
-Often there is a need for periodicity detection to find whether and how frequent a periodic/aperiodic pattern is repeated within the series.
--->
-
-<!-- The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity.
-The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.
--->
-
-<!-- CONTENT CUT -->
-<!-- It can be noted here that if a linear granularity $G'$ is quasi-periodic (Bettini2000-vy) <!--(Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$.<!--(by Definition \autoref{def:quasi}).-->
-
-<!-- \begin{equation}\label{eq6} -->
-<!-- A_{B, G'}(z) = L(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert - \sum_{u=0}^{e_k}\vert E_{u} \vert) -->
-<!-- \end{equation} -->
-
-<!-- for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$. -->
+Panel b depicts the intuition through a measuring tape analogy where the folds are not of same length but has same length over a period.", fig.pos="!htb"}
+```
 
@@
 
-<!--Periodic and Quasi-periodic behavior can be defined by a repeating pattern except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.--> Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
+Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
 Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant.
-A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.<!--U.S labor day is the first Monday in September and U.S.
-Memorial Day is the last day in May.
---> In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
-Examples of aperiodic granularity may also include school holidays or a scheduled event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns.
+A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
+In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
+Examples of aperiodic granularity may also include school holidays or a scheduled event.
+All of these are recurring events, but with non-periodic patterns.
 As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
 
-\noindent Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ such that Definition \ref{def:periodical} does not hold true with respect to $B$.
-<!--that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity.--> However, $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$.
+Aperiodic cyclic granularities are defined using aperiodic linear granularities.
+Consider $n$ aperiodic linear granularities $M_i$,  $i \in \{1, 2, \dots, n\}$, such that Definition \ref{def:periodical} does not hold true with respect to $B$.
+However, $B \trianglelefteq M_i$.
 Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$.
@@
 
-```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap= "(ref:aperiodic-example)"}
-
-```
+```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap= "(ref:aperiodic-example)", fig.pos="!htb"}
+```
+
 (ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
@@
 \begin{definition}\label{def:aperiodic}
-An {\bf aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if
+An \textbf{aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if
 \begin{equation}\label{eq:aperiodic}
@@
 
-<!-- \forall i \in {1, 2, \dots, n}\\ -->
-
 \begin{example}
@@
 
-<!-- In Section \ref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed.
-In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities --><!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated.
-Since, our method caters to computation of cyclic granularities, we shall refer to it as--> <!-- through cyclic calendar algebra operations.-->
-
 The hierarchical structure of time creates a natural nested ordering which can be used in the computation of relative pairs of granularities.
@@
 \begin{definition}\label{def:hierarchy}
-The nested ordering of linear granularities can be organized into a {\bf hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order.
+The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order.
 It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$.
 can be defined relative to the linear granularities.
-Let $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$.
-Then $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$, if $C_{G_{l}, G_{m}}$ is a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$, if $C_{G_{l}, G_{m}}$ is quasi-circular.
-\end{definition}
-
-\begin{example}
-{\rm  \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar.
+Let $G_{\ell}$ and $G_{m}$ represent the linear granularity of order $\ell$ and $m$ respectively with $\ell<m$.
+Then $K \equiv P(\ell,m)$ represents the period length of the grouping $(G_{\ell}, G_{m})$, if $C_{G_{\ell}, G_{m}}$ is a circular granularity and $K \equiv k(\ell,m)$ represents the operation to obtain $G_{m}$ from $G_{\ell}$, if $C_{G_{\ell}, G_{m}}$ is quasi-circular.
+\end{definition}
+
+For example, \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar.
 In the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun.
-}
-\end{example}
 
@@
 
- <!--So far we have used examples of linear granularities from Gregorian calendar as it is the most widely used calendar. However, it is far from being the only one.-->
-\noindent It is interesting to note, that in the Mayan calendar, like most calendars, that day is the basic unit of time [@Reingold2001-kf].
+It is interesting to note, that in the Mayan calendar, like most calendars, that day is the basic unit of time [@Reingold2001-kf].
 The structuring of larger units, weeks, months, years and cycle of years, though, varies substantially.
@@
 Note that it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
-Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.
+Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(\ell),G(m)}$ is either circular or quasi-circular.
 
@@
 
- <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->
-
 \begin{example}
@@
 \end{example}
-
-<!-- \begin{definition}\label{def:norderup}  -->
-
-<!-- \end{definition} -->
 
@@
 
-<!-- Thus the first linear granularity can either be the bottom granularity (finest possible) or the finest granularity within the set of linear granularities and the other granularities are subsequently more "coarse" than the earlier one.
--->
-
-<!-- If linear granularities are not provided or bottom granularity is not known, any circular granularities of hierarchy order m can be computed, provided the circular granularities till hierarchy order (m-1) are provided.
--->
-
-<!-- some discussions on primary units of calendar -->
-<!-- **Example:**  -->
-
 ### Single-to-multiple {#sec:single-to-multiple}
 
-<!-- #### All circular single order-up granularities {#sec:all-circular}   -->
-
-<!-- Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity. -->
-
-\noindent Methods to obtain multiple-order-up granularity will be different based on if the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
+Methods to obtain multiple-order-up granularity will be different based on if the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
 The methods require the use of integer arithmetic, and hence it is important that the label mapping of the individual single-order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
 The label mapping of the resultant multiple-order-up granularity can be chosen depending on the context.
-Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using Equation \ref{eq:eq7} for $l < m - 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-l-1\}$.
-
+Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using Equation \ref{eq:eq7} for $\ell < m - 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-\ell-1\}$.
 \begin{equation} \label{eq:eq7}
 \begin{split}
-C_{G_l,G_m}(z) & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_m}(z)\\
-  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)( C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)C_{G_{l+2}, G_m}(z)) \\
-  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_{l+2}}(z) + P(l, l+1)P(l+1, l+2)C_{G_{l+2}, G_{m}}\\
-  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_{l+2}}(z) + P(l, l+2)C_{G_{l+2},G_{l+m}}(z)\\
+C_{G_\ell,G_m}(z) & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)C_{G_{\ell+1},G_m}(z)\\
+  & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)( C_{G_{\ell+1},G_{\ell+2}}(z) + P(\ell+1, \ell+2)C_{G_{\ell+2}, G_m}(z)) \\
+  & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)C_{G_{\ell+1},G_{\ell+2}}(z) + P(\ell, \ell+1)P(\ell+1, \ell+2)C_{G_{\ell+2}, G_{m}}\\
+  & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)C_{G_{\ell+1},G_{\ell+2}}(z) + P(\ell, \ell+2)C_{G_{\ell+2},G_{\ell+m}}(z)\\
   &\vdots\\
-  & = \sum_{i=0}^{m - l - 1} P(l, l+i)C_{G_{l+i},G_{l+i+1}}(z)\\
+  & = \sum_{i=0}^{m - \ell - 1} P(\ell, \ell+i)C_{G_{\ell+i},G_{\ell+i+1}}(z)\\
 \end{split}
 \end{equation}
-
-\noindent For example, the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar <!--with the index set  $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ mapping "kins" to a set of positive integers--> could be obtained using Equation \ref{eq:eq8}.
-
+For example, the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar could be obtained using Equation \ref{eq:eq8}.
 \begin{equation} \label{eq:eq8}
@@
 \end{equation}
-
-<!-- #### Circular or quasi-circular single order-up granularities {#sec:notall-circular}   -->
-
-<!-- CONTENT CUT -->
-<!-- There can be three scenarios for obtaining multiple-order-granularities here:   -->
-<!--  - cyclic granularities relating two linear granularities whose orders are less than day   -->
-<!--  - cyclic granularities relating two linear granularities whose orders are more than month   -->
-<!--  - granularities relating two linear granularities with order at most day and another with order at least month -->
-
-<!-- The multiple-order-up cyclic granularites resulting from the first two cases are circular and has been handled in \autoref{sec:all-circular}.
-Any cyclic granularity resulting from the last case are quasi-circular.
-Examples might include hour-of-month or day-of-year.
--->
-
-\noindent We consider the case of only one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
-Any multiple-order-up quasi-circular granularity $C_{l, m}(z)$ could then be obtained as a dicrete combination of circular and quasi-circular granularities.
-<!--function of a single (or multiple) order-up circular granularity $C_{l,m'}(z)$ and single order-up quasi-circular granularity $C_{m',m}(z)$ or vice versa.
---> Depending on the order of the combination, again two different approaches need to be employed leading to the following cases:
-
-<!-- For example, with $C_{l, m}(z) = f(C_{l,m'}(z), C_{m',m}(z))$, -->
-- $C_{l,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+We consider the case of only one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
+Any multiple-order-up quasi-circular granularity $C_{\ell, m}(z)$ could then be obtained as a dicrete combination of circular and quasi-circular granularities.
+
+Depending on the order of the combination, again two different approaches need to be employed leading to the following cases:
+
+  * $C_{\ell,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
 \begin{equation} \label{eq:multifromsingle-quasi1}
 \begin{split}
-C_{G_l,G_{m}}(z) & = C_{G_{l},G_{m'}}(z) + P(l, m')C_{G_{m'},G_{m}}(z) \\
+C_{G_\ell,G_{m}}(z) & = C_{G_{\ell},G_{m'}}(z) + P(\ell, m')C_{G_{m'},G_{m}}(z) \\
 \end{split}
@@
 
-- $C_{l,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
+  * $C_{\ell,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
 \begin{equation} \label{eq:multifromsingle-quasi2}
 \begin{split}
-C_{G_l,G_{m}}(z) & = C_{G_{l},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
+C_{G_\ell,G_{m}}(z) & = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
 \end{split}
 \end{equation}
-
-\noindent where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
-
-<!-- Let us consider the computation of $C_{hour, month}(z)$ for a $z$ such that $C_{hour, day}(z) = 12$, $C_{day, month}(z) = 20$ and $C_{month, year}(z) = 9$. -->
-
-\noindent Consider a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
-<!--Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days.--> This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities with day-of-month as the only single-order-up quasi-circular granularity.
-
-\noindent Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we then have:
-
- $$C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)$$
- $$C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert)$$, where $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$
+where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{\ell}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+
+Consider a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
+This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities with day-of-month as the only single-order-up quasi-circular granularity.
+
+Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we then have:
+$$
+  C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)
+$$
+$$
+  C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert),
+$$
+where $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$.
 
@@
 
-<!-- #### Multiple order-up circular granularities {#sec:all-circular-multiple}   -->
-
-\noindent Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
-For a hierarchy table $H_n: (G, C, K)$ with only  circular single-order-up granularities in the hierarchy and $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, multiple-order-up granularities could be obtained using \ref{eq:all-circular-multiple}.
-
+Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
+For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities in the hierarchy and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities could be obtained using \ref{eq:all-circular-multiple}.
 \begin{equation} \label{eq:all-circular-multiple}
-C_{G_{l_1}, G_{m_1}}(z) = \lfloor C_{G_{l_2}, G_{m_2}}(z)/P(l_2,l_1) \rfloor \mod P(l_1, m_1)
+C_{G_{\ell_1}, G_{m_1}}(z) = \lfloor C_{G_{\ell_2}, G_{m_2}}(z)/P(\ell_2,\ell_1) \rfloor \mod P(\ell_1, m_1)
 \end{equation}
-
-\noindent For example, in Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) =  \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.
+For example, in Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) = \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.
 
@@
 
-<!-- \begin{equation} \label{eq:quasi-circular-multiple} -->
-<!-- C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod P(m_1, m_2)) -->
-<!-- \end{equation} -->
-
-<!-- CONTENT CHANGE -->
-<!-- The representation of single order-up quasi-circular granularities using multiple-order-up quasi-circular granularities is not discussed in this paper and is left for future work. -->
-
 Single-order-up quasi-circular granularity could be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}.
@@
 
-<!-- How volumes of continuously collected data can be visualized to ease further decision making?  -->
-
-<!-- # Why data structure and basis for cyclic granularities exploration -->
-
-\noindent Effective exploration and visualization benefits from well-organized data structures.
+Effective exploration and visualization benefits from well-organized data structures.
 @wang2020tsibble introduced a tidy data structure, tsibble, to support exploration and modeling of temporal data.
@@
 A bottom granularity is represented by the index of the tsibble.
-<!-- It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.
-Moreover, one can use calendars based on application domain that define contextual system of linear and cyclic granularities.-->
-
-<!-- #recent data structures for temporal exploration -->
-
-<!--# our proposal of extension of tsibble data structure and -->
-<!--# why looking at pair of granularities instead of all -->
-
-\noindent All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure).
+
+All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure).
 This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.
@@
 
-<!-- CONTENT SHORT -->
-<!-- Cyclic time granularities can be implemented through the tsibble [@wang2019tsibble] (Time Series taBBLE) data structure.
-A tsibble consists of an index, key and measured variables.
-An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
-<!--In a tsibble, each observation (row) is uniquely identified by index and key.A time domain, as defined by @Bettini1998-ed, is essentially a mapping of the index set to the index of a tsibble.
-A linear granularity is a mapping of the index set to subsets of the time domain.
-For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
-A bottom granularity is represented by the index of the tsibble.<!--Consistent with the assumption of @Bettini2000-vy, all linear granularities can be generated from the bottom granularity by calendar algebraic operations.\todo[inline]{Would it be worth to start this discussion about data structures earlier, with the linear granularity, and follow it through when appropriate; or put all discussion on implementation in the data structures section.} -->
-
-<!-- CONTENT SHORT -->
-<!-- across bivariate cyclic granularities.  -->
-<!-- Let us consider two cyclic granularities $C_1$ and $C_j$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
-Data sets of the form <$C_1$, $C_j$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
--->
-
-<!--Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities.-->
-<!-- Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.
--->
-
-<!--  These subsets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities. -->
-
-<!-- This data representation can lead to comprehensive exploration by conditioning the levels taken by one or more of the cyclic granularities. -->
-
-<!-- This can be an useful data structure for exploratory data analysis as we can rapidly compare patterns in different parts of the data by creating same plots using data subsets corresponding to each level of the conditioned variable.
--->
-
-<!-- These are based on conditioning the values taken on by one or more of the variables in a data set.
-This means creating same plots using data subsets corresponding to each level of the conditioned variable.
-This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
--->
-
-```{r data-structure, fig.cap = "(ref:data-structure)", fig.pos="htb"}
+```{r data-structure, fig.cap = "(ref:data-structure)", fig.pos="!htb"}
 ```
@@
 
-<!-- Let us consider two cyclic granularities $C_1$ and $C_j$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\} and $v$ denotes the measurement variable.
-Data sets of the form <$C_1$, $C_j$, $v$> can then form the basis for exploration and analysis of the measured variable across bivariate cyclic granularities.
--->
-
 The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
-Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k | k \in \mathbb{N}, k \leq K\}$ and $C_j$ maps index set to a set $\{B_l | l \in \mathbb{N}, l \leq L\}$.
-Here, $A_k$'s or $B_l$'s are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
-Let $S_{kl}$ be a subset of the index set such that for all $s \in S_{kl}$, $C_i(s) = A_k$ and $C_j(s) = B_l$.
-Data subsets for each combination of levels ($A_k$, $B_l$) like <$A_k$, $B_l$, $v(s)$> can be obtained for all $k \in {1, 2, \dots, K}$ and $l \in {1, 2, \dots, L}$  which will lead to $KL$ data subsets.
+Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k \in \mathbb{N}, k \leq K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell \in \mathbb{N}, \ell \leq L\}$.
+Here, $A_k$'s or $B_\ell$'s are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
+Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$.
+Data subsets for each combination of levels ($A_k$, $B_\ell$) like <$A_k$, $B_\ell$, $v(s)$> can be obtained for all $k \in {1, 2, \dots, K}$ and $\ell \in {1, 2, \dots, L}$ which will lead to $KL$ data subsets.
 Now, some situations can lead to few or many of these sets being empty.
 We will discuss few cases, where one or more of these $KL$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
-<!--The implication of these empty sets on visualization will be discussed in Section \ref{sec:vis-interaction}.
--->
 
 \begin{definition}\label{def:harmony}
-A {\bf clash} is a pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
+A \textbf{clash} is a pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
 \end{definition}
@@
 \begin{definition}\label{def:clash}
-A {\bf harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
-\end{definition}
-
-<!--To consider a general framework, we will define pairs of cyclic granularities as harmonies or clashes.
-
-<!-- The cyclic time granularities which are essentially categorizations of time are discrete variable, mapped as a factor since they have an inherent order.
-Since two cyclic granularities are mapped to the the fixed and conditioned variable, the way in which they are related becomes important in the data structure.
-To define a general framework, we define harmony and clashes.
--->
-
-<!-- From the above examples, we can see that the choices of the circular granularities are harmonies while some are clashes.
-Also since some choices work and others don't, it must be the attributes of the circular granularities or their relationships which are in play in deciding if the resulting plot would be a good candidate for exploratory analysis.
--->
-
-<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable.
--->
-
-\noindent Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
+A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
+\end{definition}
+
+First, empty combinations can arise due to the structure of the calendar or hierarchy.
 These are called "structurally" empty combinations.
 Let us take a specific example, where $C_i$ be day-of-month with 31 levels and $C_j$ be week-of-month with 5 levels.
-There will be $31\times 5=155$ sets $S_{kl}$ corresponding to possible combinations of $C_i$ and $C_j$.
+There will be $31\times 5=155$ sets $S_{k\ell}$ corresponding to possible combinations of $C_i$ and $C_j$.
 Many of these like $S_{1,5}$, $S_{21,2}$ are empty.
@@
 
-<!--  In fact, most of these 155 sets will be empty, making the combination of $C_i$ and $C_j$ in a graph unhelpful. -->
-
-\noindent Secondly, empty combinations can turn up due to differences in event location or duration in a calendar.
+Second, empty combinations can turn up due to differences in event location or duration in a calendar.
 These are called "event-driven" empty combinations.
 Let us consider $C_i$ be day-of-week with 7 levels and $C_j$ be WorkingDay/NonWorkingDay with 2 levels.
-While potentially all of these 14 sets $S_{kl}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
+While potentially all of these 14 sets $S_{k\ell}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
 For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
 
-<!-- However, this is different from the first example because these are not structurally empty.
-Also, the graphic may still be useful if we want to look at whether the interaction of $C_i$ and $C_j$ affects the distribution we are plotting.
--->
-
-\noindent Thirdly, empty combinations can be a result of how granularities are constructed.
+Third, empty combinations can be a result of how granularities are constructed.
 These are called "build-based" empty combinations.
@@
 
-\noindent An example when there will be no empty combinations could be where $C_i$ and $C_j$ maps index set to day-of-week and month-of-year respectively.
+An example when there will be no empty combinations could be where $C_i$ and $C_j$ maps index set to day-of-week and month-of-year respectively.
 Here $C_i$ can have 7 levels while $C_j$ can have 12 levels.
-So there are $12\times7=84$ sets $S_{kl}$.
+So there are $12\times7=84$ sets $S_{k\ell}$.
 All of these are non-empty because every day-of-week can occur in every month.
@@
 
-<!-- Why we looking at probability measure -->
-
-<!-- For a tsibble with index ranging from ${1, 2, \dots, (T + 1)}$, that is (T+1) observations -->
-
-<!-- In linear granularities, if we consider a data set of size T where each time point corresponds to an unique value of the measured variable.  -->
-
 Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
-<!--This is unlike the case of linear granularities where each time point corresponds to an unique value of the measured variable.--> It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
+It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
 However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
 Summarizing the distribution of the measured variable using these multiple observations could be a potential way to explore and bring forward different features of the data.
-<!-- compare and contrast periodicities in the data.<!--Any attempt to encode all or many cyclic granularities together to develop insights on periodicity might become clumsy.
-Instead, the big problem is broken down by focusing on two cyclic granularities at a time Summarizing properties of the distribution of the measured variable can give valuable indication of features.-->
-
-<!-- using some summary statistics, each of which can highlight different features of the data or through probability densities -->
-
-<!-- T+1 observation not T -->
+
 However, we need to consider the effect of number of observations on the summarization even for harmonies.
@@
 Let $nobs$ be the number of observations for a combination of levels from $(C_i, C_j)$.
-Now, $nobs$ might be equal for all combinations on an average so that $nobs = (T + 1)/(KL)$ as $T\rightarrow\infty$ or ideally take any value between $\{0, 1, 2, \dots, T\}$.
+Now, $nobs$ might be equal for all combinations on an average so that $nobs = (T + 1)/(KL)$ as $T\rightarrow\infty$ or ideally take any value between $\{0, 1, \dots, T\}$.
 The statistical transformation used for summarizing the measured variable should be chosen in line with $nobs$.
@@
 
-<!-- ## Summarizing the distribution of a measured variable {#sec:measured-variable} -->
-
-<!-- <!-- Moreover, the measured variable in the considered data structure can be dicrete or continuous. -->
-
-<!-- Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation.
-Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries.
-Another way of summarizing could be estimating the probability density function (continuous) or probability mass function (dicrete).
-The approach can be parametric or non-parametric.
-A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself.
-These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses.
-Several forms of non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form in @Silverman1986-cl<!--,  @Jones1992-pj)is as follows:-->
-
-<!-- <!-- The most commonly used density estimation method is the kernel density estimator.  -->
-
-<!-- If we assume that $X_1, X_2, \dots, X_n \sim F$ are independent and identically distributed observations from an univariate distribution with probability density/mass function $f(.)$, <!--and $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$,the Kernel density estimator $\hat{f(x)}$ with Kernel $K$ is defined by $\hat{f(x)} = (1/nh)\sum_{i=1}^{n} K(x-X_i)/h$, where $h$ is the smoothing parameter or the bandwidth.
-The MSE of the estimator is given by:
-<!-- \begin{equation}\label{eq:MSE} -->
-<!-- \begin{split} -->
-<!-- MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa -->
-<!-- \end{split} -->
-<!-- \end{equation} -->
-
-<!-- where $\kappa = \int k^2(y)dy$. -->
-
-<!-- CONTENT CUT -->
-<!-- The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by: -->
-
-<!-- \begin{equation} -->
-<!-- \hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u) -->
-<!-- \end{equation} -->
-
-<!-- where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv Q'$. -->
-<!--  -->
-<!-- MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\ -->
-
-<!-- other non parametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990).
-Also sample quantiles can be of the form that are not defined for all p including hinges, letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).
--->
-
-<!-- would involve a faceting approach with one temporal granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis. -->
-
- <!-- Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space. -->
-
 # Visualization {#sec:visualization}
-
-<!-- Why visualize -->
-<!-- Analysts often want to fit their data to statistical models, either to test hypotheses or predict future values.
-However, improper choice of models can lead to wrong predictions.
-One important use of visualization is exploratory data analysis, which is gaining insight into how data is distributed to inform data transformation and modeling decisions.
--->
-
-<!-- When the data for entire population is available at our disposal, then there is no inference that is to be made for the unobserved population. We can just examine what is in front of us. -->
-
-<!-- Include why distribution plot, why 2D, why faceting in introduction -->
-<!-- Why distribution plot -->
-<!-- Huge data difficult to summaries using one statistic-->
-
-<!-- With huge amount of data being available, mean, median or any one summary statistic might not enough to understand a data set. Soon enough following questions become more interesting:   -->
-
-<!--  - Are values clustered around mean/median or mostly around tails? In other words, what is the combined weight of tails relative to the rest of the distribution?  -->
-
-<!--  - Does values rise very quickly between 25th percentile and median but not as quickly between median and 75th percentile?
-More generally, how the variation is the data set changes across different percentiles/deciles?
--->
-
-<!--   - Is the tail on the left hand side longer than that on the right side? Or are they equally balanced around mean/median? -->
-
-<!-- This is when displaying a probability distribution becomes a potentially useful approach. -->
-<!-- The entire distribution can be visualized or some contextual summary statistics can be visualized to emphasize certain properties of the distribution.
-These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anomalies in the data set.
--->
-
-<!-- ## Distributions conditional on bivariate granularities -->
-
-<!-- Why 2D -->
-
-<!-- Graphical output devices are flat.
-If there are only two variables, a Cartesian graph of one variable against the other can show the configuration of points.
-As soon as the data move to three variables, we need to infer multi-dimensional structure by a two-dimensional medium[Cleveland, elements of graphing data].
-Any attempt to encode all these temporal granularities together to develop insights on periodicity might fail or become clumsy.
-Instead, the big problem can be broken down into smaller pieces by focusing on specific tasks per visual representation in a 2D space.
--->
-
-<!-- Why facet-ing -->
-
-<!-- `ggplot2` (Wickham, 2016) is an excellent tool that facilitates the process of mapping different variables to a 2D frame through grammar of graphics.
-One aspect of ggplot2 is faceting, inspired by Trellis plots (Cleveland, 1993).
-These are based on conditioning the values taken on by one or more of the variables in a data set.
-This means creating same plots using data subsets corresponding to each level of the conditioned variable.
-This is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data.
--->
-
-<!-- What are we visualizing -->
-
-<!-- We will see the distribution of the time series across bivariate temporal granularities through faceting approach with one temporal granularity plotted along the x-axis, the other one across facets and the dependent time series variable on the y-axis.
--->
-
-<!-- A well designed visual representation can replace cognitive calculations with simple perpetual inferences and improve speed of comprehension and decision making.
-However, creating a visualization requires a number of nuanced judgements.
-The challenge being given a data set, the number of visual encodings and hence the space of possible visual design is extremely huge.
--->
 
@@
 Drawing from the grammar of graphics, if <$C_i$, $C_j$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
-
-<!-- CONTENT CUT -->
-<!-- There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements.
-Data are observations made on a number of variables, typically thought of as a sample from a population distribution.
-Aesthetic mapping describes how variables are mapped to elements of the plot, such as horizontal axis or color.
-Geometries determine the physical representation of the data in the plot.
-The remaining grammatical elements of a graphic namely facets, statistics, coordinates and themes are optional and contains detail of the plot.
-Facets dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
--->
-
-<!-- Briefly, the grammatical elements of any plot are as follows: -->
-
-<!-- - **Layers**:  made up of data, mapping, statistical transformation, geometric objects and position adjustments -->
-
-<!--   - **Data**: observations made on a number of variables, typically thought of as a sample from a population distribution -->
-
-<!--   - **Mappings** describing how variables are mapped to elements of the plot, such as horizontal axis or color -->
-
-<!--   - **Statistical transformation** can summarize the data in many useful ways. The geometric object determines the physical representation of the data in the graph. -->
-
-<!--   - **Geometric objects** control the type of plot that is created and are classified by their dimensions, for example points (0 dimension), line (1 dimension), polygons(2 dimensions) and so on.
--->
-
-<!--   - **Position adjustment** of elements on the plot are needed for dense data, otherwise data points might obscure one another. -->
-
-<!-- - **Facet** dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.
--->
-
-<!-- - **Scale** controls how data is mapped to aesthetic attributes. For every aesthetic property employed in a layer, we need one scale.  -->
-
-<!-- - **A coordinate system** maps the position of objects onto the plane of the plot, and controls how the axes and grid lines are drawn.  -->
 
@@
 
-<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
-
-<!-- The three essential component layers are data, aesthetic mapping and the geometric objects. -->
-<!-- Data are observations made on a number of variables, typically thought of as a sample from a population distribution.
-Mappings describe how variables are mapped to elements of the plot, such as horizontal axis or color -->
-
-<!-- the second layer is the aesthetic mapping, which refers to the scales on to which we want to map our data -->
-<!-- The third essential layer is the geometry which decides how the plot will look. After the essential layers, the next layer being used is the facet which dictates how to split up our plot -->
-
-<!-- Aesthetics refers to the scales on to which we want to map our data. we create plots by layering grammatical elements on top of one another and use aesthetic maps -->
-
-<!-- The grammar of graphics establish the building blocks for creating meaningful visualizations. It -->
-<!-- This means we have a framework that allows us to communicate in a way.  -->
-
-<!-- .
-allows us to produce graphics using the same structured thinking that is used to design analysis.
-Briefly, any graphics are made up of **Layers**, **Facet**, **Scale** and a  **coordinate system**.
-**Layers** are made up of **data**, **mapping**, **statistical transformation**, **geometric objects** and **position adjustments**.
--->
-
-<!-- The three key components of every visualization is: data, aesthetic mappings between variables in the data and visual attributes, and at least one layer which describes how to render each observation.
-Layers are usually created with a geom function.
-Considering the data structure <$A_k$, $B_l(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, visualizations can be defined by: -->
-<!-- - Data: <$A_k$, $B_l(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ -->
-<!-- - Aesthetic mapping: $A_k$ mapped to $x$ position and $v(s)$ to $y$ position. -->
-<!-- - Layer: any geom that summarizes distribution -->
-
-<!-- To add additional variables to a plot, we can use other aesthetics like colour, shape, and size or faceting. Faceting creates tables of graphics by displaying the same graph for each subset. -->
-<!-- - Facet:  $B_l(s)$ -->
-
-<!-- We exploit some features of layers and faceting [@Wilkinson1999-nk] and [@Wickham2009pk] to come up with a recommendation system while visualizing the distribution of the "dependent" variable across bivariate cyclic temporal granularities with the data structure <$A_k$, $B_l(s)$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$.
-The framework involves faceting approach with one cyclic granularity plotted along the x-axis, the other one across facet and the dependent time series variable on the y-axis.
--->
-
-<!-- Hence, the ingredients for the plots are subsets of numbers (or categorical measurements), and the choice of faceting or axis alignment for the subset strata.
-The basic plot choice is display of density.
-Different plots might be appropriate depending on which features of the density we arae interested to look at, the levels of time granularities being plotted, how granularities interact and number of observations available.
-We will look at each of these aspects separately and analyze the effect of each on visualization.
-While we accomplish that, we need to bear in mind that any customized species of visualization might be constructed by varying the set of mappings between data properties and visual attributes such as position, shape and color.
-We are trying to address problems where even with good aesthetic qualities and good data, the graph might be confusing or misleading because of how people process the information.
--->
-
-<!-- Now, due to the hierarchical arrangement of the granularities, there are certain granularities which when plotted together do not give us the layout to do exploration, for example, structurally empty combinations (clashes) are not recommended to plot together.
-The harmonies when plotted together can help exploration.
-But still the question remains that which distribution plot should be chosen to bring out the best of exploratory data analysis.
-This is a function of which features of the distribution we are interested to look at, how much display space is available to us and also if the number of observations are enough for that distribution plot.
--->
-
 ## Choice of statistical transformations and geometric objects
 
-<!-- Common benefits and challenges of empirical and density based visualization -->
-
 Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization.
-The basic plot choice for our data structure is the one that can display distributions<!--In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated.-->.
+The basic plot choice for our data structure is the one that can display distributions.
 We will discuss few conventional and recent ways to plot distributions using both Kernel density estimates and descriptive statistics.
@@
 
-<!-- CONTENT CUT -->
-<!-- Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set.
-The descriptive statistics based methods do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former.
-However, they avoid clutter and help us to focus on some specific properties of the data.
-Also, the summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.
--->
-
-<!--  Furthermore, these methods are subject to the same sample size restrictions and challenges that apply to any density estimation.  -->
-<!-- The probabilities displayed are based on actual data, which is aligned to principles that governed Tukey’s original boxplot.
-They can be useful in spotting multimodality, however, they might become obscuring with too many categories and information on the entire distribution can add to cognitive load.
-Also, the probabilities in density plot are estimated through kernel density estimation and thus makes assumptions when selecting kernel or bandwidth.
-As a result, the plot shows smooth summaries based on the assumptions and not only on actual observations.
--->
-
-<!-- CONTENT CUT -->
-<!-- ### Descriptive statistics based displays -->
-
-<!-- Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers at a glance.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.
-Due to wide spread popularity and simplicity in implementation, a number of variations like Notched box plots [@Mcgill1978-hg] are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers.<!--has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.
-Challenge of boxplot.
-The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance.
-However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy.
-Also, the number of outliers is large for larger data sets.
--->
-
-<!-- <!-- Description and lit review of letter-value plots -->
-<!-- The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets.<!--“outliers” in letter value plots are those observations beyond the most extreme letter value.
-The letter values are displayed till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.
--->
-
-<!-- <!-- Description and lit review of quantile plots -->
-
-<!-- Quantile plots display quantiles instead of quartiles in a traditional boxplot.
-These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities.
-While in a boxplot and letter value plots are well defined <!--an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g.
-outside 1.5 times the interquartile range above the upper quartile and below the lower quartile), in quantile plots outliers are open to interpretation and not shown.
--->
-
-<!-- <!-- Description and lit review of density plots -->
-
-<!-- <!-- The vase plot [@Benjamini1988-io, 1988] was a major revision from the concept of box plots where the width of box at each point is proportional to estimated density.  -->
-
-<!-- ### Kernel density based displays -->
-
-<!-- Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi].
-The shape of the violin represents the density estimate of the variable.
-<!--The more data points in a specific range, the larger the violin is for that range.
-Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.
-It becomes less useful when there are too many levels or categories to display.
--->
-
-<!-- The summary plot [@Potter2010-qc] combines a minimal box plot with glyphs representing the first five moments and a sectioned density plot crossed with a violin plot and an overlay of a reference distribution.
-This suffers from the same problem as boxplots or violin plot, as it is combination of those two.
--->
-
-<!-- <!-- Description and lit review of density plots -->
-<!--  A Ridge line plot <!--(sometimes called Joy plot) shows the distribution for several groups.
-Distribution can be represented using density plots, all aligned to the same horizontal scale and presented with a slight overlap.
-Like other density based displays, these plots allow us to see multimodality in the distribution.
-However, these plots can be obscuring when there is overlap of distribution for two or more categories of the y-axis.
-Also, with lot of categories, it is difficult to compare the height of the densities across categories.
--->
-
-<!-- <!-- Description and lit review of HDR plots -->
-<!-- The highest density region (HDR) box plot proposed by [@Hyndman1996-ft] displays a probability density region that contains points of relatively highest density.
-The probabilities for which the summarization is required can be chosen based on the requirement.
-These regions do not need to be contiguous and help identify multi-modality.
--->
-
-<!-- Each type of density display has different parameters, that need to be estimated given the data.
-Each is equipped with some benefits and challenges which should be borne in mind while using them for exploration.
-<!--For example, if we have too many categories in the variables mapped to facet or x-axis, the quantile plots might be useful for comparing patterns, whereas, other more involved methods of plotting are useful for studying anomalies, outlier or multimodal behavior-->
-
 ## Facet and aesthetic variables
@@
 
-The levels of cyclic granularities has a role to play on the choice of plots <!--Recall that for cyclic granularities $C_i$ and $C_j$ such that $C_i$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_j$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}.
-$A_k$'s and $B_l$'s are referred to as the levels of $C_i$ and $C_j$ for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$ respectively.The levels have an impact on the--> since space and resolution might become a problem with too many levels.
+The levels of cyclic granularities has a role to play on the choice of plots since space and resolution might become a problem with too many levels.
 A potential approach could be to categorize the levels as very high/high/medium/low for each cyclic granularity and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
-Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight or days of the week.<!--For example, any levels above 31 can be considered as very high, any levels between 14 to 31 can be taken as high and that between 7 to 14 can be taken as medium and below 7 as low.>
-
-<!-- The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us.  -->
-
- <!-- - If levels of both the granularities are low, then any statistical distribution plot might be chosen depending on which feature of the distribution one is interested to look at.  -->
-
-<!-- CONTENT CUT -->
-<!-- The following principles are useful while choosing distribution plots given two temporal granularities: -->
-
-<!--   - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus. -->
-
-<!--  - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories. -->
-
-<!--   - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred. -->
-
-<!--   - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.
--->
-
-<!-- Let us consider case by case and see which plots are better suitable in which scenarios. -->
-
-<!--  Assumptions should be made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.  -->
-<!-- Moreover, the recommendation system ensures that there are just enough observations before choosing a distribution plot. -->
-
-<!-- Levels are categorized as very high/high/medium/low each for the facet variable and the x-axis variable.
-Default values for these levels are based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week.
-For example, any levels above 31 is considered as very high, any levels between 14 to 31 are taken as high and that between 7 to 14 is taken as medium and below 7 is low.
-31, 14 and 7 are the levels of days-of-month, days-of-fortnight and days-of week respectively.
-These default values are decided based on usual cognitive power while comparing across facets and display size available to us.
-Let us consider case by case and see which plots are better suitable in which scenarios.
--->
-
-<!-- Faceting creates small multiples of data each representing a different subset of the data.
-Small multiples are powerful for exploratory data analysis where we can rapidly compare patterns in different parts of the data and see whether they are the same or different.
-But space and resolution quickly become a problem if the number of levels of facet variables is too high.
-Also, Cleveland suggested that a graphic should display as much information as it can, with the lowest possible cognitive strain to the viewer.
--->
-
-<!-- ## Choosing harmonies to show -->
-
-<!-- Even after screening harmonies from the list of all possible pairs of cyclic time granularities, the options could be too many and it might be useful to rank these harmonies.
-Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities.
-For each facet level, the most important display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum.
-And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data.
-Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leibler divergence measure.
-Probability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.
--->
-<!-- Suppose we have ${x_1, x_2, \dots, x_n} which are random sample from some underlying population. --> <!-- Let $p(x)$ and $q(x)$ be two probability distributions of a random variable x. -->
-<!-- We compute Kullback-Liebler distance based on quantiles $q_1$ and $q_2$ of $v$ at probabilities $p$ corresponding to $C_i$ and $C_j$ respectively. -->
-
-<!-- \begin{equation} -->
-<!-- D(q_1,q_2) = \int^{\infty}_{-\infty}q_1(x)logq_1(x)/q_2(x)\,dx -->
-<!-- \end{equation} -->
-
-<!-- But since this is not symmetric, Jensen-Shanon distance is calculated which is based on Kullback-Liebler divergence. -->
-
-<!-- $$JS(q_1, q_2) = [D(q_1, r) + D(q_2, r)]/2 where r = (q_1 + q_2)/2$$ -->
+Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight or days of the week.
 
@@
 ```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
-
 ```
@@
 We will discuss the effect of the mapping of cyclic granularities in this section.
-When we consider data sets of the form <$C_i$, $C_j$, $v$> with $C_i$ mapped to $x$ position and $C_j$ to facets, then $A_k$'s are placed in close proximity and each $B_l$ represent a group/facet.
+When we consider data sets of the form <$C_i$, $C_j$, $v$> with $C_i$ mapped to $x$ position and $C_j$ to facets, then $A_k$'s are placed in close proximity and each $B_\ell$ represent a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case $A_k$'s are compared against each other within each group.<!-- When $C_j$ is mapped to facets and $C_i$ is mapped to $x$ axis, comparisons would ideally happen across all $B_l$ within each $A_k$--> With the mapping of $C_i$ and $C_j$ reversed, emphasis will shift to different behavior of the variables.
+Hence, in this case $A_k$'s are compared against each other within each group.
+With the mapping of $C_i$ and $C_j$ reversed, emphasis will shift to different behavior of the variables.
 \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
@@
 
-<!-- For Kernel density based plots $nobs>30$ is suggested, for summary statistics based plots like boxplots $nobs = 10$ is acceptable and for quantile plots $nobs$ should at least be equal to the number of quantiles plotted.
-If we require at least $k$ observations per combination to create a meaningful summarization, then only $T \ge lmk$ will be acceptable.
-The value of $k$ will depend on the statistical transformation we are producing.
-For computing deciles, $k=10$ may be acceptable, but for density estimates, $k \ge 30$ is suggested.
--->
-
-<!-- CONTENT CUT -->
-<!-- ### Rarely occurring events -->
-
-<!-- Suppose we have $T$ observations, and two cyclic granularities $C_i$ with $n$ categories and $C_2$ with $m$ categories.
-Each element of $C_i$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times.
-There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$.
-If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable.
-The value of $k$ will depend on the statistical transformation we are producing.
-For computing deciles, $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$.
-Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.
--->
-
-<!-- The above discussion shows that problems will occur when $n_1n_2 > T/k$, even without structurally empty combinations. -->
-
-<!-- (see \autoref{eq:MSE})  shows that the variance of both $\hat{f(x)}$ <!--and $\hat{q(u)}$is a function of the sample size $n$.-->
-
-<!-- the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$,  -->
-
-<!-- Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels. -->
-
 # Applications {#sec:application}
@@
 
-<!-- # how data looks -->
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
 One of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
-<!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to visualize the distribution of the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
-
-<!-- and identify the extreme and regular households. -->
-
-<!-- ### Data structure for visualization: -->
-
-<!-- The data structure for a sample of 50 households from that trial is depicted in \autoref{tab:smart-data} where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
--->
-<!-- <!-- We take a random sample of 50 households to see how we can achieve that.  -->
-
-```{r smart-data}
-# load("data/sm_cust50.Rdata")
-#
-# smart_data <- smart_meter <- sm_cust50 %>%
-#   as_tsibble() %>%
-#   dplyr::select(
-#     customer_id,
-#     reading_datetime,
-#     general_supply_kwh
-#   ) %>% head(n = 5)
-#
-#   knitr::kable(smart_data,
-#                format = "latex",
-#                booktabs = TRUE,
-#                caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
-```
-
-<!--of the tsibble and represents the time variable.
-The variable denoting different households represent the key of the tsibble and is the time series variable which needs to be analyzed.Data sets of the form <$C1$, $C2$, `general_supply_kwh`> will form the basis of exploration for each key in the data set, where the cyclic granularity $C1$ would be mapped to $x$ axis and `general_supply_kwh` to $y$ axis with the cyclic granularity $C2$ mapped to facet.
-Any geometry displaying the probability distribution of `general_supply_kwh` like boxplot, letter value, violin,
-ridge or highest density region plots can then be used to explore the data set.-->
+The idea here is to show how to visualize the distribution of the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
 
@@
 Considering $6$ linear granularities half-hour, hour, day, week, month and year in the hierarchy table, $N_C = (6*5/2) = 15$.
-<!--(via Section \ref{sec:data-structure}) circular or quasi-circular granularities that could be formed relating two linear granularities at a time.--> If $N_C$ seem too large, the smallest and largest linear granularities could be considered to be removed for analysis.
+If $N_C$ seem too large, the smallest and largest linear granularities could be considered to be removed for analysis.
 Granularities half-year and year are removed to have $N_C = (4*3/2) = 6$ and obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
@@
 Now that we have a list of cyclic granularities to look at, we should be able to compute them using Section \ref{sec:cyclic-calendar}.
-<!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
-with the index set mapping half-hourly time periods to a set of positive integers.-->
-
-<!-- R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively.
-Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
-Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
-Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed.
-If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped.
-Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".
-Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Sections \ref{sec:cyclic-calendar}.
-<!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).
--->
-<!-- Here, the index set $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ maps the half-hourly time instants to a set of positive integers.
-Then, multiple-order-up circular granularity $hour\_week$ and quasi-circular granularity $hour\_month$ could be computed from single-order-up granularities as follows: -->
-<!-- <!-- $$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular}) -->
-<!-- \begin{align} -->
-<!--   hour\_week(z) =  day\_week(z) + 24*hour\_day(z) -->
-<!-- \end{align} -->
-
-<!-- <!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) -->
-<!-- \begin{align} -->
-<!-- hour\_month(z) =  hour\_day(z) + 24*day\_month(z) -->
-<!-- \end{align} -->
-
-<!-- These cyclic granularities are a function of the index variable and are computed based on if they are circular <!--(Section \ref{sec:circular-gran-def}), quasi-circular.
-<!--(Section \ref{sec:quasi- circular-gran-def}) or aperiodic (Section \ref{sec:aperiodic-gran-def})-->
 
@@
 
-<!-- #smart_smart_search <- smart_meter %>% search_gran() -->
-<!-- #knitr::kable(smart_smart_search, format = "markdown") -->
-
-<!-- CONTENT CUT -->
-<!-- # ```{r search_limit1, echo=TRUE} -->
-<!-- # smart_meter10 %>% -->
-<!-- #   search_gran(highest_unit = "month") -->
-<!-- # ``` -->
-
-<!-- # smart_meter_limit1 <- smart_meter10 %>%  -->
-<!-- #   search_gran(highest_unit = "month") -->
-<!-- # knitr::kable(smart_meter_limit1, format = "markdown") -->
-
 ```{r search_gran_limit2, echo = FALSE, eval = FALSE}
 ```
-
-<!-- # smart_meter_limit2 <- smart_meter10 %>% search_gran(highest_unit = "month", -->
-<!-- #                               filter_out = c("hhour", "fortnight") -->
-<!-- # ) -->
-<!-- # knitr::kable(smart_meter_limit2, format = "markdown") -->
-
-<!-- Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:circular-gran-def}, \ref{sec:quasi-circular-gran-def} and \ref{sec:aperiodic-gran-def}.
--->
 
@@
 
-<!--The MMPD is also shown along with indicator (*) only when variation of measured variable across the harmony pair significant.
-Starting from $42$ possible pairs of cyclic granularities to visualize, we are finally left with only $6$, which is a very sizable number of displays for exploration.-->
-
-\noindent Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the <!--significance of MMPD, threshold and the -->impact of different distribution plots and reverse mapping.
+Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the impact of different distribution plots and reverse mapping.
 For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
@@
 \autoref{fig:bothcust} (a) displays reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour of the day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.
-In \autoref{fig:bothcust} (b), <!--the distribution of energy consumption is plotted across the harmony pair (weekday/weekend, hour of the day) through an area quantile plot-->the black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
+In \autoref{fig:bothcust} (b), the black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
 The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
@@
 
-\noindent Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust}(c) for some more insights.
+Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust}(c) for some more insights.
 There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
-<!--Also the hours from 7 to 13 look most volatile.--> If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
+If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
 For example in \autoref{fig:bothcust}(a), it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends.
@@
 
-<!-- From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior.
-Given the data structure <$C1$, $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet.
-Thus the problem is equivalent to taking 2 granularities at a time from six, which essentially is equivalent to having 30 data subsets for visualization.
-However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together.
-\autoref{tab:harmony-tab} shows 13 harmonies pairs, each of which can be plotted together to look at the energy behavior from different perspectives.
--->
-
-<!-- We will look at few harmony pairs and see how the proposed workflow can be utilized.  -->
-
-<!-- From the Gestalt theory with the same data  <$C1$ $C2$, `general_supply_kwh`>, mapping $C1$ on the x-axis vs facet might lead to different insights. -->
-
 ```{r harmony-tab, echo=FALSE}
@@
 
-<!-- # smart_meter10 %>% gran_advice("wknd_wday", "hour_day") -->
-
-<!-- The median consumption or width of other bands doesn't vary much across seasons, implying extreme behavior or extreme customers does not vary across seasons much.  -->
-
-<!-- No need to share the same for 50 customers - cut down text and anyway nothing much interesting is coming -->
-
-<!-- ```{r wknd-wday50, fig.cap = "Quantile plot of energy consumption of 50 households across different hours of the day faceted by weekday and weekend.
-The quantiles are not different for weekdays and weekends implying either the behavior balances out among these customers or most of them do not behave differently for weekdays and weekends."} -->
-
-<!-- smart_meter <-  -->
-<!--   read_rds("data/sm_cust50.rds") -->
-
-<!-- smart_meter %>%  -->
-<!--   prob_plot("wknd_wday", -->
-<!--                             "hour_day", -->
-<!--                             response = "general_supply_kwh", -->
-<!--                             plot_type = "quantile", -->
-<!--             symmetric = FALSE,  -->
-<!--             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
-<!--   scale_y_sqrt() -->
-<!-- ``` -->
-
-<!-- # ```{r wknd-wday50_df} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   prob_plot("month_year", -->
-<!-- #                             "day_fortnight", -->
-<!-- #                             response = "general_supply_kwh", -->
-<!-- #                             plot_type = "quantile", -->
-<!-- #             symmetric = TRUE,  -->
-<!-- #             quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)) + -->
-<!-- #   scale_y_sqrt(breaks = c(0.1, 0.25, 0.5, 1:3)) -->
-<!-- # ``` -->
-
-<!-- ```{r my-hd, fig.cap="Area quantile plots of energy consumption across hours of the day faceted by months of the year.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-The median and quartile consumption increase in winter months, but extreme behavior represented by higher bands mostly stay the same across seasons.
-"} -->
-
-<!-- library(gravitas) -->
-<!-- library(tsibble) -->
-
-<!-- library(ggplot2) -->
-<!-- smart_meter %>% -->
-<!--   prob_plot("month_year", "hour_day", -->
-<!--     response = "general_supply_kwh", -->
-<!--     plot_type = "quantile", -->
-<!--     quantile_prob = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99), -->
-<!--     symmetric = TRUE) + -->
-<!--   ggtitle("") + scale_y_sqrt(breaks = c(0.01, 0.1, 0.5, 1:3)) -->
-<!-- ``` -->
-
-<!-- Figure \ref{fig:dw-hd} shows that on morning peaks for Saturday and Sunday are higher and occurs in a later hour of the day compared to weekdays.
-This is true for both top 50% and 75% of the customers.
-Hence, this is indicative of the fact that at least 50% of the households in this subset has traditional work days (Mon-Fri) and weekends(Sat-Sun).
-All the percentiles peaks up in the morning hours and then drops in the afternoon, before again peaking up in the evening.
--->
-
-<!-- ```{r dw-hd} -->
-
-<!-- # Hourly usage across days of the week -->
-<!-- sm_cust50 %>% -->
-<!--   prob_plot("day_week",  -->
-<!--            "hour_day", -->
-<!--            "general_supply_kwh",  -->
-<!--            plot_type = "quantile",  -->
-<!--           quantile_prob = c(0.25, 0.5, 0.75), -->
-<!--            response = "general_supply_kwh", overlay = FALSE) + -->
-<!--   ggtitle("") -->
-
-<!-- ``` -->
-
-<!-- on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.  -->
-
-<!-- typical and extreme behaviors of 50 households from that trial.  -->
-
-<!-- , and indeed many parts of the world.
-Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
--->
-
-<!-- <!-- Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
-<!-- It is common to see aggregates of usage across households, total kwh used each half-hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
--->
-
-<!-- We want to look at the distribution of energy across coarser temporal granularities and then deep dive into finer temporal granularities. -->
-
-<!-- In Figure \ref{fig:day-fortnight} and \ref{fig:day-fortnight-agg} , the smart meter data is filtered for two customers to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
--->
-
-<!-- can be added later -->
-
-<!-- In Figure \ref{fig:day-fortnight},to illustrate what kinds of insights can be drawn for the energy behavior of these two customers.
-it can be seen that for most days of the fortnight, the second household has much less consumption than the first one.
-However, there is additional information that we can derive looking at the distribution.
-If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
-However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
-This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
--->
-
-<!-- # ```{r day-fortnight_agg, echo = FALSE} -->
-<!-- # library(lvplot) -->
-<!-- # library(ggplot2) -->
-<!-- # library(dplyr) -->
-<!-- # library(tibble) -->
-<!-- # library(gravitas) -->
-<!-- # library(tidyr) -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!-- #   create_gran("day_fortnight") %>% -->
-<!-- #   as_tibble() %>%  -->
-<!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarize(median= log(median(general_supply_kwh)), -->
-<!-- #             total = log(sum(general_supply_kwh))) %>%  -->
-<!-- #   gather("daily_consumption", "value", - c(customer_id, day_fortnight)) %>%  -->
-<!-- #   ggplot2::ggplot(aes( -->
-<!-- #     x = as.factor(day_fortnight), -->
-<!-- #     y = value,  -->
-<!-- #     color = daily_consumption, -->
-<!-- #     group = daily_consumption)) + -->
-<!-- #   geom_line() + -->
-<!-- #   xlab("day_fortnight") + -->
-<!-- #   ylab("general_supply_kwh") +  -->
-<!-- #   facet_wrap(~customer_id) +  -->
-<!-- #   theme_bw()  -->
-
-<!-- # smart_meter10 %>%  -->
-<!-- #   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!-- #   create_gran("day_fortnight") %>% -->
-<!-- #   as_tibble() %>%  -->
-<!-- #   group_by(customer_id, day_fortnight) %>%  -->
-<!-- #   summarize(median_consumption= median(general_supply_kwh)) %>% -->
-<!-- #   ggplot2::ggplot(aes( -->
-<!-- #     x = as.factor(day_fortnight), -->
-<!-- #     y = median_consumption, -->
-<!-- #     color =  customer_id, -->
-<!-- #     group = customer_id)) + -->
-<!-- #   geom_line() + -->
-<!-- #   xlab("day_fortnight") + -->
-<!-- #   ylab("general_supply_kwh") +  -->
-<!-- #   theme_bw()  -->
-<!-- ``` -->
-
-<!-- ```{r day-fortnight, fig.cap="Letter value plot of two households across days of the fortnight.
-M, F, E, D and C represents the letter values of energy consumption.
-Regular behavior (within letter value F) is more variable for the 2nd household, whereas extreme behavior (beyond letter value F) is varying more for the first household."} -->
-<!-- library(lvplot) -->
-<!-- library(ggplot2) -->
-<!-- library(dplyr) -->
-<!-- library(tibble) -->
-<!-- library(gravitas) -->
-<!-- smart_meter10 %>% -->
-<!--   filter(customer_id %in% c(10006704, 10017936)) %>%  -->
-<!--   create_gran("day_fortnight") %>% -->
-<!--   ggplot2::ggplot(aes( -->
-<!--     x = as.factor(day_fortnight), -->
-<!--     y = general_supply_kwh)) + -->
-<!--   xlab("day_fortnight") + -->
-<!--   geom_lv( -->
-<!--     outlier.colour = "red", -->
-<!--     aes(fill = ..LV..), -->
-<!--     k = 5) + -->
-<!--   facet_wrap(~customer_id) +  -->
-<!--   scale_fill_lv() + -->
-<!--   theme_bw() + -->
-<!--   scale_y_sqrt()  -->
-<!-- ``` -->
-
-<!-- Figure \autoref{fig:my-hd} shows the overlay quantile plot of energy consumption of 50 households.
-The black line is the median, whereas the pink band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-Each facet represents a month and energy consumption across each hours of the day is shown inside each facet.
-It can be observed that the median is very close (and hence not distinctly visible) to the lower boundaries of all the other bands implying energy consumption for these households are extremely left skewed.75% of the households has distinct higher evening peaks compared to morning peaks, especially in the winter months.
-The top 5% households behave very differently than even the top 10% households in terms of total energy consumption across hours of the day.
--->
-
-<!-- # ```{r is_harmony, echo= TRUE} -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   is_harmony(gran1 = "hour_day",  -->
-<!-- #              gran2 = "day_week") -->
-<!-- #  -->
-<!-- # smart_meter10 %>% -->
-<!-- #   is_harmony(gran1 = "hour_day",  -->
-<!-- #              gran2 = "day_week",  -->
-<!-- #              facet_h = 14) -->
-<!-- #  -->
-<!-- # smart_meter10 %>%  -->
-<!-- #   is_harmony(gran1 = "day_month", -->
-<!-- #              gran2 = "week_month") -->
-<!-- # ``` -->
-
-<!-- Boxplot of energy consumption is shown across `wknd_wday` (facet) and `hour_day` (x-axis) for the same two households.
-For the first household, the shape of the distribution does not change significantly for weekdays and weekends.
-The distribution is roughly smooth with a morning and an evening peak.
--->
-
-<!-- If we consider letter value F as a regular behavior and letter values beyond F as not-so-regular behavior, we can conclude that the regular behavior of the first household is more stable than the second household.
-However, the distribution of tail of the first household is more variable, observed through distinct letter values, implying that their not-so-regular behavior is quite extreme.
-This shows, how looking at the distribution of the dependent variable can throw more light on the energy behavior of the customers, which are lost using aggregate or summary statistics.
--->
-
-```{r bothcust, fig.cap = "(ref:bothcust)", out.width="90%"}
+```{r bothcust, fig.cap = "(ref:bothcust)", out.width="90%", fig.pos="!htb"}
 ```
@@
 
-<!-- # ```{r cust-ridge, fig.cap ="Boxplot of energy consumption across hours of the day faceted by weekday/weekend for customer id: 10017936.
-The change in mapping of the cyclic granularities leads to easier comparison within weekdays and weekends for each hour.
-Median consumption till 6am is higher for weekends compared to weekdays although the quartiles look similar.
-Outliers appear more during the end of the day implying more uncertain behavior in latter hours of the day compared to early morning hours."} -->
-<!-- #  -->
-<!-- # smart_meter10 %>% -->
-<!-- #   filter(customer_id %in% c(10017936)) %>% -->
-<!-- #   prob_plot("hour_day", -->
-<!-- #     "wknd_wday", -->
-<!-- #     response = "general_supply_kwh", -->
-<!-- #     plot_type = "boxplot" -->
-<!-- #   ) + -->
-<!-- #   scale_y_sqrt() -->
-<!-- # ``` -->
-
-<!-- Now, we would like to see how these customers' behavior relate to the rest of the 50 households across these two measures -   -->
-
-<!-- - if energy distribution is skewed towards extreme? -->
-
-<!-- - if the weekend and weekday behavior are different for most of these households? -->
-
 If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one observation shows a particular behavior, but this behavior paradoxically becomes obscured by aggregation.
@@
 This is beyond the scope of the current work.
-
-<!-- For the purpose of this paper, the smart meter data is filtered for two customers to illustrate kinds of insights that can be drawn after proceeding with the systematic exploration. -->
-
-<!-- For example, Figure \ref{fig:wknd-wday50} shows the quantile plot of energy consumption of a sample of 50 households.
-Quantiles for weekends and weekdays for these customers look exactly the same.
-It is not clear if the behaviors of most of these customers do not alter between weekdays and weekends or looking at all customers together balances out the behaviors.
-Consider another example in Figure \ref{fig:my-hd} showing the area quantile plots of 50 customers.
-The median is very close to the lower boundaries of all the other bands implying energy consumption for these households are left skewed.
-The pink band changes significantly in winter months (May - August) implying that much of the behavioral changes occur in the quartiles, that too in peak hours of the day.
-It is to be noted here that the the level of all quantiles increased too in winter months (increased energy usage in weather conditions), the interesting part is to notice that the relationship between bands other than quartile stayed same across seasons.
--->
-
-<!--This case study shows systematic exploration of energy behavior for a household to gain exhaustive insights on periodic behavior of the households. -->
-
-<!--First, it helps us to find the list of cyclic granularities to look at, then shrinks the number of possible visualizations by identifying harmonies, visualize a harmony pair and shows the effect of different distribution plots or reverse mapping.-->
 
@@
 
-<!-- However, irrespective of the type of cyclic granularities, it can be interesting to visualize the distribution of a measured variable across these cyclic granularities to throw light on the periodic behavior of a non-temporal data set similar to any temporal data set.
--->
-
-\noindent The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
+The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
 The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
@@
 However, even if the situation is not periodic, it can be interesting to visualize the distribution of a measured variable to shed light on the aperiodic behavior of a non-temporal data set similar to aperiodic events like formal meetings, workshops, conferences, school semesters in a temporal set up.
-<!--There are many interesting questions that could possibly be answered with such a data set irrespective of the type of cyclic granularities.-->
 
@@
 
- <!--We will explore a few interesting questions and understand how the proposed approach in the paper can help answer some of the questions.-->
-
-<!-- the two teams have a single innings each, which is restricted to a maximum of 20 overs.
-Hence, in this format of cricket, a match will consist of 2 innings, an innings will consist of 20 overs, an over will consist of 6 balls with some exceptions.
--->
-
-<!-- ```{r hierarchy, echo=FALSE} -->
-<!-- hierarchy_model <- tibble::tibble(linear gran = c("ball", "over", "inning", "match"),  -->
-<!--                                   convert_fct = c(6, 20, 2, 1)) -->
-<!-- knitr::kable(hierarchy_model, caption = " A hierarchy table for T20 cricket") -->
-<!-- ``` -->
-
-<!-- Each team is given a two-and-a-half-minute "strategic timeout" during each innings; one must be taken by the bowling team between the ends of the 6th and 9th overs, and one by the batting team between the ends of the 13th and 16th overs.
--->
-
-<!-- Suppose, we are interested to see how the distribution of scores vary from the start to the end of the game. Let us brainstorm some of the questions that might help us comprehend that. -->
-
-<!-- a) What is the most common pattern for batting and bowling teams across balls, overs, innings and matches? Which are the teams which are typical and which are exceptions? -->
-
-<!-- How the scores per over vary across granularities/categorizations like innings of a match or matches of a season?
-Are these different for the winning teams in that season and the teams that couldn't qualify for the playoffs?
--->
-
-<!-- Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings? -->
-
-<!-- We will look at the ball by ball data for all batting teams.
-Since we want a periodic world, where each over consists of 6 balls and each match consists of two innings, we shall filter out the matches or overs for which that is not true.
-Also, we look at runs per over as that would have more variability compared to runs per ball and it be easier to observe the strategies of the winning team through that.
--->
-
-<!-- Making it short -->
-
-<!-- First, we look at the distribution of runs (measured variable) across over-of-inning (circular granularity) and match-of-season (aperiodic cyclic granularity) in \autoref{fig:seas-over-inning}.
-The distribution of runs per over has not significantly changed from 2008 to 2016.
-There is no clear pattern/trend that runs per over is increasing or decreasing across seasons.
-Hence, we work with subsets of seasons to answer some of the questions: -->
-
-\noindent There are many interesting questions that could possibly be answered with the `cricket` data set.
+There are many interesting questions that could possibly be answered with the `cricket` data set.
 For example, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second innings.
-The Mumbai Indians (MI) and Chennai Super kings (CSK) <!--are considered one of the best teams in IPL with multiple winning titles and --> appeared in final playoffs from 2010 to 2015.
+The Mumbai Indians (MI) and Chennai Super kings (CSK) appeared in final playoffs from 2010 to 2015.
 We take their example in order to dive deeper into this question.
-<!--Circular granularities "over-of-inning" and "inning-of-match" can be computed using \ref{sec:circular-gran-def} with over as index of the tsibble.--> From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
+From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
 This seem to indicate that players feel mounting pressure to score more runs as they approach towards the end of the first inning.
@@
 
-<!-- Also longer and more distinct letter values in the second innings suggests that the variability is more in the second innings. -->
-<!-- - Q1: How their run rates vary depending on if they bat first or 2nd? Is there a chance that they are more likely to win if they bat first? -->
-
-<!-- - Q2: Which team is more consistent in their approach in terms of run rate across different overs of the innings? -->
-
-<!-- # ```{r seas-over-inning, fig.cap = "Quantile plot of runs per over across overs of different seasons. There is no pattern on increase or decrease of runs across overs for seasons."} -->
-<!-- #   cricket_tsibble_all %>%  -->
-<!-- #   prob_plot("over",  -->
-<!-- #             "season", -->
-<!-- #             hierarchy_model, -->
-<!-- #             response = "runs_per_over", -->
-<!-- #             plot_type = "quantile", -->
-<!-- #             quantile_prob = c(0.25, 0.5, 0.75),  -->
-<!-- #             symmetric = FALSE)  + -->
-<!-- #      theme( # remove the vertical grid lines -->
-<!-- #             panel.grid.major.x = element_blank() , -->
-<!-- #             # explicitly set the horizontal lines (or they will disappear too) -->
-<!-- #             panel.grid.major.y = element_blank()) + scale_x_discrete(breaks = seq(2008, 2016, 3)) -->
-<!-- # ``` -->
-
-\noindent Another fascinating exploration would be to examine if runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
+Another fascinating exploration would be to examine if runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
 For establishing the fielding quality, we apply an indicator function on dismissals (1 if there was at least one wicket in the previous over due to run out or catch, 0 otherwise).
 Runs in the current over is then the observation variable.
-<!--If a batsman is bowled out, it does not necessarily signify good fielding.
-So we only include number of catches and run out in an over as a measure of good fielding.
-Difference in runs across overs are likely to be negative if good fielding has an impact on the runs scored in the subsequent overs.
---> Dismissals in the previous over can lead to a batsman adopting a more defensive play style.
+Dismissals in the previous over can lead to a batsman adopting a more defensive play style.
 Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over as compared to the case when there has been at least one dismissal in the previous over.
@@
 
-<!-- Difference in runs across over should be negative if good fielding has an impact on the runs scored in the subsequent overs.  -->
-
-<!-- fetching raw data since dot and fielding information not available in dismissal type not available in cricket data gravitas. Also `cricket` is aggregated across overs -->
-
 ```{r cricex, fig.cap= "(ref:cricex)", warning = FALSE, message = FALSE, out.width = "90%"}
-
 ```
@@
 
-<!-- Q2: Among good fielding and bowling - which affect the runs of the subsequent overs more? -->
-
-<!-- Q3: Are runs set to reduce in the next over for dot balls in the previous over? -->
-
-<!-- A dot ball is a delivery bowled without any runs scored off it.
-The number of dot balls is reflective of the quality of bowling in the game.
-Run rate of an over should ideally decrease if the number of dot balls increase.
-However, what is the effect of dot balls on runs scored in the subsequent over.
-Will players batsman likely to go for big shots because they couldn't score good runs in the previous over?
-Or they should play consistently and avoid scoring high?
-Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
--->
-<!-- With at least one dot balls per over, the distribution of run rates in facet 1 increase slower compared to that in facet 2.
-This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
--->
-
-<!-- ```{r exdot, fig.cap="25th, 50th, 75th quantiles of runs per over are drawn across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
-For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.", fig.pos="ht"} -->
-
-<!-- cricket_data %>% -->
-<!--   # filter(dot_balls %in% c(0, 1, 2)) %>% -->
-<!--   prob_plot("lag_dot", -->
-<!--     "over", -->
-<!--     hierarchy_model, -->
-<!--     response = "run_rate", -->
-<!--     plot_type = "quantile", -->
-<!--     quantile_prob = c(0.25, 0.5, 0.75), -->
-<!--     symmetric = FALSE -->
-<!--   ) + ggtitle("") -->
-<!-- ``` -->
-
-<!-- CONTENT CUT -->
-<!--While any special event that corresponds to a time domain can be treated as an aperiodic linear granularity in a temporal case, dot balls or wickets that corresponds to certain balls (index) could be treated as aperiodic events in cricket.-->
-
 # Discussion {#sec:discussion}
@@
 
-<!--The infrastructure of systematic exploration of observed variable across pair of cyclic time granularities has been implemented in the
-R package `gravitas`.
---> A missing piece in the package `gravitas` is the  computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
+A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
 A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
@@
 
-The authors would like to thank the cohort [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages and Dr.
-Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application.
+The authors would like to thank the cohort [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience in developing R packages, and Dr Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of the smart meter application.
 The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
 We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
-More details about the package can be found on the package website [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
+More details about the package can be found at [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
 This article was created with `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
-This paper's Github repository, [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-
-<!-- # Appendix -->
-
-<!-- ```{r granadvice, echo = TRUE, warning=FALSE, message=FALSE} -->
-<!-- smart_meter10 %>% -->
-<!--   filter(customer_id %in% c(10017936)) %>% -->
-<!--   gran_advice( -->
-<!--     "wknd_wday", -->
-<!--     "hour_day" -->
-<!--   ) -->
-<!-- ``` -->
-
-<!-- Q3: Are runs set to reduce in the next over for dot balls in the previous over? -->
-
-<!-- A dot ball is a delivery bowled without any runs scored off it.
-The number of dot balls is reflective of the quality of bowling in the game.
-Run rate of an over should ideally decrease if the number of dot balls increase.
-However, what is the effect of dot balls on runs scored in the subsequent over.
-Will players batsman likely to go for big shots because they couldn't score good runs in the previous over?
-Or they should play consistently and avoid scoring high?
-Figure \ref{fig:exdot} shows the quantile plot of runs across overs for at least one dot ball per over (facet 1) or no dot balls per over (facet 2).
--->
-<!-- With at least one dot balls per over, the distribution of run rates in facet 1 increase slower compared to that in facet 2.
-This implies that run rates are likely to decrease in the subsequent over as a result of dot balls in the previous over.
--->
-
-<!-- ```{r exdot, fig.cap="25th, 50th, 75th quantiles of runs per over are drawn across overs of the innings with no (facet 2), more than zero (facet 1) dot balls per over.
-For all quantiles, run rates mostly increase at a higher rate in facet 2 compared to facet 1 implying run rates decrease with at least one dot ball in the previous over.", fig.pos="ht"} -->
-
-<!-- cricket_data %>% -->
-<!--   # filter(dot_balls %in% c(0, 1, 2)) %>% -->
-<!--   prob_plot("lag_dot", -->
-<!--     "over", -->
-<!--     hierarchy_model, -->
-<!--     response = "run_rate", -->
-<!--     plot_type = "quantile", -->
-<!--     quantile_prob = c(0.25, 0.5, 0.75), -->
-<!--     symmetric = FALSE -->
-<!--   ) + ggtitle("") -->
-<!-- ``` -->
-
-# Supplemental Materials
+The Github repository, [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+
+# Supplementary Materials
 

### 2020-08-27 main.Rmd
Minor wording improvements in Section 1

@@
 Temporal data are available at various resolutions depending on the context.
-Social and economic data like GDP is often collected and reported at coarse temporal scales such as monthly, quarterly or annually.
+Social and economic data are often collected and reported at coarse temporal scales such as monthly, quarterly or annually.
 With recent advancement in technology, more and more data are recorded at much finer temporal scales.
@@
 It is important to be able to navigate through all of these time granularities to have multiple perspectives on the periodicity of the observed data.
-This idea aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+This aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
 Visualizing probability distributions conditional on one or more granularities is an indispensable tool for exploration.
 Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly becomes overwhelming.
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
 
@@
 
-Linear time granularities respect the linear progression of time and are non-repeating such as hours, days, weeks and months.
+Linear time granularities (such as hours, days, weeks and months) respect the linear progression of time and are non-repeating.
 One of the first attempts to characterize these granularities is due to @Bettini1998-ed.
-However, the definitions and rules defined are inadequate for describing cyclic or repeating granularities.
-Hence, there is a need to define some new cyclic time granularities, that can be useful in visualizations.
+However, the definitions and rules defined are inadequate for describing non-linear granularities.
+Hence, there is a need to define some new time granularities, that can be useful in visualizations.
 Cyclic time granularities can be circular, quasi-circular or aperiodic.
@@
 The motivation for this work stems from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
-Smart meters currently provide half-hourly use in kWh for each household, from the time that they were installed, some as early as 2012.
+Smart meters currently provide half-hourly use in kWh for each household, from the time they were installed, some as early as 2012.
 Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+The behavioral patterns in households vary substantially; for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
 It is common to see aggregates [see @Goodwin_2012] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
 But studying overall energy use hides the distribution of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.
-We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns/anomalies.
-Although, the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.
+We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns and anomalies.
+Although we were motivated by the smart meter example, the problem and the solutions we propose are relevant to any temporal data observed more than once per year.
 
@@
   * develop an approach to check the feasibility of creating plots or drawing inferences for any two cyclic granularities;
-  * recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pair of granularities.
+  * recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pairs of granularities.
 
-The remainder of the paper is organized as follows: Section \ref{sec:linear-time} provides some background material on linear granularities and calendar algebra for computing different linear granularities.
-Section \ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities.
-Section \ref{sec:cyclic-calendar} introduces cyclic calendar algebra for computing cyclic time granularities.
-Section \ref{sec:data-structure} discusses the data structure for exploring the conditional distributions of the associated time series across pairs of cyclic time granularities.
-Section \ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
-Section \ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application.
-Section \ref{sec:discussion} summarizes this paper and discusses possible future direction.
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:linear-time} provides some background material on linear granularities and calendar algebra for computing different linear granularities.
+Section&nbsp;\ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities, and introducing cyclic calendar algebra for computing cyclic time granularities.
+The data structure for exploring the conditional distributions of the associated time series across pairs of cyclic time granularities is discussed in Section&nbsp;\ref{sec:data-structure}.
+Section&nbsp;\ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
+Section&nbsp;\ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application.
+Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
 

### 2020-08-28 main.Rmd
Reduced and improved Section 2

@@
 
-Discrete abstraction of time such as weeks, months or holidays can be thought of as "time granularities".
+Discrete abstractions of time such as weeks, months or holidays can be thought of as "time granularities".
 Time granularities are **linear** if they respect the linear progression of time.
@@
 \begin{definition}\label{def:index set}
-There is a unique \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, which map the time instants to a set of non-negative integers. Often, this is thought of as $t=0, \dots, T$.
+The \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, uniquely maps the time instants to the set of non-negative integers.
 \end{definition}
@@
 \begin{definition}\label{def:linear}
-A \textbf{linear granularity} is a mapping $G$ from the integers (the index set, $Z$) to subsets of the time domain such that:
+A \textbf{linear granularity} is a mapping $G$ from the index set, $Z$, to subsets of the time domain such that:
   (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less than all elements of $G(j)$; and
@@
 
-\noindent This implies that the granules in a linear granularity are non-overlapping, continuous and ordering is maintained.
-The indexing for each granule can also be associated with textual representation, called the label.
+\noindent This implies that the granules in a linear granularity are non-overlapping, continuous and ordered.
+The indexing for each granule can also be associated with a textual representation, called the label.
 A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
-\autoref{fig:linear-time} illustrates the linear time granularities.
-Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities which maps from index set to subsets of the hourly time domain.
+\autoref{fig:linear-time} illustrates some common linear time granularities.
+Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities formed by mapping the index set to subsets of the hourly time domain.
 If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
-These linear granularities are ordered with ordering guided by the index set which is a set of integers.
-Hence, they are uni-directional and non-repeating.
+These linear granularities are uni-directional and non-repeating.
 
@@
 \noindent For example, both $day \trianglelefteq week$ and $day \preceq week$ hold, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
-Consider another example where $day \trianglelefteq month$.
-This relationship however is incomplete without its association to periodicity.
-Each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year, if leap years are ignored.
-This grouping period becomes 4 and 400 years with the inclusion of leap years and leap seconds respectively.
+The relationship has period 7.
+
+The relationship $day \trianglelefteq month$ has a more complicated period.
+If leap years are ignored, each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year.
+With the inclusion of leap years, the grouping period is 400 years.
 
@@
 
-If ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
-$H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$.
-Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way, called **quasi-periodic** relationships by @Bettini2000-vy.
+For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$.
 
-For example, in a Gregorian calendar without leap years we could say $day$ groups periodically into month with the period $P = 365$ and the number of granules of $month$ in each period given by $R=12$.
-In a Gregorian calendar with leap years, $day$ groups quasi-periodically into $month$ with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
+Granularities can also be periodical with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
 
@@
 For example, if the bottom granularity is day, then granularity week will have lower order than month since each week consist of fewer days than each month.
-Here neither week *groups into* month or week *finer than* month, but their relative order can be determined.
 
-## Computation
-
-Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity.
+Granules in any granularity may be aggregated to form a coarser granularity.
 A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson_2000.
-Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning_2002].
-It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one.
-Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and quasi-circular granularities.
-
-\begin{definition}\label{def:grouping_operation}
-Let $G_1$ be a full-integer labelled granularity, and $m$ a positive integer.
-\textbf{The grouping operation} Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity.
-More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labelled granularity such that for each integer $i$,
-$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
-The grouping operation Group$_m(G)$, provides a new granularity whose granules are the integer part of the quotient when the integer set spanned by linear granularity $G$ is divided by $m$, such that, $G_2(i) = \lfloor G_1(j)/m \rfloor$.
-\end{definition}
-
-\begin{example}
-{\rm Due to even length of $day$ and $week$, we can derive them from $hour$ using the grouping operation as follows: day = G$_{24}(\text{hour})$, week = G$_{24*7}(\text{hour})$.}
-\end{example}
-
-\noindent For more variations of calendar algebra operations, see @Ning_2002.
+Linear time granularities are computed through "calendar algebra" operations [@Ning_2002] designed to generate new granularities recursively from the bottom granularity.
+For example, due to the constant length of day and week, we can derive them from hour using
+$$
+  D(j) = \lfloor H(i)/24\rfloor, \qquad W(k) = \lfloor H(i)/(24*7)\rfloor,
+$$
+where $H$, $D$ and $W$ denote hours, days and weeks respectively.
 

### 2020-08-28 main.Rmd
Working on Section 3

@@
   \newtheorem{definition}{Definition}
-  \newtheorem*{example}{Example}
   \usepackage{mathptmx}
@@
 They can be thought of as additional categorizations of time that are not linear.
-Cyclic granularities can be constructed by operating on two linear granularities.
-Cycles can be either _regular_, called **circular**, or _irregular_, **quasi-circular** when these two linear granularities relate periodically.
+Cyclic granularities can be constructed from two linear granularities, that relate periodically; the resulting cycles can be either _regular_ (**circular**), or _irregular_ (**quasi-circular**).
 
@@
 \begin{definition}\label{def:circular}
-A \textbf{circular granularity} $C_{B, G}$ relates a linear granularity $G$ to the bottom granularity B, if
+A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$, if
 \begin{equation} \label{eq:circular-gran}
@@
 where
-z denotes the index set,
-B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, and $P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$, also called the period of the grouping $(B, G)$.
+$z$ denotes the index set,
+$B$ groups periodically into $G$ with regular mapping and period $P(B, G)$.
 \end{definition}
-
-For example, \autoref{fig:circular-dow} illustrates the linear and the corresponding cyclical granularities.
-Cyclical granularities can be considered to be cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
-$B, G, H$ (day, week, fortnight, respectively) are linear granularities.
-The circular granularity $C_{B , G}$ (day-of-week)is constructed from $B$ and $G$.
-The circular granularity $C_{B, H}$ (day-of-fortnight) is constructed from $B$ and $H$.
-These overlapping cyclical granularities share elements from the linear granularity.
-Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns $\{0, 1, \dots, 6\}$ and $\{0, 1, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
-Each circular granularity can use descriptive label mappings.
-Suppose ${L}$ is a label mapping that defines an unique label for each index $\ell \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined (for example) as
-\begin{align*}
-  L: \{0,1, \dots, 6\} & \longmapsto\ \{\text{Sun}, \text{Mon}, \dots, \text{Sat}\}\\
-\text{or}\qquad
-  L: \{0,1, \dots, 6\} & \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}.
-\end{align*}
 
@@
 
-(ref:circular-dow) Illustration of circular relative to linear granularities (a).
-Circular granularities can be considered to be cutting the linear granularity into pieces and stacking them (b).
-The circular granularity creates repeated integer sequences.
+(ref:circular-dow) a. Index sets for some linear and circular granularities.  b. Circular granularities can be constructed by slicing the linear granularity into pieces and stacking them.
 
-In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H)$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping and $P(G,H)$ is the period of the grouping $(G, H)$.
-Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with period $P_i$ and minutes as the bottom granularity.
+\autoref{fig:circular-dow} illustrates some linear and cyclical granularities.
+Cyclical granularities are constructed be cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
+$B, G, H$ (day, week, fortnight, respectively) are linear granularities.
+The circular granularity $C_{B, G}$ (day-of-week) is constructed from $B$ and $G$, while  circular granularity $C_{B, H}$ (day-of-fortnight) is constructed from $B$ and $H$.
+These overlapping cyclical granularities share elements from the linear granularity.
+Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns $\{0, 1, \dots, 6\}$ and $\{0, 1, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
+
+Suppose ${L}$ is a label mapping that defines a unique label for each index $\ell \in \{ 0,1,\dots, (P-1)\}$. For example, the label mapping $L$ for $C_{B, G}$ can be defined as
+$$
+  L: \{0,1, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}.
+$$
+
+In general, any circular granularity relating two linear granularities can be expressed as
+$$
+  C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H),
+$$
+where $H$ is periodic with respect to $G$ with regular mapping and period $P(G,H)$.
+Table&nbsp;\ref{tab:definitions} shows several circular granularities constructed using minutes as the bottom granularity.
 
@@
 \toprule
-circular granularity & expression & period \\
+Circular granularity & Expression & Period \\
 \midrule
@@
 
-A **quasi-circular** granularity can not be defined using modular arithmetic since they are formed using two linear granularities with irregular mapping.
-However, they are still formed with linear granularities, one of which "groups periodically into" the other.
-\autoref{tab:quasi} shows some example of quasi-circular granularities ($Q_i$) with ($P_i$) denoting the plausible choices of period of the grouping of two linear granularities.
+A **quasi-circular** granularity cannot be defined using modular arithmetic  because of the irregular mapping.
+However, they are still formed with linear granularities, one of which groups periodically into the other.
+\autoref{tab:quasi} shows some examples of quasi-circular granularities.
 
 \begin{table}[ht]
-\begin{center}
+\centering
 \begin{tabular}{lr@{~}lr@{~}r}
 \toprule
-quasi-circular granularity && potential period lengths\\
+Quasi-circular granularity && Possible period lengths\\
 \midrule
 $Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\
-$Q_2 =$ hour-of-month && $P_2 = 24*31, 24*30, 24*29, 24*28$\\
+$Q_2 =$ hour-of-month && $P_2 = 24\times 31, 24\times 30, 24\times 29, 24\times 28$\\
 $Q_3 =$ day-of-year && $P_3 = 366, 365$\\
@@
 \end{tabular}
-\end{center}
-\caption{Examples of quasi-circular granularities $Q_i$ with potential period lengths $P_i$.
-These quasi-circular granularities relate two linear granularities one of which groups periodically into the other with irregular mapping leading to many potential period lengths and hence can not be expressed through modular arithmetic.
-}
+\caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.}
 \label{tab:quasi}
@@
 \begin{definition}\label{def:quasicircular}
-A {\bf quasi-circular granularity} $Q_{B, G'}$ relates linear granularities $G'$ and bottom granularity $B$, if
+A \textbf{quasi-circular granularity} $Q_{B, G'}$ relates linear granularity $G'$ and bottom granularity $B$ if
 \begin{equation}\label{eq:quasi}
-\begin{split}
 Q_{B, G'}(z) =
-z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, & \quad \text{for}\quad z \in T_{k}
-\end{split}
+z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, \quad \text{for}\quad z \in T_{k},
 \end{equation}
-where,
-$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping, $P'$ and $R'$ are the period of the grouping $(B, G')$ and the number of granules of $G'$ in $P'$, $T_w$ are the sets of indices of $B$ describing $G'(w)$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+where
+$z$ denotes the index set,
+$B$ groups periodically into $G'$ with irregular mapping,
+$P'$ and $R'$ are the period of the grouping $(B, G')$ and the number of granules of $G'$ in $P'$,
+$T_w$ are the sets of indices of $B$ describing $G'(w)$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
+and $\vert T_w \vert$ is the cardinality of set $T_w$.
 \end{definition}
 
-Consider $G'$ such that every two consecutive granules of $G'$ are made up of $7$ and $5$ granules of $B$ respectively within each period of the grouping ($B$, $G'$).
-Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$.
+For example, suppose consecutive granules of $G'$ are alternately made up of $7$ and $5$ granules of $B$ within each period of the grouping ($B$, $G'$).
+Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
 Here, $T_0 = \{0, 1, \dots, 6\}$ and $T_1 = \{7, \dots 11\}$.
@@
 
-If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical}) Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
+If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'} B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical}).
+Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
 The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
@@
 
-\begin{example}
-{\rm Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
+Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
 Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
@@
 \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}.
-It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.}
-\end{example}
+It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
 
@@
 \begin{definition}\label{def:norderup}
-The hierarchy table contains {\bf multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
-A {\bf single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
+The hierarchy table contains \textbf{multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
+A \textbf{single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
 \end{definition}
 
-\begin{example}
-{\rm In \autoref{tab:tab-mayan}, kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.}
-\end{example}
+In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
 
@@
 
-Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section \ref{sec:summarize-measured}).
+Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section&nbsp;\ref{sec:summarize-measured}).
 Even when there are no rarely occurring events, number of observations might vary hugely within or across each facet.
@@
 Further, we add cyclic granularity day-type( "wknd_wday") to capture weekend and weekday behavior.
-Now that we have a list of cyclic granularities to look at, we should be able to compute them using Section \ref{sec:cyclic-calendar}.
+Now that we have a list of cyclic granularities to look at, we should be able to compute them using Section&nbsp;\ref{sec:cyclic-calendar}.
 
@@
 Harmonies can be identified among those $42$ possibilities to narrow the search.
-\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective exploration becomes difficult with many levels (Section \ref{sec:levels}).
+\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:levels}).
+
+```{r harmony-tab, echo=FALSE}
+```
+
+```{r bothcust, fig.cap = "(ref:bothcust)", out.width="90%", fig.pos="!htb"}
+```
+
+(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
+Two granularities are used: hour of the day (I) and weekday/weekend (II).
+Plot (a) shows granularity I faceted by granularity II, and plots (b), (c) shows the converse mapping.
+Plot (a) makes a comparison of usage by workday within each hour of the day using side-by-side boxplots.
+Generally, on a work day there is more consumption early in the day.
+Plots (b) and (c) examine the temporal trend of consumption over the course of a day, separately for the type of day.
+Plot (b) uses an area quantile to put the emphasis on the time series, for example, the median consumption over time shows prolonged usage in the morning on weekdays.
+Plot (c) uses a violin plot to place emphasis on distributional differences across hours.
+It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.
 
@@
 All of these indicate that looking at different distribution geometry or changing the mapping might shed lights on different aspect of the energy behavior for the same sample population.
-
-```{r harmony-tab, echo=FALSE}
-```
-
-```{r bothcust, fig.cap = "(ref:bothcust)", out.width="90%", fig.pos="!htb"}
-```
-
-(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
-Two granularities are used: hour of the day (I) and weekday/weekend (II).
-Plot (a) shows granularity I faceted by granularity II, and plots (b), (c) shows the converse mapping.
-Plot (a) makes a comparison of usage by workday within each hour of the day using side-by-side boxplots.
-Generally, on a work day there is more consumption early in the day.
-Plots (b) and (c) examine the temporal trend of consumption over the course of a day, separately for the type of day.
-Plot (b) uses an area quantile to put the emphasis on the time series, for example, the median consumption over time shows prolonged usage in the morning on weekdays.
-Plot (c) uses a violin plot to place emphasis on distributional differences across hours.
-It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.
 

### 2020-08-28 scripts/main.R

@@ -165,7 +165,7 @@ harmonies %>%
     caption = "Harmonies with a pair of cyclic granularity one placed on facet and the other on x-axis. Out of 42 possible combinations of cyclic granularities, only 16 are harmony pairs."
   ) %>%
   # row_spec(0, bold = TRUE) %>%
-  kable_styling()
+  kable_styling(latex_options = "hold_position")
 
 ## ----bothcust
 cust2_quantile <- smart_meter10 %>%

### 2020-08-28 main.Rmd
Some questions

@@
 preamble: >
-  \usepackage{mathtools,amssymb,booktabs,amsthm}
+  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes}
   \def\mod{~\text{mod}~}
@@
 \begin{definition}\label{def:quasicircular}
-A \textbf{quasi-circular granularity} $Q_{B, G'}$ relates linear granularity $G'$ and bottom granularity $B$ if
+A \textbf{quasi-circular granularity} $Q_{B, G'}$ relates linear granularity $G'$ and bottom granularity $B$ if\todo{This is not a condition. What is $Q$ used for?}
 \begin{equation}\label{eq:quasi}
@@
 $B$ groups periodically into $G'$ with irregular mapping,
-$P'$ and $R'$ are the period of the grouping $(B, G')$ and the number of granules of $G'$ in $P'$,
-$T_w$ are the sets of indices of $B$ describing $G'(w)$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
+$R'$ is the number of granules of $G'$ in each repetition of the grouping,
+$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
 and $\vert T_w \vert$ is the cardinality of set $T_w$.
@@
 
-For example, suppose consecutive granules of $G'$ are alternately made up of $7$ and $5$ granules of $B$ within each period of the grouping ($B$, $G'$).
-Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
+For example, suppose consecutive granules of $G'$ are alternately made up of $7$ and $5$ granules of $B$ within each period, so that $R'=2$ Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
 Here, $T_0 = \{0, 1, \dots, 6\}$ and $T_1 = \{7, \dots 11\}$.
@@
 $$
-  Q_{B,H'}(10)  = 10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert =3
+  Q_{B,H'}(10)  = 10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert = 10 - 7 = 3
 $$
@@
 
-Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules.
+Aperiodic time granularities are those that can not be specified as a periodic repetition of a pattern of granules.
 Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant.
@@
 \begin{definition}\label{def:aperiodic}
-An \textbf{aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if
+An \textbf{aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if\todo{This is not a condition. What is $A$ used for?}
 \begin{equation}\label{eq:aperiodic}
@@
 \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}.
-It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
+Here, $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
 
@@
 
-It is interesting to note, that in the Mayan calendar, like most calendars, that day is the basic unit of time [@Reingold2001-kf].
-The structuring of larger units, weeks, months, years and cycle of years, though, varies substantially.
-An exception is the French revolutionary calendar, which divides each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
+Like most calendars, the Mayan calender used the day as the basic unit of time [@Reingold2001-kf].
+The structuring of larger units, weeks, months, years and cycle of years, though, varies substantially between calendars.
+For example,  the French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
 Nevertheless, for any calendar a hierarchy table can be defined.
 Note that it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
-Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(\ell),G(m)}$ is either circular or quasi-circular.
+Hence, we assume that the hierarchy table consists of periodic linear granularities only, and that the cyclic granularity $C_{G(\ell),G(m)}$ is either circular or quasi-circular.
 
@@
 
-In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
+\noindent In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
 
@@
 
-In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities.
-This is similar to calendar algebra in @Ning_2002 for linear granularities but caters to computation of cyclic granularities.
-We shall refer to it as through cyclic calendar algebra operations.
-The cyclic calendar algebra consists of broadly two kinds of operations:
-(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively.
+Following the calendar algebra of @Ning_2002 for linear granularities, we can define cyclic calendar algebra to compute cyclic granularities. Cyclic calendar algebra comprises two kinds of operations:
+(1) **single-to-multiple** (the calculation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities) and (2) **multiple-to-single** (the reverse).
 

### 2020-08-31 main.Rmd
Working on Section 3

@@
 - calendar algebra
-- periodicties
+- periodicities
 - grammar of graphics
@@
 \begin{definition}\label{def:quasicircular}
-A \textbf{quasi-circular granularity} $Q_{B, G'}$ relates linear granularity $G'$ and bottom granularity $B$ if\todo{This is not a condition. What is $Q$ used for?}
+A \textbf{quasi-circular granularity} $Q_{B, G'}$ is formed when bottom granularity $B$ groups periodically into linear granularity $G'$ with irregular mapping such that the granularities are given by
 \begin{equation}\label{eq:quasi}
@@
 $z$ denotes the index set,
-$B$ groups periodically into $G'$ with irregular mapping,
 $R'$ is the number of granules of $G'$ in each repetition of the grouping,
@@
 
-For example, suppose consecutive granules of $G'$ are alternately made up of $7$ and $5$ granules of $B$ within each period, so that $R'=2$ Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
-Here, $T_0 = \{0, 1, \dots, 6\}$ and $T_1 = \{7, \dots 11\}$.
-Hence using Definition \autoref{def:quasicircular} we will have:
-$$
-  Q_{B,H'}(10)  = 10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert = 10 - 7 = 3
-$$
-since $10 \in T_{1}$.
-
-If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'} B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical}).
-Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
-The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
-The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$.
-The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ in Definition \autoref{def:quasicircular} denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$.
-Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.
-
-```{r quasi-circular-example, echo = FALSE, out.width = "100%",fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
-The linear display shows linear granularities days, weeks, semester weeks, semester Terminology and intuition of quasi-circular granularities are shown.
-Panel a shows the time domain distributed as linear granularities $B$ (bottom granularity), $G$, $H'$ and quasi-circular granularities $Q(B, G)$ and $Q(B, H')$.
-Both $Q(B, G)$ and $Q(B, H')$ have repeating patterns unlike $G$ or $H'$ and the patterns are guided by the distribution of $B$ within $G$ and $H'$ and remains constant within a period shown by the hashed arrows.
-Panel b depicts the intuition through a measuring tape analogy where the folds are not of same length but has same length over a period.", fig.pos="!htb"}
-```
+For example, day-of-year is quasi-periodic with either 365 or 366 granules of $B$ (days) within each period of $G'$ (years).
+The pattern repeats every $R'=400$ years.
+So $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
 
@@
 
-Aperiodic time granularities are those that can not be specified as a periodic repetition of a pattern of granules.
+Aperiodic time granularities are those that cannot be specified as a periodic repetition of a pattern of granules.
 Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant.
-A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.
+A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years.
 In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
@@
 All of these are recurring events, but with non-periodic patterns.
-As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.
+Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.
 
-Aperiodic cyclic granularities are defined using aperiodic linear granularities.
-Consider $n$ aperiodic linear granularities $M_i$,  $i \in \{1, 2, \dots, n\}$, such that Definition \ref{def:periodical} does not hold true with respect to $B$.
-However, $B \trianglelefteq M_i$.
-Then according to Definition \ref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$.
-Suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.
-Here, index ${\{i_j\}}$ stands for the $j^{\text{th}}$ granule of the $i^{\text{th}}$ linear aperiodic granularity.
+\begin{definition}\label{def:aperiodic}
+An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups aperiodically into linear granularity $M$ such that the granularities are given by
+\begin{equation}\label{eq:aperiodic}
+A_{B, M}(z) = \begin{cases}
+                  i, & \text{for}\quad z \in T_{i_j} \\
+                  0  & \text{otherwise},
+                \end{cases}
+\end{equation}
+where
+$z$ denotes the index set,
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
+\end{definition}
 
@@
 
-(ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time.
-The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future.
-Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks.
-This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 18.3 weeks will be a quasi-circular granularity with repeating patterns.
+(ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated through (a) linear and (b) stacked displays of time.
+The linear display shows granularities days, weeks, semester weeks, semester week type distributed over linear time.
+Here a semester lasts for 18 weeks and 2 days, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week, an in-session period of 7 weeks, a 1-week study break, before final exams which continue for 16 days.
+This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 128 days will be a quasi-circular granularity with repeating patterns.
 $A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years.
 
-\begin{definition}\label{def:aperiodic}
-An \textbf{aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if\todo{This is not a condition. What is $A$ used for?}
-\begin{equation}\label{eq:aperiodic}
-\begin{split}
-A_{B, M}(z) & =
-\begin{cases}
-i, & \text{for}\quad z \in T_{i_j} \\
-0  & \text{otherwise}
-\end{cases}
-\end{split}
-\end{equation}
-where,
-$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
-$B$ denotes a full-integer labelled bottom granularity which groups into linear granularity $M$ but not periodically, $T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
-\end{definition}
-
-Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 16 days.
-Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively.
-Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$).
-Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity.
-This is because the placement of the semester within an year would vary across years.
-\autoref{fig:aperiodic-example} b shows the stack display of time with granules representing same categories (semester break/in-session) stacked on top of one another.
-\autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}.
-Here, $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
+For example, consider the school semester shown in \autoref{fig:aperiodic-example}.
+Let the linear granularities $M_1$ and $M_2$ denote the in-session semester period and semester break period respectively.
+Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the "semester week type" are aperiodic with respect to days ($B$) or weeks ($G$).
+Hence $A_{B, M}$ denoting day-of-the-"semester week type" would be an aperiodic cyclic granularity, because the placement of the semester within an year would vary across years.
+Here, $Q_{H, M}$ denoting week-of-the-"semester week type" would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
 
@@
 The hierarchical structure of time creates a natural nested ordering which can be used in the computation of relative pairs of granularities.
-This is illustrated using two examples, a Mayan calendar, which is simple, in addition to the more complex Gregorian calendar.
 
@@
 The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order.
-It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$.
-can be defined relative to the linear granularities.
+It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$, can be defined relative to the linear granularities.
 Let $G_{\ell}$ and $G_{m}$ represent the linear granularity of order $\ell$ and $m$ respectively with $\ell<m$.
@@
 For example, \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar.
-In the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun.
+In the Mayan calendar, one day was referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun.
 
@@
 
-Methods to obtain multiple-order-up granularity will be different based on if the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
-The methods require the use of integer arithmetic, and hence it is important that the label mapping of the individual single-order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$.
+Methods to obtain multiple-order-up granularity will depend on whether the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
+The methods require the use of integer arithmetic, and hence it is important that the label mapping of the individual single-order-up granularity is an identity function, $L(x) = x$ for all $x$.
 The label mapping of the resultant multiple-order-up granularity can be chosen depending on the context.
@@
 We consider the case of only one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
-Any multiple-order-up quasi-circular granularity $C_{\ell, m}(z)$ could then be obtained as a dicrete combination of circular and quasi-circular granularities.
+Any multiple-order-up quasi-circular granularity $C_{\ell, m}(z)$ could then be obtained as a discrete combination of circular and quasi-circular granularities.
 
@@
 
-Consider a hierarchy using linear granularties from a Gregorian calendar in \autoref{tab:tab-gregorian}.
+Consider a hierarchy using linear granularities from a Gregorian calendar in \autoref{tab:tab-gregorian}.
 This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities with day-of-month as the only single-order-up quasi-circular granularity.
@@
 It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
-However, this approach hides the distributions of the measured variable induced by the re-organised data structure.
+However, this approach hides the distributions of the measured variable induced by the re-organized data structure.
 Summarizing the distribution of the measured variable using these multiple observations could be a potential way to explore and bring forward different features of the data.

### 2020-08-31 main.Rmd
Finished Section 3

@@
 
-Like most calendars, the Mayan calender used the day as the basic unit of time [@Reingold2001-kf].
+Like most calendars, the Mayan calendar used the day as the basic unit of time [@Reingold2001-kf].
 The structuring of larger units, weeks, months, years and cycle of years, though, varies substantially between calendars.
@@
 Methods to obtain multiple-order-up granularity will depend on whether the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
-The methods require the use of integer arithmetic, and hence it is important that the label mapping of the individual single-order-up granularity is an identity function, $L(x) = x$ for all $x$.
-The label mapping of the resultant multiple-order-up granularity can be chosen depending on the context.
-Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using Equation \ref{eq:eq7} for $\ell < m - 1$ and $P(i, j) = 1, \quad \forall i = j \in \{0, 1, \dots, m-\ell-1\}$.
+Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using
 \begin{equation} \label{eq:eq7}
-\begin{split}
-C_{G_\ell,G_m}(z) & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)C_{G_{\ell+1},G_m}(z)\\
-  & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)( C_{G_{\ell+1},G_{\ell+2}}(z) + P(\ell+1, \ell+2)C_{G_{\ell+2}, G_m}(z)) \\
-  & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)C_{G_{\ell+1},G_{\ell+2}}(z) + P(\ell, \ell+1)P(\ell+1, \ell+2)C_{G_{\ell+2}, G_{m}}\\
-  & = C_{G_\ell,G_{\ell+1}}(z) + P(\ell, \ell+1)C_{G_{\ell+1},G_{\ell+2}}(z) + P(\ell, \ell+2)C_{G_{\ell+2},G_{\ell+m}}(z)\\
-  &\vdots\\
-  & = \sum_{i=0}^{m - \ell - 1} P(\ell, \ell+i)C_{G_{\ell+i},G_{\ell+i+1}}(z)\\
-\end{split}
+C_{G_\ell,G_m}(z)
+  = \sum_{i=0}^{m - \ell - 1} P(\ell, \ell+i)C_{G_{\ell+i},G_{\ell+i+1}}(z),
 \end{equation}
-For example, the multiple-order-up granularity $C_{uinal,katun}$ for Mayan calendar could be obtained using Equation \ref{eq:eq8}.
-\begin{equation} \label{eq:eq8}
-\begin{split}
-C_{uinal, baktun}(z)  = & C_{uinal, tun}(z) + P(uinal, tun)C_{tun,katun}(z) + P(uinal, katun)C_{katun, baktun}(z) \\
-               = & \lfloor z/20\rfloor \mod 18 + 18*\lfloor z/20*18\rfloor \mod 20  \\
-               &~~~ + 18*20\lfloor z/20*18*20\rfloor \mod 20 \\
-\end{split}
-\end{equation}
-We consider the case of only one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
+where $\ell < m - 1$ and $P(i, i) = 1$ for $i=0,1,\dots,m-\ell-1$, and
+$C_{B, G}(z)  = z\mod P(B, G)$ as per Equation \eqref{eq:circular-gran}.
+
+For example, the multiple-order-up granularity $C_{\text{uinal},\text{katun}}$ for the Mayan calendar could be obtained using
+\begin{align*}
+C_{\text{uinal}, \text{baktun}}(z)  &=  C_{\text{uinal}, \text{tun}}(z) + P(\text{uinal}, \text{tun})C_{\text{tun},\text{katun}}(z) + P(\text{uinal},\text{katun})C_{\text{katun}, \text{baktun}}(z) \\
+               &=  \lfloor z/20\rfloor \mod 18 + 18\lfloor 18z/20\rfloor \mod 20
+                + 18\times20\lfloor 18\times20\times z/20\rfloor \mod 20.
+\end{align*}
+
+Now consider the case where there is one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
 Any multiple-order-up quasi-circular granularity $C_{\ell, m}(z)$ could then be obtained as a discrete combination of circular and quasi-circular granularities.
 
-Depending on the order of the combination, again two different approaches need to be employed leading to the following cases:
+Depending on the order of the combination, two different approaches need to be employed leading to the following cases:
 
@@
 
-Consider a hierarchy using linear granularities from a Gregorian calendar in \autoref{tab:tab-gregorian}.
-This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities with day-of-month as the only single-order-up quasi-circular granularity.
+```{r tab-gregorian}
+```
 
-Using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}, we then have:
+For example, the Gregorian calendar (\autoref{tab:tab-gregorian}) has day-of-month as a single-order-up quasi-circular granularity, with the other granularities being circular.
+Using Equations \eqref{eq:multifromsingle-quasi1} and \eqref{eq:multifromsingle-quasi2}, we then have:
 $$
@@
 
-```{r tab-gregorian}
-```
-
 ### Multiple-to-single {#sec:multiple-to-single}
@@
 Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
-For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities in the hierarchy and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities could be obtained using \ref{eq:all-circular-multiple}.
+For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities can be obtained using \eqref{eq:all-circular-multiple}.
 \begin{equation} \label{eq:all-circular-multiple}
@@
 \end{equation}
-For example, in Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) = \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.
+For example, in the Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) = \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.
 
@@
 
-Single-order-up quasi-circular granularity could be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}.
+Single-order-up quasi-circular granularity can be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \eqref{eq:multifromsingle-quasi1} and \eqref{eq:multifromsingle-quasi2}.
 
@@
 
-Restructuring time from linear to cyclic time granularities leads to re-organisation of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
+Restructuring time from linear to cyclic time granularities leads to re-organization of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
 It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
@@
 The idea of cyclic granularities can be likewise mapped to this hierarchy.
-Example granularites then include ball of the over, over of the innings and ball of the innings.
+Example granularities then include ball of the over, over of the innings and ball of the innings.
 Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularities are aperiodic.
@@
 
-# Acknowledgements {-}
+# Acknowledgments {-}
 

### 2020-08-31 main.Rmd
Working on section 4

@@
 preamble: >
-  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes}
+  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
   \def\mod{~\text{mod}~}
@@
   \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
+  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
+  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
 output:
@@
 Effective exploration and visualization benefits from well-organized data structures.
-@wang2020tsibble introduced a tidy data structure, tsibble, to support exploration and modeling of temporal data.
+@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data.
 This forms the basis of the structure for cyclic granularities.
-A tsibble consists of an index, key and measured variables.
+A tsibble comprises an index, optional key(s), and measured variables.
 An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
@@
 
-All cyclic granularities can be expressed in terms of the index set, and hence, we introduce the data structure in Figure \@ref(fig:data-structure).
-This basic structure of a tsibble (index, key, measurements) is augmented by columns of cyclic granularities.
-The total number of cyclic granularities would be based on the number of linear granularities considered in the hierarchy table and presence of any aperiodic cyclic granularities, for example, if we have $n$ periodic linear granularities in the hierarchy table, $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.
-Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that could originate from $N_L$ periodic and aperiodic linear granularities.
-Any attempt to encode all or many of these cyclic granularities at the same time to develop insights on periodicity might fail or otherwise become too numerous for comprehensive human consumption.
-Instead, this big problem is broken down by focusing on pair of cyclic granularities ($C_i$, $C_j$) at a time for $i, j \in \{1, 2, \dots, N_C\}$.
-Data sets of the form <$C_i$, $C_j$, $v$> then forms the basis for exploration and analysis of the measured variable $v$.
+\novspacing
 
-```{r data-structure, fig.cap = "(ref:data-structure)", fig.pos="!htb"}
+```{r data-structure, fig.pos="!htb"}
+library(dplyr)
+as.data.frame(matrix(" ", ncol=7, nrow=1)) %>%
+  mutate(
+    V1 = cell_spec(V1, "latex", background="#dbe3f1"),
+    V2 = cell_spec(V2, "latex", background="#dbe3f1"),
+    V3 = cell_spec(V3, "latex", background="#dbe3f1"),
+    V4 = cell_spec(V4, "latex", background="#fdf2d0"),
+    V5 = cell_spec(V5, "latex", background="#fdf2d0"),
+    V6 = cell_spec(V6, "latex", background="#fdf2d0"),
+    V7 = cell_spec(V7, "latex", background="#fdf2d0"),
+  ) %>%
+  knitr::kable(format='latex',
+               booktabs=TRUE,
+               escape=FALSE,
+               caption = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.",
+               col.names = c("\\cellcolor[HTML]{dbe3f1}{index}","\\cellcolor[HTML]{dbe3f1}{key}","\\cellcolor[HTML]{dbe3f1}{measurements}","$C_1$","$C_2$","$\\cdots$","$C_{N_C}$")
+  ) %>%
+  row_spec(0, bold=TRUE,background="#fdf2d0") %>%
+  kable_styling(position = "center")
 ```
 
-(ref:data-structure) The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.
+\vspacing
+
+All cyclic granularities can be expressed in terms of the index set.
+\autoref{tab:data-structure} shows the tsibble structure (index, key, measurements) augmented by columns of cyclic granularities.
+The total number of cyclic granularities depends on the number of linear granularities considered in the hierarchy table and the presence of any aperiodic cyclic granularities.
+For example, if we have $n$ periodic linear granularities in the hierarchy table, then $n(n-1)/2$ circular or quasi-circular cyclic granularities can be constructed.
+Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that can originate from the underlying periodic and aperiodic linear granularities.
+Simultaneously encoding more than a few of these cyclic granularities when visualizing the data overwhelms human comprehension.
+Instead, we focus on visualizing the data split by pairs of cyclic granularities ($C_i$, $C_j$).
+Data sets of the form <$C_i$, $C_j$, $v$> then forms the basis for exploration and analysis of the measured variable $v$.
 
@@
 
-The way cyclic granularities relate become important when we consider the data structure in Figure \@ref(fig:data-structure).
+The way cyclic granularities relate become important when we consider the data structure in \autoref{tab:data-structure}.
 Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k \in \mathbb{N}, k \leq K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell \in \mathbb{N}, \ell \leq L\}$.

### 2020-08-31 scripts/main.R

@@ -51,9 +51,6 @@ table_greg %>%
   # row_spec(0, bold = TRUE) %>%
   kable_styling()
 
-## ----data-structure
-include_graphics("Figs/data-struc-diffcol.png")
-
 ## ----allFig
 VIC <- gravitas::smart_meter10 %>%
   filter(customer_id == "10006486")

### 2020-09-02 main.Rmd
Updated 4.1

@@
 \begin{definition}\label{def:circular}
-A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$, if
+A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$ if
 \begin{equation} \label{eq:circular-gran}
@@
 \begin{equation} \label{eq:multifromsingle-quasi1}
-\begin{split}
-C_{G_\ell,G_{m}}(z) & = C_{G_{\ell},G_{m'}}(z) + P(\ell, m')C_{G_{m'},G_{m}}(z) \\
-\end{split}
+C_{G_\ell,G_{m}}(z) = C_{G_{\ell},G_{m'}}(z) + P(\ell, m')C_{G_{m'},G_{m}}(z)
 \end{equation}
@@
 \begin{equation} \label{eq:multifromsingle-quasi2}
-\begin{split}
-C_{G_\ell,G_{m}}(z) & = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
-\end{split}
+C_{G_\ell,G_{m}}(z)  = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)
 \end{equation}
@@
 Instead, we focus on visualizing the data split by pairs of cyclic granularities ($C_i$, $C_j$).
-Data sets of the form <$C_i$, $C_j$, $v$> then forms the basis for exploration and analysis of the measured variable $v$.
+Data sets of the form <$C_i$, $C_j$, $v$> then allow exploration and analysis of the measured variable $v$.
 
@@
 
-The way cyclic granularities relate become important when we consider the data structure in \autoref{tab:data-structure}.
-Let us consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k \in \mathbb{N}, k \leq K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell \in \mathbb{N}, \ell \leq L\}$.
-Here, $A_k$'s or $B_\ell$'s are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
+The way granularities are related is important when we consider data visualizations.
+Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$.
+Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
 Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$.
-Data subsets for each combination of levels ($A_k$, $B_\ell$) like <$A_k$, $B_\ell$, $v(s)$> can be obtained for all $k \in {1, 2, \dots, K}$ and $\ell \in {1, 2, \dots, L}$ which will lead to $KL$ data subsets.
-Now, some situations can lead to few or many of these sets being empty.
-We will discuss few cases, where one or more of these $KL$ sets will be empty either due to the structure of the calendar, duration and location of events in a calendar or just by the construction of the cyclic granularities.
+There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$).
+Some of these sets may be empty due to the structure of the calendar, or because of the duration and location of events in a calendar, or because of the construction of the cyclic granularities.
 
 \begin{definition}\label{def:harmony}
-A \textbf{clash} is a pair of cyclic granularities which contains structurally, event-driven or build-based empty combinations of its categories.
+A \textbf{clash} is a pair of cyclic granularities that contains structurally, event-driven or build-based empty combinations of categories.
 \end{definition}
@@
 
-First, empty combinations can arise due to the structure of the calendar or hierarchy.
-These are called "structurally" empty combinations.
-Let us take a specific example, where $C_i$ be day-of-month with 31 levels and $C_j$ be week-of-month with 5 levels.
+Structurally empty combinations can arise due to the structure of the calendar or hierarchy.
+For example, let $C_i$ be day-of-month with 31 levels and $C_j$ be week-of-month with 5 levels.
 There will be $31\times 5=155$ sets $S_{k\ell}$ corresponding to possible combinations of $C_i$ and $C_j$.
-Many of these like $S_{1,5}$, $S_{21,2}$ are empty.
-This is also intuitive since the first day of the month can never correspond to fifth week of the month.
+Many of these are empty.
+For example, $S_{1,5}$ is empty because the first day of the month can never correspond to the fifth week of the month.
 Hence the pair (day-of-month, week-of-month) is a clash.
 
-Second, empty combinations can turn up due to differences in event location or duration in a calendar.
-These are called "event-driven" empty combinations.
-Let us consider $C_i$ be day-of-week with 7 levels and $C_j$ be WorkingDay/NonWorkingDay with 2 levels.
+Event-driven empty combinations arise due to differences in event location or duration in a calendar.
+For example, let $C_i$ be day-of-week with 7 levels and $C_j$ be WorkingDay/NonWorkingDay with 2 levels.
 While potentially all of these 14 sets $S_{k\ell}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
@@
 
-Third, empty combinations can be a result of how granularities are constructed.
-These are called "build-based" empty combinations.
-Let $C_i$ be Business-days, which are days from Monday to Friday except holidays and $C_j$ be day-of-month.
-Then the days denoting weekends in a month would not correspond to any Business days.
-This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularities does.
+Build-based empty combinations are a result of how granularities are constructed.
+For example, let $C_i$ be business-days (days from Monday to Friday other than holidays) and let $C_j$ be day-of-month.
+Then the days denoting weekends in a month would not correspond to any business days.
+This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularities does.\todo{This seems like it is related to the calendar structure to me.
+I'm not sure of the distinction.}
 

### 2020-09-02 main.Rmd
Better section numbers/titles

@@
 
-## Circular {#sec:circular-gran-def}
+## Circular granularities {#sec:circular-gran-def}
 
@@
 
-## Quasi-circular {#sec:quasi-circular-gran-def}
+## Quasi-circular granularities {#sec:quasi-circular-gran-def}
 
@@
 
-## Aperiodic {#sec:aperiodic-gran-def}
+## Aperiodic granularities {#sec:aperiodic-gran-def}
 
@@
 
-### Single-to-multiple {#sec:single-to-multiple}
+### Single-to-multiple order-up {-}
 
@@
 
-### Multiple-to-single {#sec:multiple-to-single}
+### Multiple-to-single order-up {-}
 
@@
 
-#### Multiple order-up quasi-circular granularities {#sec:quasi-circular-multiple}
+### Multiple order-up quasi-circular granularities {-}
 
@@
 
-# Data structure{#sec:data-structure}
+# Data structure {#sec:data-structure}
 
@@
 
-### Cyclic granularities search and computation:
+### Cyclic granularities search and computation {-}
 

### 2020-09-02 main.Rmd
Section 4

@@
 There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$).
-Some of these sets may be empty due to the structure of the calendar, or because of the duration and location of events in a calendar, or because of the construction of the cyclic granularities.
+Some of these sets may be empty due to the structure of the calendar, or because of the duration and location of events in a calendar.
+
+\begin{definition}\label{def:clash}
+A \textbf{clash} is a pair of cyclic granularities that contains empty combinations of categories.
+\end{definition}
 
 \begin{definition}\label{def:harmony}
-A \textbf{clash} is a pair of cyclic granularities that contains structurally, event-driven or build-based empty combinations of categories.
-\end{definition}
-
-\begin{definition}\label{def:clash}
 A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
@@
 Event-driven empty combinations arise due to differences in event location or duration in a calendar.
-For example, let $C_i$ be day-of-week with 7 levels and $C_j$ be WorkingDay/NonWorkingDay with 2 levels.
+For example, let $C_i$ be day-of-week with 7 levels and $C_j$ be working-day/non-working-day with 2 levels.
 While potentially all of these 14 sets $S_{k\ell}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
@@
 
-Build-based empty combinations are a result of how granularities are constructed.
-For example, let $C_i$ be business-days (days from Monday to Friday other than holidays) and let $C_j$ be day-of-month.
-Then the days denoting weekends in a month would not correspond to any business days.
-This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularities does.\todo{This seems like it is related to the calendar structure to me.
-I'm not sure of the distinction.}
-
-An example when there will be no empty combinations could be where $C_i$ and $C_j$ maps index set to day-of-week and month-of-year respectively.
-Here $C_i$ can have 7 levels while $C_j$ can have 12 levels.
-So there are $12\times7=84$ sets $S_{k\ell}$.
+An example of a harmony is where $C_i$ and $C_j$ denote day-of-week and month-of-year respectively.
+So $C_i$ will have 7 levels while $C_j$ will have 12 levels, giving $12\times7=84$ sets $S_{k\ell}$.
 All of these are non-empty because every day-of-week can occur in every month.
@@
 
-## Summarizing the measured variable {#sec:summarize-measured}
+## Near-clashes {#sec:near-clashes}
 
-Restructuring time from linear to cyclic time granularities leads to re-organization of the data structure, where each level of a cyclic granularity corresponds to multiple values of the measured variable.
-It is common to see summarization of these multiple values through aggregation or an unique summary statistic like mean or median to have each level of cyclic granularity correspond to an unique value of the measured variable.
-However, this approach hides the distributions of the measured variable induced by the re-organized data structure.
-Summarizing the distribution of the measured variable using these multiple observations could be a potential way to explore and bring forward different features of the data.
-
-However, we need to consider the effect of number of observations on the summarization even for harmonies.
-Consider a data set with $(T + 1)$ observations, and two cyclic granularities $C_i$ and $C_j$ with $K$ and $L$ categories respectively.
-Let $nobs$ be the number of observations for a combination of levels from $(C_i, C_j)$.
-Now, $nobs$ might be equal for all combinations on an average so that $nobs = (T + 1)/(KL)$ as $T\rightarrow\infty$ or ideally take any value between $\{0, 1, \dots, T\}$.
-The statistical transformation used for summarizing the measured variable should be chosen in line with $nobs$.
-For example, it might be useful to compute deciles for summarization only when $nobs \ge 10$.
-Rarely occurring categories such as the 366th day of the year, or the 31st day of the month are likely to suffer from this problem.
+Suppose $C_i$ denotes day-of-year and $C_j$ denotes day-of-week.
+While any day of the week can occur on any day of the year, some combinations will be very rare.
+For example, the 366th day of the year will only coincide with a Wednesday approximately every 28 years on average.
+We refer to these as "near-clashes".
 

### 2020-09-02 main.Rmd
Section 5

@@
 
-The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by @Wickham2009pk, which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements.
-Drawing from the grammar of graphics, if <$C_i$, $C_j$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:
+The grammar of graphics introduced a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk].
+The layered grammar of graphics proposed by @Wickham2009pk gives an alternative and modified parametrization of the grammar, and suggests that graphics are made up of distinct layers of grammatical elements.
 
-- Data: <$C_i$, $C_j$, $v$>
-- Aesthetic mapping (mapping of variables to elements of the plot): $C_i$ mapped to $x$ position and $v$ to $y$ position
-- Statistical transformation (data summarization): any descriptive or smoothing statistics that summarizes distribution of $v$
-- Geometric objects (physical representation of the data): any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
-- Facet (split plots): $C_j$
+Drawing from the grammar of graphics, we consider visualizing the distribution of the measure variable $v$ conditional on the values of two granularities, $C_i$ and $C_j$.
+The following layers can be specified:
+
+- Data: <$C_i$, $C_j$, $v$>;
+- Aesthetic mapping (mapping of variables to elements of the plot): $C_i$ mapped to $x$ position and $v$ to $y$ position;
+- Facet (split plots): $C_j$;
+- Statistical transformation (data summarization): any descriptive or smoothing statistics that summarizes the distribution of $v$;
+- Geometric objects (physical representation of the data): any geometry displaying the distribution; for example, boxplot, letter value, violin, ridge or highest density region plots.
 
@@
 
-Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization.
-The basic plot choice for our data structure is the one that can display distributions.
-We will discuss few conventional and recent ways to plot distributions using both Kernel density estimates and descriptive statistics.
-Descriptive statistics based displays include box plots [@Tukey1977-jx] or different variations of it like notched box plots [@Mcgill1978-hg].
-More recent ways are the letter-value box plot [@Hofmann2017-sg] or quantile plots which display quantiles instead of quartiles in a traditional boxplot.
-Kernel density based plots for displaying a distribution include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, highest density regions (HDR) box [@Hyndman1996-ft].
-Each type of density display has different parameters, that need to be estimated given the data.
-Each is equipped with some benefits and challenges which should be borne in mind while using them for exploration.
+Plot selection is dictated by the statistical transformations and geometric objects used for the visualization.
+The basic plot choice for our data structure is one that can display distributions using kernel density estimates or descriptive statistics.
+Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@Mcgill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
+Plots based on kernel density estiatmes include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, and highest density region (HDR) boxplots [@Hyndman1996-ft].
+Each type of density display has parameters that need to be estimated from the data.
+Each is equipped with some benefits and challenges that should be borne in mind while using them for exploration.
 
-## Facet and aesthetic variables
+## Facet and aesthetic variables {#sec:aesthetics}
 
-### Levels {#sec:levels}
+### Levels {-}
 
-The levels of cyclic granularities has a role to play on the choice of plots since space and resolution might become a problem with too many levels.
-A potential approach could be to categorize the levels as very high/high/medium/low for each cyclic granularity and define some criteria based on usual cognitive power, display size available and the aesthetic mappings.
-Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight or days of the week.
+The levels of cyclic granularities affect plotting choices since space and resolution may be problematic with too many levels.
+A potential approach could be to categorize the number of levels as low/medium/high/very high for each cyclic granularity and define some criteria based on human cognitive power, available display size and the aesthetic mappings.
+Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight, or days of the week.
 
-### Synergy of cyclic granularities {#sec:vis-interaction}
+### Synergy of cyclic granularities {-}
 
-For data sets of the form ($C_i$, $C_j$, $v$) $\forall i, j \in N_C$, if there are levels of $C_i$ (or $C_j$) not spanned by levels of $C_j$ (or $C_i$), empty sets are formed leading to potential ineffective graphs.
-The conjecture is that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.
-Harmonies are pairs of granularities that do not contain empty combinations and could possibly be useful for exploring patterns.
-However, plotting clashes should be avoided since they might contain combinations which are empty or have too few observations.
-For illustration, \autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plot across months of the year faceted by quarters of the year.
-This plot does not work because every quarter does not correspond to all months of the year, for example, the first quarter of the year would never correspond to the month December of an year.
+The synergy of the two cyclic granularities will affect plotting choices for exploratory analysis.
+Cyclic granularities that form clashes (Section \@ref(sec:synergy)) or near-clashes lead to potentially ineffective graphs.
+Harmonies tend to be more useful for exploring patterns.
 
-```{r allFig, out.width="100%", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
+\autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plots across months of the year faceted by quarters of the year.
+This plot does not work because quarter-of-the-year clashes with month-of-the-year, leading to empty subsets.
+For example, the first quarter never corresponds to December.
+
+```{r allFig, out.width="100%", fig.pos="!htb", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
 ```
 
-(ref:allFig) Distribution of energy consumption displayed through letter value plots.
-Plot (a) displays consumption across month-of-year faceted by quarters of the year, (b) across weekday/weekend faceted by quarters of the year and (c) across quarters of the year faceted by weekday/weekend.
-Plots (b) and (c) both show harmonies since each quarter consists of both weekdays and weekends.
-Conversely, weekend and weekday can occur in every quarter.
-Analysts should avoid plotting such clashes.
-Plot (b) helps to compare weekends and weekdays for each quarter.
-It can be seen that for every quarter weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently, whereas, Plot (c) helps to compare quarters within weekdays and weekends.
-For example, the quartile spread of consumption shrinks/lowers from first to fourth quarter for weekdays, whereas this pattern is not true for weekends.
-Plot (a) shows a clash since all quarters do not correspond to all months of the year.
+(ref:allFig) Distribution of energy consumption displayed through letter value plots: (a) across month-of-year faceted by quarter-of-year; (b) across weekday/weekend faceted by quarter-of-year; (c) across quarter-of-year faceted by weekday/weekend.
+Plot (a) shows a clash since there are empty combinations.
+Plots (b) and (c) show harmonies since each quarter includes both weekdays and weekends.
+It can be seen in (b) that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
+Plot (c) switch the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends.
 
-### Interchangeability of mappings
+### Interchangeability of mappings {-}
 
-We will discuss the effect of the mapping of cyclic granularities in this section.
-When we consider data sets of the form <$C_i$, $C_j$, $v$> with $C_i$ mapped to $x$ position and $C_j$ to facets, then $A_k$'s are placed in close proximity and each $B_\ell$ represent a group/facet.
+When $C_i$ is mapped to the $x$ position and $C_j$ to facets, then the $A_k$ levels are juxtaposed and each $B_\ell$ represent a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case $A_k$'s are compared against each other within each group.
-With the mapping of $C_i$ and $C_j$ reversed, emphasis will shift to different behavior of the variables.
-\autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
+Hence, in this case the $A_k$s are compared against each other within each group.
+With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels.
+
+For example, \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
 \autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
@@
 
-Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section&nbsp;\ref{sec:summarize-measured}).
-Even when there are no rarely occurring events, number of observations might vary hugely within or across each facet.
-This might happen due to missing observations in the data or uneven locations of events in time domain.
-In such cases, the statistical transformations should be used with caution as sample sizes would directly affect both the variance and consequently the confidence interval of the estimators.
+Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
+Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
+In such cases, the statistical transformations should be used with caution as sample sizes will directly affect the accuracy of the estimated quantities being displayed.
 
@@
 
-### Screening and visualizing harmonies
+### Screening and visualizing harmonies {-}
 
@@
 Harmonies can be identified among those $42$ possibilities to narrow the search.
-\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:levels}).
+\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:aesthetics}).
 

### 2020-09-02 main.Rmd
Section 6

@@
 
-```{r linear-time, echo=FALSE, out.width = "100%", fig.cap="(ref:linear-time)", fig.pos="!htb"}
+```{r linear-time, echo=FALSE, fig.cap="(ref:linear-time)", fig.pos="!htb"}
 ```
@@
 
-```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap= "(ref:circular-dow)",fig.pos="!htb"}
+```{r circular-dow, echo=FALSE, fig.cap= "(ref:circular-dow)",fig.pos="!htb"}
 ```
@@
 
-```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap= "(ref:aperiodic-example)", fig.pos="!htb"}
+```{r aperiodic-example, echo=FALSE, fig.cap= "(ref:aperiodic-example)", fig.pos="!htb"}
 ```
@@
 
-```{r allFig, out.width="100%", fig.pos="!htb", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
+```{r allFig, fig.pos="!htb", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
 ```
@@
 Smart meters provide large quantities of measurements on energy usage for households across Australia.
-One of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014.
-The idea here is to show how to visualize the distribution of the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
+One of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle and parts of Sydney provides customer level data on energy consumption for every half hour from February 2012 to March 2014.
+We can use this data set to visualize the distribution of energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
 
@@
 
-The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] consisting of `reading_datetime`, `customer_id` and `general_supply_kwh` denoting the index, key and measured variable of the tsibble is used to facilitate the systematic exploration.
-While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing to consider is which cyclic time granularities we can look at exhaustively.
-Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year).
-Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year.
-Considering $6$ linear granularities half-hour, hour, day, week, month and year in the hierarchy table, $N_C = (6*5/2) = 15$.
-If $N_C$ seem too large, the smallest and largest linear granularities could be considered to be removed for analysis.
-Granularities half-year and year are removed to have $N_C = (4*3/2) = 6$ and obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
-Further, we add cyclic granularity day-type( "wknd_wday") to capture weekend and weekday behavior.
-Now that we have a list of cyclic granularities to look at, we should be able to compute them using Section&nbsp;\ref{sec:cyclic-calendar}.
+The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] includes the variables `reading_datetime`, `customer_id` and `general_supply_kwh` denoting the index, key and measured variable respectively.
+The interval of this tsibble is 30 minutes.
 
-```{r search, echo=FALSE, eval = FALSE}
-```
-
-```{r search_gran_limit2, echo = FALSE, eval = FALSE}
-```
+To identify the available cyclic time granularities, consider the conventional time deconstructions for a Gregorian calendar that can be formed from the 30-minute time index: half-hour, hour, day, week, month, quarter, half-year, year.
+In this example, we will consider the granularities hour, day, week and month giving six cyclic granularities "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
+To these we add day-type ("wknd_wday") to capture weekend and weekday behavior.
+Now that we have a list of cyclic granularities to look at, we can compute them using the results in Section&nbsp;\ref{sec:cyclic-calendar}.
 
@@
 
-From the search list, $N_C = 7$ cyclic granularities are chosen for which we would like to derive insights of energy behavior.
-Recalling the data structure <$C_i$, $C_j$, `general_supply_kwh`> for exploration $\forall i, j \in \{1, 2, \ldots, 7\}$, each of these $7$ cyclic granularities can either be mapped to x-axis or to facet.
-Choosing $2$ of the possible $7$ granularities, which is equivalent to having $^{7}P_2 = 42$ candidates for visualization.
+Using these seven cyclic granularities, we want to explore patterns of energy behavior.
+Each of these seven cyclic granularities can either be mapped to the x-axis or to facets.
+Choosing $2$ of the possible $7$ granularities, gives $^{7}P_2 = 42$ candidates for visualization.
 Harmonies can be identified among those $42$ possibilities to narrow the search.
-\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with levels more than $31$, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:aesthetics}).
+\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with more than $31$ levels, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:aesthetics}).
 
@@
 
-```{r bothcust, fig.cap = "(ref:bothcust)", out.width="90%", fig.pos="!htb"}
+```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!htb"}
 ```
@@
 (ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
-Two granularities are used: hour of the day (I) and weekday/weekend (II).
-Plot (a) shows granularity I faceted by granularity II, and plots (b), (c) shows the converse mapping.
-Plot (a) makes a comparison of usage by workday within each hour of the day using side-by-side boxplots.
-Generally, on a work day there is more consumption early in the day.
-Plots (b) and (c) examine the temporal trend of consumption over the course of a day, separately for the type of day.
-Plot (b) uses an area quantile to put the emphasis on the time series, for example, the median consumption over time shows prolonged usage in the morning on weekdays.
-Plot (c) uses a violin plot to place emphasis on distributional differences across hours.
-It can be seen that the morning use on weekdays is bimodal, some work days there is low usage, which might indicate the person is working from home and also having a late start.
+Two granularities are used: 1.
+hour of the day; and 2.
+weekday/weekend.
+Plot a shows granularity 1 faceted by granularity 2; plots b and c shows the reverse mapping.
+Plot a allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
+Plots b and c examine the temporal trend of consumption over the course of a day, separately for the type of day.
+Plot b uses an area quantile to emphasise the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
+Plot c uses a violin plot emphasise subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
 
-Few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the impact of different distribution plots and reverse mapping.
-For each of \autoref{fig:bothcust} (b) and (c), $C_i$ is the circular granularity day-type (weekday/weekend) and $C_j$ is hour of the day.
-The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust} (b and c respectively).
-\autoref{fig:bothcust} (a) displays reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour of the day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.
-In \autoref{fig:bothcust} (b), the black line is the median, whereas the purple band covers 25th to 75th percentile, the orange band covers 10th to 90th percentile and the green band covers 1st to 99th percentile.
-The first facet represents the weekday behavior while the second one displays the weekend behavior and energy consumption across each hours of the day is shown inside each facet.
-The energy consumption is extremely (positive- or right-) skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
+A few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the impact of different distribution plots and reverse mapping.
+For each of \autoref{fig:bothcust}b and c, $C_i$ denotes day-type (weekday/weekend) and $C_j$ is hour-of-day.
+The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust}b and c respectively.
+\autoref{fig:bothcust}a shows the reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour-of-day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.
+
+In \autoref{fig:bothcust}b, the black line is the median, whereas the purple band covers the 25th to 75th percentile, the orange band covers the 10th to 90th percentile, and the green band covers the 1st to 99th percentile.
+The first facet represents the weekday behavior while the second facet displays the weekend behavior; energy consumption across each hour of the day is shown inside each facet.
+The energy consumption is extremely skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
 This is common across both weekdays and weekends.
-For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer as compared to weekends.
+For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer compared to weekends.
 
-Consider looking at violin plots instead of quantile plots to look at the same data in \autoref{fig:bothcust}(c) for some more insights.
-There is bimodality in the early hours of the day, implying both low and high energy consumption is probable in the early hours of the day both for weekdays and weekends.
-If we visualize the same data with reverse mapping of the cyclic granularities, then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
-For example in \autoref{fig:bothcust}(a), it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends.
+The same data is shown using violin plots instead of quantile plots in \autoref{fig:bothcust}c.
+There is bimodality in the early hours of the day for weekdays and weekends.
+If we visualize the same data with reverse mapping of the cyclic granularities (\autoref{fig:bothcust}a), then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
+Then it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends.
 Also, outliers are more prominent in the latter hours of the day.
-All of these indicate that looking at different distribution geometry or changing the mapping might shed lights on different aspect of the energy behavior for the same sample population.
+All of these indicate that looking at different distribution geometry or changing the mapping can shed light on different aspects of energy behavior for the same sample.
 
-If the data for all keys are visualized together, it might lead to Simpson's paradox, which occurs when one observation shows a particular behavior, but this behavior paradoxically becomes obscured by aggregation.
-For example in a particular neighborhood one household may have the least daily power consumption for a full week, yet still not be the household with the minimum weekly power consumption.
-This is an intuitive possibility, because heterogeneous `customer_id`'s with very different occupation or demographics will tend to have very different energy behavior and combining them together will somehow weaken any typical or extreme behavior.
-A strategy for analyzing periodicities of multiple keys together could be to group the keys basis their energy behavior across different cyclic granularities.
-This is beyond the scope of the current work.
+## T20 cricket data of Indian Premier League {#sec:cricket}
 
-## T20 cricket data of Indian Premiere League {#sec:cricket}
-
-The method is not only restricted to temporal data, and can be generalized to many hierarchical granularities (with continuous and uni-directional nature).
-We illustrate this with an application to the sport cricket.
-Although there is no conventional time component in cricket, each ball can be thought to represent an ordering from past to future with the game progressing forward with each ball.
+Our proposed approach can be generalized to other hierarchical granularities where there is an underlying ordered index.
+We illustrate this with data from the sport cricket.
+Although there is no conventional time component in cricket, each ball can be thought to represent an ordering over the course of the game.
 In the Twenty20 format, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
-Thus, similar to time, there is a hierarchy where ball is nested within overs, overs nested within innings and innings within matches.
-The idea of cyclic granularities can be likewise mapped to this hierarchy.
-Example granularities then include ball of the over, over of the innings and ball of the innings.
-Although most of these cyclic granularities are circular in design of the hierarchy, in application of the rules some granularities are aperiodic.
-For example, in most cases an over will consist of 6 balls with some exceptions like wide balls or when an innings finishes before the over finishes.
-Thus, the cyclic granularity ball-of-over will be circular in most cases and aperiodic in others.
-
-The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
-The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).
-The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data across overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each innings has 20 overs and each match has 2 innings.
-This could be useful in a periodic world when we wish to compute any circular/quasi-circular granularity based on a hierarchy table which look like \autoref{tab:hierarchy-cric}.
-However, even if the situation is not periodic, it can be interesting to visualize the distribution of a measured variable to shed light on the aperiodic behavior of a non-temporal data set similar to aperiodic events like formal meetings, workshops, conferences, school semesters in a temporal set up.
+Thus, there is a hierarchy where ball is nested within overs, overs nested within innings, and innings within matches.
+Cyclic granularities can be constructed using this hierarchy.
+Example granularities include ball of the over, over of the innings, and ball of the innings.
+The hierarchy table is given in  \autoref{tab:hierarchy-cric}.
 
@@
 
-There are many interesting questions that could possibly be answered with the `cricket` data set.
-For example, it would be interesting to see if the distribution of total runs vary depending on if a team bats in the first or second innings.
-The Mumbai Indians (MI) and Chennai Super kings (CSK) appeared in final playoffs from 2010 to 2015.
-We take their example in order to dive deeper into this question.
-From Figure \ref{fig:cricex}(a), it can be observed that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the teams batting in the second inning.
-This seem to indicate that players feel mounting pressure to score more runs as they approach towards the end of the first inning.
-Whereas teams batting in the second innings have a set target in mind and are not subjected to such mounting pressure and may adopt a more conservative strategy, to score runs.
-Thus winning teams like CSK and MI seem to employ different inning strategies when it comes to their batting order.
+Although most of these cyclic granularities are circular by design of the hierarchy, in practice some granularities are aperiodic.
+For example, most overs will consist of 6 balls, but there are exceptions due to wide balls, no-balls, or when an innings finishes before the over finishes.
+Thus, the cyclic granularity ball-of-over may be aperiodic.
 
-Another fascinating exploration would be to examine if runs per over decrease in the subsequent over if fielding (defending) was good in the previous over?
-For establishing the fielding quality, we apply an indicator function on dismissals (1 if there was at least one wicket in the previous over due to run out or catch, 0 otherwise).
-Runs in the current over is then the observation variable.
-Dismissals in the previous over can lead to a batsman adopting a more defensive play style.
-Figure \ref{fig:cricex}(b) shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over as compared to the case when there has been at least one dismissal in the previous over.
-Wickets per over are considered as an aperiodic cyclic granularity with wickets as an aperiodic linear granularity.
-These granularities do not appear in the hierarchy table since it is difficult to position them in a hierarchy.
-These are similar to holidays or special events in temporal data.
+The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
+The IPL ball-by-ball data is provided in the `cricket` data set in the `gravitas` package for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each innings has 20 overs and each match has 2 innings.
 
-```{r cricex, fig.cap= "(ref:cricex)", warning = FALSE, message = FALSE, out.width = "90%"}
+There are many interesting questions that could be addressed with the `cricket` data set.
+For example, does the distribution of total runs vary depending on if a team bats in the first or second innings?
+The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in final playoffs from 2010 to 2015.
+Using  data from these two teams, it can be observed (\autoref{fig:cricex}a) that that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the team batting in the second innings after the first few overs.
+This suggests that players feel mounting pressure to score more runs as they approach the end of the first innings, while teams batting second have a set target in mind and are not subjected to such mounting pressure and therefore may adopt a more conservative run-scoring strategy.
+
+Another question that can be addressed is if good fielding (defending) affects the scoring rate in the subsequent over?
+To measure the fielding quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over due to a run out or catch, 0 otherwise).
+The scoring rate is measured by runs per over.
+\autoref{fig:cricex}b shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over compared to the case when there has been at least one dismissal in the previous over.
+Thus, dismissals lead a batsman to adopt a more defensive play style.
+This seems to be unaffected by the over of the innings (the facet variable).
+
+Here, wickets per over is an aperiodic cyclic granularity, so it does not  appear in the hierarchy table. These are similar to holidays or special events in temporal data.
+
+```{r cricex, fig.cap= "(ref:cricex)", warning = FALSE, message = FALSE, fig.height=4, fig.width=8}
 ```
@@
 (ref:cricex) Runs per over shown with different distribution displays, and granularities.
-Plot (a) shows letter value plot across overs faceted by innings.
+Plot a shows letter value plots across overs faceted by innings.
 For the team batting in the first innings there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
-Plot (b) shows quantile plot of runs per over across an indicator of wickets in previous over faceted by current over.
-This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over.
+Plot b shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
+This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.
 

### 2020-09-02 scripts/main.R

@@ -159,7 +159,7 @@ harmonies %>%
   kable(
     format = "latex",
     booktabs = TRUE,
-    caption = "Harmonies with a pair of cyclic granularity one placed on facet and the other on x-axis. Out of 42 possible combinations of cyclic granularities, only 16 are harmony pairs."
+    caption = "Harmonies with pairs of cyclic granularities, one mapped to facets and the other to the x-axis. Only 16 of 42 possible combinations of cyclic granularities are harmony pairs."
   ) %>%
   # row_spec(0, bold = TRUE) %>%
   kable_styling(latex_options = "hold_position")

### 2020-09-02 main.Rmd
Discussion, Acknowledgements and bibliography

@@
 Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@Mcgill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
-Plots based on kernel density estiatmes include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, and highest density region (HDR) boxplots [@Hyndman1996-ft].
+Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, and highest density region (HDR) boxplots [@Hyndman1996-ft].
 Each type of density display has parameters that need to be estimated from the data.
@@
 
-Exploratory data analysis and data analysis in general involve many iterations of finding and summarizing patterns.
+Exploratory data analysis involve many iterations of finding and summarizing patterns.
 With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible granularities to explore.
-This work provides a framework to systematically explore distribution of an univariate measured variable across two cyclic time granularities by creating any cyclic granularity, ranking a list of harmonies and thereby identifying possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.
+This work provides tools to classify and compute possible cyclic granularities from an ordered (usually temporal) index.
+We also provide a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities using visualizations based on the synergy and levels of the cyclic granularities.
+
+The `gravitas` package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities.
 
 A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
-A few R packages like `almanac` and `gs` provide functionality to create recurring events that are not periodic.
-These functions can be imported in the `gravitas` package to accommodate for aperiodic cyclic granularities.
-Another future direction of work could be to further refine the search of harmonies by only selecting pairs of cyclic granularities for which variation in measured variable is significant and rating these selected harmony pairs in order of importance for exploration.
+A few R packages including `almanac` and `gs` provide the tools to create recurring aperiodic events.
+These functions can be used with the `gravitas` package to accommodate aperiodic cyclic granularities.
+
+We propose producing plots based on pairs of cyclic granularities that form harmonies rather than clashes or near-clashes.
+A future direction of work could be to further refine the selection of appropriate pairs of granularities by identifying those for which the differences between the displayed distributions is greatest, and rating these selected harmony pairs in order of importance for exploration.
 
@@
 
-The authors would like to thank the cohort [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience in developing R packages, and Dr Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of the smart meter application.
+Thanks to the [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team) at Monash University for sharing their wisdom and experience in developing R packages, and Dr Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of the smart meter application.
 The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
@@
 
+\noindent
 **R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
@@
 Version 0.1.2 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).
-
-**R-packages:** Each of the R packages used in this article are available online with URLs provided in the bibliography.
 

### 2020-09-03 main.Rmd
check should be checks, recommend should be recommends

@@
 **R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
-The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, check feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommend prospective probability distributions through factors described in the article.
+The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends prospective probability distributions through factors described in the article.
 Version 0.1.2 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).

### 2020-09-03 main.Rmd
p25 "that that" should be "that"

@@
 The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in final playoffs from 2010 to 2015.
-Using  data from these two teams, it can be observed (\autoref{fig:cricex}a) that that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the team batting in the second innings after the first few overs.
+Using  data from these two teams, it can be observed (\autoref{fig:cricex}a) that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the team batting in the second innings after the first few overs.
 This suggests that players feel mounting pressure to score more runs as they approach the end of the first innings, while teams batting second have a set target in mind and are not subjected to such mounting pressure and therefore may adopt a more conservative run-scoring strategy.
@@
 A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
-A few R packages including `almanac` and `gs` provide the tools to create recurring aperiodic events.
+A few R packages including `almanac`[@R-almanac] and `gs` provide the tools to create recurring aperiodic events.
 These functions can be used with the `gravitas` package to accommodate aperiodic cyclic granularities.

### 2020-09-03 main.Rmd
p28 The current version on CRAN is 0.1.3, should that not be used?
p18 Displays are not "equipped with benefits". How about "Each has its own strengths and weaknesses and that should be..."

@@
 Each type of density display has parameters that need to be estimated from the data.
-Each is equipped with some benefits and challenges that should be borne in mind while using them for exploration.
+Each has its own strengths and weaknesses that should be borne in mind while using them for exploration.
 
@@
 The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends prospective probability distributions through factors described in the article.
-Version 0.1.2 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).
+Version 0.1.3 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).
 

### 2020-09-03 main.Rmd
p18 and in references: Mcgill should be McGill

@@
 The basic plot choice for our data structure is one that can display distributions using kernel density estimates or descriptive statistics.
-Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@Mcgill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
+Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
 Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, and highest density region (HDR) boxplots [@Hyndman1996-ft].

### 2020-09-03 main.Rmd
p10 "A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years." Fascinating, do you have a reference?
These are more physical properties. Demographic properties would be more about the people living there. Both need to be mentioned.

@@
 Smart meters currently provide half-hourly use in kWh for each household, from the time they were installed, some as early as 2012.
-Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+Households are distributed geographically and have different demographic properties as well as physical properties such as the existence of solar panels, central heating or air conditioning.
 The behavioral patterns in households vary substantially; for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
@@
 Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant.
-A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years.
+A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
 In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.

### 2020-09-03 main.Rmd
p27 the package gs is not easy to find. Should almanac and gs both be referenced?
grammar

@@
 It can be seen in (b) that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
-Plot (c) switch the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+Plot (c) switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
 For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends.
@@
 A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
-A few R packages including `almanac`[@R-almanac] and `gs` provide the tools to create recurring aperiodic events.
+A few R packages including `almanac`[@R-almanac] and `gs`[@R-gs] provide the tools to create recurring aperiodic events.
 These functions can be used with the `gravitas` package to accommodate aperiodic cyclic granularities.

### 2020-09-10 main.Rmd
p5 What are "prospective probability distributions" ?
removed from intro
changed to possible visual summaries in discussion

@@
   * facilitate manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra;
-  * develop an approach to check the feasibility of creating plots or drawing inferences for any two cyclic granularities;
-  * recommend prospective probability distributions for exploring distributions of a univariate dependent variable across pairs of granularities.
+  * develop an approach to check the feasibility of creating plots or drawing inferences for any two cyclic granularities.
 
@@
 **R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
-The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends prospective probability distributions through factors described in the article.
+The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends possible visual summaries through factors described in the article.
 Version 0.1.3 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).

### 2020-09-10 main.Rmd
Antony's remark on this could be applied on data by years, decades

@@
 We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns and anomalies.
-Although we were motivated by the smart meter example, the problem and the solutions we propose are relevant to any temporal data observed more than once per year.
+Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.
+In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be in weather or astronomical data.
 

### 2020-09-10 cover-letter/cover-letter.Rmd
cover letter added

---
from: Sayani Gupta
fromqualifications: M.Stat
fromtitle: PhD student
fromWWW: sayani.netlify.app
fromemail: Sayani.Gupta@monash.edu
fromphone: +61 4 2388 9998
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick
closing: Best wishes
spacing: 1.4
output: MonashEBSTemplates::letter
---
I am very excited to submit an original research article entitled “ Visualizing probability distributions across bivariate cyclic temporal granularities” by Sayani Gupta, Rob J Hyndman, Dianne Cook, and Antony Unwin for your consideration for publication in the Journal of Computational and Graphical Statistics.

We believe that this manuscript is appropriate for publication by the Journal of Computational and Graphical Statistics because it deals with the interaction of the disintegration of time innovatively by exploiting the hierarchical structure of time and provides direction on systematically exploring periodicities of large amounts of high-frequency time series data. The method is also applicable for non-temporal data, which has a similar nested hierarchical structure as time.

The manuscript is original, and no part has been published before, nor is any part of it under consideration for publication at another journal. The preprint has been uploaded to arXiv and submitted to be added to Monash EBS working papers.

Thank you for your consideration!

### 2020-09-16 cover-letter/cover-letter.Rmd
cover letter fixed  using Di's comments

@@
 toaddress: Journal of Computational and Graphical Statistics
-opening: Dear Dr McCormick
+opening: Dear Dr McCormick,
 closing: Best wishes
@@
 ---
-I am very excited to submit an original research article entitled “ Visualizing probability distributions across bivariate cyclic temporal granularities” by Sayani Gupta, Rob J Hyndman, Dianne Cook, and Antony Unwin for your consideration for publication in the Journal of Computational and Graphical Statistics.
 
-We believe that this manuscript is appropriate for publication by the Journal of Computational and Graphical Statistics because it deals with the interaction of the disintegration of time innovatively by exploiting the hierarchical structure of time and provides direction on systematically exploring periodicities of large amounts of high-frequency time series data.
-The method is also applicable for non-temporal data, which has a similar nested hierarchical structure as time.
+Please consider my article entitled “Visualizing probability distributions across bivariate cyclic temporal granularities” co-authored by Rob J Hyndman, Dianne Cook, and Antony Unwin for publication in the Journal of Computational and Graphical Statistics.
+
+We believe that this manuscript is appropriate for publication by the Journal of Computational and Graphical Statistics because it makes unique computational and graphical contributions for analyzing temporal data and also non-temporal data with a nested hierarchical structure.
 
 The manuscript is original, and no part has been published before, nor is any part of it under consideration for publication at another journal.
-The preprint has been uploaded to arXiv and submitted to be added to Monash EBS working papers.
+A preprint has been made available at arXiv and submitted to be added to Monash EBS working papers.
 

### 2020-09-16 main.Rmd
cricket data example 2 write up updated

@@
 
-Another question that can be addressed is if good fielding (defending) affects the scoring rate in the subsequent over?
-To measure the fielding quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over due to a run out or catch, 0 otherwise).
+Another question that can be addressed is if good fielding or bowling (defending) in the previous over affects the scoring rate in the subsequent over?
+To measure the fielding and bowling quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over, 0 otherwise).
 The scoring rate is measured by runs per over.
 \autoref{fig:cricex}b shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over compared to the case when there has been at least one dismissal in the previous over.
-Thus, dismissals lead a batsman to adopt a more defensive play style.
 This seems to be unaffected by the over of the innings (the facet variable).
+This might be because the new batsman needs to play himself in or the dismissals lead the (not-dismissed) batsman to adopt a more defensive play style.
+Run rates will also vary depending on which player is facing the next over and when the wicket falls in the previous over.
 

### 2020-09-16 scripts/main.R
all dismissals included rather than just fielding

@@ -296,7 +296,7 @@ cricket_dot_field <- cricket_season %>%
   summarise(
     runs_per_over = sum(total_runs),
     run_rate = sum(total_runs) * 6 / length(total_runs),
-    fielding_wckts = sum(fielding_proxy),
+    wckts = sum(wicket_proxy),
     dot_balls = sum(dot_ball_proxy)
   ) %>%
   mutate(diff_run_rate = c(0, diff(run_rate)))
@@ -308,7 +308,7 @@ cricket_tsibble <- cricket_dot_field %>%
 
 cricket_data <- cricket_tsibble %>%
   mutate(
-    field = if_else(fielding_wckts == 0, "0", "1+"),
+    field = if_else(wckts == 0, "0", "1+"),
     dot = if_else(dot_balls == 0, "no dot balls", ">0 dot balls"),
     lag_field = lag(field),
     lag_dot = lag(dot)

### 2020-09-16 main.Rmd
A recommendation of appropriate distribution displays is provided  - deleted from abstract
Antony's comments on§5.3 addressed

@@
 Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
-A recommendation of appropriate distribution displays is provided.
 
@@
 - Facet (split plots): $C_j$;
-- Statistical transformation (data summarization): any descriptive or smoothing statistics that summarizes the distribution of $v$;
+- Data summarization: any descriptive or smoothing statistics that summarizes the distribution of $v$;
 - Geometric objects (physical representation of the data): any geometry displaying the distribution; for example, boxplot, letter value, violin, ridge or highest density region plots.
 
-## Choice of statistical transformations and geometric objects
+## Data summarization and geometric objects
 
@@
 Each has its own strengths and weaknesses that should be borne in mind while using them for exploration.
+
+\noindent Visualizing distributions can be uninformative or potentially misleading if data summarization are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
+Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
+In such cases, data summarization should be used with caution as sample sizes will directly affect the accuracy of the estimated quantities being displayed.
 

### 2020-09-16 main.Rmd
emphasize not emphasise

@@
 Plots b and c examine the temporal trend of consumption over the course of a day, separately for the type of day.
-Plot b uses an area quantile to emphasise the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
-Plot c uses a violin plot emphasise subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
+Plot b uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
+Plot c uses a violin plot emphasize subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
 

### 2020-09-16 scripts/main.R
not adding over as a fct gives the warning

@@ -316,6 +316,7 @@ cricket_data <- cricket_tsibble %>%
   filter(lag_field != 0, lag_dot != 0)
 
 cricket_data$lag_field <- factor(cricket_data$field, levels = c("0", "1+"))
+#cricket_data$over <- factor(cricket_data$over, levels = seq(1:20))
 
 cricket_data %>%
   filter(over != 1) %>%

### 2020-09-16 main.Rmd
all statistical transformation changed to data summarization

@@
 
-Plot selection is dictated by the statistical transformations and geometric objects used for the visualization.
+Plot selection is dictated by the choice of data summarization and geometric objects used for the visualization.
 The basic plot choice for our data structure is one that can display distributions using kernel density estimates or descriptive statistics.
@@
 Each has its own strengths and weaknesses that should be borne in mind while using them for exploration.
-
-\noindent Visualizing distributions can be uninformative or potentially misleading if data summarization are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
+Visualizing distributions can be uninformative or potentially misleading if data summarization are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
 Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
@@
 \autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
-
-## Number of observations and statistical transformations
-
-Visualizing distributions can be misleading if statistical transformations are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
-Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
-In such cases, the statistical transformations should be used with caution as sample sizes will directly affect the accuracy of the estimated quantities being displayed.
 
@@
 
-**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.Rmd).
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
 

### 2020-09-16 scripts/main.R
last plot title fixed and pdf update

@@ -332,7 +332,7 @@ cricket_data %>%
   xlab("number of wickets in previous over") +
   ggtitle("b") +
   theme(
-    plot.title = element_text(face = "bold"),
+    plot.title = element_text(hjust = 0),
     axis.ticks = element_blank(),
     legend.background = element_blank(),
     legend.key = element_blank(),

### 2020-09-22 main.Rmd
ridge plot reference added and minor correction

@@
 
-Drawing from the grammar of graphics, we consider visualizing the distribution of the measure variable $v$ conditional on the values of two granularities, $C_i$ and $C_j$.
+Drawing from the grammar of graphics, we consider visualizing the distribution of the measured variable $v$ conditional on the values of two granularities, $C_i$ and $C_j$.
 The following layers can be specified:
@@
 Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
-Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots, and highest density region (HDR) boxplots [@Hyndman1996-ft].
+Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) boxplots [@Hyndman1996-ft].
 Each type of density display has parameters that need to be estimated from the data.
@@
 Another question that can be addressed is if good fielding or bowling (defending) in the previous over affects the scoring rate in the subsequent over?
-To measure the fielding and bowling quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over, 0 otherwise).
+To measure the defending quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over, 0 otherwise).
 The scoring rate is measured by runs per over.

### 2020-09-22 main.Rmd
why second quarter consumption is lower
plot (a) changed to plot a in Figure 4

@@
 
-(ref:allFig) Distribution of energy consumption displayed through letter value plots: (a) across month-of-year faceted by quarter-of-year; (b) across weekday/weekend faceted by quarter-of-year; (c) across quarter-of-year faceted by weekday/weekend.
-Plot (a) shows a clash since there are empty combinations.
-Plots (b) and (c) show harmonies since each quarter includes both weekdays and weekends.
-It can be seen in (b) that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
-Plot (c) switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+(ref:allFig) Distribution of energy consumption displayed through letter value plots: plot a across month-of-year faceted by quarter-of-year; plot b across weekday/weekend faceted by quarter-of-year; plot c across quarter-of-year faceted by weekday/weekend.
+Plot a shows a clash since there are empty combinations.
+Plots b and c show harmonies since each quarter includes both weekdays and weekends.
+It can be seen in plot b that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
+This is probably because temperatures are lower in the second quarter compared to the first quarter (summer in Australia), leading to lower consumption of energy in the second quarter.
+Plot c switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
 For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends.

### 2020-09-22 scripts/main.R
Figure 4 each plot sarrounded by a light margin to have the labels separated

@@ -69,7 +69,10 @@ scene1 <- VIC %>%
   theme(legend.position = "right") +
   ggtitle("") +
   scale_y_log10() +
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    #panel.grid.major = element_line(colour = "#E0E0E0"),
+    panel.border = element_rect(colour = "#E0E0E0", fill = NA))
 
 scene2 <- VIC %>%
   prob_plot(
@@ -84,7 +87,10 @@ scene2 <- VIC %>%
   theme(legend.position = "right") +
   ggtitle("") +
   scale_y_log10() +
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    #panel.grid.major = element_line(colour = "#E0E0E0"),
+    panel.border = element_rect(colour = "#E0E0E0", fill = NA))
 
 scene3 <- VIC %>%
   prob_plot(
@@ -99,7 +105,10 @@ scene3 <- VIC %>%
   scale_fill_brewer(palette = "Set2") +
   ggtitle("") +
   scale_y_log10() +
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    #panel.grid.major = element_line(colour = "#E0E0E0"),
+    panel.border = element_rect(colour = "#E0E0E0", fill = NA))
 
 gg_fig <- ggarrange(
   scene3,

### 2020-09-22 scripts/main.R
borders of plots fixed

@@ -188,7 +188,9 @@ cust2_quantile <- smart_meter10 %>%
   # ylab("electricity demand [KWh]") +
   xlab("hours of the day") +
   ggtitle("") +
-  theme_minimal()
+  theme_minimal() +
+  theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
+  )
 
 cust2_violin <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@ -202,7 +204,9 @@ cust2_violin <- smart_meter10 %>%
   xlab("hours of the day") +
   ggtitle("") +
   scale_x_discrete(breaks = seq(0, 23, 5)) +
-  theme_minimal()
+  theme_minimal() +
+  theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
+  )
 
 cust2_box <- smart_meter10 %>%
   filter(customer_id %in% c(10017936)) %>%
@@ -217,7 +221,9 @@ cust2_box <- smart_meter10 %>%
   scale_x_discrete(labels = c("wday", "wend")) +
   scale_y_sqrt() +
   theme(axis.text.x = element_text(size = 7)) +
-  theme_minimal()
+  theme_minimal() +
+ theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
+)
 
 gg_fig <- ggarrange(
   cust2_box,
@@ -280,7 +286,9 @@ cricket_tsibble %>%
   theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))
   ) +
-  theme_minimal()
+  theme_minimal() +
+  theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
+)
 
 cricket_all <- read_csv("data-raw/deliveries_all.csv")
 matches_all <- read_csv("data-raw/matches_all.csv")

### 2020-09-22 main.Rmd
instead of 18z write 18\timesz

@@
 C_{\text{uinal}, \text{baktun}}(z)  &=  C_{\text{uinal}, \text{tun}}(z) + P(\text{uinal}, \text{tun})C_{\text{tun},\text{katun}}(z) + P(\text{uinal},\text{katun})C_{\text{katun}, \text{baktun}}(z) \\
-               &=  \lfloor z/20\rfloor \mod 18 + 18\lfloor 18z/20\rfloor \mod 20
+               &=  \lfloor z/20\rfloor \mod 18 + 18\lfloor 18\times z/20\rfloor \mod 20
                 + 18\times20\lfloor 18\times20\times z/20\rfloor \mod 20.

### 2020-09-22 main.Rmd
revisions done

@@
 It can be seen in plot b that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
-This is probably because temperatures are lower in the second quarter compared to the first quarter (summer in Australia), leading to lower consumption of energy in the second quarter.
+This is probably because of lower temperatures in the second quarter compared to the first quarter (summer in Australia).
 Plot c switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.

### 2020-09-23 main.Rmd
shorten abstract to 200 words

@@
 This paper describes classes of time deconstructions using linear and cyclic time granularities.
-Linear time granularities respect the linear progression of time such as hours, days, weeks and months.
-Cyclic time granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
+Linear granularities respect the linear progression of time such as hours, days, weeks and months.
+Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
 The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up.
@@
 
- The methods are implemented in the open source R package `gravitas`, providing functions for creating granularities and exploring the associated time series which are consistent with a tidy workflow (@Grolemund2018-po), and the probability distributions can be examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
+ The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.
+
 bibliography: [bibliography.bib]
@@
 
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow. In particular, we
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@Grolemund2018-po). In particular, we
 

### 2020-09-23 cover-letter/cover-letter.Rmd
minor changes

@@
 
-The manuscript is original, and no part has been published before, nor is any part of it under consideration for publication at another journal.
-A preprint has been made available at arXiv and submitted to be added to Monash EBS working papers.
+The manuscript is original.
+A preprint has been made available at arXiv and Monash EBS working paper series, but no part has been published before, nor is any part of it under consideration for publication at another journal.
 

### 2020-09-23 main.Rmd
numbats link update

@@
 
-Thanks to the [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team) at Monash University for sharing their wisdom and experience in developing R packages, and Dr Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of the smart meter application.
+Thanks to the [NUMBATS](https://numbat.space/) at Monash University for sharing their wisdom and experience in developing R packages, and Dr Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of the smart meter application.
 The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).

### 2020-09-23 main.Rmd
acknowledgement section updated

@@
 
-Thanks to the [NUMBATS](https://numbat.space/) at Monash University for sharing their wisdom and experience in developing R packages, and Dr Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of the smart meter application.
+The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+Thanks to [Data61 CSIRO](https://data61.csiro.au/) for partially funding Sayani's research and Dr Peter Toscas for providing useful inputs on improving the analysis of the smart meter application.
 The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).

### 2020-09-23 scripts/main.R
change Figs to img to solve Arxiv error

@@ -12,13 +12,13 @@ library(ggpubr)
 library(kableExtra)
 
 ## ---- linear-time
-include_graphics("Figs/linear-ex.png")
+include_graphics("img/linear-ex.png")
 
 ## ---- circular-dow
-include_graphics("Figs/circular-ex.png")
+include_graphics("img/circular-ex.png")
 
 ## ----aperiodic-example
-include_graphics("Figs/aperiodic-ex3.png")
+include_graphics("img/aperiodic-ex3.png")
 
 ## ----tab-mayan
 table_mayan <- tibble(

### 2020-09-25 main.Rmd
acknowledgement margin fixed
bold label for last figure and for figure 4 no brackets for labels in text

@@
   fig.show = "hold",
-  cache = TRUE,
+  cache = FALSE,
   cache.path = "cache/",
@@
 
-\autoref{fig:allFig} (a) shows the distribution of half-hourly electricity consumption through letter value plots across months of the year faceted by quarters of the year.
+\autoref{fig:allFig}a shows the distribution of half-hourly electricity consumption through letter value plots across months of the year faceted by quarters of the year.
 This plot does not work because quarter-of-the-year clashes with month-of-the-year, leading to empty subsets.
@@
 
-For example, \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed.
-\autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.
+For example, \autoref{fig:allFig}b shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig}c shows the same two cyclic granularities with their mapping reversed.
+\autoref{fig:allFig}b helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig}c helps to compare quarters within weekend and weekday.
 
@@
 Thanks to [Data61 CSIRO](https://data61.csiro.au/) for partially funding Sayani's research and Dr Peter Toscas for providing useful inputs on improving the analysis of the smart meter application.
+We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
 The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
-We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
 More details about the package can be found at [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
+The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
 This article was created with `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
-The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
 

### 2020-09-25 scripts/main.R

@@ -281,14 +281,13 @@ cricket_tsibble %>%
   theme(legend.position = "right") +
   ggtitle("a") +
   ylab("runs per over") +
-  xlab("overs of the innings") +
-  theme(plot.title = element_text(face = "bold")) +
-  theme(
-    strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))
-  ) +
+  xlab("overs of the innings")  +
   theme_minimal() +
   theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
-)
+) + 
+  theme(plot.title = element_text(face = "bold"),
+        strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))
+  )
 
 cricket_all <- read_csv("data-raw/deliveries_all.csv")
 matches_all <- read_csv("data-raw/matches_all.csv")
@@ -349,7 +348,7 @@ cricket_data %>%
   xlab("number of wickets in previous over") +
   ggtitle("b") +
   theme(
-    plot.title = element_text(hjust = 0),
+    plot.title = element_text(hjust = 0, face = "bold"),
     axis.ticks = element_blank(),
     legend.background = element_blank(),
     legend.key = element_blank(),

### 2020-09-25 main_wo_fig_tbl.Rmd
delete without figure version

--- 2020-09-23 main.Rmd
@@
   message = FALSE,
+  eval = FALSE,
   comment = "#>",
@@
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lll}
-\toprule
-Circular granularity & Expression & Period \\
-\midrule
-minute-of-hour                               &
-  $C_1 = z \mod 60$                     &
-  $P_1 = \phantom{99}60$ \\
-minute-of-day                                &
-  $C_j = z \mod 60*24$                  &
-  $P_2= `r 60*24`$\\
-hour-of-day                                  &
-  $C_3 = \lfloor z/60\rfloor\mod 24$    &
-  $P_3 = \phantom{99}24$ \\
-hour-of-week                                 &
-  $C_4 = \lfloor z/60\rfloor\mod 24*7$  &
-  $P_4= \phantom{9}`r 24*7`$\\
-day-of-week                                  &
-  $C_5 = \lfloor z/24*60\rfloor \mod 7$ &
-  $P_5= \phantom{999}7$\\
-\bottomrule
-\end{tabular}
-\end{center}
-\caption{Examples of circular granularities with bottom granularity minutes.
+<!-- \begin{table}[ht] -->
+<!-- \begin{center} -->
+<!-- \begin{tabular}{lll} -->
+<!-- \toprule -->
+<!-- Circular granularity & Expression & Period \\ -->
+<!-- \midrule -->
+<!-- minute-of-hour                               & -->
+<!--   $C_1 = z \mod 60$                     & -->
+<!--   $P_1 = \phantom{99}60$ \\ -->
+<!-- minute-of-day                                & -->
+<!--   $C_j = z \mod 60*24$                  & -->
+<!--   $P_2= `r 60*24`$\\ -->
+<!-- hour-of-day                                  & -->
+<!--   $C_3 = \lfloor z/60\rfloor\mod 24$    & -->
+<!--   $P_3 = \phantom{99}24$ \\ -->
+<!-- hour-of-week                                 & -->
+<!--   $C_4 = \lfloor z/60\rfloor\mod 24*7$  & -->
+<!--   $P_4= \phantom{9}`r 24*7`$\\ -->
+<!-- day-of-week                                  & -->
+<!--   $C_5 = \lfloor z/24*60\rfloor \mod 7$ & -->
+<!--   $P_5= \phantom{999}7$\\ -->
+<!-- \bottomrule -->
+<!-- \end{tabular} -->
+<!-- \end{center} -->
+<!-- \caption{Examples of circular granularities with bottom granularity minutes.
 Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
 Circular granularities can be expressed using modular arithmetic due to their regular mapping.
-}
-\label{tab:definitions}
-\end{table}
+} -->
+<!-- \label{tab:definitions} -->
+<!-- \end{table} -->
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{lr@{~}lr@{~}r}
-\toprule
-Quasi-circular granularity && Possible period lengths\\
-\midrule
-$Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\
-$Q_2 =$ hour-of-month && $P_2 = 24\times 31, 24\times 30, 24\times 29, 24\times 28$\\
-$Q_3 =$ day-of-year && $P_3 = 366, 365$\\
-$Q_4 =$ week-of-month && $P_4 = 5, 4$\\
-\bottomrule
-\end{tabular}
-\caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.}
-\label{tab:quasi}
-\end{table}
+<!-- \begin{table}[ht] -->
+<!-- \centering -->
+<!-- \begin{tabular}{lr@{~}lr@{~}r} -->
+<!-- \toprule -->
+<!-- Quasi-circular granularity && Possible period lengths\\ -->
+<!-- \midrule -->
+<!-- $Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\ -->
+<!-- $Q_2 =$ hour-of-month && $P_2 = 24\times 31, 24\times 30, 24\times 29, 24\times 28$\\ -->
+<!-- $Q_3 =$ day-of-year && $P_3 = 366, 365$\\ -->
+<!-- $Q_4 =$ week-of-month && $P_4 = 5, 4$\\ -->
+<!-- \bottomrule -->
+<!-- \end{tabular} -->
+<!-- \caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.} -->
+<!-- \label{tab:quasi} -->
+<!-- \end{table} -->
 

### 2021-03-24 reviews/response.Rmd
reviewer's comments added code change

---
title: Response to reviewers
author: JCGS-20-344
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their helpful feedback which has led to many improvements in the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

This manuscript has undergone an overhaul to address reviewer's comments:

1. 
2. 
3. 
4.

Along with these major changes, we have done lots of minor tweaks to improve the manuscript.

# Reviewer 1 comments {-}

> comment 1.

response

> comment 2

response

> comment 3

response

> comment 4

response

### 2021-03-24 reviews/reviewer1.Rmd
reviewer's comments added code change

---
title: "Review on Visualizing probability distributions across bivariate cyclic temporal granularities"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Comments to the Author
Short Summary:
The paper introduces a calendar algebra approach to specify classes of deconstructing time into granularities and relate pairs of granularities to each other. The definitions and computations allow to construct linear and cyclic time granularities. For these granularity pairs, the paper introduces ways to visualise the descriptive statistics in different plots and illustrates with two example datasets how the visualisation is constructed, explored, interpreted and analysed. The authors also provide an implementation of the introduced computations and visualisations in an R package.

In the following I will give an overview on the major strengths and weaknesses, as well as a more detailed discussion on some shortcomings and give pointers to possible ways to improve the paper.

Strengths:
- Introduces a more generalised definition and characterisation of granularities to allow single- and multiple-order-up combinations
- allows granularities for different time structures, like linear and different periodic structures, like cyclic (circular, quasi-circular, and aperiodic).
- Defines relativities for pairs of granularities, the necessary calendar algebra to do the computation, and relations between granularities.
- Provides an R package for applying and computing the proposed ideas and visualise the results in order to analyse the data (missing only the computation of cyclic aperiodic granularities)
- Illustrated the applicability and usefulness using two example applications and datasets.

Weaknesses:
- Cricket sport argumentation a bit difficult to follow for people not knowing the sport at all.
- Embedding the visualisation into the grammar of graphics is not convincing or may not be necessary at all. Either extend or remove.
- Not clear which visualisations for descriptive statistics known from literature are implemented and supported.
- Letter-value plots are widely used in the paper, but not explained in the necessary detail to understand the visualisation adequately.

Overall the paper is very well written, all necessary details provided. The paper introduces an interesting and challenging problem of formalising more general hierarchical structures, like granularities, for time and introduce the necessary computations to allow the application of these concepts to different data. The visualisation and provided discussion allows to comprehend arguments provided and how it turns out in practical applications. The provided R package is useful for applying granularities to ones own data and use for visually analyse the data on different combinations of granularity levels. Below there are some more detailed remarks to improve the identified weaknesses in the paper.

1. _I am not convinced that it is required to introduce the concept of grammar of graphics and embed the used visualization in this framework on such a level of detail like in Section 5. Of course some elements of the grammar of graphics are used in Section 5.2 to argue about facets, levels, aesthetic variables, and mappings, but if the reader does not know about this framework, the introduction and explanations about this concept/framework provided in Section 5 is not enough to follow the later argumentation. I would like to see some more details on the concept/framework on the grammar of graphics, or not using the concepts in the parts of Section 5.2 and paraphrasing it without using the terms from the grammar of graphics. I think that it is doable to explain the construction of the visualization without using this concept/framework, because the mappings are quite straight forward. I would like to see (a) either extend the introduction to grammar of graphics and the basics needed to argue in more detail about the embedding of your visualisations using this concept/framework, or (b) remove it._

2. _It is not clear from the paragraph in Section 5.1 what basic plot choices are implemented and used and supported for visualisation. There is only the list of plots to display descriptive statistics provided. It would be important to describe which are implemented and possible to use. The later examples use primarily letter-value plots, then also box-plots, violin plots and line charts with area quantile plots. The letter-value plots are used in major examples of the paper, therefore I would argue to provide some more details on this visual representation and what the 5 colours and letters mean (e.g. LV=letter value, M = median, F = fourths, E = eighths, D = sixteenths, C = thirty-seconds), as it is not as widely known like standard box-plots. This would help greatly in better following the argumentation and explanations._

3. _I have to confess that I do not know anything about the sport cricket and it may be unthinkable for people from countries where cricket is a national sport, but I really had difficulties to follow the example in Section 6.2. The basic hierarchical concept of over—inning--match--season was understandable, but I think that some parts need to be revisited and clarified for also reaching readers not knowledgeable in cricket. Especially the discussion on the interesting questions raised are a bit difficult to follow without knowing much about cricket. For example some terms should be at least briefly explained or removed if not necessary for the explanation, e.g., Twenty20 format, fielding, bowling, dismissal, wicket. I do know a lot about different sports and also the basic idea and structure in many sports, but cricket was never on my radar. Apologise my ignorance about cricket._

### 2021-03-25 reviews/response.Rmd
cricket description added to address comments

@@
 title: Response to reviewers
-author: JCGS-20-344
+author: "JCGS-20-344"
 output:
@@
 
-We thank the reviewers for their helpful feedback which has led to many improvements in the paper.
+We would really like to thank the editor and the reviewers for their encouragement and detailed feedback which has helped to improve the paper in terms of both concision and precision and as well as reaching out to more readers.
 The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
-
-This manuscript has undergone an overhaul to address reviewer's comments:
-
-1. 
-2. 
-3. 
-4.
-
-Along with these major changes, we have done lots of minor tweaks to improve the manuscript.
 
@@
 
-> comment 3
+> I have to confess that I do not know anything about the sport cricket and it may be unthinkable for people from countries where cricket is a national sport, but I really had difficulties to follow the example in Section 6.2.
+The basic hierarchical concept of over—inning--match--season was understandable, but I think that some parts need to be revisited and clarified for also reaching readers not knowledgeable in cricket.
+Especially the discussion on the interesting questions raised are a bit difficult to follow without knowing much about cricket.
+For example some terms should be at least briefly explained or removed if not necessary for the explanation, e.g., Twenty20 format, fielding, bowling, dismissal, wicket.
+I do know a lot about different sports and also the basic idea and structure in many sports, but cricket was never on my radar.
+Apologise my ignorance about cricket.
 
-response
+Thanks for the advice.
+We have added a description of the game cricket in the first paragraph of Section 6.2 so that each new terminology applicable for cricket (and relevant for the following analysis) is highlighted in italics and briefly explained.
+We believe that this will help the readers not knowledgeable in cricket to also follow the presented analysis.
 

### 2021-03-25 main.Rmd
cricket description added to address comments

@@
 We illustrate this with data from the sport cricket.
+Cricket is played with two teams of 11 players each, with each team taking turns batting and playing the field.
+This is similar to baseball, wherein the _batsman_ and _bowler_ in cricket are analogous to a batter and pitcher in baseball.
+A _wicket_ is a structure with three sticks, stuck into the ground, placed at the end of the cricket pitch behind the batsman.
+A batsman tries to prevent the bowler from hitting the wicket by hitting the _ball_ using a _bat_.
+Two batsmen of a team are on the pitch at the same time.
+There are nine _fielders_ on the opposing team, with one player out of the 11 acting as the bowler, while another takes up the role of the _wicket-keeper_.
+To score a _run_, a batsman can run after hitting the ball and changing places on the pitch or also by hitting the ball out of the field on a fly (6 runs) or on a bounce (4 runs).
+A batsman is _dismissed_ when the period of batting is brought to an end either due to good bowling of the bowler or good fielding of the rest of the opposing team.
+
+Cricket is played in various formats and Twenty20 cricket (T20) is a shortened format, where the two teams have a single _innings_ each, which is restricted to a maximum of 20 _overs_.
+An over will consist of 6 balls (with some exceptions).
+A single _match_ will consist of 2 innings and a _season_ consists of several matches.
 Although there is no conventional time component in cricket, each ball can be thought to represent an ordering over the course of the game.
-In the Twenty20 format, an over will consist of 6 balls (with some exceptions), an innings is restricted to a maximum of 20 overs, a match will consist of 2 innings and a season consists of several matches.
-Thus, there is a hierarchy where ball is nested within overs, overs nested within innings, and innings within matches.
+Then, we can conceive a hierarchy where the ball is nested within overs, overs nested within innings, innings within matches, and matches within seasons.
 Cyclic granularities can be constructed using this hierarchy.
@@
 The hierarchy table is given in  \autoref{tab:hierarchy-cric}.
+Although most of these cyclic granularities are circular by the design of the hierarchy, in practice some granularities are aperiodic.
+For example, most overs will consist of 6 balls, but there are exceptions due to wide balls, no-balls, or when an innings finishes before the over finishes.
+Thus, the cyclic granularity ball-of-over may be aperiodic.
 
@@
 
-Although most of these cyclic granularities are circular by design of the hierarchy, in practice some granularities are aperiodic.
-For example, most overs will consist of 6 balls, but there are exceptions due to wide balls, no-balls, or when an innings finishes before the over finishes.
-Thus, the cyclic granularity ball-of-over may be aperiodic.
-
-The Indian Premier League (IPL) is a professional Twenty20 cricket league in India contested by eight teams representing eight different cities in India.
+The Indian Premier League (IPL) is a professional T20 cricket league in India contested by eight teams representing eight different cities in India.
 The IPL ball-by-ball data is provided in the `cricket` data set in the `gravitas` package for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each innings has 20 overs and each match has 2 innings.
 
-There are many interesting questions that could be addressed with the `cricket` data set.
+Many interesting questions could be addressed with the `cricket` data set.
 For example, does the distribution of total runs vary depending on if a team bats in the first or second innings?
-The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in final playoffs from 2010 to 2015.
-Using  data from these two teams, it can be observed (\autoref{fig:cricex}a) that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in median and quartile deviation of runs for the team batting in the second innings after the first few overs.
+The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in the final playoffs from 2010 to 2015.
+Using data from these two teams, it can be observed (\autoref{fig:cricex}a) that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in the median and quartile deviation of runs for the team batting in the second innings after the first few overs.
 This suggests that players feel mounting pressure to score more runs as they approach the end of the first innings, while teams batting second have a set target in mind and are not subjected to such mounting pressure and therefore may adopt a more conservative run-scoring strategy.
@@
 \autoref{fig:cricex}b shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over compared to the case when there has been at least one dismissal in the previous over.
-This seems to be unaffected by the over of the innings (the facet variable).
-This might be because the new batsman needs to play himself in or the dismissals lead the (not-dismissed) batsman to adopt a more defensive play style.
+This seems to be unaffected by the over of the innings (the faceting variable).
+This might be because the new batsman needs to play himself in or the dismissals lead the (not-dismissed) batsman to adopt a more defensive playstyle.
 Run rates will also vary depending on which player is facing the next over and when the wicket falls in the previous over.
@@
 Plot a shows letter value plots across overs faceted by innings.
-For the team batting in the first innings there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
+For the team batting in the first innings, there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
 Plot b shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
-This indicates that at least one wicket in the previous over leads to lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.
+This indicates that at least one wicket in the previous over leads to a lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.
 

### 2021-04-07 reviews/response.Rmd
reponse update

@@
 
-We would really like to thank the editor and the reviewers for their encouragement and detailed feedback which has helped to improve the paper in terms of both concision and precision and as well as reaching out to more readers.
+We would really like to thank the editor and the reviewers for their encouragement and detailed feedback which has helped to improve the paper in terms of both concision and precision as well as reaching out to more readers.
 The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
@@
 
-> comment 1.
+> I am not convinced that it is required to introduce the concept of grammar of graphics and embed the used visualization in this framework on such a level of detail like in Section 5.
+Of course some elements of the grammar of graphics are used in Section 5.2 to argue about facets, levels, aesthetic variables, and mappings, but if the reader does not know about this framework, the introduction and explanations about this concept/framework provided in Section 5 is not enough to follow the later argumentation.
+I would like to see some more details on the concept/framework on the grammar of graphics, or not using the concepts in the parts of Section 5.2 and paraphrasing it without using the terms from the grammar of graphics.
+I think that it is doable to explain the construction of the visualization without using this concept/framework, because the mappings are quite straight forward.
+I would like to see (a) either extend the introduction to grammar of graphics and the basics needed to argue in more detail about the embedding of your visualisations using this concept/framework, or (b) remove it.
 
@@
 response
+
+# Reviewer 2 comments {-}
+
+> What is the definition of "period"?
+Is the period defined on the grouping or on the granularity?Is the period denoted by P or (R, P) in Definition 6?
+From Pages 6 Line 55 to Page 7 Line 7, period was referred to an integer (7), and two time granularities (one year and 400 years).
+Are they consistent?
+
+Period is defined for a pair of granularities $(G, H)$.
+In Definition 6, the period is defined by $P$, which is the number of granules of $G$ contained in $R$ granules of $H$..
+Thanks for pointing out the inconsistencies.
+We have now made it consistent.

### 2021-04-07 main.Rmd
Reviewer 2 - comment 1 addressed

@@
 \noindent For example, both $day \trianglelefteq week$ and $day \preceq week$ hold, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
-The relationship has period 7.
 
-The relationship $day \trianglelefteq month$ has a more complicated period.
+<!-- # talk about period after definition 6 after you have explained period -->
+<!-- The relationship has period 7. -->
+<!-- The relationship $day \trianglelefteq month$ has a more complicated period.
 If leap years are ignored, each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year.
 With the inclusion of leap years, the grouping period is 400 years.
+-->
 
@@
 
-For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$.
+If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$.
+Condition (2) in Definition \ref{def:periodical} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
+In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
+$P$ is called the \textbf{Period} (@Bettini2000-qk).
+
+For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
+Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).
 

### 2021-04-07 reviews/response.Rmd
Reviewer 2 - comment 1 addressed
Reviewer 2 + other minor issues - addressed

@@
 Period is defined for a pair of granularities $(G, H)$.
-In Definition 6, the period is defined by $P$, which is the number of granules of $G$ contained in $R$ granules of $H$..
+In Definition 6, the period is defined by $P$, which is briefly defined while explaining Definition 6.
 Thanks for pointing out the inconsistencies.
 We have now made it consistent.
+
+> What is the definition of "grouping"? Though "G groups into H" was defined in Definition 5, there is not a notation of grouping $(G, H)$ until we see grouping $(day, month)$ at
+Page 7 Line 7. Please rephrase the sentence each month is a grouping of the same number of days over years on Page 7 Line 5.
+
+We have used the term "pair" instead "grouping" to avoid conflict with the definition of "groups into" and rephrased the sentence.
+
+> What is the definition of aperiodic linear granularities?
+On Page 11, Definition 10, $M_i$ are aperiodic linear granularities.
+If $M_i$ are aperiodic, then is $M$ also an aperiodic linear granularity?
+What is the relationship between aperiodic linear granularity and aperiodic
+cyclic granularity?

### 2021-04-07 main.Rmd
Reviewer 2 + other minor issues - addressed

@@
 $$
-  C_{(G, H)}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H),
+  C_{G, H}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H),
 $$
@@
 
-  * $C_{\ell,m'}(z)$ is circular and $C_{m',m}(z)$ is quasi-circular
+  * $C_{G_{\ell},G_{m'}}(z)$ is circular and $C_{G_{m'},G_{m}}(z)$ is quasi-circular
 \begin{equation} \label{eq:multifromsingle-quasi1}
@@
 
-  * $C_{\ell,m'}(z)$ is quasi-circular and $C_{m',m}(z)$ is circular
+  * $C_{G_{\ell},G_{m'}}(z)$ is quasi-circular and $C_{G_{m'},G_{m}}(z)$ is circular
 \begin{equation} \label{eq:multifromsingle-quasi2}
-C_{G_\ell,G_{m}}(z)  = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)
+C_{G_\ell,G_{m}}(z)  = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{G_{m'},G_{m}}(z) -1}(\vert T_{w} \vert)
 \end{equation}
@@
 weekday/weekend.
-Plot a shows granularity 1 faceted by granularity 2; plots b and c shows the reverse mapping.
+Plot a shows granularity 2 faceted by granularity 1; plots b and c shows the reverse mapping.
 Plot a allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.

### 2021-04-08 main.Rmd
reviewer 1- comment 1 addressed

@@
 
-The grammar of graphics introduced a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk].
-The layered grammar of graphics proposed by @Wickham2009pk gives an alternative and modified parametrization of the grammar, and suggests that graphics are made up of distinct layers of grammatical elements.
+We consider visualizing the distribution of the continuous variable ($v$) conditional on the values of two granularities , $C_i$ and $C_j$.
+Since $C_i$ and $C_j$ are factors or categorical variables, data subsets corresponding to each combination of their levels form a subgroup and the visualization amounts to having displays of distributions for different subgroups.
+A potential way to have such a display is to plot the response variable $(v)$ on the y-axis and the levels of $C_i (C_j)$ on the x-axis, conditional on the levels of $C_j (C_i)$.
+This means, carrying out the same plot corresponding to each level of the conditioning variable.
 
-Drawing from the grammar of graphics, we consider visualizing the distribution of the measured variable $v$ conditional on the values of two granularities, $C_i$ and $C_j$.
-The following layers can be specified:
+<!-- When a data is made up of different groups, it is often informative to compare the distribution of the variables across the groups. -->
 
-- Data: <$C_i$, $C_j$, $v$>;
-- Aesthetic mapping (mapping of variables to elements of the plot): $C_i$ mapped to $x$ position and $v$ to $y$ position;
-- Facet (split plots): $C_j$;
-- Data summarization: any descriptive or smoothing statistics that summarizes the distribution of $v$;
-- Geometric objects (physical representation of the data): any geometry displaying the distribution; for example, boxplot, letter value, violin, ridge or highest density region plots.
+## Plot options
 
-## Data summarization and geometric objects
-
-Plot selection is dictated by the choice of data summarization and geometric objects used for the visualization.
-The basic plot choice for our data structure is one that can display distributions using kernel density estimates or descriptive statistics.
+Plotting each point for each combination through dot plots or rug plots is not useful for our data structure since that leads to a blob of points for each subgroup.
+Further, they are anyway a poor choice for large datasets.
+Hence, the basic plot choice for our data structure is one that can display distributions.
+For displaying the distribution of continuous data graphically, many options are available.
 Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
 Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) boxplots [@Hyndman1996-ft].
-Each type of density display has parameters that need to be estimated from the data.
-Each has its own strengths and weaknesses that should be borne in mind while using them for exploration.
-Visualizing distributions can be uninformative or potentially misleading if data summarization are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
+One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices.
+For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
+
+## Data summarization
+
+Each plot option has parameters that need to be estimated or summarized from the data.
+For example, for histogram, empirical distribution is estimated, for boxplot or its variation, number summaries are computed, for violin or any density-based plots, the density of the variable's distribution needs to be estimated.
+This data estimation or summarization could be potentially misleading if they are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
 Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
@@
 
-## Facet and aesthetic variables {#sec:aesthetics}
+## Comparisons by sub-groups
+
+<!-- While it is often beneficial to compare distributions in this way, there are some caveats that we should be kept in mind while exploring these displays, subject to our data structure as follows: -->
 
@@
 Harmonies tend to be more useful for exploring patterns.
-
 \autoref{fig:allFig}a shows the distribution of half-hourly electricity consumption through letter value plots across months of the year faceted by quarters of the year.
@@
 
-### Interchangeability of mappings {-}
+### Conditioning variable {-}
 
@@
 With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels.
-
 For example, \autoref{fig:allFig}b shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig}c shows the same two cyclic granularities with their mapping reversed.

### 2021-04-08 reviews/response.Rmd
Reviewer 1 - Comment 2 addressed

@@
 We would really like to thank the editor and the reviewers for their encouragement and detailed feedback which has helped to improve the paper in terms of both concision and precision as well as reaching out to more readers.
-The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
+The point by point description of changes are below:
 
@@
 
-> I am not convinced that it is required to introduce the concept of grammar of graphics and embed the used visualization in this framework on such a level of detail like in Section 5.
+Q1.
+I am not convinced that it is required to introduce the concept of grammar of graphics and embed the used visualization in this framework on such a level of detail like in Section 5.
 Of course some elements of the grammar of graphics are used in Section 5.2 to argue about facets, levels, aesthetic variables, and mappings, but if the reader does not know about this framework, the introduction and explanations about this concept/framework provided in Section 5 is not enough to follow the later argumentation.
@@
 
-response
+A1. removed.
 
-> comment 2
+Q2.
+It is not clear from the paragraph in Section 5.1 what basic plot choices are implemented and used and supported for visualisation.
+There is only the list of plots to display descriptive statistics provided.
+It would be important to describe which are implemented and possible to use.
+The later examples use primarily letter-value plots, then also box-plots, violin plots and line charts with area quantile plots.
+The letter-value plots are used in major examples of the paper, therefore I would argue to provide some more details on this visual representation and what the 5 colours and letters mean (e.g.
+LV=letter value, M = median, F = fourths, E = eighths, D = sixteenths, C = thirty-seconds), as it is not as widely known like standard box-plots.
+This would help greatly in better following the argumentation and explanations.
 
-response
+A2. added in Section 5.1
 
-> I have to confess that I do not know anything about the sport cricket and it may be unthinkable for people from countries where cricket is a national sport, but I really had difficulties to follow the example in Section 6.2.
+Q3.
+I have to confess that I do not know anything about the sport cricket and it may be unthinkable for people from countries where cricket is a national sport, but I really had difficulties to follow the example in Section 6.2.
 The basic hierarchical concept of over—inning--match--season was understandable, but I think that some parts need to be revisited and clarified for also reaching readers not knowledgeable in cricket.
@@
 
+A3.
 Thanks for the advice.
@@
 
-> comment 4
+# Reviewer 2 comments
 
-response
+## Notation/Definition clarification
 
-# Reviewer 2 comments {-}
-
-> What is the definition of "period"?
+Q1.What is the definition of "period"?
 Is the period defined on the grouping or on the granularity?Is the period denoted by P or (R, P) in Definition 6?
@@
 
+A1.
 Period is defined for a pair of granularities $(G, H)$.
@@
 
-> What is the definition of "grouping"? Though "G groups into H" was defined in Definition 5, there is not a notation of grouping $(G, H)$ until we see grouping $(day, month)$ at
+Q2. What is the definition of "grouping"? Though "G groups into H" was defined in Definition 5, there is not a notation of grouping $(G, H)$ until we see grouping $(day, month)$ at
 Page 7 Line 7. Please rephrase the sentence each month is a grouping of the same number of days over years on Page 7 Line 5.
 
-We have used the term "pair" instead "grouping" to avoid conflict with the definition of "groups into" and rephrased the sentence.
+A2. We have used the term "pair" instead "grouping" to avoid conflict with the definition of "groups into" and rephrased the sentence.
 
-> What is the definition of aperiodic linear granularities?
+Q3.
+What is the definition of aperiodic linear granularities?
 On Page 11, Definition 10, $M_i$ are aperiodic linear granularities.
@@
 cyclic granularity?
+
+A3.
+
+Q4.
+$w$ in $T_w$ is not well defined for circular granularities in Section 3.2 & 3.5.
+The definition of $T_w$ uses $T_w$  itself in the formula making it hard to understand.
+Relevantly, what is the interpretation of $k$ where w = 0; 1; :::; k-1?
+In my opinion, the number of cycles
+should be defined for cyclic granularities, for example, $w_{B,G}(z) = \lfloor z/P(B;G) \rfloor$, which
+represent the cycle that the granule is in.
+
+A4.
+
+## Places that need clarification
+
+## Other questions of interest.
+
+Q1. Page 6, Definition 5, should S be a contiguous set of integers or not necessary?
+
+A1. Not necessary. 
+<!-- Consider an example, where $B$ is the bottom granularity denoting day, $G$ denotes every alternate business days (Monday, Wednesday and Friday) and $H$ denotes weeks.
+Here, $G \trianglelefteq H$.
+-->
+
+Q2. Page 6, what is the difference between finer than and group into in Definitions 4 and 5? On Page 7, Line 34, they are used as alternative choices. Can you give an example
+that $G \preceq H$ does not imply $G  \trianglelefteq H$, or vice versa? "finer than" is not used a lot in this paper. If the two concepts are similar, why not pick one of them?
+
+A2.
+If $G \trianglelefteq H$, then every granule of $H$ can be expressed as an union of some set of granules in $G$.
+If $G \preceq H$, then every granule of $G$ is a subset of some granule of $H$.
+These are not equivalent.
+Consider an example, where $G_1$ denotes "weekend" and $H_1$ denotes "week".
+Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$.
+Consider another example, with $G_2$ denotes "days" and $H_2$ denotes "business-week".
+Then, $G_2 \npreceq H_2$, but $G_2 \ntrianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
+Further, consider an example with  $H_3$ denoting "public holidays", then $G \npreceq H_3$ and $G \ntrianglelefteq H_3$.
+
+Q3. Page 16, Line 13, multiple order-up quasi-circular granularities. The equations for this
+operation are not trivial to be ignored. Should be added if possible.
+
+A3.
+
+Q4. Can the time granularity concept be extended to continuous time domain?
+
+A4.
+Linear granularities are defined for continuous time domain.
+For continuous time domains, the dependent variable is specified for every value of time $t$, whereas, in applications we generally have the values of the dependent variable for discrete time intervals.
+Hence, we will restrict ourselves to explore only discrete time domain.
+
+## Other minor issues.
+
+corrected.

### 2021-04-08 main.Rmd
Reviewer 1 - Comment 2 addressed

@@
 For displaying the distribution of continuous data graphically, many options are available.
-Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg], letter-value plots [@Hofmann2017-sg] or quantile plots.
+Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg] or quantile plots.
+Letter-Value plots [@Hofmann2017-sg] are an useful advancement to boxplots that defines boxes recursively to visualize the different partitions of a dataset.
+Letter values are order statistics with specific depths.
+The median ($M$) divides the data set into half halves.
+Each of the next steps splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
+are obtained.
+They are useful for displaying the distribution of the data beyond the quantiles specially for large data, when boxplots could hide the interesting patterns at the tails of the dataset by tagging them as outliers.
 Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) boxplots [@Hyndman1996-ft].

### 2021-04-13 main.Rmd
Reviewer 1 - Comment 1 improved

@@
 
-We consider visualizing the distribution of the continuous variable ($v$) conditional on the values of two granularities , $C_i$ and $C_j$.
+The purpose is to visualize the distribution of the continuous variable ($v$) conditional on the values of two granularities , $C_i$ and $C_j$.
 Since $C_i$ and $C_j$ are factors or categorical variables, data subsets corresponding to each combination of their levels form a subgroup and the visualization amounts to having displays of distributions for different subgroups.
-A potential way to have such a display is to plot the response variable $(v)$ on the y-axis and the levels of $C_i (C_j)$ on the x-axis, conditional on the levels of $C_j (C_i)$.
+The response variable $(v)$ is plotted on the y-axis and the levels of $C_i (C_j)$ on the x-axis, conditional on the levels of $C_j (C_i)$.
 This means, carrying out the same plot corresponding to each level of the conditioning variable.
+This is consistent with the widely used grammar of graphics
+which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk, @Wickham2009pk].
 
@@
 
-## Plot options
-
-Plotting each point for each combination through dot plots or rug plots is not useful for our data structure since that leads to a blob of points for each subgroup.
-Further, they are anyway a poor choice for large datasets.
-Hence, the basic plot choice for our data structure is one that can display distributions.
-For displaying the distribution of continuous data graphically, many options are available.
-Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg] or quantile plots.
-Letter-Value plots [@Hofmann2017-sg] are an useful advancement to boxplots that defines boxes recursively to visualize the different partitions of a dataset.
-Letter values are order statistics with specific depths.
-The median ($M$) divides the data set into half halves.
-Each of the next steps splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
-are obtained.
-They are useful for displaying the distribution of the data beyond the quantiles specially for large data, when boxplots could hide the interesting patterns at the tails of the dataset by tagging them as outliers.
-Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) boxplots [@Hyndman1996-ft].
-One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices.
-For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
-
 ## Data summarization
 
-Each plot option has parameters that need to be estimated or summarized from the data.
-For example, for histogram, empirical distribution is estimated, for boxplot or its variation, number summaries are computed, for violin or any density-based plots, the density of the variable's distribution needs to be estimated.
+There are several ways to summarize a distribution like estimating empirical distribution or density of the variable's distribution, computing few number summaries or quantiles.
 This data estimation or summarization could be potentially misleading if they are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
@@
 
-## Comparisons by sub-groups
+<!-- Each plot option has parameters that need to be estimated or summarized from the data. -->
+<!-- For example, for histogram, empirical distribution is estimated, for boxplot or its variation, number summaries are computed, for violin or any density-based plots, the density of the variable's distribution needs to be estimated.
+-->
+
+## Display choices for univariate distributions
+
+<!-- Plotting each point for each combination through dot plots or rug plots is not useful for our data structure since that leads to a blob of points for each subgroup.
+Further, they are anyway a poor choice for large datasets.
+Hence, -->
+
+The basic plot choice for our data structure is one that can display distributions.
+For displaying the distribution of continuous univariate variable graphically, many options are available.
+Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg] or quantile plots.
+Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) boxplots [@Hyndman1996-ft].
+The less commonly used Letter-Value plots [@Hofmann2017-sg] is a midway between boxplots and density plots.
+Letter values are order statistics with specific depths.
+The median ($M$) divides the data set into halves.
+Each of the next steps splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
+are obtained.
+They display the distributions beyond the quantiles specially for large data,
+where boxplots mislabel data points as outliers.
+One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices.
+For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
+
+## Comparison across sub-groups induced by conditioning
 

### 2021-04-14 reviews/response.Rmd
reviewer 2 - clarification 3 addressed

@@
 A3.
+Aperiodic linear granularities are those that can not be defined by Definition 6.
+If $M_i$ is periodic, then $M = \bigcup M_i$ is periodic, but if each $M_i$'s are aperiodic, then $M$ could be periodic or aperiodic.
+In Definition 10, we want $M$ to be aperiodic.
+Definition 10 and introduction of Section 3.3 briefly rephrased to improve clarity.
 

### 2021-04-14 main.Rmd
reviewer 2 - clarification 3 addressed

@@
 
-Aperiodic time granularities are those that cannot be specified as a periodic repetition of a pattern of granules.
-Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant.
+Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodical}.
+Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
+Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant.
 A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
@@
 \begin{definition}\label{def:aperiodic}
-An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups aperiodically into linear granularity $M$ such that the granularities are given by
+An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups into an aperiodic linear granularity $M$ such that the granularities are given by
 \begin{equation}\label{eq:aperiodic}

### 2021-04-14 reviews/response.Rmd
Reviewer 2 - places that need clarification points update

@@
     includes:
-      in_header: preamble2.tex
 fontfamily: mathpazo
@@
 
+Q1.
+Page 7, Definition 6.
+Is "the number of granules of H" finite or infinite?
+If each granule $G(i)$ has the index $i \in Z_{>=0}$, then the phrase is similar as "the number of non-negative integers."
+
+A1. Number of granules of $H$ should be finite. Question
+
+Q2.
+Page 7, Lines 19-20.
+It is claimed that quasi-periodic relationship has a finite number of periods where granularities behave in an anomalous way.
+What does the finite number of periods mean and why is it finite?
+Does it mean a finite number of granularity groupings?
+Does it mean a finite number of irregular cycles?
+By Definition 4 of Bettini & De Sibi (2000), $E_1, .
+.
+.
+, E_z$ are the granularity exceptions, but the number of granularity exceptions does not have to be finite.
+
+A2.
+"Number of periods" rephrased as "spans of time".
+Bettini & De Sibi (2000) states in P57 L17 while introducing quasi-periodic relationships that " except for a finite number of spans of time where they behave in an anomalous way."
+
+Q3. Page 8, Definition 8, what is the "period" P(B,G)? Is it same as P in Definition 6? Is it a mapping or function?
+
+It is the same but stressed to make it clear that the period will be a "function" of two granularities $B$ and $G$.
+
+Q4.
+Page 10, Line 14, Q4 week-of-month is probably not a good example for quasi-circular granularities.
+By Definition 4, week cannot be grouped into month, because there could exist $week(i) \subset month(j)$ when the $i^{th}$ week crosses two months.
+Similarly, week is not finer than month by Definition 5.
+Since periodicity is defined on top of grouping, and circular is defined on top of periodicity, the scenario of Q4 cannot be defined.
+On the other hand, the period length for week-of-month is not an integer, but it is an integer for Q1 - Q3.
+
+A4.
+That's right.
+Alternatively, if we define the linear granularity week such that a week ends when a month ends, then week groups into months in a quasi-periodic way where the period length of week-of-month would be either 4 or 5, since any month could only be composed of either 4 or 5 weeks.
+week-of-month is a cyclic granularity that could be important for certain contexts and hence it makes sense
+to define weeks like that. The first, second, third week-of-month corresponds to first, second, third quarter of the month and so on.
+
+Q5.  Page 10, Definition 9, is R' similar as the period P of the grouping (B,G')? Is R' a constant or variable with irregular mapping?
+
+A5.
+R' is the number of granules of G' in the period P.
+For example, the grouping (day, month) has period 365 (ignoring leap years), then R = 12, since there are 12 months in 365 days.
+Given a period, R' is constant.
+
+Q6.  Page 11, Definition 10, what is the interpretation of n where i = 1, ..., n?
+
+n is the number of Aperiodic linear granularities.
+
+Q7.
+Page 14, Lines 26-27.
+Is z the original index for the bottom granularity?
+If yes, why $C_{katun, baktun}(z) = \lfloor 18 × 20 × z/20 \rfloor mod 20$?
+Shouldn't it be $z mod 20$?
+If not, then what is $z$?
+By Definition 8, $z \in Z$ is the original time index based on the bottom granularity, but here $z$ seems to be defined as the time index based on granularity katun.
+
+A7.
+$z$ is the original index for the bottom granularity $kin$.
+So, $C_{kin, uinal}(z) = z mod 20$.
+For $C_{katun,baktun}(z)$, we first need to have katun in terms of kin through the operation $18*20*z/20$
+
+Q8.
+Page 11, lines 22-33.
+The explanation of Figure 3 is confusing.
+If the week-of-the- semester week type is a quasi-circular granularity, then the day-of-the-semester week type can be also quasi-circular granularity, as all days are nested in weeks.
+Actually the notation should not be "week-of-the-semester week type", but "semester week-of-the- semester week type".
+The relationship between G and H should be clarified.
+It will be clearer if the authors can make a granularity "semester day" (U) from B.
+Therefore we have the following granularities.
+*$Q_{H,M}$ : quasi-circular granularity, 
+semester week -of-the-semester week type
+*$Q_{U,M}$ : quasi-circular granularity, semester day-of-the-semester week type.
+
+Page 20, Line 47, "for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends" is hard to be justified.
+By Figure the IQR is given by the height of blue (LV=F) boxes.
+They are the largest in Quarter 2 and smallest in Quarter 4 for both weekdays and weekends.
+
+rephrased.
+
 ## Other questions of interest.

### 2021-04-14 reviews/response.Rmd
notation/definition clarification

@@
 
-A4.
+A4. R' is the number of granules of $G'$ in each cycle.
+If $z$ belongs to $k^{th}$ granule of $G'$, then $z \in T_k$.
+Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
+The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one cycle, we can find the composition of any granule of $G'$ beyond a cycle since the "pattern" repeats itself along the time domain due to the periodic property.
+The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$.
+The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$.
+Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition 9.
 

### 2021-04-20 main.Rmd
definition 6 - added finite granularity and Z_>=0

@@
 \begin{definition}\label{def:periodical}
-A granularity $G$ is \textbf{periodical} with respect to a granularity $H$ if:
+A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
 (1) $G \trianglelefteq H$; and
-(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
+(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
 \end{definition}

### 2021-04-20 main.Rmd
except in application section week-of-month removed

@@
 With two linear granularities $G$ and $H$, if $G$ *groups into* or *finer than* $H$ then $G$ is of lower order than $H$.
-For example, if the bottom granularity is day, then granularity week will have lower order than month since each week consist of fewer days than each month.
+For example, if the bottom granularity is hour, then granularity day will have lower order than week since each day consist of fewer hours than each week.
 
@@
 $Q_3 =$ day-of-year && $P_3 = 366, 365$\\
-$Q_4 =$ week-of-month && $P_4 = 5, 4$\\
 \bottomrule
@@
 Structurally empty combinations can arise due to the structure of the calendar or hierarchy.
-For example, let $C_i$ be day-of-month with 31 levels and $C_j$ be week-of-month with 5 levels.
-There will be $31\times 5=155$ sets $S_{k\ell}$ corresponding to possible combinations of $C_i$ and $C_j$.
+For example, let $C_i$ be day-of-month with 31 levels and $C_j$ be day-of-year with 365 levels.
+There will be $31\times 365=11315$ sets $S_{k\ell}$ corresponding to possible combinations of $C_i$ and $C_j$.
 Many of these are empty.
-For example, $S_{1,5}$ is empty because the first day of the month can never correspond to the fifth week of the month.
-Hence the pair (day-of-month, week-of-month) is a clash.
+For example, $S_{1,5}$ is empty because the first day of the month can never correspond to the fifth day of the year.
+Hence the pair (day-of-month, day-of-year) is a clash.
 

### 2021-04-21 main.Rmd
viz section consistency

@@
 $z$ denotes the index set,
-$R'$ is the number of granules of $G'$ in each repetition of the grouping,
+$w$ denotes the index of $G'$,
+$R'$ is the number of granules of $G'$ in each period of $(B, G')$,
 $T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
@@
 For example, day-of-year is quasi-periodic with either 365 or 366 granules of $B$ (days) within each period of $G'$ (years).
-The pattern repeats every $R'=400$ years.
-So $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
+The pattern repeats every $4$ years (ignoring leap seconds).
+Hence $R'= 4$.
+$Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
 
@@
 $z$ denotes the index set,
-$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$, $n$ being the number of  aperiodic linear granularities in consideration.
 \end{definition}
@@
 The basic plot choice for our data structure is one that can display distributions.
-For displaying the distribution of continuous univariate variable graphically, many options are available.
-Displays based on descriptive statistics include variations of box plots [@Tukey1977-jx] such as notched box plots [@McGill1978-hg] or quantile plots.
-Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) boxplots [@Hyndman1996-ft].
+For displaying the distribution of continuous univariate variable, many options are available.
+Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in [@stryjewski201040].
+They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.
+Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft].
 The less commonly used Letter-Value plots [@Hofmann2017-sg] is a midway between boxplots and density plots.
@@
 The median ($M$) divides the data set into halves.
-Each of the next steps splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
+Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
 are obtained.
-They display the distributions beyond the quantiles specially for large data,
+They are useful for displaying the distributions beyond the quartiles specially for large data,
 where boxplots mislabel data points as outliers.
@@
 Harmonies tend to be more useful for exploring patterns.
-\autoref{fig:allFig}a shows the distribution of half-hourly electricity consumption through letter value plots across months of the year faceted by quarters of the year.
-This plot does not work because quarter-of-the-year clashes with month-of-the-year, leading to empty subsets.
+\autoref{fig:allFig}a shows the distribution of half-hourly electricity consumption through letter value plots across months of the year conditional on quarters of the year.
+This plot does not work because quarter-of-year clashes with month-of-year, leading to empty subsets.
 For example, the first quarter never corresponds to December.
@@
 With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels.
-For example, \autoref{fig:allFig}b shows the letter value plot across weekday/weekend faceted by quarters of the year and \autoref{fig:allFig}c shows the same two cyclic granularities with their mapping reversed.
+For example, \autoref{fig:allFig}b shows the letter value plot across weekday/weekend partitioned by quarters of the year and \autoref{fig:allFig}c shows the same two cyclic granularities with their mapping reversed.
 \autoref{fig:allFig}b helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig}c helps to compare quarters within weekend and weekday.

### 2021-04-21 reviews/response.Rmd
Places that need clarification section
fine tuning responses for comments by Reviewer 2

@@
 Q1.What is the definition of "period"?
-Is the period defined on the grouping or on the granularity?Is the period denoted by P or (R, P) in Definition 6?
+Is the period defined on the grouping or on the granularity?
+Is the period denoted by P or (R, P) in Definition 6?
 From Pages 6 Line 55 to Page 7 Line 7, period was referred to an integer (7), and two time granularities (one year and 400 years).
@@
 In Definition 6, the period is defined by $P$, which is briefly defined while explaining Definition 6.
-Thanks for pointing out the inconsistencies.
-We have now made it consistent.
+We have now made the notations consistent.
 
@@
 
-A1. Number of granules of $H$ should be finite. Question
+A1. Number of granules of $H$ should be finite.
 
@@
 
-It is the same but stressed to make it clear that the period will be a "function" of two granularities $B$ and $G$.
+A3. It is the same but stressed to make it clear that the period will be a "function" of two granularities $B$ and $G$.
 
@@
 
-A4.
-That's right.
-Alternatively, if we define the linear granularity week such that a week ends when a month ends, then week groups into months in a quasi-periodic way where the period length of week-of-month would be either 4 or 5, since any month could only be composed of either 4 or 5 weeks.
-week-of-month is a cyclic granularity that could be important for certain contexts and hence it makes sense
-to define weeks like that. The first, second, third week-of-month corresponds to first, second, third quarter of the month and so on.
+A4. removed
 
@@
 
-n is the number of Aperiodic linear granularities.
+A6. n is the number of Aperiodic linear granularities. Added in Definition 10.
 
@@
 $z$ is the original index for the bottom granularity $kin$.
-So, $C_{kin, uinal}(z) = z mod 20$.
-For $C_{katun,baktun}(z)$, we first need to have katun in terms of kin through the operation $18*20*z/20$
+But the second line is replaced since the point is to show how the value of a multiple-order-up granularity can be obtained as a function of single-order-up granularities.
+It is sufficient to know the values of the single order granularities at $z$ and not the exact value of $z$, so we chose to not expand it using $z$.
 
@@
 
+A8. done (to be)
+
+Q9.
 Page 20, Line 47, "for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends" is hard to be justified.
@@
 
-rephrased.
+A9. rephrased.
 
@@
 
-corrected.
+done.

### 2021-04-21 main.Rmd
fine tuning responses for comments by Reviewer 2

@@
 C_{\text{uinal}, \text{baktun}}(z)  &=  C_{\text{uinal}, \text{tun}}(z) + P(\text{uinal}, \text{tun})C_{\text{tun},\text{katun}}(z) + P(\text{uinal},\text{katun})C_{\text{katun}, \text{baktun}}(z) \\
-               &=  \lfloor z/20\rfloor \mod 18 + 18\lfloor 18\times z/20\rfloor \mod 20
-                + 18\times20\lfloor 18\times20\times z/20\rfloor \mod 20.
+         &=  C_{\text{uinal}, \text{tun}}(z) + 18 \times{ C_{\text{tun},\text{katun}}(z)} + 18 \times 20 \times{C_{\text{katun}, \text{baktun}}(z)}
 \end{align*}
+
+<!--                &=  \lfloor \frac{z}{20}\rfloor \mod 18 + 18\lfloor  \frac{z}{20 \times18} \rfloor \mod 20 -->
+<!--                 + 18 \times20 \lfloor \frac{z}{(20\times18\times20} \rfloor \mod 20 -->
 
@@
 
-  * $C_{G_{\ell},G_{m'}}(z)$ is circular and $C_{G_{m'},G_{m}}(z)$ is quasi-circular
+  * $C_{G_{\ell},G_{m'}}$ is circular and $C_{G_{m'},G_{m}}$ is quasi-circular
 \begin{equation} \label{eq:multifromsingle-quasi1}
@@
 
-  * $C_{G_{\ell},G_{m'}}(z)$ is quasi-circular and $C_{G_{m'},G_{m}}(z)$ is circular
+  * $C_{G_{\ell},G_{m'}}$ is quasi-circular and $C_{G_{m'},G_{m}}$ is circular
 \begin{equation} \label{eq:multifromsingle-quasi2}
@@
 Plot c switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
-For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends.
+<!-- For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends. -->
 

### 2021-04-21 reviews/response.Rmd
comments revisited

@@
 
-A4. R' is the number of granules of $G'$ in each cycle.
+A4.
+$w$ denotes the index set of $G'$.
+$R'$ is the number of granules of $G'$ in each period.$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ does not use the definition of $T_w$.
+For example, if $B$ is `day` and $G'$ is `month`, then $G'(0) = {B(0)\cup B(2)\cup \dots \cup B(30)}$, $G'(1) = {B(31)\cup B(32)\cup \dots \cup B(58)}$ corresponding to January and February.
+Then $T_0 = \{0, 1, 2, \dots, 30\}$ and $T_1 = \{31, 32, \dots, 58\}$.
 If $z$ belongs to $k^{th}$ granule of $G'$, then $z \in T_k$.
-Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
-The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one cycle, we can find the composition of any granule of $G'$ beyond a cycle since the "pattern" repeats itself along the time domain due to the periodic property.
+For example, if $z = 31$ then $z \in T_1$ and thus $k = 1$.
+(Should this be added to paper?)
+
+<!-- Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$.
+The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property.
 The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$.
@@
 Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition 9.
+-->
 
@@
 
-A1. Number of granules of $H$ should be finite.
+A1. Number of granules of $H$ should be finite. Definition rephrased to account for that.
 
@@
 For example, the grouping (day, month) has period 365 (ignoring leap years), then R = 12, since there are 12 months in 365 days.
-Given a period, R' is constant.
+Given a period P, R' is constant.
 
@@
 
-A9. rephrased.
+A9. removed.
 
@@
 
-Q1. Page 6, Definition 5, should S be a contiguous set of integers or not necessary?
+Q1. Page 6, Definition 5, should $S$ be a contiguous set of integers or not necessary?
 
@@
 Consider another example, with $G_2$ denotes "days" and $H_2$ denotes "business-week".
-Then, $G_2 \npreceq H_2$, but $G_2 \ntrianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
-Further, consider an example with  $H_3$ denoting "public holidays", then $G \npreceq H_3$ and $G \ntrianglelefteq H_3$.
+Then, $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
+Further, consider an example with  $H_3$ denoting "public holidays", then $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.    
+(Should this be added to paper?)
 
@@
 
-A3.
+A3. Exactly equations (5) and (6) are to be utilized for this operation (with sides interchanged) and hence not added again.
 
@@
 A4.
-Linear granularities are defined for continuous time domain.
-For continuous time domains, the dependent variable is specified for every value of time $t$, whereas, in applications we generally have the values of the dependent variable for discrete time intervals.
-Hence, we will restrict ourselves to explore only discrete time domain.
+For continuous time domains, the dependent variable is specified for every value of time $t$, whereas, in applications we generally have the values of the dependent variable for discrete time domain Hence, we will restrict ourselves to explore only discrete time domain.
 

### 2021-04-21 scripts/main.R

@@ -65,7 +65,7 @@ scene1 <- VIC %>%
   ylab("") +
   xlab("") +
   # xlab("weekend/weekday") +
-  scale_fill_brewer(palette = "Dark2") +
+  scale_fill_brewer(type = "seq", direction = -1) +
   theme(legend.position = "right") +
   ggtitle("") +
   scale_y_log10() +
@@ -83,14 +83,16 @@ scene2 <- VIC %>%
   ) +
   ylab("") +
   xlab("quarters of the year") +
-  scale_fill_brewer(palette = "Paired") +
+  scale_fill_brewer(type = "seq", direction = -1) +
   theme(legend.position = "right") +
   ggtitle("") +
   scale_y_log10() +
   theme_minimal() +
   theme(
     #panel.grid.major = element_line(colour = "#E0E0E0"),
-    panel.border = element_rect(colour = "#E0E0E0", fill = NA))
+    panel.border = element_rect(colour = "#E0E0E0",
+                                fill = NA),
+    panel.grid.major.x = element_blank())
 
 scene3 <- VIC %>%
   prob_plot(
@@ -102,13 +104,15 @@ scene3 <- VIC %>%
   ylab("") +
   xlab("months of the year") +
   theme(legend.position = "right") +
-  scale_fill_brewer(palette = "Set2") +
+  scale_fill_brewer(type = "seq", direction = -1) +
   ggtitle("") +
   scale_y_log10() +
   theme_minimal() +
   theme(
     #panel.grid.major = element_line(colour = "#E0E0E0"),
-    panel.border = element_rect(colour = "#E0E0E0", fill = NA))
+    panel.border = element_rect(colour = "#E0E0E0",
+                                fill = NA),
+    panel.grid.major.x = element_blank())
 
 gg_fig <- ggarrange(
   scene3,
@@ -276,7 +280,7 @@ cricket_tsibble %>%
     hierarchy_model,
     plot_type = "lv"
   ) +
-  scale_fill_brewer(palette = "Dark2") +
+  scale_fill_brewer(type = "seq", palette = "Blues", direction = -1) +
   # ggtitle("(a) Runs per over across over faceted by inning") +
   theme(legend.position = "right") +
   ggtitle("a") +
@@ -286,8 +290,9 @@ cricket_tsibble %>%
   theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
 ) + 
   theme(plot.title = element_text(face = "bold"),
[…]

### 2021-04-22 main.Rmd
comments revisited
made periodic instead of periodical throughout

@@
 
-\begin{definition}\label{def:periodical}
-A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
+\begin{definition}\label{def:periodic}
+A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
 (1) $G \trianglelefteq H$; and
@@
 If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$.
-Condition (2) in Definition \ref{def:periodical} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
+Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
 In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
@@
 
-For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
+For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
 Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).
 
-Granularities can also be periodical with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+Granularities can also be periodic with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy].
 In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
@@
 
-Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodical}.
+Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}.
 Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
-Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant.
+These are events which repeat but there is no reasonably small period within which their behavior remains constant.
 A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
@@
 For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
+In R package `gravitas` [@R-gravitas], boxplots, violin, ridge, letter-value, line and area quantile plots are implemented, but it is potentially possible to use any plots which can display the distribution of the data.
 

### 2021-04-22 main.Rmd
had to remove the definition line in the yaml because it was duplicated in one of the packages used
typo in Dfinition 6

@@
 (1) $G \trianglelefteq H$; and
-(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
+(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z}_{\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
 \end{definition}
@@
 
-Granularities can also be periodic with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+Granularities can also be periodic with respect to other granularities, "except for a finite number of spans of
+time where they behave in an anomalous way"; these are called **quasi-periodic** relationships [@Bettini2000-vy].
 In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

### 2021-04-22 main.Rmd
comments update
definition without section

@@
   \def\mod{~\text{mod}~}
-  \newtheorem{definition}{Definition}
   \usepackage{mathptmx}
   \usepackage{caption}
+  \newtheorem{defn}{Definition}
   \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
@@
 
-\begin{definition}\label{def:definition}
+\begin{defn}\label{def:timedom}
 A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:index set}
+\begin{defn}\label{def:index set}
 The \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, uniquely maps the time instants to the set of non-negative integers.
-\end{definition}
+\end{defn}
 
-\begin{definition}\label{def:linear}
+\begin{defn}\label{def:linear}
 A \textbf{linear granularity} is a mapping $G$ from the index set, $Z$, to subsets of the time domain such that:
@@
 Each non-empty subset $G(i)$ is called a \textbf{granule}.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:finerthan}
+\begin{defn}\label{def:finerthan}
 A linear granularity $G$ is \textbf{finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
 $G(i) \subset H(j).$
-\end{definition}
+\end{defn}
 
-\begin{definition}\label{def:groupsinto}
+\begin{defn}\label{def:groupsinto}
 A linear granularity $G$ \textbf{groups into} a linear granularity $H$, denoted
 $G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:periodic}
+\begin{defn}\label{def:periodic}
 A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
@@
 (2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:order}
+\begin{defn}\label{def:order}
 The \textbf{order} of a linear granularity is the level of coarseness associated with a linear granularity.
 A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:circular}
+\begin{defn}\label{def:circular}
 A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$ if
@@
 $B$ groups periodically into $G$ with regular mapping and period $P(B, G)$.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:quasicircular}
+\begin{defn}\label{def:quasicircular}
 A \textbf{quasi-circular granularity} $Q_{B, G'}$ is formed when bottom granularity $B$ groups periodically into linear granularity $G'$ with irregular mapping such that the granularities are given by
@@
 and $\vert T_w \vert$ is the cardinality of set $T_w$.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:aperiodic}
+\begin{defn}\label{def:aperiodic}
 An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups into an aperiodic linear granularity $M$ such that the granularities are given by
@@
 $T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$, $n$ being the number of  aperiodic linear granularities in consideration.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:hierarchy}
+\begin{defn}\label{def:hierarchy}
 The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order.
@@
 Then $K \equiv P(\ell,m)$ represents the period length of the grouping $(G_{\ell}, G_{m})$, if $C_{G_{\ell}, G_{m}}$ is a circular granularity and $K \equiv k(\ell,m)$ represents the operation to obtain $G_{m}$ from $G_{\ell}$, if $C_{G_{\ell}, G_{m}}$ is quasi-circular.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:norderup}
+\begin{defn}\label{def:norderup}
 The hierarchy table contains \textbf{multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
 A \textbf{single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:clash}
+\begin{defn}\label{def:clash}
 A \textbf{clash} is a pair of cyclic granularities that contains empty combinations of categories.
-\end{definition}
+\end{defn}
 
-\begin{definition}\label{def:harmony}
+\begin{defn}\label{def:harmony}
 A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
-\end{definition}
+\end{defn}
 

### 2021-04-22 main.Rmd
section and article reference fixed

@@
 
-Granularities can also be periodic with respect to other granularities, "except for a finite number of spans of
-time where they behave in an anomalous way"; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+Granularities can also be periodic with respect to other granularities, _"except for a finite number of spans of time where they behave in an anomalous way"_; these are called **quasi-periodic** relationships [@Bettini2000-vy].
 In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
@@
 For displaying the distribution of continuous univariate variable, many options are available.
-Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in [@stryjewski201040].
+Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in [@Wickham2011-yl].
 They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.
@@
 
-## Comparison across sub-groups induced by conditioning
+## Comparison across sub-groups induced by conditioning {#sec:aesthetics}
 

### 2021-04-22 reviews/response.Rmd
reviews rephrased

@@
 
-A1. removed.
+A1. Removed.
 
@@
 
-A2. added in Section 5.1
+A2. Added in Section 5.1
 
@@
 
-A4.
-$w$ denotes the index set of $G'$.
-$R'$ is the number of granules of $G'$ in each period.$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ does not use the definition of $T_w$.
+A4. Definition of $w$ and $R'$ included in Definition 9.
+
+$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ does not use the definition of $T_w$.
 For example, if $B$ is `day` and $G'$ is `month`, then $G'(0) = {B(0)\cup B(2)\cup \dots \cup B(30)}$, $G'(1) = {B(31)\cup B(32)\cup \dots \cup B(58)}$ corresponding to January and February.
@@
 For example, if $z = 31$ then $z \in T_1$ and thus $k = 1$.
-(Should this be added to paper?)
 
@@
 Given a period P, R' is constant.
+Another example included in the paper.
 
@@
 $z$ is the original index for the bottom granularity $kin$.
-But the second line is replaced since the point is to show how the value of a multiple-order-up granularity can be obtained as a function of single-order-up granularities.
-It is sufficient to know the values of the single order granularities at $z$ and not the exact value of $z$, so we chose to not expand it using $z$.
+The equation is modified since we removed the line expressing the multiple-order-up granularities in terms of $z$.
+The point is to show how the value of a multiple-order-up granularity can be obtained when values of single-order-up granularities are available.
+It is sufficient to know the values of the single order-up-granularities at $z$ and not the exact value of $z$, so we chose to not expand it using $z$.
 
@@
 
-A4.
-For continuous time domains, the dependent variable is specified for every value of time $t$, whereas, in applications we generally have the values of the dependent variable for discrete time domain Hence, we will restrict ourselves to explore only discrete time domain.
+A4. It is not clear how to extend the granularity concepts to continuous time.
 

### 2021-04-27 main.Rmd
changing the order of quantiles and size

--- 2021-04-22 main.Rmd
@@
   \def\mod{~\text{mod}~}
-  \newtheorem{definition}{Definition}
   \usepackage{mathptmx}
@@
 
-\begin{definition}\label{def:periodic}
-A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
+\begin{definition}\label{def:periodical}
+A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
 (1) $G \trianglelefteq H$; and
@@
 If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$.
-Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
+Condition (2) in Definition \ref{def:periodical} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
 In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
@@
 
-For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
+For example, day is periodical with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodical with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
 Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).
 
-Granularities can also be periodic with respect to other granularities, "except for a finite number of spans of
-time where they behave in an anomalous way"; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+Granularities can also be periodical with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy].
 In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
@@
 
-Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}.
+Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodical}.
 Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
-These are events which repeat but there is no reasonably small period within which their behavior remains constant.
+Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant.
 A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
@@
 For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
-In R package `gravitas` [@R-gravitas], boxplots, violin, ridge, letter-value, line and area quantile plots are implemented, but it is potentially possible to use any plots which can display the distribution of the data.
 

### 2021-04-27 scripts/main.R

@@ -365,4 +365,7 @@ cricket_data %>%
     complete = TRUE,
     panel.grid.major = element_line(colour = "#E0E0E0"),
     panel.border = element_rect(colour = "#E0E0E0", fill = NA)
-  ) + scale_colour_manual(values = c("#7093DB", 	"#3A5FCD", "#22316C"))
+  ) + scale_colour_manual(values = c("#7093DB", 	"#3A5FCD", "#22316C")) +
+  guides(color = guide_legend(reverse = TRUE)) +
+  scale_x_discrete(labels = c("0", "1+")) +
+  geom_line(size = 0.8)

### 2021-04-27 reviews/response.Rmd
groups into and finer than nlot equivalent definition added

@@
 These are not equivalent.
-Consider an example, where $G_1$ denotes "weekend" and $H_1$ denotes "week".
-Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$.
-Consider another example, with $G_2$ denotes "days" and $H_2$ denotes "business-week".
-Then, $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
-Further, consider an example with  $H_3$ denoting "public holidays", then $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.    
-(Should this be added to paper?)
+Examples added in the paper after their definition in Page 6.
 

### 2021-04-27 main.Rmd
groups into and finer than nlot equivalent definition added

@@
 \noindent For example, both $day \trianglelefteq week$ and $day \preceq week$ hold, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
+These definitions are not equivalent.
+Consider another example, where $G_1$ denotes "weekend" and $H_1$ denotes "week".
+Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$.
+Further, with $G_2$ denoting "days" and $H_2$ denoting "business-week", $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
+Moreover, with  $H_3$ denoting "public holidays", $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.
 
@@
 For displaying the distribution of continuous univariate variable, many options are available.
-Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in [@stryjewski201040].
+Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in @Wickham2012-vr.
 They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.

### 2021-04-28 main.Rmd
added aperiodic new example

@@
 (ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated through (a) linear and (b) stacked displays of time.
-The linear display shows granularities days, weeks, semester weeks, semester week type distributed over linear time.
-Here a semester lasts for 18 weeks and 2 days, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week, an in-session period of 7 weeks, a 1-week study break, before final exams which continue for 16 days.
-This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 128 days will be a quasi-circular granularity with repeating patterns.
-$A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years.
+The linear display shows granularities days, weeks, semester days, and stages of a semester distributed over linear time.
+Here a semester lasts for 18 weeks and 2 days, each consisting of 5 stages.
+It starts with an orientation stage, followed by a first stage of in-session period (6 weeks), a break (1 week), a second stage of in-session period (7 weeks), a 1-week study break stage, before final exams stage which continue for 16 days.
+This distribution of semester days within a semester remains the same for every semesters and hence $Q_{B', M}$ with $P$ = 128 days will be a quasi-circular granularity with repeating patterns.
+$A_{B, M}$ will be an aperiodic cyclic granularity since
+the placement of the semester within an year varies across years and there is no definite number of days after which the semester begins every year.
 
 For example, consider the school semester shown in \autoref{fig:aperiodic-example}.
-Let the linear granularities $M_1$ and $M_2$ denote the in-session semester period and semester break period respectively.
-Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the "semester week type" are aperiodic with respect to days ($B$) or weeks ($G$).
-Hence $A_{B, M}$ denoting day-of-the-"semester week type" would be an aperiodic cyclic granularity, because the placement of the semester within an year would vary across years.
-Here, $Q_{H, M}$ denoting week-of-the-"semester week type" would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years.
+Let the linear granularities $M_1$ and $M_2$ denote the teaching and non-teaching stages of the semester respectively.
+Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the "stages" of the semester are aperiodic with respect to days ($B$) or weeks ($G$).
+Hence $A_{B, M}$ denoting day-of-the-stage would be an aperiodic cyclic granularity, because the placement of the semester within an year would vary across years.
+Here, $Q_{B', M}$ denoting semester-day-of-the-stage would be a quasi-circular granularity since the distribution of semester days within a semester is assumed to remain constant over years.
 

### 2021-04-28 scripts/main.R

@@ -18,7 +18,8 @@ include_graphics("img/linear-ex.png")
 include_graphics("img/circular-ex.png")
 
 ## ----aperiodic-example
-include_graphics("img/aperiodic-ex3.png")
+include_graphics("img/aperiodic_revised.png")
+#include_graphics("img/aperiodic-ex3.png")
 
 ## ----tab-mayan
 table_mayan <- tibble(

### 2021-04-30 reviews/response.Rmd
proof read viz

@@
 $T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ does not use the definition of $T_w$.
-For example, if $B$ is `day` and $G'$ is `month`, then $G'(0) = {B(0)\cup B(2)\cup \dots \cup B(30)}$, $G'(1) = {B(31)\cup B(32)\cup \dots \cup B(58)}$ corresponding to January and February.
+For example, if $B$ is `day` and $G'$ is `month`, then $G'(0) = {B(0)\cup B(1)\cup \dots \cup B(30)}$, $G'(1) = {B(31)\cup B(32)\cup \dots \cup B(58)}$ corresponding to January and February.
 Then $T_0 = \{0, 1, 2, \dots, 30\}$ and $T_1 = \{31, 32, \dots, 58\}$.

### 2021-04-30 main.Rmd
proof read viz

@@
 The less commonly used Letter-Value plots [@Hofmann2017-sg] is a midway between boxplots and density plots.
-Letter values are order statistics with specific depths.
-The median ($M$) divides the data set into halves.
+Letter values are order statistics with specific depths, for example, the median ($M$) is a letter value which divides the data set into halves.
 Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
@@
 One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices.
-For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
+For example, boxplots might give an idea of the distribution at a glance while occupying less space by design, but they obscure multimodality.
+Interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
 
-## Comparison across sub-groups induced by conditioning
+## Comparison across sub-groups induced by conditioning {#sec:aesthetics}
 
@@
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case the $A_k$s are compared against each other within each group.
+Hence, in this case the $A_k$'s are compared against each other within each group.
 With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels.

### 2021-05-01 main.Rmd
Definition without section

@@
   \usepackage{caption}
+  \newtheorem{defn}{Definition}
   \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
@@
 
-\begin{definition}\label{def:definition}
+\begin{defn}\label{def:definition}
 A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
-\end{definition}
+\end{defn}
 
@@
 
-\begin{definition}\label{def:periodical}
+\begin{defn}\label{def:periodical}
 A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
@@
 One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices.
-For example, boxplots might give an idea of the distribution at a glance while occupying less space by design, but they obscure multimodality.
-Interpretation of density estimates and histograms may change depending on the bandwidth and bin widths respectively.
+For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and binwidths respectively.
+In R package `gravitas` [@R-gravitas], boxplots, violin, ridge, letter-value, line and area quantile plots are implemented, but it is potentially possible to use any plots which can display the distribution of the data.
 

### 2021-05-01 main.Rmd
periodical to periodic

@@
 
-\begin{defn}\label{def:periodical}
-A granularity $G$ is \textbf{periodical} with respect to a finite granularity $H$ if:
+\begin{defn}\label{def:periodic}
+A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
 (1) $G \trianglelefteq H$; and
@@
 
-Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodical}.
+Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}.
 Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.

### 2021-05-01 main.Rmd
notation Z>=0 fixed

@@
 (1) $G \trianglelefteq H$; and
-(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z_\geq 0}$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
+(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z}_{\ge {0}}$ , if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
 \end{defn}

### 2021-05-01 reviews/response.Rmd
minor changes

@@
 The definition of $T_w$ uses $T_w$ itself in the formula making it hard to understand.
-Relevantly, what is the interpretation of $k$ where w = 0; 1; :::; k-1?
+Relevantly, what is the interpretation of $k$ where $w = 0, 1, \dots k-1$?
 In my opinion, the number of cycles
@@
 
-A8. done (to be)
+A8.
+Sem week is removed and only "semester day" abbreviated as "sem day" is introduced.
+"Day"($B$) and "Sem day"($B'$) differs in their structure as granules of $B'$ corresponds to semesters only.
+"Stage" is used instead of "Semester week type" to refer to different stages of a semester.
+Then $A_{B,M}$ is aperiodic, whereas $Q_{B', M}$ is quasi-periodic.
 

### 2021-05-01 main.Rmd
minor changes

@@
 
-Granularities can also be periodic with respect to other granularities, except for a finite number of periods where they behave in an anomalous way; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+Granularities can also be periodic with respect to other granularities, _"except for a finite span of time where they behave in an anomalous way"_; these are called **quasi-periodic** relationships [@Bettini2000-vy].
 In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
@@
 Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
-Examples include public holidays which repeat every year, but there is no reasonably small period within which their behavior remains constant.
+Examples include public holidays which repeat every year, but there is no reasonably small span of time within which their behavior remains constant.
 A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
@@
 (ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated through (a) linear and (b) stacked displays of time.
-The linear display shows granularities days, weeks, semester days, and stages of a semester distributed over linear time.
+The linear display shows granularities days ($B$), weeks ($G$), semester days ($B'$), and stages of a semester ($M$) distributed over linear time.
+Granules of $B'$ is only defined for days when the semester is running.
 Here a semester lasts for 18 weeks and 2 days, each consisting of 5 stages.
@@
 
+The linear display shows granularities days ($B$), weeks ($G$), , and stages of a semester ($M$) distributed over linear time.
+
 For example, consider the school semester shown in \autoref{fig:aperiodic-example}.
@@
 Hence $A_{B, M}$ denoting day-of-the-stage would be an aperiodic cyclic granularity, because the placement of the semester within an year would vary across years.
-Here, $Q_{B', M}$ denoting semester-day-of-the-stage would be a quasi-circular granularity since the distribution of semester days within a semester is assumed to remain constant over years.
+Here, $Q_{B', M}$ denoting semester-day-of-the-stage would be a quasi-circular granularity
+since the distribution of semester days within a semester is assumed to remain constant over years.
+Here semester-day is denoted by "sem day" ($B'$) and its granules are only defined for the span of the semesters.
 

### 2021-05-01 main.Rmd
gram and spell check

@@
 Social and economic data are often collected and reported at coarse temporal scales such as monthly, quarterly or annually.
-With recent advancement in technology, more and more data are recorded at much finer temporal scales.
+With recent advancements in technology, more and more data are recorded at much finer temporal scales.
 Energy consumption may be collected every half an hour, energy supply may be collected every minute, and web search data might be recorded every second.
@@
 
-Calendar-based graphics [@wang2020calendar] are useful in visualizing patterns in the weekly and monthly structure, and are helpful when checking for the effects of weekends or special days.
+Calendar-based graphics [@wang2020calendar] are useful in visualizing patterns in the weekly and monthly structure and are helpful when checking for the effects of weekends or special days.
 Any temporal data at sub-daily resolution can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year, or another sub-daily deconstruction of time.
@@
 
-For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of same number of days across years.
+For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of the same number of days across years.
 Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).
@@
 \autoref{fig:circular-dow} illustrates some linear and cyclical granularities.
-Cyclical granularities are constructed be cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
+Cyclical granularities are constructed by cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
 $B, G, H$ (day, week, fortnight, respectively) are linear granularities.
@@
 Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
-Circular granularities can be expressed using modular arithmetic due to their regular mapping.
-}
+Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
 \label{tab:definitions}
@@
 Here a semester lasts for 18 weeks and 2 days, each consisting of 5 stages.
-It starts with an orientation stage, followed by a first stage of in-session period (6 weeks), a break (1 week), a second stage of in-session period (7 weeks), a 1-week study break stage, before final exams stage which continue for 16 days.
-This distribution of semester days within a semester remains the same for every semesters and hence $Q_{B', M}$ with $P$ = 128 days will be a quasi-circular granularity with repeating patterns.
+It starts with an orientation stage, followed by a first stage of the in-session period (6 weeks), a break (1 week), a second stage of the in-session period (7 weeks), a 1-week study break stage, before final exams stage which continue for 16 days.
+This distribution of semester days within a semester remains the same for every semester and hence $Q_{B', M}$ with $P$ = 128 days will be a quasi-circular granularity with repeating patterns.
 $A_{B, M}$ will be an aperiodic cyclic granularity since
-the placement of the semester within an year varies across years and there is no definite number of days after which the semester begins every year.
-
-The linear display shows granularities days ($B$), weeks ($G$), , and stages of a semester ($M$) distributed over linear time.
+the placement of the semester within a year varies across years and there is no definite number of days after which the semester begins every year.
 
@@
 Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the "stages" of the semester are aperiodic with respect to days ($B$) or weeks ($G$).
-Hence $A_{B, M}$ denoting day-of-the-stage would be an aperiodic cyclic granularity, because the placement of the semester within an year would vary across years.
+Hence $A_{B, M}$ denoting day-of-the-stage would be an aperiodic cyclic granularity because the placement of the semester within a year would vary across years.
 Here, $Q_{B', M}$ denoting semester-day-of-the-stage would be a quasi-circular granularity
@@
 For example,  the French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
-Nevertheless, for any calendar a hierarchy table can be defined.
+Nevertheless, for any calendar, a hierarchy table can be defined.
 Note that it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
@@
 
-Effective exploration and visualization benefits from well-organized data structures.
+Effective exploration and visualization benefit from well-organized data structures.
 @wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data.
@@
 
-An example of a harmony is where $C_i$ and $C_j$ denote day-of-week and month-of-year respectively.
+An example of harmony is where $C_i$ and $C_j$ denote day-of-week and month-of-year respectively.
 So $C_i$ will have 7 levels while $C_j$ will have 12 levels, giving $12\times7=84$ sets $S_{k\ell}$.
@@
 
-The purpose is to visualize the distribution of the continuous variable ($v$) conditional on the values of two granularities , $C_i$ and $C_j$.
+The purpose is to visualize the distribution of the continuous variable ($v$) conditional on the values of two granularities, $C_i$ and $C_j$.
 Since $C_i$ and $C_j$ are factors or categorical variables, data subsets corresponding to each combination of their levels form a subgroup and the visualization amounts to having displays of distributions for different subgroups.
@@
 
-There are several ways to summarize a distribution like estimating empirical distribution or density of the variable's distribution, computing few number summaries or quantiles.
+There are several ways to summarize a distribution like estimating empirical distribution or density of the variable's distribution, computing a few number summaries or quantiles.
 This data estimation or summarization could be potentially misleading if they are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
@@
 The basic plot choice for our data structure is one that can display distributions.
-For displaying the distribution of continuous univariate variable, many options are available.
-Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variation such as notched box plots [@McGill1978-hg] or other variations as mentioned in @Wickham2012-vr.
+For displaying the distribution of a continuous univariate variable, many options are available.
+Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variants such as notched box plots [@McGill1978-hg] or other variations as mentioned in @Wickham2012-vr.
 They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.
 Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft].
-The less commonly used Letter-Value plots [@Hofmann2017-sg] is a midway between boxplots and density plots.
-Letter values are order statistics with specific depths, for example, the median ($M$) is a letter value which divides the data set into halves.
-Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$) etc.
+The less commonly used Letter-Value plots [@Hofmann2017-sg] is midway between boxplots and density plots.
+Letter values are order statistics with specific depths, for example, the median ($M$) is a letter value that divides the data set into halves.
+Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$), etc.
 are obtained.
-They are useful for displaying the distributions beyond the quartiles specially for large data,
+They are useful for displaying the distributions beyond the quartiles especially for large data,
 where boxplots mislabel data points as outliers.
@@
 Plot a shows a clash since there are empty combinations.
-Plots b and c show harmonies since each quarter includes both weekdays and weekends.
+Plots b and c show harmonies since each quarter include both weekdays and weekends.
 It can be seen in plot b that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
@@
 
-When $C_i$ is mapped to the $x$ position and $C_j$ to facets, then the $A_k$ levels are juxtaposed and each $B_\ell$ represent a group/facet.
+When $C_i$ is mapped to the $x$ position and $C_j$ to facets, then the $A_k$ levels are juxtaposed and each $B_\ell$ represents a group/facet.
 Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
-Hence, in this case the $A_k$'s are compared against each other within each group.
+Hence, in this case, the $A_k$'s are compared against each other within each group.
 With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels.
@@
 In this example, we will consider the granularities hour, day, week and month giving six cyclic granularities "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
-To these we add day-type ("wknd_wday") to capture weekend and weekday behavior.
+To these, we add day-type ("wknd_wday") to capture weekend and weekday behavior.
 Now that we have a list of cyclic granularities to look at, we can compute them using the results in Section&nbsp;\ref{sec:cyclic-calendar}.
@@
 Plot b uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
-Plot c uses a violin plot emphasize subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
+Plot c uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
 
@@
 Then it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends.
-Also, outliers are more prominent in the latter hours of the day.
+Also, outliers are more prominent in the later hours of the day.
 All of these indicate that looking at different distribution geometry or changing the mapping can shed light on different aspects of energy behavior for the same sample.
@@
 
-Exploratory data analysis involve many iterations of finding and summarizing patterns.
+Exploratory data analysis involves many iterations of finding and summarizing patterns.
 With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible granularities to explore.
@@
 We propose producing plots based on pairs of cyclic granularities that form harmonies rather than clashes or near-clashes.
-A future direction of work could be to further refine the selection of appropriate pairs of granularities by identifying those for which the differences between the displayed distributions is greatest, and rating these selected harmony pairs in order of importance for exploration.
+A future direction of work could be to further refine the selection of appropriate pairs of granularities by identifying those for which the differences between the displayed distributions is greatest and rating these selected harmony pairs in order of importance for exploration.
 

### 2021-05-01 reviews/response.Rmd
respone file spell check

@@
 I think that it is doable to explain the construction of the visualization without using this concept/framework, because the mappings are quite straight forward.
-I would like to see (a) either extend the introduction to grammar of graphics and the basics needed to argue in more detail about the embedding of your visualisations using this concept/framework, or (b) remove it.
+I would like to see (a) either extend the introduction to grammar of graphics and the basics needed to argue in more detail about the embedding of your visualizations using this concept/framework, or (b) remove it.
 
@@
 Q2.
-It is not clear from the paragraph in Section 5.1 what basic plot choices are implemented and used and supported for visualisation.
+It is not clear from the paragraph in Section 5.1 what basic plot choices are implemented and used and supported for visualization.
 There is only the list of plots to display descriptive statistics provided.
@@
 The later examples use primarily letter-value plots, then also box-plots, violin plots and line charts with area quantile plots.
-The letter-value plots are used in major examples of the paper, therefore I would argue to provide some more details on this visual representation and what the 5 colours and letters mean (e.g.
+The letter-value plots are used in major examples of the paper, therefore I would argue to provide some more details on this visual representation and what the 5 colors and letters mean (e.g.
 LV=letter value, M = median, F = fourths, E = eighths, D = sixteenths, C = thirty-seconds), as it is not as widely known like standard box-plots.
@@
 I do know a lot about different sports and also the basic idea and structure in many sports, but cricket was never on my radar.
-Apologise my ignorance about cricket.
+Apologize my ignorance about cricket.
 

### 2021-05-09 main.Rmd
figure 4 correction

@@
 
-(ref:allFig) Distribution of energy consumption displayed through letter value plots: plot a across month-of-year faceted by quarter-of-year; plot b across weekday/weekend faceted by quarter-of-year; plot c across quarter-of-year faceted by weekday/weekend.
-Plot a shows a clash since there are empty combinations.
-Plots b and c show harmonies since each quarter include both weekdays and weekends.
-It can be seen in plot b that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter where the letter values below D and E behave differently.
+(ref:allFig) Distribution of energy consumption displayed through letter value plots used to illustrate harmonies and clashes: (a) across month-of-year faceted by quarter-of-year; (b) across weekday/weekend faceted by quarter-of-year; (c) across quarter-of-year faceted by weekday/weekend.
+(a) shows a clash since there are empty combinations.
+(b) and (c) show harmonies since each quarter include both weekdays and weekends.
+It can be seen in (b) that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter.
+For example, the letter value D for weekdays mostly corresponds to letter value E for weekends, implying there are lower values of energy consumption during weekdays in the second quarter.
 This is probably because of lower temperatures in the second quarter compared to the first quarter (summer in Australia).
-Plot c switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+(c) switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+For example, the lower D and E letter values are most different across different quarters for weekdays and weekends respectively.
 <!-- For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends. -->

### 2021-05-09 scripts/main.R

@@ -73,7 +73,9 @@ scene1 <- VIC %>%
   theme_minimal() +
   theme(
     #panel.grid.major = element_line(colour = "#E0E0E0"),
-    panel.border = element_rect(colour = "#E0E0E0", fill = NA))
+    panel.border = element_rect(colour = "#E0E0E0",
+                                fill = NA),
+    panel.grid.major.x = element_blank())
 
 scene2 <- VIC %>%
   prob_plot(
@@ -112,16 +114,16 @@ scene3 <- VIC %>%
   theme(
     #panel.grid.major = element_line(colour = "#E0E0E0"),
     panel.border = element_rect(colour = "#E0E0E0",
-                                fill = NA),
-    panel.grid.major.x = element_blank())
+                                fill = NA))
+    #panel.grid.major.x = element_blank())
 
 gg_fig <- ggarrange(
   scene3,
   ggarrange(
     scene1, scene2,
-    ncol = 2, labels = c("b", "c")
+    ncol = 2, labels = c("(b)", "(c)")
   ),
-  nrow = 2, labels = "a"
+  nrow = 2, labels = "(a)", common.legend = TRUE
 )
 # label.y = "electricity demand [KWh]"
 

### 2021-05-10 main.Rmd
Improved wording

@@
 
-There are several ways to summarize a distribution like estimating empirical distribution or density of the variable's distribution, computing a few number summaries or quantiles.
-This data estimation or summarization could be potentially misleading if they are performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
+There are several ways to summarize the distribution of a data set such as estimating the empirical distribution or density of the data, or computing a few quantiles or other statistics.
+This estimation or summarization could be potentially misleading if it is performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
 Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
@@
 weekday/weekend.
-Plot a shows granularity 2 faceted by granularity 1; plots b and c shows the reverse mapping.
-Plot a allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
-Plots b and c examine the temporal trend of consumption over the course of a day, separately for the type of day.
-Plot b uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
-Plot c uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
+Plot **a** shows granularity 2 faceted by granularity 1; plots **b** and **c** shows the reverse mapping.
+Plot **a** allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
+Plots **b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.
+Plot **b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
+Plot **c** uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
 
@@
 We illustrate this with data from the sport cricket.
-Cricket is played with two teams of 11 players each, with each team taking turns batting and playing the field.
+Cricket is played with two teams of 11 players each, with each team taking turns batting and fielding.
 This is similar to baseball, wherein the _batsman_ and _bowler_ in cricket are analogous to a batter and pitcher in baseball.
-A _wicket_ is a structure with three sticks, stuck into the ground, placed at the end of the cricket pitch behind the batsman.
-A batsman tries to prevent the bowler from hitting the wicket by hitting the _ball_ using a _bat_.
-Two batsmen of a team are on the pitch at the same time.
-There are nine _fielders_ on the opposing team, with one player out of the 11 acting as the bowler, while another takes up the role of the _wicket-keeper_.
-To score a _run_, a batsman can run after hitting the ball and changing places on the pitch or also by hitting the ball out of the field on a fly (6 runs) or on a bounce (4 runs).
-A batsman is _dismissed_ when the period of batting is brought to an end either due to good bowling of the bowler or good fielding of the rest of the opposing team.
+A _wicket_ is a structure with three sticks, stuck into the ground at the end of the cricket pitch behind the batsman.
+One player from the fielding team acts as the bowler, while another takes up the role of the _wicket-keeper_ (similar to a catcher in baseball).
+The bowler tries to hit the wicket with a _ball_, and the batsman defends the wicket using a _bat_.
+At any one time, two of the batting team and all of the fielding team are on the field.
+The batting team aims to score as many _runs_ as possible, while the fielding team aims to successively _dismiss_ 10 players from the batting team.
+The team with the highest number of runs wins the match.
 
@@
 The Indian Premier League (IPL) is a professional T20 cricket league in India contested by eight teams representing eight different cities in India.
-The IPL ball-by-ball data is provided in the `cricket` data set in the `gravitas` package for a sample of 214 matches spanning 9 seasons (2008 to 2016) such that each over has 6 balls, each innings has 20 overs and each match has 2 innings.
+The IPL ball-by-ball data is provided in the `cricket` data set in the `gravitas` package for a sample of 214 matches spanning 9 seasons (2008 to 2016).
 
 Many interesting questions could be addressed with the `cricket` data set.
-For example, does the distribution of total runs vary depending on if a team bats in the first or second innings?
+For example, does the distribution of total runs depend on whether a team bats in the first or second innings?
 The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in the final playoffs from 2010 to 2015.
@@
 This seems to be unaffected by the over of the innings (the faceting variable).
-This might be because the new batsman needs to play himself in or the dismissals lead the (not-dismissed) batsman to adopt a more defensive playstyle.
+This might be because the new batsman needs to "play himself in" or the dismissals lead the (not-dismissed) batsman to adopt a more defensive playstyle.
 Run rates will also vary depending on which player is facing the next over and when the wicket falls in the previous over.
@@
 (ref:cricex) Runs per over shown with different distribution displays, and granularities.
-Plot a shows letter value plots across overs faceted by innings.
+Plot **a** shows letter value plots across overs faceted by innings.
 For the team batting in the first innings, there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
-Plot b shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
+Plot **b** shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
 This indicates that at least one wicket in the previous over leads to a lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.

### 2021-05-10 reviews/response.Rmd
Updated response

@@
 
+\newpage
+
 # Reviewer 2 comments
@@
 
-$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$ does not use the definition of $T_w$.
+$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$.
 For example, if $B$ is `day` and $G'$ is `month`, then $G'(0) = {B(0)\cup B(1)\cup \dots \cup B(30)}$, $G'(1) = {B(31)\cup B(32)\cup \dots \cup B(58)}$ corresponding to January and February.
@@
 Does it mean a finite number of irregular cycles?
-By Definition 4 of Bettini & De Sibi (2000), $E_1, .
-.
-.
-, E_z$ are the granularity exceptions, but the number of granularity exceptions does not have to be finite.
+By Definition 4 of Bettini & De Sibi (2000), $E_1, \dots, E_z$ are the granularity exceptions, but the number of granularity exceptions does not have to be finite.
 
@@
 "Number of periods" rephrased as "spans of time".
-Bettini & De Sibi (2000) states in P57 L17 while introducing quasi-periodic relationships that " except for a finite number of spans of time where they behave in an anomalous way."
+Bettini & De Sibi (2000) states in P57 L17 while introducing quasi-periodic relationships that "except for a finite number of spans of time where they behave in an anomalous way".
 
@@
 
-A3. Exactly equations (5) and (6) are to be utilized for this operation (with sides interchanged) and hence not added again.
+A3. Equations (5) and (6) are used for this operation (with sides interchanged) and hence are not added again.
 

### 2021-05-10 main.Rmd
made graph label consistent and deleted word plot

@@
 
-(ref:allFig) Distribution of energy consumption displayed through letter value plots used to illustrate harmonies and clashes: (a) across month-of-year faceted by quarter-of-year; (b) across weekday/weekend faceted by quarter-of-year; (c) across quarter-of-year faceted by weekday/weekend.
-(a) shows a clash since there are empty combinations.
-(b) and (c) show harmonies since each quarter include both weekdays and weekends.
-It can be seen in (b) that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter.
+(ref:allFig) Distribution of energy consumption displayed through letter value plots used to illustrate harmonies and clashes: **a** across month-of-year faceted by quarter-of-year; **b** across weekday/weekend faceted by quarter-of-year; **c** across quarter-of-year faceted by weekday/weekend.
+**a** shows a clash since there are empty combinations.
+**b** and **c** show harmonies since each quarter include both weekdays and weekends.
+It can be seen in **b** that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter.
 For example, the letter value D for weekdays mostly corresponds to letter value E for weekends, implying there are lower values of energy consumption during weekdays in the second quarter.
 This is probably because of lower temperatures in the second quarter compared to the first quarter (summer in Australia).
-(c) switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+**c** switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
 For example, the lower D and E letter values are most different across different quarters for weekdays and weekends respectively.
@@
 weekday/weekend.
-Plot **a** shows granularity 2 faceted by granularity 1; plots **b** and **c** shows the reverse mapping.
-Plot **a** allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
-Plots **b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.
-Plot **b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
-Plot **c** uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
+**a** shows granularity 2 faceted by granularity 1; **b** and **c** shows the reverse mapping.
+**a** allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
+**b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.**b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
+**c** uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
 
@@
 
-(ref:cricex) Runs per over shown with different distribution displays, and granularities.
-Plot **a** shows letter value plots across overs faceted by innings.
+(ref:cricex) Distributions of runs per over shown with different distribution displays, and granularities.
+**a** shows letter value plots across overs faceted by innings.
 For the team batting in the first innings, there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
-Plot **b** shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
+**b** shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
 This indicates that at least one wicket in the previous over leads to a lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.

### 2021-05-10 scripts/main.R

@@ -18,7 +18,7 @@ include_graphics("img/linear-ex.png")
 include_graphics("img/circular-ex.png")
 
 ## ----aperiodic-example
-include_graphics("img/aperiodic_revised.png")
+include_graphics("img/aperiodic-revised2.png")
 #include_graphics("img/aperiodic-ex3.png")
 
 ## ----tab-mayan
@@ -121,9 +121,9 @@ gg_fig <- ggarrange(
   scene3,
   ggarrange(
     scene1, scene2,
-    ncol = 2, labels = c("(b)", "(c)")
+    ncol = 2, labels = c("b", "c")
   ),
-  nrow = 2, labels = "(a)", common.legend = TRUE
+  nrow = 2, labels = "a", common.legend = TRUE
 )
 # label.y = "electricity demand [KWh]"
 

### 2021-05-12 main.Rmd
changed the layout of Figure 4 to fix aspect ratio

@@
 **a** allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
-**b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.**b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
+**b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.
+**b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
 **c** uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.

### 2021-05-12 scripts/main.R

@@ -4,6 +4,8 @@ library(readr)
 library(dplyr)
 library(ggplot2)
 library(tsibble)
+library(patchwork)
+library(lvplot)
 if (requireNamespace("gravitas") == FALSE) {
   remotes::install_github("Sayani07/gravitas")
 }
@@ -57,15 +59,14 @@ VIC <- gravitas::smart_meter10 %>%
   filter(customer_id == "10006486")
 
 scene1 <- VIC %>%
-  prob_plot(
-    "quarter_year", "wknd_wday",
-    response = "general_supply_kwh",
-    plot_type = "lv",
-    symmetric = FALSE
-  ) +
+  create_gran("quarter_year") %>% 
+  create_gran("wknd_wday") %>% 
+  ggplot(aes(x = wknd_wday, 
+             y = general_supply_kwh)) +
+  facet_wrap(~quarter_year, nrow = 1) +
+  geom_lv(aes(fill = ..LV..), k = 5) +
   ylab("") +
   xlab("") +
-  # xlab("weekend/weekday") +
   scale_fill_brewer(type = "seq", direction = -1) +
   theme(legend.position = "right") +
   ggtitle("") +
@@ -75,7 +76,31 @@ scene1 <- VIC %>%
     #panel.grid.major = element_line(colour = "#E0E0E0"),
     panel.border = element_rect(colour = "#E0E0E0",
                                 fill = NA),
-    panel.grid.major.x = element_blank())
+    panel.grid.major.x = element_blank(),
+    legend.position = "top")
+  
+
+
+# VIC %>%
+#   prob_plot(
+#     "quarter_year", "wknd_wday",
+#     response = "general_supply_kwh",
+#     plot_type = "lv",
+#     symmetric = FALSE
+#   ) +
+#   ylab("") +
+#   xlab("") +
+#   scale_fill_brewer(type = "seq", direction = -1) +
+#   theme(legend.position = "right") +
+#   ggtitle("") +
+#   scale_y_log10() +
+#   theme_minimal() +
+#   theme(
+#     #panel.grid.major = element_line(colour = "#E0E0E0"),
+#     panel.border = element_rect(colour = "#E0E0E0",
+#                                 fill = NA),
+#     panel.grid.major.x = element_blank(),
+#     legend.position = "top")
 
 scene2 <- VIC %>%
   prob_plot(
@@ -86,8 +111,7 @@ scene2 <- VIC %>%
   ) +
   ylab("") +
[…]

### 2021-05-12 main.Rmd
rephrase fig 4 caption

@@
 
-(ref:allFig) Distribution of energy consumption displayed through letter value plots used to illustrate harmonies and clashes: **a** across month-of-year faceted by quarter-of-year; **b** across weekday/weekend faceted by quarter-of-year; **c** across quarter-of-year faceted by weekday/weekend.
+(ref:allFig) Distribution of energy consumption displayed through letter value plots to illustrate harmonies and clashes and how change in x-axis and facet mappings change emphasis.
+**c** shown across month-of-year faceted by quarter-of-year is a clash, as there are empty sets.
+For example, November can never correspond to Quarter 1.
+**a** shown across quarter-of-year faceted by weekday/weekend do not include any empty sets and hence is a harmony.
+The same is true for **b**, which shows the same two cyclic granularities as **a** with reverse mapping.
+There are no empty combination since each quarter contains both weekdays and weekends.
+In **a**, it is easier to compare across weekday type (x-axis) within a quarter (facet); within quarter 2, more mass occupied the lower tail on the weekends (letter value E with depth 1/8) relative to that of the weekdays (letter value D 1/16).
+Coversely, **b** exaggerates differences across quarters (x-axis now) within a specific weekday type; within weekends, the lower letter values C and D for the first and forth quarter are higher than the other quarters.
+This is probably because of temperatures in those two quarters are higher as they correspond to summer and spring in Australia.
+
+<!-- ; **b** across weekday/weekend faceted by quarter-of-year; **c** across quarter-of-year faceted by weekday/weekend.
 **a** shows a clash since there are empty combinations.
@@
 It can be seen in **b** that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter.
-For example, the letter value D for weekdays mostly corresponds to letter value E for weekends, implying there are lower values of energy consumption during weekdays in the second quarter.
+-->
+<!-- For example, the letter value D for weekdays mostly corresponds to letter value E for weekends, implying there are lower values of energy consumption during weekdays in the second quarter.
 This is probably because of lower temperatures in the second quarter compared to the first quarter (summer in Australia).
@@
 For example, the lower D and E letter values are most different across different quarters for weekdays and weekends respectively.
+-->
 <!-- For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends. -->

### 2021-05-12 main.Rmd
Fig 3 caption revised

@@
 
-(ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated through (a) linear and (b) stacked displays of time.
-The linear display shows granularities days ($B$), weeks ($G$), semester days ($B'$), and stages of a semester ($M$) distributed over linear time.
-Granules of $B'$ is only defined for days when the semester is running.
-Here a semester lasts for 18 weeks and 2 days, each consisting of 5 stages.
-It starts with an orientation stage, followed by a first stage of the in-session period (6 weeks), a break (1 week), a second stage of the in-session period (7 weeks), a 1-week study break stage, before final exams stage which continue for 16 days.
-This distribution of semester days within a semester remains the same for every semester and hence $Q_{B', M}$ with $P$ = 128 days will be a quasi-circular granularity with repeating patterns.
-$A_{B, M}$ will be an aperiodic cyclic granularity since
-the placement of the semester within a year varies across years and there is no definite number of days after which the semester begins every year.
+(ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated by linear (a) and 
+stacked-displays (b) progression of time.
+The linear display shows the granularities days ($B$), weeks ($G$), semester days ($B'$), and stages of a semester ($M$) indexed over a linear representation of time.
+The granules of $B'$ is only defined for days when the semester is running.
+Here a semester spans 18 weeks and 2 days, and consists of 6 stages.
+It starts with a week of orientation, followed by an in-session period (6 weeks), a break (1 week), the second half of semester (7 weeks), a 1-week study break before final exams, which spans the next 16 days.
+This distribution of semester days remains relatively similar for every semester.
+$Q_{B',M}$ with P = 128 is a quasi-circular granularity with repeating patterns, while $A_{B,M}$ is an aperiodic cyclic granularity as the placement of the semester within a year varies from year to year with no fixed start and end dates.
 

### 2021-05-13 main.Rmd
captions for last three figures re-arranged

@@
 
-(ref:allFig) Distribution of energy consumption displayed through letter value plots to illustrate harmonies and clashes and how change in x-axis and facet mappings change emphasis.
-**c** shown across month-of-year faceted by quarter-of-year is a clash, as there are empty sets.
-For example, November can never correspond to Quarter 1.
-**a** shown across quarter-of-year faceted by weekday/weekend do not include any empty sets and hence is a harmony.
-The same is true for **b**, which shows the same two cyclic granularities as **a** with reverse mapping.
-There are no empty combination since each quarter contains both weekdays and weekends.
-In **a**, it is easier to compare across weekday type (x-axis) within a quarter (facet); within quarter 2, more mass occupied the lower tail on the weekends (letter value E with depth 1/8) relative to that of the weekdays (letter value D 1/16).
-Coversely, **b** exaggerates differences across quarters (x-axis now) within a specific weekday type; within weekends, the lower letter values C and D for the first and forth quarter are higher than the other quarters.
-This is probably because of temperatures in those two quarters are higher as they correspond to summer and spring in Australia.
+(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empy sets and white space.
+Placement within a facet should be done for primary comparisons.
+For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet).
+It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E with depth 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
 
@@
 
-(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements.
-Two granularities are used: 1.
-hour of the day; and 2.
-weekday/weekend.
-**a** shows granularity 2 faceted by granularity 1; **b** and **c** shows the reverse mapping.
-**a** allows comparison of usage by workday within each hour of the day using side-by-side boxplots, showing that on work days there is more consumption early in the day.
-**b** and **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.
-**b** uses an area quantile to emphasize the time component; e.g., median consumption shows prolonged high usage in the morning on weekdays.
-**c** uses a violin plot emphasizes subtler distributional differences across hours, showing that morning use on weekdays is bimodal, and on some work days there is low usage possibly indicating the person is working from home or having a late start.
+(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements: hour of the day; and weekday/weekend.
+**a** The side-by-side boxplots make the comparison between day types easier, and suggest that there is generally lower energy use on the weekend.
+Interestingly, this is the opposite to what might be expected.
+Plots **b**, **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.
+The area quantile emphasizes time, and indicates that median consumption shows prolonged high usage in the morning on weekdays.
+The violin plot emphasizes subtler distributional differences across hours: morning use on weekdays is bimodal.
 
@@
 
-(ref:cricex) Distributions of runs per over shown with different distribution displays, and granularities.
-**a** shows letter value plots across overs faceted by innings.
-For the team batting in the first innings, there is an upward trend of runs per over, while there is no such pattern of runs for the teams batting in the second innings.
-**b** shows quantile plots of runs per over across an indicator of wickets in the previous over faceted by current over.
-This indicates that at least one wicket in the previous over leads to a lower median run rate and quartile spread in the subsequent over, regardless of the over of the innings.
+(ref:cricex) Examining distribution of runs per by innings, over of the innings and number of wickets in previous innings.
+Plot **a** displays distribution using letter value plots.
+A gradual upward trend in runs per over can be seen in innings 1, which is not present in innings 2.
+Plot **b** shows quantile plots of runs per over across an indicator of wickets in the previous over, faceted by current over.
+When a wicket occurred in the previous over, the runs per over tends to be lower throughout the innings.
 

### 2021-05-13 main.Rmd
changed fig.pos

--- 2021-05-12 main.Rmd
@@
 
-```{r linear-time, echo=FALSE, fig.cap="(ref:linear-time)", fig.pos="!htb"}
+```{r linear-time, echo=FALSE, fig.cap="(ref:linear-time)", fig.pos="!htbp"}
 ```
@@
 
-```{r allFig, fig.pos="!htb", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
+```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
 ```
@@
 
-```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!htb"}
+```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!p"}
 ```

### 2021-05-13 main.Rmd
spell check

@@
 
-(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empy sets and white space.
+(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space.
 Placement within a facet should be done for primary comparisons.

### 2021-05-16 main.Rmd
example of quasi rephrased - within each granule not period

@@
 With two linear granularities $G$ and $H$, if $G$ *groups into* or *finer than* $H$ then $G$ is of lower order than $H$.
-For example, if the bottom granularity is hour, then granularity day will have lower order than week since each day consist of fewer hours than each week.
+For example, if the bottom granularity is hour, then granularity $day$ will have lower order than $week$ since each day consist of fewer hours than each week.
 
@@
 
-For example, day-of-year is quasi-periodic with either 365 or 366 granules of $B$ (days) within each period of $G'$ (years).
+For example, day-of-year is quasi-periodic with either 365 or 366 granules of $B$ (days) within each granule of $G'$ (years).
 The pattern repeats every $4$ years (ignoring leap seconds).

### 2021-05-16 main.Rmd
where $z$ is the index of the bottom granularity $kin$
finite number of spans of time

@@
 
-Granularities can also be periodic with respect to other granularities, _"except for a finite span of time where they behave in an anomalous way"_; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+Granularities can also be periodic with respect to other granularities, _"except for a finite number of spans of time where they behave in an anomalous way"_; these are called **quasi-periodic** relationships [@Bettini2000-vy].
 In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
@@
 \end{align*}
+
+, where $z$ is the index of the bottom granularity $kin$.
 

### 2021-05-16 main.Rmd
replace EDA by “exploratory data analysis (EDA)”,

@@
 It is important to be able to navigate through all of these time granularities to have multiple perspectives on the periodicity of the observed data.
-This aligns with the notion of EDA [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+This aligns with the notion of exploratory data analysis (EDA) [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
 Visualizing probability distributions conditional on one or more granularities is an indispensable tool for exploration.

### 2021-05-16 main.Rmd
(a) and (b) for both circular and quasi-circular figures

@@
 
-(ref:circular-dow) a. Index sets for some linear and circular granularities. b. Circular granularities can be constructed by slicing the linear granularity into pieces and stacking them.
+(ref:circular-dow) Index sets for some linear and circular granularities (a). Circular granularities can be constructed by slicing the linear granularity into pieces and stacking them (b).
 

### 2021-05-16 main.Rmd
The term “time instants” briefly described

@@
 \begin{defn}\label{def:definition}
-A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
+A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants (equivalently, moments or points) and $\le$ is a total order on $T$.
 \end{defn}

### 2021-05-16 scripts/main.R
especially in Figure 3 a darker reddish to better distinguish the colours - removed redundant figs

@@ -20,7 +20,7 @@ include_graphics("img/linear-ex.png")
 include_graphics("img/circular-ex.png")
 
 ## ----aperiodic-example
-include_graphics("img/aperiodic-revised2.png")
+include_graphics("img/aperiodic-ex.png")
 #include_graphics("img/aperiodic-ex3.png")
 
 ## ----tab-mayan

### 2021-05-16 main.Rmd
all figures are still interpretable in black and white.

@@
 For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet).
-It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E with depth 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
+It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
 
@@
 
-In \autoref{fig:bothcust}b, the black line is the median, whereas the purple band covers the 25th to 75th percentile, the orange band covers the 10th to 90th percentile, and the green band covers the 1st to 99th percentile.
+In \autoref{fig:bothcust}b, the black line is the median, whereas the purple (narrow) band covers the 25th to 75th percentile, the orange (middle) band covers the 10th to 90th percentile, and the green (broad) band covers the 1st to 99th percentile.
 The first facet represents the weekday behavior while the second facet displays the weekend behavior; energy consumption across each hour of the day is shown inside each facet.

### 2021-05-16 scripts/main.R
remove redundant code

@@ -81,27 +81,6 @@ scene1 <- VIC %>%
   
 
 
-# VIC %>%
-#   prob_plot(
-#     "quarter_year", "wknd_wday",
-#     response = "general_supply_kwh",
-#     plot_type = "lv",
-#     symmetric = FALSE
-#   ) +
-#   ylab("") +
-#   xlab("") +
-#   scale_fill_brewer(type = "seq", direction = -1) +
-#   theme(legend.position = "right") +
-#   ggtitle("") +
-#   scale_y_log10() +
-#   theme_minimal() +
-#   theme(
-#     #panel.grid.major = element_line(colour = "#E0E0E0"),
-#     panel.border = element_rect(colour = "#E0E0E0",
-#                                 fill = NA),
-#     panel.grid.major.x = element_blank(),
-#     legend.position = "top")
-
 scene2 <- VIC %>%
   prob_plot(
     "wknd_wday", "quarter_year",
@@ -142,21 +121,6 @@ scene3 <- VIC %>%
   theme(legend.position = "none",
         axis.text.x=element_text(angle = 90, vjust = 0.5)) 
     #panel.grid.major.x = element_blank())
-
-# gg_fig <- ggarrange(
-#   scene3,
-#   ggarrange(
-#     scene1, scene2,
-#     ncol = 2, labels = c("b", "c")
-#   ),
-#   nrow = 2, labels = "a", common.legend = TRUE
-# )
-
-# 
-# gg_fig <- scene1/ (scene2 + scene3) 
-# 
-# gg_fig + annotate(label = "something")
-#   
 
 gg_fig <- ggarrange(
    scene1,

### 2021-06-22 main.Rmd
Table 1: minute-of-day (C_2 not C_j)

@@
 minute-of-day                                &
-  $C_j = z \mod 60*24$                  &
+  $C_2 = z \mod 60*24$                  &
   $P_2= `r 60*24`$\\

### 2021-06-22 main.Rmd
Multiple-to-single order-up section (Equation 7 not only 7)

@@
 Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
-For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities can be obtained using \eqref{eq:all-circular-multiple}.
+For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities can be obtained using Equation \eqref{eq:all-circular-multiple}.
 \begin{equation} \label{eq:all-circular-multiple}

### 2021-06-22 main.Rmd
Earlier: Morning use on weekdays is bimodal (Fig 5)Revised: Morning use is bimodal
bibliography Q1 and Q2 UCGS-production fix

@@
 
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@Grolemund2018-po). In particular, we
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@wickham2016r). In particular, we
 
@@
 For displaying the distribution of a continuous univariate variable, many options are available.
-Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variants such as notched box plots [@McGill1978-hg] or other variations as mentioned in @Wickham2012-vr.
+Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variants such as notched box plots [@McGill1978-hg] or other variations as mentioned in @boxplots.
 They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.
@@
 The area quantile emphasizes time, and indicates that median consumption shows prolonged high usage in the morning on weekdays.
-The violin plot emphasizes subtler distributional differences across hours: morning use on weekdays is bimodal.
+The violin plot emphasizes subtler distributional differences across hours: morning use is bimodal.
 
